### 2019-07-10 paper.tex
Add files via upload

% !TeX TXS-program:compile = txs:///pdflatex/[--shell-escape]
%
% acm templates v1.57 (sigconf template)
%
%\documentclass[sigconf, screen, review, anonymous, natbib=false]{acmart}
\documentclass[sigconf, screen, natbib=false, dvipsnames, table]{acmart}
\settopmatter{printacmref=true}
% mandatory for CCS'19
\usepackage{balance}
% for creating a balanced last page (usually last page with references)

% defining the \BibTeX command - from Oren Patashnik's original BibTeX documentation.
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08emT\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

% Rights management information. 
% This information is sent to you when you complete the rights form.
% These commands have SAMPLE values in them; it is your responsibility as an author to replace
% the commands and values with those provided to you when you complete the rights form.
%
% These commands are for a PROCEEDINGS abstract or paper.

\copyrightyear{2019} 
\acmYear{2019} 
\setcopyright{acmcopyright}
\acmConference[CCS '19]{2019 ACM SIGSAC Conference on Computer and Communications Security}{November 11--15, 2019}{London, United Kingdom}
\acmBooktitle{2019 ACM SIGSAC Conference on Computer and Communications Security (CCS '19), November 11--15, 2019, London, United Kingdom}
\acmPrice{15.00}
\acmDOI{10.1145/3319535.3339818}
\acmISBN{978-1-4503-6747-9/19/11}

%
% These commands are for a JOURNAL article.
%\setcopyright{acmcopyright}
%\acmJournal{TOG}
%\acmYear{2018}\acmVolume{37}\acmNumber{4}\acmArticle{111}\acmMonth{8}
%\acmDOI{10.1145/1122445.1122456}

%
% Submission ID. 
% Use this when submitting an article to a sponsored event. You'll receive a unique submission ID from the organizers
% of the event, and this ID should be used as the parameter to this command.
%\acmSubmissionID{123-A56-BU3}

%
% The majority of ACM publications use numbered citations and references. If you are preparing content for an event
% sponsored by ACM SIGGRAPH, you must use the "author year" style of citations and references. Uncommenting
% the next command will enable that style.
%\citestyle{acmauthoryear}

 \makeatletter
 % \renewcommand{\section}{\abovedisplayskip 10\p@ \@plus3\p@ \@minus1\p@%
 %                       \belowdisplayskip 5\p@ \@plus3\p@ \@minus1\p@%
 %                       \abovedisplayshortskip 0pt \@plus2\p@%
 %                       \belowdisplayshortskip 0pt \@plus2\p@ \@minus0\p@%
 %                       \@startsection{section}{1}{\z@}%
 %                        {-17\p@ \@plus -4\p@ \@minus -4\p@}%
 %                        {6\p@ \@plus 4\p@ \@minus 4\p@}%
 %                        {\normalfont\large\bfseries\boldmath
 %                         \rightskip=\z@ \@plus 8em\pretolerance=10000 }}
 \renewcommand{\subsection}{\@startsection{subsection}{2}{\z@}%
                        {-8\p@ \@plus -4\p@ \@minus -4\p@}%
                        {5\p@ \@plus 2\p@ \@minus 2\p@}%
                        {\normalfont\Large\bfseries\boldmath
                         \rightskip=\z@ \@plus 3em\pretolerance=10000 }}
  \renewcommand{\subsubsection}{\@startsection{subsubsection}{3}{\z@}%
                        {-6\p@ \@plus -4\p@ \@minus -4\p@}%
                        {1\p@ \@plus 1\p@ \@minus 0\p@}%
                        {\normalfont\normalsize\bfseries\boldmath}}
% \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\z@}%
%                       {-8\p@ \@plus -4\p@ \@minus -4\p@}%
%                       {-2\p@ \@plus -0.22em \@minus -0.1em}%
%                       {\normalfont\normalsize\bfseries}}
 \makeatother

% *** GRAPHICS RELATED PACKAGES ***
\usepackage{graphicx}
\usepackage{comment}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{listings}
% declare the path(s) where your graphic files are
\graphicspath{{../figures/}}
% and their extensions so you won't have to specify these with
% every instance of \includegraphics
\DeclareGraphicsExtensions{.pdf,.jpeg,.png}

%\usepackage[outputdir=/Volumes/ramdisk/]{minted}
\usepackage{minted}
% *** MATH PACKAGES ***
\usepackage{amsmath}

% *** SPECIALIZED LIST PACKAGES ***
\usepackage{algorithmic}

% *** ALIGNMENT PACKAGES ***
\usepackage{array}

\usepackage{url}

\usepackage[utf8]{inputenc}
\usepackage[american]{babel}
\usepackage[backend=biber,style=ACM-Reference-Format]{biblatex}
\addbibresource{../library.bib} 
\addbibresource{../cryptobib/abbrev2.bib}
\addbibresource{../cryptobib/crypto_crossref.bib}

% Used for Theorems and Definitions
\usepackage{amsthm}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{definition}
\newtheorem{remark}{Remark}

\usepackage{xcolor} 
\usepackage{xspace}
\usepackage{xstring} % required by IfEqCase

\usepackage{booktabs} % better tables

% ishaq: adding our macros
\input{macros.tex}

\sloppy
\begin{document}
\fancyhead{}
% do not delete this code.

%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title[Scheduling and Parallelization for MPC]{Scheduling and Parallelization for MPC}

%
% The "author" command and its associated commands are used to define the authors and their affiliations.
% Of note is the shared affiliation of the first two authors, and the "authornote" and "authornotemark" commands
% used to denote shared contribution to the research.
\author{Muhammad Ishaq}
\authornote{This work was done in part while the author was at RPI.}
\email{m.ishaq@ed.ac.uk}
%\orcid{1234-5678-9012}
%\author{G.K.M. Tobin}
%\authornotemark[1]
%\email{webmaster@marysville-ohio.com}
\affiliation{%
    \institution{University of Edinburgh}
    %\streetaddress{P.O. Box 1212}
    \city{Edinburgh}
    \state{Scotland}
%    \postcode{43017-6221}
}

\author{Lindsey Kennard}
\email{kennal@rpi.edu}
\affiliation{%
    \institution{Rensselaer Polytechnic Institute}
    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
    \city{Troy}
    \state{New York}
    %\country{Iceland}
}

\author{Ana L. Milanova}
\email{milanova@cs.rpi.edu}
\affiliation{%
    \institution{Rensselaer Polytechnic Institute}
    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
    \city{Troy}
    \state{New York}
    %\country{Iceland}
}

\author{Vassilis Zikas}
\authornote{This work was done in part while the author was visiting UCLA and supported in part by DARPA and SPAWAR under contract N66001-15-C-4065 and by a SICSA Cyber Nexus Research Exchanges grant.}
\email{vzikas@inf.ed.ac.uk}
%\orcid{1234-5678-9012}
%\author{G.K.M. Tobin}
%\authornotemark[1]
%\email{webmaster@marysville-ohio.com}
\affiliation{%
    \institution{University of Edinburgh} 
    %\streetaddress{P.O. Box 1212}
    \city{Edinburgh}
    \state{Scotland}
    %    \postcode{43017-6221}
}

%
% By default, the full list of authors will be used in the page headers. Often, this list is too long, and will overlap
% other information printed in the page headers. This command allows the author to define a more concise list
% of authors' names for this purpose.
%\renewcommand{\shortauthors}{Trovato and Tobin, et al.}

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}

\end{abstract}

%
% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
% Please copy and paste the code instead of the example below.
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10003752.10010124.10010138.10010143</concept_id>
<concept_desc>Theory of computation~Program analysis</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003752.10003777.10003789</concept_id>
<concept_desc>Theory of computation~Cryptographic protocols</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10002978.10002979</concept_id>
<concept_desc>Security and privacy~Cryptography</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Theory of computation~Program analysis}
\ccsdesc[500]{Theory of computation~Cryptographic protocols}
\ccsdesc[300]{Security and privacy~Cryptography}

%
% Keywords. The author(s) should pick words that accurately describe the work being
% presented. Separate the keywords with commas.
\keywords{protocol mixing; linear programming; multiparty computation; program analysis; cryptography}

\begin{comment}
%
% A "teaser" image appears between the author and affiliation information and the body 
% of the document, and typically spans the page. 
\begin{teaserfigure}
    \includegraphics[width=\textwidth]{sampleteaser}
    \caption{Seattle Mariners at Spring Training, 2010.}
    \Description{Enjoying the baseball game from the third-base seats. Ichiro Suzuki preparing to bat.}
    \label{fig:teaser}
\end{teaserfigure}
\end{comment}
%
% This command processes the author and affiliation and title information and builds
% the first part of the formatted document.
\maketitle

\section{Introduction}
\label{sec:introduction}

In this short paper we define the problem of \emph{optimal parallelization} for Multi-party Computation (MPC), 
set it apart from classical work on parallelization for High-performance Computing (HPC), present preliminary 
results, and set directions for future research.

\ana{1. Intro to MPC (from a PL point of view). Define the problem.}

\ana{2. Difference with HPC. Why "optimal HPC" = "minimal delay" is NOT "optimal MPC"}

\ana{3. Results statement.}

\ana{4. Future directions.}

% no \IEEEPARstart

\section{Scheduling in MPC}

For this treatment we make the following simplifying assumptions:

\begin{enumerate}
\item All statements in the program execute using the same protocol (sharing). That is, there is no share conversion.
\item All MPC instructions have the same unit cost, 1 unit.
\item There is unlimited bandwidth---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely 1 unit.
\item MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption, 
2 MUL instructions scheduled in parallel benefit from amortization and cost 1, however a MUL and a MUX instructions scheduled 
in parallel still cost 2.
\end{enumerate}

\paragraph{Problem 1:} Given a dependence graph of MPC instructions, find the \emph{loop-body} schedule 
with \emph{minimal cost}.

%\begin{figure*}[tbhp]
\begin{figure*}[tbhp]
\begin{tabular}{llll}
\begin{minipage}[b]{4.25cm}

\begin{minted}[fontsize=\footnotesize, linenos, numbersep=5pt, escapeinside=||]{java}
for (int i=1; i<=N; i++) {
  A: A[i] = f_1(B[i]);
  B: B[i] = f_2(A[i],D[i-1]);
  C: C[i] = f_3(A[i],D[i-1]);
  D: D[i] = f_4(B[i],C[i]);
}
\end{minted}
\end{minipage}

&

\begin{minipage}[b]{4.25cm}
\includegraphics[width=0.5\textwidth]{figs/dependenceGraph.pdf}
\end{minipage}

&

\begin{minipage}[b]{4.25cm}
\includegraphics[height=0.15\textheight]{figs/MPCloop.pdf}
\end{minipage}

&

\begin{minipage}[b]{4.25cm}
\includegraphics[height=0.20\textheight]{figs/MPCaccrossLoop.pdf}
\end{minipage}
\\

(a) Source 
&
(b) Dependence graph
&
(c) Loop-body schedule
&
(d) Across-loop schedule

\end{tabular}
\caption{(a) shows the source code, taken from Aiken and Nicolau~\cite{Aiken:1988}, 
(b) shows the dependence graph. The solid edges denote within-iteration ordering constraints, e.g., A must run before B
because B depends on A. The dashed edges denote across-iteration ordering constraints, e.g., B(2), i.e., B in iteration 2
depends on D(1), i.e., D in iteration 1. (c) shows the optimal loop-body schedule, and (d) shows the
optimal across-loop schedule. A(1), B(1), etc. denote instruction A in iteration 1, etc.}
\label{fig:MPCexample}\vspace{-2ex}
\end{figure*}

\paragraph{Problem 2:} Given a dependence graph of MPC instructions representing a loop body (or given a loop-body schedule), 
find the \emph{across-loop} schedule with  \emph{minimal cost}.

\figref{fig:MPCexample} illustrates the problem with a code example from the classical high-performance computing (HPC) literature. 
\figref{fig:MPCexample}(c) shows the optimal loop-body schedule (answer to Problem 1). Instructions B and C can be scheduled in parallel, 
however, the schedule does not necessarily benefit form parallelization/amortization. If B and C (functions $f_1$ and $f_2$) are the \emph{same}
MPC instruction (e.g., both are MUX, MUL, ADD, or CMP, etc.), then they \emph{can} benefit from parallelization/amortization. Given
our assumptions stated above, the cost of running the loop-body schedule is 3. However, if B and C are different MPC-instructions,
e.g., B is MUL, but C is MUX, then the cost of running the schedule is 4.

\figref{fig:MPCexample}(d) shows the optimal across-loop schedule (answer to Problem 2). We can schedule all A-instructions, across all iterations, 
in parallel, as A depends only on the initial value of B. We cannot do better but schedule B, C, and D sequentially, as they each depend on values
computed in the previous iteration. Again, if B and C are the same instruction, the optimal across-loop schedule will have cost of $1 + 2N$, 
and if B and C are NOT the same instruction, then it will have cost $1+3N$.

\section{Scheduling in HPC}

In this section, we state the classical HPC scheduling problem, and argue the MPC scheduling problem poses unique challenges.

Similar problems arise in the HPC setting.

%\begin{figure*}[tbhp]
\begin{figure*}[tbhp]
\begin{tabular}{lll}

\begin{minipage}[b]{4.25cm}
\includegraphics[width=0.4\textwidth]{figs/HPCloop.pdf}
\end{minipage}

&

\begin{minipage}[b]{4.25cm}
\includegraphics[width=0.4\textwidth]{figs/HPCdoacrossLoop.pdf}
\end{minipage}

&

\begin{minipage}[b]{4.25cm}
\includegraphics[width=0.6\textwidth]{figs/HPCgreedyLoop.pdf}
\end{minipage}

\\

(a) Loop-body schedule
&
(b) Doacross across-loop schedule
& 
(c) Greedy across-loop schedule

\end{tabular}
\caption{Optimal HPC schedules.}
\label{fig:HPCexample}\vspace{-2ex}
\end{figure*}

\paragraph{Problem 1:} Given a loop body (i.e., dependence graph), find the loop-body schedule with \emph{minimal delay}.

\paragraph{Problem 2:} Given a loop body (or a loop-body schedule), find the \emph{doacross} schedule with minimal delay.

\paragraph{Problem 3:} Given a loop body (or a loop-body schedule), find the \emph{greedy schedule} with minimal delay.

\figref{fig:HPCexample} illustrates the problem in HPC. \figref{fig:HPCexample}(a) shows the optimal loop-body schedule
which results in cost of 3. Again, we assume that all instructions have the same cost, a unit of 1. The key difference with the 
MPC schedule is that regardless of whether B and C are the same instruction opcode or not, B and C can be scheduled 
to run in parallel and the overall running time benefits from parallelization.

The HPC literature has considered different kinds of across-loop scheduling. One scheduling, called \emph{doacross}~\cite{aiken,cytron},
requires that all instructions in an iteration are scheduled on the same processor. The optimal doacross schedule for our example
is shown in \figref{fig:HPCexample}(b) (taken directly from~\cite{aiken}). Aiken and Nicolau proposed a new schedule,
which they called \emph{greedy} schedule, where instructions in an iteration can be scheduled on different processors. 
The optimal greedy schedule for our example, which achieves better minimal delay than the optimal doaccross schedule, 
is shown in \figref{fig:HPCexample}(c) (again, taken directly from~\cite{aiken}).

\ana{Here we have to make the argument that MPC is different from HPC.}
The example illustrates that optimal MPC scheduling and optimal HPC scheduling are different problems.

%\section{\bf Preliminaries}
%\label{sec:preliminaries}

\ana{Relevant papers/books:

\begin{enumerate}

\item Book: "Scheduling and automatic parallelization", Springer 2000, by Alain Darte, Yves Robert and Frederic Vivien. 
This book describes scheduling and parallelization in HPC: given a dependence graph and unlimited 
number of processors, one can compute an optimal (minimal delay) schedule in polynomial time. 
Given a finite number of processors, computing an optimal schedule becomes NP-hard.

\item https://theory.stanford.edu/~aiken/publications/papers/pldi88.pdf. "Optimal loop parallelization", PLDI 1988, by
Alex Aiken and Alex Nicolau. 
Given a dependence graph of the loop body, computes an optimal parallelization across loop iterations. 
The optimality metric is minimal delay. E.g., assuming 2 iterations, the delay is measured by the number 
of instructions in iteration 2 that are lagging behind iteration 1.

\item https://www.cs.indiana.edu/~achauhan/Teaching/B629/2006-Fall/CourseMaterial/1998-popl-knobe-arrayssa.pdf.
"Array SSA", POPL 1998, by Kathleen Knobe and Vivek Sarkar. This paper defines an SSA form for arrays which I think 
is crucial for parallelization across loop iterations.

\end{enumerate}
}

\section{Loop Body Parallelization}
\label{sec:parallelization_within_loop}
%\input{../sections/analysis.tex}

We first consider the answer to Problem 1 (in MPC). 
We argue that optimal parallelization within a loop body is NP-complete (under the above stated assumptions). 
Therefore, compilers must resort to heuristics to compute a schedule for the instructions within a loop body.

We consider two operations, call them $A$ and $M$. Each instruction in the program is either an $A$-instruction or
an $M$-instruction. In order to benefit from parallelization/amortization, we must schedule two or more 
$A$-instructions in the same parallel node (or two or more $M$-instructions in the same parallel node). 
Scheduling $A$-instructions in parallel with $M$-instruction does not benefit from amortization.
It incurs the exact same cost as scheduling the $A$-instructions in a node $P_A$, scheduling the $M$-instructions 
in a node $P_M$, and having $P_A$ precede $P_M$ in the parallel schedule. This is the difference 
between classical scheduling, as studied in parallel computing, and MPC scheduling.

We make the following assumptions:

\begin{enumerate}

\item $A$ and $M$ are of equal cost, 1 unit.
\item There is unlimited bandwidth---i.e., a single $A$-instrution (or $M$-instruction) costs as much as $N$ amortized $A$-instructions 
(or $M$-instructions), namely 1 unit.

\end{enumerate}

Consider a loop body that consists of $n$ sequences: $S_1$, ... $S_n$ of $A$ and $M$ instructions. 
More precisely, the loop body is such that its instructions can be grouped into such sequences. 
$S_1$, ... $S_n$ can execute in parallel, however, all instructions within a sequence must 
execute sequentially. For example, consider the three sequences (the right arrow indicates a \emph{dependence},
meaning that the source node must execute before the target node): 
\begin{enumerate}
\item $A \rightarrow M \rightarrow A$
\item $A \rightarrow A \rightarrow A$
\item $M \rightarrow A \rightarrow M$
\end{enumerate}

A \emph{schedule} $P: P_1 \rightarrow P_2 \dots \rightarrow P_k$ is such that for each sequence 
$S_i$ in the set, if $S_i[k]$ precedes $S_i[k']$ in $S_i$ then $S_i[k]$ is scheduled in node $P_l$, $S_i[k]$ 
is scheduled in node $P_{l'}$, and $P_l$ precedes $P_{l'}$ in $P$.

The cost of a schedule $P$ is 
\[\mathit{cost}(P) = \sum_{i=1}^k \mathit{cost}(P_i)\]
where $\mathit{cost}(P_i) = 1$ if $P_i$ consists of $A$-instructions only, or $M$-instructions only, 
and $\mathit{cost}(P_i) = 2$ if $P_i$ mixes $A$-instructions and $M$-instructions.

The problem is to find a schedule $P$ with \emph{minimal cost}. For example, 
a schedule with minimal cost for the sequences above is
\[ A(1), A(2) \rightarrow M(1), A(2), M(3) \rightarrow A(1), A(2), A(3) \rightarrow M(3) \]
(The parentheses above indicate the sequence where the instruction comes from: (1), (2), or (3).)
The cost of this schedule is 5.

The problem of finding a schedule with minimal cost 
easily reduces to the problem of finding a \emph{shortest common supersequence}, 
a known NP-complete problem \ana{citation needed}. To see the reduction, 
suppose $P$ is a schedule with minimal cost (computed by a black-box algorithm). 
We can derive a schedule $P'$ with the same cost as $P$, by mapping each mixed node $P_i \in P$ 
to two consecutive nodes in $P'$: an $A$-instruction node followed by an $M$-instruction node.
Clearly, $P'$, which now is a sequence of $A$'s and $M$'s, is a supersequence of each sequence 
$S_i$, i.e., $P'$ is a common supersequence 
of $S_1 \dots S_n$. It is also a shortest common supersequence. (To see this, suppose, a 
shorter common supersequence, $P''$, exists. $P''$ is a schedule of $S_1 \dots S_n$
and the $\mathit{cost}(P'')$ equals the length of $P''$. Since $P'$ is longer than
$P''$, and $\mathit{cost}(P') = \mathit{cost}(P)$, that means $\mathit{cost}(P'') < \mathit{cost}(P)$, 
which is a contradiction since $P$ is a schedule with minimal cost.)

\ana{Question: are these two assumptions too strong?}

\ana{Assumption 1 can be relaxed --- e.g., if $A$ is 5 times cheaper than $M$, 
we can construct a program where all $A$'s are sequences of five consecutive 
$A$'s, which reduces to the previous problem.}

\ana{Assumption 2, I am not sure if this is acceptable, or maybe it is too strong. 
My intuition is that incorporating into the cost a dependence on the number 
of instructions in each parallel node, is going to make the problem harder, not easier...}

\section{Optimal Loop Parallelization} 
\label{sec:optimal_loop_parallelization}
%\input{../sections/optimal_loop_parallelization.tex}

\ana{Aiken's and Array SSA to define the notion of optimal loop parallelization for MPC.}

We now consider the answer to Problem 2 (in MPC).

\subsection{Problem Overview}

Recall Problem 2, given a loop-body schedule, compute the across-loop schedule
with minimal cost. The following examples illustrate. The first example is the inner 
product computation, and the second is Aiken and Nicolau's example.

\begin{tabular}{lll}
\begin{lstlisting}[language=Java]
A = ...
B = ...
ip = 0;
for (i=1; i<=N; i++) {
  t=A[i]*B[i];
  ip=ip+t;
}

\end{lstlisting}
& 
~
&
\begin{lstlisting}[language=Java]
A = ...
B = ...
ip = 0;
for (i=1; i<=N; i++) {
  if (i%step5==1) para(t[i],...,t[i+step])
  if (i%step6==1) ip = ip + t[i];
}  
\end{lstlisting}

\end{tabular}

\ana{More to introduce what the schedule is.}

In the above example {\sf step5} is N, meaning that all N multiplications are executed in parallel.
In contrast, {\sf step6} is 1, meaning that the summation statements are executed sequentially.

A more interesting example involves array reads and writes. We return to the Aiken and Nicolau
example:

\begin{tabular}{lll}
\begin{lstlisting}[language=Java]
for (i=1; i<=N; i++) {
  A[i]=f1(B[i]);
  B[i]=f2(A[i],D[i-1]);
  C[i]=f3(A[i],D[i-1]);
  D[i]=f4(B[i],C[i]);
}

\end{lstlisting}
& 
~
&
\begin{lstlisting}[language=Java]
for (i=1; i<=N; i++) {
  if (i%N==1) para(A[i]=f1(B[i]),...);
  if (i%1==1) para(B[i]=f2(A[i],D[i-1]),...);
  if (i%1==1) para(C[i]=f3(A[i],D[i-1]),...);
  if (i%1==1) para(D[i]=f4(B[i],C[i]),...);
}	
\end{lstlisting}

\end{tabular}

The above across-loop schedule means that the first instruction is parallelized---all
N instructions A can be computed at once because each depends only on {\sf B[i]}, which
is assigned outside of the loop. In contrast, subsequent instructions are executed 
sequentially. Take B, for example. There is a dependence ``cycle'' from {\sf B[i]} 
to {\sf D[i]} then {\sf B[i+1]} in the subsequent iteration. (The cycle is shown in~\figref{fig:MPCexample}(b).) 
{\sf B[i]}'s cannot be computed in parallel because each one depends on the {\sf B[i]} 
value computed in the previous iteration.

Now consider the above code but slightly modified. {\sf B[i]} and {\sf C[i]} are now 
computed based on the value of {\sf D[i]} from two iterations ago.

\begin{tabular}{lll}
\begin{lstlisting}[language=Java]
for (i=1; i<=N; i++) {
  A[i]=f1(B[i]);
  B[i]=f2(A[i],D[i-2]);
  C[i]=f3(A[i],D[i-2]);
  D[i]=f4(B[i],C[i]);
}

\end{lstlisting}
& 
~
&
\begin{lstlisting}[language=Java]
for (i=1; i<=N; i++) {
  if (i%N==1) para(A[i]=f1(B[i]));
  if (i%2==1) para(B[i]=f2(A[i],D[i-2]));
  if (i%2==1) para(C[i]=f3(A[i],D[i-2]));
  if (i%2==1) para(D[i]=f4(B[i],C[i]));
}	
\end{lstlisting}

\end{tabular}

{\sf B[i]} still depends on values computed in earlier iterations, but one of the values, D[i-2] 
is two iterations earlier. Thus, we can take parallelization steps of 2, e.g., computing 
{\sf B[1]} and {\sf B[2]} in parallel, then {\sf B[3]} and {\sf B[4]} in parallel, etc.
Note however, that we cannot take a parallelization step larger than 2.
The value of B will depend on a value being computed in parallel.

To simplify, we assume that a loop-body schedule is a straight line. \ana{What does this mean?}
\ana{Will we relax?}
The problem therefore becomes: compute the maximal parallelization 
step allowed for each statement. This will account for an MPC schedule with
minimal cost. \ana{Why?}

\subsection{Preliminaries}

\ana{1. MPC-Source}

%\begin{figure*}[tbhp]
\begin{figure}[tbhp]
\begin{tabular}{ll}
\begin{minipage}[b]{4.25cm}

\begin{minted}[fontsize=\footnotesize, linenos, numbersep=5pt, escapeinside=||]{java}
// returns val%mod    
int rem(int val, int mod) {
  int rem = 0;
  for (int j = LEN-1; j |$\ge$| 0; j--) 
  {
    rem = rem << 1;
    // rem[0] = val[j]                                                                                                            
    rem = rem + ((val>>j)&1);   
    if (rem |$\ge$| mod) 
    {
      rem = rem - mod;
    }
  }    
  return rem;
}
\end{minted}
\end{minipage}

&

\begin{minipage}[b]{4.25cm}
\begin{minted}[fontsize=\footnotesize, linenos, numbersep=5pt, escapeinside=||]{java}
// begin inlined rem
int rem0 = 0;
for (int j = LEN-1; j >= 0; j--) 
{
  rem1 = (j==LEN-1) ? rem0 : rem5;
  rem2 = rem1 << 1;
  rem3 = rem2 + (x1>>j)&1; 
  rem4 = rem3 - y1;
  cnd1 = CMP(rem3 >= y1);
  rem5 = MUX(rem3,rem4,cnd1);
}  
// end inline rem        
\end{minted}
\end{minipage}
\\

(a) Source 
&
(c) MPC-source

\end{tabular}
\caption{(a) shows standard (IMP-like) source for an MPC-friendly implementation 
of integer division~\cite{Demmler}, and (b) shows MPC-source. Line 5 {\sf rem1 = (j==LEN-1) ? rem0 : rem5;} 
is what we call a {\em pseudo-$\phi$} node. Note that the {\sf if}-statement on left
becomes straight-line code in MPC. The program \emph{runs both branches}, and 
at the end runs the conditional of the if-statement followed by the multiplexer. 
If {\sf cnd1} is {\sf true}, then {\sf rem} gets assigned the value along the left branch;
otherwise, it gets assigned the value along the right branch.}
    \label{fig:example}\vspace{-2ex}
\end{figure}

\paragraph{MPC-Source.} MPC-Source is an SSA-based intermediate representation
of MPC programs~\cite{ccs paper}. MPC-Source is, essentially, a straight-line sequence
of ``components'', where each ``component'' is either a nested for-loop sequence (called for-loop block), 
or a simple MPC-statement. \figref{fig:example} shows an example. MPC-Source can be constructed
in a standard way from an SSA-translation of the source---each branch of an 
{\sf if}-statement is added to the sequence, the conditional becomes an MPC comparison (CMP), 
and the $\phi$-node becomes an MPC multiplexer (MUX).

\paragraph{Array SSA.} The other essential concept we build upon is Array SSA. Originally proposed

\ana{3. MPC-Source with Array SSA}

\begin{lstlisting}[language=Java,escapeinside=`']
for (i=1; i<=N; i++) {

  Invariant: @A1[j] == j if `$j < i$', @A1[j] == 0 otherwise

  A1 = `$\phi$'(A3,A0); @A1 = max(A3,A0);  
  B1 = `$\phi$'(B3,B0); @B1 = max(B3,B0);
  C1 = `$\phi$'(C3,C0); @C1 = max(C3,C0);
  D1 = `$\phi$'(D3,D0); @D1 = max(D3,D0);

  A2[i] = f1(B1[i]); @A2[i] = i;
  A3 = `$\phi$'(A2,A1); @A3 = max(A2, A1);
  B2[i] = f2(A3[i],D1[i-1]); @B2[i] = i;
  B3 = `$\phi$'(B2,B1); @B3 = max(B2, B1);
  C3[i] = f3(A3[i],D1[i-1]); @C2[i] = i;
  C3 = `$\phi$'(C2,C1); @C3 = max(C2, C1);
  D3[i] = f4(B3[i],C3[i]); @D2[i] = i;
  D3 = `$\phi$'(D2,D1); @D3 = max(D2, D1); 
}
\end{lstlisting}

"Index invariant" makes an assertion about when A[i] is written as a function of i.
\ana{Figure out how to reason about cycles based on Array SSA.}

\subsection{Algorithm}

for a statement x = f(y,z), follow the reverse def-use link.

\section{Future Work}
\label{sec:implementation_and_benchmarks}
%\input{../sections/implementation_and_benchmarks.tex}
%\input{../sections/evaluation.tex}

\section{Conclusions}
\label{sec:conclusion}
%\input{../sections/conclusion.tex}

%\begin{acks}
%    \ishaq{TODO}
%\end{acks}

\printbibliography

% that's all folks
\end{document}

### 2019-11-20 paper.tex
Just checking if this works.
some edits

@@
 We now consider the answer to Problem 2 (in MPC). 
+\ana{Make a note that the problem is easy assuming NO array Writes. We consider array writes.}
 
-\subsection{Problem Overview}
+\subsection{Overview and Problem Statement}
 
-Recall Problem 2, given a loop-body schedule, compute the across-loop schedule
+The problem is, given a loop-body schedule, compute the across-loop schedule
 with minimal cost. The following examples illustrate. The first example is the inner 
@@
 for (i=1; i<=N; i++) {
-  if (i%step5==1) para(t[i],...,t[i+step])
-  if (i%step6==1) ip = ip + t[i];
+  if (i%d(5)==1) para(t[i],...,t[i+d(5)])
+  if (i%d(6)==1) ip = ip + t[i];
 }  
@@
 
-In the above example {\sf step5} is N, meaning that all N multiplications are executed in parallel.
-In contrast, {\sf step6} is 1, meaning that the summation statements are executed sequentially.
+In the above example {\sf d(5)} is N, meaning that all N multiplications are executed in parallel.
+In contrast, {\sf d(6)} is 1, meaning that the summation statements are executed sequentially.
 
@@
 
-The above across-loop schedule means that the first instruction is parallelized---all
-N instructions A can be computed at once because each depends only on {\sf B[i]}, which
-is assigned outside of the loop. In contrast, subsequent instructions are executed 
-sequentially. Take B, for example. There is a dependence ``cycle'' from {\sf B[i]} 
-to {\sf D[i]} then {\sf B[i+1]} in the subsequent iteration. (The cycle is shown in~\figref{fig:MPCexample}(b).) 
-{\sf B[i]}'s cannot be computed in parallel because each one depends on the {\sf B[i]} 
-value computed in the previous iteration.
+The above across-loop schedule entails that the first instruction is parallelized---all
+N instructions A can be computed at once because each depends only on {\sf B[i]}, 
+which at the point {\sf A[i]} is computed, refers to a value assigned outside of the loop. 
+In contrast, subsequent instructions are executed sequentially. Take B, for example. 
+There is a dependence ``cycle'' from {\sf B[i]} to {\sf D[i]} then {\sf B[i+1]} in the subsequent 
+iteration. (The cycle is shown in~\figref{fig:MPCexample}(b).) {\sf B[i]}'s cannot be computed 
+in parallel because each one depends on the {\sf B[i]} value computed in the previous iteration.
 
 Now consider the above code but slightly modified. {\sf B[i]} and {\sf C[i]} are now 
-computed based on the value of {\sf D[i]} from two iterations ago.
+computed based on the value of {\sf D[i-2]} assigned two iterations earlier.
 
@@
 is two iterations earlier. Thus, we can take parallelization steps of 2, e.g., computing 
-{\sf B[1]} and {\sf B[2]} in parallel, then {\sf B[3]} and {\sf B[4]} in parallel, etc.
+{\sf B[1],B[2]} in parallel, then {\sf B[3],B[4]}, etc.
 Note however, that we cannot take a parallelization step larger than 2.
-The value of B will depend on a value being computed in parallel.
+%The value of B will depend on a value being computed in parallel.
 
@@
 \ana{Will we relax?}
-The problem therefore becomes: compute the maximal parallelization 
-step allowed for each statement. This will account for an MPC schedule with
+The problem becomes: for each step, compute the maximal parallelization 
+step allowed by program dependences. This will account for an MPC schedule with 
 minimal cost. \ana{Why?}
@@
 \subsection{Preliminaries}
-
-\ana{1. MPC-Source}
 
@@
 
-(a) Source 
+(a) GCD Source 
 &
@@
 
-\paragraph{MPC-Source.} MPC-Source is an SSA-based intermediate representation
+\subsubsection{MPC-Source}
+
+MPC-Source is an SSA-based intermediate representation
 of MPC programs~\cite{ccs paper}. MPC-Source is, essentially, a straight-line sequence
@@
 
-\paragraph{Array SSA.} The other essential concept we build upon is Array SSA. Originally proposed
+\ana{Add a paragraph on handling of arrays in standard SSA.}
+
+\subsubsection{Array SSA}
+
+The other essential concept we build upon is Array SSA. 
+Originally proposed by Knobe and Sarkar, Array SSA has been used for loop parallelization in HPC.
+We adapt Array SSA for MPC-source and build our optimal loop parallelization algorithm
+based on the adaptation.
 
@@
 
+Array SSA is best illustrated by an example. Consider Aiken and Nicolau's example in Array SSA:
+
+\begin{figure}
 \begin{lstlisting}[language=Java,escapeinside=`']
@@
 \end{lstlisting}
+\caption{Aiken and Nicolau's example using Array SSA.}
+\label{fig:ArraySSAAikenNicolau}
+\end{figure}
+
+There are three kinds of statements added by Array SSA. First, 
+\[
+{\sf A2[i] = f1(B1[i]); @A2[i] = i;}
+\]
+Array SSA statement {\sf @A2[i] = i;} records the iteration, {\sf i} when {\sf A2[i]} is written.
+Note that the index is not necessarily {\sf i}, it can be any $f(i)$, however, we assume
+that the loop always iterates from 1 to $N$ with step 1. Second and third,
+\[ 
+{\sf A3 = }\phi{\sf (A2,A1); @A3 = max(A2, A1);}
+\]
+Statement ${\sf A3 = }\phi{\sf (A2,A1);}$ is a \emph{define} $\phi$ node where arrays 
+{\sf A2} and {\sf A1} are ``merged''---in this case, each value at index other than $i$ 
+is taken from array {\sf A1}, and index $i$ is taken from {\sf A2}. Statement 
+{\sf @A3 = max(A2, A1);} is a timestamp statement; it picks the latest iteration 
+to assign for each index $i$---in this case, when the phi- and max- statements 
+appear immediately under an array write, it assigns {\sf @A3[j] = i} if $j=i$, 
+and {\sf @A3[j] = @A2[j]} otherwise.
+
+\ana{Add merge $\phi$}
+
+\subsubsection{MPC-Source with Array SSA}
+
+The phi-node at if-then-else statements, e.g., ${\sf A3 = \phi(A2,A1), @A3 = max(A2, A1);}$ 
+works for MPC-source without loss of privacy. ${\sf A3 = \phi(A2,A1)}$ is translated as 
+point-wise MUX for each index $i$. The timestamp statement ${\sf @A3 = max(A2, A1);}$ 
+has the same meaning as in standard Array SSA---{\sf @A3[i] = max(@A2[i],@A1[i])}.
+Since MPC-source expands and executes each arm of the if-then-else statement,
+if the then-arm modifies the array at index $i$, but the else arm does not, 
+{\sf @A3[i] = max(@A2[i],@A1[i]) = i}. Similarly, if both arms modify the array at $i$,
+the index again will be $i$. We note that in MPC-source the timestamp arrays {\sf @A} 
+are known \emph{statically}, unlike in standard Array SSA, where they are inherently \emph{dynamic}. 
+\ana{Double check? Make a statement that an adversary monitoring execution does not
+infer anything from SSA phi nodes.}
+
+\subsection{Analysis Algorithm}
+
+\subsubsection{Restrictions on MPC-Source}
+
+To make the problem tractable we adopt certain constraints on the loop body and loop body schedule.
+These constraints hold in classical works on optimal loop parallelization, e.g., Aiken and Nicolau.
+
+Array writes.~\ana{We have to see what the distinction was, regular vs. irregular loops}:
+\begin{enumerate}
+\item For each array read of $A$: $... = ... A[f(i)]$ and corresponding write to $A$: $A[f'(i)] = ...$, 
+we have that $f(i) = f'(i)$ for every $0 \le i < N$. Here $i$ is the induction variable and $N$ is
+the loop bound.
+
+\item Let $A[f_1(i)]...A[f_k(i)]$ be a sequence of writes to $A$ in the loop body sequence 
+\ana{How is the sequence determined?} with $A[f_k(i)]$ being the most recent one. Here
+$i$ denotes the induction variable. We require that for every $0 \le i < N, j,k$, $f_j(i) \neq f_k(i)$, 
+or in other words, those writes access different elements of the array. Furthermore, we require 
+that $f_k(i) \neq f_l(j)$, or in other words, writes across different iterations access different array 
+elements. \ana{Needs rewording. Can be simplified.}
+\end{enumerate}
+
+Constant loop schedule. Let $S$ be a loop schedule, i.e., a directed acyclic graph that defines "executes before"
+scheduling constraints among the statements of the loop body. We require that the 
+parallelization of the loop schedule obeys the constant loop schedule constraint: each instance of $S$
+in the loop linearization obeys the scheduling constraints defined by the DAG of $S$. \ana{Justify why!}
+
+\subsubsection{Computing the Index Invariant}
 
 "Index invariant" makes an assertion about when A[i] is written as a function of i.
-\ana{Figure out how to reason about cycles based on Array SSA.}
+\ana{Can we compute this index easily: A[f(j)] = i, if j>=i, A[f(j)] = 0 otherwise?
+Under what assumptions this is true???}. \ana{We will use the Array SSA stuff to compute index; 
+can get rid of it once we.}
 
-\subsection{Algorithm}
+\subsubsection{Computing $d(s)$}
+We assume a loop body that consists of three-address MPC statements:
+\[ v1 = v2 \oplus v3 \]
+MUX statements
+\[ v1 = MUX(v2, v3, v4) \]
+and
+phi-nodes
+\[ v1 = \phi(v2, v3) \]
+where $v1,v2$, $v3$, and $v4$ are either scalar variables, e.g., {\sf x} or array elements {\sf A[i]}.
+Array-write phi- and max- statements are elided from the representation, as
+their only purpose is to compute the invariant. Control-merge phi- are translated
+into point-wise MUX statements, as is standard in MPC. \ana{Are they translated during analysis?}
+
+\ana{Pseudo phi nodes!}
+
+\begin{figure}
+\begin{lstlisting}[language=Java,escapeinside=`']
+for (i=1; i<=N; i++) {
+
+  // Index Invariant: @A1[j] == j if `$j < i$', @A1[j] == 0 otherwise
+
+  A1 = `$\phi$'(A3,A0); 
+  B1 = `$\phi$'(B3,B0); 
+  C1 = `$\phi$'(C3,C0);
+  D1 = `$\phi$'(D3,D0);
+
+  A2[i] = f1(B1[i]); 
+  B2[i] = f2(A3[i],D1[i-1]); 
+  C3[i] = f3(A3[i],D1[i-1]); 
+  D3[i] = f4(B3[i],C3[i]); 
+}
+\end{lstlisting}
+\caption{Aiken and Nicolau's example using Array SSA.}
+\label{fig:ArraySSAMPCAikenNicolau}
+\end{figure}
+
+%\ana{Figure out how to reason about cycles based on Array SSA.}
+
+Recall that the loop body is a linear sequence of statement. To compute $d(s)$ of statement $s$, 
+we follow def-use chains backwards collecting all reachable pseudo-phi node. If there is a psudo-phi 
+node of a scalar, then $d(s) = 1$; for an array we retrieve the value $A[f(i)]$ from the invariant. \ana{formalize, polish!}
+
+\ana{Define Schedule $S$.}
+
+\ana{Define "constant loop body schedule"}
+
+\ana{Prove main theorem, that this schedule is of minimal cost. I.e., under the restriction of "constant loop body schedule", there
+does not exist a schedule $S'$}
 
 for a statement x = f(y,z), follow the reverse def-use link.
+
+\section{Extensions}
+
+\subsection{Multi Dimensional Arrays}
+
+\subsection{Nested Loops}
 

### 2019-12-27 paper.tex
Changes.

@@
 \newtheorem{remark}{Remark}
+\newtheorem{assumption}{Assumption}
 
@@
 
+\ana{Scope of paper: Should we expand towards 1) nested loops, 2) conversions, 3) implementation (what will we be saying with an implementation?)}
+
 % no \IEEEPARstart
@@
 because B depends on A. The dashed edges denote across-iteration ordering constraints, e.g., B(2), i.e., B in iteration 2
-depends on D(1), i.e., D in iteration 1. (c) shows the optimal loop-body schedule, and (d) shows the
-optimal across-loop schedule. A(1), B(1), etc. denote instruction A in iteration 1, etc.}
+depends on D(1), i.e., D in iteration 1; in the HPC literature these are known as loop-carried dependencies. 
+(c) shows the optimal loop-body schedule, and (d) shows the optimal across-loop schedule. A(1), B(1), etc. denote 
+instruction A in iteration 1, etc.}
 \label{fig:MPCexample}\vspace{-2ex}
@@
 %\input{../sections/optimal_loop_parallelization.tex}
+
+The ``constant rate'' assumption states that a statement $A$ depends on 
+itself at a ``constant rate'', i.e., there cannot be a case where $A(i)$ depends 
+on $A(j), j<i$ and on $A(k), k<i$ s.t. $i \neq j$. \ana{Why the constant rate assumption?}
+
+\begin{assumption} Let $A(j) \dashrightarrow \cdot \longrightarrow A(i)$ be a dependence chain,
+where the dashed arrow denotes the loop carried dependence. For every 
+$A(j) \dashrightarrow \cdot \longrightarrow A(i)$ we have $i = j+m$ for some constant $m$.
+\end{assumption}
 

### 2020-02-16 notes/notes.tex
Added some notes on the algorithm and implementation.

\documentclass{article}

\begin{document}

\section{Examples of SSA with Arrays}

Start with a few examples of "SSA with Arrays". In the beginning we'll start with manually transforming our code into "SSA with Arrays". We will eventually augment with some kind of automatic transformation. By "SSA with Arrays" I mean just the standard treatment of arrays in the classical SSA paper by Cytron et al. not the "Array SSA" form by Knobe and Sarkar that we were looking at some time ago. For example:

\begin{verbatim}
for (int i=1; i<=N; i++) {
  A[i] = f1(B[i]);
  B[i] = f2(A[i], D[i-1]);
  C[i] = f3(A[i], D[i-1]);
  D[i] = f4(B[i], C[i]);
}
\end{verbatim}

becomes

\begin{verbatim}

// initialize A, B, C, and D

for (int i=1; i<=N; i++) {
  // This is the phi node that SSA would insert
  A1 = phi(A, A2); B1 = phi(B, B2); C1 = phi(C, C2); D1 = phi(D, D2);

  // This is a "fake" arraycopy as SSA would do
  // In standard SSA these 2 lines will be expressed by
  // A2 = Update(A1,i,f1(B1[i]));
  // I think the 2 lines are clearer and maybe easier to analyze
  A2 = arraycpy(A1); 
  A2[i] = f1(B1[i]);

  B2 = arraycpy(B1);
  B2[i] = f2(A2[i], D1[i-1]);

  C2 = arraycpy(C1);
  C2[i] = f3(A2[i], D1[i-1]);

  D2 = arraycpy(D1);
  D2[i] = f4(B2[i], C2[i]);
}
\end{verbatim}

Another example, with if-then-else:

\begin{verbatim}
for (int i=1; i<=N; i++) {
  if (C[i]) then
    X[i] = ...
  else 
    X[i] = ...   
}
\end{verbatim}

becomes

\begin{verbatim}
for (int i=1; i<=N; i++) {
  X1 = phi(X,X4);

  if (C[i]) then {
    X2 = arraycpy(X1);
    X2[i] = ...
  }  
  else {
    X3 = arraycpy(X2);
    X3[i] = ...
  }     
  X4 = phi(X2,X3);
}
\end{verbatim}

You can think of other examples too. Look at the ArraySSA paper for more examples.

\section{Implementation}

Start with Soot's Shimple. Ishaq's OPA repo is based on Shimple, so you can start off from Ishaq's code but in the beginning build on Shimple only. 
Shimple is Soot's SSA but according to Ishaq it's buggy, and it does not have support for arrays. So my suggestion is to write the code with those
\texttt{phi} and \texttt{arraycpy} transforms, i.e., in the beginning we'll just manually translate into SSA with arrays. Run on several examples just to see what Shimple does.

\section{Array Def-Use and the Dependence Graph}

Next, implement def-use. Scalar def-use is trivial in SSA, there is probably already API to get the def of a given use. We have to implement array def-use:
\begin{itemize}

\item Let $A_j[f(i)] = ...$ be a definition at some node $d$, and let $ ... = A_k[f'(i)]$ be a use at some node $u$. There is a dependence edge in our graph (i.e., a def-use chain) from $d$ to $u$ iff $A_j[f(i)]$ reaches $A_k[f'(i)]$ and $f(i) = f'(i)$. Here ``reaches'' is in the standard sense of reaching definitions: there is at least one path from $d$ to $u$ without a redefinition of $A_j[f(i)]$. For example, there is a def-use chain from node $A2[i] = f1(B1[i])$ to node $B2[i] = f2(A2[i], D1[i-1])$ in our first example. In most cases the indices $j$ and $k$ will be the same, but that is not necessarily so.

\item Let $A_j[f(i)] = ...$ be a definition at some node $d$, and let $ ... = A_1[f'(i)]$. Fix an index $j$. If there is some $j' < j$ such that $f'(j') = f(j)$ (if there is a $j'$ then it is unique), and $A_j[f(i)]$ is such that it reaches the loop body, then there is a \emph{loop-carried} dependence edge from $d$ to $u$. We'll annotate the loop-carried dependence edge with the difference $d = j - j'$; this indicates that node $u$ depends on the computation at node $d$ that happened $d$ iterations earlier.

In our first example, there is a loop-carried dependence edge from $D2[i] = f4(B2[i], C2[i])$ to $B2[i] = f2(A2[i], D1[i-1])$. Here $f(i) = i-1$ and $f'(i) = i$, which is the most common case. So the $j'$ is $j-1$, and the difference is $j - (j - 1) = 1$.

\end{itemize}

In order for the above to work we will make the following assumptions about array accesses:

\begin{itemize}

\item Let $f(i,x)$ be a "index function", i.e., function we use to access array elements. Here $i$ is the induction variable, and $x$ is a vector of other variables and constants. (Most of the time this is just $i$ but I'd like to make things more general with an eye on extending with nested loops.) The index functions are such that $x$ is a constant throughout the duration of the entire loop. I.e., $f(i,x)$ depends only on $i$.

\item We have ``disjointness of writes''. For each pair of writes $A_j[f(i)] = ... $ and $A_k[f'(i')] = ... $ where $i$ and $i'$ denote values of the induction variable, if $i \neq i'$ then $f(i) \neq f'(i')$. In other words, we cannot have two distinct iterations modifying the same array location.

\end{itemize}

\section{Find SCCs}

The above will construct the graph (nodes are statements in the loop body) and edges are the def-use dependences, including within-iteration dependences and loop-carried dependences. From here we can construct those SCCs and form a DAG of SCCs, where each SCC comes with a delay $d$. We make use of the \emph{constant-rate assumption} and we'll have to take the $d$'s of a bunch of loops to verify that indeed the constant-rate assumption holds. In practice $d$ is either 1 or infinity, i.e., we either have a sequential schedule for the instruction or the instruction can be vectorized.

\end{document}

### 2020-08-26 paper.tex
Some changes.

--- 2019-07-10 paper.tex
@@
 \usepackage{balance}
+\usepackage{mathtools}
 % for creating a balanced last page (usually last page with references)
@@
 \usepackage{booktabs} % better tables
+\usepackage{algorithm}
+\usepackage{algorithmic}
 
@@
 % Do not put math or special symbols in the title.
-\title[Scheduling and Parallelization for MPC]{Scheduling and Parallelization for MPC}
+\title[Scheduling and Amortization for MPC]{Scheduling and Amortization for MPC}
 
@@
 %\renewcommand{\shortauthors}{Trovato and Tobin, et al.}
+
+\renewcommand{\A}{{\sf A}}
+\renewcommand{\B}{{\sf B}}
+\renewcommand{\C}{{\sf C}}
+\renewcommand{\D}{{\sf D}}
 
@@
 
+\ana{Scope of paper: Should we expand towards 1) nested loops, 2) conversions, 3) implementation (what will we be saying with an implementation?)}
+
 % no \IEEEPARstart
 
-\section{Scheduling in MPC}
+\section{Background and Problem Statement}
+\label{sec:background}
+
+\subsection{Scheduling in MPC}
+
+\ana{Here we need more on MPC before jumping to scheduling. MPC-source, basic assumptions about static loop bounds, etc.}
 
@@
 
-\section{Scheduling in HPC}
+\ana{Mention difficulties. We start from IMP-like general-purpose PL syntax with array reads and writes. We translate the program into a dependence 
+graph of MPC instructions. Then we proceed to compute the schedule.}
+
+\subsection{Scheduling in HPC}
+\label{sec:hpc}
 
@@
 The example illustrates that optimal MPC scheduling and optimal HPC scheduling are different problems.
+To summarize, scheduling and amortization in MPC is a different problem due to two key differences: (1) there 
+is unlimited bandwidth while in HPC, we typically have a limited number of processors we can use to 
+schedule instructions in parallel, and (2) we benefit from amortization only if the set of instructions scheduled
+in parallel are the same; in contrast, in HPC any two instructions can be scheduled in parallel.
+
+\ana{Have to make the argument that in MPC all stmts are executed. Thus, the vectorization problem is
+immediately simpler!}
 
@@
 }
+
+\subsection{Precise Problem Statement}
+\ana{TODO}
 
@@
 
-\section{Optimal Loop Parallelization} 
+\section{Across-Loop Parallelization} 
 \label{sec:optimal_loop_parallelization}
@@
 
-\ana{Aiken's and Array SSA to define the notion of optimal loop parallelization for MPC.}
-
 We now consider the answer to Problem 2 (in MPC).
 
-\subsection{Problem Overview}
-
-Recall Problem 2, given a loop-body schedule, compute the across-loop schedule
+\begin{comment}
+\subsection{Overview and Problem Statement}
+
+The problem is, given a loop-body schedule, compute the across-loop schedule
 with minimal cost. The following examples illustrate. The first example is the inner 
@@
 
-\begin{tabular}{lll}
+\begin{center}
 \begin{lstlisting}[language=Java]
@@
 ip = 0;
-for (i=1; i<=N; i++) {
-  t=A[i]*B[i];
-  ip=ip+t;
-}
-
+for (i=0; i<N; i++) {
+  if (i%d(5)==0) para{t[i]=A[i]*B[i],...,t[i+d(5)-1]=A[i+d(5)-1]*B[i+d(5)-1]}
+  if (i%d(6)==0) para{ip = ip + t[i]};
+}  
 \end{lstlisting}
-& 
-~
-&
-\begin{lstlisting}[language=Java]
-A = ...
-B = ...
-ip = 0;
-for (i=1; i<=N; i++) {
-  if (i%step5==1) para(t[i],...,t[i+step])
-  if (i%step6==1) ip = ip + t[i];
-}  
-\end{lstlisting}
-
-\end{tabular}
-
-\ana{More to introduce what the schedule is.}
-
-In the above example {\sf step5} is N, meaning that all N multiplications are executed in parallel.
-In contrast, {\sf step6} is 1, meaning that the summation statements are executed sequentially.
+\end{center}
+
+The above notation is a convenient way to specify a schedule. Each statement $s$ in the loop body sequence 
+is parallelized at a rate $d(s)$ (we assume that the loop body is a straight-line sequence which is the case in MPC). 
+Notation  
+\[{\sf if \; (i\%d(s)\!==\!0) \; para\{s(i),s(i+1),...,s(i+d(s)-1)\}}\]
+means that the first $d(s)$ iterations of statement $s$, denoted by $s(i),...s(i+d(s)-1)$, are executed in a parallel block; 
+then the following $d(s)$ iterations are executed in a parallel block, and so on. The execution of $d(s)$ statements $s$ in 
+a block benefits from amortization of the operation at $s$.
+If {\sf i\%d(s)==0} evaluates to false, then the $i$-th iteration skips $s$, because $s(i)$ has executed earlier in a parallel block.
+
+In the above example {\sf d(5)} is N, meaning that all N multiplications are executed in parallel.
+In contrast, {\sf d(6)} is 1, meaning that the summation statements are executed sequentially.
 
@@
 
-\begin{tabular}{lll}
 \begin{lstlisting}[language=Java]
-for (i=1; i<=N; i++) {
-  A[i]=f1(B[i]);
-  B[i]=f2(A[i],D[i-1]);
-  C[i]=f3(A[i],D[i-1]);
-  D[i]=f4(B[i],C[i]);
-}
-
+for (i=0; i<N; i++) {
+  if (i%N==0) para{A[i]=f1(B[i]),...,A[i+N-1]=f1(B[i+N-1])};
+  if (i%1==0) para{B[i]=f2(A[i],D[i-1])};
+  if (i%1==0) para{C[i]=f3(A[i],D[i-1])};
+  if (i%1==0) para{D[i]=f4(B[i],C[i])};
+}	
 \end{lstlisting}
-& 
-~
-&
-\begin{lstlisting}[language=Java]
-for (i=1; i<=N; i++) {
-  if (i%N==1) para(A[i]=f1(B[i]),...);
-  if (i%1==1) para(B[i]=f2(A[i],D[i-1]),...);
-  if (i%1==1) para(C[i]=f3(A[i],D[i-1]),...);
-  if (i%1==1) para(D[i]=f4(B[i],C[i]),...);
-}	
-\end{lstlisting}
-
-\end{tabular}
-
-The above across-loop schedule means that the first instruction is parallelized---all
-N instructions A can be computed at once because each depends only on {\sf B[i]}, which
-is assigned outside of the loop. In contrast, subsequent instructions are executed 
-sequentially. Take B, for example. There is a dependence ``cycle'' from {\sf B[i]} 
-to {\sf D[i]} then {\sf B[i+1]} in the subsequent iteration. (The cycle is shown in~\figref{fig:MPCexample}(b).) 
-{\sf B[i]}'s cannot be computed in parallel because each one depends on the {\sf B[i]} 
-value computed in the previous iteration.
+
+The above across-loop schedule entails that the first instruction is parallelized---all
+N instructions A can be computed at once because each depends only on {\sf B[i]}, 
+which at the point {\sf A[i]} is computed, refers to a value assigned outside of the loop. 
+In contrast, subsequent instructions are executed sequentially. Take B, for example. 
+There is a dependence ``cycle'' from {\sf B[i]} to {\sf D[i]} then {\sf B[i+1]} in the subsequent 
+iteration. (The cycle is shown in~\figref{fig:MPCexample}(b).) {\sf B[i]}'s cannot be computed 
+in parallel because each one depends on the {\sf B[i]} value computed in the previous iteration.
 
 Now consider the above code but slightly modified. {\sf B[i]} and {\sf C[i]} are now 
-computed based on the value of {\sf D[i]} from two iterations ago.
+computed based on the value of {\sf D[i-2]} assigned two iterations earlier.
 
@@
 \begin{lstlisting}[language=Java]
-for (i=1; i<=N; i++) {
-  if (i%N==1) para(A[i]=f1(B[i]));
-  if (i%2==1) para(B[i]=f2(A[i],D[i-2]));
-  if (i%2==1) para(C[i]=f3(A[i],D[i-2]));
-  if (i%2==1) para(D[i]=f4(B[i],C[i]));
+for (i=0; i<N; i++) {
+  if (i%N==1) para{A[i]=f1(B[i]),...};
+  if (i%2==1) para{B[i]=f2(A[i],D[i-2]),...};
+  if (i%2==1) para{C[i]=f3(A[i],D[i-2]),...};
+  if (i%2==1) para{D[i]=f4(B[i],C[i]),...};
 }	
@@
 
-{\sf B[i]} still depends on values computed in earlier iterations, but one of the values, D[i-2] 
-is two iterations earlier. Thus, we can take parallelization steps of 2, e.g., computing 
-{\sf B[1]} and {\sf B[2]} in parallel, then {\sf B[3]} and {\sf B[4]} in parallel, etc.
-Note however, that we cannot take a parallelization step larger than 2.
-The value of B will depend on a value being computed in parallel.
-
-To simplify, we assume that a loop-body schedule is a straight line. \ana{What does this mean?}
-\ana{Will we relax?}
-The problem therefore becomes: compute the maximal parallelization 
-step allowed for each statement. This will account for an MPC schedule with
-minimal cost. \ana{Why?}
+{\sf B[i]} still depends on a value computed in earlier iterations, i.e., there are loop-carried dependences, 
+but the value, D[i-2] is two iterations earlier. Thus, we can take parallelization blocks of 2, e.g., computing 
+{\sf B[0],B[1]} in a parallel block, then {\sf B[2],B[3]}, etc.
+Note however, that we cannot take a parallelization block larger than 2.
+%The value of B will depend on a value being computed in parallel.
+
+The key problem that we address in this paper is computing the maximal rate $d(s)$ allowed by program dependences 
+for each statement $s$. Program dependences give rise to a partitioning of the sequence of {\sf para} blocks into a DAG of
+strongly connected components. (We elaborate on this and define the partitioning precisely later in the paper.)
+In the above example, there are 2 SCCs---one component consists of just the {\sf para} block that computes 
+{\sf f1(B[i])} and the other component consists of the remaining three {\sf para} blocks. The three para blocks
+are interdependent and hence, they are in the same component.
+
+It is an invariant of our approach that the rates of all blocks in a SCC are the same, $d(s_1) = d(s_2) = d(\mathit{SCC})$
+for each $s_1, s_2 \in \mathit{SCC}$. Unfortunately, as we argued earlier scheduling within a SCC is NP-complete, 
+and also, scheduling of the components in the DAG is NP-complete. Assuming that we have an optimal schedule 
+for each SCC, the sum of the costs of all SCCs is an upper bound on the optimal cost and the sum of the longest chain
+in the DAG is a lower bound on the optimal cost. In the special case when the DAG is a straight line, which is the case 
+is the typical case \ana{how many of the actual benchmarks???} the sum the costs of all SCCs is the \emph{minimal} cost.
+
+\ana{Because the results are somewhat negative, i.e., thing is NP-complete, we'd need some experiments, at least 
+analysis of loop structure in different benchmarks. I am pretty sure the structure is straight-line DAG with very simple SCCs, 
+hence, we do have an optimal schedule anyway.}
 
 \subsection{Preliminaries}
-
-\ana{1. MPC-Source}
 
@@
 
-\paragraph{Array SSA.} The other essential concept we build upon is Array SSA. Originally proposed
-
-\ana{3. MPC-Source with Array SSA}
-
+MPC-Source uses standard SSA~\cite{Cytron:1991TOPLAS} which treats array writes as scalar writes, 
+e.g., {\sf A[i] = f1(B[i])} is a write of a ``scalar'' variable $A$. Each {\sf A[i] = ...} and {\sf A[j] =...} will be treated 
+as separate assignments to variable $A$, thus inducing {\sf A1 = ... } and {\sf A2 = ... } static single assignment
+nodes. The semantics of these array assignments is different than scalar assignments in
+the sense that they are ``vector assignments''; they absorb the write to the corresponding index
+and copy the values at the remaining indices. However, from the point of view of analysis, including
+our analysis for the purposes of scheduling and amortization, the array and scalar assignments are 
+treated the same way.
+
+The assignment to {\sf A[i]} (and to {\sf B[i], C[i]} and {\sf D[i]}) in Aiken and Nicolau's 
+example will turn into {\sf A2 = f1(B1,A1)} ({\sf A1} is included in the operation because the
+elements of all but index $i$ are used in the assignment) and the loop gives rise to several pseudo-$\phi$ nodes:
+
+\begin{lstlisting}[language=Java,escapeinside=`']
+// initialization of A0,B0,C0, and D0
+for (i=1; i<=N; i++) {
+  A1 = `$\phi$'(A0,A2);
+  B1 = `$\phi$'(B0,B2);
+  C1 = `$\phi$'(C0,C2);
+  D1 = `$\phi$'(D0,D2);
+  A2 = f1(B1,A1) // A[i]=f1(B[i]);
+  B2 = f2(A1,D1,B1); // B[i]=f2(A[i],D[i-1]);
+  C2 = f3(A1,D1,C1); // C[i]=f3(A[i],D[i-1]);
+  D2 = f4(B2,C2,D1); // D[i]=f4(B[i],C[i]);
+}	
+\end{lstlisting}
+
+The problem with standard SSA is that it looses fine-grain index information that is necessary for parallelization. 
+In the above example, the analysis determines that assignment A2 depends on the previous iteration,
+which entails parallelization rate of 1.
+
+\subsubsection{Array SSA}
+
+Therefore, we make use of Array SSA. 
+Originally proposed by Knobe and Sarkar, Array SSA has been used for loop parallelization in HPC.
+We adapt Array SSA for MPC-source and build our optimal loop parallelization algorithm
+based on this adaptation.
+
+Array SSA introduces a parallel array structure, denoted $@A$, associated to each array $A$. $@A$ is a ``timestamp'' 
+that tracks the specific loop iteration when each $A[i]$ was last modified. We can use this information to compute a more
+precise parallelization rate. 
+Consider Aiken and Nicolau's example, now using Array SSA:
+
+\begin{figure}
 \begin{lstlisting}[language=Java,escapeinside=`']
@@
 \end{lstlisting}
-
-"Index invariant" makes an assertion about when A[i] is written as a function of i.
-\ana{Figure out how to reason about cycles based on Array SSA.}
-
-\subsection{Algorithm}
+\caption{Aiken and Nicolau's example using Array SSA.}
+\label{fig:ArraySSAAikenNicolau}
+\end{figure}
+
+There are three kinds of statements added by Array SSA. First, 
+\[
+{\sf A2[i] = f1(B1[i]); @A2[i] = i;}
+\]
+Array SSA statement {\sf @A2[i] = i;} records the iteration, $i$ when {\sf A2[i]} is written.
+The index is not always $i$, it can be any $f(i)$, function of $i$, however, we assume
+that the loop always iterates from 1 to $N$ with step 1. Second and third, there are
+the \emph{define} $\phi$ statement  and corresponding timestamp recording statement:
+\[ 
+{\sf A3 = }\phi{\sf (A2,A1); @A3 = max(A2, A1);}
+\]
+Statement ${\sf A3 = }\phi{\sf (A2,A1);}$ is a \emph{define} $\phi$ node that merges arrays 
+{\sf A2} and {\sf A1}---the semantics of the define $\phi$ is that the value at the index that is
+modified is taken from ${\sf A2}$ and the values at indeces other than $i$ 
+are taken from array {\sf A1}. Statement 
+{\sf @A3 = max(A2, A1);} is a timestamp recording statement; it picks the latest iteration 
+to assign for each index $i$---in this case, when the phi- and max- statements 
+appear immediately under an array write, it assigns {\sf @A3[j] = i} if $j=i$, 
+and {\sf @A3[j] = @A2[j]} otherwise.
+
+Array SSA allows for the standard \emph{merge} $\phi$ statement at if-then-else statements. 
+We may have 
+\[ 
+{\sf A3 = }\phi{\sf (A2,A1); @A3 = max(A2, A1);}
+\]
+in which case ${\sf A2}$ is the array long the true branch and ${\sf A1}$ is the array
+along the false branch. 
+%\ana{Add merge $\phi$}
+
+\subsubsection{MPC-Source with Array SSA}
+
+Merge phi nodes  e.g., ${\sf A3 = \phi(A2,A1), @A3 = max(A2, A1);}$ 
+can be added to MPC-source without privacy leaks. ${\sf A3 = \phi(A2,A1)}$ is naturally translated as 
+point-wise MUX for each index $i$. The timestamp statement ${\sf @A3 = max(A2, A1);}$ 
+has the same meaning as in standard Array SSA---{\sf @A3[i] = max(@A2[i],@A1[i])}.
+MPC-source expands and executes each arm of the if-then-else statement.
+If only one arm modified the array at index $i$ {\sf @A3[i] = max(@A2[i],@A1[i]) = i}. 
+Similarly, if both arms modified the array at $i$, the timestamp index again will be again $i$. 
+%We note that in MPC-source the timestamp arrays {\sf @A} are known \emph{statically}, 
+%unlike in standard Array SSA, where they are inherently \emph{dynamic}. 
+%\ana{Double check? Make a statement that an adversary monitoring execution does not
+%infer anything from SSA phi nodes.}
+
+\subsection{Analysis Algorithm}
+
+\subsubsection{Assumptions on MPC-Source}
+
+To make the problem tractable we state several assumption about the loop body.
+These assumptions are natural and they have been adopted in classical works as 
+well~\cite{Aiken:PLDI1988}. 
+These assumptions are natural and hold for all loop body sequences we have examined,
+both in MPC and HPC. This is because loops that violate these assumptions would be
+very difficult to understand and maintain. Moreover, if loops use arrays in readonly way, 
+which is the case in the vast majority of cases particularly in MPC, the assumptions automatically 
+hold true.
+~\ana{This statement is true for all loops I have seen but we do need 
+a more systematic way of studying loops and reporting on structure.}
+
+%Array writes.~\ana{We have to see what the distinction was, regular vs. irregular loops}:
+\begin{enumerate}
+\item For each array read of $A$: $... = g(A[f(i)])$ and corresponding write to $A$: $A[f'(i)] = ...$
+we require that $f(i) = f'(i)$ for every $0 \le i < N$. Here $i$ is the induction variable and $N$ is
+the loop bound.
+
+\item Let $A[f_1(i)]...A[f_k(i)]$ be a sequence of writes to $A$ in the loop body sequence. 
+%\ana{How is the sequence determined?} with $A[f_k(i)]$ being the most recent one. 
+As usual, here $i$ denotes the induction variable. We require that for every $0 \le i < N, k,l$, $f_k(i) \neq f_l(i)$, 
+or in other words, those writes access different elements of the array. Furthermore, we require 
+that $f_k(i) \neq f_l(j)$, or in other words, writes across different iterations access different array 
+elements. %\ana{Needs rewording. Can be simplified.}
+
+\item Our last assumption is the ``constant rate'' assumption. It states that a statement 
+$s$ in the loop body sequence depends on itself at a ``constant rate'', i.e., there cannot be a 
+case where $s(i)$ depends on $s(j), j<i$ and on $s(k), k<i$ s.t. $i \neq j$. (Here $s(i)$
+denotes the execution of $s$ in iteration $i$.)
+More formally, let $s(j) \dashrightarrow \cdot \longrightarrow s(i)$ be a dependence chain,
+where the dashed arrow denotes the loop carried dependence. For every 
+$s(j) \dashrightarrow \cdot \longrightarrow s(i)$ we have $i = j+m$ for some constant $m$.
+
+\end{enumerate}
+
+%Constant loop schedule. Let $S$ be a loop schedule, i.e., a directed acyclic graph that defines "executes before"
+%scheduling constraints among the statements of the loop body. We require that the 
+%parallelization of the loop schedule obeys the constant loop schedule constraint: each instance of $S$
+%in the loop linearization obeys the scheduling constraints defined by the DAG of $S$. \ana{Justify why!}
+
+\subsubsection{Index Invariant}
+
+The \emph{index invariant} makes an assertion about when {\sf A[i]} is written as a function of $i$.
+It is essential in computing the rate of parallelization $d(s)$. \figref{fig:ArraySSAMPCAikenNicolau}
+shows the index invariant on ${\sf @A1}$ (all other arrays have the exact same index invariant). 
+The index invariant can be specified by the programmer, as it is a form of a loop invariant or it can 
+be inferred. \ana{Can we infer the invariant???}
+
+%\ana{Can we compute this index easily: A[f(j)] = i, if j>=i, A[f(j)] = 0 otherwise?
+%Under what assumptions this is true???}. \ana{We will use the Array SSA stuff to compute index; 
+%can get rid of it once we.}
+
+\subsubsection{Computing $d(s)$}
+
+We assume a loop body that consists of three-address MPC statements:
+\[ v1 = v2 \oplus v3 \]
+MUX statements:
+\[ v1 = MUX(v2, v3, v4) \]
+and $\phi$-nodes:
+\[ v1 = \phi(v2, v3) \]
+where $v1,v2$, $v3$, and $v4$ are either scalar variables, e.g., {\sf x} or array elements {\sf A[f(i)]}.
+Define phi- and max- statements are elided from the representation, as
+their only purpose is to compute the invariant. Merge phi-statements are translated
+into point-wise MUX statements, as is standard in MPC.
+
+\ana{Pseudo phi nodes!}
+
+\begin{figure}
+\begin{lstlisting}[language=Java,escapeinside=`']
+for (i=1; i<=N; i++) {
+
+  // Index Invariant: @A1[j] == j if `$j < i$', @A1[j] == 0 otherwise
+
+  A1 = `$\phi$'(A3,A0); 
+  B1 = `$\phi$'(B3,B0); 
+  C1 = `$\phi$'(C3,C0);
+  D1 = `$\phi$'(D3,D0);
+
+  A2[i] = f1(B1[i]); 
+  B2[i] = f2(A3[i],D1[i-1]); 
+  C3[i] = f3(A3[i],D1[i-1]); 
+  D3[i] = f4(B3[i],C3[i]); 
+}
+\end{lstlisting}
+\caption{Aiken and Nicolau's example using Array SSA.}
+\label{fig:ArraySSAMPCAikenNicolau}
+\end{figure}
+
+%\ana{Figure out how to reason about cycles based on Array SSA.}
+
+Recall that the loop body is a linear sequence of statements. To compute $d(s)$ of statement $s$, 
+we follow def-use chains backwards collecting all reachable pseudo-phi node. If there is a psudo-phi 
+node of a scalar, then $d(s) = 1$; for an array we retrieve the value $A[f(i)]$ from the invariant. \ana{formalize, polish!}
+
+\ana{Define Schedule $S$.}
+
+\ana{Add Algorithm}
+
+\ana{Prove main theorem, that this schedule is of minimal cost. I.e., under the restriction of "constant loop body schedule", there
+does not exist a schedule $S'$}
 
 for a statement x = f(y,z), follow the reverse def-use link.
+\end{comment}
+
+\subsection{High-level CFG to MPC $G$}
+\label{CFG}
+
+\subsubsection{From $\phi$-nodes to MUX-nodes}
+
+\ana{Difficulties: Translate to MUX nodes, adapt classical SSA algorithm, arrays, reads and writes!}
+
+\begin{figure}
+\small
+\[
+\begin{array}{l@{~~~~~}l}
+  \begin{array}{l@{~}l@{~~~}l}
+  s & ::= s ; s \\ 
+  & \mid \x = \y \; \mid \x=\y \; \op \; \z \mid {\sf A[f(i)] = x} \; \mid {\sf x = A[f(i)]} & \mathit{assignment} \\
+%  & \mid \x = \phi(\y,\z) \\
+  & \mid {\sf for} \; (i = 0; i \le \mathit{Const}; \mathit{i\!+\!+}) \; \{ \; s \; \} \\
+  & \mid {\sf if} \; (\x \; \cop \; \y) \; \{ \; s \; \} \; {\sf else} \; \{ \; s \; \} \\
+ % \bool &::= \x \; \cop \; \y & \mathit{comparison} \\
+  \op &::= + \mid - \mid * \mid / & \mathit{arithmetic} \; \mathit{operator} \\
+  \cop &::=  \; == \; \mid \; != \; \mid \; < \; \mid \; \le & \mathit{comparison} \; \mathit{operator}
+
+  \end{array}
+\end{array}
+\]
+%\vspace{-0.2in}
+\caption{IMP syntax as defined in \cite{Ishaq2019}: $s$ is a sequence of statements. 
+\code{x}, \code{y}, \code{z}, and $i$, are variables.
+} 
+\label{fig:syntax}
+\end{figure}
+
+We consider a standard IMP-like imperative language. The syntax is given in~\figref{fig:syntax}. 
+Without loss of generality, we assume that the syntax gives rise to a CFG and we can apply the 
+Cytron et al.'s Algorithm for the efficient computation of SSA over this CFG~\cite{Cytron:1998}.
+
+In particular, we are interested in a sequence of statements that form a loop body, i.e., while 
+the outermost statement is a {\sf for} the remaining statements are assignments and {\sf if}s. 
+There are two principal hurdles. One is to translate the CFG into SSA form, where $\phi$ nodes 
+are turned into MUX nodes. One cannot simply reuse the classical algorithm by Cytron et. al. 
+because $V_3 = \phi(V_1,V_2)$ nodes do not retain information about the predicate that triggers 
+$V_1$ vs. $V_2$; all $\phi$ nodes represent is conditional control flow, i.e., that $V_3$ is either 
+$V_1$ or $V_2$. Another issue is that SSA $\phi$ nodes can combine multiple conditionals. 
+To illustrate, consider the code
+
+\begin{minted}[fontsize=\footnotesize, linenos, numbersep=5pt, escapeinside=||]{java}
+z = 0;
+...
+if (x > 0) {
+  if (y > 0) {
+    z = 1;
+  }  else {
+    z = -1;
+  }
+}
+print z
+\end{minted}
+Line 10 is the dominance frontier of the assignments to \z, and Cytron et. al.'s algorithm
+creates a $\phi$ node at 10: $z_3 = \phi(z_0, z_1, z_2)$. There is no information retained
+that $z_0$ is triggered when $x \le 0$, $z_1$ is triggered when $x > 0 \; \wedge \; y > 0$, 
+and $z_2$ is triggered when $x > 0 \; \wedge \; y \le 0$. The Multiplexer at the join point
+needs this information however in order to assign the correct value to $z_3$ at runtime.
+
+We propose a simple extension to the classical algorithm. Recall from Cytron et. al.'s 
+that efficient static single assignment computation has two phases, phase one is the 
+placement of phi-nodes at dominance frontiers of assignments, and phase two is the 
+renaming of variables $V$ into $V_1$, $V_2$, etc.
+
+\ana{Here we need an explanation for lines 23-45. Turning n-ary phi-nodes into binary MUX nodes, 
+just the same classical algorithm with minor extension.}
+
+\ana{Example of CFGs for above example.}
+
+Once we have the graph in SSA form with \ana{MORE}
+
+\begin{algorithm}
+\caption{Determine Graph edges}
+\begin{algorithmic}[1]
+%\REQUIRE $c \neq null$
+%\STATE $L \leftarrow get\_nodes(c)$
+%\FORALL{$n$ in $L$}
+%\STATE $...$
+%\ENDFOR
+\FORALL {variables $V$}
+\STATE $C(V) \leftarrow 0$
+\STATE $S(V) \leftarrow \mathit{EmptyStack}$
+\STATE $P \leftarrow \mathit{EmptyStack}$
+\ENDFOR
+\STATE call SEARCH($\mathit{Entry}$) \\
+\STATE ... \\
+SEARCH(X) : \\
+\FORALL {statements $A$ in $X$} 
+\IF {$A$ is a $\phi$-assignment}
+   \STATE {remove $A$ and continue}
+\ENDIF
+\FORALL {variables $V$ in $\mathit{RHS}(A)$}
+   \STATE {replace $V$ with $V_i$ where $i = \mathit{Top}(S(V))$}
+\ENDFOR
+
+\FORALL {variable $V$ in $\mathit{LHS}(A)$}
+   \STATE { $i \gets C(V)$} 
+   \STATE {replace $V$ with $V_i$ in $\mathit{LHS}(A)$}
+   \STATE {push $i$ onto $S(V)$}
+   \STATE {$C(V) \gets i+1$}
+\ENDFOR
+\ENDFOR
+
+\FORALL {$Y \in Succ(X)$}
+\IF {$X$ is a conditional node}
+   \STATE {push $\mathit{Cond}, \mathit{Branch}$ onto $P$ // $\mathit{Branch}$ is $T$ if $X \rightarrow Y$ is the True branch, $F$ otherwise}     
+\ENDIF   
+     \STATE{// $Y$ is not done if $Y$ has an incomplete MUX predecessor or an original CFG predecessor \ana{Better?}}
+      \IF {$Y$ has $\phi$-assignments and $Y$ is not done} 
+          \STATE {$\mathit{Cond}, \mathit{Branch} \gets$ pop $P$}
+          \FORALL {$\phi$-assignments in $Y$}
+               \IF {$Z\!: V = \mathit{MUX}(Cond, ..., ...) \in \mathit{Pred}(Y)$}
+                  %\STATE {replace $V$ in $RHS(Z)$ with $V_i$, where $i = Top(S(V)$}
+                  \STATE {remove edge $X\rightarrow Y$, add edge $X\rightarrow Z$ to $\mathit{CFG}$}
+                  \STATE {call SEARCH($Z$) // updates $V$'s in LHS and RHS} 
+               \ELSE
+                   \STATE {$i = Top(S(V))$}
+                   \IF {$\mathit{Branch} == T$}
+                      \STATE {create new node $Z\!: V = \mathit{MUX}(\mathit{Cond}, V_i, V)$}
+                   \ELSE 
+                      \STATE {create new node $Z\!: V = \mathit{MUX}(\mathit{Cond}, V, V_i)$}
+                   \ENDIF   
+                   \STATE {remove edge $X\rightarrow Y$, add edge $X\rightarrow Z$ to $\mathit{CFG}$} 
+               \ENDIF
+          \ENDFOR 
+      \ENDIF
+\ENDFOR
+\FORALL {$Y \in \mathit{Children}(X)$}
+   \STATE {call SEARCH($Y$)}
+\ENDFOR
+\FORALL {assignments $A$ in $X$}
+   \FORALL {variables $V$ in oldLHS($A$)}
+       \STATE {pop S(V)}
+   \ENDFOR
+\ENDFOR
+\end{algorithmic}
+\label{alg:algorithm1}
+\end{algorithm}
+
+\subsubsection{Handling of Arrays}
+\label{???}
+
+So far we focused on the classical SSA analysis and the extension to create MUX nodes. 
+Classical SSA treats arrays as scalars, i.e., an array write, e.g., {\sf A[i] = x+y} is considered
+a modification of {\sf A} and warrants its own indexed variable ${\sf A}_k$. A later use of {\sf A}, 
+even one such as {\sf x = A[i+1]} where there is no def-use link from {\sf A[i] = x+y}, will be 
+considered a use of ${\sf A}_k$. This hinders parallelization. To illustrate, consider our running
+example in~\figref{fig:MPCexample}:
+\begin{minted}[fontsize=\footnotesize, numbersep=5pt, escapeinside=||]{java}
+for (...) {
+  A: A[i] = f_1(B[i]);
+  B: B[i] = f_2(A[i]);
+  ...
+}
+\end{minted}
+Since {\sf A} and {\sf B} are treated as standard scalars, the dependence analysis will record
+a dependence from {\sf B} to {\sf A} via the back edge. As a result, statement {\sf A} wouldn't 
+be vectorized because each (i+1)-th iteration depends on {\sf B} in the i-th iteration,
+which depends on {\sf A} in the i-th iteration.
+
+To mitigate against this, we propose a technique, similar to Array-SSA though much simpler and 
+less general, which breaks certain dependence edges and allows for better vectorization. 
+\ana{Have to make a claim how 
+often this technique resolves dependences perfectly, and how often it defaults to standard SSA.}
+
+\ana{TODO: Summary of the technique}
+
+\paragraph {Equivalence and Non-equivalence} Let $f(i)$ and $f'(i)$ be two index functions on induction variables $i$, 
+e.g., we have array accesses ${\A}[f(i)]$ and ${\A}[f'(i)]$. We say that $f(i) \cong f'(i)$ iff for every $1 \le i \le N$,
+$f(i) = f'(i)$. We say that $f(i) \ncong f'(i)$ iff for every $1 \le i \le N$, $f(i) \neq f'(i)$.
+
+\paragraph{Array Well-formedness Conditions} Let ${\A}$ be an array mentioned in the loop body. 
+We will treat array accesses ${\A}[f(i)]$ as distinct scalar variables if we can show that the loop body is
+well-formed with respect to \A. Otherwise, we treat array {\A} as a single variable, thus defaulting to the 
+case of standard SSA. The well-formedness conditions are as follows: \ana{This needs some work!}
+
+\begin{enumerate}
+\item For each pair of array writes to \A, $n_1\!: A[f(i)] = ... $ and $n_2\!: A[f'(i)] = ... $, we have either $f(i) \cong f'(i)$ or 
+$f(i) \ncong f'(i)$.
+
+\item For each pair $n_1\!: A[f(i)] = ... $ and $n_2\!: ... = A[f'(i)]$ such that $n_1$ reaches $n_2$ on a forward path, 
+we have either $f(i) \cong f'(i)$ or $f(i) \ncong f'(i)$.
+
+%\item For each pair $n_1\!: A[f(i)] = ... $ and $n_2\!: A[f'(i)] = ... $ either $f(i-d) \cong f'(i)$ for some unique $d > 0$, 
+%or $f(i-d) \ncong f'(i)$ for every $d > 1$.
+
+\item For each pair $n_1\!: A[f(i)] = ... $ and $n_2\!: ... = A[f'(i)]$ %such that $n_1$ reaches the loop exit (we assume that there is
+%a unique loop exit node) on a forward path, and there is  $n_2$ on a forward path, 
+we have either $f(i-d) \cong f'(i)$ or $f(i-d) \ncong f'(i)$. \ana{FIX!}
+
+\end{enumerate}
+
+To see the intuition behind these rules, consider rule (1). 
+In other words, the two array writes are either equivalent, i.e., at each iteration they modify the same
+location, or they are non-equivalent, i.e., at each iteration they modify distinct locations. Note that the two array writes
+``clash'' in one way or another. If $n_1$ reaches $n_2$ (or vice versa, $n_2$ reaches $n_1$), then if $f(i)$ and $f'(i)$ 
+are neither equivalent nor non-equivalent, there would be a kill in some iterations and not in other; in contrast, if
+they are equivalent, there is always a kill, and if they are non-equivalent, there is never a kill. If neither $n_1$ reaches
+$n_2$ nor the other way around, then the two array writes will ''clash'' at a join node. If $f(i)$ and $f'(i)$ are neither 
+equivalent nor non-equivalent, then there would be one kind of MUX for certain iterations and a different kind of MUX 
+for the rest; if $f(i)$ is equivalent to $f'(i)$, then the MUX will be $... = \mathit{MUX}(\mathit{cond}, {\A}[f(i)], {\A[f'(i)]})$, and 
+$f(i)$ and $f'(i)$ are non-equivalent then there will be two MUX nodes: $... = \mathit{MUX}(\mathit{cond}, {\A}[f(i)], {\A'[f(i)]})$
+and $... = \mathit{MUX}(..., {\A}[f'(i)], {\A'[f'(i)]})$. (Note that after SSA, the A's will be properly indexed.)
+\ana{I need explanations of (2) and (3).}
+
+Consider our running example in~\figref{fig:MPCexample}(a). There is a single write to each array in the loop body, 
+so there are no write-write pairs and no tests for (1). There are the following write-read pairs: 2 and 3 (access of \A), 
+2 and 4 (access of \A), 3 and 5 (access of \B), 4 and 5 (access of \C), 3 and 2 (access of \B), 5 and 3 (access of \D) and 5 and 4 (access of \D).
+We apply the forward test, (2), only on the first three pairs, as for the remaining ones, there is no forward path from the write to the read. 
+Test (2) clearly passes because $i$ is obviously equivalent to $i$. For the remaining pairs, we only apply the backward test, test (3). 
+Pair 3 and 2 passes as well because $i-1$ and $i$ are nonequivalent for every $d > 0$. 5 and 3, and analogously 5 and 4, pass test (3). 
+In these two cases, the unique $d = 1$ makes $i - d$ and $i-1$ equivalent.
+
+\ana{Where do we put this? Theorem: simple rule is not restrictive, and Theorem 2: program equivalent to transformed one where 
+each A[f(i)] is replaced with scalar.}
+
+\subsubsection{Constructing G}
+
+Construction proceeds as follows:
+
+\begin{enumerate}
+
+\item We consider the forward flow graph, and group array accesses $\A[f(i)]$ into equivalence classes. If $f(i) \cong f'(i)$
+then accesses $\A[f(i)]$ and $\A[f'(i)]$ are in the same equivalence class, i.e., they are equivalent accesses. We write $\A[f(i)]_k$ 
+(just as $V_k$) to denote the SSA-indexed versions of the equivalent accesses. (Although we give a general treatment, the reader 
+may just think of the equivalence accesses as just one $\A[f(i)]$, e.g., $\A[i]$).
+
+\item We run \ref{alg:algorithm1} to compute SSA of the loop body with MUX nodes. Equivalent accesses get their own indexed versions.
+
+\item Constructing the ``forward'' portion of the graph: add a forward edge $n_1 \rightarrow n_2$ if for some $V_k$, 
+$V_k \in LHS(n_1)$ and $V_k \in RHS(n_2)$.
+
+\item For each scalar $V$, let $n_1$ be the node such that $V_\mathit{max} \in LHS(n_1)$, 
+i.e., this is the node that defines the $V$ exposed at the exit of the loop. Add an edge $n_1 \stackrel{-1}{\dasharrow} n_2$ 
+for each $n_2$ such that $V_0 \in RHS(n_2)$. In other words, $n_2$ has an upward exposed use of $V$, i.e., one that comes
+from. For each array write equivalence class, let $n_1$ be the node such that $A[f(i)]_\mathit{max} \in LHS(n_1)$. Add an edge 
+$n_1 \stackrel{-d}{\dasharrow} n_2$ for each $n_2$ such that $A[f'(i)]_0 \in RHS(n_2)$ and $f(i-d) \cong f'(i)$.
+
+\end{enumerate}
+
+As an example of these, return to~\figref{fig:MPCexample}(a). Phase (1) groups the array accesses into equivalence classes:
+$\A[i], \B[i], \C[i], \D[i-1], \D[i]$. Treating the above array accesses as scalars, phase (2) transforms the code into the following:
+
+\begin{minted}[fontsize=\footnotesize, linenos, numbersep=5pt, escapeinside=||]{java}
+for (int i=1; i<=N; i++) {
+  A[i]|$_1$| = f|$_1$|(B[i]|$_0$|);
+  B[i]|$_1$| = f|$_2$|(A[i]|$_1$|,D[i-1]|$_0$|);
+  C[i]|$_1$| = f|$_3$|(A[i]|$_1$|,D[i-1]|$_0$|);
+  D[i]|$_1$| = f|$_4$|(B[i]|$_1$|,C[i]|$_1$|);
+}
+\end{minted}
+
+There are no MUX-nodes due as there are no conditionals in the loop body. Phase (3) adds the forward edges:
+$2 \rightarrow 3$, $2 \rightarrow 3$, $3 \rightarrow 5$, and $4 \rightarrow 5$. Phase (4) adds the backward edges:
+$5 \stackrel{-1}{\dasharrow} 3$ and $5 \stackrel{-1}{\dasharrow} 4$. The graph is shown in~\figref{fig:MPCexample}(b).
+
+\ana{Another example, with conditional?}
+
+\subsection{SCCs and Amortized Schedule}
+\label{sec:scc}
+
+The next phase is the actual scheduling algorithm. Recall that backward edges had
+the negative weight $-1$ in all our examples, however, the weight on backward edges can be any value from $-1$ 
+to $-N+1$. Forward edges have weight 0.
+
+Consider a cycle $c \in G$. $d(c)$ is the absolute value of the sum of the weights of the edges in $c$. Consider a 
+node $n \in G$. We say $n$ has \emph{constant rate} if all cycles $c$ through $n$ have the same $d(c)$. We write 
+$d(n) = d(c)$. The $d(c)$ means that the computation at node $n$ at iteration $i$ depends on the computation 
+at the same node at iteration $i-d(c)$. If $n$ is constant-rate with value $d(n)$, then for every $i$, $n(i)$
+depends on the value computed $d(n)$ iterations ago; but if the $d(c)$'s are not the same, then $n(i)$ depends on 
+values computed at $n$ at different earlier iterations. We extend the definition of constant-rate to $G$. We say that 
+$G$ is constant-rate if all $n \in G$ are constant-rate. If $n$ is a node, such that there are no cycles through $n$ in $G$, 
+then $d(n) = N$, or in other words, $n$ can be vectorized. If a node $n$ is not constant rate, we assign a default $d(n) = 1$ 
+for the purpose of amortization. For the rest of the text, we are interested in constant-rate $G$'s and consider optimality 
+results on constant-rate $G$'s. In practice, nodes have $d(n)$ values of either $N$ or $1$ even though other values of 
+$d(n)$ are possible in theory. \ana{HERE HAVE TO GIVE empirical evidence, how good/bad it is.}
+
+%We will consider Strongly Connected Components (SCCs) in $G$. In the running example in~\figref{fig:MPCexample}(b), 
+%we have two SCCs, $SCC_1$ consists of node 2, and $SCC_2$ consists of nodes $3,4,5$. 
+Without loss of generality, we may assume that each cycle has the form $n_1 \rightarrow n_2 \rightarrow ... \rightarrow n_k \stackrel{-d}{\dasharrow} n_1$, 
+or in other words, there is a single backward edge in the entire cycle. Even though cycles with $n \ge 2$ backward 
+edges are possible, such cycles can be transformed into $n$ disjoint cycles of the above form. The $-d$ on the
+single backward edge in each of the $n$ cycles, is the sum of the $-d$'s of all backward edges in the original cycle. 
+\ana{We probably need more here.}
+
+We now compute the amortized schedule $\mathcal{A}$.
+
+\begin{enumerate}
+
+\item We begin by decomposing $G$ into Strongly Connected Components (SCCs). It is a theorem, that if $G$ is constant-rate, 
+then for each $\mathit{SCC}$, and each pair of nodes $n_1, n_2 \in \mathit{SCC}, d(n_1) = d(n_2)$. We write $d(\mathit{SCC})$.
+For the running example in~\figref{fig:MPCexample}(b), we have two SCCs, $SCC_1$ consists of node 2, and $SCC_2$ 
+consists of nodes $3,4,5$. $d(\mathit{SCC}_1) = N$ and $d(\mathit{SCC}_2) = 1$.
+
+\item For each SCC, assume that there is a schedule $S: P_1 \rightarrow P_2 ... \rightarrow P_k$ given 
+as input. $S$ takes into account only the forward edges in $G$: if $n_1, n_2 \in SCC$ and $n_1 \rightarrow n_2 \in G$, 
+then $S(n_1)$ precedes $S(n_2)$ in the linear schedule. Note that computing an optimal schedule is NP-complete, 
+as we showed earlier, however, given a small SCC, one can brute force an optimal $S$. For the running example in 
+in~\figref{fig:MPCexample}(b), assuming that nodes 3 and 4 perform the same operation, we work with schedule 
+$S: \{3,4\} \rightarrow \{5\}$
+
+\item Given $S$ for a given SCC, we then compute the amortized schedule $\mathcal{A}_{\mathit{SCC}}$; 
+this is the schedule across all iterations of the loop. For each $P_i$ in $S$, there are $\lceil{\frac{N}{d(SCC)}}\rceil$
+nodes in $\mathcal{A}_{\mathit{SCC}}$. $P_i(1),P_i(2),...P_i(d(SCC))$ go into node $P^1_i$, $P_i(d(SCC)+1),...P_i(2\cdot d(SCC))$ 
+go into node $P^2_i$, etc. The parenthesized number denotes the iteration, e.g., $P_i(1)$ denotes $P_i$ at the first iteration 
+of the loop. There are dependence edges as follows:
+\begin{enumerate}
+\item $P^j_i \rightarrow P^j_{i'}$ if there is an edge $P_i \rightarrow P_{i'}$ in $S$ (intra-loop edges).
+\item $P^{\lceil{\frac{N}{d(SCC)}}\rceil}_i \rightarrow P^1_{i+1}$ for each $1 \le i < \lceil{\frac{N}{d(SCC)}}\rceil$ (across-loop edges).
+\end{enumerate}
+In the running example, $SCC_1$'s schedule is just the single vectorized node that executes all A's at once.
+$SCC_2$'s schedule repeats $S: \{3,4\} \rightarrow \{5\}$ $N$ times: we have $\{3,4\}(1) \rightarrow \{5\}(1) \rightarrow \{3,4\}(2) \rightarrow \{5\}(2) ... \{3,4\}(N) \rightarrow \{5\}(N)$, 
+as each iteration depends on the previous-iteration value computed at the same node.
+This is illustrated graphically in~\figref{fig:MPCexample}(d).
+
+\item User gives an SCC schedule $SCC_1 \rightarrow SCC_2 \rightarrow ... \rightarrow SCC_m$. Again, an optimal schedule 
+is NP-complete. The final schedule is $\mathcal{A} = \mathcal{A}_{\mathit{SCC}_1} \rightarrow \mathcal{A}_{\mathit{SCC}_2} ... \rightarrow \mathcal{A}_{\mathit{SCC}_m}$.
+In our running example, there is only one possible ordering of the two components, and the final schedule is as depicted in~\figref{fig:MPCexample}(d).
+%dependence edges $P(n)_1 \dasharrow P(n)_2 ... \dasharrow P(n)_{\lceil{\frac{N}{d(n)}}\rceil}$, reflecting the across-loop dependence 
+%edges we saw earlier. The rest of the edges for every edge $n_1 \rightarrow n_2 \in G$ and every $1 \le i \le N$, add an edge 
+%$G'(n_1(i)) \rightarrow G'(n_2(i))$ to $G'$ where $G'(n(i))$ returns the $P$ node in $G'$ that contains $n(i)$.
+
+\end{enumerate}
+
+%We assign a parallelization factor of $N$ to $n$, meaning that 
+%$n$ can be fully vectorized. A more interesting case is a non-singleton $SCC$. Then there is at least one cycle 
+%from $n$ to itself. We consider \emph{cycles} $c: n \rightsquigarrow n \in SCC$
+%such that there is no repeated node in $c$, and sum up the edge weights along the edges in $c$. If all 
+%cycles including $n$ have the same value $-d$, then we assign a parallelization factor of $d$ to $n$. That 
+%means that $n$ depends on the value computed at $n$ exactly $d$ iterations earlier, and it does not depend 
+%on node $n$ of iterations in-between. In contrast, if there are two distinct cycles with two different weights, 
+%then we assign the default value of 1 to $n$. Intuitively, two distinct weights on cycles, say $-d_1$ and $-d_2$ 
+%means that $n$ depends on the value computed at $n$ $d_1$ iterations ago and on the value computed 
+%$d_2$ iterations ago. Taking all nodes in the SCC, if all nodes have the same parallelization factor $d$, 
+%then we do nothing, otherwise we assign to all nodes factor -1. \ana{More intuition here, why is this important etc}
+
+\ana{Have to make some kind of claim that certain \% of bench falls under this assumptions or otherwise our results don't count for much.}
+
+\subsection{Theoretical Guarantees}
+\label{sec:results}
+
+Unfortunately, finding an optimal across-loop amortized schedule is NP-complete, just as finding an optimal loop-body schedule is.
+Our key premise is that one can still compute an optimal across-loop amortized schedule for real-world loops. Given certain 
+properties of the loop, the amortized schedule we compute in the previous section turns out to be correct and optimal.
+
+The first theorem below establishes correctness, i.e., that $\mathcal{A}$ computes the same result as $\mathcal{S}$ where
+$\mathcal{S}$ is the standard sequential schedule of the loop.
+
+\begin{theorem}(Correctness)
+Let $\mathcal{A}$ be the amortized schedule (computed as in~\secref{sec:scc}) and let $\mathcal{S}$ be the standard sequential 
+schedule of the loop. $\mathcal{A}$ preserves dependences, i.e., for every def-use chain in $(d,u)$ in $\mathcal{S}$, we have that
+$\mathcal{A}(d)$ precedes $\mathcal{A}(u)$ in $\mathcal{A}$.
+\end{theorem}
+
+\begin{proof}
+The proof is straight-forward by case analysis. Case 1 considers when the variable in the def-use chain is a scalar, and Case 2 considers when 
+it is an array location. 
+\end{proof}
+
+Correctness of execution of $\mathcal{A}$ is a simple corollary of the theorem. This follows by induction. Assume that all statements that 
+execute before a given statement $s$ in $\mathcal{A}$ compute the same result in $\mathcal{A}$ as in $\mathcal{S}$. We will consider 
+the execution of $s$ and show that $s$ in $\mathcal{A}$ computes the same result as $s$ in $\mathcal{S}$. This is obvious from the 
+theorem: the computation of the operands of $s$ precedes $s$ in $\mathcal{A}$ and by the inductive hypothesis, they compute the
+same value as in $\mathcal{S}$; therefore, $s$ computes the same result.
+
+%\ana{Correctness theorem should come here...}
+
+\begin{theorem}(Optimality)
+Let $G$ be a constant-rate graph that entails a DAG $D$, where the nodes of $D$ are the strongly connected components of $G$. 
+If $D$ is linear, and each $SCC \in D$ entails a loop-body schedule $S: P_1 \rightarrow P_2 ... P_k$, s.t. (1) $S$ is optimal 
+and (2) for every $P_i, i<k$, for every $n_1 \in P_i, \exists n_2 \in P_{i+1}$, s.t. $n_1 \rightarrow n_2 \in G$, then $\mathcal{A}$ is optimal.
+\end{theorem}
+
+\begin{proof} (Sketch)
+Each SCC schedule is optimal because 1) dependences are the same across the loop iterations, and 2) there is dependence between 
+iteration $i$ and iteration $i+1$, no node in $i+1$ can be computed before the last node, i.e., $P_k(i)$ has been computed, as all 
+nodes in $i+1$ depend on $P_k(i)$. \ana{Expand here.}
+\end{proof}
+
+One can also use Aiken's result to compute an optimal greedy schedule using $n^2$ iterations, where $n$ is the number of nodes 
+in the loop-body schedule. \ana{Add citations.}
+
+\section{Divide and Conquer Parallelization}
+\label{sec:divide}
+
+\ana{BIG TODO...}
+
+\section{Nested Loops}
+\ana{Wish list :)}
+
+\section{Implementation and Evaluation}
+\label{sec:implementation}
 

### 2021-02-06 chapterfour/analysis.tex
added MPC loop material, from chapter four of Lindseys Dissertation

\section{Loop-body Analysis}
\label{sec:loopbodyanalysis}

We first consider the answer to \problemref{problem:prob1} (in MPC). 
%\ana{Lindsey, you have to define Problem 1. This is the problem to compute the optimal 
%schedule taking into account just intra-loop dependences.}
We argue that optimal parallelization within a loop body is NP-Hard (under the above stated assumptions). 
Therefore, compilers must resort to heuristics to compute a schedule for the instructions within a loop body.

We consider two operations, call them $A$ and $M$. $A$ and $M$ are two abstract MPC instruction, 
but as an example, $A$ stands for the ADD MPC instruction 
and $M$ stands for the MUL instruction. Each instruction in the program is either an $A$-instruction or
an $M$-instruction. In order to benefit from parallelization/amortization, we must schedule two or more 
$A$-instructions in the same parallel node (or two or more $M$-instructions in the same parallel node). 
Scheduling $A$-instructions in parallel with $M$-instruction does not benefit from amortization.
It incurs the exact same cost as scheduling the $A$-instructions in a node $P_A$, scheduling the $M$-instructions 
in a node $P_M$, and having $P_A$ precede $P_M$ in the parallel schedule. This is the difference 
between classical scheduling, as studied in parallel computing, and MPC scheduling.

In our theoretical treatment, we make the following assumptions:

\begin{enumerate}

\item $A$ and $M$ are of equal cost, 1 unit.
\item There is unlimited bandwidth---i.e., a single $A$-instruction (or $M$-instruction) costs as much as $N$ amortized $A$-instructions 
(or $M$-instructions), namely 1 unit.

\end{enumerate}

%  Furthermore the run time cost of 
% computing operations of the same type remains the same regardless of the number of operations that are run.
% Roughly speaking, 1 A operation costs the same as 100 A operations and 1000 A operations.

Consider a loop body that consists of $n$ sequences: $S_1$, ... $S_n$ of $A$ and $M$ instructions. 
More precisely, the loop body is such that its instructions can be grouped into such sequences. 
$S_1$, ... $S_n$ can execute in parallel, however, all instructions within a sequence must 
execute sequentially. For example, consider the three sequences (the right arrow indicates a \emph{dependence},
meaning that the source node must execute before the target node): 
\begin{enumerate}
\item $A \rightarrow M \rightarrow A$
\item $A \rightarrow A \rightarrow A$
\item $M \rightarrow A \rightarrow M$
\end{enumerate}

A \emph{schedule} $P: P_1 \rightarrow P_2 \dots \rightarrow P_k$ is such that for each sequence 
$S_i$ in the set, if $S_i[k]$ precedes $S_i[k']$ in $S_i$ then $S_i[k]$ is scheduled in node $P_l$, $S_i[k]$ 
is scheduled in node $P_{l'}$, and $P_l$ precedes $P_{l'}$ in $P$.

The cost of a schedule $P$ is

\begin{equation}
\mathit{cost}(P) = \sum_{i=1}^k \mathit{cost}(P_i)
\end{equation}

where $\mathit{cost}(P_i) = 1$ if $P_i$ consists of $A$-instructions only, or $M$-instructions only, 
and $\mathit{cost}(P_i) = 2$ if $P_i$ mixes $A$-instructions and $M$-instructions.

The problem is to find a schedule $P$ with \emph{minimal cost}. For example, 
a schedule with minimal cost for the sequences above is
\[ A(1), A(2) \rightarrow M(1), A(2), M(3) \rightarrow A(1), A(2), A(3) \rightarrow M(3) \]
(The parentheses above indicate the sequence where the instruction comes from: (1), (2), or (3).)
The cost of this schedule is 5.

The problem of finding a schedule $P$ with a minimal $cost(P)$ for a given loop body has been shown
to be an NP-Hard problem, as it can be reduced to the problem of finding a \emph{shortest common supersequence}, 
a known NP-Hard problem\cite{Maier1978},\cite{Vazirani2010}. The shortest common supersequence problem is as follows: 
given two or more sequences find the the shortest sequence that contains all of the original sequences. This can be solved
in $O(n^k)$ time, where $n$ is the cardinality of the longest sequence and $k$ is the 
number of sequences. For our problem $n$ is the maximum length of a node and $k$ is the number
of total number of nodes. %\ana{I think it is clear without additional explanation...}
%\ana{Also, add the full statement of the problem here. Now it's not clear what's NP-hard, what's n, what's k, etc.?}
%\lindsey{not exactly sure what need to go it there, but does that make it clearer?}

To see the reduction, 
suppose $P$ is a schedule with minimal cost (computed by a black-box algorithm). 
We can derive a schedule $P'$ with the same cost as $P$, by mapping each mixed node $P_i \in P$ 
to two consecutive nodes in $P'$: an $A$-instruction node followed by an $M$-instruction node.
Clearly, $P'$, which now is a sequence of $A$'s and $M$'s, is a supersequence of each sequence 
$S_i$, i.e., $P'$ is a common supersequence 
of $S_1 \dots S_n$. It is also a shortest common supersequence. (To see this, suppose, a 
shorter common supersequence, $P''$, exists. $P''$ is a schedule of $S_1 \dots S_n$
and the $\mathit{cost}(P'')$ equals the length of $P''$. Since $P'$ is longer than
$P''$, and $\mathit{cost}(P') = \mathit{cost}(P)$, that means $\mathit{cost}(P'') < \mathit{cost}(P)$, 
which is a contradiction since $P$ is a schedule with minimal cost.)

%\ana{Question: are these two assumptions too strong?} \lindsey{I don't think so? makes sense to me}

%\ana{Assumption 1 can be relaxed --- e.g., if $A$ is 5 times cheaper than $M$, 
%we can construct a program where all $A$'s are sequences of five consecutive 
%$A$'s, which reduces to the previous problem.} \lindsey{it doesn't matter I think... as long as it is a static 
%difference.}

%\ana{Assumption 2, I am not sure if this is acceptable, or maybe it is too strong. My intuition is that incorporating into the cost a dependence on the number of instructions in each parallel node, is going to make the problem harder, not easier...} 
%\lindsey{I think that the important note is that it is mannnnny magnitudes of 10 cheaper}
%\lindsey{Ana, Is this now a stale comment?}

\section{Across-loop Analysis}
\label{sec:analysis}

We have shown that in general scheduling a loop body is NP-hard. However, in practice loop bodies consist
of relatively small number of statements. Multiple parallel sequences as in the proofs above, are uncommon.
Therefore, one can compute the shortest common supersequence using a brute force approach.

In this section, we describe the computation of the across-loop schedule. The algorithm proceeds in several phases. 
First, we extend Cytron et al.'s algorithms, outlined in \secref{sec:ssa}
to construct the MPC-MUX nodes. Even though there is significant overlap between SSA construction 
and translation to MPC target code, the problem of placing MUX nodes has not been previously addressed.
We define an algorithm that addresses this problem in~\secref{sec:cfgtompc}. Second, we handle arrays in a way that
improves parallelization. We describe this handling in~\secref{sec:arrayhandling}. These two parts give rise to a graph where  
each node is an MPC target code statement. The edges represent the intra-loop control flow and largely 
overlap with the original CFG edges. The statements are in SSA form and $\phi$-nodes are replaced with MUX nodes. 
We call this graph the MPC $G$.

Once we have constructed the MPC $G$ for the loop body, we construct the dependence graph $G$, which reflects
intra-loop dependences and across-loop dependences. We describe the construction of $G$ 
in~\secref{sec:depgraph}. % and~\secref{sec:depgraphalgo}. 
%\ana{Broken link}\lindsey{removed the second one. links to old section of paper that is commented out.}
This graph gives rise of the across-loop schedule, which we describe in~\secref{sec:loopscheduling}. 
%\ana{Lindsey, can you fix these references?}

\subsection{High-level CFG to MPC $G$} 
\label{sec:cfgtompc}

\subsubsection{From $\phi$-nodes to MUX-nodes}

%\ana{Difficulties: Translate to MUX nodes, adapt classical SSA algorithm, arrays, reads and writes!}

% \begin{figure}
% \small
% \[
% \begin{array}{l@{~~~~~}l}
%   \begin{array}{l@{~}l@{~~~}l}
%   s & ::= s ; s \\ 
%   & \mid \x = \y \; \mid \x=\y \; \op \; \z \mid {\sf A[f(i)] = x} \; \mid {\sf x = A[f(i)]} & \mathit{assignment} \\
% %  & \mid \x = \phi(\y,\z) \\
%   & \mid {\sf for} \; (i = 0; i \le \mathit{Const}; \mathit{i\!+\!+}) \; \{ \; s \; \} \\
%   & \mid {\sf if} \; (\x \; \cop \; \y) \; \{ \; s \; \} \; {\sf else} \; \{ \; s \; \} \\
%  % \bool &::= \x \; \cop \; \y & \mathit{comparison} \\
%   \op &::= + \mid - \mid * \mid / & \mathit{arithmetic} \; \mathit{operator} \\
%   \cop &::=  \; == \; \mid \; != \; \mid \; < \; \mid \; \le & \mathit{comparison} \; \mathit{operator}

%   \end{array}
% \end{array}
% \]
% %\vspace{-0.2in}
% \caption{IMP syntax as defined in \cite{Ishaq2019}: $s$ is a sequence of statements. 
% \code{x}, \code{y}, \code{z}, and $i$, are variables. \ana{Remove Fig. 16 because you have Fig. 14. But take f(i) for array indices in 14.}
% } 
% \label{fig:syntax}
% \end{figure}

Again, we consider a standard IMP-like\cite{Nipkow2014} imperative language.
Without loss of generality, we assume that the syntax gives rise to a CFG and we can apply the 
Cytron et al.'s Algorithm for the efficient computation of SSA over this CFG~\cite{Cytron1991}
(shown in \algref{alg:cytronphiplacement} and \algref{alg:cytronlabeling}).

% Once we have the code in 
% SSA form with we can proceed to transform if statements into MUX nodes.

\begin{figure}[h]
\centering
\begin{minipage}{0.7\textwidth}
\begin{javacode}
z = 0;
...
if (x > 0) {
  if (y > 0) {
    z = 1;
  }  else {
    z = -1;
  }
}
print(z);
\end{javacode}
\end{minipage}
\caption{Branching Code.}
\label{fig:examplecfgcode}
\end{figure}

In particular, we are interested in a sequence of statements that form a loop body, i.e., while 
the outermost statement is a {\sf for} the remaining statements are assignments and {\sf if}s.
There are two principal hurdles. One is to translate the CFG into SSA form, where $\phi$ nodes 
are turned into MUX nodes. One cannot simply reuse the classical algorithm by Cytron et. al. 
because $V_3 = \phi(V_1,V_2)$ nodes do not retain information about the predicate that triggers 
$V_1$ vs. $V_2$; all $\phi$ nodes represent is conditional control flow, i.e., that $V_3$ is either 
$V_1$ or $V_2$. Another issue is that SSA $\phi$ nodes can combine multiple conditionals.

\begin{figure}[H]
\centering
\begin{minipage}{0.30\textwidth}
\includegraphics[width=\textwidth]{cfgphinode}    
\end{minipage}
\caption{CGF after \algref{alg:cytronphiplacement} is run.
The $\phi$ node is added but no labeling has taken place.}
\label{fig:cfgwithphinode}
\end{figure}

\begin{figure}[H]
\centering
\begin{minipage}{0.38\textwidth}
\includegraphics[width=\textwidth]{cfgmuxnode}    
\end{minipage}
\caption{Result of running \algref{alg:cytronextension}. 
MUX nodes are added at merges and refer to conditionals.}
\label{fig:cfgwithmuxnode}
\end{figure}

To illustrate, consider the code shown in \figref{fig:examplecfgcode}.
Line 10 is the dominance frontier of the assignments to 
\z, and Cytron et. al.'s algorithm creates a $\phi$ node at 10: $z_3 = \phi(z_0, z_1, z_2)$. 
There is no information retained that $z_0$ is triggered when $x \le 0$, $z_1$ is triggered when 
$x > 0 \; \wedge \; y > 0$, and $z_2$ is triggered when $x > 0 \; \wedge \; y \le 0$. The Multiplexer at the join point
needs this information in order to assign the correct value to $z_3$ at runtime.

We propose a simple extension to the classical algorithm. Recall from Cytron et. al.'s 
that efficient static single assignment computation has two phases, phase one is the 
placement of phi-nodes at dominance frontiers of assignments, and phase two is the 
renaming of variables $V$ into $V_1$, $V_2$, etc. \figref{fig:cfgwithphinode}
shows the CFG after \algref{alg:cytronphiplacement} is run. The $\phi$ node is placed but the 
variables have not been labeled correctly. Instead of transforming the phi node completley
we instead create MUX nodes that reference the conditionals that are associated with each variable.
MUX nodes are then placed at the merge points as shown in, and all the variables are labeled,
\figref{fig:cfgwithmuxnode} shows the CFG after this step. The algorithm for this 
extension is shown in \algref{alg:cytronextension}.

%\lindsey{changed font size in algo to footnotesize.... is that too small?} \ana{It's good.}
\begin{algorithm}
\footnotesize
\begin{algorithmic}[1]
\ForAll {variables $V$}
\State $C(V) \leftarrow 0$
\State $S(V) \leftarrow \mathit{EmptyStack}$
\State $P \leftarrow \mathit{EmptyStack}$
\EndFor
\State call SEARCH($\mathit{Entry}$) \\
\State ... \\
SEARCH(X) : \\
\ForAll {statements $A$ in $X$} 
\If {$A$ is a $\phi$-assignment}
   \State {remove $A$ and continue}
\EndIf
\ForAll {variables $V$ in $\mathit{RHS}(A)$}
   \State {replace $V$ with $V_i$ where $i = \mathit{Top}(S(V))$}
\EndFor

\ForAll {variable $V$ in $\mathit{LHS}(A)$}
   \State { $i \gets C(V)$} 
   \State {replace $V$ with $V_i$ in $\mathit{LHS}(A)$}
   \State {push $i$ onto $S(V)$}
   \State {$C(V) \gets i+1$}
\EndFor
\EndFor
\ForAll {$Y \in Succ(X)$}
\If {$X$ is a conditional node}
   \State {push $\mathit{Cond}, \mathit{Branch}$ onto $P$ // $\mathit{Branch}$ is $T$ if $X \rightarrow Y$ is the True branch, $F$ otherwise}     
\EndIf   
     \State{// $Y$ is not done if $Y$ has an incomplete MUX predecessor or an original CFG predecessor}
      \If {$Y$ has $\phi$-assignments and $Y$ is not done} 
          \State {$\mathit{Cond}, \mathit{Branch} \gets$ pop $P$}
          \ForAll {$\phi$-assignments in $Y$}
               \If {$Z\!: V = \mathit{MUX}(Cond, ..., ...) \in \mathit{Pred}(Y)$}
                  %\STATE {replace $V$ in $RHS(Z)$ with $V_i$, where $i = Top(S(V)$}
                  \State {remove edge $X\rightarrow Y$, add edge $X\rightarrow Z$ to $\mathit{CFG}$}
                  \State {call SEARCH($Z$) // updates $V$'s in LHS and RHS} 
               \Else
                   \State {$i = Top(S(V))$}
                   \If {$\mathit{Branch} == T$}
                      \State {create new node $Z\!: V = \mathit{MUX}(\mathit{Cond}, V_i, V)$}
                   \Else 
                      \State {create new node $Z\!: V = \mathit{MUX}(\mathit{Cond}, V, V_i)$}
                   \EndIf   
                   \State {remove edge $X\rightarrow Y$, add edge $X\rightarrow Z$ to $\mathit{CFG}$} 
               \EndIf
          \EndFor 
      \EndIf
\EndFor
\ForAll {$Y \in \mathit{Children}(X)$}
   \State {call SEARCH($Y$)}
\EndFor
\ForAll {assignments $A$ in $X$}
   \ForAll {variables $V$ in oldLHS($A$)}
       \State {pop S(V)}
   \EndFor
\EndFor
\end{algorithmic}
\caption{Extension of \algref{alg:cytronphiplacement} replacing $\phi$ nodes with MUX nodes}
\label{alg:cytronextension}
\end{algorithm}

%\lindsey{Ana, I think I got it now! Check the figures!} \ana{Yes, excellent!}
% \ana{TODO: Ana: Here we need an explanation for lines 23-45. Turning n-ary phi-nodes into binary MUX nodes, 
% just the same classical algorithm with minor extension.} \lindsey{which lines/figure?}\ana{Lines in the Algo. I'll do it.}
% \ana{TODO: Lindsey: Example of CFGs for above example.} \lindsey{I made the CFG.... but am still confused, there
% are only 10 lines in the example?} \ana{Yes, that one's almost there. But this is the MUX graph, each MERGE becomes a MUX.}
% \ana{This graph is the After graph, with each MERGE node now a MUX node. In the after graph, show the MUX what arguments it takes.} 
% \ana{We also need the Before graph which has just a single Dominance Frontier node that merges control from all the conditionals.}
% \lindsey{now I am really confused.... how many graphs should there be?}

\subsubsection{Handling of Arrays}
\label{sec:arrayhandling}

So far we focused on the classical SSA analysis and the extension to create MUX nodes. 
Classical SSA treats arrays as scalars, i.e., an array write, e.g., {\sf A[i] = x+y} is considered
a modification of {\sf A} and warrants its own indexed variable ${\sf A}_k$. A later use of {\sf A}, 
even one such as {\sf x = A[i+1]} where there is no def-use link from {\sf A[i] = x+y}, will be 
considered a use of ${\sf A}_k$. This hinders parallelization. To illustrate, consider 
what is shown in \figref{fig:arrayhand}.
%\begin{minted}[fontsize=\footnotesize, numbersep=5pt, escapeinside=||]{java}

\begin{figure}[H]
\centering
\begin{minipage}{0.7\textwidth}
\begin{javacode}
for (...) {
  A: A[i] = f_1(B[i]);
  B: B[i] = f_2(A[i]);
  ...
}
\end{javacode}
\end{minipage}
\caption{Array handling.}
\label{fig:arrayhand}
\end{figure}

%\end{minted}
Since {\sf A} and {\sf B} are treated as standard scalars, the dependence analysis will record
a dependence from {\sf B} to {\sf A} via the back edge. As a result, statement {\sf A} wouldn't 
be vectorized because each (i+1)-th iteration depends on {\sf B} in the i-th iteration,
which depends on {\sf A} in the i-th iteration.

To mitigate against this, we propose a technique, similar to Array-SSA though much simpler and 
less general, which breaks certain dependence edges and allows for better vectorization. 
We term this technique \emph{scalar substitution}.
%\ana{Have to make a claim how often this technique resolves dependences perfectly, and how often it defaults to standard SSA.}

%\ana{TODO: Summary of the technique}

The essence of scalar substitution is that in certain cases, one can treat array accesses as distinct scalar accesses.
For example, reads and writes of $\A[i]$ are treated as reads and writes to a scalar variable $\A[i]$.
Our SSA transformation assigns versions $\A[i]_0$, $\A[i]_1$, etc. when $\A[i]$ is written in the loop body. 
The key issue is that we need to properly account for across-loop dependences. When there is an array write 
$\A[f(i)] = ...$ and an array read $ ... = \A[f'(i)]$ and the array references have been replaced with scalar variables, 
our analysis must properly account for across-loop dependences from one scalar variable to another.

In summary, our technique first outlines conditions on arrays that will allow us to safely carry such transformation.
These are the Array Well-formedness conditions defined below. If these conditions hold for array \A, then when 
we do scalar substitution for \A: we replace 
array accesses of {\A} with scalars. Otherwise, we leave {\A} as is, defaulting to the standard treatment in Cytron et al. 
The standard treatment treats writes to different indices of {\A} as writes to a single variable {\A}, and similarly, reads
to different indices of {\A} as reads of that single variable. This is safe, however, it may miss opportunities for
parallelization. In our experience, our Array Well-formedness conditions are not restrictive, as in the majority of benchmarks
(typical MPC benchmarks), arrays are read-only. However, more experiments are needed, particularly 
with benchmarks that are standard in HPC, that we hope to transfer to our setting.

Our technique next constructs the dependence graph taking into account the replacement with scalar variables. 
This is described in~\secref{sec:depgraph}. The algorithm constructs a correct dependence graph taking into 
account loop-carried dependences between scalars in the cases when Array Well-formedness holds.

\paragraph {Equivalence and Non-equivalence} Let $f(i)$ and $f'(i)$ be two index functions on induction variables $i$, 
e.g., we have array accesses ${\A}[f(i)]$ and ${\A}[f'(i)]$. We say that $f(i) \cong f'(i)$ iff for every $1 \le i \le N$,
$f(i) = f'(i)$. We say that $f(i) \ncong f'(i)$ iff for every $1 \le i \le N$, $f(i) \neq f'(i)$. 
%\ana{TODO Ana. This needs to be expanded
%with consideration of equivalence classes of variables in f(i). 
%For some set of variables functions are equivalent for another set they are non-equivalent, etc.
%Have to say something, for a fixed set of constants...}

Strictly, equivalence and nonequivalence is defined under conditions on variables in $f(i)$ and $f'(i)$ 
other than the induction variable $i$. For example if $f(i) = x*i$ and $f'(i) = x*(i-1)$, equivalence and non-equivalence 
depends on $x$. If $x$ is 0 then the two formulas are equivalent and if $x \neq 0$ then they are non-equivalent. 
When we write $f(i) \cong f'(i)$ we mean that the following logical formula is valid:

\begin{equation}
1 \le i \le N \Rightarrow f(i) = f'(i) 
\end{equation}
and similarly, 
\begin{equation}
1 \le i \le N \Rightarrow f(i) \neq f'(i) 
\end{equation}
In future work we will consider conditional equivalence and non-equivalence, i.e, in the above example, we will 
consider formulas
\begin{equation}
x = 0 \Rightarrow (1 \le i \le N \Rightarrow f(i) = f'(i)) 
\end{equation}
and
\begin{equation}
x \ne 0 \Rightarrow (1 \le i \le N \Rightarrow f(i) \neq f'(i) ) 
\end{equation}

\paragraph{Array Well-formedness Conditions} Let ${\A}$ be an array mentioned in the loop body. 
We will treat array accesses ${\A}[f(i)]$ as distinct scalar variables if we can show that the loop body is
well-formed with respect to \A. Otherwise, we treat array {\A} as a single variable, thus defaulting to the 
case of standard SSA. The array well-formedness conditions for array {\A} are defined as follows:

\begin{enumerate}
\item For each pair of array writes to \A, $n_1\!: \A[f(i)] = ... $ and $n_2\!: \A[f'(i)] = ... $, we have either $f(i) \cong f'(i)$ or 
$f(i) \ncong f'(i)$.

\item In addition, for each pair of array writes to \A, $n_1\!: \A[f(i)] = ... $ and $n_2\!: \A[f'(i)] = ... $, there does not exist 
a pair $1 \le i, j \le N$, $i \neq j$ such that $f(i) = f'(j)$, or in other words array writes are disjoint, i.e., the same array location 
cannot be written in two different iterations. 
%\ana{Check whether one of these is sufficient.}

\item For each pair $n_1\!: \A[f(i)] = ... $ and $n_2\!: ... = \A[f'(i)]$ such that $n_1$ reaches $n_2$ on a forward path, 
we have either $f(i) \cong f'(i)$ or $f(i) \ncong f'(i)$.

%\item For each pair $n_1\!: A[f(i)] = ... $ and $n_2\!: A[f'(i)] = ... $ either $f(i-d) \cong f'(i)$ for some unique $d > 0$, 
%or $f(i-d) \ncong f'(i)$ for every $d > 1$.

\item For each pair $n_1\!: \A[f(i)] = ... $ and $n_2\!: ... = \A[f'(i)]$ %such that $n_1$ reaches the loop exit (we assume that there is
%a unique loop exit node) on a forward path, and there is  $n_2$ on a forward path, 
we have either $f(i-d) \cong f'(i)$ for some unique $d$ s.t. $1 \le d < N$, or $f(i-d) \ncong f'(i)$ for every $d$ s. t. $1 \le d \le N$.

In this $d$ is the distance between the iterations in the loops i.e. For if statement A has a $d$ value of 2 from statement B, 
statement A was computed 2 loop iterations before statement B.

\end{enumerate}

To see the intuition behind these rules, consider rule (1). 
In other words, the two array writes are either equivalent, i.e., at each iteration they modify the same
location, or they are non-equivalent, i.e., at each iteration they modify distinct locations. Note that the two array writes
``clash'' in one way or another. If $n_1$ reaches $n_2$ (or vice versa, $n_2$ reaches $n_1$), then if $f(i)$ and $f'(i)$ 
are neither equivalent nor non-equivalent, there would be a kill in some iterations and not in others; in contrast, if
they are equivalent, there is always a kill, and if they are non-equivalent, there is never a kill. If neither $n_1$ reaches
$n_2$ nor the other way around, then the two array writes will ''clash'' at a join node. If $f(i)$ and $f'(i)$ are neither 
equivalent nor non-equivalent, then there would be one kind of MUX for certain iterations and a different kind of MUX 
for the rest; if $f(i)$ is equivalent to $f'(i)$, then the MUX will be $... = \mathit{MUX}(\mathit{cond}, {\A}[f(i)], {\A[f'(i)]})$, and 
$f(i)$ and $f'(i)$ are non-equivalent then there will be two MUX nodes: $... = \mathit{MUX}(\mathit{cond}, {\A}[f(i)], {\A'[f(i)]})$
and $... = \mathit{MUX}(..., {\A}[f'(i)], {\A'[f'(i)]})$. (Note that after SSA, the A's will be properly indexed.)
%\ana{I need explanations of (2) and (3).}

Condition (2), disjointness of writes, is necessary for our graph construction algorithm to properly account
for loop-carried dependences. We have to make sure that a write in a later iteration $i$ does not ''kill'' a write in an earlier iteration $j$.

Similarly, for (3) if an array write reaches an array read they either reference 
the same location or never reference the same location. In other words, there is either an intra-loop 
def-use in each iteration, or there is no intra-loop def-use in any iteration. 
More concretely if we have $A[f(i)] = ...$ flow into $... = A[g(i)]$ then for $1 \le i \le N, f(i) = g(i) ~ | ~ \mbox{ for } 1 \le i \le N, f(i) \neq g(i)$.

Finally, (4) is a little more involved, if there is an array write and then a read and they always differ, 
they always differ by a static constant $d$. This would mean that the read always uses the write $d$ iterations before.
%So then if we again have $A[f(i)] = ...$ flow into $... = A[g(i)]$ and 
%$\forall i \in \Z, f(i) \neq g(i)$ then $\forall i \in \Z, f(i) - g(i) = d$ where $d \in \Z$. \lindsey{Ana, doe this make sense}

\paragraph{Example} Consider our running example in \figref{fig:arrayloop}. 
There is a single write to each array in the loop body,  so there are no write-write pairs and no 
tests for (1). There are the following write-read pairs: 2 and 3 (access of \A), 
2 and 4 (access of \A), 3 and 5 (access of \B), 4 and 5 (access of \C), 3 and 2 (access of \B), 5 and 3 (access of \D) 
and 5 and 4 (access of \D).
We apply the forward test, (2), only on the first three pairs, as for the remaining ones, there is no forward path from 
the write to the read. 
Test (2) clearly passes because $i$ is obviously equivalent to $i$. For the remaining pairs, we only apply the backward 
test, test (3). 
Pair 3 and 2 passes as well because $i-1$ and $i$ are nonequivalent for every $d > 0$. 5 and 3, and analogously 5 and 
4, pass test (3). 
In these two cases, the unique $d = 1$ makes $i - d$ and $i-1$ equivalent.

\paragraph{Checking well-formedness}

%\ana{TODO Ana: Add paragraph on checking.}

We ensure the above conditions as follows:
For conditions 1. and 3. we check the validity of the following formula: 
\begin{equation}
(1 \le i \le N \Rightarrow f(i) = f'(i)) \vee (1 \le i \le N \Rightarrow f(i) \neq f'(i))
\end{equation}
For condition 2 we check validity of
\begin{equation}
(1 \le i,j \le N \wedge i\neq j \Rightarrow f(i) \neq f'(j))
\end{equation}
Finally, for condition 4 we find $d$ by asking satisfiability of
\begin{equation}
1 \le i \le N \Rightarrow f(i-d) = f'(i)
\end{equation}
and then ensure validity of
\begin{equation}
1 \le i \le N \wedge 1 \le d' < N \wedge d' \neq d \Rightarrow f(i-d) \neq f'(i)
\end{equation}

This checks can be done with off the shelf SMT solvers such as Z3.

%\ana{Where do we put this? Theorem: simple rule is not restrictive} \ana{This one can be left out.}
%\ana{TODO Ana: Theorem 2: program is equivalent to transformed one where each A[f(i)] is replaced with a scalar.}  
%\ana{The idea is that we replace each A[f(i)] with scalar, i.e., we treat A[f(i)] like a scalar. E.g., if we have A[i] in our SSA algorithms we are treating it as $A_i$ , a variable. We'll compute versions of Ai like Ai_1, Ai_2, etc., just like for scalars.} 
%\lindsey{I think we cover this above now? Is this a stale comment?}

\subsection{Across-loop Dependence Graph $G$: Overview}
\label{sec:depgraph}

Our next step is to determine dependences, including intra-loop dependences and more complex 
across-loop dependencies through the creation of the dependence graph $G$ and analysis and 
reasoning over $G$. Our transformation analyzes arrays as described earlier, and does scalar substitution
for well-formed arrays, while treating non-well-formed arrays in the standard way. Then we use \algref{alg:cytronextension}
on the loop body code to add MUX nodes.

Each node of the $G$ represents an MPC target code statement in a loop body, while each edge
represents a def-use dependency between the two connected nodes. The analysis considers
each statement of the loop. If the statement contains an array reference (Note, 
that SSA guarantees that there can be only one array reference in any single
statement) it is classified as a read, or a write. The statement is a write only
if the statement is an assignment statement, and the array reference is on
the left-hand side of the assignment, otherwise it is classified as a read.
This statement, its original source line number, and index of the array
reference are stored as a node. Scalars that change per iteration are also
handled. Every scalar that changes in the loop body is defined in a $\phi$
node at the head of the loop, and the changed variable is defined
at the end of the loop. This makes tracing reads and writes of 
changing variables as easy as tracing induction variables through the loop
and noting the node that defines the changed variable.
%\ana{Lindsey, you talk about array references, but how about scalar reads and writes?}
%\lindsey{not explained perfectly, work in progress, but does that make more sense?}
%\ana{Somewhere here we need to say that we apply Alg. 3 to produce the SSA.}
%\lindsey{blunt, but effective? Is that OK?}

%\ana{This was repeated from above, removed.}
%Each node of the SCC represents a statement in a loop body, while each 
%edge represents a dependency between the two connected nodes. The parser 
%analyzes each statement of the loop. If the statement contains an array 
%reference (Note, that SSA guarantees that there can be only one array reference 
%in any single statement) it is classified as a read, or a write. The statement 
%is a write only if the statement is an assignment statement, and the array reference 
%is on the left-hand side of the assignment, otherwise it is classified as a read. 
%This statement, its original source line number, and index of the array 
%reference are stored as a node.

After all the nodes have been found, edge creation and labeling begins. 
For each node, $c$, in the graph the following steps are taken to determine 
the distance between each pair of def/use statements. This value determines 
the parallelization schedule for each pair interaction 
in the context of the larger loop body. The full algorithm is shown in 
\algref{alg:edgecreation}.

\begin{algorithm}
\begin{algorithmic}
\State // Step 1: Forward edge construction 
\ForAll{$n_1 \in \mathit{Nodes}(G)$}
\ForAll{$n_2 \in \mathit{Nodes}(G)$}
\If{$n_1 \neq n_2$}
\ForAll{$v \in \mathit{Scalars}(G)$}
%\State $MaxNode \leftarrow GetMaxNodeScalar(v)$
\If{$v \in \mathit{LHS}(n_1)$ \&\& $v \in \mathit{RHS}(n_2)$}
\State $\mathit{Edges} \leftarrow \mathit{Edges} \cup \mathit{ForwardEdge}(n_1, n_2)$
\EndIf
\EndFor
\ForAll{$\A \in \mathit{Arrays}(G)$}
\If{$\A[f(i)] \in \mathit{LHS}(n_1)$ \&\& $\A[f'(i)] \in \mathit{RHS}(n_2)$ \&\& $f(i) \cong f'(i)$} %$\forall i \in \Z, f(i) == f'(i)$}
\State $\mathit{Edges} \leftarrow \mathit{Edges} \cup \mathit{ForwardEdge}(n_1, n_2)$
\EndIf
\EndFor
\EndIf
\EndFor
\EndFor
\State // Step 2: Backward edges for scalars (including non well-formed arrays)
\ForAll{$v \in \mathit{Scalars}(G)$}
\State $\mathit{MaxNode} \leftarrow \mathit{GetMaxNodeScalar}(v)$
\ForAll{$n \in \mathit{Nodes}(G)$}
\If{$n \neq \mathit{MaxNode}$ \&\& $V_0 \in \mathit{RHS}(n)$}
\State $\mathit{Edges} \leftarrow \mathit{Edges} \cup \mathit{BackwardEdge}(\mathit{MaxNode}, n, -1)$
\EndIf
\EndFor
\EndFor
\State // Step 3: Backward edges for well-formed arrays
\ForAll{$\A \in \mathit{Arrays}(G)$}
\State $\mathit{MaxNode} \leftarrow \mathit{GetMaxNodeArray}(\A)$
\State $f(i) \leftarrow \mathit{ArrayWriteIndex}(\mathit{MaxNode})$
\ForAll{$n \in \mathit{Nodes}(G)$}
\If{$\mathit{IsArrayRead}(n)$ \&\& $n \neq \mathit{MaxNode}$ \&\& $A_0 \in \mathit{RHS}(n)$}
\State $f'(i) \leftarrow \mathit{ArrayReadIndex}(\mathit{MaxNode})$
\State $d \leftarrow \mathit{Z3}(f(i), f'(i))$
\State $\mathit{Edges} \leftarrow \mathit{Edges} \cup \mathit{BackwardEdge}(\mathit{MaxNode}, n, d)$
\EndIf
\EndFor
\EndFor
%\EndFor
\end{algorithmic}
\caption{Edge Creation}
\label{alg:edgecreation}
\end{algorithm}

%\ana{The above text is good, we may make some edits later on.}

The following constructs are used in the algorithm

\begin{itemize}
	\item $\mathit{Edges}$ is a set of edges
	\item $\mathit{Nodes}(G)$ returns all nodes in the graph $G$
	\item $\mathit{Scalars}(G)$ returns all scalars in the program including non-well-formed arrays
	\item $\mathit{Arrays}(G)$ returns all well-formed arrays
	\item $\mathit{IsArrayWrite}(N)$ returns true iff $\mathit{LHS}(N)$ is an array write
	\item $\mathit{ArrayWriteIndex}(N)$ returns an index $f(i)$ in an array write, where $i$ is the induction variable
	\item $\mathit{ArrayReadIndex}(N)$ returns an index $f(i)$ in an array read, where $i$ is the induction variable
	\item $\mathit{GetMaxNodeScalar}(V)$ returns a node $N$ such that $V_{\mathit{max}} \in N$
	\item $\mathit{GetMaxNodeArray}(A_{\mathit{max}})$ returns a node $N$ such that $A_{\mathit{max}} \in \mathit{LHS}(N)$
	\item $\mathit{ForwardEdge}(N_1, N_2)$ returns a new node between $N_1$ and $N_2$
	\item $\mathit{BackwardEdge}(N_1, N_2, d)$ returns a new node between $N_1$ and $N_2$ with a distance of $d$
	\item $\mathit{Z3}(f(i), g(i))$ returns the distance $d$ between two indices based on the induction variable $i$
\end{itemize}

%%% DO NOT DELETE! OLD VERSION THAT IS CORRECT!

% \begin{algorithm}
% \begin{algorithmic}
% \ForAll{$n_1 \in Nodes(G)$}
% \ForAll{$n_2 \in Nodes(G)$}
% \If{$n_1 \neq \n_2$}
% \ForAll{$v \in Scalars(G)$}
% \If{$v \in LHS(n_1)$ \&\& $v \in RHS(n_2)$}
% \State $Edges \leftarrow Edges \cup ForwardEdge(n_1, n_2)$
% \EndIf
% \EndFor
% \ForAll{$a \in Arrays(G)$}
% \If{$a[f(i)] \in LHS(n_1)$ \&\& $a[f'(i)] \in RHS(n_2)$ \&\& $\forall i \in \Z, f(i) == f'(i)$}
% \State $Edges \leftarrow Edges \cup edge(n_1, n_2)$
% \EndIf
% \EndFor
% \EndIf
% \EndFor
% \EndFor
% \ForAll{$v \in Scalars(G)$}
% \State $MaxNode \leftarrow GetMaxNodeScaler(v)$
% \ForAll{$n \in Nodes(G)$}
% \If{$n \neq MaxNode$ \&\& $V_0 \in RHS(n)$}
% \State $Edges \leftarrow Edges \cup BackwardEdge(MaxNode, n, -1)$
% \EndIf
% \EndFor
% \EndFor
% \ForAll{$a \in Arrays(G)$}
% \State $MaxNode \leftarrow GetMaxNodeArray(a)$
% \State $f(i) \leftarrow ArrayWriteIndex(MaxNode)$
% \ForAll{$n \in Nodes(G)$}
% \If{$IsArrayWrite(n)$ \&\& $n \neq MaxNode$ \&\& $A_0 \in RHS(n)$}
% \State $f'(i) \leftarrow ArrayReadIndex(MaxNode)$
% \State $d \leftarrow Z3(f(i), f'(i))$
% \State $Edges \leftarrow Edges \cup BackwardEdge(MaxNode, n, d)$
% \EndIf
% \EndFor
% \EndFor
% \end{algorithmic}
% \caption{Edge Creation}
% \label{alg:edgecreation}
% \end{algorithm}

%%% DO NOT DELETE! OLD VERSION THAT IS CORRECT!
\begin{figure}[h]
\centering
\begin{minipage}{0.7\textwidth}
\begin{javacode}
     label1:
        i13_1 = Phi(i13 #0, i13_2 #1);
        if i13_1 > b0 goto label2;
        $i1 = r3[i13_1];
        r1[i13_1] = $i1;
        $i4 = r1[i13_1];
        $i2 = i13_1 - 1;
        $i3 = r7[$i2];
        $i5 = $i4 * $i3;
        r3[i13_1] = $i5;
        $i8 = r1[i13_1];
        $i6 = i13_1 - 1;
        $i7 = r7[$i6];
        $i9 = $i8 * $i7;
        r5[i13_1] = $i9;
        $i11 = r3[i13_1];
        $i10 = r5[i13_1];
        $i12 = $i11 * $i10;
        r7[i13_1] = $i12;
        i13_2 = i13_1 + 1;
(1)     goto label1;
\end{javacode}
\end{minipage}
\caption{Transformed loop body from \figref{fig:arrayloop}.} 
\label{fig:transformedaikenexample}
\end{figure}

\begin{figure}[h]
\centering
\begin{minipage}{0.7\textwidth}
\begin{pythoncode}
from z3 import *
# i2 = i13_1 - 1
# i3 = r7[i2]
i2 = Int('i2')
i13_1 = Int('i13_1')
F = [i2 != 0, i13_1 != 0]
s = Solver()
s.add(F)
s.add(i2 == i13_1 - 1)
print(s.check())
if s.check() == z3.sat:
    m = s.model()
    for el in m:
        print(el, m[el])
\end{pythoncode}
\end{minipage}
\caption{Example of Z3 solving for a $d$ value in \figref{fig:transformedaikenexample}.}
\label{fig:z3example}
\end{figure}

As an example of Z3 solving for a $d$ value consider the code in \figref{fig:z3example}.
This code determines a $d$ value for one of the array accesses in our running 
example taken from Aiken et. al \cite{Aiken1988} and shown in~\figref{fig:arrayloop}. 
% The original equation defining an index variable is flattened and all 
% non-induction variables on the right hand side of the equation are replaced with their induction 
% definitions. \ana{I'm not quite sure what you mean here, can you expand/rewrite?} 
In this case the index variable $i2$ is comparted with a its definition $i13\_1 - 1$ where 
$i13\_1$ is an induction varaible that changes every loop iteration. From this we can create an equation to 
determine distance between the index variable we are interested in (in this case $i2$) and other 
and the induction variable (in this case $i13\_1$). Z3 then finds the smallest possible values that 
to satify this equation. By taking the difference of the generated values we determine the $d$ value (in
this case $d = -1$).

We also assume the \emph{constant rate assumption} for every statement that we analyze. This ensures
that any one statement cannot have differing d-values regardless of the number of array accesses on the 
right hand side of the assignment. Line 6 in \figref{fig:fibconstantrate} violates this assumption,
as it simultaneously has a d-values of 1 and 2. In other words, the statement at line 6 depends on itself
1 iteration ago and 2 iterations ago.

\begin{figure}[h]
\centering
\begin{minipage}{0.7\textwidth}
\begin{javacode}
int fib(int n) {
    int[] arr = new int[n];
    arr[0] = 1;
    arr[1] = 1;
    for(int i = 2; i < n; i++) {
        arr[i] = arr[i - 1] + arr[i - 2];
    }
    return arr[n - 1];
}
\end{javacode}
\end{minipage}
\caption{The dynamic program solution for solving the $n$-th Fibonacci number violates the Constant Rate
assumption.}
\label{fig:fibconstantrate}
\end{figure}

\begin{figure}[h]
\centering
\begin{minipage}{0.50\textwidth}
\includegraphics[width=\textwidth]{aikenscc}
\end{minipage}
\caption{SCC for \figref{fig:arrayloop}.}  
\label{fig:aikenscc}  
\end{figure}

Consider our running example shown in~\figref{fig:arrayloop}. 
The output of our SCC transformation is shown in~\figref{fig:aikenscc}. The intra-loop dependencies are shown
as solid black lines, while the cross loop dependencies are shown in dashed black lines. In this case 
we also have multiple cycles: one between lines 3 and 4 and the other between lines 2 and 4. This is caused
by the combination of intra-loop dependencies ($B[i] \rightarrow D[i]$ and $C[i] \rightarrow D[i]$) and
cross-loop ($D[i - 1] \rightarrow B[i]$ and $D[i - 1] \rightarrow C[i]$) dependencies on the same line.

%\ana{Lindsey, here you can give a few examples from the one you have with your code, and draw the graphs. But not the graphs that we construct, but the graphs that we want to compute the SCCs over.} \lindsey{I am not sure which ones you mean? Should I construct them by hand?} \ana{Sorry, yes, this was unclear :). I meant the graphs that we want to compute the SSCs over (not the schedules that we construct). The graph where we have forward edges and backward edges annotated with d, i.e., the graph created by our Alg. 2.} \lindsey{I think it is better to just stick with the aiken example!}

%\subsection{Across-loop Dependence Graph: Algorithm}
%\label{sec:depgraphalgo}

% \ana{Now I think it will be easiest to describe the above algorithm as an attribute grammar, similarly to what we did in Ishaq's paper to construct the MPC code out of the SSA code.}

% \ana{Try rephrasing the above algorithm in terms of an attribute grammar, and we'll see what happens...}

% \ana{Looking at Ishaq's paper, he must have taken this out from the published version, but here it is I found it from the drafts:}

%Once the SCC graph is complete, it is partitioned into two sub-graphs. 
%One is the MPC sub-graph (original computation of loop, based on hidden values), 
%and the other is the index sub-graph  (which is plain-text). By parsing the 
%index sub-graph the individual components of every index can be placed in a 
%single equation. This is how we determine the d values for every pair of 
%Strongly connected components. However if we assume that \emph{all} values
%are sensetive there is no need to partition the graph, as the plain-text sub
%graph would be empty.

%Once the analysis creates the SCC graph along with the d values, it then applies  
%compiler optimizations on each def/use cycle to create a heuristic based schedule. As stated 
%in \secref{sec:hpcparallelization} and \secref{sec:mpccomp} identical operations can be infinitely
%amortized, avoiding the hard processor limit that restricts loop scheduling in HPC applications. 
%This only applies to operations on secret values, which do not include looping variables.

\begin{figure}[h]
\centering
\begin{minipage}{0.70\textwidth}
\includegraphics[width=\textwidth]{unrolledloopdep}
\end{minipage}
\caption{Dependencies for running example. %\figref{fig:flattenedloop}
%\lindsey{Ana, you can copy and paste this for the comment in
%\secref{sec:theoreticalguarentees}!}
}
\label{fig:unrolledloopdep}  
\end{figure}

\begin{figure}[h]
\centering
\begin{minipage}{\textwidth}
\includegraphics[width=\textwidth]{unrolledloopschedule}
\end{minipage}
\caption{Schedule for running example. %\figref{fig:flattenedloop} 
%\lindsey{Ana, you can copy and paste this for the comment in
%\secref{sec:theoreticalguarentees}!}
}
\label{fig:unrolledloopschedule}  
\end{figure}

### 2021-02-06 chapterfour/introduction.tex
added MPC loop material, from chapter four of Lindseys Dissertation

\section{Introduction}
\label{sec:intro}

%\ana{Lindsey, look at the figures to format some of them and put the period at the end of the caption if it's missing.}

As the demand for cloud computation has increased so has the demand for secure solutions 
for cloud platforms. In the last decade cloud serves have become an almost \$100 billion industry. 
\cite{Synergy2019}. This high consumer demand, as well as the need for high security within these
frameworks has fueled research into varying cryptographic concepts like Multiparty Computation and 
Homomorphic Encryption.

Researchers have proposed different encryption schemes to support 
a variety of operations, allowing for larger and larger subsets of
programs to be run this way. Again, Fully Homomorphic Encryption (FHE) \cite{Gentry:2010},\cite{Gentry:20102}  
can perform arbitrary operations on encrypted data, but in its current state FHE
is still prohibitively expensive \cite{Gentry:2011}. An alternative to FHE 
is Partially Homomorphic Encryption (PHE), which scales substantially better, but cannot 
perform arbitrary operations, and requires the conversions that were discussed in \chapref{chapter:chapterthree}.

As we discussed earlier, Multiparty Computation (MPC) describe groups of protocols that can be used in secure collaborative
computation. MPC benefits from having high security stemming from extensive theoretical guarantees. 
%without having to resort to expensive
%solutions (like FHE) or solutions with lower security (PHE conversion).

Within MPC certain optimizations can be made in the evaluation of loops bodies. Due to the transformation of problems into an 
intermediate representation, similar operations within an MPC problem can be amortized. %, or performed in parallel, 
%with little drawbacks as long as certain assumptions hold. 
We present a technique for automatically locating opportunities for amortization to improve performance in compiled MPC programs.

The rests of this chapter is organized as follows: \secref{sec:mpcoverview}, provides a brief over view of our work. 
\secref{sec:language} shows our language specification.
\secref{sec:parallelization} briefly outlines outlines dependenceies, and describes HPC Parallelizion as well as 
MPC Amortization in \secref{sec:hpcparallelization} and \secref{sec:mpcamortization}. 
\secref{sec:loopbodyanalysis} and \secref{sec:analysis} show our analysis. \secref{sec:loopscheduling} 
presents our scheduling algorithms. \secref{sec:results} presents our empirical results. 
% Finally, \secref{sec:conclusions} concludes. 
%\ana{Broken links again here. I think this needs to be rewritten as we moved stuff into Chapter 2.} \lindsey{fixed!}

### 2021-02-06 chapterfour/language.tex
added MPC loop material, from chapter four of Lindseys Dissertation

\section{Language Specification}
\label{sec:language}

We start with an IMP-like \cite{Nipkow2014} language, syntactically similar to imperative
C-style languages (C, C++, Java, C\#, etc.). %\ana{What's this ref to?}\lindsey{my mistake, to nothing removed.} 
We restrict our language by disallowing recursion, and assuming that all loop bounds are both known and static. 
% This leaves us with a general language syntax as shown in \figref{fig:impsyntax}.
Next the program is converted to SSA using a modified version of the algorithms of 
Cytron et al. \cite{Cytron1991}. Those algorithms have two main steps: the creation
of $\phi()$ functions at merge points in the program, and the renaming of variables
to ensure that every variable is only on the left hand side of an assignment statement once.
Crucially during our transformation merges between two blocks that contain different variable
versions are not transformed into $\phi()$ nodes but are instead converted to MUX gates. This
prevents data leakage due to observations of program execution. Once in this form loop 
bodies can be unrolled, allowing the MPC compiler to create a circuit 
representation of the program. Our language is similar to recent work
in the MPC literature that optimizes protocol selection during 
compilation \cite{Ishaq2019}. \figref{fig:muximpsyntax} shows our output syntax. 
We assume standard scalar variables (i.e. $\x$, $\y$, ${\sf c}$) and array members (i.e. ${\A[i]}$) are sensitive 
values, while looping variables and variables that represent loop bounds (i.e. $N$, $i$) are known at compile time.
All functions in our language can be vectorized. This means that any function in our language 
(i.e. ADD) can be computed on two arrays for the same cost as computing the same 
function of two scalar values.
%\ana{This last sentence is unclear, try to rewrite.}
%\lindsey{better?}
% However, as stated above, do not include $\phi()$ functions directly 
% and convert if statements to MUX nodes.

% \begin{figure}
% \small
% \[
% \begin{array}{l@{~~~~~}l}
%   \begin{array}{l@{~}l@{~~~}l}
%   s & ::= s ; s \\ 
%   & \mid \x = \y  \\
%   & \mid \x=\y \; \op \; \z \\
%   & \mid {\sf A}[i] = \x \\
%   & \mid \x = \A[i] \\
%   & \mid {\sf for} \; (i = 0; i \le n; \mathit{i\!+\!+}) \; \{ \; s \; \} \\
%   & \mid {\sf if} \; (\x \; \cop \; \y) \; \{ \; s \; \} \; {\sf else} \; \{ \; s \; \} \; & \mathit{statement} \\ 
%   \op &::= + \mid - \mid * \mid / & \mathit{arithmetic} \; \mathit{operator} \\
%   \cop &::=  \; == \; \mid \; != \; \mid \; < \; \mid \; \le & \mathit{comparison} \; \mathit{operator}

%   \end{array}
% \end{array}
% \]
% %\vspace{-0.2in}
% \caption{Source IMP syntax as defined in \cite{Nipkow2014}: \sss is a sequence of statements. 
% \x, \y, \z, \iii, and \nnn are variables.} 
% \label{fig:impsyntax}
% \end{figure}

\begin{figure}
\small
\[
\begin{array}{l@{~~~~~}l}
  \begin{array}{l@{~}l@{~~~}l}
  s & ::= s ; s \\ 
  & \mid \x = \y  \mid \x = p \mid \A[i] = \x \mid \x = \A[i] \\
  & \mid \A[f(i)] = \x \; \mid \x = \A[f(i)] 
  \end{array}
\end{array}
\begin{array}{l@{~~~~~}l}
  \begin{array}{l@{~}l@{~~~}l}  
  p &  ::= {\sf ADD(x, y) \mid ADD\_N(X, Y, n)} & \mathit{primitive} \\ 
  & \mid {\sf SUB(x, y) \mid SUB\_N(X, Y, n)} &  \\ 
  & \mid {\sf MUL(x, y) \mid MUL\_N(X, Y, n)} &  \\ 
  & \mid {\sf CMP(x, y) \mid CMP\_N(X, Y, n)} &  \\ 
  & \mid {\sf XOR(x, y) \mid XOR\_N(X, Y, n)} &  \\ 
  & \mid {\sf SHL(x, y) \mid SHL\_N(X, Y, n)} &  \\ 
  & \mid {\sf SHR(x, y) \mid SHR\_N(X, Y, n)} &  \\ 
  & \mid {\sf REM(x, y) \mid REM\_N(X, Y, n)} & \\ 
  & \mid {\sf OR(x, y)  \mid OR\_N(X, Y, n)} &  \\ 
  & \mid {\sf AND(x, y) \mid AND\_N(X, Y, n)} &  \\ 
  & \mid {\sf MUX(x, y, c) \mid MUX\_N(X, Y, c, n)} &  \\ 
  \end{array}
\end{array}
\]
%\vspace{-0.2in}
\caption{Target MPC syntax: $s$ is a straight-line sequence of statements. 
\x, \y, and {\sf i} are variables, {\A}, \X, and \Y are arrays. $p$ is what we call an MPC primitive. 
\n is the length of the arrays passed to parallel functions. This means that
ADD, SUB, MUL, CMP, XOR, SHL, SHR, REM, OR, AND, and MUX are functions on scalar values, 
while ADD_N, SUB_N, MUL_N, CMP_N, XOR_N, SHL_N, SHR_N, REM_N, OR_N, AND_N, and MUX_N are parallel functions on arrays.
Our analysis takes a high-level C-like program and outputs a (potentially vectorized) in this low-level target language.
%\ana{Change font for x and y, as in the other syntax figure.}\lindsey{fixed!}
%\ana{I'm having hard time remembering what we had in mind with this figure... This was the target MPC code, right?}
%\ana{If yes, then remove the for loop. There is no for loop in MPC code, we just unrol.}
%\ana{TODO, ANA: I think here we might have to introduce the parallel MUX, AND, etc.,} \lindsey{fixed!}
}
\label{fig:muximpsyntax}
\end{figure}

### 2021-02-06 chapterfour/overview.tex
added MPC loop material, from chapter four of Lindseys Dissertation

% \section{Overview and Background}
% \label{sec:overviewbackground}

% In this section, we consider background on Multi-Party Computation (MPC) and program analysis.

% \input{cryptoback}

% \input{progback}

% \subsection{Related Work}

\section{Overview of Our Work}
\label{sec:mpcoverview}

Our work aims to create a novel program analysis that improves the performance of loops within the 
context of MPC. We posit that we can leverage classical optimizations for loops from the 
High Performance Computing (HPC) literature. 
%\ana{I'm not sure what you mean with the above sentnences...}\lindsey{yes, that was not good
%is it clearer now?} 
However there are several caveats. 
Classically, HPC optimizations can perform in parallel any two operations as long as there are processors to perform them. In 
MPC this is not the case. Parallelization for MPC works similarly to operations in PHE- operations 
on secret values can only be performed in parallel with other operations of the same type. 
By using classical techniques a program loop can be segmented to take advantage of this, 
and the overall running time can be greatly reduced.

In summary, our analysis takes a loop, written using high-level C-like sytnax, and transforms it into 
into a straight-line vectorized MPC target code. We specify the input syntax and the MPC target syntax shortly.
Consider the example shown in \figref{fig:squarearray}: in this
code every member of an array is squared, this takes $n$ operations where $n$ is the size of the array $A$.
In MPC  this can be optimized as shown in \figref{fig:squarearraympc}, each operation can be put in a single
"thread" and can be performed in approximate constant 
time regardless of the number of operations being performed. This is only achievable on operations done
on secret values (in this case MPC shares of values). This is further expanded on in 
\secref{sec:hpcparallelization} and \secref{sec:mpcamortization}.

\begin{figure}[h]
\centering
\begin{minipage}{0.7\textwidth}
% \begin{figure}[h]
\begin{javacode}
int sq_arr(int A[]) {
    for(int i = 0; i < A.length; i++) {
        A[i] = A[i] * A[i];
    }
    return A;
}
\end{javacode}
% \end{figure}
\end{minipage}
\caption{Code to square each member of an array.}
\label{fig:squarearray}
\end{figure}

\begin{figure}[h]
\centering
\begin{minipage}{0.7\textwidth}
% \begin{figure}[h]
\begin{javacode}
int sq_arr(Share A[]) {
    B = MUL_N(A, A);
    return B;
}
\end{javacode}
% \end{figure}
\end{minipage}
\caption{Simple MPC transformation to square an array.}
\label{fig:squarearraympc}
\end{figure}

% \ana{Yes, I agree with the structure you have. Start with an Overview, then Crypto primitives, particularly MPC, then Program Analysis.}
% \ana{I agree with the general subsections you've outlined too.}

### 2021-02-06 chapterfour/parallelization.tex
added MPC loop material, from chapter four of Lindseys Dissertation

\section{Parallelization}
\label{sec:parallelization}

\subsection{Loop Dependencies} 
\label{sec:loopdependencies}

Array loop optimization is a crucial component of compiler design. More specifically
the optimizations that can be done on arrays \emph{within} a loop make better use of 
parallelization which results in decreased running times. The effectiveness of the 
optimization is entirely dependent on two factors: the number of processing threads
available, and the dependencies between the statements of the loop. These dependencies 
come in two types: Control-dependencies and data-dependencies. Control dependencies
exist between statements where the control flow of the program determines
program execution. Data-dependencies are dependencies that exist 
between nodes based on their interaction with shared data.

Aiken et al. \cite{Aiken1988} provides an illustration of data-dependencies. 
Consider the code in \figref{fig:arrayloop}. Line 2 is not dependent on any other 
line calculated in the same iteration. Line 5 is dependent on the calculation 
of lines 3, and 4. More interestingly lines 3, 4 are not only dependent on line 2, 
but also on a \emph{previous calculation} of line 5.

\begin{figure}[h]
\centering
\begin{minipage}{0.70\textwidth}
\begin{javacode}
	for(int i = 0; i < N; i++) {
		A[i] = B[i];
		B[i] = A[i] * D[i - 1];
		C[i] = A[i] * D[i - 1];
		D[i] = B[i] * C[i]; 
	}
\end{javacode}
\end{minipage}
\caption{Loop containing array dependencies, a modified example from Aiken et al. \cite{Aiken1988}.}
\label{fig:arrayloop}
\end{figure}

\begin{figure}[h]
\centering
\begin{minipage}{0.70\textwidth}
\begin{javacode}
  if(<Cond1>) {
      <Stmt1>
  } else if(<Cond2>) {
      <Stmt2>
  } else {
      <Stmt3>
  }
\end{javacode}
\end{minipage}
\caption{If-Then-Else dependency.}
\label{fig:ifthenelsedep}
\end{figure}

Every data-dependency either effects variables within the current loop iteration
(intra-loop dependency) or effects variables on a different loop iteration
(across-loop dependency). One type of simple across-loop dependency occurs when a single
variable is modified by a previous calculation of itself. This is a trivial case,
but still counts as an across-loop dependency. Data dependences are preserved in the 
transformation to target MPC code.

The most common type of control-dependency is the If-Then-Else dependency shown
in \figref{fig:ifthenelsedep}. The execution of each statement is dependent on the various conditionals 
present. In this example all three statements are dependent on the outer if conditional while only
statements 2 and 3 are dependent on inner conditional. Control dependences 
are ``inverted'' in the target MPC code. Instead of dependencies between the branch statements
to the conditional, there will be a dependence from statement 2 and 3 to the MUX 
node, as well as a dependecy between the MUX node and the conditional.

\subsection{HPC Parallelization}
\label{sec:hpcparallelization}

In the High performance computing (HPC) world a schedule is created to minimize the number
of sequential instructions of a program by utilizing a fixed number of threads. There 
are two main factors that determine the extent to which optimization can take place.
The first are control and data dependencies discussed above. The second is the number
of parallel processes that can be executed (i.e. the number of threads available). 
Importantly, each thread can perform arbitrary operations that are not reliant on the operations 
run on any other thread. The optimal schedule is then that which achieves the minimum number 
of sequential steps, while both preserving inter-loop and cross-loop dependencies 
and maximizing resource use. 
% In HPC optimizations \ana{Incomplete sentence?}.

Schedules are evaluated based off of their maximum \emph{delay}. The delay of a given 
schedule represents the number of steps between loop iterations that are 
carried out to preserve all looping dependencies \cite{Aiken1988}. 
For example, a variable modifying itself within a loop has a static distance of 1
iteration between the definition and usage. Creating a schedule with an optimal minimal delay
(steps between loop iterations) with a restricted number of processors has been shown to be
NP-hard \cite{Cytron1984},\cite{Graham1971},\cite{Darte2000}. 
Modern compilers often use heuristics to achieve a schedule
that is as close to optimal as possible. The classical HPC literature 
has considered different types of across-loop scheduling schemes. 
One scheduling scheme, called \emph{doacross}~\cite{Aiken1988},\cite{Cytron1986}, requires that all instructions in an 
iteration are scheduled on the same processor. 
The optimal doacross schedule for \figref{fig:arrayloop} (from \cite{Aiken1988}) is shown in \figref{fig:HPCexample} (b). 
Next, Aiken and Nicolau \cite{Aiken1988} proposed a new schedule, which they called
\emph{greedy} schedule, where instructions in an iteration can be scheduled on different processors. 
The optimal greedy schedule for our example, which achieves better minimal delay than the optimal 
doaccross schedule, is shown in \figref{fig:HPCexample}(c) (again, taken directly from~\cite{Aiken1988}).

\begin{figure*}[tbhp]
\small
\begin{tabular}{lll}

\begin{minipage}[b]{4.20cm}
\centering
\includegraphics[width=0.4\textwidth]{images/HPCloop.pdf}
\end{minipage}

&

\begin{minipage}[b]{4.20cm}
\centering
\includegraphics[width=0.4\textwidth]{images/HPCdoacrossLoop.pdf}
\end{minipage}

&

\begin{minipage}[b]{4.20cm}
\centering
\includegraphics[width=0.6\textwidth]{images/HPCgreedyLoop.pdf}
\end{minipage}

\\

(a) Loop-body schedule
&
(b) Doacross across-loop schedule
& 
(c) Greedy across-loop schedule

\end{tabular}
\caption{Optimal HPC schedules. Different statements are each given letters. 
In (a) only one loop iteration is shown. In (b) and (c) variables 
are labeled according to both their statement (i.e A, B, C) and the loop iteration that is being 
computed which is shown in parentheses (A(1) represents the statement A on iteration 1).}
\label{fig:HPCexample}\vspace{-2ex}
\end{figure*}

\subsection{MPC Amortization}
\label{sec:mpcamortization}

In MPC, the loop scheduling problem is different. Instead of having a fixed number of 
processors that carry out arbitrary independent operations, in MPC schedules the we assume that we can run
infinitely many operations (same operations) in parallel. We note that the relaxation of allowing infinite bandwidth 
is often made in the HPC setting as well and it is not necessarily valid in either case. 
Our current relies on this assumption however, in future work we plan to relax the 
assumption and extend our study and results.

%\lindsey{Ana, does something like this work? ``While there is still
%a practial limit to the number of operations that can be run in parallel it is not directly defined by the
%number of physical processors avaialable. We can also assume infinite processors in HPC, but this 
%fundamentally removes all constraits from the problem.'' This is to address wes's questions: We can do this in HPC waht is the diff}
%threads that can only run similar operations simultaneously.

For example, consider two separate operations A and B. We assume that both operations have 
the same constant running time. In classic HPC optimization both A and B can be run simultaneously with no 
restrictions as long as two processors are available to compute them. In MPC scheduling,
there is no restriction based on number of processors, but there is one on the types of
operations that are run in parallel. This means that there is only a benefit from amortization if A is run in parallel 
with another A operation, or B is run with another B operation. Our work aims to create a novel analysis 
that can be used to create schedules that improve the performance of MPC loops.

% All similar operations  (in this case equality, 
% comparison, and subtraction) can be performed at the same cost as a single operation of that type.

% \ana{Don't use the GCD example because it does not lend to parallelization. 
% Use the Aiken paper example, or the inner product example.}
% \figref{fig:gcdparallel}, shows a Java 8 examaple of how gcd() could be run in 
% parallel so each operation could benefit from amoritization.

\begin{figure}[h]
\centering
\begin{minipage}{0.7\textwidth}
\begin{javacode}
  InnerProduct(A, B) {
      sum = 0;
      for(int i = 0; i < N; i++) {
          sum += A[i] * B[i];
      }
      return sum;
  }
\end{javacode}
\end{minipage}
\caption{Original inner product code.}
\label{fig:innerproduct}
\end{figure}

\begin{figure}[h]
\centering
\begin{minipage}{0.7\textwidth}
\begin{javacode}
  InnerProduct(A, B) {
      sum = 0
      for(int i = 0; i < N; i++) {
          tmp = MUL(A[i], B[i]);
          sum = ADD(sum, tmp);
      }
      return sum;
  }
\end{javacode}
\end{minipage}
\caption{MPC source for inner product.}
\label{fig:innerproductmpc}
\end{figure}

\begin{figure}[h]
\centering
\begin{minipage}{0.7\textwidth}
\begin{javacode}
  InnerProduct(A, B) {
      C = MUL(A, B, N)
      sum = 0
      for(int i = 0; i < N; i++) {
          sum = ADD(sum, C[i]);
      }
  }
\end{javacode}
\end{minipage}
\caption{Vectorized MPC source for inner product.}
\label{fig:innerproductmpcvec}
\end{figure}

\begin{figure}[h]
\centering
\begin{minipage}{0.7\textwidth}
\begin{javacode}
  int[][]  split_list(int[] A) {
      A = pad_array(A); // adds 0 to an odd length array
      int[][] B = new int[2][A.length / 2];
      for(int i = 0; i < A.length / 2; i++) {
          B[0][i] = A[2 * i];
          B[1][i] = A[(2 * i) + 1];
      }
      return B;
  }

  private static int  SUM(int[] A) {
      int[][] W = split_list(A);
      while(A.length > 1) {
          A = ADD_N(W[0], W[1]);
          W = split_list(A);
      }
      return A[0];
  }
  InnerProduct(A, B) {
      C = MUL(A, B, N)
      return SUM(C)
  }
\end{javacode}
\end{minipage}
\caption{Vectorized MPC source for inner product using divide and conquer.}
\label{fig:innerproductmpcvecdandc}
\end{figure}

\begin{figure}[h]
\centering
\begin{minipage}{0.7\textwidth}
\includegraphics[width=1\textwidth]{divideconquer}
\end{minipage}
\caption{An example of \figref{fig:innerproductmpcvecdandc} for $n = 8$. 
Each level of tree is executed in parallel. This reduces the run time 
complexity from $O(n)$ to $O(log_{2}(n))$.}
\label{fig:treeexample}
\end{figure}

\figref{fig:innerproduct}, \figref{fig:innerproductmpc}, and \figref{fig:innerproductmpcvec} 
show a full transformation from high-level C-like source that computes the inner produce of 2 arrays {\sf A} and {\sf B}, 
to a fully vectorized MPC source that can be turned into an optimal MPC schedule. \figref{fig:treeexample}
shows a visualization of the execution performed for $n = 8$.
Initially in \figref{fig:innerproduct} and \figref{fig:innerproductmpc} $N$ 
multiplications and additions are performed. However in \figref{fig:innerproductmpcvec} 
the multiplications can be flattened to one operation, while the
additions are still performed sequentially with a \emph{for} loop. In  \figref{fig:innerproduct} the time
complexity of the algorithm is $O(n)$ where $n$ is the length of the array. Removing
the multiplication from the loop body does not reduce the asymptotic complexity, however,
in practice this has significant impact as in MPC multiplication is an expensive operation that
requires remote computation of the parties while addition is local and inexpensive. 
In addition, vectorization of multiplication allows for the addition operation to be isolated and 
additionally optimized, using divide-and-conquer techniques, which reduces the complexity to
$O(log_2(n))$ \cite{Farzan2017}. The transformation for inner product is shown in \figref{fig:innerproductmpcvecdandc}.

The great benefit of MPC amortizations is, due to the circuit representation, there is no inherent 
limit of the scale of parallelization. As we mentioned in \secref{sec:hpcparallelization} general
scheduling is a NP-Hard problem \cite{Cytron1984},\cite{Graham1971} however this is due to the restrictions
on the number of processors, with an unlimited number of processors the problem can be solved in polynomial time
\cite{Darte2000}. We claim that our problem is substantially different. In summary, the MPC amortization problem 
is different from the HPC parallelization problem because: in MPC we can assume unlimited bandwidth, and also 
we can only ``schedule in parallel'' two operations that are the same; in HPC we have a fixed number of processors, 
and we can schedule in parallel any two operations. Thus, we avoid some of the limitations that are present in HPC 
scheduling, while MPC imposes its own limitations and constraints.

This leads to our two main problems.

\begin{problem}
Given a loop body, find a schedule for the execution of statements in that body within a single iteration
of the loop taking into account intra-loop dependencies. This schedule should improve or not effect the
performance of the loop.
\label{problem:prob1}
\end{problem}

\begin{problem}
Given a loop, compute a schedule for the execution of the loop that takes into account intra-loop and across-loop dependencies.
\label{problem:prob2}
\end{problem}

Our work computes a schedule for a given a loop. A schedule is a sequence of MPC-primitives, as shown in~\figref{fig:muximpsyntax}.

%For the general case with arbitrary across-loop dependences, our vectorization may not be optimal.
%However, it is optimal under certain restrictions on across-loop dependences and certain assumptions about the loop body schedule. We elaborate 
%on these restrictions in \secref{sec:theoreticalguarentees}. Additionally in \secref{sec:results} we show that in practice, many of the 
%programs we saw from the HyCC benchmark did not violate these assumptions. %\ana{THIS MAY NEED TO CHANGE!!!} 
%\ana{Lindsey, check if this is true :) or we need to tone this down.} \lindsey{toned down!}

%\lindsey{not sure where to put this}
%\begin{itemize} 
%  \item Every dependency exists regardless of which iteration is being computed
%  \item Every dependency is remains unchanged regardless of which iteration is being computed
%\end{itemize}

### 2021-02-06 chapterfour/results.tex
added MPC loop material, from chapter four of Lindseys Dissertation

\section{Experimental Results}
\label{sec:results}

\input{table1.tex}

We compared 49 benchmarks, 23 of which are loops taken from
the test code in HyCC \cite{Buscher2018}, while the other 26 are part of a micro benchmark suite cases 
we developed for this purpose. \tabref{table:resultstablehycc} shows our results from the HyCC loops 
and \tabref{table:resultstablemicro} shows our results on our own benchmark suite. We found that
our assumptions held for all the test cases. 
The majority of our programs benefited from our analysis with $\nicefrac{40}{49}$ programs showing
a run time improvement.

\input{table2.tex}

As stated at the end of \secref{sec:loopscheduling} most of the loops we analyzed from HyCC were manipulations
of vectorized multi-dimensional arrays into output arrays or values. Since the multi-dimensional arrays
were represented in single dimensional format, indexes on these arrays were almost always very simple operations.
\figref{fig:MNIST2HyCC} shows a typical example of this; the arrays {\sf kernel} and {\sf image} are used to modify the 
loop output variable {\sf tmp}. The only across-loop dependency
is {\sf tmp} on itself which can be vectorized much like the inner-product example in 
\secref{sec:mpcamortization} using divide-and-conquer. In many cases we also found that 
as parallelization of independent operations increased, the cost most often 
decreased from $O(n)$ to either $O(log_{2}(n))$ or $O(1)$. The results
for this are  shown in \tabref{table:runtimeresults}. Not all problems showed changed;
For example, simple assignments incur no cost meaning that any loop body that consists of 
\emph{only} assignments shows no improvement. This is the case for DB_JOIN_TWO1 (shown in \figref{fig:DBJOINTWO1HyCC}).
In general our hand crafted benchmarks were significantly more difficult than those of HyCC. 
Consider Test6 (shown in \figref{fig:Example6}). The use of many different
conditional statements and complex dependencies lead to the high vectorized cost. 
Whereas in the HyCC benchmarks we found that most dependencies were quite simple.
An important note is that none of the the number of sequential opereations in our tests 
ever increased, they only decreased or remained costant. 
This means that existing techniques for loop optimizations only benefit from our work.

Any schedule of SCC components that consists of one SCC node of any complexity flowing into a fully 
parallelizable node (performing $n$ operations simultaneously) is optimal when executed linearly. We found this
to be the case in the majority of the HyCC loop bodies. All of the benchmarks in the HyCC suite were found to be optimal.
On the other hand any, as we stated in \secref{sec:theoreticalguarentees}, any SCC DAG that contains edges 
that connect a non constant rate node to another non-constant rate node node is not known to be optimal. 
Again, this is because our analysis does not provide a way to optimally schedule the statements within the 
individual iterations in parallel. Node pairs of this type have unique dependencies, and running them linearly may miss 
places where statements from one node can be run in parallel with statements from the other
node. We did find that the majority of our tests were indeed optimal, this is shown in
\tabref{table:optimalclassification}. This interleaving of statements between SCC nodes is most certainly an area 
for future work, and possible proofs for optimality. Despite a lack of a formal optimality result we saw
that almost all of our test cases showed a drastic improvement in run time complexity.

\input{table3.tex}

\begin{figure}[h]
\centering
\begin{minipage}{0.70\textwidth}
\begin{javacode}
...
for(wx = 0; wx < window_size; wx++) {
    int convPos = wx+wy*window_size;
    tmp += kernel[convPos] 
    	* image[(y*stride + wy) 
    	* image_width 
    	+ (x*stride + wx)];
}
...
\end{javacode}
\end{minipage}
\caption{Example taken from MNIST2 in HyCC. %\ana{This is O(logN), no? Just a readonly array and a sum.}
%\lindsey{yes! I reference it in the second paragraph of \secref{sec:results} to illustrate the fact that most loops are manipulating vectorized multi-dimensional arrays}
}
\label{fig:MNIST2HyCC}
\end{figure}

\begin{figure}[h]
\centering
\begin{minipage}{0.70\textwidth}
\begin{javacode}
...
for(int j = from; j < n; j++) {
    int tmp = m[from*n+j];
    m[from*n+j] = m[to*n+j];
    m[to*n+j] = tmp;
}
...
\end{javacode}
\end{minipage}
\caption{Example taken from DB_JOIN_TWO in HyCC. %\ana{This is just 0, no? It's local computation since it's just}
%\lindsey{referenced above to show that not all programs change since this is just a cost of 0}
}
\label{fig:DBJOINTWO1HyCC}
\end{figure}

\begin{figure}[h]
\centering
\begin{minipage}{0.70\textwidth}
\begin{javacode}
...
k = i;
z = z + i;
a[i] = i;
a[k] = i + a[i];
int j = i - 1;
if(a[i] > 3) {
    a[k] = p * 10;
    if(p > 20) {
        a[i] = p * 20;
    } else {
        a[k] = p * 30;
    }
    z = z + a[i];
} else {
    a[i] = p * 40;
}
q = q + a[i];
i++;
...
\end{javacode}
\end{minipage}
\caption{Test6 from our handcrafted benchmarks.}
\label{fig:Example6}
\end{figure}

### 2021-02-06 chapterfour/schedule.tex
added MPC loop material, from chapter four of Lindseys Dissertation

\subsection{Loop Scheduling}
\label{sec:loopscheduling}

After the creation of the graph $G$ is complete, the scheduling step takes
place, scheduling independent operations in the most parallel way possible. In the 
best case there are no conflicting dependencies. For example in the calculation 
of the inner-product of two vectors (shown below) every operation is independent, 
as there are no data-dependency cycles, this is shown in the trasformation from
\figref{fig:innerproduct} to \figref{fig:innerproductmpcvec} in \secref{sec:mpcamortization}.
This independence means that, given the proper resources, every multiplication 
can be done in one step and vectorized into a temporary array. The summation of 
this temporary array is more complex, but only requires $\log_2N$ steps (again 
assuming adequate resources) in a manner similar to that shown
in \figref{fig:innerproductmpcvecdandc}. 
More commonly, dependent statements, especially across-loop dependencies, 
limit the amount of parallelization that can be done. The maximum parallelization 
between any two dependent statements depends on the $d$ value between those statements
that is calculated during the creation of the $G$ graph.

The next phase is the actual scheduling algorithm. Recall that backward edges had
the negative weight $-1$ in all our examples, however, the weight on backward edges can be any value from $-1$ 
to $-(N-1)$. Forward edges have weight 0. Our algorithm proceeds in two phases: (1) strongly-connected component (SCC)
computation, and (2) scheduling. The first phase computes the strongly connected components in the graph $G$, intuitively,
each strongly connected component (SCC) consists of nodes that are interdependent and need to be scheduled accordingly. 
Once the SCC decomposition is done, we proceed to compute the schedule.

\subsubsection{SCC Decomposition}
\label{sec:SCC}

Consider a cycle $c \in G$. $d(c)$ is the absolute value of the sum of the weights of the edges in $c$. Consider a 
node $n \in G$. We say $n$ has \emph{constant rate} if all cycles $c$ through $n$ have the same $d(c)$. We write 
$d(n) = d(c)$. The $d(c)$ means that the computation at node $n$ at iteration $i$ depends on the computation 
at the same node $n$ at iteration $i-d(n)$. If $n$ is constant-rate with value $d(n)$, then for every $i$, $n(i)$
depends on the value computed at $n$ $d(n)$ iterations ago; but if the $d(c)$'s are not the same, then $n(i)$ depends on 
values computed at $n$ at different earlier iterations. This situation was illustrated earlier in~\figref{fig:fibconstantrate}.

We extend the definition of constant-rate to $G$. We say that 
$G$ is constant-rate if all $n \in G$ are constant-rate. If $n$ is a node, such that there are no cycles through $n$ in $G$, 
then $d(n) = N$, or in other words, $n$ can be fully vectorized. If a node $n$ is not constant rate, we assign a default $d(n) = 1$ 
for the purpose of amortization. For the rest of the text, we are interested in constant-rate $G$'s and consider  
results on constant-rate $G$'s. In practice, nodes have $d(n)$ values of either $N$ or $1$ even though other values of 
$d(n)$ are possible in theory. We found that of the 23 loops we analyzed from the HYCC \cite{Buscher2018} 
benchmark suite none violated the assumption. A typical example is shown in \figref{fig:dbjoin2nonconstant}.
%only 1 did not follow our constant-rate assumption, this is shown in \figref{fig:dbjoin2nonconstant}. 
%\ana{TODO Ana: What are the implications of non-constant rate.}
% \ana{TODO Lindsey: can you show the example that is not constant rate?}
%\ana{Also, can you show here some stats on values of $d$ in those benchmarks? We don't have any values of d other than 1 or N, correct?}
%\ana{HERE HAVE TO GIVE empirical evidence, how good/bad it is.} \lindsey{does that work?}
\begin{figure}[h]
\centering
\begin{minipage}{0.7\textwidth}
\begin{javacode}
...
for(int i = 0; i < len; i++) {
    sum[i] = db[i*att+1] + db[i*att+2];
}
...
\end{javacode}
\end{minipage}
\caption{DB_JOIN2 from the HyCC benchmarks. There are no across-loop dependencies and the constant-rate assumption trivially holds.} 
\label{fig:dbjoin2nonconstant}
\end{figure}

%We will consider Strongly Connected Components (SCCs) in $G$. In the running example in~\figref{fig:MPCexample}(b), 
%we have two SCCs, $SCC_1$ consists of node 2, and $SCC_2$ consists of nodes $3,4,5$. 
Without loss of generality, we may assume that each cycle has the form $n_1 \rightarrow n_2 \rightarrow ... \rightarrow n_k \stackrel{-d}{\dasharrow} n_1$, 
or in other words, there is a single backward edge in the entire cycle. Even though cycles with $2$ or more backward 
edges are possible, such cycles can be transformed into $n$ disjoint cycles of the above form. The $-d$ on the
single backward edge in each of the $n$ cycles, is the sum of the $-d$'s of all backward edges in the original cycle. 
%\ana{TODO ANA: We probably need more here.}

\subsubsection{Amortized Schedule}
\label{sec:schedule}

We now compute the amortized schedule $\mathcal{A}$.

%\lindsey{I am not sure what the figure fig:MPCexample is.... Ana, was that from the other paper? Or did I accedentally delete it or something?}

\begin{enumerate}

\item We begin by decomposing $G$ into Strongly Connected Components (SCCs). It is a theorem, that if $G$ is constant-rate, 
then for each $\mathit{SCC}$, and each pair of nodes $n_1, n_2 \in \mathit{SCC}, d(n_1) = d(n_2)$. We write $d(\mathit{SCC})$.
For the running example (code in~\figref{fig:ifthenelsedep} and graph $G$ shown in~\figref{fig:aikenscc}), we have two SCCs, 
$SCC_1$ consists of node 1, and $SCC_2$ consists of nodes $2,3,4$. $d(\mathit{SCC}_1) = N$ and $d(\mathit{SCC}_2) = 1$.

\item For each SCC, assume that there is a linear schedule $S: P_1 \rightarrow P_2 ... \rightarrow P_k$ given 
as input. $S$ takes into account only the forward edges in $G$: if $n_1, n_2 \in SCC$ and $n_1 \rightarrow n_2 \in G$, 
then $S(n_1)$ precedes $S(n_2)$ in the linear schedule. Note that computing an optimal schedule is NP-Hard, 
as we showed earlier, however, given a small SCC, one can brute force an optimal $S$. For the running example (graph $G$ shown in~\figref{fig:aikenscc}), 
we assume that nodes 2 and 3 perform the same operation. We work with schedule 
$S: \{2,3\} \rightarrow \{4\}$. It is important to note that with our treatment of arrays we have the same optimal schedule
for each iteration, as we have the exact same intra-loop dependences for each iteration.

%\ana{Also note, that due to our treatment of arrays we have the same schedule for each iteration of the loop, i.e., for each iteration, we either have a dependence edge or we don't.}

\item Given $S$ for a given SCC, we then compute the amortized schedule $\mathcal{A}({\mathit{SCC}})$; 
this is the schedule across all iterations of the loop for the given SCC. For each $P_i$ in $S$, there are $\lceil{\frac{N}{d(SCC)}}\rceil$
nodes in $\mathcal{A}_{\mathit{SCC}}$. $P_i(1),P_i(2),...P_i(d(SCC))$ go into node $P^1_i$, $P_i(d(SCC)+1),...P_i(2\cdot d(SCC))$ 
go into node $P^2_i$, etc. Recall that the parenthesized number denotes the iteration, e.g., $P_i(1)$ denotes $P_i$ at the first iteration 
of the loop. There are dependence edges as follows:
\begin{enumerate}
\item $P^j_i \rightarrow P^j_{i'}$ if there is an edge $P_i \rightarrow P_{i'}$ in $S$ (intra-loop edges).
\item $P^{\lceil{\frac{N}{d(SCC)}}\rceil}_i \rightarrow P^1_{i+1}$ for each $1 \le i < \lceil{\frac{N}{d(SCC)}}\rceil$ (across-loop edges).
\end{enumerate}
In the running example, $SCC_1$'s schedule is just the single vectorized node that executes all A's at once.
$SCC_2$'s schedule repeats $S: \{2,3\} \rightarrow \{4\}$ $N$ times: we have 
$\{2,3\}(1) \rightarrow \{4\}(1) \rightarrow \{2,3\}(2) \rightarrow \{4\}(2) ... \{2,3\}(N) \rightarrow \{4\}(N)$, 
as each iteration depends on the previous-iteration value computed at the same node.
This is illustrated graphically in~\figref{fig:unrolledloopdep}.   %\ana{Lindsey, can you draw this graph? It's basically just repeating A $->$ B,C $->$ B,C $->$ B,C etc.}

\item The decomposition of $G$ into SCCs gives rise to a DAG where the nodes in the DAG are the individual SCCs. 
If there is an intra-loop or an across-loop dependence edge from a node in $\mathit{SCC}_1$ to a node in $\mathit{SCC}_2$, 
then there is an edge from $\mathit{SCC}_1$ to $\mathit{SCC}_2$. The edges arise from intra-loop and from across-loop dependences, 
but note that across-loop dependences are rare, as they are typically accounted for within SCCs. We call this graph the SCC DAG.
The user gives a linearization of this DAG: $SCC_1 \rightarrow SCC_2 \rightarrow ... \rightarrow SCC_m$. 
The final schedule is as follows: $\mathcal{A} = \mathcal{A}({\mathit{SCC}_1}) \rightarrow \mathcal{A}({\mathit{SCC}_2}) ... \rightarrow \mathcal{A}({\mathit{SCC}_m})$.
In our running example, there is only one possible ordering of the two components, 
and the final schedule is as depicted in~\figref{fig:unrolledloopschedule}. %\ana{Lindsey, add the reference here to that figure you draw.}

%dependence edges $P(n)_1 \dasharrow P(n)_2 ... \dasharrow P(n)_{\lceil{\frac{N}{d(n)}}\rceil}$, reflecting the across-loop dependence 
%edges we saw earlier. The rest of the edges for every edge $n_1 \rightarrow n_2 \in G$ and every $1 \le i \le N$, add an edge 
%$G'(n_1(i)) \rightarrow G'(n_2(i))$ to $G'$ where $G'(n(i))$ returns the $P$ node in $G'$ that contains $n(i)$.

\end{enumerate}

The test cases taken from the HyCC benchmark suite involved the manipulation of multi-dimentional arrays into output structures. 
This meant that almost all array accesses were independent - and those
that were dependent had simple constant rate dependencies. In addition most did not modify their own structure but were dependent 
on other arrays. None of the cases taken from the HyCC benchmark suite violated our assumptions 
this is further expanded on in \secref{sec:results}.

%We assign a parallelization factor of $N$ to $n$, meaning that 
%$n$ can be fully vectorized. A more interesting case is a non-singleton $SCC$. Then there is at least one cycle 
%from $n$ to itself. We consider \emph{cycles} $c: n \rightsquigarrow n \in SCC$
%such that there is no repeated node in $c$, and sum up the edge weights along the edges in $c$. If all 
%cycles including $n$ have the same value $-d$, then we assign a parallelization factor of $d$ to $n$. That 
%means that $n$ depends on the value computed at $n$ exactly $d$ iterations earlier, and it does not depend 
%on node $n$ of iterations in-between. In contrast, if there are two distinct cycles with two different weights, 
%then we assign the default value of 1 to $n$. Intuitively, two distinct weights on cycles, say $-d_1$ and $-d_2$ 
%means that $n$ depends on the value computed at $n$ $d_1$ iterations ago and on the value computed 
%$d_2$ iterations ago. Taking all nodes in the SCC, if all nodes have the same parallelization factor $d$, 
%then we do nothing, otherwise we assign to all nodes factor -1. \ana{More intuition here, why is this important etc}

%\ana{Have to make some kind of claim that certain \% of bench falls under this assumptions or otherwise 
%our results don't count for much.} \lindsey{added}

\subsection{On the Correctness and Optimality of Schedules $\mathcal{A}$}
\label{sec:theoreticalguarentees}

In \secref{sec:mpcamortization} we stated our two main problems \problemref{problem:prob1}, to compute an optimal loop body schedule, 
and \problemref{problem:prob2}, to compute an optimal across-loop schedule. We note that the notion of optimality is restricted by the assumptions in
our model, outlined in \secref{sec:loopbodyanalysis}. Recall that we assume unlimited bandwidth as well as identical cost of execution, e.g.,
executing 1 MPC primitive costs 1, and executing 1000 primitives at once costs 1 as well. In our model, an optimal schedule is the shortest 
sequence of MPC primitives, some possibly vectorized, such that the sequence observes \emph{all} inter-loop and across-loop dependencies.

\subsubsection{On Correctness of Schedules $\mathcal{A}$}

Problem 2 can be formulated on an unrolled loop as follows. 
Let $L$ be a loop body within an MPC program. %\ana{This needs to be revised. Take into account the MUX.} Due to the restrictions on MPC outlined in
As stated in \secref{sec:mpc} the loop bounds of $L$ are both \emph{static} and \emph{known at compile time}. 
$L$ can be unrolled while preserving program semantics. Each iteration is placed in line with
 the previous, and all variables of a specific iteration are renamed to be associated
only with that specific iteration. Dependencies between variables of different iterations can now be rewritten 
as dependencies between these newly renamed variables. Say, $L$ contains an 
across-loop dependency between variables $x$ and $y$ where $d$ is the number of iterations between 
the computation of $x$ and the computation of $y$. After unrolling the loop; the value of $x$ for 
iteration $i$ can be computed: $\forall i \in L, x_{i} = y_{i - d}$. This also holds for array 
accesses. 
%This creates a body of code that only executes once, the analysis of which
%is described in \problemref{problem:prob1}. 
Consider our running example in \figref{fig:arrayloop}. 
If we bound $i$ to 2 iterations we can unroll the loop and determine the dependencies, this is shown
in \figref{fig:flattenedloop}. This unrolling presents the \emph{sequential schedule} of the loop which we denote by  $\mathcal{S}$.
Importantly, the unrolling is done on the SSA-form of the loop body after scalar substitution. There are no MUX nodes
in our example, but in general there would be.
This sequential schedule $\mathcal{S}$ is the base line that we use to argue correctness
for the schedule our algorithm computes (after making use of the well-formedness property of arrays).

%\lindsey{Ana, do you think that is too jarring?} \ana{Needs some work... So how are we distinguishing this.}\ana{This is great!}

\begin{figure}[h]
\centering
\begin{minipage}{0.7\textwidth}
\begin{javacode}
A1_1[1] = B0[1];              // A1 = update(A0,1,B0)
B1_1[1] = A1_1[1] * D0[0];    // B1 = update(B0,1,A1*D0)
C1_1[1] = A1_1[1] * D0[0];    // C1 = update(C0,1,A1*D0)
D1_1[1] = B1_1[1] * C1_1[1];  // D1 = update(D0,1,B1*C1)

A1_2[2] = B1_1[2];            // A1 = update(A1,2,B1)
B1_2[2] = A1_2[2] * D1_1[1]); // B1 = update(B1,2,A1*D1)
C1_2[2] = A1_2[2] * D1_1[1]); // C1 = update(C1,2,A1*D1)
D1_2[2] = B1_1[2] * C1_2[2]); // D1 = update(D1,2,B1*C1)
\end{javacode}
\end{minipage}
\caption{Unrolled loop of \figref{fig:arrayloop} with a static bound on the loop variable $1 \le i \le 2$. 
This is the sequential schedule $S$ for the example. After unrolling the across-loop def-use dependency 
between {\sf D[1] = ...} and {\sf B[2] = A[2] * D[1];} becomes explicit. We simplify the SSA of array variables 
slightly in order to avoid notational clutter.} 
%\ana{Lindsey, I think you can draw the dependence graph too? One part is the code and the other part is the graph.}
%\lindsey{Ana, did you not like this example?}\ana{I do!}
\label{fig:flattenedloop}
\end{figure}

Unfortunately, as we have stated, finding an optimal loop-body amortized schedule is NP-hard. 
Therefore, finding an optimal across-loop amortized schedule is NP-hard, just as is finding an optimal loop-body schedule.
However, we posit that loop bodies are typically small and therefore, it would be feasible to compute the shortest common 
supersequence using a brute force approach. It would not be feasible of course to use a brute force approach to compute
the shortest common supersequence for the sequential schedule.
Therefore, in our work we are interested whether certain regularity of across-loop dependencies and the availability of
an optimal loop-body schedule can lead to efficient computation of an optimal across-loop schedule.
%Our key premise is that one can an across-loop amortized schedule that always either improves
%or does not effect real-world loops the run time of real-world loops. Given certain properties of the loop, the amortized 
%schedule we compute in the previous section is always correct and can also be optimal depending on the
%nature of the SCC graph.

The first theorem below establishes correctness, i.e., that $\mathcal{A}$ computes the same result as $\mathcal{S}$ where
$\mathcal{S}$ is the standard sequential schedule of the loop. The correctness result takes into account
the well-formedness conditions and parallelization using well-formed arrays.

\begin{theorem}(Correctness)
Let $\mathcal{A}$ be the amortized schedule (computed as in~\secref{sec:loopscheduling}) and let $\mathcal{S}$ be the standard sequential 
schedule of the loop. $\mathcal{A}$ preserves dependences, i.e., for every def-use chain $(d,u) \in \mathcal{S}$, we have that
$\mathcal{A}(d)$ precedes $\mathcal{A}(u)$ in $\mathcal{A}$ and there is no redefinition of $\mathit{LHS}(d)$ in any 
of the nodes from $\mathcal{A}(d)$ to $\mathcal{A}(u)$. (Here $\mathcal{A}(n)$ gives the node in the schedule that contains $n$).
\end{theorem}

\begin{proof} (Sketch)
The proof is by case analysis. Case 1 considers when the variable in the def-use 
chain is a scalar, and Case 2 considers when it is an array location. 
The key is to prove that if there is a def-use dependence in $\mathcal{S}$, then the same def-use dependence
is represented in $G$.
Case 1 is straightforward, as dependencies between scalar variables are handled in the standard way. 
For Case 2, there are again two cases, the first case is when the array is reduced to standard SSA, in which 
case case 2 reduces to case 1, and the second case when the array is well-formed under our well-formedness
conditions.

Suppose we have an array write $d: \A[f(i)] = ...$ and an array read $u: ... = \A[f'(i)]$ that form a def-use relation, i.e., 
the location that is written at the array write is used at the array read. If the def and the use happen in the same 
iteration, say $k$, that means that $f(k) = f'(k)$ and therefore, $f(i) \cong f'(i)$. If there are other writes to the same
location, they would happen earlier in the iteration, and the writes $d': \A[f(i)]$ have been numbered accordingly so
that the latest one is $d$ and the only def-use link into $u$ is from $d$. Now suppose that $d$ happens in iteration $i$
and $u$ happens in iteration $i+d$. That means $f(i-d) = f'(i)$ and therefore $f(i-d) \cong f'(i)$ in the analysis. Again, 
without loss of generality we may assume that $d$ is the last write to that location in iteration $i$ and our analysis 
will add an appropriate dependence edge. There will be no other def-use link into $u$ because that would
violate the disjointness assumption.

Since our schedule respects def-use links, it follows that $d$ is scheduled ahead of $u$. Since there is a unique def-use
link, it follows that there is no write into that location that may happen between the $d$ and the $u$. 
%\ana{TODO: Ana, more here!}
\end{proof}

Correctness of execution of $\mathcal{A}$ is a simple corollary of the theorem. This follows by induction. Assume that all statements that 
execute before a given statement $s$ in $\mathcal{A}$ compute the same result in $\mathcal{A}$ as in $\mathcal{S}$. We will consider 
the execution of $s$ and show that $s$ in $\mathcal{A}$ computes the same result as $s$ in $\mathcal{S}$. This is obvious from the 
theorem: the computation of the operands of $s$ precedes $s$ in $\mathcal{A}$ and by the inductive hypothesis, they compute the
same value as in $\mathcal{S}$; therefore, $s$ computes the same result. %\ana{TODO: Ana, more here.}

\subsubsection{On Optimality}

Unfortunately, optimality is more difficult to establish. Again, we stress that the notion of optimality is limited by our model. 
In the general case, computing an optimal across loop schedule will be NP-hard,
using the same argument as for the optimal loop-body schedule. We study different common schedules, and we claim that if the code
obeys our assumptions (well-formedness of arrays and constant-rate across-loop dependencies) typical schedules that arise in practice
will be optimal. We first prove that each SCC schedule, as computed by our analysis, is optimal. Then we consider a class of commonly occurring schedules
and we prove that the schedule $\mathcal{A}$ computed by our analysis is optimal. The classes of loops that we observed that exhibit optimal schedule
have a MapReduce-like structure --- there are one ore more map-like computations followed by a reduce-like computation. An example is Inner Product,
where point-wise multiplications can be vectorized, producing an array of products, and computing the sum amounts to a reduce of that array around 
the plus operation. In future work we plan to explore additional loops and build a larger and more comprehensive categorization of the classes
of loops that occur in practice.

%\lindsey{WES: call ut classes for which we can demonstrate optimality,
%Something like: We found that optimal scheudes occured mostly for simple SCCs....? not sure.} 
%Finally, we consider some preliminary results and conjectures for schedules of different shape.

\begin{lemma}(Optimality of SCC Schedules) Assuming an optimal schedule of the loop-body component of the SCC, well-formed arrays, 
and constant-rate $G$, then the $\mathcal{A}(SCC)$ computed in~\secref{sec:loopscheduling} is optimal.
\end{lemma}

\begin{proof} (Sketch)
Recall that the $\mathcal{A}(SCC)$ schedule has cost $n \cdot \lceil{\frac{N}{d(\mathit{SCC})}}\rceil$, where $n$ is the height (i.e., length) of the optimal 
schedule of the loop-body component. Since we have a strongly connected component and constant-rate across-loop dependencies, 
then for each pair of statements $s_1, s_2 \in SCC$ 
we have that $s_1$ in iteration $i+d$ depends on $s_2$ in iteration $i$, and vice versa,  $s_2$ in iteration $i+d$ depends on $s_1$ in iteration $i$.
Let $s_1$ be a statement in the first node in $\mathcal{A}(SCC)$, and let $s_2$ be a statement in the last node in $\mathcal{A}(SCC)$. 
Therefore we have
\[s_1 (1) \rightarrow ... \rightarrow s_2 (1) \dasharrow s_1(d+1) \rightarrow ... \rightarrow s_2 (d+1) \dasharrow ...\]
where the length of each solid arrow chain is optimal and of length $n$, and the length of the dashed arrow chain is $\lceil{\frac{N}{d(\mathit{SCC})}}\rceil$.
\end{proof}

%\ana{Need a sketch of proof}

The most common schedules involve a MapReduce-like pattern. All our running examples follow this pattern. The intra-loop dependence graph has the form $s_1 \rightarrow s_2 \rightarrow ... \rightarrow \mathit{SCC}_k \rightarrow s_{k+1} \rightarrow $. Here $s_1, s_2, ... s_{k-1}, s_{k+1},...$ give rise to special SCCs that can be fully vectorized. $s_1, s_2, ... s_{k-1}, s_{k+1}, ... $ can be fully vectorized, while $\mathit{SCC}_k$ can be scheduled optimally according to the $d(\mathit{SCC}_k)$. Intuitively, since all $s_1, s_2, ...s_{k-1}$ must execute before $\mathit{SCC}_k$, then ${k-1}$ is a lower bound on their execution. An additional constraint is that none of $s_1, s_2$, etc., can be executed in parallel with some of the iterations of $\mathit{SCC}_k$, in which case ${k-1} + \mathit{cost}(\mathit{SCC}_k)$ is a lower bound on the execution of $s_1 \rightarrow s_2 \rightarrow ... \rightarrow \mathit{SCC}_k$. 
This is formalized by the following theorem.

\begin{theorem}(Optimality of Common Schedules) Let the SCC DAG be a straight-line DAG: $\mathit{SCC}_1 \rightarrow \mathit{SCC}_2 \rightarrow ... \rightarrow \mathit{SCC}_k \rightarrow
... \rightarrow \mathit{SCC}_{n-1} \rightarrow \mathit{SCC}_n$ where $d(\mathit{SCC}_k) \le N$, while all other $d(\mathit{SCC})=N$. Let all dependence edges in the SCC DAG be triggered by intra-loop dependence edges, let there be intra-loop dependences from $s_{k-1}$ to every $s_k \in \mathit{SCC}_k$ and let there be intra-loop dependences from every $s_k \in \mathit{SCC}_k$ to $s_{k+1}$. The amortized schedule computed in~\secref{sec:loopscheduling} is optimal. %\ana{Still not done.}
\end{theorem}

\begin{proof} (Sketch)
The proof follows directly from the lemma. From the lemma, the schedule of $\mathcal{A}(\mathit{SCC}_k)$ is optimal. Since SCCs $1, 2, ...,k-1$ must be scheduled ahead of $\mathit{SCC}_k$
and none of them can be scheduled in parallel with an iteration of $\mathcal{A}(\mathit{SCC}_k)$. This is because there are intra-loop dependences from every $s_1, s_2, ... s_{k-1}$ to every $s_k$ in the $\mathit{SCC}_k$. Similarly, $s_{k+1},...$ must be scheduled after $\mathit{SCC}_k$ due to the corresponding intra-loop dependences. Therefore, the schedule we compute is an optimal schedule. 
\end{proof}

Even though we capture most common loop structures, and compute optimal schedules for those loop structures, we would like to relax our assumptions and produce optimality results for a larger class of loop structures. 
%Clearly the problem is NP-hard, however, structuring the loop into SCCs naturally partitions the loop into smaller components. We conjecture that computing optimal schedules for each SCC can be feasible due to the small size of the SCCs.

%The principal problem when we have a SCC DAG where more than one of the SCCs are not fully vectorized, i.e., have we have some $d(\mathit{SCC}) < N$, is that different iterations can be scheduled in parallel, even in the restricted case when there are intra-loop dependences from every node in one SCC to every node in the other SCC. We consider the following problem and conjecture. 
%Consider two strongly connected components, each with $d=1$ and the following SCC DAG: $\mathit{SCC}_1 \rightarrow \mathit{SCC}_2$. 
%We compute a schedule as follows. Let the intra-loop graph of $\mathit{SCC}_1$ be $A$, and the intra-loop graph of the $\mathit{SCC}_2$ be $B$. 
%We compute the Shortest Common Supersequence of $A$ and $B$ denoted by $S = \mathit{SCS}(A,B)$. Let us partition $S$ into $A'S'B'$
%where $A'$ is the longest common prefix of $A$ and $S$ such that $S'B'$ is a supersequence of $B$.

%Let $A'$ be the prefix and $B'$ be the suffix of $S$, 
%such that $A'$ is the longest common prefix of $S$ and of $A$, and $B'$ is the longest common suffix of $S$ and $B$. 
%If $A'$ or $B'$ is empty, then we compute the schedule by chaining $S$ $N$ times. Otherwise, we compute the Shortest Common Supersequence of $A'$ and $B'$
%and so on, appropriately chaining these supersequences to form a schedule. We conjecture that this algorithm yields an optimal schedule. \ana{Show this in the case when A' or B' is empty.}
%\ana{Needs work.}
%and consider A'' and B'' as before, and so on. If we chain the SCSs we get an optimal schedule.

%\ana{State conjecture. 2 SCCs with d = 1. We can compute an optimal schedule as follows. Compute the SCS of A and B. Let A' be the prefix and B' be the suffix, 
%such that A' is a common prefix of the SCS and of A, and B' is the longest common suffix of the SCS and of B. If A' or B' is empty, we are done, otherwise compute the SCS of A' and B'
%and consider A'' and B'' as before, and so on. If we chain the SCSs we get an optimal schedule.}

%\ana{TODO: Here opine on the differences between the vectorization problem and the problem in MPC.}

%\ana{Correctness theorem should come here...}

%\paragraph{Creating The SCC Schedule}
%\label{sec:sccschedulecreation}
%To create the amoritized schedule (the execution of SCC within G) we use the following steps. 
%Let $G$ be a constant-rate graph that entails a DAG $D$, where the nodes of $D$ are the strongly 
%connected components of $G$. If $D$ is linear, and each $SCC \in D$ entails a loop-body schedule 
%$S: P_1 \rightarrow P_2 ... P_k$ with a cost of $d$. The schedule is created by running each 
%$SCC \in D$ sequentially. Futhermore if $\nexists SCC_i \rightarrow SCC_j$, s.t. $d_i \neq d_j \land d_j \neq 1$
%then the schedule is optimal. If the above is \emph{not} true then an optimal scheule of the two 
%SCC nodes is not known (and possibly NP-Hard to discover) as it depends on which
%specific instructions within each SCC can be scheduled together. One can also use Aiken's result to compute an 
%optimal greedy schedule using $n^2$ iterations, where $n$ is the number of nodes 
%in the loop-body schedule \cite{Aiken1988}. \ana{Add citations.} \lindsey{is this just aiken? or does someone else also say this}

%\begin{theorem} (Optimal Conditions)
%Let $G$ be a constant-rate graph that entails a DAG $D$, where the nodes of $D$ are the strongly connected components of $G$. 
%If $D$ is linear, and each $SCC \in D$ entails a loop-body schedule $S: P_1 \rightarrow P_2 ... P_k$, s.t. (1) $S$ is optimal 
%and (2) for every $P_i, i<k$, for every $n_1 \in P_i, \exists n_2 \in P_{i+1}$, s.t. $n_1 \rightarrow n_2 \in G$, (3)
%$\nexists SCC_i \rightarrow SCC_j$, s.t. $d_i \neq d_j \land d_j \neq 1$ then $\mathcal{A}$ is optimal.
%\end{theorem}

%\begin{proof} (Sketch)
%This SCC schedule is optimal because 1) dependences are the same across the loop iterations, and 2) there is dependence between 
%iteration $i$ and iteration $i+1$, no node in $i+1$ can be computed before the last node, i.e., $P_k(i)$ has been computed, as all 
%nodes in $i+1$ depend on $P_k(i)$. Furthermore each SCC dependency $SCC_i \rightarrow SCC_j$ respects 
%$d_i = d_j \lor d_j = 1$ meaning there exists no schedule that will outperform linear execution. \lindsey{this last part seems wrong....}
%\end{proof}

\begin{figure}[H]
\centering
\begin{minipage}{0.50\textwidth}
\includegraphics[width=\textwidth]{aikensccdag} 
\end{minipage}
\caption{SCC DAG for \figref{fig:arrayloop} formed from \figref{fig:aikenscc}.}
\label{fig:aikensccdag} 
\end{figure}

\subsubsection{Example} Consider our running example shown in \figref{fig:arrayloop}. There are three SCC nodes for this program. 
The first consists of line 1, while the second of lines 2 and 3, and the third line 4. This creates 
a DAG as shown in \figref{fig:aikensccdag}. The SCC node containing line 1 would be executed first 
with a $d$ value of 0 (assignments incur no cost). Next the node lines 2 and 3 are executed- both multiplications 
are performed at once incurring a cost and therefore $d$ of 1. The last node is also executed for a cost 
(and again a $d$ value) of 1. This is an optimal schedule for this exeuction as all of the 
$d$ values involved are only 1.

\subsection{Open Problems}

So far, we have identified a way of reasoning about loops as well as a way of computing schedules that turn out to be optimal 
for many loops in practice. Our work leaves a lot of open problems and future directions of research.

\paragraph{Optimality results} First, we believe that our results about optimality can be extended. A principal problem is, how do we schedule SCCs in the DAG that 
have $d$s different than $N$? The problem when we have a SCC DAG where more than one of the SCCs are not fully vectorized, i.e., we have some $d(\mathit{SCC}) < N$, is that different iterations can be scheduled in parallel, even in the restricted case when there are intra-loop dependences from every node in one SCC to every node in the other SCC.
We believe that we can compute optimal schedules per SCC and then combine those schedules to arrive at an optimal amortized schedule for the entire loop body.
%Even in the case when we have SCC1 connected to SCC2 via intra-loop edges, these two 
%SCCs can be scheduled in parallel as we can interleave the iteratio
%\ana{1. Extended proofs of optimality}

%\ana{2. Extension of results to nested loops}
\paragraph{Nested loops} Second, a further extension will be made for nested loops. The process begins at the innermost loop 
computing a vectorization, then inlining the vectorization within the enclosing loop.
Then the entire process can begin again.
%If there are no changes the variables that are in the outer loop dependecies  
%scheduling the nested loop this is trivial, as all input variables can be simply treated as constants. 
%However if any change does occur they must be reflected upon re-entering
%the outer loop, which may change the entire depedency. 
We leave this for future work.  %\lindsey{something like that, we should talk about it}.

%\ana{3. Divide and Conquer Parallelization}

\paragraph{Divide-and-conquer parallelization}
\label{sec:divideconquer}
% \lindsey{optomize certain operations that have a d value that does not equal 0. Do a divide and conquer}
Once the $d$ values are determined there are in general a lot of SCCs with $d=1$. 
%and all operations with $d = 0$ are fully amortized the next step is 
%to vectorized the remaining operations. 
Following the dependences as we did so far yields that full vectorization is not possible. For example, 
the sum in the Inner Product example will be $O(N)$. However, a divide and conquer 
algorithm can still reduce the overall running time when the aggregation operation has certain properties. 
For example consider (c) in \figref{fig:innerproduct}.
It is easy to see that the $d = 1$ in the addition on line 6, \texttt{sum} is dependent on its own previous value. 
If computed in the way shown, obeying all intra-loop and across-loop dependences, it will take $N$ steps, where $N = \vert C \vert$. 
This can be greatly reduced using a divide and conquer method, which is possible because the addition is an associative
operation. Consider the code in \figref{fig:innerproductmpcvecdandc}.
this shows how the number of operations in inner product can be reduce to $1 + \log_2 N$ steps. %More formally 
%this is shown in \lemref{lem:vecotrizationlemma}. 
By using this technique dependent operations can be computed
faster reducing the overall run time of the vectorized program. We plan to explore this direction in future work.
The principal problem is proving associativity of the loop computation.

%\begin{lemma}
%Given an vector $V$ and an associative operation $O$ with $K$ operands all
%operations can be completed in $O(log_K \vert V \vert)$ time.
%\label{lem:vecotrizationlemma}
%\end{lemma}

% 1. Add an edge from the conditional node, Cond, to the corresponding MUX node. This is just a forward edge.

% 2. Constructing the ``forward'' portion of the graph. If V is a scalar, add a forward edge $n_1 \rightarrow n_2$ such that for some $V_k$,
% $V_k \in LHS(n_1)$ and $V_k \in RHS(n_2)$. For arrays, add a forward edge $n_1 \rightarrow n_2$ such that there is $A_k[f(i)] \in LHS(n_1)$ 
% and $A_k[f'(i)] \in RHS(n_2)$ and $f(i)$ and $f'(i)$ are "equivalent", i.e., the same for all i. (Usually these equivalent accesses are just
% A[i]).

% 3. Constructing the "backward" portion of the graph. For each scalar $V$, let $n_1$ be the node such that $V_\mathit{max} \in LHS(n_1)$,
% i.e., this is the node that defines the $V$ exposed at the exit of the loop. Add an edge $n_1 \stackrel{-1}{\dasharrow} n_2$
% for each $n_2$ such that $V_0 \in RHS(n_2)$. In other words, $n_2$ has an upward exposed use of $V$, i.e., one that comes
% from. For each array write, let $n_1$ be the node such that $A[f(i)]_\mathit{max} \in LHS(n_1)$. Add an edge
% $n_1 \stackrel{-d}{\dasharrow} n_2$ for each $n_2$ such that $A_0[f'(i)] \in RHS(n_2)$ and $f(i-d) \cong f'(i)$.

### 2021-02-08 chapterfour/introduction.tex
Made chapterfour.tex standalone.

@@
 is Partially Homomorphic Encryption (PHE), which scales substantially better, but cannot 
-perform arbitrary operations, and requires the conversions that were discussed in \chapref{chapter:chapterthree}.
+perform arbitrary operations, and requires the conversions. %that were discussed in \chapref{chapter:chapterthree}.
 

### 2021-02-08 chapterfour/overview.tex
Made chapterfour.tex standalone.

@@
 % \begin{figure}[h]
-\begin{javacode}
-int sq_arr(int A[]) {
+\begin{code}
+int sq\_arr(int A[]) {
     for(int i = 0; i < A.length; i++) {
@@
 }
-\end{javacode}
+\end{code}
 % \end{figure}
@@
 % \begin{figure}[h]
-\begin{javacode}
-int sq_arr(Share A[]) {
-    B = MUL_N(A, A);
+\begin{code}
+int sq\_arr(Share A[]) {
+    B = MUL\_N(A, A);
     return B;
 }
-\end{javacode}
+\end{code}
 % \end{figure}

### 2021-11-20 compiler/compiler/ast_shared.py
Fix mypy errors

@@ -3,7 +3,7 @@ from typing import Generic, TypeVar, Union
 from dataclasses import dataclass
 from textwrap import indent
 
-import networkx
+import networkx  # type: ignore
 
 
 @dataclass(frozen=True)

### 2021-11-20 compiler/compiler/ast_to_restricted_ast.py

@@ -374,6 +374,7 @@ class _StatementConverter(_StrictNodeVisitor):
     def visit_Expr(self, node: ast.Expr) -> Optional[restricted_ast.Statement]:
         # Ignore docstrings
         _DocstringExpector(self.source_code_info).visit(node.value)
+        return None
 
 
 class _ReturnValueGetter(_StrictNodeVisitor):

### 2021-11-20 compiler/compiler/restricted_ast_to_tac_cfg.py

@@ -4,7 +4,7 @@ a three-address code control flow graph
 """
 
 
-import networkx
+import networkx  # type: ignore
 
 from . import restricted_ast
 from . import tac_cfg
@@ -178,11 +178,12 @@ def _build_expression_as_update_value(
 
 def _build_assignment(assignment: restricted_ast.Assign, builder: _CFGBuilder):
     lhs = assignment.lhs
-    rhs = _build_expression_as_update_value(assignment.rhs, builder)
+    rhs_update_value = _build_expression_as_update_value(assignment.rhs, builder)
+    rhs: tac_cfg.AssignRHS
     if isinstance(lhs, restricted_ast.Var):
-        pass
+        rhs = rhs_update_value
     elif isinstance(lhs, restricted_ast.Subscript):
-        rhs = tac_cfg.Update(array=lhs.array, index=lhs.index, value=rhs)
+        rhs = tac_cfg.Update(array=lhs.array, index=lhs.index, value=rhs_update_value)
         lhs = lhs.array
     else:
         assert_never(lhs)

### 2021-11-20 compiler/compiler/ssa.py

@@ -1,7 +1,8 @@
 """Data types representing code in static single assignment form"""
 
-import networkx
 from dataclasses import dataclass
+
+import networkx  # type: ignore
 
 from .ast_shared import (
     Var,

### 2021-11-20 compiler/compiler/ssa_phi_to_mux.py

@@ -2,8 +2,8 @@ from . import ssa
 
 
 def replace_phi_with_mux(result: ssa.Function) -> None:
+    block: ssa.Block
     for block in result.body.nodes:
-        block: ssa.Block
         if block.merge_condition is not None:
             mux_calls = []
             for phi in block.phi_functions:

### 2021-11-20 compiler/compiler/tac_cfg_to_ssa.py

@@ -7,7 +7,7 @@ from collections import Counter
 import itertools
 from typing import Optional, Union
 
-import networkx
+import networkx  # type: ignore
 
 from . import tac_cfg
 from . import ssa
@@ -20,8 +20,8 @@ def _compute_blocks_setting_vars(
 
     result: dict[ssa.Var, set[ssa.Block]] = dict()
 
+    block: ssa.Block
     for block in function.body.nodes:
-        block: ssa.Block = block
         for assignment in block.assignments:
             var = assignment.lhs
 
@@ -44,8 +44,8 @@ def _tac_cfg_to_ssa_struct(tac_cfg_function: tac_cfg.Function) -> ssa.Function:
     cfg = networkx.DiGraph()
     mapping: dict[tac_cfg.Block, ssa.Block] = dict()
 
+    tac_cfg_block: tac_cfg.Block
     for tac_cfg_block in tac_cfg_function.body.nodes:
-        tac_cfg_block: tac_cfg.Block = tac_cfg_block
         ssa_block = ssa.Block(
             phi_functions=[],
             assignments=tac_cfg_block.assignments,
@@ -55,11 +55,10 @@ def _tac_cfg_to_ssa_struct(tac_cfg_function: tac_cfg.Function) -> ssa.Function:
         cfg.add_node(ssa_block)
         mapping[tac_cfg_block] = ssa_block
 
+    source_block: tac_cfg.Block
+    dest_block: tac_cfg.Block
+    label: ssa.BranchKind
     for source_block, dest_block, label in tac_cfg_function.body.edges(data="label"):
-        source_block: tac_cfg.Block = source_block
-        dest_block: tac_cfg.Block = dest_block
-        label: tac_cfg.BranchKind = label
-        label: ssa.BranchKind = label
         cfg.add_edge(
             u_of_edge=mapping[source_block],
             v_of_edge=mapping[dest_block],
@@ -95,8 +94,8 @@ def place_phi_functions(result: ssa.Function) -> None:
 
         while W != set():
             X = W.pop()
+            Y: ssa.Block
             for Y in dominance_frontiers[X]:
-                Y: ssa.Block = Y
                 if has_already[Y] < iter_count:
                     num_predecessors = sum(1 for _ in result.body.predecessors(Y))
                     Y.phi_functions.append(ssa.Phi(lhs=V, rhs=[V] * num_predecessors))
@@ -235,7 +234,16 @@ def rename_variables(result: ssa.Function) -> None:
[…]

### 2021-11-21 paper_SIMD.tex
Basic vectorization, cleaner version.

--- 2019-11-20 paper.tex
@@
 % *** SPECIALIZED LIST PACKAGES ***
+%\usepackage{algorithm2e}
 \usepackage{algorithmic}
@@
 % used to denote shared contribution to the research.
-\author{Muhammad Ishaq}
-\authornote{This work was done in part while the author was at RPI.}
-\email{m.ishaq@ed.ac.uk}
-%\orcid{1234-5678-9012}
-%\author{G.K.M. Tobin}
-%\authornotemark[1]
-%\email{webmaster@marysville-ohio.com}
+
+\author{Benjamin Levy}
+\email{levyb3@rpi.edu}
 \affiliation{%
-    \institution{University of Edinburgh}
-    %\streetaddress{P.O. Box 1212}
-    \city{Edinburgh}
-    \state{Scotland}
-%    \postcode{43017-6221}
+    \institution{Rensselaer Polytechnic Institute}
+    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
+    \city{Troy}
+    \state{New York}
+    %\country{Iceland}
+}
+
+\author{Benjamin Sherman}
+\email{shermb@rpi.edu}
+\affiliation{%
+    \institution{Rensselaer Polytechnic Institute}
+    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
+    \city{Troy}
+    \state{New York}
+    %\country{Iceland}
 }
@@
 
+\author{Muhammad Ishaq}
+\authornote{This work was done in part while the author was at RPI.}
+\email{m.ishaq@ed.ac.uk}
+%\orcid{1234-5678-9012}
+%\author{G.K.M. Tobin}
+%\authornotemark[1]
+%\email{webmaster@marysville-ohio.com}
+\affiliation{%
+    \institution{University of Edinburgh}
+    %\streetaddress{P.O. Box 1212}
+    \city{Edinburgh}
+    \state{Scotland}
+%    \postcode{43017-6221}
+}
+
 \author{Vassilis Zikas}
@@
 
-In this short paper we define the problem of \emph{optimal parallelization} for Multi-party Computation (MPC), 
-set it apart from classical work on parallelization for High-performance Computing (HPC), present preliminary 
-results, and set directions for future research.
+\section{Background and Problem Statement}
+\label{sec:background}
 
-\ana{1. Intro to MPC (from a PL point of view). Define the problem.}
+\subsection{Scheduling in MPC}
 
-\ana{2. Difference with HPC. Why "optimal HPC" = "minimal delay" is NOT "optimal MPC"}
-
-\ana{3. Results statement.}
-
-\ana{4. Future directions.}
-
-% no \IEEEPARstart
-
-\section{Scheduling in MPC}
+\ana{Here we need more on MPC before jumping to scheduling. MPC-source, basic assumptions about static loop bounds, etc.}
 
@@
 
-\paragraph{Problem 1:} Given a dependence graph of MPC instructions, find the \emph{loop-body} schedule 
-with \emph{minimal cost}.
+\subsection{Scheduling in HPC}
+\label{sec:hpc}
 
-%\begin{figure*}[tbhp]
-\begin{figure*}[tbhp]
-\begin{tabular}{llll}
-\begin{minipage}[b]{4.25cm}
+\section{\bf Preliminaries}
+\label{sec:preliminaries}
 
-\begin{minted}[fontsize=\footnotesize, linenos, numbersep=5pt, escapeinside=||]{java}
-for (int i=1; i<=N; i++) {
-  A: A[i] = f_1(B[i]);
-  B: B[i] = f_2(A[i],D[i-1]);
-  C: C[i] = f_3(A[i],D[i-1]);
-  D: D[i] = f_4(B[i],C[i]);
-}
-\end{minted}
-\end{minipage}
+We assume arbitrarily nested loops in the MPC-source IR and read-only arrays.
+We assume that loops range from 0 to some constant N.
+Arrays are linearized (row-major order as in MOTION) and accesses are via functions of the induction variables of the enclosing loops.
+We write $i_1,i_2,...i_k$ to denote the loop nest: $i_1$ is the outermost loop, $i_2$, is immediately nested in $i_1$, and so on until $i_k$.
+As an example, a statement nested in $i_1,i_2,...i_k$ can access array $A[f(i_1,i_2,...i_k)]$.
+We write $A[i_1,i_2,...i_k]$ interchangeably.
 
-&
+The dependence graph computes def-use edges:
 
-\begin{minipage}[b]{4.25cm}
-\includegraphics[width=0.5\textwidth]{figs/dependenceGraph.pdf}
-\end{minipage}
+\begin{itemize}
+\item same-level $X\rightarrow Y$ where $X = ...$ and $Y = ...$ are in the same loop nest $i_1,i_2,...i_k$.
+E.g., $\texttt{d} = \mathit{SUM}(\texttt{S}[i,j], \texttt{C}[j])$ to $\texttt{p} = \mathit{MUL}(\texttt{d},\texttt{d})$ in Biometric is a same-level edge.
+\item outer-to-inner $X\rightarrow Y$ where $X$ is in an outer loop nest, $i_1,i_2,...i_j$, and $Y$ is in an inner one, $i_1,i_2,...i_j,...,i_k$.
+\item inner-to-outer $X\rightarrow Y$ where $X$ is in an inner loop nest, $i_1,i_2,...i_j,...,i_k$, and $Y$ is in an outer one, $i_1,i_2,...i_j$.
+E.g.
+$\texttt{sum}_1 = \texttt{ADD}(\texttt{sum}_0,\texttt{p})$ to
+$\texttt{c} = \mathit{CMP}(\texttt{sum}_1,\texttt{min}_0)$ is an inner-to-outer edge.
+\item same-level back-edge $X\rightarrow Y$.
+$Y$ is a phi-node in the header of the loop and $X$ is a definition of the variable in the loop body.
+E.g., $\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{sum}_1,\texttt{min}_1)$ to $\texttt{min}_0 = \phi(\texttt{min}_1, 10000)$ in Biometric is a same-level back-edge.
+\item outer-to-inner back-edge $X\rightarrow Y$: $X$ and $Y$ are both $\phi$-nodes for some variable. The source $X$ is in an immediately nested loop of $Y$'s loop. 
+ \end{itemize}
 
-&
+We also define the $\mathit{closure}(n)$ where $n$ is a phi-node. Intuitively, it computes the set of statements that form a cycle in the loop with $n$. The closure of $n$ is defined as follows:
+\begin{itemize}
+\item $n$ is in $\mathit{closure}(n)$
+\item $X$ is in $\mathit{closure}(n)$ if there is a same-level path from $n$ to $X$, and $X \rightarrow n$ is a same-level back-edge. 
+\item $Y$ is in $\mathit{closure}(n)$ if there is a same-level path from $n$ to $Y$ and there is a same-level path from $Y$ to some $X$ in $\mathit{closure}(n)$.
+\end{itemize}
 
-\begin{minipage}[b]{4.25cm}
-\includegraphics[height=0.15\textheight]{figs/MPCloop.pdf}
-\end{minipage}
+\section{Basic Vectorization}
 
-&
+\ana{Add back-edges into Phase 1.
+A back-edge from a non-phi-node in loop $i$ to a phi-node in loop $i$'s header is a same-level edge.
+The only difference with the handling of normal def-use is that the operand index will become $i-1$.
+A back edge from a phi-node in loop $j$ to a phi-node in loop $i$, where $j$ is immediately nested in $i$ is an inner-to-outer edge and will require dropping dimension.
+We can show these are the only kinds of back-edges that may occur.}
 
-\begin{minipage}[b]{4.25cm}
-\includegraphics[height=0.20\textheight]{figs/MPCaccrossLoop.pdf}
-\end{minipage}
-\\
+%\subsection{Algorithm (Still work in progress.)}
 
-(a) Source 
-&
-(b) Dependence graph
-&
-(c) Loop-body schedule
-&
-(d) Across-loop schedule
+\begin{algorithmic}
 
-\end{tabular}
-\caption{(a) shows the source code, taken from Aiken and Nicolau~\cite{Aiken:1988}, 
-(b) shows the dependence graph. The solid edges denote within-iteration ordering constraints, e.g., A must run before B
-because B depends on A. The dashed edges denote across-iteration ordering constraints, e.g., B(2), i.e., B in iteration 2
-depends on D(1), i.e., D in iteration 1. (c) shows the optimal loop-body schedule, and (d) shows the
-optimal across-loop schedule. A(1), B(1), etc. denote instruction A in iteration 1, etc.}
-\label{fig:MPCexample}\vspace{-2ex}
-\end{figure*}
+\STATE \COMMENT {Phase 1: Raise dimension of scalar variables to corresponding loop nest. We may assume linear traversal of the MPC-source. \ana{Double check!}}
 
-\paragraph{Problem 2:} Given a dependence graph of MPC instructions representing a loop body (or given a loop-body schedule), 
-find the \emph{across-loop} schedule with  \emph{minimal cost}.
+\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i_1,...,i_k$}
+\FOR {each $Y_i$}
+\STATE {case def-use edge $Y_i \rightarrow X$ of} \\
+\STATE \hspace{0.25cm} {same-level $\rightarrow$ add $Y_i = Y_i$} \\
+\STATE \hspace{0.25cm} {outer-to-inner $\rightarrow$ add $Y'_i[i_1,...,i_k] = \mathit{raise\_dim}(Y_i)$} 
+\STATE \hspace{0.25cm} {inner-to-outer $\rightarrow$ add $Y'_i[i_1,...,i_k] = \mathit{drop\_dim}(Y_i)$} 
+\ENDFOR
 
-\figref{fig:MPCexample} illustrates the problem with a code example from the classical high-performance computing (HPC) literature. 
-\figref{fig:MPCexample}(c) shows the optimal loop-body schedule (answer to Problem 1). Instructions B and C can be scheduled in parallel, 
-however, the schedule does not necessarily benefit form parallelization/amortization. If B and C (functions $f_1$ and $f_2$) are the \emph{same}
-MPC instruction (e.g., both are MUX, MUL, ADD, or CMP, etc.), then they \emph{can} benefit from parallelization/amortization. Given
-our assumptions stated above, the cost of running the loop-body schedule is 3. However, if B and C are different MPC-instructions,
-e.g., B is MUL, but C is MUX, then the cost of running the schedule is 4.
+\COMMENT {Optimistically vectorize all. $\stackrel{\rightarrow}{i}$ means vectorized dimension.}
+\STATE {change to $X[\stackrel{\rightarrow}{i_1},...\stackrel{\rightarrow}{i_k}] = \mathit{Op}(Y'_1[i_1,...,i_k],Y'_2[i_1,...,i_k])$ }
+\ENDFOR
 
-\figref{fig:MPCexample}(d) shows the optimal across-loop schedule (answer to Problem 2). We can schedule all A-instructions, across all iterations, 
-in parallel, as A depends only on the initial value of B. We cannot do better but schedule B, C, and D sequentially, as they each depend on values
-computed in the previous iteration. Again, if B and C are the same instruction, the optimal across-loop schedule will have cost of $1 + 2N$, 
-and if B and C are NOT the same instruction, then it will have cost $1+3N$.
+\COMMENT {Phase 2: Recreating FOR loops for cycles; vectorizable statements hoisted up.}
 
-\section{Scheduling in HPC}
+\FOR {each dimension $d$ from highest to 0}
+\FOR {each $\phi$-node $n$ in loop $i_1,...,i_k$}
+\STATE {compute $\mathit{closure}(n)$}
+\WHILE {there are closure $cl_1$ and $cl_2$ that intersect}
+\STATE {merge $cl_1$ and $cl_2$}
+\ENDWHILE 
+\FOR {each closure $cl$ (after merge)}
+\STATE {create FOR $i_k = 0; ...$ loop}  
+\STATE {add $\phi$-nodes in $cl$ to header block} 
+\STATE {add statements in $cl$ to loop body in some order of dependences} 
+\STATE \COMMENT {Dimension is not vectorizable: }
+\STATE {change $\stackrel{\rightarrow}{i_k}$ to $i_k$ in all statements in loop}
+\STATE {treat FOR loop as monolith node: change def-use edges accordingly.}
+\ENDFOR
+\ENDFOR
+\ENDFOR
 
-In this section, we state the classical HPC scheduling problem, and argue the MPC scheduling problem poses unique challenges.
+\COMMENT {Phase 3:}
+\STATE {add SIMD for simdified dimensions}
 
-Similar problems arise in the HPC setting.
+\end{algorithmic}
 
-%\begin{figure*}[tbhp]
-\begin{figure*}[tbhp]
-\begin{tabular}{lll}
+\subsection{Example: Biometric}
 
-\begin{minipage}[b]{4.25cm}
-\includegraphics[width=0.4\textwidth]{figs/HPCloop.pdf}
-\end{minipage}
+\subsubsection{MPC-source}
 
-&
+Below is the MPC-source resulting from Benjamin's analysis. It gives rise to a corresponding dependence graph. (Not shown.)
 
-\begin{minipage}[b]{4.25cm}
-\includegraphics[width=0.4\textwidth]{figs/HPCdoacrossLoop.pdf}
-\end{minipage}
+\begin{algorithmic}
 
-&
+\STATE \COMMENT {Begin of outer loop $i$.}
+\STATE {$\texttt{index}_0 = \phi(\texttt{index}_1, -1)$} 
+\STATE {$\texttt{min}_0 = \phi(\texttt{min}_1, 10000)$}
 
-\begin{minipage}[b]{4.25cm}
-\includegraphics[width=0.6\textwidth]{figs/HPCgreedyLoop.pdf}
-\end{minipage}
+\STATE \COMMENT {Begin of inner loop $j$.}
+\STATE {$\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$}
+\STATE {$\texttt{d} = \mathit{SUM}(\texttt{S}[i,j], \texttt{C}[j])$}
+\STATE {$\texttt{p} = \mathit{MUL}(\texttt{d},\texttt{d})$}
+\STATE {$\texttt{sum}_1 = \texttt{ADD}(\texttt{sum}_0,\texttt{p})$}
+\STATE \COMMENT {End of inner loop $j$.}
+\STATE {$\texttt{c} = \mathit{CMP}(\texttt{sum}_1,\texttt{min}_0)$}
+\STATE {$\texttt{index}_1 = \mathit{MUX}(\texttt{c},\texttt{index}_1,i)$}
+\STATE {$\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{sum}_1,\texttt{min}_1)$}
+\STATE \COMMENT {End of outer loop $i$.}
 
-\\
+\end{algorithmic}
 
-(a) Loop-body schedule
-&
-(b) Doacross across-loop schedule
-& 
-(c) Greedy across-loop schedule
+\subsubsection{Phase 1 of Vectorization algorithm:}
 
-\end{tabular}
-\caption{Optimal HPC schedules.}
-\label{fig:HPCexample}\vspace{-2ex}
-\end{figure*}
+The transformation preserves the dependence edges.
+It raises the dimensions of scalars and optimistically vectorizes all operations.
+The last phase discovers loop-carried dependences and removes affected vectorization.
 
-\paragraph{Problem 1:} Given a loop body (i.e., dependence graph), find the loop-body schedule with \emph{minimal delay}.
+\begin{algorithmic}
 
-\paragraph{Problem 2:} Given a loop body (or a loop-body schedule), find the \emph{doacross} schedule with minimal delay.
+\STATE \COMMENT {Begin of outer loop $i$.}
+\STATE {$\texttt{index}_0[i] = \phi(\texttt{index}_1[i-1], -1)$} 
+\STATE {$\texttt{min}_0[i] = \phi(\texttt{min}_1[i-1], 10000)$}
 
-\paragraph{Problem 3:} Given a loop body (or a loop-body schedule), find the \emph{greedy schedule} with minimal delay.
+\STATE \COMMENT {Begin of inner loop $j$.}
+\STATE {$\texttt{sum}_0[i,j] = \phi(\texttt{sum}_1[i,j-1], [0,0,...])$}
+\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
+\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUM}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
+\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
+\STATE {$\texttt{sum}_1[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
+\STATE \COMMENT {End of inner loop $j$.}
+\STATE {$\texttt{sum}'_1[i] = \mathit{drop\_dim}(\texttt{sum}_1[i,j],j)$}
+\STATE {$\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$}
+\STATE {$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_1[\stackrel{\rightarrow}{i}],[0,1,...N])$}
+\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_1[\stackrel{\rightarrow}{i}])$}
+\STATE \COMMENT {End of outer loop $i$.}
 
-\figref{fig:HPCexample} illustrates the problem in HPC. \figref{fig:HPCexample}(a) shows the optimal loop-body schedule
-which results in cost of 3. Again, we assume that all instructions have the same cost, a unit of 1. The key difference with the 
-MPC schedule is that regardless of whether B and C are the same instruction opcode or not, B and C can be scheduled 
-to run in parallel and the overall running time benefits from parallelization.
+\end{algorithmic}
 
-The HPC literature has considered different kinds of across-loop scheduling. One scheduling, called \emph{doacross}~\cite{aiken,cytron},
-requires that all instructions in an iteration are scheduled on the same processor. The optimal doacross schedule for our example
-is shown in \figref{fig:HPCexample}(b) (taken directly from~\cite{aiken}). Aiken and Nicolau proposed a new schedule,
-which they called \emph{greedy} schedule, where instructions in an iteration can be scheduled on different processors. 
-The optimal greedy schedule for our example, which achieves better minimal delay than the optimal doaccross schedule, 
-is shown in \figref{fig:HPCexample}(c) (again, taken directly from~\cite{aiken}).
+\subsubsection{Phase 2}
 
-\ana{Here we have to make the argument that MPC is different from HPC.}
-The example illustrates that optimal MPC scheduling and optimal HPC scheduling are different problems.
+This phase analyzes statements from the innermost loop to the outermost. The key point is to discover loop-carried dependencies and re-introduce loops whenever dependencies make this necessary.
 
-%\section{\bf Preliminaries}
-%\label{sec:preliminaries}
+Starting at the inner phi-node $\texttt{sum}_0[i,j] = \phi(\texttt{sum}_1[i,j], [0,0,...])$, the algorithm first computes it's closure.
+The closure amounts to the phi-node itself and $\texttt{sum}_1[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$, accounting for the loop-carried dependency of the computation of \texttt{sum}.
+The algorithm replaces this closure with a FOR loop on $j$ removing vectoriaztion on $j$.
+Note that the SUB and MUL computations remain outside of the loop as they do not depend on phi-nodes that are part of cycles.
+The algorithm adds same-level edges, one from $\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = ...$ to the monolithic FOR-loop node, and one from the FOR-loop node to $\texttt{c}[\stackrel{\rightarrow}{i}] = ...$.
 
-\ana{Relevant papers/books:
+\begin{algorithmic}
 
-\begin{enumerate}
+\STATE \COMMENT {Begin of outer loop $i$.}
+\STATE {$\texttt{index}_0[i] = \phi(\texttt{index}_1[i-1], -1)$} 
+\STATE {$\texttt{min}_0[i] = \phi(\texttt{min}_1[i-1], 10000)$}
 
-\item Book: "Scheduling and automatic parallelization", Springer 2000, by Alain Darte, Yves Robert and Frederic Vivien. 
-This book describes scheduling and parallelization in HPC: given a dependence graph and unlimited 
-number of processors, one can compute an optimal (minimal delay) schedule in polynomial time. 
-Given a finite number of processors, computing an optimal schedule becomes NP-hard.
+\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
+\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUM}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
+\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
 
-\item https://theory.stanford.edu/~aiken/publications/papers/pldi88.pdf. "Optimal loop parallelization", PLDI 1988, by
-Alex Aiken and Alex Nicolau. 
-Given a dependence graph of the loop body, computes an optimal parallelization across loop iterations. 
-The optimality metric is minimal delay. E.g., assuming 2 iterations, the delay is measured by the number 
-of instructions in iteration 2 that are lagging behind iteration 1.
+\STATE \COMMENT {Begin of inner loop $j$. Loop is now much "shorter"!}
+\STATE {$\texttt{sum}_0[i,j] = \phi(\texttt{sum}_1[i,j], [0,0,...])$}
+\STATE {FOR j=0; j<=D; j++} \COMMENT{ Will turn into a MOTION loop.}
+\STATE \hspace{0.25cm}{$\texttt{sum}_1[\stackrel{\rightarrow}{i},{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},{j}],\texttt{p}[\stackrel{\rightarrow}{i},{j}])$}
+\STATE \COMMENT {End of inner loop $j$.}
 
-\item https://www.cs.indiana.edu/~achauhan/Teaching/B629/2006-Fall/CourseMaterial/1998-popl-knobe-arrayssa.pdf.
-"Array SSA", POPL 1998, by Kathleen Knobe and Vivek Sarkar. This paper defines an SSA form for arrays which I think 
-is crucial for parallelization across loop iterations.
+\STATE {$\texttt{sum}'_1[i] = \mathit{drop\_dim}(\texttt{sum}_1[i,j],j)$}
+\STATE {$\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$}
+\STATE {$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_1[\stackrel{\rightarrow}{i}],[0,1,...N])$}
+\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_1[\stackrel{\rightarrow}{i}])$}
+\STATE \COMMENT {End of outer loop $i$.}
 
-\end{enumerate}
-}
+\end{algorithmic}
 
-\section{Loop Body Parallelization}
-\label{sec:parallelization_within_loop}
-%\input{../sections/analysis.tex}
+Next, analysis moves to outer dimension $i$. There are two phi-nodes, $\texttt{index}_0[i] = \phi(\texttt{index}_1[i-1], -1)$ and $\texttt{min}_0[i] = \phi(\texttt{min}_1[i-1], 10000)$.
+The closure of the first is 
+\[
+\begin{array}{l}
+\texttt{index}_0[i] = \phi(\texttt{index}_1[i-1], -1), \\ 
+\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
+\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_1[\stackrel{\rightarrow}{i}],[0,1,...N]) 
+\end{array}
+\]
+and the closure of the second one is
+\[
+\begin{array}{l}
+\texttt{min}_0[i] = \phi(\texttt{min}_1[i-1], 10000) \\
+\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
+\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_1[\stackrel{\rightarrow}{i}])
+\end{array}
+\]
+Since the two closures intersect, we merge them into a single loop.
+We can now rewrite those statements, getting rid of the vectorization on $i$ in second FOR loop.
+Notably, the SUB and MUL computations are fully vectorizable and the ADD computation is vectorizable across the $i$-dimension.
 
-We first consider the answer to Problem 1 (in MPC). 
-We argue that optimal parallelization within a loop body is NP-complete (under the above stated assumptions). 
-Therefore, compilers must resort to heuristics to compute a schedule for the instructions within a loop body.
+\begin{algorithmic}
 
-We consider two operations, call them $A$ and $M$. Each instruction in the program is either an $A$-instruction or
-an $M$-instruction. In order to benefit from parallelization/amortization, we must schedule two or more 
-$A$-instructions in the same parallel node (or two or more $M$-instructions in the same parallel node). 
-Scheduling $A$-instructions in parallel with $M$-instruction does not benefit from amortization.
-It incurs the exact same cost as scheduling the $A$-instructions in a node $P_A$, scheduling the $M$-instructions 
-in a node $P_M$, and having $P_A$ precede $P_M$ in the parallel schedule. This is the difference 
-between classical scheduling, as studied in parallel computing, and MPC scheduling.
+\STATE 
+\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
+\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUM\_SIMD}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
+\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL\_SIMD}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$} \\
+\STATE {} 
+\STATE {$\texttt{sum}_0[i,j] = \phi(\texttt{sum}_1[i,j-1], [0,0,...])$}
+\STATE {FOR j=0; j<=D; j++} \COMMENT{ Turns into a MOTION loop, was $j$-loop.}
+\STATE \hspace{0.25cm}{$\texttt{sum}_1[\stackrel{\rightarrow}{i},{j}] = \mathit{ADD\_SIMD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},{j}],\texttt{p}[\stackrel{\rightarrow}{i},{j}])$}
+\STATE {} 
+\STATE {$\texttt{sum}'_1[i] = \mathit{drop\_dim}(\texttt{sum}_1[i,j],j)$}
+\STATE {} 
+\STATE {$\texttt{index}_0[i] = \phi(\texttt{index}_1[i-1], -1)$} 
+\STATE {$\texttt{min}_0[i] = \phi(\texttt{min}_1[i-1], 10000)$}
+\STATE {FOR i=0; i<=N; i++} \COMMENT{ Turns into a MOTION loop, was $i$-loop.}
+\STATE \hspace{0.25cm}{$\texttt{c}[{i}] = \mathit{CMP}(\texttt{sum}'_1[{i}],\texttt{min}_0[{i}])$}
+\STATE \hspace{0.25cm}{$\texttt{index}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{index}_1[{i}],[0,1,...N])$}
+\STATE \hspace{0.25cm}{$\texttt{min}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{sum}'_1[{i}],\texttt{min}_1[{i}])$}
 
-We make the following assumptions:
+\end{algorithmic}
 
-\begin{enumerate}
+\subsection{Correctness Argument}
 
-\item $A$ and $M$ are of equal cost, 1 unit.
-\item There is unlimited bandwidth---i.e., a single $A$-instrution (or $M$-instruction) costs as much as $N$ amortized $A$-instructions 
-(or $M$-instructions), namely 1 unit.
+\subsection{Extensions of Basic Vectorization}
 
-\end{enumerate}
+\ana{TODO: Will extend with read/write arrays and removal of (certain) infeasible edges.}
 
-Consider a loop body that consists of $n$ sequences: $S_1$, ... $S_n$ of $A$ and $M$ instructions. 
-More precisely, the loop body is such that its instructions can be grouped into such sequences. 
-$S_1$, ... $S_n$ can execute in parallel, however, all instructions within a sequence must 
-execute sequentially. For example, consider the three sequences (the right arrow indicates a \emph{dependence},
-meaning that the source node must execute before the target node): 
-\begin{enumerate}
-\item $A \rightarrow M \rightarrow A$
-\item $A \rightarrow A \rightarrow A$
-\item $M \rightarrow A \rightarrow M$
-\end{enumerate}
+\ana{Conjecture: Basic Vectorization still works with minor extensions and correctness argument still hods. }
 
-A \emph{schedule} $P: P_1 \rightarrow P_2 \dots \rightarrow P_k$ is such that for each sequence 
-$S_i$ in the set, if $S_i[k]$ precedes $S_i[k']$ in $S_i$ then $S_i[k]$ is scheduled in node $P_l$, $S_i[k]$ 
-is scheduled in node $P_{l'}$, and $P_l$ precedes $P_{l'}$ in $P$.
+\section{Divide-and-Conquer}
 
-The cost of a schedule $P$ is 
-\[\mathit{cost}(P) = \sum_{i=1}^k \mathit{cost}(P_i)\]
-where $\mathit{cost}(P_i) = 1$ if $P_i$ consists of $A$-instructions only, or $M$-instructions only, 
-and $\mathit{cost}(P_i) = 2$ if $P_i$ mixes $A$-instructions and $M$-instructions.
+\ana{TODO: Now that we have broken FOR loops into smaller chunks, we can add Divide-and-conquer reasoning with Z3 and implement this additional transform.}
 
-The problem is to find a schedule $P$ with \emph{minimal cost}. For example, 
-a schedule with minimal cost for the sequences above is
-\[ A(1), A(2) \rightarrow M(1), A(2), M(3) \rightarrow A(1), A(2), A(3) \rightarrow M(3) \]
-(The parentheses above indicate the sequence where the instruction comes from: (1), (2), or (3).)
-The cost of this schedule is 5.
-
-The problem of finding a schedule with minimal cost 
-easily reduces to the problem of finding a \emph{shortest common supersequence}, 
-a known NP-complete problem \ana{citation needed}. To see the reduction, 
-suppose $P$ is a schedule with minimal cost (computed by a black-box algorithm). 
-We can derive a schedule $P'$ with the same cost as $P$, by mapping each mixed node $P_i \in P$ 
-to two consecutive nodes in $P'$: an $A$-instruction node followed by an $M$-instruction node.
-Clearly, $P'$, which now is a sequence of $A$'s and $M$'s, is a supersequence of each sequence 
-$S_i$, i.e., $P'$ is a common supersequence 
-of $S_1 \dots S_n$. It is also a shortest common supersequence. (To see this, suppose, a 
-shorter common supersequence, $P''$, exists. $P''$ is a schedule of $S_1 \dots S_n$
-and the $\mathit{cost}(P'')$ equals the length of $P''$. Since $P'$ is longer than
-$P''$, and $\mathit{cost}(P') = \mathit{cost}(P)$, that means $\mathit{cost}(P'') < \mathit{cost}(P)$, 
-which is a contradiction since $P$ is a schedule with minimal cost.)
-
-\ana{Question: are these two assumptions too strong?}
-
-\ana{Assumption 1 can be relaxed --- e.g., if $A$ is 5 times cheaper than $M$, 
-we can construct a program where all $A$'s are sequences of five consecutive 
-$A$'s, which reduces to the previous problem.}
-
-\ana{Assumption 2, I am not sure if this is acceptable, or maybe it is too strong. 
-My intuition is that incorporating into the cost a dependence on the number 
-of instructions in each parallel node, is going to make the problem harder, not easier...}
-
-\section{Optimal Loop Parallelization} 
-\label{sec:optimal_loop_parallelization}
-%\input{../sections/optimal_loop_parallelization.tex}
-
-\ana{Aiken's and Array SSA to define the notion of optimal loop parallelization for MPC.}
-
-We now consider the answer to Problem 2 (in MPC). 
-\ana{Make a note that the problem is easy assuming NO array Writes. We consider array writes.}
-
-\subsection{Overview and Problem Statement}
-
-The problem is, given a loop-body schedule, compute the across-loop schedule
-with minimal cost. The following examples illustrate. The first example is the inner 
-product computation, and the second is Aiken and Nicolau's example.
-
-\begin{tabular}{lll}
-\begin{lstlisting}[language=Java]
-A = ...
-B = ...
-ip = 0;
-for (i=1; i<=N; i++) {
-  t=A[i]*B[i];
-  ip=ip+t;
-}
-
-\end{lstlisting}
-& 
-~
-&
-\begin{lstlisting}[language=Java]
-A = ...
-B = ...
-ip = 0;
-for (i=1; i<=N; i++) {
-  if (i%d(5)==1) para(t[i],...,t[i+d(5)])
-  if (i%d(6)==1) ip = ip + t[i];
-}  
-\end{lstlisting}
-
-\end{tabular}
-
-\ana{More to introduce what the schedule is.}
-
-In the above example {\sf d(5)} is N, meaning that all N multiplications are executed in parallel.
-In contrast, {\sf d(6)} is 1, meaning that the summation statements are executed sequentially.
-
-A more interesting example involves array reads and writes. We return to the Aiken and Nicolau
-example:
-
-\begin{tabular}{lll}
-\begin{lstlisting}[language=Java]
-for (i=1; i<=N; i++) {
-  A[i]=f1(B[i]);
-  B[i]=f2(A[i],D[i-1]);
-  C[i]=f3(A[i],D[i-1]);
-  D[i]=f4(B[i],C[i]);
-}
-
-\end{lstlisting}
-& 
-~
-&
-\begin{lstlisting}[language=Java]
-for (i=1; i<=N; i++) {
-  if (i%N==1) para(A[i]=f1(B[i]),...);
-  if (i%1==1) para(B[i]=f2(A[i],D[i-1]),...);
-  if (i%1==1) para(C[i]=f3(A[i],D[i-1]),...);
-  if (i%1==1) para(D[i]=f4(B[i],C[i]),...);
-}	
-\end{lstlisting}
-
-\end{tabular}
-
-The above across-loop schedule entails that the first instruction is parallelized---all
-N instructions A can be computed at once because each depends only on {\sf B[i]}, 
-which at the point {\sf A[i]} is computed, refers to a value assigned outside of the loop. 
-In contrast, subsequent instructions are executed sequentially. Take B, for example. 
-There is a dependence ``cycle'' from {\sf B[i]} to {\sf D[i]} then {\sf B[i+1]} in the subsequent 
-iteration. (The cycle is shown in~\figref{fig:MPCexample}(b).) {\sf B[i]}'s cannot be computed 
-in parallel because each one depends on the {\sf B[i]} value computed in the previous iteration.
-
-Now consider the above code but slightly modified. {\sf B[i]} and {\sf C[i]} are now 
-computed based on the value of {\sf D[i-2]} assigned two iterations earlier.
-
-\begin{tabular}{lll}
-\begin{lstlisting}[language=Java]
-for (i=1; i<=N; i++) {
-  A[i]=f1(B[i]);
-  B[i]=f2(A[i],D[i-2]);
-  C[i]=f3(A[i],D[i-2]);
-  D[i]=f4(B[i],C[i]);
-}
-
-\end{lstlisting}
-& 
-~
-&
-\begin{lstlisting}[language=Java]
-for (i=1; i<=N; i++) {
-  if (i%N==1) para(A[i]=f1(B[i]));
-  if (i%2==1) para(B[i]=f2(A[i],D[i-2]));
-  if (i%2==1) para(C[i]=f3(A[i],D[i-2]));
-  if (i%2==1) para(D[i]=f4(B[i],C[i]));
-}	
-\end{lstlisting}
-
-\end{tabular}
-
-{\sf B[i]} still depends on values computed in earlier iterations, but one of the values, D[i-2] 
-is two iterations earlier. Thus, we can take parallelization steps of 2, e.g., computing 
-{\sf B[1],B[2]} in parallel, then {\sf B[3],B[4]}, etc.
-Note however, that we cannot take a parallelization step larger than 2.
-%The value of B will depend on a value being computed in parallel.
-
-To simplify, we assume that a loop-body schedule is a straight line. \ana{What does this mean?}
-\ana{Will we relax?}
-The problem becomes: for each step, compute the maximal parallelization 
-step allowed by program dependences. This will account for an MPC schedule with 
-minimal cost. \ana{Why?}
-
-\subsection{Preliminaries}
-
-%\begin{figure*}[tbhp]
-\begin{figure}[tbhp]
-\begin{tabular}{ll}
-\begin{minipage}[b]{4.25cm}
-
-\begin{minted}[fontsize=\footnotesize, linenos, numbersep=5pt, escapeinside=||]{java}
-// returns val%mod    
-int rem(int val, int mod) {
-  int rem = 0;
-  for (int j = LEN-1; j |$\ge$| 0; j--) 
-  {
-    rem = rem << 1;
-    // rem[0] = val[j]                                                                                                            
-    rem = rem + ((val>>j)&1);   
-    if (rem |$\ge$| mod) 
-    {
-      rem = rem - mod;
-    }
-  }    
-  return rem;
-}
-\end{minted}
-\end{minipage}
-
-&
-
-\begin{minipage}[b]{4.25cm}
-\begin{minted}[fontsize=\footnotesize, linenos, numbersep=5pt, escapeinside=||]{java}
-// begin inlined rem
-int rem0 = 0;
-for (int j = LEN-1; j >= 0; j--) 
-{
-  rem1 = (j==LEN-1) ? rem0 : rem5;
-  rem2 = rem1 << 1;
-  rem3 = rem2 + (x1>>j)&1; 
-  rem4 = rem3 - y1;
-  cnd1 = CMP(rem3 >= y1);
-  rem5 = MUX(rem3,rem4,cnd1);
-}  
-// end inline rem        
-\end{minted}
-\end{minipage}
-\\
-
-(a) GCD Source 
-&
-(c) MPC-source
-
-\end{tabular}
-\caption{(a) shows standard (IMP-like) source for an MPC-friendly implementation 
-of integer division~\cite{Demmler}, and (b) shows MPC-source. Line 5 {\sf rem1 = (j==LEN-1) ? rem0 : rem5;} 
-is what we call a {\em pseudo-$\phi$} node. Note that the {\sf if}-statement on left
-becomes straight-line code in MPC. The program \emph{runs both branches}, and 
-at the end runs the conditional of the if-statement followed by the multiplexer. 
-If {\sf cnd1} is {\sf true}, then {\sf rem} gets assigned the value along the left branch;
-otherwise, it gets assigned the value along the right branch.}
-    \label{fig:example}\vspace{-2ex}
-\end{figure}
-
-\subsubsection{MPC-Source}
-
-MPC-Source is an SSA-based intermediate representation
-of MPC programs~\cite{ccs paper}. MPC-Source is, essentially, a straight-line sequence
-of ``components'', where each ``component'' is either a nested for-loop sequence (called for-loop block), 
-or a simple MPC-statement. \figref{fig:example} shows an example. MPC-Source can be constructed
-in a standard way from an SSA-translation of the source---each branch of an 
-{\sf if}-statement is added to the sequence, the conditional becomes an MPC comparison (CMP), 
-and the $\phi$-node becomes an MPC multiplexer (MUX).
-
-\ana{Add a paragraph on handling of arrays in standard SSA.}
-
-\subsubsection{Array SSA}
-
-The other essential concept we build upon is Array SSA. 
-Originally proposed by Knobe and Sarkar, Array SSA has been used for loop parallelization in HPC.
-We adapt Array SSA for MPC-source and build our optimal loop parallelization algorithm
-based on the adaptation.
-
-\ana{3. MPC-Source with Array SSA}
-
-Array SSA is best illustrated by an example. Consider Aiken and Nicolau's example in Array SSA:
-
-\begin{figure}
-\begin{lstlisting}[language=Java,escapeinside=`']
-for (i=1; i<=N; i++) {
-
-  Invariant: @A1[j] == j if `$j < i$', @A1[j] == 0 otherwise
-
-  A1 = `$\phi$'(A3,A0); @A1 = max(A3,A0);  
-  B1 = `$\phi$'(B3,B0); @B1 = max(B3,B0);
-  C1 = `$\phi$'(C3,C0); @C1 = max(C3,C0);
-  D1 = `$\phi$'(D3,D0); @D1 = max(D3,D0);
-
-  A2[i] = f1(B1[i]); @A2[i] = i;
-  A3 = `$\phi$'(A2,A1); @A3 = max(A2, A1);
-  B2[i] = f2(A3[i],D1[i-1]); @B2[i] = i;
-  B3 = `$\phi$'(B2,B1); @B3 = max(B2, B1);
-  C3[i] = f3(A3[i],D1[i-1]); @C2[i] = i;
-  C3 = `$\phi$'(C2,C1); @C3 = max(C2, C1);
-  D3[i] = f4(B3[i],C3[i]); @D2[i] = i;
-  D3 = `$\phi$'(D2,D1); @D3 = max(D2, D1); 
-}
-\end{lstlisting}
-\caption{Aiken and Nicolau's example using Array SSA.}
-\label{fig:ArraySSAAikenNicolau}
-\end{figure}
-
-There are three kinds of statements added by Array SSA. First, 
-\[
-{\sf A2[i] = f1(B1[i]); @A2[i] = i;}
-\]
-Array SSA statement {\sf @A2[i] = i;} records the iteration, {\sf i} when {\sf A2[i]} is written.
-Note that the index is not necessarily {\sf i}, it can be any $f(i)$, however, we assume
-that the loop always iterates from 1 to $N$ with step 1. Second and third,
-\[ 
-{\sf A3 = }\phi{\sf (A2,A1); @A3 = max(A2, A1);}
-\]
-Statement ${\sf A3 = }\phi{\sf (A2,A1);}$ is a \emph{define} $\phi$ node where arrays 
-{\sf A2} and {\sf A1} are ``merged''---in this case, each value at index other than $i$ 
-is taken from array {\sf A1}, and index $i$ is taken from {\sf A2}. Statement 
-{\sf @A3 = max(A2, A1);} is a timestamp statement; it picks the latest iteration 
-to assign for each index $i$---in this case, when the phi- and max- statements 
-appear immediately under an array write, it assigns {\sf @A3[j] = i} if $j=i$, 
-and {\sf @A3[j] = @A2[j]} otherwise.
-
-\ana{Add merge $\phi$}
-
-\subsubsection{MPC-Source with Array SSA}
-
-The phi-node at if-then-else statements, e.g., ${\sf A3 = \phi(A2,A1), @A3 = max(A2, A1);}$ 
-works for MPC-source without loss of privacy. ${\sf A3 = \phi(A2,A1)}$ is translated as 
-point-wise MUX for each index $i$. The timestamp statement ${\sf @A3 = max(A2, A1);}$ 
-has the same meaning as in standard Array SSA---{\sf @A3[i] = max(@A2[i],@A1[i])}.
-Since MPC-source expands and executes each arm of the if-then-else statement,
-if the then-arm modifies the array at index $i$, but the else arm does not, 
-{\sf @A3[i] = max(@A2[i],@A1[i]) = i}. Similarly, if both arms modify the array at $i$,
-the index again will be $i$. We note that in MPC-source the timestamp arrays {\sf @A} 
-are known \emph{statically}, unlike in standard Array SSA, where they are inherently \emph{dynamic}. 
-\ana{Double check? Make a statement that an adversary monitoring execution does not
-infer anything from SSA phi nodes.}
-
-\subsection{Analysis Algorithm}
-
-\subsubsection{Restrictions on MPC-Source}
-
-To make the problem tractable we adopt certain constraints on the loop body and loop body schedule.
-These constraints hold in classical works on optimal loop parallelization, e.g., Aiken and Nicolau.
-
-Array writes.~\ana{We have to see what the distinction was, regular vs. irregular loops}:
-\begin{enumerate}
-\item For each array read of $A$: $... = ... A[f(i)]$ and corresponding write to $A$: $A[f'(i)] = ...$, 
-we have that $f(i) = f'(i)$ for every $0 \le i < N$. Here $i$ is the induction variable and $N$ is
-the loop bound.
-
-\item Let $A[f_1(i)]...A[f_k(i)]$ be a sequence of writes to $A$ in the loop body sequence 
-\ana{How is the sequence determined?} with $A[f_k(i)]$ being the most recent one. Here
-$i$ denotes the induction variable. We require that for every $0 \le i < N, j,k$, $f_j(i) \neq f_k(i)$, 
-or in other words, those writes access different elements of the array. Furthermore, we require 
-that $f_k(i) \neq f_l(j)$, or in other words, writes across different iterations access different array 
-elements. \ana{Needs rewording. Can be simplified.}
-\end{enumerate}
-
-Constant loop schedule. Let $S$ be a loop schedule, i.e., a directed acyclic graph that defines "executes before"
-scheduling constraints among the statements of the loop body. We require that the 
-parallelization of the loop schedule obeys the constant loop schedule constraint: each instance of $S$
-in the loop linearization obeys the scheduling constraints defined by the DAG of $S$. \ana{Justify why!}
-
-\subsubsection{Computing the Index Invariant}
-
-"Index invariant" makes an assertion about when A[i] is written as a function of i.
-\ana{Can we compute this index easily: A[f(j)] = i, if j>=i, A[f(j)] = 0 otherwise?
-Under what assumptions this is true???}. \ana{We will use the Array SSA stuff to compute index; 
-can get rid of it once we.}
-
-\subsubsection{Computing $d(s)$}
-We assume a loop body that consists of three-address MPC statements:
-\[ v1 = v2 \oplus v3 \]
-MUX statements
-\[ v1 = MUX(v2, v3, v4) \]
-and
-phi-nodes
-\[ v1 = \phi(v2, v3) \]
-where $v1,v2$, $v3$, and $v4$ are either scalar variables, e.g., {\sf x} or array elements {\sf A[i]}.
-Array-write phi- and max- statements are elided from the representation, as
-their only purpose is to compute the invariant. Control-merge phi- are translated
-into point-wise MUX statements, as is standard in MPC. \ana{Are they translated during analysis?}
-
-\ana{Pseudo phi nodes!}
-
-\begin{figure}
-\begin{lstlisting}[language=Java,escapeinside=`']
-for (i=1; i<=N; i++) {
-
-  // Index Invariant: @A1[j] == j if `$j < i$', @A1[j] == 0 otherwise
-
-  A1 = `$\phi$'(A3,A0); 
-  B1 = `$\phi$'(B3,B0); 
-  C1 = `$\phi$'(C3,C0);
-  D1 = `$\phi$'(D3,D0);
-
-  A2[i] = f1(B1[i]); 
-  B2[i] = f2(A3[i],D1[i-1]); 
-  C3[i] = f3(A3[i],D1[i-1]); 
-  D3[i] = f4(B3[i],C3[i]); 
-}
-\end{lstlisting}
-\caption{Aiken and Nicolau's example using Array SSA.}
-\label{fig:ArraySSAMPCAikenNicolau}
-\end{figure}
-
-%\ana{Figure out how to reason about cycles based on Array SSA.}
-
-Recall that the loop body is a linear sequence of statement. To compute $d(s)$ of statement $s$, 
-we follow def-use chains backwards collecting all reachable pseudo-phi node. If there is a psudo-phi 
-node of a scalar, then $d(s) = 1$; for an array we retrieve the value $A[f(i)]$ from the invariant. \ana{formalize, polish!}
-
-\ana{Define Schedule $S$.}
-
-\ana{Define "constant loop body schedule"}
-
-\ana{Prove main theorem, that this schedule is of minimal cost. I.e., under the restriction of "constant loop body schedule", there
-does not exist a schedule $S'$}
-
-for a statement x = f(y,z), follow the reverse def-use link.
-
-\section{Extensions}
-
-\subsection{Multi Dimensional Arrays}
-
-\subsection{Nested Loops}
+\section{Implementation and Evaluation}
+\label{sec:implementation}
 

### 2021-11-21 paper_SIMD.tex
Fixed typo.

@@
 
-The dependence graph computes def-use edges:
+The dependence graph has the following def-use edges:
 
 \begin{itemize}
-\item same-level $X\rightarrow Y$ where $X = ...$ and $Y = ...$ are in the same loop nest $i_1,i_2,...i_k$.
+\item same-level (forward) $X\rightarrow Y$ where $X = ...$ and $Y = ...$ are in the same loop nest $i_1,i_2,...i_k$.
 E.g., $\texttt{d} = \mathit{SUM}(\texttt{S}[i,j], \texttt{C}[j])$ to $\texttt{p} = \mathit{MUL}(\texttt{d},\texttt{d})$ in Biometric is a same-level edge.
-\item outer-to-inner $X\rightarrow Y$ where $X$ is in an outer loop nest, $i_1,i_2,...i_j$, and $Y$ is in an inner one, $i_1,i_2,...i_j,...,i_k$.
-\item inner-to-outer $X\rightarrow Y$ where $X$ is in an inner loop nest, $i_1,i_2,...i_j,...,i_k$, and $Y$ is in an outer one, $i_1,i_2,...i_j$.
+\item outer-to-inner (forward) $X\rightarrow Y$ where $X$ is in an outer loop nest, $i_1,i_2,...i_j$, and $Y$ is in an inner one, $i_1,i_2,...i_j,...,i_k$.
+\item inner-to-outer (forward) $X\rightarrow Y$ where $X$ is in an inner loop nest, $i_1,i_2,...i_j,...,i_k$, and $Y$ is in an outer one, $i_1,i_2,...i_j$.
 E.g.
-$\texttt{sum}_1 = \texttt{ADD}(\texttt{sum}_0,\texttt{p})$ to
-$\texttt{c} = \mathit{CMP}(\texttt{sum}_1,\texttt{min}_0)$ is an inner-to-outer edge.
+$\texttt{sum}_1 = \texttt{ADD}(\texttt{sum}_0,\texttt{p})$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_1,\texttt{min}_0)$ is an inner-to-outer edge.
 \item same-level back-edge $X\rightarrow Y$.
@@
 
-We also define the $\mathit{closure}(n)$ where $n$ is a phi-node. Intuitively, it computes the set of statements that form a cycle in the loop with $n$. The closure of $n$ is defined as follows:
+We also define $\mathit{closure}(n)$ where $n$ is a phi-node.
+Intuitively, it computes the set of nodes (i.e., statements) that form a dependence cycle with $n$.
+\ana{Cycle(n) is probably a better name.} The closure of $n$ is defined as follows:
 \begin{itemize}
@@
 A back-edge from a non-phi-node in loop $i$ to a phi-node in loop $i$'s header is a same-level edge.
-The only difference with the handling of normal def-use is that the operand index will become $i-1$.
+The only difference with the handling of normal same-level forward def-use is that the operand index will become $i-1$.
 A back edge from a phi-node in loop $j$ to a phi-node in loop $i$, where $j$ is immediately nested in $i$ is an inner-to-outer edge and will require dropping dimension.
@@
 \STATE {change $\stackrel{\rightarrow}{i_k}$ to $i_k$ in all statements in loop}
-\STATE {treat FOR loop as monolith node: change def-use edges accordingly.}
+\STATE {treat FOR loop as monolith node: change def-use edges accordingly. Some edges become same-level.}
 \ENDFOR
@@
 \STATE {$\texttt{c} = \mathit{CMP}(\texttt{sum}_1,\texttt{min}_0)$}
-\STATE {$\texttt{index}_1 = \mathit{MUX}(\texttt{c},\texttt{index}_1,i)$}
-\STATE {$\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{sum}_1,\texttt{min}_1)$}
+\STATE {$\texttt{index}_1 = \mathit{MUX}(\texttt{c},\texttt{index}_0,i)$}
+\STATE {$\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{min}_0,\texttt{sum}_1)$}
 \STATE \COMMENT {End of outer loop $i$.}
@@
 It raises the dimensions of scalars and optimistically vectorizes all operations.
-The last phase discovers loop-carried dependences and removes affected vectorization.
+The next phase discovers loop-carried dependences and removes affected vectorization.
 
@@
 \STATE {$\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$}
-\STATE {$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_1[\stackrel{\rightarrow}{i}],[0,1,...N])$}
-\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_1[\stackrel{\rightarrow}{i}])$}
+\STATE {$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$}
+\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_1[\stackrel{\rightarrow}{i}])$}
 \STATE \COMMENT {End of outer loop $i$.}
@@
 
-Starting at the inner phi-node $\texttt{sum}_0[i,j] = \phi(\texttt{sum}_1[i,j], [0,0,...])$, the algorithm first computes it's closure.
+Starting at the inner phi-node $\texttt{sum}_0[i,j] = \phi(\texttt{sum}_1[i,j], [0,0,...])$, the algorithm first computes its closure.
 The closure amounts to the phi-node itself and $\texttt{sum}_1[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$, accounting for the loop-carried dependency of the computation of \texttt{sum}.
@@
 \texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
-\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_1[\stackrel{\rightarrow}{i}],[0,1,...N]) 
+\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N]) 
 \end{array}
@@
 \texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
-\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_1[\stackrel{\rightarrow}{i}])
+\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_1[\stackrel{\rightarrow}{i}])
 \end{array}
@@
 \STATE \hspace{0.25cm}{$\texttt{c}[{i}] = \mathit{CMP}(\texttt{sum}'_1[{i}],\texttt{min}_0[{i}])$}
-\STATE \hspace{0.25cm}{$\texttt{index}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{index}_1[{i}],[0,1,...N])$}
-\STATE \hspace{0.25cm}{$\texttt{min}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{sum}'_1[{i}],\texttt{min}_1[{i}])$}
+\STATE \hspace{0.25cm}{$\texttt{index}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{index}_0[{i}],[0,1,...N])$}
+\STATE \hspace{0.25cm}{$\texttt{min}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{min}_0[{i}],\texttt{sum}'_1[{i}])$}
 

### 2021-11-22 paper_SIMD.tex
Fixed typos.
added my comments/questions

@@
 E.g., $\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{sum}_1,\texttt{min}_1)$ to $\texttt{min}_0 = \phi(\texttt{min}_1, 10000)$ in Biometric is a same-level back-edge.
-\item outer-to-inner back-edge $X\rightarrow Y$: $X$ and $Y$ are both $\phi$-nodes for some variable. The source $X$ is in an immediately nested loop of $Y$'s loop. 
+\item inner-to-outer back-edge $X\rightarrow Y$: $X$ and $Y$ are both $\phi$-nodes for some variable. The source $X$ is in an immediately nested loop of $Y$'s loop. 
  \end{itemize}
@@
 
-\section{Basic Vectorization}
+\section{Vectorization}
 
@@
 
+\subsection{Basic Vectorization}
+
 \begin{algorithmic}
@@
 \STATE {case def-use edge $Y_i \rightarrow X$ of} \\
-\STATE \hspace{0.25cm} {same-level $\rightarrow$ add $Y_i = Y_i$} \\
+\STATE \hspace{0.25cm} {same-level $\rightarrow$ add $Y'_i = Y_i$} \\
 \STATE \hspace{0.25cm} {outer-to-inner $\rightarrow$ add $Y'_i[i_1,...,i_k] = \mathit{raise\_dim}(Y_i)$} 
@@
 \COMMENT {Optimistically vectorize all. $\stackrel{\rightarrow}{i}$ means vectorized dimension.}
-\STATE {change to $X[\stackrel{\rightarrow}{i_1},...\stackrel{\rightarrow}{i_k}] = \mathit{Op}(Y'_1[i_1,...,i_k],Y'_2[i_1,...,i_k])$ }
+\STATE {change to $X[\stackrel{\rightarrow}{i_1},...\stackrel{\rightarrow}{i_k}] = \mathit{Op}(Y'_1[\stackrel{\rightarrow}{i_1},...,\stackrel{\rightarrow}{i_k}],Y'_2[\stackrel{\rightarrow}{i_1},...,\stackrel{\rightarrow}{i_k}])$ }
 \ENDFOR
@@
 \end{algorithmic}
+
+\ishaq{We need to think of a way to handle 4th case for def-use, when a nested block B def is followed by another, different, nested block B' use.
+Note: The handwritten notes say we can break it into 2 edges.
+One idea to this end is to insert a def' node which is a copy of the original def.
+The def' should be placed in a block that contains both B and B'.}
 
@@
 
+\ishaq{At least for input variables: when raising dimensions, we should keep some meta data around so we can later write code that could use 1 share if it knows the other N shares are just duplicates of it.}
+
+\ishaq{Naive question: Which index of $\texttt{min}_{1}[\stackrel{\rightarrow}{i}]$ would contain the minimum, how are we reducing it?}
+
 \subsubsection{Phase 2}
@@
 The algorithm adds same-level edges, one from $\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = ...$ to the monolithic FOR-loop node, and one from the FOR-loop node to $\texttt{c}[\stackrel{\rightarrow}{i}] = ...$.
+
+\ishaq{For our rader: Such dependencies will not involve phi nodes only. e.g. $a[i] = a[i] + a[i-1]$}
 
@@
 \]
+
+\ishaq{I don't think I understand closures.
+Why does the closure for $\texttt{index}_0[i]$ include $\texttt{c}[\stackrel{\rightarrow}{i}]$ but the closure for $\texttt{index}_0[i] = \phi(\texttt{index}_1[i-1], -1)$ not include $\texttt{p}$.}
+
 Since the two closures intersect, we merge them into a single loop.

### 2021-11-23 paper_SIMD.tex
Fixed some typos in paper.

@@
 
+A pseudo $\phi$-node $X_1 = \phi(X_0,X_2)$ in a loop header is evaluated during circuit generation.
+If it is the 0-th iteration, then the $\phi$-node evaluates to $X_0$, otherwise, it evaluates to $X_2$.
+
+\ana{Have to double check and make sure we capture all the cases.}
 The dependence graph has the following def-use edges:
@@
 \begin{itemize}
-\item same-level (forward) $X\rightarrow Y$ where $X = ...$ and $Y = ...$ are in the same loop nest $i_1,i_2,...i_k$.
+\item same-level forward $X\rightarrow Y$ where $X$ and $Y$ are in the same loop nest $i_1,i_2,...i_k$.
 E.g., $\texttt{d} = \mathit{SUM}(\texttt{S}[i,j], \texttt{C}[j])$ to $\texttt{p} = \mathit{MUL}(\texttt{d},\texttt{d})$ in Biometric is a same-level edge.
-\item outer-to-inner (forward) $X\rightarrow Y$ where $X$ is in an outer loop nest, $i_1,i_2,...i_j$, and $Y$ is in an inner one, $i_1,i_2,...i_j,...,i_k$.
-\item inner-to-outer (forward) $X\rightarrow Y$ where $X$ is in an inner loop nest, $i_1,i_2,...i_j,...,i_k$, and $Y$ is in an outer one, $i_1,i_2,...i_j$.
+A $\phi$ node can be a source of a same-level forward edge but not a target.
+\item outer-to-inner forward $X\rightarrow Y$ where $X$ is in an outer loop nest, $i_1,i_2,...i_j$, and $Y$ is in an inner one, $i_1,i_2,...i_j,...,i_k$.
+A $\phi$-node can be a source or a target of an outer-to-inner forward edge.
+\item inner-to-outer forward $X\rightarrow Y$ where $X$ is a $phi$-node in an inner loop nest, $i_1,i_2,...i_k,i_{k+1}$, and $Y$ is in the enclosing loop nest $i_1,i_2,...i_k$.
 E.g.
-$\texttt{sum}_1 = \texttt{ADD}(\texttt{sum}_0,\texttt{p})$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_1,\texttt{min}_0)$ is an inner-to-outer edge.
+$\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$ is an inner-to-outer forward edge.
+Note that the source is \emph{always} a $\phi$-node in the immediately enclosing loop.
+The interpretation of this edge is that the use node $Y$ uses the definition made in the last iteration of the inner loop.
 \item same-level back-edge $X\rightarrow Y$.
@@
 E.g., $\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{sum}_1,\texttt{min}_1)$ to $\texttt{min}_0 = \phi(\texttt{min}_1, 10000)$ in Biometric is a same-level back-edge.
-\item inner-to-outer back-edge $X\rightarrow Y$: $X$ and $Y$ are both $\phi$-nodes for some variable. The source $X$ is in an immediately nested loop of $Y$'s loop. 
+\item inner-to-outer back-edge $X\rightarrow Y$: $X$ and $Y$ are both $\phi$-nodes for some variable. The source $X$ is in a loop nest immediately nested into $Y$'s loop nest. 
  \end{itemize}
@@
 \FOR {each $Y_i$}
-\STATE {case def-use edge $Y_i \rightarrow X$ of} \\
+\STATE {case def-use edge $\mathit{stmt}' (\mbox{def of } Y_i) \rightarrow \mathit{stmt} (\mbox{def of } X)$ of} \\
 \STATE \hspace{0.25cm} {same-level $\rightarrow$ add $Y'_i = Y_i$} \\
@@
 \STATE \COMMENT {End of inner loop $j$.}
-\STATE {$\texttt{c} = \mathit{CMP}(\texttt{sum}_1,\texttt{min}_0)$}
+\STATE {$\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$}
 \STATE {$\texttt{index}_1 = \mathit{MUX}(\texttt{c},\texttt{index}_0,i)$}
-\STATE {$\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{min}_0,\texttt{sum}_1)$}
+\STATE {$\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{min}_0,\texttt{sum}_0)$}
 \STATE \COMMENT {End of outer loop $i$.}
@@
 \STATE \COMMENT {End of inner loop $j$.}
-\STATE {$\texttt{sum}'_1[i] = \mathit{drop\_dim}(\texttt{sum}_1[i,j],j)$}
-\STATE {$\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$}
+\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
+\STATE {$\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$}
 \STATE {$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$}
-\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_1[\stackrel{\rightarrow}{i}])$}
+\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])$}
 \STATE \COMMENT {End of outer loop $i$.}
@@
 
-\STATE {$\texttt{sum}'_1[i] = \mathit{drop\_dim}(\texttt{sum}_1[i,j],j)$}
-\STATE {$\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$}
+\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
+\STATE {$\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$}
 \STATE {$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$}
-\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_1[\stackrel{\rightarrow}{i}])$}
+\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])$}
 \STATE \COMMENT {End of outer loop $i$.}
@@
 
-Next, analysis moves to outer dimension $i$. There are two phi-nodes, $\texttt{index}_0[i] = \phi(\texttt{index}_1[i-1], -1)$ and $\texttt{min}_0[i] = \phi(\texttt{min}_1[i-1], 10000)$.
+Next, analysis moves to outer dimension $i$. There are two phi-nodes, $\texttt{min}_0[i] = \phi(\texttt{min}_1[i-1], 10000)$ and $\texttt{index}_0[i] = \phi(\texttt{index}_1[i-1], -1)$.
 The closure of the first is 
@@
 \begin{array}{l}
-\texttt{index}_0[i] = \phi(\texttt{index}_1[i-1], -1), \\ 
-\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
-\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N]) 
+\texttt{min}_0[i] = \phi(\texttt{min}_1[i-1], 10000) \\
+\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
+\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])
 \end{array}
@@
 \begin{array}{l}
-\texttt{min}_0[i] = \phi(\texttt{min}_1[i-1], 10000) \\
-\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
-\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_1[\stackrel{\rightarrow}{i}])
+\texttt{index}_0[i] = \phi(\texttt{index}_1[i-1], -1), \\ 
+%\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
+\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N]) 
 \end{array}
@@
 
-Since the two closures intersect, we merge them into a single loop.
-We can now rewrite those statements, getting rid of the vectorization on $i$ in second FOR loop.
+Since the two closures \emph{do not} intersect, we have two distinct FOR-loops on $i$.
+The first FOR loop (on \texttt{min}) is scheduled first, because of the dependence edge from $\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$ to
+$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$.
+We can now rewrite those statements, getting rid of the vectorization on $i$ in the two FOR loops.
 Notably, the SUB and MUL computations are fully vectorizable and the ADD computation is vectorizable across the $i$-dimension.
@@
 \STATE {} 
-\STATE {$\texttt{sum}'_1[i] = \mathit{drop\_dim}(\texttt{sum}_1[i,j],j)$}
+\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
+\STATE {} 
+\STATE {$\texttt{min}_0[i] = \phi(\texttt{min}_1[i-1], 10000)$}
+\STATE {FOR i=0; i<=N; i++} \COMMENT{ Turns into a MOTION loop, was $i$-loop.}
+\STATE \hspace{0.25cm}{$\texttt{c}[{i}] = \mathit{CMP}(\texttt{sum}'_0[{i}],\texttt{min}_0[{i}])$}
+\STATE \hspace{0.25cm}{$\texttt{min}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{min}_0[{i}],\texttt{sum}'_1[{i}])$}
 \STATE {}
 \STATE {$\texttt{index}_0[i] = \phi(\texttt{index}_1[i-1], -1)$} 
-\STATE {$\texttt{min}_0[i] = \phi(\texttt{min}_1[i-1], 10000)$}
 \STATE {FOR i=0; i<=N; i++} \COMMENT{ Turns into a MOTION loop, was $i$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{c}[{i}] = \mathit{CMP}(\texttt{sum}'_1[{i}],\texttt{min}_0[{i}])$}
-\STATE \hspace{0.25cm}{$\texttt{index}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{index}_0[{i}],[0,1,...N])$}
-\STATE \hspace{0.25cm}{$\texttt{min}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{min}_0[{i}],\texttt{sum}'_1[{i}])$}
+\STATE \hspace{0.25cm}{$\texttt{index}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{index}_0[{i}],i)$}
 

### 2021-11-29 benchmarks/psi.py
MPC-correct version of PSI.

@@ -5,16 +5,22 @@ import typing
 # requires: no repetition of elements in either A or B
 # requires: len(A) = SA, len(B) = SB
 def psi(A: list[int], SA, B: list[int], SB) -> list[int]:
+  dummy : int = 0
   result: list[int] = []
   for i in range(0,SA):
+    flag : Bool = False
     for j in range(0,SB):
       if A[i] == B[j]:
-        #overloaded +. This is append actually.
-        result = result + [A[i]]
+        flag = True
+    val : int = dummy
+    if flag:
+      val : int = A[i]
+    #overloaded +. This is append actually.
+    result = result + [val]
   return result
 
-A = [1,2,3]
-B = [2]
-intersect = psi(A,3,B,1)
+A = [4,2,3,1,10]
+B = [2,10,3,4,5,6,7]
+intersect = psi(A,5,B,7)
 print(intersect)
 

### 2021-11-29 paper_SIMD.tex
Fixed bugs in paper.
Added removal of edges.

@@
 
-\ana{Have to double check and make sure we capture all the cases.}
+\ana{We need to double check that the following captures all cases of def-use edges.}
 The dependence graph has the following def-use edges:
@@
 
+\subsubsection{Removal of Infeasible Edges}
+
+Array writes limit vectorization as they introduce sometimes infeasible loop-carried dependencies. Consider the following example: \ana{Have to add citation to Aiken's paper}
+
+\begin{algorithmic}
+\STATE for i in range(N):
+\STATE \hspace{0.25cm} A[i] = B[i] + 10;
+\STATE \hspace{0.25cm} B[i] = A[i] * D[i-1];
+\STATE \hspace{0.25cm} C[i] = A[i] * D[i-1];
+\STATE \hspace{0.25cm} D[i] = B[i] * C[i];
+\end{algorithmic}
+
+In Cytron's SSA this code (roughly) translates into
+
+\begin{algorithmic}
+\STATE for i in range(N):
+\STATE 1. $\texttt{A}_0$ = $\phi(\texttt{A},\texttt{A}_1)$ 
+\STATE 2. $\texttt{B}_0$ = $\phi(\texttt{B},\texttt{B}_1)$ 
+\STATE 3. $\texttt{C}_0$ = $\phi(\texttt{C},\texttt{C}_1)$ 
+\STATE 4. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$ 
+\STATE 5. $\texttt{A}_1$ = $\texttt{A}_0$ 
+\STATE 6. $\texttt{A}_1$[i] = $\texttt{B}_0$[i] + 10;
+\STATE 7. $\texttt{B}_1$ = $\texttt{B}_0$ 
+\STATE 8. $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
+\STATE 9. $\texttt{C}_1$ = $\texttt{C}_0$ 
+\STATE 10. $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
+\STATE 11. $\texttt{D}_1$ = $\texttt{D}_0$ 
+\STATE 12. $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
+\end{algorithmic}
+
+There is a cycle around $\texttt{B}_0$ = $\phi(\texttt{B},\texttt{B}_1)$ that includes statement $\texttt{A}_1$[i] = $\texttt{B}_0$[i] + 10;
+and that statement won't be vectorized even though in fact there is no loop-carried dependency from the write of $\texttt{B}_1$[i] at 8 to the 
+read of $ ... = \texttt{B}_0$[i] at 6.
+
+The following algorithm removes certain infeasible loop-carried dependencies that are due to array writes. Consider a loop with index $0 \le j < J$
+nested at $i,j,k$. Here $i$ represents the enclosing loops of $j$ and $k$ represents the enclosed loops in $j$.
+
+\begin{algorithmic}
+\FOR {each array $\texttt{A}$ written in loop $j$} 
+\STATE \COMMENT { including enclosed loops in $j$ }
+\STATE dep = False
+\FOR {each pair def: $\texttt{A}_m[f(i,j,k)] = ... $, and use: $ ... = \texttt{A}_n[f'(i,j,k)]$ in loop $j$}
+\IF {$\exists \underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$, s.t.
+$0\!
+\le\!
+\underline{i}\!
+<\!
+I,\: 0\!
+\le\!
+\underline{j}, \underline{j}' \!
+<\!
+J,\: 0\!
+\le\!
+\underline{k}, \underline{k}' \!
+<\!
+K,\: \underline{j}<\underline{j}'$, and $f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$}
+\STATE dep = True
+\ENDIF
+\ENDFOR
+\IF {dep == False}
+\STATE remove back edge into $\texttt{A}$'s $\phi$-node in loop $j$.
+\ENDIF
+\ENDFOR
+\end{algorithmic}
+
+Consider a loop $j$ enclosed in some fixed $\underline{i}$. Only if an update (definition) $\texttt{A}_m[f(i,j,k)] = ... $ at some iteration $\underline{j}$ 
+references the \emph{same} array element as an use $ ... = \texttt{A}_n[f'(i,j,k)]$ at some later iteration $\underline{j}$, 
+we may have a loop-carried dependence for $\texttt{A}$ due to this def-use pair. (In contrast, Cytron's algorithm inserts a loop-carried dependency every time there is an array update.)
+The algorithm above examines all def-use pairs in loop $j$, including defs and uses in nested loops, searching for values $\underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$ that satisfy
+$f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$. If such values exist for some def-use pair, then there is a potential 
+loop-carried dependence on $\texttt{A}$; otherwise there is not and we can remove the spurious backward edge thus ``freeing up'' statements for vectorization. \ana{Double check this!!!}
+
+Consider the earlier example.
+There is a single loop, $i$.
+Clearly, there is no pair $\underline{i}$ and $\underline{i}'$, where $\underline{i} < \underline{i}'$ that make $\underline{i} = \underline{i}'$ (due to the def-use pairs of \texttt{A} 6-8 and 6-10).
+Therefore, we remove the back edge from 6 to 1.
+Analogously, we remove the back edges from 8 to 2 and 10 to 3.
+However, there are many values $\underline{i} < \underline{i'}$ that make $\underline{i} = \underline{i'}-1$ and the back edge from 12 to 4 remains (def-use pairs for \texttt{D}).
+As a result of removing these spurious edges, Basic Vectorization will find that statement 6 is vectorizable.
+Statements 8, 10 and 12 will correctly appear in the FOR loop.
+
+\subsubsection{Array MUX refinement}
+
+Next, the algorithm refines array MUX statement. 
+MPC-source after Cytron's SSA may result in statements $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$, which imply that any index 
+of A can be written at this point and therefore there is a loop-carried dependency. In some cases the MUX can be refined to just a single 
+index or a pair of indices, e.g., $\texttt{A}_j[i] = \emph{MUX}(c, \texttt{A}_k[i], \texttt{A}_l[i])$. \ana{Clarify why this helps.}
+
+\begin{algorithmic}
+\FOR {each $\mathit{stmt}$: $\texttt{A}_j = \emph{MUX}(c, \texttt{A}_k, \texttt{A}_l)$ in the MPC-source seq.}
+\STATE $i_1 = \mathit{find\_update}(A_k)$ \COMMENT { Is null when $A_k = \phi(...)$ }
+\STATE $i_2 = \mathit{find\_update}(A_l)$ \COMMENT { Is null when $A_l = \phi(...)$ }
+\IF {$i_1 \cong i_2$} \STATE \COMMENT{ Also $i_1 \cong i_2$ when one is null. WLG assume $i_2$ is null } 
+\STATE replace $\mathit{stmt}$ with \\
+\STATE $\texttt{A}_j = \texttt{A}_k$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
+%\ELSIF {$i_1 \ncong i_2$}  
+%\STATE replace $\mathit{stmt}$ with \\
+%\STATE $\texttt{A}_j = \texttt{A}_k$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
+\ELSE \STATE \COMMENT{ Also, not $\cong$ when one is $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$}
+\STATE $\mathit{stmt}$ stays as is
+\ENDIF
+\ENDFOR
+\end{algorithmic}
+
 \ana{TODO: Will extend with read/write arrays and removal of (certain) infeasible edges.}

### 2021-11-29 paper_SIMD.tex
Removed a TODO.
Fixed typos.

@@
 Consider a loop $j$ enclosed in some fixed $\underline{i}$. Only if an update (definition) $\texttt{A}_m[f(i,j,k)] = ... $ at some iteration $\underline{j}$ 
-references the \emph{same} array element as an use $ ... = \texttt{A}_n[f'(i,j,k)]$ at some later iteration $\underline{j}$, 
+references the \emph{same} array element as an use $ ... = \texttt{A}_n[f'(i,j,k)]$ at some later iteration $\underline{j}'$, 
 we may have a loop-carried dependence for $\texttt{A}$ due to this def-use pair. (In contrast, Cytron's algorithm inserts a loop-carried dependency every time there is an array update.)
@@
 of A can be written at this point and therefore there is a loop-carried dependency. In some cases the MUX can be refined to just a single 
-index or a pair of indices, e.g., $\texttt{A}_j[i] = \emph{MUX}(c, \texttt{A}_k[i], \texttt{A}_l[i])$. \ana{Clarify why this helps.}
+index or a pair of indices, e.g., $\texttt{A}_j[i] = \emph{MUX}(c, \texttt{A}_k[i], \texttt{A}_l[i])$. \ana{Clarify why this helps. Still work in progress.}
 
@@
 \end{algorithmic}
-
-\ana{TODO: Will extend with read/write arrays and removal of (certain) infeasible edges.}
 

### 2021-12-13 compiler/compiler/ast_shared.py
Add typing to function parameters

@@ -1,9 +1,14 @@
 from enum import Enum
-from typing import Generic, TypeVar, Union
+from typing import Generic, TypeVar, Union, Optional
 from dataclasses import dataclass
 from textwrap import indent
 
 import networkx  # type: ignore
+
+
+class VarType(str, Enum):
+    PLAINTEXT_INT = "plaintext"
+    SHARED_INT = "shared"
 
 
 @dataclass(frozen=True)
@@ -11,9 +16,13 @@ class Var:
     """A variable named `name`"""
 
     name: str
+    var_type: Optional[VarType] = None
 
     def __str__(self) -> str:
-        return self.name
+        str_rep = self.name
+        if self.var_type is not None:
+            str_rep += f": {self.var_type}"
+        return str_rep
 
     def name_without_ssa_rename(self) -> str:
         return self.name.strip("!")[0]

### 2021-12-13 compiler/compiler/ast_to_restricted_ast.py

@@ -8,6 +8,7 @@ from dataclasses import dataclass
 from typing import Optional, final, NoReturn
 
 from . import restricted_ast
+from .ast_shared import VarType
 
 
 @dataclass
@@ -397,7 +398,15 @@ class _FunctionConverter(_StrictNodeVisitor):
         return restricted_ast.Function(
             # TODO: Exclude other kinds of arguments
             name=node.name,
-            parameters=[restricted_ast.Var(name=arg.arg) for arg in node.args.args],
+            parameters=[
+                restricted_ast.Var(
+                    name=arg.arg,
+                    var_type=VarType.PLAINTEXT_INT
+                    if arg.annotation is None
+                    else VarType.SHARED_INT,
+                )
+                for arg in node.args.args
+            ],
             body=_convert_statements(self.source_code_info, node.body[:-1]),
             return_value=_ReturnValueGetter(self.source_code_info).visit(node.body[-1]),
         )

### 2021-12-13 paper_SIMD.tex
minor comments

@@
 The def' should be placed in a block that contains both B and B'.}
+\ishaq{THE ABOVE COMMENT IS RESOLVED.
+After transformation to SSA, there should be an intermediate variable (phi node) that is assigned a value depending on whether control branched or not.
+Should we put this into the writeup?}
 
@@
 
-\ishaq{Naive question: Which index of $\texttt{min}_{1}[\stackrel{\rightarrow}{i}]$ would contain the minimum, how are we reducing it?}
-
 \subsubsection{Phase 2}
@@
 
-\ishaq{For our rader: Such dependencies will not involve phi nodes only. e.g. $a[i] = a[i] + a[i-1]$}
+\ishaq{For our radar: Such dependencies will not involve phi nodes only. e.g. $a[i] = a[i] + a[i-1]$}
 
@@
 \]
-
-\ishaq{I don't think I understand closures.
-Why does the closure for $\texttt{index}_0[i]$ include $\texttt{c}[\stackrel{\rightarrow}{i}]$ but the closure for $\texttt{index}_0[i] = \phi(\texttt{index}_1[i-1], -1)$ not include $\texttt{p}$.}
 
@@
 
+\ishaq{Note to self: This algorithm is an instantiation for $j$ loop, the one for $k$ loop will be exactly the same, modulo variable name.}.
+
 Consider a loop $j$ enclosed in some fixed $\underline{i}$. Only if an update (definition) $\texttt{A}_m[f(i,j,k)] = ... $ at some iteration $\underline{j}$ 
-references the \emph{same} array element as an use $ ... = \texttt{A}_n[f'(i,j,k)]$ at some later iteration $\underline{j}'$, 
+references the \emph{same} array element as a use $ ... = \texttt{A}_n[f'(i,j,k)]$ at some later iteration $\underline{j}'$, 
 we may have a loop-carried dependence for $\texttt{A}$ due to this def-use pair. (In contrast, Cytron's algorithm inserts a loop-carried dependency every time there is an array update.)

### 2021-12-17 paper_SIMD.tex
Work in progress commit. Added handling of "mixed" def-use edges. Still need add raise_dim and drop_dim and handling of array accesses A[f(i)].

@@
 
-\ana{We need to double check that the following captures all cases of def-use edges.}
+%\ana{We need to double check that the following captures all cases of def-use edges.}
 The dependence graph has the following def-use edges:
@@
 E.g., $\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{sum}_1,\texttt{min}_1)$ to $\texttt{min}_0 = \phi(\texttt{min}_1, 10000)$ in Biometric is a same-level back-edge.
-\item inner-to-outer back-edge $X\rightarrow Y$: $X$ and $Y$ are both $\phi$-nodes for some variable. The source $X$ is in a loop nest immediately nested into $Y$'s loop nest. 
+\item inner-to-outer back-edge $X\rightarrow Y$: $X$ and $Y$ are both $\phi$-nodes for some variable. The source $X$ is in a loop nested into $Y$'s loop (not necessarily immediately).
+
+\item mixed forward edge $X\rightarrow Y$.
+$X$ is a $phi$-node in some loop $i_1,i_2,...i_k,i_{k+1}$ and $Y$ is a node in a loop nested into $i_1,i_2,...i_k$.
+We transform mixed forward edges as follows.
+Let $\x_j$ be the variable defined at the phi-node $X$.
+We add a variable and assignment $\x'_j = \x_j$ immediately after the $i_1,i_2,...i_k$ loop.
+Then we replace the use of $\x_j$ at $Y$ with $\x'_j$.
+This transforms a mixed forward edge into an "inner-to-outer" forward edge followed by an outer-to-inner forward edge.
+Thus Basic Vectorization handles one of "same-level", "inner-to-outer", or "outer-to-inner" def-use edges.
  \end{itemize}
@@
 The only difference with the handling of normal same-level forward def-use is that the operand index will become $i-1$.
-A back edge from a phi-node in loop $j$ to a phi-node in loop $i$, where $j$ is immediately nested in $i$ is an inner-to-outer edge and will require dropping dimension.
+A back edge from a phi-node in loop $j$ to a phi-node in loop $i$, where $j$ is nested in $i$ is an inner-to-outer edge and will require dropping dimension.
 We can show these are the only kinds of back-edges that may occur.}
@@
 
-\STATE \COMMENT {Phase 1: Raise dimension of scalar variables to corresponding loop nest. We may assume linear traversal of the MPC-source. \ana{Double check!}}
+\STATE \COMMENT {Phase 1: Raise dimension of scalar variables to corresponding loop nest. We may assume linear traversal of the MPC-source.}
 
@@
 
-\subsubsection{Phase 1 of Vectorization algorithm:}
+\subsubsection{Phase 1 of Basic Vectorization:}
 
@@
 
-\ana{Conjecture: Basic Vectorization still works with minor extensions and correctness argument still hods. }
-
 \section{Divide-and-Conquer}

### 2021-12-17 paper_SIMD.tex
Added raise_dim and drop_dim. Still need to add handling of A[f(i)].

@@
 
+\subsection{Pseudo $\phi$-nodes}
+
 A pseudo $\phi$-node $X_1 = \phi(X_0,X_2)$ in a loop header is evaluated during circuit generation.
@@
 %\ana{We need to double check that the following captures all cases of def-use edges.}
+
+\subsection{Def-use Edges}
+
 The dependence graph has the following def-use edges:
@@
 
-We also define $\mathit{closure}(n)$ where $n$ is a phi-node.
+\subsection{Helper Functions}
+
+We define $\mathit{closure}(n)$ where $n$ is a phi-node.
 Intuitively, it computes the set of nodes (i.e., statements) that form a dependence cycle with $n$.
@@
 \end{itemize}
+
+We define the $\mathit{raise\_dim}$ (raise dimensions) and $\mathit{drop\_dim}$ (drop dimension) functions.
+Raise dimension "lifts" a lower-dimension array (\ana{The right term here is tensor, I am pretty sure!}) into a higher dimension one.
+This is necessary when a lower-dimensional array is used in a higher dimensional loop and, essentially, is just copying of values.
+For example, Biometric contains the statement $d = \mathit{SUB}(S[i, j], C[j])$, in the $j$-loop which is nested into the $i$-loop.
+Here $C$ is a one-dimensional array, however, to vectorize across both loops it is necessary to turn it into a two-dimensional array: $C[i,j]$ becomes $[C[0],C[1],...C[J],C[0],C[1],...C[J], ...
+C[0],C[1],...C[J]]$, which turns the row into a matrix of $I$ identical rows.
+(We use capital letters to denote the upper bounds of loops, e.g., $J$ is the upper bound of the $j$-loop and $I$ is the upper bound of the $i$-loop.)
+
+$\mathit{raise\_dim}(A[i_1,...,i_k],i_{k+1})$ is defined as follows.
+It results in a new {k+1}-dimensional array $A'$ where for every $0\le i_{k+1} < I_{k+1}$, $A'[i_1,...,i_k,i_{k+1}] = A[i_1,...,i_k]$.
+We add the new dimension at the end arbitrarily, it can be added at any position, e.g., before $i_1$, etc.
+Adding $n$ dimensions is trivially extended as a composition of $n$ $\mathit{raise\_dim}$ that each adds a single dimension.
+
+As expected, drop dimension turns a higher-dimensional array into a lower-dimensional one.
+The key use case is an inner-to-outer def-use edge.
+The code may define a variable, e.g., $\x$ in an inner loop, say $j$, then use this variable in an enclosing loop, say $i$.
+Our algorithm may vectorize the computation of $\x$ in the $j$-loop thus producing a vector $\x[j]$ where $\x[1]$ is the value of \x after the 1st iteration and so on.
+Drop dimension states that the outer loop will use the value of the variable at the last iteration.
+
+$\mathit{drop\_dim}(A[i_1,...,i_k,i_{k+1}])$ produces a k-dimensional array $A'$ where $A'[i_1,...,i_k] = A'[i_1,...,i_k,I_{k+1}-1]$.
+In our analysis, dimensions are always dropped at the end, and again, one can define dropping $n$ dimensions as a composition of $n$ $\mathit{drop\_dim}$.
+
+\ana{This section is still informal. Needs work to make more precise.}
 
@@
 \STATE {$\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$}
-\STATE {$\texttt{d} = \mathit{SUM}(\texttt{S}[i,j], \texttt{C}[j])$}
+\STATE {$\texttt{d} = \mathit{SUB}(\texttt{S}[i,j], \texttt{C}[j])$}
 \STATE {$\texttt{p} = \mathit{MUL}(\texttt{d},\texttt{d})$}
@@
 \STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUM}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
+\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
 \STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
@@
 \STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUM\_SIMD}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
+\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB\_SIMD}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
 \STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL\_SIMD}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$} \\
@@
 
+\subsubsection{Array rebase}
+
+\ana{TODO, add here.}
+
 \subsubsection{Array MUX refinement}

### 2021-12-20 paper_SIMD.tex
Edits to Extension of Basic Vectorization.

@@
 $f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$. If such values exist for some def-use pair, then there is a potential 
-loop-carried dependence on $\texttt{A}$; otherwise there is not and we can remove the spurious backward edge thus ``freeing up'' statements for vectorization. \ana{Double check this!!!}
+loop-carried dependence on $\texttt{A}$; otherwise there is not and we can remove the spurious backward edge thus ``freeing up'' statements for vectorization. %\ana{Double check this!!!}
 
@@
 
-\subsubsection{Array rebase}
-
-\ana{TODO, add here.}
+\ana{TODO: We may remove phi-nodes that end up without incoming back-edges. This can free some cycles that we'd waste on bookkeeping arrays.}
 
@@
 
-Next, the algorithm refines array MUX statement. 
+\ana{TODO: I think we should implement this.}
+
+Next, the algorithm refines array MUX statements. 
 MPC-source after Cytron's SSA may result in statements $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$, which imply that any index 
 of A can be written at this point and therefore there is a loop-carried dependency. In some cases the MUX can be refined to just a single 
-index or a pair of indices, e.g., $\texttt{A}_j[i] = \emph{MUX}(c, \texttt{A}_k[i], \texttt{A}_l[i])$. \ana{Clarify why this helps. Still work in progress.}
+index or a pair of indices, e.g., $\texttt{A}_j[i] = \emph{MUX}(c, \texttt{A}_k[i], \texttt{A}_l[i])$.
+
+%\ana{Clarify why this helps. Still work in progress.}
+This is to reduce the dimensionality of simd-ified computation. Technically, $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$ is 
+a simdified operation that can be carried out in parallel "in one round". However, if \texttt{A} is a multi-dimensional array, this would lead 
+to substantial increase in the size of the arrays (vectors) we send to SIMD operations. Refining to an update to a specific index would reduce 
+the size of those vectors. Note that this is a heuristic that handle a common case, but not all cases of array updates.
 
@@
 \STATE $i_2 = \mathit{find\_update}(A_l)$ \COMMENT { Is null when $A_l = \phi(...)$ }
-\IF {$i_1 \cong i_2$} \STATE \COMMENT{ Also $i_1 \cong i_2$ when one is null. WLG assume $i_2$ is null } 
+\IF {$i_1 == i_2$ or $i_1$ is null or $i_2$ is null} \STATE \COMMENT{ With our restrictions on writes we must have $i_1 = i_2$. } 
 \STATE replace $\mathit{stmt}$ with \\
@@
 %\STATE $\texttt{A}_j = \texttt{A}_k$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
-\ELSE \STATE \COMMENT{ Also, not $\cong$ when one is $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$}
+\ELSE \STATE
 \STATE $\mathit{stmt}$ stays as is
@@
 \end{algorithmic}
+
+\subsubsection{Extending Basic Vectorization with Array Writes}
+
+\ana{TODO, add the weekend notes here.
+The big things are 1) Restricting array writes to "canonical writes" and 2) change of index, e.g., if accessing A[i] in loop j, have to change to accessing A[i,j].}
 

### 2021-12-20 compiler/compiler/type_analysis.py
Add type inference for array updates

@@ -64,9 +64,18 @@ def _type_assign_rhs(
     elif isinstance(rhs, Mux):
         return VarType.SHARED
     elif isinstance(rhs, Update):
-        # TODO: what is the return value of Update()?
-        # TODO: do Update statements need to update the type of the array?
-        return None
+        if (
+            _type_assign_rhs(rhs.value, type_env) == VarType.SHARED
+            or type_env.get(rhs.array.name) == VarType.SHARED
+        ):
+            return VarType.SHARED
+        elif (
+            _type_assign_rhs(rhs.value, type_env) == VarType.PLAINTEXT
+            and type_env.get(rhs.array.name) == VarType.PLAINTEXT
+        ):
+            return VarType.PLAINTEXT
+        else:
+            return None
 
     raise AssertionError(f"Unexpected type {type(rhs)} passed to _type_assign_rhs")
 

### 2021-12-21 paper_SIMD.tex
Work on extension with array writes: restrictions, changes and examples. Still need to iron out all changes to Basic vectorization algorithm.

@@
 We write $A[i_1,i_2,...i_k]$ interchangeably.
+\ana{Need to state this more precisely.}
 
@@
 \FOR {each dimension $d$ from highest to 0}
-\FOR {each $\phi$-node $n$ in loop $i_1,...,i_k$}
+\FOR {each $\phi$-node $n$ in loop $i_1,...,i_d$} 
 \STATE {compute $\mathit{closure}(n)$}
@@
 \FOR {each closure $cl$ (after merge)}
-\STATE {create FOR $i_k = 0; ...$ loop}  
+\STATE {create FOR $i_d = 0; ...$ loop}  
 \STATE {add $\phi$-nodes in $cl$ to header block} 
@@
 \STATE \COMMENT {Dimension is not vectorizable: }
-\STATE {change $\stackrel{\rightarrow}{i_k}$ to $i_k$ in all statements in loop}
+\STATE {change $\stackrel{\rightarrow}{i_d}$ to $i_d$ in all statements in loop}
 \STATE {treat FOR loop as monolith node: change def-use edges accordingly. Some edges become same-level.}
@@
 
-\subsection{Extensions of Basic Vectorization}
+\subsection{Towards Extension of Basic Vectorization}
 
@@
 This is to reduce the dimensionality of simd-ified computation. Technically, $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$ is 
-a simdified operation that can be carried out in parallel "in one round". However, if \texttt{A} is a multi-dimensional array, this would lead 
-to substantial increase in the size of the arrays (vectors) we send to SIMD operations. Refining to an update to a specific index would reduce 
-the size of those vectors. Note that this is a heuristic that handle a common case, but not all cases of array updates.
+a simdified operation that can be carried out in parallel "in one round". However, particularly when \texttt{A} is a multi-dimensional array, there is
+substantial increase in the size of the arrays (vectors) we send to SIMD operations. Refining to an update to a specific index would reduce 
+the size of those vectors. Note that this is a heuristic that handles a common case, but not all cases of array updates. 
+\ana{It may actually be handling \_all\_ cases, with the restriction to canonical updates.}
 
@@
 
-\subsubsection{Extending Basic Vectorization with Array Writes}
+\subsection{Extension of Basic Vectorization with Array Writes}
 
-\ana{TODO, add the weekend notes here.
-The big things are 1) Restricting array writes to "canonical writes" and 2) change of index, e.g., if accessing A[i] in loop j, have to change to accessing A[i,j].}
+\subsubsection{Restricting Array Writes}
+
+For now, we restrict array updates to \emph{canonical updates}. Assume (for simplicity) a two-dimensional array $\texttt{A}[I,J]$. A canonical update is the following:
+\begin{algorithmic}
+\STATE for i in range(I):
+\STATE \hspace{0.25cm} for j in range(J):
+\STATE \hspace{0.55cm} ... 
+\STATE \hspace{0.55cm} A[i,j] = ... 
+\STATE \hspace{0.55cm} ... 
+\end{algorithmic}
+
+The update $\texttt{A}[i,j]$ can be nested into an inner loop and there may be multiple updates.
+However, updates such as $\texttt{A}[i-1,j] = ...$ or $\texttt{A}[i-1,j-1] = ...$, etc., are not allowed.
+This simplifies reasoning about correctness in this early stage of the compiler; we will look to relax this restriction in future work.
+Reads through an arbitrary formula, such as $\texttt{A}[i-1]$ for example, are allowed, however, we assume the programmer ensures that the formula is within the bounds of the array.
+
+\subsubsection{Changes to Basic Vectorization}
+
+One change to Basic vectorization is the expansion of dimension if the array write or read occurs in a nested loop.
+That is, if there is an update to $\texttt{A}[i,j]$ that occurs in loop nest $i,j,k$, $\texttt{A}[i,j]$ will be rewritten into $\texttt{A}[i,j,k]$.
+Similarly, a read $\texttt{A}[f(i,j)]$ will be rewritten into $\texttt{A}[f(i,j),k]$.
+\ana{This notion of $f(i,j)$ needs a precise definition.
+For now I'm hoping the hand waving works.}.
+This can be a preprocessing step or it may be incorporated into the Basic Vectorization algorithm.
+Basic Vectorization will not expand array accesses at array writes the way it does with scalars now, as they are already expanded.
+
+\ana{TODO: Another change. What about the A = phi(A1,A2), and A = A1. We have to expand on what happens with those.}
+
+\ana{TODO: Third change. "Mixed" outer -> edge with arrays!}
+
+\subsubsection{Examples with Array Writes}
+
+\paragraph{Example 1}
+
+After removal of infeasible edges, the Aiken's array write example will be (roughly) as follows:
+
+\begin{algorithmic}
+\STATE for i in range(N):
+\STATE 1. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$ 
+\STATE 2. $\texttt{A}$[i] = $\texttt{B}$[i] + 10
+\STATE 3. $\texttt{B}$[i] = $\texttt{A}$[i] * $\texttt{D}_0$[i-1]
+\STATE 4. $\texttt{C}$[i] = $\texttt{A}$[i] * $\texttt{D}_0$[i-1]
+\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$ 
+\STATE 6. $\texttt{D}_1$[i] = $\texttt{B}$[i] * $\texttt{C}$[i]
+\end{algorithmic}
+
+There are no nested loops, thus, array accesses remain as is.
+
+After Phase 1 of Basic vectorization we have the following code:
+\begin{algorithmic}
+\STATE for i in range(N):
+\STATE 1. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$ 
+\STATE 2. $\texttt{A}[\stackrel{\rightarrow}{i}] = \texttt{B}[\stackrel{\rightarrow}i] + [10,...]$
+\STATE 3. $\texttt{B}[\stackrel{\rightarrow}i] = \texttt{A}[\stackrel{\rightarrow}i] * \texttt{D}_0[\stackrel{\rightarrow}{i-1}]$
+\STATE 4. $\texttt{C}[\stackrel{\rightarrow}i] = \texttt{A}[\stackrel{\rightarrow}i] * \texttt{D}_0[\stackrel{\rightarrow}{i-1}]$
+\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$ 
+\STATE 6. $\texttt{D}_1[\stackrel{\rightarrow}i] = \texttt{B}[\stackrel{\rightarrow}i] * \texttt{C}[\stackrel{\rightarrow}i]$
+\end{algorithmic}
+
+In Phase 2 we get rid of the arrows on statements 3, 4, and 6 as they are not vectorizable and in Phase 3 we add SIMD appropriately:
+
+\begin{algorithmic}
+\STATE 1. $\texttt{A}[\stackrel{\rightarrow}{i}] = \mathit{ADD\_SIMD}(\texttt{B}[\stackrel{\rightarrow}i],[10,...])$ \COMMENT{Fully vectorized size N.} 
+\STATE FOR i=0; i<N; i++; \COMMENT{ MOTION loop }
+\STATE 2. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$ 
+\STATE 3. $\texttt{B}[i] = \mathit{MUL}(\texttt{A}[i], \texttt{D}_0[{i-1}])$
+\STATE 4. $\texttt{C}[i] = \mathit{MUL}(\texttt{A}[i], \texttt{D}_0[{i-1}])$
+\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$
+\STATE 6. $\texttt{D}_1[i] = \mathit{MUL}(\texttt{B}[i], \texttt{C}[i])$
+\end{algorithmic}
+
+\paragraph{Example 2}
+
+Consider MPC-source of Histrogram after removal of infeasible edges and MUX refinement (copied from Repo with some modifications):
+
+\begin{algorithmic}
+
+\STATE res!1 = []
+\STATE for i: plaintext in range(0, num\_bins):
+\STATE \hspace{0.25cm} res!2 = $\phi$(res!1, res!3)
+\STATE  \hspace{0.25cm} ... 
+\STATE res!4 = res!2 \COMMENT{Added due to the "mixed" def-use edge.} 
+\STATE for i: plaintext in range(0, num\_bins):
+%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
+\STATE \hspace{0.25cm}        for j: plaintext in range(0, N):
+\STATE \hspace{0.5cm}            res!5 = $\phi$(res!4, res!7)
+\STATE \hspace{0.5cm}            !2!3 = (A[j] == i)
+\STATE \hspace{0.5cm}            !3!3 = (res!5[i] + B[j])
+\STATE \hspace{0.5cm}            res!6 = res!5
+\STATE \hspace{0.5cm}            res!6[i] = !3!3
+\STATE \hspace{0.5cm}            res!7[i] = MUX(!2!3, res!6[i], res!5[i])
+
+\end{algorithmic}
+
+After Phase 1 of Basic Vectorization:
+
+\begin{algorithmic}
+\STATE res!1 = []
+\STATE for i: plaintext in range(0, num\_bins):
+\STATE \hspace{0.25cm} res!2 = $\phi$(res!1, res!3)
+\STATE  \hspace{0.25cm} ... 
+\STATE res!4 = res!2 \COMMENT{Added due to the "mixed" def-use edge.} 
+\STATE A[$i,j$] = $\mathit{raise\_dim}$(A[$j$],$i$) \COMMENT{Copying row A num\_bin times. Reusing name.} 
+\STATE B[$i,j$] = $\mathit{raise\_dim}$(B[$i$],$j$) \COMMENT{Copying column B N times. Reusing name.} 
+\STATE res!4[$i,j$] = $\mathit{raise\_dim}$(res!4[$i$],$j$) \COMMENT{Reusing name.} 
+\STATE for i: plaintext in range(0, num\_bins):
+%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
+\STATE \hspace{0.25cm}        for j: plaintext in range(0, N):
+\STATE \hspace{0.5cm}            res!5 = $\phi$(res!4, res!7) \COMMENT{Do nothing. res!4 has been raised to proper dimension.}
+\STATE \hspace{0.5cm}            !2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = EQ(A[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], [[0,0...], [1,1...],...[num\_bins-1,num\_bins-1...])
+\STATE \hspace{0.5cm}            !3!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = ADD(res!5[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$],B[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$])
+\STATE \hspace{0.5cm}            res!6 = res!5 \COMMENT{Do nothing, just copy.}
+\STATE \hspace{0.5cm}            res!6[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = !3!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$]
+\STATE \hspace{0.5cm}            res!7[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = MUX(!2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], res!6[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], res!5[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$])
+\end{algorithmic}
+
+After Phase 2 and Phase 3. The EQ operation is vectorizable across both dimensions, and the rest of the computation is vectorizable across the $i$-dimension.
+
+\begin{algorithmic}
+\STATE ...
+\STATE !2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = EQ(A[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], [[0,0...], [1,1...],...[num\_bins-1,num\_bins-1...])
+%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
+\STATE FOR j=0; j<N; j++ \COMMENT{MOTION loop}
+\STATE \hspace{0.25cm}            res!5 = $\phi$(res!4, res!7) \COMMENT{Do nothing, just copy.}
+\STATE \hspace{0.25cm}            !3!3[$\stackrel{\rightarrow}i,j$] = $\mathit{ADD\_SIMD}$(res!5[$\stackrel{\rightarrow}i,j$],B[$\stackrel{\rightarrow}i,j$])
+\STATE \hspace{0.25cm}            res!6 = res!5 \COMMENT{Do nothing, just copy.}
+\STATE \hspace{0.25cm}            res!6[$\stackrel{\rightarrow}i,j$] = !3!3[$\stackrel{\rightarrow}i,j$]
+\STATE \hspace{0.25cm}            res!7[$\stackrel{\rightarrow}i,j$] = MUX(!2!3[$\stackrel{\rightarrow}i,j$], res!6[$\stackrel{\rightarrow}i,j$], res!5[$\stackrel{\rightarrow}i,j$])
+\end{algorithmic}
 

### 2021-12-22 paper_SIMD.tex
Fixed typos in extension to basic vectorization.
Draft 1 "Extension of Basic Vectorization with Array Writes".

@@
 \STATE replace $\mathit{stmt}$ with \\
-\STATE $\texttt{A}_j = \texttt{A}_k$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
+\STATE $\texttt{A}_j = \texttt{A}_{j-1}$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
 %\ELSIF {$i_1 \ncong i_2$}  
@@
 
-The update $\texttt{A}[i,j]$ can be nested into an inner loop and there may be multiple updates.
-However, updates such as $\texttt{A}[i-1,j] = ...$ or $\texttt{A}[i-1,j-1] = ...$, etc., are not allowed.
-This simplifies reasoning about correctness in this early stage of the compiler; we will look to relax this restriction in future work.
+The update $\texttt{A}[i,j]$ can be nested into an inner loop and there may be multiple updates, i.e., writes to $\texttt{A}[i,j]$.
+However, update such as $\texttt{A}[i-1,j] = ...$ or $\texttt{A}[i-1,j-1] = ...$, etc., is not allowed.
+Additionally, while there could be several different loops that perform canonical updates, they must be of the same dimensionality, i.e., an update of higher or lower dimension, e.g., $A[i,j,k] = ...$ is not allowed.
+We assign the \emph{canonical dimensionality} to each array in the obvious way.
+This restriction simplifies reasoning in this early stage of the compiler; we will look to relax the restriction in future work.
+
 Reads through an arbitrary formula, such as $\texttt{A}[i-1]$ for example, are allowed, however, we assume the programmer ensures that the formula is within the bounds of the array.
@@
 One change to Basic vectorization is the expansion of dimension if the array write or read occurs in a nested loop.
-That is, if there is an update to $\texttt{A}[i,j]$ that occurs in loop nest $i,j,k$, $\texttt{A}[i,j]$ will be rewritten into $\texttt{A}[i,j,k]$.
+That is, if there is an update $\texttt{A}[i,j] = ...$ that occurs in loop nest $i,j,k$, $\texttt{A}[i,j]$ will be rewritten into $\texttt{A}[i,j,k]$.
 Similarly, a read $\texttt{A}[f(i,j)]$ will be rewritten into $\texttt{A}[f(i,j),k]$.
-\ana{This notion of $f(i,j)$ needs a precise definition.
-For now I'm hoping the hand waving works.}.
-This can be a preprocessing step or it may be incorporated into the Basic Vectorization algorithm.
-Basic Vectorization will not expand array accesses at array writes the way it does with scalars now, as they are already expanded.
+\ana{TODO: This notion of $f(i,j)$ needs a precise definition.
+For now I'm hoping the hand waving works.} This can be a preprocessing step or it can be incorporated into the Basic Vectorization algorithm.
 
-\ana{TODO: Another change. What about the A = phi(A1,A2), and A = A1. We have to expand on what happens with those.}
+The other change concerns def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable (e.g., the definition can be a propagation $A_2 = A_1$ or a pseudo $\phi$-node $A_2 = \phi(A_0,A_1)$).
+These edges are not handled in the same way as in Basic Vectorization, specifically, we do not expand dimension as we do for scalars in Basic Vectorization according to the loop nest.
+A key invariant is that the dimension of an array $A$ cannot drop below $A$'s canonical dimensionality.
+Below we enumerate the cases of def-use edges.
 
-\ana{TODO: Third change. "Mixed" outer -> edge with arrays!}
+\begin{enumerate}
+
+\item same-level $X \rightarrow Y$.
+We do nothing, just propagate the array, which happens to be of the right dimension.
+\ana{There might be some opportunities to do copy propagation optimization and save some cycles, but let's leave this for later.}
+
+\item inner-to-outer $X \rightarrow Y$.
+Let $A$ be the array defined at $X$.
+If the dimensionality $A$ is greater than its canonical dimensionality, then add \emph{drop\_dim(...)} at $Y$, as in Basic Vectorization.
+Otherwise, do nothing.
+
+\item outer-to-inner $X \rightarrow Y$. Add \emph{raise\_dim(...)} (at $X$) as in Basic Vectorization.
+
+\item "mixed" $X \rightarrow Y$. We assume that the mixed edge is transformed into an inner-to-outer followed by outer-to-inner edge before we perform vectorization, just as with Basic vectorization.
+
+\end{enumerate}
 
@@
 \begin{algorithmic}
-\STATE 1. $\texttt{A}[\stackrel{\rightarrow}{i}] = \mathit{ADD\_SIMD}(\texttt{B}[\stackrel{\rightarrow}i],[10,...])$ \COMMENT{Fully vectorized size N.} 
+\STATE 1. $\texttt{A}[\stackrel{\rightarrow}{i}] = \mathit{ADD\_SIMD}(\texttt{B}[\stackrel{\rightarrow}i],[10,...])$ \COMMENT{Fully vectorized, size N.} 
 \STATE FOR i=0; i<N; i++; \COMMENT{ MOTION loop }
@@
 
-Consider MPC-source of Histrogram after removal of infeasible edges and MUX refinement (copied from Repo with some modifications):
+Consider MPC-source of Histogram after removal of infeasible edges and MUX refinement:
 
@@
 \STATE \hspace{0.5cm}            res!6[i] = !3!3 
+\STATE \hspace{0.5cm}            res!7 = res!6 
 \STATE \hspace{0.5cm}            res!7[i] = MUX(!2!3, res!6[i], res!5[i])
@@
 \STATE  \hspace{0.25cm} ... 
-\STATE res!4 = res!2 \COMMENT{Added due to the "mixed" def-use edge.} 
+\STATE res!4 = res!2 \COMMENT{Added due to the "mixed" def-use edge. Because the canonical dimensionality of res is 1, we DO NOT drop dimension of res!2.} 
 \STATE A[$i,j$] = $\mathit{raise\_dim}$(A[$j$],$i$) \COMMENT{Copying row A num\_bin times. Reusing name.} 
@@
 \STATE \hspace{0.5cm}            res!5 = $\phi$(res!4, res!7) \COMMENT{Do nothing. res!4 has been raised to proper dimension.}
-\STATE \hspace{0.5cm}            !2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = EQ(A[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], [[0,0...], [1,1...],...[num\_bins-1,num\_bins-1...])
-\STATE \hspace{0.5cm}            !3!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = ADD(res!5[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$],B[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$])
+\STATE \hspace{0.5cm}            !2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{EQ}$(A[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], [[0,0...], [1,1...],...[num\_bins-1,num\_bins-1...]])
+\STATE \hspace{0.5cm}            !3!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{ADD}$(res!5[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$],B[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$])
 \STATE \hspace{0.5cm}            res!6 = res!5 \COMMENT{Do nothing, just copy.}
 \STATE \hspace{0.5cm}            res!6[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = !3!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$]
-\STATE \hspace{0.5cm}            res!7[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = MUX(!2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], res!6[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], res!5[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$])
+\STATE \hspace{0.5cm}            res!7 = res!6
+\STATE \hspace{0.5cm}            res!7[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{MUX}$(!2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], res!6[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], res!5[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$])
 \end{algorithmic}
@@
 \STATE ...
-\STATE !2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = EQ(A[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], [[0,0...], [1,1...],...[num\_bins-1,num\_bins-1...])
+\STATE !2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{EQ\_SIMD}$(A[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], [[0,0...], [1,1...],...[num\_bins-1,num\_bins-1...]])
 %\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE FOR j=0; j<N; j++ \COMMENT{MOTION loop}
+\STATE FOR j=0; j<N; j++; \COMMENT{MOTION loop}
 \STATE \hspace{0.25cm}            res!5 = $\phi$(res!4, res!7) \COMMENT{Do nothing, just copy.}
@@
 \STATE \hspace{0.25cm}            res!6[$\stackrel{\rightarrow}i,j$] = !3!3[$\stackrel{\rightarrow}i,j$]
-\STATE \hspace{0.25cm}            res!7[$\stackrel{\rightarrow}i,j$] = MUX(!2!3[$\stackrel{\rightarrow}i,j$], res!6[$\stackrel{\rightarrow}i,j$], res!5[$\stackrel{\rightarrow}i,j$])
+\STATE \hspace{0.25cm}            res!7 = res!6
+\STATE \hspace{0.25cm}            res!7[$\stackrel{\rightarrow}i,j$] = $\mathit{MUX\_SIMD}$(!2!3[$\stackrel{\rightarrow}i,j$], res!6[$\stackrel{\rightarrow}i,j$], res!5[$\stackrel{\rightarrow}i,j$])
 \end{algorithmic}

### 2021-12-27 paper_SIMD.tex
add comments on assumptions

@@
 \begin{enumerate}
-\item All statements in the program execute using the same protocol (sharing). That is, there is no share conversion.
-\item All MPC instructions have the same unit cost, 1 unit.
-\item There is unlimited bandwidth---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely 1 unit.
+\item All statements in the program execute using the same protocol (sharing).
+That is, there is no share conversion.
+\ishaq{We should say somewhere (in a footnote?) that we know this is not ideal, otherwise reviewers will complain.}
+\item All MPC instructions have the same unit cost, 1 unit. \ishaq{Don't we need to distinguish between local (cheap) vs. interactive (expensive) instructions? I don't see how we could assume this.}
+\item There is unlimited bandwidth---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely 1 unit.\ishaq{This is fair for theoretical model, although in some real life applications this is not true, e.g.
+L2 norm of large vectors.}
 \item MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption, 
 2 MUL instructions scheduled in parallel benefit from amortization and cost 1, however a MUL and a MUX instructions scheduled 
-in parallel still cost 2. 
+in parallel still cost 2. \ishaq{This is not true, can we still assume it?} 
 \end{enumerate}

### 2021-12-31 paper_SIMD.tex
minor edit about Vassilis' suggestion about cost
minor edit
Minor edits to Extension of Basic Vectorization

@@
 \ishaq{We should say somewhere (in a footnote?) that we know this is not ideal, otherwise reviewers will complain.}
-\item All MPC instructions have the same unit cost, 1 unit. \ishaq{Don't we need to distinguish between local (cheap) vs. interactive (expensive) instructions? I don't see how we could assume this.}
+\item All MPC instructions have the same unit cost, 1 unit.
+\ishaq{Don't we need to distinguish between local (cheap) vs.
+interactive (expensive) instructions?
+I don't see how we could assume this.
+-- Past suggestion from Vassilis: Ideally we should assume cost to be a convex function.}
 \item There is unlimited bandwidth---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely 1 unit.\ishaq{This is fair for theoretical model, although in some real life applications this is not true, e.g.
@@
 2 MUL instructions scheduled in parallel benefit from amortization and cost 1, however a MUL and a MUX instructions scheduled 
-in parallel still cost 2. \ishaq{This is not true, can we still assume it?} 
+in parallel still cost 2. \ishaq{This is not true, can we still assume it? -- YES.} 
 \end{enumerate}
@@
 
-Array writes limit vectorization as they introduce sometimes infeasible loop-carried dependencies. Consider the following example: \ana{Have to add citation to Aiken's paper}
+Array writes limit vectorization as they sometimes introduce infeasible loop-carried dependencies. Consider the following example: \ana{Have to add citation to Aiken's paper}
 
@@
 \STATE 4. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$ 
-\STATE 5. $\texttt{A}_1$ = $\texttt{A}_0$ 
-\STATE 6. $\texttt{A}_1$[i] = $\texttt{B}_0$[i] + 10;
-\STATE 7. $\texttt{B}_1$ = $\texttt{B}_0$ 
-\STATE 8. $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 9. $\texttt{C}_1$ = $\texttt{C}_0$ 
-\STATE 10. $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 11. $\texttt{D}_1$ = $\texttt{D}_0$ 
-\STATE 12. $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
+\STATE 5. $\texttt{A}_1$ = $\texttt{A}_0$; $\texttt{A}_1$[i] = $\texttt{B}_0$[i] + 10; \COMMENT{equiv. to Update}
+\STATE 6. $\texttt{B}_1$ = $\texttt{B}_0$; $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
+\STATE 7. $\texttt{C}_1$ = $\texttt{C}_0$; $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
+\STATE 8. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
 \end{algorithmic}
@@
 
-\ana{TODO: We may remove phi-nodes that end up without incoming back-edges. This can free some cycles that we'd waste on bookkeeping arrays.}
+\ana{TODO: We need to remove phi-nodes that end up without incoming back-edges.}
+
+The final step is to remove array phi-nodes that are left without incoming back edges after removal of infeasible edges.
+Clearly, such a phi-node is redundant when there are no loop-carried dependences that are due to writes to the corresponding array.
+In our running example, the phi-nodes for \texttt{A}, \texttt{B}, and \texttt{C}, are all redundant.
+
+Removing redundant phi-nodes is as follows. 
+Denote a redundant phi-node by $\texttt{A}_1 = \phi(\texttt{A}_0,\texttt{A}_k)$. $\texttt{A}_0$ is the array at the beginning of the loop
+and $\texttt{A}_k$ is the array at the end of an iteration (including the last iteration). For each outer-to-inner or same-level def-use edge 
+that originates at the node, replace $\texttt{A}_1$ at the use node by $\texttt{A}_0$. This is correct, i.e., it is guaranteed that the use 
+references $\texttt{A}_0$; had the use been referencing an array modified in an earlier iteration, the back-edge to the phi-node would
+not have been removed and the phi-node would not have been classified as redundant. 
+For each inner-to-outer def-use edge, replace $\texttt{A}_1$ by $\texttt{A}_k$. In this case, the use references the array
+that results from the computation in the loop.
+
+Our running example is transformed as follows. Removing the phi-nodes for \texttt{A}, \texttt{B}, and \texttt{C} transforms the 
+replaces the references to $\texttt{A}_0$, $\texttt{B}_0$, and $\texttt{C}_0$, with references to $\texttt{A}$, $\texttt{B}$, and $\texttt{C}$
+respectively, i.e., the arrays before the loop.
+
+\begin{algorithmic}
+\STATE for i in range(N):
+\STATE 1. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$ 
+\STATE 2. $\texttt{A}_1$ = $\texttt{A}$; $\texttt{A}_1$[i] = $\texttt{B}$[i] + 10;
+\STATE 3. $\texttt{B}_1$ = $\texttt{B}$; $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
+\STATE 4. $\texttt{C}_1$ = $\texttt{C}$; $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
+\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
+\end{algorithmic}
 
@@
 the size of those vectors. Note that this is a heuristic that handles a common case, but not all cases of array updates. 
-\ana{It may actually be handling \_all\_ cases, with the restriction to canonical updates.}
+\ana{TODO: Simplify this algorithm, taking into account the restriction to canonical updates. It should be handling \_all\_ cases.}
 
@@
 %\STATE $\texttt{A}_j = \texttt{A}_k$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
-\ELSE \STATE
+\ELSE 
 \STATE $\mathit{stmt}$ stays as is
@@
 Let $A$ be the array defined at $X$.
-If the dimensionality $A$ is greater than its canonical dimensionality, then add \emph{drop\_dim(...)} at $Y$, as in Basic Vectorization.
+If the dimensionality of $A$ is greater than its canonical dimensionality, then add \emph{drop\_dim(...)} at $Y$, as in Basic Vectorization.
 Otherwise, do nothing.
@@
 
-After removal of infeasible edges, the Aiken's array write example will be (roughly) as follows:
+Recall that after removal of infeasible edges and redundant phi-nodes, the Aiken's array write example will be (roughly) as follows:
 
@@
 \STATE 1. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$ 
-\STATE 2. $\texttt{A}$[i] = $\texttt{B}$[i] + 10
-\STATE 3. $\texttt{B}$[i] = $\texttt{A}$[i] * $\texttt{D}_0$[i-1]
-\STATE 4. $\texttt{C}$[i] = $\texttt{A}$[i] * $\texttt{D}_0$[i-1]
-\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$ 
-\STATE 6. $\texttt{D}_1$[i] = $\texttt{B}$[i] * $\texttt{C}$[i]
+\STATE 2. $\texttt{A}_1$ = $\texttt{A}$; $\texttt{A}_1$[i] = $\texttt{B}$[i] + 10;
+\STATE 3. $\texttt{B}_1$ = $\texttt{B}$; $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
+\STATE 4. $\texttt{C}_1$ = $\texttt{C}$; $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
+\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
 \end{algorithmic}
@@
 \STATE 1. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$ 
-\STATE 2. $\texttt{A}[\stackrel{\rightarrow}{i}] = \texttt{B}[\stackrel{\rightarrow}i] + [10,...]$
-\STATE 3. $\texttt{B}[\stackrel{\rightarrow}i] = \texttt{A}[\stackrel{\rightarrow}i] * \texttt{D}_0[\stackrel{\rightarrow}{i-1}]$
-\STATE 4. $\texttt{C}[\stackrel{\rightarrow}i] = \texttt{A}[\stackrel{\rightarrow}i] * \texttt{D}_0[\stackrel{\rightarrow}{i-1}]$
-\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$ 
-\STATE 6. $\texttt{D}_1[\stackrel{\rightarrow}i] = \texttt{B}[\stackrel{\rightarrow}i] * \texttt{C}[\stackrel{\rightarrow}i]$
+\STATE 2. $\texttt{A}_1$ = $\texttt{A}$; $\texttt{A}_1[\stackrel{\rightarrow}{i}] = \texttt{B}[\stackrel{\rightarrow}i] + [10,...]$
+\STATE 3. $\texttt{B}_1$ = $\texttt{B}$; $\texttt{B}_1[\stackrel{\rightarrow}i] = \texttt{A}_1[\stackrel{\rightarrow}i] * \texttt{D}_0[\stackrel{\rightarrow}{i-1}]$
+\STATE 4. $\texttt{C}_1$ = $\texttt{C}$; $\texttt{C}_1[\stackrel{\rightarrow}i] = \texttt{A}_1[\stackrel{\rightarrow}i] * \texttt{D}_0[\stackrel{\rightarrow}{i-1}]$
+\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1[\stackrel{\rightarrow}i] = \texttt{B}_1[\stackrel{\rightarrow}i] * \texttt{C}_1[\stackrel{\rightarrow}i]$
 \end{algorithmic}
 
-In Phase 2 we get rid of the arrows on statements 3, 4, and 6 as they are not vectorizable and in Phase 3 we add SIMD appropriately:
+In Phase 2 we get rid of the arrows on statements 3, 4, and 5 as they are not vectorizable and in Phase 3 we add SIMD appropriately:
 
 \begin{algorithmic}
-\STATE 1. $\texttt{A}[\stackrel{\rightarrow}{i}] = \mathit{ADD\_SIMD}(\texttt{B}[\stackrel{\rightarrow}i],[10,...])$ \COMMENT{Fully vectorized, size N.} 
+\STATE 1. $\texttt{A}_1$ = $\texttt{A}$; $\texttt{A}_1[\stackrel{\rightarrow}{i}] = \mathit{ADD\_SIMD}(\texttt{B}[\stackrel{\rightarrow}i],[10,...])$ \COMMENT{Fully vectorized, size N.} 
 \STATE FOR i=0; i<N; i++; \COMMENT{ MOTION loop }
 \STATE 2. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$ 
-\STATE 3. $\texttt{B}[i] = \mathit{MUL}(\texttt{A}[i], \texttt{D}_0[{i-1}])$
-\STATE 4. $\texttt{C}[i] = \mathit{MUL}(\texttt{A}[i], \texttt{D}_0[{i-1}])$
-\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$
-\STATE 6. $\texttt{D}_1[i] = \mathit{MUL}(\texttt{B}[i], \texttt{C}[i])$
+\STATE 3. $\texttt{B}_1$ = $\texttt{B}$; $\texttt{B}_1[i] = \mathit{MUL}(\texttt{A}_1[i], \texttt{D}_0[{i-1}])$
+\STATE 4. $\texttt{C}_1$ = $\texttt{C}$; $\texttt{C}_1[i] = \mathit{MUL}(\texttt{A}_1[i], \texttt{D}_0[{i-1}])$
+\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1[i] = \mathit{MUL}(\texttt{B}_1[i], \texttt{C}_1[i])$
 \end{algorithmic}
@@
 
-Consider MPC-source of Histogram after removal of infeasible edges and MUX refinement:
+Consider MPC-source of Histogram after removal of infeasible edges and redundant phi-nodes, and MUX refinement:
 
@@
 \STATE \hspace{0.5cm}            !3!3 = (res!5[i] + B[j])
-\STATE \hspace{0.5cm}            res!6 = res!5
-\STATE \hspace{0.5cm}            res!6[i] = !3!3 
-\STATE \hspace{0.5cm}            res!7 = res!6 
-\STATE \hspace{0.5cm}            res!7[i] = MUX(!2!3, res!6[i], res!5[i])
+\STATE \hspace{0.5cm}            res!6 = res!5; res!6[i] = !3!3 
+\STATE \hspace{0.5cm}            res!7 = res!6; res!7[i] = MUX(!2!3, res!6[i], res!5[i])
 

### 2021-12-31 paper_SIMD.tex
Fixed some typos in Vectorization seciton.

@@
 \STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0 = \phi(\texttt{index}_1, -1)$} 
-\STATE {$\texttt{min}_0 = \phi(\texttt{min}_1, 10000)$}
+\STATE {$\texttt{index}_0 = \phi(-1, \texttt{index}_1)$} 
+\STATE {$\texttt{min}_0 = \phi(10000,\texttt{min}_1)$}
 
 \STATE \COMMENT {Begin of inner loop $j$.}
-\STATE {$\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$}
+\STATE {$\texttt{sum}_0 = \phi(0, \texttt{sum}_1)$}
 \STATE {$\texttt{d} = \mathit{SUB}(\texttt{S}[i,j], \texttt{C}[j])$}
@@
 \STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0[i] = \phi(\texttt{index}_1[i-1], -1)$} 
-\STATE {$\texttt{min}_0[i] = \phi(\texttt{min}_1[i-1], 10000)$}
+\STATE {$\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$} 
+\STATE {$\texttt{min}_0[i] = \phi(10000, \texttt{min}_1[i-1])$}
 
 \STATE \COMMENT {Begin of inner loop $j$.}
-\STATE {$\texttt{sum}_0[i,j] = \phi(\texttt{sum}_1[i,j-1], [0,0,...])$}
+\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...], \texttt{sum}_1[i,j-1])$}
 \STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
@@
 
-Starting at the inner phi-node $\texttt{sum}_0[i,j] = \phi(\texttt{sum}_1[i,j], [0,0,...])$, the algorithm first computes its closure.
+Starting at the inner phi-node $\texttt{sum}_0[i,j] = \phi([0,0,...],\texttt{sum}_1[i,j-1])$, the algorithm first computes its closure.
 The closure amounts to the phi-node itself and $\texttt{sum}_1[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$, accounting for the loop-carried dependency of the computation of \texttt{sum}.
@@
 \STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0[i] = \phi(\texttt{index}_1[i-1], -1)$} 
-\STATE {$\texttt{min}_0[i] = \phi(\texttt{min}_1[i-1], 10000)$}
+\STATE {$\texttt{index}_0[i] = \phi(-1, \texttt{index}_1[i-1])$} 
+\STATE {$\texttt{min}_0[i] = \phi(10000, \texttt{min}_1[i-1])$}
 
@@
 \STATE \COMMENT {Begin of inner loop $j$. Loop is now much "shorter"!}
-\STATE {$\texttt{sum}_0[i,j] = \phi(\texttt{sum}_1[i,j], [0,0,...])$}
-\STATE {FOR j=0; j<=D; j++} \COMMENT{ Will turn into a MOTION loop.}
+\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...], \texttt{sum}_1[i,j-1])$}
+\STATE {FOR j=0; j<D; j++} \COMMENT{ Will turn into a MOTION loop.}
 \STATE \hspace{0.25cm}{$\texttt{sum}_1[\stackrel{\rightarrow}{i},{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},{j}],\texttt{p}[\stackrel{\rightarrow}{i},{j}])$}
@@
 
-Next, analysis moves to outer dimension $i$. There are two phi-nodes, $\texttt{min}_0[i] = \phi(\texttt{min}_1[i-1], 10000)$ and $\texttt{index}_0[i] = \phi(\texttt{index}_1[i-1], -1)$.
+Next, analysis moves to outer dimension $i$. There are two phi-nodes, $\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1])$ and $\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$.
 The closure of the first is 
@@
 \begin{array}{l}
-\texttt{min}_0[i] = \phi(\texttt{min}_1[i-1], 10000) \\
+\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1]) \\
 \texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
@@
 \begin{array}{l}
-\texttt{index}_0[i] = \phi(\texttt{index}_1[i-1], -1), \\ 
+\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1]), \\ 
 %\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
@@
 \STATE {} 
-\STATE {$\texttt{sum}_0[i,j] = \phi(\texttt{sum}_1[i,j-1], [0,0,...])$}
-\STATE {FOR j=0; j<=D; j++} \COMMENT{ Turns into a MOTION loop, was $j$-loop.}
+\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...],\texttt{sum}_1[i,j-1])$}
+\STATE {FOR j=0; j<D; j++} \COMMENT{ Turns into a MOTION loop, was $j$-loop.}
 \STATE \hspace{0.25cm}{$\texttt{sum}_1[\stackrel{\rightarrow}{i},{j}] = \mathit{ADD\_SIMD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},{j}],\texttt{p}[\stackrel{\rightarrow}{i},{j}])$}
@@
 \STATE {} 
-\STATE {$\texttt{min}_0[i] = \phi(\texttt{min}_1[i-1], 10000)$}
-\STATE {FOR i=0; i<=N; i++} \COMMENT{ Turns into a MOTION loop, was $i$-loop.}
+\STATE {$\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1])$}
+\STATE {FOR i=0; i<N; i++} \COMMENT{ Turns into a MOTION loop, was $i$-loop.}
 \STATE \hspace{0.25cm}{$\texttt{c}[{i}] = \mathit{CMP}(\texttt{sum}'_0[{i}],\texttt{min}_0[{i}])$}
@@
 \STATE {}
-\STATE {$\texttt{index}_0[i] = \phi(\texttt{index}_1[i-1], -1)$} 
-\STATE {FOR i=0; i<=N; i++} \COMMENT{ Turns into a MOTION loop, was $i$-loop.}
+\STATE {$\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$} 
+\STATE {FOR i=0; i<N; i++} \COMMENT{ Turns into a MOTION loop, was $i$-loop.}
 \STATE \hspace{0.25cm}{$\texttt{index}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{index}_0[{i}],i)$}

### 2021-12-31 compiler/compiler/ast_shared.py
Make Var class more type safe

@@ -46,13 +46,16 @@ PLAINTEXT_INT = VarType(VarVisibility.PLAINTEXT, 0)
 class Var:
     """A variable named `name`"""
 
-    name: str
+    # This is a string for user-provided variables,
+    # and an integer for temporary variables the compiler automatically generates
+    name: Union[str, int]
+
+    rename_subscript: Optional[int] = None
 
     def __str__(self) -> str:
-        return self.name
-
-    def name_without_ssa_rename(self) -> str:
-        return self.name.strip("!")[0]
+        name = self.name if isinstance(self.name, str) else f"!{self.name}"
+        subscript = "" if self.rename_subscript is None else f"!{self.rename_subscript}"
+        return name + subscript
 
 
 @dataclass

### 2021-12-31 compiler/compiler/restricted_ast_to_tac_cfg.py

@@ -24,7 +24,7 @@ class _CFGBuilder:
 
     def generate_variable(self) -> tac_cfg.Var:
         self._tmp_name_counter += 1
-        return tac_cfg.Var(name=f"!{self._tmp_name_counter}")
+        return tac_cfg.Var(self._tmp_name_counter)
 
     def make_empty_block(self) -> tac_cfg.Block:
         block = tac_cfg.Block(assignments=[], terminator=None, merge_condition=None)

### 2021-12-31 compiler/compiler/tac_cfg_to_ssa.py

@@ -125,7 +125,7 @@ def rename_variables(result: ssa.Function) -> None:
             except KeyError:
                 return V
 
-        return ssa.Var(f"{V.name}!{i}")
+        return ssa.Var(V.name, i)
 
     def rename_subscript_index(index: ssa.SubscriptIndex) -> ssa.SubscriptIndex:
         if isinstance(index, ssa.Var):

### 2021-12-31 compiler/compiler/type_analysis.py

@@ -232,7 +232,7 @@ def type_check(func: loop_linear_code.Function, dep_graph: DepGraph) -> TypeEnv:
     type_env = TypeEnv(
         {
             # TODO: fix the below hack once parameter ssa renaming is properly implemented
-            Var(f"{param.var.name}!0"): cast(VarType, param.var_type)
+            Var(param.var.name, 0): cast(VarType, param.var_type)
             for param in func.parameters
         }
     )

### 2021-12-31 compiler/compiler/vectorize.py

@@ -3,6 +3,7 @@ from .dep_graph import DepGraph, PhiOrAssign
 from .util import assert_never
 
 from dataclasses import dataclass
+from typing import Union
 
 import z3  # type: ignore
 
@@ -31,11 +32,13 @@ def _arrays_written_in_loop(loop: llc.For) -> list[llc.Assign]:
     return result
 
 
-def _arrays_read_in_rhs(rhs: llc.AssignRHS, array_name: str) -> list[llc.Subscript]:
+def _arrays_read_in_rhs(
+    rhs: llc.AssignRHS, array_name: Union[str, int]
+) -> list[llc.Subscript]:
     if isinstance(rhs, llc.Var) or isinstance(rhs, llc.ConstantInt):
         return []
     elif isinstance(rhs, llc.Subscript):
-        if rhs.array.name_without_ssa_rename() == array_name:
+        if rhs.array.name == array_name:
             return [rhs]
         else:
             return []
@@ -64,7 +67,7 @@ def _arrays_read_in_rhs(rhs: llc.AssignRHS, array_name: str) -> list[llc.Subscri
 
 
 def _arrays_read_in_loop(
-    parent_loops: list[llc.For], array_name: str
+    parent_loops: list[llc.For], array_name: Union[str, int]
 ) -> list[ArrayRead]:
     result: list[ArrayRead] = []
     for statement in parent_loops[-1].body:
@@ -184,23 +187,23 @@ def _get_k(j: llc.For, A_use: ArrayRead) -> list[llc.For]:
     return k[1:]
 
 
-def _remove_back_edge(A_name: str, dep_graph: DepGraph, j: llc.For) -> None:
+def _remove_back_edge(A_name: Union[str, int], dep_graph: DepGraph, j: llc.For) -> None:
     """Remove back edge to Phi node in loop j"""
     for A_use in j.body:
-        if isinstance(A_use, llc.Phi) and A_use.lhs.name_without_ssa_rename() == A_name:
+        if isinstance(A_use, llc.Phi) and A_use.lhs.name == A_name:
             assert dep_graph.enclosing_loops[A_use][-1] == j
             A_defs: list[PhiOrAssign] = list(
                 dep_graph.def_use_graph.predecessors(A_use)
             )
             for A_def in A_defs:
-                assert A_def.lhs.name_without_ssa_rename() == A_name
+                assert A_def.lhs.name == A_name
                 if dep_graph.is_back_edge(A_def, A_use):
[…]

### 2022-01-01 compiler/compiler/__init__.py
Implement array MUX refinement

@@ -72,6 +72,14 @@ def main():
     print(dep_graph)
     print()
 
+    (linear, dep_graph) = vectorize.refine_array_mux(linear, dep_graph)
+    print("Array MUX refinement:")
+    print(linear)
+    print()
+    print("Array MUX refinement (dependence graph):")
+    print(dep_graph)
+    print()
+
     type_env = type_check(linear, dep_graph)
     print("Type environment:")
     print(type_env)

### 2022-01-01 compiler/compiler/vectorize.py

@@ -3,7 +3,7 @@ from .dep_graph import DepGraph, PhiOrAssign
 from .util import assert_never
 
 from dataclasses import dataclass
-from typing import Union
+from typing import Iterable, Union, Optional
 
 import z3  # type: ignore
 
@@ -12,6 +12,12 @@ import z3  # type: ignore
 class ArrayRead:
     subscript: llc.Subscript
     parent_loops: list[llc.For]
+
+
+def _new_temp_var(dep_graph: DepGraph):
+    nodes: Iterable[PhiOrAssign] = dep_graph.def_use_graph.nodes
+    number = max([0 if isinstance(a.lhs.name, str) else a.lhs.name for a in nodes]) + 1
+    return llc.Var(number, 0)
 
 
 def _arrays_written_in_loop(loop: llc.For) -> list[llc.Assign]:
@@ -229,3 +235,88 @@ def _prune_edges_from_statement(
 def remove_infeasible_edges(function: llc.Function, dep_graph: DepGraph) -> None:
     for statement in function.body:
         _prune_edges_from_statement([], statement, dep_graph)
+
+
+def _find_update(
+    A: llc.Var, A_use: llc.Assign, dep_graph: DepGraph
+) -> Optional[llc.SubscriptIndex]:
+    A_def: PhiOrAssign
+    [A_def] = [
+        A_def for A_def in dep_graph.def_use_graph.predecessors(A_use) if A_def.lhs == A
+    ]
+    return A_def.rhs.index if isinstance(A_def.rhs, llc.Update) else None
+
+
+def _refine_array_mux_statement(
+    statement: llc.Statement, dep_graph: DepGraph
+) -> list[llc.Statement]:
+    if isinstance(statement, llc.Assign) and isinstance(statement.rhs, llc.Mux):
+        Aj = statement.lhs
+        mux = statement.rhs
+        c = mux.condition
+        Ak = mux.false_value
+        Al = mux.true_value
+        if (
+            isinstance(Aj, llc.Var)
+            and isinstance(Ak, llc.Var)
+            and isinstance(Al, llc.Var)
+        ):
+            i1 = _find_update(Ak, statement, dep_graph)
+            i2 = _find_update(Al, statement, dep_graph)
+            if i1 == i2 and i1 is not None and i2 is not None:
+                name = Aj.name
+                assert Aj.name == Ak.name == Al.name
+                k = Ak.rename_subscript
+                l = Al.rename_subscript
[…]

### 2022-01-01 compiler/make_results_markdown.py

@@ -159,6 +159,17 @@ def main():
         md += "### Removal of infeasible edges\n"
         md += f"![]({filename})\n"
 
+        (loop_linear_code, dep_graph) = compiler.vectorize.refine_array_mux(
+            loop_linear_code, dep_graph
+        )
+        md += "### Array MUX refinement\n"
+        md += f"```python\n{loop_linear_code}\n```\n"
+        filename = f"{test_case_dir.name}_array_mux_refinement_dep_graph.png"
+        path = os.path.join(args.path, filename)
+        dep_graph_to_image(dep_graph, loop_linear_code, path)
+        md += "### Array MUX refinement (dependence graph)\n"
+        md += f"![]({filename})\n"
+
         type_env = compiler.type_check(loop_linear_code, dep_graph)
         md += "### Type environment\n"
         md += f"{type_env_to_table(type_env)}\n"

### 2022-01-01 compiler/tests/test_stages.py

@@ -48,6 +48,12 @@ class StagesTestCase(unittest.TestCase):
                 str(dep_graph), stages["dep_graph_remove_infeasible_edges.txt"]
             )
 
+            (loop_linear, dep_graph) = compiler.vectorize.refine_array_mux(
+                loop_linear, dep_graph
+            )
+            self.assertEqual(str(loop_linear), stages["refine_array_mux.txt"])
+            self.assertEqual(str(dep_graph), stages["refine_array_mux_dep_graph.txt"])
+
             type_env = compiler.type_check(loop_linear, dep_graph)
             self.assertEqual(str(type_env), stages["type_env.txt"])
 
@@ -90,6 +96,16 @@ def regenerate_stages():
         compiler.vectorize.remove_infeasible_edges(loop_linear, dep_graph)
         with open(
             os.path.join(test_case_dir, "dep_graph_remove_infeasible_edges.txt"), "w"
+        ) as f:
+            f.write(f"{dep_graph}\n")
+
+        (loop_linear, dep_graph) = compiler.vectorize.refine_array_mux(
+            loop_linear, dep_graph
+        )
+        with open(os.path.join(test_case_dir, "refine_array_mux.txt"), "w") as f:
+            f.write(f"{loop_linear}\n")
+        with open(
+            os.path.join(test_case_dir, "refine_array_mux_dep_graph.txt"), "w"
         ) as f:
             f.write(f"{dep_graph}\n")
 

### 2022-01-01 compiler/compiler/ast_shared.py
Change variable strings to use Unicode subscripts

@@ -54,7 +54,16 @@ class Var:
 
     def __str__(self) -> str:
         name = self.name if isinstance(self.name, str) else f"!{self.name}"
-        subscript = "" if self.rename_subscript is None else f"!{self.rename_subscript}"
+
+        # Convert subscript int to subscripted Unicode string
+        subscript = (
+            ""
+            if self.rename_subscript is None
+            else "".join(
+                [chr(0x2080 + int(digit)) for digit in str(self.rename_subscript)]
+            )
+        )
+
         return name + subscript
 
 

### 2022-01-01 compiler/compiler/ast_shared.py
Revert "Change variable strings to use Unicode subscripts"
This reverts commit c127a5e67a1a449e08427af7e6332a3797f2ba8a.

@@ -54,16 +54,7 @@ class Var:
 
     def __str__(self) -> str:
         name = self.name if isinstance(self.name, str) else f"!{self.name}"
-
-        # Convert subscript int to subscripted Unicode string
-        subscript = (
-            ""
-            if self.rename_subscript is None
-            else "".join(
-                [chr(0x2080 + int(digit)) for digit in str(self.rename_subscript)]
-            )
-        )
-
+        subscript = "" if self.rename_subscript is None else f"!{self.rename_subscript}"
         return name + subscript
 
 

### 2022-01-01 compiler/compiler/dead_code_elim.py
Refactor or'ed isinstance() calls

@@ -45,7 +45,7 @@ def _accessed_vars_assign_rhs(rhs: ssa.AssignRHS) -> list[ssa.Var]:
         )
     elif isinstance(rhs, ssa.UnaryOp):
         return _accessed_vars_assign_rhs(rhs.operand)
-    elif isinstance(rhs, ssa.List) or isinstance(rhs, ssa.Tuple):
+    elif isinstance(rhs, (ssa.List, ssa.Tuple)):
         return [var for item in rhs.items for var in _accessed_vars_assign_rhs(item)]
     elif isinstance(rhs, ssa.Mux):
         return (
@@ -99,7 +99,7 @@ def _compute_statements_setting_vars(
             assignments.append(block.terminator)
 
         for assignment in assignments:
-            if isinstance(assignment, ssa.Phi) or isinstance(assignment, ssa.Assign):
+            if isinstance(assignment, (ssa.Phi, ssa.Assign)):
                 var = assignment.lhs
             elif isinstance(assignment, ssa.For):
                 var = assignment.counter

### 2022-01-01 compiler/compiler/dep_graph.py

@@ -23,7 +23,7 @@ class DepGraph:
             statements: list[llc.Statement], enclosing_loops: list[llc.For]
         ):
             for statement in statements:
-                if isinstance(statement, llc.Phi) or isinstance(statement, llc.Assign):
+                if isinstance(statement, (llc.Phi, llc.Assign)):
                     all_assignments.append((statement, enclosing_loops))
                 elif isinstance(statement, llc.For):
                     loop = statement
@@ -109,7 +109,7 @@ def _vars_in_rhs(rhs: llc.AssignRHS) -> list[llc.Var]:
         return _vars_in_rhs(rhs.left) + _vars_in_rhs(rhs.right)
     elif isinstance(rhs, llc.UnaryOp):
         return _vars_in_rhs(rhs.operand)
-    elif isinstance(rhs, llc.List) or isinstance(rhs, llc.Tuple):
+    elif isinstance(rhs, (llc.List, llc.Tuple)):
         return [lhs for rhs_item in rhs.items for lhs in _vars_in_rhs(rhs_item)]
     elif isinstance(rhs, llc.Mux):
         return [

### 2022-01-01 compiler/compiler/restricted_ast_to_tac_cfg.py

@@ -121,9 +121,7 @@ def _build_expression_as_atom(
     expression: restricted_ast.Expression, builder: _CFGBuilder
 ) -> tac_cfg.Atom:
     assign_rhs = _build_expression_as_rhs(expression, builder)
-    if isinstance(assign_rhs, tac_cfg.Var) or isinstance(
-        assign_rhs, tac_cfg.ConstantInt
-    ):
+    if isinstance(assign_rhs, (tac_cfg.Var, tac_cfg.ConstantInt)):
         return assign_rhs
     else:
         var = builder.generate_variable()

### 2022-01-01 compiler/compiler/tac_cfg_to_ssa.py

@@ -247,7 +247,7 @@ def rename_variables(result: ssa.Function) -> None:
             else:
                 assert_never(A)
 
-            if isinstance(A, ssa.Phi) or isinstance(A, ssa.Assign):
+            if isinstance(A, (ssa.Phi, ssa.Assign)):
                 V = A.lhs
                 i = C[V]
                 old_lhs[A] = A.lhs

### 2022-01-01 compiler/compiler/vectorize.py

@@ -41,7 +41,7 @@ def _arrays_written_in_loop(loop: llc.For) -> list[llc.Assign]:
 def _arrays_read_in_rhs(
     rhs: llc.AssignRHS, array_name: Union[str, int]
 ) -> list[llc.Subscript]:
-    if isinstance(rhs, llc.Var) or isinstance(rhs, llc.ConstantInt):
+    if isinstance(rhs, (llc.Var, llc.ConstantInt)):
         return []
     elif isinstance(rhs, llc.Subscript):
         if rhs.array.name == array_name:
@@ -54,7 +54,7 @@ def _arrays_read_in_rhs(
         return left + right
     elif isinstance(rhs, llc.UnaryOp):
         return _arrays_read_in_rhs(rhs.operand, array_name)
-    elif isinstance(rhs, llc.List) or isinstance(rhs, llc.Tuple):
+    elif isinstance(rhs, (llc.List, llc.Tuple)):
         return [
             subscript
             for item in rhs.items

### 2022-01-01 compiler/make_results_markdown.py

@@ -36,7 +36,7 @@ def dep_graph_to_image(
 
         for statement in statements:
             result.append((indent, statement))
-            if isinstance(statement, llc.Phi) or isinstance(statement, llc.Assign):
+            if isinstance(statement, (llc.Phi, llc.Assign)):
                 pass
             elif isinstance(statement, llc.For):
                 result += search(statement.body, indent + 1)
@@ -49,7 +49,7 @@ def dep_graph_to_image(
 
     main_label = ""
     for i, (indent, statement) in enumerate(all_statements):
-        if isinstance(statement, llc.Phi) or isinstance(statement, llc.Assign):
+        if isinstance(statement, (llc.Phi, llc.Assign)):
             s = str(statement)
         elif isinstance(statement, llc.For):
             s = statement.heading_str()

### 2022-01-02 paper_SIMD.tex
Add note about pseudo-phi node placement
Add note to paper about biometric example

@@
 The interpretation of this edge is that the use node $Y$ uses the definition made in the last iteration of the inner loop.
+\ben{The current representation of pseudo $\phi$-nodes shows them attached to the loop header (i.e. in the loop).  We may want to clarify that they are also evaluated at the loop's termination.}
 \item same-level back-edge $X\rightarrow Y$.
@@
 \STATE \hspace{0.25cm}{$\texttt{index}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{index}_0[{i}],i)$}
+
+\ben{The above code implicitly stores the final values of min\_0 and index\_0 in their last indices, but there is no primitive like drop\_dim to make it clear.
+This will have to be added in the SSA code generation (though it might not be necessary to include in the paper?).
+It also is pretty wishy-washy about the size and layout of each vector (again more of an implementation issue), but maybe we should have a different syntax for array accesses for read/write (e.g.
+the ADD\_SIMD steps) vs the array definitions (e.g.
+the SUB\_SIMD and raise/drop\_dim statements)}
 

### 2022-01-06 paper_SIMD.tex
updated assumption comments

@@
 That is, there is no share conversion.
-\ishaq{We should say somewhere (in a footnote?) that we know this is not ideal, otherwise reviewers will complain.}
+\ishaq{This is not an assumption, this is our setting.
+We work in the single protocol world.}
 \item All MPC instructions have the same unit cost, 1 unit.
@@
 -- Past suggestion from Vassilis: Ideally we should assume cost to be a convex function.}
-\item There is unlimited bandwidth---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely 1 unit.\ishaq{This is fair for theoretical model, although in some real life applications this is not true, e.g.
-L2 norm of large vectors.}
+\item There is unlimited bandwidth---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely 1 unit.\ishaq{Comment from Vassilis: We assume there are infinite parallel capacities, this is the standard assumption in PRAM (Cryptographic Parallel RAM).}
 \item MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption,
 2 MUL instructions scheduled in parallel benefit from amortization and cost 1, however a MUL and a MUX instructions scheduled
-in parallel still cost 2. \ishaq{This is not true, can we still assume it? -- YES.}
+in parallel still cost 2.
+\ishaq{We need to reword this assumption to something like "$K$ parallel MUL costs much less than $K$ (because they get amortized), but any mix of $K$  MUL and MUX still cost roughly $K$.
+Specifically, we should take away the low constants (2 and 1) because we know for low constants this is not true.}
 \end{enumerate}

### 2022-01-12 paper_SIMD.tex
Commented out incorrect handling of targetless phi nodes. Added a TODO in Extension to Basic Vectorization.

@@
 
-\ana{TODO: We need to remove phi-nodes that end up without incoming back-edges.}
+\ana{Commented out previous writeup. Removal/handling of targetless phi-nodes will go into the Extension to Basic vectorization.}
 
+\begin{comment}
 The final step is to remove array phi-nodes that are left without incoming back edges after removal of infeasible edges.
-Clearly, such a phi-node is redundant when there are no loop-carried dependences that are due to writes to the corresponding array.
-In our running example, the phi-nodes for \texttt{A}, \texttt{B}, and \texttt{C}, are all redundant.
+Such a phi-node is redundant when there are no loop-carried dependences that are due to writes to the corresponding array.
+In our running example, the phi-nodes for \texttt{A}, \texttt{B}, and \texttt{C}, are redundant.
 
-Removing redundant phi-nodes is as follows.
+Removing redundant phi-nodes proceeds from \emph{innermost to outermost} loops as follows.
 Denote a redundant phi-node by $\texttt{A}_1 = \phi(\texttt{A}_0,\texttt{A}_k)$. $\texttt{A}_0$ is the array at the beginning of the loop
-and $\texttt{A}_k$ is the array at the end of an iteration (including the last iteration). For each outer-to-inner or same-level def-use edge
-that originates at the node, replace $\texttt{A}_1$ at the use node by $\texttt{A}_0$. This is correct, i.e., it is guaranteed that the use
+and $\texttt{A}_k$ is the array at the end of an iteration (including the last iteration). We consider two cases:
+\begin{enumerate}
+\item[Case 1] Inner-to-outer or same-level def-use edge
+\begin{itemize}
+\item If use is an array read $ ... = \texttt{A}_1[f(i)]$, then replace $\texttt{A}_1$ with $\texttt{A}_0$. This is correct, i.e., it is guaranteed that the use
 references $\texttt{A}_0$; had the use been referencing an array modified in an earlier iteration, the back-edge to the phi-node would
 not have been removed and the phi-node would not have been classified as redundant.
+\item If use is an array write (update) $\texttt{A}_2 = \texttt{A}_1; \texttt{A}_2[i] = ... $ equiv. to $\texttt{A}_2 = \mathit{update}(\texttt{A}_1, i, ...)$ then handling becomes more tricky.
+The semantics of this update is that when $i$ is 0, then $\texttt{A}_0$ is being read.
+But when $i > 0$, then $\texttt{A}_k$ is being read.
+If we are at this case, this means that the def-use edge is same-level and we are removing a redundant node in the loop enclosing the update.
+The removal of the original phi-node entails pushing the phi-node at the update:
+$\texttt{A}_2 = \phi(\texttt{A}_0,\texttt{A}_k); \texttt{A}_2[i] = ... $ equiv. to $\texttt{A}_2 = \mathit{update}(\phi(\texttt{A}_0,\texttt{A}_k), i, ...)$.
+\item If use is a phi-node $... = \phi(\texttt{A}_1,...)$, that means we have an outer-to-inner edge. The
+
+\end{itemize}
+\end{enumerate}
+For each outer-to-inner or same-level def-use edge
+that originates at the node, replace $\texttt{A}_1$ at the use node by $\texttt{A}_0$. 
 For each inner-to-outer def-use edge, replace $\texttt{A}_1$ by $\texttt{A}_k$. In this case, the use references the array
@@
 \end{algorithmic}
+\end{comment}
 
@@
 \subsection{Extension of Basic Vectorization with Array Writes}
+
+\ana{TODO: Handling of Targetless phi nodes should be done here, as an extension to Basic Vectorization.}
 

### 2022-01-12 paper_SIMD.tex
Edits to assumptions as discussed today in meeting.

@@
 I don't see how we could assume this.
--- Past suggestion from Vassilis: Ideally we should assume cost to be a convex function.}
-\item There is unlimited bandwidth---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely 1 unit.\ishaq{Comment from Vassilis: We assume there are infinite parallel capacities, this is the standard assumption in PRAM (Cryptographic Parallel RAM).}
+-- Past suggestion from Vassilis: Ideally we should assume cost to be a convex function.}\ana{After discussion 1/12: replace single costs 1 with two levels of costs: $\alpha$ for non-local operations, e.g., MUL, and $\beta$ for local ops, e.g., ADD.}
+\item There is unlimited bandwidth---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely 1 unit.\ishaq{Comment from Vassilis: We assume there are infinite parallel capacities, this is the standard assumption in PRAM (Cryptographic Parallel RAM).}\ana{PRAM assumption stays.
+We have to replace the unit cost of 1 with a suitable amortization function $f(n)$, which I don't think changes much in the cost modeling and analysis.}
 \item MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption,
@@
 \ishaq{We need to reword this assumption to something like "$K$ parallel MUL costs much less than $K$ (because they get amortized), but any mix of $K$  MUL and MUX still cost roughly $K$.
-Specifically, we should take away the low constants (2 and 1) because we know for low constants this is not true.}
+Specifically, we should take away the low constants (2 and 1) because we know for low constants this is not true.} \ana{Again, core assumption that MUL and MUX don't benefit from amortization stays.
+We have to replace constant costs with functions, as in (3).}
 \end{enumerate}

### 2022-01-22 compiler/compiler/motion_backend/reference_implementations.py
Add implementations of raise_dim and drop_dim

@@ -1,0 +1,99 @@
+def raise_dim(
+    arr: list[int], dims_in_arr: list[int], all_dims: list[int], dim_to_raise: int
+) -> list[int]:
+    """
+    NOTE: This implementation is a recreation of the C++ implementation, so it could probably
+    be made more pythonic.
+
+    Raises the dimensionality of `arr` by extending it over dimension `dim_to_raise`.
+    Example: raise_dim([0, 1, 2], [0], [3, 2], 1) will return [0, 0, 1, 1, 2, 2]
+    Example: raise_dim([0, 1, 2], [1], [2, 3], 0) will return [0, 1, 2, 0, 1, 2]
+
+    :param arr           Input array.  This array's elements should be ordered in row-major
+                           order according to `dims_in_arr`.
+    :param dims_in_arr   Order sequence of which dimensions are in `arr`.  Each number in
+                           this array refers to an index in `all_dims` corresponding to
+                           the relevant dimension. THIS ARRAY MUST BE SORTED IN ASCENDING
+                           ORDER.
+    :param all_dims      Array sizes of each dimsnion.  For example, the array [2, 1, 3]
+                           means that dimension 0 has size 2, dimension 1 has size 1, and
+                           dimension 2 has size 3.
+    :param dim_to_raise  Index of the dimension `all_dims` to extend `arr` over.  This index
+                           cannot be in `dims_in_arr`.
+
+    :returns The extended version of `arr`
+    """
+    raised_size = all_dims[dim_to_raise]
+    for dim_idx in dims_in_arr:
+        assert dim_idx != dim_to_raise
+        raised_size *= all_dims[dim_idx]
+
+    raised = [0] * raised_size
+
+    dims_in_raised = sorted([*dims_in_arr, dim_to_raise])
+
+    def RaisedArrIdx(idx_in_arr: int, idx_in_dim: int) -> int:
+        dim_idxs = [0] * len(dims_in_arr)
+        for dim_idx in reversed(range(len(dims_in_arr))):
+            dim_size = all_dims[dims_in_arr[dim_idx]]
+
+            dim_idxs[dim_idx] = idx_in_arr % dim_size
+            # Drop the most recent dimension from the array index
+            idx_in_arr //= dim_size
+
+        raised_idx = 0
+        dim_idx = 0
+        # Go over dimensions until we reach the one we're raising over
+        while (
[…]

### 2022-01-22 paper_SIMD.tex
Update writeup description of raise_dim() and drop_dim()

@@
 
-$\mathit{raise\_dim}(A[i_1,...,i_k],i_{k+1})$ is defined as follows.
-It results in a new {k+1}-dimensional array $A'$ where for every $0\le i_{k+1} < I_{k+1}$, $A'[i_1,...,i_k,i_{k+1}] = A[i_1,...,i_k]$.
-We add the new dimension at the end arbitrarily, it can be added at any position, e.g., before $i_1$, etc.
+$\mathit{raise\_dim}(A[i_1,...,i_k],i_j)$ is defined as follows.
+It results in a new {k+1}-dimensional array $A'$ where for every $0 \le i_{k} < I_{k+1}$, $A'[i_1, \ldots, i_j, \ldots, i_k] = A[i_1, \ldots, i_k]$.
 Adding $n$ dimensions is trivially extended as a composition of $n$ $\mathit{raise\_dim}$ that each adds a single dimension.
@@
 
-$\mathit{drop\_dim}(A[i_1,...,i_k,i_{k+1}])$ produces a k-dimensional array $A'$ where $A'[i_1,...,i_k] = A'[i_1,...,i_k,I_{k+1}-1]$.
+$\mathit{drop\_dim}(A[i_1,\ldots, i_j, \ldots,i_k], i_j)$ produces a k-dimensional array $A'$ where $A'[i_1,\ldots,i_k] = A'[i_1,\ldots, I_j-1, \ldots,i_k]$.
 In our analysis, dimensions are always dropped at the end, and again, one can define dropping $n$ dimensions as a composition of $n$ $\mathit{drop\_dim}$.
+
+\ben{I'm not sure if I wrote up the above paragraphs correctly -- the idea is that we're just copying values to extend over dimension $j$ or only retain its last elements.
+The current writeup implies that $1 < j < k$ though, which isn't always true.
+Python implementations of $\mathit{raise\_dim}()$ and $\mathit{drop\_dim}()$ can be found here: \url{https://github.com/milana2/ParallelizationForMPC/blob/master/compiler/compiler/motion_backend/reference_implementations.py}}
 

### 2022-01-22 compiler/make_results_markdown.py
Fix issues with markdown generation

@@ -11,7 +11,7 @@ import compiler
 import compiler.loop_linear_code as llc
 from compiler.type_analysis import TypeEnv
 from compiler.util import assert_never
-from tests import STAGES_DIR
+from tests.context import STAGES_DIR
 
 
 def cfg_to_image(G: networkx.DiGraph, path: str):

### 2022-01-30 paper_SIMD.tex
(Draft) extension of Basic vectorization for handling of targetless phi-nodes.

@@
 \STATE {compute $\mathit{closure}(n)$}
+\ENDFOR
+
+\COMMENT {{\color{red} $cl_1$ and $cl_2$ intersect if they have common statement, or update same array; "intersect" definition can be expanded}}
+
 \WHILE {there are closure $cl_1$ and $cl_2$ that intersect}
@@
 \STATE {add $\phi$-nodes in $cl$ to header block}
+\STATE {{\color{red}{add target-less $\phi$-node for A if $cl$ updates array A}}}
 \STATE {add statements in $cl$ to loop body in some order of dependences}
@@
 \ENDFOR
+{
+\color{red}
+\FOR {each target-less $\phi$-node $A_1 = \phi(A_0,A_k)$}
+\STATE {in vectorizable stmts, replace use of $A_1$ with $A_0$}
+\STATE {discard $\phi$-node if not used in any $cl$}
 \ENDFOR
+}
 \ENDFOR

### 2022-01-30 paper_SIMD.tex
Sec. 4.5 Work in progress.

@@
 
-\ana{Commented out previous writeup. Removal/handling of targetless phi-nodes will go into the Extension to Basic vectorization.}
+%\ana{Commented out previous writeup. Removal/handling of targetless phi-nodes will go into the Extension to Basic vectorization.}
 
@@
 
-\ana{TODO: Handling of Targetless phi nodes should be done here, as an extension to Basic Vectorization.}
+\ana{TODO: This section is work in progress!}
 
@@
 
-The other change concerns def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable (e.g., the definition can be a propagation $A_2 = A_1$ or a pseudo $\phi$-node $A_2 = \phi(A_0,A_1)$).
+The other change concerns def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable (e.g., the definition can be an update $A_2 = update(A_1,...)$ or a pseudo $\phi$-node $A_2 = \phi(A_0,A_1)$).
 These edges are not handled in the same way as in Basic Vectorization, specifically, we do not expand dimension as we do for scalars in Basic Vectorization according to the loop nest.
@@
 
-Recall that after removal of infeasible edges and redundant phi-nodes, the Aiken's array write example will be (roughly) as follows:
+Recall that after removal of infeasible edges, the Aiken's array write example will be (roughly) as follows:
 
@@
 \STATE for i in range(N):
-\STATE 1. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$
-\STATE 2. $\texttt{A}_1$ = $\texttt{A}$; $\texttt{A}_1$[i] = $\texttt{B}$[i] + 10;
-\STATE 3. $\texttt{B}_1$ = $\texttt{B}$; $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 4. $\texttt{C}_1$ = $\texttt{C}$; $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
+\STATE 1. $\texttt{A}_1$ = $\phi(\texttt{A}_0,\texttt{A}_2)$
+\STATE 2. $\texttt{B}_1$ = $\phi(\texttt{B}_0,\texttt{B}_2)$
+\STATE 3. $\texttt{C}_1$ = $\phi(\texttt{C}_0,\texttt{C}_2)$
+\STATE 4. $\texttt{D}_2$ = $\phi(\texttt{D}_0,\texttt{D}_2)$
+\STATE 5. $\texttt{A}_2$ = update($\texttt{A}_1$, i, $\texttt{B}_1$[i] + 10);
+\STATE 6. $\texttt{B}_2$ = update($\texttt{B}_1$, i, $\texttt{A}_2$[i] * $\texttt{D}_1$[i-1]);
+\STATE 7. $\texttt{C}_2$ = update($\texttt{C}_1$, i, $\texttt{A}_2$[i] * $\texttt{D}_1$[i-1]);
+\STATE 8. $\texttt{D}_2$ = update($\texttt{D}_1$; i, $\texttt{B}_2$[i] * $\texttt{C}_2$[i]);
 \end{algorithmic}
 
-There are no nested loops, thus, array accesses remain as is.
+There are no nested loops, thus, array accesses remain as is. The closure of 4 includes 4,6,7, and 8, 
+while 5 is vectorizable.
 
+Removal of infeasible edges leaves $\phi$-nodes 1,2, and 3, i.e., A,B, and C as targetless. 
 After Phase 1 of Basic vectorization we have the following code:

### 2022-01-30 paper_SIMD.tex
Lock Sec. 4.5. Updates on Sec. 4.5

@@
 
-Recall that after removal of infeasible edges, the Aiken's array write example will be (roughly) as follows:
+After Phase 1 of Basic Vectorization the Aiken's array write example will be (roughly) as follows:
 
@@
 \STATE 4. $\texttt{D}_2$ = $\phi(\texttt{D}_0,\texttt{D}_2)$
-\STATE 5. $\texttt{A}_2$ = update($\texttt{A}_1$, i, $\texttt{B}_1$[i] + 10);
-\STATE 6. $\texttt{B}_2$ = update($\texttt{B}_1$, i, $\texttt{A}_2$[i] * $\texttt{D}_1$[i-1]);
-\STATE 7. $\texttt{C}_2$ = update($\texttt{C}_1$, i, $\texttt{A}_2$[i] * $\texttt{D}_1$[i-1]);
-\STATE 8. $\texttt{D}_2$ = update($\texttt{D}_1$; i, $\texttt{B}_2$[i] * $\texttt{C}_2$[i]);
+\STATE 5. $\texttt{A}_2 = update(\texttt{A}_1, \stackrel{\rightarrow}{i}, \texttt{B}_1[\stackrel{\rightarrow}{i}] + 10)$;
+\STATE 6. $\texttt{B}_2 = update(\texttt{B}_1, \stackrel{\rightarrow}{i}, \texttt{A}_2[\stackrel{\rightarrow}{i}] * \texttt{D}_1[\stackrel{\rightarrow}{i-1}])$;
+\STATE 7. $\texttt{C}_2 = update(\texttt{C}_1, \stackrel{\rightarrow}{i}, \texttt{A}_2[\stackrel{\rightarrow}{i}] * \texttt{D}_1[\stackrel{\rightarrow}{i-1}])$;
+\STATE 8. $\texttt{D}_2 = update(\texttt{D}_1; \stackrel{\rightarrow}{i}, \texttt{B}_2[\stackrel{\rightarrow}{i}] * \texttt{C}_2[\stackrel{\rightarrow}{i}])$;
 \end{algorithmic}
 
-There are no nested loops, thus, array accesses remain as is. The closure of 4 includes 4,6,7, and 8, 
-while 5 is vectorizable.
-
-Removal of infeasible edges leaves $\phi$-nodes 1,2, and 3, i.e., A,B, and C as targetless. 
-After Phase 1 of Basic vectorization we have the following code:
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$
-\STATE 2. $\texttt{A}_1$ = $\texttt{A}$; $\texttt{A}_1[\stackrel{\rightarrow}{i}] = \texttt{B}[\stackrel{\rightarrow}i] + [10,...]$
-\STATE 3. $\texttt{B}_1$ = $\texttt{B}$; $\texttt{B}_1[\stackrel{\rightarrow}i] = \texttt{A}_1[\stackrel{\rightarrow}i] * \texttt{D}_0[\stackrel{\rightarrow}{i-1}]$
-\STATE 4. $\texttt{C}_1$ = $\texttt{C}$; $\texttt{C}_1[\stackrel{\rightarrow}i] = \texttt{A}_1[\stackrel{\rightarrow}i] * \texttt{D}_0[\stackrel{\rightarrow}{i-1}]$
-\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1[\stackrel{\rightarrow}i] = \texttt{B}_1[\stackrel{\rightarrow}i] * \texttt{C}_1[\stackrel{\rightarrow}i]$
-\end{algorithmic}
-
-In Phase 2 we get rid of the arrows on statements 3, 4, and 5 as they are not vectorizable and in Phase 3 we add SIMD appropriately:
+There are no nested loops, thus, array accesses remain as is. Phase 2 computes the closure of 4; $cl = \{4,6,7, 8\}$ 
+while 5 is vectorizable. Recall that 1,2, and 3 are targetless phi-nodes. Since the closure $cl$ includes updates to $B$ and $C$,
+the corresponding phi-nodes are added to the closure, i.e., the FOR loop. The uses of $A_1$ and $B_1$ in the vectorized statement
+turn into uses of $A_0$ and $B_0$ respectively. (But note that $A_0$ is irrelevant; the updates writes into array $A_2$ in parallel.)
 
 \begin{algorithmic}
-\STATE 1. $\texttt{A}_1$ = $\texttt{A}$; $\texttt{A}_1[\stackrel{\rightarrow}{i}] = \mathit{ADD\_SIMD}(\texttt{B}[\stackrel{\rightarrow}i],[10,...])$ \COMMENT{Fully vectorized, size N.}
+\STATE 1. $\texttt{A}_2 = update(\texttt{A}_0; \stackrel{\rightarrow}{i}, \mathit{ADD\_SIMD}(\texttt{B}_0[\stackrel{\rightarrow}i],[10,...])$ \COMMENT{Fully vectorized, size N.}
 \STATE FOR i=0; i<N; i++; \COMMENT{ MOTION loop }
-\STATE 2. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$
-\STATE 3. $\texttt{B}_1$ = $\texttt{B}$; $\texttt{B}_1[i] = \mathit{MUL}(\texttt{A}_1[i], \texttt{D}_0[{i-1}])$
-\STATE 4. $\texttt{C}_1$ = $\texttt{C}$; $\texttt{C}_1[i] = \mathit{MUL}(\texttt{A}_1[i], \texttt{D}_0[{i-1}])$
-\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1[i] = \mathit{MUL}(\texttt{B}_1[i], \texttt{C}_1[i])$
+\STATE 2. $\texttt{B}_1$ = $\phi(\texttt{B}_0,\texttt{B}_2)$
+\STATE 3. $\texttt{C}_1$ = $\phi(\texttt{C}_0,\texttt{C}_2)$
+\STATE 4. $\texttt{D}_1$ = $\phi(\texttt{D}_0,\texttt{D}_2)$
+\STATE 5. $\texttt{B}_2 = update(\texttt{B}_1, i, \mathit{MUL}(\texttt{A}_2[i], \texttt{D}_1[{i-1}])$
+\STATE 6. $\texttt{C}_2 = update(\texttt{C}_1, i, \mathit{MUL}(\texttt{A}_2[i], \texttt{D}_1[{i-1}])$
+\STATE 7. $\texttt{D}_2 = update(\texttt{D}_1, i, \mathit{MUL}(\texttt{B}_2[i], \texttt{C}_1[i])$
 \end{algorithmic}

### 2022-02-01 biometric_vectorization_example.tex
Changed biometric example to literal code.

%\documentclass{article}

%\begin{document}

\subsection{Example: biometric}

\subsubsection{Benjamin's code with linear loops (MPC Source)}

\begin{verbatim}
min_sum!1 = 10000
min_index!1 = 0
for i in range(0, N!0):
   min_sum!2 = PHI(min_sum!1, min_sum!4)
   min_index!2 = PHI(min_index!1, min_index!4)
   sum!2 = 0
   for j in range(0, D!0):
     sum!3 = PHI(sum!2, sum!4)
     d!3 = (S!0[((i * D!0) + j)] - C!0[j])
     p!3 = (d!3 * d!3)
     sum!4 = (sum!3 + p!3)
   !1!2 = (sum!3 < min_sum!2)
   min_sum!3 = sum!3
   min_index!3 = i
   min_sum!4 = MUX(!1!2, min_sum!3, min_sum!2)
   min_index!4 = MUX(!1!2, min_index!3, min_index!2)
\end{verbatim}

\subsubsection{Phase 1 of Basic Vectorization}

The transformation preserves the dependence edges. It raises the dimensions of scalars and optimistically vectorizes all operations. 
The next phase discovers loop-carried dependences and removes affected vectorization.

In the code below, all initializations (e.g., \texttt{min\_sum!3 = i}), operations, and PHI nodes are \emph{implicitly vectorized}. 
$\mathit{raise\_dim}$ and $\mathit{drop\_dim}$ statements, as well as propagation statements (e.g., \texttt{min\_sum!3 = sum!3\^}) 
may have slightly different interpretation. \ana{This is vague but I think it's clear.}

Note the two different versions of $\mathit{raise\_dim}$. \ana{More here! One just adds a least-significant dimension. The other one may reshape an input array.}

\begin{verbatim}
min_sum!1 = 10000
min_sum!1^ = raise_dim(min_sum!1, (i:N!0))
min_index!1 = 0
min_index!1^ = raise_dim(min_index!1, (i:N!0))
S!0^ = raise_dim(S!0, ((i * D!0) + j), (i:N!0,j:D!0))
C!0^ = raise_dim(C!0, j, (i:N!0,j:D!0))
for i in range(0, N!0):
   min_sum!2 = PHI(min_sum!1^, min_sum!4)
   min_index!2 = PHI(min_index!1^, min_index!4) 
   sum!2 = 0 // Will lift, when hoisted
   sum!2^ = raise_dim(sum!2, (j:D!0)) // Special form?
   for j in range(0, D!0):
      sum!3 = PHI(sum!2^, sum!4)
      d!3 = S!0^ - C!0^
      p!3 = (d!3 * d!3) 
      sum!4 = (sum!3 + p!3)
   sum!3^ = drop_dim(sum!3)     
   !1!2 = (sum!3^ < min_sum!2)
   min_sum!3 = sum!3^
   min_index!3 = i  // Same-level, will lift when hoisted
   min_sum!4 = MUX(!1!2, min_sum!3, min_sum!2)
   min_index!4 = MUX(!1!2, min_index!3, min_index!2)
\end{verbatim}

\subsubsection{Phase 2 of Basic Vectorization}

This phase analyzes statements from the innermost loop to the outermost. The key point is to discover loop-carried dependencies and re-introduce loops whenever dependencies make this necessary.

Starting at the inner phi-node \texttt{sum!3 = PHI(sum!2\^, sum!4)}, the algorithm first computes its closure. The closure amounts to the phi-node itself and the addition node \texttt{sum!4 = (sum!3 + p!3)}, accounting for the loop-carried dependency of the computation of \texttt{sum}. The algorithm replaces this closure with a FOR loop on $j$ removing vectoriaztion on $j$. Note that the SUB and MUL computations remain outside of the loop as they do not depend on phi-nodes that are part of cycles. The dependences are from \texttt{p!3[I,J] = (d!3[I,J] * d!3[I,J])} to the monolithic FOR loop and from the FOR loop to \texttt{sum!3\^ = drop\_dim(sum!3)}. (Lower case index, e.g., $i,j$ indicates non-vectorized dimension, while uppercase index, e.g., $I$ indicates vectorized dimenion.)

After processing inner loop code becomes:

\begin{verbatim}
min_sum!1 = 10000
min_sum!1^ = raise_dim(min_sum!1, (i:N!0))
min_index!1 = 0
min_index!1^ = raise_dim(min_index!1, (i:N!0))
S!0^ = raise_dim(S!0, ((i * D!0) + j), (i:N!0,j:D!0))
C!0^ = raise_dim(C!0, j, (i:N!0,,j:D!0))
for i in range(0, N!0):
  min_sum!2[I] = PHI(min_sum!1^[I], min_sum!4[I]) 
   min_index!2[I] = PHI(min_index!1^[I], min_index!4[I])  
  sum!2 = [0,..,0] 
  sum!2^ = raise_dim(sum!2, (j:D!0))
  d!3[I,J] = S!0^[I,J] - C!0^[I,J]
  p!3[I,J] = (d!3[I,J] * d!3[I,J])
  for j in range(0, D!0):
    sum!3[I,j] = PHI(sum!2^[I,j], sum!4[I,j-1])       
    sum!4[I,j] = (sum!3[I,j] + p!3[I,j])
  sum!3^ = drop_dim(sum!3)     
  !1!2[I] = (sum!3^[I] < min_sum!2[I])
  min_sum!3 = sum!3^
  min_index!3 = [0,1,2,...,N!0-1]
   min_sum!4[I] = MUX(!1!2[I], min_sum!3[I], min_sum!2[I])
   min_index!4[I] = MUX(!1!2[I], min_index!3[I], min_index!2[I])
\end{verbatim}

\noindent When processing the outer loop two closures arise, one for \texttt{min\_sum!2[I] = PHI(...)} and one 
for \texttt{min\_index!2[I] = PHI(...)}. Since the two closures \emph{do not} intersect, we have two distinct FOR-loops on $i$:

\begin{verbatim}
min_sum!1 = 10000
min_sum!1^ = raise_dim(min_sum!1, (i:N!0))
min_index!1 = 0
min_index!1^ = raise_dim(min_index!1, (i:N!0))
S!0^ = raise_dim(S!0, ((i * D!0) + j), (i:N!0,j:D!0))
C!0^ = raise_dim(C!0, j, (i:N!0))

sum!2 = [0,..,0]
sum!2^ = raise_dim(sum!2, (j:D!0))
d!3[I,J] = S!0^[I,J] - C!0^[I,J]
p!3[I,J] = (d!3[I,J] * d!3[I,J])

for j in range(0, D!0):
  sum!3[I,j] = PHI(sum!2^[I,j], sum!4[I,j-1])       
  sum!4[I,j] = (sum!3[I,j] + p!3[I,j])

sum!3^ = drop_dim(sum!3)     
min_index!3 = [1,2,...N!0]   
min_sum!3 = sum!3^

for i in range(0, N!0):
  min_sum!2[i] = PHI(min_sum!1^[i], min_sum!4[i-1]) 
  !1!2[i] = (sum!3^[i] < min_sum!2[i])
   min_sum!4[i] = MUX(!1!2[i], min_sum!3[i], min_sum!2[i])

for i in range(0, N!0):
  min_index!2[i] = PHI(min_index!1^[i], min_index!4[i-1])  
  min_index!4[i] = MUX(!1!2[i], min_index!3[i], min_index!2[i])

\end{verbatim}

\subsubsection{Phase 3 of Basic Vectorization}

And this phase adds SIMD operations:
\begin{verbatim}
min_sum!1 = 10000
min_sum!1^ = raise_dim(min_sum!1, (i:N!0))
min_index!1 = 0
min_index!1^ = raise_dim(min_index!1, (i:N!0))
S!0^ = raise_dim(S!0, ((i * D!0) + j), (i:N!0,j:D!0))
C!0^ = raise_dim(C!0, j, (i:N!0,j:D!0))

sum!2 = [0,..,0]
sum!2^ = raise_dim(sum!2, (j:D!0))
d!3[I,J] = SUB_SIMD(S!0^[I,J],C!0^[I,J])
p!3[I,J] = MUL_SIMD(d!3[I,J] * d!3[I,J])

for j in range(0, D!0):
  sum!3[I,j] = PHI(sum!2^[I,j], sum!4[I,j-1])       
  sum!4[I,j] = ADD_SIMD(sum!3[I,j],p!3[I,j])

sum!3^ = drop_dim(sum!3)     
min_index!3 = [0,1,...N!0-1]   
min_sum!3 = sum!3^

for i in range(0, N!0):
  min_sum!2[i] = PHI(min_sum!1^[i], min_sum!4[i-1]) 
  !1!2[i] = CMP(sum!3^[i],min_sum!2[i])
   min_sum!4[i] = MUX(!1!2[i], min_sum!3[i], min_sum!2[i])

for i in range(0, N!0):
  min_index!2[i] = PHI(min_index!1^[i], min_index!4[i-1])  
  min_index!4[i] = MUX(!1!2[i], min_index!3[i], min_index!2[i])
\end{verbatim}

\ana{Begin Work-in-progress notes on implementation. Just some thoughts...}

For now, I suggest we go with a "naive" implementation and try to do only minimal optimizations. 
All arrays are stored in the program as one-dimensional arrays. Arrays lifted from scalars have dimensionality of the loop enclosure in 
the original code, e.g., \texttt{sum!3} is two dimensional \texttt{[N!0,D!0]} (but represented linearly in row-major order; all these "copies" are 
arrays of references to shares). Naturally, all writes into these arrays are canonical writes, i.e., $A[i,j] = ...$ and no things like $A[i+1,j-1] = ...$.

The loop "for j in range..." is perhaps the most interesting. First, consider "projection", e.g., p!3[I,j]. 
We fix j (the current iteration) and iterate over I using the row-major formula; it returns the j-th column, 
a one-dimensional column vector. I think this can be generalized across any mix of vectorized vs. 
non-vectorized dimensions but I'm not 100\% sure.

\begin{verbatim}
// project the (j-1)-st column vector:
A = sum!4[I,j-1]
// project j-th column from p!3: 
B = p!3[I,j] 
// we get a N!0-size vector of references to NEW shares:
C = ADD_SIMD(A,B) 
// Writes j-th column of sum!4 with the new references:
sum!4[I,j] = C 
// may eventually optimize since C is A
\end{verbatim}

\ana{End Work-in-progress notes.}

%\end{document}

### 2022-02-01 biometric_vectorization_example.tex
Some more notes on arrays, raise_dim and drop_dim.

@@
 
+There are three kinds of arrays (for now all kept internally as one-dimensional arrays, but that's under discussion).
+\begin{itemize}
+
+\item Scalars: These are scalar variables that we lift into arrays for the purpose of vectorization. 
+For those, all writes are canonical writes and all reads are canonical reads. We may apply both raise dimension 
+and drop dimension on these.
+
+\item Read-only input arrays: Read-only inputs. There are NO writes while we may have non-canonical reads, $f(i,j,k)$. 
+Vectorization will lift the raise dimension operation at the beginning of the function and raise dimension may reshape 
+arrays. If there are multiple "views" of the input array, there could be multiple raise dimension statements to create
+these views. The invariant is that at reads in loops, the reads of "views" of the original input array will be canonical.
+Only raise-dimension applies.
+
+ \item Read-write output arrays: Writes are canonical (by restriction) but reads can be non-canonical. Conjecture:
+ do nothing. Dependence analysis takes care of limiting vectorization so non-canonical access will work. 
+Still work in progress, but we may apply both raise and drop dimension.
+
+\end{itemize}
+
 \ana{End Work-in-progress notes.}

### 2022-02-02 biometric_vectorization_example.tex
Changed fonts to fit code.

@@
 
+{\small
 \begin{verbatim}
@@
 \end{verbatim}
+}
 
@@
 \end{verbatim}
+}
 
-\noindent When processing the outer loop two closures arise, one for \texttt{min\_sum!2[I] = PHI(...)} and one 
+When processing the outer loop two closures arise, one for \texttt{min\_sum!2[I] = PHI(...)} and one 
 for \texttt{min\_index!2[I] = PHI(...)}. Since the two closures \emph{do not} intersect, we have two distinct FOR-loops on $i$:
 
+{\small
 \begin{verbatim}
@@
    min_index!4[i] = MUX(!1!2[i], min_index!3[i], min_index!2[i])
-
 \end{verbatim}
+}
 
@@
 And this phase adds SIMD operations:
+
+{\small
 \begin{verbatim}
@@
 
-\item Scalars: These are scalar variables that we lift into arrays for the purpose of vectorization. 
+\item Scalars: These are scalar variables we lift into arrays for the purposes of vectorization. 
 For those, all writes are canonical writes and all reads are canonical reads. We may apply both raise dimension 
@@
 
-\item Read-only input arrays: Read-only inputs. There are NO writes while we may have non-canonical reads, $f(i,j,k)$. 
-Vectorization will lift the raise dimension operation at the beginning of the function and raise dimension may reshape 
-arrays. If there are multiple "views" of the input array, there could be multiple raise dimension statements to create
-these views. The invariant is that at reads in loops, the reads of "views" of the original input array will be canonical.
-Only raise-dimension applies.
+\item Read-only input arrays: Read-only inputs. There are NO writes, while we may have non-canonical reads, $f(i,j,k)$. 
+Phase 1 of Basic vectorization will add raise dimension operation at the beginning of the function and raise dimension 
+may reshape arrays. If there are multiple "views" of the input array, there could be multiple raise dimension statements to create
+each one of these views. The invariant is that at reads in loops, the reads of "views" of the original input array will be canonical.
+Only raise dimension applies.
 
@@
  do nothing. Dependence analysis takes care of limiting vectorization so non-canonical access will work. 
-Still work in progress, but we may apply both raise and drop dimension.
+Still work in progress. We may apply both raise and drop dimension.
 

### 2022-02-02 biometric_vectorization_example.tex
Fixed typos in code.

@@
 $\mathit{raise\_dim}$ and $\mathit{drop\_dim}$ statements, as well as propagation statements (e.g., \texttt{min\_sum!3 = sum!3\^}) 
-may have slightly different interpretation. \ana{This is vague but I think it's clear.}
+may have slightly different interpretation. \ana{This is vague, may need elaboration.}
 
@@
 The dependences are from \texttt{p!3[I,J] = (d!3[I,J] * d!3[I,J])} to the monolithic FOR loop and from the FOR loop to \texttt{sum!3\^ = drop\_dim(sum!3)}.
-(Lower case index, e.g., $i,j$ indicates non-vectorized dimension, while uppercase index, e.g., $I$ indicates vectorized dimenion.)
+(Lower case index, e.g., \texttt{i}, indicates non-vectorized dimension, while uppercase index, e.g., \texttt{I} indicates vectorized dimension.)
 
@@
 S!0^ = raise_dim(S!0, ((i * D!0) + j), (i:N!0,j:D!0))
-C!0^ = raise_dim(C!0, j, (i:N!0,,j:D!0))
+C!0^ = raise_dim(C!0, j, (i:N!0,j:D!0))
 for i in range(0, N!0):
@@
 S!0^ = raise_dim(S!0, ((i * D!0) + j), (i:N!0,j:D!0))
-C!0^ = raise_dim(C!0, j, (i:N!0))
+C!0^ = raise_dim(C!0, j, (i:N!0,j:D!0))
 
@@
 sum!3^ = drop_dim(sum!3)     
-min_index!3 = [1,2,...N!0]   
+min_index!3 = [0,1,2,...N!0-1]   
 min_sum!3 = sum!3^

### 2022-02-04 compiler/compiler/motion_backend/reference_implementations.py
Update raise_dim and drop_dim implementations

@@ -1,99 +1,57 @@
+from typing import Callable
+
+
 def raise_dim(
-    arr: list[int], dims_in_arr: list[int], all_dims: list[int], dim_to_raise: int
+    arr: list[int], access_pattern: Callable[[list[int]], int], dim_sizes: list[int]
 ) -> list[int]:
     """
-    NOTE: This implementation is a recreation of the C++ implementation, so it could probably
-    be made more pythonic.
+    Reshapes @p arr to the shape provided by @p dim_sizes.  @p access_pattern provides
+      the mapping between the original and the reshaped array.
 
-    Raises the dimensionality of `arr` by extending it over dimension `dim_to_raise`.
-    Example: raise_dim([0, 1, 2], [0], [3, 2], 1) will return [0, 0, 1, 1, 2, 2]
-    Example: raise_dim([0, 1, 2], [1], [2, 3], 0) will return [0, 1, 2, 0, 1, 2]
+    :param arr             The array to reshape.
+    :param access_pattern  Mapping of an index in the output array (broken down
+    :                        by dimension) to an index in the input array.
+    :param dim_sizes       The size of each dimension in the output array.
 
-    :param arr           Input array.  This array's elements should be ordered in row-major
-                           order according to `dims_in_arr`.
-    :param dims_in_arr   Order sequence of which dimensions are in `arr`.  Each number in
-                           this array refers to an index in `all_dims` corresponding to
-                           the relevant dimension. THIS ARRAY MUST BE SORTED IN ASCENDING
-                           ORDER.
-    :param all_dims      Array sizes of each dimsnion.  For example, the array [2, 1, 3]
-                           means that dimension 0 has size 2, dimension 1 has size 1, and
-                           dimension 2 has size 3.
-    :param dim_to_raise  Index of the dimension `all_dims` to extend `arr` over.  This index
-                           cannot be in `dims_in_arr`.
-
-    :returns The extended version of `arr`
+    :returns The reshaped array.
     """
-    raised_size = all_dims[dim_to_raise]
-    for dim_idx in dims_in_arr:
-        assert dim_idx != dim_to_raise
-        raised_size *= all_dims[dim_idx]
+
+    raised_size = 1
+    for dim_size in dim_sizes:
+        raised_size *= dim_size
 
[…]

### 2022-02-04 paper_SIMD.tex
Changed biometric example to literal code.
Edits to 4.5 Extension to Basic Vectorization.

@@
 
+\input{biometric_vectorization_example.tex}
+
+\begin{comment}
 \subsection{Example: Biometric}
@@
 Additionally, while there could be several different loops that perform canonical updates, they must be of the same dimensionality, i.e., an update of higher or lower dimension, e.g., $A[i,j,k] = ...$ is not allowed.
-We assign the \emph{canonical dimensionality} to each array in the obvious way.
+We compute the \emph{canonical dimensionality} of each write array in the obvious way \emph{before basic vectorization}.
 This restriction simplifies reasoning in this early stage of the compiler; we will look to relax the restriction in future work.
 
-Reads through an arbitrary formula, such as $\texttt{A}[i-1]$ for example, are allowed, however, we assume the programmer ensures that the formula is within the bounds of the array.
+Another restriction/assumption is that we assume the output array is given as input with initial values, and it is of size consistent with its canonical dimensionality.
+\ana{This restriction will require that we update our benchmarks that come with output arrays.}
+
+Reads through an arbitrary formula, such as $\texttt{A}[i-1]$ for example, are allowed; currently, our projection function returns dummy values if the read formula is out of bounds; we assume the programmer ensures that the program still computes correct output in this case.
 
@@
 One change to Basic vectorization is the expansion of dimension if the array write or read occurs in a nested loop.
-That is, if there is an update $\texttt{A}[i,j] = ...$ that occurs in loop nest $i,j,k$, $\texttt{A}[i,j]$ will be rewritten into $\texttt{A}[i,j,k]$.
+That is, if there is an update $\texttt{A}[i,j] = ...$ that occurs in loop dimensionality $i,j,k$, $\texttt{A}[i,j]$ will be rewritten into $\texttt{A}[i,j,k]$.
 Similarly, a read $\texttt{A}[f(i,j)]$ will be rewritten into $\texttt{A}[f(i,j),k]$.
@@
 
-The other change concerns def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable (e.g., the definition can be an update $A_2 = update(A_1,...)$ or a pseudo $\phi$-node $A_2 = \phi(A_0,A_1)$).
-These edges are not handled in the same way as in Basic Vectorization, specifically, we do not expand dimension as we do for scalars in Basic Vectorization according to the loop nest.
-A key invariant is that the dimension of an array $A$ cannot drop below $A$'s canonical dimensionality.
-Below we enumerate the cases of def-use edges.
+The other change concerns def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable, e.g., the definition can be an update $A_2 = update(A_1,...)$ or a pseudo $\phi$-node $A_2 = \phi(A_0,A_1)$.
+Note that $\phi$ nodes for arrays are no subscript operations on those $\phi$-nodes the way there are in scalar arrays.
+These edges are not handled in the same way as in Basic Vectorization, specifically, we do not raise and drop dimension as we do for scalars in Basic Vectorization.
+A key invariant is that the dimension of an array $A$ cannot go above or drop below $A$'s canonical dimensionality.
+We enumerate the cases of def-use edges.
 
@@
 \item inner-to-outer $X \rightarrow Y$.
-Let $A$ be the array defined at $X$.
-If the dimensionality of $A$ is greater than its canonical dimensionality, then add \emph{drop\_dim(...)} at $Y$, as in Basic Vectorization.
+If dimensionality of the loop enclosure of $X$ is greater than the canonical dimensionality of the array, then add \emph{drop\_dim(...)} at $Y$, as in Basic Vectorization.
 Otherwise, do nothing.
 
-\item outer-to-inner $X \rightarrow Y$. Add \emph{raise\_dim(...)} (at $X$) as in Basic Vectorization.
+\item outer-to-inner $X \rightarrow Y$.
+If dimensionality of loop enclosure of $Y$ is greater than the canonical dimensionality of the array, then add \emph{raise\_dim(...)} (at $X$) as in Basic Vectorization.
+Otherwise, do nothing.
 
@@
 
-After Phase 1 of Basic Vectorization the Aiken's array write example will be (roughly) as follows:
+First, the canonical dimensionality of all \texttt{A,B,C} and \texttt{D} is 1. 
+Thus, there is no addition of extra dimension for inner loops. After Phase 1 
+of Basic Vectorization the Aiken's array write example will be (roughly) as follows:
 
@@
 \STATE 3. $\texttt{C}_1$ = $\phi(\texttt{C}_0,\texttt{C}_2)$
-\STATE 4. $\texttt{D}_2$ = $\phi(\texttt{D}_0,\texttt{D}_2)$
+\STATE 4. $\texttt{D}_1$ = $\phi(\texttt{D}_0,\texttt{D}_2)$
 \STATE 5. $\texttt{A}_2 = update(\texttt{A}_1, \stackrel{\rightarrow}{i}, \texttt{B}_1[\stackrel{\rightarrow}{i}] + 10)$;
@@
 
-There are no nested loops, thus, array accesses remain as is. Phase 2 computes the closure of 4; $cl = \{4,6,7, 8\}$ 
+Phase 2 computes the closure of 4; $cl = \{4,6,7, 8\}$ 
 while 5 is vectorizable. Recall that 1,2, and 3 are targetless phi-nodes. Since the closure $cl$ includes updates to $B$ and $C$,
@@
 
+\ana{TODO: Format either Histogram or Matrix Multiplication. Both are interesting examples that use output arrays.}
+
+\begin{comment}
 \paragraph{Example 2}

### 2022-02-05 paper_SIMD.tex
Removed a TODO in 4.5.
Removed and added TODOs in Section 4.5.

@@
 
-\ana{TODO: This section is work in progress!}
-
 \subsubsection{Restricting Array Writes}
@@
 
+There are two changes to Basic Vectorization to account for write arrays. We can account for them during processing of raise dimension/drop dimension in Phase 1 of Basic Vectorization.
+
 One change to Basic vectorization is the expansion of dimension if the array write or read occurs in a nested loop.
@@
 Similarly, a read $\texttt{A}[f(i,j)]$ will be rewritten into $\texttt{A}[f(i,j),k]$.
-\ana{TODO: This notion of $f(i,j)$ needs a precise definition.
-For now I'm hoping the hand waving works.} This can be a preprocessing step or it can be incorporated into the Basic Vectorization algorithm.
+($\texttt{A}[f(i,j),k]$ will be reshaped as $f(i,j)*I*J+k$ during Phase 1 of Basic Vectorization.
+\ana{TODO: Double check.})
 
-The other change concerns def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable, e.g., the definition can be an update $A_2 = update(A_1,...)$ or a pseudo $\phi$-node $A_2 = \phi(A_0,A_1)$.
+The other change concerns def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable.
+The definition can be an update $A_2 = update(A_1,...)$ or a pseudo $\phi$-node $A_2 = \phi(A_0,A_1)$.
 Note that $\phi$ nodes for arrays are no subscript operations on those $\phi$-nodes the way there are in scalar arrays.

### 2022-02-09 biometric_vectorization_example.tex
Added a TODO in Basic Vectorization.

@@
 
+\ana{TODO: It will be good to add a pass in Phase 3 to get rid of extra dimensionality. 
+Once we have vectorized as much as possible, there is no need to keep higher 
+dimensionality. E.g., there is no need for the 2-dimensional array sum in the first for loop, 
+and similarly, there is no need for 1-dimensional arrays for min\_sum and min\_index.
+We should get rid of these before generating MOTION code.}
+
 \ana{Begin Work-in-progress notes on implementation. Just some thoughts...}

### 2022-02-09 biometric_vectorization_example.tex
Minor edits of TODO item.

@@
 
-\ana{TODO: It will be good to add a pass in Phase 3 to get rid of extra dimensionality. 
+\ana{TODO: It will be good to add a pass in Phase 3, or post Phase 3, to get rid of extra dimensionality. 
 Once we have vectorized as much as possible, there is no need to keep higher 
@@
 We should get rid of these before generating MOTION code.}
+
+\ana{I don't have a crisp algorithm in mind. I don't think it will be difficult but we have to 
+think this through. Put on our TODO list for us to think about since it will simplify MOTION 
+code tremendously.}
 

### 2022-02-11 biometric_vectorization_example.tex
Minor edit to Biometric code in Basic vectorization example.

@@
    min_index!4 = MUX(!1!2, min_index!3, min_index!2)
+!2!1 = (min_sum!2, min_index!2)   
 \end{verbatim}
@@
    min_index!4 = MUX(!1!2, min_index!3, min_index!2)
+min_sum!2^ = drop_dim(min_sum!2)
+min_index!2^ = drop_dim(min_index!2)   
+!2!1 = (min_sum!2^, min_index!2^)     
 \end{verbatim}
@@
    min_index!4[I] = MUX(!1!2[I], min_index!3[I], min_index!2[I])
+min_sum!2^ = drop_dim(min_sum!2)
+min_index!2^ = drop_dim(min_index!2)   
+!2!1 = (min_sum!2^, min_index!2^)
 \end{verbatim}
@@
    min_index!4[i] = MUX(!1!2[i], min_index!3[i], min_index!2[i])
+
+min_sum!2^ = drop_dim(min_sum!2)
+min_index!2^ = drop_dim(min_index!2)   
+!2!1 = (min_sum!2^, min_index!2^)
 \end{verbatim}
@@
    min_index!4[i] = MUX(!1!2[i], min_index!3[i], min_index!2[i])
+
+min_sum!2^ = drop_dim(min_sum!2)
+min_index!2^ = drop_dim(min_index!2)   
+!2!1 = (min_sum!2^, min_index!2^)   
 \end{verbatim}

### 2022-02-11 biometric_vectorization_example.tex
Added draft of Phase 4 of Basic Vectorization (Dimensionality reduction). Still work in progress.

@@
 
+\subsubsection{Phase 4 of Basic Vectorization}
+
+And this phase gets rid of redundant dimensionality. Code moves back to 
+a version close to the original one.
+
+{\small
+\begin{verbatim}
+min_sum!1 = 10000
+min_index!1 = 0
+S!0^ = raise_dim(S!0, ((i * D!0) + j), (i:N!0,j:D!0))
+C!0^ = raise_dim(C!0, j, (i:N!0,j:D!0))
+
+sum!2 = [0,..,0]
+d!3[I,J] = SUB_SIMD(S!0^[I,J],C!0^[I,J])
+p!3[I,J] = MUL_SIMD(d!3[I,J] * d!3[I,J])
+
+for j in range(0, D!0):
+  // I dim is a noop. sum is already a one-dimensional vector
+  sum!3[I] = PHI(sum!2^[I], sum!4[I])       
+  sum!4[I] = ADD_SIMD(sum!3[I],p!3[I,j])
+
+min_index!3 = [0,1,...N!0-1]
+
+for i in range(0, N!0):
+  min_sum!2 = PHI(min_sum!1, min_sum!4) 
+  !1!2[i] = CMP(sum!3[i],min_sum!2)
+  min_sum!4 = MUX(!1!2[i], sum!3[i], min_sum!2)
+
+for i in range(0, N!0):
+  min_index!2 = PHI(min_index!1, min_index!4)  
+  min_index!4 = MUX(!1!2[i], min_index!3[i], min_index!2)
+
+!2!1 = (min_sum!2, min_index!2)   
+\end{verbatim}
+}
+
 \ana{TODO: It will be good to add a pass in Phase 3, or post Phase 3, to get rid of extra dimensionality. 

### 2022-02-11 paper_SIMD.tex
Added draft of Phase 4 of Basic Vectorization (Dimensionality reduction). Still work in progress.

@@
 \STATE {add SIMD for simdified dimensions}
+
+\COMMENT {Phase 4: Dimensionality reduction: remove unnecessary dimensionality. \ana{Draft. Needs work.}}
+\FOR {each FOR i = 0; ... loop from innermost to outermost}
+\FOR {each stmt v[...i...] = ..., including in nested loops}
+\STATE {if def-use from stmt v[...i...] to a non drop\_dim(v,i) use outside of loop, mark "live dimension i"}
+\STATE {otherwise, mark "dead dimension i"}
+\STATE {for all variables v[...i...] (defs and uses) that are "dead dimension i", remove dimension i}
+\COMMENT {Two cases: forward edge will use i, backward edge will use i-1}
+\ENDFOR
+
+\COMMENT {Now clean up drop\_dim and raise\_dim}
+\FOR {each "dead dimension i" def: v[...i...] in loop to use: v' =  drop\_dim(v,i) }
+\STATE {replace v' = drop\_dim(v,i) with v' = v in use}
+\ENDFOR
+\FOR {each "def: v' = raise\_dim(v,i) in outer loop to use: v'[...i...] in loop}
+\STATE {replace v' with v in use statement}
+\ENDFOR
+\ENDFOR
+\STATE {do copy propagation and dead code elimination to get rid of redundant variables and raise dimension statements}
 

### 2022-02-16 paper_SIMD.tex
Edits to Demensionality reduction.

--- 2022-02-05 paper_SIMD.tex
@@
 \STATE {add SIMD for simdified dimensions}
+
+\COMMENT {Phase 4: Dimensionality reduction: remove unnecessary dimensionality. \ana{Better draft! Needs a bit more work.}}
+
+\COMMENT { A dimension i is dead on exit from stmt v[...i...] = ... if all def-uses with targets outside of the enclosing 
+FOR i=0 ... MOTION loop end at target (use) v' = drop\_dim(v, i). } 
+\FOR {each stmt and dimension v[...i...] = ...}
+\STATE {if i is a dead dimension on exit from stmt v[...i...] = ..., remove i from v[...i...] (at def and at all uses)} 
+%\COMMENT{ A stmt v[...i] is a "dead dimension i" if all edges out of FOR i = 0; ... loop end at drop\_dim(v, i)}
+\ENDFOR
+
+\COMMENT {Now clean up drop\_dim and raise\_dim}
+\FOR {each v' =  drop\_dim(v,i) }
+\STATE {replace with v' = v if i is dead in v.}
+\ENDFOR
+\FOR {each v' = raise\_dim(v,i) in outer loop to use: v'[...i...] in loop}
+\STATE {replace v' with v in use statement}
+\ENDFOR
+\STATE {do copy propagation and dead code elimination to get rid of redundant variables and raise dimension statements}
 

### 2022-02-19 biometric_vectorization_example.tex
Minor edit to Biometric example section formatting.
Minor edit to Basic Vectorization.

@@
 
-\subsubsection{Benjamin's code with linear loops (MPC Source)}
+We start from Benjamin's code with linear loops (MPC Source):
 
@@
   min_sum!3 = sum!3^
-  min_index!3 = [0,1,2,...,N!0-1]
+  min_index!3 = i
   min_sum!4[I] = MUX(!1!2[I], min_sum!3[I], min_sum!2[I])
@@
 sum!3^ = drop_dim(sum!3)     
-min_index!3 = [0,1,2,...N!0-1]   
+min_index!3 = [0,1,2,...N!0-1] // or min_index!3 = [i, (i:N!0)]
 min_sum!3 = sum!3^

### 2022-02-19 biometric_vectorization_example.tex
Phases 3-4 of Basic Vectorization.

@@
 
-And this phase adds SIMD operations:
+This phase removes redundant dimensionality. 
+It starts by removing redundant dimensions in MOTION loops followed
+by removal of redundant drop dimension statements.
+It then does (extended) constant propagation 
+to "bypass" raise statements, copy propagation
+and dead code elimination.
+
+The code becomes closer to what we started with:
 
@@
 min_sum!1 = 10000
-min_sum!1^ = raise_dim(min_sum!1, (i:N!0))
 min_index!1 = 0
-min_index!1^ = raise_dim(min_index!1, (i:N!0))
 S!0^ = raise_dim(S!0, ((i * D!0) + j), (i:N!0,j:D!0))
@@
 sum!2 = [0,..,0]
-sum!2^ = raise_dim(sum!2, (j:D!0))
-d!3[I,J] = SUB_SIMD(S!0^[I,J],C!0^[I,J])
-p!3[I,J] = MUL_SIMD(d!3[I,J] * d!3[I,J])
+d!3[I,J] = S!0^[I,J] - C!0^[I,J]
+p!3[I,J] = (d!3[I,J] * d!3[I,J])
 
+// j is redundant for sum!3 and sum!4
 for j in range(0, D!0):
-  sum!3[I,j] = PHI(sum!2^[I,j], sum!4[I,j-1])       
-  sum!4[I,j] = ADD_SIMD(sum!3[I,j],p!3[I,j])
+  sum!3[I] = PHI(sum!2^[I], sum!4[I])       
+  sum!4[I] = (sum!3[I] + p!3[I,j])
 
-sum!3^ = drop_dim(sum!3)     
-min_index!3 = [0,1,...N!0-1]   
-min_sum!3 = sum!3^
+// drop_dim is redundant, removing
+// then copy propagation and dead code elimination 
+min_index!3 = [0,1,2,...N!0-1] // or min_index!3 = [i, (i:N!0)]
 
+// i is redundant for min_sum!2, min_sum!4 but not for !12![i]!
 for i in range(0, N!0):
-  min_sum!2[i] = PHI(min_sum!1^[i], min_sum!4[i-1]) 
-  !1!2[i] = CMP(sum!3^[i],min_sum!2[i])
-  min_sum!4[i] = MUX(!1!2[i], min_sum!3[i], min_sum!2[i])
+  min_sum!2 = PHI(min_sum!1, min_sum!4) 
+  !1!2[i] = (sum!3[i] < min_sum!2)
+  min_sum!4 = MUX(!1!2[i], sum!3[i], min_sum!2)
 
+// same, i is redundant for min_index!2, min_index!4
 for i in range(0, N!0):
-  min_index!2[i] = PHI(min_index!1^[i], min_index!4[i-1])  
-   min_index!4[i] = MUX(!1!2[i], min_index!3[i], min_index!2[i])
+  min_index!2 = PHI(min_index!1, min_index!4)  
+   min_index!4 = MUX(!1!2[i], min_index!3[i], min_index!2)
 
-min_sum!2^ = drop_dim(min_sum!2)
-min_index!2^ = drop_dim(min_index!2)   
-!2!1 = (min_sum!2^, min_index!2^)   
+// drop_dim becomes redundant
+!2!1 = (min_sum!2, min_index!2)
 \end{verbatim}
@@
 
-And this phase gets rid of redundant dimensionality. Code moves back to 
-a version close to the original one.
+And this phase adds SIMD operations:
 

### 2022-02-19 paper_SIMD.tex
Edits on Phase 4.
Phases 3-4 of Basic Vectorization.

--- 2022-02-05 paper_SIMD.tex
@@
 
-\COMMENT {{\color{red} $cl_1$ and $cl_2$ intersect if they have common statement, or update same array; "intersect" definition can be expanded}}
+\COMMENT {{\color{red} $cl_1$ and $cl_2$ intersect if they have common statement or update same array; "intersect" definition can be expanded}}
 
@@
 \STATE {change $\stackrel{\rightarrow}{i_d}$ to $i_d$ in all statements in loop}
-\STATE {treat FOR loop as monolith node: change def-use edges accordingly. Some edges become same-level.}
+\STATE {treat FOR loop as monolith node: some def-use edges become same-level.}
 \ENDFOR
@@
 
-\COMMENT {Phase 3:}
+\COMMENT {Phase 3: Dimensionality reduction: remove unnecessary dimensionality. \ana{Better draft! Needs a bit more work.}}
+
+\COMMENT { A dimension $i$ is dead on exit from stmt $X[...i...] = ...$ if all def-uses with targets outside of the enclosing 
+FOR $i=0 ...$ MOTION loop end at target (use) $X' = \mathit{drop\_dim}(X, i)$. } 
+\FOR {each stmt and dimension $X[...i...] = ...$}
+\STATE {if $i$ is a dead dimension on exit from stmt $X[...i...] = ...$, remove $i$ from $X$} 
+%\COMMENT{ A stmt v[...i] is a "dead dimension i" if all edges out of FOR i = 0; ... loop end at drop\_dim(v, i)}
+\ENDFOR
+
+\COMMENT {Now clean up drop\_dim and raise\_dim}
+\FOR {each $X' =  \mathit{drop\_dim(X,i)}$ }
+\STATE {replace with $X' = X$ if $i$ is dead in $X$.}
+\ENDFOR
+\STATE {do (1) (extended) constant propagation, (2) copy propagation and (3) dead code elimination to get rid of redundant variables and raise and drop dimension statements}
+
+\COMMENT {Phase 4:}
 \STATE {add SIMD for simdified dimensions}

### 2022-02-19 paper_SIMD.tex
Cleaned up Basic Vectorization a bit.

@@
 
-\STATE \COMMENT {Phase 1: Raise dimension of scalar variables to corresponding loop nest. We may assume linear traversal of the MPC-source.}
+\STATE \COMMENT { Phase 1: Raise dimension of scalar variables to corresponding loop nest. We can traverse stmts linearly in MPC-source. }
 
-\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i_1,...,i_k$}
-\FOR {each $Y_i$}
-\STATE {case def-use edge $\mathit{stmt}' (\mbox{def of } Y_i) \rightarrow \mathit{stmt} (\mbox{def of } X)$ of} \\
-\STATE \hspace{0.25cm} {same-level $\rightarrow$ add $Y'_i = Y_i$} \\
-\STATE \hspace{0.25cm} {outer-to-inner $\rightarrow$ add $Y'_i[i_1,...,i_k] = \mathit{raise\_dim}(Y_i)$}
-\STATE \hspace{0.25cm} {inner-to-outer $\rightarrow$ add $Y'_i[i_1,...,i_k] = \mathit{drop\_dim}(Y_i)$}
+%\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i,j,k$}
+%\FOR {each $Y_n$}
+\FOR {each def-use edge $\mathit{stmt}' (\mbox{def of } Y) \rightarrow \mathit{stmt} (\mbox{def of } X)$}
+\STATE {case def-use edge $\mathit{stmt}' (\mbox{def of } Y) \rightarrow \mathit{stmt} (\mbox{def of } X)$ of} \\
+\STATE \hspace{0.25cm} {same-level: $Y'$ is $Y$} \\
+\STATE \hspace{0.25cm} {outer-to-inner: add $Y'[i,j,k] = \mathit{raise\_dim}(Y)$ at $\mathit{stmt'}$}
+\STATE \hspace{0.25cm} {inner-to-outer: add $Y'[i,j,k] = \mathit{drop\_dim}(Y)$ at $\mathit{stmt}$}
 \ENDFOR
 
-\COMMENT {Optimistically vectorize all. $\stackrel{\rightarrow}{i}$ means vectorized dimension.}
-\STATE {change to $X[\stackrel{\rightarrow}{i_1},...\stackrel{\rightarrow}{i_k}] = \mathit{Op}(Y'_1[\stackrel{\rightarrow}{i_1},...,\stackrel{\rightarrow}{i_k}],Y'_2[\stackrel{\rightarrow}{i_1},...,\stackrel{\rightarrow}{i_k}])$ }
+\COMMENT { Optimistically vectorize all. $\stackrel{\rightarrow}{i}$ means vectorized dimension. }
+\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i,j,k$}
+\STATE {change to $X[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j},\stackrel{\rightarrow}{k}] = \mathit{Op}(Y'_1[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j},\stackrel{\rightarrow}{k}],Y'_2[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j},\stackrel{\rightarrow}{k}])$ }
 \ENDFOR
 
-\COMMENT {Phase 2: Recreating FOR loops for cycles; vectorizable statements hoisted up.}
+\COMMENT { Phase 2: Recreating FOR loops for cycles; vectorizable statements hoisted up. }
 
@@
 \STATE {add statements in $cl$ to loop body in some order of dependences}
-\STATE \COMMENT {Dimension is not vectorizable: }
+\STATE \COMMENT { Dimension is not vectorizable: }
 \STATE {change $\stackrel{\rightarrow}{i_d}$ to $i_d$ in all statements in loop}
@@
 
-\COMMENT {Phase 3: Dimensionality reduction: remove unnecessary dimensionality. \ana{Better draft! Needs a bit more work.}}
+\COMMENT { Phase 3: Dimensionality reduction: removing unnecessary dimensionality. }
 
@@
 \FOR {each stmt and dimension $X[...i...] = ...$}
-\STATE {if $i$ is a dead dimension on exit from stmt $X[...i...] = ...$, remove $i$ from $X$} 
+\STATE {if $i$ is a dead dimension on exit from stmt $X[...i...] = ...$, remove $i$ from $X$ (all defs and uses)} 
 %\COMMENT{ A stmt v[...i] is a "dead dimension i" if all edges out of FOR i = 0; ... loop end at drop\_dim(v, i)}
@@
 
-\COMMENT {Now clean up drop\_dim and raise\_dim}
+\COMMENT { Now clean up drop\_dim and raise\_dim }
 \FOR {each $X' =  \mathit{drop\_dim(X,i)}$ }
@@
 
-\COMMENT {Phase 4:}
+\COMMENT { Phase 4: }
 \STATE {add SIMD for simdified dimensions}

### 2022-02-19 benchmarks/histogram.py
Update to histogram.py. Result array is now input.

@@ -16,20 +16,21 @@ from UTIL import shared
 # But we were first to suggest this as a benchmark :).
 # requires: len(A) == len(B) = N
 def histogram(
-    A: shared[list[int]], B: shared[list[int]], N: int, num_bins: int
+    A: shared[list[int]], B: shared[list[int]], N: int, num_bins: int, result: list[int] 
 ) -> shared[list[int]]:
-    result: list[int] = []
-    # initialize result to 0
-    for i in range(num_bins):
-        result = result + [0]
     for i in range(num_bins):
         for j in range(N):
             if A[j] == i:
-                result[i] = result[i] + B[j]
+                val = result[i] + B[j]
+            else:
+                val = result[i]
+            result[i] = val
     return result
 
 
 A = [0, 2, 1, 0, 3, 4, 2, 3]
 B = [10, 1, 5, 2, 15, 0, 10, 1000]
 N = 8  # len(A)
-print(histogram(A, B, N, 5))
+#R = [12, 5, 11, 1015, 0]
+R = [0, 0, 0, 0, 0]
+print(histogram(A, B, N, 5, R))

### 2022-02-20 paper_SIMD.tex
Cleanup writeup.
Made a plan. Added an overview section.

@@
 
-\section{Background and Problem Statement}
-\label{sec:background}
+\begin{itemize}
+\item We define the scheduling problem for MPC. 
+We present an analytical model to reason about cost of schedules
+and show that scheduling is NP-hard via a reduction to the shortest 
+common supersequence problem.
+
+\item We present a compiler that takes a Python-like high-level 
+program and produces amortized (i.e., vectorized) low-level cryptographic 
+code in the MOTION framework. Central to our approach is a new 
+vectorization algorithm that produces optimal schedules for a large 
+number of MPC programs.
+
+\item We present an implementation and evaluation in the MOTION framework. 
+\ana{Fill in with final results. Mention benchmarks (standard + new ones + HyCC). }
+
+\end{itemize}
+
+\section{Overview}
+\label{sec:overview}
+
+\subsection{Source}
+
+As a running example, we assume a standard MPC benchmark, Biometric matching.
+An intuitive implementation is as follows:
+
+{\small
+\begin{verbatim}
+def biometric(C: shared[list[int]], D: int, 
+              S: shared[list[int]], N: int) -> 
+                      tuple[shared[int], shared[int]]:
+    min_sum = 10000
+    min_index = 0  
+    for i in range(N):
+        sum = 0
+        for j in range(D):
+            d = S[i * D + j] - C[j]
+            p = d * d
+            sum = sum + p
+        if sum < min_sum:
+            min_sum = sum
+            min_index = i
+    return (min_sum, min_index)
+\end{verbatim}
+}
+
+Array \texttt{C} is the feature vector of \texttt{D} 
+features that we wish to match and array \texttt{S} is the database of \texttt{N} vectors of size \texttt{D}
+that we match against. The programmer annotates the inputs and outputs as \emph{shared}.
+The program iterates over the entries in the database and computes
+the sum of squares of the differences of individual features. The program returns the 
+index \texttt{i} of the vector that gives the best match plus the corresponding sum of squares.
+
+Our compiler imposes the following restrictions. We note that in most cases, the restrictions 
+can be easily lifted and we plan to do so in future iterations of our compiler.
+
+\begin{enumerate}
+\item The program contains arbitrarily nested loops, however, loop bounds are fixed: \texttt{0 <= i < N}.
+A standard restriction in MPC is that the bounds must be known at circuit-generation time.
+\item Arrays are one-dimensional. N-dimensional arrays are linearized and accessed 
+in row-major order and at this point the programmer is responsible for linearization
+and access. 
+\item Array subscrpts are plaintext values.
+\item Our compiler allows for output (write) arrays, however it restricts
+write access to \emph{canonical writes} along the dimensions of the array (i.e., \texttt{A[i,j] = ...}
+where \texttt{i} and \texttt{j} loop over the two dimensions of \texttt{A} is allowed, 
+but \texttt{A[i,j+2] = ...} is not allowed). Read access is arbitrary. 
+\end{enumerate}
+
+\subsection{MPC Source and Cost of Schedule}
+
+The compiler generates an IR, MPC source:
+{\small
+\begin{verbatim}
+1. min_sum!1 = 10000
+2. min_index!1 = 0
+3. for i in range(0, N!0):
+4.    min_sum!2 = PHI(min_sum!1, min_sum!4)
+5.    min_index!2 = PHI(min_index!1, min_index!4)
+6.    sum!2 = 0
+7.    for j in range(0, D!0):
+8.      sum!3 = PHI(sum!2, sum!4)
+9.      d!3 = (S!0[((i * D!0) + j)] - C!0[j]) // MPC
+10.     p!3 = (d!3 * d!3) // MPC
+11.     sum!4 = (sum!3 + p!3) // MPC
+12.  !1!2 = (sum!3 < min_sum!2) // MPC
+13.  min_sum!3 = sum!3
+14.  min_index!3 = i
+15.  min_sum!4 = MUX(!1!2, min_sum!3, min_sum!2) // MPC
+16.  min_index!4 = MUX(!1!2, min_index!3, min_index!2) // MPC
+17.!2!1 = (min_sum!2, min_index!2)   
+\end{verbatim}
+}
+
+We linearize the source turning conditionals into MUX statements. The PHI nodes
+are remnants of the SSA IR; our compiler generates code that picks the correct value 
+when producing MOTION output and the MOTION framework in turn linearizes 
+MPC source loops when it generates the circuit.
+
+We turn to our analytical model to compute the cost of this program. Assuming fixed 
+cost $\beta$ for a local MPC operation (essentially just ADD) and cost $\alpha$ for a remote
+MPC operation (e.g., MUX, CMP, and remaining operations), the cost of the iterative schedule 
+will be $N*D*(2*\alpha+\beta) + N*3*\alpha$.
+
+Our key contribution is the vectorizing transformation. We can compute all $N*D$ 
+subtraction operations (line 9) in a single SIMD instruction; similarly we can compute 
+all multiplication operations (line 10) in a single SIMD instruction. And while we cannot
+vectorize computation of the N individual sums, we can compute the N sums in parallel. 
+Our compiler automatically detects these opportunities and transforms the program. 
+It is standard that MPC researchers write vectorized versions of the Biometric
+program by hand; we are the first (to the best of our knowledge) to automatically 
+transform an intuitive, iterative MPC program into an unintuitive vectorized one.
+
+\subsection{Vectorization and Cost of Schedule}
+
+Our compiler produces the following vectorized program. (Note that this is
+still higher-level IR, our compiler turns this code into MOTION variables 
+loops and SIMD primitives, which MOTION then uses to generate the circuit.)
+
+{\small
+\begin{verbatim}
+min_sum!1 = 10000
+min_index!1 = 0
+// S!0^ is same as S!0. C!0 replicates C!0 N-times: 
+S!0^ = raise_dim(S!0, ((i * D!0) + j), (i:N!0,j:D!0))
+C!0^ = raise_dim(C!0, j, (i:N!0,j:D!0))
+
+sum!2 = [0,..,0]
+// Computes all differences and all products "at once"
+d!3[I,J] = SUB_SIMD(S!0^[I,J],C!0^[I,J])
+p!3[I,J] = MUL_SIMD(d!3[I,J] * d!3[I,J])
+
+for j in range(0, D!0):
+  // sum!2[I], sum!3[I], sum!4[I] are vectors of size N  
+  // Computes N intermediate sums "at once"
+  sum!3[I] = PHI(sum!2^[I], sum!4[I])       
+  sum!4[I] = ADD_SIMD(sum!3[I],p!3[I,j])
+
+min_index!3 = [0,1,...N!0-1]
+
+for i in range(0, N!0):
+  min_sum!2 = PHI(min_sum!1, min_sum!4) 
+  !1!2[i] = CMP(sum!3[i],min_sum!2)
+  min_sum!4 = MUX(!1!2[i], sum!3[i], min_sum!2)
+
+for i in range(0, N!0):
+  min_index!2 = PHI(min_index!1, min_index!4)  
+  min_index!4 = MUX(!1!2[i], min_index!3[i], min_index!2)
+!2!1 = (min_sum!2, min_index!2)   
+\end{verbatim}
+}
+
+In MPC compilers a vectorized operation that computing M operations "at once" costs 
+essentially the same ($\alpha$ or $\beta$) as an individual operation. We elaborate 
+on these in the following section. Thus, the vectorized program costs 
+$2*\alpha$ + $D * \beta + N*3*\alpha$. The first term in the sum corresponds to the vectorized
+subtraction and multiplication, the second term corresponds to the for loop on $j$ and the third 
+one corresponds to the remaining for loops on $i$. Clearly, $2*\alpha + D * \beta + N*3*\alpha << N*D*(2*\alpha+\beta) + N*3*\alpha$.
+Our experimental results illustrate this as well. \ana{Add numbers.}
+
+\section{Analytical Model}
+\label{sec:model}
 
 \subsection{Scheduling in MPC}
-
-\ana{Here we need more on MPC before jumping to scheduling. MPC-source, basic assumptions about static loop bounds, etc.}
+\label{sec:mpc}
+
+%\ana{Here we need more on MPC before jumping to scheduling. MPC-source, basic assumptions about static loop bounds, etc.}
 
@@
 
-\subsection{Scheduling in HPC}
-\label{sec:hpc}
-
-\section{\bf Preliminaries}
+\subsection{Problem Statement}
+\label{sec:problem}
+
+\subsection{Scheduling is NP-hard}
+\label{sec:np}
+
+\section{\bf Compiler Framework}
 \label{sec:preliminaries}
@@
 
+\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i,j,k$}
+%\FOR {each $Y_n$}
+\FOR {each argument $Y_n$} %def-use edge $\mathit{stmt}' (\mbox{def of } Y) \rightarrow \mathit{stmt} (\mbox{def of } X)$}
+\STATE {case def-use edge $\mathit{stmt}' (\mbox{def of } Y_n) \rightarrow \mathit{stmt} (\mbox{def of } X)$ of} \\
+\STATE \hspace{0.25cm} {same-level: $Y'_n$ is $Y_n$} \\
+\STATE \hspace{0.25cm} {outer-to-inner: add $Y'_n[i,j,k] = \mathit{raise\_dim}(Y_n)$ at $\mathit{stmt'}$}
+\STATE \hspace{0.25cm} {inner-to-outer: add $Y'_n[i,j,k] = \mathit{drop\_dim}(Y_n)$ at $\mathit{stmt}$}
+\ENDFOR
+
+\COMMENT { Optimistically vectorize all. $I$ means vectorized dimension. }
 %\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i,j,k$}
-%\FOR {each $Y_n$}
-\FOR {each def-use edge $\mathit{stmt}' (\mbox{def of } Y) \rightarrow \mathit{stmt} (\mbox{def of } X)$}
-\STATE {case def-use edge $\mathit{stmt}' (\mbox{def of } Y) \rightarrow \mathit{stmt} (\mbox{def of } X)$ of} \\
-\STATE \hspace{0.25cm} {same-level: $Y'$ is $Y$} \\
-\STATE \hspace{0.25cm} {outer-to-inner: add $Y'[i,j,k] = \mathit{raise\_dim}(Y)$ at $\mathit{stmt'}$}
-\STATE \hspace{0.25cm} {inner-to-outer: add $Y'[i,j,k] = \mathit{drop\_dim}(Y)$ at $\mathit{stmt}$}
-\ENDFOR
-
-\COMMENT { Optimistically vectorize all. $\stackrel{\rightarrow}{i}$ means vectorized dimension. }
-\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i,j,k$}
-\STATE {change to $X[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j},\stackrel{\rightarrow}{k}] = \mathit{Op}(Y'_1[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j},\stackrel{\rightarrow}{k}],Y'_2[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j},\stackrel{\rightarrow}{k}])$ }
+\STATE {change to $X[I,J,K] = \mathit{Op}(Y'_1[I,J,K],Y'_2[I,J,K])$ }
 \ENDFOR
@@
 \STATE \COMMENT { Dimension is not vectorizable: }
-\STATE {change $\stackrel{\rightarrow}{i_d}$ to $i_d$ in all statements in loop}
+\STATE {change ${I_d}$ to $i_d$ in all statements in loop}
 \STATE {treat FOR loop as monolith node: some def-use edges become same-level.}
@@
 Another restriction/assumption is that we assume the output array is given as input with initial values, and it is of size consistent with its canonical dimensionality. 
-\ana{This restriction will require that we update our benchmarks that come with output arrays.}
+%\ana{This restriction will require that we update our benchmarks that come with output arrays.}
 

### 2022-02-20 biometric_vectorization_example.tex
Made a plan. Added an overview section.
Reorganized, removed TODOS, added other TODOs.

@@
 
-\subsection{Example: biometric}
+\subsection{Example: Biometric}
 
@@
 It then does (extended) constant propagation 
-to "bypass" raise statements, copy propagation
+to "bypass" raise statements, followed by copy propagation
 and dead code elimination.
@@
 for j in range(0, D!0):
-  sum!3[I] = PHI(sum!2^[I], sum!4[I])       
+  sum!3[I] = PHI(sum!2[I], sum!4[I])       
   sum!4[I] = (sum!3[I] + p!3[I,j])
@@
 
-\ana{TODO: It will be good to add a pass in Phase 3, or post Phase 3, to get rid of extra dimensionality. 
-Once we have vectorized as much as possible, there is no need to keep higher 
-dimensionality. E.g., there is no need for the 2-dimensional array sum in the first for loop, 
-and similarly, there is no need for 1-dimensional arrays for min\_sum and min\_index.
+%\ana{TODO: It will be good to add a pass in Phase 3, or post Phase 3, to get rid of extra dimensionality.
+Once we have vectorized as much as possible, there is no need to keep higher dimensionality.
+E.g., there is no need for the 2-dimensional array sum in the first for loop, and similarly, there is no need for 1-dimensional arrays for min\_sum and min\_index.
 We should get rid of these before generating MOTION code.}
 
-\ana{I don't have a crisp algorithm in mind. I don't think it will be difficult but we have to 
-think this through. Put on our TODO list for us to think about since it will simplify MOTION 
-code tremendously.}
+%\ana{I don't have a crisp algorithm in mind.
+I don't think it will be difficult but we have to think this through.
+Put on our TODO list for us to think about since it will simplify MOTION code tremendously.}
 
 \ana{Begin Work-in-progress notes on implementation. Just some thoughts...} 
-
+\begin{comment}
 For now, I suggest we go with a "naive" implementation and try to do only minimal optimizations. 
@@
 \end{verbatim}
-
+\end{comment}
 There are three kinds of arrays (for now all kept internally as one-dimensional arrays, but that's under discussion).

### 2022-02-20 paper_SIMD.tex
Reorganized, removed TODOS, added other TODOs.

@@
 Our compiler produces the following vectorized program. (Note that this is
-still higher-level IR, our compiler turns this code into MOTION variables 
+still higher-level IR, our compiler turns this code into MOTION variables, 
 loops and SIMD primitives, which MOTION then uses to generate the circuit.)
@@
 min_index!1 = 0
-// S!0^ is same as S!0. C!0 replicates C!0 N-times: 
+// S!0^ is same as S!0. C!0^ replicates C!0 N-times: 
 S!0^ = raise_dim(S!0, ((i * D!0) + j), (i:N!0,j:D!0))
@@
   // Computes N intermediate sums "at once"
-  sum!3[I] = PHI(sum!2^[I], sum!4[I])       
+  sum!3[I] = PHI(sum!2[I], sum!4[I])       
   sum!4[I] = ADD_SIMD(sum!3[I],p!3[I,j])
@@
 \begin{enumerate}
-\item All statements in the program execute using the same protocol (sharing).
-That is, there is no share conversion.
-\ishaq{This is not an assumption, this is our setting.
-We work in the single protocol world.}
-\item All MPC instructions have the same unit cost, 1 unit.
-\ishaq{Don't we need to distinguish between local (cheap) vs.
+\item All statements in the program execute using the same protocol (sharing). That is, there is no share conversion. 
+%\ishaq{This is not an assumption, this is our setting. We work in the single protocol world.}
+\item There are two tiers of MPC instructions, local and remote. A local instruction (essentially just ADD) has cost $\beta$ 
+and a remote instruction (e.g., MUX, MUL, SHL, etc.) has cost $\alpha$, where $\alpha >> \beta$. We assume that all remote 
+instructions have the same cost.
+%\ishaq{Don't we need to distinguish between local (cheap) vs.
 interactive (expensive) instructions?
@@
 -- Past suggestion from Vassilis: Ideally we should assume cost to be a convex function.}\ana{After discussion 1/12: replace single costs 1 with two levels of costs: $\alpha$ for non-local operations, e.g., MUL, and $\beta$ for local ops, e.g., ADD.}
-\item There is unlimited bandwidth---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely 1 unit.\ishaq{Comment from Vassilis: We assume there are infinite parallel capacities, this is the standard assumption in PRAM (Cryptographic Parallel RAM).}\ana{PRAM assumption stays.
+\item We assume infinite parallel capacity---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely $\alpha$ or $\beta$. 
+This is a standard assumption in Cryptographic Parallel RAM. ABY presents empirical support for this assumption~\ana{Add citations. PRAM, ABY.}
+%\ishaq{Comment from Vassilis: We assume there are infinite parallel capacities, this is the standard assumption in PRAM (Cryptographic Parallel RAM).}\ana{PRAM assumption stays.
 We have to replace the unit cost of 1 with a suitable amortization function $f(n)$, which I don't think changes much in the cost modeling and analysis.}
 \item MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption,
-2 MUL instructions scheduled in parallel benefit from amortization and cost 1, however a MUL and a MUX instructions scheduled
-in parallel still cost 2.
-\ishaq{We need to reword this assumption to something like "$K$ parallel MUL costs much less than $K$ (because they get amortized), but any mix of $K$  MUL and MUX still cost roughly $K$.
+2 MUL instructions scheduled in parallel benefit from amortization and cost $\alpha$, however a MUL and a MUX instructions scheduled
+in parallel still cost $2\alpha$. 
+%\ishaq{We need to reword this assumption to something like "$K$ parallel MUL costs much less than $K$ (because they get amortized), but any mix of $K$  MUL and MUX still cost roughly $K$.
 Specifically, we should take away the low constants (2 and 1) because we know for low constants this is not true.} \ana{Again, core assumption that MUL and MUX don't benefit from amortization stays.
@@
 
+\ana{Ishaq?
+Basically, define sequential schedule, then define an equivalent parallel schedule.
+A parallel schedule is equivalent if it preserves def-use relations in sequential schedule, or in other words, schedules def ahead of the use.
+Problem is to minimize cost of Parallel schedule.}
+
 \subsection{Scheduling is NP-hard}
@@
 
+\ana{Ishaq?}
+
 \section{\bf Compiler Framework}
@@
 
-We assume arbitrarily nested loops in the MPC-source IR and read-only arrays.
+\begin{figure*}
+  \includegraphics[width=1.0\linewidth]{figs_paper_SIMD/compiler_framework.png}
+  \caption{Compiler Framework.} 
+  \label{fig:compiler_framework}
+\end{figure*}
+
+%We assume arbitrarily nested loops in the MPC-source IR and read-only arrays.
 We assume that loops range from 0 to some constant N.
 Arrays are linearized (row-major order as in MOTION) and accesses are via functions of the induction variables of the enclosing loops.
-We write $i_1,i_2,...i_k$ to denote the loop nest: $i_1$ is the outermost loop, $i_2$, is immediately nested in $i_1$, and so on until $i_k$.
-As an example, a statement nested in $i_1,i_2,...i_k$ can access array $A[f(i_1,i_2,...i_k)]$.
-We write $A[i_1,i_2,...i_k]$ interchangeably.
-\ana{Need to state this more precisely.}
+
+\figref{fig:compiler_framework} presents an overview of our compiler. In this section, we describe several of the phases of the compiler. 
+Sections~\secref{sec:vectorization} and~\secref{sec:div_and_conq} describe vectorization and divide-and-conquer.
+We write $i,j,k$ to denote the loop nest: $i$ is the outermost loop, $j$, is immediately nested in $i$, and so on until $k$. 
+We write $A[i,j,k]$ to denote canonical access to an array element. \ana{Need to state this more precisely.}
+
+\subsection{Semantic Analysis}
+
+\ana{Benjamin?}
 
@@
 
-\subsection{Def-use Edges}
+\subsection{Dependence Analysis}
+
+\ana{TOTO Ana: Shorten this section. Lots of this is unnecessary.}
+
+\subsubsection{Def-use Edges}
 
@@
 \item same-level forward $X\rightarrow Y$ where $X$ and $Y$ are in the same loop nest $i_1,i_2,...i_k$.
-E.g., $\texttt{d} = \mathit{SUM}(\texttt{S}[i,j], \texttt{C}[j])$ to $\texttt{p} = \mathit{MUL}(\texttt{d},\texttt{d})$ in Biometric is a same-level edge.
+E.g., $\texttt{d} = \mathit{SUB}(\texttt{S}[i,j], \texttt{C}[j])$ to $\texttt{p} = \mathit{MUL}(\texttt{d},\texttt{d})$ in Biometric is a same-level edge.
 A $\phi$ node can be a source of a same-level forward edge but not a target.
@@
 
-\subsection{Helper Functions}
+\subsubsection{Closures}
 
@@
 
+\subsection{Taint Analysis}
+
+\ana{Ben?}
+
+\section{Vectorization}
+\label{sec:vectorization}
+
+\ana{TODO: Still work in progress, all over.}
+
+%\ana{Add back-edges into Phase 1.
+A back-edge from a non-phi-node in loop $i$ to a phi-node in loop $i$'s header is a same-level edge.
+The only difference with the handling of normal same-level forward def-use is that the operand index will become $i-1$.
+A back edge from a phi-node in loop $j$ to a phi-node in loop $i$, where $j$ is nested in $i$ is an inner-to-outer edge and will require dropping dimension.
+We can show these are the only kinds of back-edges that may occur.}
+
+%\subsection{Algorithm (Still work in progress.)}
+
+\subsection{Raise Dimension and Drop Dimension}
+
 We define the $\mathit{raise\_dim}$ (raise dimensions) and $\mathit{drop\_dim}$ (drop dimension) functions.
@@
 \ana{This section is still informal. Needs work to make more precise.}
-
-\section{Vectorization}
-
-\ana{Add back-edges into Phase 1.
-A back-edge from a non-phi-node in loop $i$ to a phi-node in loop $i$'s header is a same-level edge.
-The only difference with the handling of normal same-level forward def-use is that the operand index will become $i-1$.
-A back edge from a phi-node in loop $j$ to a phi-node in loop $i$, where $j$ is nested in $i$ is an inner-to-outer edge and will require dropping dimension.
-We can show these are the only kinds of back-edges that may occur.}
-
-%\subsection{Algorithm (Still work in progress.)}
 
@@
 
-\ishaq{We need to think of a way to handle 4th case for def-use, when a nested block B def is followed by another, different, nested block B' use.
+%\ishaq{We need to think of a way to handle 4th case for def-use, when a nested block B def is followed by another, different, nested block B' use.
 Note: The handwritten notes say we can break it into 2 edges.
@@
 The def' should be placed in a block that contains both B and B'.}
-\ishaq{THE ABOVE COMMENT IS RESOLVED.
+%\ishaq{THE ABOVE COMMENT IS RESOLVED.
 After transformation to SSA, there should be an intermediate variable (phi node) that is assigned a value depending on whether control branched or not.
@@
 the size of those vectors. Note that this is a heuristic that handles a common case, but not all cases of array updates.
-\ana{TODO: Simplify this algorithm, taking into account the restriction to canonical updates. It should be handling \_all\_ cases.}
+%\ana{TODO: Simplify this algorithm, taking into account the restriction to canonical updates. It should be handling \_all\_ cases.}
 
@@
 
-\item same-level $X \rightarrow Y$.
-We do nothing, just propagate the array, which happens to be of the right dimension.
-\ana{There might be some opportunities to do copy propagation optimization and save some cycles, but let's leave this for later.}
+\item same-level $X \rightarrow Y$. We do nothing, just propagate the array, which happens to be of the right dimension. 
+%\ana{There might be some opportunities to do copy propagation optimization and save some cycles, but let's leave this for later.}
 
@@
 \section{Divide-and-Conquer}
+\label{sec:div_and_conq}
 

### 2022-02-21 paper_SIMD.tex
Minor edits.

@@
 % Do not put math or special symbols in the title.
-\title[Scheduling and Amortization for MPC]{Scheduling and Amortization for MPC}
+\title[Scheduling and Vectorization for MPC]{Scheduling and Vectorization for MPC}
 
@@
 % presented. Separate the keywords with commas.
-\keywords{protocol mixing; linear programming; multiparty computation; program analysis; cryptography}
+\keywords{multiparty computation; compilers; cryptography}
 
@@
 
-As a running example, we assume a standard MPC benchmark, Biometric matching.
+As a running example, consider Biometric matching, a standard MPC benchmark.
+Array \texttt{C} is the feature vector of \texttt{D} features that we wish to match and array \texttt{S} 
+is the database of \texttt{N} vectors of size \texttt{D} that we match against.
 An intuitive implementation is as follows:
@@
     min_index = 0  
-    for i in range(N):
+    for i in range(N): #loop over database
         sum = 0
-        for j in range(D):
-            d = S[i * D + j] - C[j]
+        for j in range(D): #loop over features
+            d = S[i * D + j] - C[j] #i.e., d = S[i,j] - C[j]
             p = d * d
@@
 
-Array \texttt{C} is the feature vector of \texttt{D} 
-features that we wish to match and array \texttt{S} is the database of \texttt{N} vectors of size \texttt{D}
-that we match against. The programmer annotates the inputs and outputs as \emph{shared}.
-The program iterates over the entries in the database and computes
+Our compiler takes (essentially) standard IMP syntax.
+The programmer can write intuitive iterative programs as the one above.
+They annotate certain inputs and outputs as \emph{shared}.
+Here the code iterates over the entries in the database and computes
 the sum of squares of the differences of individual features. The program returns the 
@@
 
-Our compiler imposes the following restrictions. We note that in most cases, the restrictions 
+Our compiler imposes the following restrictions. We note that in some cases, the restrictions 
 can be easily lifted and we plan to do so in future iterations of our compiler.
@@
 \item Our compiler allows for output (write) arrays, however it restricts
-write access to \emph{canonical writes} along the dimensions of the array (i.e., \texttt{A[i,j] = ...}
+write access to \emph{canonical writes} along the dimensions of the array. I.e., \texttt{A[i,j] = ...}
 where \texttt{i} and \texttt{j} loop over the two dimensions of \texttt{A} is allowed, 
-but \texttt{A[i,j+2] = ...} is not allowed). Read access is arbitrary. 
+but \texttt{A[i,j+2] = ...} is not allowed. Read access is arbitrary. 
 \end{enumerate}
@@
 
-We linearize the source turning conditionals into MUX statements. The PHI nodes
-are remnants of the SSA IR; our compiler generates code that picks the correct value 
-when producing MOTION output and the MOTION framework in turn linearizes 
-MPC source loops when it generates the circuit.
+The compiler linearizes the source turning conditionals into MUX statements. The PHI nodes
+are remnants of the SSA IR; the compiler generates code that picks the correct value 
+when producing MOTION output; the MOTION framework in turn linearizes 
+loops when it generates the circuit.
 
@@
 
-Our key contribution is the vectorizing transformation. We can compute all $N*D$ 
+A key contribution is the vectorizing transformation. We can compute all $N*D$ 
 subtraction operations (line 9) in a single SIMD instruction; similarly we can compute 
@@
 vectorize computation of the N individual sums, we can compute the N sums in parallel. 
-Our compiler automatically detects these opportunities and transforms the program. 
+Our compiler \emph{automatically detects these opportunities and transforms the program}. 
 It is standard that MPC researchers write vectorized versions of the Biometric
@@
 
-\subsection{Vectorization and Cost of Schedule}
+\subsection{Vectorized MPC Source and Cost of Schedule}
 
 Our compiler produces the following vectorized program. (Note that this is
-still higher-level IR, our compiler turns this code into MOTION variables, 
+still higher-level IR, Vectorized MPC Source. Our compiler turns this code into MOTION variables, 
 loops and SIMD primitives, which MOTION then uses to generate the circuit.)

### 2022-02-23 paper_SIMD.tex
Minor TODO.
wip: problem statement

@@
 
+We have def-use dependency graph $G(V, E)$ for the input program.
+Each vertex/node in this graph represents an operation/instruction in the program and an edge represents dependence i.e.
+an edge from $V_1$ to $V_2$ means the result of operation in $V_1$ is used in the operation in $V_2$.
+To build a parallel schedule, we translate this graph $G(V, E)$ into another dependency graph $G'(V', E')$ such that:
+
+\begin{enumerate}
+    \item We collapse multiple vertices of the same type (i.e. same operation) in $V$ into a single vertex in $V'$.
+    \item dependencies are not broken i.e.
+If there is an edge from $V_1$ to $V_2$ ($V_1, V_2 \in V$), and we put them in vertices $V'_1, V'_2 \in G'$ respectively, then an edge must exist from $V'_1$ to $V'_2$.
+\end{enumerate}
+
+We want to build a graph $G'$ with fewest possible vertices since fewer vertices means more parallelization.
+\ishaq{TODO: make it more concrete, pick some variables e.g.
+$\alpha, \beta$ like above and show 1) parallel schedule will have lower cost (this should be easy).
+and 2) some sort of optimality?
+this would be hard}
+
 \subsection{Scheduling is NP-hard}
@@
 
-\ana{Ishaq?}
+\ana{Ishaq?}\ishaq{TODO: As a first step, pull the super-sequence proof from Lindsay's thesis? }
 
@@
 
-\subsection{Pseudo $\phi$-nodes}
+\subsection{MUX Nodes and Pseudo $\phi$-nodes}
+
+\ana{Benjamin?}
+
+\ana{I think we need to add more here. Conversion from SSA to MPC Source is not so trivial.}
 

### 2022-02-27 paper_SIMD.tex
Improved Extension of Baisic Vectorization with Array Writes.

@@
 
-\subsection{Towards Extension of Basic Vectorization}
+\subsection{Extension of Basic Vectorization}
 
@@
 
+\begin{verbatim}
+for i in range(N):
+  A[i] = B[i] + 10;
+  B[i] = A[i] * D[i-1];
+  C[i] = A[i] * D[i-1];
+  D[i] = B[i] * C[i];
+\end{verbatim}
+
+In Cytron's SSA this code (roughly) translates into
+
+\begin{verbatim}
+for i in range(N):
+1. A_1 = PHI(A_0,A_2)
+2. B_1 = PHI(B_0,B_2)
+3. C_1 = PHI(C_0,C_2)
+4. D_1 = PHI(D_0,D_2)
+5. A_2 = update(A_1, i, B_1[i] + 10); 
+6. B_2 = update(B_1, i, A_2[i] * D_1[i-1]);
+7. C_2 = update(C_1, i, A_2[i] * D_1[i-1]);
+8. D_2 = update(D_1, i, B_2[i] * C_2[i]);
+\end{verbatim}
+
+\begin{comment}
 \begin{algorithmic}
@@
 \end{algorithmic}
-
-In Cytron's SSA this code (roughly) translates into
 
@@
 \end{algorithmic}
-
-There is a cycle around $\texttt{B}_0$ = $\phi(\texttt{B},\texttt{B}_1)$ that includes statement $\texttt{A}_1$[i] = $\texttt{B}_0$[i] + 10;
-and that statement won't be vectorized even though in fact there is no loop-carried dependency from the write of $\texttt{B}_1$[i] at 8 to the
-read of $ ... = \texttt{B}_0$[i] at 6.
+\end{comment}
+
+There is a cycle around \texttt{B\_1 = PHI(B\_0,B\_2)} that includes statement 
+\texttt{A\_1 = update(A\_0, i, B\_1[i] + 10)} and that statement won't be vectorized even 
+though in fact there is no loop-carried dependency from the write of \texttt{B\_1[i]} at 6 to the
+read of \texttt{ ... = B\_1[i]} at 8.
 
@@
 There is a single loop, $i$.
-Clearly, there is no pair $\underline{i}$ and $\underline{i}'$, where $\underline{i} < \underline{i}'$ that make $\underline{i} = \underline{i}'$ (due to the def-use pairs of \texttt{A} 6-8 and 6-10).
-Therefore, we remove the back edge from 6 to 1.
-Analogously, we remove the back edges from 8 to 2 and 10 to 3.
-However, there are many values $\underline{i} < \underline{i'}$ that make $\underline{i} = \underline{i'}-1$ and the back edge from 12 to 4 remains (def-use pairs for \texttt{D}).
-As a result of removing these spurious edges, Basic Vectorization will find that statement 6 is vectorizable.
-Statements 8, 10 and 12 will correctly appear in the FOR loop.
+Clearly, there is no pair $\underline{i}$ and $\underline{i}'$, where $\underline{i} < \underline{i}'$ that make $\underline{i} = \underline{i}'$
+due to the def-use pairs of \texttt{A} 5-6 and 5-7.
+Therefore, we remove the back edge from 5 to the phi-node 1.
+Analogously, we remove the back edges from 6 to 2 and from 7 to 3.
+However, there are many values $\underline{i} < \underline{i'}$ that make $\underline{i} = \underline{i'}-1$ and the back edge from 8 to 4 remains (def-use pairs for \texttt{D}).
+As a result of removing these spurious edges, Basic Vectorization will find that statement 5 is vectorizable.
+Statements 6, 7 and 8 will correctly appear in the FOR loop.
+
+Note however, that this step renders some array phi-nodes target-less. We handle target-less phi-nodes with a minor extension of Basic Vectorization. 
+First, we merge closures that update the same array. \ana{Why, intuition?} Second, we add the target-less node of array \texttt{A} back to the closure that updates \texttt{A} ---
+the intuition is, even if there is no loop-carried dependence from writes to reads on \texttt{A}, \texttt{A} is written and the write (i.e., update) cannot be vectorized; 
+therefore, the updated array has to carry to the next iteration of the loop. Third, in cases when the phi-node remains target-less, i.e., cases when the
+array write can be vectorized, we have to properly remove the phi-node replacing uses of the left-hand side of the phi-node with its arguments.
 
@@
 
-\ana{TODO: I think we should implement this.}
+\ana{TODO: WIP, get to BV without MUX refinement for now? For now, skip MUX refinement.}
 
@@
 
-\subsection{Extension of Basic Vectorization with Array Writes}
-
 \subsubsection{Restricting Array Writes}
 
-For now, we restrict array updates to \emph{canonical updates}. Assume (for simplicity) a two-dimensional array $\texttt{A}[I,J]$. A canonical update is the following:
-\begin{algorithmic}
-\STATE for i in range(I):
-\STATE \hspace{0.25cm} for j in range(J):
-\STATE \hspace{0.55cm} ...
-\STATE \hspace{0.55cm} A[i,j] = ...
-\STATE \hspace{0.55cm} ...
-\end{algorithmic}
-
-The update $\texttt{A}[i,j]$ can be nested into an inner loop and there may be multiple updates, i.e., writes to $\texttt{A}[i,j]$.
-However, update such as $\texttt{A}[i-1,j] = ...$ or $\texttt{A}[i-1,j-1] = ...$, etc., is not allowed.
-Additionally, while there could be several different loops that perform canonical updates, they must be of the same dimensionality, i.e., an update of higher or lower dimension, e.g., $A[i,j,k] = ...$ is not allowed.
-We compute the \emph{canonical dimensionality} of each write array in the obvious way \emph{before basic vectorization}.
+For now, we restrict array updates to \emph{canonical updates}. Assume (for simplicity) a two-dimensional array \texttt{A[I,J]}. A canonical update is the following:
+\begin{verbatim}
+for i in range(I):
+  for j in range(J):
+    ...
+    A[i,j] = ...
+    ...
+\end{verbatim}
+
+The update \texttt{A[i,j]} can be nested into an inner loop and there may be multiple updates, i.e., writes to \texttt{A[i,j]}.
+However, update such as \texttt{A[i-1,j] = ...} or \texttt{A[i-1,j-1] = ...}, etc., is not allowed.
+Additionally, while there could be several different loops that perform canonical updates, they must be of the same dimensionality, i.e., an update of higher or lower dimension, e.g., \texttt{A[i,j,k] = ...} is not allowed.
+We compute the \emph{canonical dimensionality} of each write array by examining the array writes in the original program and rejecting programs that violate the canonical write restriction.
 This restriction simplifies reasoning in this early stage of the compiler; we will look to relax the restriction in future work.
@@
 
-Reads through an arbitrary formula, such as $\texttt{A}[i-1]$ for example, are allowed; currently, our projection function returns dummy values if the read formula is out of bounds; we assume the programmer ensures that the program still computes correct output in this case.
+Reads through an arbitrary formula, such as \texttt{A[i-1]} for example, are allowed; currently, the projection function returns dummy values if the read formula is out of bounds; we assume the programmer ensures that the program still computes correct output in this case.
 
@@
 
+\begin{comment}
 There are two changes to Basic Vectorization to account for write arrays. We can account for them during processing of raise dimension/drop dimension in Phase 1 of Basic Vectorization.
@@
 \ana{TODO: Double check.})
-
-The other change concerns def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable.
-The definition can be an update $A_2 = update(A_1,...)$ or a pseudo $\phi$-node $A_2 = \phi(A_0,A_1)$.
-Note that $\phi$ nodes for arrays are no subscript operations on those $\phi$-nodes the way there are in scalar arrays.
-These edges are not handled in the same way as in Basic Vectorization, specifically, we do not raise and drop dimension as we do for scalars in Basic Vectorization.
-A key invariant is that the dimension of an array $A$ cannot go above or drop below $A$'s canonical dimensionality.
-We enumerate the cases of def-use edges.
+\end{comment}
+
+In addition to the changes for the handling of target-less phi-nodes, Basic Vectorization has to handle differently def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable.
+The definition can be an update \texttt{A\_2 = update(A\_1,i,...)} or a pseudo $\phi$-node \texttt{A\_2 = PHI(A\_0,A\_1)}, etc..
+Note that $\phi$-nodes for arrays have no subscript operations the way there are subscript operations in analysis-introduced arrays representing scalars.
+In addition, array variables are different from scalar variables.
+Raising and dropping dimension still apply, however, the dimension cannot be raised or dropped to a dimension lower than the canonical one:
 
@@
 First, the canonical dimensionality of all \texttt{A,B,C} and \texttt{D} is 1. 
-Thus, there is no addition of extra dimension for inner loops. After Phase 1 
-of Basic Vectorization the Aiken's array write example will be (roughly) as follows:
-
+After Phase 1 of Basic Vectorization the Aiken's array write example will be (roughly) as follows:
+
+\begin{verbatim}
+for i in range(N):
+1. A_1 = PHI(A_0,A_2)
+2. B_1 = PHI(B_0,B_2)
+3. C_1 = PHI(C_0,C_2)
+4. D_1 = PHI(D_0,D_2)
+5. A_2 = update(A_1, I, B_1[I] + 10); 
+6. B_2 = update(B_1, I, A_2[I] * D_1[I-1]);
+7. C_2 = update(C_1, I, A_2[I] * D_1[I-1]);
+8. D_2 = update(D_1, I, B_2[I] * C_2[I]);
+\end{verbatim}
+
+Note that since all def-uses are same-level (i.e., reads and writes of the array elements) 
+no raise dimension or drop dimension happens.
+
+\begin{comment}
 \begin{algorithmic}
@@
 \end{algorithmic}
+\end{comment}
 
 Phase 2 computes the closure of 4; $cl = \{4,6,7, 8\}$ 
-while 5 is vectorizable. Recall that 1,2, and 3 are targetless phi-nodes. Since the closure $cl$ includes updates to $B$ and $C$,
-the corresponding phi-nodes are added to the closure, i.e., the FOR loop. The uses of $A_1$ and $B_1$ in the vectorized statement
-turn into uses of $A_0$ and $B_0$ respectively. (But note that $A_0$ is irrelevant; the updates writes into array $A_2$ in parallel.)
-
+while 5 is vectorizable. Recall that 1,2, and 3 are target-less phi-nodes. Since the closure $cl$ includes updates to \texttt{B} and \texttt{C},
+the corresponding phi-nodes are added back to the closure and the def-use edges are added back to the target-less nodes. The uses of \texttt{A\_1} and \texttt{B\_1} in the vectorized statement
+turn into uses of \texttt{A\_0} and \texttt{B\_0} respectively; this is done for all original target-less phi-node.
+(But note that \texttt{A\_0} is irrelevant; the update writes into array \texttt{A\_2} in parallel.)
+Finally, the target-less phi-node for \texttt{A} is discarded.
+
+\begin{verbatim}
+1. A_2 = update(A_0, I, ADD_SIMD(B_0[I],10));
+   equiv. to A_2[I] = ADD_SIMD(B_0[I],10)
+FOR i=0; i<N; i++; // MOTION loop
+2. B_1 = PHI(B_0,B_2)
+3. C_1 = PHI(C_0,C_2)
+4. D_1 = PHI(D_0,D_2)
+5. B_2 = update(B_1, i, A_2[i] * D_1[i-1]);
+   equiv. to B_2 = B_1; B_2[i] = A_2[i] * D_1[i-1];
+6. C_2 = update(C_1, i, A_2[i] * D_1[i-1]);
+7. D_2 = update(D_1, i, B_2[i] * C_2[i]);
+\end{verbatim}
+
+\begin{comment}
 \begin{algorithmic}
@@
 \end{algorithmic}
-
-\ana{TODO: Format either Histogram or Matrix Multiplication. Both are interesting examples that use output arrays.}
+\end{comment}
+
+\paragraph{Example 2} Now consider the MPC Source of Histogram:
+
+\begin{verbatim}
+for i in range(0, num_bins):
+   res1 = PHI(res, res2)
+   for j in range(0, N):
+      res2 = PHI(res1, res3)
+      tmp1 = (A[j] == i)
+      tmp2 = (res2[i] + B[j])
+      tmp3 = MUX(tmp1, res2[i], tmp2)
+      res3 = Update(res2, i, tmp3)     
+return res1
+\end{verbatim}
+
+The canonical dimensionality of \texttt{res} is 1.
+Also, the phi-node \texttt{res2 = PHI(res1, res3)} is a target-less phi-node (the implication being that the inner for loop can be vectorized across $i$).
+After Phase 1, Basic vectorization produces the following (everything is implicitly vectorized along $i$ and $j$ and the update is turned into a regular write):
+
+\begin{verbatim}
+A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
+B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
+I = raise_dim(i, ((i:num_bins),(j:N)))
+for i in range(0, num_bins):
+   res1 = PHI(res, res2) // target-less, no drop_dim
+   res1' = raise_dim(res1, (j:N))
+   for j in range(0, N):
+      res2 = PHI(res1', res3)
+      tmp1 = (A1 == I)
+      tmp2 = (res2 + B1)
+      tmp3 = MUX(tmp1, res2, tmp2)
+      res3 = res2 // copy propagation: res2 into res3
+      res3 = tmp3 // "normal": res3[I,J] = tmp3[I,J]    
+return res1
+\end{verbatim}
+
+Processing the inner loop in Phase 2 vectorizes \texttt{tmp1 = (A1 == I)} along the $j$ dimension but leaves the rest of the statements in a MOTION loop.
+Processing the outer loop is interesting.
+This is because the PHI node is a target-less node, and therefore, there are no closures!
+Thus everything can be vectorized along the $i$ dimension:
+
+\begin{verbatim}
+A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
+B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
+I1 = raise_dim(i, ((i:num_bins),(j:N)))
+
+tmp1[I,J] = (A1[I,J] == I1[I,J])
+
+res1' = raise_dim(res, (j:N)) // raising original array to a 2-dimensional array
+for j in range(0, N):
+   res2 = PHI(res1', res3)
+   tmp2[I,j] = (res2[I,j] + B1[I,j])
+   tmp3[I,j] = MUX(tmp1[I,j], res2[I,j], tmp2[I,j])
+   res3 = res2
+   res3[I,j] = tmp3[I,j]
+res1 = drop_dim(res2) // Have to add drop_dim before removing redundant phi-node      
+return res1
+\end{verbatim}
+
+The cleanup phase can then get rid of $j$ in the res variables and tmp variables, so the loop operates over the num\_bins-size vectors. 
+This will lead to dropping raise\_dim and drop\_dim as well:
+
+\begin{verbatim}
+A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
+B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
+I1 = raise_dim(i, ((i:num_bins),(j:N)))
+
+tmp1[I,J] = (A1[I,J] == I1[I,J])
+
+for j in range(0, N):
+   res2 = PHI(res1', res3) // now 1-dim array
+   tmp2[I] = (res2[I] + B1[I,j])
+   tmp3[I] = MUX(tmp1[I,j], res2[I], tmp2[I])
+   res3 = res2
+   res3[I] = tmp3[I]
+return res2
+\end{verbatim}
 

### 2022-02-28 paper_SIMD.tex
Minor formatting.

@@
 
-res1' = raise_dim(res, (j:N)) // raising original array to a 2-dimensional array
+res1' = raise_dim(res, (j:N)) // raising res to a 2-dim. array
 for j in range(0, N):

### 2022-03-01 paper_SIMD.tex
Minor edits to Array section.

@@
 
-\COMMENT { Phase 3: Dimensionality reduction: removing unnecessary dimensionality. }
+\COMMENT { Phase 3: Remove unnecessary dimensionality. }
 
@@
 Note however, that this step renders some array phi-nodes target-less. We handle target-less phi-nodes with a minor extension of Basic Vectorization. 
-First, we merge closures that update the same array. \ana{Why, intuition?} Second, we add the target-less node of array \texttt{A} back to the closure that updates \texttt{A} ---
+First, we merge closures that update the same array. This simplifies handling of array $\phi$-nodes: if each closure is turned into a separate loop
+each loop will need to have its own array phi-node to account for the update and this would complicate the analysis.
+Second, we add the target-less node of array \texttt{A} back to the closure that updates \texttt{A} ---
 the intuition is, even if there is no loop-carried dependence from writes to reads on \texttt{A}, \texttt{A} is written and the write (i.e., update) cannot be vectorized; 

### 2022-03-01 paper_SIMD.tex
Removed todos.

@@
 Sections~\secref{sec:vectorization} and~\secref{sec:div_and_conq} describe vectorization and divide-and-conquer.
-We write $i,j,k$ to denote the loop nest: $i$ is the outermost loop, $j$, is immediately nested in $i$, and so on until $k$. 
-We write $A[i,j,k]$ to denote canonical access to an array element. \ana{Need to state this more precisely.}
+We write $i,j,k$ to denote the loop nest: $i$ is the outermost loop, $j$, is immediately nested in $i$, and so on until $k$
+and we use $I,J,K$ to denote the corresponding upper bounds. For simplicity, we write $A[i,j,k]$ to denote canonical access 
+to an array element. In the program, canonical access is achieved via the standard row-major order formula: $(J*K)*i + K*j + k$.
+To simplify the presentation we describe our algorithms in terms of three-element tuples $i,j,k$.
+All discussion generalizes to arbitrarily large loop nests.
+%\ana{Need to state this more precisely.}
 
@@
 
-\ana{TODO: Still work in progress, all over.}
+An important component of our algorithm is the "lifting" of scalars to the corresponding loop dimensionality. 
+For example, \texttt{d = S[i * D + j] - C[j]} equiv. to \texttt{d = S[i,j] - C[j]}, which gave rise to \texttt{N*D} subtraction operations in the sequential schedule, 
+is lifted. The argument arrays \texttt{S} and \texttt{C} are lifted and the scaler \texttt{d} is lifted: \texttt{d[i,j] = S[i,j] - C[i,j]}.
+The algorithm then detects that the statement can be vectorized.
+
+There are three kinds of arrays (for now all kept internally as one-dimensional arrays, but that's under discussion).
+\begin{itemize}
+
+\item Scalars: These are scalar variables we lift into arrays for the purposes of vectorization. 
+For those, all writes are canonical writes and all reads are canonical reads. We will apply raise dimension when 
+a scalar is used in an inner loop (e.g., \texttt{sum0} in line 6 of the MPC source code will be raised to a 1-dimensional 
+array since \texttt{sum0} is used in the inner $j$-loop). Drop dimension applies as well; this happens when a scalar
+written in an inner loop is used outside of the loop (e.g., \texttt{sum0} for which the lifted inner loop computes 
+\texttt{D} values, but the outer loop only needs the last one.)
+
+\item Read-only input arrays: Read-only inputs. There are NO writes, while we may have non-canonical reads, $f(i,j,k)$. 
+Phase 1 of Basic vectorization will add raise dimension operation at the beginning of the function to lift these arrays, 
+and raise dimension may \emph{reshape} arrays. If there are multiple "views" of the input array, there would be multiple raise dimension statements to create
+each one of these views. The invariant is that at reads in loops, the reads of "views" of the original input array are canonical.
+Only raise dimension applies to these arrays, and only in the beginning of the program. For example, 1-dimensional array 
+\texttt{C} is lifted into 2-dimensional array \texttt{N x D} by copying the row \texttt{N} times.
+
+ \item Read-write output arrays: Writes are canonical (by restriction) but reads can be non-canonical. 
+ Dependence analysis limits vectorization when non-canonical read access refers to array writes in
+ previous iterations, thus creating loop-carried dependences. We may apply both raise and drop dimension, however, 
+ they respect the fixed dimensionality of the output array. The array cannot be raised to a dimension lower than
+ its canonical (fixed) dimensionality and it cannot be dropped lower. In addition, non-canonical reads may 
+ require lifting (i.e., reshaping) of the array after the most recent write, rather than in the beginning of the program, 
+ in order to reduce a non-canonical read to a canonical one.
+
+\end{itemize}
+
+In the sections below we detail the $\mathit{raise\_dim}$ (raise dimension) and $\mathit{drop\_dim}$ (drop dimension) operations,
+followed by our vectorization algorithm.
 
@@
 
+There are two conceptual versions of $\mathit{raise\_dim}$. One applies on read-only input arrays and reshapes those arrays 
+when necessary to ensure canonical read access in the corresponding loop. The signature of $\mathit{raise\_dim}$
+is as follows. It takes the original array \texttt{C}, the access pattern function $f(i,j,k)$ in loop nest $i,j,k$ and the loop bounds
+\texttt{((i:I),(j:J),(k:K))}: 
+\[ \mathit{raise\_dim}(\texttt{A}, f(i,j,k), \texttt{((i:I),(j:J),(k:K))}) \]
+It produces a new 3-dimensional array \texttt{A'} by iterating over $i,j,k$ and setting each element of  \texttt{A'} as follows:
+\[ \texttt{A'[i,j,k] =  A[f(i,j,k)]} \]
+The end result is that uses of \texttt{[A[f(i,jk)]} in loop nest $i,j,k$ are replaced with canonical read-accesses to \texttt{A'[i,j,k]}
+that can be vectorized. In the running Biometric example, \texttt{C'} = $\mathit{raise\_dim}$\texttt{(C, j, (i:N,j:D))} lifts the 
+1-dimensional array \texttt{C} into a 2-dimensional array. The $i,j$ loop now accesses \texttt{C'} in the canonical way, \texttt{C'[i,j]}. 
+Similarly, \texttt{S'} = $\mathit{raise\_dim}$\texttt{(S, i*D+j, (i:N,j:D))} tries to lift \texttt{S}, but the operation turns into a no-op 
+because \texttt{S} is already a 2-dimensional array and the read access is canonical.
+
+The other version of $\mathit{raise\_dim}$ applies on scalars and read-write arrays. It lifts a lower-dimension array into a 
+higher-dimension for access in a nested loop. Here \texttt{A} is an $i$ array and raise dimension adds two additional dimensions:
+\[ \mathit{raise\_dim}(\texttt{A}, \texttt{(j:J,k:K)}) \]
+This version is reduced to the above version by adding the access pattern function, which is just $i$: 
+\[ \mathit{raise\_dim}(\texttt{A}, i, \texttt{(j:J,k:K)}) \]
+
+The corresponding $\mathit{drop\_dim}$ is carried out when an array written in an inner loop is used in an enclosing loop. 
+It takes a higher dimensional array, say $i,j,k$ and removes trailing dimensions, say $j,k$:
+\[ \mathit{drop\_dim}(\texttt{A}, \texttt{(j:J,k:K)}) \]
+It iterates over $i$ and takes the result at the maximal index of $j$ and $k$, i.e., the result at the last iterations of $j$ and $k$:
+\[ \texttt{A'[i] = A[i,J-1,K-1]} \]
+
+\begin{comment}
 We define the $\mathit{raise\_dim}$ (raise dimensions) and $\mathit{drop\_dim}$ (drop dimension) functions.
@@
 \ana{This section is still informal. Needs work to make more precise.}
+\end{comment}
 

### 2022-03-01 paper_SIMD.tex
More edits, removed more TODOs.

@@
 
-\ana{TOTO Ana: Shorten this section. Lots of this is unnecessary.}
-
 \subsubsection{Def-use Edges}
@@
 \begin{itemize}
-\item same-level forward $X\rightarrow Y$ where $X$ and $Y$ are in the same loop nest $i_1,i_2,...i_k$.
-E.g., $\texttt{d} = \mathit{SUB}(\texttt{S}[i,j], \texttt{C}[j])$ to $\texttt{p} = \mathit{MUL}(\texttt{d},\texttt{d})$ in Biometric is a same-level edge.
-A $\phi$ node can be a source of a same-level forward edge but not a target.
-\item outer-to-inner forward $X\rightarrow Y$ where $X$ is in an outer loop nest, $i_1,i_2,...i_j$, and $Y$ is in an inner one, $i_1,i_2,...i_j,...,i_k$.
-A $\phi$-node can be a source or a target of an outer-to-inner forward edge.
-\item inner-to-outer forward $X\rightarrow Y$ where $X$ is a $phi$-node in an inner loop nest, $i_1,i_2,...i_k,i_{k+1}$, and $Y$ is in the enclosing loop nest $i_1,i_2,...i_k$.
+\item same-level edge $X\rightarrow Y$ where $X$ and $Y$ are in the same loop nest, say $i,j,k$.
+E.g., the def-use edge from \texttt{d = S[i,j] - C[j]} to \texttt{p = d*d} in the Biometric MPC-source is a same-level edge.
+A same-level edge can be a back-edge in which case a $\phi$ node is the target of the edge.
+E.g., $\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{sum}_1,\texttt{min}_1)$ to $\texttt{min}_0 = \phi(\texttt{min}_1, 10000)$ in Biometric is a same-level back-edge.
+\item outer-to-inner $X\rightarrow Y$ where $X$ is in an outer loop nest, say $i$, and $Y$ is in an inner one, say $i,j,k$. 
+%A $\phi$-node can be a source or a target of an outer-to-inner forward edge.
+\item inner-to-outer $X\rightarrow Y$ where $X$ is a $phi$-node in an inner loop nest, $i,j,k$, and $Y$ is in the enclosing loop nest $i,j$.
 E.g.
-$\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$ is an inner-to-outer forward edge.
-Note that the source is \emph{always} a $\phi$-node in the immediately enclosing loop.
+$\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$ is an inner-to-outer edge.
+An inner-to-outer edge can be a back-edge as well in which case both $X$ and $Y$ are phi-nodes with the source $X$ in a loop nested into $Y$'s loop (not necessarily immediately).
+
+%Note that the source is \emph{always} a $\phi$-node of the $i,j,k$ loop and the target is in the immediately enclosing loop.
 The interpretation of this edge is that the use node $Y$ uses the definition made in the last iteration of the inner loop.
-\ben{The current representation of pseudo $\phi$-nodes shows them attached to the loop header (i.e. in the loop).  We may want to clarify that they are also evaluated at the loop's termination.}
-\item same-level back-edge $X\rightarrow Y$.
+%\ben{The current representation of pseudo $\phi$-nodes shows them attached to the loop header (i.e. in the loop).  We may want to clarify that they are also evaluated at the loop's termination.}
+%\item same-level back-edge $X\rightarrow Y$.
 $Y$ is a phi-node in the header of the loop and $X$ is a definition of the variable in the loop body.
 E.g., $\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{sum}_1,\texttt{min}_1)$ to $\texttt{min}_0 = \phi(\texttt{min}_1, 10000)$ in Biometric is a same-level back-edge.
-\item inner-to-outer back-edge $X\rightarrow Y$: $X$ and $Y$ are both $\phi$-nodes for some variable. The source $X$ is in a loop nested into $Y$'s loop (not necessarily immediately).
+%\item inner-to-outer back-edge $X\rightarrow Y$: $X$ and $Y$ are both $\phi$-nodes for some variable. The source $X$ is in a loop nested into $Y$'s loop (not necessarily immediately).
 
 \item mixed forward edge $X\rightarrow Y$.
-$X$ is a $phi$-node in some loop $i_1,i_2,...i_k,i_{k+1}$ and $Y$ is a node in a loop nested into $i_1,i_2,...i_k$.
+$X$ is in some loop $i,j,k$ and $Y$ is in a loop nested into $i,j,k'$.
 We transform mixed forward edges as follows.
-Let $\x_j$ be the variable defined at the phi-node $X$.
-We add a variable and assignment $\x'_j = \x_j$ immediately after the $i_1,i_2,...i_k$ loop.
-Then we replace the use of $\x_j$ at $Y$ with $\x'_j$.
+Let $\x$ be the variable defined at $X$.
+We add a variable and assignment $\x' = \x$ immediately after the $i,j,k$ loop.
+Then we replace the use of $\x$ at $Y$ with $\x'$.
 This transforms a mixed forward edge into an "inner-to-outer" forward edge followed by an outer-to-inner forward edge.
-Thus Basic Vectorization handles one of "same-level", "inner-to-outer", or "outer-to-inner" def-use edges.
+Thus, Basic Vectorization handles one of "same-level", "inner-to-outer", or "outer-to-inner" def-use edges.
  \end{itemize}
@@
 
-We define $\mathit{closure}(n)$ where $n$ is a phi-node.
-Intuitively, it computes the set of nodes (i.e., statements) that form a dependence cycle with $n$.
-\ana{Cycle(n) is probably a better name.} The closure of $n$ is defined as follows:
+We define $\mathit{closure}(n)$ where $n$ is a phi-node. Intuitively, it computes the set of nodes (i.e., statements) that form a dependence cycle with $n$. The closure of $n$ is defined as follows:
 \begin{itemize}
@@
 
-\subsection{Extension of Basic Vectorization}
+\ana{Add the AI-based proof here.}
+
+\subsection{Extension of Basic Vectorization with Array Writes}
 

### 2022-03-01 paper_SIMD.tex
wip on problem statement and costs

--- 2022-03-01 paper_SIMD.tex
@@
 \usepackage[backend=biber,style=ACM-Reference-Format]{biblatex}
-\addbibresource{../library.bib}
+\addbibresource{library.bib}
 \addbibresource{../cryptobib/abbrev2.bib}
@@
 \usepackage{algorithmic}
+
+% for references, use it like this: \cref{some_label}
+% it will automatically put Appendix, Figure, Section, etc before the reference
+\usepackage[capitalise]{cleveref}
 
@@
 
-We have def-use dependency graph $G(V, E)$ for the input program.
+At the lowest level, we have two types of MPC instructions 1) local/non-interactive instruction (i.e.
+\add) and 2) remote/interactive instruction (i.e.
+\mul).
+Let the cost of a single \add instruction be given by a monotonically decreasing $f(n)$, where the argument $n$ is the number of \add instructions being executed in parallel.
+Similarly the cost of a single \mul is given by a monotonically decreasing function $g(n)$.
+
+Given a serial schedule (a linear graph) of an MPC program i.e.
+a sequence of instructions $\mathcal{S}:=(S_1, \dots, S_n)$, where $S_i \in \{\text{\add, \mul}\}, 1 \leq i \leq n$, and a def-use dependency graph $G(V, E)$ corresponding to $\mathcal{S}$, our task is to construct a parallel schedule (another linear graph) $\mathcal{P} := (P_1, \dots, P_n)$ observing the following conditions:
+
+\begin{enumerate}
+    \item Multiple, not necessarily continuous instructions of the same type (i.e.
+either \add or \mul) from $\mathcal{S}$ can be grouped into a single $P_i$  However, all such instructions must be of the same type (either \add or \mul).
+    \item Def-use dependencies of the graph $G(V, E)$ are should be preserved i.e.
+if instructions $S_i, S_j, i < j$ are a def-use (an edge exists from $S_i$ to $S_j$ in $G$), then they can only be mapped to $P_{i'}, P_{j'}, i' < j'$.
+\end{enumerate}
+
+Our goal is to construct minimize the height of the graph $\mathcal{P}$. Indeed, a graph $\mathcal{P}$ with minimum height will maximize parallelization.
+
+\paragraph{Correctness} Correctness of $\mathcal{P}$ is guaranteed by definition. Since def-use dependencies are preserved, the function (being computed) remains the same.
+
+\paragraph{Cost Comparison} For the sequential schedule $\mathcal{S}$ consisting of $L$ local and $R$ remote instructions, the total cost is $\mathit{cost}(\mathcal{S}) = L \cdot f(1) + R \cdot g(1)$.
+In the extreme case where all $L$ and all $R$ instructions can be parallelized, the cost of $\mathcal{P}$ is $\mathit{cost}(\mathcal{P}) = L \cdot f(L) + R \cdot g(R)$.
+Since both $f$ and $g$ are monotonically decreasing, $\mathit{cost}(\mathcal{P}) < \mathit{cost}(\mathcal{S})$.
+Cost of all other parallel schedules lies between the extremes of $\mathit{cost}(\mathcal{S})$ and $\mathit{cost}(\mathcal{P})$.
+
+Note that we use an MPC-Source control flow graph (CFG) $G'(V', E')$ along with def-use graph $G(V,E)$ to construct $\mathcal{P}$.
+We consider a linearized MPC schedule $\mathcal{S}$ above for ease of exposition.
+The argument becomes slightly more involved when dealing with a graph $G'$ that may contain cycles.
+
+% Given a def-use dependency graph $G(V, E)$ that is constructed from for the input program.
 Each vertex/node in this graph represents an operation/instruction in the program and an edge represents dependence i.e.
@@
 
-\begin{enumerate}
-    \item We collapse multiple vertices of the same type (i.e. same operation) in $V$ into a single vertex in $V'$.
-    \item dependencies are not broken i.e.
+% \begin{enumerate}
+%     \item We collapse multiple vertices of the same type (i.e. same operation) in $V$ into a single vertex in $V'$.
+%     \item dependencies are not broken i.e.
 If there is an edge from $V_1$ to $V_2$ ($V_1, V_2 \in V$), and we put them in vertices $V'_1, V'_2 \in G'$ respectively, then an edge must exist from $V'_1$ to $V'_2$.
-\end{enumerate}
-
-We want to build a graph $G'$ with fewest possible vertices since fewer vertices means more parallelization.
-\ishaq{TODO: make it more concrete, pick some variables e.g.
-$\alpha, \beta$ like above and show 1) parallel schedule will have lower cost (this should be easy).
-and 2) some sort of optimality?
-this would be hard}
+% \end{enumerate}
+
+% We want to build a graph $G'$ with fewest possible vertices since fewer vertices means more parallelization.
+
+\ishaq{TODO: @Ana: I am still confused about what prevents a reviewer from saying, "you argue correctness of the schedule from unrolled MPC, you don't argue it for the pre-unrolled CFG that you use".
+Please comment.}
 
@@
 
-\ana{Ishaq?}\ishaq{TODO: As a first step, pull the super-sequence proof from Lindsay's thesis? }
+\ana{Ishaq?}
+
+\ishaq{The follow text is all from Linsay's thesis}.
+\brevised{
+We consider two operations, call them $A$ and $M$. $A$ and $M$ are two abstract MPC instruction, 
+but as an example, $A$ stands for the ADD MPC instruction 
+and $M$ stands for the MUL instruction. Each instruction in the program is either an $A$-instruction or
+an $M$-instruction. In order to benefit from parallelization/amortization, we must schedule two or more 
+$A$-instructions in the same parallel node (or two or more $M$-instructions in the same parallel node). 
+Scheduling $A$-instructions in parallel with $M$-instruction does not benefit from amortization.
+It incurs the exact same cost as scheduling the $A$-instructions in a node $P_A$, scheduling the $M$-instructions 
+in a node $P_M$, and having $P_A$ precede $P_M$ in the parallel schedule. This is the difference 
+between classical scheduling, as studied in parallel computing, and MPC scheduling.
+
+In our theoretical treatment, we make the following assumptions:
+
+\begin{enumerate}
+
+\item $A$ and $M$ are of equal cost, 1 unit.
+\item There is unlimited bandwidth---i.e., a single $A$-instruction (or $M$-instruction) costs as much as $N$ amortized $A$-instructions 
+(or $M$-instructions), namely 1 unit.
+
+\end{enumerate}
+
+%  Furthermore the run time cost of 
+% computing operations of the same type remains the same regardless of the number of operations that are run.
+% Roughly speaking, 1 A operation costs the same as 100 A operations and 1000 A operations.
+
+Consider a loop body that consists of $n$ sequences: $S_1$, ... $S_n$ of $A$ and $M$ instructions. 
+More precisely, the loop body is such that its instructions can be grouped into such sequences. 
+$S_1$, ... $S_n$ can execute in parallel, however, all instructions within a sequence must 
+execute sequentially. For example, consider the three sequences (the right arrow indicates a \emph{dependence},
+meaning that the source node must execute before the target node): 
+\begin{enumerate}
+\item $A \rightarrow M \rightarrow A$
+\item $A \rightarrow A \rightarrow A$
+\item $M \rightarrow A \rightarrow M$
+\end{enumerate}
+
+A \emph{schedule} $P: P_1 \rightarrow P_2 \dots \rightarrow P_k$ is such that for each sequence 
+$S_i$ in the set, if $S_i[k]$ precedes $S_i[k']$ in $S_i$ then $S_i[k]$ is scheduled in node $P_l$, $S_i[k]$ 
+is scheduled in node $P_{l'}$, and $P_l$ precedes $P_{l'}$ in $P$.
+
+The cost of a schedule $P$ is
+
+\begin{equation}
+\mathit{cost}(P) = \sum_{i=1}^k \mathit{cost}(P_i)
+\end{equation}
+
+where $\mathit{cost}(P_i) = 1$ if $P_i$ consists of $A$-instructions only, or $M$-instructions only, 
+and $\mathit{cost}(P_i) = 2$ if $P_i$ mixes $A$-instructions and $M$-instructions.
+
+The problem is to find a schedule $P$ with \emph{minimal cost}. For example, 
+a schedule with minimal cost for the sequences above is
+\[ A(1), A(2) \rightarrow M(1), A(2), M(3) \rightarrow A(1), A(2), A(3) \rightarrow M(3) \]
+(The parentheses above indicate the sequence where the instruction comes from: (1), (2), or (3).)
+The cost of this schedule is 5.
+
+The problem of finding a schedule $P$ with a minimal $cost(P)$ for a given loop body has been shown
+to be an NP-Hard problem, as it can be reduced to the problem of finding a \emph{shortest common supersequence}, 
+a known NP-Hard problem\cite{Maier1978},\cite{Vazirani2010}. The shortest common supersequence problem is as follows: 
+given two or more sequences find the the shortest sequence that contains all of the original sequences. This can be solved
+in $O(n^k)$ time, where $n$ is the cardinality of the longest sequence and $k$ is the 
+number of sequences. For our problem $n$ is the maximum length of a node and $k$ is the number
+of total number of nodes. %\ana{I think it is clear without additional explanation...}
+%\ana{Also, add the full statement of the problem here. Now it's not clear what's NP-hard, what's n, what's k, etc.?}
+%\lindsey{not exactly sure what need to go it there, but does that make it clearer?}
+
+To see the reduction, 
+suppose $P$ is a schedule with minimal cost (computed by a black-box algorithm). 
+We can derive a schedule $P'$ with the same cost as $P$, by mapping each mixed node $P_i \in P$ 
+to two consecutive nodes in $P'$: an $A$-instruction node followed by an $M$-instruction node.
+Clearly, $P'$, which now is a sequence of $A$'s and $M$'s, is a supersequence of each sequence 
+$S_i$, i.e., $P'$ is a common supersequence 
+of $S_1 \dots S_n$. It is also a shortest common supersequence. (To see this, suppose, a 
+shorter common supersequence, $P''$, exists. $P''$ is a schedule of $S_1 \dots S_n$
+and the $\mathit{cost}(P'')$ equals the length of $P''$. Since $P'$ is longer than
+$P''$, and $\mathit{cost}(P') = \mathit{cost}(P)$, that means $\mathit{cost}(P'') < \mathit{cost}(P)$, 
+which is a contradiction since $P$ is a schedule with minimal cost.)
+}
 
@@
 
-\COMMENT { Phase 3: Remove unnecessary dimensionality. }
+\COMMENT { Phase 3: Dimensionality reduction: removing unnecessary dimensionality. }
 
@@
 
-\subsection{Extension of Basic Vectorization}
+\subsection{Towards Extension of Basic Vectorization}
 
@@
 
-{\small
-\begin{verbatim}
-for i in range(N):
-  A[i] = B[i] + 10;
-  B[i] = A[i] * D[i-1];
-  C[i] = A[i] * D[i-1];
-  D[i] = B[i] * C[i];
-\end{verbatim}
-}
-
-In Cytron's SSA this code (roughly) translates into
-
-{\small
-\begin{verbatim}
-for i in range(N):
-1. A_1 = PHI(A_0,A_2)
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. A_2 = update(A_1, i, B_1[i] + 10); 
-6. B_2 = update(B_1, i, A_2[i] * D_1[i-1]);
-7. C_2 = update(C_1, i, A_2[i] * D_1[i-1]);
-8. D_2 = update(D_1, i, B_2[i] * C_2[i]);
-\end{verbatim}
-}
-
-\begin{comment}
 \begin{algorithmic}
@@
 \end{algorithmic}
+
+In Cytron's SSA this code (roughly) translates into
 
@@
 \end{algorithmic}
-\end{comment}
-
-There is a cycle around \texttt{B\_1 = PHI(B\_0,B\_2)} that includes statement 
-\texttt{A\_1 = update(A\_0, i, B\_1[i] + 10)} and that statement won't be vectorized even 
-though in fact there is no loop-carried dependency from the write of \texttt{B\_1[i]} at 6 to the
-read of \texttt{ ... = B\_1[i]} at 8.
+
+There is a cycle around $\texttt{B}_0$ = $\phi(\texttt{B},\texttt{B}_1)$ that includes statement $\texttt{A}_1$[i] = $\texttt{B}_0$[i] + 10;
+and that statement won't be vectorized even though in fact there is no loop-carried dependency from the write of $\texttt{B}_1$[i] at 8 to the
+read of $ ... = \texttt{B}_0$[i] at 6.
 
@@
 There is a single loop, $i$.
-Clearly, there is no pair $\underline{i}$ and $\underline{i}'$, where $\underline{i} < \underline{i}'$ that make $\underline{i} = \underline{i}'$
-due to the def-use pairs of \texttt{A} 5-6 and 5-7.
-Therefore, we remove the back edge from 5 to the phi-node 1.
-Analogously, we remove the back edges from 6 to 2 and from 7 to 3.
-However, there are many values $\underline{i} < \underline{i'}$ that make $\underline{i} = \underline{i'}-1$ and the back edge from 8 to 4 remains (def-use pairs for \texttt{D}).
-As a result of removing these spurious edges, Basic Vectorization will find that statement 5 is vectorizable.
-Statements 6, 7 and 8 will correctly appear in the FOR loop.
-
-Note however, that this step renders some array phi-nodes target-less. We handle target-less phi-nodes with a minor extension of Basic Vectorization. 
-First, we merge closures that update the same array. This simplifies handling of array $\phi$-nodes: if each closure is turned into a separate loop
-each loop will need to have its own array phi-node to account for the update and this would complicate the analysis.
-Second, we add the target-less node of array \texttt{A} back to the closure that updates \texttt{A} ---
-the intuition is, even if there is no loop-carried dependence from writes to reads on \texttt{A}, \texttt{A} is written and the write (i.e., update) cannot be vectorized; 
-therefore, the updated array has to carry to the next iteration of the loop. Third, in cases when the phi-node remains target-less, i.e., cases when the
-array write can be vectorized, we have to properly remove the phi-node replacing uses of the left-hand side of the phi-node with its arguments.
+Clearly, there is no pair $\underline{i}$ and $\underline{i}'$, where $\underline{i} < \underline{i}'$ that make $\underline{i} = \underline{i}'$ (due to the def-use pairs of \texttt{A} 6-8 and 6-10).
+Therefore, we remove the back edge from 6 to 1.
+Analogously, we remove the back edges from 8 to 2 and 10 to 3.
+However, there are many values $\underline{i} < \underline{i'}$ that make $\underline{i} = \underline{i'}-1$ and the back edge from 12 to 4 remains (def-use pairs for \texttt{D}).
+As a result of removing these spurious edges, Basic Vectorization will find that statement 6 is vectorizable.
+Statements 8, 10 and 12 will correctly appear in the FOR loop.
 
@@
 
-\ana{TODO: WIP, get to BV without MUX refinement for now? For now, skip MUX refinement.}
+\ana{TODO: I think we should implement this.}
 
@@
 
+\subsection{Extension of Basic Vectorization with Array Writes}
+
 \subsubsection{Restricting Array Writes}
 
-For now, we restrict array updates to \emph{canonical updates}. Assume (for simplicity) a two-dimensional array \texttt{A[I,J]}. A canonical update is the following:
-\begin{verbatim}
-for i in range(I):
-  for j in range(J):
-    ...
-    A[i,j] = ...
-    ...
-\end{verbatim}
-
-The update \texttt{A[i,j]} can be nested into an inner loop and there may be multiple updates, i.e., writes to \texttt{A[i,j]}.
-However, update such as \texttt{A[i-1,j] = ...} or \texttt{A[i-1,j-1] = ...}, etc., is not allowed.
-Additionally, while there could be several different loops that perform canonical updates, they must be of the same dimensionality, i.e., an update of higher or lower dimension, e.g., \texttt{A[i,j,k] = ...} is not allowed.
-We compute the \emph{canonical dimensionality} of each write array by examining the array writes in the original program and rejecting programs that violate the canonical write restriction.
+For now, we restrict array updates to \emph{canonical updates}. Assume (for simplicity) a two-dimensional array $\texttt{A}[I,J]$. A canonical update is the following:
+\begin{algorithmic}
+\STATE for i in range(I):
+\STATE \hspace{0.25cm} for j in range(J):
+\STATE \hspace{0.55cm} ...
+\STATE \hspace{0.55cm} A[i,j] = ...
+\STATE \hspace{0.55cm} ...
+\end{algorithmic}
+
+The update $\texttt{A}[i,j]$ can be nested into an inner loop and there may be multiple updates, i.e., writes to $\texttt{A}[i,j]$.
+However, update such as $\texttt{A}[i-1,j] = ...$ or $\texttt{A}[i-1,j-1] = ...$, etc., is not allowed.
+Additionally, while there could be several different loops that perform canonical updates, they must be of the same dimensionality, i.e., an update of higher or lower dimension, e.g., $A[i,j,k] = ...$ is not allowed.
+We compute the \emph{canonical dimensionality} of each write array in the obvious way \emph{before basic vectorization}.
 This restriction simplifies reasoning in this early stage of the compiler; we will look to relax the restriction in future work.
@@
 
-Reads through an arbitrary formula, such as \texttt{A[i-1]} for example, are allowed; currently, the projection function returns dummy values if the read formula is out of bounds; we assume the programmer ensures that the program still computes correct output in this case.
+Reads through an arbitrary formula, such as $\texttt{A}[i-1]$ for example, are allowed; currently, our projection function returns dummy values if the read formula is out of bounds; we assume the programmer ensures that the program still computes correct output in this case.
 
@@
 
-\begin{comment}
 There are two changes to Basic Vectorization to account for write arrays. We can account for them during processing of raise dimension/drop dimension in Phase 1 of Basic Vectorization.
@@
 \ana{TODO: Double check.})
-\end{comment}
-
-In addition to the changes for the handling of target-less phi-nodes, Basic Vectorization has to handle differently def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable.
-The definition can be an update \texttt{A\_2 = update(A\_1,i,...)} or a pseudo $\phi$-node \texttt{A\_2 = PHI(A\_0,A\_1)}, etc..
-Note that $\phi$-nodes for arrays have no subscript operations the way there are subscript operations in analysis-introduced arrays representing scalars.
-In addition, array variables are different from scalar variables.
-Raising and dropping dimension still apply, however, the dimension cannot be raised or dropped to a dimension lower than the canonical one:
+
+The other change concerns def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable.
+The definition can be an update $A_2 = update(A_1,...)$ or a pseudo $\phi$-node $A_2 = \phi(A_0,A_1)$.
+Note that $\phi$ nodes for arrays are no subscript operations on those $\phi$-nodes the way there are in scalar arrays.
+These edges are not handled in the same way as in Basic Vectorization, specifically, we do not raise and drop dimension as we do for scalars in Basic Vectorization.
+A key invariant is that the dimension of an array $A$ cannot go above or drop below $A$'s canonical dimensionality.
+We enumerate the cases of def-use edges.
 
@@
 First, the canonical dimensionality of all \texttt{A,B,C} and \texttt{D} is 1. 
-After Phase 1 of Basic Vectorization the Aiken's array write example will be (roughly) as follows:
-
-{\small
-\begin{verbatim}
-for i in range(N):
-1. A_1 = PHI(A_0,A_2)
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. A_2 = update(A_1, I, B_1[I] + 10); 
-6. B_2 = update(B_1, I, A_2[I] * D_1[I-1]);
-7. C_2 = update(C_1, I, A_2[I] * D_1[I-1]);
-8. D_2 = update(D_1, I, B_2[I] * C_2[I]);
-\end{verbatim}
-}
-
-Note that since all def-uses are same-level (i.e., reads and writes of the array elements) 
-no raise dimension or drop dimension happens.
-
-\begin{comment}
+Thus, there is no addition of extra dimension for inner loops. After Phase 1 
+of Basic Vectorization the Aiken's array write example will be (roughly) as follows:
+
 \begin{algorithmic}
@@
 \end{algorithmic}
-\end{comment}
 
 Phase 2 computes the closure of 4; $cl = \{4,6,7, 8\}$ 
-while 5 is vectorizable. Recall that 1,2, and 3 are target-less phi-nodes. Since the closure $cl$ includes updates to \texttt{B} and \texttt{C},
-the corresponding phi-nodes are added back to the closure and the def-use edges are added back to the target-less nodes. The uses of \texttt{A\_1} and \texttt{B\_1} in the vectorized statement
-turn into uses of \texttt{A\_0} and \texttt{B\_0} respectively; this is done for all original target-less phi-node.
-(But note that \texttt{A\_0} is irrelevant; the update writes into array \texttt{A\_2} in parallel.)
-Finally, the target-less phi-node for \texttt{A} is discarded.
-
-{\small
-\begin{verbatim}
-1. A_2 = update(A_0, I, ADD_SIMD(B_0[I],10));
-   equiv. to A_2[I] = ADD_SIMD(B_0[I],10)
-FOR i=0; i<N; i++; // MOTION loop
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. B_2 = update(B_1, i, A_2[i] * D_1[i-1]);
-   equiv. to B_2 = B_1; B_2[i] = A_2[i] * D_1[i-1];
-6. C_2 = update(C_1, i, A_2[i] * D_1[i-1]);
-7. D_2 = update(D_1, i, B_2[i] * C_2[i]);
-\end{verbatim}
-}
-
-\begin{comment}
+while 5 is vectorizable. Recall that 1,2, and 3 are targetless phi-nodes. Since the closure $cl$ includes updates to $B$ and $C$,
+the corresponding phi-nodes are added to the closure, i.e., the FOR loop. The uses of $A_1$ and $B_1$ in the vectorized statement
+turn into uses of $A_0$ and $B_0$ respectively. (But note that $A_0$ is irrelevant; the updates writes into array $A_2$ in parallel.)
+
 \begin{algorithmic}
@@
 \end{algorithmic}
-\end{comment}
-
-\paragraph{Example 2} Now consider the MPC Source of Histogram:
-
-{\small
-\begin{verbatim}
-for i in range(0, num_bins):
-   res1 = PHI(res, res2)
-   for j in range(0, N):
-      res2 = PHI(res1, res3)
-      tmp1 = (A[j] == i)
-      tmp2 = (res2[i] + B[j])
-      tmp3 = MUX(tmp1, res2[i], tmp2)
-      res3 = Update(res2, i, tmp3)     
-return res1
-\end{verbatim}
-}
-
-The canonical dimensionality of \texttt{res} is 1.
-Also, the phi-node \texttt{res2 = PHI(res1, res3)} is a target-less phi-node (the implication being that the inner for loop can be vectorized across $i$).
-After Phase 1, Basic vectorization produces the following (everything is implicitly vectorized along $i$ and $j$ and the update is turned into a regular write):
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I = raise_dim(i, ((i:num_bins),(j:N)))
-for i in range(0, num_bins):
-   res1 = PHI(res, res2) // target-less, no drop_dim
-   res1' = raise_dim(res1, (j:N))
-   for j in range(0, N):
-      res2 = PHI(res1', res3)
-      tmp1 = (A1 == I)
-      tmp2 = (res2 + B1)
-      tmp3 = MUX(tmp1, res2, tmp2)
-      res3 = res2 // copy propagation: res2 into res3
-      res3 = tmp3 // "normal": res3[I,J] = tmp3[I,J]    
-return res1
-\end{verbatim}
-}
-
-Processing the inner loop in Phase 2 vectorizes \texttt{tmp1 = (A1 == I)} along the $j$ dimension but leaves the rest of the statements in a MOTION loop.
-Processing the outer loop is interesting.
-This is because the PHI node is a target-less node, and therefore, there are no closures!
-Thus everything can be vectorized along the $i$ dimension:
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I1 = raise_dim(i, ((i:num_bins),(j:N)))
-
-tmp1[I,J] = (A1[I,J] == I1[I,J])
-
-res1' = raise_dim(res, (j:N)) // raising res to a 2-dim. array
-for j in range(0, N):
-   res2 = PHI(res1', res3)
-   tmp2[I,j] = (res2[I,j] + B1[I,j])
-   tmp3[I,j] = MUX(tmp1[I,j], res2[I,j], tmp2[I,j])
-   res3 = res2
-   res3[I,j] = tmp3[I,j]
-res1 = drop_dim(res2) // Have to add drop_dim before removing redundant phi-node      
-return res1
-\end{verbatim}
-}
-
-The cleanup phase can then get rid of $j$ in the res variables and tmp variables, so the loop operates over the num\_bins-size vectors. 
-This will lead to dropping raise\_dim and drop\_dim as well:
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I1 = raise_dim(i, ((i:num_bins),(j:N)))
-
-tmp1[I,J] = (A1[I,J] == I1[I,J])
-
-for j in range(0, N):
-   res2 = PHI(res1', res3) // now 1-dim array
-   tmp2[I] = (res2[I] + B1[I,j])
-   tmp3[I] = MUX(tmp1[I,j], res2[I], tmp2[I])
-   res3 = res2
-   res3[I] = tmp3[I]
-return res2
-\end{verbatim}
-}
+
+\ana{TODO: Format either Histogram or Matrix Multiplication. Both are interesting examples that use output arrays.}
 

### 2022-03-02 paper_SIMD.tex
fixed the mix-up, Ana's work restored.

--- 2022-03-01 paper_SIMD.tex
@@
 \end{algorithmic}
-
-%\ishaq{We need to think of a way to handle 4th case for def-use, when a nested block B def is followed by another, different, nested block B' use.
-Note: The handwritten notes say we can break it into 2 edges.
-One idea to this end is to insert a def' node which is a copy of the original def.
-The def' should be placed in a block that contains both B and B'.}
-%\ishaq{THE ABOVE COMMENT IS RESOLVED.
-After transformation to SSA, there should be an intermediate variable (phi node) that is assigned a value depending on whether control branched or not.
-Should we put this into the writeup?}
 
@@
 
-\ishaq{Note to self: This algorithm is an instantiation for $j$ loop, the one for $k$ loop will be exactly the same, modulo variable name.}.
-
 Consider a loop $j$ enclosed in some fixed $\underline{i}$. Only if an update (definition) $\texttt{A}_m[f(i,j,k)] = ... $ at some iteration $\underline{j}$

### 2022-03-02 benchmarks/longest_even_0.py
Get MOTION output into a compileable state

@@ -24,3 +24,6 @@ def longest_even_0(Seq: shared[list[int]], N: int, Sym: shared[int]) -> shared[i
             max_length = tmp_max_len
 
     return max_length
+
+
+print(longest_even_0([0, 0, 1, 0, 0, 0], 6, 0))

### 2022-03-02 compiler/compiler/dep_graph.py

@@ -172,8 +172,7 @@ class DepGraph:
             return True
         else:
             return any(
-                self.def_use_graph.has_edge(src, use_stmt)
-                and self.has_same_level_path(use_stmt, dest, visited | {src})
+                self.has_same_level_path(use_stmt, dest, visited | {src})
                 and self.edge_kind(src, use_stmt) == EdgeKind.SAME_LEVEL
                 for use_stmt in self.def_use_graph.successors(src)
                 if use_stmt not in visited

### 2022-03-02 compiler/compiler/motion_backend/__init__.py

@@ -120,6 +120,17 @@ def render_function(func: Function, type_env: TypeEnv) -> str:
             render_type(var_type, plaintext=False)
             + " "
             + render_expr(var, dt.replace(render_ctx, plaintext=False))
+            # Initialize vectorized arrays with a size
+            + (
+                "(("
+                + ") * (".join(
+                    render_expr(bound, dt.replace(render_ctx, plaintext=True))
+                    for _, bound in var_type.dim_sizes
+                )
+                + "))"
+                if var_type.dim_sizes
+                else ""
+            )
             + ";"
             for var, var_type in sorted(type_env.items(), key=lambda x: str(x[0]))
             if not any(
@@ -133,7 +144,21 @@ def render_function(func: Function, type_env: TypeEnv) -> str:
     plaintext_var_definitions = (
         "// Plaintext variable declarations\n"
         + "\n".join(
-            f"{render_type(var_type, plaintext=True)} {render_expr(var, dt.replace(render_ctx, plaintext=True))};"
+            render_type(var_type, plaintext=True)
+            + " "
+            + render_expr(var, dt.replace(render_ctx, plaintext=True))
+            # Initialize vectorized arrays with a size
+            + (
+                "(("
+                + ") * (".join(
+                    render_expr(bound, dt.replace(render_ctx, plaintext=True))
+                    for _, bound in var_type.dim_sizes
+                )
+                + "))"
+                if var_type.dim_sizes
+                else ""
+            )
+            + ";"
             for var, var_type in sorted(type_env.items(), key=lambda x: str(x[0]))
             if var_type.is_plaintext()
             if not any(

### 2022-03-02 compiler/compiler/motion_backend/low_level_rendering.py

@@ -32,6 +32,7 @@ from ..tac_cfg import (
 )
 from ..util import assert_never
 from ..loop_linear_code import For, Phi, Return
+from ..type_analysis import type_assign_expr
 
 
 @dc.dataclass(frozen=True)
@@ -103,6 +104,13 @@ def render_stmt(stmt: Union[Assign, For, Return], type_env: TypeEnv) -> str:
         # If we're assigning to a vectorized value, use a specialized function for this.
         if isinstance(stmt.lhs, VectorizedAccess):
             ctx = RenderContext(type_env, plaintext=False)
+            val_expr = render_expr(stmt.rhs, ctx)
+
+            # If this isn't a true vectorized access, just subscript normally
+            if all(vectorized == False for vectorized in stmt.lhs.vectorized_dims):
+                lhs = render_expr(stmt.lhs, ctx)
+                return f"{lhs} = {val_expr};"
+
             dim_sizes = (
                 "{"
                 + ", ".join(
@@ -126,7 +134,8 @@ def render_stmt(stmt: Union[Assign, For, Return], type_env: TypeEnv) -> str:
                 )
                 + "}"
             )
-            return f"vectorized_assign({render_expr(stmt.lhs.array, ctx)}, {dim_sizes}, {vectorized_dims}, {idxs}, {render_expr(stmt.rhs, ctx)});"
+
+            return f"vectorized_assign({render_expr(stmt.lhs.array, ctx)}, {dim_sizes}, {vectorized_dims}, {idxs}, {val_expr});"
         else:
             shared_assignment = (
                 render_expr(stmt.lhs, RenderContext(type_env, plaintext=False))
@@ -150,6 +159,14 @@ def render_stmt(stmt: Union[Assign, For, Return], type_env: TypeEnv) -> str:
                 return shared_assignment + "\n" + plaintext_assignment
 
     elif isinstance(stmt, For):
+        ctr_initializer = (
+            "// Initialize loop counter\n"
+            + render_expr(stmt.counter, RenderContext(type_env, plaintext=True))
+            + " = "
+            + render_expr(stmt.bound_low, RenderContext(type_env, plaintext=True))
+            + ";"
+        )
+
         phi_initializations = "// Initialize phi values\n" + "\n".join(
             render_stmt(Assign(phi.lhs, phi.rhs_false), type_env)
             for phi in stmt.body
[…]

### 2022-03-02 compiler/compiler/tac_cfg.py

@@ -133,7 +133,7 @@ def assign_rhs_accessed_vars(rhs: AssignRHS) -> list[Var]:
         return [var for item in rhs.items for var in assign_rhs_accessed_vars(item)]
     elif isinstance(rhs, Mux):
         return (
-            [rhs.condition]
+            assign_rhs_accessed_vars(rhs.condition)
             + assign_rhs_accessed_vars(rhs.false_value)
             + assign_rhs_accessed_vars(rhs.true_value)
         )

### 2022-03-02 compiler/compiler/type_analysis.py

@@ -22,7 +22,7 @@ from .ast_shared import (
 from .tac_cfg import AssignRHS, List, Tuple, Mux, Update, LiftExpr, DropDim
 
 
-def _type_assign_expr(
+def type_assign_expr(
     expr: Union[AssignRHS, SubscriptIndex, VectorizedAccess], type_env: TypeEnv
 ) -> VarType:
     """
@@ -40,7 +40,7 @@ def _type_assign_expr(
         return VarType(VarVisibility.PLAINTEXT, 0, expr.datatype)
 
     elif isinstance(expr, Subscript):
-        index_type = _type_assign_expr(expr.index, type_env)
+        index_type = type_assign_expr(expr.index, type_env)
         if not index_type.could_become(PLAINTEXT_INT):
             raise TypeError(
                 f"Array subscript index {expr.index} is not a plaintext int"
@@ -53,8 +53,8 @@ def _type_assign_expr(
         return type_env[expr.array].drop_dim()
 
     elif isinstance(expr, BinOp):
-        lhs_type = _type_assign_expr(expr.left, type_env)
-        rhs_type = _type_assign_expr(expr.right, type_env)
+        lhs_type = type_assign_expr(expr.left, type_env)
+        rhs_type = type_assign_expr(expr.right, type_env)
 
         expr_type = VarType.merge(lhs_type, rhs_type)
 
@@ -75,7 +75,7 @@ def _type_assign_expr(
 
     elif isinstance(expr, UnaryOp):
         op_datatype = expr.operator.get_ret_datatype()
-        expr_type = _type_assign_expr(expr.operand, type_env)
+        expr_type = type_assign_expr(expr.operand, type_env)
 
         if expr_type.datatype not in [*expr.operator.get_operand_datatypes(), None]:
             raise TypeError(
@@ -95,14 +95,14 @@ def _type_assign_expr(
             return VarType(VarVisibility.PLAINTEXT, 1, DataType.INT)
 
         elem_type = VarType.merge(
-            *[_type_assign_expr(item, type_env) for item in expr.items],
+            *[type_assign_expr(item, type_env) for item in expr.items],
             mixed_shared_plaintext_allowed=True,
         )
 
         return elem_type.add_dim()
 
     elif isinstance(expr, Tuple):
-        elem_types = [_type_assign_expr(item, type_env) for item in expr.items]
+        elem_types = [type_assign_expr(item, type_env) for item in expr.items]
         return VarType(
             VarVisibility.PLAINTEXT,  # Tuples are always plaintext
[…]

### 2022-03-02 compiler/compiler/vectorize.py

@@ -1,8 +1,8 @@
 from collections import defaultdict
 import functools
 
-from .ast_shared import TypeEnv, Var, VectorizedAccess, Subscript
-from .tac_cfg import DropDim, LiftExpr, Assign
+from .ast_shared import TypeEnv, Var, VectorizedAccess, Subscript, DataType
+from .tac_cfg import DropDim, LiftExpr, Assign, BinOp, BinOpKind
 from . import loop_linear_code as llc
 from .dep_graph import DepGraph, DepNode, EdgeKind
 from . import util
@@ -367,89 +367,88 @@ def _basic_vectorization_phase_1(
     inside_loop_result: list[llc.Statement] = []
 
     for stmt in stmts:
+
+        def replace_var(
+            var: llc.Var,
+            stmt: DepNode,
+            access_pattern: Optional[llc.SubscriptIndex] = None,
+        ) -> llc.Var:
+            def_var = dep_graph.var_to_assignment[var]
+            edge_kind = dep_graph.edge_kind(def_var, stmt)
+            if edge_kind is EdgeKind.SAME_LEVEL:
+                return var
+            elif edge_kind is EdgeKind.OUTER_TO_INNER:
+                var_prime = tmp_var_gen.get()
+                above_loop_result.append(
+                    llc.Assign(
+                        lhs=var_prime,
+                        rhs=LiftExpr(
+                            Subscript(var, access_pattern) if access_pattern else var,
+                            tuple(
+                                (loop.counter, loop.bound_high)
+                                for loop in dep_graph.enclosing_loops[stmt]
+                            ),
+                        ),
+                    )
+                )
+                return var_prime
+            elif edge_kind is EdgeKind.INNER_TO_OUTER:
+                var_prime = tmp_var_gen.get()
+                inside_loop_result.append(
+                    llc.Assign(
+                        lhs=var_prime,
+                        rhs=DropDim(
+                            var,
+                            tuple(
+                                loop.bound_high
+                                for loop in dep_graph.enclosing_loops[def_var]
+                            ),
+                        ),
+                    )
+                )
+                return var_prime
+            else:
+                assert_never(edge_kind)
+
+        def replace_atom(atom: llc.Atom, stmt: DepNode) -> llc.Atom:
+            if isinstance(atom, llc.Var):
+                return replace_var(atom, stmt)
+            elif isinstance(atom, llc.Constant):
+                return atom
+            else:
[…]

### 2022-03-02 compiler/tests/test_stages.py

@@ -111,7 +111,7 @@ class StagesTestCase(unittest.TestCase):
 
             # Collect expected output
             proc = subprocess.run(
-                ["python3", input_fname],
+                ["python3.9", input_fname],
                 check=True,
                 stdout=subprocess.PIPE,
                 text=True,
@@ -144,8 +144,12 @@ class StagesTestCase(unittest.TestCase):
                 text=True,
             )
 
+            print("Waiting...")
+
             party0.wait(30)
             party1.wait(30)
+
+            print("wait complete")
 
             received_output = party1.stdout.read()
 

### 2022-03-02 paper_SIMD.tex
Edits to Arrays section. Still need to double check some.

@@
 Note that $\phi$-nodes for arrays have no subscript operations the way there are subscript operations in analysis-introduced arrays representing scalars.
-In addition, array variables are different from scalar variables.
 Raising and dropping dimension still apply, however, the dimension cannot be raised or dropped to a dimension lower than the canonical one:
@@
 The canonical dimensionality of \texttt{res} is 1.
-Also, the phi-node \texttt{res2 = PHI(res1, res3)} is a target-less phi-node (the implication being that the inner for loop can be vectorized across $i$).
-After Phase 1, Basic vectorization produces the following (everything is implicitly vectorized along $i$ and $j$ and the update is turned into a regular write):
+Also, the phi-node \texttt{res1 = PHI(res, res2)} is a target-less phi-node (the implication being that the inner for loop can be vectorized across $i$).
+After Phase 1, Basic vectorization produces the following code (statements are implicitly vectorized along $i$ and $j$).
+\ana{Define the interpretation of vectorized Update!}
 
@@
       tmp3 = MUX(tmp1, res2, tmp2)
-      res3 = res2 // copy propagation: res2 into res3
-      res3 = tmp3 // "normal": res3[I,J] = tmp3[I,J]    
+      res3 = Update(res2, (I,J), tmp3) 
 return res1
@@
 This is because the PHI node is a target-less node, and therefore, there are no closures!
-Thus everything can be vectorized along the $i$ dimension:
+Thus, everything can be vectorized along the $i$ dimension:
 
@@
    tmp3[I,j] = MUX(tmp1[I,j], res2[I,j], tmp2[I,j])
-   res3 = res2
-   res3[I,j] = tmp3[I,j]
-res1 = drop_dim(res2) // Have to add drop_dim before removing redundant phi-node      
+   res3 = Update(res2, (I,j), tmp3)
+res1 = drop_dim(res2) // Add drop_dim before removing redundant phi-node!      
 return res1
@@
 
-The cleanup phase can then get rid of $j$ in the res variables and tmp variables, so the loop operates over the num\_bins-size vectors. 
+The cleanup phase can then get rid of $j$ in the \texttt{res} variables and \texttt{tmp} variables, so the loop operates over the \texttt{num\_bins}-size vectors. 
 This will lead to dropping raise\_dim and drop\_dim as well:
@@
 for j in range(0, N):
-   res2 = PHI(res1', res3) // now 1-dim array
-   tmp2[I] = (res2[I] + B1[I,j])
-   tmp3[I] = MUX(tmp1[I,j], res2[I], tmp2[I])
+   res2 = PHI(res1, res3) // now 1-dim array
+   tmp2[I] = ADD_SIMD(res2[I],B1[I,j])
+   tmp3[I] = MUX_SIMD(tmp1[I,j], res2[I], tmp2[I])
    res3 = res2

### 2022-03-02 paper_SIMD.tex
Minor change to code example.

@@
    tmp3[I] = MUX_SIMD(tmp1[I,j], res2[I], tmp2[I])
-   res3 = res2
-   res3[I] = tmp3[I]
+   res3 = Update(res2, I, tmp3[I])
 return res2

### 2022-03-02 paper_SIMD.tex
Yet another minor change, removed a TODO.
Answered Ishaq's question.

@@
 \ishaq{TODO: @Ana: I am still confused about what prevents a reviewer from saying, "you argue correctness of the schedule from unrolled MPC, you don't argue it for the pre-unrolled CFG that you use".
-Please comment.}
+Please comment.} \ana{The correctness proof answer this question.
+The unrolled schedule is the \emph{concretization} of MPC source.
+We define (1) the obvious $\gamma$, (2) we define partial order over MPC source codes (this is our abstract domain $A$), and (3) a partial order over unrolled schedules.
+Then we prove two theorems that $A \le A' \Rightarrow \gamma(A) \subseteq \gamma(A')$ and that the MPC source $\le$ Vectorized MPC source.}
 
@@
 After Phase 1, Basic vectorization produces the following code (statements are implicitly vectorized along $i$ and $j$).
-\ana{Define the interpretation of vectorized Update!}
+In a vectorized update statement, we can ignore the incoming array, \texttt{res2} in this case.
+The update writes (in parallel) all locations of the 2-dimensional array, in this case it sets up each \texttt{res3[i,j] = tmp3[i,j]}.
 

### 2022-03-05 paper_SIMD.tex
Commented out Longest Common SuperSeq stuff to test whether latex github action works.
Commented out section to have paper compile on Github.
Added first draft of correctness argument section.

@@
 
+\ana{Commenting out this section to have paper compile on Github. Will revert!}
+\begin{comment}
+
 At the lowest level, we have two types of MPC instructions 1) local/non-interactive instruction (i.e.
@@
 
+\end{comment}
+
 \subsection{Scheduling is NP-hard}
@@
 
-\ana{Ishaq?}
-
+\ana{Commented out this section to have paper compile on Github. Will revert!}
+
+\begin{comment}
 \ishaq{The follow text is all from Linsay's thesis}.
@@
 }
+\end{comment}
 
@@
 
-\ana{Add the AI-based proof here.}
+We build a correctness argument that loosely follows the theory of Abstract Interpretation. 
+We define the syntax of MPC Source programs. The domain of MPC Source programs
+expressible in the syntax (with certain semantic restrictions) is the abstract domain $A$. 
+We then define the \emph{linearization} of an MPC Source program as an interpretation 
+over the syntax. The linearization, which is a \emph{schedule}, is the concrete domain $C$. 
+Since we reason over def-use graphs in $A$ we define a partial order relation over elements of $A$ 
+in terms of def-use relations. We define a partial order over elements of $C$ as well,
+in terms of def-use relations in the concrete domain $C$. We prove two theorems
+that state (informally) that the schedule corresponding to the original program computes the same 
+result as the schedule corresponding to the vectorized program.
+
+\begin{figure}%[!b]
+%\small
+\[
+\begin{array}{l@{~~~~~}l}
+  \begin{array}{l@{~}l@{~~~}l}
+  s &::= s;s & sequence \\
+     & \mid \x[i,J,k] = \y[i,J,k] \; \mathit{op\_SIMD} \; \z[i,J,k] & operation \\ 
+     & \mid \x[i,J,k]  = \texttt{PHI}(\x_1[i,J,k],\x_2[i,J,k\!-\!1]) & \mathit{phi} \; \mathit{node} \\
+     & \mid \x[i,J,k] = \mathit{raise\_dim}(\x'[i],(J\!:\!J,k\!:\!K)) & \mathit{raise} \; \mathit{dimension(s)} \\
+     & \mid \x[i,J] = \mathit{drop\_dim}(\x'[i,J,k],k)  & \mathit{drop} \; \mathit{dimension(s)} \\
+     & \mid \x = \y & \mathit{propagation} \\
+     & \mid \texttt{FOR} \; 0 \le i < I: s  & \mathit{loop} \\
+  \end{array}    \\
+\end{array}
+\]
+\vspace{-3mm}
+\caption{MPC Source Syntax}\label{fig:MPC_Source_syntax}
+\vspace{-5mm}
+\end{figure}
+
+\paragraph{MPC Source Syntax}
+
+\figref{fig:MPC_Source_syntax} defines the syntax for our intermediate representation, MPC Source. 
+There are semantics restrictions over the syntax as well: a variable $\x[i,j,k]$ is a 3-dimensional array 
+$(i:I,j:J,k:K)$ and also, a statement $\x[i,J,k] = ...$ is enclosed in loops over $i$ and $k$ as shown below. 
+Thus, $i$ and $k$ are in scope.
+\begin{verbatim}
+FOR 0 <= i < I:
+    ...
+    FOR 0 <= k < K:
+        x[i,J,k] = ...
+\end{verbatim}
+
+Statements $\mathit{operation, phi, raise \; dimension(s), drop \; dimension(s)}$ are \emph{base} statements, 
+and $\mathit{sequence, loop}$ are \emph{compound statements}.
+
+\paragraph{Linearization}
+
+Linearization is the concretization operation, which, as we mentioned earlier computes a schedule.
+The concretization function $\gamma: A \rightarrow C$ is defined as an interpretation of
+MPC Source syntax, as it is standard. The concretization of each one 
+of the base statements is as follows: 
+\[\begin{array}{l}
+\gamma(\x[i,J,k] = \y[i,J,k] \; \mathit{op\_SIMD} \; \z[i,J,k]) = \\ 
+\x[i,0,k] = \y[i,0,k] \; \mathit{op} \; \z[i,0,k] \;\; || \\
+\x[i,1,k] = \y[i,1,k] \; \mathit{op} \; \z[i,1,k] \;\; || \; ...\; || \\
+\x[i,I-1,k] = \y[i,I-1,k] \; \mathit{op} \; \z[i,I-1,k]
+\end{array}\]
+meaning that the vectorized dimension(s) are expanded into parallel statements. || introduces SIMD (parallel) execution.
+
+The concretization of the FOR statement is as follows:
+\[ \gamma(\texttt{FOR} \; 0 \le i < I: s) = \gamma(s)[0/i] ; \gamma(s)[1/i] ; ... \gamma(s)[I-1/i] \]
+$\gamma$ simply unrolls the loop substituting $i$ with 0, 1, etc. Here ; denotes sequential execution.
+
+\paragraph{Partial Orders}
+
+For each MPC Source program $a$ we compute the def-use edges in the standard way: if base statement $s1 \in a$ defines variable \x, e.g., $\x[i,j,k] = ...$, 
+and base statement $s2 \in a$ uses \x, e.g., $... = ... \x[i,j,k]$ and there is a path in the trivial CFG from $s1$ to $s2$, then there is a def-use edge from
+ $s1$ to $s2$. We extend the dimensionality of a statement into $s1[i,j,k]$ where $s1[i,j,k]$ inherits the dimensionality of the left-hand-side of the assignment.
+
+ Let $a_0, a_1$ be two MPC Source programs in $A$. 
+ Two base statements, $s_0 \in a_0$ and $s_1 \in a_1$ are \emph{same}, written $s_0 \equiv s_1$ if they are of the same operation and they operate on 
+ the same variables: same variable name and same dimensionality. Recall that dimensions in MPC Source are either iterative, lower case, or vectorized, upper case. 
+ Two statements are same even if one operates on an iterative dimension and the other one operates on a vectorized one, e.g., $s_0[i,j,k] \equiv s_1[I,j,K]$.
+
+ \begin{definition} Let $a_0, a_1 \in A$. We say that $a_0 \le a_1$ iff for every def-use edge $s1 \rightarrow s2$ in $a_0$ there is an edge
+ $s1' \rightarrow s2'$ where $s1 \equiv s1'$, $s2 \equiv s2'$ and the two edges of either both forward or both backward.
+ \end{definition}
+
+ The def-use edges in the concrete schedule are as expected. There is a def use edge from statement $s1$ that defines $\x[\underline{i},\underline{j},\underline{k}]$
+ to statement $s2$ that uses $\x[\underline{i},\underline{j},\underline{k}]$ if $s1$ is scheduled ahead of $s2$ in the linear schedule. We note that the underlined
+ indices, e.g., $\underline{i}$, refer to fixed values, not iterative or vectorized dimensions since in the concrete schedule all induction variables are expanded. 
+ E.g., there is a def-use edge from the statement that defines $\x[0,1,2]$ and a statement that uses $\x[0,1,2]$.
+
+\paragraph{Theorems}
+
+\begin{theorem} $a_0 \le a_1 \Rightarrow \gamma(a_0) \subseteq \gamma(a_1)$.
+\end{theorem}
+
+\begin{theorem} Let $a_0$ be the iterative MPC Source and let $a_1$ be the vectorized MPC Source computed by the vectorization algorithm. 
+We have that $a_0 \le a_1$. 
+\end{theorem}
+
+\ana{Write the proof sketch and final argument, etc.}
 

### 2022-03-05 paper_SIMD.tex
Uncommented Ishaq's section.

@@
 \ana{Commenting out this section to have paper compile on Github. Will revert!}
-\begin{comment}
+%\begin{comment}
 
@@
 
-\ishaq{TODO: @Ana: I am still confused about what prevents a reviewer from saying, "you argue correctness of the schedule from unrolled MPC, you don't argue it for the pre-unrolled CFG that you use".
+%\ishaq{TODO: @Ana: I am still confused about what prevents a reviewer from saying, "you argue correctness of the schedule from unrolled MPC, you don't argue it for the pre-unrolled CFG that you use".
 Please comment.} \ana{The correctness proof answer this question.
@@
 
-\end{comment}
+%\end{comment}
 

### 2022-03-05 paper_SIMD.tex
Minor edit.

@@
 
-\item We present a compiler that takes a Python-like high-level 
+\item We present a compiler that takes an IMP-like high-level 
 program and produces amortized (i.e., vectorized) low-level cryptographic 
-code in the MOTION framework. Central to our approach is a new 
-vectorization algorithm that produces optimal schedules for a large 
-number of MPC programs.
+code in the MOTION framework. Central contributions are 1) a novel compiler framework, 
+2) a vectorization algorithm that produces optimal schedules for a large 
+number of MPC programs, and 3) reasoning over output arrays without Array SSA; we remove 
+infeasible loop-carried dependences introduced by standard SSA to improve vectorization.
 

### 2022-03-09 prob_stmt_np_hardness.tex
extract sec 3.2 to its own file

--- 2022-03-05 paper_SIMD.tex
@@
-% !TeX TXS-program:compile = txs:///pdflatex/[--shell-escape]
-%
-% acm templates v1.57 (sigconf template)
-%
-%\documentclass[sigconf, screen, review, anonymous, natbib=false]{acmart}
-\documentclass[sigconf, screen, natbib=false, dvipsnames, table]{acmart}
-\settopmatter{printacmref=true}
-% mandatory for CCS'19
-\usepackage{balance}
-\usepackage{mathtools}
-% for creating a balanced last page (usually last page with references)
-
-% defining the \BibTeX command - from Oren Patashnik's original BibTeX documentation.
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08emT\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-
-% Rights management information.
-% This information is sent to you when you complete the rights form.
-% These commands have SAMPLE values in them; it is your responsibility as an author to replace
-% the commands and values with those provided to you when you complete the rights form.
-%
-% These commands are for a PROCEEDINGS abstract or paper.
-
-\copyrightyear{2019}
-\acmYear{2019}
-\setcopyright{acmcopyright}
-\acmConference[CCS '19]{2019 ACM SIGSAC Conference on Computer and Communications Security}{November 11--15, 2019}{London, United Kingdom}
-\acmBooktitle{2019 ACM SIGSAC Conference on Computer and Communications Security (CCS '19), November 11--15, 2019, London, United Kingdom}
-\acmPrice{15.00}
-\acmDOI{10.1145/3319535.3339818}
-\acmISBN{978-1-4503-6747-9/19/11}
-
-%
-% These commands are for a JOURNAL article.
-%\setcopyright{acmcopyright}
-%\acmJournal{TOG}
-%\acmYear{2018}\acmVolume{37}\acmNumber{4}\acmArticle{111}\acmMonth{8}
-%\acmDOI{10.1145/1122445.1122456}
-
-%
-% Submission ID.
-% Use this when submitting an article to a sponsored event. You'll receive a unique submission ID from the organizers
-% of the event, and this ID should be used as the parameter to this command.
-%\acmSubmissionID{123-A56-BU3}
-
-%
-% The majority of ACM publications use numbered citations and references. If you are preparing content for an event
-% sponsored by ACM SIGGRAPH, you must use the "author year" style of citations and references. Uncommenting
-% the next command will enable that style.
-%\citestyle{acmauthoryear}
-
- \makeatletter
- % \renewcommand{\section}{\abovedisplayskip 10\p@ \@plus3\p@ \@minus1\p@%
- %                       \belowdisplayskip 5\p@ \@plus3\p@ \@minus1\p@%
- %                       \abovedisplayshortskip 0pt \@plus2\p@%
- %                       \belowdisplayshortskip 0pt \@plus2\p@ \@minus0\p@%
- %                       \@startsection{section}{1}{\z@}%
- %                        {-17\p@ \@plus -4\p@ \@minus -4\p@}%
- %                        {6\p@ \@plus 4\p@ \@minus 4\p@}%
- %                        {\normalfont\large\bfseries\boldmath
- %                         \rightskip=\z@ \@plus 8em\pretolerance=10000 }}
- \renewcommand{\subsection}{\@startsection{subsection}{2}{\z@}%
-                        {-8\p@ \@plus -4\p@ \@minus -4\p@}%
-                        {5\p@ \@plus 2\p@ \@minus 2\p@}%
-                        {\normalfont\Large\bfseries\boldmath
-                         \rightskip=\z@ \@plus 3em\pretolerance=10000 }}
-  \renewcommand{\subsubsection}{\@startsection{subsubsection}{3}{\z@}%
-                        {-6\p@ \@plus -4\p@ \@minus -4\p@}%
-                        {1\p@ \@plus 1\p@ \@minus 0\p@}%
-                        {\normalfont\normalsize\bfseries\boldmath}}
-% \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\z@}%
-%                       {-8\p@ \@plus -4\p@ \@minus -4\p@}%
-%                       {-2\p@ \@plus -0.22em \@minus -0.1em}%
-%                       {\normalfont\normalsize\bfseries}}
- \makeatother
-
-% *** GRAPHICS RELATED PACKAGES ***
-\usepackage{graphicx}
-\usepackage{comment}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{listings}
-% declare the path(s) where your graphic files are
-\graphicspath{{../figures/}}
-% and their extensions so you won't have to specify these with
-% every instance of \includegraphics
-\DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-
-%\usepackage[outputdir=/Volumes/ramdisk/]{minted}
-\usepackage{minted}
-% *** MATH PACKAGES ***
-\usepackage{amsmath}
-
-% *** SPECIALIZED LIST PACKAGES ***
-%\usepackage{algorithm2e}
-\usepackage{algorithmic}
-
-% *** ALIGNMENT PACKAGES ***
-\usepackage{array}
-
-\usepackage{url}
-
-\usepackage[utf8]{inputenc}
-\usepackage[american]{babel}
-\usepackage[backend=biber,style=ACM-Reference-Format]{biblatex}
-\addbibresource{library.bib}
-\addbibresource{../cryptobib/abbrev2.bib}
-\addbibresource{../cryptobib/crypto_crossref.bib}
-
-% Used for Theorems and Definitions
-\usepackage{amsthm}
-\newtheorem{definition}{Definition}
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}{Lemma}
-\theoremstyle{definition}
-\newtheorem{remark}{Remark}
-\newtheorem{assumption}{Assumption}
-
-\usepackage{xcolor}
-\usepackage{xspace}
-\usepackage{xstring} % required by IfEqCase
-
-\usepackage{booktabs} % better tables
-\usepackage{algorithm}
-\usepackage{algorithmic}
-
-% for references, use it like this: \cref{some_label}
-% it will automatically put Appendix, Figure, Section, etc before the reference
-\usepackage[capitalise]{cleveref}
-
-% ishaq: adding our macros
-\input{macros.tex}
-
-\sloppy
-\begin{document}
-\fancyhead{}
-% do not delete this code.
-
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
-\title[Scheduling and Vectorization for MPC]{Scheduling and Vectorization for MPC}
-
-%
-% The "author" command and its associated commands are used to define the authors and their affiliations.
-% Of note is the shared affiliation of the first two authors, and the "authornote" and "authornotemark" commands
-% used to denote shared contribution to the research.
-
-\author{Benjamin Levy}
-\email{levyb3@rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Benjamin Sherman}
-\email{shermb@rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Lindsey Kennard}
-\email{kennal@rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Ana L. Milanova}
-\email{milanova@cs.rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Muhammad Ishaq}
-\authornote{This work was done in part while the author was at RPI.}
-\email{m.ishaq@ed.ac.uk}
-%\orcid{1234-5678-9012}
-%\author{G.K.M. Tobin}
-%\authornotemark[1]
-%\email{webmaster@marysville-ohio.com}
-\affiliation{%
-    \institution{University of Edinburgh}
-    %\streetaddress{P.O. Box 1212}
-    \city{Edinburgh}
-    \state{Scotland}
-%    \postcode{43017-6221}
-}
-
-\author{Vassilis Zikas}
-\authornote{This work was done in part while the author was visiting UCLA and supported in part by DARPA and SPAWAR under contract N66001-15-C-4065 and by a SICSA Cyber Nexus Research Exchanges grant.}
-\email{vzikas@inf.ed.ac.uk}
-%\orcid{1234-5678-9012}
-%\author{G.K.M. Tobin}
-%\authornotemark[1]
-%\email{webmaster@marysville-ohio.com}
-\affiliation{%
-    \institution{University of Edinburgh}
-    %\streetaddress{P.O. Box 1212}
-    \city{Edinburgh}
-    \state{Scotland}
-    %    \postcode{43017-6221}
-}
-
-%
-% By default, the full list of authors will be used in the page headers. Often, this list is too long, and will overlap
-% other information printed in the page headers. This command allows the author to define a more concise list
-% of authors' names for this purpose.
-%\renewcommand{\shortauthors}{Trovato and Tobin, et al.}
-
-\renewcommand{\A}{{\sf A}}
-\renewcommand{\B}{{\sf B}}
-\renewcommand{\C}{{\sf C}}
-\renewcommand{\D}{{\sf D}}
-
-% As a general rule, do not put math, special symbols or citations
-% in the abstract
-\begin{abstract}
-
-\end{abstract}
-
-%
-% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-% Please copy and paste the code instead of the example below.
-%
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10003752.10010124.10010138.10010143</concept_id>
-<concept_desc>Theory of computation~Program analysis</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10003752.10003777.10003789</concept_id>
-<concept_desc>Theory of computation~Cryptographic protocols</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10002978.10002979</concept_id>
-<concept_desc>Security and privacy~Cryptography</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Theory of computation~Program analysis}
-\ccsdesc[500]{Theory of computation~Cryptographic protocols}
-\ccsdesc[300]{Security and privacy~Cryptography}
-
-%
-% Keywords. The author(s) should pick words that accurately describe the work being
-% presented. Separate the keywords with commas.
-\keywords{multiparty computation; compilers; cryptography}
-
-\begin{comment}
-%
-% A "teaser" image appears between the author and affiliation information and the body
-% of the document, and typically spans the page.
-\begin{teaserfigure}
-    \includegraphics[width=\textwidth]{sampleteaser}
-    \caption{Seattle Mariners at Spring Training, 2010.}
-    \Description{Enjoying the baseball game from the third-base seats. Ichiro Suzuki preparing to bat.}
-    \label{fig:teaser}
-\end{teaserfigure}
-\end{comment}
-%
-% This command processes the author and affiliation and title information and builds
-% the first part of the formatted document.
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-
-\begin{itemize}
-\item We define the scheduling problem for MPC. 
-We present an analytical model to reason about cost of schedules
-and show that scheduling is NP-hard via a reduction to the shortest 
-common supersequence problem.
-
-\item We present a compiler that takes a Python-like high-level 
-program and produces amortized (i.e., vectorized) low-level cryptographic 
-code in the MOTION framework. Central to our approach is a new 
-vectorization algorithm that produces optimal schedules for a large 
-number of MPC programs.
-
-\item We present an implementation and evaluation in the MOTION framework. 
-\ana{Fill in with final results. Mention benchmarks (standard + new ones + HyCC). }
-
-\end{itemize}
-
-\section{Overview}
-\label{sec:overview}
-
-\subsection{Source}
-
-As a running example, consider Biometric matching, a standard MPC benchmark.
-Array \texttt{C} is the feature vector of \texttt{D} features that we wish to match and array \texttt{S} 
-is the database of \texttt{N} vectors of size \texttt{D} that we match against.
-An intuitive implementation is as follows:
-
-{\small
-\begin{verbatim}
-def biometric(C: shared[list[int]], D: int, 
-              S: shared[list[int]], N: int) -> 
-                      tuple[shared[int], shared[int]]:
-    min_sum = 10000
-    min_index = 0  
-    for i in range(N): #loop over database
-        sum = 0
-        for j in range(D): #loop over features
-            d = S[i * D + j] - C[j] #i.e., d = S[i,j] - C[j]
-            p = d * d
-            sum = sum + p
-        if sum < min_sum:
-            min_sum = sum
-            min_index = i
-    return (min_sum, min_index)
-\end{verbatim}
-}
-
-Our compiler takes (essentially) standard IMP syntax.
-The programmer can write intuitive iterative programs as the one above.
-They annotate certain inputs and outputs as \emph{shared}.
-Here the code iterates over the entries in the database and computes
-the sum of squares of the differences of individual features. The program returns the 
-index \texttt{i} of the vector that gives the best match plus the corresponding sum of squares.
-
-Our compiler imposes the following restrictions. We note that in some cases, the restrictions 
-can be easily lifted and we plan to do so in future iterations of our compiler.
-
-\begin{enumerate}
-\item The program contains arbitrarily nested loops, however, loop bounds are fixed: \texttt{0 <= i < N}.
-A standard restriction in MPC is that the bounds must be known at circuit-generation time.
-\item Arrays are one-dimensional. N-dimensional arrays are linearized and accessed 
-in row-major order and at this point the programmer is responsible for linearization
-and access. 
-\item Array subscrpts are plaintext values.
-\item Our compiler allows for output (write) arrays, however it restricts
-write access to \emph{canonical writes} along the dimensions of the array. I.e., \texttt{A[i,j] = ...}
-where \texttt{i} and \texttt{j} loop over the two dimensions of \texttt{A} is allowed, 
-but \texttt{A[i,j+2] = ...} is not allowed. Read access is arbitrary. 
-\end{enumerate}
-
-\subsection{MPC Source and Cost of Schedule}
-
-The compiler generates an IR, MPC source:
-{\small
-\begin{verbatim}
-1. min_sum!1 = 10000
-2. min_index!1 = 0
-3. for i in range(0, N!0):
-4.    min_sum!2 = PHI(min_sum!1, min_sum!4)
-5.    min_index!2 = PHI(min_index!1, min_index!4)
-6.    sum!2 = 0
-7.    for j in range(0, D!0):
-8.      sum!3 = PHI(sum!2, sum!4)
-9.      d!3 = (S!0[((i * D!0) + j)] - C!0[j]) // MPC
-10.     p!3 = (d!3 * d!3) // MPC
-11.     sum!4 = (sum!3 + p!3) // MPC
-12.  !1!2 = (sum!3 < min_sum!2) // MPC
-13.  min_sum!3 = sum!3
-14.  min_index!3 = i
-15.  min_sum!4 = MUX(!1!2, min_sum!3, min_sum!2) // MPC
-16.  min_index!4 = MUX(!1!2, min_index!3, min_index!2) // MPC
-17.!2!1 = (min_sum!2, min_index!2)   
-\end{verbatim}
-}
-
-The compiler linearizes the source turning conditionals into MUX statements. The PHI nodes
-are remnants of the SSA IR; the compiler generates code that picks the correct value 
-when producing MOTION output; the MOTION framework in turn linearizes 
-loops when it generates the circuit.
-
-We turn to our analytical model to compute the cost of this program. Assuming fixed 
-cost $\beta$ for a local MPC operation (essentially just ADD) and cost $\alpha$ for a remote
-MPC operation (e.g., MUX, CMP, and remaining operations), the cost of the iterative schedule 
-will be $N*D*(2*\alpha+\beta) + N*3*\alpha$.
-
-A key contribution is the vectorizing transformation. We can compute all $N*D$ 
-subtraction operations (line 9) in a single SIMD instruction; similarly we can compute 
-all multiplication operations (line 10) in a single SIMD instruction. And while we cannot
-vectorize computation of the N individual sums, we can compute the N sums in parallel. 
-Our compiler \emph{automatically detects these opportunities and transforms the program}. 
-It is standard that MPC researchers write vectorized versions of the Biometric
-program by hand; we are the first (to the best of our knowledge) to automatically 
-transform an intuitive, iterative MPC program into an unintuitive vectorized one.
-
-\subsection{Vectorized MPC Source and Cost of Schedule}
-
-Our compiler produces the following vectorized program. (Note that this is
-still higher-level IR, Vectorized MPC Source. Our compiler turns this code into MOTION variables, 
-loops and SIMD primitives, which MOTION then uses to generate the circuit.)
-
-{\small
-\begin{verbatim}
-min_sum!1 = 10000
-min_index!1 = 0
-// S!0^ is same as S!0. C!0^ replicates C!0 N-times: 
-S!0^ = raise_dim(S!0, ((i * D!0) + j), (i:N!0,j:D!0))
-C!0^ = raise_dim(C!0, j, (i:N!0,j:D!0))
-
-sum!2 = [0,..,0]
-// Computes all differences and all products "at once"
-d!3[I,J] = SUB_SIMD(S!0^[I,J],C!0^[I,J])
-p!3[I,J] = MUL_SIMD(d!3[I,J] * d!3[I,J])
-
-for j in range(0, D!0):
-  // sum!2[I], sum!3[I], sum!4[I] are vectors of size N  
-  // Computes N intermediate sums "at once"
-  sum!3[I] = PHI(sum!2[I], sum!4[I])       
-  sum!4[I] = ADD_SIMD(sum!3[I],p!3[I,j])
-
-min_index!3 = [0,1,...N!0-1]
-
-for i in range(0, N!0):
-  min_sum!2 = PHI(min_sum!1, min_sum!4) 
-  !1!2[i] = CMP(sum!3[i],min_sum!2)
-  min_sum!4 = MUX(!1!2[i], sum!3[i], min_sum!2)
-
-for i in range(0, N!0):
-  min_index!2 = PHI(min_index!1, min_index!4)  
-  min_index!4 = MUX(!1!2[i], min_index!3[i], min_index!2)
-!2!1 = (min_sum!2, min_index!2)   
-\end{verbatim}
-}
-
-In MPC compilers a vectorized operation that computing M operations "at once" costs 
-essentially the same ($\alpha$ or $\beta$) as an individual operation. We elaborate 
-on these in the following section. Thus, the vectorized program costs 
-$2*\alpha$ + $D * \beta + N*3*\alpha$. The first term in the sum corresponds to the vectorized
-subtraction and multiplication, the second term corresponds to the for loop on $j$ and the third 
-one corresponds to the remaining for loops on $i$. Clearly, $2*\alpha + D * \beta + N*3*\alpha << N*D*(2*\alpha+\beta) + N*3*\alpha$.
-Our experimental results illustrate this as well. \ana{Add numbers.}
-
-\section{Analytical Model}
-\label{sec:model}
-
-\subsection{Scheduling in MPC}
-\label{sec:mpc}
-
-%\ana{Here we need more on MPC before jumping to scheduling. MPC-source, basic assumptions about static loop bounds, etc.}
-
-For this treatment we make the following simplifying assumptions:
-
-\begin{enumerate}
-\item All statements in the program execute using the same protocol (sharing). That is, there is no share conversion. 
-%\ishaq{This is not an assumption, this is our setting. We work in the single protocol world.}
-\item There are two tiers of MPC instructions, local and remote. A local instruction (essentially just ADD) has cost $\beta$ 
-and a remote instruction (e.g., MUX, MUL, SHL, etc.) has cost $\alpha$, where $\alpha >> \beta$. We assume that all remote 
-instructions have the same cost.
-%\ishaq{Don't we need to distinguish between local (cheap) vs.
-interactive (expensive) instructions?
-I don't see how we could assume this.
--- Past suggestion from Vassilis: Ideally we should assume cost to be a convex function.}\ana{After discussion 1/12: replace single costs 1 with two levels of costs: $\alpha$ for non-local operations, e.g., MUL, and $\beta$ for local ops, e.g., ADD.}
-\item We assume infinite parallel capacity---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely $\alpha$ or $\beta$. 
-This is a standard assumption in Cryptographic Parallel RAM. ABY presents empirical support for this assumption~\ana{Add citations. PRAM, ABY.}
-%\ishaq{Comment from Vassilis: We assume there are infinite parallel capacities, this is the standard assumption in PRAM (Cryptographic Parallel RAM).}\ana{PRAM assumption stays.
-We have to replace the unit cost of 1 with a suitable amortization function $f(n)$, which I don't think changes much in the cost modeling and analysis.}
-\item MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption,
-2 MUL instructions scheduled in parallel benefit from amortization and cost $\alpha$, however a MUL and a MUX instructions scheduled
-in parallel still cost $2\alpha$. 
-%\ishaq{We need to reword this assumption to something like "$K$ parallel MUL costs much less than $K$ (because they get amortized), but any mix of $K$  MUL and MUX still cost roughly $K$.
-Specifically, we should take away the low constants (2 and 1) because we know for low constants this is not true.} \ana{Again, core assumption that MUL and MUX don't benefit from amortization stays.
-We have to replace constant costs with functions, as in (3).}
-\end{enumerate}
-
 \subsection{Problem Statement}
@@
 \end{comment}
-
-\section{\bf Compiler Framework}
-\label{sec:preliminaries}
-
-\begin{figure*}
-  \includegraphics[width=1.0\linewidth]{figs_paper_SIMD/compiler_framework.png}
-  \caption{Compiler Framework.} 
-  \label{fig:compiler_framework}
-\end{figure*}
-
-%We assume arbitrarily nested loops in the MPC-source IR and read-only arrays.
-We assume that loops range from 0 to some constant N.
-Arrays are linearized (row-major order as in MOTION) and accesses are via functions of the induction variables of the enclosing loops.
-
-\figref{fig:compiler_framework} presents an overview of our compiler. In this section, we describe several of the phases of the compiler. 
-Sections~\secref{sec:vectorization} and~\secref{sec:div_and_conq} describe vectorization and divide-and-conquer.
-We write $i,j,k$ to denote the loop nest: $i$ is the outermost loop, $j$, is immediately nested in $i$, and so on until $k$
-and we use $I,J,K$ to denote the corresponding upper bounds. For simplicity, we write $A[i,j,k]$ to denote canonical access 
-to an array element. In the program, canonical access is achieved via the standard row-major order formula: $(J*K)*i + K*j + k$.
-To simplify the presentation we describe our algorithms in terms of three-element tuples $i,j,k$.
-All discussion generalizes to arbitrarily large loop nests.
-%\ana{Need to state this more precisely.}
-
-\subsection{Semantic Analysis}
-
-\ana{Benjamin?}
-
-\subsection{MUX Nodes and Pseudo $\phi$-nodes}
-
-\ana{Benjamin?}
-
-\ana{I think we need to add more here. Conversion from SSA to MPC Source is not so trivial.}
-
-A pseudo $\phi$-node $X_1 = \phi(X_0,X_2)$ in a loop header is evaluated during circuit generation.
-If it is the 0-th iteration, then the $\phi$-node evaluates to $X_0$, otherwise, it evaluates to $X_2$.
-
-%\ana{We need to double check that the following captures all cases of def-use edges.}
-
-\subsection{Dependence Analysis}
-
-\subsubsection{Def-use Edges}
-
-The dependence graph has the following def-use edges:
-
-\begin{itemize}
-\item same-level edge $X\rightarrow Y$ where $X$ and $Y$ are in the same loop nest, say $i,j,k$.
-E.g., the def-use edge from \texttt{d = S[i,j] - C[j]} to \texttt{p = d*d} in the Biometric MPC-source is a same-level edge.
-A same-level edge can be a back-edge in which case a $\phi$ node is the target of the edge.
-E.g., $\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{sum}_1,\texttt{min}_1)$ to $\texttt{min}_0 = \phi(\texttt{min}_1, 10000)$ in Biometric is a same-level back-edge.
-\item outer-to-inner $X\rightarrow Y$ where $X$ is in an outer loop nest, say $i$, and $Y$ is in an inner one, say $i,j,k$. 
-%A $\phi$-node can be a source or a target of an outer-to-inner forward edge.
-\item inner-to-outer $X\rightarrow Y$ where $X$ is a $phi$-node in an inner loop nest, $i,j,k$, and $Y$ is in the enclosing loop nest $i,j$.
-E.g.
-$\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$ is an inner-to-outer edge.
-An inner-to-outer edge can be a back-edge as well in which case both $X$ and $Y$ are phi-nodes with the source $X$ in a loop nested into $Y$'s loop (not necessarily immediately).
-
-%Note that the source is \emph{always} a $\phi$-node of the $i,j,k$ loop and the target is in the immediately enclosing loop.
-The interpretation of this edge is that the use node $Y$ uses the definition made in the last iteration of the inner loop.
-%\ben{The current representation of pseudo $\phi$-nodes shows them attached to the loop header (i.e. in the loop).  We may want to clarify that they are also evaluated at the loop's termination.}
-%\item same-level back-edge $X\rightarrow Y$.
-$Y$ is a phi-node in the header of the loop and $X$ is a definition of the variable in the loop body.
-E.g., $\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{sum}_1,\texttt{min}_1)$ to $\texttt{min}_0 = \phi(\texttt{min}_1, 10000)$ in Biometric is a same-level back-edge.
-%\item inner-to-outer back-edge $X\rightarrow Y$: $X$ and $Y$ are both $\phi$-nodes for some variable. The source $X$ is in a loop nested into $Y$'s loop (not necessarily immediately).
-
-\item mixed forward edge $X\rightarrow Y$.
-$X$ is in some loop $i,j,k$ and $Y$ is in a loop nested into $i,j,k'$.
-We transform mixed forward edges as follows.
-Let $\x$ be the variable defined at $X$.
-We add a variable and assignment $\x' = \x$ immediately after the $i,j,k$ loop.
-Then we replace the use of $\x$ at $Y$ with $\x'$.
-This transforms a mixed forward edge into an "inner-to-outer" forward edge followed by an outer-to-inner forward edge.
-Thus, Basic Vectorization handles one of "same-level", "inner-to-outer", or "outer-to-inner" def-use edges.
- \end{itemize}
-
-\subsubsection{Closures}
-
-We define $\mathit{closure}(n)$ where $n$ is a phi-node. Intuitively, it computes the set of nodes (i.e., statements) that form a dependence cycle with $n$. The closure of $n$ is defined as follows:
-\begin{itemize}
-\item $n$ is in $\mathit{closure}(n)$
-\item $X$ is in $\mathit{closure}(n)$ if there is a same-level path from $n$ to $X$, and $X \rightarrow n$ is a same-level back-edge.
-\item $Y$ is in $\mathit{closure}(n)$ if there is a same-level path from $n$ to $Y$ and there is a same-level path from $Y$ to some $X$ in $\mathit{closure}(n)$.
-\end{itemize}
-
-\subsection{Taint Analysis}
-
-\ana{Ben?}
-
-\section{Vectorization}
-\label{sec:vectorization}
-
-An important component of our algorithm is the "lifting" of scalars to the corresponding loop dimensionality. 
-For example, \texttt{d = S[i * D + j] - C[j]} equiv. to \texttt{d = S[i,j] - C[j]}, which gave rise to \texttt{N*D} subtraction operations in the sequential schedule, 
-is lifted. The argument arrays \texttt{S} and \texttt{C} are lifted and the scaler \texttt{d} is lifted: \texttt{d[i,j] = S[i,j] - C[i,j]}.
-The algorithm then detects that the statement can be vectorized.
-
-There are three kinds of arrays (for now all kept internally as one-dimensional arrays, but that's under discussion).
-\begin{itemize}
-
-\item Scalars: These are scalar variables we lift into arrays for the purposes of vectorization. 
-For those, all writes are canonical writes and all reads are canonical reads. We will apply raise dimension when 
-a scalar is used in an inner loop (e.g., \texttt{sum0} in line 6 of the MPC source code will be raised to a 1-dimensional 
-array since \texttt{sum0} is used in the inner $j$-loop). Drop dimension applies as well; this happens when a scalar
-written in an inner loop is used outside of the loop (e.g., \texttt{sum0} for which the lifted inner loop computes 
-\texttt{D} values, but the outer loop only needs the last one.)
-
-\item Read-only input arrays: Read-only inputs. There are NO writes, while we may have non-canonical reads, $f(i,j,k)$. 
-Phase 1 of Basic vectorization will add raise dimension operation at the beginning of the function to lift these arrays, 
-and raise dimension may \emph{reshape} arrays. If there are multiple "views" of the input array, there would be multiple raise dimension statements to create
-each one of these views. The invariant is that at reads in loops, the reads of "views" of the original input array are canonical.
-Only raise dimension applies to these arrays, and only in the beginning of the program. For example, 1-dimensional array 
-\texttt{C} is lifted into 2-dimensional array \texttt{N x D} by copying the row \texttt{N} times.
-
- \item Read-write output arrays: Writes are canonical (by restriction) but reads can be non-canonical. 
- Dependence analysis limits vectorization when non-canonical read access refers to array writes in
- previous iterations, thus creating loop-carried dependences. We may apply both raise and drop dimension, however, 
- they respect the fixed dimensionality of the output array. The array cannot be raised to a dimension lower than
- its canonical (fixed) dimensionality and it cannot be dropped lower. In addition, non-canonical reads may 
- require lifting (i.e., reshaping) of the array after the most recent write, rather than in the beginning of the program, 
- in order to reduce a non-canonical read to a canonical one.
-
-\end{itemize}
-
-In the sections below we detail the $\mathit{raise\_dim}$ (raise dimension) and $\mathit{drop\_dim}$ (drop dimension) operations,
-followed by our vectorization algorithm.
-
-%\ana{Add back-edges into Phase 1.
-A back-edge from a non-phi-node in loop $i$ to a phi-node in loop $i$'s header is a same-level edge.
-The only difference with the handling of normal same-level forward def-use is that the operand index will become $i-1$.
-A back edge from a phi-node in loop $j$ to a phi-node in loop $i$, where $j$ is nested in $i$ is an inner-to-outer edge and will require dropping dimension.
-We can show these are the only kinds of back-edges that may occur.}
-
-%\subsection{Algorithm (Still work in progress.)}
-
-\subsection{Raise Dimension and Drop Dimension}
-
-There are two conceptual versions of $\mathit{raise\_dim}$. One applies on read-only input arrays and reshapes those arrays 
-when necessary to ensure canonical read access in the corresponding loop. The signature of $\mathit{raise\_dim}$
-is as follows. It takes the original array \texttt{C}, the access pattern function $f(i,j,k)$ in loop nest $i,j,k$ and the loop bounds
-\texttt{((i:I),(j:J),(k:K))}: 
-\[ \mathit{raise\_dim}(\texttt{A}, f(i,j,k), \texttt{((i:I),(j:J),(k:K))}) \]
-It produces a new 3-dimensional array \texttt{A'} by iterating over $i,j,k$ and setting each element of  \texttt{A'} as follows:
-\[ \texttt{A'[i,j,k] =  A[f(i,j,k)]} \]
-The end result is that uses of \texttt{[A[f(i,jk)]} in loop nest $i,j,k$ are replaced with canonical read-accesses to \texttt{A'[i,j,k]}
-that can be vectorized. In the running Biometric example, \texttt{C'} = $\mathit{raise\_dim}$\texttt{(C, j, (i:N,j:D))} lifts the 
-1-dimensional array \texttt{C} into a 2-dimensional array. The $i,j$ loop now accesses \texttt{C'} in the canonical way, \texttt{C'[i,j]}. 
-Similarly, \texttt{S'} = $\mathit{raise\_dim}$\texttt{(S, i*D+j, (i:N,j:D))} tries to lift \texttt{S}, but the operation turns into a no-op 
-because \texttt{S} is already a 2-dimensional array and the read access is canonical.
-
-The other version of $\mathit{raise\_dim}$ applies on scalars and read-write arrays. It lifts a lower-dimension array into a 
-higher-dimension for access in a nested loop. Here \texttt{A} is an $i$ array and raise dimension adds two additional dimensions:
-\[ \mathit{raise\_dim}(\texttt{A}, \texttt{(j:J,k:K)}) \]
-This version is reduced to the above version by adding the access pattern function, which is just $i$: 
-\[ \mathit{raise\_dim}(\texttt{A}, i, \texttt{(j:J,k:K)}) \]
-
-The corresponding $\mathit{drop\_dim}$ is carried out when an array written in an inner loop is used in an enclosing loop. 
-It takes a higher dimensional array, say $i,j,k$ and removes trailing dimensions, say $j,k$:
-\[ \mathit{drop\_dim}(\texttt{A}, \texttt{(j:J,k:K)}) \]
-It iterates over $i$ and takes the result at the maximal index of $j$ and $k$, i.e., the result at the last iterations of $j$ and $k$:
-\[ \texttt{A'[i] = A[i,J-1,K-1]} \]
-
-\begin{comment}
-We define the $\mathit{raise\_dim}$ (raise dimensions) and $\mathit{drop\_dim}$ (drop dimension) functions.
-Raise dimension "lifts" a lower-dimension array (\ana{The right term here is tensor, I am pretty sure!}) into a higher dimension one.
-This is necessary when a lower-dimensional array is used in a higher dimensional loop and, essentially, is just copying of values.
-For example, Biometric contains the statement $d = \mathit{SUB}(S[i, j], C[j])$, in the $j$-loop which is nested into the $i$-loop.
-Here $C$ is a one-dimensional array, however, to vectorize across both loops it is necessary to turn it into a two-dimensional array: $C[i,j]$ becomes $[C[0],C[1],...C[J],C[0],C[1],...C[J], ...
-C[0],C[1],...C[J]]$, which turns the row into a matrix of $I$ identical rows.
-(We use capital letters to denote the upper bounds of loops, e.g., $J$ is the upper bound of the $j$-loop and $I$ is the upper bound of the $i$-loop.)
-
-$\mathit{raise\_dim}(A[i_1,...,i_k],i_j)$ is defined as follows.
-It results in a new {k+1}-dimensional array $A'$ where for every $0 \le i_{k} < I_{k+1}$, $A'[i_1, \ldots, i_j, \ldots, i_k] = A[i_1, \ldots, i_k]$.
-Adding $n$ dimensions is trivially extended as a composition of $n$ $\mathit{raise\_dim}$ that each adds a single dimension.
-
-As expected, drop dimension turns a higher-dimensional array into a lower-dimensional one.
-The key use case is an inner-to-outer def-use edge.
-The code may define a variable, e.g., $\x$ in an inner loop, say $j$, then use this variable in an enclosing loop, say $i$.
-Our algorithm may vectorize the computation of $\x$ in the $j$-loop thus producing a vector $\x[j]$ where $\x[1]$ is the value of \x after the 1st iteration and so on.
-Drop dimension states that the outer loop will use the value of the variable at the last iteration.
-
-$\mathit{drop\_dim}(A[i_1,\ldots, i_j, \ldots,i_k], i_j)$ produces a k-dimensional array $A'$ where $A'[i_1,\ldots,i_k] = A'[i_1,\ldots, I_j-1, \ldots,i_k]$.
-In our analysis, dimensions are always dropped at the end, and again, one can define dropping $n$ dimensions as a composition of $n$ $\mathit{drop\_dim}$.
-
-\ben{I'm not sure if I wrote up the above paragraphs correctly -- the idea is that we're just copying values to extend over dimension $j$ or only retain its last elements.
-The current writeup implies that $1 < j < k$ though, which isn't always true.
-Python implementations of $\mathit{raise\_dim}()$ and $\mathit{drop\_dim}()$ can be found here: \url{https://github.com/milana2/ParallelizationForMPC/blob/master/compiler/compiler/motion_backend/reference_implementations.py}}
-
-\ana{This section is still informal. Needs work to make more precise.}
-\end{comment}
-
-\subsection{Basic Vectorization}
-
-\begin{algorithmic}
-
-\STATE \COMMENT { Phase 1: Raise dimension of scalar variables to corresponding loop nest. We can traverse stmts linearly in MPC-source. }
-
-\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i,j,k$}
-%\FOR {each $Y_n$}
-\FOR {each argument $Y_n$} %def-use edge $\mathit{stmt}' (\mbox{def of } Y) \rightarrow \mathit{stmt} (\mbox{def of } X)$}
-\STATE {case def-use edge $\mathit{stmt}' (\mbox{def of } Y_n) \rightarrow \mathit{stmt} (\mbox{def of } X)$ of} \\
-\STATE \hspace{0.25cm} {same-level: $Y'_n$ is $Y_n$} \\
-\STATE \hspace{0.25cm} {outer-to-inner: add $Y'_n[i,j,k] = \mathit{raise\_dim}(Y_n)$ at $\mathit{stmt'}$}
-\STATE \hspace{0.25cm} {inner-to-outer: add $Y'_n[i,j,k] = \mathit{drop\_dim}(Y_n)$ at $\mathit{stmt}$}
-\ENDFOR
-
-\COMMENT { Optimistically vectorize all. $I$ means vectorized dimension. }
-%\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i,j,k$}
-\STATE {change to $X[I,J,K] = \mathit{Op}(Y'_1[I,J,K],Y'_2[I,J,K])$ }
-\ENDFOR
-
-\COMMENT { Phase 2: Recreating FOR loops for cycles; vectorizable statements hoisted up. }
-
-\FOR {each dimension $d$ from highest to 0}
-\FOR {each $\phi$-node $n$ in loop $i_1,...,i_d$}
-\STATE {compute $\mathit{closure}(n)$}
-\ENDFOR
-
-\COMMENT {{\color{red} $cl_1$ and $cl_2$ intersect if they have common statement or update same array; "intersect" definition can be expanded}}
-
-\WHILE {there are closure $cl_1$ and $cl_2$ that intersect}
-\STATE {merge $cl_1$ and $cl_2$}
-\ENDWHILE
-\FOR {each closure $cl$ (after merge)}
-\STATE {create FOR $i_d = 0; ...$ loop}
-\STATE {add $\phi$-nodes in $cl$ to header block}
-\STATE {{\color{red}{add target-less $\phi$-node for A if $cl$ updates array A}}}
-\STATE {add statements in $cl$ to loop body in some order of dependences}
-\STATE \COMMENT { Dimension is not vectorizable: }
-\STATE {change ${I_d}$ to $i_d$ in all statements in loop}
-\STATE {treat FOR loop as monolith node: some def-use edges become same-level.}
-\ENDFOR
-{
-\color{red}
-\FOR {each target-less $\phi$-node $A_1 = \phi(A_0,A_k)$}
-\STATE {in vectorizable stmts, replace use of $A_1$ with $A_0$}
-\STATE {discard $\phi$-node if not used in any $cl$}
-\ENDFOR
-}
-\ENDFOR
-
-\COMMENT { Phase 3: Remove unnecessary dimensionality. }
-
-\COMMENT { A dimension $i$ is dead on exit from stmt $X[...i...] = ...$ if all def-uses with targets outside of the enclosing 
-FOR $i=0 ...$ MOTION loop end at target (use) $X' = \mathit{drop\_dim}(X, i)$. } 
-\FOR {each stmt and dimension $X[...i...] = ...$}
-\STATE {if $i$ is a dead dimension on exit from stmt $X[...i...] = ...$, remove $i$ from $X$ (all defs and uses)} 
-%\COMMENT{ A stmt v[...i] is a "dead dimension i" if all edges out of FOR i = 0; ... loop end at drop\_dim(v, i)}
-\ENDFOR
-
-\COMMENT { Now clean up drop\_dim and raise\_dim }
-\FOR {each $X' =  \mathit{drop\_dim(X,i)}$ }
-\STATE {replace with $X' = X$ if $i$ is dead in $X$.}
-\ENDFOR
-\STATE {do (1) (extended) constant propagation, (2) copy propagation and (3) dead code elimination to get rid of redundant variables and raise and drop dimension statements}
-
-\COMMENT { Phase 4: }
-\STATE {add SIMD for simdified dimensions}
-
-\end{algorithmic}
-
-\input{biometric_vectorization_example.tex}
-
-\begin{comment}
-\subsection{Example: Biometric}
-
-\subsubsection{MPC-source}
-
-Below is the MPC-source resulting from Benjamin's analysis. It gives rise to a corresponding dependence graph. (Not shown.)
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0 = \phi(-1, \texttt{index}_1)$}
-\STATE {$\texttt{min}_0 = \phi(10000,\texttt{min}_1)$}
-
-\STATE \COMMENT {Begin of inner loop $j$.}
-\STATE {$\texttt{sum}_0 = \phi(0, \texttt{sum}_1)$}
-\STATE {$\texttt{d} = \mathit{SUB}(\texttt{S}[i,j], \texttt{C}[j])$}
-\STATE {$\texttt{p} = \mathit{MUL}(\texttt{d},\texttt{d})$}
-\STATE {$\texttt{sum}_1 = \texttt{ADD}(\texttt{sum}_0,\texttt{p})$}
-\STATE \COMMENT {End of inner loop $j$.}
-\STATE {$\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$}
-\STATE {$\texttt{index}_1 = \mathit{MUX}(\texttt{c},\texttt{index}_0,i)$}
-\STATE {$\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{min}_0,\texttt{sum}_0)$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-\subsubsection{Phase 1 of Basic Vectorization:}
-
-The transformation preserves the dependence edges.
-It raises the dimensions of scalars and optimistically vectorizes all operations.
-The next phase discovers loop-carried dependences and removes affected vectorization.
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$}
-\STATE {$\texttt{min}_0[i] = \phi(10000, \texttt{min}_1[i-1])$}
-
-\STATE \COMMENT {Begin of inner loop $j$.}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...], \texttt{sum}_1[i,j-1])$}
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{sum}_1[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE \COMMENT {End of inner loop $j$.}
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {$\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$}
-\STATE {$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$}
-\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-\ishaq{At least for input variables: when raising dimensions, we should keep some meta data around so we can later write code that could use 1 share if it knows the other N shares are just duplicates of it.}
-
-\subsubsection{Phase 2}
-
-This phase analyzes statements from the innermost loop to the outermost. The key point is to discover loop-carried dependencies and re-introduce loops whenever dependencies make this necessary.
-
-Starting at the inner phi-node $\texttt{sum}_0[i,j] = \phi([0,0,...],\texttt{sum}_1[i,j-1])$, the algorithm first computes its closure.
-The closure amounts to the phi-node itself and $\texttt{sum}_1[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$, accounting for the loop-carried dependency of the computation of \texttt{sum}.
-The algorithm replaces this closure with a FOR loop on $j$ removing vectoriaztion on $j$.
-Note that the SUB and MUL computations remain outside of the loop as they do not depend on phi-nodes that are part of cycles.
-The algorithm adds same-level edges, one from $\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = ...$ to the monolithic FOR-loop node, and one from the FOR-loop node to $\texttt{c}[\stackrel{\rightarrow}{i}] = ...$.
-
-\ishaq{For our radar: Such dependencies will not involve phi nodes only. e.g. $a[i] = a[i] + a[i-1]$}
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0[i] = \phi(-1, \texttt{index}_1[i-1])$}
-\STATE {$\texttt{min}_0[i] = \phi(10000, \texttt{min}_1[i-1])$}
-
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-
-\STATE \COMMENT {Begin of inner loop $j$. Loop is now much "shorter"!}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...], \texttt{sum}_1[i,j-1])$}
-\STATE {FOR j=0; j<D; j++} \COMMENT{ Will turn into a MOTION loop.}
-\STATE \hspace{0.25cm}{$\texttt{sum}_1[\stackrel{\rightarrow}{i},{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},{j}],\texttt{p}[\stackrel{\rightarrow}{i},{j}])$}
-\STATE \COMMENT {End of inner loop $j$.}
-
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {$\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$}
-\STATE {$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$}
-\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-Next, analysis moves to outer dimension $i$. There are two phi-nodes, $\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1])$ and $\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$.
-The closure of the first is
-\[
-\begin{array}{l}
-\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1]) \\
-\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
-\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])
-\end{array}
-\]
-and the closure of the second one is
-\[
-\begin{array}{l}
-\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1]), \\
-%\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
-\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])
-\end{array}
-\]
-
-Since the two closures \emph{do not} intersect, we have two distinct FOR-loops on $i$.
-The first FOR loop (on \texttt{min}) is scheduled first, because of the dependence edge from $\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$ to
-$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$.
-We can now rewrite those statements, getting rid of the vectorization on $i$ in the two FOR loops.
-Notably, the SUB and MUL computations are fully vectorizable and the ADD computation is vectorizable across the $i$-dimension.
-
-\begin{algorithmic}
-
-\STATE
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB\_SIMD}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL\_SIMD}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$} \\
-\STATE {}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...],\texttt{sum}_1[i,j-1])$}
-\STATE {FOR j=0; j<D; j++} \COMMENT{ Turns into a MOTION loop, was $j$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{sum}_1[\stackrel{\rightarrow}{i},{j}] = \mathit{ADD\_SIMD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},{j}],\texttt{p}[\stackrel{\rightarrow}{i},{j}])$}
-\STATE {}
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {}
-\STATE {$\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1])$}
-\STATE {FOR i=0; i<N; i++} \COMMENT{ Turns into a MOTION loop, was $i$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{c}[{i}] = \mathit{CMP}(\texttt{sum}'_0[{i}],\texttt{min}_0[{i}])$}
-\STATE \hspace{0.25cm}{$\texttt{min}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{min}_0[{i}],\texttt{sum}'_1[{i}])$}
-\STATE {}
-\STATE {$\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$}
-\STATE {FOR i=0; i<N; i++} \COMMENT{ Turns into a MOTION loop, was $i$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{index}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{index}_0[{i}],i)$}
-
-\ben{The above code implicitly stores the final values of min\_0 and index\_0 in their last indices, but there is no primitive like drop\_dim to make it clear.
-This will have to be added in the SSA code generation (though it might not be necessary to include in the paper?).
-It also is pretty wishy-washy about the size and layout of each vector (again more of an implementation issue), but maybe we should have a different syntax for array accesses for read/write (e.g.
-the ADD\_SIMD steps) vs the array definitions (e.g.
-the SUB\_SIMD and raise/drop\_dim statements)}
-
-\end{algorithmic}
-\end{comment}
-
-\subsection{Correctness Argument}
-
-We build a correctness argument that loosely follows the theory of Abstract Interpretation. 
-We define the syntax of MPC Source programs. The domain of MPC Source programs
-expressible in the syntax (with certain semantic restrictions) is the abstract domain $A$. 
-We then define the \emph{linearization} of an MPC Source program as an interpretation 
-over the syntax. The linearization, which is a \emph{schedule}, is the concrete domain $C$. 
-Since we reason over def-use graphs in $A$ we define a partial order relation over elements of $A$ 
-in terms of def-use relations. We define a partial order over elements of $C$ as well,
-in terms of def-use relations in the concrete domain $C$. We prove two theorems
-that state (informally) that the schedule corresponding to the original program computes the same 
-result as the schedule corresponding to the vectorized program.
-
-\begin{figure}%[!b]
-%\small
-\[
-\begin{array}{l@{~~~~~}l}
-  \begin{array}{l@{~}l@{~~~}l}
-  s &::= s;s & sequence \\
-     & \mid \x[i,J,k] = \y[i,J,k] \; \mathit{op\_SIMD} \; \z[i,J,k] & operation \\ 
-     & \mid \x[i,J,k]  = \texttt{PHI}(\x_1[i,J,k],\x_2[i,J,k\!-\!1]) & \mathit{phi} \; \mathit{node} \\
-     & \mid \x[i,J,k] = \mathit{raise\_dim}(\x'[i],(J\!:\!J,k\!:\!K)) & \mathit{raise} \; \mathit{dimension(s)} \\
-     & \mid \x[i,J] = \mathit{drop\_dim}(\x'[i,J,k],k)  & \mathit{drop} \; \mathit{dimension(s)} \\
-     & \mid \x = \y & \mathit{propagation} \\
-     & \mid \texttt{FOR} \; 0 \le i < I: s  & \mathit{loop} \\
-  \end{array}    \\
-\end{array}
-\]
-\vspace{-3mm}
-\caption{MPC Source Syntax}\label{fig:MPC_Source_syntax}
-\vspace{-5mm}
-\end{figure}
-
-\paragraph{MPC Source Syntax}
-
-\figref{fig:MPC_Source_syntax} defines the syntax for our intermediate representation, MPC Source. 
-There are semantics restrictions over the syntax as well: a variable $\x[i,j,k]$ is a 3-dimensional array 
-$(i:I,j:J,k:K)$ and also, a statement $\x[i,J,k] = ...$ is enclosed in loops over $i$ and $k$ as shown below. 
-Thus, $i$ and $k$ are in scope.
-\begin{verbatim}
-FOR 0 <= i < I:
-    ...
-    FOR 0 <= k < K:
-        x[i,J,k] = ...
-\end{verbatim}
-
-Statements $\mathit{operation, phi, raise \; dimension(s), drop \; dimension(s)}$ are \emph{base} statements, 
-and $\mathit{sequence, loop}$ are \emph{compound statements}.
-
-\paragraph{Linearization}
-
-Linearization is the concretization operation, which, as we mentioned earlier computes a schedule.
-The concretization function $\gamma: A \rightarrow C$ is defined as an interpretation of
-MPC Source syntax, as it is standard. The concretization of each one 
-of the base statements is as follows: 
-\[\begin{array}{l}
-\gamma(\x[i,J,k] = \y[i,J,k] \; \mathit{op\_SIMD} \; \z[i,J,k]) = \\ 
-\x[i,0,k] = \y[i,0,k] \; \mathit{op} \; \z[i,0,k] \;\; || \\
-\x[i,1,k] = \y[i,1,k] \; \mathit{op} \; \z[i,1,k] \;\; || \; ...\; || \\
-\x[i,I-1,k] = \y[i,I-1,k] \; \mathit{op} \; \z[i,I-1,k]
-\end{array}\]
-meaning that the vectorized dimension(s) are expanded into parallel statements. || introduces SIMD (parallel) execution.
-
-The concretization of the FOR statement is as follows:
-\[ \gamma(\texttt{FOR} \; 0 \le i < I: s) = \gamma(s)[0/i] ; \gamma(s)[1/i] ; ... \gamma(s)[I-1/i] \]
-$\gamma$ simply unrolls the loop substituting $i$ with 0, 1, etc. Here ; denotes sequential execution.
-
-\paragraph{Partial Orders}
-
-For each MPC Source program $a$ we compute the def-use edges in the standard way: if base statement $s1 \in a$ defines variable \x, e.g., $\x[i,j,k] = ...$, 
-and base statement $s2 \in a$ uses \x, e.g., $... = ... \x[i,j,k]$ and there is a path in the trivial CFG from $s1$ to $s2$, then there is a def-use edge from
- $s1$ to $s2$. We extend the dimensionality of a statement into $s1[i,j,k]$ where $s1[i,j,k]$ inherits the dimensionality of the left-hand-side of the assignment.
-
- Let $a_0, a_1$ be two MPC Source programs in $A$. 
- Two base statements, $s_0 \in a_0$ and $s_1 \in a_1$ are \emph{same}, written $s_0 \equiv s_1$ if they are of the same operation and they operate on 
- the same variables: same variable name and same dimensionality. Recall that dimensions in MPC Source are either iterative, lower case, or vectorized, upper case. 
- Two statements are same even if one operates on an iterative dimension and the other one operates on a vectorized one, e.g., $s_0[i,j,k] \equiv s_1[I,j,K]$.
-
- \begin{definition} Let $a_0, a_1 \in A$. We say that $a_0 \le a_1$ iff for every def-use edge $s1 \rightarrow s2$ in $a_0$ there is an edge
- $s1' \rightarrow s2'$ where $s1 \equiv s1'$, $s2 \equiv s2'$ and the two edges of either both forward or both backward.
- \end{definition}
-
- The def-use edges in the concrete schedule are as expected. There is a def use edge from statement $s1$ that defines $\x[\underline{i},\underline{j},\underline{k}]$
- to statement $s2$ that uses $\x[\underline{i},\underline{j},\underline{k}]$ if $s1$ is scheduled ahead of $s2$ in the linear schedule. We note that the underlined
- indices, e.g., $\underline{i}$, refer to fixed values, not iterative or vectorized dimensions since in the concrete schedule all induction variables are expanded. 
- E.g., there is a def-use edge from the statement that defines $\x[0,1,2]$ and a statement that uses $\x[0,1,2]$.
-
-\paragraph{Theorems}
-
-\begin{theorem} $a_0 \le a_1 \Rightarrow \gamma(a_0) \subseteq \gamma(a_1)$.
-\end{theorem}
-
-\begin{theorem} Let $a_0$ be the iterative MPC Source and let $a_1$ be the vectorized MPC Source computed by the vectorization algorithm. 
-We have that $a_0 \le a_1$. 
-\end{theorem}
-
-\ana{Write the proof sketch and final argument, etc.}
-
-\subsection{Extension of Basic Vectorization with Array Writes}
-
-\subsubsection{Removal of Infeasible Edges}
-
-Array writes limit vectorization as they sometimes introduce infeasible loop-carried dependencies. Consider the following example: \ana{Have to add citation to Aiken's paper}
-
-{\small
-\begin{verbatim}
-for i in range(N):
-  A[i] = B[i] + 10;
-  B[i] = A[i] * D[i-1];
-  C[i] = A[i] * D[i-1];
-  D[i] = B[i] * C[i];
-\end{verbatim}
-}
-
-In Cytron's SSA this code (roughly) translates into
-
-{\small
-\begin{verbatim}
-for i in range(N):
-1. A_1 = PHI(A_0,A_2)
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. A_2 = update(A_1, i, B_1[i] + 10); 
-6. B_2 = update(B_1, i, A_2[i] * D_1[i-1]);
-7. C_2 = update(C_1, i, A_2[i] * D_1[i-1]);
-8. D_2 = update(D_1, i, B_2[i] * C_2[i]);
-\end{verbatim}
-}
-
-\begin{comment}
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE \hspace{0.25cm} A[i] = B[i] + 10;
-\STATE \hspace{0.25cm} B[i] = A[i] * D[i-1];
-\STATE \hspace{0.25cm} C[i] = A[i] * D[i-1];
-\STATE \hspace{0.25cm} D[i] = B[i] * C[i];
-\end{algorithmic}
-
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{A}_0$ = $\phi(\texttt{A},\texttt{A}_1)$
-\STATE 2. $\texttt{B}_0$ = $\phi(\texttt{B},\texttt{B}_1)$
-\STATE 3. $\texttt{C}_0$ = $\phi(\texttt{C},\texttt{C}_1)$
-\STATE 4. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$
-\STATE 5. $\texttt{A}_1$ = $\texttt{A}_0$; $\texttt{A}_1$[i] = $\texttt{B}_0$[i] + 10; \COMMENT{equiv. to Update}
-\STATE 6. $\texttt{B}_1$ = $\texttt{B}_0$; $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 7. $\texttt{C}_1$ = $\texttt{C}_0$; $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 8. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
-\end{algorithmic}
-\end{comment}
-
-There is a cycle around \texttt{B\_1 = PHI(B\_0,B\_2)} that includes statement 
-\texttt{A\_1 = update(A\_0, i, B\_1[i] + 10)} and that statement won't be vectorized even 
-though in fact there is no loop-carried dependency from the write of \texttt{B\_1[i]} at 6 to the
-read of \texttt{ ... = B\_1[i]} at 8.
-
-The following algorithm removes certain infeasible loop-carried dependencies that are due to array writes. Consider a loop with index $0 \le j < J$
-nested at $i,j,k$. Here $i$ represents the enclosing loops of $j$ and $k$ represents the enclosed loops in $j$.
-
-\begin{algorithmic}
-\FOR {each array $\texttt{A}$ written in loop $j$}
-\STATE \COMMENT { including enclosed loops in $j$ }
-\STATE dep = False
-\FOR {each pair def: $\texttt{A}_m[f(i,j,k)] = ... $, and use: $ ... = \texttt{A}_n[f'(i,j,k)]$ in loop $j$}
-\IF {$\exists \underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$, s.t.
-$0\!
-\le\!
-\underline{i}\!
-<\!
-I,\: 0\!
-\le\!
-\underline{j}, \underline{j}' \!
-<\!
-J,\: 0\!
-\le\!
-\underline{k}, \underline{k}' \!
-<\!
-K,\: \underline{j}<\underline{j}'$, and $f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$}
-\STATE dep = True
-\ENDIF
-\ENDFOR
-\IF {dep == False}
-\STATE remove back edge into $\texttt{A}$'s $\phi$-node in loop $j$.
-\ENDIF
-\ENDFOR
-\end{algorithmic}
-
-Consider a loop $j$ enclosed in some fixed $\underline{i}$. Only if an update (definition) $\texttt{A}_m[f(i,j,k)] = ... $ at some iteration $\underline{j}$
-references the \emph{same} array element as a use $ ... = \texttt{A}_n[f'(i,j,k)]$ at some later iteration $\underline{j}'$,
-we may have a loop-carried dependence for $\texttt{A}$ due to this def-use pair. (In contrast, Cytron's algorithm inserts a loop-carried dependency every time there is an array update.)
-The algorithm above examines all def-use pairs in loop $j$, including defs and uses in nested loops, searching for values $\underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$ that satisfy
-$f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$. If such values exist for some def-use pair, then there is a potential
-loop-carried dependence on $\texttt{A}$; otherwise there is not and we can remove the spurious backward edge thus ``freeing up'' statements for vectorization. %\ana{Double check this!!!}
-
-Consider the earlier example.
-There is a single loop, $i$.
-Clearly, there is no pair $\underline{i}$ and $\underline{i}'$, where $\underline{i} < \underline{i}'$ that make $\underline{i} = \underline{i}'$
-due to the def-use pairs of \texttt{A} 5-6 and 5-7.
-Therefore, we remove the back edge from 5 to the phi-node 1.
-Analogously, we remove the back edges from 6 to 2 and from 7 to 3.
-However, there are many values $\underline{i} < \underline{i'}$ that make $\underline{i} = \underline{i'}-1$ and the back edge from 8 to 4 remains (def-use pairs for \texttt{D}).
-As a result of removing these spurious edges, Basic Vectorization will find that statement 5 is vectorizable.
-Statements 6, 7 and 8 will correctly appear in the FOR loop.
-
-Note however, that this step renders some array phi-nodes target-less. We handle target-less phi-nodes with a minor extension of Basic Vectorization. 
-First, we merge closures that update the same array. This simplifies handling of array $\phi$-nodes: if each closure is turned into a separate loop
-each loop will need to have its own array phi-node to account for the update and this would complicate the analysis.
-Second, we add the target-less node of array \texttt{A} back to the closure that updates \texttt{A} ---
-the intuition is, even if there is no loop-carried dependence from writes to reads on \texttt{A}, \texttt{A} is written and the write (i.e., update) cannot be vectorized; 
-therefore, the updated array has to carry to the next iteration of the loop. Third, in cases when the phi-node remains target-less, i.e., cases when the
-array write can be vectorized, we have to properly remove the phi-node replacing uses of the left-hand side of the phi-node with its arguments.
-
-%\ana{Commented out previous writeup. Removal/handling of targetless phi-nodes will go into the Extension to Basic vectorization.}
-
-\begin{comment}
-The final step is to remove array phi-nodes that are left without incoming back edges after removal of infeasible edges.
-Such a phi-node is redundant when there are no loop-carried dependences that are due to writes to the corresponding array.
-In our running example, the phi-nodes for \texttt{A}, \texttt{B}, and \texttt{C}, are redundant.
-
-Removing redundant phi-nodes proceeds from \emph{innermost to outermost} loops as follows.
-Denote a redundant phi-node by $\texttt{A}_1 = \phi(\texttt{A}_0,\texttt{A}_k)$. $\texttt{A}_0$ is the array at the beginning of the loop
-and $\texttt{A}_k$ is the array at the end of an iteration (including the last iteration). We consider two cases:
-\begin{enumerate}
-\item[Case 1] Inner-to-outer or same-level def-use edge
-\begin{itemize}
-\item If use is an array read $ ... = \texttt{A}_1[f(i)]$, then replace $\texttt{A}_1$ with $\texttt{A}_0$. This is correct, i.e., it is guaranteed that the use
-references $\texttt{A}_0$; had the use been referencing an array modified in an earlier iteration, the back-edge to the phi-node would
-not have been removed and the phi-node would not have been classified as redundant.
-\item If use is an array write (update) $\texttt{A}_2 = \texttt{A}_1; \texttt{A}_2[i] = ... $ equiv. to $\texttt{A}_2 = \mathit{update}(\texttt{A}_1, i, ...)$ then handling becomes more tricky.
-The semantics of this update is that when $i$ is 0, then $\texttt{A}_0$ is being read.
-But when $i > 0$, then $\texttt{A}_k$ is being read.
-If we are at this case, this means that the def-use edge is same-level and we are removing a redundant node in the loop enclosing the update.
-The removal of the original phi-node entails pushing the phi-node at the update:
-$\texttt{A}_2 = \phi(\texttt{A}_0,\texttt{A}_k); \texttt{A}_2[i] = ... $ equiv. to $\texttt{A}_2 = \mathit{update}(\phi(\texttt{A}_0,\texttt{A}_k), i, ...)$.
-\item If use is a phi-node $... = \phi(\texttt{A}_1,...)$, that means we have an outer-to-inner edge. The
-
-\end{itemize}
-\end{enumerate}
-For each outer-to-inner or same-level def-use edge
-that originates at the node, replace $\texttt{A}_1$ at the use node by $\texttt{A}_0$.
-For each inner-to-outer def-use edge, replace $\texttt{A}_1$ by $\texttt{A}_k$. In this case, the use references the array
-that results from the computation in the loop.
-
-Our running example is transformed as follows. Removing the phi-nodes for \texttt{A}, \texttt{B}, and \texttt{C} transforms the
-replaces the references to $\texttt{A}_0$, $\texttt{B}_0$, and $\texttt{C}_0$, with references to $\texttt{A}$, $\texttt{B}$, and $\texttt{C}$
-respectively, i.e., the arrays before the loop.
-
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$
-\STATE 2. $\texttt{A}_1$ = $\texttt{A}$; $\texttt{A}_1$[i] = $\texttt{B}$[i] + 10;
-\STATE 3. $\texttt{B}_1$ = $\texttt{B}$; $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 4. $\texttt{C}_1$ = $\texttt{C}$; $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
-\end{algorithmic}
-\end{comment}
-
-\subsubsection{Array MUX refinement}
-
-\ana{TODO: WIP, get to BV without MUX refinement for now? For now, skip MUX refinement.}
-
-Next, the algorithm refines array MUX statements.
-MPC-source after Cytron's SSA may result in statements $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$, which imply that any index
-of A can be written at this point and therefore there is a loop-carried dependency. In some cases the MUX can be refined to just a single
-index or a pair of indices, e.g., $\texttt{A}_j[i] = \emph{MUX}(c, \texttt{A}_k[i], \texttt{A}_l[i])$.
-
-%\ana{Clarify why this helps. Still work in progress.}
-This is to reduce the dimensionality of simd-ified computation. Technically, $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$ is
-a simdified operation that can be carried out in parallel "in one round". However, particularly when \texttt{A} is a multi-dimensional array, there is
-substantial increase in the size of the arrays (vectors) we send to SIMD operations. Refining to an update to a specific index would reduce
-the size of those vectors. Note that this is a heuristic that handles a common case, but not all cases of array updates.
-%\ana{TODO: Simplify this algorithm, taking into account the restriction to canonical updates. It should be handling \_all\_ cases.}
-
-\begin{algorithmic}
-\FOR {each $\mathit{stmt}$: $\texttt{A}_j = \emph{MUX}(c, \texttt{A}_k, \texttt{A}_l)$ in the MPC-source seq.}
-\STATE $i_1 = \mathit{find\_update}(A_k)$ \COMMENT { Is null when $A_k = \phi(...)$ }
-\STATE $i_2 = \mathit{find\_update}(A_l)$ \COMMENT { Is null when $A_l = \phi(...)$ }
-\IF {$i_1 == i_2$ or $i_1$ is null or $i_2$ is null} \STATE \COMMENT{ With our restrictions on writes we must have $i_1 = i_2$. }
-\STATE replace $\mathit{stmt}$ with \\
-\STATE $\texttt{A}_j = \texttt{A}_{j-1}$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
-%\ELSIF {$i_1 \ncong i_2$}
-%\STATE replace $\mathit{stmt}$ with \\
-%\STATE $\texttt{A}_j = \texttt{A}_k$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
-\ELSE
-\STATE $\mathit{stmt}$ stays as is
-\ENDIF
-\ENDFOR
-\end{algorithmic}
-
-\subsubsection{Restricting Array Writes}
-
-For now, we restrict array updates to \emph{canonical updates}. Assume (for simplicity) a two-dimensional array \texttt{A[I,J]}. A canonical update is the following:
-\begin{verbatim}
-for i in range(I):
-  for j in range(J):
-    ...
-    A[i,j] = ...
-    ...
-\end{verbatim}
-
-The update \texttt{A[i,j]} can be nested into an inner loop and there may be multiple updates, i.e., writes to \texttt{A[i,j]}.
-However, update such as \texttt{A[i-1,j] = ...} or \texttt{A[i-1,j-1] = ...}, etc., is not allowed.
-Additionally, while there could be several different loops that perform canonical updates, they must be of the same dimensionality, i.e., an update of higher or lower dimension, e.g., \texttt{A[i,j,k] = ...} is not allowed.
-We compute the \emph{canonical dimensionality} of each write array by examining the array writes in the original program and rejecting programs that violate the canonical write restriction.
-This restriction simplifies reasoning in this early stage of the compiler; we will look to relax the restriction in future work.
-
-Another restriction/assumption is that we assume the output array is given as input with initial values, and it is of size consistent with its canonical dimensionality. 
-%\ana{This restriction will require that we update our benchmarks that come with output arrays.}
-
-Reads through an arbitrary formula, such as \texttt{A[i-1]} for example, are allowed; currently, the projection function returns dummy values if the read formula is out of bounds; we assume the programmer ensures that the program still computes correct output in this case.
-
-\subsubsection{Changes to Basic Vectorization}
-
-\begin{comment}
-There are two changes to Basic Vectorization to account for write arrays. We can account for them during processing of raise dimension/drop dimension in Phase 1 of Basic Vectorization.
-
-One change to Basic vectorization is the expansion of dimension if the array write or read occurs in a nested loop.
-That is, if there is an update $\texttt{A}[i,j] = ...$ that occurs in loop dimensionality $i,j,k$, $\texttt{A}[i,j]$ will be rewritten into $\texttt{A}[i,j,k]$.
-Similarly, a read $\texttt{A}[f(i,j)]$ will be rewritten into $\texttt{A}[f(i,j),k]$.
-($\texttt{A}[f(i,j),k]$ will be reshaped as $f(i,j)*I*J+k$ during Phase 1 of Basic Vectorization.
-\ana{TODO: Double check.})
-\end{comment}
-
-In addition to the changes for the handling of target-less phi-nodes, Basic Vectorization has to handle differently def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable.
-The definition can be an update \texttt{A\_2 = update(A\_1,i,...)} or a pseudo $\phi$-node \texttt{A\_2 = PHI(A\_0,A\_1)}, etc..
-Note that $\phi$-nodes for arrays have no subscript operations the way there are subscript operations in analysis-introduced arrays representing scalars.
-Raising and dropping dimension still apply, however, the dimension cannot be raised or dropped to a dimension lower than the canonical one:
-
-\begin{enumerate}
-
-\item same-level $X \rightarrow Y$. We do nothing, just propagate the array, which happens to be of the right dimension. 
-%\ana{There might be some opportunities to do copy propagation optimization and save some cycles, but let's leave this for later.}
-
-\item inner-to-outer $X \rightarrow Y$.
-If dimensionality of the loop enclosure of $X$ is greater than the canonical dimensionality of the array, then add \emph{drop\_dim(...)} at $Y$, as in Basic Vectorization.
-Otherwise, do nothing.
-
-\item outer-to-inner $X \rightarrow Y$.
-If dimensionality of loop enclosure of $Y$ is greater than the canonical dimensionality of the array, then add \emph{raise\_dim(...)} (at $X$) as in Basic Vectorization.
-Otherwise, do nothing.
-
-\item "mixed" $X \rightarrow Y$. We assume that the mixed edge is transformed into an inner-to-outer followed by outer-to-inner edge before we perform vectorization, just as with Basic vectorization.
-
-\end{enumerate}
-
-\subsubsection{Examples with Array Writes}
-
-\paragraph{Example 1}
-
-First, the canonical dimensionality of all \texttt{A,B,C} and \texttt{D} is 1. 
-After Phase 1 of Basic Vectorization the Aiken's array write example will be (roughly) as follows:
-
-{\small
-\begin{verbatim}
-for i in range(N):
-1. A_1 = PHI(A_0,A_2)
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. A_2 = update(A_1, I, B_1[I] + 10); 
-6. B_2 = update(B_1, I, A_2[I] * D_1[I-1]);
-7. C_2 = update(C_1, I, A_2[I] * D_1[I-1]);
-8. D_2 = update(D_1, I, B_2[I] * C_2[I]);
-\end{verbatim}
-}
-
-Note that since all def-uses are same-level (i.e., reads and writes of the array elements) 
-no raise dimension or drop dimension happens.
-
-\begin{comment}
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{A}_1$ = $\phi(\texttt{A}_0,\texttt{A}_2)$
-\STATE 2. $\texttt{B}_1$ = $\phi(\texttt{B}_0,\texttt{B}_2)$
-\STATE 3. $\texttt{C}_1$ = $\phi(\texttt{C}_0,\texttt{C}_2)$
-\STATE 4. $\texttt{D}_1$ = $\phi(\texttt{D}_0,\texttt{D}_2)$
-\STATE 5. $\texttt{A}_2 = update(\texttt{A}_1, \stackrel{\rightarrow}{i}, \texttt{B}_1[\stackrel{\rightarrow}{i}] + 10)$;
-\STATE 6. $\texttt{B}_2 = update(\texttt{B}_1, \stackrel{\rightarrow}{i}, \texttt{A}_2[\stackrel{\rightarrow}{i}] * \texttt{D}_1[\stackrel{\rightarrow}{i-1}])$;
-\STATE 7. $\texttt{C}_2 = update(\texttt{C}_1, \stackrel{\rightarrow}{i}, \texttt{A}_2[\stackrel{\rightarrow}{i}] * \texttt{D}_1[\stackrel{\rightarrow}{i-1}])$;
-\STATE 8. $\texttt{D}_2 = update(\texttt{D}_1; \stackrel{\rightarrow}{i}, \texttt{B}_2[\stackrel{\rightarrow}{i}] * \texttt{C}_2[\stackrel{\rightarrow}{i}])$;
-\end{algorithmic}
-\end{comment}
-
-Phase 2 computes the closure of 4; $cl = \{4,6,7, 8\}$ 
-while 5 is vectorizable. Recall that 1,2, and 3 are target-less phi-nodes. Since the closure $cl$ includes updates to \texttt{B} and \texttt{C},
-the corresponding phi-nodes are added back to the closure and the def-use edges are added back to the target-less nodes. The uses of \texttt{A\_1} and \texttt{B\_1} in the vectorized statement
-turn into uses of \texttt{A\_0} and \texttt{B\_0} respectively; this is done for all original target-less phi-node.
-(But note that \texttt{A\_0} is irrelevant; the update writes into array \texttt{A\_2} in parallel.)
-Finally, the target-less phi-node for \texttt{A} is discarded.
-
-{\small
-\begin{verbatim}
-1. A_2 = update(A_0, I, ADD_SIMD(B_0[I],10));
-   equiv. to A_2[I] = ADD_SIMD(B_0[I],10)
-FOR i=0; i<N; i++; // MOTION loop
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. B_2 = update(B_1, i, A_2[i] * D_1[i-1]);
-   equiv. to B_2 = B_1; B_2[i] = A_2[i] * D_1[i-1];
-6. C_2 = update(C_1, i, A_2[i] * D_1[i-1]);
-7. D_2 = update(D_1, i, B_2[i] * C_2[i]);
-\end{verbatim}
-}
-
-\begin{comment}
-\begin{algorithmic}
-\STATE 1. $\texttt{A}_2 = update(\texttt{A}_0; \stackrel{\rightarrow}{i}, \mathit{ADD\_SIMD}(\texttt{B}_0[\stackrel{\rightarrow}i],[10,...])$ \COMMENT{Fully vectorized, size N.}
-\STATE FOR i=0; i<N; i++; \COMMENT{ MOTION loop }
-\STATE 2. $\texttt{B}_1$ = $\phi(\texttt{B}_0,\texttt{B}_2)$
-\STATE 3. $\texttt{C}_1$ = $\phi(\texttt{C}_0,\texttt{C}_2)$
-\STATE 4. $\texttt{D}_1$ = $\phi(\texttt{D}_0,\texttt{D}_2)$
-\STATE 5. $\texttt{B}_2 = update(\texttt{B}_1, i, \mathit{MUL}(\texttt{A}_2[i], \texttt{D}_1[{i-1}])$
-\STATE 6. $\texttt{C}_2 = update(\texttt{C}_1, i, \mathit{MUL}(\texttt{A}_2[i], \texttt{D}_1[{i-1}])$
-\STATE 7. $\texttt{D}_2 = update(\texttt{D}_1, i, \mathit{MUL}(\texttt{B}_2[i], \texttt{C}_1[i])$
-\end{algorithmic}
-\end{comment}
-
-\paragraph{Example 2} Now consider the MPC Source of Histogram:
-
-{\small
-\begin{verbatim}
-for i in range(0, num_bins):
-   res1 = PHI(res, res2)
-   for j in range(0, N):
-      res2 = PHI(res1, res3)
-      tmp1 = (A[j] == i)
-      tmp2 = (res2[i] + B[j])
-      tmp3 = MUX(tmp1, res2[i], tmp2)
-      res3 = Update(res2, i, tmp3)     
-return res1
-\end{verbatim}
-}
-
-The canonical dimensionality of \texttt{res} is 1.
-Also, the phi-node \texttt{res1 = PHI(res, res2)} is a target-less phi-node (the implication being that the inner for loop can be vectorized across $i$).
-After Phase 1, Basic vectorization produces the following code (statements are implicitly vectorized along $i$ and $j$).
-In a vectorized update statement, we can ignore the incoming array, \texttt{res2} in this case.
-The update writes (in parallel) all locations of the 2-dimensional array, in this case it sets up each \texttt{res3[i,j] = tmp3[i,j]}.
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I = raise_dim(i, ((i:num_bins),(j:N)))
-for i in range(0, num_bins):
-   res1 = PHI(res, res2) // target-less, no drop_dim
-   res1' = raise_dim(res1, (j:N))
-   for j in range(0, N):
-      res2 = PHI(res1', res3)
-      tmp1 = (A1 == I)
-      tmp2 = (res2 + B1)
-      tmp3 = MUX(tmp1, res2, tmp2)
-      res3 = Update(res2, (I,J), tmp3) 
-return res1
-\end{verbatim}
-}
-
-Processing the inner loop in Phase 2 vectorizes \texttt{tmp1 = (A1 == I)} along the $j$ dimension but leaves the rest of the statements in a MOTION loop.
-Processing the outer loop is interesting.
-This is because the PHI node is a target-less node, and therefore, there are no closures!
-Thus, everything can be vectorized along the $i$ dimension:
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I1 = raise_dim(i, ((i:num_bins),(j:N)))
-
-tmp1[I,J] = (A1[I,J] == I1[I,J])
-
-res1' = raise_dim(res, (j:N)) // raising res to a 2-dim. array
-for j in range(0, N):
-   res2 = PHI(res1', res3)
-   tmp2[I,j] = (res2[I,j] + B1[I,j])
-   tmp3[I,j] = MUX(tmp1[I,j], res2[I,j], tmp2[I,j])
-   res3 = Update(res2, (I,j), tmp3)
-res1 = drop_dim(res2) // Add drop_dim before removing redundant phi-node!      
-return res1
-\end{verbatim}
-}
-
-The cleanup phase can then get rid of $j$ in the \texttt{res} variables and \texttt{tmp} variables, so the loop operates over the \texttt{num\_bins}-size vectors. 
-This will lead to dropping raise\_dim and drop\_dim as well:
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I1 = raise_dim(i, ((i:num_bins),(j:N)))
-
-tmp1[I,J] = (A1[I,J] == I1[I,J])
-
-for j in range(0, N):
-   res2 = PHI(res1, res3) // now 1-dim array
-   tmp2[I] = ADD_SIMD(res2[I],B1[I,j])
-   tmp3[I] = MUX_SIMD(tmp1[I,j], res2[I], tmp2[I])
-   res3 = Update(res2, I, tmp3[I])
-return res2
-\end{verbatim}
-}
-
-\begin{comment}
-\paragraph{Example 2}
-
-Consider MPC-source of Histogram after removal of infeasible edges and redundant phi-nodes, and MUX refinement:
-
-\begin{algorithmic}
-
-\STATE res!1 = []
-\STATE for i: plaintext in range(0, num\_bins):
-\STATE \hspace{0.25cm} res!2 = $\phi$(res!1, res!3)
-\STATE  \hspace{0.25cm} ...
-\STATE res!4 = res!2 \COMMENT{Added due to the "mixed" def-use edge.}
-\STATE for i: plaintext in range(0, num\_bins):
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE \hspace{0.25cm}        for j: plaintext in range(0, N):
-\STATE \hspace{0.5cm}            res!5 = $\phi$(res!4, res!7)
-\STATE \hspace{0.5cm}            !2!3 = (A[j] == i)
-\STATE \hspace{0.5cm}            !3!3 = (res!5[i] + B[j])
-\STATE \hspace{0.5cm}            res!6 = res!5; res!6[i] = !3!3
-\STATE \hspace{0.5cm}            res!7 = res!6; res!7[i] = MUX(!2!3, res!6[i], res!5[i])
-
-\end{algorithmic}
-
-After Phase 1 of Basic Vectorization:
-
-\begin{algorithmic}
-\STATE res!1 = []
-\STATE for i: plaintext in range(0, num\_bins):
-\STATE \hspace{0.25cm} res!2 = $\phi$(res!1, res!3)
-\STATE  \hspace{0.25cm} ...
-\STATE res!4 = res!2 \COMMENT{Added due to the "mixed" def-use edge. Because the canonical dimensionality of res is 1, we DO NOT drop dimension of res!2.}
-\STATE A[$i,j$] = $\mathit{raise\_dim}$(A[$j$],$i$) \COMMENT{Copying row A num\_bin times. Reusing name.}
-\STATE B[$i,j$] = $\mathit{raise\_dim}$(B[$i$],$j$) \COMMENT{Copying column B N times. Reusing name.}
-\STATE res!4[$i,j$] = $\mathit{raise\_dim}$(res!4[$i$],$j$) \COMMENT{Reusing name.}
-\STATE for i: plaintext in range(0, num\_bins):
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE \hspace{0.25cm}        for j: plaintext in range(0, N):
-\STATE \hspace{0.5cm}            res!5 = $\phi$(res!4, res!7) \COMMENT{Do nothing. res!4 has been raised to proper dimension.}
-\STATE \hspace{0.5cm}            !2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{EQ}$(A[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], [[0,0...], [1,1...],...[num\_bins-1,num\_bins-1...]])
-\STATE \hspace{0.5cm}            !3!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{ADD}$(res!5[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$],B[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$])
-\STATE \hspace{0.5cm}            res!6 = res!5 \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.5cm}            res!6[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = !3!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$]
-\STATE \hspace{0.5cm}            res!7 = res!6
-\STATE \hspace{0.5cm}            res!7[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{MUX}$(!2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], res!6[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], res!5[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$])
-\end{algorithmic}
-
-After Phase 2 and Phase 3. The EQ operation is vectorizable across both dimensions, and the rest of the computation is vectorizable across the $i$-dimension.
-
-\begin{algorithmic}
-\STATE ...
-\STATE !2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{EQ\_SIMD}$(A[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], [[0,0...], [1,1...],...[num\_bins-1,num\_bins-1...]])
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE FOR j=0; j<N; j++; \COMMENT{MOTION loop}
-\STATE \hspace{0.25cm}            res!5 = $\phi$(res!4, res!7) \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.25cm}            !3!3[$\stackrel{\rightarrow}i,j$] = $\mathit{ADD\_SIMD}$(res!5[$\stackrel{\rightarrow}i,j$],B[$\stackrel{\rightarrow}i,j$])
-\STATE \hspace{0.25cm}            res!6 = res!5 \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.25cm}            res!6[$\stackrel{\rightarrow}i,j$] = !3!3[$\stackrel{\rightarrow}i,j$]
-\STATE \hspace{0.25cm}            res!7 = res!6
-\STATE \hspace{0.25cm}            res!7[$\stackrel{\rightarrow}i,j$] = $\mathit{MUX\_SIMD}$(!2!3[$\stackrel{\rightarrow}i,j$], res!6[$\stackrel{\rightarrow}i,j$], res!5[$\stackrel{\rightarrow}i,j$])
-\end{algorithmic}
-\end{comment}
-
-\section{Divide-and-Conquer}
-\label{sec:div_and_conq}
-
-\ana{TODO: Now that we have broken FOR loops into smaller chunks, we can add Divide-and-conquer reasoning with Z3 and implement this additional transform.}
-
-\section{Implementation and Evaluation}
-\label{sec:implementation}
-
-\section{Future Work}
-\label{sec:implementation_and_benchmarks}
-%\input{../sections/implementation_and_benchmarks.tex}
-%\input{../sections/evaluation.tex}
-
-\section{Conclusions}
-\label{sec:conclusion}
-%\input{../sections/conclusion.tex}
-
-%\begin{acks}
-%    \ishaq{TODO}
-%\end{acks}
-
-\printbibliography
-
-% that's all folks
-\end{document}

### 2022-03-09 paper_SIMD.tex
removed section 3.2 from main latex file

@@
 
-\subsection{Problem Statement}
-\label{sec:problem}
-
-\ana{Ishaq?
-Basically, define sequential schedule, then define an equivalent parallel schedule.
-A parallel schedule is equivalent if it preserves def-use relations in sequential schedule, or in other words, schedules def ahead of the use.
-Problem is to minimize cost of Parallel schedule.}
-
-\ana{Commenting out this section to have paper compile on Github. Will revert!}
-%\begin{comment}
-
-At the lowest level, we have two types of MPC instructions 1) local/non-interactive instruction (i.e.
-\add) and 2) remote/interactive instruction (i.e.
-\mul).
-Let the cost of a single \add instruction be given by a monotonically decreasing $f(n)$, where the argument $n$ is the number of \add instructions being executed in parallel.
-Similarly the cost of a single \mul is given by a monotonically decreasing function $g(n)$.
-
-Given a serial schedule (a linear graph) of an MPC program i.e.
-a sequence of instructions $\mathcal{S}:=(S_1, \dots, S_n)$, where $S_i \in \{\text{\add, \mul}\}, 1 \leq i \leq n$, and a def-use dependency graph $G(V, E)$ corresponding to $\mathcal{S}$, our task is to construct a parallel schedule (another linear graph) $\mathcal{P} := (P_1, \dots, P_n)$ observing the following conditions:
-
-\begin{enumerate}
-    \item Multiple, not necessarily continuous instructions of the same type (i.e.
-either \add or \mul) from $\mathcal{S}$ can be grouped into a single $P_i$  However, all such instructions must be of the same type (either \add or \mul).
-    \item Def-use dependencies of the graph $G(V, E)$ are should be preserved i.e.
-if instructions $S_i, S_j, i < j$ are a def-use (an edge exists from $S_i$ to $S_j$ in $G$), then they can only be mapped to $P_{i'}, P_{j'}, i' < j'$.
-\end{enumerate}
-
-Our goal is to construct minimize the height of the graph $\mathcal{P}$. Indeed, a graph $\mathcal{P}$ with minimum height will maximize parallelization.
-
-\paragraph{Correctness} Correctness of $\mathcal{P}$ is guaranteed by definition. Since def-use dependencies are preserved, the function (being computed) remains the same.
-
-\paragraph{Cost Comparison} For the sequential schedule $\mathcal{S}$ consisting of $L$ local and $R$ remote instructions, the total cost is $\mathit{cost}(\mathcal{S}) = L \cdot f(1) + R \cdot g(1)$.
-In the extreme case where all $L$ and all $R$ instructions can be parallelized, the cost of $\mathcal{P}$ is $\mathit{cost}(\mathcal{P}) = L \cdot f(L) + R \cdot g(R)$.
-Since both $f$ and $g$ are monotonically decreasing, $\mathit{cost}(\mathcal{P}) < \mathit{cost}(\mathcal{S})$.
-Cost of all other parallel schedules lies between the extremes of $\mathit{cost}(\mathcal{S})$ and $\mathit{cost}(\mathcal{P})$.
-
-Note that we use an MPC-Source control flow graph (CFG) $G'(V', E')$ along with def-use graph $G(V,E)$ to construct $\mathcal{P}$.
-We consider a linearized MPC schedule $\mathcal{S}$ above for ease of exposition.
-The argument becomes slightly more involved when dealing with a graph $G'$ that may contain cycles.
-
-% Given a def-use dependency graph $G(V, E)$ that is constructed from for the input program.
-Each vertex/node in this graph represents an operation/instruction in the program and an edge represents dependence i.e.
-an edge from $V_1$ to $V_2$ means the result of operation in $V_1$ is used in the operation in $V_2$.
-To build a parallel schedule, we translate this graph $G(V, E)$ into another dependency graph $G'(V', E')$ such that:
-
-% \begin{enumerate}
-%     \item We collapse multiple vertices of the same type (i.e. same operation) in $V$ into a single vertex in $V'$.
-%     \item dependencies are not broken i.e.
-If there is an edge from $V_1$ to $V_2$ ($V_1, V_2 \in V$), and we put them in vertices $V'_1, V'_2 \in G'$ respectively, then an edge must exist from $V'_1$ to $V'_2$.
-% \end{enumerate}
-
-% We want to build a graph $G'$ with fewest possible vertices since fewer vertices means more parallelization.
-
-%\ishaq{TODO: @Ana: I am still confused about what prevents a reviewer from saying, "you argue correctness of the schedule from unrolled MPC, you don't argue it for the pre-unrolled CFG that you use".
-Please comment.} \ana{The correctness proof answer this question.
-The unrolled schedule is the \emph{concretization} of MPC source.
-We define (1) the obvious $\gamma$, (2) we define partial order over MPC source codes (this is our abstract domain $A$), and (3) a partial order over unrolled schedules.
-Then we prove two theorems that $A \le A' \Rightarrow \gamma(A) \subseteq \gamma(A')$ and that the MPC source $\le$ Vectorized MPC source.}
-
-%\end{comment}
-
-\subsection{Scheduling is NP-hard}
-\label{sec:np}
-
-\ana{Commented out this section to have paper compile on Github. Will revert!}
-
-\begin{comment}
-\ishaq{The follow text is all from Linsay's thesis}.
-\brevised{
-We consider two operations, call them $A$ and $M$. $A$ and $M$ are two abstract MPC instruction, 
-but as an example, $A$ stands for the ADD MPC instruction 
-and $M$ stands for the MUL instruction. Each instruction in the program is either an $A$-instruction or
-an $M$-instruction. In order to benefit from parallelization/amortization, we must schedule two or more 
-$A$-instructions in the same parallel node (or two or more $M$-instructions in the same parallel node). 
-Scheduling $A$-instructions in parallel with $M$-instruction does not benefit from amortization.
-It incurs the exact same cost as scheduling the $A$-instructions in a node $P_A$, scheduling the $M$-instructions 
-in a node $P_M$, and having $P_A$ precede $P_M$ in the parallel schedule. This is the difference 
-between classical scheduling, as studied in parallel computing, and MPC scheduling.
-
-In our theoretical treatment, we make the following assumptions:
-
-\begin{enumerate}
-
-\item $A$ and $M$ are of equal cost, 1 unit.
-\item There is unlimited bandwidth---i.e., a single $A$-instruction (or $M$-instruction) costs as much as $N$ amortized $A$-instructions 
-(or $M$-instructions), namely 1 unit.
-
-\end{enumerate}
-
-%  Furthermore the run time cost of 
-% computing operations of the same type remains the same regardless of the number of operations that are run.
-% Roughly speaking, 1 A operation costs the same as 100 A operations and 1000 A operations.
-
-Consider a loop body that consists of $n$ sequences: $S_1$, ... $S_n$ of $A$ and $M$ instructions. 
-More precisely, the loop body is such that its instructions can be grouped into such sequences. 
-$S_1$, ... $S_n$ can execute in parallel, however, all instructions within a sequence must 
-execute sequentially. For example, consider the three sequences (the right arrow indicates a \emph{dependence},
-meaning that the source node must execute before the target node): 
-\begin{enumerate}
-\item $A \rightarrow M \rightarrow A$
-\item $A \rightarrow A \rightarrow A$
-\item $M \rightarrow A \rightarrow M$
-\end{enumerate}
-
-A \emph{schedule} $P: P_1 \rightarrow P_2 \dots \rightarrow P_k$ is such that for each sequence 
-$S_i$ in the set, if $S_i[k]$ precedes $S_i[k']$ in $S_i$ then $S_i[k]$ is scheduled in node $P_l$, $S_i[k]$ 
-is scheduled in node $P_{l'}$, and $P_l$ precedes $P_{l'}$ in $P$.
-
-The cost of a schedule $P$ is
-
-\begin{equation}
-\mathit{cost}(P) = \sum_{i=1}^k \mathit{cost}(P_i)
-\end{equation}
-
-where $\mathit{cost}(P_i) = 1$ if $P_i$ consists of $A$-instructions only, or $M$-instructions only, 
-and $\mathit{cost}(P_i) = 2$ if $P_i$ mixes $A$-instructions and $M$-instructions.
-
-The problem is to find a schedule $P$ with \emph{minimal cost}. For example, 
-a schedule with minimal cost for the sequences above is
-\[ A(1), A(2) \rightarrow M(1), A(2), M(3) \rightarrow A(1), A(2), A(3) \rightarrow M(3) \]
-(The parentheses above indicate the sequence where the instruction comes from: (1), (2), or (3).)
-The cost of this schedule is 5.
-
-The problem of finding a schedule $P$ with a minimal $cost(P)$ for a given loop body has been shown
-to be an NP-Hard problem, as it can be reduced to the problem of finding a \emph{shortest common supersequence}, 
-a known NP-Hard problem\cite{Maier1978},\cite{Vazirani2010}. The shortest common supersequence problem is as follows: 
-given two or more sequences find the the shortest sequence that contains all of the original sequences. This can be solved
-in $O(n^k)$ time, where $n$ is the cardinality of the longest sequence and $k$ is the 
-number of sequences. For our problem $n$ is the maximum length of a node and $k$ is the number
-of total number of nodes. %\ana{I think it is clear without additional explanation...}
-%\ana{Also, add the full statement of the problem here. Now it's not clear what's NP-hard, what's n, what's k, etc.?}
-%\lindsey{not exactly sure what need to go it there, but does that make it clearer?}
-
-To see the reduction, 
-suppose $P$ is a schedule with minimal cost (computed by a black-box algorithm). 
-We can derive a schedule $P'$ with the same cost as $P$, by mapping each mixed node $P_i \in P$ 
-to two consecutive nodes in $P'$: an $A$-instruction node followed by an $M$-instruction node.
-Clearly, $P'$, which now is a sequence of $A$'s and $M$'s, is a supersequence of each sequence 
-$S_i$, i.e., $P'$ is a common supersequence 
-of $S_1 \dots S_n$. It is also a shortest common supersequence. (To see this, suppose, a 
-shorter common supersequence, $P''$, exists. $P''$ is a schedule of $S_1 \dots S_n$
-and the $\mathit{cost}(P'')$ equals the length of $P''$. Since $P'$ is longer than
-$P''$, and $\mathit{cost}(P') = \mathit{cost}(P)$, that means $\mathit{cost}(P'') < \mathit{cost}(P)$, 
-which is a contradiction since $P$ is a schedule with minimal cost.)
-}
-\end{comment}
+\input{prob_stmt_np_hardness.tex}
 

### 2022-03-14 paper_SIMD.tex
Minor local changes?
Add taint analysis to writeup

@@
 
-\ana{Ben?}
+We require that all inputs are marked as either shared or plaintext.
+We then determine if intermediate variables are shared through taint analysis with "taintedness" referring to the shared attribute.
+Specifically, our compiler follows the following rules:
+\begin{itemize}
+\item If any variable on the right-hand side of an assignment is shared, then the assigned variable is shared
+\item If all variables on the right-hand side of an assignment are plaintext, then the assigned variable is plaintext
+\item Loop counters are always plaintext
+\item Any variables which cannot be determined as shared or plaintext via the above rules are plaintext
+\end{itemize}
+
+The first two rules are standard for taint analysis, and the third rule follows from the MPC problem statement.
+\ben{Is the explanation for the third rule correct?}  The final rule is needed to handle cycles of plaintext values.
+For example, in the below snippet \texttt{sum!2} and \texttt{sum!3} form a dependency cycle and cannot be marked as plaintext through simple taint analysis:
+{\small
+\begin{verbatim}
+plaintext_array = [0, 1, 2, ...]
+sum!1 = 0
+for i in range(0, N):
+    sum!2 = PHI(sum!1, sum!3)
+    sum!3 = sum!2 + plaintext_array[i]
+\end{verbatim}
+}
+\ben{I think the above example is unnecessary and could be replaced by an explanation of how "untainted" variables are implicitly plaintext, but I couldn't think of a way to phrase that.}
+
+When converting to MOTION code, any plaintext value used in the right-hand side of a shared assignment is implicitly converted to a shared value for that expression.
+\ben{Is this necessary to include?}
 
@@
 \[\begin{array}{l}
-\gamma(\x[i,J,k] = \y[i,J,k] \; \mathit{op\_SIMD} \; \z[i,J,k]) = \\ 
+\gamma(\x[i,J,k] = \mathit{op\_SIMD}(\y[i,J,k], \z[i,J,k])) = \\ 
 \x[i,0,k] = \y[i,0,k] \; \mathit{op} \; \z[i,0,k] \;\; || \\

### 2022-03-15 prob_stmt_np_hardness.tex
first iteration of NP-Hardness proof

@@
 
-\ana{Commenting out this section to have paper compile on Github. Will revert!}
-%\begin{comment}
+\ishaq{TODO: make it consistent with the next section.}
 
@@
 
-\ana{Commented out this section to have paper compile on Github. Will revert!}
+\ishaq{TODO: need make amortized cost a function (like in the problem statement above), it is proving to be tricky.}
 
-\begin{comment}
-\ishaq{The follow text is all from Linsay's thesis}.
-\brevised{
-We consider two operations, call them $A$ and $M$. $A$ and $M$ are two abstract MPC instruction, 
-but as an example, $A$ stands for the ADD MPC instruction 
-and $M$ stands for the MUL instruction. Each instruction in the program is either an $A$-instruction or
-an $M$-instruction. In order to benefit from parallelization/amortization, we must schedule two or more 
-$A$-instructions in the same parallel node (or two or more $M$-instructions in the same parallel node). 
-Scheduling $A$-instructions in parallel with $M$-instruction does not benefit from amortization.
-It incurs the exact same cost as scheduling the $A$-instructions in a node $P_A$, scheduling the $M$-instructions 
-in a node $P_M$, and having $P_A$ precede $P_M$ in the parallel schedule. This is the difference 
-between classical scheduling, as studied in parallel computing, and MPC scheduling.
+We consider two operations, call them $A$ and $M$.
+$A$ and $M$ are two abstract MPC instruction, but as an example, $A$ stands for the ADD MPC instruction and $M$ stands for the MUL instruction.
+Each instruction in the program is either an $A$-instruction or an $M$-instruction.
+In order to benefit from parallelization/amortization, we must schedule two or more $A$-instructions in the same parallel node (or two or more $M$-instructions in the same parallel node).
+We also assume that scheduling $A$-instructions in parallel with $M$-instruction does not benefit from amortization\footnote{this is not strictly true, but assuming it, e.g.
+as in \cite{Ishaq2019, Demmler2015ABYA, Mohassel2018}, helps with the exposition.}.
+It incurs the exact same cost as scheduling the $A$-instructions in a node $P_A$, scheduling the $M$-instructions in a node $P_M$, and having $P_A$ precede $P_M$ in the parallel schedule.
 
-In our theoretical treatment, we make the following assumptions:
+We use the following cost model:
 
 \begin{enumerate}
-
-\item $A$ and $M$ are of equal cost, 1 unit.
-\item There is unlimited bandwidth---i.e., a single $A$-instruction (or $M$-instruction) costs as much as $N$ amortized $A$-instructions 
-(or $M$-instructions), namely 1 unit.
-
+    \item $A$ costs $\alpha$ units and $M$ costs $\beta$ unites.
+    \item There is unlimited bandwidth i.e.
+a single $A$-instruction (or $M$-instruction) costs as much as $N$ amortized $A$-instructions (or $M$-instructions), concretely either $\alpha$ unites or $\beta$ units.
 \end{enumerate}
 
-%  Furthermore the run time cost of 
-% computing operations of the same type remains the same regardless of the number of operations that are run.
-% Roughly speaking, 1 A operation costs the same as 100 A operations and 1000 A operations.
+Consider a loop body that consists of $n$ sequences: $S_1$, ...
+$S_n$ of $A$ and $M$ instructions.
+More precisely, the loop body is such that its instructions can be grouped into such sequences.
+$S_1$, ...
+$S_n$ can execute in parallel, however, all instructions within a sequence must execute sequentially.
+For example, consider the three sequences (the right arrow indicates a \emph{dependence}, meaning that the source node must execute before the target node):
 
-Consider a loop body that consists of $n$ sequences: $S_1$, ... $S_n$ of $A$ and $M$ instructions. 
-More precisely, the loop body is such that its instructions can be grouped into such sequences. 
-$S_1$, ... $S_n$ can execute in parallel, however, all instructions within a sequence must 
-execute sequentially. For example, consider the three sequences (the right arrow indicates a \emph{dependence},
-meaning that the source node must execute before the target node): 
 \begin{enumerate}
@@
 
-A \emph{schedule} $P: P_1 \rightarrow P_2 \dots \rightarrow P_k$ is such that for each sequence 
-$S_i$ in the set, if $S_i[k]$ precedes $S_i[k']$ in $S_i$ then $S_i[k]$ is scheduled in node $P_l$, $S_i[k]$ 
-is scheduled in node $P_{l'}$, and $P_l$ precedes $P_{l'}$ in $P$.
+A \emph{schedule} $P: P_1 \rightarrow P_2 \dots \rightarrow P_k$ is such that for each sequence $S_i$ in the set, if $S_i[j]$ precedes $S_i[j']$ in $S_i$ then $S_i[j]$ is scheduled in node $P_\ell$, $S_i[j']$ is scheduled in node $P_{\ell'}$, and $P_\ell$ precedes $P_{l\ell'}$ in $P$.
 
@@
 
-where $\mathit{cost}(P_i) = 1$ if $P_i$ consists of $A$-instructions only, or $M$-instructions only, 
-and $\mathit{cost}(P_i) = 2$ if $P_i$ mixes $A$-instructions and $M$-instructions.
+where $\mathit{cost}(P_i) = \alpha$ if $P_i$ consists of $A$-instructions only, $\beta$ if $P_i$ consists of $M$-instructions only, and $\mathit{cost}(P_i) = \alpha + \beta$ if $P_i$ mixes $A$-instructions and $M$-instructions.
 
-The problem is to find a schedule $P$ with \emph{minimal cost}. For example, 
-a schedule with minimal cost for the sequences above is
-\[ A(1), A(2) \rightarrow M(1), A(2), M(3) \rightarrow A(1), A(2), A(3) \rightarrow M(3) \]
-(The parentheses above indicate the sequence where the instruction comes from: (1), (2), or (3).)
-The cost of this schedule is 5.
+The problem is to find a schedule $P$ with \emph{minimal cost}.
+For example, a schedule with minimal cost for the sequences above is \[ A(1), A(2) \rightarrow M(1), A(2), M(3) \rightarrow A(1), A(2), A(3) \rightarrow M(3) \] (The parentheses above indicate the sequence where the instruction comes from: (1), (2), or (3).) The cost of this schedule is $3\alpha + 2\beta$.
 
-The problem of finding a schedule $P$ with a minimal $cost(P)$ for a given loop body has been shown
-to be an NP-Hard problem, as it can be reduced to the problem of finding a \emph{shortest common supersequence}, 
-a known NP-Hard problem\cite{Maier1978},\cite{Vazirani2010}. The shortest common supersequence problem is as follows: 
-given two or more sequences find the the shortest sequence that contains all of the original sequences. This can be solved
-in $O(n^k)$ time, where $n$ is the cardinality of the longest sequence and $k$ is the 
-number of sequences. For our problem $n$ is the maximum length of a node and $k$ is the number
-of total number of nodes. %\ana{I think it is clear without additional explanation...}
-%\ana{Also, add the full statement of the problem here. Now it's not clear what's NP-hard, what's n, what's k, etc.?}
-%\lindsey{not exactly sure what need to go it there, but does that make it clearer?}
+The problem of finding a schedule $P$ with a minimal $cost(P)$ for a given loop body has been shown to be an NP-Hard problem, as it can be reduced to the problem of finding a \emph{shortest common supersequence}, a known NP-Hard problem\cite{Maier1978},\cite{Vazirani2010}.
+The shortest common supersequence problem is as follows: {\it given two or more sequences find the the shortest sequence that contains all of the original sequences.} This can be solved in $O(n^k)$ time, where $n$ is the cardinality of the longest sequence and $k$ is the number of sequences.
+For our problem $n$ is the maximum length of a node and $k$ is the number of total number of nodes.
 
-To see the reduction, 
-suppose $P$ is a schedule with minimal cost (computed by a black-box algorithm). 
-We can derive a schedule $P'$ with the same cost as $P$, by mapping each mixed node $P_i \in P$ 
-to two consecutive nodes in $P'$: an $A$-instruction node followed by an $M$-instruction node.
-Clearly, $P'$, which now is a sequence of $A$'s and $M$'s, is a supersequence of each sequence 
-$S_i$, i.e., $P'$ is a common supersequence 
-of $S_1 \dots S_n$. It is also a shortest common supersequence. (To see this, suppose, a 
-shorter common supersequence, $P''$, exists. $P''$ is a schedule of $S_1 \dots S_n$
-and the $\mathit{cost}(P'')$ equals the length of $P''$. Since $P'$ is longer than
-$P''$, and $\mathit{cost}(P') = \mathit{cost}(P)$, that means $\mathit{cost}(P'') < \mathit{cost}(P)$, 
-which is a contradiction since $P$ is a schedule with minimal cost.)
-}
-\end{comment}
+To see the reduction, suppose $P$ is a schedule with minimal cost (computed by a black-box algorithm).
+We can derive a schedule $P'$ with the same cost as $P$, by mapping each mixed node $P_i \in P$ to two consecutive nodes in $P'$: an $A$-instruction node followed by an $M$-instruction node.
+Clearly, $P'$, which now is a sequence of $A$'s and $M$'s, is a supersequence of each sequence $S_i$, i.e., $P'$ is a common supersequence of $S_1 \dots S_n$.
+It is also a shortest common supersequence.
+To see this, let $X$ and $Y$ denote, respectively, the number of $A$ and $M$ nodes in $P'$.
+The cost of $P'$, and $P$, is $X \cdot \alpha + Y \cdot \beta$.
+Now suppose, there exists a shorter common supersequence, $P''$ that consists of $X'$ nodes of type $A$-instructions $Y'$ nodes of type $M$-instructions.
+Since $P''$ is shorter than $P'$, therefore $X' + Y' < X + Y$, and $X' \cdot \alpha + Y' \cdot \beta < X \cdot \alpha + Y \cdot \beta$ i.e.
+$\mathit{cost}(P'') < \mathit{cost}(P')$.
+But $\mathit{cost}(P') = \mathit{cost}(P)$ and $\mathit{cost}(P)$ is the optimal cost.
+Therefore $\mathit{cost}(P'') < \mathit{cost}(P')$ is contradiction and no such $P''$ exists.

### 2022-03-16 paper_SIMD.tex
Minor edits to paper?
Add initial semantic analysis description

@@
 
-\ana{Benjamin?}
+Our compiler performs the following semantic analysis steps:
+
+\begin{enumerate}
+  \item
+        \textbf{Parsing}:
+        Use Python's \texttt{ast} module to parse the input source code to a Python AST.
+  \item
+        \textbf{Syntax checking}:
+        Ensure that the AST matches a restricted subset that our compiler supports.
+        This step outputs an instance of the \texttt{restricted\_ast.Function} class, which represents our restricted subset of the Python AST.
+  \item
+        \textbf{3-address CFG conversion}: TODO
+  \item
+        \textbf{SSA conversion}:
+        Convert the 3-address CFG to SSA with Cytron's algorithm.
+\end{enumerate}
 
@@
 \gamma(\x[i,J,k] = \mathit{op\_SIMD}(\y[i,J,k], \z[i,J,k])) = \\ 
-\x[i,0,k] = \y[i,0,k] \; \mathit{op} \; \z[i,0,k] \;\; || \\
-\x[i,1,k] = \y[i,1,k] \; \mathit{op} \; \z[i,1,k] \;\; || \; ...\; || \\
-\x[i,I-1,k] = \y[i,I-1,k] \; \mathit{op} \; \z[i,I-1,k]
+~~\x[i,0,k] = \y[i,0,k] \; \mathit{op} \; \z[i,0,k] \;\; || \\
+~~~~\x[i,1,k] = \y[i,1,k] \; \mathit{op} \; \z[i,1,k] \;\; || \; ...\; || \\
+~~~~~~\x[i,I-1,k] = \y[i,I-1,k] \; \mathit{op} \; \z[i,I-1,k]
 \end{array}\]
@@
 The concretization of the FOR statement is as follows:
-\[ \gamma(\texttt{FOR} \; 0 \le i < I: s) = \gamma(s)[0/i] ; \gamma(s)[1/i] ; ... \gamma(s)[I-1/i] \]
+\[ \gamma(\texttt{FOR} \; 0 \le i < I: s) = \gamma(s)[0/i] \; ; \; \gamma(s)[1/i] \; ; \; ... \gamma(s)[I-1/i] \]
 $\gamma$ simply unrolls the loop substituting $i$ with 0, 1, etc. Here ; denotes sequential execution.

### 2022-03-16 compiler/compiler/vectorize.py
Remove always-false isinstance() checks

@@ -472,20 +472,12 @@ def _basic_vectorization_phase_1(
                 elif isinstance(stmt.rhs, llc.UnaryOp):
                     stmt.rhs.operand = replace_operand(stmt.rhs.operand, stmt)
                 elif isinstance(stmt.rhs, llc.List):
-                    replaced = [replace_atom(atom, stmt) for atom in stmt.rhs.items]
                     stmt.rhs = llc.List(
-                        [
-                            v.array if isinstance(v, VectorizedAccess) else v
-                            for v in replaced
-                        ]
+                        [replace_atom(atom, stmt) for atom in stmt.rhs.items]
                     )
                 elif isinstance(stmt.rhs, llc.Tuple):
-                    replaced = [replace_atom(atom, stmt) for atom in stmt.rhs.items]
                     stmt.rhs = llc.Tuple(
-                        [
-                            v.array if isinstance(v, VectorizedAccess) else v
-                            for v in replaced
-                        ]
+                        [replace_atom(atom, stmt) for atom in stmt.rhs.items]
                     )
                 elif isinstance(stmt.rhs, llc.Mux):
                     stmt.rhs.false_value = replace_operand(stmt.rhs.false_value, stmt)

### 2022-03-16 compiler/compiler/vectorize.py
Revert "Remove always-false isinstance() checks"
This reverts commit dfb8b9757fbf583cf85d8935c10953497a03bbd1.

@@ -472,12 +472,20 @@ def _basic_vectorization_phase_1(
                 elif isinstance(stmt.rhs, llc.UnaryOp):
                     stmt.rhs.operand = replace_operand(stmt.rhs.operand, stmt)
                 elif isinstance(stmt.rhs, llc.List):
+                    replaced = [replace_atom(atom, stmt) for atom in stmt.rhs.items]
                     stmt.rhs = llc.List(
-                        [replace_atom(atom, stmt) for atom in stmt.rhs.items]
+                        [
+                            v.array if isinstance(v, VectorizedAccess) else v
+                            for v in replaced
+                        ]
                     )
                 elif isinstance(stmt.rhs, llc.Tuple):
+                    replaced = [replace_atom(atom, stmt) for atom in stmt.rhs.items]
                     stmt.rhs = llc.Tuple(
-                        [replace_atom(atom, stmt) for atom in stmt.rhs.items]
+                        [
+                            v.array if isinstance(v, VectorizedAccess) else v
+                            for v in replaced
+                        ]
                     )
                 elif isinstance(stmt.rhs, llc.Mux):
                     stmt.rhs.false_value = replace_operand(stmt.rhs.false_value, stmt)

### 2022-03-16 compiler/compiler/tac_cfg.py
Add VectorizedAccess to Atom

@@ -26,9 +26,9 @@ from .ast_shared import (
 from .util import assert_never
 
 
-Atom = Union[Var, Constant]
+Atom = Union[Var, Constant, VectorizedAccess]
 
-Operand = Union[Atom, Subscript, "BinOp", "UnaryOp", VectorizedAccess]
+Operand = Union[Atom, Subscript, "BinOp", "UnaryOp"]
 
 
 class BinOp(_BinOp[Operand]):
@@ -89,7 +89,6 @@ AssignRHS = Union[
     Update,
     "LiftExpr",
     "DropDim",
-    VectorizedAccess,
 ]
 
 

### 2022-03-16 compiler/compiler/restricted_ast_to_tac_cfg.py
Fix two type errors

@@ -120,6 +120,7 @@ def _build_expression_as_var(
         builder.add_assignment(tac_cfg.Assign(lhs=var, rhs=atom))
         return var
     else:
+        assert not isinstance(atom, tac_cfg.VectorizedAccess)
         assert_never(atom)
 
 

### 2022-03-16 compiler/compiler/tac_cfg_to_ssa.py

@@ -170,6 +170,7 @@ def rename_variables(result: ssa.Function) -> None:
         elif isinstance(atom, ssa.Constant):
             return atom
         else:
+            assert not isinstance(atom, ssa.VectorizedAccess)
             assert_never(atom)
 
     def rename_operand(operand: ssa.Operand) -> ssa.Operand:

### 2022-03-16 compiler/compiler/vectorize.py
Fix another type error

@@ -416,6 +416,7 @@ def _basic_vectorization_phase_1(
             elif isinstance(atom, llc.Constant):
                 return atom
             else:
+                assert not isinstance(atom, llc.VectorizedAccess)
                 assert_never(atom)
 
         def replace_subscript(

### 2022-03-16 paper_SIMD.tex
Minor change in Extension with arrays.

@@
 
-Our compiler performs the following semantic analysis steps:
-
-\begin{enumerate}
-  \item
-        \textbf{Parsing}:
-        Use Python's \texttt{ast} module to parse the input source code to a Python AST.
-  \item
-        \textbf{Syntax checking}:
-        Ensure that the AST matches a restricted subset that our compiler supports.
-        This step outputs an instance of the \texttt{restricted\_ast.Function} class, which represents our restricted subset of the Python AST.
-  \item
-        \textbf{3-address CFG conversion}: TODO
-  \item
-        \textbf{SSA conversion}:
-        Convert the 3-address CFG to SSA with Cytron's algorithm.
-\end{enumerate}
+\ana{Benjamin?}
 
@@
 Otherwise, do nothing.
+\ana{I think it will be better to add drop dim and raise dim as in the previous case but change the semantics of both as written here.}
 

### 2022-03-16 overleaf/biometric_vectorization_example.tex
put paper latex sources in 'overleaf' folder

--- 2022-02-20 biometric_vectorization_example.tex
@@
 In the code below, all initializations (e.g., \texttt{min\_sum!3 = i}), operations, and PHI nodes are \emph{implicitly vectorized}. 
-$\mathit{raise\_dim}$ and $\mathit{drop\_dim}$ statements, as well as propagation statements (e.g., \texttt{min\_sum!3 = sum!3\^}) 
-may have slightly different interpretation. \ana{This is vague, may need elaboration.}
+$\mathit{raise\_dim}$ and $\mathit{drop\_dim}$ statements have slightly different interpretation. %\ana{This is vague, may need elaboration.}
 
-Note the two different versions of $\mathit{raise\_dim}$. \ana{More here! One just adds a least-significant dimension. The other one may reshape an input array.}
+The example illustrates the two different versions of $\mathit{raise\_dim}$. \texttt{C!0' = raise\_dim(C!0, j, (i:N!0,j:D!0))} reshapes
+the read-only input array, while \texttt{sum!3' = drop\_dim(sum!3)} removes the $j$ dimension of \texttt{sum!3}.
+%\ana{More here! One just adds a least-significant dimension. The other one may reshape an input array.}
 
@@
   // I dim is a noop. sum is already a one-dimensional vector
-  sum!3[I] = PHI(sum!2^[I], sum!4[I])       
+  sum!3[I] = PHI(sum!2[I], sum!4[I])       
   sum!4[I] = ADD_SIMD(sum!3[I],p!3[I,j])
@@
 
-\ana{Begin Work-in-progress notes on implementation. Just some thoughts...} 
+%\ana{Begin Work-in-progress notes on implementation. Just some thoughts...} 
 \begin{comment}
@@
 \end{comment}
-There are three kinds of arrays (for now all kept internally as one-dimensional arrays, but that's under discussion).
-\begin{itemize}
 
-\item Scalars: These are scalar variables we lift into arrays for the purposes of vectorization. 
-For those, all writes are canonical writes and all reads are canonical reads. We may apply both raise dimension 
-and drop dimension on these.
-
-\item Read-only input arrays: Read-only inputs. There are NO writes, while we may have non-canonical reads, $f(i,j,k)$. 
-Phase 1 of Basic vectorization will add raise dimension operation at the beginning of the function and raise dimension 
-may reshape arrays. If there are multiple "views" of the input array, there could be multiple raise dimension statements to create
-each one of these views. The invariant is that at reads in loops, the reads of "views" of the original input array will be canonical.
-Only raise dimension applies.
-
- \item Read-write output arrays: Writes are canonical (by restriction) but reads can be non-canonical. Conjecture:
- do nothing. Dependence analysis takes care of limiting vectorization so non-canonical access will work. 
-Still work in progress. We may apply both raise and drop dimension.
-
-\end{itemize}
-
-\ana{End Work-in-progress notes.}
+%\ana{End Work-in-progress notes.}
 

### 2022-03-16 overleaf/prob_stmt_np_hardness.tex
cleaned up the writeup for section 3

--- 2022-03-15 prob_stmt_np_hardness.tex
@@
 
-\ishaq{TODO: make it consistent with the next section.}
+At the lowest level, we have two types of MPC instructions (also called \emph{gates} in similar works) 1) local/non-interactive instruction (i.e.
+an addition instruction $A$) and 2) remote/interactive instruction (i.e.
+a multiplication instruction $M$).
+Each instruction in the program is either an $A$-instruction or an $M$-instruction.
 
-At the lowest level, we have two types of MPC instructions 1) local/non-interactive instruction (i.e.
-\add) and 2) remote/interactive instruction (i.e.
-\mul).
-Let the cost of a single \add instruction be given by a monotonically decreasing $f(n)$, where the argument $n$ is the number of \add instructions being executed in parallel.
-Similarly the cost of a single \mul is given by a monotonically decreasing function $g(n)$.
-
-Given a serial schedule (a linear graph) of an MPC program i.e.
-a sequence of instructions $\mathcal{S}:=(S_1, \dots, S_n)$, where $S_i \in \{\text{\add, \mul}\}, 1 \leq i \leq n$, and a def-use dependency graph $G(V, E)$ corresponding to $\mathcal{S}$, our task is to construct a parallel schedule (another linear graph) $\mathcal{P} := (P_1, \dots, P_n)$ observing the following conditions:
-
-\begin{enumerate}
-    \item Multiple, not necessarily continuous instructions of the same type (i.e.
-either \add or \mul) from $\mathcal{S}$ can be grouped into a single $P_i$  However, all such instructions must be of the same type (either \add or \mul).
-    \item Def-use dependencies of the graph $G(V, E)$ are should be preserved i.e.
-if instructions $S_i, S_j, i < j$ are a def-use (an edge exists from $S_i$ to $S_j$ in $G$), then they can only be mapped to $P_{i'}, P_{j'}, i' < j'$.
-\end{enumerate}
-
-Our goal is to construct minimize the height of the graph $\mathcal{P}$. Indeed, a graph $\mathcal{P}$ with minimum height will maximize parallelization.
-
-\paragraph{Correctness} Correctness of $\mathcal{P}$ is guaranteed by definition. Since def-use dependencies are preserved, the function (being computed) remains the same.
-
-\paragraph{Cost Comparison} For the sequential schedule $\mathcal{S}$ consisting of $L$ local and $R$ remote instructions, the total cost is $\mathit{cost}(\mathcal{S}) = L \cdot f(1) + R \cdot g(1)$.
-In the extreme case where all $L$ and all $R$ instructions can be parallelized, the cost of $\mathcal{P}$ is $\mathit{cost}(\mathcal{P}) = L \cdot f(L) + R \cdot g(R)$.
-Since both $f$ and $g$ are monotonically decreasing, $\mathit{cost}(\mathcal{P}) < \mathit{cost}(\mathcal{S})$.
-Cost of all other parallel schedules lies between the extremes of $\mathit{cost}(\mathcal{S})$ and $\mathit{cost}(\mathcal{P})$.
-
-Note that we use an MPC-Source control flow graph (CFG) $G'(V', E')$ along with def-use graph $G(V,E)$ to construct $\mathcal{P}$.
-We consider a linearized MPC schedule $\mathcal{S}$ above for ease of exposition.
-The argument becomes slightly more involved when dealing with a graph $G'$ that may contain cycles.
-
-% Given a def-use dependency graph $G(V, E)$ that is constructed from for the input program.
-Each vertex/node in this graph represents an operation/instruction in the program and an edge represents dependence i.e.
-an edge from $V_1$ to $V_2$ means the result of operation in $V_1$ is used in the operation in $V_2$.
-To build a parallel schedule, we translate this graph $G(V, E)$ into another dependency graph $G'(V', E')$ such that:
-
-% \begin{enumerate}
-%     \item We collapse multiple vertices of the same type (i.e. same operation) in $V$ into a single vertex in $V'$.
-%     \item dependencies are not broken i.e.
-If there is an edge from $V_1$ to $V_2$ ($V_1, V_2 \in V$), and we put them in vertices $V'_1, V'_2 \in G'$ respectively, then an edge must exist from $V'_1$ to $V'_2$.
-% \end{enumerate}
-
-% We want to build a graph $G'$ with fewest possible vertices since fewer vertices means more parallelization.
-
-%\ishaq{TODO: @Ana: I am still confused about what prevents a reviewer from saying, "you argue correctness of the schedule from unrolled MPC, you don't argue it for the pre-unrolled CFG that you use".
-Please comment.} \ana{The correctness proof answer this question.
-The unrolled schedule is the \emph{concretization} of MPC source.
-We define (1) the obvious $\gamma$, (2) we define partial order over MPC source codes (this is our abstract domain $A$), and (3) a partial order over unrolled schedules.
-Then we prove two theorems that $A \le A' \Rightarrow \gamma(A) \subseteq \gamma(A')$ and that the MPC source $\le$ Vectorized MPC source.}
-
-%\end{comment}
-
-\subsection{Scheduling is NP-hard}
-\label{sec:np}
-
-\ishaq{TODO: need make amortized cost a function (like in the problem statement above), it is proving to be tricky.}
-
-We consider two operations, call them $A$ and $M$.
-$A$ and $M$ are two abstract MPC instruction, but as an example, $A$ stands for the ADD MPC instruction and $M$ stands for the MUL instruction.
-Each instruction in the program is either an $A$-instruction or an $M$-instruction.
 In order to benefit from parallelization/amortization, we must schedule two or more $A$-instructions in the same parallel node (or two or more $M$-instructions in the same parallel node).
@@
 It incurs the exact same cost as scheduling the $A$-instructions in a node $P_A$, scheduling the $M$-instructions in a node $P_M$, and having $P_A$ precede $P_M$ in the parallel schedule.
-
 We use the following cost model:
@@
 
-Consider a loop body that consists of $n$ sequences: $S_1$, ...
-$S_n$ of $A$ and $M$ instructions.
-More precisely, the loop body is such that its instructions can be grouped into such sequences.
-$S_1$, ...
-$S_n$ can execute in parallel, however, all instructions within a sequence must execute sequentially.
-For example, consider the three sequences (the right arrow indicates a \emph{dependence}, meaning that the source node must execute before the target node):
+% Let the cost of a single \add ~instruction be given by a monotonically decreasing $f(n)$, where the argument $n$ is the number of \add ~instructions being executed in parallel.
+Similarly the cost of a single \mul ~is given by a monotonically decreasing function $g(n)$.
+
+% Given a serial schedule (a linear graph) of an MPC program i.e.
+a sequence of instructions $\mathcal{S}:=(S_1, \dots, S_n)$, where $S_i \in \{A, M\}, 1 \leq i \leq n$, and a def-use dependency graph $G(V, E)$ corresponding to $\mathcal{S}$, our task is to construct a parallel schedule (another linear graph) $\mathcal{P} := (P_1, \dots, P_n)$ observing the following conditions:
+
+% \begin{enumerate}
+%     \item Multiple, not necessarily continuous instructions of the same type (i.e. either $A$ or $M$) from $\mathcal{S}$ can be grouped into a single $P_i$.
+%     \item Def-use dependencies of the graph $G(V, E)$ should be preserved i.e.
+if instructions $S_i, S_j, i < j$ are a def-use (an edge exists from $S_i$ to $S_j$ in $G$), then they can only be mapped to $P_{i'}, P_{j'}, i' < j'$.
+% \end{enumerate}
+
+% Our goal is to construct minimize the height of the graph $\mathcal{P}$. Indeed, a graph $\mathcal{P}$ with minimum height will maximize parallelization.
+
+\ishaq{Should be better to model this as a tree, and the branches (before they merge) can be executed in parallel, Right now it is too imprecise. Ideas please?}
+Consider an MPC program as a set of $n$ sequences: $\{S_1, \dots, S_n\}$.
+Each $S_i$ is a tuple of $A$ and $M$ instructions.\footnote{Every MPC program can be transformed into such a set of sequences.} All elements of the set $\{S_1, \dots, S_n\}$ can execute in parallel, however, all instructions within an element (i.e.
+sequence) must execute sequentially.
+For example, consider the three sequences, right arrow indicates a \emph{dependence}, meaning that the source node must execute before the target node:
 
@@
 
-A \emph{schedule} $P: P_1 \rightarrow P_2 \dots \rightarrow P_k$ is such that for each sequence $S_i$ in the set, if $S_i[j]$ precedes $S_i[j']$ in $S_i$ then $S_i[j]$ is scheduled in node $P_\ell$, $S_i[j']$ is scheduled in node $P_{\ell'}$, and $P_\ell$ precedes $P_{l\ell'}$ in $P$.
+A \emph{schedule} $P: P_1 \rightarrow P_2 \dots \rightarrow P_k$ is such that for each sequence $S_i$ in the set, if $S_i[j]$ precedes $S_i[j']$ in $S_i$ then $S_i[j]$ is scheduled in node $P_\ell$, $S_i[j']$ is scheduled in node $P_{\ell'}$, and $P_\ell$ precedes $P_{\ell'}$ in $P$.
 
@@
 
-where $\mathit{cost}(P_i) = \alpha$ if $P_i$ consists of $A$-instructions only, $\beta$ if $P_i$ consists of $M$-instructions only, and $\mathit{cost}(P_i) = \alpha + \beta$ if $P_i$ mixes $A$-instructions and $M$-instructions.
+where $\mathit{cost}(P_i) = \alpha$ if $P_i$ consists of $A$-instructions only, it is $\beta$ if $P_i$ consists of $M$-instructions only, and it is $(\alpha + \beta)$ if $P_i$ mixes $A$-instructions and $M$-instructions.
 
-The problem is to find a schedule $P$ with \emph{minimal cost}.
-For example, a schedule with minimal cost for the sequences above is \[ A(1), A(2) \rightarrow M(1), A(2), M(3) \rightarrow A(1), A(2), A(3) \rightarrow M(3) \] (The parentheses above indicate the sequence where the instruction comes from: (1), (2), or (3).) The cost of this schedule is $3\alpha + 2\beta$.
+The problem is to find a schedule $P$ with \emph{minimal cost}. For example, a schedule with minimal cost for the sequences above is
+
+$$ 
+A(1), A(2) \rightarrow M(1), A(2), M(3) \rightarrow A(1), A(2), A(3) \rightarrow M(3)
+$$
+
+The parentheses above indicate the sequence where the instruction comes from: (1), (2), or (3). The cost of this schedule is $3\alpha + 2\beta$.
+
+\paragraph{Correctness} Correctness of $\mathcal{P}$ is guaranteed by definition. Since \emph{dependencies} are preserved, the computed function remains the same.
+
+% \paragraph{Cost Comparison} For the sequential schedule $\mathcal{S}$ consisting of $L$ local and $R$ remote instructions, the total cost is $\mathit{cost}(\mathcal{S}) = L \cdot f(1) + R \cdot g(1)$.
+In the extreme case where all $L$ and all $R$ instructions can be parallelized, the cost of $\mathcal{P}$ is $\mathit{cost}(\mathcal{P}) = L \cdot f(L) + R \cdot g(R)$.
+Since both $f$ and $g$ are monotonically decreasing, $\mathit{cost}(\mathcal{P}) < \mathit{cost}(\mathcal{S})$.
+Cost of all other parallel schedules lies between the extremes of $\mathit{cost}(\mathcal{S})$ and $\mathit{cost}(\mathcal{P})$.
+
+% Note that we use an MPC-Source control flow graph (CFG) $G'(V', E')$ along with def-use graph $G(V,E)$ to construct $\mathcal{P}$.
+We consider a linearized MPC schedule $\mathcal{S}$ above for ease of exposition.
+The argument becomes slightly more involved when dealing with a graph $G'$ that may contain cycles.
+
+\subsection{Scheduling is NP-hard}
+\label{sec:np}
+
+\ishaq{TODO: rather than multiple amortized instructions costing the same as a single non-amortized instruction, Assume a function monotonically decreasing function for costs and prove the theorem using that function.}
 

### 2022-03-20 benchmarks/floyd_warshall.py
Adding HyCC benchmarks.

@@ -1,0 +1,23 @@
+from UTIL import shared
+
+# Floyd-Warshall's algorithm for computing shortest distance between vertices 
+# Array Dist is the N*N distance matrix
+# Dist = [0,10000,-2,10000,4,0,3,10000,10000,10000,0,2,10000,-1,10000,0]
+# N is number of vertices
+
+# requires: len(Dist) == N*N
+# Dist[i,i] = 0, Dist[i,j] = weights(i,j) if edge i->j, Dist[i,j] = infinity otherwise 
+def floyd_warshall(
+    Dist: shared[list[int]], N: int) -> shared[list[int]]:
+    for k in range(N):
+        for i in range(N):
+            for j in range(N):
+                if Dist[i*N+j] > Dist[i*N+k]+Dist[k*N+j]:
+                    Dist[i*N+j] = Dist[i*N+k]+Dist[k*N+j]
+    return Dist
+
+# Wishful. Array write extension may require modification to handle!!!
+# 10000 = infinity
+Dist = [0,10000,-2,10000,4,0,3,10000,10000,10000,0,2,10000,-1,10000,0]
+N = 4
+print(floyd_warshall(Dist, N))

### 2022-03-20 benchmarks/matrix_multiply.py

@@ -1,0 +1,27 @@
+from UTIL import shared
+
+# Array A contains linearization of a MxK matrix
+# Array B contains linearization of a K*N matrix
+# Array C of 0's is the result M*N matrix
+# A = [0,2,1,0,3,4,2,3,10] 3x3 matrix
+# B = [10,1,5,2,15,0,10,1000,1,1,2,3] 3x4 matrix
+# C = [0,0,0,0,0,0,0,0,0,0,0,0] 3x4 matrix of 0's
+
+# requires: len(A) == M*K, len(B) == K*N, len(C) = M*N, C is an array of 0's
+def matrix_multiply(
+    A: shared[list[int]], B: shared[list[int]], C: shared[list[int]], M: int, N: int, K: int) -> shared[list[int]]:
+    for i in range(M):
+        for j in range(N):
+            for k in range(K):
+                # C[i,j] = C[i,j] + A[i,k]*B[k,j]
+                C[i*N+j] = C[i*N+j] + A[i*K+k]*B[k*N+j]
+    return C
+
+
+A = [0,2,1,0,3,4,2,3,10]
+B = [10,1,5,2,15,0,10,1000,1,1,2,3]
+C = [0,0,0,0,0,0,0,0,0,0,0,0]
+M = 3
+N = 4
+K = 3
+print(matrix_multiply(A, B, C, M, N, K))

### 2022-03-20 overleaf/paper_SIMD.tex
put paper latex sources in 'overleaf' folder
cleaned up the writeup for section 3
Minor tweaks to Extension with Array Writes. (Made more precise.)

--- 2022-03-16 paper_SIMD.tex
@@
 
-% defining the \BibTeX command - from Oren Patashnik's original BibTeX documentation.
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08emT\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+% % defining the \BibTeX command - from Oren Patashnik's original BibTeX documentation.
+% \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08emT\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
 
@@
 
+% Minted is used for source code highlighting
 %\usepackage[outputdir=/Volumes/ramdisk/]{minted}
-\usepackage{minted}
+% \usepackage{minted}
 % *** MATH PACKAGES ***
@@
 \usepackage[backend=biber,style=ACM-Reference-Format]{biblatex}
-\addbibresource{library.bib}
-\addbibresource{../cryptobib/abbrev2.bib}
-\addbibresource{../cryptobib/crypto_crossref.bib}
+\addbibresource{./library.bib}
+\addbibresource{./cryptobib/abbrev2.bib}
+\addbibresource{./cryptobib/crypto_crossref.bib}
 
@@
 d!3[I,J] = SUB_SIMD(S!0^[I,J],C!0^[I,J])
-p!3[I,J] = MUL_SIMD(d!3[I,J] * d!3[I,J])
+p!3[I,J] = MUL_SIMD(d!3[I,J],d!3[I,J])
 
@@
 \STATE \hspace{0.25cm} {outer-to-inner: add $Y'_n[i,j,k] = \mathit{raise\_dim}(Y_n)$ at $\mathit{stmt'}$}
+\STATE \hspace{0.25cm} {(more precisely, right after $\mathit{stmt'}$)}
 \STATE \hspace{0.25cm} {inner-to-outer: add $Y'_n[i,j,k] = \mathit{drop\_dim}(Y_n)$ at $\mathit{stmt}$}
+\STATE \hspace{0.25cm} {(more precisely, in loop of $\mathit{stmt}$ right after loop of $\mathit{stmt'}$)}
 \ENDFOR
@@
 \STATE {in vectorizable stmts, replace use of $A_1$ with $A_0$}
-\STATE {discard $\phi$-node if not used in any $cl$}
+\STATE {discard $\phi$-node if not used in any $cl$, replacing $A_1$ with $A_0$ or $A_k$ appropriately}
 \ENDFOR
@@
 
-Note however, that this step renders some array phi-nodes target-less. We handle target-less phi-nodes with a minor extension of Basic Vectorization. 
+Note however, that this step renders some array phi-nodes target-less. We handle target-less phi-nodes with a minor extension of Basic Vectorization (Phase 2). 
 First, we merge closures that update the same array. This simplifies handling of array $\phi$-nodes: if each closure is turned into a separate loop
@@
 
-In addition to the changes for the handling of target-less phi-nodes, Basic Vectorization has to handle differently def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable.
-The definition can be an update \texttt{A\_2 = update(A\_1,i,...)} or a pseudo $\phi$-node \texttt{A\_2 = PHI(A\_0,A\_1)}, etc..
+In addition to the changes for the handling of target-less phi-nodes, Basic Vectorization has to handle def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable.
+The definition can be an update \texttt{A\_2 = update(A\_1,i,...)}, a pseudo $\phi$-node \texttt{A\_2 = PHI(A\_0,A\_1)}, etc..
 Note that $\phi$-nodes for arrays have no subscript operations the way there are subscript operations in analysis-introduced arrays representing scalars.
-Raising and dropping dimension still apply, however, the dimension cannot be raised or dropped to a dimension lower than the canonical one:
+While there are variations, the most intuitive implementation will perform Basic Vectorization Phase 1 as is, inserting
+$\mathit{raise\_dim}$ and $\mathit{drop\_dim}$ in the same way.
+However, the implementation of raise dimension and drop dimension will be adapted because the dimension cannot be raised or drop to a dimension lower than the canonical one.
+Consider a def-use edge $X \rightarrow Y$ for an array $\texttt{A}$.
 
@@
 
-\item same-level $X \rightarrow Y$. We do nothing, just propagate the array, which happens to be of the right dimension. 
+\item same-level $X \rightarrow Y$. We propagate the array, which happens to be of the right dimension. 
 %\ana{There might be some opportunities to do copy propagation optimization and save some cycles, but let's leave this for later.}
 
-\item inner-to-outer $X \rightarrow Y$.
-If dimensionality of the loop enclosure of $X$ is greater than the canonical dimensionality of the array, then add \emph{drop\_dim(...)} at $Y$, as in Basic Vectorization.
-Otherwise, do nothing.
-
-\item outer-to-inner $X \rightarrow Y$.
-If dimensionality of loop enclosure of $Y$ is greater than the canonical dimensionality of the array, then add \emph{raise\_dim(...)} (at $X$) as in Basic Vectorization.
-Otherwise, do nothing.
+\item inner-to-outer $X \rightarrow Y$ triggers the addition of $\mathit{drop\_dim}$.
+However, if the dimensionality of the loop enclosure of $X$ is equal or less than the canonical dimensionality of the array, then $\mathit{drop\_dim}$ has no effect, otherwise its behavior is as in Basic Vectorization.
+
+\item outer-to-inner $X \rightarrow Y$ riggers  $\mathit{raise\_dim}$.
+Again, if the dimensionality of the loop enclosure of $Y$ is smaller or same as the canonical dimensionality of the array, then it has no effect, otherwise, if dimensionality is greater than the canonical dimensionality, $\mathit{raise\_dim(...)}$ (at $X$) is the same as in Basic Vectorization.
 
@@
 \end{enumerate}
+
+If the use of the array is a read $\texttt{A}[f(i,j,k)]$ different than a canonical read $\texttt{A}[i,j,k]$, then we need to add a reshape operation as 
+all arrays are $\texttt{A}[i,j,k]$. It can be added after raise\_dim/drop\_dim or incorporated in these operations.
+The bulk of the change is in Phase 2 of Basic Vectorization as outlined earlier.
 
@@
 for i in range(0, num_bins):
-   res1 = PHI(res, res2) // target-less, no drop_dim
+   res1 = PHI(res, res2') // target-less phi-node
    res1' = raise_dim(res1, (j:N))
@@
       res3 = Update(res2, (I,J), tmp3) 
-return res1
+   res2' = drop_dim(res2)    
+res1'' = drop_dim(res1)
+return res1''
 \end{verbatim}
@@
 This is because the PHI node is a target-less node, and therefore, there are no closures!
-Thus, everything can be vectorized along the $i$ dimension:
+Several things happen.
+(1) Everything can be vectorized along the $i$ dimension.
+(2) We remove the target-less PHI node, however, we must update uses of \texttt{res1} appropriately: the use at raise\_dim goes to the first argument of the PHI function and the use at drop\_dim goes to the second argument.
 
@@
 
-res1' = raise_dim(res, (j:N)) // raising res to a 2-dim. array
+res1' = raise_dim(res, (j:N)) // replacing res1 with res, 1st arg
 for j in range(0, N):
@@
    res3 = Update(res2, (I,j), tmp3)
-res1 = drop_dim(res2) // Add drop_dim before removing redundant phi-node!      
+   equiv. to res3 = res2; res3[I,j] = tmp3[I,j]
+res2' = drop_dim(res2)   
+res1 = drop_dim(res2') // replacing with res2', 2nd arg. NOOP
 return res1
@@
 
+\begin{comment}
 The cleanup phase can then get rid of $j$ in the \texttt{res} variables and \texttt{tmp} variables, so the loop operates over the \texttt{num\_bins}-size vectors. 

### 2022-03-20 overleaf/paper_SIMD.tex
Minor fix (typos and rewrites).

@@
 $\mathit{raise\_dim}$ and $\mathit{drop\_dim}$ in the same way.
-However, the implementation of raise dimension and drop dimension will be adapted because the dimension cannot be raised or drop to a dimension lower than the canonical one.
+However, the implementation of raise dimension and drop dimension will be adapted because the dimension cannot be raised or dropped to a dimension lower than the canonical one.
 Consider a def-use edge $X \rightarrow Y$ for an array $\texttt{A}$.
@@
 
-\item same-level $X \rightarrow Y$. We propagate the array, which happens to be of the right dimension. 
+\item same-level $X \rightarrow Y$. Do nothing, propagate the array, which happens to be of the right dimension. 
 %\ana{There might be some opportunities to do copy propagation optimization and save some cycles, but let's leave this for later.}
@@
 \item inner-to-outer $X \rightarrow Y$ triggers the addition of $\mathit{drop\_dim}$.
-However, if the dimensionality of the loop enclosure of $X$ is equal or less than the canonical dimensionality of the array, then $\mathit{drop\_dim}$ has no effect, otherwise its behavior is as in Basic Vectorization.
+However, the dimensionality cannot be dropped below the canonical dimensionality of the array.
+E.g., if the dimensionality of the loop enclosure $X$ is already at the canonical one, then $\mathit{drop\_dim}$ has no effect.
 

### 2022-03-22 overleaf/prob_stmt_np_hardness.tex
updated NP-Hardness proof, added "loud thinking" section

@@
 
+% Let the cost of a single \add ~instruction be given by a monotonically decreasing $f(n)$, where the argument $n$ is the number of \add ~instructions being executed in parallel.
+Similarly the cost of a single \mul ~is given by a monotonically decreasing function $g(n)$.
+
+Given a serial schedule (a linear graph) of an MPC program i.e.
+a sequence of instructions $S := (S_1, \dots, S_n)$, where $S_i \in \{A, M\}, 1 \leq i \leq n$, and a def-use dependency graph $G(V, E)$ corresponding to $S$, our task is to construct a parallel schedule (another linear graph) $P := (P_1, \dots, P_m)$ observing the following conditions:
+
+\begin{enumerate}
+    \item Multiple, not necessarily continuous instructions from $S$ can be grouped into a single $P_i$.
+    \item Def-use dependencies of the graph $G(V, E)$ should be preserved i.e.
+if instructions $S_i, S_j, i < j$ form a def-use i.e.
+an edge exists from $S_i$ to $S_j$ in $G$), then they can only be mapped to $P_{i'}, P_{j'}$ such that $i' < j'$.
+\end{enumerate}
+
+\paragraph{Correctness} Correctness of $\mathcal{P}$ is guaranteed by definition. Preserving def-use \emph{dependencies} means the computed function remains the same in both $S$ and $P$.
+
 In order to benefit from parallelization/amortization, we must schedule two or more $A$-instructions in the same parallel node (or two or more $M$-instructions in the same parallel node).
 We also assume that scheduling $A$-instructions in parallel with $M$-instruction does not benefit from amortization\footnote{this is not strictly true, but assuming it, e.g.
-as in \cite{Ishaq2019, Demmler2015ABYA, Mohassel2018}, helps with the exposition.}.
+as in \cite{Ishaq2019, Demmler2015ABYA, Mohassel2018}, helps simplify the problem.}.
 It incurs the exact same cost as scheduling the $A$-instructions in a node $P_A$, scheduling the $M$-instructions in a node $P_M$, and having $P_A$ precede $P_M$ in the parallel schedule.
@@
 
-% Let the cost of a single \add ~instruction be given by a monotonically decreasing $f(n)$, where the argument $n$ is the number of \add ~instructions being executed in parallel.
-Similarly the cost of a single \mul ~is given by a monotonically decreasing function $g(n)$.
+The cost of schedule $S$ is
+\begin{equation}
+    \mathit{cost}(S) = \sum_{i=1}^n \mathit{cost}(S_i)
+\end{equation}
 
-% Given a serial schedule (a linear graph) of an MPC program i.e.
-a sequence of instructions $\mathcal{S}:=(S_1, \dots, S_n)$, where $S_i \in \{A, M\}, 1 \leq i \leq n$, and a def-use dependency graph $G(V, E)$ corresponding to $\mathcal{S}$, our task is to construct a parallel schedule (another linear graph) $\mathcal{P} := (P_1, \dots, P_n)$ observing the following conditions:
+where $cost(S_i) = \alpha$ if $S_i$ is an $A$-instruction and $cost(S_i) = \beta$ if $S_i$ is an $M$-instruction. Similarly, the cost of schedule $P$ is
 
-% \begin{enumerate}
-%     \item Multiple, not necessarily continuous instructions of the same type (i.e. either $A$ or $M$) from $\mathcal{S}$ can be grouped into a single $P_i$.
-%     \item Def-use dependencies of the graph $G(V, E)$ should be preserved i.e.
-if instructions $S_i, S_j, i < j$ are a def-use (an edge exists from $S_i$ to $S_j$ in $G$), then they can only be mapped to $P_{i'}, P_{j'}, i' < j'$.
-% \end{enumerate}
+\begin{equation}\label{eq_par_schedule_cost}
+    \mathit{cost}(P) = \sum_{i=1}^m \mathit{cost}(P_i)
+\end{equation}
 
-% Our goal is to construct minimize the height of the graph $\mathcal{P}$. Indeed, a graph $\mathcal{P}$ with minimum height will maximize parallelization.
+Each $P_i$ may contain multiple instructions, thus $\mathit{cost}(P_i) = \alpha$ if $P_i$ consists of $A$-instructions only, it is $\beta$ if $P_i$ consists of $M$-instructions only, and it is $(\alpha + \beta)$ if $P_i$ mixes $A$-instructions and $M$-instructions.
+Our goal is to construct a parallel schedule $P$ that reduces the program cost (when compared to cost of $S$).
 
-\ishaq{Should be better to model this as a tree, and the branches (before they merge) can be executed in parallel, Right now it is too imprecise. Ideas please?}
-Consider an MPC program as a set of $n$ sequences: $\{S_1, \dots, S_n\}$.
-Each $S_i$ is a tuple of $A$ and $M$ instructions.\footnote{Every MPC program can be transformed into such a set of sequences.} All elements of the set $\{S_1, \dots, S_n\}$ can execute in parallel, however, all instructions within an element (i.e.
-sequence) must execute sequentially.
-For example, consider the three sequences, right arrow indicates a \emph{dependence}, meaning that the source node must execute before the target node:
+% \paragraph{Cost Comparison} For the sequential schedule $\mathcal{S}$ consisting of $L$ local and $R$ remote instructions, the total cost is $\mathit{cost}(\mathcal{S}) = L \cdot f(1) + R \cdot g(1)$.
+In the extreme case where all $L$ and all $R$ instructions can be parallelized, the cost of $\mathcal{P}$ is $\mathit{cost}(\mathcal{P}) = L \cdot f(L) + R \cdot g(R)$.
+Since both $f$ and $g$ are monotonically decreasing, $\mathit{cost}(\mathcal{P}) < \mathit{cost}(\mathcal{S})$.
+Cost of all other parallel schedules lies between the extremes of $\mathit{cost}(\mathcal{S})$ and $\mathit{cost}(\mathcal{P})$.
+
+Note that we consider a linearized MPC schedule $S$ above for ease of exposition only.
+In our tool-chain we use an MPC-Source control flow graph (CFG) $G'(V', E')$ along with def-use graph $G(V,E)$ to construct $P$.
+The argument becomes slightly more involved when dealing with a graph $G'$ that may contain cycles.
+
+\subsection{Scheduling is NP-hard}
+\label{sec:np}
+
+To prove that optimal scheduling is an NP-Hard problem, we consider the following convenient representation.
+An MPC program is represented as a set of sequences $S = \{S_1, \dots, S_n\}$.
+Each element $S_i \in S$ is a tuple.
+The items of the tuple $S_i$ are operations, i.e.
+$A$ or $M$ instructions, that have to be executed in order (operations depend on previous operations i.e.
+$S_i[j], j > 1$ depends $S_i[j-1]$).
+However, the sequences $S_i, 1 \leq i \leq n$ themselves, can overlap each other in any way i.e.
+distinct sequences can be executed in parallel.
+We argue that an MPC program can be transformed into such collection of sequences by traversing the circuit for each pair of input and output values.
+\ishaq{I am not sure how this will be done, see \cref{fig:sequence_construction}}
+
+As an example, consider the MPC program consisting of the following three sequences, right arrow indicates a \emph{dependence}, meaning that the source node must execute before the target node:
 
@@
 
-The cost of a schedule $P$ is
+Cost of schedule $P$ is computed using \cref{eq_par_schedule_cost} above.
+% The cost of a schedule $P$ is
 
-\begin{equation}
-    \mathit{cost}(P) = \sum_{i=1}^k \mathit{cost}(P_i)
-\end{equation}
+% \begin{equation}
+%     \mathit{cost}(P) = \sum_{i=1}^k \mathit{cost}(P_i)
+% \end{equation}
 
-where $\mathit{cost}(P_i) = \alpha$ if $P_i$ consists of $A$-instructions only, it is $\beta$ if $P_i$ consists of $M$-instructions only, and it is $(\alpha + \beta)$ if $P_i$ mixes $A$-instructions and $M$-instructions.
+% where $\mathit{cost}(P_i) = \alpha$ if $P_i$ consists of $A$-instructions only, it is $\beta$ if $P_i$ consists of $M$-instructions only, and it is $(\alpha + \beta)$ if $P_i$ mixes $A$-instructions and $M$-instructions.
 
@@
 
-\paragraph{Correctness} Correctness of $\mathcal{P}$ is guaranteed by definition. Since \emph{dependencies} are preserved, the computed function remains the same.
-
-% \paragraph{Cost Comparison} For the sequential schedule $\mathcal{S}$ consisting of $L$ local and $R$ remote instructions, the total cost is $\mathit{cost}(\mathcal{S}) = L \cdot f(1) + R \cdot g(1)$.
-In the extreme case where all $L$ and all $R$ instructions can be parallelized, the cost of $\mathcal{P}$ is $\mathit{cost}(\mathcal{P}) = L \cdot f(L) + R \cdot g(R)$.
-Since both $f$ and $g$ are monotonically decreasing, $\mathit{cost}(\mathcal{P}) < \mathit{cost}(\mathcal{S})$.
-Cost of all other parallel schedules lies between the extremes of $\mathit{cost}(\mathcal{S})$ and $\mathit{cost}(\mathcal{P})$.
-
-% Note that we use an MPC-Source control flow graph (CFG) $G'(V', E')$ along with def-use graph $G(V,E)$ to construct $\mathcal{P}$.
-We consider a linearized MPC schedule $\mathcal{S}$ above for ease of exposition.
-The argument becomes slightly more involved when dealing with a graph $G'$ that may contain cycles.
-
-\subsection{Scheduling is NP-hard}
-\label{sec:np}
-
-\ishaq{TODO: rather than multiple amortized instructions costing the same as a single non-amortized instruction, Assume a function monotonically decreasing function for costs and prove the theorem using that function.}
-
-The problem of finding a schedule $P$ with a minimal $cost(P)$ for a given loop body has been shown to be an NP-Hard problem, as it can be reduced to the problem of finding a \emph{shortest common supersequence}, a known NP-Hard problem\cite{Maier1978},\cite{Vazirani2010}.
+The problem of finding a schedule $P$ with a minimal $cost(P)$ for a given loop body has been shown to be an NP-Hard problem, as it can be reduced to the problem of finding a \emph{shortest common supersequence}, a known NP-Hard problem\cite{Maier1978, Vazirani2010}.
 The shortest common supersequence problem is as follows: {\it given two or more sequences find the the shortest sequence that contains all of the original sequences.} This can be solved in $O(n^k)$ time, where $n$ is the cardinality of the longest sequence and $k$ is the number of sequences.
@@
 Therefore $\mathit{cost}(P'') < \mathit{cost}(P')$ is contradiction and no such $P''$ exists.
+\qed
+
+\subsection{Loud Thinking}
+\ishaq{I will delete this section after we have decided on how to handle various issues raised in this section}.
+
+\begin{figure}
+  \includegraphics[width=0.6\linewidth]{overleaf/figs_paper_SIMD/sequence_construction.pdf}
+  \caption{How do we construct sequences from this circuit?} 
+  \label{fig:sequence_construction}
+\end{figure}
+
+The more realistic cost model is as under:
+\begin{enumerate}
+    \item The cost of an $A$-instruction is given by a function $\alpha(\cdot)$ where the only parameter to the function is the number of $A$ instructions that will be executed in parallel.
+    \item Similarly, the cost of an $M$-instruction is given by function $\beta(\cdot)$.
+    \item Both $\alpha$ and $\beta$ are amortization function e.g. $\alpha(n) \leq \alpha(\ell) + \alpha(m); n = \ell + m$. Same condition applies for $\beta$.
+\end{enumerate}
+
+\begin{figure}
+  \includegraphics[width=0.6\linewidth]{overleaf/figs_paper_SIMD/cost_inequality.pdf}
+  \caption{It is unclear whether right super sequence cost less/more/the-same as left one.} 
+  \label{fig:cost_comparison}
+\end{figure}
+
+I am not using this cost model (for now) because NP-Hardness proof is tricky (see \cref{fig:cost_comparison}).
+Consider a super sequence that has nodes (of only one type) with following weights: $SS_1 = (1, 1, 4)$ where $Length(SS_1) = 3$ and $cost(SS_1) = \alpha(1) + \alpha(1) + \alpha(4)$.
+
+Using $\alpha(4) \leq \alpha(1) + \alpha(3)$, we can say
+
+\begin{equation}\label{ineq_cost_1}
+    \alpha(1) + \alpha(1) + \alpha(4) \leq \alpha(1) + \alpha(1) + \alpha(1) + \alpha(3)
+\end{equation}
+
+Now suppose, for the same schedule, another super sequence with weights $SS_2 = (3, 3)$ exists, $Length(SS_2) = 2$ and $cost(SS_2) = \alpha(3) + \alpha(3)$.
+Using $\alpha(3) \leq \alpha(1) + \alpha(1) + \alpha(1)$, we can say
+
+\begin{equation}\label{ineq_cost_2}
+    \alpha(3) + \alpha(3) \leq \alpha(1) + \alpha(1) + \alpha(1) + \alpha(3)
+\end{equation}
+
+The problem is that while RHS is the same in \cref{ineq_cost_1} and \cref{ineq_cost_2}, we cannot say anything about the relationship between their LHS.
+It could be possible that cost of $SS_2$, LHS of  \cref{ineq_cost_2}, is more than cost of $SS_1$, LHS of \cref{ineq_cost_1}.
+Thus, we have a shorter super sequence $SS_2$ that does not contradict that the optimality of the schedule used to generate $SS_1$.

### 2022-03-22 overleaf/prob_stmt_np_hardness.tex
add missing files

@@
 \begin{figure}
-  \includegraphics[width=0.6\linewidth]{overleaf/figs_paper_SIMD/sequence_construction.pdf}
+  \includegraphics[width=0.6\linewidth]{./figs_paper_SIMD/sequence_construction.pdf}
   \caption{How do we construct sequences from this circuit?} 
@@
 \begin{figure}
-  \includegraphics[width=0.6\linewidth]{overleaf/figs_paper_SIMD/cost_inequality.pdf}
+  \includegraphics[width=0.6\linewidth]{./figs_paper_SIMD/cost_inequality.pdf}
   \caption{It is unclear whether right super sequence cost less/more/the-same as left one.} 

### 2022-03-22 compiler/compiler/restricted_ast_to_tac_cfg.py
Remove unneeded current_block_done()

@@ -39,9 +39,6 @@ class _CFGBuilder:
         self._current_block = block
         assert self._current_block.terminator is None
 
-    def current_block_done(self) -> bool:
-        return self._current_block.terminator is not None
-
     def add_assignment(self, assignment: tac_cfg.Assign):
         assert self._current_block.terminator is None
         self._current_block.assignments.append(assignment)
@@ -212,13 +209,11 @@ def _build_if(if_statement: restricted_ast.If, builder: _CFGBuilder):
 
     builder.set_current_block(then_block)
     _build_statements(if_statement.then_body, builder)
-    if not builder.current_block_done():
-        builder.add_jump(after_block)
+    builder.add_jump(after_block)
 
     builder.set_current_block(else_block)
     _build_statements(if_statement.else_body, builder)
-    if not builder.current_block_done():
-        builder.add_jump(after_block)
+    builder.add_jump(after_block)
 
     builder.set_current_block(after_block)
     assert isinstance(condition_block.terminator, tac_cfg.ConditionalJump)

### 2022-03-22 overleaf/paper_SIMD.tex
Add write-up for 3-address CFG conversion

@@
   \item
-        \textbf{3-address CFG conversion}: TODO
+        \textbf{3-address CFG conversion}:
+        TODO: Is this a good amount of detail?
+
+        Convert the restricted-syntax AST to a three-address control-flow graph.
+        To do this, first,
+        add an empty basic block to the CFG and mark it as current.
+        Next, for each statement in the restricted AST's function body,
+        process the statement.
+        Statements can either be for-loops, if-statements, or assignments.
+        Rules for processing each kind of statement are given below:
+        \begin{enumerate}
+          \item
+                \textbf{For-loops}:
+                Create new basic blocks for
+                the loop condition
+                (the condition block),
+                the loop body
+                (the body block),
+                and the code after the loop
+                (the after-block).
+                Insert a jump from the end of the current block to the condition block.
+                Then, mark the condition block as the current block.
+                Insert a for-instruction at the end of the current block with the loop counter variable and bounds from the AST.
+                Next, add an edge from the current block to the after-block labeled ``FALSE'' and
+                an edge from the current block to the
+                body block labeled ``TRUE''.
+                Then, set the body block to be the current block
+                and process all statements in the AST's loop body.
+                Finally, insert a jump to the condition block and set the after-block as current.
+          \item
+                \textbf{If-statements}:
+                Create new basic blocks for
+                the ``then'' statements of the if-statement
+                (the then-block),
+                the ``else'' statements of the if-statement
+                (the else-block),
+                and the code after the if-statement
+                (the after-block).
+                At the end of the current block,
+                insert a conditional jump to the then-block or else-block
+                depending on the if-statement condition in the AST.
+                Next,
+                mark the then-block as current,
+                process all then-statements,
+                and add a jump to the after-block.
+                Similarly,
+                mark the else-block as current,
+                process all else-statements,
+                and add a jump to the after-block.
+                Finally, set the after-block to be the current block,
+                and give it a ``merge condition'' property equal to the condition of the if-statement.
+          \item
+                \textbf{Assignments}:
+                In the restricted-syntax AST,
+                the left-hand side of assignments
+                can be a variable or an array subscript.
+                If it is an array subscript such as \texttt{A[i] = x},
+                change the statement to \texttt{A = Update(A, i, x)}.
+                If the statement is not already three-address code,
+                for each sub-expression in the right-hand side of the assignment,
+                insert an assignment to a temporary variable.
+        \end{enumerate}
   \item

### 2022-04-08 overleaf/paper_SIMD.tex
Add LaTeX \benjamin command

@@
         \textbf{3-address CFG conversion}:
-        TODO: Is this a good amount of detail?
+        \benjamin{TODO: Is this a good amount of detail?}
 
@@
 
+Once the compiler converts the code to SSA,
+it transforms $\phi$-nodes that correspond to if-statements into MUX nodes.
+From the 3-address CFG conversion step,
+$\phi$-nodes corresponding to if-statements will be in a basic block
+with the ``merge condition'' property.
+For example, if $X_3 = \phi(X_1,X_2)$ is in a block with merge condition $C$,
+the compiler transforms it into $X_3 = \mathit{MUX}(C, X_1, X_2)$.
+Next, the compiler runs the dead code elimination algorithm from Cytron's SSA paper.
+Then, the control-flow graph is \emph{linearized} into MPC source,
+which has loops but no if-statements.
+\benjamin{TODO: Add more here}
+
 \ana{I think we need to add more here. Conversion from SSA to MPC Source is not so trivial.}

### 2022-04-08 compiler/paper_benchmarks.py
add continue_on_error

@@ -191,18 +191,24 @@ def run_paper_benchmarks(filename):
 
         compile = True
         i = 0
+        non_vec_failed = False
         for args in all_args:
             log.info("\n{} - arguments: {}".format(test_case_dir.name, args.args));
 
             gmw_p0 = gmw_p1 = None
-            if( i < non_vec_up_to):
+            if (i < non_vec_up_to) and not non_vec_failed:
                 log.info("Running GMW Non Vectorized {} {}".format(test_case_dir.name, args.label));           
                 gmw_p0, gmw_p1 = run_benchmark(
-                    test_case_dir.name, test_case_dir.path, GMW_PROTOCOL, False, None, args.args, compile
+                    test_case_dir.name, test_case_dir.path, GMW_PROTOCOL, False, None, args.args, compile,
+                    continue_on_error=True
                 )
-                log.info("GMW Non Vectorized output is {}".format(gmw_p0.output.strip()))
-                assert gmw_p0.output.strip() == gmw_p1.output.strip(), \
-                    (gmw_p0.output.strip(), gmw_p1.output.strip())
+                if gmw_p0 is not None and gmw_p1 is not None:
+                    log.info("GMW Non Vectorized output is {}".format(gmw_p0.output.strip()))
+                    assert gmw_p0.output.strip() == gmw_p1.output.strip(), \
+                        (gmw_p0.output.strip(), gmw_p1.output.strip())
+                else:
+                    log.info("GMW Non Vectorized FAILED! Will not run Non-Vectorized from now.")
+                    non_vec_failed = True
             
             log.info("Running GMW Vectorized {} {}".format(test_case_dir.name, args.label));
             gmw_vec_p0, gmw_vec_p1 = run_benchmark(
@@ -213,14 +219,19 @@ def run_paper_benchmarks(filename):
                 (gmw_vec_p0.output.strip(), gmw_vec_p1.output.strip())
 
             bmr_p0 = bmr_p1 = None
-            if( i < non_vec_up_to):
+            if (i < non_vec_up_to) and not non_vec_failed:
                 log.info("Running BMR Non Vectorized {} {}".format(test_case_dir.name, args.label));
                 bmr_p0, bmr_p1 = run_benchmark(
-                    test_case_dir.name, test_case_dir.path, BMR_PROTOCOL, False, None, args.args, compile
[…]

### 2022-04-08 compiler/tests/benchmark.py

@@ -37,7 +37,7 @@ class BenchmarkOutput:
 
 def run_benchmark(
     benchmark_name: str, benchmark_path: str, protocol: str, vectorized=True, timeout=600,
-    cmd_args = [], compile=True
+    cmd_args = [], compile=True, continue_on_error = False
 ) -> tuple[BenchmarkOutput, BenchmarkOutput]:
     input_fname = os.path.join(benchmark_path, "input.py")
 
@@ -117,6 +117,10 @@ def run_benchmark(
             with open(os.path.join(party0_dir, "stderr"), "w") as f:
                 f.write(party0_stderr)
 
+            if(party0.returncode != 0 and party1.returncode != 0 and continue_on_error):
+                return (
+                    None, None
+                )
             assert party1.stdout is not None
             assert party1.stderr is not None
             party1_stdout_raw = party1.stdout.read()

### 2022-04-08 compiler/paper_benchmarks.py
bug fixes

@@ -106,13 +106,62 @@ def get_inner_product_inputs()-> tuple[list[InputArgs], int]:
         all_args.append(InputArgs(label, args))
     return (all_args, non_vec_up_to)
 
+def get_convex_hull_inputs():
+    all_args = []
+    non_vec_up_to = 100
+    for N in [4, 8, 16, 32, 64, 128, 256, 512, 1024, 4096]:
+        args = [
+        "--N", "{}".format(N),
+        ]
+        X_coords = get_rand_ints(N)
+        Y_coords = get_rand_ints(N)
+        result_X = get_rand_ints(N)
+        result_Y = get_rand_ints(N)
+        args.append("--X_coords")
+        args.extend(list(map(str, X_coords)))
+        args.append("--Y_coords")
+        args.extend(list(map(str, Y_coords)))
+        args.append("--result_X")
+        args.extend(list(map(str, result_X)))
+        args.append("--result_Y")
+        args.extend(list(map(str, result_Y)))
+        label = "N: {}".format(N)
+        all_args.append(InputArgs(label, args))
+    return (all_args, non_vec_up_to)
+def get_cryptonets_convolution_naive_inputs():
+    all_args = []
+    non_vec_up_to = 100
+    for N in [4, 8, 16, 32, 64, 128, 256, 512, 1024, 4096]:
+        args = [
+        "--N", "{}".format(N),
+        ]
+        X_coords = get_rand_ints(N)
+        Y_coords = get_rand_ints(N)
+        result_X = get_rand_ints(N)
+        result_Y = get_rand_ints(N)
+        args.append("--X_coords")
+        args.extend(list(map(str, X_coords)))
+        args.append("--Y_coords")
+        args.extend(list(map(str, Y_coords)))
+        args.append("--result_X")
+        args.extend(list(map(str, result_X)))
+        args.append("--result_Y")
+        args.extend(list(map(str, result_Y)))
+        label = "N: {}".format(N)
+        all_args.append(InputArgs(label, args))
+    return (all_args, non_vec_up_to)
+
 def get_inputs(name: str) -> tuple[list[InputArgs], int]:
-    # if name == "biometric":
-    #      return get_biometric_inputs()
-    # if name == "psi":
-    #     return get_psi_inputs()
+    if name == "biometric" or name == "biometric_fast":
+         return get_biometric_inputs()
+    if name == "psi":
+        return get_psi_inputs()
     if name == "inner_product":
[…]

### 2022-04-12 compiler/compiler/motion_backend/low_level_rendering.py
Make vectorization and motion transformation more robust

@@ -44,6 +44,7 @@ class RenderContext:
     as_motion_input: bool = False
     int_type: str = "std::uint32_t"
     var_mappings: dict[Var, str] = dc.field(default_factory=dict)
+    enclosing_loops: list[For] = dc.field(default_factory=list)
 
 
 def render_type(var_type: VarType, plaintext: Optional[bool] = None) -> str:
@@ -102,12 +103,12 @@ def render_param(param: Parameter, type_env: TypeEnv) -> str:
 
 
 def render_stmt(
-    stmt: Union[Assign, For, Return], type_env: TypeEnv, ran_vectorization: bool
+    stmt: Union[Assign, For, Return], type_env: TypeEnv, ran_vectorization: bool, enclosing_loops=[]
 ) -> str:
     if isinstance(stmt, Assign):
         # If we're assigning to a vectorized value, use a specialized function for this.
         if isinstance(stmt.lhs, VectorizedAccess):
-            ctx = RenderContext(type_env, plaintext=False)
+            ctx = RenderContext(type_env, plaintext=False, enclosing_loops=enclosing_loops)
             val_expr = render_expr(stmt.rhs, ctx)
 
             # If this isn't a true vectorized access, just subscript normally
@@ -147,33 +148,33 @@ def render_stmt(
         if isinstance(stmt.rhs, Update):
             shared_assignment = (
                 (
-                    render_expr(stmt.rhs, RenderContext(type_env, plaintext=False))
+                    render_expr(stmt.rhs, RenderContext(type_env, plaintext=False, enclosing_loops=enclosing_loops))
                     + ";\n"
                 )
-                + render_expr(stmt.lhs, RenderContext(type_env, plaintext=False))
+                + render_expr(stmt.lhs, RenderContext(type_env, plaintext=False, enclosing_loops=enclosing_loops))
                 + " = "
-                + render_expr(stmt.rhs.array, RenderContext(type_env, plaintext=False))
+                + render_expr(stmt.rhs.array, RenderContext(type_env, plaintext=False, enclosing_loops=enclosing_loops))
                 + ";"
             )
             plaintext_assignment = (
-                (render_expr(stmt.rhs, RenderContext(type_env, plaintext=True)) + ";\n")
-                + render_expr(stmt.lhs, RenderContext(type_env, plaintext=True))
+                (render_expr(stmt.rhs, RenderContext(type_env, plaintext=True, enclosing_loops=enclosing_loops)) + ";\n")
[…]

### 2022-04-12 compiler/compiler/vectorize.py

@@ -1185,6 +1185,11 @@ def _basic_vectorization_phase_2(
                         setattr(stmt, field.name, val)
                     unvectorize_loop_dim(val)
 
+                # In case the left side of this assignment was updated above, update the right side
+                if isinstance(stmt, llc.Assign) and isinstance(stmt.lhs, llc.VectorizedAccess) and isinstance(stmt.rhs, llc.LiftExpr):
+                    new_dims = tuple((var, dim_size) for var, dim_size, vectorized in zip(stmt.lhs.idx_vars, stmt.lhs.dim_sizes, stmt.lhs.vectorized_dims) if vectorized)
+                    stmt.rhs.dims = new_dims
+
         for orig_var, lifted in lifted_vars.items():
             monolithic_for = util.replace_pattern(monolithic_for, orig_var, lifted)
         monolithic_for = util.replace_pattern(

### 2022-04-12 compiler/tests/stages/mnist_relu/input.py

@@ -1,1 +1,0 @@
-../../../../benchmarks/mnist_relu.py

### 2022-04-12 compiler/tests/stages/mnist_relu/restricted_ast.py

@@ -1,8 +1,0 @@
-def mnist_decomposed_relu(input: shared[list[int; ?]], OUTPUT_res: shared[list[int; ?]], len_outer: plaintext[int], len_inner: plaintext[int]) -> shared[list[int; ?]]:
-    for i: plaintext[int] in range(0, len_outer):
-        for j: plaintext[int] in range(0, len_inner):
-            val = 0
-            if (input[((i * len_inner) + j)] > 0):
-                val = input[((i * len_inner) + j)]
-            OUTPUT_res[((i * len_inner) + j)] = val
-    return OUTPUT_res

### 2022-04-12 benchmarks/mnist_relu.py
Enable mnist_relu

@@ -1,25 +1,28 @@
 from UTIL import shared
 
-# input is a 2-d matrix of shared integer values and 
+# input is a 2-d matrix of shared integer values and
 # OUTPUT_res is the result 2-d matrix of non-negative values
 # It turns all negative values in input matrix into 0s
 
 # requires: len(input)==len(OUTPUT_res)==len_outer*len_inner
 # OUTPUT_res is array of 0's
-def mnist_decomposed_relu(
-    input: shared[list[int]], OUTPUT_res: shared[list[int]], len_outer: int, len_inner: int
+def mnist_relu(
+    input: shared[list[int]],
+    OUTPUT_res: shared[list[int]],
+    len_outer: int,
+    len_inner: int,
 ) -> shared[list[int]]:
-    for i in range(len_outer): 
+    for i in range(len_outer):
         for j in range(len_inner):
-            val = 0
-            if input[i*len_inner+j] > 0:
-              val = input[i*len_inner+j]
-            OUTPUT_res[i*len_inner+j] = val
+            val = 1
+            if input[i * len_inner + j] > 1:
+                val = input[i * len_inner + j]
+            OUTPUT_res[i * len_inner + j] = val
     return OUTPUT_res
+
 
 len_inner = 10
 len_outer = 20
-input = [i if i%2 == 0 else -2*i for i in range(len_inner*len_outer)]
-OUTPUT_res = [0 for i in range(len_inner*len_outer)]
-print(mnist_decomposed_relu(input, OUTPUT_res, len_outer, len_inner))
-
+input = [i if i % 2 == 0 else 0 for i in range(len_inner * len_outer)]
+OUTPUT_res = [0 for i in range(len_inner * len_outer)]
+print(mnist_relu(input, OUTPUT_res, len_outer, len_inner))

### 2022-04-12 compiler/compiler/vectorize.py

@@ -479,10 +479,15 @@ def _basic_vectorization_phase_1(
                 return var
             elif edge_kind is EdgeKind.OUTER_TO_INNER:
                 var_prime = tmp_var_gen.get()
-                dims = tuple(
-                    (loop.counter, loop.bound_high)
-                    for loop in dep_graph.enclosing_loops[stmt]
-                )
+                if lhs_dims is not None and len(lhs_dims) > len(dep_graph.enclosing_loops[stmt]):
+                    dims = tuple(
+                        (Var("_"), dim_size) for dim_size in lhs_dims
+                    )
+                else:
+                    dims = tuple(
+                        (loop.counter, loop.bound_high)
+                        for loop in dep_graph.enclosing_loops[stmt]
+                    )
 
                 above_loop_result.append(
                     llc.Assign(

### 2022-04-12 compiler/tests/context.py

@@ -15,7 +15,6 @@ SKIPPED_TESTS = [
     # TODO: figure out why these tests are failing:
     "chapterfour_figure_12",
     "longest_odd_10",
-    "mnist_relu",
     # The following tests cause issues since they don't loop from 0
     "longest_1s",
     "longest_even_0",

### 2022-04-12 compiler/tests/stages/mnist_relu/input.py

@@ -1,0 +1,1 @@
+../../../../benchmarks/mnist_relu.py

### 2022-04-12 compiler/tests/stages/mnist_relu/restricted_ast.py

@@ -1,0 +1,8 @@
+def mnist_relu(input: shared[list[int; ?]], OUTPUT_res: shared[list[int; ?]], len_outer: plaintext[int], len_inner: plaintext[int]) -> shared[list[int; ?]]:
+    for i: plaintext[int] in range(0, len_outer):
+        for j: plaintext[int] in range(0, len_inner):
+            val = 1
+            if (input[((i * len_inner) + j)] > 1):
+                val = input[((i * len_inner) + j)]
+            OUTPUT_res[((i * len_inner) + j)] = val
+    return OUTPUT_res

### 2022-04-12 compiler/compiler/motion_backend/low_level_rendering.py
Run black on compiler directory

@@ -103,12 +103,17 @@ def render_param(param: Parameter, type_env: TypeEnv) -> str:
 
 
 def render_stmt(
-    stmt: Union[Assign, For, Return], type_env: TypeEnv, ran_vectorization: bool, enclosing_loops=[]
+    stmt: Union[Assign, For, Return],
+    type_env: TypeEnv,
+    ran_vectorization: bool,
+    enclosing_loops=[],
 ) -> str:
     if isinstance(stmt, Assign):
         # If we're assigning to a vectorized value, use a specialized function for this.
         if isinstance(stmt.lhs, VectorizedAccess):
-            ctx = RenderContext(type_env, plaintext=False, enclosing_loops=enclosing_loops)
+            ctx = RenderContext(
+                type_env, plaintext=False, enclosing_loops=enclosing_loops
+            )
             val_expr = render_expr(stmt.rhs, ctx)
 
             # If this isn't a true vectorized access, just subscript normally
@@ -148,33 +153,86 @@ def render_stmt(
         if isinstance(stmt.rhs, Update):
             shared_assignment = (
                 (
-                    render_expr(stmt.rhs, RenderContext(type_env, plaintext=False, enclosing_loops=enclosing_loops))
+                    render_expr(
+                        stmt.rhs,
+                        RenderContext(
+                            type_env, plaintext=False, enclosing_loops=enclosing_loops
+                        ),
+                    )
                     + ";\n"
                 )
-                + render_expr(stmt.lhs, RenderContext(type_env, plaintext=False, enclosing_loops=enclosing_loops))
+                + render_expr(
+                    stmt.lhs,
+                    RenderContext(
+                        type_env, plaintext=False, enclosing_loops=enclosing_loops
+                    ),
+                )
                 + " = "
-                + render_expr(stmt.rhs.array, RenderContext(type_env, plaintext=False, enclosing_loops=enclosing_loops))
+                + render_expr(
+                    stmt.rhs.array,
+                    RenderContext(
+                        type_env, plaintext=False, enclosing_loops=enclosing_loops
+                    ),
+                )
                 + ";"
             )
             plaintext_assignment = (
-                (render_expr(stmt.rhs, RenderContext(type_env, plaintext=True, enclosing_loops=enclosing_loops)) + ";\n")
-                + render_expr(stmt.lhs, RenderContext(type_env, plaintext=True, enclosing_loops=enclosing_loops))
+                (
+                    render_expr(
+                        stmt.rhs,
+                        RenderContext(
[…]

### 2022-04-12 compiler/compiler/vectorize.py

@@ -479,10 +479,10 @@ def _basic_vectorization_phase_1(
                 return var
             elif edge_kind is EdgeKind.OUTER_TO_INNER:
                 var_prime = tmp_var_gen.get()
-                if lhs_dims is not None and len(lhs_dims) > len(dep_graph.enclosing_loops[stmt]):
-                    dims = tuple(
-                        (Var("_"), dim_size) for dim_size in lhs_dims
-                    )
+                if lhs_dims is not None and len(lhs_dims) > len(
+                    dep_graph.enclosing_loops[stmt]
+                ):
+                    dims = tuple((Var("_"), dim_size) for dim_size in lhs_dims)
                 else:
                     dims = tuple(
                         (loop.counter, loop.bound_high)
@@ -1191,8 +1191,20 @@ def _basic_vectorization_phase_2(
                     unvectorize_loop_dim(val)
 
                 # In case the left side of this assignment was updated above, update the right side
-                if isinstance(stmt, llc.Assign) and isinstance(stmt.lhs, llc.VectorizedAccess) and isinstance(stmt.rhs, llc.LiftExpr):
-                    new_dims = tuple((var, dim_size) for var, dim_size, vectorized in zip(stmt.lhs.idx_vars, stmt.lhs.dim_sizes, stmt.lhs.vectorized_dims) if vectorized)
+                if (
+                    isinstance(stmt, llc.Assign)
+                    and isinstance(stmt.lhs, llc.VectorizedAccess)
+                    and isinstance(stmt.rhs, llc.LiftExpr)
+                ):
+                    new_dims = tuple(
+                        (var, dim_size)
+                        for var, dim_size, vectorized in zip(
+                            stmt.lhs.idx_vars,
+                            stmt.lhs.dim_sizes,
+                            stmt.lhs.vectorized_dims,
+                        )
+                        if vectorized
+                    )
                     stmt.rhs.dims = new_dims
 
         for orig_var, lifted in lifted_vars.items():

### 2022-04-12 overleaf/paper_SIMD.tex
Small math mode paper fix

@@
 MPC-source after Cytron's SSA may result in statements $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$, which imply that any index
-of A can be written at this point and therefore there is a loop-carried dependency. In some cases the MUX can be refined to just a single
+of $A$ can be written at this point and therefore there is a loop-carried dependency. In some cases the MUX can be refined to just a single
 index or a pair of indices, e.g., $\texttt{A}_j[i] = \emph{MUX}(c, \texttt{A}_k[i], \texttt{A}_l[i])$.

### 2022-04-13 overleaf/paper_SIMD.tex
Add to paper section 4.2 (MUX and Pseudo PHI)

@@
 
-\ana{Benjamin?}
-
 Once the compiler converts the code to SSA,
@@
 Next, the compiler runs the dead code elimination algorithm from Cytron's SSA paper.
+
 Then, the control-flow graph is \emph{linearized} into MPC source,
 which has loops but no if-statements.
-\benjamin{TODO: Add more here}
-
-\ana{I think we need to add more here. Conversion from SSA to MPC Source is not so trivial.}
-
+This means that both branches of all if-statements are executed,
+and the MUX nodes determine whether to use results from the then-block or from the else-block.
+The compiler linearizes the control-flow graph with a variation of breadth-first search.
+Blocks with the ``merge condition'' property are only considered
+the second time they are visited,
+since that will be after both branches of the if-statement are visited.
+Each time the compiler visits a block,
+it adds the block's statements to the MPC source.
+If the block ends in a for-instruction,
+the compiler recursively converts the body and code after the loop to MPC source
+and adds the for-loop and code after the loop to the main MPC source.
+If the block does not end in a for-instruction,
+the compiler recursively converts all successor branches to MPC source and
+appends these to the main MPC source.
+
+Now, the remaining $\phi$-nodes in MPC source are \emph{pseudo} $\phi$-nodes.
 A pseudo $\phi$-node $X_1 = \phi(X_0,X_2)$ in a loop header is evaluated during circuit generation.

### 2022-04-13 compiler/compiler/restricted_ast_to_tac_cfg.py
Fix Phi node argument ordering bug

@@ -17,6 +17,7 @@ class _CFGBuilder:
     _entry_block: tac_cfg.Block
     _current_block: tac_cfg.Block
     _tmp_name_counter: int = 0
+    _edge_id_counter: int = 0
 
     def __init__(self):
         self._cfg = networkx.DiGraph()
@@ -47,25 +48,22 @@ class _CFGBuilder:
         assert self._current_block.terminator is None
         self._current_block.terminator = terminator
 
-    def add_jump(self, target_block: tac_cfg.Block):
-        self._add_terminator(tac_cfg.Jump())
+    def _add_branch(self, target_block: tac_cfg.Block, branch_kind: tac_cfg.BranchKind):
+        self._edge_id_counter += 1
         self._cfg.add_edge(
             u_of_edge=self._current_block,
             v_of_edge=target_block,
-            label=tac_cfg.BranchKind.UNCONDITIONAL,
+            label=branch_kind,
+            ident=self._edge_id_counter,
         )
 
+    def add_jump(self, target_block: tac_cfg.Block):
+        self._add_terminator(tac_cfg.Jump())
+        self._add_branch(target_block, tac_cfg.BranchKind.UNCONDITIONAL)
+
     def _add_conditional_jump_edges(self, false_block, true_block):
-        self._cfg.add_edge(
-            u_of_edge=self._current_block,
-            v_of_edge=false_block,
-            label=tac_cfg.BranchKind.FALSE,
-        )
-        self._cfg.add_edge(
-            u_of_edge=self._current_block,
-            v_of_edge=true_block,
-            label=tac_cfg.BranchKind.TRUE,
-        )
+        self._add_branch(false_block, tac_cfg.BranchKind.FALSE)
+        self._add_branch(true_block, tac_cfg.BranchKind.TRUE)
 
     def add_conditional_jump(
         self,

### 2022-04-13 compiler/compiler/util.py
Enable longest_odd_10 benchmark

@@ -1,5 +1,6 @@
 import dataclasses as dc
-from typing import NoReturn, cast, TypeVar
+from typing import NoReturn, cast, TypeVar, Callable
+import networkx
 
 
 # https://github.com/python/mypy/issues/5818#issue-372451340
@@ -38,3 +39,18 @@ def replace_pattern(expr: E, pattern: P, replacement: P, include_return=True) ->
             for field in dc.fields(expr)
         }
         return dc.replace(expr, **new_params)
+
+
+def partially_ordered_sort(elems: list[E], ordering: Callable[[E, E], int]) -> list[E]:
+    graph = networkx.DiGraph()
+    for elem in elems:
+        graph.add_node(elem)
+
+    for i, e1 in enumerate(elems):
+        for e2 in elems[i + 1 :]:
+            if ordering(e1, e2) < 0:
+                graph.add_edge(e1, e2)
+            elif ordering(e2, e1) > 0:
+                graph.add_edge(e2, e1)
+
+    return list(networkx.topological_sort(graph))

### 2022-04-13 compiler/compiler/vectorize.py

@@ -1310,13 +1310,19 @@ def _basic_vectorization_phase_2(
             assert_never(stmt)
 
     def ordering(
-        s1: Union[llc.Statement, list[llc.Statement]],
-        s2: Union[llc.Statement, list[llc.Statement]],
+        s1: Union[llc.Statement, tuple[llc.Statement]],
+        s2: Union[llc.Statement, tuple[llc.Statement]],
     ) -> int:
+        if isinstance(s1, llc.For):
+            s1 = tuple(s1.body)
+
+        if isinstance(s2, llc.For):
+            s2 = tuple(s2.body)
+
         # If s1 is a closure
-        if isinstance(s1, list):
+        if isinstance(s1, tuple):
             # If s2 is a closure
-            if isinstance(s2, list):
+            if isinstance(s2, tuple):
                 orderings = set(ordering(ss1, s2) for ss1 in s1)
                 orderings.discard(0)
                 if len(orderings) == 2:
@@ -1341,7 +1347,7 @@ def _basic_vectorization_phase_2(
         # If s1 is a statement
         else:
             # If s2 is a closure
-            if isinstance(s2, list):
+            if isinstance(s2, tuple):
                 # If there is a dependency from s1 to s2, s1 must be placed first
                 if any(dep_graph.has_same_level_path(s1, ss2) for ss2 in s2):
                     return -1
@@ -1364,14 +1370,14 @@ def _basic_vectorization_phase_2(
                     return 0
 
     stmts_and_closures: list[
-        Union[llc.Statement, list[llc.Statement]]
+        Union[llc.Statement, tuple[llc.Statement]]
     ] = vectorizeable + [
-        closure for _, closure in closures  # type: ignore
+        tuple(closure) for _, closure in closures  # type: ignore
     ]  # type: ignore
 
-    stmts_and_closures = sorted(
-        stmts_and_closures, key=functools.cmp_to_key(ordering)
-    )  # Sort by index to retain def-use order
+    stmts_and_closures = util.partially_ordered_sort(
+        stmts_and_closures, ordering
+    )  # Sort to retain def-use order
 
     # Generate output
     output: list[llc.Statement] = []
@@ -1406,8 +1412,8 @@ def _basic_vectorization_phase_2(
     phi_renames = {}
 
     for i in range(len(stmts_and_closures)):
-        if isinstance(stmts_and_closures[i], list):
-            closure = cast(list, stmts_and_closures[i])
[…]

### 2022-04-13 compiler/tests/context.py

@@ -14,7 +14,6 @@ SKIPPED_TESTS = [
     "convex_hull",
     # TODO: figure out why these tests are failing:
     "chapterfour_figure_12",
-    "longest_odd_10",
     # The following tests cause issues since they don't loop from 0
     "longest_1s",
     "longest_even_0",

### 2022-04-13 compiler/tests/context.py
Enable chapterfour_figure_12 benchmark

@@ -12,8 +12,6 @@ SKIPPED_TESTS = [
     "biometric_vectorized",
     # These benchmarks fail because they involve appending to arrays
     "convex_hull",
-    # TODO: figure out why these tests are failing:
-    "chapterfour_figure_12",
     # The following tests cause issues since they don't loop from 0
     "longest_1s",
     "longest_even_0",

### 2022-04-13 compiler/tests/stages/convex_hull/restricted_ast.py
Enable convex hull benchmark

@@ -1,6 +1,4 @@
-def convex_hull(X_coords: shared[list[int]], Y_coords: shared[list[int]], N: plaintext[int]) -> tuple[shared[list[int]], shared[list[int]]]:
-    hull_X = []
-    hull_Y = []
+def convex_hull(X_coords: shared[list[int; ?]], Y_coords: shared[list[int; ?]], N: plaintext[int], result_X: shared[list[int; ?]], result_Y: shared[list[int; ?]]) -> tuple[shared[list[int; ?]], shared[list[int; ?]]]:
     for i: plaintext[int] in range(0, N):
         is_hull = True
         p1_X = X_coords[i]
@@ -11,7 +9,11 @@ def convex_hull(X_coords: shared[list[int]], Y_coords: shared[list[int]], N: pla
                 p2_Y = Y_coords[j]
                 if not ((p1_X <= p2_X) or (p1_Y >= p2_Y)):
                     is_hull = False
+        val_X = result_X[i]
+        val_Y = result_Y[i]
         if is_hull:
-            hull_X = (hull_X + [p1_X])
-            hull_Y = (hull_Y + [p1_Y])
-    return (hull_X, hull_Y)
+            val_X = p1_X
+            val_Y = p1_Y
+        result_X[i] = val_X
+        result_Y[i] = val_Y
+    return (result_X, result_Y)

### 2022-04-14 overleaf/paper_SIMD.tex
Added Intro.

@@
 \usepackage[american]{babel}
-\usepackage[backend=biber,style=ACM-Reference-Format]{biblatex}
-\addbibresource{./library.bib}
-\addbibresource{./cryptobib/abbrev2.bib}
-\addbibresource{./cryptobib/crypto_crossref.bib}
+%\usepackage[backend=biber,style=ACM-Reference-Format]{biblatex}
+%\addbibresource{./library.bib}
+%\addbibresource{./cryptobib/abbrev2.bib}
+%\addbibresource{./cryptobib/crypto_crossref.bib}
 
@@
 \email{shermb@rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Lindsey Kennard}
-\email{kennal@rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Ana L. Milanova}
-\email{milanova@cs.rpi.edu}
 \affiliation{%
@@
 %    \postcode{43017-6221}
+}
+
+\author{Lindsey Kennard}
+\email{kennal@rpi.edu}
+\affiliation{%
+    \institution{Rensselaer Polytechnic Institute}
+    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
+    \city{Troy}
+    \state{New York}
+    %\country{Iceland}
+}
+
+\author{Ana L. Milanova}
+\email{milanova@cs.rpi.edu}
+\affiliation{%
+    \institution{Rensselaer Polytechnic Institute}
+    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
+    \city{Troy}
+    \state{New York}
+    %\country{Iceland}
 }
@@
 
+Multi-party computation (MPC) allows $N$ parties $p_1,\dots, p_N$ to perform a computation on their private inputs securely.
+Informally, security means that the secure computation protocol computes the correct output (correctness) and it does not leak any information about the individual party inputs, other than what can be deduced from the output (privacy).
+
+% The first results from the early 80's~\cite{FOCS:Yao82b,STOC:GolMicWig87,STOC:BenGolWig88,STOC:ChaCreDam88} introduced the problem and formally defined its security, and investigate feasibility in different models---e.g., passive or malicious adversaries, unconditional or computational security, with or without a broadcast channel.
+This spawned a rich theoretical literature of beautiful works improving and extending in various manners the original ideas.
+
+MPC theory dates back to the early 1980-ies~\cite{FOCS:Yao82b,STOC:GolMicWig87,STOC:BenGolWig88,STOC:ChaCreDam88}. 
+Long the realm of theoretical cryptography, MPC has seen significant advances in programming technology in recent years. 
+These advances bring MPC closer to practice and wider applicability --- 
+MPC technology has been employed in real-world scenarios such as auctions~\cite{FC:BCDGJK09}, biometric identification~\cite{ESORICS:BlaGas11}, and privacy-preserving machine learning~\cite{SP:MohZha17,CCS:MohRin18}.
+The goal is to bring the technology to a level where programmers can write \emph{secure} and \emph{efficient} programs without 
+commanding extensive knowledge of cryptographic primitives.
+
+The problem, therefore, is to build a high-level programming language and a compiler, and there has been significant advance in 
+this space, e.g., ~\cite{CCS:BenNisPin08,ESORICS:BogLauWil08,CCS:ZhaSteBla13,Hastings:2019,Keller:2020,Acay:2021,Braun:2022} among other work. 
+Current research largely falls at the two ends 
+of the classical compiler: (1) work on \emph{front-end} language design and (2) work on \emph{back-end} protocol implementation. Work on language design 
+focuses on high-level constructs necessary to express multiple parties, computation by different parties, and information flow from one party 
+to another~\cite{SP:RasHamHic14, Acay:2021}. On the other end, work on protocol implementation focuses on cryptographic foundations and their efficient circuit-level
+implementation~\cite{NDSS:DemSchZoh15, CCS:ABFKLO18, Braun:2022}, e.g., implementation of operations (e.g., MUL, ADD) using different sharing protocols 
+(Boolean or Arithmetic GMW~\cite{STOC:GolMicWig87} or Yao's garbled circuits~\cite{FOCS:Yao82b}), as well as efficient share conversion from one representation 
+to another.
+
+Earlier compilers did both back-end and front-end translation without a specific focus on either, as their aim was to demonstrate applicability of MPC on real-world
+programming problems. As the field advanced, works have focused more closely on front-end language design (e.g., Wysteria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021}) 
+or back-end ``circuit-level'' design and implementation (e.g., MOTION~\cite{Braun:2022}).
+
+In this work we focus on an intermediate language and what we call \emph{circuit-independent optimizations}, in a close analogy to \emph{machine-independent} optimizations
+in the classical compiler. The following figure summarizes our key idea:
+
+{\begin{center}
+\includegraphics[width=0.8\linewidth]{figs_paper_SIMD/focus.png}
+\end{center}
+}
+
+We emphasize the MPC Source intermediate representation and optimization over MPC Source. As in classical compilers, we envision different front ends (e.g., Wysteria,
+our front end IMP-MPC) compiling into MPC Source. MPC Source is particularly suitable for optimizations such as protocol mixing~\cite{Ishaq:2019, Fang:2022} and SIMD-vectorization, 
+which takes advantage of amortization at the circuit level. The MPC Source IR exposes the linear structure of MPC programs, which simplifies program analysis; this is
+in contrast to source, which has if-then-else constructs. In the same time, MPC Source is sufficiently ``high-level''  to support analyses and optimizations that take into account 
+control and data flow in a specific program. MPC Source is small in size and analysis is tractable, as opposed to analysis over an unrolled circuit~\cite{Ishaq:2019}. Again as
+in classical compilers, we envision translation of MPC Source (optimized or unoptimized) into MOTION, SPDZ, or other back-end code.
+
+%In contrast, the high-level program contains if-then-else and array writes, or the low-level circuit . 
+%The compilation problem, i.e., translation from a high-level language to circuits, is difficult and therefore, breaking the problem into sub-problems can  tractable.
+
+\subsection{Our contribution} In this paper, we develop a compiler framework that takes a Python-like routine and produces MOTION code: we describe (a) the IMP-MPC language, its syntax and semantic restrictions, (b) translation from IMP-MPC into MPC Source, (c) a specific circuit-independent optimization: novel SIMD-vectorization on MPC Source, and (d) translation from MPC Source into MOTION code.
+\ana{I think we have to add more on MOTION, specifically mixing and amortization.}
+
+Our second contribution is an analytical model for cost estimation of amortized schedules. Originally, we hoped that optimal scheduling (under our model) was tractable, as the problem
+appeared simpler than the classical scheduling problem. Unfortunately, we show that optimal scheduling is NP-hard via a reduction to the Shortest Common Supersequence (SCS) problem. 
+Cost estimation is important as it drives scheduling and protocol mixing optimizations~\cite{Ishaq:2019, Fang:2022}.
+
+Our most important contribution is the implementation and evaluation of the compiler framework.
+We demonstrate expressivity of the source language by running the compiler on \ana{X} programs; these include classical MPC benchmarks such as PSI and Biometric matching, as well as kMeans, Histogram, and others.
+Vectorization leads to \ana{X}x improvement in number of gates on average, \ana{Y}x speedup in circuit generation time, \ana{Z}x speedup in the LAN setting, and \ana{Z}x speedup in the WAN setting using the Boolean GMW protocol.
+Results form BMR are similar.
+
+Our results emphasize the importance of ``circuit-independent'' optimizations --- vectorization (described in this work) and protocol mixing (tackled in previous works~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022}) are two optimizations readily available at the level of MPC Source.
+We believe that our work can lead to future work on ``circuit-independent'' compilation and optimization, bringing in new MPC optimizations and combinations of optimizations, much in the vein of standard compilers, and bring MPC programming technology closer to practice and wider applicability.
+
+\subsection{Outline}
+
+The rest of the paper is organized as follows.
+\secref{sec:overview} presents an overview of the compiler.
+\secref{sec:model} describes our model for cost estimation and argues NP-hardness of optimal scheduling and \secref{sec:compiler} describes the compiler.
+\secref{sec:experiments} presents implementation and evaluation.
+\secref{sec:related} discusses related work and \secref{sec:conclusion} concludes.
+
+\begin{comment}
 \begin{itemize}
@@
 \end{itemize}
+\end{comment}
 
@@
 
+\begin{comment}
 \subsubsection{Array MUX refinement}
@@
 
-\section{Divide-and-Conquer}
-\label{sec:div_and_conq}
-
-\ana{TODO: Now that we have broken FOR loops into smaller chunks, we can add Divide-and-conquer reasoning with Z3 and implement this additional transform.}
+%\section{Divide-and-Conquer}
+%\label{sec:div_and_conq}
+%\ana{TODO: Now that we have broken FOR loops into smaller chunks, we can add Divide-and-conquer reasoning with Z3 and implement this additional transform.}
 
@@
 
-\section{Future Work}
-\label{sec:implementation_and_benchmarks}
+\section{Related Work}
+\label{sec:related}
+
+%\section{Future Work}
+%\label{sec:implementation_and_benchmarks}
 %\input{../sections/implementation_and_benchmarks.tex}
@@
 
-\section{Conclusions}
+\section{Conclusion and Future Work}
 \label{sec:conclusion}
@@
 
-\printbibliography
+%\printbibliography
+
+\bibliographystyle{plain}
+%\bibliography{c:/trink/Doc/MyTexFiles/texrefs}
+%\bibliography{mybib,library,library1,library2,library3,library_ana}
+\bibliography{library_ana,crypto_crossref}
 

### 2022-04-26 overleaf/paper_SIMD.tex
Added draft of Related work section.

@@
 
+\paragraph{MPC languages and compilers.}
+
+Languages and compilers for secure computation have seen significant attention and advances in recent years. The early MPC compilers Fairplay~\cite{CCS:BenNisPin08}, 
+and Sharemind~\cite{ESORICS:BogLauWil08} were followed by PICCO~\cite{CCS:ZhaSteBla13}, Obliv-C~\cite{Zahur:2015}, TinyGarble~\cite{SP:SHSSK15}, 
+Wystiria~\cite{SP:RasHamHic14}, and others. A new generation of MPC compilers 
+includes SPDZ/SCALE-MAMBA/MP-SPDZ~\cite{Keller:2020} and the ABY/HyCC/MOTION~\cite{NDSS:DemSchZoh15,CCS:BDKKS18,Braun:2022} frameworks. 
+These two families are the state-of-the art and
+are actively developed. Another recent development is Viaduct, a functional language and compiler that supports a range of secure computation 
+frameworks, including MPC and ZKP. Hastings et el. present a review of compiler frameworks~\cite{Hastings:2019}.
+
+While each of these languages and compilers brings in new ideas and advances, none addresses the problem of ``circuit independent'' intermediate 
+representation and optimization. We envision a classical compiler structure: (1) a Wysteria, Viaduct, Obliv-C, or IMP Source front end, including rich type systems and 
+AST-level semantic analysis, compile into the MPC Source IR, (2) MPC Source-level optimizations take place, followed by (3) back-end compilers into circuits. 
+Our focus is at the intermediate level.
+
+%{\begin{center}
+%\includegraphics[width=0.5\linewidth]{figs/focus.png}
+%\end{center}
+%}
+
+Many works focus on the implementation of MPC protocols exposing an API to the programmer. For example, the ABY/MOTION line of 
+compiler frameworks provides a library of MPC primitives; the programmer writes MPC programs in C++ on top of the library. These back ends implement 
+different protocols and allow for mixing, but notably, they leave it to the programer to assign different protocols to different parts of the 
+computation and perform share conversion accordingly. In addition, MOTION provides SIMD primitives, which allows for efficient execution 
+of MPC operations, but again, using SIMD primitives is the responsibility of the programer. There is interest in frameworks for automatic mixing, 
+e.g., \cite{CCS:BDKKS18,Ishaq:2019,Fang:2022}.
+
+Other works, e.g., Obliv-C~\cite{Zahur:2015}, Wystiria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021} focus on higher-level language design, particularly information-flow systems that 
+restrict flow between secure and insecure parts of the program. 
+%Key ideas such as ${\sf insecure} <: {\sf secure}$ qualifier systems are similar to the ideas we 
+%present in~\Secref{sec:interprocedural} and~\Secref{sec:zkp} and we intend to draw on these work. While we will design a high-level language expanding our 
+%intraprocedural IMP-MPC syntax and semantics into the \emph{interprocedural} case, our interest and focus lie primarily in the ``machine-independent'' optimizations 
+%thrust of the proposal.
+
+%There is significant development in ZKP languages as well. ~\ana{Alex Ozdemir's talk, ZoKrates, Cairo, others} Our work focuses on ZKP via MPC and our key idea
+%is the separation of the higher-level ``backend-independent'' compilation and optimization and the lower-level compilation into circuits.
+
+\paragraph{Classical HPC compilers.}
+Automatic vectorization is a longstanding problem in high-performance computing (HPC). 
+There are thousands of works in this area reflecting over 40 years of research. We presented a vectorization
+algorithm for MPC Source, essentially extending classical loop vectorization \cite{Allen:1987}. In HPC vectorization, conditional control flow 
+presents a challenge --- one cannot estimate the cost of a schedule or vectorize branches in a straightforward manner --- in contrast to 
+MPC Source vectorization. 
+We view Karrenberg's work on Whole function vectorization~\cite{Karrenberg:2015} as most closely related to ours --- it linearizes the program and vectorizes 
+both branches of a conditional applying masking to avoid execution of the branch-not-taken code, and selection (similar to MUX) to select the correct value based on the 
+result of the condition at runtime. 
+The problem is that masking and selection, or more generally, handling control predicates~\cite{Benabderrahmane:2010,Karrenberg:2015}, 
+can lead to \emph{slowdown}. 
+%As a result, vectorizing compilers and polyhedral compilers tend to focus on straight-line code due to the unpredictable 
+%overhead conditionals may introduce. \ana{add Louis noel pochet's benchmarks. citation on vectorizing compilers???}
+
+We argue that vectorization over linear MPC Source is a different problem, one that warrants a new look, while drawing from 
+results in HPC.
+Since both branches of the conditional and the multiplexer \emph{always} execute, not only can we apply aggressive vectorization on linear code, but (perhaps more importantly) 
+we can also build analytical models that accurately predict execution time. These models in turn would drive optimizations such as vectorization, protocol mixing, and others. 
+Vectorization meshes in with those additional optimizations in non-trivial ways.
+
+Furthermore, extensions of classical loop vectorization with array writes, arbitrary indexing, including non-affine indexing, and interaction with SSA are non-trivial 
+and present novel challenges and opportunities for contribution. Polyhedral parallelization~\cite{Benabderrahmane:2010} considers a higher-level source (typically AST) 
+representation, while our work takes advantage of linear MPC Source and SSA form. The work by Karrenberg~\cite{Karrenberg:2015} is rare in that space, in the sense that it considers 
+vectorization over SSA form, which has similarities to MPC Source. We consider different array representation, notion of dependence, and reasoning 
+about dependence, which we conjecture is more suitable for MPC Source. Buscher~\cite{Buscher:2018} considers SIMD-vectorization 
+at the level of source code, which then combines with circuit-level optimizations in the TinyGarble compiler. 
+He proposes using an off-the-shelf polyhedral compiler, however, application is limited to only two routines, 
+essentially just inner product and euclidian distance; it is unclear how effective the off-the-shelf compiler is. 
+In contrast, we consider vectorization at the level of MPC Source separating ``backend-independent'' vectorization
+and circuit-level amortization (done by MOTION). We apply our compiler on a wide range of routines.
+
 %\section{Future Work}

### 2022-04-26 overleaf/paper_SIMD.tex
Changed title (feel free to suggest better ones) and affiliations.

@@
 % Do not put math or special symbols in the title.
-\title[Scheduling and Vectorization for MPC]{Scheduling and Vectorization for MPC}
+\title[Compilation and Backend-Independent Vectorization for Multi-Party Computation]{Compilation and Backend-Independent Vectorization for Multi-Party Computation}
 
@@
 \author{Muhammad Ishaq}
-\authornote{This work was done in part while the author was at RPI.}
-\email{m.ishaq@ed.ac.uk}
+%\authornote{This work was done in part while the author was at RPI.}
+\email{ishaqm@purdue.edu}
 %\orcid{1234-5678-9012}
@@
 \affiliation{%
-    \institution{University of Edinburgh}
+    \institution{Purdue University}
     %\streetaddress{P.O. Box 1212}
-    \city{Edinburgh}
-    \state{Scotland}
+    \city{West Lafayette}
+    \state{Indiana}
 %    \postcode{43017-6221}
@@
 \author{Lindsey Kennard}
-\email{kennal@rpi.edu}
+\email{fire.elemental@gmail.com}
 \affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
+    \institution{STR}
     %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
+    \city{Boston}
+    \state{Massachusets}
     %\country{Iceland}
@@
 \author{Vassilis Zikas}
-\authornote{This work was done in part while the author was visiting UCLA and supported in part by DARPA and SPAWAR under contract N66001-15-C-4065 and by a SICSA Cyber Nexus Research Exchanges grant.}
-\email{vzikas@inf.ed.ac.uk}
+%\authornote{This work was done in part while the author was visiting UCLA and supported in part by DARPA and SPAWAR under contract N66001-15-C-4065 and by a SICSA Cyber Nexus Research Exchanges grant.}
+\email{vzikas@purdue.edu}
 %\orcid{1234-5678-9012}
@@
 \affiliation{%
-    \institution{University of Edinburgh}
+    \institution{Purdue University}
     %\streetaddress{P.O. Box 1212}
-    \city{Edinburgh}
-    \state{Scotland}
+    \city{West Lafayette}
+    \state{Indiana}
     %    \postcode{43017-6221}

### 2022-04-26 compiler/paper_benchmarks.py
setup time in graphs is sum of preprocessing and setup.

@@ -353,7 +353,7 @@ def get_max_dist_between_syms_inputs():
 def get_mnist_relu_inputs()-> tuple[list[InputArgs], int]:
     all_args = []
     non_vec_up_to = 6
-    for config in [[16, 16], [16, 32], [16, 64], [16, 128], [16, 256], [16, 512], [16, 1024]]:#[2048, 2084], [4096, 4096]]:
+    for config in [[16, 16], [16, 32], [16, 64], [16, 128], [16, 256], [16, 512], [16, 1024]]:
         len_inner = config[0]
         len_outer = config[1]
         args = [
@@ -394,38 +394,38 @@ def get_psi_inputs()-> tuple[list[InputArgs], int]:
     return (all_args, non_vec_up_to)
 
 def get_inputs(name: str) -> tuple[list[InputArgs], int]:
-    # if name == "biometric" or name == "biometric_fast": # --- need to get inputs at 2048
-    #     return get_biometric_inputs()
-    # if name == "chapterfour_figure_12": # new ---
-    #     return get_chapterfour_figure_12_inputs()
-    if name == "convex_hull":# or name == "minimal_points": #convex hull is new
+    if name == "biometric" or name == "biometric_fast":
+        return get_biometric_inputs()
+    if name == "chapterfour_figure_12": # millionaire's problem, not interesting
+        return get_chapterfour_figure_12_inputs()
+    if name == "convex_hull" or name == "minimal_points":
         return get_convex_hull_inputs()
-    # if name == "count_102" or name == "longest_102":
-    #     return get_count_102_inputs()
-    # if name == "count_10s":
-    #     return get_count_10s_inputs()
-    # if name == "count_123":
-    #     return get_count_123_inputs()
-    # if name == "cryptonets_max_pooling":
-    #     return get_cryptonets_max_pooling_inputs()
-    # if name == "db_cross_join_trivial": # could only do up to 64
-    #     return get_db_cross_join_trivial_inputs()
-    # if name == "db_variance":
-    #     return get_db_variance_inputs()
-    # if name == "histogram":
-    #     return get_histogram_inputs()
-    # if name == "inner_product":
-    #     return get_inner_product_inputs()
-    # if name == "kmeans_iteration":
-    #     return get_kmeans_iteration_inputs()
-    # if name == "longest_odd_10": # need to run for 4096
-    #     return get_longest_odd_10_inputs()
[…]

### 2022-04-27 compiler/compiler/motion_backend/low_level_rendering.py
Only convert plaintext values to shares when they're needed

@@ -102,6 +102,41 @@ def render_param(param: Parameter, type_env: TypeEnv) -> str:
     )
 
 
+def collect_plaintext_conversions(expr: AssignRHS, type_env: TypeEnv) -> list[Var]:
+    if isinstance(expr, Var):
+        if type_env[expr].is_plaintext():
+            return [expr]
+        else:
+            return []
+    elif isinstance(expr, UnaryOp):
+        return collect_plaintext_conversions(expr.operand, type_env)
+    elif isinstance(expr, BinOp):
+        return collect_plaintext_conversions(
+            expr.left, type_env
+        ) + collect_plaintext_conversions(expr.right, type_env)
+    elif isinstance(expr, (List, Tuple)):
+        return [
+            var
+            for elem in expr.items
+            for var in collect_plaintext_conversions(elem, type_env)
+        ]
+    elif isinstance(expr, (Update, VectorizedUpdate)):
+        return collect_plaintext_conversions(expr.value, type_env)
+    elif isinstance(expr, Mux):
+        return (
+            collect_plaintext_conversions(expr.condition, type_env)
+            + collect_plaintext_conversions(expr.true_value, type_env)
+            + collect_plaintext_conversions(expr.false_value, type_env)
+        )
+    elif isinstance(expr, LiftExpr):
+        return collect_plaintext_conversions(expr.expr, type_env)
+    # The below expressions never require conversion to plaintext
+    elif isinstance(expr, (Constant, Subscript, VectorizedAccess, DropDim)):
+        return []
+    else:
+        assert_never(expr)
+
+
 def render_stmt(
     stmt: Union[Assign, For, Return],
     type_env: TypeEnv,
@@ -109,6 +144,37 @@ def render_stmt(
     enclosing_loops=[],
 ) -> str:
     if isinstance(stmt, Assign):
+        # Convert any plaintext assignments
+        plaintext_conversions = ""
+        if (
+            type_assign_expr(stmt.lhs, type_env, stmt, None).is_shared()
+            # lift expressions are always shared, even if they're lifting to "plaintext" arrays
+            or isinstance(stmt.rhs, LiftExpr)
+        ):
+            vars_needing_conversions = collect_plaintext_conversions(stmt.rhs, type_env)
+            plaintext_conversions = "\n".join(
+                render_expr(
+                    var,
+                    RenderContext(
+                        type_env, plaintext=False, enclosing_loops=enclosing_loops
[…]

### 2022-04-27 compiler/compiler/type_analysis.py

@@ -238,6 +238,8 @@ def type_assign_expr(
             expr_type,
             _dims=0,
             dim_sizes=[loop_bound for _idx_var, loop_bound in expr.dims],
+            # Vectorized arrays are always shared
+            visibility=VarVisibility.SHARED,
         )
 
     elif isinstance(expr, DropDim):

### 2022-04-27 compiler/compiler/vectorize.py
Fix type error

@@ -1314,10 +1314,10 @@ def _basic_vectorization_phase_2(
         s2: Union[llc.Statement, tuple[llc.Statement]],
     ) -> int:
         if isinstance(s1, llc.For):
-            s1 = tuple(s1.body)
+            s1 = tuple(s1.body) # type: ignore
 
         if isinstance(s2, llc.For):
-            s2 = tuple(s2.body)
+            s2 = tuple(s2.body) # type: ignore
 
         # If s1 is a closure
         if isinstance(s1, tuple):

### 2022-04-27 overleaf/paper_SIMD.tex

@@
 
+\ana{We need an abstract.}
+
 \end{abstract}
@@
 These advances bring MPC closer to practice and wider applicability --- 
-MPC technology has been employed in real-world scenarios such as auctions~\cite{FC:BCDGJK09}, biometric identification~\cite{ESORICS:BlaGas11}, and privacy-preserving machine learning~\cite{SP:MohZha17,CCS:MohRin18}.
+MPC technology has been employed in real-world scenarios such as auctions~\cite{FC:BCDGJK09}, biometric identification~\cite{ESORICS:BlaGas11}, 
+and privacy-preserving machine learning~\cite{SP:MohZha17,CCS:MohRin18}. 
 The goal is to bring the technology to a level where programmers can write \emph{secure} and \emph{efficient} programs without 
@@
 
-\subsection{Our contribution} In this paper, we develop a compiler framework that takes a Python-like routine and produces MOTION code: we describe (a) the IMP-MPC language, its syntax and semantic restrictions, (b) translation from IMP-MPC into MPC Source, (c) a specific circuit-independent optimization: novel SIMD-vectorization on MPC Source, and (d) translation from MPC Source into MOTION code.
-\ana{I think we have to add more on MOTION, specifically mixing and amortization.}
+\subsection{Our Contribution} In this paper, we develop a compiler framework that takes a Python-like routine and produces MOTION code: we describe (a) the IMP-MPC language, its syntax and semantic restrictions, (b) translation from IMP-MPC into MPC Source, (c) a specific circuit-independent optimization: novel SIMD-vectorization on MPC Source, and (d) translation from MPC Source into MOTION code.
+%\ana{I think we have to add more on MOTION, specifically mixing and amortization.}
+
+We focus on the MOTION framework as our back-end for several reasons.
+First, it is the state of the art in terms of performance~\cite{Braun:2022}.
+Second, it provides an API over efficient implementation for a wide variety of cryptographic operations in three different protocols ---  Arithmetic GMW, Boolean GMW, and BMR --- which allows for protocol mixing~\cite{Ishaq:2019, Fang:2022}, a known backend-independent optimization.
+Third, MOTION provides API for SIMD-level operations, which amortize cost and lead to significant improvement in memory footprint and
+throughput~\cite{NDSS:DemSchZoh15, CCS:ABFKLO18, Braun:2022}. It also enables MPC Source-level optimizations such SIMD vectorization.
 
@@
 appeared simpler than the classical scheduling problem. Unfortunately, we show that optimal scheduling is NP-hard via a reduction to the Shortest Common Supersequence (SCS) problem. 
-Cost estimation is important as it drives scheduling and protocol mixing optimizations~\cite{Ishaq:2019, Fang:2022}.
+Cost estimation is important as it drives not only vectorization but other optimizations such as protocol mixing optimizations and scheduling as well~\cite{Ishaq:2019, Fang:2022}.
 
@@
 
+\begin{table*}
+\begin{tabular}{ccc}
+\begin{minipage}{0.3\textwidth}
+{\small
+\begin{pythonn}
+def biometric(C: shared[list[int]], D: int, 
+      S: shared[list[int]], N: int) -> 
+      shared[tuple[int,int]]:
+   min_sum = MAX_INT
+   min_idx = 0  
+   for i in range(N): 
+      sum = 0
+      for j in range(D): 
+         # d = S[i,j] - C[j]
+         d = S[i * D + j] - C[j] 
+         p = d * d
+         sum = sum + p
+      if sum < min_sum:
+         min_sum = sum
+         min_idx = i
+   return (min_sum, min_idx)
+\end{pythonn}
+}
+\end{minipage}
+
+&
+
+\begin{minipage}{0.33\textwidth}
+{\tiny
+\begin{pythonn}
+min_sum!1 = MAX_INT
+min_idx!1 = 0
+for i in range(0, N):
+   min_sum!2 = PHI(min_sum!1, min_sum!4)
+   min_idx!2 = PHI(min_idx!1, min_idx!4)
+   sum!2 = 0
+   for j in range(0, D):
+      sum!3 = PHI(sum!2, sum!4)
+      d = SUB(S[((i * D) + j)],C[j]) 
+      p = MUL(d,d) 
+      sum!4 = ADD(sum!3,p) 
+   t = CMP(sum!3,min_sum!2) 
+   min_sum!3 = sum!3
+   min_idx!3 = i
+   min_sum!4 = MUX(t, min_sum!3, min_sum!2) 
+   min_idx!4 = MUX(t, min_idx!3, min_idx!2) 
+return (min_sum!2, min_idx!2)   
+\end{pythonn}
+}
+\end{minipage}
+
+&
+
+\begin{minipage}{0.33\textwidth}
+{\small
+\begin{pythonn}
+min_sum!1 = MAX_INT
+min_idx!1 = 0
+# S^ is same as S. C^ replicates C N times: 
+S^ = lift(S, ((i * D) + j), (i:N,j:D)) # S^[i,j] = S[i,j]
+C^ = lift(C, j, (i:N,j:D)) # C^[i,j] = C[j]
+
+sum!2[I] = [0,..,0]
+# computes _all_  "at once"
+d[I,J] = SUB_SIMD(S^[I,J],C^[I,J])
+p[I,J] = MUL_SIMD(d[I,J],d[I,J])
+
+for j in range(0, D):
+   # sum!2[I], sum!3[I], sum!4[I] are size-N vectors 
+   # computes N intermediate sums "at once"
+   sum!3[I] = PHI(sum!2[I], sum!4[I])       
+   sum!4[I] = ADD_SIMD(sum!3[I],p[I,j])
+
+min_idx!3[I] = [0,1,...N-1]   
+for i in range(0, N):
+   min_sum!2 = PHI(min_sum!1, min_sum!4) 
+   t[i] = CMP(sum!3[i],min_sum!2)
+   min_sum!4 = MUX(t[i], sum!3[i], min_sum!2)  
+for i in range(0, N):
+   min_idx!2 = PHI(min_idx!1, min_idx!4)    
+   min_idx!4 = MUX(t[i], min_idx!3[i], min_idx!2)
+return (min_sum!2, min_idx!2)   
+\end{pythonn}
+}
+\end{minipage}
+
+\\
+
+(a) IMP Source & (b) MPC Source & (c) Optimized MPC Source
+\end{tabular}
+\caption{Biometric Matching: From IMP Source to MPC Source to Optimized MPC Source. }
+\label{tab:source_and_MPC_source_and_optimized_MPC_source}
+\end{table*}
+
 \subsection{Source}
@@
 As a running example, consider Biometric matching, a standard MPC benchmark.
-Array \texttt{C} is the feature vector of \texttt{D} features that we wish to match and array \texttt{S} 
-is the database of \texttt{N} vectors of size \texttt{D} that we match against.
-An intuitive implementation is as follows:
-
+An intuitive (and naive) implementation is as shown in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(a).
+Array \python{C} is the feature vector of \texttt{D} features that we wish to match and \python{S} 
+is the database of \python{N} size-\python{D} vectors that we match against.
+
+\begin{comment}
 {\small
@@
 }
-
-Our compiler takes (essentially) standard IMP syntax.
-The programmer can write intuitive iterative programs as the one above.
-They annotate certain inputs and outputs as \emph{shared}.
-Here the code iterates over the entries in the database and computes
-the sum of squares of the differences of individual features. The program returns the 
-index \texttt{i} of the vector that gives the best match plus the corresponding sum of squares.
-
-Our compiler imposes the following restrictions. We note that in some cases, the restrictions 
-can be easily lifted and we plan to do so in future iterations of our compiler.
-
-\begin{enumerate}
-\item The program contains arbitrarily nested loops, however, loop bounds are fixed: \texttt{0 <= i < N}.
-A standard restriction in MPC is that the bounds must be known at circuit-generation time.
-\item Arrays are one-dimensional. N-dimensional arrays are linearized and accessed 
-in row-major order and at this point the programmer is responsible for linearization
-and access. 
-\item Array subscrpts are plaintext values.
-\item Our compiler allows for output (write) arrays, however it restricts
-write access to \emph{canonical writes} along the dimensions of the array. I.e., \texttt{A[i,j] = ...}
-where \texttt{i} and \texttt{j} loop over the two dimensions of \texttt{A} is allowed, 
-but \texttt{A[i,j+2] = ...} is not allowed. Read access is arbitrary. 
-\end{enumerate}
+\end{comment}
+
+Our compiler takes essentially standard IMP~\ana{Need citation here.} 
+syntax and imposes certain semantic restrictions. (We detail the restrictions in the following sections.)
+The programmer writes an iterative program and annotate certain inputs 
+and outputs as \emph{shared}. In the example arrays \python{C} and \python{S}
+are shared, meaning that they store shares, however, the array sizes \python{D} and
+\python{N} respectively are plaintext. The code iterates over the entries in the 
+database and computes the Euclidean distance (its square actually) of the current 
+entry \python{S[i]} and \python{C}. The program returns the index of the vector that gives 
+the best match plus the corresponding sum of squares.
 
@@
 
-The compiler generates an IR, MPC source:
+Our compiler generates an intermediate representation, MPC Source. MPC Source 
+for Biometric Matching is shown in 
+Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(b). First, MPC Source is an SSA form. 
+Second, it is linear. The conditional in lines 13-14 in IMP Source turns into the linear code in lines 12-16 in MPC Source. 
+The test turns into a CMP operation (\python{t = CMP(sum!3,min\_sum!2)}), followed by the 
+true-branch sequence, followed by the MUX operations. The first MUX operation selects the value
+of \python{min\_sum}: if \python{t} is true, then \python{min\_sum} gets the value of the second multiplexer
+argument,  \python{min\_sum!3}, otherwise it takes the value of the third argument, \python{min\_sum!2}.  
+Third, MPC Source is a special form of SSA. The SSA $\phi$-nodes at the if-then-else (lines 13-15) turn into 
+MUX operations, while the $\phi$-nodes at for loops turn into what we call pseudo PHI nodes with a straightforward semantics.
+
+\begin{comment}
 {\small
-\begin{verbatim}
+\begin{pythonn}
 1. min_sum!1 = 10000
@@
 17.!2!1 = (min_sum!2, min_index!2)   
-\end{verbatim}
-}
-
-The compiler linearizes the source turning conditionals into MUX statements. The PHI nodes
-are remnants of the SSA IR; the compiler generates code that picks the correct value 
-when producing MOTION output; the MOTION framework in turn linearizes 
-loops when it generates the circuit.
-
-We turn to our analytical model to compute the cost of this program. Assuming fixed 
+\end{pythonn}
+}
+\end{comment}
+
+%The compiler linearizes the source turning conditionals into MUX statements. The PHI nodes
+%are remnants of the SSA IR; the compiler generates code that picks the correct value 
+%when producing MOTION output; the MOTION framework in turn linearizes 
+%loops when it generates the circuit.
+
+We turn to our analytical model to compute the \emph{cost} of the program. Assuming fixed 
 cost $\beta$ for a local MPC operation (essentially just ADD) and cost $\alpha$ for a remote
 MPC operation (e.g., MUX, CMP, and remaining operations), the cost of the iterative schedule 
-will be $N*D*(2*\alpha+\beta) + N*3*\alpha$.
+will be $ND(2\alpha+\beta) + N(3\alpha$).
 
@@
 program by hand; we are the first (to the best of our knowledge) to automatically 
-transform an intuitive, iterative MPC program into an unintuitive vectorized one.
+transform a naive, iterative MPC program into an unintuitive vectorized one.
 
@@
 
-Our compiler produces the following vectorized program. (Note that this is
-still higher-level IR, Vectorized MPC Source. Our compiler turns this code into MOTION variables, 
-loops and SIMD primitives, which MOTION then uses to generate the circuit.)
-
+Our compiler produces the vectorized program shown in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(c). 
+Note that this is still our intermediate representation, Optimized MPC Source. Subsequently,
+the compiler turns this code into MOTION variables, loops and SIMD primitives, which MOTION then 
+uses to generate the circuit.
+
+The compiler determines that SUB and MUL in ``naive'' MPC Source (lines 9 and 10 in (b))
+can be fully vectorized into the SIMD SUB and MUL in optimized MPC Source (lines 9 and 10 in (c)). 
+Notation \python{p[I,J]} denotes a 2-dimensional array with fully vectorized dimensions.
+The computation of sum (line 11 in (b))
+is sequential across the $j$-dimension, but it is parallel across the $i$-dimension. 
+The loop in lines 12-16 in (c) illustrates; here \python{p[I,j]} refers to the $j$-th column in \python{p}. 
+Unfortunately, CMP and MUX remain sequential.
+
+\begin{comment}
 {\small
@@
 }
-
-In MPC compilers a vectorized operation that computing M operations "at once" costs 
-essentially the same ($\alpha$ or $\beta$) as an individual operation. We elaborate 
-on these in the following section. Thus, the vectorized program costs 
-$2*\alpha$ + $D * \beta + N*3*\alpha$. The first term in the sum corresponds to the vectorized
-subtraction and multiplication, the second term corresponds to the for loop on $j$ and the third 
-one corresponds to the remaining for loops on $i$. Clearly, $2*\alpha + D * \beta + N*3*\alpha << N*D*(2*\alpha+\beta) + N*3*\alpha$.
-Our experimental results illustrate this as well. \ana{Add numbers.}
+\end{comment}
+
+In MPC backends, executing $n$ operations ``at once'' in a single SIMD operation costs a lot less than executing those $n$ operations one by one.
+This is particularly important when there is communication (i.e., in remote), since many 1-bit values are sent at once rather than sequentially. 
+%In MPC compilers a vectorized operation that computing M operations "at once" costs 
+%essentially the same ($\alpha$ or $\beta$) as an individual operation. 
+We elaborate on the cost model in the following section but for now consider that 
+each operation has a \emph{fixed} portion (does benefits from amortization) and 
+a \emph{variable} portion (does not benefit from amortization): $\alpha = \alpha_\mathit{fix} + \alpha_\mathit{var}$. 
+This gives rise to the following formula for amortized cost: $f(n) = \alpha_\mathit{fix} + n\alpha_\mathit{var}$, 
+as opposed to unamortized cost $g(n) = n\alpha_\mathit{fix} + n\alpha_\mathit{var}$. (We extend the same reasoning to 
+$\beta$-instructions.)
+
+Thus, the fixed cost of the vectorized program amounts to
+$2\alpha_\mathit{fix}$ + $D \beta_\mathit{fix} + N(3\alpha_\mathit{fix})$. (The variable cost is the same in both the vectorized and un-vectorized programs.) 
+The first term in the sum corresponds to the vectorized
+subtraction and multiplication, the second term corresponds to the FOR loop on $j$ and the third 
+one corresponds to the remaining FOR loops on $i$.
+Clearly, $2\alpha_\mathit{fix} + D\beta_\mathit{fix} + N(3\alpha_\mathit{fix}) << ND(2\alpha_\mathit{fix}+\beta_\mathit{fix}) + N3\alpha_\mathit{fix}$.
+Empirically, we observe 10x to 50x improvement over un-vectorized Biometric Matching in circuit generation time and setup time and 5x to 30x improvement in online time.
+Additionally, the un-vectorized version runs out of memory for N=256, while the vectorized one continues to run up to standard maximal input size N=4096.
+Our experimental results illustrate this as well. \ana{Check numbers.}
 
@@
 %\ana{Need to state this more precisely.}
+
+\subsection{Syntax}
+
+Our compiler imposes the following restrictions. We note that in some cases, the restrictions 
+can be easily lifted and we plan to do so in future iterations of our compiler.
+
+\begin{enumerate}
+\item The program contains arbitrarily nested loops, however, loop bounds are fixed: \texttt{0 <= i < N}.
+A standard restriction in MPC is that the bounds must be known at circuit-generation time.
+\item Arrays are one-dimensional. N-dimensional arrays are linearized and accessed 
+in row-major order and at this point the programmer is responsible for linearization
+and access. 
+\item Array subscrpts are plaintext values.
+\item Our compiler allows for output (write) arrays, however it restricts
+write access to \emph{canonical writes} along the dimensions of the array. I.e., \texttt{A[i,j] = ...}
+where \texttt{i} and \texttt{j} loop over the two dimensions of \texttt{A} is allowed, 
+but \texttt{A[i,j+2] = ...} is not allowed. Read access is arbitrary. 
+\end{enumerate}
 

### 2022-04-27 overleaf/paper_SIMD.tex
Edits to Analytical Model.

@@
 
-\author{Ana L. Milanova}
+\author{Ana Milanova}
 \email{milanova@cs.rpi.edu}
@@
 
-In this work we focus on an intermediate language and what we call \emph{circuit-independent optimizations}, in a close analogy to \emph{machine-independent} optimizations
+In this work we focus on an intermediate language and what we call \emph{backend-independent optimizations}, in a close analogy to \emph{machine-independent} optimizations
 in the classical compiler. The following figure summarizes our key idea:
@@
 
-\subsection{Our Contribution} In this paper, we develop a compiler framework that takes a Python-like routine and produces MOTION code: we describe (a) the IMP-MPC language, its syntax and semantic restrictions, (b) translation from IMP-MPC into MPC Source, (c) a specific circuit-independent optimization: novel SIMD-vectorization on MPC Source, and (d) translation from MPC Source into MOTION code.
+\subsection{Our Contribution} In this paper, we develop a compiler framework that takes a Python-like routine and produces MOTION code: we describe (a) the IMP-MPC language, its syntax and semantic restrictions, (b) translation from IMP-MPC into MPC Source, (c) a specific backend-independent optimization: novel SIMD-vectorization on MPC Source, and (d) translation from MPC Source into MOTION code.
 %\ana{I think we have to add more on MOTION, specifically mixing and amortization.}
@@
 
-Our results emphasize the importance of ``circuit-independent'' optimizations --- vectorization (described in this work) and protocol mixing (tackled in previous works~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022}) are two optimizations readily available at the level of MPC Source.
-We believe that our work can lead to future work on ``circuit-independent'' compilation and optimization, bringing in new MPC optimizations and combinations of optimizations, much in the vein of standard compilers, and bring MPC programming technology closer to practice and wider applicability.
+Our results emphasize the importance of backend-independent optimizations --- vectorization (described in this work) and protocol mixing (tackled in previous works~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022}) are two optimizations readily available at the level of MPC Source.
+We believe that our work can lead to future work on backend-independent compilation and optimization, bringing in new MPC optimizations and combinations of optimizations, much in the vein of standard compilers, and bring MPC programming technology closer to practice and wider applicability.
 
@@
 \begin{tabular}{ccc}
-\begin{minipage}{0.3\textwidth}
+\begin{minipage}{0.275\textwidth}
 {\small
@@
 # S^ is same as S. C^ replicates C N times: 
-S^ = lift(S, ((i * D) + j), (i:N,j:D)) # S^[i,j] = S[i,j]
-C^ = lift(C, j, (i:N,j:D)) # C^[i,j] = C[j]
+S^ = raise_dim(S, ((i * D) + j), (i:N,j:D)) #S^[i,j] = S[i,j]
+C^ = raise_dim(C, j, (i:N,j:D)) #C^[i,j] = C[j]
 
@@
 Empirically, we observe 10x to 50x improvement over un-vectorized Biometric Matching in circuit generation time and setup time and 5x to 30x improvement in online time.
-Additionally, the un-vectorized version runs out of memory for N=256, while the vectorized one continues to run up to standard maximal input size N=4096.
-Our experimental results illustrate this as well. \ana{Check numbers.}
+Additionally, the un-vectorized version runs out of memory for $N=256$, while the vectorized one runs with the standard maximal input size $N=4,096$.
+\ana{Edit numbers with final experiments.}
 
@@
 \label{sec:model}
+
+\ana{We need an intro to section here.}
 
@@
 -- Past suggestion from Vassilis: Ideally we should assume cost to be a convex function.}\ana{After discussion 1/12: replace single costs 1 with two levels of costs: $\alpha$ for non-local operations, e.g., MUL, and $\beta$ for local ops, e.g., ADD.}
-\item We assume infinite parallel capacity---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely $\alpha$ or $\beta$. 
-This is a standard assumption in Cryptographic Parallel RAM. ABY presents empirical support for this assumption~\ana{Add citations. PRAM, ABY.}
+\item  In MPC frameworks, executing $n$ operations ``at once'' in a single SIMD operation costs a lot less than executing those $n$ operations one by one.
+Following Amdahl's law~\ana{citation!}, we write $\alpha = \frac{1}{s}p\alpha + (1-p)\alpha$, where $p$ is the fraction of execution time that benefits from amortization and $(1-p)$ 
+is the fraction that does not, and $s$ is the available resource. Thus, $n\alpha = \frac{n}{s}p\alpha + n(1-p)\alpha$. 
+For the purpose of the model we assume that $s$ is large enough and the term $\frac{n}{s}p\alpha$ amounts to a \emph{fixed cost} incurred regardless of
+whether $n$ is $10,000$ or just $1$. (This models the cost of sending a packet from party A to party B for example.) The bottom line is that amortized execution 
+of $n$ operations costs $\alpha_\mathit{fix} + n\alpha_{var}$ in contrast to unamortized execution, which costs $n\alpha_\mathit{fix} + n\alpha_{var}$ and, of course,
+$\alpha_\mathit{fix} << n\alpha_\mathit{fix}$.
+%We assume infinite parallel capacity---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely $\alpha$ or $\beta$. 
+%This is a standard assumption in Cryptographic Parallel RAM. ABY presents empirical support for this assumption~\ana{Add citations. PRAM, ABY.}
 %\ishaq{Comment from Vassilis: We assume there are infinite parallel capacities, this is the standard assumption in PRAM (Cryptographic Parallel RAM).}\ana{PRAM assumption stays.
@@
 \item MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption,
-2 MUL instructions scheduled in parallel benefit from amortization and cost $\alpha$, however a MUL and a MUX instructions scheduled
-in parallel still cost $2\alpha$. 
+2 MUL instructions can be amortized in a single SIMD instruction that costs $\alpha_\mathit{fix} + 2\alpha_\mathit{var}$, however a MUL and a MUX instruction 
+still cost $2\alpha_\mathit{fix} + 2\alpha_\mathit{var}$ even when scheduled ``in parallel''. 
+%MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption,
+%2 MUL instructions scheduled in parallel benefit from amortization and cost $\alpha$, however a MUL and a MUX instructions scheduled
+%in parallel still cost $2\alpha$. 
 %\ishaq{We need to reword this assumption to something like "$K$ parallel MUL costs much less than $K$ (because they get amortized), but any mix of $K$  MUL and MUX still cost roughly $K$.

### 2022-04-27 overleaf/prob_stmt_np_hardness.tex
Edits to Analytical Model.

@@
 
-\ana{Ishaq?
+%\ana{Ishaq?
 Basically, define sequential schedule, then define an equivalent parallel schedule.
@@
 
-At the lowest level, we have two types of MPC instructions (also called \emph{gates} in similar works) 1) local/non-interactive instruction (i.e.
+As mentioned earlier, at the lowest level, we have two types of MPC instructions (also called \emph{gates} in similar works) 1) local/non-interactive instruction (i.e.
 an addition instruction $A$) and 2) remote/interactive instruction (i.e.
 a multiplication instruction $M$).
-Each instruction in the program is either an $A$-instruction or an $M$-instruction.
+%Each instruction in the program is either an $A$-instruction or an $M$-instruction.
 
@@
 Given a serial schedule (a linear graph) of an MPC program i.e.
-a sequence of instructions $S := (S_1, \dots, S_n)$, where $S_i \in \{A, M\}, 1 \leq i \leq n$, and a def-use dependency graph $G(V, E)$ corresponding to $S$, our task is to construct a parallel schedule (another linear graph) $P := (P_1, \dots, P_m)$ observing the following conditions:
+a sequence of instructions $S := (S_1; \dots; S_n)$, where $S_i \in \{A, M\}, 1 \leq i \leq n$, and a def-use dependency graph $G(V, E)$ corresponding to $S$, our task is to construct a parallel schedule (another linear graph) $P := (P_1; \dots; P_m)$ observing the following conditions:
 
 \begin{enumerate}
-    \item Multiple, not necessarily continuous instructions from $S$ can be grouped into a single $P_i$.
+    \item All $P$'s consist of MPC instructions of the same kind, e.g., all MUL, or ADD, etc. 
     \item Def-use dependencies of the graph $G(V, E)$ should be preserved i.e.
 if instructions $S_i, S_j, i < j$ form a def-use i.e.
-an edge exists from $S_i$ to $S_j$ in $G$), then they can only be mapped to $P_{i'}, P_{j'}$ such that $i' < j'$.
+an edge exists from $S_i$ to $S_j$ in $G$, then they can only be mapped to $P_{i'}, P_{j'}$ such that $i' < j'$.
 \end{enumerate}
 
-\paragraph{Correctness} Correctness of $\mathcal{P}$ is guaranteed by definition. Preserving def-use \emph{dependencies} means the computed function remains the same in both $S$ and $P$.
+\paragraph{Correctness} Correctness of $P$ is guaranteed by definition. Preserving def-use \emph{dependencies} means the computed function remains the same in both $S$ and $P$.
 
 In order to benefit from parallelization/amortization, we must schedule two or more $A$-instructions in the same parallel node (or two or more $M$-instructions in the same parallel node).
-We also assume that scheduling $A$-instructions in parallel with $M$-instruction does not benefit from amortization\footnote{this is not strictly true, but assuming it, e.g.
+Recall that we also assume that scheduling $A$-instructions in parallel with $M$-instruction does not benefit from amortization\footnote{this is not strictly true, but assuming it, e.g.
 as in \cite{Ishaq2019, Demmler2015ABYA, Mohassel2018}, helps simplify the problem.}.
@@
 
-\begin{enumerate}
-    \item $A$ costs $\alpha$ units and $M$ costs $\beta$ unites.
-    \item There is unlimited bandwidth i.e.
+%\begin{enumerate}
+%    \item $A$ costs $\alpha$ units and $M$ costs $\beta$ unites.
+%    \item There is unlimited bandwidth i.e.
 a single $A$-instruction (or $M$-instruction) costs as much as $N$ amortized $A$-instructions (or $M$-instructions), concretely either $\alpha$ unites or $\beta$ units.
-\end{enumerate}
+%\end{enumerate}
 
@@
 
-where $cost(S_i) = \alpha$ if $S_i$ is an $A$-instruction and $cost(S_i) = \beta$ if $S_i$ is an $M$-instruction. Similarly, the cost of schedule $P$ is
+where $cost(S_i) = \alpha$ or $\beta$. %if $S_i$ is an $A$-instruction and $cost(S_i) = \beta$ if $S_i$ is an $M$-instruction. 
+Similarly, the cost of schedule $P$ is
 
@@
 
-Each $P_i$ may contain multiple instructions, thus $\mathit{cost}(P_i) = \alpha$ if $P_i$ consists of $A$-instructions only, it is $\beta$ if $P_i$ consists of $M$-instructions only, and it is $(\alpha + \beta)$ if $P_i$ mixes $A$-instructions and $M$-instructions.
-Our goal is to construct a parallel schedule $P$ that reduces the program cost (when compared to cost of $S$).
+Each $P_i$ may contain multiple instructions, and $\mathit{cost}(P_i)$ is amortized.
+Thus, according to our model $\mathit{cost}(P_i) = \alpha_\mathit{fix} + |P_i|\alpha_\mathit{var}$ if $P_i$ stores $\alpha$-instructions, 
+or $\mathit{cost}(P_i) = \beta_\mathit{fix} + |P_i|\beta_\mathit{var}$ if it stores $\beta$-instructions.
+%if $P_i$ consists of $A$-instructions only, it is $\beta$ if $P_i$ consists of $M$-instructions only, and it is $(\alpha + \beta)$ if $P_i$ mixes $A$-instructions and $M$-instructions. 
+Our goal is to construct a parallel schedule $P$ that reduces the program cost (when compared to cost of $S$), possibly an optimal schedule. 
+Originally we hoped that the problem is simper and computation of the optimal schedule is tractable. Unfortunately, the optimal schedule turns out to be NP-hard via a reduction 
+to the Shortest Common Supersequence problem.
 
@@
 Note that we consider a linearized MPC schedule $S$ above for ease of exposition only.
-In our tool-chain we use an MPC-Source control flow graph (CFG) $G'(V', E')$ along with def-use graph $G(V,E)$ to construct $P$.
-The argument becomes slightly more involved when dealing with a graph $G'$ that may contain cycles.
+In our tool-chain we use an MPC Source control flow graph (CFG) $G'(V', E')$ along with def-use graph $G(V,E)$ to construct $P$.
+%The argument becomes slightly more involved when dealing with a graph $G'$ that may contain cycles.
 
@@
 
-As an example, consider the MPC program consisting of the following three sequences, right arrow indicates a \emph{dependence}, meaning that the source node must execute before the target node:
+As an example, consider the MPC program consisting of the following three sequences, all made up of two distinct $\alpha$-instructions $M_1$ and $M_2$.
+The right arrow indicates a def-use \emph{dependence}, meaning that the source node must execute before the target node:
 
 \begin{enumerate}
-    \item $A \rightarrow M \rightarrow A$
-    \item $A \rightarrow A \rightarrow A$
-    \item $M \rightarrow A \rightarrow M$
+    \item $M_1 \rightarrow M_2 \rightarrow M_1$
+    \item $M_1 \rightarrow M_1 \rightarrow M_1$
+    \item $M_2 \rightarrow M_1 \rightarrow M_2$
 \end{enumerate}
 
-A \emph{schedule} $P: P_1 \rightarrow P_2 \dots \rightarrow P_k$ is such that for each sequence $S_i$ in the set, if $S_i[j]$ precedes $S_i[j']$ in $S_i$ then $S_i[j]$ is scheduled in node $P_\ell$, $S_i[j']$ is scheduled in node $P_{\ell'}$, and $P_\ell$ precedes $P_{\ell'}$ in $P$.
+%A \emph{schedule} $P: (P_1 ; P_2 \dots ; P_m)$ is such that for each sequence $S_i$ in the set, if $S_i[j]$ precedes $S_i[j']$ in $S_i$ then $S_i[j]$ is scheduled in node $P_\ell$, $S_i[j']$ is scheduled in node $P_{\ell'}$, and $P_\ell$ precedes $P_{\ell'}$ in $P$.
 
-Cost of schedule $P$ is computed using \cref{eq_par_schedule_cost} above.
 % The cost of a schedule $P$ is
@@
 $$ 
-A(1), A(2) \rightarrow M(1), A(2), M(3) \rightarrow A(1), A(2), A(3) \rightarrow M(3)
+M_1(1), M_1(2) \; ;\; M_1(2) \; ; \; M_2(1), M_2(3) \; ; \; M_1(1), M_1(2), M_1(3) \; ; \;  M_2(3)
 $$
 
-The parentheses above indicate the sequence where the instruction comes from: (1), (2), or (3). The cost of this schedule is $3\alpha + 2\beta$.
+The parentheses above indicate the sequence where the instruction comes from: (1), (2), or (3). 
+Cost of schedule $P$ is computed using \cref{eq_par_schedule_cost} above and it amounts to $5\alpha_\mathit{fix} + 9\alpha_\mathit{var}$.
 
-The problem of finding a schedule $P$ with a minimal $cost(P)$ for a given loop body has been shown to be an NP-Hard problem, as it can be reduced to the problem of finding a \emph{shortest common supersequence}, a known NP-Hard problem\cite{Maier1978, Vazirani2010}.
+The problem of finding a schedule $P$ with a minimal $cost(P)$ is shown to be NP-Hard problem, as it can be reduced to the problem of finding a \emph{shortest common supersequence}, a known NP-Hard problem\cite{Maier1978, Vazirani2010}.
 The shortest common supersequence problem is as follows: {\it given two or more sequences find the the shortest sequence that contains all of the original sequences.} This can be solved in $O(n^k)$ time, where $n$ is the cardinality of the longest sequence and $k$ is the number of sequences.
-For our problem $n$ is the maximum length of a node and $k$ is the number of total number of nodes.
+For our problem $n$ is the maximum length of a sequence and $k$ is the number of total number of sequences.
+We can immediately see that the optimal schedule is the shortest schedule, since the shortest schedule minimizes the fixed cost while the variable cost remains the same.
 
-To see the reduction, suppose $P$ is a schedule with minimal cost (computed by a black-box algorithm).
-We can derive a schedule $P'$ with the same cost as $P$, by mapping each mixed node $P_i \in P$ to two consecutive nodes in $P'$: an $A$-instruction node followed by an $M$-instruction node.
-Clearly, $P'$, which now is a sequence of $A$'s and $M$'s, is a supersequence of each sequence $S_i$, i.e., $P'$ is a common supersequence of $S_1 \dots S_n$.
+\ana{This paragraph needs work.} To formalize the reduction, suppose $P$ is a schedule with minimal cost (computed by a black-box algorithm). 
+%We can derive a schedule $P'$ with the same cost as $P$, by mapping each mixed node $P_i \in P$ to two consecutive nodes in $P'$: an $A$-instruction node followed by an $M$-instruction node. 
+Clearly, $P'$, which now is a sequence of $M_1$ and $M_2$ nodes, is a supersequence of each sequence $S_i$, i.e., $P'$ is a common supersequence of $S_1 \dots S_n$.
 It is also a shortest common supersequence.
-To see this, let $X$ and $Y$ denote, respectively, the number of $A$ and $M$ nodes in $P'$.
-The cost of $P'$, and $P$, is $X \cdot \alpha + Y \cdot \beta$.
-Now suppose, there exists a shorter common supersequence, $P''$ that consists of $X'$ nodes of type $A$-instructions $Y'$ nodes of type $M$-instructions.
+%To see this, let $X$ and $Y$ denote, respectively, the number of $A$ and $M$ nodes in $P'$. 
+The cost of $P$, is $L\alpha_\mathit{fix}+N\alpha_\mathit{var}$ where $L$ is the length of $P$ and $N$ is the total number of instructions across all sequences. %$X \cdot \alpha + Y \cdot \beta$.
+
+Now suppose, there exists a shorter common supersequence $P'$. 
+that consists of $X'$ nodes of type $A$-instructions $Y'$ nodes of type $M$-instructions.
 Since $P''$ is shorter than $P'$, therefore $X' + Y' < X + Y$, and $X' \cdot \alpha + Y' \cdot \beta < X \cdot \alpha + Y \cdot \beta$ i.e.
@@
 
+\begin{comment}
 \subsection{Loud Thinking}
@@
 Thus, we have a shorter super sequence $SS_2$ that does not contradict that the optimality of the schedule used to generate $SS_1$.
+\end{comment}

### 2022-04-27 compiler/compiler/motion_backend/low_level_rendering.py
Only add input gates for loop counter plaintext conversions

@@ -102,34 +102,35 @@ def render_param(param: Parameter, type_env: TypeEnv) -> str:
     )
 
 
-def collect_plaintext_conversions(expr: AssignRHS, type_env: TypeEnv) -> list[Var]:
+# Collect counter uses for conversion to shared variables
+def collect_counter_uses(expr: AssignRHS, enclosing_loops: list[For]) -> list[Var]:
     if isinstance(expr, Var):
-        if type_env[expr].is_plaintext():
+        if any(loop.counter == expr for loop in enclosing_loops):
             return [expr]
         else:
             return []
     elif isinstance(expr, UnaryOp):
-        return collect_plaintext_conversions(expr.operand, type_env)
+        return collect_counter_uses(expr.operand, enclosing_loops)
     elif isinstance(expr, BinOp):
-        return collect_plaintext_conversions(
-            expr.left, type_env
-        ) + collect_plaintext_conversions(expr.right, type_env)
+        return collect_counter_uses(
+            expr.left, enclosing_loops
+        ) + collect_counter_uses(expr.right, enclosing_loops)
     elif isinstance(expr, (List, Tuple)):
         return [
             var
             for elem in expr.items
-            for var in collect_plaintext_conversions(elem, type_env)
+            for var in collect_counter_uses(elem, enclosing_loops)
         ]
     elif isinstance(expr, (Update, VectorizedUpdate)):
-        return collect_plaintext_conversions(expr.value, type_env)
+        return collect_counter_uses(expr.value, enclosing_loops)
     elif isinstance(expr, Mux):
         return (
-            collect_plaintext_conversions(expr.condition, type_env)
-            + collect_plaintext_conversions(expr.true_value, type_env)
-            + collect_plaintext_conversions(expr.false_value, type_env)
+            collect_counter_uses(expr.condition, enclosing_loops)
+            + collect_counter_uses(expr.true_value, enclosing_loops)
+            + collect_counter_uses(expr.false_value, enclosing_loops)
         )
     elif isinstance(expr, LiftExpr):
-        return collect_plaintext_conversions(expr.expr, type_env)
+        return collect_counter_uses(expr.expr, enclosing_loops)
     # The below expressions never require conversion to plaintext
     elif isinstance(expr, (Constant, Subscript, VectorizedAccess, DropDim)):
         return []
@@ -145,35 +146,29 @@ def render_stmt(
 ) -> str:
     if isinstance(stmt, Assign):
[…]

### 2022-04-27 overleaf/prob_stmt_np_hardness.tex
Finished pass on Analytical model section.

@@
 
-\ana{This paragraph needs work.} To formalize the reduction, suppose $P$ is a schedule with minimal cost (computed by a black-box algorithm). 
+To formalize the reduction, suppose $P$ is a schedule with minimal cost (computed by a black-box algorithm). 
 %We can derive a schedule $P'$ with the same cost as $P$, by mapping each mixed node $P_i \in P$ to two consecutive nodes in $P'$: an $A$-instruction node followed by an $M$-instruction node. 
-Clearly, $P'$, which now is a sequence of $M_1$ and $M_2$ nodes, is a supersequence of each sequence $S_i$, i.e., $P'$ is a common supersequence of $S_1 \dots S_n$.
+Clearly, $P$, which now is a sequence of $M_1$ and $M_2$ nodes, is a supersequence of each sequence $S_i$, i.e., $P$ is a common supersequence of $S_1 \dots S_n$.
 It is also a shortest common supersequence.
 %To see this, let $X$ and $Y$ denote, respectively, the number of $A$ and $M$ nodes in $P'$. 
-The cost of $P$, is $L\alpha_\mathit{fix}+N\alpha_\mathit{var}$ where $L$ is the length of $P$ and $N$ is the total number of instructions across all sequences. %$X \cdot \alpha + Y \cdot \beta$.
-
-Now suppose, there exists a shorter common supersequence $P'$. 
-that consists of $X'$ nodes of type $A$-instructions $Y'$ nodes of type $M$-instructions.
+The cost of $\mathit{cost}(P) = L\alpha_\mathit{fix}+N\alpha_\mathit{var}$ where $L$ is the length of $P$ and $N$ is the total number of instructions across all sequences.
+%$X \cdot \alpha + Y \cdot \beta$.
+Now suppose, there exist a shorter common supersequence $P'$ of length $L'$.
+$\mathit{cost}(P') < \mathit{cost}(P)$ since $L'\alpha_\mathit{var} + N\alpha_\mathit{var} < L\alpha_\mathit{var} + N\alpha_\mathit{var}$, however, this is a contradiction to the assumption that $P$ has the lowest cost.
+\qed
+%that consists of $X'$ nodes of type $A$-instructions $Y'$ nodes of type $M$-instructions.
 Since $P''$ is shorter than $P'$, therefore $X' + Y' < X + Y$, and $X' \cdot \alpha + Y' \cdot \beta < X \cdot \alpha + Y \cdot \beta$ i.e.

### 2022-04-27 compiler/compiler/motion_backend/low_level_rendering.py
Fix issues in vectorized_update type signature

@@ -783,7 +783,7 @@ def render_expr(expr: Union[AssignRHS, SubscriptIndex], ctx: RenderContext) -> s
         )
         idxs = "{}"
         update_array = expr.array
-        if isinstance(update_array, VectorizedAccess):
+        if isinstance(update_array, VectorizedAccess) and all(vectorized for vectorized in update_array.vectorized_dims):
             update_array = update_array.array
         return f"vectorized_update({render_expr(update_array, ctx)}, {dim_sizes}, {vectorized_dims}, {idxs}, {render_expr(expr.value, ctx)})"
 

### 2022-04-27 overleaf/paper_SIMD.tex
Some more edits on Secs. 1-3. Still holding lock.

@@
 whether $n$ is $10,000$ or just $1$. (This models the cost of sending a packet from party A to party B for example.) The bottom line is that amortized execution 
-of $n$ operations costs $\alpha_\mathit{fix} + n\alpha_{var}$ in contrast to unamortized execution, which costs $n\alpha_\mathit{fix} + n\alpha_{var}$ and, of course,
+of $n$ operations is $f(n) = \alpha_\mathit{fix} + n\alpha_{var}$ in contrast to unamortized execution $g(n) = n\alpha_\mathit{fix} + n\alpha_{var}$ and, of course,
 $\alpha_\mathit{fix} << n\alpha_\mathit{fix}$.
@@
 
+\begin{figure*}
+  \includegraphics[width=1.0\linewidth]{figs_paper_SIMD/compiler_framework.png}
+  \caption{Compiler Framework. \ana{Change figure since we are not doing div-and-conquer.}}
+  \label{fig:compiler_framework}
+\end{figure*}
+
 \section{\bf Compiler Framework}
 \label{sec:preliminaries}
-
-\begin{figure*}
-  \includegraphics[width=1.0\linewidth]{figs_paper_SIMD/compiler_framework.png}
-  \caption{Compiler Framework.} 
-  \label{fig:compiler_framework}
-\end{figure*}
 
@@
 
-\figref{fig:compiler_framework} presents an overview of our compiler. In this section, we describe several of the phases of the compiler. 
-Sections~\secref{sec:vectorization} and~\secref{sec:div_and_conq} describe vectorization and divide-and-conquer.
+\figref{fig:compiler_framework} presents an overview of our compiler. This section describes the front-end phases of the compiler
+and Section~\secref{sec:vectorization} describes our vectorization algorithm. \ana{We'll need a section on MOTION code gen too.}
+%\ana{Need to state this more precisely.}
+
+\subsection{Syntax and Semantic Restrictions}
+
+Source syntax is essentially standard IMP syntax as shown below:
+%\begin{figure}[t]
+{\small
+\[
+\begin{array}{l@{~~~~~}l}
+  \begin{array}{l@{~}l@{~~~}l}
+ e & ::= e \; \mathit{op} \; e \mid e \mid \x \mid  \texttt{const} \mid \texttt{A[$e$]} & \mathit{expression} \\
+ % s &::= t \;\f & \mathit{field} \\
+ %\md &::= t\;{\m}(t \;{\this}, \; t\;\x) \; \{\; \ol{t\;\y}\;\s;\;\return\;\y \; \} 
+ %                 & \mathit{method} \\ 
+  s & ::= s;s \mid  & \mathit{sequence} \\
+  & \x = e \mid \texttt{A[$e$]} = e \mid & \mathit{assignment \; stmt} \\
+  & \texttt{for $i$ in range($I$)}: s \mid & \mathit{for \; stmt} \\
+  & \texttt{if $e$ then $s$ else $s$} & \mathit{if \; stmt} \\ 
+%   t &::= q \; {\sf C} & \mathit{qualified \; type}\\
+%   q &::= \high \mid \poly \mid \low & \mathit{FlowCFL \; qualifier}\\
+%    t' &::= q' \; \C & \mathit{field \; qualified \; type}\\
+%   q' &::= \tainted \mid \poly & \mathit{field \; qualifier}\\
+
+  \end{array}
+\end{array}
+\]
+}
+%\end{figure}
+
+The syntax allows for array accesses, arbitrarily nested loops, and if-then-else control flow. 
+Expressions are typed $q \; \tau$, where qualifier $q$ and base type $\tau$ are:
+
+{\small
+\[
+\begin{array}{l@{~~~~~}l}
+  \begin{array}{l@{~}l@{~~~}l}
+ \tau & ::= \texttt{int}  \mid \texttt{bool} \mid \texttt{list[int]} \mid  \texttt{list[bool]} & \mathit{base \; types} \\
+  q & ::= \texttt{shared} \mid \texttt{plain} & \mathit{qualifiers} \\
+  \end{array}
+\end{array}
+\]
+}
+
+The type system is mostly standard, and in our experience, a sweet spot between simplicity and expressivity.
+The \texttt{shared} qualifier denotes shared values, i.e., ones shared among the parties and computed upon 
+under secure computation protocols vs. \texttt{plain} which denotes plaintext values. Shared lists denote 
+shared elements, though the bounds of the list are plaintext. \ana{add subtyping}
+
+Our compiler imposes certain semantic restrictions that it enforces throughout the various 
+phased of compilation. We note that in some cases, the restrictions 
+can be easily lifted and we plan to do so in future iterations of our compiler.
+
+\begin{enumerate}
+\item Loops are of the form \texttt{0 <= i < I} and bounds are fixed at compile time. 
+It is a standard restriction in MPC that the bounds must be known at circuit-generation time.
+\item Arrays are one-dimensional. N-dimensional arrays are linearized and accessed 
+in row-major order and at this point the programmer is responsible for linearization
+and access. (This restriction can be easily lifted.)
+\item Array subscrpts are plaintext values. The typing rule for array access \texttt{A[$e$]} \ana{???}
+Our compiler allows for output (write) arrays, however it restricts
+write access to \emph{canonical writes} along the dimensions of the array. I.e., \texttt{A[i,j] = ...}
+where \texttt{i} and \texttt{j} loop over the two dimensions of \texttt{A} is allowed, 
+but \texttt{A[i,j+2] = ...} is not allowed. Read access can be an arbitrary function of the 
+indices of the enclosing loops. 
+\item \ana{Add rules for logical ops}
+\item \ana{Add rules for arithmetic ops}
+\item \ana{Add rules for MUX}
+\end{enumerate}
+
 We write $i,j,k$ to denote the loop nest: $i$ is the outermost loop, $j$, is immediately nested in $i$, and so on until $k$
@@
 All discussion generalizes to arbitrarily large loop nests.
-%\ana{Need to state this more precisely.}
-
-\subsection{Syntax}
-
-Our compiler imposes the following restrictions. We note that in some cases, the restrictions 
-can be easily lifted and we plan to do so in future iterations of our compiler.
-
-\begin{enumerate}
-\item The program contains arbitrarily nested loops, however, loop bounds are fixed: \texttt{0 <= i < N}.
-A standard restriction in MPC is that the bounds must be known at circuit-generation time.
-\item Arrays are one-dimensional. N-dimensional arrays are linearized and accessed 
-in row-major order and at this point the programmer is responsible for linearization
-and access. 
-\item Array subscrpts are plaintext values.
-\item Our compiler allows for output (write) arrays, however it restricts
-write access to \emph{canonical writes} along the dimensions of the array. I.e., \texttt{A[i,j] = ...}
-where \texttt{i} and \texttt{j} loop over the two dimensions of \texttt{A} is allowed, 
-but \texttt{A[i,j+2] = ...} is not allowed. Read access is arbitrary. 
-\end{enumerate}
 
@@
         \textbf{3-address CFG conversion}:
-        \benjamin{TODO: Is this a good amount of detail?}
+        \benjamin{TODO: Is this a good amount of detail?} \ana{This is good amount of detail for the Tech report. In the actual paper we'll shorten a lot and add pointers to the Tech report.}
 
@@
 
-Then, the control-flow graph is \emph{linearized} into MPC source,
-which has loops but no if-statements.
+Then, the control-flow graph is \emph{linearized} into MPC Source,
+which has loops but no if-then-else-statements.
 This means that both branches of all if-statements are executed,
@@
 
-An important component of our algorithm is the "lifting" of scalars to the corresponding loop dimensionality. 
+An important component of our algorithm is the scalar expansion to the corresponding loop dimensionality. 
 For example, \texttt{d = S[i * D + j] - C[j]} equiv. to \texttt{d = S[i,j] - C[j]}, which gave rise to \texttt{N*D} subtraction operations in the sequential schedule, 

### 2022-04-27 overleaf/paper_SIMD.tex
Edits to Section 3.

@@
       shared[tuple[int,int]]:
-   min_sum = MAX_INT
-   min_idx = 0  
+   min_sum : int = MAX_INT
+   min_idx : int = 0  
    for i in range(N): 
-      sum = 0
+      sum : int = 0
       for j in range(D): 
          # d = S[i,j] - C[j]
-         d = S[i * D + j] - C[j] 
-         p = d * d
+         d : int = S[i * D + j] - C[j] 
+         p : int = d * d
          sum = sum + p
       if sum < min_sum:
-         min_sum = sum
-         min_idx = i
+         min_sum : int = sum
+         min_idx : int = i
    return (min_sum, min_idx)
@@
   & \texttt{for $i$ in range($I$)}: s \mid & \mathit{for \; stmt} \\
-  & \texttt{if $e$ then $s$ else $s$} & \mathit{if \; stmt} \\ 
+  & \texttt{if $e$: $s$ else: $s$} & \mathit{if \; stmt} \\ 
 %   t &::= q \; {\sf C} & \mathit{qualified \; type}\\
@@
 
-The type system is mostly standard, and in our experience, a sweet spot between simplicity and expressivity.
+The type system is mostly standard, and in our experience, a sweet spot between readability and expressivity.
 The \texttt{shared} qualifier denotes shared values, i.e., ones shared among the parties and computed upon 
-under secure computation protocols vs. \texttt{plain} which denotes plaintext values. Shared lists denote 
-shared elements, though the bounds of the list are plaintext. \ana{add subtyping}
+under secure computation protocols; the \texttt{plain} qualifier denotes plaintext values. Shared lists denote 
+shared elements, but plaintext length. Subtyping is again as expected: $q_1 \; \tau <: q_2 \; \tau$ 
+iff $q_1 <: q_2$ where $\texttt{plain} <: \texttt{shared}$. Thus, we can convert a plaintext value into a shared one, 
+but not vice versa.
 
 Our compiler imposes certain semantic restrictions that it enforces throughout the various 
-phased of compilation. We note that in some cases, the restrictions 
-can be easily lifted and we plan to do so in future iterations of our compiler.
+phases of compilation. We note that in some cases, the restrictions 
+can be easily lifted and we plan to do so in future iterations of the work.
 
 \begin{enumerate}
-\item Loops are of the form \texttt{0 <= i < I} and bounds are fixed at compile time. 
+\item Loops are of the form $0 <= i < I$ and bounds are fixed at compile time. 
 It is a standard restriction in MPC that the bounds must be known at circuit-generation time.
@@
 and access. (This restriction can be easily lifted.)
-\item Array subscrpts are plaintext values. The typing rule for array access \texttt{A[$e$]} \ana{???}
-Our compiler allows for output (write) arrays, however it restricts
-write access to \emph{canonical writes} along the dimensions of the array. I.e., \texttt{A[i,j] = ...}
-where \texttt{i} and \texttt{j} loop over the two dimensions of \texttt{A} is allowed, 
-but \texttt{A[i,j+2] = ...} is not allowed. Read access can be an arbitrary function of the 
-indices of the enclosing loops. 
-\item \ana{Add rules for logical ops}
-\item \ana{Add rules for arithmetic ops}
-\item \ana{Add rules for MUX}
+\item Array subscrpts are plaintext values as specified by the rule:
+\begin{semantics}
+    \ntyperule{Array Access}{
+%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
+  \Gamma \vdash e : \texttt{plain int} \quad \Gamma \vdash \texttt{A} : q \; \texttt{list}[\tau] \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
+}{
+  \Gamma \vdash \texttt{A[$e$]} : q \; \tau
+}
+\end{semantics}
+The subscript $e$ is a function of the indices of the enclosing loops. 
+For read access, the compiler allows an arbitrary such function. 
+However, it restricts write access to \emph{canonical writes}, i.e., \texttt{A[i,j,k]} = ...
+where \texttt{i}, \texttt{j} and \texttt{k} loop over the three dimensions of \texttt{A}.  
+Write access such as for example \texttt{A[i,j+2]} = ... is not allowed. 
+%\item \ana{Add rules for logical ops}
+%\item \ana{Add rules for arithmetic ops}
+\item The final restriction involves MUX as expressed by the rule: 
+\begin{semantics}
+    \ntyperule{MUX}{
+%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
+  \Gamma \vdash e_1 : q_1 \; \texttt{bool} \quad \Gamma \vdash e_2 : q_2 \; \tau \quad \Gamma \vdash e_2 : q_2 \; \tau \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
+}{
+  \Gamma \vdash \texttt{MUX($e_1$,$e_2$,$e_3$)} : q_1 \vee q_2 \vee q_3 \; \tau
+}
+\end{semantics}
 \end{enumerate}
-
-We write $i,j,k$ to denote the loop nest: $i$ is the outermost loop, $j$, is immediately nested in $i$, and so on until $k$
-and we use $I,J,K$ to denote the corresponding upper bounds. For simplicity, we write $A[i,j,k]$ to denote canonical access 
+Importantly, the arguments of MUX are restricted to base types, as passing array arguments 
+is not valid MOTION MPC. \ana{Check this.} This causes a slight inconvenience as we could not write 
+\begin{pythonn}
+if e: A[i] = val
+\end{pythonn}
+Instead we had to write
+\begin{pythonn}
+if not(e): val = A[i];
+A[i] = val;
+\end{pythonn}
+
+We note that while the programmer writes 
+annotations at the level of IMP Source (recall \ana{ref to overview section}), the 
+annotations propagate through the transformations and are checked at the 
+level of MPC Source.
+
+%\ana{Disadvantage of MUX}
+
+For the rest of this section we write $i,j,k$ to denote the loop nest: $i$ is the outermost loop, $j$, is immediately nested in $i$, and so on until $k$
+and we use $I,J,K$ to denote the corresponding upper bounds. We write $A[i,j,k]$ to denote canonical access 
 to an array element. In the program, canonical access is achieved via the standard row-major order formula: $(J*K)*i + K*j + k$.
-To simplify the presentation we describe our algorithms in terms of three-element tuples $i,j,k$.
-All discussion generalizes to arbitrarily large loop nests.
-
-\subsection{Semantic Analysis}
+To simplify the presentation we describe our algorithms in terms of three-element tuples $i,j,k$, however, discussion easily generalizes to 
+arbitrarily large loop nests.
+
+\subsection{From IMP Source to SSA}
 
@@
 
-\subsection{MUX Nodes and Pseudo $\phi$-nodes}
+\subsection{From SSA to MPC Source: MUX Nodes and Pseudo $\phi$-nodes}
 

### 2022-04-27 compiler/paper_benchmarks.py
WIP on benchmarks on separate machines

@@ -9,13 +9,18 @@ import logging
 import logging.handlers
 import random
 import json
+import socket
 
 import compiler
 
 from tests import context as test_context
-from tests.benchmark import run_benchmark
+from tests.benchmark import run_benchmark, compile_benchmark, run_benchmark_for_party
 
-from utils import json_serialize, json_deserialize, StatsForInputConfig, StatsForTask
+from utils import json_serialize, json_deserialize, StatsForInputConfig, StatsForTask, RunBenchmarkArgs
+from utils import read_message, write_message
+
+SERVER_PORT = 42143
+CONNECTION_TIMEOUT = 3000
 
 GMW_PROTOCOL = "BooleanGmw"
 BMR_PROTOCOL = "Bmr"
@@ -500,8 +505,6 @@ def run_paper_benchmarks():
             continue;
 
         task_stats = StatsForTask(test_case_dir.name, [])
-        input_fname = os.path.join(test_case_dir.path, "input.py")
-
         compile = True
         i = 0
         non_vec_failed = False
@@ -585,6 +588,138 @@ def run_paper_benchmarks():
         all_stats.append(task_stats)
         log.info("task {} DONE".format(task_stats.label))
     print_benchmark_data(all_stats)
+
+def compile_all_benchmarks():
+    for test_case_dir in os.scandir(test_context.STAGES_DIR):
+        if test_case_dir.name in test_context.SKIPPED_TESTS:
+                continue
+        log.info("Compiling {} ...".format(test_case_dir.name))        
+        compile_benchmark(test_case_dir.name, test_case_dir.path, GMW_PROTOCOL, False)
+        compile_benchmark(test_case_dir.name, test_case_dir.path, GMW_PROTOCOL, True)
+        compile_benchmark(test_case_dir.name, test_case_dir.path, BMR_PROTOCOL, False)
+        compile_benchmark(test_case_dir.name, test_case_dir.path, BMR_PROTOCOL, True)
+
+def run_server_role(address):
+    log.info("Compiling All benchmarks")
+    compile_all_benchmarks()
+    s = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
+    s.bind((address, int(SERVER_PORT)))
+    s.listen()
+    log.info("Server started at address {} port {}".format(address, SERVER_PORT))
+    while True:
+        conn, addr = s.accept()
+        conn.settimeout(CONNECTION_TIMEOUT)
+        with conn:
+            log.debug('Connected by ', addr)
[…]

### 2022-04-27 compiler/tests/benchmark.py

@@ -104,13 +104,6 @@ def run_benchmark(
             text=True,
             cwd=party1_dir,
         ) as party1:
-            # party0.wait(timeout)
-            # party1.wait(timeout)
-            # assert party0.stdout is not None
-            # assert party0.stderr is not None
-            # party0_stdout_raw = party0.stdout.read()
-            # party0_stderr = party0.stderr.read()
-
             try:
                 party0_stdout_raw, party0_stderr = party0.communicate(timeout)
                 party1_stdout_raw, party1_stderr = party1.communicate(timeout)
@@ -167,3 +160,90 @@ def run_benchmark(
                     circuit_stats=party1_circuit_stats,
                 ),
             )
+
+
+def run_benchmark_for_party(
+    myid: str, party0_mpc_addr: str, party1_mpc_addr: str, app_path: str, timeout:int, cmd_args: list[str]
+) -> BenchmarkOutput:
+    # Create directories for output and MOTION logs
+    party_dir = os.path.join(app_path, "party" + myid)
+    if os.path.exists(party_dir):
+        shutil.rmtree(party_dir)
+    os.makedirs(party_dir, exist_ok=True)
+
+    exe_name = os.path.join(app_path, "build", benchmark_name)
+    with subprocess.Popen(
+        [
+            exe_name,
+            "--parties",
+            party0_mpc_addr,
+            party1_mpc_addr,
+            "--my-id",
+            myid,
+        ] + cmd_args,
+        stdout=subprocess.PIPE,
+        stderr=subprocess.PIPE,
+        text=True,
+        cwd=party0_dir,
+    ) as party:
+        try:
+            party_stdout_raw, party_stderr = party.communicate(timeout)
+        except subprocess.TimeoutExpired:
+            party.kill()
+            party_stdout_raw, party_stderr = party.communicate(timeout)
+
+        with open(os.path.join(party_dir, "stdout"), "w") as f:
+            f.write(party_stdout_raw)
+        with open(os.path.join(party_dir, "stderr"), "w") as f:
+            f.write(party_stderr)
+
+        if(party.returncode != 0):
+            return None
+
+        party_timing_stats = statistics.parse_timing_data(
+            party_stderr.split("\n")
+        )
+
+        party_stdout_lines = party_stdout_raw.split("\n")
+        party_output = party_stdout_lines[0]
+        party_circuit_stats = statistics.parse_circuit_data(
+            party_stdout_lines[1:]
[…]

### 2022-04-27 compiler/utils.py

@@ -1,7 +1,10 @@
 from dataclasses import dataclass
-import json
 from tests.benchmark import BenchmarkOutput
 from tests.statistics import CircuitStatistics, TimingStatistics, CommunicationStatistics, TimingDatapoint
+
+import struct
+import json
+import socket
 
 @dataclass
 class StatsForInputConfig:
@@ -73,19 +76,82 @@ class StatsForTask:
                 }
 
 
+@dataclass
+class RunBenchmarkArgs:
+    party0_mpc_addr: str
+    party1_mpc_addr: str
+    cmd_args: list[str]
+    benchmark_name: str
+    benchmark_path: str
+    protocol: str
+    vectorized: bool
+
+    @classmethod
+    def from_dictionary(cls, params):
+        party0_mpc_addr = params['party0_mpc_addr']
+        party1_mpc_addr = params['party1_mpc_addr']
+        cmd_args = params['cmd_args']
+        benchmark_name = params['benchmark_name']
+        benchmark_path = params['benchmark_path']
+        protocol = params['protocol']
+        vectorized = params['vectorized']
+
+        return cls(
+            party0_mpc_addr=party0_mpc_addr,
+            party1_mpc_addr=party1_mpc_addr,
+            cmd_args=cmd_args,
+            benchmark_name=benchmark_name,
+            benchmark_path=benchmark_path,
+            protocol=protocol,
+            vectorized=vectorized
+            )
+
+    def to_dictionary(self):
+        return {
+            'party0_mpc_addr': self.party0_mpc_addr,
+            'party1_mpc_addr': self.party1_mpc_addr,
+            'cmd_args': cmd_args,
+            'benchmark_name': benchmark_name,
+            'benchmark_path': benchmark_path,
+            'protocol': protocol,
+            'vectorized': vectorized
+        }
+
+
 def json_serialize(obj):
     return json.dumps(obj, default=_to_json)
 
 def json_deserialize(json_str):
     return json.loads(json_str, object_hook=_from_json)
 
-def _to_json(python_object):
-    """
-    serializes data objects used in the benchmark script
-    :param python_object: object to serialize
-    :return: json serialization `python_object`
-    """
 
+def read_message(conn):
+    buf = _read_socket_buf(conn, 4)
+    if not buf:
+        print("Error: Unable to read message size")
+        return None
+
+    msg_size = struct.unpack("!i", buf)[0]
+    print("Length of the message is: {}".format(msg_size))
+
[…]

### 2022-04-27 overleaf/paper_SIMD.tex
Edits to Section 4.

@@
 Thus, the fixed cost of the vectorized program amounts to
-$2\alpha_\mathit{fix}$ + $D \beta_\mathit{fix} + N(3\alpha_\mathit{fix})$. (The variable cost is the same in both the vectorized and un-vectorized programs.) 
+$2\alpha_\mathit{fix}$ + $D \beta_\mathit{fix} + N(3\alpha_\mathit{fix})$. (The variable cost is the same in both the vectorized and non-vectorized programs.) 
 The first term in the sum corresponds to the vectorized
@@
 \subsection{Syntax and Semantic Restrictions}
+\label{sec:syntax}
 
@@
 
-Our compiler performs the following semantic analysis steps:
+Our compiler translates from Source to SSA as follows:
 
@@
         \textbf{Syntax checking}:
-        Ensure that the AST matches a restricted subset that our compiler supports.
+        Ensure that the AST matches the restricted subset defined in Section~\secref{sec:syntax}.
         This step outputs an instance of the \texttt{restricted\_ast.Function} class, which represents our restricted subset of the Python AST.
@@
         \textbf{3-address CFG conversion}:
-        \benjamin{TODO: Is this a good amount of detail?} \ana{This is good amount of detail for the Tech report. In the actual paper we'll shorten a lot and add pointers to the Tech report.}
+        %\benjamin{TODO: Is this a good amount of detail?} \ana{This is good amount of detail for the Tech report. In the actual paper we'll shorten a lot and add pointers to the Tech report.}
 
@@
         process the statement.
-        Statements can either be for-loops, if-statements, or assignments.
+        Statements can either be for-loops, if-statements, or assignments (as in~\secref{sec:syntax}).
         Rules for processing each kind of statement are given below:
@@
                 the loop condition
-                (the condition block),
+                (the \emph{condition-block}),
                 the loop body
-                (the body block),
+                (the \emph{body-block}),
                 and the code after the loop
-                (the after-block).
-                Insert a jump from the end of the current block to the condition block.
-                Then, mark the condition block as the current block.
+                (the \emph{after-block}).
+                Insert a jump from the end of the current block to the condition-block.
+                Then, mark the condition-block as the current block.
                 Insert a for-instruction at the end of the current block with the loop counter variable and bounds from the AST.
@@
                 an edge from the current block to the
-                body block labeled ``TRUE''.
-                Then, set the body block to be the current block
+                body-block labeled ``TRUE''.
+                Then, set the body-block to be the current block
                 and process all statements in the AST's loop body.
-                Finally, insert a jump to the condition block and set the after-block as current.
+                Finally, insert a jump to the condition-block and set the after-block as current.
           \item
@@
                 the ``then'' statements of the if-statement
-                (the then-block),
+                (the \emph{then-block}),
                 the ``else'' statements of the if-statement
-                (the else-block),
+                (the \emph{else-block}),
                 and the code after the if-statement
-                (the after-block).
+                (the \emph{after-block}).
                 At the end of the current block,
@@
                 Finally, set the after-block to be the current block,
-                and give it a ``merge condition'' property equal to the condition of the if-statement.
+                and give it a \emph{merge condition} property equal to the condition of the if-statement.
           \item
@@
                 can be a variable or an array subscript.
-                If it is an array subscript such as \texttt{A[i] = x},
+                If it is an array subscript, e.g., \texttt{A[i] = x},
                 change the statement to \texttt{A = Update(A, i, x)}.
@@
 
-\subsection{From SSA to MPC Source: MUX Nodes and Pseudo $\phi$-nodes}
+\subsection{From SSA to MPC Source} %: MUX Nodes and Pseudo $\phi$-nodes}
 
@@
 $\phi$-nodes corresponding to if-statements will be in a basic block
-with the ``merge condition'' property.
+with the merge condition property.
 For example, if $X_3 = \phi(X_1,X_2)$ is in a block with merge condition $C$,
@@
 
-Then, the control-flow graph is \emph{linearized} into MPC Source,
+Next, the control-flow graph is \emph{linearized} into MPC Source,
 which has loops but no if-then-else-statements.
@@
 since that will be after both branches of the if-statement are visited. 
+(The Python AST naturally gives rise to a translation where each conditional has exactly two targets, 
+and each ``merge condition'' block has exactly two incoming edges, a TRUE and a FALSE edge.
+Thus, each $\phi$-node has exactly two multiplexer arguments, which dovetails into MUX.
+This is in contrast with Cytron's algorithm which operates at the level of the CFG and allows for 
+$\phi$-nodes with multiple arguments.)
 Each time the compiler visits a block,
@@
 
-Now, the remaining $\phi$-nodes in MPC source are \emph{pseudo} $\phi$-nodes.
-A pseudo $\phi$-node $X_1 = \phi(X_0,X_2)$ in a loop header is evaluated during circuit generation.
-If it is the 0-th iteration, then the $\phi$-node evaluates to $X_0$, otherwise, it evaluates to $X_2$.
+Now, the remaining $\phi$-nodes in MPC source are the loop header nodes. We call these nodes \emph{pseudo} $\phi$-nodes 
+and we write \python{PHI} in MPC Source. A pseudo $\phi$-node $X_1 = $\python{PHI}$(X_0,X_2)$ in a loop header is evaluated 
+during circuit generation. If it is the 0-th iteration, then the $\phi$-node evaluates to $X_0$, otherwise, it evaluates to $X_2$.
 
@@
 \subsection{Dependence Analysis}
+
+The next steps construct the dependence graph and ???.
 
@@
 \item same-level edge $X\rightarrow Y$ where $X$ and $Y$ are in the same loop nest, say $i,j,k$.
-E.g., the def-use edge from \texttt{d = S[i,j] - C[j]} to \texttt{p = d*d} in the Biometric MPC-source is a same-level edge.
+E.g., the def-use edge from \python{d = S[i,j] - C[j]} to \python{p = d*d} in the Biometric MPC-source is a same-level edge.
 A same-level edge can be a back-edge in which case a $\phi$ node is the target of the edge.
@@
 \ben{Is this necessary to include?}
+
+\subsection{From (Optimized) MPC Source to MOTION}
+\label{sec:MOTION}
+\ana{Ben, add the MOTION section here.}
 

### 2022-04-27 overleaf/paper_SIMD.tex
Did a pass on Section 4. Moving on to Seciton 5.

@@
 \begin{pythonn}
-if not(e): val = A[i];
-A[i] = val;
+if not(e): val = A[i]
+A[i] = val
 \end{pythonn}
@@
 
-The next steps construct the dependence graph and ???.
+The next steps construct the dependence graph and \ana{Add an intro here.}
 
@@
 \item same-level edge $X\rightarrow Y$ where $X$ and $Y$ are in the same loop nest, say $i,j,k$.
-E.g., the def-use edge from \python{d = S[i,j] - C[j]} to \python{p = d*d} in the Biometric MPC-source is a same-level edge.
+E.g., the def-use edge 9 to 10 in the Biometric MPC Source in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source} is a same-level edge.
 A same-level edge can be a back-edge in which case a $\phi$ node is the target of the edge.
-E.g., $\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{sum}_1,\texttt{min}_1)$ to $\texttt{min}_0 = \phi(\texttt{min}_1, 10000)$ in Biometric is a same-level back-edge.
-\item outer-to-inner $X\rightarrow Y$ where $X$ is in an outer loop nest, say $i$, and $Y$ is in an inner one, say $i,j,k$. 
+E.g., 15 to 4 in Biometric is a same-level back-edge.
+\item outer-to-inner $X\rightarrow Y$ where $X$ is in an outer loop nest, say $i$, and $Y$ is in an inner one, say $i,j,k$. E.g., 1 to 4 in Biometric forms is an outer-to-inner edge.
 %A $\phi$-node can be a source or a target of an outer-to-inner forward edge.
 \item inner-to-outer $X\rightarrow Y$ where $X$ is a $phi$-node in an inner loop nest, $i,j,k$, and $Y$ is in the enclosing loop nest $i,j$.
-E.g.
-$\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$ is an inner-to-outer edge.
-An inner-to-outer edge can be a back-edge as well in which case both $X$ and $Y$ are phi-nodes with the source $X$ in a loop nested into $Y$'s loop (not necessarily immediately).
+E.g., the def-use from 8 to 12 gives rise to an inner-to-outer edge.
+%E.g. $\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$ is an inner-to-outer edge. 
+An inner-to-outer edge can be a back-edge as well, in which case both $X$ and $Y$ are phi-nodes with the source $X$ in a loop nested into $Y$'s loop (not necessarily immediately).
 
@@
 
-We require that all inputs are marked as either shared or plaintext.
-We then determine if intermediate variables are shared through taint analysis with "taintedness" referring to the shared attribute.
-Specifically, our compiler follows the following rules:
-\begin{itemize}
-\item If any variable on the right-hand side of an assignment is shared, then the assigned variable is shared
-\item If all variables on the right-hand side of an assignment are plaintext, then the assigned variable is plaintext
-\item Loop counters are always plaintext
-\item Any variables which cannot be determined as shared or plaintext via the above rules are plaintext
-\end{itemize}
-
-The first two rules are standard for taint analysis, and the third rule follows from the MPC problem statement.
-\ben{Is the explanation for the third rule correct?}  The final rule is needed to handle cycles of plaintext values.
-For example, in the below snippet \texttt{sum!2} and \texttt{sum!3} form a dependency cycle and cannot be marked as plaintext through simple taint analysis:
+We require that all inputs are marked as either shared or plaintext, however, we do not require any other qualifier annotations in the program.
+Then we infer if intermediate variables and expressions are shared through taint analysis with "taintedness" referring to the shared attribute.
+Specifically, the compiler follows the following rules, which are standard in positive-negative qualifier systems (here \texttt{shared} is the positive qualifier and \texttt{plain} is the negative one):
+\begin{enumerate}
+\item Loop counters and array upper bounds are always \texttt{plain}.
+\item If any variable on the right-hand side of an assignment is shared, then the assigned variable is \texttt{shared}.
+%\item If all variables on the right-hand side of an assignment are plaintext, then the assigned variable is plaintext \ana{Will be covered by the last rule.}
+\item Any variables that cannot be determined as shared via the above rules are \texttt{plain}.
+\end{enumerate}
+
+%The first two rules are standard for taint analysis, and the third rule follows from the MPC problem statement. \ben{Is the explanation for the third rule correct?}  \ana{Yes, that is exactly it!}
+The final rule is needed in examples like the following. In the below snippet \texttt{sum!2} and \texttt{sum!3} form a dependency cycle and there is no \texttt{shared} value that flows to either one. 
+They are inferred as plaintext. 
+%cannot be marked as plaintext through simple taint analysis:
 {\small
@@
 }
-\ben{I think the above example is unnecessary and could be replaced by an explanation of how "untainted" variables are implicitly plaintext, but I couldn't think of a way to phrase that.}
-
-When converting to MOTION code, any plaintext value used in the right-hand side of a shared assignment is implicitly converted to a shared value for that expression.
-\ben{Is this necessary to include?}
+%\ben{I think the above example is unnecessary and could be replaced by an explanation of how "untainted" variables are implicitly plaintext, but I couldn't think of a way to phrase that.}\ana{I left the example, we can leave it in the TR, remove from paper.}
+
+When converting to MOTION code, any plaintext value used in the right-hand side of a shared assignment is converted to a shared value for that expression. 
+%\ben{Is this necessary to include?} \ana{Yes, that's important.}
+
+Importantly, the taint analysis works on MPC Source, which lacks if-then-else control flow. 
+This significantly simplifies the treatment as there is no need to handle conditionals and implicit flow.
 

### 2022-04-28 benchmarks/biometric_fast.py
Enable biometric_fast benchmark

@@ -1,7 +1,7 @@
 from UTIL import shared
 
 
-def biometric_matching_fast(
+def biometric_fast(
     D: int,
     N: int,
     C: shared[list[int]],
@@ -9,6 +9,7 @@ def biometric_matching_fast(
     two_C: shared[list[int]],
     S: shared[list[int]],
     S_sqr_sum: shared[list[int]],
+    differences: shared[list[int]],
 ) -> tuple[shared[int], shared[int]]:
     """
     Computes biometric matching
@@ -29,12 +30,7 @@ def biometric_matching_fast(
 
     """
 
-    differences: list[int] = []
-    for i in range(D):
-        differences = differences + [0]
-
     min_index: int = 0
-    min_diff: int = differences[0]
     for i in range(N):
         a_sqr_plus_b_sqr: int = S_sqr_sum[i] + C_sqr_sum
         two_a_b: int = 0
@@ -46,35 +42,26 @@ def biometric_matching_fast(
         this_diff: int = a_sqr_plus_b_sqr - two_a_b
         differences[i] = this_diff
 
-        min_diff = differences[0]
         min_index = 0
 
-        for k in range(N):
-            if differences[k] < min_diff:
-                min_diff = differences[k]
-                min_index = k
+    min_diff: int = 99999
+    for i in range(N):
+        if differences[i] < min_diff:
+            min_diff = differences[i]
+            min_index = i
 
     return (min_diff, min_index)
 
 
-def test_biometric_matching_fast(D, N, C, S):
-    """
-    just a convenience method for testing, computes the pre-processing data for the actual call
-    """
-    two_C = [0] * D
-    C_sqr_sum = 0
-    S_sqr_sum = [0] * N
-    for i in range(D):
-        two_C[i] = 2 * C[i]
-        C_sqr_sum = C_sqr_sum + (C[i] * C[i])
-
-    for i in range(N):
-        for j in range(D):
-            S_sqr_sum[i] = S_sqr_sum[i] + (S[i * D + j] * S[i * D + j])
-
-    print(biometric_matching_fast(D, N, C, C_sqr_sum, two_C, S, S_sqr_sum))
-
-
+D = 4
+N = 4
 C = [1, 2, 3, 4]
 S = [4, 5, 2, 10, 2, 120, 4, 10, 99, 88, 77, 66, 55, 44, 33, 22]
[…]

### 2022-04-28 compiler/compiler/motion_backend/low_level_rendering.py

@@ -112,9 +112,9 @@ def collect_counter_uses(expr: AssignRHS, enclosing_loops: list[For]) -> list[Va
     elif isinstance(expr, UnaryOp):
         return collect_counter_uses(expr.operand, enclosing_loops)
     elif isinstance(expr, BinOp):
-        return collect_counter_uses(
-            expr.left, enclosing_loops
-        ) + collect_counter_uses(expr.right, enclosing_loops)
+        return collect_counter_uses(expr.left, enclosing_loops) + collect_counter_uses(
+            expr.right, enclosing_loops
+        )
     elif isinstance(expr, (List, Tuple)):
         return [
             var
@@ -178,7 +178,7 @@ def render_stmt(
             val_expr = render_expr(stmt.rhs, ctx)
 
             # If this isn't a true vectorized access, just subscript normally
-            if all(vectorized == False for vectorized in stmt.lhs.vectorized_dims):
+            if all(not vectorized for vectorized in stmt.lhs.vectorized_dims):
                 lhs = render_expr(stmt.lhs, ctx)
                 return plaintext_conversions + f"{lhs} = {val_expr};"
 
@@ -306,7 +306,9 @@ def render_stmt(
         ):
             return plaintext_conversions + shared_assignment
         else:
-            return plaintext_conversions + shared_assignment + "\n" + plaintext_assignment
+            return (
+                plaintext_conversions + shared_assignment + "\n" + plaintext_assignment
+            )
 
     elif isinstance(stmt, For):
         ctr_initializer = (
@@ -666,7 +668,9 @@ def render_expr(expr: Union[AssignRHS, SubscriptIndex], ctx: RenderContext) -> s
         # If we're lifting a vectorized array, don't render a vectorized access or else the lifted array
         # will hold too many values
         expr_to_render = expr.expr
-        if isinstance(expr_to_render, VectorizedAccess):
+        if isinstance(expr_to_render, VectorizedAccess) and any(
+            vectorized for vectorized in expr_to_render.vectorized_dims
+        ):
             post_expr = ".Unsimdify()"
         else:
             post_expr = ""
@@ -720,7 +724,7 @@ def render_expr(expr: Union[AssignRHS, SubscriptIndex], ctx: RenderContext) -> s
 
     elif isinstance(expr, VectorizedAccess):
         # If this isn't a true vectorized access, just subscript normally
-        if all(vectorized == False for vectorized in expr.vectorized_dims):
[…]

### 2022-04-28 compiler/compiler/vectorize.py

@@ -1314,10 +1314,10 @@ def _basic_vectorization_phase_2(
         s2: Union[llc.Statement, tuple[llc.Statement]],
     ) -> int:
         if isinstance(s1, llc.For):
-            s1 = tuple(s1.body) # type: ignore
+            s1 = tuple(s1.body)  # type: ignore
 
         if isinstance(s2, llc.For):
-            s2 = tuple(s2.body) # type: ignore
+            s2 = tuple(s2.body)  # type: ignore
 
         # If s1 is a closure
         if isinstance(s1, tuple):

### 2022-04-28 compiler/tests/context.py

@@ -8,7 +8,6 @@ RUN_EXAMPLE_APPS = False
 
 SKIPPED_TESTS = [
     # These benchmarks will always be skipped (they're essentially pseudocode)
-    "biometric_fast",
     "biometric_vectorized",
     # The following tests cause issues since they don't loop from 0
     "longest_1s",

### 2022-04-28 compiler/tests/stages/biometric_fast/restricted_ast.py

@@ -1,9 +1,5 @@
-def biometric_matching_fast(D: plaintext[int], N: plaintext[int], C: shared[list[int; ?]], C_sqr_sum: shared[int], two_C: shared[list[int; ?]], S: shared[list[int; ?]], S_sqr_sum: shared[list[int; ?]]) -> tuple[shared[int], shared[int]]:
-    differences = []
-    for i: plaintext[int] in range(0, D):
-        differences = (differences + [0])
+def biometric_fast(D: plaintext[int], N: plaintext[int], C: shared[list[int; ?]], C_sqr_sum: shared[int], two_C: shared[list[int; ?]], S: shared[list[int; ?]], S_sqr_sum: shared[list[int; ?]], differences: shared[list[int; ?]]) -> tuple[shared[int], shared[int]]:
     min_index = 0
-    min_diff = differences[0]
     for i: plaintext[int] in range(0, N):
         a_sqr_plus_b_sqr = (S_sqr_sum[i] + C_sqr_sum)
         two_a_b = 0
@@ -12,10 +8,10 @@ def biometric_matching_fast(D: plaintext[int], N: plaintext[int], C: shared[list
             two_a_b = (two_a_b + tmp)
         this_diff = (a_sqr_plus_b_sqr - two_a_b)
         differences[i] = this_diff
-        min_diff = differences[0]
         min_index = 0
-        for k: plaintext[int] in range(0, N):
-            if (differences[k] < min_diff):
-                min_diff = differences[k]
-                min_index = k
+    min_diff = 99999
+    for i: plaintext[int] in range(0, N):
+        if (differences[i] < min_diff):
+            min_diff = differences[i]
+            min_index = i
     return (min_diff, min_index)

### 2022-04-28 compiler/paper_benchmarks.py
updated benchmark script to run on network

@@ -14,13 +14,17 @@ import socket
 import compiler
 
 from tests import context as test_context
-from tests.benchmark import run_benchmark, compile_benchmark, run_benchmark_for_party
+from tests.benchmark import run_benchmark, compile_benchmark, run_benchmark_for_party, BenchmarkOutput
 
-from utils import json_serialize, json_deserialize, StatsForInputConfig, StatsForTask, RunBenchmarkArgs
+from utils import json_serialize, json_deserialize, StatsForInputConfig, StatsForTask, RunBenchmarkReq
+from utils import GetAddressReq, GetAddressResp
 from utils import read_message, write_message
 
-SERVER_PORT = 42143
+SERVER_PORT = 42142
 CONNECTION_TIMEOUT = 3000
+MPC_PARTY_SERVER_ID = "0"
+MPC_PARTY_CLIENT_ID = "1"
+NUM_ITERS = 5
 
 GMW_PROTOCOL = "BooleanGmw"
 BMR_PROTOCOL = "Bmr"
@@ -73,7 +77,7 @@ def get_rand_ints(n, min=1, max=100):
 def get_biometric_inputs() -> tuple[list[InputArgs], int]:
     all_args = []
     non_vec_up_to = 6 # Only run non-vectorized benchmark upto this index
-    for config in [[4, 4], [4, 8], [4, 16], [4, 32], [4, 64], [4, 128], [4, 256], [4, 512], [4, 1024], [4, 2048], [4, 4096]]:
+    for config in [[4, 4]]:#, [4, 8], [4, 16], [4, 32], [4, 64], [4, 128], [4, 256], [4, 512], [4, 1024], [4, 2048], [4, 4096]]:
         D = config[0]
         N = config[1]
         args = [
@@ -401,36 +405,36 @@ def get_psi_inputs()-> tuple[list[InputArgs], int]:
 def get_inputs(name: str) -> tuple[list[InputArgs], int]:
     if name == "biometric" or name == "biometric_fast":
         return get_biometric_inputs()
-    if name == "chapterfour_figure_12": # millionaire's problem, not interesting
-        return get_chapterfour_figure_12_inputs()
-    if name == "convex_hull" or name == "minimal_points":
-        return get_convex_hull_inputs()
-    if name == "count_102" or name == "longest_102":
-        return get_count_102_inputs()
-    if name == "count_10s":
-        return get_count_10s_inputs()
-    if name == "count_123":
-        return get_count_123_inputs()
[…]

### 2022-04-28 compiler/tests/benchmark.py

@@ -13,6 +13,7 @@ class BenchmarkOutput:
     output: str
     timing_stats: statistics.TimingStatistics
     circuit_stats: statistics.CircuitStatistics
+    
     @classmethod
     def from_dictionary(cls, params):
         name = params['name']
@@ -33,6 +34,23 @@ class BenchmarkOutput:
                 'timing_stats': self.timing_stats,
                 'circuit_stats': self.circuit_stats,
                 }
+
+    @classmethod
+    def sum(cls, a, b):
+        return cls(
+           name = a.name,
+           output = a.output,
+           timing_stats = statistics.TimingStatistics.sum(a.timing_stats, b.timing_stats),
+           circuit_stats = statistics.CircuitStatistics.sum(a.circuit_stats, b.circuit_stats)
+        )
+    @classmethod
+    def div(cls, obj, d):
+        return cls(
+            name = obj.name,
+            output = obj.output,
+            timing_stats = statistics.TimingStatistics.div(obj.timing_stats, d),
+            circuit_stats = statistics.CircuitStatistics.div(obj.circuit_stats, d)   
+        )
 
 
 def run_benchmark(
@@ -163,9 +181,13 @@ def run_benchmark(
 
 
 def run_benchmark_for_party(
-    myid: str, party0_mpc_addr: str, party1_mpc_addr: str, app_path: str, timeout:int, cmd_args: list[str]
+    myid: str, party0_mpc_addr: str, party1_mpc_addr: str, benchmark_name: str, benchmark_path: str, protocol: str, vectorized,
+    timeout:int, cmd_args: list[str]
 ) -> BenchmarkOutput:
     # Create directories for output and MOTION logs
+    app_path = os.path.join(
+        benchmark_path, "motion_app" + "-" + protocol + ("-vectorized" if vectorized else "")
+    )
     party_dir = os.path.join(app_path, "party" + myid)
     if os.path.exists(party_dir):
         shutil.rmtree(party_dir)
@@ -184,7 +206,7 @@ def run_benchmark_for_party(
         stdout=subprocess.PIPE,
         stderr=subprocess.PIPE,
         text=True,
-        cwd=party0_dir,
+        cwd=party_dir,
     ) as party:
         try:
             party_stdout_raw, party_stderr = party.communicate(timeout)
@@ -198,6 +220,7 @@ def run_benchmark_for_party(
             f.write(party_stderr)
 
         if(party.returncode != 0):
[…]

### 2022-04-28 compiler/utils.py

@@ -77,12 +77,38 @@ class StatsForTask:
 
 
 @dataclass
-class RunBenchmarkArgs:
+class GetAddressReq:
+
+    @classmethod
+    def from_dictionary(cls, params):
+        return cls()
+
+    def to_dictionary(self):
+        return {}
+
+@dataclass
+class GetAddressResp:
+    client_address: (str, int) # (address, port)
+
+    @classmethod
+    def from_dictionary(cls, params):
+        client_address = params['client_address']
+
+        return cls(
+            client_address=client_address,
+            )
+
+    def to_dictionary(self):
+        return {
+            'client_address': self.client_address,
+        }
+
+@dataclass
+class RunBenchmarkReq:
     party0_mpc_addr: str
     party1_mpc_addr: str
     cmd_args: list[str]
     benchmark_name: str
-    benchmark_path: str
     protocol: str
     vectorized: bool
 
@@ -92,7 +118,6 @@ class RunBenchmarkArgs:
         party1_mpc_addr = params['party1_mpc_addr']
         cmd_args = params['cmd_args']
         benchmark_name = params['benchmark_name']
-        benchmark_path = params['benchmark_path']
         protocol = params['protocol']
         vectorized = params['vectorized']
 
@@ -101,7 +126,6 @@ class RunBenchmarkArgs:
             party1_mpc_addr=party1_mpc_addr,
             cmd_args=cmd_args,
             benchmark_name=benchmark_name,
-            benchmark_path=benchmark_path,
             protocol=protocol,
             vectorized=vectorized
             )
@@ -110,11 +134,10 @@ class RunBenchmarkArgs:
         return {
             'party0_mpc_addr': self.party0_mpc_addr,
             'party1_mpc_addr': self.party1_mpc_addr,
-            'cmd_args': cmd_args,
-            'benchmark_name': benchmark_name,
-            'benchmark_path': benchmark_path,
-            'protocol': protocol,
-            'vectorized': vectorized
+            'cmd_args': self.cmd_args,
+            'benchmark_name': self.benchmark_name,
+            'protocol': self.protocol,
+            'vectorized': self.vectorized
         }
 
 
@@ -128,15 +151,15 @@ def json_deserialize(json_str):
 def read_message(conn):
     buf = _read_socket_buf(conn, 4)
     if not buf:
-        print("Error: Unable to read message size")
+        # print("Error: Unable to read message size")
         return None
 
     msg_size = struct.unpack("!i", buf)[0]
[…]

### 2022-04-28 overleaf/paper_SIMD.tex
One line typo fix.

@@
 
+\ana{Convention: use \texttt{shared, plain}, {\sf code}, $i$ and $I$ for counters outside of code.}
+
 %We assume arbitrarily nested loops in the MPC-source IR and read-only arrays.
@@
 For read access, the compiler allows an arbitrary such function. 
-However, it restricts write access to \emph{canonical writes}, i.e., \texttt{A[i,j,k]} = ...
-where \texttt{i}, \texttt{j} and \texttt{k} loop over the three dimensions of \texttt{A}.  
-Write access such as for example \texttt{A[i,j+2]} = ... is not allowed. 
+However, it restricts write access to \emph{canonical writes}, i.e., {\sf A[i,j,k] = ...}
+where $i$, $j$ and $k$ loop over the three dimensions of {\sf A}.  
+Write access such as for example {\sf A[i,j+2] = ...} is not allowed. 
 %\item \ana{Add rules for logical ops}
@@
                 can be a variable or an array subscript.
-                If it is an array subscript, e.g., \texttt{A[i] = x},
-                change the statement to \texttt{A = Update(A, i, x)}.
+                If it is an array subscript, e.g., {\sf A[i] = x},
+                change the statement to {\sf A = Update(A, i, x)}.
                 If the statement is not already three-address code,
@@
 {\small
-\begin{verbatim}
+\begin{pythonn}
 plaintext_array = [0, 1, 2, ...]
@@
     sum!3 = sum!2 + plaintext_array[i]
-\end{verbatim}
+\end{pythonn}
 }
@@
 Importantly, the taint analysis works on MPC Source, which lacks if-then-else control flow. 
-This significantly simplifies the treatment as there is no need to handle conditionals and implicit flow.
+This significantly simplifies treatment as there is no need to handle conditionals and implicit flow.
 
@@
 
+\ana{Need an intro here!}
+
+\subsection{Scalar Expansion}
+
 An important component of our algorithm is the scalar expansion to the corresponding loop dimensionality. 
-For example, \texttt{d = S[i * D + j] - C[j]} equiv. to \texttt{d = S[i,j] - C[j]}, which gave rise to \texttt{N*D} subtraction operations in the sequential schedule, 
-is lifted. The argument arrays \texttt{S} and \texttt{C} are lifted and the scaler \texttt{d} is lifted: \texttt{d[i,j] = S[i,j] - C[i,j]}.
+For example, {\sf d = S[i*D+j] - C[j]} equiv. to {\sf d = S[i,j] - C[j]}, which gave rise to {\sf N*D} subtraction operations in the sequential schedule, 
+is lifted. The argument arrays {\sf S} and {\sf C} are lifted and the scaler {\sf d} is lifted: {\sf d[i,j] = S[i,j] - C[i,j]}.
 The algorithm then detects that the statement can be vectorized.
@@
 For those, all writes are canonical writes and all reads are canonical reads. We will apply raise dimension when 
-a scalar is used in an inner loop (e.g., \texttt{sum0} in line 6 of the MPC source code will be raised to a 1-dimensional 
-array since \texttt{sum0} is used in the inner $j$-loop). Drop dimension applies as well; this happens when a scalar
-written in an inner loop is used outside of the loop (e.g., \texttt{sum0} for which the lifted inner loop computes 
-\texttt{D} values, but the outer loop only needs the last one.)
+a scalar is used in an inner loop (e.g., {\sf sum!2} in line 6 of the MPC source code will be raised to a 1-dimensional 
+array since {\sf sum!2} is used in the inner $j$-loop). Drop dimension applies as well; this happens when a scalar
+written in an inner loop is used outside of the loop (e.g., {\sf sum!2} for which the lifted inner loop computes 
+{\sf D} values, but the outer loop only needs the last one.)
 
@@
 Only raise dimension applies to these arrays, and only in the beginning of the program. For example, 1-dimensional array 
-\texttt{C} is lifted into 2-dimensional array \texttt{N x D} by copying the row \texttt{N} times.
+\python{C} is lifted into 2-dimensional array {\sf N x D} by copying the row {\sf N} times.
 
@@
 is as follows. It takes the original array \texttt{C}, the access pattern function $f(i,j,k)$ in loop nest $i,j,k$ and the loop bounds
-\texttt{((i:I),(j:J),(k:K))}: 
+$((i:I),(j:J),(k:K))$: 
 \[ \mathit{raise\_dim}(\texttt{A}, f(i,j,k), \texttt{((i:I),(j:J),(k:K))}) \]

### 2022-04-28 compiler/paper_benchmarks.py
added input for biometric_fast

@@ -90,6 +90,40 @@ def get_biometric_inputs() -> tuple[list[InputArgs], int]:
         args.extend(list(map(str, C)))
         args.append("--S")
         args.extend(list(map(str, S)))
+
+        two_C = [2 * C[i] for i in range(D)]
+        args.append("--two_C")
+        args.extend(list(map(str, two_C)))
+        C_sqr_sum = sum(val * val for val in C)
+        args.append("--C_sqr_sum")
+        args.extend(list(map(str, C_sqr_sum)))
+        S_sqr_sum = [sum(S[i * D + j] * S[i * D + j] for j in range(D)) for i in range(N)]
+        args.append("--S_sqr_sum")
+        args.extend(list(map(str, S_sqr_sum)))
+        differences = [0] * D
+        args.append("--differences")
+        args.extend(list(map(str, differences)))
+
+        label = "D: {}, N: {}".format(D, N)
+        all_args.append(InputArgs(label, args))
+    return (all_args, non_vec_up_to)
+
+def get_biometric_fast_inputs() -> tuple[list[InputArgs], int]:
+    all_args = []
+    non_vec_up_to = 6 # Only run non-vectorized benchmark upto this index
+    for config in [[4, 4]]:#, [4, 8], [4, 16], [4, 32], [4, 64], [4, 128], [4, 256], [4, 512], [4, 1024], [4, 2048], [4, 4096]]:
+        D = config[0]
+        N = config[1]
+        args = [
+        "--D", "{}".format(D),
+        "--N", "{}".format(N),
+        ]
+        C = get_rand_ints(D)
+        S = get_rand_ints(D * N)
+        args.append("--C")
+        args.extend(list(map(str, C)))
+        args.append("--S")
+        args.extend(list(map(str, S)))
         label = "D: {}, N: {}".format(D, N)
         all_args.append(InputArgs(label, args))
     return (all_args, non_vec_up_to)
@@ -403,8 +437,10 @@ def get_psi_inputs()-> tuple[list[InputArgs], int]:
     return (all_args, non_vec_up_to)
 
 def get_inputs(name: str) -> tuple[list[InputArgs], int]:
-    if name == "biometric" or name == "biometric_fast":
-        return get_biometric_inputs()
+    # if name == "biometric":
+    #     return get_biometric_inputs()
+    if name == "biometric_fast":
+        return get_biometric_fast_inputs()
     # if name == "chapterfour_figure_12": # millionaire's problem, not interesting
[…]

### 2022-04-28 overleaf/paper_SIMD.tex
Add translation to MOTION to paper

@@
 \label{sec:MOTION}
-\ana{Ben, add the MOTION section here.}
+
+MOTION supports FOR loops and SIMD operations, so translation from MPC source to MOTION C++ code is relatively straightforward.
+% Variable naming
+Our generated C++ uses the following variable-naming scheme: shared variables are named the same as in the MPC source with the \texttt{!} replaced with an underscore (e.g.
+\texttt{sum!2} would be translated to \texttt{sum\_2}).
+Plaintext variables follow the same naming convention as shared variables but are prefixed with \texttt{\_MPC\_PLAINTEXT\_}.
+The shared representation of constants are named \texttt{\_MPC\_CONSTANT\_} followed by the literal constant (e.g.
+the shared constant \texttt{0} would be named \texttt{\_MPC\_CONSTANT\_0}).
+
+% Function preamble
+The generated MOTION code begins with the declaration of all variables used in the function, including loop counters.
+If a variable is a vectorized array, it is initialized to a correctly-sized array of empty MOTION shares.
+Additionally, each plaintext variable and parameter has a shared counterpart declared.
+Next, all constant values which are used as part of shared expressions are initialized as a shared input from party 0.
+Finally, all plaintext parameters are converted used as shared inputs from party 0 to initialize their shared counterparts.
+
+% Non-vectorized assignments and returns
+Once the function preamble is complete, the MPC source is translated into C++ one statement at a time.
+If there is no vectorization present in a statement, translation to C++ is straightforward: outside of Mux statements and array updates, non-vectorized assignments, expressions, and returns directly translated into their C++ equivalents.
+Unvectorized Mux statements are converted to MOTION's Mux member function on the condition variable.
+Array updates are translated into two C++ assignments: one to update the value in the original array and one to assign the new array.
+
+\begin{table*}
+\begin{tabular}{ccc}
+\begin{minipage}{0.33\textwidth}
+{\small
+\begin{pythonn}
+A[i] = val
+\end{pythonn}
+}
+\end{minipage}
+
+&
+
+\begin{minipage}{0.33\textwidth}
+{\small
+\begin{pythonn}
+A!2 = update(A!1, i, val)
+\end{pythonn}
+}
+\end{minipage}
+
+&
+
+\begin{minipage}{0.33\textwidth}
+{\small
+\begin{cppp}
+A_1[i] = val;
+A_2 = A_1;
+\end{cppp}
+}
+\end{minipage}
+
+\\
+
+IMP Source & MPC Source & MOTION Code
+\end{tabular}
+\caption{MOTION Translation: Array Updates}
+\label{tab:motion_translation_array_updates}
+\end{table*}
+
+% Phi nodes and FOR loops
+Phi nodes are broken into two components: the ``false'' branch which assigns the initial value of the phi node and the ``true'' branch which assigns the phi node's backedge.
+The assignment of the false branch occurs right before the Phi node's enclosing loop.
+Inside of the Phi node's enclosing loop, a C++ \texttt{if} statement is inserted to only assign the true branch of the phi node after the first iteration.
+MPC FOR loops are converted to C++ FOR loops which iterate the loop counter over the specified range.
+The loop counter is initialized before any phi nodes in the loop as they could rely on it; there is no other special handling of FOR loops.
+
+\begin{table*}
+\begin{tabular}{cc}
+\begin{minipage}{0.5\textwidth}
+{\small
+\begin{pythonn}
+for i in range(N):
+   tmp = PHI(arr[i], val!0)
+	 ...
+\end{pythonn}
+}
+\end{minipage}
+
+&
+
+\begin{minipage}{0.5\textwidth}
+{\small
+\begin{cppp}
+_MPC_PLAINTEXT_i = 0;
+tmp = arr[_MPC_PLAINTEXT_i];
+for (; _MPC_PLAINTEXT_i < _MPC_PLAINTEXT_N; _MPC_PLAINTEXT_i++) {
+   if (_MPC_PLAINTEXT_i != 0) {
+      tmp = val_0;
+	 }
+	 ...
+}
+\end{cppp}
+}
+\end{minipage}
+
+\\
+
+MPC Source & MOTION Code
+\end{tabular}
+\caption{MOTION Translation: FOR loop with Phi nodes}
+\label{tab:motion_translation_for_loop}
+\end{table*}
+
+% SIMD representation
+Vectorization is handled with utility functions to manage accessing and updating slices of arrays.
+All SIMD values are stored in non-vectorized form as 1-dimensional \texttt{std::vector}s in row-major order.
+Whenever a SIMD value is used in an expression, the utility function \texttt{vectorized\_access()} takes the unvectorized representation of a SIMD value, along with the size of each dimension and the requested slice's indices, and converts that slice to a MOTION SIMD value.
+Because MOTION supports SIMD operations using the same C++ operators as non-SIMD operations, we do not need to perform any other transformations to the expression.
+Therefore the translation of an expression containing SIMD values is identical to that of expressions without SIMD values once vectorized accesses are inserted.
+
+Similarly, the \texttt{vectorized\_assign()} function assigns a (potentially SIMD) value to a slice of a vectorized array.
+This operation cannot be done with a simple subscript as SIMD assignments will update a range of values in the non-vectorized representation.
+\ben{Is the next sentence needed?  They can always check the implementation of vectorized\_assign() if they want to see specifics.}
+The function takes the array being assigned to as a reference to update it in place so that any existing values are not lost.  SIMD values are never directly assigned to non-SIMD values.
+
+\begin{table*}
+\begin{tabular}{cc}
+\begin{minipage}{0.5\textwidth}
+{\small
+\begin{pythonn}
+sum!4[I] = ADD_SIMD(sum!3[I], p[I, j])
+\end{pythonn}
+}
+\end{minipage}
+
+&
+
+\begin{minipage}{0.5\textwidth}
+{\small
+% Ben: This line is too long but I don't know how to break it well
+\begin{cppp}
+vectorized_assign(sum_4, {_MPC_PLAINTEXT_N}, {true}, {}, vectorized_access(sum_3, {_MPC_PLAINTEXT_N}, {true}, {}) + vectorized_access(p, {_MPC_PLAINTEXT_N, _MPC_PLAINTEXT_D}, {true, false}, {_MPC_PLAINTEXT_j}));
+\end{cppp}
+}
+\end{minipage}
+
+\\
+
+MPC Source & MOTION Code
+\end{tabular}
+\caption{MOTION Translation: Assignment to SIMD value}
+\label{tab:motion_translation_simd_assignment}
+\end{table*}
+
+% vectorized_update()
+\ben{Technically this is semantically the same as non-vectorized update but we use a utility function to combine the steps into a single C++ statement - should it still be included?}
+Updating SIMD arrays is also implemented differently from updating non-vectorized arrays.
+Instead of separating the array update from the assignment of the new array, these steps are combined with the \texttt{vectorized\_update()} utility function.
+This function operates identically to \texttt{vectorized\_assign()}, however it additionally returns the array after the assignment occurs.
+This value is then used for the assignment to the new variable.
+
+% raise_dim() implementation
+\ben{The compiler and C++ internally uses ``lift'' as the terminology for raise\_dim(), but the paper does not :(...}
+Raising the dimensions of an expression uses the \texttt{lift()} utility function which takes a lambda for the raised expression and the dimensions of the output.
+The lambda accepts an array of integers representing the index along each of the dimensions being raised, and the translation of the expression which is being raised replaces each of the dimension index variables with the relevant subscript of this array.
+There is also a special case of the \texttt{lift()} function which occurs when we are raising an array.
+This case only occurs when the dimensionality of the array is equal to the first dimensions of the array (any dimension which is being lifted to which is not in the array is after every dimension in the array) \ben{I don't know how to phrase the previous sentence more concisely}.
+In this case, the array is simply extended along the new dimensions (if the array is already correctly sized it is returned as is).
+
+% Ben (note as comment since tables aren't fixed placement): I only include the translation of the expression itself (instead of an assignment) since the assignment includes the extra junk from a vectorized_assign.
+Also none of the raise_dims in biometric use the dimension index which I think is the most interesting/important part of the lift expression translation so I made up a toy example.
+\begin{table*}
+\begin{tabular}{cc}
+\begin{minipage}{0.5\textwidth}
+{\small
+\begin{pythonn}
+raise_dim(i + j, (i:N, j:M))
+\end{pythonn}
+}
+\end{minipage}
+
+&
+
+\begin{minipage}{0.5\textwidth}
+{\small
+\begin{cppp}
+lift(std::function([&](const std::vector<std::uint32_t> &idxs){return idxs[0] + idxs[1];}), {_MPC_PLAINTEXT_N, _MPX_PLAINTEXT_M})
+\end{cppp}
+}
+\end{minipage}
+
+\\
+
+MPC Source & MOTION Code
+\end{tabular}
+\caption{MOTION Translation: Raising dimensions}
+\label{tab:motion_translation_raise_dim}
+\end{table*}
+
+% drop_dim() implementation
+Dropping dimensions use the \texttt{drop\_dim()} and \texttt{drop\_dim\_monoreturn()} utility functions.
+They function identically but the latter returns a scalar for the case when the final dimension of an array is dropped.
+These functions take the non-vectorized representation of an array, along with the dimensions of that array, and return the array with the final dimension dropped.
+\ben{drop\_dim() taking the true branch of phi node values doesn't need to be covered here, right?}
+
+% Shared/plaintext conversions
+\ben{Ishaq or Vassilis: please check over this paragraph... I don't want to make some mistake about MOTION's capabilities.}
+Currently, our compiler only supports the \texttt{Bmr} and \texttt{BooleanGMW} protocols as MOTION does not implement all operations for other protocols \ben{Should that go here?}.
+MOTION does not support publicly-known constants for these protocols, so all conversions from plaintext values to shares are performed by providing the plaintext value as a shared input from party 0.
+Due to this limitation, our translation to MOTION code attempts to minimize the number of conversions into a plaintext value.
+This is accomplished by creating a shared copy of each plaintext variable and updating that copy in lock-step with the plaintext variable.
+Since variables are often initialized to a common constant value (e.g.
+0), this approach decreases the number of input gates by only creating a shared input for each constant.
+Loop counters must still be converted to a shared value on each iteration that they are used, however we only output this conversion when necessary to prevent loops where the counter is not used in a shared expression from increasing the number of input gates.
+
+% Tons of copies != runtime costs
+Due to the SSA translation phase as well as the various conversions to and from SIMD values which our utility functions perform, our generated MOTION code often includes many copies of arrays and scalar values.
+These copies do not incur a runtime cost as the arrays simply hold pointers to the underlying shares, so no new shares or gates are created as a result of this copying.
+\ben{Is there anything else to say here?
+Should this be placed somewhere else?}
 

### 2022-04-28 overleaf/biometric_vectorization_example.tex
Minor tweaks to Extension with Array Writes. (Made more precise.)
Changed language=cppp to language=C++ in macros.

@@
 
-We start from Benjamin's code with linear loops (MPC Source):
+We start from MPC Source:
 
 {\small
-\begin{verbatim}
-min_sum!1 = 10000
+\begin{pythonn}
+min_sum!1 = MAX_INT
 min_index!1 = 0
@@
 !2!1 = (min_sum!2, min_index!2)   
-\end{verbatim}
+\end{pythonn}
 }
@@
 
-In the code below, all initializations (e.g., \texttt{min\_sum!3 = i}), operations, and PHI nodes are \emph{implicitly vectorized}. 
-$\mathit{raise\_dim}$ and $\mathit{drop\_dim}$ statements have slightly different interpretation. %\ana{This is vague, may need elaboration.}
-
-The example illustrates the two different versions of $\mathit{raise\_dim}$. \texttt{C!0' = raise\_dim(C!0, j, (i:N!0,j:D!0))} reshapes
-the read-only input array, while \texttt{sum!3' = drop\_dim(sum!3)} removes the $j$ dimension of \texttt{sum!3}.
+In the code below statements (e.g., {\sf min\_sum!3 = i}) are implicitly vectorized. 
+%$\mathit{raise\_dim}$ and $\mathit{drop\_dim}$ statements have slightly different interpretation. %\ana{This is vague, may need elaboration.}
+The example illustrates the two different versions of $\mathit{raise\_dim}$. {\sf raise\_dim(C!0, j, (i:N!0,j:D!0))} reshapes
+the read-only input array, while {\sf drop\_dim(sum!3)} removes the $j$ dimension of {\sf sum!3}.
 %\ana{More here! One just adds a least-significant dimension. The other one may reshape an input array.}
@@
 {\small
-\begin{verbatim}
-min_sum!1 = 10000
+\begin{pythonn}
+min_sum!1 = MAX_INT
 min_sum!1^ = raise_dim(min_sum!1, (i:N!0))
@@
    sum!2 = 0 // Will lift, when hoisted
-   sum!2^ = raise_dim(sum!2, (j:D!0)) // Special form?
+   sum!2^ = raise_dim(sum!2, (j:D!0)) 
    for j in range(0, D!0):
@@
 !2!1 = (min_sum!2^, min_index!2^)     
-\end{verbatim}
+\end{pythonn}
 }
@@
 
-Starting at the inner phi-node \texttt{sum!3 = PHI(sum!2\^, sum!4)}, the algorithm first computes its closure.
-The closure amounts to the phi-node itself and the addition node \texttt{sum!4 = (sum!3 + p!3)}, accounting for the loop-carried dependency of the computation of \texttt{sum}.
+Starting at the inner phi-node {\sf sum!3 = PHI(...)}, the algorithm first computes its closure.
+The closure amounts to the phi-node itself and the addition node {\sf sum!4 = (sum!3 + p!3)}, accounting for the loop-carried dependency of the computation of {\sf sum}.
 The algorithm replaces this closure with a FOR loop on $j$ removing vectoriaztion on $j$.
 Note that the SUB and MUL computations remain outside of the loop as they do not depend on phi-nodes that are part of cycles.
-The dependences are from \texttt{p!3[I,J] = (d!3[I,J] * d!3[I,J])} to the monolithic FOR loop and from the FOR loop to \texttt{sum!3\^ = drop\_dim(sum!3)}.
-(Lower case index, e.g., \texttt{i}, indicates non-vectorized dimension, while uppercase index, e.g., \texttt{I} indicates vectorized dimension.)
+The dependences are from {\sf p!3[I,J] = (d!3[I,J] * d!3[I,J])} to the monolithic FOR loop and from the FOR loop to {\sf sum!3\^ = drop\_dim(sum!3)}.
+Lower case index, e.g., \texttt{i}, indicates non-vectorized dimension, while uppercase index, e.g., \texttt{I} indicates vectorized dimension.
 
@@
 !2!1 = (min_sum!2^, min_index!2^)
-\end{verbatim}
+\end{pythonn}
 }
 
-When processing the outer loop two closures arise, one for \texttt{min\_sum!2[I] = PHI(...)} and one 
-for \texttt{min\_index!2[I] = PHI(...)}. Since the two closures \emph{do not} intersect, we have two distinct FOR-loops on $i$:
+When processing the outer loop two closures arise, one for {\sf min\_sum!2[I] = PHI(...)} and one 
+for {\sf min\_index!2[I] = PHI(...)}. Since the two closures \emph{do not} intersect, we have two distinct FOR-loops on $i$:
 
 {\small
-\begin{verbatim}
-min_sum!1 = 10000
+\begin{pythonn}
+min_sum!1 = MAX_INT
 min_sum!1^ = raise_dim(min_sum!1, (i:N!0))
@@
 !2!1 = (min_sum!2^, min_index!2^)
-\end{verbatim}
+\end{pythonn}
 }
@@
 {\small
-\begin{verbatim}
-min_sum!1 = 10000
+\begin{pythonn}
+min_sum!1 = MAX_INT
 min_index!1 = 0
@@
 !2!1 = (min_sum!2, min_index!2)
-\end{verbatim}
+\end{pythonn}
 }
@@
 d!3[I,J] = SUB_SIMD(S!0^[I,J],C!0^[I,J])
-p!3[I,J] = MUL_SIMD(d!3[I,J] * d!3[I,J])
+p!3[I,J] = MUL_SIMD(d!3[I,J], d!3[I,J])
 

### 2022-04-28 overleaf/compiler.tex
Split paper into sections.

--- 2022-04-28 overleaf/paper_SIMD.tex
@@
-% !TeX TXS-program:compile = txs:///pdflatex/[--shell-escape]
-%
-% acm templates v1.57 (sigconf template)
-%
-%\documentclass[sigconf, screen, review, anonymous, natbib=false]{acmart}
-\documentclass[sigconf, screen, natbib=false, dvipsnames, table]{acmart}
-\settopmatter{printacmref=true}
-% mandatory for CCS'19
-\usepackage{balance}
-\usepackage{mathtools}
-% for creating a balanced last page (usually last page with references)
-
-% % defining the \BibTeX command - from Oren Patashnik's original BibTeX documentation.
-% \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08emT\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-
-% Rights management information.
-% This information is sent to you when you complete the rights form.
-% These commands have SAMPLE values in them; it is your responsibility as an author to replace
-% the commands and values with those provided to you when you complete the rights form.
-%
-% These commands are for a PROCEEDINGS abstract or paper.
-
-\copyrightyear{2019}
-\acmYear{2019}
-\setcopyright{acmcopyright}
-\acmConference[CCS '19]{2019 ACM SIGSAC Conference on Computer and Communications Security}{November 11--15, 2019}{London, United Kingdom}
-\acmBooktitle{2019 ACM SIGSAC Conference on Computer and Communications Security (CCS '19), November 11--15, 2019, London, United Kingdom}
-\acmPrice{15.00}
-\acmDOI{10.1145/3319535.3339818}
-\acmISBN{978-1-4503-6747-9/19/11}
-
-%
-% These commands are for a JOURNAL article.
-%\setcopyright{acmcopyright}
-%\acmJournal{TOG}
-%\acmYear{2018}\acmVolume{37}\acmNumber{4}\acmArticle{111}\acmMonth{8}
-%\acmDOI{10.1145/1122445.1122456}
-
-%
-% Submission ID.
-% Use this when submitting an article to a sponsored event. You'll receive a unique submission ID from the organizers
-% of the event, and this ID should be used as the parameter to this command.
-%\acmSubmissionID{123-A56-BU3}
-
-%
-% The majority of ACM publications use numbered citations and references. If you are preparing content for an event
-% sponsored by ACM SIGGRAPH, you must use the "author year" style of citations and references. Uncommenting
-% the next command will enable that style.
-%\citestyle{acmauthoryear}
-
- \makeatletter
- % \renewcommand{\section}{\abovedisplayskip 10\p@ \@plus3\p@ \@minus1\p@%
- %                       \belowdisplayskip 5\p@ \@plus3\p@ \@minus1\p@%
- %                       \abovedisplayshortskip 0pt \@plus2\p@%
- %                       \belowdisplayshortskip 0pt \@plus2\p@ \@minus0\p@%
- %                       \@startsection{section}{1}{\z@}%
- %                        {-17\p@ \@plus -4\p@ \@minus -4\p@}%
- %                        {6\p@ \@plus 4\p@ \@minus 4\p@}%
- %                        {\normalfont\large\bfseries\boldmath
- %                         \rightskip=\z@ \@plus 8em\pretolerance=10000 }}
- \renewcommand{\subsection}{\@startsection{subsection}{2}{\z@}%
-                        {-8\p@ \@plus -4\p@ \@minus -4\p@}%
-                        {5\p@ \@plus 2\p@ \@minus 2\p@}%
-                        {\normalfont\Large\bfseries\boldmath
-                         \rightskip=\z@ \@plus 3em\pretolerance=10000 }}
-  \renewcommand{\subsubsection}{\@startsection{subsubsection}{3}{\z@}%
-                        {-6\p@ \@plus -4\p@ \@minus -4\p@}%
-                        {1\p@ \@plus 1\p@ \@minus 0\p@}%
-                        {\normalfont\normalsize\bfseries\boldmath}}
-% \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\z@}%
-%                       {-8\p@ \@plus -4\p@ \@minus -4\p@}%
-%                       {-2\p@ \@plus -0.22em \@minus -0.1em}%
-%                       {\normalfont\normalsize\bfseries}}
- \makeatother
-
-% *** GRAPHICS RELATED PACKAGES ***
-\usepackage{graphicx}
-\usepackage{comment}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{listings}
-% declare the path(s) where your graphic files are
-\graphicspath{{../figures/}}
-% and their extensions so you won't have to specify these with
-% every instance of \includegraphics
-\DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-
-% Minted is used for source code highlighting
-%\usepackage[outputdir=/Volumes/ramdisk/]{minted}
-% \usepackage{minted}
-% *** MATH PACKAGES ***
-\usepackage{amsmath}
-
-% *** SPECIALIZED LIST PACKAGES ***
-%\usepackage{algorithm2e}
-\usepackage{algorithmic}
-
-% *** ALIGNMENT PACKAGES ***
-\usepackage{array}
-
-\usepackage{url}
-
-\usepackage[utf8]{inputenc}
-\usepackage[american]{babel}
-%\usepackage[backend=biber,style=ACM-Reference-Format]{biblatex}
-%\addbibresource{./library.bib}
-%\addbibresource{./cryptobib/abbrev2.bib}
-%\addbibresource{./cryptobib/crypto_crossref.bib}
-
-% Used for Theorems and Definitions
-\usepackage{amsthm}
-\newtheorem{definition}{Definition}
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}{Lemma}
-\theoremstyle{definition}
-\newtheorem{remark}{Remark}
-\newtheorem{assumption}{Assumption}
-
-\usepackage{xcolor}
-\usepackage{xspace}
-\usepackage{xstring} % required by IfEqCase
-
-\usepackage{booktabs} % better tables
-\usepackage{algorithm}
-\usepackage{algorithmic}
-
-% for references, use it like this: \cref{some_label}
-% it will automatically put Appendix, Figure, Section, etc before the reference
-\usepackage[capitalise]{cleveref}
-
-% ishaq: adding our macros
-\input{macros.tex}
-
-\sloppy
-\begin{document}
-\fancyhead{}
-% do not delete this code.
-
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
-\title[Compilation and Backend-Independent Vectorization for Multi-Party Computation]{Compilation and Backend-Independent Vectorization for Multi-Party Computation}
-
-%
-% The "author" command and its associated commands are used to define the authors and their affiliations.
-% Of note is the shared affiliation of the first two authors, and the "authornote" and "authornotemark" commands
-% used to denote shared contribution to the research.
-
-\author{Benjamin Levy}
-\email{levyb3@rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Benjamin Sherman}
-\email{shermb@rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Muhammad Ishaq}
-%\authornote{This work was done in part while the author was at RPI.}
-\email{ishaqm@purdue.edu}
-%\orcid{1234-5678-9012}
-%\author{G.K.M. Tobin}
-%\authornotemark[1]
-%\email{webmaster@marysville-ohio.com}
-\affiliation{%
-    \institution{Purdue University}
-    %\streetaddress{P.O. Box 1212}
-    \city{West Lafayette}
-    \state{Indiana}
-%    \postcode{43017-6221}
-}
-
-\author{Lindsey Kennard}
-\email{fire.elemental@gmail.com}
-\affiliation{%
-    \institution{STR}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Boston}
-    \state{Massachusets}
-    %\country{Iceland}
-}
-
-\author{Ana Milanova}
-\email{milanova@cs.rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Vassilis Zikas}
-%\authornote{This work was done in part while the author was visiting UCLA and supported in part by DARPA and SPAWAR under contract N66001-15-C-4065 and by a SICSA Cyber Nexus Research Exchanges grant.}
-\email{vzikas@purdue.edu}
-%\orcid{1234-5678-9012}
-%\author{G.K.M. Tobin}
-%\authornotemark[1]
-%\email{webmaster@marysville-ohio.com}
-\affiliation{%
-    \institution{Purdue University}
-    %\streetaddress{P.O. Box 1212}
-    \city{West Lafayette}
-    \state{Indiana}
-    %    \postcode{43017-6221}
-}
-
-%
-% By default, the full list of authors will be used in the page headers. Often, this list is too long, and will overlap
-% other information printed in the page headers. This command allows the author to define a more concise list
-% of authors' names for this purpose.
-%\renewcommand{\shortauthors}{Trovato and Tobin, et al.}
-
-\renewcommand{\A}{{\sf A}}
-\renewcommand{\B}{{\sf B}}
-\renewcommand{\C}{{\sf C}}
-\renewcommand{\D}{{\sf D}}
-
-% As a general rule, do not put math, special symbols or citations
-% in the abstract
-\begin{abstract}
-
-\ana{We need an abstract.}
-
-\end{abstract}
-
-%
-% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-% Please copy and paste the code instead of the example below.
-%
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10003752.10010124.10010138.10010143</concept_id>
-<concept_desc>Theory of computation~Program analysis</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10003752.10003777.10003789</concept_id>
-<concept_desc>Theory of computation~Cryptographic protocols</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10002978.10002979</concept_id>
-<concept_desc>Security and privacy~Cryptography</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Theory of computation~Program analysis}
-\ccsdesc[500]{Theory of computation~Cryptographic protocols}
-\ccsdesc[300]{Security and privacy~Cryptography}
-
-%
-% Keywords. The author(s) should pick words that accurately describe the work being
-% presented. Separate the keywords with commas.
-\keywords{multiparty computation; compilers; cryptography}
-
-\begin{comment}
-%
-% A "teaser" image appears between the author and affiliation information and the body
-% of the document, and typically spans the page.
-\begin{teaserfigure}
-    \includegraphics[width=\textwidth]{sampleteaser}
-    \caption{Seattle Mariners at Spring Training, 2010.}
-    \Description{Enjoying the baseball game from the third-base seats. Ichiro Suzuki preparing to bat.}
-    \label{fig:teaser}
-\end{teaserfigure}
-\end{comment}
-%
-% This command processes the author and affiliation and title information and builds
-% the first part of the formatted document.
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-
-Multi-party computation (MPC) allows $N$ parties $p_1,\dots, p_N$ to perform a computation on their private inputs securely.
-Informally, security means that the secure computation protocol computes the correct output (correctness) and it does not leak any information about the individual party inputs, other than what can be deduced from the output (privacy).
-
-% The first results from the early 80's~\cite{FOCS:Yao82b,STOC:GolMicWig87,STOC:BenGolWig88,STOC:ChaCreDam88} introduced the problem and formally defined its security, and investigate feasibility in different models---e.g., passive or malicious adversaries, unconditional or computational security, with or without a broadcast channel.
-This spawned a rich theoretical literature of beautiful works improving and extending in various manners the original ideas.
-
-MPC theory dates back to the early 1980-ies~\cite{FOCS:Yao82b,STOC:GolMicWig87,STOC:BenGolWig88,STOC:ChaCreDam88}.
-Long the realm of theoretical cryptography, MPC has seen significant advances in programming technology in recent years.
-These advances bring MPC closer to practice and wider applicability ---
-MPC technology has been employed in real-world scenarios such as auctions~\cite{FC:BCDGJK09}, biometric identification~\cite{ESORICS:BlaGas11},
-and privacy-preserving machine learning~\cite{SP:MohZha17,CCS:MohRin18}.
-The goal is to bring the technology to a level where programmers can write \emph{secure} and \emph{efficient} programs without
-commanding extensive knowledge of cryptographic primitives.
-
-The problem, therefore, is to build a high-level programming language and a compiler, and there has been significant advance in
-this space, e.g., ~\cite{CCS:BenNisPin08,ESORICS:BogLauWil08,CCS:ZhaSteBla13,Hastings:2019,Keller:2020,Acay:2021,Braun:2022} among other work.
-Current research largely falls at the two ends
-of the classical compiler: (1) work on \emph{front-end} language design and (2) work on \emph{back-end} protocol implementation. Work on language design
-focuses on high-level constructs necessary to express multiple parties, computation by different parties, and information flow from one party
-to another~\cite{SP:RasHamHic14, Acay:2021}. On the other end, work on protocol implementation focuses on cryptographic foundations and their efficient circuit-level
-implementation~\cite{NDSS:DemSchZoh15, CCS:ABFKLO18, Braun:2022}, e.g., implementation of operations (e.g., MUL, ADD) using different sharing protocols
-(Boolean or Arithmetic GMW~\cite{STOC:GolMicWig87} or Yao's garbled circuits~\cite{FOCS:Yao82b}), as well as efficient share conversion from one representation
-to another.
-
-Earlier compilers did both back-end and front-end translation without a specific focus on either, as their aim was to demonstrate applicability of MPC on real-world
-programming problems. As the field advanced, works have focused more closely on front-end language design (e.g., Wysteria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021})
-or back-end ``circuit-level'' design and implementation (e.g., MOTION~\cite{Braun:2022}).
-
-In this work we focus on an intermediate language and what we call \emph{backend-independent optimizations}, in a close analogy to \emph{machine-independent} optimizations
-in the classical compiler. The following figure summarizes our key idea:
-
-{\begin{center}
-\includegraphics[width=0.8\linewidth]{figs_paper_SIMD/focus.png}
-\end{center}
-}
-
-We emphasize the MPC Source intermediate representation and optimization over MPC Source. As in classical compilers, we envision different front ends (e.g., Wysteria,
-our front end IMP-MPC) compiling into MPC Source. MPC Source is particularly suitable for optimizations such as protocol mixing~\cite{Ishaq:2019, Fang:2022} and SIMD-vectorization,
-which takes advantage of amortization at the circuit level. The MPC Source IR exposes the linear structure of MPC programs, which simplifies program analysis; this is
-in contrast to source, which has if-then-else constructs. In the same time, MPC Source is sufficiently ``high-level''  to support analyses and optimizations that take into account
-control and data flow in a specific program. MPC Source is small in size and analysis is tractable, as opposed to analysis over an unrolled circuit~\cite{Ishaq:2019}. Again as
-in classical compilers, we envision translation of MPC Source (optimized or unoptimized) into MOTION, SPDZ, or other back-end code.
-
-%In contrast, the high-level program contains if-then-else and array writes, or the low-level circuit .
-%The compilation problem, i.e., translation from a high-level language to circuits, is difficult and therefore, breaking the problem into sub-problems can  tractable.
-
-\subsection{Our Contribution} In this paper, we develop a compiler framework that takes a Python-like routine and produces MOTION code: we describe (a) the IMP-MPC language, its syntax and semantic restrictions, (b) translation from IMP-MPC into MPC Source, (c) a specific backend-independent optimization: novel SIMD-vectorization on MPC Source, and (d) translation from MPC Source into MOTION code.
-%\ana{I think we have to add more on MOTION, specifically mixing and amortization.}
-
-We focus on the MOTION framework as our back-end for several reasons.
-First, it is the state of the art in terms of performance~\cite{Braun:2022}.
-Second, it provides an API over efficient implementation for a wide variety of cryptographic operations in three different protocols ---  Arithmetic GMW, Boolean GMW, and BMR --- which allows for protocol mixing~\cite{Ishaq:2019, Fang:2022}, a known backend-independent optimization.
-Third, MOTION provides API for SIMD-level operations, which amortize cost and lead to significant improvement in memory footprint and
-throughput~\cite{NDSS:DemSchZoh15, CCS:ABFKLO18, Braun:2022}. It also enables MPC Source-level optimizations such SIMD vectorization.
-
-Our second contribution is an analytical model for cost estimation of amortized schedules. Originally, we hoped that optimal scheduling (under our model) was tractable, as the problem
-appeared simpler than the classical scheduling problem. Unfortunately, we show that optimal scheduling is NP-hard via a reduction to the Shortest Common Supersequence (SCS) problem.
-Cost estimation is important as it drives not only vectorization but other optimizations such as protocol mixing optimizations and scheduling as well~\cite{Ishaq:2019, Fang:2022}.
-
-Our most important contribution is the implementation and evaluation of the compiler framework.
-We demonstrate expressivity of the source language by running the compiler on \ana{X} programs; these include classical MPC benchmarks such as PSI and Biometric matching, as well as kMeans, Histogram, and others.
-Vectorization leads to \ana{X}x improvement in number of gates on average, \ana{Y}x speedup in circuit generation time, \ana{Z}x speedup in the LAN setting, and \ana{Z}x speedup in the WAN setting using the Boolean GMW protocol.
-Results form BMR are similar.
-
-Our results emphasize the importance of backend-independent optimizations --- vectorization (described in this work) and protocol mixing (tackled in previous works~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022}) are two optimizations readily available at the level of MPC Source.
-We believe that our work can lead to future work on backend-independent compilation and optimization, bringing in new MPC optimizations and combinations of optimizations, much in the vein of standard compilers, and bring MPC programming technology closer to practice and wider applicability.
-
-\subsection{Outline}
-
-The rest of the paper is organized as follows.
-\secref{sec:overview} presents an overview of the compiler.
-\secref{sec:model} describes our model for cost estimation and argues NP-hardness of optimal scheduling and \secref{sec:compiler} describes the compiler.
-\secref{sec:experiments} presents implementation and evaluation.
-\secref{sec:related} discusses related work and \secref{sec:conclusion} concludes.
-
-\begin{comment}
-\begin{itemize}
-\item We define the scheduling problem for MPC.
-We present an analytical model to reason about cost of schedules
-and show that scheduling is NP-hard via a reduction to the shortest
-common supersequence problem.
-
-\item We present a compiler that takes an IMP-like high-level
-program and produces amortized (i.e., vectorized) low-level cryptographic
-code in the MOTION framework. Central contributions are 1) a novel compiler framework,
-2) a vectorization algorithm that produces optimal schedules for a large
-number of MPC programs, and 3) reasoning over output arrays without Array SSA; we remove
-infeasible loop-carried dependences introduced by standard SSA to improve vectorization.
-
-\item We present an implementation and evaluation in the MOTION framework.
-\ana{Fill in with final results. Mention benchmarks (standard + new ones + HyCC). }
-
-\end{itemize}
-\end{comment}
-
-\section{Overview}
-\label{sec:overview}
-
-\begin{table*}
-\begin{tabular}{ccc}
-\begin{minipage}{0.275\textwidth}
-{\small
-\begin{pythonn}
-def biometric(C: shared[list[int]], D: int,
-      S: shared[list[int]], N: int) ->
-      shared[tuple[int,int]]:
-   min_sum : int = MAX_INT
-   min_idx : int = 0
-   for i in range(N):
-      sum : int = 0
-      for j in range(D):
-         # d = S[i,j] - C[j]
-         d : int = S[i * D + j] - C[j]
-         p : int = d * d
-         sum = sum + p
-      if sum < min_sum:
-         min_sum : int = sum
-         min_idx : int = i
-   return (min_sum, min_idx)
-\end{pythonn}
-}
-\end{minipage}
-
-&
-
-\begin{minipage}{0.33\textwidth}
-{\tiny
-\begin{pythonn}
-min_sum!1 = MAX_INT
-min_idx!1 = 0
-for i in range(0, N):
-   min_sum!2 = PHI(min_sum!1, min_sum!4)
-   min_idx!2 = PHI(min_idx!1, min_idx!4)
-   sum!2 = 0
-   for j in range(0, D):
-      sum!3 = PHI(sum!2, sum!4)
-      d = SUB(S[((i * D) + j)],C[j])
-      p = MUL(d,d)
-      sum!4 = ADD(sum!3,p)
-   t = CMP(sum!3,min_sum!2)
-   min_sum!3 = sum!3
-   min_idx!3 = i
-   min_sum!4 = MUX(t, min_sum!3, min_sum!2)
-   min_idx!4 = MUX(t, min_idx!3, min_idx!2)
-return (min_sum!2, min_idx!2)
-\end{pythonn}
-}
-\end{minipage}
-
-&
-
-\begin{minipage}{0.33\textwidth}
-{\small
-\begin{pythonn}
-min_sum!1 = MAX_INT
-min_idx!1 = 0
-# S^ is same as S. C^ replicates C N times:
-S^ = raise_dim(S, ((i * D) + j), (i:N,j:D)) #S^[i,j] = S[i,j]
-C^ = raise_dim(C, j, (i:N,j:D)) #C^[i,j] = C[j]
-
-sum!2[I] = [0,..,0]
-# computes _all_  "at once"
-d[I,J] = SUB_SIMD(S^[I,J],C^[I,J])
-p[I,J] = MUL_SIMD(d[I,J],d[I,J])
-
-for j in range(0, D):
-   # sum!2[I], sum!3[I], sum!4[I] are size-N vectors
-   # computes N intermediate sums "at once"
-   sum!3[I] = PHI(sum!2[I], sum!4[I])
-   sum!4[I] = ADD_SIMD(sum!3[I],p[I,j])
-
-min_idx!3[I] = [0,1,...N-1]
-for i in range(0, N):
-   min_sum!2 = PHI(min_sum!1, min_sum!4)
-   t[i] = CMP(sum!3[i],min_sum!2)
-   min_sum!4 = MUX(t[i], sum!3[i], min_sum!2)
-for i in range(0, N):
-   min_idx!2 = PHI(min_idx!1, min_idx!4)
-   min_idx!4 = MUX(t[i], min_idx!3[i], min_idx!2)
-return (min_sum!2, min_idx!2)
-\end{pythonn}
-}
-\end{minipage}
-
-\\
-
-(a) IMP Source & (b) MPC Source & (c) Optimized MPC Source
-\end{tabular}
-\caption{Biometric Matching: From IMP Source to MPC Source to Optimized MPC Source. }
-\label{tab:source_and_MPC_source_and_optimized_MPC_source}
-\end{table*}
-
-\subsection{Source}
-
-As a running example, consider Biometric matching, a standard MPC benchmark.
-An intuitive (and naive) implementation is as shown in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(a).
-Array \python{C} is the feature vector of \texttt{D} features that we wish to match and \python{S}
-is the database of \python{N} size-\python{D} vectors that we match against.
-
-\begin{comment}
-{\small
-\begin{verbatim}
-def biometric(C: shared[list[int]], D: int,
-              S: shared[list[int]], N: int) ->
-                      tuple[shared[int], shared[int]]:
-    min_sum = 10000
-    min_index = 0
-    for i in range(N): #loop over database
-        sum = 0
-        for j in range(D): #loop over features
-            d = S[i * D + j] - C[j] #i.e., d = S[i,j] - C[j]
-            p = d * d
-            sum = sum + p
-        if sum < min_sum:
-            min_sum = sum
-            min_index = i
-    return (min_sum, min_index)
-\end{verbatim}
-}
-\end{comment}
-
-Our compiler takes essentially standard IMP~\ana{Need citation here.}
-syntax and imposes certain semantic restrictions. (We detail the restrictions in the following sections.)
-The programmer writes an iterative program and annotate certain inputs
-and outputs as \emph{shared}. In the example arrays \python{C} and \python{S}
-are shared, meaning that they store shares, however, the array sizes \python{D} and
-\python{N} respectively are plaintext. The code iterates over the entries in the
-database and computes the Euclidean distance (its square actually) of the current
-entry \python{S[i]} and \python{C}. The program returns the index of the vector that gives
-the best match plus the corresponding sum of squares.
-
-\subsection{MPC Source and Cost of Schedule}
-
-Our compiler generates an intermediate representation, MPC Source. MPC Source
-for Biometric Matching is shown in
-Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(b). First, MPC Source is an SSA form.
-Second, it is linear. The conditional in lines 13-14 in IMP Source turns into the linear code in lines 12-16 in MPC Source.
-The test turns into a CMP operation (\python{t = CMP(sum!3,min\_sum!2)}), followed by the
-true-branch sequence, followed by the MUX operations. The first MUX operation selects the value
-of \python{min\_sum}: if \python{t} is true, then \python{min\_sum} gets the value of the second multiplexer
-argument,  \python{min\_sum!3}, otherwise it takes the value of the third argument, \python{min\_sum!2}.
-Third, MPC Source is a special form of SSA. The SSA $\phi$-nodes at the if-then-else (lines 13-15) turn into
-MUX operations, while the $\phi$-nodes at for loops turn into what we call pseudo PHI nodes with a straightforward semantics.
-
-\begin{comment}
-{\small
-\begin{pythonn}
-1. min_sum!1 = 10000
-2. min_index!1 = 0
-3. for i in range(0, N!0):
-4.    min_sum!2 = PHI(min_sum!1, min_sum!4)
-5.    min_index!2 = PHI(min_index!1, min_index!4)
-6.    sum!2 = 0
-7.    for j in range(0, D!0):
-8.      sum!3 = PHI(sum!2, sum!4)
-9.      d!3 = (S!0[((i * D!0) + j)] - C!0[j]) // MPC
-10.     p!3 = (d!3 * d!3) // MPC
-11.     sum!4 = (sum!3 + p!3) // MPC
-12.  !1!2 = (sum!3 < min_sum!2) // MPC
-13.  min_sum!3 = sum!3
-14.  min_index!3 = i
-15.  min_sum!4 = MUX(!1!2, min_sum!3, min_sum!2) // MPC
-16.  min_index!4 = MUX(!1!2, min_index!3, min_index!2) // MPC
-17.!2!1 = (min_sum!2, min_index!2)
-\end{pythonn}
-}
-\end{comment}
-
-%The compiler linearizes the source turning conditionals into MUX statements. The PHI nodes
-%are remnants of the SSA IR; the compiler generates code that picks the correct value
-%when producing MOTION output; the MOTION framework in turn linearizes
-%loops when it generates the circuit.
-
-We turn to our analytical model to compute the \emph{cost} of the program. Assuming fixed
-cost $\beta$ for a local MPC operation (essentially just ADD) and cost $\alpha$ for a remote
-MPC operation (e.g., MUX, CMP, and remaining operations), the cost of the iterative schedule
-will be $ND(2\alpha+\beta) + N(3\alpha$).
-
-A key contribution is the vectorizing transformation. We can compute all $N*D$
-subtraction operations (line 9) in a single SIMD instruction; similarly we can compute
-all multiplication operations (line 10) in a single SIMD instruction. And while we cannot
-vectorize computation of the N individual sums, we can compute the N sums in parallel.
-Our compiler \emph{automatically detects these opportunities and transforms the program}.
-It is standard that MPC researchers write vectorized versions of the Biometric
-program by hand; we are the first (to the best of our knowledge) to automatically
-transform a naive, iterative MPC program into an unintuitive vectorized one.
-
-\subsection{Vectorized MPC Source and Cost of Schedule}
-
-Our compiler produces the vectorized program shown in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(c).
-Note that this is still our intermediate representation, Optimized MPC Source. Subsequently,
-the compiler turns this code into MOTION variables, loops and SIMD primitives, which MOTION then
-uses to generate the circuit.
-
-The compiler determines that SUB and MUL in ``naive'' MPC Source (lines 9 and 10 in (b))
-can be fully vectorized into the SIMD SUB and MUL in optimized MPC Source (lines 9 and 10 in (c)).
-Notation \python{p[I,J]} denotes a 2-dimensional array with fully vectorized dimensions.
-The computation of sum (line 11 in (b))
-is sequential across the $j$-dimension, but it is parallel across the $i$-dimension.
-The loop in lines 12-16 in (c) illustrates; here \python{p[I,j]} refers to the $j$-th column in \python{p}.
-Unfortunately, CMP and MUX remain sequential.
-
-\begin{comment}
-{\small
-\begin{verbatim}
-min_sum!1 = 10000
-min_index!1 = 0
-// S!0^ is same as S!0. C!0^ replicates C!0 N-times:
-S!0^ = raise_dim(S!0, ((i * D!0) + j), (i:N!0,j:D!0))
-C!0^ = raise_dim(C!0, j, (i:N!0,j:D!0))
-
-sum!2 = [0,..,0]
-// Computes all differences and all products "at once"
-d!3[I,J] = SUB_SIMD(S!0^[I,J],C!0^[I,J])
-p!3[I,J] = MUL_SIMD(d!3[I,J],d!3[I,J])
-
-for j in range(0, D!0):
-  // sum!2[I], sum!3[I], sum!4[I] are vectors of size N
-  // Computes N intermediate sums "at once"
-  sum!3[I] = PHI(sum!2[I], sum!4[I])
-  sum!4[I] = ADD_SIMD(sum!3[I],p!3[I,j])
-
-min_index!3 = [0,1,...N!0-1]
-
-for i in range(0, N!0):
-  min_sum!2 = PHI(min_sum!1, min_sum!4)
-  !1!2[i] = CMP(sum!3[i],min_sum!2)
-  min_sum!4 = MUX(!1!2[i], sum!3[i], min_sum!2)
-
-for i in range(0, N!0):
-  min_index!2 = PHI(min_index!1, min_index!4)
-  min_index!4 = MUX(!1!2[i], min_index!3[i], min_index!2)
-!2!1 = (min_sum!2, min_index!2)
-\end{verbatim}
-}
-\end{comment}
-
-In MPC backends, executing $n$ operations ``at once'' in a single SIMD operation costs a lot less than executing those $n$ operations one by one.
-This is particularly important when there is communication (i.e., in remote), since many 1-bit values are sent at once rather than sequentially.
-%In MPC compilers a vectorized operation that computing M operations "at once" costs
-%essentially the same ($\alpha$ or $\beta$) as an individual operation.
-We elaborate on the cost model in the following section but for now consider that
-each operation has a \emph{fixed} portion (does benefits from amortization) and
-a \emph{variable} portion (does not benefit from amortization): $\alpha = \alpha_\mathit{fix} + \alpha_\mathit{var}$.
-This gives rise to the following formula for amortized cost: $f(n) = \alpha_\mathit{fix} + n\alpha_\mathit{var}$,
-as opposed to unamortized cost $g(n) = n\alpha_\mathit{fix} + n\alpha_\mathit{var}$. (We extend the same reasoning to
-$\beta$-instructions.)
-
-Thus, the fixed cost of the vectorized program amounts to
-$2\alpha_\mathit{fix}$ + $D \beta_\mathit{fix} + N(3\alpha_\mathit{fix})$. (The variable cost is the same in both the vectorized and non-vectorized programs.)
-The first term in the sum corresponds to the vectorized
-subtraction and multiplication, the second term corresponds to the FOR loop on $j$ and the third
-one corresponds to the remaining FOR loops on $i$.
-Clearly, $2\alpha_\mathit{fix} + D\beta_\mathit{fix} + N(3\alpha_\mathit{fix}) << ND(2\alpha_\mathit{fix}+\beta_\mathit{fix}) + N3\alpha_\mathit{fix}$.
-Empirically, we observe 10x to 50x improvement over un-vectorized Biometric Matching in circuit generation time and setup time and 5x to 30x improvement in online time.
-Additionally, the un-vectorized version runs out of memory for $N=256$, while the vectorized one runs with the standard maximal input size $N=4,096$.
-\ana{Edit numbers with final experiments.}
-
-\section{Analytical Model}
-\label{sec:model}
-
-\ana{We need an intro to section here.}
-
-\subsection{Scheduling in MPC}
-\label{sec:mpc}
-
-%\ana{Here we need more on MPC before jumping to scheduling. MPC-source, basic assumptions about static loop bounds, etc.}
-
-For this treatment we make the following simplifying assumptions:
-
-\begin{enumerate}
-\item All statements in the program execute using the same protocol (sharing). That is, there is no share conversion.
-%\ishaq{This is not an assumption, this is our setting. We work in the single protocol world.}
-\item There are two tiers of MPC instructions, local and remote. A local instruction (essentially just ADD) has cost $\beta$
-and a remote instruction (e.g., MUX, MUL, SHL, etc.) has cost $\alpha$, where $\alpha >> \beta$. We assume that all remote
-instructions have the same cost.
-%\ishaq{Don't we need to distinguish between local (cheap) vs.
-interactive (expensive) instructions?
-I don't see how we could assume this.
--- Past suggestion from Vassilis: Ideally we should assume cost to be a convex function.}\ana{After discussion 1/12: replace single costs 1 with two levels of costs: $\alpha$ for non-local operations, e.g., MUL, and $\beta$ for local ops, e.g., ADD.}
-\item  In MPC frameworks, executing $n$ operations ``at once'' in a single SIMD operation costs a lot less than executing those $n$ operations one by one.
-Following Amdahl's law~\ana{citation!}, we write $\alpha = \frac{1}{s}p\alpha + (1-p)\alpha$, where $p$ is the fraction of execution time that benefits from amortization and $(1-p)$
-is the fraction that does not, and $s$ is the available resource. Thus, $n\alpha = \frac{n}{s}p\alpha + n(1-p)\alpha$.
-For the purpose of the model we assume that $s$ is large enough and the term $\frac{n}{s}p\alpha$ amounts to a \emph{fixed cost} incurred regardless of
-whether $n$ is $10,000$ or just $1$. (This models the cost of sending a packet from party A to party B for example.) The bottom line is that amortized execution
-of $n$ operations is $f(n) = \alpha_\mathit{fix} + n\alpha_{var}$ in contrast to unamortized execution $g(n) = n\alpha_\mathit{fix} + n\alpha_{var}$ and, of course,
-$\alpha_\mathit{fix} << n\alpha_\mathit{fix}$.
-%We assume infinite parallel capacity---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely $\alpha$ or $\beta$.
-%This is a standard assumption in Cryptographic Parallel RAM. ABY presents empirical support for this assumption~\ana{Add citations. PRAM, ABY.}
-%\ishaq{Comment from Vassilis: We assume there are infinite parallel capacities, this is the standard assumption in PRAM (Cryptographic Parallel RAM).}\ana{PRAM assumption stays.
-We have to replace the unit cost of 1 with a suitable amortization function $f(n)$, which I don't think changes much in the cost modeling and analysis.}
-\item MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption,
-2 MUL instructions can be amortized in a single SIMD instruction that costs $\alpha_\mathit{fix} + 2\alpha_\mathit{var}$, however a MUL and a MUX instruction
-still cost $2\alpha_\mathit{fix} + 2\alpha_\mathit{var}$ even when scheduled ``in parallel''.
-%MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption,
-%2 MUL instructions scheduled in parallel benefit from amortization and cost $\alpha$, however a MUL and a MUX instructions scheduled
-%in parallel still cost $2\alpha$.
-%\ishaq{We need to reword this assumption to something like "$K$ parallel MUL costs much less than $K$ (because they get amortized), but any mix of $K$  MUL and MUX still cost roughly $K$.
-Specifically, we should take away the low constants (2 and 1) because we know for low constants this is not true.} \ana{Again, core assumption that MUL and MUX don't benefit from amortization stays.
-We have to replace constant costs with functions, as in (3).}
-\end{enumerate}
-
-\input{prob_stmt_np_hardness.tex}
-
 \begin{figure*}
@@
 Should this be placed somewhere else?}
-
-\section{Vectorization}
-\label{sec:vectorization}
-
-\ana{Need an intro here!}
-
-\subsection{Scalar Expansion}
-
-An important component of our algorithm is the scalar expansion to the corresponding loop dimensionality.
-For example, {\sf d = S[i*D+j] - C[j]} equiv. to {\sf d = S[i,j] - C[j]}, which gave rise to {\sf N*D} subtraction operations in the sequential schedule,
-is lifted. The argument arrays {\sf S} and {\sf C} are lifted and the scaler {\sf d} is lifted: {\sf d[i,j] = S[i,j] - C[i,j]}.
-The algorithm then detects that the statement can be vectorized.
-
-There are three kinds of arrays (for now all kept internally as one-dimensional arrays, but that's under discussion).
-\begin{itemize}
-
-\item Scalars: These are scalar variables we lift into arrays for the purposes of vectorization.
-For those, all writes are canonical writes and all reads are canonical reads. We will apply raise dimension when
-a scalar is used in an inner loop (e.g., {\sf sum!2} in line 6 of the MPC source code will be raised to a 1-dimensional
-array since {\sf sum!2} is used in the inner $j$-loop). Drop dimension applies as well; this happens when a scalar
-written in an inner loop is used outside of the loop (e.g., {\sf sum!2} for which the lifted inner loop computes
-{\sf D} values, but the outer loop only needs the last one.)
-
-\item Read-only input arrays: Read-only inputs. There are NO writes, while we may have non-canonical reads, $f(i,j,k)$.
-Phase 1 of Basic vectorization will add raise dimension operation at the beginning of the function to lift these arrays,
-and raise dimension may \emph{reshape} arrays. If there are multiple "views" of the input array, there would be multiple raise dimension statements to create
-each one of these views. The invariant is that at reads in loops, the reads of "views" of the original input array are canonical.
-Only raise dimension applies to these arrays, and only in the beginning of the program. For example, 1-dimensional array
-\python{C} is lifted into 2-dimensional array {\sf N x D} by copying the row {\sf N} times.
-
- \item Read-write output arrays: Writes are canonical (by restriction) but reads can be non-canonical.
- Dependence analysis limits vectorization when non-canonical read access refers to array writes in
- previous iterations, thus creating loop-carried dependences. We may apply both raise and drop dimension, however,
- they respect the fixed dimensionality of the output array. The array cannot be raised to a dimension lower than
- its canonical (fixed) dimensionality and it cannot be dropped lower. In addition, non-canonical reads may
- require lifting (i.e., reshaping) of the array after the most recent write, rather than in the beginning of the program,
- in order to reduce a non-canonical read to a canonical one.
-
-\end{itemize}
-
-In the sections below we detail the $\mathit{raise\_dim}$ (raise dimension) and $\mathit{drop\_dim}$ (drop dimension) operations,
-followed by our vectorization algorithm.
-
-%\ana{Add back-edges into Phase 1.
-A back-edge from a non-phi-node in loop $i$ to a phi-node in loop $i$'s header is a same-level edge.
-The only difference with the handling of normal same-level forward def-use is that the operand index will become $i-1$.
-A back edge from a phi-node in loop $j$ to a phi-node in loop $i$, where $j$ is nested in $i$ is an inner-to-outer edge and will require dropping dimension.
-We can show these are the only kinds of back-edges that may occur.}
-
-%\subsection{Algorithm (Still work in progress.)}
-
-\subsection{Raise Dimension and Drop Dimension}
-
-There are two conceptual versions of $\mathit{raise\_dim}$. One applies on read-only input arrays and reshapes those arrays
-when necessary to ensure canonical read access in the corresponding loop. The signature of $\mathit{raise\_dim}$
-is as follows. It takes the original array \texttt{C}, the access pattern function $f(i,j,k)$ in loop nest $i,j,k$ and the loop bounds
-$((i:I),(j:J),(k:K))$:
-\[ \mathit{raise\_dim}(\texttt{A}, f(i,j,k), \texttt{((i:I),(j:J),(k:K))}) \]
-It produces a new 3-dimensional array \texttt{A'} by iterating over $i,j,k$ and setting each element of  \texttt{A'} as follows:
-\[ \texttt{A'[i,j,k] =  A[f(i,j,k)]} \]
-The end result is that uses of \texttt{[A[f(i,jk)]} in loop nest $i,j,k$ are replaced with canonical read-accesses to \texttt{A'[i,j,k]}
-that can be vectorized. In the running Biometric example, \texttt{C'} = $\mathit{raise\_dim}$\texttt{(C, j, (i:N,j:D))} lifts the
-1-dimensional array \texttt{C} into a 2-dimensional array. The $i,j$ loop now accesses \texttt{C'} in the canonical way, \texttt{C'[i,j]}.
-Similarly, \texttt{S'} = $\mathit{raise\_dim}$\texttt{(S, i*D+j, (i:N,j:D))} tries to lift \texttt{S}, but the operation turns into a no-op
-because \texttt{S} is already a 2-dimensional array and the read access is canonical.
-
-The other version of $\mathit{raise\_dim}$ applies on scalars and read-write arrays. It lifts a lower-dimension array into a
-higher-dimension for access in a nested loop. Here \texttt{A} is an $i$ array and raise dimension adds two additional dimensions:
-\[ \mathit{raise\_dim}(\texttt{A}, \texttt{(j:J,k:K)}) \]
-This version is reduced to the above version by adding the access pattern function, which is just $i$:
-\[ \mathit{raise\_dim}(\texttt{A}, i, \texttt{(j:J,k:K)}) \]
-
-The corresponding $\mathit{drop\_dim}$ is carried out when an array written in an inner loop is used in an enclosing loop.
-It takes a higher dimensional array, say $i,j,k$ and removes trailing dimensions, say $j,k$:
-\[ \mathit{drop\_dim}(\texttt{A}, \texttt{(j:J,k:K)}) \]
-It iterates over $i$ and takes the result at the maximal index of $j$ and $k$, i.e., the result at the last iterations of $j$ and $k$:
-\[ \texttt{A'[i] = A[i,J-1,K-1]} \]
-
-\begin{comment}
-We define the $\mathit{raise\_dim}$ (raise dimensions) and $\mathit{drop\_dim}$ (drop dimension) functions.
-Raise dimension "lifts" a lower-dimension array (\ana{The right term here is tensor, I am pretty sure!}) into a higher dimension one.
-This is necessary when a lower-dimensional array is used in a higher dimensional loop and, essentially, is just copying of values.
-For example, Biometric contains the statement $d = \mathit{SUB}(S[i, j], C[j])$, in the $j$-loop which is nested into the $i$-loop.
-Here $C$ is a one-dimensional array, however, to vectorize across both loops it is necessary to turn it into a two-dimensional array: $C[i,j]$ becomes $[C[0],C[1],...C[J],C[0],C[1],...C[J], ...
-C[0],C[1],...C[J]]$, which turns the row into a matrix of $I$ identical rows.
-(We use capital letters to denote the upper bounds of loops, e.g., $J$ is the upper bound of the $j$-loop and $I$ is the upper bound of the $i$-loop.)
-
-$\mathit{raise\_dim}(A[i_1,...,i_k],i_j)$ is defined as follows.
-It results in a new {k+1}-dimensional array $A'$ where for every $0 \le i_{k} < I_{k+1}$, $A'[i_1, \ldots, i_j, \ldots, i_k] = A[i_1, \ldots, i_k]$.
-Adding $n$ dimensions is trivially extended as a composition of $n$ $\mathit{raise\_dim}$ that each adds a single dimension.
-
-As expected, drop dimension turns a higher-dimensional array into a lower-dimensional one.
-The key use case is an inner-to-outer def-use edge.
-The code may define a variable, e.g., $\x$ in an inner loop, say $j$, then use this variable in an enclosing loop, say $i$.
-Our algorithm may vectorize the computation of $\x$ in the $j$-loop thus producing a vector $\x[j]$ where $\x[1]$ is the value of \x after the 1st iteration and so on.
-Drop dimension states that the outer loop will use the value of the variable at the last iteration.
-
-$\mathit{drop\_dim}(A[i_1,\ldots, i_j, \ldots,i_k], i_j)$ produces a k-dimensional array $A'$ where $A'[i_1,\ldots,i_k] = A'[i_1,\ldots, I_j-1, \ldots,i_k]$.
-In our analysis, dimensions are always dropped at the end, and again, one can define dropping $n$ dimensions as a composition of $n$ $\mathit{drop\_dim}$.
-
-\ben{I'm not sure if I wrote up the above paragraphs correctly -- the idea is that we're just copying values to extend over dimension $j$ or only retain its last elements.
-The current writeup implies that $1 < j < k$ though, which isn't always true.
-Python implementations of $\mathit{raise\_dim}()$ and $\mathit{drop\_dim}()$ can be found here: \url{https://github.com/milana2/ParallelizationForMPC/blob/master/compiler/compiler/motion_backend/reference_implementations.py}}
-
-\ana{This section is still informal. Needs work to make more precise.}
-\end{comment}
-
-\subsection{Basic Vectorization}
-
-\begin{algorithmic}
-
-\STATE \COMMENT { Phase 1: Raise dimension of scalar variables to corresponding loop nest. We can traverse stmts linearly in MPC-source. }
-
-\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i,j,k$}
-%\FOR {each $Y_n$}
-\FOR {each argument $Y_n$} %def-use edge $\mathit{stmt}' (\mbox{def of } Y) \rightarrow \mathit{stmt} (\mbox{def of } X)$}
-\STATE {case def-use edge $\mathit{stmt}' (\mbox{def of } Y_n) \rightarrow \mathit{stmt} (\mbox{def of } X)$ of} \\
-\STATE \hspace{0.25cm} {same-level: $Y'_n$ is $Y_n$} \\
-\STATE \hspace{0.25cm} {outer-to-inner: add $Y'_n[i,j,k] = \mathit{raise\_dim}(Y_n)$ at $\mathit{stmt'}$}
-\STATE \hspace{0.25cm} {(more precisely, right after $\mathit{stmt'}$)}
-\STATE \hspace{0.25cm} {inner-to-outer: add $Y'_n[i,j,k] = \mathit{drop\_dim}(Y_n)$ at $\mathit{stmt}$}
-\STATE \hspace{0.25cm} {(more precisely, in loop of $\mathit{stmt}$ right after loop of $\mathit{stmt'}$)}
-\ENDFOR
-
-\COMMENT { Optimistically vectorize all. $I$ means vectorized dimension. }
-%\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i,j,k$}
-\STATE {change to $X[I,J,K] = \mathit{Op}(Y'_1[I,J,K],Y'_2[I,J,K])$ }
-\ENDFOR
-
-\COMMENT { Phase 2: Recreating FOR loops for cycles; vectorizable statements hoisted up. }
-
-\FOR {each dimension $d$ from highest to 0}
-\FOR {each $\phi$-node $n$ in loop $i_1,...,i_d$}
-\STATE {compute $\mathit{closure}(n)$}
-\ENDFOR
-
-\COMMENT {{\color{red} $cl_1$ and $cl_2$ intersect if they have common statement or update same array; "intersect" definition can be expanded}}
-
-\WHILE {there are closure $cl_1$ and $cl_2$ that intersect}
-\STATE {merge $cl_1$ and $cl_2$}
-\ENDWHILE
-\FOR {each closure $cl$ (after merge)}
-\STATE {create FOR $i_d = 0; ...$ loop}
-\STATE {add $\phi$-nodes in $cl$ to header block}
-\STATE {{\color{red}{add target-less $\phi$-node for A if $cl$ updates array A}}}
-\STATE {add statements in $cl$ to loop body in some order of dependences}
-\STATE \COMMENT { Dimension is not vectorizable: }
-\STATE {change ${I_d}$ to $i_d$ in all statements in loop}
-\STATE {treat FOR loop as monolith node: some def-use edges become same-level.}
-\ENDFOR
-{
-\color{red}
-\FOR {each target-less $\phi$-node $A_1 = \phi(A_0,A_k)$}
-\STATE {in vectorizable stmts, replace use of $A_1$ with $A_0$}
-\STATE {discard $\phi$-node if not used in any $cl$, replacing $A_1$ with $A_0$ or $A_k$ appropriately}
-\ENDFOR
-}
-\ENDFOR
-
-\COMMENT { Phase 3: Remove unnecessary dimensionality. }
-
-\COMMENT { A dimension $i$ is dead on exit from stmt $X[...i...] = ...$ if all def-uses with targets outside of the enclosing
-FOR $i=0 ...$ MOTION loop end at target (use) $X' = \mathit{drop\_dim}(X, i)$. }
-\FOR {each stmt and dimension $X[...i...] = ...$}
-\STATE {if $i$ is a dead dimension on exit from stmt $X[...i...] = ...$, remove $i$ from $X$ (all defs and uses)}
-%\COMMENT{ A stmt v[...i] is a "dead dimension i" if all edges out of FOR i = 0; ... loop end at drop\_dim(v, i)}
-\ENDFOR
-
-\COMMENT { Now clean up drop\_dim and raise\_dim }
-\FOR {each $X' =  \mathit{drop\_dim(X,i)}$ }
-\STATE {replace with $X' = X$ if $i$ is dead in $X$.}
-\ENDFOR
-\STATE {do (1) (extended) constant propagation, (2) copy propagation and (3) dead code elimination to get rid of redundant variables and raise and drop dimension statements}
-
-\COMMENT { Phase 4: }
-\STATE {add SIMD for simdified dimensions}
-
-\end{algorithmic}
-
-\input{biometric_vectorization_example.tex}
-
-\begin{comment}
-\subsection{Example: Biometric}
-
-\subsubsection{MPC-source}
-
-Below is the MPC-source resulting from Benjamin's analysis. It gives rise to a corresponding dependence graph. (Not shown.)
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0 = \phi(-1, \texttt{index}_1)$}
-\STATE {$\texttt{min}_0 = \phi(10000,\texttt{min}_1)$}
-
-\STATE \COMMENT {Begin of inner loop $j$.}
-\STATE {$\texttt{sum}_0 = \phi(0, \texttt{sum}_1)$}
-\STATE {$\texttt{d} = \mathit{SUB}(\texttt{S}[i,j], \texttt{C}[j])$}
-\STATE {$\texttt{p} = \mathit{MUL}(\texttt{d},\texttt{d})$}
-\STATE {$\texttt{sum}_1 = \texttt{ADD}(\texttt{sum}_0,\texttt{p})$}
-\STATE \COMMENT {End of inner loop $j$.}
-\STATE {$\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$}
-\STATE {$\texttt{index}_1 = \mathit{MUX}(\texttt{c},\texttt{index}_0,i)$}
-\STATE {$\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{min}_0,\texttt{sum}_0)$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-\subsubsection{Phase 1 of Basic Vectorization:}
-
-The transformation preserves the dependence edges.
-It raises the dimensions of scalars and optimistically vectorizes all operations.
-The next phase discovers loop-carried dependences and removes affected vectorization.
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$}
-\STATE {$\texttt{min}_0[i] = \phi(10000, \texttt{min}_1[i-1])$}
-
-\STATE \COMMENT {Begin of inner loop $j$.}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...], \texttt{sum}_1[i,j-1])$}
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{sum}_1[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE \COMMENT {End of inner loop $j$.}
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {$\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$}
-\STATE {$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$}
-\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-\ishaq{At least for input variables: when raising dimensions, we should keep some meta data around so we can later write code that could use 1 share if it knows the other N shares are just duplicates of it.}
-
-\subsubsection{Phase 2}
-
-This phase analyzes statements from the innermost loop to the outermost. The key point is to discover loop-carried dependencies and re-introduce loops whenever dependencies make this necessary.
-
-Starting at the inner phi-node $\texttt{sum}_0[i,j] = \phi([0,0,...],\texttt{sum}_1[i,j-1])$, the algorithm first computes its closure.
-The closure amounts to the phi-node itself and $\texttt{sum}_1[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$, accounting for the loop-carried dependency of the computation of \texttt{sum}.
-The algorithm replaces this closure with a FOR loop on $j$ removing vectoriaztion on $j$.
-Note that the SUB and MUL computations remain outside of the loop as they do not depend on phi-nodes that are part of cycles.
-The algorithm adds same-level edges, one from $\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = ...$ to the monolithic FOR-loop node, and one from the FOR-loop node to $\texttt{c}[\stackrel{\rightarrow}{i}] = ...$.
-
-\ishaq{For our radar: Such dependencies will not involve phi nodes only. e.g. $a[i] = a[i] + a[i-1]$}
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0[i] = \phi(-1, \texttt{index}_1[i-1])$}
-\STATE {$\texttt{min}_0[i] = \phi(10000, \texttt{min}_1[i-1])$}
-
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-
-\STATE \COMMENT {Begin of inner loop $j$. Loop is now much "shorter"!}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...], \texttt{sum}_1[i,j-1])$}
-\STATE {FOR j=0; j<D; j++} \COMMENT{ Will turn into a MOTION loop.}
-\STATE \hspace{0.25cm}{$\texttt{sum}_1[\stackrel{\rightarrow}{i},{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},{j}],\texttt{p}[\stackrel{\rightarrow}{i},{j}])$}
-\STATE \COMMENT {End of inner loop $j$.}
-
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {$\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$}
-\STATE {$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$}
-\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-Next, analysis moves to outer dimension $i$. There are two phi-nodes, $\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1])$ and $\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$.
-The closure of the first is
-\[
-\begin{array}{l}
-\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1]) \\
-\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
-\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])
-\end{array}
-\]
-and the closure of the second one is
-\[
-\begin{array}{l}
-\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1]), \\
-%\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
-\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])
-\end{array}
-\]
-
-Since the two closures \emph{do not} intersect, we have two distinct FOR-loops on $i$.
-The first FOR loop (on \texttt{min}) is scheduled first, because of the dependence edge from $\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$ to
-$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$.
-We can now rewrite those statements, getting rid of the vectorization on $i$ in the two FOR loops.
-Notably, the SUB and MUL computations are fully vectorizable and the ADD computation is vectorizable across the $i$-dimension.
-
-\begin{algorithmic}
-
-\STATE
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB\_SIMD}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL\_SIMD}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$} \\
-\STATE {}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...],\texttt{sum}_1[i,j-1])$}
-\STATE {FOR j=0; j<D; j++} \COMMENT{ Turns into a MOTION loop, was $j$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{sum}_1[\stackrel{\rightarrow}{i},{j}] = \mathit{ADD\_SIMD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},{j}],\texttt{p}[\stackrel{\rightarrow}{i},{j}])$}
-\STATE {}
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {}
-\STATE {$\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1])$}
-\STATE {FOR i=0; i<N; i++} \COMMENT{ Turns into a MOTION loop, was $i$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{c}[{i}] = \mathit{CMP}(\texttt{sum}'_0[{i}],\texttt{min}_0[{i}])$}
-\STATE \hspace{0.25cm}{$\texttt{min}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{min}_0[{i}],\texttt{sum}'_1[{i}])$}
-\STATE {}
-\STATE {$\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$}
-\STATE {FOR i=0; i<N; i++} \COMMENT{ Turns into a MOTION loop, was $i$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{index}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{index}_0[{i}],i)$}
-
-\ben{The above code implicitly stores the final values of min\_0 and index\_0 in their last indices, but there is no primitive like drop\_dim to make it clear.
-This will have to be added in the SSA code generation (though it might not be necessary to include in the paper?).
-It also is pretty wishy-washy about the size and layout of each vector (again more of an implementation issue), but maybe we should have a different syntax for array accesses for read/write (e.g.
-the ADD\_SIMD steps) vs the array definitions (e.g.
-the SUB\_SIMD and raise/drop\_dim statements)}
-
-\end{algorithmic}
-\end{comment}
-
-\subsection{Correctness Argument}
-
-We build a correctness argument that loosely follows the theory of Abstract Interpretation.
-We define the syntax of MPC Source programs. The domain of MPC Source programs
-expressible in the syntax (with certain semantic restrictions) is the abstract domain $A$.
-We then define the \emph{linearization} of an MPC Source program as an interpretation
-over the syntax. The linearization, which is a \emph{schedule}, is the concrete domain $C$.
-Since we reason over def-use graphs in $A$ we define a partial order relation over elements of $A$
-in terms of def-use relations. We define a partial order over elements of $C$ as well,
-in terms of def-use relations in the concrete domain $C$. We prove two theorems
-that state (informally) that the schedule corresponding to the original program computes the same
-result as the schedule corresponding to the vectorized program.
-
-\begin{figure}%[!b]
-%\small
-\[
-\begin{array}{l@{~~~~~}l}
-  \begin{array}{l@{~}l@{~~~}l}
-  s &::= s;s & sequence \\
-     & \mid \x[i,J,k] = \y[i,J,k] \; \mathit{op\_SIMD} \; \z[i,J,k] & operation \\
-     & \mid \x[i,J,k]  = \texttt{PHI}(\x_1[i,J,k],\x_2[i,J,k\!-\!1]) & \mathit{phi} \; \mathit{node} \\
-     & \mid \x[i,J,k] = \mathit{raise\_dim}(\x'[i],(J\!:\!J,k\!:\!K)) & \mathit{raise} \; \mathit{dimension(s)} \\
-     & \mid \x[i,J] = \mathit{drop\_dim}(\x'[i,J,k],k)  & \mathit{drop} \; \mathit{dimension(s)} \\
-     & \mid \x = \y & \mathit{propagation} \\
-     & \mid \texttt{FOR} \; 0 \le i < I: s  & \mathit{loop} \\
-  \end{array}    \\
-\end{array}
-\]
-\vspace{-3mm}
-\caption{MPC Source Syntax}\label{fig:MPC_Source_syntax}
-\vspace{-5mm}
-\end{figure}
-
-\paragraph{MPC Source Syntax}
-
-\figref{fig:MPC_Source_syntax} defines the syntax for our intermediate representation, MPC Source.
-There are semantics restrictions over the syntax as well: a variable $\x[i,j,k]$ is a 3-dimensional array
-$(i:I,j:J,k:K)$ and also, a statement $\x[i,J,k] = ...$ is enclosed in loops over $i$ and $k$ as shown below.
-Thus, $i$ and $k$ are in scope.
-\begin{verbatim}
-FOR 0 <= i < I:
-    ...
-    FOR 0 <= k < K:
-        x[i,J,k] = ...
-\end{verbatim}
-
-Statements $\mathit{operation, phi, raise \; dimension(s), drop \; dimension(s)}$ are \emph{base} statements,
-and $\mathit{sequence, loop}$ are \emph{compound statements}.
-
-\paragraph{Linearization}
-
-Linearization is the concretization operation, which, as we mentioned earlier computes a schedule.
-The concretization function $\gamma: A \rightarrow C$ is defined as an interpretation of
-MPC Source syntax, as it is standard. The concretization of each one
-of the base statements is as follows:
-\[\begin{array}{l}
-\gamma(\x[i,J,k] = \mathit{op\_SIMD}(\y[i,J,k], \z[i,J,k])) = \\
-~~\x[i,0,k] = \y[i,0,k] \; \mathit{op} \; \z[i,0,k] \;\; || \\
-~~~~\x[i,1,k] = \y[i,1,k] \; \mathit{op} \; \z[i,1,k] \;\; || \; ...\; || \\
-~~~~~~\x[i,I-1,k] = \y[i,I-1,k] \; \mathit{op} \; \z[i,I-1,k]
-\end{array}\]
-meaning that the vectorized dimension(s) are expanded into parallel statements. || introduces SIMD (parallel) execution.
-
-The concretization of the FOR statement is as follows:
-\[ \gamma(\texttt{FOR} \; 0 \le i < I: s) = \gamma(s)[0/i] \; ; \; \gamma(s)[1/i] \; ; \; ... \gamma(s)[I-1/i] \]
-$\gamma$ simply unrolls the loop substituting $i$ with 0, 1, etc. Here ; denotes sequential execution.
-
-\paragraph{Partial Orders}
-
-For each MPC Source program $a$ we compute the def-use edges in the standard way: if base statement $s1 \in a$ defines variable \x, e.g., $\x[i,j,k] = ...$,
-and base statement $s2 \in a$ uses \x, e.g., $... = ... \x[i,j,k]$ and there is a path in the trivial CFG from $s1$ to $s2$, then there is a def-use edge from
- $s1$ to $s2$. We extend the dimensionality of a statement into $s1[i,j,k]$ where $s1[i,j,k]$ inherits the dimensionality of the left-hand-side of the assignment.
-
- Let $a_0, a_1$ be two MPC Source programs in $A$.
- Two base statements, $s_0 \in a_0$ and $s_1 \in a_1$ are \emph{same}, written $s_0 \equiv s_1$ if they are of the same operation and they operate on
- the same variables: same variable name and same dimensionality. Recall that dimensions in MPC Source are either iterative, lower case, or vectorized, upper case.
- Two statements are same even if one operates on an iterative dimension and the other one operates on a vectorized one, e.g., $s_0[i,j,k] \equiv s_1[I,j,K]$.
-
- \begin{definition} Let $a_0, a_1 \in A$. We say that $a_0 \le a_1$ iff for every def-use edge $s1 \rightarrow s2$ in $a_0$ there is an edge
- $s1' \rightarrow s2'$ where $s1 \equiv s1'$, $s2 \equiv s2'$ and the two edges of either both forward or both backward.
- \end{definition}
-
- The def-use edges in the concrete schedule are as expected. There is a def use edge from statement $s1$ that defines $\x[\underline{i},\underline{j},\underline{k}]$
- to statement $s2$ that uses $\x[\underline{i},\underline{j},\underline{k}]$ if $s1$ is scheduled ahead of $s2$ in the linear schedule. We note that the underlined
- indices, e.g., $\underline{i}$, refer to fixed values, not iterative or vectorized dimensions since in the concrete schedule all induction variables are expanded.
- E.g., there is a def-use edge from the statement that defines $\x[0,1,2]$ and a statement that uses $\x[0,1,2]$.
-
-\paragraph{Theorems}
-
-\begin{theorem} $a_0 \le a_1 \Rightarrow \gamma(a_0) \subseteq \gamma(a_1)$.
-\end{theorem}
-
-\begin{theorem} Let $a_0$ be the iterative MPC Source and let $a_1$ be the vectorized MPC Source computed by the vectorization algorithm.
-We have that $a_0 \le a_1$.
-\end{theorem}
-
-\ana{Write the proof sketch and final argument, etc.}
-
-\subsection{Extension of Basic Vectorization with Array Writes}
-
-\subsubsection{Removal of Infeasible Edges}
-
-Array writes limit vectorization as they sometimes introduce infeasible loop-carried dependencies. Consider the following example: \ana{Have to add citation to Aiken's paper}
-
-{\small
-\begin{verbatim}
-for i in range(N):
-  A[i] = B[i] + 10;
-  B[i] = A[i] * D[i-1];
-  C[i] = A[i] * D[i-1];
-  D[i] = B[i] * C[i];
-\end{verbatim}
-}
-
-In Cytron's SSA this code (roughly) translates into
-
-{\small
-\begin{verbatim}
-for i in range(N):
-1. A_1 = PHI(A_0,A_2)
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. A_2 = update(A_1, i, B_1[i] + 10);
-6. B_2 = update(B_1, i, A_2[i] * D_1[i-1]);
-7. C_2 = update(C_1, i, A_2[i] * D_1[i-1]);
-8. D_2 = update(D_1, i, B_2[i] * C_2[i]);
-\end{verbatim}
-}
-
-\begin{comment}
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE \hspace{0.25cm} A[i] = B[i] + 10;
-\STATE \hspace{0.25cm} B[i] = A[i] * D[i-1];
-\STATE \hspace{0.25cm} C[i] = A[i] * D[i-1];
-\STATE \hspace{0.25cm} D[i] = B[i] * C[i];
-\end{algorithmic}
-
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{A}_0$ = $\phi(\texttt{A},\texttt{A}_1)$
-\STATE 2. $\texttt{B}_0$ = $\phi(\texttt{B},\texttt{B}_1)$
-\STATE 3. $\texttt{C}_0$ = $\phi(\texttt{C},\texttt{C}_1)$
-\STATE 4. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$
-\STATE 5. $\texttt{A}_1$ = $\texttt{A}_0$; $\texttt{A}_1$[i] = $\texttt{B}_0$[i] + 10; \COMMENT{equiv. to Update}
-\STATE 6. $\texttt{B}_1$ = $\texttt{B}_0$; $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 7. $\texttt{C}_1$ = $\texttt{C}_0$; $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 8. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
-\end{algorithmic}
-\end{comment}
-
-There is a cycle around \texttt{B\_1 = PHI(B\_0,B\_2)} that includes statement
-\texttt{A\_1 = update(A\_0, i, B\_1[i] + 10)} and that statement won't be vectorized even
-though in fact there is no loop-carried dependency from the write of \texttt{B\_1[i]} at 6 to the
-read of \texttt{ ... = B\_1[i]} at 8.
-
-The following algorithm removes certain infeasible loop-carried dependencies that are due to array writes. Consider a loop with index $0 \le j < J$
-nested at $i,j,k$. Here $i$ represents the enclosing loops of $j$ and $k$ represents the enclosed loops in $j$.
-
-\begin{algorithmic}
-\FOR {each array $\texttt{A}$ written in loop $j$}
-\STATE \COMMENT { including enclosed loops in $j$ }
-\STATE dep = False
-\FOR {each pair def: $\texttt{A}_m[f(i,j,k)] = ... $, and use: $ ... = \texttt{A}_n[f'(i,j,k)]$ in loop $j$}
-\IF {$\exists \underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$, s.t.
-$0\!
-\le\!
-\underline{i}\!
-<\!
-I,\: 0\!
-\le\!
-\underline{j}, \underline{j}' \!
-<\!
-J,\: 0\!
-\le\!
-\underline{k}, \underline{k}' \!
-<\!
-K,\: \underline{j}<\underline{j}'$, and $f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$}
-\STATE dep = True
-\ENDIF
-\ENDFOR
-\IF {dep == False}
-\STATE remove back edge into $\texttt{A}$'s $\phi$-node in loop $j$.
-\ENDIF
-\ENDFOR
-\end{algorithmic}
-
-Consider a loop $j$ enclosed in some fixed $\underline{i}$. Only if an update (definition) $\texttt{A}_m[f(i,j,k)] = ... $ at some iteration $\underline{j}$
-references the \emph{same} array element as a use $ ... = \texttt{A}_n[f'(i,j,k)]$ at some later iteration $\underline{j}'$,
-we may have a loop-carried dependence for $\texttt{A}$ due to this def-use pair. (In contrast, Cytron's algorithm inserts a loop-carried dependency every time there is an array update.)
-The algorithm above examines all def-use pairs in loop $j$, including defs and uses in nested loops, searching for values $\underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$ that satisfy
-$f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$. If such values exist for some def-use pair, then there is a potential
-loop-carried dependence on $\texttt{A}$; otherwise there is not and we can remove the spurious backward edge thus ``freeing up'' statements for vectorization. %\ana{Double check this!!!}
-
-Consider the earlier example.
-There is a single loop, $i$.
-Clearly, there is no pair $\underline{i}$ and $\underline{i}'$, where $\underline{i} < \underline{i}'$ that make $\underline{i} = \underline{i}'$
-due to the def-use pairs of \texttt{A} 5-6 and 5-7.
-Therefore, we remove the back edge from 5 to the phi-node 1.
-Analogously, we remove the back edges from 6 to 2 and from 7 to 3.
-However, there are many values $\underline{i} < \underline{i'}$ that make $\underline{i} = \underline{i'}-1$ and the back edge from 8 to 4 remains (def-use pairs for \texttt{D}).
-As a result of removing these spurious edges, Basic Vectorization will find that statement 5 is vectorizable.
-Statements 6, 7 and 8 will correctly appear in the FOR loop.
-
-Note however, that this step renders some array phi-nodes target-less. We handle target-less phi-nodes with a minor extension of Basic Vectorization (Phase 2).
-First, we merge closures that update the same array. This simplifies handling of array $\phi$-nodes: if each closure is turned into a separate loop
-each loop will need to have its own array phi-node to account for the update and this would complicate the analysis.
-Second, we add the target-less node of array \texttt{A} back to the closure that updates \texttt{A} ---
-the intuition is, even if there is no loop-carried dependence from writes to reads on \texttt{A}, \texttt{A} is written and the write (i.e., update) cannot be vectorized;
-therefore, the updated array has to carry to the next iteration of the loop. Third, in cases when the phi-node remains target-less, i.e., cases when the
-array write can be vectorized, we have to properly remove the phi-node replacing uses of the left-hand side of the phi-node with its arguments.
-
-%\ana{Commented out previous writeup. Removal/handling of targetless phi-nodes will go into the Extension to Basic vectorization.}
-
-\begin{comment}
-The final step is to remove array phi-nodes that are left without incoming back edges after removal of infeasible edges.
-Such a phi-node is redundant when there are no loop-carried dependences that are due to writes to the corresponding array.
-In our running example, the phi-nodes for \texttt{A}, \texttt{B}, and \texttt{C}, are redundant.
-
-Removing redundant phi-nodes proceeds from \emph{innermost to outermost} loops as follows.
-Denote a redundant phi-node by $\texttt{A}_1 = \phi(\texttt{A}_0,\texttt{A}_k)$. $\texttt{A}_0$ is the array at the beginning of the loop
-and $\texttt{A}_k$ is the array at the end of an iteration (including the last iteration). We consider two cases:
-\begin{enumerate}
-\item[Case 1] Inner-to-outer or same-level def-use edge
-\begin{itemize}
-\item If use is an array read $ ... = \texttt{A}_1[f(i)]$, then replace $\texttt{A}_1$ with $\texttt{A}_0$. This is correct, i.e., it is guaranteed that the use
-references $\texttt{A}_0$; had the use been referencing an array modified in an earlier iteration, the back-edge to the phi-node would
-not have been removed and the phi-node would not have been classified as redundant.
-\item If use is an array write (update) $\texttt{A}_2 = \texttt{A}_1; \texttt{A}_2[i] = ... $ equiv. to $\texttt{A}_2 = \mathit{update}(\texttt{A}_1, i, ...)$ then handling becomes more tricky.
-The semantics of this update is that when $i$ is 0, then $\texttt{A}_0$ is being read.
-But when $i > 0$, then $\texttt{A}_k$ is being read.
-If we are at this case, this means that the def-use edge is same-level and we are removing a redundant node in the loop enclosing the update.
-The removal of the original phi-node entails pushing the phi-node at the update:
-$\texttt{A}_2 = \phi(\texttt{A}_0,\texttt{A}_k); \texttt{A}_2[i] = ... $ equiv. to $\texttt{A}_2 = \mathit{update}(\phi(\texttt{A}_0,\texttt{A}_k), i, ...)$.
-\item If use is a phi-node $... = \phi(\texttt{A}_1,...)$, that means we have an outer-to-inner edge. The
-
-\end{itemize}
-\end{enumerate}
-For each outer-to-inner or same-level def-use edge
-that originates at the node, replace $\texttt{A}_1$ at the use node by $\texttt{A}_0$.
-For each inner-to-outer def-use edge, replace $\texttt{A}_1$ by $\texttt{A}_k$. In this case, the use references the array
-that results from the computation in the loop.
-
-Our running example is transformed as follows. Removing the phi-nodes for \texttt{A}, \texttt{B}, and \texttt{C} transforms the
-replaces the references to $\texttt{A}_0$, $\texttt{B}_0$, and $\texttt{C}_0$, with references to $\texttt{A}$, $\texttt{B}$, and $\texttt{C}$
-respectively, i.e., the arrays before the loop.
-
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$
-\STATE 2. $\texttt{A}_1$ = $\texttt{A}$; $\texttt{A}_1$[i] = $\texttt{B}$[i] + 10;
-\STATE 3. $\texttt{B}_1$ = $\texttt{B}$; $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 4. $\texttt{C}_1$ = $\texttt{C}$; $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
-\end{algorithmic}
-\end{comment}
-
-\begin{comment}
-\subsubsection{Array MUX refinement}
-
-\ana{TODO: WIP, get to BV without MUX refinement for now? For now, skip MUX refinement.}
-
-Next, the algorithm refines array MUX statements.
-MPC-source after Cytron's SSA may result in statements $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$, which imply that any index
-of $A$ can be written at this point and therefore there is a loop-carried dependency. In some cases the MUX can be refined to just a single
-index or a pair of indices, e.g., $\texttt{A}_j[i] = \emph{MUX}(c, \texttt{A}_k[i], \texttt{A}_l[i])$.
-
-%\ana{Clarify why this helps. Still work in progress.}
-This is to reduce the dimensionality of simd-ified computation. Technically, $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$ is
-a simdified operation that can be carried out in parallel "in one round". However, particularly when \texttt{A} is a multi-dimensional array, there is
-substantial increase in the size of the arrays (vectors) we send to SIMD operations. Refining to an update to a specific index would reduce
-the size of those vectors. Note that this is a heuristic that handles a common case, but not all cases of array updates.
-%\ana{TODO: Simplify this algorithm, taking into account the restriction to canonical updates. It should be handling \_all\_ cases.}
-
-\begin{algorithmic}
-\FOR {each $\mathit{stmt}$: $\texttt{A}_j = \emph{MUX}(c, \texttt{A}_k, \texttt{A}_l)$ in the MPC-source seq.}
-\STATE $i_1 = \mathit{find\_update}(A_k)$ \COMMENT { Is null when $A_k = \phi(...)$ }
-\STATE $i_2 = \mathit{find\_update}(A_l)$ \COMMENT { Is null when $A_l = \phi(...)$ }
-\IF {$i_1 == i_2$ or $i_1$ is null or $i_2$ is null} \STATE \COMMENT{ With our restrictions on writes we must have $i_1 = i_2$. }
-\STATE replace $\mathit{stmt}$ with \\
-\STATE $\texttt{A}_j = \texttt{A}_{j-1}$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
-%\ELSIF {$i_1 \ncong i_2$}
-%\STATE replace $\mathit{stmt}$ with \\
-%\STATE $\texttt{A}_j = \texttt{A}_k$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
-\ELSE
-\STATE $\mathit{stmt}$ stays as is
-\ENDIF
-\ENDFOR
-\end{algorithmic}
-\end{comment}
-
-\subsubsection{Restricting Array Writes}
-
-For now, we restrict array updates to \emph{canonical updates}. Assume (for simplicity) a two-dimensional array \texttt{A[I,J]}. A canonical update is the following:
-\begin{verbatim}
-for i in range(I):
-  for j in range(J):
-    ...
-    A[i,j] = ...
-    ...
-\end{verbatim}
-
-The update \texttt{A[i,j]} can be nested into an inner loop and there may be multiple updates, i.e., writes to \texttt{A[i,j]}.
-However, update such as \texttt{A[i-1,j] = ...} or \texttt{A[i-1,j-1] = ...}, etc., is not allowed.
-Additionally, while there could be several different loops that perform canonical updates, they must be of the same dimensionality, i.e., an update of higher or lower dimension, e.g., \texttt{A[i,j,k] = ...} is not allowed.
-We compute the \emph{canonical dimensionality} of each write array by examining the array writes in the original program and rejecting programs that violate the canonical write restriction.
-This restriction simplifies reasoning in this early stage of the compiler; we will look to relax the restriction in future work.
-
-Another restriction/assumption is that we assume the output array is given as input with initial values, and it is of size consistent with its canonical dimensionality.
-%\ana{This restriction will require that we update our benchmarks that come with output arrays.}
-
-Reads through an arbitrary formula, such as \texttt{A[i-1]} for example, are allowed; currently, the projection function returns dummy values if the read formula is out of bounds; we assume the programmer ensures that the program still computes correct output in this case.
-
-\subsubsection{Changes to Basic Vectorization}
-
-\begin{comment}
-There are two changes to Basic Vectorization to account for write arrays. We can account for them during processing of raise dimension/drop dimension in Phase 1 of Basic Vectorization.
-
-One change to Basic vectorization is the expansion of dimension if the array write or read occurs in a nested loop.
-That is, if there is an update $\texttt{A}[i,j] = ...$ that occurs in loop dimensionality $i,j,k$, $\texttt{A}[i,j]$ will be rewritten into $\texttt{A}[i,j,k]$.
-Similarly, a read $\texttt{A}[f(i,j)]$ will be rewritten into $\texttt{A}[f(i,j),k]$.
-($\texttt{A}[f(i,j),k]$ will be reshaped as $f(i,j)*I*J+k$ during Phase 1 of Basic Vectorization.
-\ana{TODO: Double check.})
-\end{comment}
-
-In addition to the changes for the handling of target-less phi-nodes, Basic Vectorization has to handle def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable.
-The definition can be an update \texttt{A\_2 = update(A\_1,i,...)}, a pseudo $\phi$-node \texttt{A\_2 = PHI(A\_0,A\_1)}, etc..
-Note that $\phi$-nodes for arrays have no subscript operations the way there are subscript operations in analysis-introduced arrays representing scalars.
-While there are variations, the most intuitive implementation will perform Basic Vectorization Phase 1 as is, inserting
-$\mathit{raise\_dim}$ and $\mathit{drop\_dim}$ in the same way.
-However, the implementation of raise dimension and drop dimension will be adapted because the dimension cannot be raised or dropped to a dimension lower than the canonical one.
-Consider a def-use edge $X \rightarrow Y$ for an array $\texttt{A}$.
-
-\begin{enumerate}
-
-\item same-level $X \rightarrow Y$. Do nothing, propagate the array, which happens to be of the right dimension.
-%\ana{There might be some opportunities to do copy propagation optimization and save some cycles, but let's leave this for later.}
-
-\item inner-to-outer $X \rightarrow Y$ triggers the addition of $\mathit{drop\_dim}$.
-However, the dimensionality cannot be dropped below the canonical dimensionality of the array.
-E.g., if the dimensionality of the loop enclosure $X$ is already at the canonical one, then $\mathit{drop\_dim}$ has no effect.
-
-\item outer-to-inner $X \rightarrow Y$ riggers  $\mathit{raise\_dim}$.
-Again, if the dimensionality of the loop enclosure of $Y$ is smaller or same as the canonical dimensionality of the array, then it has no effect, otherwise, if dimensionality is greater than the canonical dimensionality, $\mathit{raise\_dim(...)}$ (at $X$) is the same as in Basic Vectorization.
-
-\item "mixed" $X \rightarrow Y$. We assume that the mixed edge is transformed into an inner-to-outer followed by outer-to-inner edge before we perform vectorization, just as with Basic vectorization.
-
-\end{enumerate}
-
-If the use of the array is a read $\texttt{A}[f(i,j,k)]$ different than a canonical read $\texttt{A}[i,j,k]$, then we need to add a reshape operation as
-all arrays are $\texttt{A}[i,j,k]$. It can be added after raise\_dim/drop\_dim or incorporated in these operations.
-The bulk of the change is in Phase 2 of Basic Vectorization as outlined earlier.
-
-\subsubsection{Examples with Array Writes}
-
-\paragraph{Example 1}
-
-First, the canonical dimensionality of all \texttt{A,B,C} and \texttt{D} is 1.
-After Phase 1 of Basic Vectorization the Aiken's array write example will be (roughly) as follows:
-
-{\small
-\begin{verbatim}
-for i in range(N):
-1. A_1 = PHI(A_0,A_2)
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. A_2 = update(A_1, I, B_1[I] + 10);
-6. B_2 = update(B_1, I, A_2[I] * D_1[I-1]);
-7. C_2 = update(C_1, I, A_2[I] * D_1[I-1]);
-8. D_2 = update(D_1, I, B_2[I] * C_2[I]);
-\end{verbatim}
-}
-
-Note that since all def-uses are same-level (i.e., reads and writes of the array elements)
-no raise dimension or drop dimension happens.
-
-\begin{comment}
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{A}_1$ = $\phi(\texttt{A}_0,\texttt{A}_2)$
-\STATE 2. $\texttt{B}_1$ = $\phi(\texttt{B}_0,\texttt{B}_2)$
-\STATE 3. $\texttt{C}_1$ = $\phi(\texttt{C}_0,\texttt{C}_2)$
-\STATE 4. $\texttt{D}_1$ = $\phi(\texttt{D}_0,\texttt{D}_2)$
-\STATE 5. $\texttt{A}_2 = update(\texttt{A}_1, \stackrel{\rightarrow}{i}, \texttt{B}_1[\stackrel{\rightarrow}{i}] + 10)$;
-\STATE 6. $\texttt{B}_2 = update(\texttt{B}_1, \stackrel{\rightarrow}{i}, \texttt{A}_2[\stackrel{\rightarrow}{i}] * \texttt{D}_1[\stackrel{\rightarrow}{i-1}])$;
-\STATE 7. $\texttt{C}_2 = update(\texttt{C}_1, \stackrel{\rightarrow}{i}, \texttt{A}_2[\stackrel{\rightarrow}{i}] * \texttt{D}_1[\stackrel{\rightarrow}{i-1}])$;
-\STATE 8. $\texttt{D}_2 = update(\texttt{D}_1; \stackrel{\rightarrow}{i}, \texttt{B}_2[\stackrel{\rightarrow}{i}] * \texttt{C}_2[\stackrel{\rightarrow}{i}])$;
-\end{algorithmic}
-\end{comment}
-
-Phase 2 computes the closure of 4; $cl = \{4,6,7, 8\}$
-while 5 is vectorizable. Recall that 1,2, and 3 are target-less phi-nodes. Since the closure $cl$ includes updates to \texttt{B} and \texttt{C},
-the corresponding phi-nodes are added back to the closure and the def-use edges are added back to the target-less nodes. The uses of \texttt{A\_1} and \texttt{B\_1} in the vectorized statement
-turn into uses of \texttt{A\_0} and \texttt{B\_0} respectively; this is done for all original target-less phi-node.
-(But note that \texttt{A\_0} is irrelevant; the update writes into array \texttt{A\_2} in parallel.)
-Finally, the target-less phi-node for \texttt{A} is discarded.
-
-{\small
-\begin{verbatim}
-1. A_2 = update(A_0, I, ADD_SIMD(B_0[I],10));
-   equiv. to A_2[I] = ADD_SIMD(B_0[I],10)
-FOR i=0; i<N; i++; // MOTION loop
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. B_2 = update(B_1, i, A_2[i] * D_1[i-1]);
-   equiv. to B_2 = B_1; B_2[i] = A_2[i] * D_1[i-1];
-6. C_2 = update(C_1, i, A_2[i] * D_1[i-1]);
-7. D_2 = update(D_1, i, B_2[i] * C_2[i]);
-\end{verbatim}
-}
-
-\begin{comment}
-\begin{algorithmic}
-\STATE 1. $\texttt{A}_2 = update(\texttt{A}_0; \stackrel{\rightarrow}{i}, \mathit{ADD\_SIMD}(\texttt{B}_0[\stackrel{\rightarrow}i],[10,...])$ \COMMENT{Fully vectorized, size N.}
-\STATE FOR i=0; i<N; i++; \COMMENT{ MOTION loop }
-\STATE 2. $\texttt{B}_1$ = $\phi(\texttt{B}_0,\texttt{B}_2)$
-\STATE 3. $\texttt{C}_1$ = $\phi(\texttt{C}_0,\texttt{C}_2)$
-\STATE 4. $\texttt{D}_1$ = $\phi(\texttt{D}_0,\texttt{D}_2)$
-\STATE 5. $\texttt{B}_2 = update(\texttt{B}_1, i, \mathit{MUL}(\texttt{A}_2[i], \texttt{D}_1[{i-1}])$
-\STATE 6. $\texttt{C}_2 = update(\texttt{C}_1, i, \mathit{MUL}(\texttt{A}_2[i], \texttt{D}_1[{i-1}])$
-\STATE 7. $\texttt{D}_2 = update(\texttt{D}_1, i, \mathit{MUL}(\texttt{B}_2[i], \texttt{C}_1[i])$
-\end{algorithmic}
-\end{comment}
-
-\paragraph{Example 2} Now consider the MPC Source of Histogram:
-
-{\small
-\begin{verbatim}
-for i in range(0, num_bins):
-   res1 = PHI(res, res2)
-   for j in range(0, N):
-      res2 = PHI(res1, res3)
-      tmp1 = (A[j] == i)
-      tmp2 = (res2[i] + B[j])
-      tmp3 = MUX(tmp1, res2[i], tmp2)
-      res3 = Update(res2, i, tmp3)
-return res1
-\end{verbatim}
-}
-
-The canonical dimensionality of \texttt{res} is 1.
-Also, the phi-node \texttt{res1 = PHI(res, res2)} is a target-less phi-node (the implication being that the inner for loop can be vectorized across $i$).
-After Phase 1, Basic vectorization produces the following code (statements are implicitly vectorized along $i$ and $j$).
-In a vectorized update statement, we can ignore the incoming array, \texttt{res2} in this case.
-The update writes (in parallel) all locations of the 2-dimensional array, in this case it sets up each \texttt{res3[i,j] = tmp3[i,j]}.
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I = raise_dim(i, ((i:num_bins),(j:N)))
-for i in range(0, num_bins):
-   res1 = PHI(res, res2') // target-less phi-node
-   res1' = raise_dim(res1, (j:N))
-   for j in range(0, N):
-      res2 = PHI(res1', res3)
-      tmp1 = (A1 == I)
-      tmp2 = (res2 + B1)
-      tmp3 = MUX(tmp1, res2, tmp2)
-      res3 = Update(res2, (I,J), tmp3)
-   res2' = drop_dim(res2)
-res1'' = drop_dim(res1)
-return res1''
-\end{verbatim}
-}
-
-Processing the inner loop in Phase 2 vectorizes \texttt{tmp1 = (A1 == I)} along the $j$ dimension but leaves the rest of the statements in a MOTION loop.
-Processing the outer loop is interesting.
-This is because the PHI node is a target-less node, and therefore, there are no closures!
-Several things happen.
-(1) Everything can be vectorized along the $i$ dimension.
-(2) We remove the target-less PHI node, however, we must update uses of \texttt{res1} appropriately: the use at raise\_dim goes to the first argument of the PHI function and the use at drop\_dim goes to the second argument.
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I1 = raise_dim(i, ((i:num_bins),(j:N)))
-
-tmp1[I,J] = (A1[I,J] == I1[I,J])
-
-res1' = raise_dim(res, (j:N)) // replacing res1 with res, 1st arg
-for j in range(0, N):
-   res2 = PHI(res1', res3)
-   tmp2[I,j] = (res2[I,j] + B1[I,j])
-   tmp3[I,j] = MUX(tmp1[I,j], res2[I,j], tmp2[I,j])
-   res3 = Update(res2, (I,j), tmp3)
-   equiv. to res3 = res2; res3[I,j] = tmp3[I,j]
-res2' = drop_dim(res2)
-res1 = drop_dim(res2') // replacing with res2', 2nd arg. NOOP
-return res1
-\end{verbatim}
-}
-
-\begin{comment}
-The cleanup phase can then get rid of $j$ in the \texttt{res} variables and \texttt{tmp} variables, so the loop operates over the \texttt{num\_bins}-size vectors.
-This will lead to dropping raise\_dim and drop\_dim as well:
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I1 = raise_dim(i, ((i:num_bins),(j:N)))
-
-tmp1[I,J] = (A1[I,J] == I1[I,J])
-
-for j in range(0, N):
-   res2 = PHI(res1, res3) // now 1-dim array
-   tmp2[I] = ADD_SIMD(res2[I],B1[I,j])
-   tmp3[I] = MUX_SIMD(tmp1[I,j], res2[I], tmp2[I])
-   res3 = Update(res2, I, tmp3[I])
-return res2
-\end{verbatim}
-}
-\end{comment}
-
-\begin{comment}
-\paragraph{Example 2}
-
-Consider MPC-source of Histogram after removal of infeasible edges and redundant phi-nodes, and MUX refinement:
-
-\begin{algorithmic}
-
-\STATE res!1 = []
-\STATE for i: plaintext in range(0, num\_bins):
-\STATE \hspace{0.25cm} res!2 = $\phi$(res!1, res!3)
-\STATE  \hspace{0.25cm} ...
-\STATE res!4 = res!2 \COMMENT{Added due to the "mixed" def-use edge.}
-\STATE for i: plaintext in range(0, num\_bins):
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE \hspace{0.25cm}        for j: plaintext in range(0, N):
-\STATE \hspace{0.5cm}            res!5 = $\phi$(res!4, res!7)
-\STATE \hspace{0.5cm}            !2!3 = (A[j] == i)
-\STATE \hspace{0.5cm}            !3!3 = (res!5[i] + B[j])
-\STATE \hspace{0.5cm}            res!6 = res!5; res!6[i] = !3!3
-\STATE \hspace{0.5cm}            res!7 = res!6; res!7[i] = MUX(!2!3, res!6[i], res!5[i])
-
-\end{algorithmic}
-
-After Phase 1 of Basic Vectorization:
-
-\begin{algorithmic}
-\STATE res!1 = []
-\STATE for i: plaintext in range(0, num\_bins):
-\STATE \hspace{0.25cm} res!2 = $\phi$(res!1, res!3)
-\STATE  \hspace{0.25cm} ...
-\STATE res!4 = res!2 \COMMENT{Added due to the "mixed" def-use edge. Because the canonical dimensionality of res is 1, we DO NOT drop dimension of res!2.}
-\STATE A[$i,j$] = $\mathit{raise\_dim}$(A[$j$],$i$) \COMMENT{Copying row A num\_bin times. Reusing name.}
-\STATE B[$i,j$] = $\mathit{raise\_dim}$(B[$i$],$j$) \COMMENT{Copying column B N times. Reusing name.}
-\STATE res!4[$i,j$] = $\mathit{raise\_dim}$(res!4[$i$],$j$) \COMMENT{Reusing name.}
-\STATE for i: plaintext in range(0, num\_bins):
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE \hspace{0.25cm}        for j: plaintext in range(0, N):
-\STATE \hspace{0.5cm}            res!5 = $\phi$(res!4, res!7) \COMMENT{Do nothing. res!4 has been raised to proper dimension.}
-\STATE \hspace{0.5cm}            !2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{EQ}$(A[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], [[0,0...], [1,1...],...[num\_bins-1,num\_bins-1...]])
-\STATE \hspace{0.5cm}            !3!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{ADD}$(res!5[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$],B[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$])
-\STATE \hspace{0.5cm}            res!6 = res!5 \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.5cm}            res!6[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = !3!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$]
-\STATE \hspace{0.5cm}            res!7 = res!6
-\STATE \hspace{0.5cm}            res!7[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{MUX}$(!2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], res!6[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], res!5[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$])
-\end{algorithmic}
-
-After Phase 2 and Phase 3. The EQ operation is vectorizable across both dimensions, and the rest of the computation is vectorizable across the $i$-dimension.
-
-\begin{algorithmic}
-\STATE ...
-\STATE !2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{EQ\_SIMD}$(A[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], [[0,0...], [1,1...],...[num\_bins-1,num\_bins-1...]])
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE FOR j=0; j<N; j++; \COMMENT{MOTION loop}
-\STATE \hspace{0.25cm}            res!5 = $\phi$(res!4, res!7) \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.25cm}            !3!3[$\stackrel{\rightarrow}i,j$] = $\mathit{ADD\_SIMD}$(res!5[$\stackrel{\rightarrow}i,j$],B[$\stackrel{\rightarrow}i,j$])
-\STATE \hspace{0.25cm}            res!6 = res!5 \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.25cm}            res!6[$\stackrel{\rightarrow}i,j$] = !3!3[$\stackrel{\rightarrow}i,j$]
-\STATE \hspace{0.25cm}            res!7 = res!6
-\STATE \hspace{0.25cm}            res!7[$\stackrel{\rightarrow}i,j$] = $\mathit{MUX\_SIMD}$(!2!3[$\stackrel{\rightarrow}i,j$], res!6[$\stackrel{\rightarrow}i,j$], res!5[$\stackrel{\rightarrow}i,j$])
-\end{algorithmic}
-\end{comment}
-
-%\section{Divide-and-Conquer}
-%\label{sec:div_and_conq}
-%\ana{TODO: Now that we have broken FOR loops into smaller chunks, we can add Divide-and-conquer reasoning with Z3 and implement this additional transform.}
-
-\section{Implementation and Evaluation}
-\label{sec:implementation}
-
-\section{Related Work}
-\label{sec:related}
-
-\paragraph{MPC languages and compilers.}
-
-Languages and compilers for secure computation have seen significant attention and advances in recent years. The early MPC compilers Fairplay~\cite{CCS:BenNisPin08},
-and Sharemind~\cite{ESORICS:BogLauWil08} were followed by PICCO~\cite{CCS:ZhaSteBla13}, Obliv-C~\cite{Zahur:2015}, TinyGarble~\cite{SP:SHSSK15},
-Wystiria~\cite{SP:RasHamHic14}, and others. A new generation of MPC compilers
-includes SPDZ/SCALE-MAMBA/MP-SPDZ~\cite{Keller:2020} and the ABY/HyCC/MOTION~\cite{NDSS:DemSchZoh15,CCS:BDKKS18,Braun:2022} frameworks.
-These two families are the state-of-the art and
-are actively developed. Another recent development is Viaduct, a functional language and compiler that supports a range of secure computation
-frameworks, including MPC and ZKP. Hastings et el. present a review of compiler frameworks~\cite{Hastings:2019}.
-
-While each of these languages and compilers brings in new ideas and advances, none addresses the problem of ``circuit independent'' intermediate
-representation and optimization. We envision a classical compiler structure: (1) a Wysteria, Viaduct, Obliv-C, or IMP Source front end, including rich type systems and
-AST-level semantic analysis, compile into the MPC Source IR, (2) MPC Source-level optimizations take place, followed by (3) back-end compilers into circuits.
-Our focus is at the intermediate level.
-
-%{\begin{center}
-%\includegraphics[width=0.5\linewidth]{figs/focus.png}
-%\end{center}
-%}
-
-Many works focus on the implementation of MPC protocols exposing an API to the programmer. For example, the ABY/MOTION line of
-compiler frameworks provides a library of MPC primitives; the programmer writes MPC programs in C++ on top of the library. These back ends implement
-different protocols and allow for mixing, but notably, they leave it to the programer to assign different protocols to different parts of the
-computation and perform share conversion accordingly. In addition, MOTION provides SIMD primitives, which allows for efficient execution
-of MPC operations, but again, using SIMD primitives is the responsibility of the programer. There is interest in frameworks for automatic mixing,
-e.g., \cite{CCS:BDKKS18,Ishaq:2019,Fang:2022}.
-
-Other works, e.g., Obliv-C~\cite{Zahur:2015}, Wystiria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021} focus on higher-level language design, particularly information-flow systems that
-restrict flow between secure and insecure parts of the program.
-%Key ideas such as ${\sf insecure} <: {\sf secure}$ qualifier systems are similar to the ideas we
-%present in~\Secref{sec:interprocedural} and~\Secref{sec:zkp} and we intend to draw on these work. While we will design a high-level language expanding our
-%intraprocedural IMP-MPC syntax and semantics into the \emph{interprocedural} case, our interest and focus lie primarily in the ``machine-independent'' optimizations
-%thrust of the proposal.
-
-%There is significant development in ZKP languages as well. ~\ana{Alex Ozdemir's talk, ZoKrates, Cairo, others} Our work focuses on ZKP via MPC and our key idea
-%is the separation of the higher-level ``backend-independent'' compilation and optimization and the lower-level compilation into circuits.
-
-\paragraph{Classical HPC compilers.}
-Automatic vectorization is a longstanding problem in high-performance computing (HPC).
-There are thousands of works in this area reflecting over 40 years of research. We presented a vectorization
-algorithm for MPC Source, essentially extending classical loop vectorization \cite{Allen:1987}. In HPC vectorization, conditional control flow
-presents a challenge --- one cannot estimate the cost of a schedule or vectorize branches in a straightforward manner --- in contrast to
-MPC Source vectorization.
-We view Karrenberg's work on Whole function vectorization~\cite{Karrenberg:2015} as most closely related to ours --- it linearizes the program and vectorizes
-both branches of a conditional applying masking to avoid execution of the branch-not-taken code, and selection (similar to MUX) to select the correct value based on the
-result of the condition at runtime.
-The problem is that masking and selection, or more generally, handling control predicates~\cite{Benabderrahmane:2010,Karrenberg:2015},
-can lead to \emph{slowdown}.
-%As a result, vectorizing compilers and polyhedral compilers tend to focus on straight-line code due to the unpredictable
-%overhead conditionals may introduce. \ana{add Louis noel pochet's benchmarks. citation on vectorizing compilers???}
-
-We argue that vectorization over linear MPC Source is a different problem, one that warrants a new look, while drawing from
-results in HPC.
-Since both branches of the conditional and the multiplexer \emph{always} execute, not only can we apply aggressive vectorization on linear code, but (perhaps more importantly)
-we can also build analytical models that accurately predict execution time. These models in turn would drive optimizations such as vectorization, protocol mixing, and others.
-Vectorization meshes in with those additional optimizations in non-trivial ways.
-
-Furthermore, extensions of classical loop vectorization with array writes, arbitrary indexing, including non-affine indexing, and interaction with SSA are non-trivial
-and present novel challenges and opportunities for contribution. Polyhedral parallelization~\cite{Benabderrahmane:2010} considers a higher-level source (typically AST)
-representation, while our work takes advantage of linear MPC Source and SSA form. The work by Karrenberg~\cite{Karrenberg:2015} is rare in that space, in the sense that it considers
-vectorization over SSA form, which has similarities to MPC Source. We consider different array representation, notion of dependence, and reasoning
-about dependence, which we conjecture is more suitable for MPC Source. Buscher~\cite{Buscher:2018} considers SIMD-vectorization
-at the level of source code, which then combines with circuit-level optimizations in the TinyGarble compiler.
-He proposes using an off-the-shelf polyhedral compiler, however, application is limited to only two routines,
-essentially just inner product and euclidian distance; it is unclear how effective the off-the-shelf compiler is.
-In contrast, we consider vectorization at the level of MPC Source separating ``backend-independent'' vectorization
-and circuit-level amortization (done by MOTION). We apply our compiler on a wide range of routines.
-
-%\section{Future Work}
-%\label{sec:implementation_and_benchmarks}
-%\input{../sections/implementation_and_benchmarks.tex}
-%\input{../sections/evaluation.tex}
-
-\section{Conclusion and Future Work}
-\label{sec:conclusion}
-%\input{../sections/conclusion.tex}
-
-%\begin{acks}
-%    \ishaq{TODO}
-%\end{acks}
-
-%\printbibliography
-
-\bibliographystyle{plain}
-%\bibliography{c:/trink/Doc/MyTexFiles/texrefs}
-%\bibliography{mybib,library,library1,library2,library3,library_ana}
-\bibliography{library_ana,crypto_crossref}
-
-% that's all folks
-\end{document}

### 2022-04-28 overleaf/introduction.tex
Split paper into sections.

--- 2022-04-27 overleaf/paper_SIMD.tex
@@
-% !TeX TXS-program:compile = txs:///pdflatex/[--shell-escape]
-%
-% acm templates v1.57 (sigconf template)
-%
-%\documentclass[sigconf, screen, review, anonymous, natbib=false]{acmart}
-\documentclass[sigconf, screen, natbib=false, dvipsnames, table]{acmart}
-\settopmatter{printacmref=true}
-% mandatory for CCS'19
-\usepackage{balance}
-\usepackage{mathtools}
-% for creating a balanced last page (usually last page with references)
-
-% % defining the \BibTeX command - from Oren Patashnik's original BibTeX documentation.
-% \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08emT\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-
-% Rights management information.
-% This information is sent to you when you complete the rights form.
-% These commands have SAMPLE values in them; it is your responsibility as an author to replace
-% the commands and values with those provided to you when you complete the rights form.
-%
-% These commands are for a PROCEEDINGS abstract or paper.
-
-\copyrightyear{2019}
-\acmYear{2019}
-\setcopyright{acmcopyright}
-\acmConference[CCS '19]{2019 ACM SIGSAC Conference on Computer and Communications Security}{November 11--15, 2019}{London, United Kingdom}
-\acmBooktitle{2019 ACM SIGSAC Conference on Computer and Communications Security (CCS '19), November 11--15, 2019, London, United Kingdom}
-\acmPrice{15.00}
-\acmDOI{10.1145/3319535.3339818}
-\acmISBN{978-1-4503-6747-9/19/11}
-
-%
-% These commands are for a JOURNAL article.
-%\setcopyright{acmcopyright}
-%\acmJournal{TOG}
-%\acmYear{2018}\acmVolume{37}\acmNumber{4}\acmArticle{111}\acmMonth{8}
-%\acmDOI{10.1145/1122445.1122456}
-
-%
-% Submission ID.
-% Use this when submitting an article to a sponsored event. You'll receive a unique submission ID from the organizers
-% of the event, and this ID should be used as the parameter to this command.
-%\acmSubmissionID{123-A56-BU3}
-
-%
-% The majority of ACM publications use numbered citations and references. If you are preparing content for an event
-% sponsored by ACM SIGGRAPH, you must use the "author year" style of citations and references. Uncommenting
-% the next command will enable that style.
-%\citestyle{acmauthoryear}
-
- \makeatletter
- % \renewcommand{\section}{\abovedisplayskip 10\p@ \@plus3\p@ \@minus1\p@%
- %                       \belowdisplayskip 5\p@ \@plus3\p@ \@minus1\p@%
- %                       \abovedisplayshortskip 0pt \@plus2\p@%
- %                       \belowdisplayshortskip 0pt \@plus2\p@ \@minus0\p@%
- %                       \@startsection{section}{1}{\z@}%
- %                        {-17\p@ \@plus -4\p@ \@minus -4\p@}%
- %                        {6\p@ \@plus 4\p@ \@minus 4\p@}%
- %                        {\normalfont\large\bfseries\boldmath
- %                         \rightskip=\z@ \@plus 8em\pretolerance=10000 }}
- \renewcommand{\subsection}{\@startsection{subsection}{2}{\z@}%
-                        {-8\p@ \@plus -4\p@ \@minus -4\p@}%
-                        {5\p@ \@plus 2\p@ \@minus 2\p@}%
-                        {\normalfont\Large\bfseries\boldmath
-                         \rightskip=\z@ \@plus 3em\pretolerance=10000 }}
-  \renewcommand{\subsubsection}{\@startsection{subsubsection}{3}{\z@}%
-                        {-6\p@ \@plus -4\p@ \@minus -4\p@}%
-                        {1\p@ \@plus 1\p@ \@minus 0\p@}%
-                        {\normalfont\normalsize\bfseries\boldmath}}
-% \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\z@}%
-%                       {-8\p@ \@plus -4\p@ \@minus -4\p@}%
-%                       {-2\p@ \@plus -0.22em \@minus -0.1em}%
-%                       {\normalfont\normalsize\bfseries}}
- \makeatother
-
-% *** GRAPHICS RELATED PACKAGES ***
-\usepackage{graphicx}
-\usepackage{comment}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{listings}
-% declare the path(s) where your graphic files are
-\graphicspath{{../figures/}}
-% and their extensions so you won't have to specify these with
-% every instance of \includegraphics
-\DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-
-% Minted is used for source code highlighting
-%\usepackage[outputdir=/Volumes/ramdisk/]{minted}
-% \usepackage{minted}
-% *** MATH PACKAGES ***
-\usepackage{amsmath}
-
-% *** SPECIALIZED LIST PACKAGES ***
-%\usepackage{algorithm2e}
-\usepackage{algorithmic}
-
-% *** ALIGNMENT PACKAGES ***
-\usepackage{array}
-
-\usepackage{url}
-
-\usepackage[utf8]{inputenc}
-\usepackage[american]{babel}
-%\usepackage[backend=biber,style=ACM-Reference-Format]{biblatex}
-%\addbibresource{./library.bib}
-%\addbibresource{./cryptobib/abbrev2.bib}
-%\addbibresource{./cryptobib/crypto_crossref.bib}
-
-% Used for Theorems and Definitions
-\usepackage{amsthm}
-\newtheorem{definition}{Definition}
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}{Lemma}
-\theoremstyle{definition}
-\newtheorem{remark}{Remark}
-\newtheorem{assumption}{Assumption}
-
-\usepackage{xcolor}
-\usepackage{xspace}
-\usepackage{xstring} % required by IfEqCase
-
-\usepackage{booktabs} % better tables
-\usepackage{algorithm}
-\usepackage{algorithmic}
-
-% for references, use it like this: \cref{some_label}
-% it will automatically put Appendix, Figure, Section, etc before the reference
-\usepackage[capitalise]{cleveref}
-
-% ishaq: adding our macros
-\input{macros.tex}
-
-\sloppy
-\begin{document}
-\fancyhead{}
-% do not delete this code.
-
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
-\title[Compilation and Backend-Independent Vectorization for Multi-Party Computation]{Compilation and Backend-Independent Vectorization for Multi-Party Computation}
-
-%
-% The "author" command and its associated commands are used to define the authors and their affiliations.
-% Of note is the shared affiliation of the first two authors, and the "authornote" and "authornotemark" commands
-% used to denote shared contribution to the research.
-
-\author{Benjamin Levy}
-\email{levyb3@rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Benjamin Sherman}
-\email{shermb@rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Muhammad Ishaq}
-%\authornote{This work was done in part while the author was at RPI.}
-\email{ishaqm@purdue.edu}
-%\orcid{1234-5678-9012}
-%\author{G.K.M. Tobin}
-%\authornotemark[1]
-%\email{webmaster@marysville-ohio.com}
-\affiliation{%
-    \institution{Purdue University}
-    %\streetaddress{P.O. Box 1212}
-    \city{West Lafayette}
-    \state{Indiana}
-%    \postcode{43017-6221}
-}
-
-\author{Lindsey Kennard}
-\email{fire.elemental@gmail.com}
-\affiliation{%
-    \institution{STR}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Boston}
-    \state{Massachusets}
-    %\country{Iceland}
-}
-
-\author{Ana Milanova}
-\email{milanova@cs.rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Vassilis Zikas}
-%\authornote{This work was done in part while the author was visiting UCLA and supported in part by DARPA and SPAWAR under contract N66001-15-C-4065 and by a SICSA Cyber Nexus Research Exchanges grant.}
-\email{vzikas@purdue.edu}
-%\orcid{1234-5678-9012}
-%\author{G.K.M. Tobin}
-%\authornotemark[1]
-%\email{webmaster@marysville-ohio.com}
-\affiliation{%
-    \institution{Purdue University}
-    %\streetaddress{P.O. Box 1212}
-    \city{West Lafayette}
-    \state{Indiana}
-    %    \postcode{43017-6221}
-}
-
-%
-% By default, the full list of authors will be used in the page headers. Often, this list is too long, and will overlap
-% other information printed in the page headers. This command allows the author to define a more concise list
-% of authors' names for this purpose.
-%\renewcommand{\shortauthors}{Trovato and Tobin, et al.}
-
-\renewcommand{\A}{{\sf A}}
-\renewcommand{\B}{{\sf B}}
-\renewcommand{\C}{{\sf C}}
-\renewcommand{\D}{{\sf D}}
-
-% As a general rule, do not put math, special symbols or citations
-% in the abstract
-\begin{abstract}
-
-\ana{We need an abstract.}
-
-\end{abstract}
-
-%
-% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-% Please copy and paste the code instead of the example below.
-%
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10003752.10010124.10010138.10010143</concept_id>
-<concept_desc>Theory of computation~Program analysis</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10003752.10003777.10003789</concept_id>
-<concept_desc>Theory of computation~Cryptographic protocols</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10002978.10002979</concept_id>
-<concept_desc>Security and privacy~Cryptography</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Theory of computation~Program analysis}
-\ccsdesc[500]{Theory of computation~Cryptographic protocols}
-\ccsdesc[300]{Security and privacy~Cryptography}
-
-%
-% Keywords. The author(s) should pick words that accurately describe the work being
-% presented. Separate the keywords with commas.
-\keywords{multiparty computation; compilers; cryptography}
-
-\begin{comment}
-%
-% A "teaser" image appears between the author and affiliation information and the body
-% of the document, and typically spans the page.
-\begin{teaserfigure}
-    \includegraphics[width=\textwidth]{sampleteaser}
-    \caption{Seattle Mariners at Spring Training, 2010.}
-    \Description{Enjoying the baseball game from the third-base seats. Ichiro Suzuki preparing to bat.}
-    \label{fig:teaser}
-\end{teaserfigure}
-\end{comment}
-%
-% This command processes the author and affiliation and title information and builds
-% the first part of the formatted document.
-\maketitle
-
 \section{Introduction}
@@
 \end{comment}
-
-\section{Overview}
-\label{sec:overview}
-
-\begin{table*}
-\begin{tabular}{ccc}
-\begin{minipage}{0.275\textwidth}
-{\small
-\begin{pythonn}
-def biometric(C: shared[list[int]], D: int, 
-      S: shared[list[int]], N: int) -> 
-      shared[tuple[int,int]]:
-   min_sum : int = MAX_INT
-   min_idx : int = 0  
-   for i in range(N): 
-      sum : int = 0
-      for j in range(D): 
-         # d = S[i,j] - C[j]
-         d : int = S[i * D + j] - C[j] 
-         p : int = d * d
-         sum = sum + p
-      if sum < min_sum:
-         min_sum : int = sum
-         min_idx : int = i
-   return (min_sum, min_idx)
-\end{pythonn}
-}
-\end{minipage}
-
-&
-
-\begin{minipage}{0.33\textwidth}
-{\tiny
-\begin{pythonn}
-min_sum!1 = MAX_INT
-min_idx!1 = 0
-for i in range(0, N):
-   min_sum!2 = PHI(min_sum!1, min_sum!4)
-   min_idx!2 = PHI(min_idx!1, min_idx!4)
-   sum!2 = 0
-   for j in range(0, D):
-      sum!3 = PHI(sum!2, sum!4)
-      d = SUB(S[((i * D) + j)],C[j]) 
-      p = MUL(d,d) 
-      sum!4 = ADD(sum!3,p) 
-   t = CMP(sum!3,min_sum!2) 
-   min_sum!3 = sum!3
-   min_idx!3 = i
-   min_sum!4 = MUX(t, min_sum!3, min_sum!2) 
-   min_idx!4 = MUX(t, min_idx!3, min_idx!2) 
-return (min_sum!2, min_idx!2)   
-\end{pythonn}
-}
-\end{minipage}
-
-&
-
-\begin{minipage}{0.33\textwidth}
-{\small
-\begin{pythonn}
-min_sum!1 = MAX_INT
-min_idx!1 = 0
-# S^ is same as S. C^ replicates C N times: 
-S^ = raise_dim(S, ((i * D) + j), (i:N,j:D)) #S^[i,j] = S[i,j]
-C^ = raise_dim(C, j, (i:N,j:D)) #C^[i,j] = C[j]
-
-sum!2[I] = [0,..,0]
-# computes _all_  "at once"
-d[I,J] = SUB_SIMD(S^[I,J],C^[I,J])
-p[I,J] = MUL_SIMD(d[I,J],d[I,J])
-
-for j in range(0, D):
-   # sum!2[I], sum!3[I], sum!4[I] are size-N vectors 
-   # computes N intermediate sums "at once"
-   sum!3[I] = PHI(sum!2[I], sum!4[I])       
-   sum!4[I] = ADD_SIMD(sum!3[I],p[I,j])
-
-min_idx!3[I] = [0,1,...N-1]   
-for i in range(0, N):
-   min_sum!2 = PHI(min_sum!1, min_sum!4) 
-   t[i] = CMP(sum!3[i],min_sum!2)
-   min_sum!4 = MUX(t[i], sum!3[i], min_sum!2)  
-for i in range(0, N):
-   min_idx!2 = PHI(min_idx!1, min_idx!4)    
-   min_idx!4 = MUX(t[i], min_idx!3[i], min_idx!2)
-return (min_sum!2, min_idx!2)   
-\end{pythonn}
-}
-\end{minipage}
-
-\\
-
-(a) IMP Source & (b) MPC Source & (c) Optimized MPC Source
-\end{tabular}
-\caption{Biometric Matching: From IMP Source to MPC Source to Optimized MPC Source. }
-\label{tab:source_and_MPC_source_and_optimized_MPC_source}
-\end{table*}
-
-\subsection{Source}
-
-As a running example, consider Biometric matching, a standard MPC benchmark.
-An intuitive (and naive) implementation is as shown in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(a).
-Array \python{C} is the feature vector of \texttt{D} features that we wish to match and \python{S} 
-is the database of \python{N} size-\python{D} vectors that we match against.
-
-\begin{comment}
-{\small
-\begin{verbatim}
-def biometric(C: shared[list[int]], D: int, 
-              S: shared[list[int]], N: int) -> 
-                      tuple[shared[int], shared[int]]:
-    min_sum = 10000
-    min_index = 0  
-    for i in range(N): #loop over database
-        sum = 0
-        for j in range(D): #loop over features
-            d = S[i * D + j] - C[j] #i.e., d = S[i,j] - C[j]
-            p = d * d
-            sum = sum + p
-        if sum < min_sum:
-            min_sum = sum
-            min_index = i
-    return (min_sum, min_index)
-\end{verbatim}
-}
-\end{comment}
-
-Our compiler takes essentially standard IMP~\ana{Need citation here.} 
-syntax and imposes certain semantic restrictions. (We detail the restrictions in the following sections.)
-The programmer writes an iterative program and annotate certain inputs 
-and outputs as \emph{shared}. In the example arrays \python{C} and \python{S}
-are shared, meaning that they store shares, however, the array sizes \python{D} and
-\python{N} respectively are plaintext. The code iterates over the entries in the 
-database and computes the Euclidean distance (its square actually) of the current 
-entry \python{S[i]} and \python{C}. The program returns the index of the vector that gives 
-the best match plus the corresponding sum of squares.
-
-\subsection{MPC Source and Cost of Schedule}
-
-Our compiler generates an intermediate representation, MPC Source. MPC Source 
-for Biometric Matching is shown in 
-Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(b). First, MPC Source is an SSA form. 
-Second, it is linear. The conditional in lines 13-14 in IMP Source turns into the linear code in lines 12-16 in MPC Source. 
-The test turns into a CMP operation (\python{t = CMP(sum!3,min\_sum!2)}), followed by the 
-true-branch sequence, followed by the MUX operations. The first MUX operation selects the value
-of \python{min\_sum}: if \python{t} is true, then \python{min\_sum} gets the value of the second multiplexer
-argument,  \python{min\_sum!3}, otherwise it takes the value of the third argument, \python{min\_sum!2}.  
-Third, MPC Source is a special form of SSA. The SSA $\phi$-nodes at the if-then-else (lines 13-15) turn into 
-MUX operations, while the $\phi$-nodes at for loops turn into what we call pseudo PHI nodes with a straightforward semantics.
-
-\begin{comment}
-{\small
-\begin{pythonn}
-1. min_sum!1 = 10000
-2. min_index!1 = 0
-3. for i in range(0, N!0):
-4.    min_sum!2 = PHI(min_sum!1, min_sum!4)
-5.    min_index!2 = PHI(min_index!1, min_index!4)
-6.    sum!2 = 0
-7.    for j in range(0, D!0):
-8.      sum!3 = PHI(sum!2, sum!4)
-9.      d!3 = (S!0[((i * D!0) + j)] - C!0[j]) // MPC
-10.     p!3 = (d!3 * d!3) // MPC
-11.     sum!4 = (sum!3 + p!3) // MPC
-12.  !1!2 = (sum!3 < min_sum!2) // MPC
-13.  min_sum!3 = sum!3
-14.  min_index!3 = i
-15.  min_sum!4 = MUX(!1!2, min_sum!3, min_sum!2) // MPC
-16.  min_index!4 = MUX(!1!2, min_index!3, min_index!2) // MPC
-17.!2!1 = (min_sum!2, min_index!2)   
-\end{pythonn}
-}
-\end{comment}
-
-%The compiler linearizes the source turning conditionals into MUX statements. The PHI nodes
-%are remnants of the SSA IR; the compiler generates code that picks the correct value 
-%when producing MOTION output; the MOTION framework in turn linearizes 
-%loops when it generates the circuit.
-
-We turn to our analytical model to compute the \emph{cost} of the program. Assuming fixed 
-cost $\beta$ for a local MPC operation (essentially just ADD) and cost $\alpha$ for a remote
-MPC operation (e.g., MUX, CMP, and remaining operations), the cost of the iterative schedule 
-will be $ND(2\alpha+\beta) + N(3\alpha$).
-
-A key contribution is the vectorizing transformation. We can compute all $N*D$ 
-subtraction operations (line 9) in a single SIMD instruction; similarly we can compute 
-all multiplication operations (line 10) in a single SIMD instruction. And while we cannot
-vectorize computation of the N individual sums, we can compute the N sums in parallel. 
-Our compiler \emph{automatically detects these opportunities and transforms the program}. 
-It is standard that MPC researchers write vectorized versions of the Biometric
-program by hand; we are the first (to the best of our knowledge) to automatically 
-transform a naive, iterative MPC program into an unintuitive vectorized one.
-
-\subsection{Vectorized MPC Source and Cost of Schedule}
-
-Our compiler produces the vectorized program shown in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(c). 
-Note that this is still our intermediate representation, Optimized MPC Source. Subsequently,
-the compiler turns this code into MOTION variables, loops and SIMD primitives, which MOTION then 
-uses to generate the circuit.
-
-The compiler determines that SUB and MUL in ``naive'' MPC Source (lines 9 and 10 in (b))
-can be fully vectorized into the SIMD SUB and MUL in optimized MPC Source (lines 9 and 10 in (c)). 
-Notation \python{p[I,J]} denotes a 2-dimensional array with fully vectorized dimensions.
-The computation of sum (line 11 in (b))
-is sequential across the $j$-dimension, but it is parallel across the $i$-dimension. 
-The loop in lines 12-16 in (c) illustrates; here \python{p[I,j]} refers to the $j$-th column in \python{p}. 
-Unfortunately, CMP and MUX remain sequential.
-
-\begin{comment}
-{\small
-\begin{verbatim}
-min_sum!1 = 10000
-min_index!1 = 0
-// S!0^ is same as S!0. C!0^ replicates C!0 N-times: 
-S!0^ = raise_dim(S!0, ((i * D!0) + j), (i:N!0,j:D!0))
-C!0^ = raise_dim(C!0, j, (i:N!0,j:D!0))
-
-sum!2 = [0,..,0]
-// Computes all differences and all products "at once"
-d!3[I,J] = SUB_SIMD(S!0^[I,J],C!0^[I,J])
-p!3[I,J] = MUL_SIMD(d!3[I,J],d!3[I,J])
-
-for j in range(0, D!0):
-  // sum!2[I], sum!3[I], sum!4[I] are vectors of size N  
-  // Computes N intermediate sums "at once"
-  sum!3[I] = PHI(sum!2[I], sum!4[I])       
-  sum!4[I] = ADD_SIMD(sum!3[I],p!3[I,j])
-
-min_index!3 = [0,1,...N!0-1]
-
-for i in range(0, N!0):
-  min_sum!2 = PHI(min_sum!1, min_sum!4) 
-  !1!2[i] = CMP(sum!3[i],min_sum!2)
-  min_sum!4 = MUX(!1!2[i], sum!3[i], min_sum!2)
-
-for i in range(0, N!0):
-  min_index!2 = PHI(min_index!1, min_index!4)  
-  min_index!4 = MUX(!1!2[i], min_index!3[i], min_index!2)
-!2!1 = (min_sum!2, min_index!2)   
-\end{verbatim}
-}
-\end{comment}
-
-In MPC backends, executing $n$ operations ``at once'' in a single SIMD operation costs a lot less than executing those $n$ operations one by one.
-This is particularly important when there is communication (i.e., in remote), since many 1-bit values are sent at once rather than sequentially. 
-%In MPC compilers a vectorized operation that computing M operations "at once" costs 
-%essentially the same ($\alpha$ or $\beta$) as an individual operation. 
-We elaborate on the cost model in the following section but for now consider that 
-each operation has a \emph{fixed} portion (does benefits from amortization) and 
-a \emph{variable} portion (does not benefit from amortization): $\alpha = \alpha_\mathit{fix} + \alpha_\mathit{var}$. 
-This gives rise to the following formula for amortized cost: $f(n) = \alpha_\mathit{fix} + n\alpha_\mathit{var}$, 
-as opposed to unamortized cost $g(n) = n\alpha_\mathit{fix} + n\alpha_\mathit{var}$. (We extend the same reasoning to 
-$\beta$-instructions.)
-
-Thus, the fixed cost of the vectorized program amounts to
-$2\alpha_\mathit{fix}$ + $D \beta_\mathit{fix} + N(3\alpha_\mathit{fix})$. (The variable cost is the same in both the vectorized and non-vectorized programs.) 
-The first term in the sum corresponds to the vectorized
-subtraction and multiplication, the second term corresponds to the FOR loop on $j$ and the third 
-one corresponds to the remaining FOR loops on $i$.
-Clearly, $2\alpha_\mathit{fix} + D\beta_\mathit{fix} + N(3\alpha_\mathit{fix}) << ND(2\alpha_\mathit{fix}+\beta_\mathit{fix}) + N3\alpha_\mathit{fix}$.
-Empirically, we observe 10x to 50x improvement over un-vectorized Biometric Matching in circuit generation time and setup time and 5x to 30x improvement in online time.
-Additionally, the un-vectorized version runs out of memory for $N=256$, while the vectorized one runs with the standard maximal input size $N=4,096$.
-\ana{Edit numbers with final experiments.}
-
-\section{Analytical Model}
-\label{sec:model}
-
-\ana{We need an intro to section here.}
-
-\subsection{Scheduling in MPC}
-\label{sec:mpc}
-
-%\ana{Here we need more on MPC before jumping to scheduling. MPC-source, basic assumptions about static loop bounds, etc.}
-
-For this treatment we make the following simplifying assumptions:
-
-\begin{enumerate}
-\item All statements in the program execute using the same protocol (sharing). That is, there is no share conversion. 
-%\ishaq{This is not an assumption, this is our setting. We work in the single protocol world.}
-\item There are two tiers of MPC instructions, local and remote. A local instruction (essentially just ADD) has cost $\beta$ 
-and a remote instruction (e.g., MUX, MUL, SHL, etc.) has cost $\alpha$, where $\alpha >> \beta$. We assume that all remote 
-instructions have the same cost.
-%\ishaq{Don't we need to distinguish between local (cheap) vs.
-interactive (expensive) instructions?
-I don't see how we could assume this.
--- Past suggestion from Vassilis: Ideally we should assume cost to be a convex function.}\ana{After discussion 1/12: replace single costs 1 with two levels of costs: $\alpha$ for non-local operations, e.g., MUL, and $\beta$ for local ops, e.g., ADD.}
-\item  In MPC frameworks, executing $n$ operations ``at once'' in a single SIMD operation costs a lot less than executing those $n$ operations one by one.
-Following Amdahl's law~\ana{citation!}, we write $\alpha = \frac{1}{s}p\alpha + (1-p)\alpha$, where $p$ is the fraction of execution time that benefits from amortization and $(1-p)$ 
-is the fraction that does not, and $s$ is the available resource. Thus, $n\alpha = \frac{n}{s}p\alpha + n(1-p)\alpha$. 
-For the purpose of the model we assume that $s$ is large enough and the term $\frac{n}{s}p\alpha$ amounts to a \emph{fixed cost} incurred regardless of
-whether $n$ is $10,000$ or just $1$. (This models the cost of sending a packet from party A to party B for example.) The bottom line is that amortized execution 
-of $n$ operations is $f(n) = \alpha_\mathit{fix} + n\alpha_{var}$ in contrast to unamortized execution $g(n) = n\alpha_\mathit{fix} + n\alpha_{var}$ and, of course,
-$\alpha_\mathit{fix} << n\alpha_\mathit{fix}$.
-%We assume infinite parallel capacity---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely $\alpha$ or $\beta$. 
-%This is a standard assumption in Cryptographic Parallel RAM. ABY presents empirical support for this assumption~\ana{Add citations. PRAM, ABY.}
-%\ishaq{Comment from Vassilis: We assume there are infinite parallel capacities, this is the standard assumption in PRAM (Cryptographic Parallel RAM).}\ana{PRAM assumption stays.
-We have to replace the unit cost of 1 with a suitable amortization function $f(n)$, which I don't think changes much in the cost modeling and analysis.}
-\item MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption,
-2 MUL instructions can be amortized in a single SIMD instruction that costs $\alpha_\mathit{fix} + 2\alpha_\mathit{var}$, however a MUL and a MUX instruction 
-still cost $2\alpha_\mathit{fix} + 2\alpha_\mathit{var}$ even when scheduled ``in parallel''. 
-%MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption,
-%2 MUL instructions scheduled in parallel benefit from amortization and cost $\alpha$, however a MUL and a MUX instructions scheduled
-%in parallel still cost $2\alpha$. 
-%\ishaq{We need to reword this assumption to something like "$K$ parallel MUL costs much less than $K$ (because they get amortized), but any mix of $K$  MUL and MUX still cost roughly $K$.
-Specifically, we should take away the low constants (2 and 1) because we know for low constants this is not true.} \ana{Again, core assumption that MUL and MUX don't benefit from amortization stays.
-We have to replace constant costs with functions, as in (3).}
-\end{enumerate}
-
-\input{prob_stmt_np_hardness.tex}
-
-\begin{figure*}
-  \includegraphics[width=1.0\linewidth]{figs_paper_SIMD/compiler_framework.png}
-  \caption{Compiler Framework. \ana{Change figure since we are not doing div-and-conquer.}}
-  \label{fig:compiler_framework}
-\end{figure*}
-
-\section{\bf Compiler Framework}
-\label{sec:preliminaries}
-
-%We assume arbitrarily nested loops in the MPC-source IR and read-only arrays.
-We assume that loops range from 0 to some constant N.
-Arrays are linearized (row-major order as in MOTION) and accesses are via functions of the induction variables of the enclosing loops.
-
-\figref{fig:compiler_framework} presents an overview of our compiler. This section describes the front-end phases of the compiler
-and Section~\secref{sec:vectorization} describes our vectorization algorithm. \ana{We'll need a section on MOTION code gen too.}
-%\ana{Need to state this more precisely.}
-
-\subsection{Syntax and Semantic Restrictions}
-\label{sec:syntax}
-
-Source syntax is essentially standard IMP syntax as shown below:
-%\begin{figure}[t]
-{\small
-\[
-\begin{array}{l@{~~~~~}l}
-  \begin{array}{l@{~}l@{~~~}l}
- e & ::= e \; \mathit{op} \; e \mid e \mid \x \mid  \texttt{const} \mid \texttt{A[$e$]} & \mathit{expression} \\
- % s &::= t \;\f & \mathit{field} \\
- %\md &::= t\;{\m}(t \;{\this}, \; t\;\x) \; \{\; \ol{t\;\y}\;\s;\;\return\;\y \; \} 
- %                 & \mathit{method} \\ 
-  s & ::= s;s \mid  & \mathit{sequence} \\
-  & \x = e \mid \texttt{A[$e$]} = e \mid & \mathit{assignment \; stmt} \\
-  & \texttt{for $i$ in range($I$)}: s \mid & \mathit{for \; stmt} \\
-  & \texttt{if $e$: $s$ else: $s$} & \mathit{if \; stmt} \\ 
-%   t &::= q \; {\sf C} & \mathit{qualified \; type}\\
-%   q &::= \high \mid \poly \mid \low & \mathit{FlowCFL \; qualifier}\\
-%    t' &::= q' \; \C & \mathit{field \; qualified \; type}\\
-%   q' &::= \tainted \mid \poly & \mathit{field \; qualifier}\\
-
-  \end{array}
-\end{array}
-\]
-}
-%\end{figure}
-
-The syntax allows for array accesses, arbitrarily nested loops, and if-then-else control flow. 
-Expressions are typed $q \; \tau$, where qualifier $q$ and base type $\tau$ are:
-
-{\small
-\[
-\begin{array}{l@{~~~~~}l}
-  \begin{array}{l@{~}l@{~~~}l}
- \tau & ::= \texttt{int}  \mid \texttt{bool} \mid \texttt{list[int]} \mid  \texttt{list[bool]} & \mathit{base \; types} \\
-  q & ::= \texttt{shared} \mid \texttt{plain} & \mathit{qualifiers} \\
-  \end{array}
-\end{array}
-\]
-}
-
-The type system is mostly standard, and in our experience, a sweet spot between readability and expressivity.
-The \texttt{shared} qualifier denotes shared values, i.e., ones shared among the parties and computed upon 
-under secure computation protocols; the \texttt{plain} qualifier denotes plaintext values. Shared lists denote 
-shared elements, but plaintext length. Subtyping is again as expected: $q_1 \; \tau <: q_2 \; \tau$ 
-iff $q_1 <: q_2$ where $\texttt{plain} <: \texttt{shared}$. Thus, we can convert a plaintext value into a shared one, 
-but not vice versa.
-
-Our compiler imposes certain semantic restrictions that it enforces throughout the various 
-phases of compilation. We note that in some cases, the restrictions 
-can be easily lifted and we plan to do so in future iterations of the work.
-
-\begin{enumerate}
-\item Loops are of the form $0 <= i < I$ and bounds are fixed at compile time. 
-It is a standard restriction in MPC that the bounds must be known at circuit-generation time.
-\item Arrays are one-dimensional. N-dimensional arrays are linearized and accessed 
-in row-major order and at this point the programmer is responsible for linearization
-and access. (This restriction can be easily lifted.)
-\item Array subscrpts are plaintext values as specified by the rule:
-\begin{semantics}
-    \ntyperule{Array Access}{
-%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
-  \Gamma \vdash e : \texttt{plain int} \quad \Gamma \vdash \texttt{A} : q \; \texttt{list}[\tau] \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
-}{
-  \Gamma \vdash \texttt{A[$e$]} : q \; \tau
-}
-\end{semantics}
-The subscript $e$ is a function of the indices of the enclosing loops. 
-For read access, the compiler allows an arbitrary such function. 
-However, it restricts write access to \emph{canonical writes}, i.e., \texttt{A[i,j,k]} = ...
-where \texttt{i}, \texttt{j} and \texttt{k} loop over the three dimensions of \texttt{A}.  
-Write access such as for example \texttt{A[i,j+2]} = ... is not allowed. 
-%\item \ana{Add rules for logical ops}
-%\item \ana{Add rules for arithmetic ops}
-\item The final restriction involves MUX as expressed by the rule: 
-\begin{semantics}
-    \ntyperule{MUX}{
-%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
-  \Gamma \vdash e_1 : q_1 \; \texttt{bool} \quad \Gamma \vdash e_2 : q_2 \; \tau \quad \Gamma \vdash e_2 : q_2 \; \tau \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
-}{
-  \Gamma \vdash \texttt{MUX($e_1$,$e_2$,$e_3$)} : q_1 \vee q_2 \vee q_3 \; \tau
-}
-\end{semantics}
-\end{enumerate}
-Importantly, the arguments of MUX are restricted to base types, as passing array arguments 
-is not valid MOTION MPC. \ana{Check this.} This causes a slight inconvenience as we could not write 
-\begin{pythonn}
-if e: A[i] = val
-\end{pythonn}
-Instead we had to write
-\begin{pythonn}
-if not(e): val = A[i]
-A[i] = val
-\end{pythonn}
-
-We note that while the programmer writes 
-annotations at the level of IMP Source (recall \ana{ref to overview section}), the 
-annotations propagate through the transformations and are checked at the 
-level of MPC Source.
-
-%\ana{Disadvantage of MUX}
-
-For the rest of this section we write $i,j,k$ to denote the loop nest: $i$ is the outermost loop, $j$, is immediately nested in $i$, and so on until $k$
-and we use $I,J,K$ to denote the corresponding upper bounds. We write $A[i,j,k]$ to denote canonical access 
-to an array element. In the program, canonical access is achieved via the standard row-major order formula: $(J*K)*i + K*j + k$.
-To simplify the presentation we describe our algorithms in terms of three-element tuples $i,j,k$, however, discussion easily generalizes to 
-arbitrarily large loop nests.
-
-\subsection{From IMP Source to SSA}
-
-Our compiler translates from Source to SSA as follows:
-
-\begin{enumerate}
-  \item
-        \textbf{Parsing}:
-        Use Python's \texttt{ast} module to parse the input source code to a Python AST.
-  \item
-        \textbf{Syntax checking}:
-        Ensure that the AST matches the restricted subset defined in Section~\secref{sec:syntax}.
-        This step outputs an instance of the \texttt{restricted\_ast.Function} class, which represents our restricted subset of the Python AST.
-  \item
-        \textbf{3-address CFG conversion}:
-        %\benjamin{TODO: Is this a good amount of detail?} \ana{This is good amount of detail for the Tech report. In the actual paper we'll shorten a lot and add pointers to the Tech report.}
-
-        Convert the restricted-syntax AST to a three-address control-flow graph.
-        To do this, first,
-        add an empty basic block to the CFG and mark it as current.
-        Next, for each statement in the restricted AST's function body,
-        process the statement.
-        Statements can either be for-loops, if-statements, or assignments (as in~\secref{sec:syntax}).
-        Rules for processing each kind of statement are given below:
-        \begin{enumerate}
-          \item
-                \textbf{For-loops}:
-                Create new basic blocks for
-                the loop condition
-                (the \emph{condition-block}),
-                the loop body
-                (the \emph{body-block}),
-                and the code after the loop
-                (the \emph{after-block}).
-                Insert a jump from the end of the current block to the condition-block.
-                Then, mark the condition-block as the current block.
-                Insert a for-instruction at the end of the current block with the loop counter variable and bounds from the AST.
-                Next, add an edge from the current block to the after-block labeled ``FALSE'' and
-                an edge from the current block to the
-                body-block labeled ``TRUE''.
-                Then, set the body-block to be the current block
-                and process all statements in the AST's loop body.
-                Finally, insert a jump to the condition-block and set the after-block as current.
-          \item
-                \textbf{If-statements}:
-                Create new basic blocks for
-                the ``then'' statements of the if-statement
-                (the \emph{then-block}),
-                the ``else'' statements of the if-statement
-                (the \emph{else-block}),
-                and the code after the if-statement
-                (the \emph{after-block}).
-                At the end of the current block,
-                insert a conditional jump to the then-block or else-block
-                depending on the if-statement condition in the AST.
-                Next,
-                mark the then-block as current,
-                process all then-statements,
-                and add a jump to the after-block.
-                Similarly,
-                mark the else-block as current,
-                process all else-statements,
-                and add a jump to the after-block.
-                Finally, set the after-block to be the current block,
-                and give it a \emph{merge condition} property equal to the condition of the if-statement.
-          \item
-                \textbf{Assignments}:
-                In the restricted-syntax AST,
-                the left-hand side of assignments
-                can be a variable or an array subscript.
-                If it is an array subscript, e.g., \texttt{A[i] = x},
-                change the statement to \texttt{A = Update(A, i, x)}.
-                If the statement is not already three-address code,
-                for each sub-expression in the right-hand side of the assignment,
-                insert an assignment to a temporary variable.
-        \end{enumerate}
-  \item
-        \textbf{SSA conversion}:
-        Convert the 3-address CFG to SSA with Cytron's algorithm.
-\end{enumerate}
-
-\subsection{From SSA to MPC Source} %: MUX Nodes and Pseudo $\phi$-nodes}
-
-Once the compiler converts the code to SSA,
-it transforms $\phi$-nodes that correspond to if-statements into MUX nodes.
-From the 3-address CFG conversion step,
-$\phi$-nodes corresponding to if-statements will be in a basic block
-with the merge condition property.
-For example, if $X_3 = \phi(X_1,X_2)$ is in a block with merge condition $C$,
-the compiler transforms it into $X_3 = \mathit{MUX}(C, X_1, X_2)$.
-Next, the compiler runs the dead code elimination algorithm from Cytron's SSA paper.
-
-Next, the control-flow graph is \emph{linearized} into MPC Source,
-which has loops but no if-then-else-statements.
-This means that both branches of all if-statements are executed,
-and the MUX nodes determine whether to use results from the then-block or from the else-block.
-The compiler linearizes the control-flow graph with a variation of breadth-first search.
-Blocks with the ``merge condition'' property are only considered
-the second time they are visited,
-since that will be after both branches of the if-statement are visited. 
-(The Python AST naturally gives rise to a translation where each conditional has exactly two targets, 
-and each ``merge condition'' block has exactly two incoming edges, a TRUE and a FALSE edge.
-Thus, each $\phi$-node has exactly two multiplexer arguments, which dovetails into MUX.
-This is in contrast with Cytron's algorithm which operates at the level of the CFG and allows for 
-$\phi$-nodes with multiple arguments.)
-Each time the compiler visits a block,
-it adds the block's statements to the MPC source.
-If the block ends in a for-instruction,
-the compiler recursively converts the body and code after the loop to MPC source
-and adds the for-loop and code after the loop to the main MPC source.
-If the block does not end in a for-instruction,
-the compiler recursively converts all successor branches to MPC source and
-appends these to the main MPC source.
-
-Now, the remaining $\phi$-nodes in MPC source are the loop header nodes. We call these nodes \emph{pseudo} $\phi$-nodes 
-and we write \python{PHI} in MPC Source. A pseudo $\phi$-node $X_1 = $\python{PHI}$(X_0,X_2)$ in a loop header is evaluated 
-during circuit generation. If it is the 0-th iteration, then the $\phi$-node evaluates to $X_0$, otherwise, it evaluates to $X_2$.
-
-%\ana{We need to double check that the following captures all cases of def-use edges.}
-
-\subsection{Dependence Analysis}
-
-The next steps construct the dependence graph and \ana{Add an intro here.}
-
-\subsubsection{Def-use Edges}
-
-The dependence graph has the following def-use edges:
-
-\begin{itemize}
-\item same-level edge $X\rightarrow Y$ where $X$ and $Y$ are in the same loop nest, say $i,j,k$.
-E.g., the def-use edge 9 to 10 in the Biometric MPC Source in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source} is a same-level edge.
-A same-level edge can be a back-edge in which case a $\phi$ node is the target of the edge.
-E.g., 15 to 4 in Biometric is a same-level back-edge.
-\item outer-to-inner $X\rightarrow Y$ where $X$ is in an outer loop nest, say $i$, and $Y$ is in an inner one, say $i,j,k$. E.g., 1 to 4 in Biometric forms is an outer-to-inner edge.
-%A $\phi$-node can be a source or a target of an outer-to-inner forward edge.
-\item inner-to-outer $X\rightarrow Y$ where $X$ is a $phi$-node in an inner loop nest, $i,j,k$, and $Y$ is in the enclosing loop nest $i,j$.
-E.g., the def-use from 8 to 12 gives rise to an inner-to-outer edge.
-%E.g. $\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$ is an inner-to-outer edge. 
-An inner-to-outer edge can be a back-edge as well, in which case both $X$ and $Y$ are phi-nodes with the source $X$ in a loop nested into $Y$'s loop (not necessarily immediately).
-
-%Note that the source is \emph{always} a $\phi$-node of the $i,j,k$ loop and the target is in the immediately enclosing loop.
-The interpretation of this edge is that the use node $Y$ uses the definition made in the last iteration of the inner loop.
-%\ben{The current representation of pseudo $\phi$-nodes shows them attached to the loop header (i.e. in the loop).  We may want to clarify that they are also evaluated at the loop's termination.}
-%\item same-level back-edge $X\rightarrow Y$.
-$Y$ is a phi-node in the header of the loop and $X$ is a definition of the variable in the loop body.
-E.g., $\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{sum}_1,\texttt{min}_1)$ to $\texttt{min}_0 = \phi(\texttt{min}_1, 10000)$ in Biometric is a same-level back-edge.
-%\item inner-to-outer back-edge $X\rightarrow Y$: $X$ and $Y$ are both $\phi$-nodes for some variable. The source $X$ is in a loop nested into $Y$'s loop (not necessarily immediately).
-
-\item mixed forward edge $X\rightarrow Y$.
-$X$ is in some loop $i,j,k$ and $Y$ is in a loop nested into $i,j,k'$.
-We transform mixed forward edges as follows.
-Let $\x$ be the variable defined at $X$.
-We add a variable and assignment $\x' = \x$ immediately after the $i,j,k$ loop.
-Then we replace the use of $\x$ at $Y$ with $\x'$.
-This transforms a mixed forward edge into an "inner-to-outer" forward edge followed by an outer-to-inner forward edge.
-Thus, Basic Vectorization handles one of "same-level", "inner-to-outer", or "outer-to-inner" def-use edges.
- \end{itemize}
-
-\subsubsection{Closures}
-
-We define $\mathit{closure}(n)$ where $n$ is a phi-node. Intuitively, it computes the set of nodes (i.e., statements) that form a dependence cycle with $n$. The closure of $n$ is defined as follows:
-\begin{itemize}
-\item $n$ is in $\mathit{closure}(n)$
-\item $X$ is in $\mathit{closure}(n)$ if there is a same-level path from $n$ to $X$, and $X \rightarrow n$ is a same-level back-edge.
-\item $Y$ is in $\mathit{closure}(n)$ if there is a same-level path from $n$ to $Y$ and there is a same-level path from $Y$ to some $X$ in $\mathit{closure}(n)$.
-\end{itemize}
-
-\subsection{Taint Analysis}
-
-We require that all inputs are marked as either shared or plaintext, however, we do not require any other qualifier annotations in the program.
-Then we infer if intermediate variables and expressions are shared through taint analysis with "taintedness" referring to the shared attribute.
-Specifically, the compiler follows the following rules, which are standard in positive-negative qualifier systems (here \texttt{shared} is the positive qualifier and \texttt{plain} is the negative one):
-\begin{enumerate}
-\item Loop counters and array upper bounds are always \texttt{plain}.
-\item If any variable on the right-hand side of an assignment is shared, then the assigned variable is \texttt{shared}.
-%\item If all variables on the right-hand side of an assignment are plaintext, then the assigned variable is plaintext \ana{Will be covered by the last rule.}
-\item Any variables that cannot be determined as shared via the above rules are \texttt{plain}.
-\end{enumerate}
-
-%The first two rules are standard for taint analysis, and the third rule follows from the MPC problem statement. \ben{Is the explanation for the third rule correct?}  \ana{Yes, that is exactly it!}
-The final rule is needed in examples like the following. In the below snippet \texttt{sum!2} and \texttt{sum!3} form a dependency cycle and there is no \texttt{shared} value that flows to either one. 
-They are inferred as plaintext. 
-%cannot be marked as plaintext through simple taint analysis:
-{\small
-\begin{verbatim}
-plaintext_array = [0, 1, 2, ...]
-sum!1 = 0
-for i in range(0, N):
-    sum!2 = PHI(sum!1, sum!3)
-    sum!3 = sum!2 + plaintext_array[i]
-\end{verbatim}
-}
-%\ben{I think the above example is unnecessary and could be replaced by an explanation of how "untainted" variables are implicitly plaintext, but I couldn't think of a way to phrase that.}\ana{I left the example, we can leave it in the TR, remove from paper.}
-
-When converting to MOTION code, any plaintext value used in the right-hand side of a shared assignment is converted to a shared value for that expression. 
-%\ben{Is this necessary to include?} \ana{Yes, that's important.}
-
-Importantly, the taint analysis works on MPC Source, which lacks if-then-else control flow. 
-This significantly simplifies the treatment as there is no need to handle conditionals and implicit flow.
-
-\subsection{From (Optimized) MPC Source to MOTION}
-\label{sec:MOTION}
-\ana{Ben, add the MOTION section here.}
-
-\section{Vectorization}
-\label{sec:vectorization}
-
-An important component of our algorithm is the scalar expansion to the corresponding loop dimensionality. 
-For example, \texttt{d = S[i * D + j] - C[j]} equiv. to \texttt{d = S[i,j] - C[j]}, which gave rise to \texttt{N*D} subtraction operations in the sequential schedule, 
-is lifted. The argument arrays \texttt{S} and \texttt{C} are lifted and the scaler \texttt{d} is lifted: \texttt{d[i,j] = S[i,j] - C[i,j]}.
-The algorithm then detects that the statement can be vectorized.
-
-There are three kinds of arrays (for now all kept internally as one-dimensional arrays, but that's under discussion).
-\begin{itemize}
-
-\item Scalars: These are scalar variables we lift into arrays for the purposes of vectorization. 
-For those, all writes are canonical writes and all reads are canonical reads. We will apply raise dimension when 
-a scalar is used in an inner loop (e.g., \texttt{sum0} in line 6 of the MPC source code will be raised to a 1-dimensional 
-array since \texttt{sum0} is used in the inner $j$-loop). Drop dimension applies as well; this happens when a scalar
-written in an inner loop is used outside of the loop (e.g., \texttt{sum0} for which the lifted inner loop computes 
-\texttt{D} values, but the outer loop only needs the last one.)
-
-\item Read-only input arrays: Read-only inputs. There are NO writes, while we may have non-canonical reads, $f(i,j,k)$. 
-Phase 1 of Basic vectorization will add raise dimension operation at the beginning of the function to lift these arrays, 
-and raise dimension may \emph{reshape} arrays. If there are multiple "views" of the input array, there would be multiple raise dimension statements to create
-each one of these views. The invariant is that at reads in loops, the reads of "views" of the original input array are canonical.
-Only raise dimension applies to these arrays, and only in the beginning of the program. For example, 1-dimensional array 
-\texttt{C} is lifted into 2-dimensional array \texttt{N x D} by copying the row \texttt{N} times.
-
- \item Read-write output arrays: Writes are canonical (by restriction) but reads can be non-canonical. 
- Dependence analysis limits vectorization when non-canonical read access refers to array writes in
- previous iterations, thus creating loop-carried dependences. We may apply both raise and drop dimension, however, 
- they respect the fixed dimensionality of the output array. The array cannot be raised to a dimension lower than
- its canonical (fixed) dimensionality and it cannot be dropped lower. In addition, non-canonical reads may 
- require lifting (i.e., reshaping) of the array after the most recent write, rather than in the beginning of the program, 
- in order to reduce a non-canonical read to a canonical one.
-
-\end{itemize}
-
-In the sections below we detail the $\mathit{raise\_dim}$ (raise dimension) and $\mathit{drop\_dim}$ (drop dimension) operations,
-followed by our vectorization algorithm.
-
-%\ana{Add back-edges into Phase 1.
-A back-edge from a non-phi-node in loop $i$ to a phi-node in loop $i$'s header is a same-level edge.
-The only difference with the handling of normal same-level forward def-use is that the operand index will become $i-1$.
-A back edge from a phi-node in loop $j$ to a phi-node in loop $i$, where $j$ is nested in $i$ is an inner-to-outer edge and will require dropping dimension.
-We can show these are the only kinds of back-edges that may occur.}
-
-%\subsection{Algorithm (Still work in progress.)}
-
-\subsection{Raise Dimension and Drop Dimension}
-
-There are two conceptual versions of $\mathit{raise\_dim}$. One applies on read-only input arrays and reshapes those arrays 
-when necessary to ensure canonical read access in the corresponding loop. The signature of $\mathit{raise\_dim}$
-is as follows. It takes the original array \texttt{C}, the access pattern function $f(i,j,k)$ in loop nest $i,j,k$ and the loop bounds
-\texttt{((i:I),(j:J),(k:K))}: 
-\[ \mathit{raise\_dim}(\texttt{A}, f(i,j,k), \texttt{((i:I),(j:J),(k:K))}) \]
-It produces a new 3-dimensional array \texttt{A'} by iterating over $i,j,k$ and setting each element of  \texttt{A'} as follows:
-\[ \texttt{A'[i,j,k] =  A[f(i,j,k)]} \]
-The end result is that uses of \texttt{[A[f(i,jk)]} in loop nest $i,j,k$ are replaced with canonical read-accesses to \texttt{A'[i,j,k]}
-that can be vectorized. In the running Biometric example, \texttt{C'} = $\mathit{raise\_dim}$\texttt{(C, j, (i:N,j:D))} lifts the 
-1-dimensional array \texttt{C} into a 2-dimensional array. The $i,j$ loop now accesses \texttt{C'} in the canonical way, \texttt{C'[i,j]}. 
-Similarly, \texttt{S'} = $\mathit{raise\_dim}$\texttt{(S, i*D+j, (i:N,j:D))} tries to lift \texttt{S}, but the operation turns into a no-op 
-because \texttt{S} is already a 2-dimensional array and the read access is canonical.
-
-The other version of $\mathit{raise\_dim}$ applies on scalars and read-write arrays. It lifts a lower-dimension array into a 
-higher-dimension for access in a nested loop. Here \texttt{A} is an $i$ array and raise dimension adds two additional dimensions:
-\[ \mathit{raise\_dim}(\texttt{A}, \texttt{(j:J,k:K)}) \]
-This version is reduced to the above version by adding the access pattern function, which is just $i$: 
-\[ \mathit{raise\_dim}(\texttt{A}, i, \texttt{(j:J,k:K)}) \]
-
-The corresponding $\mathit{drop\_dim}$ is carried out when an array written in an inner loop is used in an enclosing loop. 
-It takes a higher dimensional array, say $i,j,k$ and removes trailing dimensions, say $j,k$:
-\[ \mathit{drop\_dim}(\texttt{A}, \texttt{(j:J,k:K)}) \]
-It iterates over $i$ and takes the result at the maximal index of $j$ and $k$, i.e., the result at the last iterations of $j$ and $k$:
-\[ \texttt{A'[i] = A[i,J-1,K-1]} \]
-
-\begin{comment}
-We define the $\mathit{raise\_dim}$ (raise dimensions) and $\mathit{drop\_dim}$ (drop dimension) functions.
-Raise dimension "lifts" a lower-dimension array (\ana{The right term here is tensor, I am pretty sure!}) into a higher dimension one.
-This is necessary when a lower-dimensional array is used in a higher dimensional loop and, essentially, is just copying of values.
-For example, Biometric contains the statement $d = \mathit{SUB}(S[i, j], C[j])$, in the $j$-loop which is nested into the $i$-loop.
-Here $C$ is a one-dimensional array, however, to vectorize across both loops it is necessary to turn it into a two-dimensional array: $C[i,j]$ becomes $[C[0],C[1],...C[J],C[0],C[1],...C[J], ...
-C[0],C[1],...C[J]]$, which turns the row into a matrix of $I$ identical rows.
-(We use capital letters to denote the upper bounds of loops, e.g., $J$ is the upper bound of the $j$-loop and $I$ is the upper bound of the $i$-loop.)
-
-$\mathit{raise\_dim}(A[i_1,...,i_k],i_j)$ is defined as follows.
-It results in a new {k+1}-dimensional array $A'$ where for every $0 \le i_{k} < I_{k+1}$, $A'[i_1, \ldots, i_j, \ldots, i_k] = A[i_1, \ldots, i_k]$.
-Adding $n$ dimensions is trivially extended as a composition of $n$ $\mathit{raise\_dim}$ that each adds a single dimension.
-
-As expected, drop dimension turns a higher-dimensional array into a lower-dimensional one.
-The key use case is an inner-to-outer def-use edge.
-The code may define a variable, e.g., $\x$ in an inner loop, say $j$, then use this variable in an enclosing loop, say $i$.
-Our algorithm may vectorize the computation of $\x$ in the $j$-loop thus producing a vector $\x[j]$ where $\x[1]$ is the value of \x after the 1st iteration and so on.
-Drop dimension states that the outer loop will use the value of the variable at the last iteration.
-
-$\mathit{drop\_dim}(A[i_1,\ldots, i_j, \ldots,i_k], i_j)$ produces a k-dimensional array $A'$ where $A'[i_1,\ldots,i_k] = A'[i_1,\ldots, I_j-1, \ldots,i_k]$.
-In our analysis, dimensions are always dropped at the end, and again, one can define dropping $n$ dimensions as a composition of $n$ $\mathit{drop\_dim}$.
-
-\ben{I'm not sure if I wrote up the above paragraphs correctly -- the idea is that we're just copying values to extend over dimension $j$ or only retain its last elements.
-The current writeup implies that $1 < j < k$ though, which isn't always true.
-Python implementations of $\mathit{raise\_dim}()$ and $\mathit{drop\_dim}()$ can be found here: \url{https://github.com/milana2/ParallelizationForMPC/blob/master/compiler/compiler/motion_backend/reference_implementations.py}}
-
-\ana{This section is still informal. Needs work to make more precise.}
-\end{comment}
-
-\subsection{Basic Vectorization}
-
-\begin{algorithmic}
-
-\STATE \COMMENT { Phase 1: Raise dimension of scalar variables to corresponding loop nest. We can traverse stmts linearly in MPC-source. }
-
-\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i,j,k$}
-%\FOR {each $Y_n$}
-\FOR {each argument $Y_n$} %def-use edge $\mathit{stmt}' (\mbox{def of } Y) \rightarrow \mathit{stmt} (\mbox{def of } X)$}
-\STATE {case def-use edge $\mathit{stmt}' (\mbox{def of } Y_n) \rightarrow \mathit{stmt} (\mbox{def of } X)$ of} \\
-\STATE \hspace{0.25cm} {same-level: $Y'_n$ is $Y_n$} \\
-\STATE \hspace{0.25cm} {outer-to-inner: add $Y'_n[i,j,k] = \mathit{raise\_dim}(Y_n)$ at $\mathit{stmt'}$}
-\STATE \hspace{0.25cm} {(more precisely, right after $\mathit{stmt'}$)}
-\STATE \hspace{0.25cm} {inner-to-outer: add $Y'_n[i,j,k] = \mathit{drop\_dim}(Y_n)$ at $\mathit{stmt}$}
-\STATE \hspace{0.25cm} {(more precisely, in loop of $\mathit{stmt}$ right after loop of $\mathit{stmt'}$)}
-\ENDFOR
-
-\COMMENT { Optimistically vectorize all. $I$ means vectorized dimension. }
-%\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i,j,k$}
-\STATE {change to $X[I,J,K] = \mathit{Op}(Y'_1[I,J,K],Y'_2[I,J,K])$ }
-\ENDFOR
-
-\COMMENT { Phase 2: Recreating FOR loops for cycles; vectorizable statements hoisted up. }
-
-\FOR {each dimension $d$ from highest to 0}
-\FOR {each $\phi$-node $n$ in loop $i_1,...,i_d$}
-\STATE {compute $\mathit{closure}(n)$}
-\ENDFOR
-
-\COMMENT {{\color{red} $cl_1$ and $cl_2$ intersect if they have common statement or update same array; "intersect" definition can be expanded}}
-
-\WHILE {there are closure $cl_1$ and $cl_2$ that intersect}
-\STATE {merge $cl_1$ and $cl_2$}
-\ENDWHILE
-\FOR {each closure $cl$ (after merge)}
-\STATE {create FOR $i_d = 0; ...$ loop}
-\STATE {add $\phi$-nodes in $cl$ to header block}
-\STATE {{\color{red}{add target-less $\phi$-node for A if $cl$ updates array A}}}
-\STATE {add statements in $cl$ to loop body in some order of dependences}
-\STATE \COMMENT { Dimension is not vectorizable: }
-\STATE {change ${I_d}$ to $i_d$ in all statements in loop}
-\STATE {treat FOR loop as monolith node: some def-use edges become same-level.}
-\ENDFOR
-{
-\color{red}
-\FOR {each target-less $\phi$-node $A_1 = \phi(A_0,A_k)$}
-\STATE {in vectorizable stmts, replace use of $A_1$ with $A_0$}
-\STATE {discard $\phi$-node if not used in any $cl$, replacing $A_1$ with $A_0$ or $A_k$ appropriately}
-\ENDFOR
-}
-\ENDFOR
-
-\COMMENT { Phase 3: Remove unnecessary dimensionality. }
-
-\COMMENT { A dimension $i$ is dead on exit from stmt $X[...i...] = ...$ if all def-uses with targets outside of the enclosing 
-FOR $i=0 ...$ MOTION loop end at target (use) $X' = \mathit{drop\_dim}(X, i)$. } 
-\FOR {each stmt and dimension $X[...i...] = ...$}
-\STATE {if $i$ is a dead dimension on exit from stmt $X[...i...] = ...$, remove $i$ from $X$ (all defs and uses)} 
-%\COMMENT{ A stmt v[...i] is a "dead dimension i" if all edges out of FOR i = 0; ... loop end at drop\_dim(v, i)}
-\ENDFOR
-
-\COMMENT { Now clean up drop\_dim and raise\_dim }
-\FOR {each $X' =  \mathit{drop\_dim(X,i)}$ }
-\STATE {replace with $X' = X$ if $i$ is dead in $X$.}
-\ENDFOR
-\STATE {do (1) (extended) constant propagation, (2) copy propagation and (3) dead code elimination to get rid of redundant variables and raise and drop dimension statements}
-
-\COMMENT { Phase 4: }
-\STATE {add SIMD for simdified dimensions}
-
-\end{algorithmic}
-
-\input{biometric_vectorization_example.tex}
-
-\begin{comment}
-\subsection{Example: Biometric}
-
-\subsubsection{MPC-source}
-
-Below is the MPC-source resulting from Benjamin's analysis. It gives rise to a corresponding dependence graph. (Not shown.)
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0 = \phi(-1, \texttt{index}_1)$}
-\STATE {$\texttt{min}_0 = \phi(10000,\texttt{min}_1)$}
-
-\STATE \COMMENT {Begin of inner loop $j$.}
-\STATE {$\texttt{sum}_0 = \phi(0, \texttt{sum}_1)$}
-\STATE {$\texttt{d} = \mathit{SUB}(\texttt{S}[i,j], \texttt{C}[j])$}
-\STATE {$\texttt{p} = \mathit{MUL}(\texttt{d},\texttt{d})$}
-\STATE {$\texttt{sum}_1 = \texttt{ADD}(\texttt{sum}_0,\texttt{p})$}
-\STATE \COMMENT {End of inner loop $j$.}
-\STATE {$\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$}
-\STATE {$\texttt{index}_1 = \mathit{MUX}(\texttt{c},\texttt{index}_0,i)$}
-\STATE {$\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{min}_0,\texttt{sum}_0)$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-\subsubsection{Phase 1 of Basic Vectorization:}
-
-The transformation preserves the dependence edges.
-It raises the dimensions of scalars and optimistically vectorizes all operations.
-The next phase discovers loop-carried dependences and removes affected vectorization.
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$}
-\STATE {$\texttt{min}_0[i] = \phi(10000, \texttt{min}_1[i-1])$}
-
-\STATE \COMMENT {Begin of inner loop $j$.}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...], \texttt{sum}_1[i,j-1])$}
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{sum}_1[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE \COMMENT {End of inner loop $j$.}
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {$\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$}
-\STATE {$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$}
-\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-\ishaq{At least for input variables: when raising dimensions, we should keep some meta data around so we can later write code that could use 1 share if it knows the other N shares are just duplicates of it.}
-
-\subsubsection{Phase 2}
-
-This phase analyzes statements from the innermost loop to the outermost. The key point is to discover loop-carried dependencies and re-introduce loops whenever dependencies make this necessary.
-
-Starting at the inner phi-node $\texttt{sum}_0[i,j] = \phi([0,0,...],\texttt{sum}_1[i,j-1])$, the algorithm first computes its closure.
-The closure amounts to the phi-node itself and $\texttt{sum}_1[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$, accounting for the loop-carried dependency of the computation of \texttt{sum}.
-The algorithm replaces this closure with a FOR loop on $j$ removing vectoriaztion on $j$.
-Note that the SUB and MUL computations remain outside of the loop as they do not depend on phi-nodes that are part of cycles.
-The algorithm adds same-level edges, one from $\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = ...$ to the monolithic FOR-loop node, and one from the FOR-loop node to $\texttt{c}[\stackrel{\rightarrow}{i}] = ...$.
-
-\ishaq{For our radar: Such dependencies will not involve phi nodes only. e.g. $a[i] = a[i] + a[i-1]$}
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0[i] = \phi(-1, \texttt{index}_1[i-1])$}
-\STATE {$\texttt{min}_0[i] = \phi(10000, \texttt{min}_1[i-1])$}
-
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-
-\STATE \COMMENT {Begin of inner loop $j$. Loop is now much "shorter"!}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...], \texttt{sum}_1[i,j-1])$}
-\STATE {FOR j=0; j<D; j++} \COMMENT{ Will turn into a MOTION loop.}
-\STATE \hspace{0.25cm}{$\texttt{sum}_1[\stackrel{\rightarrow}{i},{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},{j}],\texttt{p}[\stackrel{\rightarrow}{i},{j}])$}
-\STATE \COMMENT {End of inner loop $j$.}
-
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {$\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$}
-\STATE {$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$}
-\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-Next, analysis moves to outer dimension $i$. There are two phi-nodes, $\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1])$ and $\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$.
-The closure of the first is
-\[
-\begin{array}{l}
-\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1]) \\
-\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
-\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])
-\end{array}
-\]
-and the closure of the second one is
-\[
-\begin{array}{l}
-\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1]), \\
-%\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
-\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])
-\end{array}
-\]
-
-Since the two closures \emph{do not} intersect, we have two distinct FOR-loops on $i$.
-The first FOR loop (on \texttt{min}) is scheduled first, because of the dependence edge from $\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$ to
-$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$.
-We can now rewrite those statements, getting rid of the vectorization on $i$ in the two FOR loops.
-Notably, the SUB and MUL computations are fully vectorizable and the ADD computation is vectorizable across the $i$-dimension.
-
-\begin{algorithmic}
-
-\STATE
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB\_SIMD}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL\_SIMD}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$} \\
-\STATE {}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...],\texttt{sum}_1[i,j-1])$}
-\STATE {FOR j=0; j<D; j++} \COMMENT{ Turns into a MOTION loop, was $j$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{sum}_1[\stackrel{\rightarrow}{i},{j}] = \mathit{ADD\_SIMD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},{j}],\texttt{p}[\stackrel{\rightarrow}{i},{j}])$}
-\STATE {}
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {}
-\STATE {$\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1])$}
-\STATE {FOR i=0; i<N; i++} \COMMENT{ Turns into a MOTION loop, was $i$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{c}[{i}] = \mathit{CMP}(\texttt{sum}'_0[{i}],\texttt{min}_0[{i}])$}
-\STATE \hspace{0.25cm}{$\texttt{min}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{min}_0[{i}],\texttt{sum}'_1[{i}])$}
-\STATE {}
-\STATE {$\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$}
-\STATE {FOR i=0; i<N; i++} \COMMENT{ Turns into a MOTION loop, was $i$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{index}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{index}_0[{i}],i)$}
-
-\ben{The above code implicitly stores the final values of min\_0 and index\_0 in their last indices, but there is no primitive like drop\_dim to make it clear.
-This will have to be added in the SSA code generation (though it might not be necessary to include in the paper?).
-It also is pretty wishy-washy about the size and layout of each vector (again more of an implementation issue), but maybe we should have a different syntax for array accesses for read/write (e.g.
-the ADD\_SIMD steps) vs the array definitions (e.g.
-the SUB\_SIMD and raise/drop\_dim statements)}
-
-\end{algorithmic}
-\end{comment}
-
-\subsection{Correctness Argument}
-
-We build a correctness argument that loosely follows the theory of Abstract Interpretation. 
-We define the syntax of MPC Source programs. The domain of MPC Source programs
-expressible in the syntax (with certain semantic restrictions) is the abstract domain $A$. 
-We then define the \emph{linearization} of an MPC Source program as an interpretation 
-over the syntax. The linearization, which is a \emph{schedule}, is the concrete domain $C$. 
-Since we reason over def-use graphs in $A$ we define a partial order relation over elements of $A$ 
-in terms of def-use relations. We define a partial order over elements of $C$ as well,
-in terms of def-use relations in the concrete domain $C$. We prove two theorems
-that state (informally) that the schedule corresponding to the original program computes the same 
-result as the schedule corresponding to the vectorized program.
-
-\begin{figure}%[!b]
-%\small
-\[
-\begin{array}{l@{~~~~~}l}
-  \begin{array}{l@{~}l@{~~~}l}
-  s &::= s;s & sequence \\
-     & \mid \x[i,J,k] = \y[i,J,k] \; \mathit{op\_SIMD} \; \z[i,J,k] & operation \\ 
-     & \mid \x[i,J,k]  = \texttt{PHI}(\x_1[i,J,k],\x_2[i,J,k\!-\!1]) & \mathit{phi} \; \mathit{node} \\
-     & \mid \x[i,J,k] = \mathit{raise\_dim}(\x'[i],(J\!:\!J,k\!:\!K)) & \mathit{raise} \; \mathit{dimension(s)} \\
-     & \mid \x[i,J] = \mathit{drop\_dim}(\x'[i,J,k],k)  & \mathit{drop} \; \mathit{dimension(s)} \\
-     & \mid \x = \y & \mathit{propagation} \\
-     & \mid \texttt{FOR} \; 0 \le i < I: s  & \mathit{loop} \\
-  \end{array}    \\
-\end{array}
-\]
-\vspace{-3mm}
-\caption{MPC Source Syntax}\label{fig:MPC_Source_syntax}
-\vspace{-5mm}
-\end{figure}
-
-\paragraph{MPC Source Syntax}
-
-\figref{fig:MPC_Source_syntax} defines the syntax for our intermediate representation, MPC Source. 
-There are semantics restrictions over the syntax as well: a variable $\x[i,j,k]$ is a 3-dimensional array 
-$(i:I,j:J,k:K)$ and also, a statement $\x[i,J,k] = ...$ is enclosed in loops over $i$ and $k$ as shown below. 
-Thus, $i$ and $k$ are in scope.
-\begin{verbatim}
-FOR 0 <= i < I:
-    ...
-    FOR 0 <= k < K:
-        x[i,J,k] = ...
-\end{verbatim}
-
-Statements $\mathit{operation, phi, raise \; dimension(s), drop \; dimension(s)}$ are \emph{base} statements, 
-and $\mathit{sequence, loop}$ are \emph{compound statements}.
-
-\paragraph{Linearization}
-
-Linearization is the concretization operation, which, as we mentioned earlier computes a schedule.
-The concretization function $\gamma: A \rightarrow C$ is defined as an interpretation of
-MPC Source syntax, as it is standard. The concretization of each one 
-of the base statements is as follows: 
-\[\begin{array}{l}
-\gamma(\x[i,J,k] = \mathit{op\_SIMD}(\y[i,J,k], \z[i,J,k])) = \\ 
-~~\x[i,0,k] = \y[i,0,k] \; \mathit{op} \; \z[i,0,k] \;\; || \\
-~~~~\x[i,1,k] = \y[i,1,k] \; \mathit{op} \; \z[i,1,k] \;\; || \; ...\; || \\
-~~~~~~\x[i,I-1,k] = \y[i,I-1,k] \; \mathit{op} \; \z[i,I-1,k]
-\end{array}\]
-meaning that the vectorized dimension(s) are expanded into parallel statements. || introduces SIMD (parallel) execution.
-
-The concretization of the FOR statement is as follows:
-\[ \gamma(\texttt{FOR} \; 0 \le i < I: s) = \gamma(s)[0/i] \; ; \; \gamma(s)[1/i] \; ; \; ... \gamma(s)[I-1/i] \]
-$\gamma$ simply unrolls the loop substituting $i$ with 0, 1, etc. Here ; denotes sequential execution.
-
-\paragraph{Partial Orders}
-
-For each MPC Source program $a$ we compute the def-use edges in the standard way: if base statement $s1 \in a$ defines variable \x, e.g., $\x[i,j,k] = ...$, 
-and base statement $s2 \in a$ uses \x, e.g., $... = ... \x[i,j,k]$ and there is a path in the trivial CFG from $s1$ to $s2$, then there is a def-use edge from
- $s1$ to $s2$. We extend the dimensionality of a statement into $s1[i,j,k]$ where $s1[i,j,k]$ inherits the dimensionality of the left-hand-side of the assignment.
-
- Let $a_0, a_1$ be two MPC Source programs in $A$. 
- Two base statements, $s_0 \in a_0$ and $s_1 \in a_1$ are \emph{same}, written $s_0 \equiv s_1$ if they are of the same operation and they operate on 
- the same variables: same variable name and same dimensionality. Recall that dimensions in MPC Source are either iterative, lower case, or vectorized, upper case. 
- Two statements are same even if one operates on an iterative dimension and the other one operates on a vectorized one, e.g., $s_0[i,j,k] \equiv s_1[I,j,K]$.
-
- \begin{definition} Let $a_0, a_1 \in A$. We say that $a_0 \le a_1$ iff for every def-use edge $s1 \rightarrow s2$ in $a_0$ there is an edge
- $s1' \rightarrow s2'$ where $s1 \equiv s1'$, $s2 \equiv s2'$ and the two edges of either both forward or both backward.
- \end{definition}
-
- The def-use edges in the concrete schedule are as expected. There is a def use edge from statement $s1$ that defines $\x[\underline{i},\underline{j},\underline{k}]$
- to statement $s2$ that uses $\x[\underline{i},\underline{j},\underline{k}]$ if $s1$ is scheduled ahead of $s2$ in the linear schedule. We note that the underlined
- indices, e.g., $\underline{i}$, refer to fixed values, not iterative or vectorized dimensions since in the concrete schedule all induction variables are expanded. 
- E.g., there is a def-use edge from the statement that defines $\x[0,1,2]$ and a statement that uses $\x[0,1,2]$.
-
-\paragraph{Theorems}
-
-\begin{theorem} $a_0 \le a_1 \Rightarrow \gamma(a_0) \subseteq \gamma(a_1)$.
-\end{theorem}
-
-\begin{theorem} Let $a_0$ be the iterative MPC Source and let $a_1$ be the vectorized MPC Source computed by the vectorization algorithm. 
-We have that $a_0 \le a_1$. 
-\end{theorem}
-
-\ana{Write the proof sketch and final argument, etc.}
-
-\subsection{Extension of Basic Vectorization with Array Writes}
-
-\subsubsection{Removal of Infeasible Edges}
-
-Array writes limit vectorization as they sometimes introduce infeasible loop-carried dependencies. Consider the following example: \ana{Have to add citation to Aiken's paper}
-
-{\small
-\begin{verbatim}
-for i in range(N):
-  A[i] = B[i] + 10;
-  B[i] = A[i] * D[i-1];
-  C[i] = A[i] * D[i-1];
-  D[i] = B[i] * C[i];
-\end{verbatim}
-}
-
-In Cytron's SSA this code (roughly) translates into
-
-{\small
-\begin{verbatim}
-for i in range(N):
-1. A_1 = PHI(A_0,A_2)
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. A_2 = update(A_1, i, B_1[i] + 10); 
-6. B_2 = update(B_1, i, A_2[i] * D_1[i-1]);
-7. C_2 = update(C_1, i, A_2[i] * D_1[i-1]);
-8. D_2 = update(D_1, i, B_2[i] * C_2[i]);
-\end{verbatim}
-}
-
-\begin{comment}
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE \hspace{0.25cm} A[i] = B[i] + 10;
-\STATE \hspace{0.25cm} B[i] = A[i] * D[i-1];
-\STATE \hspace{0.25cm} C[i] = A[i] * D[i-1];
-\STATE \hspace{0.25cm} D[i] = B[i] * C[i];
-\end{algorithmic}
-
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{A}_0$ = $\phi(\texttt{A},\texttt{A}_1)$
-\STATE 2. $\texttt{B}_0$ = $\phi(\texttt{B},\texttt{B}_1)$
-\STATE 3. $\texttt{C}_0$ = $\phi(\texttt{C},\texttt{C}_1)$
-\STATE 4. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$
-\STATE 5. $\texttt{A}_1$ = $\texttt{A}_0$; $\texttt{A}_1$[i] = $\texttt{B}_0$[i] + 10; \COMMENT{equiv. to Update}
-\STATE 6. $\texttt{B}_1$ = $\texttt{B}_0$; $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 7. $\texttt{C}_1$ = $\texttt{C}_0$; $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 8. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
-\end{algorithmic}
-\end{comment}
-
-There is a cycle around \texttt{B\_1 = PHI(B\_0,B\_2)} that includes statement 
-\texttt{A\_1 = update(A\_0, i, B\_1[i] + 10)} and that statement won't be vectorized even 
-though in fact there is no loop-carried dependency from the write of \texttt{B\_1[i]} at 6 to the
-read of \texttt{ ... = B\_1[i]} at 8.
-
-The following algorithm removes certain infeasible loop-carried dependencies that are due to array writes. Consider a loop with index $0 \le j < J$
-nested at $i,j,k$. Here $i$ represents the enclosing loops of $j$ and $k$ represents the enclosed loops in $j$.
-
-\begin{algorithmic}
-\FOR {each array $\texttt{A}$ written in loop $j$}
-\STATE \COMMENT { including enclosed loops in $j$ }
-\STATE dep = False
-\FOR {each pair def: $\texttt{A}_m[f(i,j,k)] = ... $, and use: $ ... = \texttt{A}_n[f'(i,j,k)]$ in loop $j$}
-\IF {$\exists \underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$, s.t.
-$0\!
-\le\!
-\underline{i}\!
-<\!
-I,\: 0\!
-\le\!
-\underline{j}, \underline{j}' \!
-<\!
-J,\: 0\!
-\le\!
-\underline{k}, \underline{k}' \!
-<\!
-K,\: \underline{j}<\underline{j}'$, and $f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$}
-\STATE dep = True
-\ENDIF
-\ENDFOR
-\IF {dep == False}
-\STATE remove back edge into $\texttt{A}$'s $\phi$-node in loop $j$.
-\ENDIF
-\ENDFOR
-\end{algorithmic}
-
-Consider a loop $j$ enclosed in some fixed $\underline{i}$. Only if an update (definition) $\texttt{A}_m[f(i,j,k)] = ... $ at some iteration $\underline{j}$
-references the \emph{same} array element as a use $ ... = \texttt{A}_n[f'(i,j,k)]$ at some later iteration $\underline{j}'$,
-we may have a loop-carried dependence for $\texttt{A}$ due to this def-use pair. (In contrast, Cytron's algorithm inserts a loop-carried dependency every time there is an array update.)
-The algorithm above examines all def-use pairs in loop $j$, including defs and uses in nested loops, searching for values $\underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$ that satisfy
-$f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$. If such values exist for some def-use pair, then there is a potential
-loop-carried dependence on $\texttt{A}$; otherwise there is not and we can remove the spurious backward edge thus ``freeing up'' statements for vectorization. %\ana{Double check this!!!}
-
-Consider the earlier example.
-There is a single loop, $i$.
-Clearly, there is no pair $\underline{i}$ and $\underline{i}'$, where $\underline{i} < \underline{i}'$ that make $\underline{i} = \underline{i}'$
-due to the def-use pairs of \texttt{A} 5-6 and 5-7.
-Therefore, we remove the back edge from 5 to the phi-node 1.
-Analogously, we remove the back edges from 6 to 2 and from 7 to 3.
-However, there are many values $\underline{i} < \underline{i'}$ that make $\underline{i} = \underline{i'}-1$ and the back edge from 8 to 4 remains (def-use pairs for \texttt{D}).
-As a result of removing these spurious edges, Basic Vectorization will find that statement 5 is vectorizable.
-Statements 6, 7 and 8 will correctly appear in the FOR loop.
-
-Note however, that this step renders some array phi-nodes target-less. We handle target-less phi-nodes with a minor extension of Basic Vectorization (Phase 2). 
-First, we merge closures that update the same array. This simplifies handling of array $\phi$-nodes: if each closure is turned into a separate loop
-each loop will need to have its own array phi-node to account for the update and this would complicate the analysis.
-Second, we add the target-less node of array \texttt{A} back to the closure that updates \texttt{A} ---
-the intuition is, even if there is no loop-carried dependence from writes to reads on \texttt{A}, \texttt{A} is written and the write (i.e., update) cannot be vectorized; 
-therefore, the updated array has to carry to the next iteration of the loop. Third, in cases when the phi-node remains target-less, i.e., cases when the
-array write can be vectorized, we have to properly remove the phi-node replacing uses of the left-hand side of the phi-node with its arguments.
-
-%\ana{Commented out previous writeup. Removal/handling of targetless phi-nodes will go into the Extension to Basic vectorization.}
-
-\begin{comment}
-The final step is to remove array phi-nodes that are left without incoming back edges after removal of infeasible edges.
-Such a phi-node is redundant when there are no loop-carried dependences that are due to writes to the corresponding array.
-In our running example, the phi-nodes for \texttt{A}, \texttt{B}, and \texttt{C}, are redundant.
-
-Removing redundant phi-nodes proceeds from \emph{innermost to outermost} loops as follows.
-Denote a redundant phi-node by $\texttt{A}_1 = \phi(\texttt{A}_0,\texttt{A}_k)$. $\texttt{A}_0$ is the array at the beginning of the loop
-and $\texttt{A}_k$ is the array at the end of an iteration (including the last iteration). We consider two cases:
-\begin{enumerate}
-\item[Case 1] Inner-to-outer or same-level def-use edge
-\begin{itemize}
-\item If use is an array read $ ... = \texttt{A}_1[f(i)]$, then replace $\texttt{A}_1$ with $\texttt{A}_0$. This is correct, i.e., it is guaranteed that the use
-references $\texttt{A}_0$; had the use been referencing an array modified in an earlier iteration, the back-edge to the phi-node would
-not have been removed and the phi-node would not have been classified as redundant.
-\item If use is an array write (update) $\texttt{A}_2 = \texttt{A}_1; \texttt{A}_2[i] = ... $ equiv. to $\texttt{A}_2 = \mathit{update}(\texttt{A}_1, i, ...)$ then handling becomes more tricky.
-The semantics of this update is that when $i$ is 0, then $\texttt{A}_0$ is being read.
-But when $i > 0$, then $\texttt{A}_k$ is being read.
-If we are at this case, this means that the def-use edge is same-level and we are removing a redundant node in the loop enclosing the update.
-The removal of the original phi-node entails pushing the phi-node at the update:
-$\texttt{A}_2 = \phi(\texttt{A}_0,\texttt{A}_k); \texttt{A}_2[i] = ... $ equiv. to $\texttt{A}_2 = \mathit{update}(\phi(\texttt{A}_0,\texttt{A}_k), i, ...)$.
-\item If use is a phi-node $... = \phi(\texttt{A}_1,...)$, that means we have an outer-to-inner edge. The
-
-\end{itemize}
-\end{enumerate}
-For each outer-to-inner or same-level def-use edge
-that originates at the node, replace $\texttt{A}_1$ at the use node by $\texttt{A}_0$.
-For each inner-to-outer def-use edge, replace $\texttt{A}_1$ by $\texttt{A}_k$. In this case, the use references the array
-that results from the computation in the loop.
-
-Our running example is transformed as follows. Removing the phi-nodes for \texttt{A}, \texttt{B}, and \texttt{C} transforms the
-replaces the references to $\texttt{A}_0$, $\texttt{B}_0$, and $\texttt{C}_0$, with references to $\texttt{A}$, $\texttt{B}$, and $\texttt{C}$
-respectively, i.e., the arrays before the loop.
-
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$
-\STATE 2. $\texttt{A}_1$ = $\texttt{A}$; $\texttt{A}_1$[i] = $\texttt{B}$[i] + 10;
-\STATE 3. $\texttt{B}_1$ = $\texttt{B}$; $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 4. $\texttt{C}_1$ = $\texttt{C}$; $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
-\end{algorithmic}
-\end{comment}
-
-\begin{comment}
-\subsubsection{Array MUX refinement}
-
-\ana{TODO: WIP, get to BV without MUX refinement for now? For now, skip MUX refinement.}
-
-Next, the algorithm refines array MUX statements.
-MPC-source after Cytron's SSA may result in statements $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$, which imply that any index
-of $A$ can be written at this point and therefore there is a loop-carried dependency. In some cases the MUX can be refined to just a single
-index or a pair of indices, e.g., $\texttt{A}_j[i] = \emph{MUX}(c, \texttt{A}_k[i], \texttt{A}_l[i])$.
-
-%\ana{Clarify why this helps. Still work in progress.}
-This is to reduce the dimensionality of simd-ified computation. Technically, $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$ is
-a simdified operation that can be carried out in parallel "in one round". However, particularly when \texttt{A} is a multi-dimensional array, there is
-substantial increase in the size of the arrays (vectors) we send to SIMD operations. Refining to an update to a specific index would reduce
-the size of those vectors. Note that this is a heuristic that handles a common case, but not all cases of array updates.
-%\ana{TODO: Simplify this algorithm, taking into account the restriction to canonical updates. It should be handling \_all\_ cases.}
-
-\begin{algorithmic}
-\FOR {each $\mathit{stmt}$: $\texttt{A}_j = \emph{MUX}(c, \texttt{A}_k, \texttt{A}_l)$ in the MPC-source seq.}
-\STATE $i_1 = \mathit{find\_update}(A_k)$ \COMMENT { Is null when $A_k = \phi(...)$ }
-\STATE $i_2 = \mathit{find\_update}(A_l)$ \COMMENT { Is null when $A_l = \phi(...)$ }
-\IF {$i_1 == i_2$ or $i_1$ is null or $i_2$ is null} \STATE \COMMENT{ With our restrictions on writes we must have $i_1 = i_2$. }
-\STATE replace $\mathit{stmt}$ with \\
-\STATE $\texttt{A}_j = \texttt{A}_{j-1}$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
-%\ELSIF {$i_1 \ncong i_2$}
-%\STATE replace $\mathit{stmt}$ with \\
-%\STATE $\texttt{A}_j = \texttt{A}_k$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
-\ELSE
-\STATE $\mathit{stmt}$ stays as is
-\ENDIF
-\ENDFOR
-\end{algorithmic}
-\end{comment}
-
-\subsubsection{Restricting Array Writes}
-
-For now, we restrict array updates to \emph{canonical updates}. Assume (for simplicity) a two-dimensional array \texttt{A[I,J]}. A canonical update is the following:
-\begin{verbatim}
-for i in range(I):
-  for j in range(J):
-    ...
-    A[i,j] = ...
-    ...
-\end{verbatim}
-
-The update \texttt{A[i,j]} can be nested into an inner loop and there may be multiple updates, i.e., writes to \texttt{A[i,j]}.
-However, update such as \texttt{A[i-1,j] = ...} or \texttt{A[i-1,j-1] = ...}, etc., is not allowed.
-Additionally, while there could be several different loops that perform canonical updates, they must be of the same dimensionality, i.e., an update of higher or lower dimension, e.g., \texttt{A[i,j,k] = ...} is not allowed.
-We compute the \emph{canonical dimensionality} of each write array by examining the array writes in the original program and rejecting programs that violate the canonical write restriction.
-This restriction simplifies reasoning in this early stage of the compiler; we will look to relax the restriction in future work.
-
-Another restriction/assumption is that we assume the output array is given as input with initial values, and it is of size consistent with its canonical dimensionality. 
-%\ana{This restriction will require that we update our benchmarks that come with output arrays.}
-
-Reads through an arbitrary formula, such as \texttt{A[i-1]} for example, are allowed; currently, the projection function returns dummy values if the read formula is out of bounds; we assume the programmer ensures that the program still computes correct output in this case.
-
-\subsubsection{Changes to Basic Vectorization}
-
-\begin{comment}
-There are two changes to Basic Vectorization to account for write arrays. We can account for them during processing of raise dimension/drop dimension in Phase 1 of Basic Vectorization.
-
-One change to Basic vectorization is the expansion of dimension if the array write or read occurs in a nested loop.
-That is, if there is an update $\texttt{A}[i,j] = ...$ that occurs in loop dimensionality $i,j,k$, $\texttt{A}[i,j]$ will be rewritten into $\texttt{A}[i,j,k]$.
-Similarly, a read $\texttt{A}[f(i,j)]$ will be rewritten into $\texttt{A}[f(i,j),k]$.
-($\texttt{A}[f(i,j),k]$ will be reshaped as $f(i,j)*I*J+k$ during Phase 1 of Basic Vectorization.
-\ana{TODO: Double check.})
-\end{comment}
-
-In addition to the changes for the handling of target-less phi-nodes, Basic Vectorization has to handle def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable.
-The definition can be an update \texttt{A\_2 = update(A\_1,i,...)}, a pseudo $\phi$-node \texttt{A\_2 = PHI(A\_0,A\_1)}, etc..
-Note that $\phi$-nodes for arrays have no subscript operations the way there are subscript operations in analysis-introduced arrays representing scalars.
-While there are variations, the most intuitive implementation will perform Basic Vectorization Phase 1 as is, inserting
-$\mathit{raise\_dim}$ and $\mathit{drop\_dim}$ in the same way.
-However, the implementation of raise dimension and drop dimension will be adapted because the dimension cannot be raised or dropped to a dimension lower than the canonical one.
-Consider a def-use edge $X \rightarrow Y$ for an array $\texttt{A}$.
-
-\begin{enumerate}
-
-\item same-level $X \rightarrow Y$. Do nothing, propagate the array, which happens to be of the right dimension. 
-%\ana{There might be some opportunities to do copy propagation optimization and save some cycles, but let's leave this for later.}
-
-\item inner-to-outer $X \rightarrow Y$ triggers the addition of $\mathit{drop\_dim}$.
-However, the dimensionality cannot be dropped below the canonical dimensionality of the array.
-E.g., if the dimensionality of the loop enclosure $X$ is already at the canonical one, then $\mathit{drop\_dim}$ has no effect.
-
-\item outer-to-inner $X \rightarrow Y$ riggers  $\mathit{raise\_dim}$.
-Again, if the dimensionality of the loop enclosure of $Y$ is smaller or same as the canonical dimensionality of the array, then it has no effect, otherwise, if dimensionality is greater than the canonical dimensionality, $\mathit{raise\_dim(...)}$ (at $X$) is the same as in Basic Vectorization.
-
-\item "mixed" $X \rightarrow Y$. We assume that the mixed edge is transformed into an inner-to-outer followed by outer-to-inner edge before we perform vectorization, just as with Basic vectorization.
-
-\end{enumerate}
-
-If the use of the array is a read $\texttt{A}[f(i,j,k)]$ different than a canonical read $\texttt{A}[i,j,k]$, then we need to add a reshape operation as 
-all arrays are $\texttt{A}[i,j,k]$. It can be added after raise\_dim/drop\_dim or incorporated in these operations.
-The bulk of the change is in Phase 2 of Basic Vectorization as outlined earlier.
-
-\subsubsection{Examples with Array Writes}
-
-\paragraph{Example 1}
-
-First, the canonical dimensionality of all \texttt{A,B,C} and \texttt{D} is 1. 
-After Phase 1 of Basic Vectorization the Aiken's array write example will be (roughly) as follows:
-
-{\small
-\begin{verbatim}
-for i in range(N):
-1. A_1 = PHI(A_0,A_2)
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. A_2 = update(A_1, I, B_1[I] + 10); 
-6. B_2 = update(B_1, I, A_2[I] * D_1[I-1]);
-7. C_2 = update(C_1, I, A_2[I] * D_1[I-1]);
-8. D_2 = update(D_1, I, B_2[I] * C_2[I]);
-\end{verbatim}
-}
-
-Note that since all def-uses are same-level (i.e., reads and writes of the array elements) 
-no raise dimension or drop dimension happens.
-
-\begin{comment}
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{A}_1$ = $\phi(\texttt{A}_0,\texttt{A}_2)$
-\STATE 2. $\texttt{B}_1$ = $\phi(\texttt{B}_0,\texttt{B}_2)$
-\STATE 3. $\texttt{C}_1$ = $\phi(\texttt{C}_0,\texttt{C}_2)$
-\STATE 4. $\texttt{D}_1$ = $\phi(\texttt{D}_0,\texttt{D}_2)$
-\STATE 5. $\texttt{A}_2 = update(\texttt{A}_1, \stackrel{\rightarrow}{i}, \texttt{B}_1[\stackrel{\rightarrow}{i}] + 10)$;
-\STATE 6. $\texttt{B}_2 = update(\texttt{B}_1, \stackrel{\rightarrow}{i}, \texttt{A}_2[\stackrel{\rightarrow}{i}] * \texttt{D}_1[\stackrel{\rightarrow}{i-1}])$;
-\STATE 7. $\texttt{C}_2 = update(\texttt{C}_1, \stackrel{\rightarrow}{i}, \texttt{A}_2[\stackrel{\rightarrow}{i}] * \texttt{D}_1[\stackrel{\rightarrow}{i-1}])$;
-\STATE 8. $\texttt{D}_2 = update(\texttt{D}_1; \stackrel{\rightarrow}{i}, \texttt{B}_2[\stackrel{\rightarrow}{i}] * \texttt{C}_2[\stackrel{\rightarrow}{i}])$;
-\end{algorithmic}
-\end{comment}
-
-Phase 2 computes the closure of 4; $cl = \{4,6,7, 8\}$ 
-while 5 is vectorizable. Recall that 1,2, and 3 are target-less phi-nodes. Since the closure $cl$ includes updates to \texttt{B} and \texttt{C},
-the corresponding phi-nodes are added back to the closure and the def-use edges are added back to the target-less nodes. The uses of \texttt{A\_1} and \texttt{B\_1} in the vectorized statement
-turn into uses of \texttt{A\_0} and \texttt{B\_0} respectively; this is done for all original target-less phi-node.
-(But note that \texttt{A\_0} is irrelevant; the update writes into array \texttt{A\_2} in parallel.)
-Finally, the target-less phi-node for \texttt{A} is discarded.
-
-{\small
-\begin{verbatim}
-1. A_2 = update(A_0, I, ADD_SIMD(B_0[I],10));
-   equiv. to A_2[I] = ADD_SIMD(B_0[I],10)
-FOR i=0; i<N; i++; // MOTION loop
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. B_2 = update(B_1, i, A_2[i] * D_1[i-1]);
-   equiv. to B_2 = B_1; B_2[i] = A_2[i] * D_1[i-1];
-6. C_2 = update(C_1, i, A_2[i] * D_1[i-1]);
-7. D_2 = update(D_1, i, B_2[i] * C_2[i]);
-\end{verbatim}
-}
-
-\begin{comment}
-\begin{algorithmic}
-\STATE 1. $\texttt{A}_2 = update(\texttt{A}_0; \stackrel{\rightarrow}{i}, \mathit{ADD\_SIMD}(\texttt{B}_0[\stackrel{\rightarrow}i],[10,...])$ \COMMENT{Fully vectorized, size N.}
-\STATE FOR i=0; i<N; i++; \COMMENT{ MOTION loop }
-\STATE 2. $\texttt{B}_1$ = $\phi(\texttt{B}_0,\texttt{B}_2)$
-\STATE 3. $\texttt{C}_1$ = $\phi(\texttt{C}_0,\texttt{C}_2)$
-\STATE 4. $\texttt{D}_1$ = $\phi(\texttt{D}_0,\texttt{D}_2)$
-\STATE 5. $\texttt{B}_2 = update(\texttt{B}_1, i, \mathit{MUL}(\texttt{A}_2[i], \texttt{D}_1[{i-1}])$
-\STATE 6. $\texttt{C}_2 = update(\texttt{C}_1, i, \mathit{MUL}(\texttt{A}_2[i], \texttt{D}_1[{i-1}])$
-\STATE 7. $\texttt{D}_2 = update(\texttt{D}_1, i, \mathit{MUL}(\texttt{B}_2[i], \texttt{C}_1[i])$
-\end{algorithmic}
-\end{comment}
-
-\paragraph{Example 2} Now consider the MPC Source of Histogram:
-
-{\small
-\begin{verbatim}
-for i in range(0, num_bins):
-   res1 = PHI(res, res2)
-   for j in range(0, N):
-      res2 = PHI(res1, res3)
-      tmp1 = (A[j] == i)
-      tmp2 = (res2[i] + B[j])
-      tmp3 = MUX(tmp1, res2[i], tmp2)
-      res3 = Update(res2, i, tmp3)     
-return res1
-\end{verbatim}
-}
-
-The canonical dimensionality of \texttt{res} is 1.
-Also, the phi-node \texttt{res1 = PHI(res, res2)} is a target-less phi-node (the implication being that the inner for loop can be vectorized across $i$).
-After Phase 1, Basic vectorization produces the following code (statements are implicitly vectorized along $i$ and $j$).
-In a vectorized update statement, we can ignore the incoming array, \texttt{res2} in this case.
-The update writes (in parallel) all locations of the 2-dimensional array, in this case it sets up each \texttt{res3[i,j] = tmp3[i,j]}.
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I = raise_dim(i, ((i:num_bins),(j:N)))
-for i in range(0, num_bins):
-   res1 = PHI(res, res2') // target-less phi-node
-   res1' = raise_dim(res1, (j:N))
-   for j in range(0, N):
-      res2 = PHI(res1', res3)
-      tmp1 = (A1 == I)
-      tmp2 = (res2 + B1)
-      tmp3 = MUX(tmp1, res2, tmp2)
-      res3 = Update(res2, (I,J), tmp3) 
-   res2' = drop_dim(res2)    
-res1'' = drop_dim(res1)
-return res1''
-\end{verbatim}
-}
-
-Processing the inner loop in Phase 2 vectorizes \texttt{tmp1 = (A1 == I)} along the $j$ dimension but leaves the rest of the statements in a MOTION loop.
-Processing the outer loop is interesting.
-This is because the PHI node is a target-less node, and therefore, there are no closures!
-Several things happen.
-(1) Everything can be vectorized along the $i$ dimension.
-(2) We remove the target-less PHI node, however, we must update uses of \texttt{res1} appropriately: the use at raise\_dim goes to the first argument of the PHI function and the use at drop\_dim goes to the second argument.
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I1 = raise_dim(i, ((i:num_bins),(j:N)))
-
-tmp1[I,J] = (A1[I,J] == I1[I,J])
-
-res1' = raise_dim(res, (j:N)) // replacing res1 with res, 1st arg
-for j in range(0, N):
-   res2 = PHI(res1', res3)
-   tmp2[I,j] = (res2[I,j] + B1[I,j])
-   tmp3[I,j] = MUX(tmp1[I,j], res2[I,j], tmp2[I,j])
-   res3 = Update(res2, (I,j), tmp3)
-   equiv. to res3 = res2; res3[I,j] = tmp3[I,j]
-res2' = drop_dim(res2)   
-res1 = drop_dim(res2') // replacing with res2', 2nd arg. NOOP
-return res1
-\end{verbatim}
-}
-
-\begin{comment}
-The cleanup phase can then get rid of $j$ in the \texttt{res} variables and \texttt{tmp} variables, so the loop operates over the \texttt{num\_bins}-size vectors. 
-This will lead to dropping raise\_dim and drop\_dim as well:
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I1 = raise_dim(i, ((i:num_bins),(j:N)))
-
-tmp1[I,J] = (A1[I,J] == I1[I,J])
-
-for j in range(0, N):
-   res2 = PHI(res1, res3) // now 1-dim array
-   tmp2[I] = ADD_SIMD(res2[I],B1[I,j])
-   tmp3[I] = MUX_SIMD(tmp1[I,j], res2[I], tmp2[I])
-   res3 = Update(res2, I, tmp3[I])
-return res2
-\end{verbatim}
-}
-\end{comment}
-
-\begin{comment}
-\paragraph{Example 2}
-
-Consider MPC-source of Histogram after removal of infeasible edges and redundant phi-nodes, and MUX refinement:
-
-\begin{algorithmic}
-
-\STATE res!1 = []
-\STATE for i: plaintext in range(0, num\_bins):
-\STATE \hspace{0.25cm} res!2 = $\phi$(res!1, res!3)
-\STATE  \hspace{0.25cm} ...
-\STATE res!4 = res!2 \COMMENT{Added due to the "mixed" def-use edge.}
-\STATE for i: plaintext in range(0, num\_bins):
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE \hspace{0.25cm}        for j: plaintext in range(0, N):
-\STATE \hspace{0.5cm}            res!5 = $\phi$(res!4, res!7)
-\STATE \hspace{0.5cm}            !2!3 = (A[j] == i)
-\STATE \hspace{0.5cm}            !3!3 = (res!5[i] + B[j])
-\STATE \hspace{0.5cm}            res!6 = res!5; res!6[i] = !3!3
-\STATE \hspace{0.5cm}            res!7 = res!6; res!7[i] = MUX(!2!3, res!6[i], res!5[i])
-
-\end{algorithmic}
-
-After Phase 1 of Basic Vectorization:
-
-\begin{algorithmic}
-\STATE res!1 = []
-\STATE for i: plaintext in range(0, num\_bins):
-\STATE \hspace{0.25cm} res!2 = $\phi$(res!1, res!3)
-\STATE  \hspace{0.25cm} ...
-\STATE res!4 = res!2 \COMMENT{Added due to the "mixed" def-use edge. Because the canonical dimensionality of res is 1, we DO NOT drop dimension of res!2.}
-\STATE A[$i,j$] = $\mathit{raise\_dim}$(A[$j$],$i$) \COMMENT{Copying row A num\_bin times. Reusing name.}
-\STATE B[$i,j$] = $\mathit{raise\_dim}$(B[$i$],$j$) \COMMENT{Copying column B N times. Reusing name.}
-\STATE res!4[$i,j$] = $\mathit{raise\_dim}$(res!4[$i$],$j$) \COMMENT{Reusing name.}
-\STATE for i: plaintext in range(0, num\_bins):
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE \hspace{0.25cm}        for j: plaintext in range(0, N):
-\STATE \hspace{0.5cm}            res!5 = $\phi$(res!4, res!7) \COMMENT{Do nothing. res!4 has been raised to proper dimension.}
-\STATE \hspace{0.5cm}            !2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{EQ}$(A[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], [[0,0...], [1,1...],...[num\_bins-1,num\_bins-1...]])
-\STATE \hspace{0.5cm}            !3!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{ADD}$(res!5[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$],B[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$])
-\STATE \hspace{0.5cm}            res!6 = res!5 \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.5cm}            res!6[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = !3!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$]
-\STATE \hspace{0.5cm}            res!7 = res!6
-\STATE \hspace{0.5cm}            res!7[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{MUX}$(!2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], res!6[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], res!5[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$])
-\end{algorithmic}
-
-After Phase 2 and Phase 3. The EQ operation is vectorizable across both dimensions, and the rest of the computation is vectorizable across the $i$-dimension.
-
-\begin{algorithmic}
-\STATE ...
-\STATE !2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{EQ\_SIMD}$(A[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], [[0,0...], [1,1...],...[num\_bins-1,num\_bins-1...]])
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE FOR j=0; j<N; j++; \COMMENT{MOTION loop}
-\STATE \hspace{0.25cm}            res!5 = $\phi$(res!4, res!7) \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.25cm}            !3!3[$\stackrel{\rightarrow}i,j$] = $\mathit{ADD\_SIMD}$(res!5[$\stackrel{\rightarrow}i,j$],B[$\stackrel{\rightarrow}i,j$])
-\STATE \hspace{0.25cm}            res!6 = res!5 \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.25cm}            res!6[$\stackrel{\rightarrow}i,j$] = !3!3[$\stackrel{\rightarrow}i,j$]
-\STATE \hspace{0.25cm}            res!7 = res!6
-\STATE \hspace{0.25cm}            res!7[$\stackrel{\rightarrow}i,j$] = $\mathit{MUX\_SIMD}$(!2!3[$\stackrel{\rightarrow}i,j$], res!6[$\stackrel{\rightarrow}i,j$], res!5[$\stackrel{\rightarrow}i,j$])
-\end{algorithmic}
-\end{comment}
-
-%\section{Divide-and-Conquer}
-%\label{sec:div_and_conq}
-%\ana{TODO: Now that we have broken FOR loops into smaller chunks, we can add Divide-and-conquer reasoning with Z3 and implement this additional transform.}
-
-\section{Implementation and Evaluation}
-\label{sec:implementation}
-
-\section{Related Work}
-\label{sec:related}
-
-\paragraph{MPC languages and compilers.}
-
-Languages and compilers for secure computation have seen significant attention and advances in recent years. The early MPC compilers Fairplay~\cite{CCS:BenNisPin08}, 
-and Sharemind~\cite{ESORICS:BogLauWil08} were followed by PICCO~\cite{CCS:ZhaSteBla13}, Obliv-C~\cite{Zahur:2015}, TinyGarble~\cite{SP:SHSSK15}, 
-Wystiria~\cite{SP:RasHamHic14}, and others. A new generation of MPC compilers 
-includes SPDZ/SCALE-MAMBA/MP-SPDZ~\cite{Keller:2020} and the ABY/HyCC/MOTION~\cite{NDSS:DemSchZoh15,CCS:BDKKS18,Braun:2022} frameworks. 
-These two families are the state-of-the art and
-are actively developed. Another recent development is Viaduct, a functional language and compiler that supports a range of secure computation 
-frameworks, including MPC and ZKP. Hastings et el. present a review of compiler frameworks~\cite{Hastings:2019}.
-
-While each of these languages and compilers brings in new ideas and advances, none addresses the problem of ``circuit independent'' intermediate 
-representation and optimization. We envision a classical compiler structure: (1) a Wysteria, Viaduct, Obliv-C, or IMP Source front end, including rich type systems and 
-AST-level semantic analysis, compile into the MPC Source IR, (2) MPC Source-level optimizations take place, followed by (3) back-end compilers into circuits. 
-Our focus is at the intermediate level.
-
-%{\begin{center}
-%\includegraphics[width=0.5\linewidth]{figs/focus.png}
-%\end{center}
-%}
-
-Many works focus on the implementation of MPC protocols exposing an API to the programmer. For example, the ABY/MOTION line of 
-compiler frameworks provides a library of MPC primitives; the programmer writes MPC programs in C++ on top of the library. These back ends implement 
-different protocols and allow for mixing, but notably, they leave it to the programer to assign different protocols to different parts of the 
-computation and perform share conversion accordingly. In addition, MOTION provides SIMD primitives, which allows for efficient execution 
-of MPC operations, but again, using SIMD primitives is the responsibility of the programer. There is interest in frameworks for automatic mixing, 
-e.g., \cite{CCS:BDKKS18,Ishaq:2019,Fang:2022}.
-
-Other works, e.g., Obliv-C~\cite{Zahur:2015}, Wystiria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021} focus on higher-level language design, particularly information-flow systems that 
-restrict flow between secure and insecure parts of the program. 
-%Key ideas such as ${\sf insecure} <: {\sf secure}$ qualifier systems are similar to the ideas we 
-%present in~\Secref{sec:interprocedural} and~\Secref{sec:zkp} and we intend to draw on these work. While we will design a high-level language expanding our 
-%intraprocedural IMP-MPC syntax and semantics into the \emph{interprocedural} case, our interest and focus lie primarily in the ``machine-independent'' optimizations 
-%thrust of the proposal.
-
-%There is significant development in ZKP languages as well. ~\ana{Alex Ozdemir's talk, ZoKrates, Cairo, others} Our work focuses on ZKP via MPC and our key idea
-%is the separation of the higher-level ``backend-independent'' compilation and optimization and the lower-level compilation into circuits.
-
-\paragraph{Classical HPC compilers.}
-Automatic vectorization is a longstanding problem in high-performance computing (HPC). 
-There are thousands of works in this area reflecting over 40 years of research. We presented a vectorization
-algorithm for MPC Source, essentially extending classical loop vectorization \cite{Allen:1987}. In HPC vectorization, conditional control flow 
-presents a challenge --- one cannot estimate the cost of a schedule or vectorize branches in a straightforward manner --- in contrast to 
-MPC Source vectorization. 
-We view Karrenberg's work on Whole function vectorization~\cite{Karrenberg:2015} as most closely related to ours --- it linearizes the program and vectorizes 
-both branches of a conditional applying masking to avoid execution of the branch-not-taken code, and selection (similar to MUX) to select the correct value based on the 
-result of the condition at runtime. 
-The problem is that masking and selection, or more generally, handling control predicates~\cite{Benabderrahmane:2010,Karrenberg:2015}, 
-can lead to \emph{slowdown}. 
-%As a result, vectorizing compilers and polyhedral compilers tend to focus on straight-line code due to the unpredictable 
-%overhead conditionals may introduce. \ana{add Louis noel pochet's benchmarks. citation on vectorizing compilers???}
-
-We argue that vectorization over linear MPC Source is a different problem, one that warrants a new look, while drawing from 
-results in HPC.
-Since both branches of the conditional and the multiplexer \emph{always} execute, not only can we apply aggressive vectorization on linear code, but (perhaps more importantly) 
-we can also build analytical models that accurately predict execution time. These models in turn would drive optimizations such as vectorization, protocol mixing, and others. 
-Vectorization meshes in with those additional optimizations in non-trivial ways.
-
-Furthermore, extensions of classical loop vectorization with array writes, arbitrary indexing, including non-affine indexing, and interaction with SSA are non-trivial 
-and present novel challenges and opportunities for contribution. Polyhedral parallelization~\cite{Benabderrahmane:2010} considers a higher-level source (typically AST) 
-representation, while our work takes advantage of linear MPC Source and SSA form. The work by Karrenberg~\cite{Karrenberg:2015} is rare in that space, in the sense that it considers 
-vectorization over SSA form, which has similarities to MPC Source. We consider different array representation, notion of dependence, and reasoning 
-about dependence, which we conjecture is more suitable for MPC Source. Buscher~\cite{Buscher:2018} considers SIMD-vectorization 
-at the level of source code, which then combines with circuit-level optimizations in the TinyGarble compiler. 
-He proposes using an off-the-shelf polyhedral compiler, however, application is limited to only two routines, 
-essentially just inner product and euclidian distance; it is unclear how effective the off-the-shelf compiler is. 
-In contrast, we consider vectorization at the level of MPC Source separating ``backend-independent'' vectorization
-and circuit-level amortization (done by MOTION). We apply our compiler on a wide range of routines.
-
-%\section{Future Work}
-%\label{sec:implementation_and_benchmarks}
-%\input{../sections/implementation_and_benchmarks.tex}
-%\input{../sections/evaluation.tex}
-
-\section{Conclusion and Future Work}
-\label{sec:conclusion}
-%\input{../sections/conclusion.tex}
-
-%\begin{acks}
-%    \ishaq{TODO}
-%\end{acks}
-
-%\printbibliography
-
-\bibliographystyle{plain}
-%\bibliography{c:/trink/Doc/MyTexFiles/texrefs}
-%\bibliography{mybib,library,library1,library2,library3,library_ana}
-\bibliography{library_ana,crypto_crossref}
-
-% that's all folks
-\end{document}

### 2022-04-28 overleaf/model.tex
Split paper into sections.

--- 2022-04-27 overleaf/paper_SIMD.tex
@@
-% !TeX TXS-program:compile = txs:///pdflatex/[--shell-escape]
-%
-% acm templates v1.57 (sigconf template)
-%
-%\documentclass[sigconf, screen, review, anonymous, natbib=false]{acmart}
-\documentclass[sigconf, screen, natbib=false, dvipsnames, table]{acmart}
-\settopmatter{printacmref=true}
-% mandatory for CCS'19
-\usepackage{balance}
-\usepackage{mathtools}
-% for creating a balanced last page (usually last page with references)
-
-% % defining the \BibTeX command - from Oren Patashnik's original BibTeX documentation.
-% \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08emT\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-
-% Rights management information.
-% This information is sent to you when you complete the rights form.
-% These commands have SAMPLE values in them; it is your responsibility as an author to replace
-% the commands and values with those provided to you when you complete the rights form.
-%
-% These commands are for a PROCEEDINGS abstract or paper.
-
-\copyrightyear{2019}
-\acmYear{2019}
-\setcopyright{acmcopyright}
-\acmConference[CCS '19]{2019 ACM SIGSAC Conference on Computer and Communications Security}{November 11--15, 2019}{London, United Kingdom}
-\acmBooktitle{2019 ACM SIGSAC Conference on Computer and Communications Security (CCS '19), November 11--15, 2019, London, United Kingdom}
-\acmPrice{15.00}
-\acmDOI{10.1145/3319535.3339818}
-\acmISBN{978-1-4503-6747-9/19/11}
-
-%
-% These commands are for a JOURNAL article.
-%\setcopyright{acmcopyright}
-%\acmJournal{TOG}
-%\acmYear{2018}\acmVolume{37}\acmNumber{4}\acmArticle{111}\acmMonth{8}
-%\acmDOI{10.1145/1122445.1122456}
-
-%
-% Submission ID.
-% Use this when submitting an article to a sponsored event. You'll receive a unique submission ID from the organizers
-% of the event, and this ID should be used as the parameter to this command.
-%\acmSubmissionID{123-A56-BU3}
-
-%
-% The majority of ACM publications use numbered citations and references. If you are preparing content for an event
-% sponsored by ACM SIGGRAPH, you must use the "author year" style of citations and references. Uncommenting
-% the next command will enable that style.
-%\citestyle{acmauthoryear}
-
- \makeatletter
- % \renewcommand{\section}{\abovedisplayskip 10\p@ \@plus3\p@ \@minus1\p@%
- %                       \belowdisplayskip 5\p@ \@plus3\p@ \@minus1\p@%
- %                       \abovedisplayshortskip 0pt \@plus2\p@%
- %                       \belowdisplayshortskip 0pt \@plus2\p@ \@minus0\p@%
- %                       \@startsection{section}{1}{\z@}%
- %                        {-17\p@ \@plus -4\p@ \@minus -4\p@}%
- %                        {6\p@ \@plus 4\p@ \@minus 4\p@}%
- %                        {\normalfont\large\bfseries\boldmath
- %                         \rightskip=\z@ \@plus 8em\pretolerance=10000 }}
- \renewcommand{\subsection}{\@startsection{subsection}{2}{\z@}%
-                        {-8\p@ \@plus -4\p@ \@minus -4\p@}%
-                        {5\p@ \@plus 2\p@ \@minus 2\p@}%
-                        {\normalfont\Large\bfseries\boldmath
-                         \rightskip=\z@ \@plus 3em\pretolerance=10000 }}
-  \renewcommand{\subsubsection}{\@startsection{subsubsection}{3}{\z@}%
-                        {-6\p@ \@plus -4\p@ \@minus -4\p@}%
-                        {1\p@ \@plus 1\p@ \@minus 0\p@}%
-                        {\normalfont\normalsize\bfseries\boldmath}}
-% \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\z@}%
-%                       {-8\p@ \@plus -4\p@ \@minus -4\p@}%
-%                       {-2\p@ \@plus -0.22em \@minus -0.1em}%
-%                       {\normalfont\normalsize\bfseries}}
- \makeatother
-
-% *** GRAPHICS RELATED PACKAGES ***
-\usepackage{graphicx}
-\usepackage{comment}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{listings}
-% declare the path(s) where your graphic files are
-\graphicspath{{../figures/}}
-% and their extensions so you won't have to specify these with
-% every instance of \includegraphics
-\DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-
-% Minted is used for source code highlighting
-%\usepackage[outputdir=/Volumes/ramdisk/]{minted}
-% \usepackage{minted}
-% *** MATH PACKAGES ***
-\usepackage{amsmath}
-
-% *** SPECIALIZED LIST PACKAGES ***
-%\usepackage{algorithm2e}
-\usepackage{algorithmic}
-
-% *** ALIGNMENT PACKAGES ***
-\usepackage{array}
-
-\usepackage{url}
-
-\usepackage[utf8]{inputenc}
-\usepackage[american]{babel}
-%\usepackage[backend=biber,style=ACM-Reference-Format]{biblatex}
-%\addbibresource{./library.bib}
-%\addbibresource{./cryptobib/abbrev2.bib}
-%\addbibresource{./cryptobib/crypto_crossref.bib}
-
-% Used for Theorems and Definitions
-\usepackage{amsthm}
-\newtheorem{definition}{Definition}
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}{Lemma}
-\theoremstyle{definition}
-\newtheorem{remark}{Remark}
-\newtheorem{assumption}{Assumption}
-
-\usepackage{xcolor}
-\usepackage{xspace}
-\usepackage{xstring} % required by IfEqCase
-
-\usepackage{booktabs} % better tables
-\usepackage{algorithm}
-\usepackage{algorithmic}
-
-% for references, use it like this: \cref{some_label}
-% it will automatically put Appendix, Figure, Section, etc before the reference
-\usepackage[capitalise]{cleveref}
-
-% ishaq: adding our macros
-\input{macros.tex}
-
-\sloppy
-\begin{document}
-\fancyhead{}
-% do not delete this code.
-
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
-\title[Compilation and Backend-Independent Vectorization for Multi-Party Computation]{Compilation and Backend-Independent Vectorization for Multi-Party Computation}
-
-%
-% The "author" command and its associated commands are used to define the authors and their affiliations.
-% Of note is the shared affiliation of the first two authors, and the "authornote" and "authornotemark" commands
-% used to denote shared contribution to the research.
-
-\author{Benjamin Levy}
-\email{levyb3@rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Benjamin Sherman}
-\email{shermb@rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Muhammad Ishaq}
-%\authornote{This work was done in part while the author was at RPI.}
-\email{ishaqm@purdue.edu}
-%\orcid{1234-5678-9012}
-%\author{G.K.M. Tobin}
-%\authornotemark[1]
-%\email{webmaster@marysville-ohio.com}
-\affiliation{%
-    \institution{Purdue University}
-    %\streetaddress{P.O. Box 1212}
-    \city{West Lafayette}
-    \state{Indiana}
-%    \postcode{43017-6221}
-}
-
-\author{Lindsey Kennard}
-\email{fire.elemental@gmail.com}
-\affiliation{%
-    \institution{STR}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Boston}
-    \state{Massachusets}
-    %\country{Iceland}
-}
-
-\author{Ana Milanova}
-\email{milanova@cs.rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Vassilis Zikas}
-%\authornote{This work was done in part while the author was visiting UCLA and supported in part by DARPA and SPAWAR under contract N66001-15-C-4065 and by a SICSA Cyber Nexus Research Exchanges grant.}
-\email{vzikas@purdue.edu}
-%\orcid{1234-5678-9012}
-%\author{G.K.M. Tobin}
-%\authornotemark[1]
-%\email{webmaster@marysville-ohio.com}
-\affiliation{%
-    \institution{Purdue University}
-    %\streetaddress{P.O. Box 1212}
-    \city{West Lafayette}
-    \state{Indiana}
-    %    \postcode{43017-6221}
-}
-
-%
-% By default, the full list of authors will be used in the page headers. Often, this list is too long, and will overlap
-% other information printed in the page headers. This command allows the author to define a more concise list
-% of authors' names for this purpose.
-%\renewcommand{\shortauthors}{Trovato and Tobin, et al.}
-
-\renewcommand{\A}{{\sf A}}
-\renewcommand{\B}{{\sf B}}
-\renewcommand{\C}{{\sf C}}
-\renewcommand{\D}{{\sf D}}
-
-% As a general rule, do not put math, special symbols or citations
-% in the abstract
-\begin{abstract}
-
-\ana{We need an abstract.}
-
-\end{abstract}
-
-%
-% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-% Please copy and paste the code instead of the example below.
-%
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10003752.10010124.10010138.10010143</concept_id>
-<concept_desc>Theory of computation~Program analysis</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10003752.10003777.10003789</concept_id>
-<concept_desc>Theory of computation~Cryptographic protocols</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10002978.10002979</concept_id>
-<concept_desc>Security and privacy~Cryptography</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Theory of computation~Program analysis}
-\ccsdesc[500]{Theory of computation~Cryptographic protocols}
-\ccsdesc[300]{Security and privacy~Cryptography}
-
-%
-% Keywords. The author(s) should pick words that accurately describe the work being
-% presented. Separate the keywords with commas.
-\keywords{multiparty computation; compilers; cryptography}
-
-\begin{comment}
-%
-% A "teaser" image appears between the author and affiliation information and the body
-% of the document, and typically spans the page.
-\begin{teaserfigure}
-    \includegraphics[width=\textwidth]{sampleteaser}
-    \caption{Seattle Mariners at Spring Training, 2010.}
-    \Description{Enjoying the baseball game from the third-base seats. Ichiro Suzuki preparing to bat.}
-    \label{fig:teaser}
-\end{teaserfigure}
-\end{comment}
-%
-% This command processes the author and affiliation and title information and builds
-% the first part of the formatted document.
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-
-Multi-party computation (MPC) allows $N$ parties $p_1,\dots, p_N$ to perform a computation on their private inputs securely.
-Informally, security means that the secure computation protocol computes the correct output (correctness) and it does not leak any information about the individual party inputs, other than what can be deduced from the output (privacy).
-
-% The first results from the early 80's~\cite{FOCS:Yao82b,STOC:GolMicWig87,STOC:BenGolWig88,STOC:ChaCreDam88} introduced the problem and formally defined its security, and investigate feasibility in different models---e.g., passive or malicious adversaries, unconditional or computational security, with or without a broadcast channel.
-This spawned a rich theoretical literature of beautiful works improving and extending in various manners the original ideas.
-
-MPC theory dates back to the early 1980-ies~\cite{FOCS:Yao82b,STOC:GolMicWig87,STOC:BenGolWig88,STOC:ChaCreDam88}. 
-Long the realm of theoretical cryptography, MPC has seen significant advances in programming technology in recent years. 
-These advances bring MPC closer to practice and wider applicability --- 
-MPC technology has been employed in real-world scenarios such as auctions~\cite{FC:BCDGJK09}, biometric identification~\cite{ESORICS:BlaGas11}, 
-and privacy-preserving machine learning~\cite{SP:MohZha17,CCS:MohRin18}. 
-The goal is to bring the technology to a level where programmers can write \emph{secure} and \emph{efficient} programs without 
-commanding extensive knowledge of cryptographic primitives.
-
-The problem, therefore, is to build a high-level programming language and a compiler, and there has been significant advance in 
-this space, e.g., ~\cite{CCS:BenNisPin08,ESORICS:BogLauWil08,CCS:ZhaSteBla13,Hastings:2019,Keller:2020,Acay:2021,Braun:2022} among other work. 
-Current research largely falls at the two ends 
-of the classical compiler: (1) work on \emph{front-end} language design and (2) work on \emph{back-end} protocol implementation. Work on language design 
-focuses on high-level constructs necessary to express multiple parties, computation by different parties, and information flow from one party 
-to another~\cite{SP:RasHamHic14, Acay:2021}. On the other end, work on protocol implementation focuses on cryptographic foundations and their efficient circuit-level
-implementation~\cite{NDSS:DemSchZoh15, CCS:ABFKLO18, Braun:2022}, e.g., implementation of operations (e.g., MUL, ADD) using different sharing protocols 
-(Boolean or Arithmetic GMW~\cite{STOC:GolMicWig87} or Yao's garbled circuits~\cite{FOCS:Yao82b}), as well as efficient share conversion from one representation 
-to another.
-
-Earlier compilers did both back-end and front-end translation without a specific focus on either, as their aim was to demonstrate applicability of MPC on real-world
-programming problems. As the field advanced, works have focused more closely on front-end language design (e.g., Wysteria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021}) 
-or back-end ``circuit-level'' design and implementation (e.g., MOTION~\cite{Braun:2022}).
-
-In this work we focus on an intermediate language and what we call \emph{backend-independent optimizations}, in a close analogy to \emph{machine-independent} optimizations
-in the classical compiler. The following figure summarizes our key idea:
-
-{\begin{center}
-\includegraphics[width=0.8\linewidth]{figs_paper_SIMD/focus.png}
-\end{center}
-}
-
-We emphasize the MPC Source intermediate representation and optimization over MPC Source. As in classical compilers, we envision different front ends (e.g., Wysteria,
-our front end IMP-MPC) compiling into MPC Source. MPC Source is particularly suitable for optimizations such as protocol mixing~\cite{Ishaq:2019, Fang:2022} and SIMD-vectorization, 
-which takes advantage of amortization at the circuit level. The MPC Source IR exposes the linear structure of MPC programs, which simplifies program analysis; this is
-in contrast to source, which has if-then-else constructs. In the same time, MPC Source is sufficiently ``high-level''  to support analyses and optimizations that take into account 
-control and data flow in a specific program. MPC Source is small in size and analysis is tractable, as opposed to analysis over an unrolled circuit~\cite{Ishaq:2019}. Again as
-in classical compilers, we envision translation of MPC Source (optimized or unoptimized) into MOTION, SPDZ, or other back-end code.
-
-%In contrast, the high-level program contains if-then-else and array writes, or the low-level circuit . 
-%The compilation problem, i.e., translation from a high-level language to circuits, is difficult and therefore, breaking the problem into sub-problems can  tractable.
-
-\subsection{Our Contribution} In this paper, we develop a compiler framework that takes a Python-like routine and produces MOTION code: we describe (a) the IMP-MPC language, its syntax and semantic restrictions, (b) translation from IMP-MPC into MPC Source, (c) a specific backend-independent optimization: novel SIMD-vectorization on MPC Source, and (d) translation from MPC Source into MOTION code.
-%\ana{I think we have to add more on MOTION, specifically mixing and amortization.}
-
-We focus on the MOTION framework as our back-end for several reasons.
-First, it is the state of the art in terms of performance~\cite{Braun:2022}.
-Second, it provides an API over efficient implementation for a wide variety of cryptographic operations in three different protocols ---  Arithmetic GMW, Boolean GMW, and BMR --- which allows for protocol mixing~\cite{Ishaq:2019, Fang:2022}, a known backend-independent optimization.
-Third, MOTION provides API for SIMD-level operations, which amortize cost and lead to significant improvement in memory footprint and
-throughput~\cite{NDSS:DemSchZoh15, CCS:ABFKLO18, Braun:2022}. It also enables MPC Source-level optimizations such SIMD vectorization.
-
-Our second contribution is an analytical model for cost estimation of amortized schedules. Originally, we hoped that optimal scheduling (under our model) was tractable, as the problem
-appeared simpler than the classical scheduling problem. Unfortunately, we show that optimal scheduling is NP-hard via a reduction to the Shortest Common Supersequence (SCS) problem. 
-Cost estimation is important as it drives not only vectorization but other optimizations such as protocol mixing optimizations and scheduling as well~\cite{Ishaq:2019, Fang:2022}.
-
-Our most important contribution is the implementation and evaluation of the compiler framework.
-We demonstrate expressivity of the source language by running the compiler on \ana{X} programs; these include classical MPC benchmarks such as PSI and Biometric matching, as well as kMeans, Histogram, and others.
-Vectorization leads to \ana{X}x improvement in number of gates on average, \ana{Y}x speedup in circuit generation time, \ana{Z}x speedup in the LAN setting, and \ana{Z}x speedup in the WAN setting using the Boolean GMW protocol.
-Results form BMR are similar.
-
-Our results emphasize the importance of backend-independent optimizations --- vectorization (described in this work) and protocol mixing (tackled in previous works~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022}) are two optimizations readily available at the level of MPC Source.
-We believe that our work can lead to future work on backend-independent compilation and optimization, bringing in new MPC optimizations and combinations of optimizations, much in the vein of standard compilers, and bring MPC programming technology closer to practice and wider applicability.
-
-\subsection{Outline}
-
-The rest of the paper is organized as follows.
-\secref{sec:overview} presents an overview of the compiler.
-\secref{sec:model} describes our model for cost estimation and argues NP-hardness of optimal scheduling and \secref{sec:compiler} describes the compiler.
-\secref{sec:experiments} presents implementation and evaluation.
-\secref{sec:related} discusses related work and \secref{sec:conclusion} concludes.
-
-\begin{comment}
-\begin{itemize}
-\item We define the scheduling problem for MPC. 
-We present an analytical model to reason about cost of schedules
-and show that scheduling is NP-hard via a reduction to the shortest 
-common supersequence problem.
-
-\item We present a compiler that takes an IMP-like high-level 
-program and produces amortized (i.e., vectorized) low-level cryptographic 
-code in the MOTION framework. Central contributions are 1) a novel compiler framework, 
-2) a vectorization algorithm that produces optimal schedules for a large 
-number of MPC programs, and 3) reasoning over output arrays without Array SSA; we remove 
-infeasible loop-carried dependences introduced by standard SSA to improve vectorization.
-
-\item We present an implementation and evaluation in the MOTION framework. 
-\ana{Fill in with final results. Mention benchmarks (standard + new ones + HyCC). }
-
-\end{itemize}
-\end{comment}
-
-\section{Overview}
-\label{sec:overview}
-
-\begin{table*}
-\begin{tabular}{ccc}
-\begin{minipage}{0.275\textwidth}
-{\small
-\begin{pythonn}
-def biometric(C: shared[list[int]], D: int, 
-      S: shared[list[int]], N: int) -> 
-      shared[tuple[int,int]]:
-   min_sum : int = MAX_INT
-   min_idx : int = 0  
-   for i in range(N): 
-      sum : int = 0
-      for j in range(D): 
-         # d = S[i,j] - C[j]
-         d : int = S[i * D + j] - C[j] 
-         p : int = d * d
-         sum = sum + p
-      if sum < min_sum:
-         min_sum : int = sum
-         min_idx : int = i
-   return (min_sum, min_idx)
-\end{pythonn}
-}
-\end{minipage}
-
-&
-
-\begin{minipage}{0.33\textwidth}
-{\tiny
-\begin{pythonn}
-min_sum!1 = MAX_INT
-min_idx!1 = 0
-for i in range(0, N):
-   min_sum!2 = PHI(min_sum!1, min_sum!4)
-   min_idx!2 = PHI(min_idx!1, min_idx!4)
-   sum!2 = 0
-   for j in range(0, D):
-      sum!3 = PHI(sum!2, sum!4)
-      d = SUB(S[((i * D) + j)],C[j]) 
-      p = MUL(d,d) 
-      sum!4 = ADD(sum!3,p) 
-   t = CMP(sum!3,min_sum!2) 
-   min_sum!3 = sum!3
-   min_idx!3 = i
-   min_sum!4 = MUX(t, min_sum!3, min_sum!2) 
-   min_idx!4 = MUX(t, min_idx!3, min_idx!2) 
-return (min_sum!2, min_idx!2)   
-\end{pythonn}
-}
-\end{minipage}
-
-&
-
-\begin{minipage}{0.33\textwidth}
-{\small
-\begin{pythonn}
-min_sum!1 = MAX_INT
-min_idx!1 = 0
-# S^ is same as S. C^ replicates C N times: 
-S^ = raise_dim(S, ((i * D) + j), (i:N,j:D)) #S^[i,j] = S[i,j]
-C^ = raise_dim(C, j, (i:N,j:D)) #C^[i,j] = C[j]
-
-sum!2[I] = [0,..,0]
-# computes _all_  "at once"
-d[I,J] = SUB_SIMD(S^[I,J],C^[I,J])
-p[I,J] = MUL_SIMD(d[I,J],d[I,J])
-
-for j in range(0, D):
-   # sum!2[I], sum!3[I], sum!4[I] are size-N vectors 
-   # computes N intermediate sums "at once"
-   sum!3[I] = PHI(sum!2[I], sum!4[I])       
-   sum!4[I] = ADD_SIMD(sum!3[I],p[I,j])
-
-min_idx!3[I] = [0,1,...N-1]   
-for i in range(0, N):
-   min_sum!2 = PHI(min_sum!1, min_sum!4) 
-   t[i] = CMP(sum!3[i],min_sum!2)
-   min_sum!4 = MUX(t[i], sum!3[i], min_sum!2)  
-for i in range(0, N):
-   min_idx!2 = PHI(min_idx!1, min_idx!4)    
-   min_idx!4 = MUX(t[i], min_idx!3[i], min_idx!2)
-return (min_sum!2, min_idx!2)   
-\end{pythonn}
-}
-\end{minipage}
-
-\\
-
-(a) IMP Source & (b) MPC Source & (c) Optimized MPC Source
-\end{tabular}
-\caption{Biometric Matching: From IMP Source to MPC Source to Optimized MPC Source. }
-\label{tab:source_and_MPC_source_and_optimized_MPC_source}
-\end{table*}
-
-\subsection{Source}
-
-As a running example, consider Biometric matching, a standard MPC benchmark.
-An intuitive (and naive) implementation is as shown in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(a).
-Array \python{C} is the feature vector of \texttt{D} features that we wish to match and \python{S} 
-is the database of \python{N} size-\python{D} vectors that we match against.
-
-\begin{comment}
-{\small
-\begin{verbatim}
-def biometric(C: shared[list[int]], D: int, 
-              S: shared[list[int]], N: int) -> 
-                      tuple[shared[int], shared[int]]:
-    min_sum = 10000
-    min_index = 0  
-    for i in range(N): #loop over database
-        sum = 0
-        for j in range(D): #loop over features
-            d = S[i * D + j] - C[j] #i.e., d = S[i,j] - C[j]
-            p = d * d
-            sum = sum + p
-        if sum < min_sum:
-            min_sum = sum
-            min_index = i
-    return (min_sum, min_index)
-\end{verbatim}
-}
-\end{comment}
-
-Our compiler takes essentially standard IMP~\ana{Need citation here.} 
-syntax and imposes certain semantic restrictions. (We detail the restrictions in the following sections.)
-The programmer writes an iterative program and annotate certain inputs 
-and outputs as \emph{shared}. In the example arrays \python{C} and \python{S}
-are shared, meaning that they store shares, however, the array sizes \python{D} and
-\python{N} respectively are plaintext. The code iterates over the entries in the 
-database and computes the Euclidean distance (its square actually) of the current 
-entry \python{S[i]} and \python{C}. The program returns the index of the vector that gives 
-the best match plus the corresponding sum of squares.
-
-\subsection{MPC Source and Cost of Schedule}
-
-Our compiler generates an intermediate representation, MPC Source. MPC Source 
-for Biometric Matching is shown in 
-Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(b). First, MPC Source is an SSA form. 
-Second, it is linear. The conditional in lines 13-14 in IMP Source turns into the linear code in lines 12-16 in MPC Source. 
-The test turns into a CMP operation (\python{t = CMP(sum!3,min\_sum!2)}), followed by the 
-true-branch sequence, followed by the MUX operations. The first MUX operation selects the value
-of \python{min\_sum}: if \python{t} is true, then \python{min\_sum} gets the value of the second multiplexer
-argument,  \python{min\_sum!3}, otherwise it takes the value of the third argument, \python{min\_sum!2}.  
-Third, MPC Source is a special form of SSA. The SSA $\phi$-nodes at the if-then-else (lines 13-15) turn into 
-MUX operations, while the $\phi$-nodes at for loops turn into what we call pseudo PHI nodes with a straightforward semantics.
-
-\begin{comment}
-{\small
-\begin{pythonn}
-1. min_sum!1 = 10000
-2. min_index!1 = 0
-3. for i in range(0, N!0):
-4.    min_sum!2 = PHI(min_sum!1, min_sum!4)
-5.    min_index!2 = PHI(min_index!1, min_index!4)
-6.    sum!2 = 0
-7.    for j in range(0, D!0):
-8.      sum!3 = PHI(sum!2, sum!4)
-9.      d!3 = (S!0[((i * D!0) + j)] - C!0[j]) // MPC
-10.     p!3 = (d!3 * d!3) // MPC
-11.     sum!4 = (sum!3 + p!3) // MPC
-12.  !1!2 = (sum!3 < min_sum!2) // MPC
-13.  min_sum!3 = sum!3
-14.  min_index!3 = i
-15.  min_sum!4 = MUX(!1!2, min_sum!3, min_sum!2) // MPC
-16.  min_index!4 = MUX(!1!2, min_index!3, min_index!2) // MPC
-17.!2!1 = (min_sum!2, min_index!2)   
-\end{pythonn}
-}
-\end{comment}
-
-%The compiler linearizes the source turning conditionals into MUX statements. The PHI nodes
-%are remnants of the SSA IR; the compiler generates code that picks the correct value 
-%when producing MOTION output; the MOTION framework in turn linearizes 
-%loops when it generates the circuit.
-
-We turn to our analytical model to compute the \emph{cost} of the program. Assuming fixed 
-cost $\beta$ for a local MPC operation (essentially just ADD) and cost $\alpha$ for a remote
-MPC operation (e.g., MUX, CMP, and remaining operations), the cost of the iterative schedule 
-will be $ND(2\alpha+\beta) + N(3\alpha$).
-
-A key contribution is the vectorizing transformation. We can compute all $N*D$ 
-subtraction operations (line 9) in a single SIMD instruction; similarly we can compute 
-all multiplication operations (line 10) in a single SIMD instruction. And while we cannot
-vectorize computation of the N individual sums, we can compute the N sums in parallel. 
-Our compiler \emph{automatically detects these opportunities and transforms the program}. 
-It is standard that MPC researchers write vectorized versions of the Biometric
-program by hand; we are the first (to the best of our knowledge) to automatically 
-transform a naive, iterative MPC program into an unintuitive vectorized one.
-
-\subsection{Vectorized MPC Source and Cost of Schedule}
-
-Our compiler produces the vectorized program shown in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(c). 
-Note that this is still our intermediate representation, Optimized MPC Source. Subsequently,
-the compiler turns this code into MOTION variables, loops and SIMD primitives, which MOTION then 
-uses to generate the circuit.
-
-The compiler determines that SUB and MUL in ``naive'' MPC Source (lines 9 and 10 in (b))
-can be fully vectorized into the SIMD SUB and MUL in optimized MPC Source (lines 9 and 10 in (c)). 
-Notation \python{p[I,J]} denotes a 2-dimensional array with fully vectorized dimensions.
-The computation of sum (line 11 in (b))
-is sequential across the $j$-dimension, but it is parallel across the $i$-dimension. 
-The loop in lines 12-16 in (c) illustrates; here \python{p[I,j]} refers to the $j$-th column in \python{p}. 
-Unfortunately, CMP and MUX remain sequential.
-
-\begin{comment}
-{\small
-\begin{verbatim}
-min_sum!1 = 10000
-min_index!1 = 0
-// S!0^ is same as S!0. C!0^ replicates C!0 N-times: 
-S!0^ = raise_dim(S!0, ((i * D!0) + j), (i:N!0,j:D!0))
-C!0^ = raise_dim(C!0, j, (i:N!0,j:D!0))
-
-sum!2 = [0,..,0]
-// Computes all differences and all products "at once"
-d!3[I,J] = SUB_SIMD(S!0^[I,J],C!0^[I,J])
-p!3[I,J] = MUL_SIMD(d!3[I,J],d!3[I,J])
-
-for j in range(0, D!0):
-  // sum!2[I], sum!3[I], sum!4[I] are vectors of size N  
-  // Computes N intermediate sums "at once"
-  sum!3[I] = PHI(sum!2[I], sum!4[I])       
-  sum!4[I] = ADD_SIMD(sum!3[I],p!3[I,j])
-
-min_index!3 = [0,1,...N!0-1]
-
-for i in range(0, N!0):
-  min_sum!2 = PHI(min_sum!1, min_sum!4) 
-  !1!2[i] = CMP(sum!3[i],min_sum!2)
-  min_sum!4 = MUX(!1!2[i], sum!3[i], min_sum!2)
-
-for i in range(0, N!0):
-  min_index!2 = PHI(min_index!1, min_index!4)  
-  min_index!4 = MUX(!1!2[i], min_index!3[i], min_index!2)
-!2!1 = (min_sum!2, min_index!2)   
-\end{verbatim}
-}
-\end{comment}
-
-In MPC backends, executing $n$ operations ``at once'' in a single SIMD operation costs a lot less than executing those $n$ operations one by one.
-This is particularly important when there is communication (i.e., in remote), since many 1-bit values are sent at once rather than sequentially. 
-%In MPC compilers a vectorized operation that computing M operations "at once" costs 
-%essentially the same ($\alpha$ or $\beta$) as an individual operation. 
-We elaborate on the cost model in the following section but for now consider that 
-each operation has a \emph{fixed} portion (does benefits from amortization) and 
-a \emph{variable} portion (does not benefit from amortization): $\alpha = \alpha_\mathit{fix} + \alpha_\mathit{var}$. 
-This gives rise to the following formula for amortized cost: $f(n) = \alpha_\mathit{fix} + n\alpha_\mathit{var}$, 
-as opposed to unamortized cost $g(n) = n\alpha_\mathit{fix} + n\alpha_\mathit{var}$. (We extend the same reasoning to 
-$\beta$-instructions.)
-
-Thus, the fixed cost of the vectorized program amounts to
-$2\alpha_\mathit{fix}$ + $D \beta_\mathit{fix} + N(3\alpha_\mathit{fix})$. (The variable cost is the same in both the vectorized and non-vectorized programs.) 
-The first term in the sum corresponds to the vectorized
-subtraction and multiplication, the second term corresponds to the FOR loop on $j$ and the third 
-one corresponds to the remaining FOR loops on $i$.
-Clearly, $2\alpha_\mathit{fix} + D\beta_\mathit{fix} + N(3\alpha_\mathit{fix}) << ND(2\alpha_\mathit{fix}+\beta_\mathit{fix}) + N3\alpha_\mathit{fix}$.
-Empirically, we observe 10x to 50x improvement over un-vectorized Biometric Matching in circuit generation time and setup time and 5x to 30x improvement in online time.
-Additionally, the un-vectorized version runs out of memory for $N=256$, while the vectorized one runs with the standard maximal input size $N=4,096$.
-\ana{Edit numbers with final experiments.}
-
 \section{Analytical Model}
@@
 \input{prob_stmt_np_hardness.tex}
-
-\begin{figure*}
-  \includegraphics[width=1.0\linewidth]{figs_paper_SIMD/compiler_framework.png}
-  \caption{Compiler Framework. \ana{Change figure since we are not doing div-and-conquer.}}
-  \label{fig:compiler_framework}
-\end{figure*}
-
-\section{\bf Compiler Framework}
-\label{sec:preliminaries}
-
-%We assume arbitrarily nested loops in the MPC-source IR and read-only arrays.
-We assume that loops range from 0 to some constant N.
-Arrays are linearized (row-major order as in MOTION) and accesses are via functions of the induction variables of the enclosing loops.
-
-\figref{fig:compiler_framework} presents an overview of our compiler. This section describes the front-end phases of the compiler
-and Section~\secref{sec:vectorization} describes our vectorization algorithm. \ana{We'll need a section on MOTION code gen too.}
-%\ana{Need to state this more precisely.}
-
-\subsection{Syntax and Semantic Restrictions}
-\label{sec:syntax}
-
-Source syntax is essentially standard IMP syntax as shown below:
-%\begin{figure}[t]
-{\small
-\[
-\begin{array}{l@{~~~~~}l}
-  \begin{array}{l@{~}l@{~~~}l}
- e & ::= e \; \mathit{op} \; e \mid e \mid \x \mid  \texttt{const} \mid \texttt{A[$e$]} & \mathit{expression} \\
- % s &::= t \;\f & \mathit{field} \\
- %\md &::= t\;{\m}(t \;{\this}, \; t\;\x) \; \{\; \ol{t\;\y}\;\s;\;\return\;\y \; \} 
- %                 & \mathit{method} \\ 
-  s & ::= s;s \mid  & \mathit{sequence} \\
-  & \x = e \mid \texttt{A[$e$]} = e \mid & \mathit{assignment \; stmt} \\
-  & \texttt{for $i$ in range($I$)}: s \mid & \mathit{for \; stmt} \\
-  & \texttt{if $e$: $s$ else: $s$} & \mathit{if \; stmt} \\ 
-%   t &::= q \; {\sf C} & \mathit{qualified \; type}\\
-%   q &::= \high \mid \poly \mid \low & \mathit{FlowCFL \; qualifier}\\
-%    t' &::= q' \; \C & \mathit{field \; qualified \; type}\\
-%   q' &::= \tainted \mid \poly & \mathit{field \; qualifier}\\
-
-  \end{array}
-\end{array}
-\]
-}
-%\end{figure}
-
-The syntax allows for array accesses, arbitrarily nested loops, and if-then-else control flow. 
-Expressions are typed $q \; \tau$, where qualifier $q$ and base type $\tau$ are:
-
-{\small
-\[
-\begin{array}{l@{~~~~~}l}
-  \begin{array}{l@{~}l@{~~~}l}
- \tau & ::= \texttt{int}  \mid \texttt{bool} \mid \texttt{list[int]} \mid  \texttt{list[bool]} & \mathit{base \; types} \\
-  q & ::= \texttt{shared} \mid \texttt{plain} & \mathit{qualifiers} \\
-  \end{array}
-\end{array}
-\]
-}
-
-The type system is mostly standard, and in our experience, a sweet spot between readability and expressivity.
-The \texttt{shared} qualifier denotes shared values, i.e., ones shared among the parties and computed upon 
-under secure computation protocols; the \texttt{plain} qualifier denotes plaintext values. Shared lists denote 
-shared elements, but plaintext length. Subtyping is again as expected: $q_1 \; \tau <: q_2 \; \tau$ 
-iff $q_1 <: q_2$ where $\texttt{plain} <: \texttt{shared}$. Thus, we can convert a plaintext value into a shared one, 
-but not vice versa.
-
-Our compiler imposes certain semantic restrictions that it enforces throughout the various 
-phases of compilation. We note that in some cases, the restrictions 
-can be easily lifted and we plan to do so in future iterations of the work.
-
-\begin{enumerate}
-\item Loops are of the form $0 <= i < I$ and bounds are fixed at compile time. 
-It is a standard restriction in MPC that the bounds must be known at circuit-generation time.
-\item Arrays are one-dimensional. N-dimensional arrays are linearized and accessed 
-in row-major order and at this point the programmer is responsible for linearization
-and access. (This restriction can be easily lifted.)
-\item Array subscrpts are plaintext values as specified by the rule:
-\begin{semantics}
-    \ntyperule{Array Access}{
-%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
-  \Gamma \vdash e : \texttt{plain int} \quad \Gamma \vdash \texttt{A} : q \; \texttt{list}[\tau] \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
-}{
-  \Gamma \vdash \texttt{A[$e$]} : q \; \tau
-}
-\end{semantics}
-The subscript $e$ is a function of the indices of the enclosing loops. 
-For read access, the compiler allows an arbitrary such function. 
-However, it restricts write access to \emph{canonical writes}, i.e., \texttt{A[i,j,k]} = ...
-where \texttt{i}, \texttt{j} and \texttt{k} loop over the three dimensions of \texttt{A}.  
-Write access such as for example \texttt{A[i,j+2]} = ... is not allowed. 
-%\item \ana{Add rules for logical ops}
-%\item \ana{Add rules for arithmetic ops}
-\item The final restriction involves MUX as expressed by the rule: 
-\begin{semantics}
-    \ntyperule{MUX}{
-%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
-  \Gamma \vdash e_1 : q_1 \; \texttt{bool} \quad \Gamma \vdash e_2 : q_2 \; \tau \quad \Gamma \vdash e_2 : q_2 \; \tau \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
-}{
-  \Gamma \vdash \texttt{MUX($e_1$,$e_2$,$e_3$)} : q_1 \vee q_2 \vee q_3 \; \tau
-}
-\end{semantics}
-\end{enumerate}
-Importantly, the arguments of MUX are restricted to base types, as passing array arguments 
-is not valid MOTION MPC. \ana{Check this.} This causes a slight inconvenience as we could not write 
-\begin{pythonn}
-if e: A[i] = val
-\end{pythonn}
-Instead we had to write
-\begin{pythonn}
-if not(e): val = A[i]
-A[i] = val
-\end{pythonn}
-
-We note that while the programmer writes 
-annotations at the level of IMP Source (recall \ana{ref to overview section}), the 
-annotations propagate through the transformations and are checked at the 
-level of MPC Source.
-
-%\ana{Disadvantage of MUX}
-
-For the rest of this section we write $i,j,k$ to denote the loop nest: $i$ is the outermost loop, $j$, is immediately nested in $i$, and so on until $k$
-and we use $I,J,K$ to denote the corresponding upper bounds. We write $A[i,j,k]$ to denote canonical access 
-to an array element. In the program, canonical access is achieved via the standard row-major order formula: $(J*K)*i + K*j + k$.
-To simplify the presentation we describe our algorithms in terms of three-element tuples $i,j,k$, however, discussion easily generalizes to 
-arbitrarily large loop nests.
-
-\subsection{From IMP Source to SSA}
-
-Our compiler translates from Source to SSA as follows:
-
-\begin{enumerate}
-  \item
-        \textbf{Parsing}:
-        Use Python's \texttt{ast} module to parse the input source code to a Python AST.
-  \item
-        \textbf{Syntax checking}:
-        Ensure that the AST matches the restricted subset defined in Section~\secref{sec:syntax}.
-        This step outputs an instance of the \texttt{restricted\_ast.Function} class, which represents our restricted subset of the Python AST.
-  \item
-        \textbf{3-address CFG conversion}:
-        %\benjamin{TODO: Is this a good amount of detail?} \ana{This is good amount of detail for the Tech report. In the actual paper we'll shorten a lot and add pointers to the Tech report.}
-
-        Convert the restricted-syntax AST to a three-address control-flow graph.
-        To do this, first,
-        add an empty basic block to the CFG and mark it as current.
-        Next, for each statement in the restricted AST's function body,
-        process the statement.
-        Statements can either be for-loops, if-statements, or assignments (as in~\secref{sec:syntax}).
-        Rules for processing each kind of statement are given below:
-        \begin{enumerate}
-          \item
-                \textbf{For-loops}:
-                Create new basic blocks for
-                the loop condition
-                (the \emph{condition-block}),
-                the loop body
-                (the \emph{body-block}),
-                and the code after the loop
-                (the \emph{after-block}).
-                Insert a jump from the end of the current block to the condition-block.
-                Then, mark the condition-block as the current block.
-                Insert a for-instruction at the end of the current block with the loop counter variable and bounds from the AST.
-                Next, add an edge from the current block to the after-block labeled ``FALSE'' and
-                an edge from the current block to the
-                body-block labeled ``TRUE''.
-                Then, set the body-block to be the current block
-                and process all statements in the AST's loop body.
-                Finally, insert a jump to the condition-block and set the after-block as current.
-          \item
-                \textbf{If-statements}:
-                Create new basic blocks for
-                the ``then'' statements of the if-statement
-                (the \emph{then-block}),
-                the ``else'' statements of the if-statement
-                (the \emph{else-block}),
-                and the code after the if-statement
-                (the \emph{after-block}).
-                At the end of the current block,
-                insert a conditional jump to the then-block or else-block
-                depending on the if-statement condition in the AST.
-                Next,
-                mark the then-block as current,
-                process all then-statements,
-                and add a jump to the after-block.
-                Similarly,
-                mark the else-block as current,
-                process all else-statements,
-                and add a jump to the after-block.
-                Finally, set the after-block to be the current block,
-                and give it a \emph{merge condition} property equal to the condition of the if-statement.
-          \item
-                \textbf{Assignments}:
-                In the restricted-syntax AST,
-                the left-hand side of assignments
-                can be a variable or an array subscript.
-                If it is an array subscript, e.g., \texttt{A[i] = x},
-                change the statement to \texttt{A = Update(A, i, x)}.
-                If the statement is not already three-address code,
-                for each sub-expression in the right-hand side of the assignment,
-                insert an assignment to a temporary variable.
-        \end{enumerate}
-  \item
-        \textbf{SSA conversion}:
-        Convert the 3-address CFG to SSA with Cytron's algorithm.
-\end{enumerate}
-
-\subsection{From SSA to MPC Source} %: MUX Nodes and Pseudo $\phi$-nodes}
-
-Once the compiler converts the code to SSA,
-it transforms $\phi$-nodes that correspond to if-statements into MUX nodes.
-From the 3-address CFG conversion step,
-$\phi$-nodes corresponding to if-statements will be in a basic block
-with the merge condition property.
-For example, if $X_3 = \phi(X_1,X_2)$ is in a block with merge condition $C$,
-the compiler transforms it into $X_3 = \mathit{MUX}(C, X_1, X_2)$.
-Next, the compiler runs the dead code elimination algorithm from Cytron's SSA paper.
-
-Next, the control-flow graph is \emph{linearized} into MPC Source,
-which has loops but no if-then-else-statements.
-This means that both branches of all if-statements are executed,
-and the MUX nodes determine whether to use results from the then-block or from the else-block.
-The compiler linearizes the control-flow graph with a variation of breadth-first search.
-Blocks with the ``merge condition'' property are only considered
-the second time they are visited,
-since that will be after both branches of the if-statement are visited. 
-(The Python AST naturally gives rise to a translation where each conditional has exactly two targets, 
-and each ``merge condition'' block has exactly two incoming edges, a TRUE and a FALSE edge.
-Thus, each $\phi$-node has exactly two multiplexer arguments, which dovetails into MUX.
-This is in contrast with Cytron's algorithm which operates at the level of the CFG and allows for 
-$\phi$-nodes with multiple arguments.)
-Each time the compiler visits a block,
-it adds the block's statements to the MPC source.
-If the block ends in a for-instruction,
-the compiler recursively converts the body and code after the loop to MPC source
-and adds the for-loop and code after the loop to the main MPC source.
-If the block does not end in a for-instruction,
-the compiler recursively converts all successor branches to MPC source and
-appends these to the main MPC source.
-
-Now, the remaining $\phi$-nodes in MPC source are the loop header nodes. We call these nodes \emph{pseudo} $\phi$-nodes 
-and we write \python{PHI} in MPC Source. A pseudo $\phi$-node $X_1 = $\python{PHI}$(X_0,X_2)$ in a loop header is evaluated 
-during circuit generation. If it is the 0-th iteration, then the $\phi$-node evaluates to $X_0$, otherwise, it evaluates to $X_2$.
-
-%\ana{We need to double check that the following captures all cases of def-use edges.}
-
-\subsection{Dependence Analysis}
-
-The next steps construct the dependence graph and \ana{Add an intro here.}
-
-\subsubsection{Def-use Edges}
-
-The dependence graph has the following def-use edges:
-
-\begin{itemize}
-\item same-level edge $X\rightarrow Y$ where $X$ and $Y$ are in the same loop nest, say $i,j,k$.
-E.g., the def-use edge 9 to 10 in the Biometric MPC Source in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source} is a same-level edge.
-A same-level edge can be a back-edge in which case a $\phi$ node is the target of the edge.
-E.g., 15 to 4 in Biometric is a same-level back-edge.
-\item outer-to-inner $X\rightarrow Y$ where $X$ is in an outer loop nest, say $i$, and $Y$ is in an inner one, say $i,j,k$. E.g., 1 to 4 in Biometric forms is an outer-to-inner edge.
-%A $\phi$-node can be a source or a target of an outer-to-inner forward edge.
-\item inner-to-outer $X\rightarrow Y$ where $X$ is a $phi$-node in an inner loop nest, $i,j,k$, and $Y$ is in the enclosing loop nest $i,j$.
-E.g., the def-use from 8 to 12 gives rise to an inner-to-outer edge.
-%E.g. $\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$ is an inner-to-outer edge. 
-An inner-to-outer edge can be a back-edge as well, in which case both $X$ and $Y$ are phi-nodes with the source $X$ in a loop nested into $Y$'s loop (not necessarily immediately).
-
-%Note that the source is \emph{always} a $\phi$-node of the $i,j,k$ loop and the target is in the immediately enclosing loop.
-The interpretation of this edge is that the use node $Y$ uses the definition made in the last iteration of the inner loop.
-%\ben{The current representation of pseudo $\phi$-nodes shows them attached to the loop header (i.e. in the loop).  We may want to clarify that they are also evaluated at the loop's termination.}
-%\item same-level back-edge $X\rightarrow Y$.
-$Y$ is a phi-node in the header of the loop and $X$ is a definition of the variable in the loop body.
-E.g., $\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{sum}_1,\texttt{min}_1)$ to $\texttt{min}_0 = \phi(\texttt{min}_1, 10000)$ in Biometric is a same-level back-edge.
-%\item inner-to-outer back-edge $X\rightarrow Y$: $X$ and $Y$ are both $\phi$-nodes for some variable. The source $X$ is in a loop nested into $Y$'s loop (not necessarily immediately).
-
-\item mixed forward edge $X\rightarrow Y$.
-$X$ is in some loop $i,j,k$ and $Y$ is in a loop nested into $i,j,k'$.
-We transform mixed forward edges as follows.
-Let $\x$ be the variable defined at $X$.
-We add a variable and assignment $\x' = \x$ immediately after the $i,j,k$ loop.
-Then we replace the use of $\x$ at $Y$ with $\x'$.
-This transforms a mixed forward edge into an "inner-to-outer" forward edge followed by an outer-to-inner forward edge.
-Thus, Basic Vectorization handles one of "same-level", "inner-to-outer", or "outer-to-inner" def-use edges.
- \end{itemize}
-
-\subsubsection{Closures}
-
-We define $\mathit{closure}(n)$ where $n$ is a phi-node. Intuitively, it computes the set of nodes (i.e., statements) that form a dependence cycle with $n$. The closure of $n$ is defined as follows:
-\begin{itemize}
-\item $n$ is in $\mathit{closure}(n)$
-\item $X$ is in $\mathit{closure}(n)$ if there is a same-level path from $n$ to $X$, and $X \rightarrow n$ is a same-level back-edge.
-\item $Y$ is in $\mathit{closure}(n)$ if there is a same-level path from $n$ to $Y$ and there is a same-level path from $Y$ to some $X$ in $\mathit{closure}(n)$.
-\end{itemize}
-
-\subsection{Taint Analysis}
-
-We require that all inputs are marked as either shared or plaintext, however, we do not require any other qualifier annotations in the program.
-Then we infer if intermediate variables and expressions are shared through taint analysis with "taintedness" referring to the shared attribute.
-Specifically, the compiler follows the following rules, which are standard in positive-negative qualifier systems (here \texttt{shared} is the positive qualifier and \texttt{plain} is the negative one):
-\begin{enumerate}
-\item Loop counters and array upper bounds are always \texttt{plain}.
-\item If any variable on the right-hand side of an assignment is shared, then the assigned variable is \texttt{shared}.
-%\item If all variables on the right-hand side of an assignment are plaintext, then the assigned variable is plaintext \ana{Will be covered by the last rule.}
-\item Any variables that cannot be determined as shared via the above rules are \texttt{plain}.
-\end{enumerate}
-
-%The first two rules are standard for taint analysis, and the third rule follows from the MPC problem statement. \ben{Is the explanation for the third rule correct?}  \ana{Yes, that is exactly it!}
-The final rule is needed in examples like the following. In the below snippet \texttt{sum!2} and \texttt{sum!3} form a dependency cycle and there is no \texttt{shared} value that flows to either one. 
-They are inferred as plaintext. 
-%cannot be marked as plaintext through simple taint analysis:
-{\small
-\begin{verbatim}
-plaintext_array = [0, 1, 2, ...]
-sum!1 = 0
-for i in range(0, N):
-    sum!2 = PHI(sum!1, sum!3)
-    sum!3 = sum!2 + plaintext_array[i]
-\end{verbatim}
-}
-%\ben{I think the above example is unnecessary and could be replaced by an explanation of how "untainted" variables are implicitly plaintext, but I couldn't think of a way to phrase that.}\ana{I left the example, we can leave it in the TR, remove from paper.}
-
-When converting to MOTION code, any plaintext value used in the right-hand side of a shared assignment is converted to a shared value for that expression. 
-%\ben{Is this necessary to include?} \ana{Yes, that's important.}
-
-Importantly, the taint analysis works on MPC Source, which lacks if-then-else control flow. 
-This significantly simplifies the treatment as there is no need to handle conditionals and implicit flow.
-
-\subsection{From (Optimized) MPC Source to MOTION}
-\label{sec:MOTION}
-\ana{Ben, add the MOTION section here.}
-
-\section{Vectorization}
-\label{sec:vectorization}
-
-An important component of our algorithm is the scalar expansion to the corresponding loop dimensionality. 
-For example, \texttt{d = S[i * D + j] - C[j]} equiv. to \texttt{d = S[i,j] - C[j]}, which gave rise to \texttt{N*D} subtraction operations in the sequential schedule, 
-is lifted. The argument arrays \texttt{S} and \texttt{C} are lifted and the scaler \texttt{d} is lifted: \texttt{d[i,j] = S[i,j] - C[i,j]}.
-The algorithm then detects that the statement can be vectorized.
-
-There are three kinds of arrays (for now all kept internally as one-dimensional arrays, but that's under discussion).
-\begin{itemize}
-
-\item Scalars: These are scalar variables we lift into arrays for the purposes of vectorization. 
-For those, all writes are canonical writes and all reads are canonical reads. We will apply raise dimension when 
-a scalar is used in an inner loop (e.g., \texttt{sum0} in line 6 of the MPC source code will be raised to a 1-dimensional 
-array since \texttt{sum0} is used in the inner $j$-loop). Drop dimension applies as well; this happens when a scalar
-written in an inner loop is used outside of the loop (e.g., \texttt{sum0} for which the lifted inner loop computes 
-\texttt{D} values, but the outer loop only needs the last one.)
-
-\item Read-only input arrays: Read-only inputs. There are NO writes, while we may have non-canonical reads, $f(i,j,k)$. 
-Phase 1 of Basic vectorization will add raise dimension operation at the beginning of the function to lift these arrays, 
-and raise dimension may \emph{reshape} arrays. If there are multiple "views" of the input array, there would be multiple raise dimension statements to create
-each one of these views. The invariant is that at reads in loops, the reads of "views" of the original input array are canonical.
-Only raise dimension applies to these arrays, and only in the beginning of the program. For example, 1-dimensional array 
-\texttt{C} is lifted into 2-dimensional array \texttt{N x D} by copying the row \texttt{N} times.
-
- \item Read-write output arrays: Writes are canonical (by restriction) but reads can be non-canonical. 
- Dependence analysis limits vectorization when non-canonical read access refers to array writes in
- previous iterations, thus creating loop-carried dependences. We may apply both raise and drop dimension, however, 
- they respect the fixed dimensionality of the output array. The array cannot be raised to a dimension lower than
- its canonical (fixed) dimensionality and it cannot be dropped lower. In addition, non-canonical reads may 
- require lifting (i.e., reshaping) of the array after the most recent write, rather than in the beginning of the program, 
- in order to reduce a non-canonical read to a canonical one.
-
-\end{itemize}
-
-In the sections below we detail the $\mathit{raise\_dim}$ (raise dimension) and $\mathit{drop\_dim}$ (drop dimension) operations,
-followed by our vectorization algorithm.
-
-%\ana{Add back-edges into Phase 1.
-A back-edge from a non-phi-node in loop $i$ to a phi-node in loop $i$'s header is a same-level edge.
-The only difference with the handling of normal same-level forward def-use is that the operand index will become $i-1$.
-A back edge from a phi-node in loop $j$ to a phi-node in loop $i$, where $j$ is nested in $i$ is an inner-to-outer edge and will require dropping dimension.
-We can show these are the only kinds of back-edges that may occur.}
-
-%\subsection{Algorithm (Still work in progress.)}
-
-\subsection{Raise Dimension and Drop Dimension}
-
-There are two conceptual versions of $\mathit{raise\_dim}$. One applies on read-only input arrays and reshapes those arrays 
-when necessary to ensure canonical read access in the corresponding loop. The signature of $\mathit{raise\_dim}$
-is as follows. It takes the original array \texttt{C}, the access pattern function $f(i,j,k)$ in loop nest $i,j,k$ and the loop bounds
-\texttt{((i:I),(j:J),(k:K))}: 
-\[ \mathit{raise\_dim}(\texttt{A}, f(i,j,k), \texttt{((i:I),(j:J),(k:K))}) \]
-It produces a new 3-dimensional array \texttt{A'} by iterating over $i,j,k$ and setting each element of  \texttt{A'} as follows:
-\[ \texttt{A'[i,j,k] =  A[f(i,j,k)]} \]
-The end result is that uses of \texttt{[A[f(i,jk)]} in loop nest $i,j,k$ are replaced with canonical read-accesses to \texttt{A'[i,j,k]}
-that can be vectorized. In the running Biometric example, \texttt{C'} = $\mathit{raise\_dim}$\texttt{(C, j, (i:N,j:D))} lifts the 
-1-dimensional array \texttt{C} into a 2-dimensional array. The $i,j$ loop now accesses \texttt{C'} in the canonical way, \texttt{C'[i,j]}. 
-Similarly, \texttt{S'} = $\mathit{raise\_dim}$\texttt{(S, i*D+j, (i:N,j:D))} tries to lift \texttt{S}, but the operation turns into a no-op 
-because \texttt{S} is already a 2-dimensional array and the read access is canonical.
-
-The other version of $\mathit{raise\_dim}$ applies on scalars and read-write arrays. It lifts a lower-dimension array into a 
-higher-dimension for access in a nested loop. Here \texttt{A} is an $i$ array and raise dimension adds two additional dimensions:
-\[ \mathit{raise\_dim}(\texttt{A}, \texttt{(j:J,k:K)}) \]
-This version is reduced to the above version by adding the access pattern function, which is just $i$: 
-\[ \mathit{raise\_dim}(\texttt{A}, i, \texttt{(j:J,k:K)}) \]
-
-The corresponding $\mathit{drop\_dim}$ is carried out when an array written in an inner loop is used in an enclosing loop. 
-It takes a higher dimensional array, say $i,j,k$ and removes trailing dimensions, say $j,k$:
-\[ \mathit{drop\_dim}(\texttt{A}, \texttt{(j:J,k:K)}) \]
-It iterates over $i$ and takes the result at the maximal index of $j$ and $k$, i.e., the result at the last iterations of $j$ and $k$:
-\[ \texttt{A'[i] = A[i,J-1,K-1]} \]
-
-\begin{comment}
-We define the $\mathit{raise\_dim}$ (raise dimensions) and $\mathit{drop\_dim}$ (drop dimension) functions.
-Raise dimension "lifts" a lower-dimension array (\ana{The right term here is tensor, I am pretty sure!}) into a higher dimension one.
-This is necessary when a lower-dimensional array is used in a higher dimensional loop and, essentially, is just copying of values.
-For example, Biometric contains the statement $d = \mathit{SUB}(S[i, j], C[j])$, in the $j$-loop which is nested into the $i$-loop.
-Here $C$ is a one-dimensional array, however, to vectorize across both loops it is necessary to turn it into a two-dimensional array: $C[i,j]$ becomes $[C[0],C[1],...C[J],C[0],C[1],...C[J], ...
-C[0],C[1],...C[J]]$, which turns the row into a matrix of $I$ identical rows.
-(We use capital letters to denote the upper bounds of loops, e.g., $J$ is the upper bound of the $j$-loop and $I$ is the upper bound of the $i$-loop.)
-
-$\mathit{raise\_dim}(A[i_1,...,i_k],i_j)$ is defined as follows.
-It results in a new {k+1}-dimensional array $A'$ where for every $0 \le i_{k} < I_{k+1}$, $A'[i_1, \ldots, i_j, \ldots, i_k] = A[i_1, \ldots, i_k]$.
-Adding $n$ dimensions is trivially extended as a composition of $n$ $\mathit{raise\_dim}$ that each adds a single dimension.
-
-As expected, drop dimension turns a higher-dimensional array into a lower-dimensional one.
-The key use case is an inner-to-outer def-use edge.
-The code may define a variable, e.g., $\x$ in an inner loop, say $j$, then use this variable in an enclosing loop, say $i$.
-Our algorithm may vectorize the computation of $\x$ in the $j$-loop thus producing a vector $\x[j]$ where $\x[1]$ is the value of \x after the 1st iteration and so on.
-Drop dimension states that the outer loop will use the value of the variable at the last iteration.
-
-$\mathit{drop\_dim}(A[i_1,\ldots, i_j, \ldots,i_k], i_j)$ produces a k-dimensional array $A'$ where $A'[i_1,\ldots,i_k] = A'[i_1,\ldots, I_j-1, \ldots,i_k]$.
-In our analysis, dimensions are always dropped at the end, and again, one can define dropping $n$ dimensions as a composition of $n$ $\mathit{drop\_dim}$.
-
-\ben{I'm not sure if I wrote up the above paragraphs correctly -- the idea is that we're just copying values to extend over dimension $j$ or only retain its last elements.
-The current writeup implies that $1 < j < k$ though, which isn't always true.
-Python implementations of $\mathit{raise\_dim}()$ and $\mathit{drop\_dim}()$ can be found here: \url{https://github.com/milana2/ParallelizationForMPC/blob/master/compiler/compiler/motion_backend/reference_implementations.py}}
-
-\ana{This section is still informal. Needs work to make more precise.}
-\end{comment}
-
-\subsection{Basic Vectorization}
-
-\begin{algorithmic}
-
-\STATE \COMMENT { Phase 1: Raise dimension of scalar variables to corresponding loop nest. We can traverse stmts linearly in MPC-source. }
-
-\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i,j,k$}
-%\FOR {each $Y_n$}
-\FOR {each argument $Y_n$} %def-use edge $\mathit{stmt}' (\mbox{def of } Y) \rightarrow \mathit{stmt} (\mbox{def of } X)$}
-\STATE {case def-use edge $\mathit{stmt}' (\mbox{def of } Y_n) \rightarrow \mathit{stmt} (\mbox{def of } X)$ of} \\
-\STATE \hspace{0.25cm} {same-level: $Y'_n$ is $Y_n$} \\
-\STATE \hspace{0.25cm} {outer-to-inner: add $Y'_n[i,j,k] = \mathit{raise\_dim}(Y_n)$ at $\mathit{stmt'}$}
-\STATE \hspace{0.25cm} {(more precisely, right after $\mathit{stmt'}$)}
-\STATE \hspace{0.25cm} {inner-to-outer: add $Y'_n[i,j,k] = \mathit{drop\_dim}(Y_n)$ at $\mathit{stmt}$}
-\STATE \hspace{0.25cm} {(more precisely, in loop of $\mathit{stmt}$ right after loop of $\mathit{stmt'}$)}
-\ENDFOR
-
-\COMMENT { Optimistically vectorize all. $I$ means vectorized dimension. }
-%\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i,j,k$}
-\STATE {change to $X[I,J,K] = \mathit{Op}(Y'_1[I,J,K],Y'_2[I,J,K])$ }
-\ENDFOR
-
-\COMMENT { Phase 2: Recreating FOR loops for cycles; vectorizable statements hoisted up. }
-
-\FOR {each dimension $d$ from highest to 0}
-\FOR {each $\phi$-node $n$ in loop $i_1,...,i_d$}
-\STATE {compute $\mathit{closure}(n)$}
-\ENDFOR
-
-\COMMENT {{\color{red} $cl_1$ and $cl_2$ intersect if they have common statement or update same array; "intersect" definition can be expanded}}
-
-\WHILE {there are closure $cl_1$ and $cl_2$ that intersect}
-\STATE {merge $cl_1$ and $cl_2$}
-\ENDWHILE
-\FOR {each closure $cl$ (after merge)}
-\STATE {create FOR $i_d = 0; ...$ loop}
-\STATE {add $\phi$-nodes in $cl$ to header block}
-\STATE {{\color{red}{add target-less $\phi$-node for A if $cl$ updates array A}}}
-\STATE {add statements in $cl$ to loop body in some order of dependences}
-\STATE \COMMENT { Dimension is not vectorizable: }
-\STATE {change ${I_d}$ to $i_d$ in all statements in loop}
-\STATE {treat FOR loop as monolith node: some def-use edges become same-level.}
-\ENDFOR
-{
-\color{red}
-\FOR {each target-less $\phi$-node $A_1 = \phi(A_0,A_k)$}
-\STATE {in vectorizable stmts, replace use of $A_1$ with $A_0$}
-\STATE {discard $\phi$-node if not used in any $cl$, replacing $A_1$ with $A_0$ or $A_k$ appropriately}
-\ENDFOR
-}
-\ENDFOR
-
-\COMMENT { Phase 3: Remove unnecessary dimensionality. }
-
-\COMMENT { A dimension $i$ is dead on exit from stmt $X[...i...] = ...$ if all def-uses with targets outside of the enclosing 
-FOR $i=0 ...$ MOTION loop end at target (use) $X' = \mathit{drop\_dim}(X, i)$. } 
-\FOR {each stmt and dimension $X[...i...] = ...$}
-\STATE {if $i$ is a dead dimension on exit from stmt $X[...i...] = ...$, remove $i$ from $X$ (all defs and uses)} 
-%\COMMENT{ A stmt v[...i] is a "dead dimension i" if all edges out of FOR i = 0; ... loop end at drop\_dim(v, i)}
-\ENDFOR
-
-\COMMENT { Now clean up drop\_dim and raise\_dim }
-\FOR {each $X' =  \mathit{drop\_dim(X,i)}$ }
-\STATE {replace with $X' = X$ if $i$ is dead in $X$.}
-\ENDFOR
-\STATE {do (1) (extended) constant propagation, (2) copy propagation and (3) dead code elimination to get rid of redundant variables and raise and drop dimension statements}
-
-\COMMENT { Phase 4: }
-\STATE {add SIMD for simdified dimensions}
-
-\end{algorithmic}
-
-\input{biometric_vectorization_example.tex}
-
-\begin{comment}
-\subsection{Example: Biometric}
-
-\subsubsection{MPC-source}
-
-Below is the MPC-source resulting from Benjamin's analysis. It gives rise to a corresponding dependence graph. (Not shown.)
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0 = \phi(-1, \texttt{index}_1)$}
-\STATE {$\texttt{min}_0 = \phi(10000,\texttt{min}_1)$}
-
-\STATE \COMMENT {Begin of inner loop $j$.}
-\STATE {$\texttt{sum}_0 = \phi(0, \texttt{sum}_1)$}
-\STATE {$\texttt{d} = \mathit{SUB}(\texttt{S}[i,j], \texttt{C}[j])$}
-\STATE {$\texttt{p} = \mathit{MUL}(\texttt{d},\texttt{d})$}
-\STATE {$\texttt{sum}_1 = \texttt{ADD}(\texttt{sum}_0,\texttt{p})$}
-\STATE \COMMENT {End of inner loop $j$.}
-\STATE {$\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$}
-\STATE {$\texttt{index}_1 = \mathit{MUX}(\texttt{c},\texttt{index}_0,i)$}
-\STATE {$\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{min}_0,\texttt{sum}_0)$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-\subsubsection{Phase 1 of Basic Vectorization:}
-
-The transformation preserves the dependence edges.
-It raises the dimensions of scalars and optimistically vectorizes all operations.
-The next phase discovers loop-carried dependences and removes affected vectorization.
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$}
-\STATE {$\texttt{min}_0[i] = \phi(10000, \texttt{min}_1[i-1])$}
-
-\STATE \COMMENT {Begin of inner loop $j$.}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...], \texttt{sum}_1[i,j-1])$}
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{sum}_1[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE \COMMENT {End of inner loop $j$.}
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {$\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$}
-\STATE {$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$}
-\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-\ishaq{At least for input variables: when raising dimensions, we should keep some meta data around so we can later write code that could use 1 share if it knows the other N shares are just duplicates of it.}
-
-\subsubsection{Phase 2}
-
-This phase analyzes statements from the innermost loop to the outermost. The key point is to discover loop-carried dependencies and re-introduce loops whenever dependencies make this necessary.
-
-Starting at the inner phi-node $\texttt{sum}_0[i,j] = \phi([0,0,...],\texttt{sum}_1[i,j-1])$, the algorithm first computes its closure.
-The closure amounts to the phi-node itself and $\texttt{sum}_1[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$, accounting for the loop-carried dependency of the computation of \texttt{sum}.
-The algorithm replaces this closure with a FOR loop on $j$ removing vectoriaztion on $j$.
-Note that the SUB and MUL computations remain outside of the loop as they do not depend on phi-nodes that are part of cycles.
-The algorithm adds same-level edges, one from $\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = ...$ to the monolithic FOR-loop node, and one from the FOR-loop node to $\texttt{c}[\stackrel{\rightarrow}{i}] = ...$.
-
-\ishaq{For our radar: Such dependencies will not involve phi nodes only. e.g. $a[i] = a[i] + a[i-1]$}
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0[i] = \phi(-1, \texttt{index}_1[i-1])$}
-\STATE {$\texttt{min}_0[i] = \phi(10000, \texttt{min}_1[i-1])$}
-
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-
-\STATE \COMMENT {Begin of inner loop $j$. Loop is now much "shorter"!}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...], \texttt{sum}_1[i,j-1])$}
-\STATE {FOR j=0; j<D; j++} \COMMENT{ Will turn into a MOTION loop.}
-\STATE \hspace{0.25cm}{$\texttt{sum}_1[\stackrel{\rightarrow}{i},{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},{j}],\texttt{p}[\stackrel{\rightarrow}{i},{j}])$}
-\STATE \COMMENT {End of inner loop $j$.}
-
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {$\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$}
-\STATE {$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$}
-\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-Next, analysis moves to outer dimension $i$. There are two phi-nodes, $\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1])$ and $\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$.
-The closure of the first is
-\[
-\begin{array}{l}
-\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1]) \\
-\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
-\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])
-\end{array}
-\]
-and the closure of the second one is
-\[
-\begin{array}{l}
-\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1]), \\
-%\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
-\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])
-\end{array}
-\]
-
-Since the two closures \emph{do not} intersect, we have two distinct FOR-loops on $i$.
-The first FOR loop (on \texttt{min}) is scheduled first, because of the dependence edge from $\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$ to
-$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$.
-We can now rewrite those statements, getting rid of the vectorization on $i$ in the two FOR loops.
-Notably, the SUB and MUL computations are fully vectorizable and the ADD computation is vectorizable across the $i$-dimension.
-
-\begin{algorithmic}
-
-\STATE
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB\_SIMD}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL\_SIMD}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$} \\
-\STATE {}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...],\texttt{sum}_1[i,j-1])$}
-\STATE {FOR j=0; j<D; j++} \COMMENT{ Turns into a MOTION loop, was $j$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{sum}_1[\stackrel{\rightarrow}{i},{j}] = \mathit{ADD\_SIMD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},{j}],\texttt{p}[\stackrel{\rightarrow}{i},{j}])$}
-\STATE {}
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {}
-\STATE {$\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1])$}
-\STATE {FOR i=0; i<N; i++} \COMMENT{ Turns into a MOTION loop, was $i$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{c}[{i}] = \mathit{CMP}(\texttt{sum}'_0[{i}],\texttt{min}_0[{i}])$}
-\STATE \hspace{0.25cm}{$\texttt{min}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{min}_0[{i}],\texttt{sum}'_1[{i}])$}
-\STATE {}
-\STATE {$\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$}
-\STATE {FOR i=0; i<N; i++} \COMMENT{ Turns into a MOTION loop, was $i$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{index}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{index}_0[{i}],i)$}
-
-\ben{The above code implicitly stores the final values of min\_0 and index\_0 in their last indices, but there is no primitive like drop\_dim to make it clear.
-This will have to be added in the SSA code generation (though it might not be necessary to include in the paper?).
-It also is pretty wishy-washy about the size and layout of each vector (again more of an implementation issue), but maybe we should have a different syntax for array accesses for read/write (e.g.
-the ADD\_SIMD steps) vs the array definitions (e.g.
-the SUB\_SIMD and raise/drop\_dim statements)}
-
-\end{algorithmic}
-\end{comment}
-
-\subsection{Correctness Argument}
-
-We build a correctness argument that loosely follows the theory of Abstract Interpretation. 
-We define the syntax of MPC Source programs. The domain of MPC Source programs
-expressible in the syntax (with certain semantic restrictions) is the abstract domain $A$. 
-We then define the \emph{linearization} of an MPC Source program as an interpretation 
-over the syntax. The linearization, which is a \emph{schedule}, is the concrete domain $C$. 
-Since we reason over def-use graphs in $A$ we define a partial order relation over elements of $A$ 
-in terms of def-use relations. We define a partial order over elements of $C$ as well,
-in terms of def-use relations in the concrete domain $C$. We prove two theorems
-that state (informally) that the schedule corresponding to the original program computes the same 
-result as the schedule corresponding to the vectorized program.
-
-\begin{figure}%[!b]
-%\small
-\[
-\begin{array}{l@{~~~~~}l}
-  \begin{array}{l@{~}l@{~~~}l}
-  s &::= s;s & sequence \\
-     & \mid \x[i,J,k] = \y[i,J,k] \; \mathit{op\_SIMD} \; \z[i,J,k] & operation \\ 
-     & \mid \x[i,J,k]  = \texttt{PHI}(\x_1[i,J,k],\x_2[i,J,k\!-\!1]) & \mathit{phi} \; \mathit{node} \\
-     & \mid \x[i,J,k] = \mathit{raise\_dim}(\x'[i],(J\!:\!J,k\!:\!K)) & \mathit{raise} \; \mathit{dimension(s)} \\
-     & \mid \x[i,J] = \mathit{drop\_dim}(\x'[i,J,k],k)  & \mathit{drop} \; \mathit{dimension(s)} \\
-     & \mid \x = \y & \mathit{propagation} \\
-     & \mid \texttt{FOR} \; 0 \le i < I: s  & \mathit{loop} \\
-  \end{array}    \\
-\end{array}
-\]
-\vspace{-3mm}
-\caption{MPC Source Syntax}\label{fig:MPC_Source_syntax}
-\vspace{-5mm}
-\end{figure}
-
-\paragraph{MPC Source Syntax}
-
-\figref{fig:MPC_Source_syntax} defines the syntax for our intermediate representation, MPC Source. 
-There are semantics restrictions over the syntax as well: a variable $\x[i,j,k]$ is a 3-dimensional array 
-$(i:I,j:J,k:K)$ and also, a statement $\x[i,J,k] = ...$ is enclosed in loops over $i$ and $k$ as shown below. 
-Thus, $i$ and $k$ are in scope.
-\begin{verbatim}
-FOR 0 <= i < I:
-    ...
-    FOR 0 <= k < K:
-        x[i,J,k] = ...
-\end{verbatim}
-
-Statements $\mathit{operation, phi, raise \; dimension(s), drop \; dimension(s)}$ are \emph{base} statements, 
-and $\mathit{sequence, loop}$ are \emph{compound statements}.
-
-\paragraph{Linearization}
-
-Linearization is the concretization operation, which, as we mentioned earlier computes a schedule.
-The concretization function $\gamma: A \rightarrow C$ is defined as an interpretation of
-MPC Source syntax, as it is standard. The concretization of each one 
-of the base statements is as follows: 
-\[\begin{array}{l}
-\gamma(\x[i,J,k] = \mathit{op\_SIMD}(\y[i,J,k], \z[i,J,k])) = \\ 
-~~\x[i,0,k] = \y[i,0,k] \; \mathit{op} \; \z[i,0,k] \;\; || \\
-~~~~\x[i,1,k] = \y[i,1,k] \; \mathit{op} \; \z[i,1,k] \;\; || \; ...\; || \\
-~~~~~~\x[i,I-1,k] = \y[i,I-1,k] \; \mathit{op} \; \z[i,I-1,k]
-\end{array}\]
-meaning that the vectorized dimension(s) are expanded into parallel statements. || introduces SIMD (parallel) execution.
-
-The concretization of the FOR statement is as follows:
-\[ \gamma(\texttt{FOR} \; 0 \le i < I: s) = \gamma(s)[0/i] \; ; \; \gamma(s)[1/i] \; ; \; ... \gamma(s)[I-1/i] \]
-$\gamma$ simply unrolls the loop substituting $i$ with 0, 1, etc. Here ; denotes sequential execution.
-
-\paragraph{Partial Orders}
-
-For each MPC Source program $a$ we compute the def-use edges in the standard way: if base statement $s1 \in a$ defines variable \x, e.g., $\x[i,j,k] = ...$, 
-and base statement $s2 \in a$ uses \x, e.g., $... = ... \x[i,j,k]$ and there is a path in the trivial CFG from $s1$ to $s2$, then there is a def-use edge from
- $s1$ to $s2$. We extend the dimensionality of a statement into $s1[i,j,k]$ where $s1[i,j,k]$ inherits the dimensionality of the left-hand-side of the assignment.
-
- Let $a_0, a_1$ be two MPC Source programs in $A$. 
- Two base statements, $s_0 \in a_0$ and $s_1 \in a_1$ are \emph{same}, written $s_0 \equiv s_1$ if they are of the same operation and they operate on 
- the same variables: same variable name and same dimensionality. Recall that dimensions in MPC Source are either iterative, lower case, or vectorized, upper case. 
- Two statements are same even if one operates on an iterative dimension and the other one operates on a vectorized one, e.g., $s_0[i,j,k] \equiv s_1[I,j,K]$.
-
- \begin{definition} Let $a_0, a_1 \in A$. We say that $a_0 \le a_1$ iff for every def-use edge $s1 \rightarrow s2$ in $a_0$ there is an edge
- $s1' \rightarrow s2'$ where $s1 \equiv s1'$, $s2 \equiv s2'$ and the two edges of either both forward or both backward.
- \end{definition}
-
- The def-use edges in the concrete schedule are as expected. There is a def use edge from statement $s1$ that defines $\x[\underline{i},\underline{j},\underline{k}]$
- to statement $s2$ that uses $\x[\underline{i},\underline{j},\underline{k}]$ if $s1$ is scheduled ahead of $s2$ in the linear schedule. We note that the underlined
- indices, e.g., $\underline{i}$, refer to fixed values, not iterative or vectorized dimensions since in the concrete schedule all induction variables are expanded. 
- E.g., there is a def-use edge from the statement that defines $\x[0,1,2]$ and a statement that uses $\x[0,1,2]$.
-
-\paragraph{Theorems}
-
-\begin{theorem} $a_0 \le a_1 \Rightarrow \gamma(a_0) \subseteq \gamma(a_1)$.
-\end{theorem}
-
-\begin{theorem} Let $a_0$ be the iterative MPC Source and let $a_1$ be the vectorized MPC Source computed by the vectorization algorithm. 
-We have that $a_0 \le a_1$. 
-\end{theorem}
-
-\ana{Write the proof sketch and final argument, etc.}
-
-\subsection{Extension of Basic Vectorization with Array Writes}
-
-\subsubsection{Removal of Infeasible Edges}
-
-Array writes limit vectorization as they sometimes introduce infeasible loop-carried dependencies. Consider the following example: \ana{Have to add citation to Aiken's paper}
-
-{\small
-\begin{verbatim}
-for i in range(N):
-  A[i] = B[i] + 10;
-  B[i] = A[i] * D[i-1];
-  C[i] = A[i] * D[i-1];
-  D[i] = B[i] * C[i];
-\end{verbatim}
-}
-
-In Cytron's SSA this code (roughly) translates into
-
-{\small
-\begin{verbatim}
-for i in range(N):
-1. A_1 = PHI(A_0,A_2)
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. A_2 = update(A_1, i, B_1[i] + 10); 
-6. B_2 = update(B_1, i, A_2[i] * D_1[i-1]);
-7. C_2 = update(C_1, i, A_2[i] * D_1[i-1]);
-8. D_2 = update(D_1, i, B_2[i] * C_2[i]);
-\end{verbatim}
-}
-
-\begin{comment}
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE \hspace{0.25cm} A[i] = B[i] + 10;
-\STATE \hspace{0.25cm} B[i] = A[i] * D[i-1];
-\STATE \hspace{0.25cm} C[i] = A[i] * D[i-1];
-\STATE \hspace{0.25cm} D[i] = B[i] * C[i];
-\end{algorithmic}
-
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{A}_0$ = $\phi(\texttt{A},\texttt{A}_1)$
-\STATE 2. $\texttt{B}_0$ = $\phi(\texttt{B},\texttt{B}_1)$
-\STATE 3. $\texttt{C}_0$ = $\phi(\texttt{C},\texttt{C}_1)$
-\STATE 4. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$
-\STATE 5. $\texttt{A}_1$ = $\texttt{A}_0$; $\texttt{A}_1$[i] = $\texttt{B}_0$[i] + 10; \COMMENT{equiv. to Update}
-\STATE 6. $\texttt{B}_1$ = $\texttt{B}_0$; $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 7. $\texttt{C}_1$ = $\texttt{C}_0$; $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 8. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
-\end{algorithmic}
-\end{comment}
-
-There is a cycle around \texttt{B\_1 = PHI(B\_0,B\_2)} that includes statement 
-\texttt{A\_1 = update(A\_0, i, B\_1[i] + 10)} and that statement won't be vectorized even 
-though in fact there is no loop-carried dependency from the write of \texttt{B\_1[i]} at 6 to the
-read of \texttt{ ... = B\_1[i]} at 8.
-
-The following algorithm removes certain infeasible loop-carried dependencies that are due to array writes. Consider a loop with index $0 \le j < J$
-nested at $i,j,k$. Here $i$ represents the enclosing loops of $j$ and $k$ represents the enclosed loops in $j$.
-
-\begin{algorithmic}
-\FOR {each array $\texttt{A}$ written in loop $j$}
-\STATE \COMMENT { including enclosed loops in $j$ }
-\STATE dep = False
-\FOR {each pair def: $\texttt{A}_m[f(i,j,k)] = ... $, and use: $ ... = \texttt{A}_n[f'(i,j,k)]$ in loop $j$}
-\IF {$\exists \underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$, s.t.
-$0\!
-\le\!
-\underline{i}\!
-<\!
-I,\: 0\!
-\le\!
-\underline{j}, \underline{j}' \!
-<\!
-J,\: 0\!
-\le\!
-\underline{k}, \underline{k}' \!
-<\!
-K,\: \underline{j}<\underline{j}'$, and $f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$}
-\STATE dep = True
-\ENDIF
-\ENDFOR
-\IF {dep == False}
-\STATE remove back edge into $\texttt{A}$'s $\phi$-node in loop $j$.
-\ENDIF
-\ENDFOR
-\end{algorithmic}
-
-Consider a loop $j$ enclosed in some fixed $\underline{i}$. Only if an update (definition) $\texttt{A}_m[f(i,j,k)] = ... $ at some iteration $\underline{j}$
-references the \emph{same} array element as a use $ ... = \texttt{A}_n[f'(i,j,k)]$ at some later iteration $\underline{j}'$,
-we may have a loop-carried dependence for $\texttt{A}$ due to this def-use pair. (In contrast, Cytron's algorithm inserts a loop-carried dependency every time there is an array update.)
-The algorithm above examines all def-use pairs in loop $j$, including defs and uses in nested loops, searching for values $\underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$ that satisfy
-$f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$. If such values exist for some def-use pair, then there is a potential
-loop-carried dependence on $\texttt{A}$; otherwise there is not and we can remove the spurious backward edge thus ``freeing up'' statements for vectorization. %\ana{Double check this!!!}
-
-Consider the earlier example.
-There is a single loop, $i$.
-Clearly, there is no pair $\underline{i}$ and $\underline{i}'$, where $\underline{i} < \underline{i}'$ that make $\underline{i} = \underline{i}'$
-due to the def-use pairs of \texttt{A} 5-6 and 5-7.
-Therefore, we remove the back edge from 5 to the phi-node 1.
-Analogously, we remove the back edges from 6 to 2 and from 7 to 3.
-However, there are many values $\underline{i} < \underline{i'}$ that make $\underline{i} = \underline{i'}-1$ and the back edge from 8 to 4 remains (def-use pairs for \texttt{D}).
-As a result of removing these spurious edges, Basic Vectorization will find that statement 5 is vectorizable.
-Statements 6, 7 and 8 will correctly appear in the FOR loop.
-
-Note however, that this step renders some array phi-nodes target-less. We handle target-less phi-nodes with a minor extension of Basic Vectorization (Phase 2). 
-First, we merge closures that update the same array. This simplifies handling of array $\phi$-nodes: if each closure is turned into a separate loop
-each loop will need to have its own array phi-node to account for the update and this would complicate the analysis.
-Second, we add the target-less node of array \texttt{A} back to the closure that updates \texttt{A} ---
-the intuition is, even if there is no loop-carried dependence from writes to reads on \texttt{A}, \texttt{A} is written and the write (i.e., update) cannot be vectorized; 
-therefore, the updated array has to carry to the next iteration of the loop. Third, in cases when the phi-node remains target-less, i.e., cases when the
-array write can be vectorized, we have to properly remove the phi-node replacing uses of the left-hand side of the phi-node with its arguments.
-
-%\ana{Commented out previous writeup. Removal/handling of targetless phi-nodes will go into the Extension to Basic vectorization.}
-
-\begin{comment}
-The final step is to remove array phi-nodes that are left without incoming back edges after removal of infeasible edges.
-Such a phi-node is redundant when there are no loop-carried dependences that are due to writes to the corresponding array.
-In our running example, the phi-nodes for \texttt{A}, \texttt{B}, and \texttt{C}, are redundant.
-
-Removing redundant phi-nodes proceeds from \emph{innermost to outermost} loops as follows.
-Denote a redundant phi-node by $\texttt{A}_1 = \phi(\texttt{A}_0,\texttt{A}_k)$. $\texttt{A}_0$ is the array at the beginning of the loop
-and $\texttt{A}_k$ is the array at the end of an iteration (including the last iteration). We consider two cases:
-\begin{enumerate}
-\item[Case 1] Inner-to-outer or same-level def-use edge
-\begin{itemize}
-\item If use is an array read $ ... = \texttt{A}_1[f(i)]$, then replace $\texttt{A}_1$ with $\texttt{A}_0$. This is correct, i.e., it is guaranteed that the use
-references $\texttt{A}_0$; had the use been referencing an array modified in an earlier iteration, the back-edge to the phi-node would
-not have been removed and the phi-node would not have been classified as redundant.
-\item If use is an array write (update) $\texttt{A}_2 = \texttt{A}_1; \texttt{A}_2[i] = ... $ equiv. to $\texttt{A}_2 = \mathit{update}(\texttt{A}_1, i, ...)$ then handling becomes more tricky.
-The semantics of this update is that when $i$ is 0, then $\texttt{A}_0$ is being read.
-But when $i > 0$, then $\texttt{A}_k$ is being read.
-If we are at this case, this means that the def-use edge is same-level and we are removing a redundant node in the loop enclosing the update.
-The removal of the original phi-node entails pushing the phi-node at the update:
-$\texttt{A}_2 = \phi(\texttt{A}_0,\texttt{A}_k); \texttt{A}_2[i] = ... $ equiv. to $\texttt{A}_2 = \mathit{update}(\phi(\texttt{A}_0,\texttt{A}_k), i, ...)$.
-\item If use is a phi-node $... = \phi(\texttt{A}_1,...)$, that means we have an outer-to-inner edge. The
-
-\end{itemize}
-\end{enumerate}
-For each outer-to-inner or same-level def-use edge
-that originates at the node, replace $\texttt{A}_1$ at the use node by $\texttt{A}_0$.
-For each inner-to-outer def-use edge, replace $\texttt{A}_1$ by $\texttt{A}_k$. In this case, the use references the array
-that results from the computation in the loop.
-
-Our running example is transformed as follows. Removing the phi-nodes for \texttt{A}, \texttt{B}, and \texttt{C} transforms the
-replaces the references to $\texttt{A}_0$, $\texttt{B}_0$, and $\texttt{C}_0$, with references to $\texttt{A}$, $\texttt{B}$, and $\texttt{C}$
-respectively, i.e., the arrays before the loop.
-
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$
-\STATE 2. $\texttt{A}_1$ = $\texttt{A}$; $\texttt{A}_1$[i] = $\texttt{B}$[i] + 10;
-\STATE 3. $\texttt{B}_1$ = $\texttt{B}$; $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 4. $\texttt{C}_1$ = $\texttt{C}$; $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
-\end{algorithmic}
-\end{comment}
-
-\begin{comment}
-\subsubsection{Array MUX refinement}
-
-\ana{TODO: WIP, get to BV without MUX refinement for now? For now, skip MUX refinement.}
-
-Next, the algorithm refines array MUX statements.
-MPC-source after Cytron's SSA may result in statements $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$, which imply that any index
-of $A$ can be written at this point and therefore there is a loop-carried dependency. In some cases the MUX can be refined to just a single
-index or a pair of indices, e.g., $\texttt{A}_j[i] = \emph{MUX}(c, \texttt{A}_k[i], \texttt{A}_l[i])$.
-
-%\ana{Clarify why this helps. Still work in progress.}
-This is to reduce the dimensionality of simd-ified computation. Technically, $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$ is
-a simdified operation that can be carried out in parallel "in one round". However, particularly when \texttt{A} is a multi-dimensional array, there is
-substantial increase in the size of the arrays (vectors) we send to SIMD operations. Refining to an update to a specific index would reduce
-the size of those vectors. Note that this is a heuristic that handles a common case, but not all cases of array updates.
-%\ana{TODO: Simplify this algorithm, taking into account the restriction to canonical updates. It should be handling \_all\_ cases.}
-
-\begin{algorithmic}
-\FOR {each $\mathit{stmt}$: $\texttt{A}_j = \emph{MUX}(c, \texttt{A}_k, \texttt{A}_l)$ in the MPC-source seq.}
-\STATE $i_1 = \mathit{find\_update}(A_k)$ \COMMENT { Is null when $A_k = \phi(...)$ }
-\STATE $i_2 = \mathit{find\_update}(A_l)$ \COMMENT { Is null when $A_l = \phi(...)$ }
-\IF {$i_1 == i_2$ or $i_1$ is null or $i_2$ is null} \STATE \COMMENT{ With our restrictions on writes we must have $i_1 = i_2$. }
-\STATE replace $\mathit{stmt}$ with \\
-\STATE $\texttt{A}_j = \texttt{A}_{j-1}$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
-%\ELSIF {$i_1 \ncong i_2$}
-%\STATE replace $\mathit{stmt}$ with \\
-%\STATE $\texttt{A}_j = \texttt{A}_k$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
-\ELSE
-\STATE $\mathit{stmt}$ stays as is
-\ENDIF
-\ENDFOR
-\end{algorithmic}
-\end{comment}
-
-\subsubsection{Restricting Array Writes}
-
-For now, we restrict array updates to \emph{canonical updates}. Assume (for simplicity) a two-dimensional array \texttt{A[I,J]}. A canonical update is the following:
-\begin{verbatim}
-for i in range(I):
-  for j in range(J):
-    ...
-    A[i,j] = ...
-    ...
-\end{verbatim}
-
-The update \texttt{A[i,j]} can be nested into an inner loop and there may be multiple updates, i.e., writes to \texttt{A[i,j]}.
-However, update such as \texttt{A[i-1,j] = ...} or \texttt{A[i-1,j-1] = ...}, etc., is not allowed.
-Additionally, while there could be several different loops that perform canonical updates, they must be of the same dimensionality, i.e., an update of higher or lower dimension, e.g., \texttt{A[i,j,k] = ...} is not allowed.
-We compute the \emph{canonical dimensionality} of each write array by examining the array writes in the original program and rejecting programs that violate the canonical write restriction.
-This restriction simplifies reasoning in this early stage of the compiler; we will look to relax the restriction in future work.
-
-Another restriction/assumption is that we assume the output array is given as input with initial values, and it is of size consistent with its canonical dimensionality. 
-%\ana{This restriction will require that we update our benchmarks that come with output arrays.}
-
-Reads through an arbitrary formula, such as \texttt{A[i-1]} for example, are allowed; currently, the projection function returns dummy values if the read formula is out of bounds; we assume the programmer ensures that the program still computes correct output in this case.
-
-\subsubsection{Changes to Basic Vectorization}
-
-\begin{comment}
-There are two changes to Basic Vectorization to account for write arrays. We can account for them during processing of raise dimension/drop dimension in Phase 1 of Basic Vectorization.
-
-One change to Basic vectorization is the expansion of dimension if the array write or read occurs in a nested loop.
-That is, if there is an update $\texttt{A}[i,j] = ...$ that occurs in loop dimensionality $i,j,k$, $\texttt{A}[i,j]$ will be rewritten into $\texttt{A}[i,j,k]$.
-Similarly, a read $\texttt{A}[f(i,j)]$ will be rewritten into $\texttt{A}[f(i,j),k]$.
-($\texttt{A}[f(i,j),k]$ will be reshaped as $f(i,j)*I*J+k$ during Phase 1 of Basic Vectorization.
-\ana{TODO: Double check.})
-\end{comment}
-
-In addition to the changes for the handling of target-less phi-nodes, Basic Vectorization has to handle def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable.
-The definition can be an update \texttt{A\_2 = update(A\_1,i,...)}, a pseudo $\phi$-node \texttt{A\_2 = PHI(A\_0,A\_1)}, etc..
-Note that $\phi$-nodes for arrays have no subscript operations the way there are subscript operations in analysis-introduced arrays representing scalars.
-While there are variations, the most intuitive implementation will perform Basic Vectorization Phase 1 as is, inserting
-$\mathit{raise\_dim}$ and $\mathit{drop\_dim}$ in the same way.
-However, the implementation of raise dimension and drop dimension will be adapted because the dimension cannot be raised or dropped to a dimension lower than the canonical one.
-Consider a def-use edge $X \rightarrow Y$ for an array $\texttt{A}$.
-
-\begin{enumerate}
-
-\item same-level $X \rightarrow Y$. Do nothing, propagate the array, which happens to be of the right dimension. 
-%\ana{There might be some opportunities to do copy propagation optimization and save some cycles, but let's leave this for later.}
-
-\item inner-to-outer $X \rightarrow Y$ triggers the addition of $\mathit{drop\_dim}$.
-However, the dimensionality cannot be dropped below the canonical dimensionality of the array.
-E.g., if the dimensionality of the loop enclosure $X$ is already at the canonical one, then $\mathit{drop\_dim}$ has no effect.
-
-\item outer-to-inner $X \rightarrow Y$ riggers  $\mathit{raise\_dim}$.
-Again, if the dimensionality of the loop enclosure of $Y$ is smaller or same as the canonical dimensionality of the array, then it has no effect, otherwise, if dimensionality is greater than the canonical dimensionality, $\mathit{raise\_dim(...)}$ (at $X$) is the same as in Basic Vectorization.
-
-\item "mixed" $X \rightarrow Y$. We assume that the mixed edge is transformed into an inner-to-outer followed by outer-to-inner edge before we perform vectorization, just as with Basic vectorization.
-
-\end{enumerate}
-
-If the use of the array is a read $\texttt{A}[f(i,j,k)]$ different than a canonical read $\texttt{A}[i,j,k]$, then we need to add a reshape operation as 
-all arrays are $\texttt{A}[i,j,k]$. It can be added after raise\_dim/drop\_dim or incorporated in these operations.
-The bulk of the change is in Phase 2 of Basic Vectorization as outlined earlier.
-
-\subsubsection{Examples with Array Writes}
-
-\paragraph{Example 1}
-
-First, the canonical dimensionality of all \texttt{A,B,C} and \texttt{D} is 1. 
-After Phase 1 of Basic Vectorization the Aiken's array write example will be (roughly) as follows:
-
-{\small
-\begin{verbatim}
-for i in range(N):
-1. A_1 = PHI(A_0,A_2)
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. A_2 = update(A_1, I, B_1[I] + 10); 
-6. B_2 = update(B_1, I, A_2[I] * D_1[I-1]);
-7. C_2 = update(C_1, I, A_2[I] * D_1[I-1]);
-8. D_2 = update(D_1, I, B_2[I] * C_2[I]);
-\end{verbatim}
-}
-
-Note that since all def-uses are same-level (i.e., reads and writes of the array elements) 
-no raise dimension or drop dimension happens.
-
-\begin{comment}
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{A}_1$ = $\phi(\texttt{A}_0,\texttt{A}_2)$
-\STATE 2. $\texttt{B}_1$ = $\phi(\texttt{B}_0,\texttt{B}_2)$
-\STATE 3. $\texttt{C}_1$ = $\phi(\texttt{C}_0,\texttt{C}_2)$
-\STATE 4. $\texttt{D}_1$ = $\phi(\texttt{D}_0,\texttt{D}_2)$
-\STATE 5. $\texttt{A}_2 = update(\texttt{A}_1, \stackrel{\rightarrow}{i}, \texttt{B}_1[\stackrel{\rightarrow}{i}] + 10)$;
-\STATE 6. $\texttt{B}_2 = update(\texttt{B}_1, \stackrel{\rightarrow}{i}, \texttt{A}_2[\stackrel{\rightarrow}{i}] * \texttt{D}_1[\stackrel{\rightarrow}{i-1}])$;
-\STATE 7. $\texttt{C}_2 = update(\texttt{C}_1, \stackrel{\rightarrow}{i}, \texttt{A}_2[\stackrel{\rightarrow}{i}] * \texttt{D}_1[\stackrel{\rightarrow}{i-1}])$;
-\STATE 8. $\texttt{D}_2 = update(\texttt{D}_1; \stackrel{\rightarrow}{i}, \texttt{B}_2[\stackrel{\rightarrow}{i}] * \texttt{C}_2[\stackrel{\rightarrow}{i}])$;
-\end{algorithmic}
-\end{comment}
-
-Phase 2 computes the closure of 4; $cl = \{4,6,7, 8\}$ 
-while 5 is vectorizable. Recall that 1,2, and 3 are target-less phi-nodes. Since the closure $cl$ includes updates to \texttt{B} and \texttt{C},
-the corresponding phi-nodes are added back to the closure and the def-use edges are added back to the target-less nodes. The uses of \texttt{A\_1} and \texttt{B\_1} in the vectorized statement
-turn into uses of \texttt{A\_0} and \texttt{B\_0} respectively; this is done for all original target-less phi-node.
-(But note that \texttt{A\_0} is irrelevant; the update writes into array \texttt{A\_2} in parallel.)
-Finally, the target-less phi-node for \texttt{A} is discarded.
-
-{\small
-\begin{verbatim}
-1. A_2 = update(A_0, I, ADD_SIMD(B_0[I],10));
-   equiv. to A_2[I] = ADD_SIMD(B_0[I],10)
-FOR i=0; i<N; i++; // MOTION loop
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. B_2 = update(B_1, i, A_2[i] * D_1[i-1]);
-   equiv. to B_2 = B_1; B_2[i] = A_2[i] * D_1[i-1];
-6. C_2 = update(C_1, i, A_2[i] * D_1[i-1]);
-7. D_2 = update(D_1, i, B_2[i] * C_2[i]);
-\end{verbatim}
-}
-
-\begin{comment}
-\begin{algorithmic}
-\STATE 1. $\texttt{A}_2 = update(\texttt{A}_0; \stackrel{\rightarrow}{i}, \mathit{ADD\_SIMD}(\texttt{B}_0[\stackrel{\rightarrow}i],[10,...])$ \COMMENT{Fully vectorized, size N.}
-\STATE FOR i=0; i<N; i++; \COMMENT{ MOTION loop }
-\STATE 2. $\texttt{B}_1$ = $\phi(\texttt{B}_0,\texttt{B}_2)$
-\STATE 3. $\texttt{C}_1$ = $\phi(\texttt{C}_0,\texttt{C}_2)$
-\STATE 4. $\texttt{D}_1$ = $\phi(\texttt{D}_0,\texttt{D}_2)$
-\STATE 5. $\texttt{B}_2 = update(\texttt{B}_1, i, \mathit{MUL}(\texttt{A}_2[i], \texttt{D}_1[{i-1}])$
-\STATE 6. $\texttt{C}_2 = update(\texttt{C}_1, i, \mathit{MUL}(\texttt{A}_2[i], \texttt{D}_1[{i-1}])$
-\STATE 7. $\texttt{D}_2 = update(\texttt{D}_1, i, \mathit{MUL}(\texttt{B}_2[i], \texttt{C}_1[i])$
-\end{algorithmic}
-\end{comment}
-
-\paragraph{Example 2} Now consider the MPC Source of Histogram:
-
-{\small
-\begin{verbatim}
-for i in range(0, num_bins):
-   res1 = PHI(res, res2)
-   for j in range(0, N):
-      res2 = PHI(res1, res3)
-      tmp1 = (A[j] == i)
-      tmp2 = (res2[i] + B[j])
-      tmp3 = MUX(tmp1, res2[i], tmp2)
-      res3 = Update(res2, i, tmp3)     
-return res1
-\end{verbatim}
-}
-
-The canonical dimensionality of \texttt{res} is 1.
-Also, the phi-node \texttt{res1 = PHI(res, res2)} is a target-less phi-node (the implication being that the inner for loop can be vectorized across $i$).
-After Phase 1, Basic vectorization produces the following code (statements are implicitly vectorized along $i$ and $j$).
-In a vectorized update statement, we can ignore the incoming array, \texttt{res2} in this case.
-The update writes (in parallel) all locations of the 2-dimensional array, in this case it sets up each \texttt{res3[i,j] = tmp3[i,j]}.
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I = raise_dim(i, ((i:num_bins),(j:N)))
-for i in range(0, num_bins):
-   res1 = PHI(res, res2') // target-less phi-node
-   res1' = raise_dim(res1, (j:N))
-   for j in range(0, N):
-      res2 = PHI(res1', res3)
-      tmp1 = (A1 == I)
-      tmp2 = (res2 + B1)
-      tmp3 = MUX(tmp1, res2, tmp2)
-      res3 = Update(res2, (I,J), tmp3) 
-   res2' = drop_dim(res2)    
-res1'' = drop_dim(res1)
-return res1''
-\end{verbatim}
-}
-
-Processing the inner loop in Phase 2 vectorizes \texttt{tmp1 = (A1 == I)} along the $j$ dimension but leaves the rest of the statements in a MOTION loop.
-Processing the outer loop is interesting.
-This is because the PHI node is a target-less node, and therefore, there are no closures!
-Several things happen.
-(1) Everything can be vectorized along the $i$ dimension.
-(2) We remove the target-less PHI node, however, we must update uses of \texttt{res1} appropriately: the use at raise\_dim goes to the first argument of the PHI function and the use at drop\_dim goes to the second argument.
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I1 = raise_dim(i, ((i:num_bins),(j:N)))
-
-tmp1[I,J] = (A1[I,J] == I1[I,J])
-
-res1' = raise_dim(res, (j:N)) // replacing res1 with res, 1st arg
-for j in range(0, N):
-   res2 = PHI(res1', res3)
-   tmp2[I,j] = (res2[I,j] + B1[I,j])
-   tmp3[I,j] = MUX(tmp1[I,j], res2[I,j], tmp2[I,j])
-   res3 = Update(res2, (I,j), tmp3)
-   equiv. to res3 = res2; res3[I,j] = tmp3[I,j]
-res2' = drop_dim(res2)   
-res1 = drop_dim(res2') // replacing with res2', 2nd arg. NOOP
-return res1
-\end{verbatim}
-}
-
-\begin{comment}
-The cleanup phase can then get rid of $j$ in the \texttt{res} variables and \texttt{tmp} variables, so the loop operates over the \texttt{num\_bins}-size vectors. 
-This will lead to dropping raise\_dim and drop\_dim as well:
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I1 = raise_dim(i, ((i:num_bins),(j:N)))
-
-tmp1[I,J] = (A1[I,J] == I1[I,J])
-
-for j in range(0, N):
-   res2 = PHI(res1, res3) // now 1-dim array
-   tmp2[I] = ADD_SIMD(res2[I],B1[I,j])
-   tmp3[I] = MUX_SIMD(tmp1[I,j], res2[I], tmp2[I])
-   res3 = Update(res2, I, tmp3[I])
-return res2
-\end{verbatim}
-}
-\end{comment}
-
-\begin{comment}
-\paragraph{Example 2}
-
-Consider MPC-source of Histogram after removal of infeasible edges and redundant phi-nodes, and MUX refinement:
-
-\begin{algorithmic}
-
-\STATE res!1 = []
-\STATE for i: plaintext in range(0, num\_bins):
-\STATE \hspace{0.25cm} res!2 = $\phi$(res!1, res!3)
-\STATE  \hspace{0.25cm} ...
-\STATE res!4 = res!2 \COMMENT{Added due to the "mixed" def-use edge.}
-\STATE for i: plaintext in range(0, num\_bins):
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE \hspace{0.25cm}        for j: plaintext in range(0, N):
-\STATE \hspace{0.5cm}            res!5 = $\phi$(res!4, res!7)
-\STATE \hspace{0.5cm}            !2!3 = (A[j] == i)
-\STATE \hspace{0.5cm}            !3!3 = (res!5[i] + B[j])
-\STATE \hspace{0.5cm}            res!6 = res!5; res!6[i] = !3!3
-\STATE \hspace{0.5cm}            res!7 = res!6; res!7[i] = MUX(!2!3, res!6[i], res!5[i])
-
-\end{algorithmic}
-
-After Phase 1 of Basic Vectorization:
-
-\begin{algorithmic}
-\STATE res!1 = []
-\STATE for i: plaintext in range(0, num\_bins):
-\STATE \hspace{0.25cm} res!2 = $\phi$(res!1, res!3)
-\STATE  \hspace{0.25cm} ...
-\STATE res!4 = res!2 \COMMENT{Added due to the "mixed" def-use edge. Because the canonical dimensionality of res is 1, we DO NOT drop dimension of res!2.}
-\STATE A[$i,j$] = $\mathit{raise\_dim}$(A[$j$],$i$) \COMMENT{Copying row A num\_bin times. Reusing name.}
-\STATE B[$i,j$] = $\mathit{raise\_dim}$(B[$i$],$j$) \COMMENT{Copying column B N times. Reusing name.}
-\STATE res!4[$i,j$] = $\mathit{raise\_dim}$(res!4[$i$],$j$) \COMMENT{Reusing name.}
-\STATE for i: plaintext in range(0, num\_bins):
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE \hspace{0.25cm}        for j: plaintext in range(0, N):
-\STATE \hspace{0.5cm}            res!5 = $\phi$(res!4, res!7) \COMMENT{Do nothing. res!4 has been raised to proper dimension.}
-\STATE \hspace{0.5cm}            !2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{EQ}$(A[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], [[0,0...], [1,1...],...[num\_bins-1,num\_bins-1...]])
-\STATE \hspace{0.5cm}            !3!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{ADD}$(res!5[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$],B[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$])
-\STATE \hspace{0.5cm}            res!6 = res!5 \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.5cm}            res!6[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = !3!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$]
-\STATE \hspace{0.5cm}            res!7 = res!6
-\STATE \hspace{0.5cm}            res!7[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{MUX}$(!2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], res!6[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], res!5[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$])
-\end{algorithmic}
-
-After Phase 2 and Phase 3. The EQ operation is vectorizable across both dimensions, and the rest of the computation is vectorizable across the $i$-dimension.
-
-\begin{algorithmic}
-\STATE ...
-\STATE !2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{EQ\_SIMD}$(A[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], [[0,0...], [1,1...],...[num\_bins-1,num\_bins-1...]])
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE FOR j=0; j<N; j++; \COMMENT{MOTION loop}
-\STATE \hspace{0.25cm}            res!5 = $\phi$(res!4, res!7) \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.25cm}            !3!3[$\stackrel{\rightarrow}i,j$] = $\mathit{ADD\_SIMD}$(res!5[$\stackrel{\rightarrow}i,j$],B[$\stackrel{\rightarrow}i,j$])
-\STATE \hspace{0.25cm}            res!6 = res!5 \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.25cm}            res!6[$\stackrel{\rightarrow}i,j$] = !3!3[$\stackrel{\rightarrow}i,j$]
-\STATE \hspace{0.25cm}            res!7 = res!6
-\STATE \hspace{0.25cm}            res!7[$\stackrel{\rightarrow}i,j$] = $\mathit{MUX\_SIMD}$(!2!3[$\stackrel{\rightarrow}i,j$], res!6[$\stackrel{\rightarrow}i,j$], res!5[$\stackrel{\rightarrow}i,j$])
-\end{algorithmic}
-\end{comment}
-
-%\section{Divide-and-Conquer}
-%\label{sec:div_and_conq}
-%\ana{TODO: Now that we have broken FOR loops into smaller chunks, we can add Divide-and-conquer reasoning with Z3 and implement this additional transform.}
-
-\section{Implementation and Evaluation}
-\label{sec:implementation}
-
-\section{Related Work}
-\label{sec:related}
-
-\paragraph{MPC languages and compilers.}
-
-Languages and compilers for secure computation have seen significant attention and advances in recent years. The early MPC compilers Fairplay~\cite{CCS:BenNisPin08}, 
-and Sharemind~\cite{ESORICS:BogLauWil08} were followed by PICCO~\cite{CCS:ZhaSteBla13}, Obliv-C~\cite{Zahur:2015}, TinyGarble~\cite{SP:SHSSK15}, 
-Wystiria~\cite{SP:RasHamHic14}, and others. A new generation of MPC compilers 
-includes SPDZ/SCALE-MAMBA/MP-SPDZ~\cite{Keller:2020} and the ABY/HyCC/MOTION~\cite{NDSS:DemSchZoh15,CCS:BDKKS18,Braun:2022} frameworks. 
-These two families are the state-of-the art and
-are actively developed. Another recent development is Viaduct, a functional language and compiler that supports a range of secure computation 
-frameworks, including MPC and ZKP. Hastings et el. present a review of compiler frameworks~\cite{Hastings:2019}.
-
-While each of these languages and compilers brings in new ideas and advances, none addresses the problem of ``circuit independent'' intermediate 
-representation and optimization. We envision a classical compiler structure: (1) a Wysteria, Viaduct, Obliv-C, or IMP Source front end, including rich type systems and 
-AST-level semantic analysis, compile into the MPC Source IR, (2) MPC Source-level optimizations take place, followed by (3) back-end compilers into circuits. 
-Our focus is at the intermediate level.
-
-%{\begin{center}
-%\includegraphics[width=0.5\linewidth]{figs/focus.png}
-%\end{center}
-%}
-
-Many works focus on the implementation of MPC protocols exposing an API to the programmer. For example, the ABY/MOTION line of 
-compiler frameworks provides a library of MPC primitives; the programmer writes MPC programs in C++ on top of the library. These back ends implement 
-different protocols and allow for mixing, but notably, they leave it to the programer to assign different protocols to different parts of the 
-computation and perform share conversion accordingly. In addition, MOTION provides SIMD primitives, which allows for efficient execution 
-of MPC operations, but again, using SIMD primitives is the responsibility of the programer. There is interest in frameworks for automatic mixing, 
-e.g., \cite{CCS:BDKKS18,Ishaq:2019,Fang:2022}.
-
-Other works, e.g., Obliv-C~\cite{Zahur:2015}, Wystiria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021} focus on higher-level language design, particularly information-flow systems that 
-restrict flow between secure and insecure parts of the program. 
-%Key ideas such as ${\sf insecure} <: {\sf secure}$ qualifier systems are similar to the ideas we 
-%present in~\Secref{sec:interprocedural} and~\Secref{sec:zkp} and we intend to draw on these work. While we will design a high-level language expanding our 
-%intraprocedural IMP-MPC syntax and semantics into the \emph{interprocedural} case, our interest and focus lie primarily in the ``machine-independent'' optimizations 
-%thrust of the proposal.
-
-%There is significant development in ZKP languages as well. ~\ana{Alex Ozdemir's talk, ZoKrates, Cairo, others} Our work focuses on ZKP via MPC and our key idea
-%is the separation of the higher-level ``backend-independent'' compilation and optimization and the lower-level compilation into circuits.
-
-\paragraph{Classical HPC compilers.}
-Automatic vectorization is a longstanding problem in high-performance computing (HPC). 
-There are thousands of works in this area reflecting over 40 years of research. We presented a vectorization
-algorithm for MPC Source, essentially extending classical loop vectorization \cite{Allen:1987}. In HPC vectorization, conditional control flow 
-presents a challenge --- one cannot estimate the cost of a schedule or vectorize branches in a straightforward manner --- in contrast to 
-MPC Source vectorization. 
-We view Karrenberg's work on Whole function vectorization~\cite{Karrenberg:2015} as most closely related to ours --- it linearizes the program and vectorizes 
-both branches of a conditional applying masking to avoid execution of the branch-not-taken code, and selection (similar to MUX) to select the correct value based on the 
-result of the condition at runtime. 
-The problem is that masking and selection, or more generally, handling control predicates~\cite{Benabderrahmane:2010,Karrenberg:2015}, 
-can lead to \emph{slowdown}. 
-%As a result, vectorizing compilers and polyhedral compilers tend to focus on straight-line code due to the unpredictable 
-%overhead conditionals may introduce. \ana{add Louis noel pochet's benchmarks. citation on vectorizing compilers???}
-
-We argue that vectorization over linear MPC Source is a different problem, one that warrants a new look, while drawing from 
-results in HPC.
-Since both branches of the conditional and the multiplexer \emph{always} execute, not only can we apply aggressive vectorization on linear code, but (perhaps more importantly) 
-we can also build analytical models that accurately predict execution time. These models in turn would drive optimizations such as vectorization, protocol mixing, and others. 
-Vectorization meshes in with those additional optimizations in non-trivial ways.
-
-Furthermore, extensions of classical loop vectorization with array writes, arbitrary indexing, including non-affine indexing, and interaction with SSA are non-trivial 
-and present novel challenges and opportunities for contribution. Polyhedral parallelization~\cite{Benabderrahmane:2010} considers a higher-level source (typically AST) 
-representation, while our work takes advantage of linear MPC Source and SSA form. The work by Karrenberg~\cite{Karrenberg:2015} is rare in that space, in the sense that it considers 
-vectorization over SSA form, which has similarities to MPC Source. We consider different array representation, notion of dependence, and reasoning 
-about dependence, which we conjecture is more suitable for MPC Source. Buscher~\cite{Buscher:2018} considers SIMD-vectorization 
-at the level of source code, which then combines with circuit-level optimizations in the TinyGarble compiler. 
-He proposes using an off-the-shelf polyhedral compiler, however, application is limited to only two routines, 
-essentially just inner product and euclidian distance; it is unclear how effective the off-the-shelf compiler is. 
-In contrast, we consider vectorization at the level of MPC Source separating ``backend-independent'' vectorization
-and circuit-level amortization (done by MOTION). We apply our compiler on a wide range of routines.
-
-%\section{Future Work}
-%\label{sec:implementation_and_benchmarks}
-%\input{../sections/implementation_and_benchmarks.tex}
-%\input{../sections/evaluation.tex}
-
-\section{Conclusion and Future Work}
-\label{sec:conclusion}
-%\input{../sections/conclusion.tex}
-
-%\begin{acks}
-%    \ishaq{TODO}
-%\end{acks}
-
-%\printbibliography
-
-\bibliographystyle{plain}
-%\bibliography{c:/trink/Doc/MyTexFiles/texrefs}
-%\bibliography{mybib,library,library1,library2,library3,library_ana}
-\bibliography{library_ana,crypto_crossref}
-
-% that's all folks
-\end{document}

### 2022-04-28 overleaf/overview.tex
Split paper into sections.

--- 2022-04-27 overleaf/paper_SIMD.tex
@@
-% !TeX TXS-program:compile = txs:///pdflatex/[--shell-escape]
-%
-% acm templates v1.57 (sigconf template)
-%
-%\documentclass[sigconf, screen, review, anonymous, natbib=false]{acmart}
-\documentclass[sigconf, screen, natbib=false, dvipsnames, table]{acmart}
-\settopmatter{printacmref=true}
-% mandatory for CCS'19
-\usepackage{balance}
-\usepackage{mathtools}
-% for creating a balanced last page (usually last page with references)
-
-% % defining the \BibTeX command - from Oren Patashnik's original BibTeX documentation.
-% \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08emT\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-
-% Rights management information.
-% This information is sent to you when you complete the rights form.
-% These commands have SAMPLE values in them; it is your responsibility as an author to replace
-% the commands and values with those provided to you when you complete the rights form.
-%
-% These commands are for a PROCEEDINGS abstract or paper.
-
-\copyrightyear{2019}
-\acmYear{2019}
-\setcopyright{acmcopyright}
-\acmConference[CCS '19]{2019 ACM SIGSAC Conference on Computer and Communications Security}{November 11--15, 2019}{London, United Kingdom}
-\acmBooktitle{2019 ACM SIGSAC Conference on Computer and Communications Security (CCS '19), November 11--15, 2019, London, United Kingdom}
-\acmPrice{15.00}
-\acmDOI{10.1145/3319535.3339818}
-\acmISBN{978-1-4503-6747-9/19/11}
-
-%
-% These commands are for a JOURNAL article.
-%\setcopyright{acmcopyright}
-%\acmJournal{TOG}
-%\acmYear{2018}\acmVolume{37}\acmNumber{4}\acmArticle{111}\acmMonth{8}
-%\acmDOI{10.1145/1122445.1122456}
-
-%
-% Submission ID.
-% Use this when submitting an article to a sponsored event. You'll receive a unique submission ID from the organizers
-% of the event, and this ID should be used as the parameter to this command.
-%\acmSubmissionID{123-A56-BU3}
-
-%
-% The majority of ACM publications use numbered citations and references. If you are preparing content for an event
-% sponsored by ACM SIGGRAPH, you must use the "author year" style of citations and references. Uncommenting
-% the next command will enable that style.
-%\citestyle{acmauthoryear}
-
- \makeatletter
- % \renewcommand{\section}{\abovedisplayskip 10\p@ \@plus3\p@ \@minus1\p@%
- %                       \belowdisplayskip 5\p@ \@plus3\p@ \@minus1\p@%
- %                       \abovedisplayshortskip 0pt \@plus2\p@%
- %                       \belowdisplayshortskip 0pt \@plus2\p@ \@minus0\p@%
- %                       \@startsection{section}{1}{\z@}%
- %                        {-17\p@ \@plus -4\p@ \@minus -4\p@}%
- %                        {6\p@ \@plus 4\p@ \@minus 4\p@}%
- %                        {\normalfont\large\bfseries\boldmath
- %                         \rightskip=\z@ \@plus 8em\pretolerance=10000 }}
- \renewcommand{\subsection}{\@startsection{subsection}{2}{\z@}%
-                        {-8\p@ \@plus -4\p@ \@minus -4\p@}%
-                        {5\p@ \@plus 2\p@ \@minus 2\p@}%
-                        {\normalfont\Large\bfseries\boldmath
-                         \rightskip=\z@ \@plus 3em\pretolerance=10000 }}
-  \renewcommand{\subsubsection}{\@startsection{subsubsection}{3}{\z@}%
-                        {-6\p@ \@plus -4\p@ \@minus -4\p@}%
-                        {1\p@ \@plus 1\p@ \@minus 0\p@}%
-                        {\normalfont\normalsize\bfseries\boldmath}}
-% \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\z@}%
-%                       {-8\p@ \@plus -4\p@ \@minus -4\p@}%
-%                       {-2\p@ \@plus -0.22em \@minus -0.1em}%
-%                       {\normalfont\normalsize\bfseries}}
- \makeatother
-
-% *** GRAPHICS RELATED PACKAGES ***
-\usepackage{graphicx}
-\usepackage{comment}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{listings}
-% declare the path(s) where your graphic files are
-\graphicspath{{../figures/}}
-% and their extensions so you won't have to specify these with
-% every instance of \includegraphics
-\DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-
-% Minted is used for source code highlighting
-%\usepackage[outputdir=/Volumes/ramdisk/]{minted}
-% \usepackage{minted}
-% *** MATH PACKAGES ***
-\usepackage{amsmath}
-
-% *** SPECIALIZED LIST PACKAGES ***
-%\usepackage{algorithm2e}
-\usepackage{algorithmic}
-
-% *** ALIGNMENT PACKAGES ***
-\usepackage{array}
-
-\usepackage{url}
-
-\usepackage[utf8]{inputenc}
-\usepackage[american]{babel}
-%\usepackage[backend=biber,style=ACM-Reference-Format]{biblatex}
-%\addbibresource{./library.bib}
-%\addbibresource{./cryptobib/abbrev2.bib}
-%\addbibresource{./cryptobib/crypto_crossref.bib}
-
-% Used for Theorems and Definitions
-\usepackage{amsthm}
-\newtheorem{definition}{Definition}
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}{Lemma}
-\theoremstyle{definition}
-\newtheorem{remark}{Remark}
-\newtheorem{assumption}{Assumption}
-
-\usepackage{xcolor}
-\usepackage{xspace}
-\usepackage{xstring} % required by IfEqCase
-
-\usepackage{booktabs} % better tables
-\usepackage{algorithm}
-\usepackage{algorithmic}
-
-% for references, use it like this: \cref{some_label}
-% it will automatically put Appendix, Figure, Section, etc before the reference
-\usepackage[capitalise]{cleveref}
-
-% ishaq: adding our macros
-\input{macros.tex}
-
-\sloppy
-\begin{document}
-\fancyhead{}
-% do not delete this code.
-
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
-\title[Compilation and Backend-Independent Vectorization for Multi-Party Computation]{Compilation and Backend-Independent Vectorization for Multi-Party Computation}
-
-%
-% The "author" command and its associated commands are used to define the authors and their affiliations.
-% Of note is the shared affiliation of the first two authors, and the "authornote" and "authornotemark" commands
-% used to denote shared contribution to the research.
-
-\author{Benjamin Levy}
-\email{levyb3@rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Benjamin Sherman}
-\email{shermb@rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Muhammad Ishaq}
-%\authornote{This work was done in part while the author was at RPI.}
-\email{ishaqm@purdue.edu}
-%\orcid{1234-5678-9012}
-%\author{G.K.M. Tobin}
-%\authornotemark[1]
-%\email{webmaster@marysville-ohio.com}
-\affiliation{%
-    \institution{Purdue University}
-    %\streetaddress{P.O. Box 1212}
-    \city{West Lafayette}
-    \state{Indiana}
-%    \postcode{43017-6221}
-}
-
-\author{Lindsey Kennard}
-\email{fire.elemental@gmail.com}
-\affiliation{%
-    \institution{STR}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Boston}
-    \state{Massachusets}
-    %\country{Iceland}
-}
-
-\author{Ana Milanova}
-\email{milanova@cs.rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Vassilis Zikas}
-%\authornote{This work was done in part while the author was visiting UCLA and supported in part by DARPA and SPAWAR under contract N66001-15-C-4065 and by a SICSA Cyber Nexus Research Exchanges grant.}
-\email{vzikas@purdue.edu}
-%\orcid{1234-5678-9012}
-%\author{G.K.M. Tobin}
-%\authornotemark[1]
-%\email{webmaster@marysville-ohio.com}
-\affiliation{%
-    \institution{Purdue University}
-    %\streetaddress{P.O. Box 1212}
-    \city{West Lafayette}
-    \state{Indiana}
-    %    \postcode{43017-6221}
-}
-
-%
-% By default, the full list of authors will be used in the page headers. Often, this list is too long, and will overlap
-% other information printed in the page headers. This command allows the author to define a more concise list
-% of authors' names for this purpose.
-%\renewcommand{\shortauthors}{Trovato and Tobin, et al.}
-
-\renewcommand{\A}{{\sf A}}
-\renewcommand{\B}{{\sf B}}
-\renewcommand{\C}{{\sf C}}
-\renewcommand{\D}{{\sf D}}
-
-% As a general rule, do not put math, special symbols or citations
-% in the abstract
-\begin{abstract}
-
-\ana{We need an abstract.}
-
-\end{abstract}
-
-%
-% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-% Please copy and paste the code instead of the example below.
-%
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10003752.10010124.10010138.10010143</concept_id>
-<concept_desc>Theory of computation~Program analysis</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10003752.10003777.10003789</concept_id>
-<concept_desc>Theory of computation~Cryptographic protocols</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10002978.10002979</concept_id>
-<concept_desc>Security and privacy~Cryptography</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Theory of computation~Program analysis}
-\ccsdesc[500]{Theory of computation~Cryptographic protocols}
-\ccsdesc[300]{Security and privacy~Cryptography}
-
-%
-% Keywords. The author(s) should pick words that accurately describe the work being
-% presented. Separate the keywords with commas.
-\keywords{multiparty computation; compilers; cryptography}
-
-\begin{comment}
-%
-% A "teaser" image appears between the author and affiliation information and the body
-% of the document, and typically spans the page.
-\begin{teaserfigure}
-    \includegraphics[width=\textwidth]{sampleteaser}
-    \caption{Seattle Mariners at Spring Training, 2010.}
-    \Description{Enjoying the baseball game from the third-base seats. Ichiro Suzuki preparing to bat.}
-    \label{fig:teaser}
-\end{teaserfigure}
-\end{comment}
-%
-% This command processes the author and affiliation and title information and builds
-% the first part of the formatted document.
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-
-Multi-party computation (MPC) allows $N$ parties $p_1,\dots, p_N$ to perform a computation on their private inputs securely.
-Informally, security means that the secure computation protocol computes the correct output (correctness) and it does not leak any information about the individual party inputs, other than what can be deduced from the output (privacy).
-
-% The first results from the early 80's~\cite{FOCS:Yao82b,STOC:GolMicWig87,STOC:BenGolWig88,STOC:ChaCreDam88} introduced the problem and formally defined its security, and investigate feasibility in different models---e.g., passive or malicious adversaries, unconditional or computational security, with or without a broadcast channel.
-This spawned a rich theoretical literature of beautiful works improving and extending in various manners the original ideas.
-
-MPC theory dates back to the early 1980-ies~\cite{FOCS:Yao82b,STOC:GolMicWig87,STOC:BenGolWig88,STOC:ChaCreDam88}. 
-Long the realm of theoretical cryptography, MPC has seen significant advances in programming technology in recent years. 
-These advances bring MPC closer to practice and wider applicability --- 
-MPC technology has been employed in real-world scenarios such as auctions~\cite{FC:BCDGJK09}, biometric identification~\cite{ESORICS:BlaGas11}, 
-and privacy-preserving machine learning~\cite{SP:MohZha17,CCS:MohRin18}. 
-The goal is to bring the technology to a level where programmers can write \emph{secure} and \emph{efficient} programs without 
-commanding extensive knowledge of cryptographic primitives.
-
-The problem, therefore, is to build a high-level programming language and a compiler, and there has been significant advance in 
-this space, e.g., ~\cite{CCS:BenNisPin08,ESORICS:BogLauWil08,CCS:ZhaSteBla13,Hastings:2019,Keller:2020,Acay:2021,Braun:2022} among other work. 
-Current research largely falls at the two ends 
-of the classical compiler: (1) work on \emph{front-end} language design and (2) work on \emph{back-end} protocol implementation. Work on language design 
-focuses on high-level constructs necessary to express multiple parties, computation by different parties, and information flow from one party 
-to another~\cite{SP:RasHamHic14, Acay:2021}. On the other end, work on protocol implementation focuses on cryptographic foundations and their efficient circuit-level
-implementation~\cite{NDSS:DemSchZoh15, CCS:ABFKLO18, Braun:2022}, e.g., implementation of operations (e.g., MUL, ADD) using different sharing protocols 
-(Boolean or Arithmetic GMW~\cite{STOC:GolMicWig87} or Yao's garbled circuits~\cite{FOCS:Yao82b}), as well as efficient share conversion from one representation 
-to another.
-
-Earlier compilers did both back-end and front-end translation without a specific focus on either, as their aim was to demonstrate applicability of MPC on real-world
-programming problems. As the field advanced, works have focused more closely on front-end language design (e.g., Wysteria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021}) 
-or back-end ``circuit-level'' design and implementation (e.g., MOTION~\cite{Braun:2022}).
-
-In this work we focus on an intermediate language and what we call \emph{backend-independent optimizations}, in a close analogy to \emph{machine-independent} optimizations
-in the classical compiler. The following figure summarizes our key idea:
-
-{\begin{center}
-\includegraphics[width=0.8\linewidth]{figs_paper_SIMD/focus.png}
-\end{center}
-}
-
-We emphasize the MPC Source intermediate representation and optimization over MPC Source. As in classical compilers, we envision different front ends (e.g., Wysteria,
-our front end IMP-MPC) compiling into MPC Source. MPC Source is particularly suitable for optimizations such as protocol mixing~\cite{Ishaq:2019, Fang:2022} and SIMD-vectorization, 
-which takes advantage of amortization at the circuit level. The MPC Source IR exposes the linear structure of MPC programs, which simplifies program analysis; this is
-in contrast to source, which has if-then-else constructs. In the same time, MPC Source is sufficiently ``high-level''  to support analyses and optimizations that take into account 
-control and data flow in a specific program. MPC Source is small in size and analysis is tractable, as opposed to analysis over an unrolled circuit~\cite{Ishaq:2019}. Again as
-in classical compilers, we envision translation of MPC Source (optimized or unoptimized) into MOTION, SPDZ, or other back-end code.
-
-%In contrast, the high-level program contains if-then-else and array writes, or the low-level circuit . 
-%The compilation problem, i.e., translation from a high-level language to circuits, is difficult and therefore, breaking the problem into sub-problems can  tractable.
-
-\subsection{Our Contribution} In this paper, we develop a compiler framework that takes a Python-like routine and produces MOTION code: we describe (a) the IMP-MPC language, its syntax and semantic restrictions, (b) translation from IMP-MPC into MPC Source, (c) a specific backend-independent optimization: novel SIMD-vectorization on MPC Source, and (d) translation from MPC Source into MOTION code.
-%\ana{I think we have to add more on MOTION, specifically mixing and amortization.}
-
-We focus on the MOTION framework as our back-end for several reasons.
-First, it is the state of the art in terms of performance~\cite{Braun:2022}.
-Second, it provides an API over efficient implementation for a wide variety of cryptographic operations in three different protocols ---  Arithmetic GMW, Boolean GMW, and BMR --- which allows for protocol mixing~\cite{Ishaq:2019, Fang:2022}, a known backend-independent optimization.
-Third, MOTION provides API for SIMD-level operations, which amortize cost and lead to significant improvement in memory footprint and
-throughput~\cite{NDSS:DemSchZoh15, CCS:ABFKLO18, Braun:2022}. It also enables MPC Source-level optimizations such SIMD vectorization.
-
-Our second contribution is an analytical model for cost estimation of amortized schedules. Originally, we hoped that optimal scheduling (under our model) was tractable, as the problem
-appeared simpler than the classical scheduling problem. Unfortunately, we show that optimal scheduling is NP-hard via a reduction to the Shortest Common Supersequence (SCS) problem. 
-Cost estimation is important as it drives not only vectorization but other optimizations such as protocol mixing optimizations and scheduling as well~\cite{Ishaq:2019, Fang:2022}.
-
-Our most important contribution is the implementation and evaluation of the compiler framework.
-We demonstrate expressivity of the source language by running the compiler on \ana{X} programs; these include classical MPC benchmarks such as PSI and Biometric matching, as well as kMeans, Histogram, and others.
-Vectorization leads to \ana{X}x improvement in number of gates on average, \ana{Y}x speedup in circuit generation time, \ana{Z}x speedup in the LAN setting, and \ana{Z}x speedup in the WAN setting using the Boolean GMW protocol.
-Results form BMR are similar.
-
-Our results emphasize the importance of backend-independent optimizations --- vectorization (described in this work) and protocol mixing (tackled in previous works~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022}) are two optimizations readily available at the level of MPC Source.
-We believe that our work can lead to future work on backend-independent compilation and optimization, bringing in new MPC optimizations and combinations of optimizations, much in the vein of standard compilers, and bring MPC programming technology closer to practice and wider applicability.
-
-\subsection{Outline}
-
-The rest of the paper is organized as follows.
-\secref{sec:overview} presents an overview of the compiler.
-\secref{sec:model} describes our model for cost estimation and argues NP-hardness of optimal scheduling and \secref{sec:compiler} describes the compiler.
-\secref{sec:experiments} presents implementation and evaluation.
-\secref{sec:related} discusses related work and \secref{sec:conclusion} concludes.
-
-\begin{comment}
-\begin{itemize}
-\item We define the scheduling problem for MPC. 
-We present an analytical model to reason about cost of schedules
-and show that scheduling is NP-hard via a reduction to the shortest 
-common supersequence problem.
-
-\item We present a compiler that takes an IMP-like high-level 
-program and produces amortized (i.e., vectorized) low-level cryptographic 
-code in the MOTION framework. Central contributions are 1) a novel compiler framework, 
-2) a vectorization algorithm that produces optimal schedules for a large 
-number of MPC programs, and 3) reasoning over output arrays without Array SSA; we remove 
-infeasible loop-carried dependences introduced by standard SSA to improve vectorization.
-
-\item We present an implementation and evaluation in the MOTION framework. 
-\ana{Fill in with final results. Mention benchmarks (standard + new ones + HyCC). }
-
-\end{itemize}
-\end{comment}
-
 \section{Overview}
@@
 
-\begin{comment}
-{\small
-\begin{verbatim}
-def biometric(C: shared[list[int]], D: int, 
-              S: shared[list[int]], N: int) -> 
-                      tuple[shared[int], shared[int]]:
-    min_sum = 10000
-    min_index = 0  
-    for i in range(N): #loop over database
-        sum = 0
-        for j in range(D): #loop over features
-            d = S[i * D + j] - C[j] #i.e., d = S[i,j] - C[j]
-            p = d * d
-            sum = sum + p
-        if sum < min_sum:
-            min_sum = sum
-            min_index = i
-    return (min_sum, min_index)
-\end{verbatim}
-}
-\end{comment}
-
 Our compiler takes essentially standard IMP~\ana{Need citation here.}
@@
 MUX operations, while the $\phi$-nodes at for loops turn into what we call pseudo PHI nodes with a straightforward semantics.
-
-\begin{comment}
-{\small
-\begin{pythonn}
-1. min_sum!1 = 10000
-2. min_index!1 = 0
-3. for i in range(0, N!0):
-4.    min_sum!2 = PHI(min_sum!1, min_sum!4)
-5.    min_index!2 = PHI(min_index!1, min_index!4)
-6.    sum!2 = 0
-7.    for j in range(0, D!0):
-8.      sum!3 = PHI(sum!2, sum!4)
-9.      d!3 = (S!0[((i * D!0) + j)] - C!0[j]) // MPC
-10.     p!3 = (d!3 * d!3) // MPC
-11.     sum!4 = (sum!3 + p!3) // MPC
-12.  !1!2 = (sum!3 < min_sum!2) // MPC
-13.  min_sum!3 = sum!3
-14.  min_index!3 = i
-15.  min_sum!4 = MUX(!1!2, min_sum!3, min_sum!2) // MPC
-16.  min_index!4 = MUX(!1!2, min_index!3, min_index!2) // MPC
-17.!2!1 = (min_sum!2, min_index!2)   
-\end{pythonn}
-}
-\end{comment}
-
-%The compiler linearizes the source turning conditionals into MUX statements. The PHI nodes
-%are remnants of the SSA IR; the compiler generates code that picks the correct value 
-%when producing MOTION output; the MOTION framework in turn linearizes 
-%loops when it generates the circuit.
 
@@
 
-\begin{comment}
-{\small
-\begin{verbatim}
-min_sum!1 = 10000
-min_index!1 = 0
-// S!0^ is same as S!0. C!0^ replicates C!0 N-times: 
-S!0^ = raise_dim(S!0, ((i * D!0) + j), (i:N!0,j:D!0))
-C!0^ = raise_dim(C!0, j, (i:N!0,j:D!0))
-
-sum!2 = [0,..,0]
-// Computes all differences and all products "at once"
-d!3[I,J] = SUB_SIMD(S!0^[I,J],C!0^[I,J])
-p!3[I,J] = MUL_SIMD(d!3[I,J],d!3[I,J])
-
-for j in range(0, D!0):
-  // sum!2[I], sum!3[I], sum!4[I] are vectors of size N  
-  // Computes N intermediate sums "at once"
-  sum!3[I] = PHI(sum!2[I], sum!4[I])       
-  sum!4[I] = ADD_SIMD(sum!3[I],p!3[I,j])
-
-min_index!3 = [0,1,...N!0-1]
-
-for i in range(0, N!0):
-  min_sum!2 = PHI(min_sum!1, min_sum!4) 
-  !1!2[i] = CMP(sum!3[i],min_sum!2)
-  min_sum!4 = MUX(!1!2[i], sum!3[i], min_sum!2)
-
-for i in range(0, N!0):
-  min_index!2 = PHI(min_index!1, min_index!4)  
-  min_index!4 = MUX(!1!2[i], min_index!3[i], min_index!2)
-!2!1 = (min_sum!2, min_index!2)   
-\end{verbatim}
-}
-\end{comment}
-
 In MPC backends, executing $n$ operations ``at once'' in a single SIMD operation costs a lot less than executing those $n$ operations one by one.
@@
 \ana{Edit numbers with final experiments.}
-
-\section{Analytical Model}
-\label{sec:model}
-
-\ana{We need an intro to section here.}
-
-\subsection{Scheduling in MPC}
-\label{sec:mpc}
-
-%\ana{Here we need more on MPC before jumping to scheduling. MPC-source, basic assumptions about static loop bounds, etc.}
-
-For this treatment we make the following simplifying assumptions:
-
-\begin{enumerate}
-\item All statements in the program execute using the same protocol (sharing). That is, there is no share conversion. 
-%\ishaq{This is not an assumption, this is our setting. We work in the single protocol world.}
-\item There are two tiers of MPC instructions, local and remote. A local instruction (essentially just ADD) has cost $\beta$ 
-and a remote instruction (e.g., MUX, MUL, SHL, etc.) has cost $\alpha$, where $\alpha >> \beta$. We assume that all remote 
-instructions have the same cost.
-%\ishaq{Don't we need to distinguish between local (cheap) vs.
-interactive (expensive) instructions?
-I don't see how we could assume this.
--- Past suggestion from Vassilis: Ideally we should assume cost to be a convex function.}\ana{After discussion 1/12: replace single costs 1 with two levels of costs: $\alpha$ for non-local operations, e.g., MUL, and $\beta$ for local ops, e.g., ADD.}
-\item  In MPC frameworks, executing $n$ operations ``at once'' in a single SIMD operation costs a lot less than executing those $n$ operations one by one.
-Following Amdahl's law~\ana{citation!}, we write $\alpha = \frac{1}{s}p\alpha + (1-p)\alpha$, where $p$ is the fraction of execution time that benefits from amortization and $(1-p)$ 
-is the fraction that does not, and $s$ is the available resource. Thus, $n\alpha = \frac{n}{s}p\alpha + n(1-p)\alpha$. 
-For the purpose of the model we assume that $s$ is large enough and the term $\frac{n}{s}p\alpha$ amounts to a \emph{fixed cost} incurred regardless of
-whether $n$ is $10,000$ or just $1$. (This models the cost of sending a packet from party A to party B for example.) The bottom line is that amortized execution 
-of $n$ operations is $f(n) = \alpha_\mathit{fix} + n\alpha_{var}$ in contrast to unamortized execution $g(n) = n\alpha_\mathit{fix} + n\alpha_{var}$ and, of course,
-$\alpha_\mathit{fix} << n\alpha_\mathit{fix}$.
-%We assume infinite parallel capacity---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely $\alpha$ or $\beta$. 
-%This is a standard assumption in Cryptographic Parallel RAM. ABY presents empirical support for this assumption~\ana{Add citations. PRAM, ABY.}
-%\ishaq{Comment from Vassilis: We assume there are infinite parallel capacities, this is the standard assumption in PRAM (Cryptographic Parallel RAM).}\ana{PRAM assumption stays.
-We have to replace the unit cost of 1 with a suitable amortization function $f(n)$, which I don't think changes much in the cost modeling and analysis.}
-\item MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption,
-2 MUL instructions can be amortized in a single SIMD instruction that costs $\alpha_\mathit{fix} + 2\alpha_\mathit{var}$, however a MUL and a MUX instruction 
-still cost $2\alpha_\mathit{fix} + 2\alpha_\mathit{var}$ even when scheduled ``in parallel''. 
-%MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption,
-%2 MUL instructions scheduled in parallel benefit from amortization and cost $\alpha$, however a MUL and a MUX instructions scheduled
-%in parallel still cost $2\alpha$. 
-%\ishaq{We need to reword this assumption to something like "$K$ parallel MUL costs much less than $K$ (because they get amortized), but any mix of $K$  MUL and MUX still cost roughly $K$.
-Specifically, we should take away the low constants (2 and 1) because we know for low constants this is not true.} \ana{Again, core assumption that MUL and MUX don't benefit from amortization stays.
-We have to replace constant costs with functions, as in (3).}
-\end{enumerate}
-
-\input{prob_stmt_np_hardness.tex}
-
-\begin{figure*}
-  \includegraphics[width=1.0\linewidth]{figs_paper_SIMD/compiler_framework.png}
-  \caption{Compiler Framework. \ana{Change figure since we are not doing div-and-conquer.}}
-  \label{fig:compiler_framework}
-\end{figure*}
-
-\section{\bf Compiler Framework}
-\label{sec:preliminaries}
-
-%We assume arbitrarily nested loops in the MPC-source IR and read-only arrays.
-We assume that loops range from 0 to some constant N.
-Arrays are linearized (row-major order as in MOTION) and accesses are via functions of the induction variables of the enclosing loops.
-
-\figref{fig:compiler_framework} presents an overview of our compiler. This section describes the front-end phases of the compiler
-and Section~\secref{sec:vectorization} describes our vectorization algorithm. \ana{We'll need a section on MOTION code gen too.}
-%\ana{Need to state this more precisely.}
-
-\subsection{Syntax and Semantic Restrictions}
-\label{sec:syntax}
-
-Source syntax is essentially standard IMP syntax as shown below:
-%\begin{figure}[t]
-{\small
-\[
-\begin{array}{l@{~~~~~}l}
-  \begin{array}{l@{~}l@{~~~}l}
- e & ::= e \; \mathit{op} \; e \mid e \mid \x \mid  \texttt{const} \mid \texttt{A[$e$]} & \mathit{expression} \\
- % s &::= t \;\f & \mathit{field} \\
- %\md &::= t\;{\m}(t \;{\this}, \; t\;\x) \; \{\; \ol{t\;\y}\;\s;\;\return\;\y \; \} 
- %                 & \mathit{method} \\ 
-  s & ::= s;s \mid  & \mathit{sequence} \\
-  & \x = e \mid \texttt{A[$e$]} = e \mid & \mathit{assignment \; stmt} \\
-  & \texttt{for $i$ in range($I$)}: s \mid & \mathit{for \; stmt} \\
-  & \texttt{if $e$: $s$ else: $s$} & \mathit{if \; stmt} \\ 
-%   t &::= q \; {\sf C} & \mathit{qualified \; type}\\
-%   q &::= \high \mid \poly \mid \low & \mathit{FlowCFL \; qualifier}\\
-%    t' &::= q' \; \C & \mathit{field \; qualified \; type}\\
-%   q' &::= \tainted \mid \poly & \mathit{field \; qualifier}\\
-
-  \end{array}
-\end{array}
-\]
-}
-%\end{figure}
-
-The syntax allows for array accesses, arbitrarily nested loops, and if-then-else control flow. 
-Expressions are typed $q \; \tau$, where qualifier $q$ and base type $\tau$ are:
-
-{\small
-\[
-\begin{array}{l@{~~~~~}l}
-  \begin{array}{l@{~}l@{~~~}l}
- \tau & ::= \texttt{int}  \mid \texttt{bool} \mid \texttt{list[int]} \mid  \texttt{list[bool]} & \mathit{base \; types} \\
-  q & ::= \texttt{shared} \mid \texttt{plain} & \mathit{qualifiers} \\
-  \end{array}
-\end{array}
-\]
-}
-
-The type system is mostly standard, and in our experience, a sweet spot between readability and expressivity.
-The \texttt{shared} qualifier denotes shared values, i.e., ones shared among the parties and computed upon 
-under secure computation protocols; the \texttt{plain} qualifier denotes plaintext values. Shared lists denote 
-shared elements, but plaintext length. Subtyping is again as expected: $q_1 \; \tau <: q_2 \; \tau$ 
-iff $q_1 <: q_2$ where $\texttt{plain} <: \texttt{shared}$. Thus, we can convert a plaintext value into a shared one, 
-but not vice versa.
-
-Our compiler imposes certain semantic restrictions that it enforces throughout the various 
-phases of compilation. We note that in some cases, the restrictions 
-can be easily lifted and we plan to do so in future iterations of the work.
-
-\begin{enumerate}
-\item Loops are of the form $0 <= i < I$ and bounds are fixed at compile time. 
-It is a standard restriction in MPC that the bounds must be known at circuit-generation time.
-\item Arrays are one-dimensional. N-dimensional arrays are linearized and accessed 
-in row-major order and at this point the programmer is responsible for linearization
-and access. (This restriction can be easily lifted.)
-\item Array subscrpts are plaintext values as specified by the rule:
-\begin{semantics}
-    \ntyperule{Array Access}{
-%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
-  \Gamma \vdash e : \texttt{plain int} \quad \Gamma \vdash \texttt{A} : q \; \texttt{list}[\tau] \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
-}{
-  \Gamma \vdash \texttt{A[$e$]} : q \; \tau
-}
-\end{semantics}
-The subscript $e$ is a function of the indices of the enclosing loops. 
-For read access, the compiler allows an arbitrary such function. 
-However, it restricts write access to \emph{canonical writes}, i.e., \texttt{A[i,j,k]} = ...
-where \texttt{i}, \texttt{j} and \texttt{k} loop over the three dimensions of \texttt{A}.  
-Write access such as for example \texttt{A[i,j+2]} = ... is not allowed. 
-%\item \ana{Add rules for logical ops}
-%\item \ana{Add rules for arithmetic ops}
-\item The final restriction involves MUX as expressed by the rule: 
-\begin{semantics}
-    \ntyperule{MUX}{
-%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
-  \Gamma \vdash e_1 : q_1 \; \texttt{bool} \quad \Gamma \vdash e_2 : q_2 \; \tau \quad \Gamma \vdash e_2 : q_2 \; \tau \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
-}{
-  \Gamma \vdash \texttt{MUX($e_1$,$e_2$,$e_3$)} : q_1 \vee q_2 \vee q_3 \; \tau
-}
-\end{semantics}
-\end{enumerate}
-Importantly, the arguments of MUX are restricted to base types, as passing array arguments 
-is not valid MOTION MPC. \ana{Check this.} This causes a slight inconvenience as we could not write 
-\begin{pythonn}
-if e: A[i] = val
-\end{pythonn}
-Instead we had to write
-\begin{pythonn}
-if not(e): val = A[i]
-A[i] = val
-\end{pythonn}
-
-We note that while the programmer writes 
-annotations at the level of IMP Source (recall \ana{ref to overview section}), the 
-annotations propagate through the transformations and are checked at the 
-level of MPC Source.
-
-%\ana{Disadvantage of MUX}
-
-For the rest of this section we write $i,j,k$ to denote the loop nest: $i$ is the outermost loop, $j$, is immediately nested in $i$, and so on until $k$
-and we use $I,J,K$ to denote the corresponding upper bounds. We write $A[i,j,k]$ to denote canonical access 
-to an array element. In the program, canonical access is achieved via the standard row-major order formula: $(J*K)*i + K*j + k$.
-To simplify the presentation we describe our algorithms in terms of three-element tuples $i,j,k$, however, discussion easily generalizes to 
-arbitrarily large loop nests.
-
-\subsection{From IMP Source to SSA}
-
-Our compiler translates from Source to SSA as follows:
-
-\begin{enumerate}
-  \item
-        \textbf{Parsing}:
-        Use Python's \texttt{ast} module to parse the input source code to a Python AST.
-  \item
-        \textbf{Syntax checking}:
-        Ensure that the AST matches the restricted subset defined in Section~\secref{sec:syntax}.
-        This step outputs an instance of the \texttt{restricted\_ast.Function} class, which represents our restricted subset of the Python AST.
-  \item
-        \textbf{3-address CFG conversion}:
-        %\benjamin{TODO: Is this a good amount of detail?} \ana{This is good amount of detail for the Tech report. In the actual paper we'll shorten a lot and add pointers to the Tech report.}
-
-        Convert the restricted-syntax AST to a three-address control-flow graph.
-        To do this, first,
-        add an empty basic block to the CFG and mark it as current.
-        Next, for each statement in the restricted AST's function body,
-        process the statement.
-        Statements can either be for-loops, if-statements, or assignments (as in~\secref{sec:syntax}).
-        Rules for processing each kind of statement are given below:
-        \begin{enumerate}
-          \item
-                \textbf{For-loops}:
-                Create new basic blocks for
-                the loop condition
-                (the \emph{condition-block}),
-                the loop body
-                (the \emph{body-block}),
-                and the code after the loop
-                (the \emph{after-block}).
-                Insert a jump from the end of the current block to the condition-block.
-                Then, mark the condition-block as the current block.
-                Insert a for-instruction at the end of the current block with the loop counter variable and bounds from the AST.
-                Next, add an edge from the current block to the after-block labeled ``FALSE'' and
-                an edge from the current block to the
-                body-block labeled ``TRUE''.
-                Then, set the body-block to be the current block
-                and process all statements in the AST's loop body.
-                Finally, insert a jump to the condition-block and set the after-block as current.
-          \item
-                \textbf{If-statements}:
-                Create new basic blocks for
-                the ``then'' statements of the if-statement
-                (the \emph{then-block}),
-                the ``else'' statements of the if-statement
-                (the \emph{else-block}),
-                and the code after the if-statement
-                (the \emph{after-block}).
-                At the end of the current block,
-                insert a conditional jump to the then-block or else-block
-                depending on the if-statement condition in the AST.
-                Next,
-                mark the then-block as current,
-                process all then-statements,
-                and add a jump to the after-block.
-                Similarly,
-                mark the else-block as current,
-                process all else-statements,
-                and add a jump to the after-block.
-                Finally, set the after-block to be the current block,
-                and give it a \emph{merge condition} property equal to the condition of the if-statement.
-          \item
-                \textbf{Assignments}:
-                In the restricted-syntax AST,
-                the left-hand side of assignments
-                can be a variable or an array subscript.
-                If it is an array subscript, e.g., \texttt{A[i] = x},
-                change the statement to \texttt{A = Update(A, i, x)}.
-                If the statement is not already three-address code,
-                for each sub-expression in the right-hand side of the assignment,
-                insert an assignment to a temporary variable.
-        \end{enumerate}
-  \item
-        \textbf{SSA conversion}:
-        Convert the 3-address CFG to SSA with Cytron's algorithm.
-\end{enumerate}
-
-\subsection{From SSA to MPC Source} %: MUX Nodes and Pseudo $\phi$-nodes}
-
-Once the compiler converts the code to SSA,
-it transforms $\phi$-nodes that correspond to if-statements into MUX nodes.
-From the 3-address CFG conversion step,
-$\phi$-nodes corresponding to if-statements will be in a basic block
-with the merge condition property.
-For example, if $X_3 = \phi(X_1,X_2)$ is in a block with merge condition $C$,
-the compiler transforms it into $X_3 = \mathit{MUX}(C, X_1, X_2)$.
-Next, the compiler runs the dead code elimination algorithm from Cytron's SSA paper.
-
-Next, the control-flow graph is \emph{linearized} into MPC Source,
-which has loops but no if-then-else-statements.
-This means that both branches of all if-statements are executed,
-and the MUX nodes determine whether to use results from the then-block or from the else-block.
-The compiler linearizes the control-flow graph with a variation of breadth-first search.
-Blocks with the ``merge condition'' property are only considered
-the second time they are visited,
-since that will be after both branches of the if-statement are visited. 
-(The Python AST naturally gives rise to a translation where each conditional has exactly two targets, 
-and each ``merge condition'' block has exactly two incoming edges, a TRUE and a FALSE edge.
-Thus, each $\phi$-node has exactly two multiplexer arguments, which dovetails into MUX.
-This is in contrast with Cytron's algorithm which operates at the level of the CFG and allows for 
-$\phi$-nodes with multiple arguments.)
-Each time the compiler visits a block,
-it adds the block's statements to the MPC source.
-If the block ends in a for-instruction,
-the compiler recursively converts the body and code after the loop to MPC source
-and adds the for-loop and code after the loop to the main MPC source.
-If the block does not end in a for-instruction,
-the compiler recursively converts all successor branches to MPC source and
-appends these to the main MPC source.
-
-Now, the remaining $\phi$-nodes in MPC source are the loop header nodes. We call these nodes \emph{pseudo} $\phi$-nodes 
-and we write \python{PHI} in MPC Source. A pseudo $\phi$-node $X_1 = $\python{PHI}$(X_0,X_2)$ in a loop header is evaluated 
-during circuit generation. If it is the 0-th iteration, then the $\phi$-node evaluates to $X_0$, otherwise, it evaluates to $X_2$.
-
-%\ana{We need to double check that the following captures all cases of def-use edges.}
-
-\subsection{Dependence Analysis}
-
-The next steps construct the dependence graph and \ana{Add an intro here.}
-
-\subsubsection{Def-use Edges}
-
-The dependence graph has the following def-use edges:
-
-\begin{itemize}
-\item same-level edge $X\rightarrow Y$ where $X$ and $Y$ are in the same loop nest, say $i,j,k$.
-E.g., the def-use edge 9 to 10 in the Biometric MPC Source in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source} is a same-level edge.
-A same-level edge can be a back-edge in which case a $\phi$ node is the target of the edge.
-E.g., 15 to 4 in Biometric is a same-level back-edge.
-\item outer-to-inner $X\rightarrow Y$ where $X$ is in an outer loop nest, say $i$, and $Y$ is in an inner one, say $i,j,k$. E.g., 1 to 4 in Biometric forms is an outer-to-inner edge.
-%A $\phi$-node can be a source or a target of an outer-to-inner forward edge.
-\item inner-to-outer $X\rightarrow Y$ where $X$ is a $phi$-node in an inner loop nest, $i,j,k$, and $Y$ is in the enclosing loop nest $i,j$.
-E.g., the def-use from 8 to 12 gives rise to an inner-to-outer edge.
-%E.g. $\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$ is an inner-to-outer edge. 
-An inner-to-outer edge can be a back-edge as well, in which case both $X$ and $Y$ are phi-nodes with the source $X$ in a loop nested into $Y$'s loop (not necessarily immediately).
-
-%Note that the source is \emph{always} a $\phi$-node of the $i,j,k$ loop and the target is in the immediately enclosing loop.
-The interpretation of this edge is that the use node $Y$ uses the definition made in the last iteration of the inner loop.
-%\ben{The current representation of pseudo $\phi$-nodes shows them attached to the loop header (i.e. in the loop).  We may want to clarify that they are also evaluated at the loop's termination.}
-%\item same-level back-edge $X\rightarrow Y$.
-$Y$ is a phi-node in the header of the loop and $X$ is a definition of the variable in the loop body.
-E.g., $\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{sum}_1,\texttt{min}_1)$ to $\texttt{min}_0 = \phi(\texttt{min}_1, 10000)$ in Biometric is a same-level back-edge.
-%\item inner-to-outer back-edge $X\rightarrow Y$: $X$ and $Y$ are both $\phi$-nodes for some variable. The source $X$ is in a loop nested into $Y$'s loop (not necessarily immediately).
-
-\item mixed forward edge $X\rightarrow Y$.
-$X$ is in some loop $i,j,k$ and $Y$ is in a loop nested into $i,j,k'$.
-We transform mixed forward edges as follows.
-Let $\x$ be the variable defined at $X$.
-We add a variable and assignment $\x' = \x$ immediately after the $i,j,k$ loop.
-Then we replace the use of $\x$ at $Y$ with $\x'$.
-This transforms a mixed forward edge into an "inner-to-outer" forward edge followed by an outer-to-inner forward edge.
-Thus, Basic Vectorization handles one of "same-level", "inner-to-outer", or "outer-to-inner" def-use edges.
- \end{itemize}
-
-\subsubsection{Closures}
-
-We define $\mathit{closure}(n)$ where $n$ is a phi-node. Intuitively, it computes the set of nodes (i.e., statements) that form a dependence cycle with $n$. The closure of $n$ is defined as follows:
-\begin{itemize}
-\item $n$ is in $\mathit{closure}(n)$
-\item $X$ is in $\mathit{closure}(n)$ if there is a same-level path from $n$ to $X$, and $X \rightarrow n$ is a same-level back-edge.
-\item $Y$ is in $\mathit{closure}(n)$ if there is a same-level path from $n$ to $Y$ and there is a same-level path from $Y$ to some $X$ in $\mathit{closure}(n)$.
-\end{itemize}
-
-\subsection{Taint Analysis}
-
-We require that all inputs are marked as either shared or plaintext, however, we do not require any other qualifier annotations in the program.
-Then we infer if intermediate variables and expressions are shared through taint analysis with "taintedness" referring to the shared attribute.
-Specifically, the compiler follows the following rules, which are standard in positive-negative qualifier systems (here \texttt{shared} is the positive qualifier and \texttt{plain} is the negative one):
-\begin{enumerate}
-\item Loop counters and array upper bounds are always \texttt{plain}.
-\item If any variable on the right-hand side of an assignment is shared, then the assigned variable is \texttt{shared}.
-%\item If all variables on the right-hand side of an assignment are plaintext, then the assigned variable is plaintext \ana{Will be covered by the last rule.}
-\item Any variables that cannot be determined as shared via the above rules are \texttt{plain}.
-\end{enumerate}
-
-%The first two rules are standard for taint analysis, and the third rule follows from the MPC problem statement. \ben{Is the explanation for the third rule correct?}  \ana{Yes, that is exactly it!}
-The final rule is needed in examples like the following. In the below snippet \texttt{sum!2} and \texttt{sum!3} form a dependency cycle and there is no \texttt{shared} value that flows to either one. 
-They are inferred as plaintext. 
-%cannot be marked as plaintext through simple taint analysis:
-{\small
-\begin{verbatim}
-plaintext_array = [0, 1, 2, ...]
-sum!1 = 0
-for i in range(0, N):
-    sum!2 = PHI(sum!1, sum!3)
-    sum!3 = sum!2 + plaintext_array[i]
-\end{verbatim}
-}
-%\ben{I think the above example is unnecessary and could be replaced by an explanation of how "untainted" variables are implicitly plaintext, but I couldn't think of a way to phrase that.}\ana{I left the example, we can leave it in the TR, remove from paper.}
-
-When converting to MOTION code, any plaintext value used in the right-hand side of a shared assignment is converted to a shared value for that expression. 
-%\ben{Is this necessary to include?} \ana{Yes, that's important.}
-
-Importantly, the taint analysis works on MPC Source, which lacks if-then-else control flow. 
-This significantly simplifies the treatment as there is no need to handle conditionals and implicit flow.
-
-\subsection{From (Optimized) MPC Source to MOTION}
-\label{sec:MOTION}
-\ana{Ben, add the MOTION section here.}
-
-\section{Vectorization}
-\label{sec:vectorization}
-
-An important component of our algorithm is the scalar expansion to the corresponding loop dimensionality. 
-For example, \texttt{d = S[i * D + j] - C[j]} equiv. to \texttt{d = S[i,j] - C[j]}, which gave rise to \texttt{N*D} subtraction operations in the sequential schedule, 
-is lifted. The argument arrays \texttt{S} and \texttt{C} are lifted and the scaler \texttt{d} is lifted: \texttt{d[i,j] = S[i,j] - C[i,j]}.
-The algorithm then detects that the statement can be vectorized.
-
-There are three kinds of arrays (for now all kept internally as one-dimensional arrays, but that's under discussion).
-\begin{itemize}
-
-\item Scalars: These are scalar variables we lift into arrays for the purposes of vectorization. 
-For those, all writes are canonical writes and all reads are canonical reads. We will apply raise dimension when 
-a scalar is used in an inner loop (e.g., \texttt{sum0} in line 6 of the MPC source code will be raised to a 1-dimensional 
-array since \texttt{sum0} is used in the inner $j$-loop). Drop dimension applies as well; this happens when a scalar
-written in an inner loop is used outside of the loop (e.g., \texttt{sum0} for which the lifted inner loop computes 
-\texttt{D} values, but the outer loop only needs the last one.)
-
-\item Read-only input arrays: Read-only inputs. There are NO writes, while we may have non-canonical reads, $f(i,j,k)$. 
-Phase 1 of Basic vectorization will add raise dimension operation at the beginning of the function to lift these arrays, 
-and raise dimension may \emph{reshape} arrays. If there are multiple "views" of the input array, there would be multiple raise dimension statements to create
-each one of these views. The invariant is that at reads in loops, the reads of "views" of the original input array are canonical.
-Only raise dimension applies to these arrays, and only in the beginning of the program. For example, 1-dimensional array 
-\texttt{C} is lifted into 2-dimensional array \texttt{N x D} by copying the row \texttt{N} times.
-
- \item Read-write output arrays: Writes are canonical (by restriction) but reads can be non-canonical. 
- Dependence analysis limits vectorization when non-canonical read access refers to array writes in
- previous iterations, thus creating loop-carried dependences. We may apply both raise and drop dimension, however, 
- they respect the fixed dimensionality of the output array. The array cannot be raised to a dimension lower than
- its canonical (fixed) dimensionality and it cannot be dropped lower. In addition, non-canonical reads may 
- require lifting (i.e., reshaping) of the array after the most recent write, rather than in the beginning of the program, 
- in order to reduce a non-canonical read to a canonical one.
-
-\end{itemize}
-
-In the sections below we detail the $\mathit{raise\_dim}$ (raise dimension) and $\mathit{drop\_dim}$ (drop dimension) operations,
-followed by our vectorization algorithm.
-
-%\ana{Add back-edges into Phase 1.
-A back-edge from a non-phi-node in loop $i$ to a phi-node in loop $i$'s header is a same-level edge.
-The only difference with the handling of normal same-level forward def-use is that the operand index will become $i-1$.
-A back edge from a phi-node in loop $j$ to a phi-node in loop $i$, where $j$ is nested in $i$ is an inner-to-outer edge and will require dropping dimension.
-We can show these are the only kinds of back-edges that may occur.}
-
-%\subsection{Algorithm (Still work in progress.)}
-
-\subsection{Raise Dimension and Drop Dimension}
-
-There are two conceptual versions of $\mathit{raise\_dim}$. One applies on read-only input arrays and reshapes those arrays 
-when necessary to ensure canonical read access in the corresponding loop. The signature of $\mathit{raise\_dim}$
-is as follows. It takes the original array \texttt{C}, the access pattern function $f(i,j,k)$ in loop nest $i,j,k$ and the loop bounds
-\texttt{((i:I),(j:J),(k:K))}: 
-\[ \mathit{raise\_dim}(\texttt{A}, f(i,j,k), \texttt{((i:I),(j:J),(k:K))}) \]
-It produces a new 3-dimensional array \texttt{A'} by iterating over $i,j,k$ and setting each element of  \texttt{A'} as follows:
-\[ \texttt{A'[i,j,k] =  A[f(i,j,k)]} \]
-The end result is that uses of \texttt{[A[f(i,jk)]} in loop nest $i,j,k$ are replaced with canonical read-accesses to \texttt{A'[i,j,k]}
-that can be vectorized. In the running Biometric example, \texttt{C'} = $\mathit{raise\_dim}$\texttt{(C, j, (i:N,j:D))} lifts the 
-1-dimensional array \texttt{C} into a 2-dimensional array. The $i,j$ loop now accesses \texttt{C'} in the canonical way, \texttt{C'[i,j]}. 
-Similarly, \texttt{S'} = $\mathit{raise\_dim}$\texttt{(S, i*D+j, (i:N,j:D))} tries to lift \texttt{S}, but the operation turns into a no-op 
-because \texttt{S} is already a 2-dimensional array and the read access is canonical.
-
-The other version of $\mathit{raise\_dim}$ applies on scalars and read-write arrays. It lifts a lower-dimension array into a 
-higher-dimension for access in a nested loop. Here \texttt{A} is an $i$ array and raise dimension adds two additional dimensions:
-\[ \mathit{raise\_dim}(\texttt{A}, \texttt{(j:J,k:K)}) \]
-This version is reduced to the above version by adding the access pattern function, which is just $i$: 
-\[ \mathit{raise\_dim}(\texttt{A}, i, \texttt{(j:J,k:K)}) \]
-
-The corresponding $\mathit{drop\_dim}$ is carried out when an array written in an inner loop is used in an enclosing loop. 
-It takes a higher dimensional array, say $i,j,k$ and removes trailing dimensions, say $j,k$:
-\[ \mathit{drop\_dim}(\texttt{A}, \texttt{(j:J,k:K)}) \]
-It iterates over $i$ and takes the result at the maximal index of $j$ and $k$, i.e., the result at the last iterations of $j$ and $k$:
-\[ \texttt{A'[i] = A[i,J-1,K-1]} \]
-
-\begin{comment}
-We define the $\mathit{raise\_dim}$ (raise dimensions) and $\mathit{drop\_dim}$ (drop dimension) functions.
-Raise dimension "lifts" a lower-dimension array (\ana{The right term here is tensor, I am pretty sure!}) into a higher dimension one.
-This is necessary when a lower-dimensional array is used in a higher dimensional loop and, essentially, is just copying of values.
-For example, Biometric contains the statement $d = \mathit{SUB}(S[i, j], C[j])$, in the $j$-loop which is nested into the $i$-loop.
-Here $C$ is a one-dimensional array, however, to vectorize across both loops it is necessary to turn it into a two-dimensional array: $C[i,j]$ becomes $[C[0],C[1],...C[J],C[0],C[1],...C[J], ...
-C[0],C[1],...C[J]]$, which turns the row into a matrix of $I$ identical rows.
-(We use capital letters to denote the upper bounds of loops, e.g., $J$ is the upper bound of the $j$-loop and $I$ is the upper bound of the $i$-loop.)
-
-$\mathit{raise\_dim}(A[i_1,...,i_k],i_j)$ is defined as follows.
-It results in a new {k+1}-dimensional array $A'$ where for every $0 \le i_{k} < I_{k+1}$, $A'[i_1, \ldots, i_j, \ldots, i_k] = A[i_1, \ldots, i_k]$.
-Adding $n$ dimensions is trivially extended as a composition of $n$ $\mathit{raise\_dim}$ that each adds a single dimension.
-
-As expected, drop dimension turns a higher-dimensional array into a lower-dimensional one.
-The key use case is an inner-to-outer def-use edge.
-The code may define a variable, e.g., $\x$ in an inner loop, say $j$, then use this variable in an enclosing loop, say $i$.
-Our algorithm may vectorize the computation of $\x$ in the $j$-loop thus producing a vector $\x[j]$ where $\x[1]$ is the value of \x after the 1st iteration and so on.
-Drop dimension states that the outer loop will use the value of the variable at the last iteration.
-
-$\mathit{drop\_dim}(A[i_1,\ldots, i_j, \ldots,i_k], i_j)$ produces a k-dimensional array $A'$ where $A'[i_1,\ldots,i_k] = A'[i_1,\ldots, I_j-1, \ldots,i_k]$.
-In our analysis, dimensions are always dropped at the end, and again, one can define dropping $n$ dimensions as a composition of $n$ $\mathit{drop\_dim}$.
-
-\ben{I'm not sure if I wrote up the above paragraphs correctly -- the idea is that we're just copying values to extend over dimension $j$ or only retain its last elements.
-The current writeup implies that $1 < j < k$ though, which isn't always true.
-Python implementations of $\mathit{raise\_dim}()$ and $\mathit{drop\_dim}()$ can be found here: \url{https://github.com/milana2/ParallelizationForMPC/blob/master/compiler/compiler/motion_backend/reference_implementations.py}}
-
-\ana{This section is still informal. Needs work to make more precise.}
-\end{comment}
-
-\subsection{Basic Vectorization}
-
-\begin{algorithmic}
-
-\STATE \COMMENT { Phase 1: Raise dimension of scalar variables to corresponding loop nest. We can traverse stmts linearly in MPC-source. }
-
-\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i,j,k$}
-%\FOR {each $Y_n$}
-\FOR {each argument $Y_n$} %def-use edge $\mathit{stmt}' (\mbox{def of } Y) \rightarrow \mathit{stmt} (\mbox{def of } X)$}
-\STATE {case def-use edge $\mathit{stmt}' (\mbox{def of } Y_n) \rightarrow \mathit{stmt} (\mbox{def of } X)$ of} \\
-\STATE \hspace{0.25cm} {same-level: $Y'_n$ is $Y_n$} \\
-\STATE \hspace{0.25cm} {outer-to-inner: add $Y'_n[i,j,k] = \mathit{raise\_dim}(Y_n)$ at $\mathit{stmt'}$}
-\STATE \hspace{0.25cm} {(more precisely, right after $\mathit{stmt'}$)}
-\STATE \hspace{0.25cm} {inner-to-outer: add $Y'_n[i,j,k] = \mathit{drop\_dim}(Y_n)$ at $\mathit{stmt}$}
-\STATE \hspace{0.25cm} {(more precisely, in loop of $\mathit{stmt}$ right after loop of $\mathit{stmt'}$)}
-\ENDFOR
-
-\COMMENT { Optimistically vectorize all. $I$ means vectorized dimension. }
-%\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i,j,k$}
-\STATE {change to $X[I,J,K] = \mathit{Op}(Y'_1[I,J,K],Y'_2[I,J,K])$ }
-\ENDFOR
-
-\COMMENT { Phase 2: Recreating FOR loops for cycles; vectorizable statements hoisted up. }
-
-\FOR {each dimension $d$ from highest to 0}
-\FOR {each $\phi$-node $n$ in loop $i_1,...,i_d$}
-\STATE {compute $\mathit{closure}(n)$}
-\ENDFOR
-
-\COMMENT {{\color{red} $cl_1$ and $cl_2$ intersect if they have common statement or update same array; "intersect" definition can be expanded}}
-
-\WHILE {there are closure $cl_1$ and $cl_2$ that intersect}
-\STATE {merge $cl_1$ and $cl_2$}
-\ENDWHILE
-\FOR {each closure $cl$ (after merge)}
-\STATE {create FOR $i_d = 0; ...$ loop}
-\STATE {add $\phi$-nodes in $cl$ to header block}
-\STATE {{\color{red}{add target-less $\phi$-node for A if $cl$ updates array A}}}
-\STATE {add statements in $cl$ to loop body in some order of dependences}
-\STATE \COMMENT { Dimension is not vectorizable: }
-\STATE {change ${I_d}$ to $i_d$ in all statements in loop}
-\STATE {treat FOR loop as monolith node: some def-use edges become same-level.}
-\ENDFOR
-{
-\color{red}
-\FOR {each target-less $\phi$-node $A_1 = \phi(A_0,A_k)$}
-\STATE {in vectorizable stmts, replace use of $A_1$ with $A_0$}
-\STATE {discard $\phi$-node if not used in any $cl$, replacing $A_1$ with $A_0$ or $A_k$ appropriately}
-\ENDFOR
-}
-\ENDFOR
-
-\COMMENT { Phase 3: Remove unnecessary dimensionality. }
-
-\COMMENT { A dimension $i$ is dead on exit from stmt $X[...i...] = ...$ if all def-uses with targets outside of the enclosing 
-FOR $i=0 ...$ MOTION loop end at target (use) $X' = \mathit{drop\_dim}(X, i)$. } 
-\FOR {each stmt and dimension $X[...i...] = ...$}
-\STATE {if $i$ is a dead dimension on exit from stmt $X[...i...] = ...$, remove $i$ from $X$ (all defs and uses)} 
-%\COMMENT{ A stmt v[...i] is a "dead dimension i" if all edges out of FOR i = 0; ... loop end at drop\_dim(v, i)}
-\ENDFOR
-
-\COMMENT { Now clean up drop\_dim and raise\_dim }
-\FOR {each $X' =  \mathit{drop\_dim(X,i)}$ }
-\STATE {replace with $X' = X$ if $i$ is dead in $X$.}
-\ENDFOR
-\STATE {do (1) (extended) constant propagation, (2) copy propagation and (3) dead code elimination to get rid of redundant variables and raise and drop dimension statements}
-
-\COMMENT { Phase 4: }
-\STATE {add SIMD for simdified dimensions}
-
-\end{algorithmic}
-
-\input{biometric_vectorization_example.tex}
-
-\begin{comment}
-\subsection{Example: Biometric}
-
-\subsubsection{MPC-source}
-
-Below is the MPC-source resulting from Benjamin's analysis. It gives rise to a corresponding dependence graph. (Not shown.)
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0 = \phi(-1, \texttt{index}_1)$}
-\STATE {$\texttt{min}_0 = \phi(10000,\texttt{min}_1)$}
-
-\STATE \COMMENT {Begin of inner loop $j$.}
-\STATE {$\texttt{sum}_0 = \phi(0, \texttt{sum}_1)$}
-\STATE {$\texttt{d} = \mathit{SUB}(\texttt{S}[i,j], \texttt{C}[j])$}
-\STATE {$\texttt{p} = \mathit{MUL}(\texttt{d},\texttt{d})$}
-\STATE {$\texttt{sum}_1 = \texttt{ADD}(\texttt{sum}_0,\texttt{p})$}
-\STATE \COMMENT {End of inner loop $j$.}
-\STATE {$\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$}
-\STATE {$\texttt{index}_1 = \mathit{MUX}(\texttt{c},\texttt{index}_0,i)$}
-\STATE {$\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{min}_0,\texttt{sum}_0)$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-\subsubsection{Phase 1 of Basic Vectorization:}
-
-The transformation preserves the dependence edges.
-It raises the dimensions of scalars and optimistically vectorizes all operations.
-The next phase discovers loop-carried dependences and removes affected vectorization.
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$}
-\STATE {$\texttt{min}_0[i] = \phi(10000, \texttt{min}_1[i-1])$}
-
-\STATE \COMMENT {Begin of inner loop $j$.}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...], \texttt{sum}_1[i,j-1])$}
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{sum}_1[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE \COMMENT {End of inner loop $j$.}
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {$\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$}
-\STATE {$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$}
-\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-\ishaq{At least for input variables: when raising dimensions, we should keep some meta data around so we can later write code that could use 1 share if it knows the other N shares are just duplicates of it.}
-
-\subsubsection{Phase 2}
-
-This phase analyzes statements from the innermost loop to the outermost. The key point is to discover loop-carried dependencies and re-introduce loops whenever dependencies make this necessary.
-
-Starting at the inner phi-node $\texttt{sum}_0[i,j] = \phi([0,0,...],\texttt{sum}_1[i,j-1])$, the algorithm first computes its closure.
-The closure amounts to the phi-node itself and $\texttt{sum}_1[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$, accounting for the loop-carried dependency of the computation of \texttt{sum}.
-The algorithm replaces this closure with a FOR loop on $j$ removing vectoriaztion on $j$.
-Note that the SUB and MUL computations remain outside of the loop as they do not depend on phi-nodes that are part of cycles.
-The algorithm adds same-level edges, one from $\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = ...$ to the monolithic FOR-loop node, and one from the FOR-loop node to $\texttt{c}[\stackrel{\rightarrow}{i}] = ...$.
-
-\ishaq{For our radar: Such dependencies will not involve phi nodes only. e.g. $a[i] = a[i] + a[i-1]$}
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0[i] = \phi(-1, \texttt{index}_1[i-1])$}
-\STATE {$\texttt{min}_0[i] = \phi(10000, \texttt{min}_1[i-1])$}
-
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-
-\STATE \COMMENT {Begin of inner loop $j$. Loop is now much "shorter"!}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...], \texttt{sum}_1[i,j-1])$}
-\STATE {FOR j=0; j<D; j++} \COMMENT{ Will turn into a MOTION loop.}
-\STATE \hspace{0.25cm}{$\texttt{sum}_1[\stackrel{\rightarrow}{i},{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},{j}],\texttt{p}[\stackrel{\rightarrow}{i},{j}])$}
-\STATE \COMMENT {End of inner loop $j$.}
-
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {$\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$}
-\STATE {$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$}
-\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-Next, analysis moves to outer dimension $i$. There are two phi-nodes, $\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1])$ and $\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$.
-The closure of the first is
-\[
-\begin{array}{l}
-\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1]) \\
-\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
-\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])
-\end{array}
-\]
-and the closure of the second one is
-\[
-\begin{array}{l}
-\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1]), \\
-%\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
-\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])
-\end{array}
-\]
-
-Since the two closures \emph{do not} intersect, we have two distinct FOR-loops on $i$.
-The first FOR loop (on \texttt{min}) is scheduled first, because of the dependence edge from $\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$ to
-$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$.
-We can now rewrite those statements, getting rid of the vectorization on $i$ in the two FOR loops.
-Notably, the SUB and MUL computations are fully vectorizable and the ADD computation is vectorizable across the $i$-dimension.
-
-\begin{algorithmic}
-
-\STATE
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB\_SIMD}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL\_SIMD}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$} \\
-\STATE {}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...],\texttt{sum}_1[i,j-1])$}
-\STATE {FOR j=0; j<D; j++} \COMMENT{ Turns into a MOTION loop, was $j$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{sum}_1[\stackrel{\rightarrow}{i},{j}] = \mathit{ADD\_SIMD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},{j}],\texttt{p}[\stackrel{\rightarrow}{i},{j}])$}
-\STATE {}
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {}
-\STATE {$\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1])$}
-\STATE {FOR i=0; i<N; i++} \COMMENT{ Turns into a MOTION loop, was $i$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{c}[{i}] = \mathit{CMP}(\texttt{sum}'_0[{i}],\texttt{min}_0[{i}])$}
-\STATE \hspace{0.25cm}{$\texttt{min}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{min}_0[{i}],\texttt{sum}'_1[{i}])$}
-\STATE {}
-\STATE {$\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$}
-\STATE {FOR i=0; i<N; i++} \COMMENT{ Turns into a MOTION loop, was $i$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{index}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{index}_0[{i}],i)$}
-
-\ben{The above code implicitly stores the final values of min\_0 and index\_0 in their last indices, but there is no primitive like drop\_dim to make it clear.
-This will have to be added in the SSA code generation (though it might not be necessary to include in the paper?).
-It also is pretty wishy-washy about the size and layout of each vector (again more of an implementation issue), but maybe we should have a different syntax for array accesses for read/write (e.g.
-the ADD\_SIMD steps) vs the array definitions (e.g.
-the SUB\_SIMD and raise/drop\_dim statements)}
-
-\end{algorithmic}
-\end{comment}
-
-\subsection{Correctness Argument}
-
-We build a correctness argument that loosely follows the theory of Abstract Interpretation. 
-We define the syntax of MPC Source programs. The domain of MPC Source programs
-expressible in the syntax (with certain semantic restrictions) is the abstract domain $A$. 
-We then define the \emph{linearization} of an MPC Source program as an interpretation 
-over the syntax. The linearization, which is a \emph{schedule}, is the concrete domain $C$. 
-Since we reason over def-use graphs in $A$ we define a partial order relation over elements of $A$ 
-in terms of def-use relations. We define a partial order over elements of $C$ as well,
-in terms of def-use relations in the concrete domain $C$. We prove two theorems
-that state (informally) that the schedule corresponding to the original program computes the same 
-result as the schedule corresponding to the vectorized program.
-
-\begin{figure}%[!b]
-%\small
-\[
-\begin{array}{l@{~~~~~}l}
-  \begin{array}{l@{~}l@{~~~}l}
-  s &::= s;s & sequence \\
-     & \mid \x[i,J,k] = \y[i,J,k] \; \mathit{op\_SIMD} \; \z[i,J,k] & operation \\ 
-     & \mid \x[i,J,k]  = \texttt{PHI}(\x_1[i,J,k],\x_2[i,J,k\!-\!1]) & \mathit{phi} \; \mathit{node} \\
-     & \mid \x[i,J,k] = \mathit{raise\_dim}(\x'[i],(J\!:\!J,k\!:\!K)) & \mathit{raise} \; \mathit{dimension(s)} \\
-     & \mid \x[i,J] = \mathit{drop\_dim}(\x'[i,J,k],k)  & \mathit{drop} \; \mathit{dimension(s)} \\
-     & \mid \x = \y & \mathit{propagation} \\
-     & \mid \texttt{FOR} \; 0 \le i < I: s  & \mathit{loop} \\
-  \end{array}    \\
-\end{array}
-\]
-\vspace{-3mm}
-\caption{MPC Source Syntax}\label{fig:MPC_Source_syntax}
-\vspace{-5mm}
-\end{figure}
-
-\paragraph{MPC Source Syntax}
-
-\figref{fig:MPC_Source_syntax} defines the syntax for our intermediate representation, MPC Source. 
-There are semantics restrictions over the syntax as well: a variable $\x[i,j,k]$ is a 3-dimensional array 
-$(i:I,j:J,k:K)$ and also, a statement $\x[i,J,k] = ...$ is enclosed in loops over $i$ and $k$ as shown below. 
-Thus, $i$ and $k$ are in scope.
-\begin{verbatim}
-FOR 0 <= i < I:
-    ...
-    FOR 0 <= k < K:
-        x[i,J,k] = ...
-\end{verbatim}
-
-Statements $\mathit{operation, phi, raise \; dimension(s), drop \; dimension(s)}$ are \emph{base} statements, 
-and $\mathit{sequence, loop}$ are \emph{compound statements}.
-
-\paragraph{Linearization}
-
-Linearization is the concretization operation, which, as we mentioned earlier computes a schedule.
-The concretization function $\gamma: A \rightarrow C$ is defined as an interpretation of
-MPC Source syntax, as it is standard. The concretization of each one 
-of the base statements is as follows: 
-\[\begin{array}{l}
-\gamma(\x[i,J,k] = \mathit{op\_SIMD}(\y[i,J,k], \z[i,J,k])) = \\ 
-~~\x[i,0,k] = \y[i,0,k] \; \mathit{op} \; \z[i,0,k] \;\; || \\
-~~~~\x[i,1,k] = \y[i,1,k] \; \mathit{op} \; \z[i,1,k] \;\; || \; ...\; || \\
-~~~~~~\x[i,I-1,k] = \y[i,I-1,k] \; \mathit{op} \; \z[i,I-1,k]
-\end{array}\]
-meaning that the vectorized dimension(s) are expanded into parallel statements. || introduces SIMD (parallel) execution.
-
-The concretization of the FOR statement is as follows:
-\[ \gamma(\texttt{FOR} \; 0 \le i < I: s) = \gamma(s)[0/i] \; ; \; \gamma(s)[1/i] \; ; \; ... \gamma(s)[I-1/i] \]
-$\gamma$ simply unrolls the loop substituting $i$ with 0, 1, etc. Here ; denotes sequential execution.
-
-\paragraph{Partial Orders}
-
-For each MPC Source program $a$ we compute the def-use edges in the standard way: if base statement $s1 \in a$ defines variable \x, e.g., $\x[i,j,k] = ...$, 
-and base statement $s2 \in a$ uses \x, e.g., $... = ... \x[i,j,k]$ and there is a path in the trivial CFG from $s1$ to $s2$, then there is a def-use edge from
- $s1$ to $s2$. We extend the dimensionality of a statement into $s1[i,j,k]$ where $s1[i,j,k]$ inherits the dimensionality of the left-hand-side of the assignment.
-
- Let $a_0, a_1$ be two MPC Source programs in $A$. 
- Two base statements, $s_0 \in a_0$ and $s_1 \in a_1$ are \emph{same}, written $s_0 \equiv s_1$ if they are of the same operation and they operate on 
- the same variables: same variable name and same dimensionality. Recall that dimensions in MPC Source are either iterative, lower case, or vectorized, upper case. 
- Two statements are same even if one operates on an iterative dimension and the other one operates on a vectorized one, e.g., $s_0[i,j,k] \equiv s_1[I,j,K]$.
-
- \begin{definition} Let $a_0, a_1 \in A$. We say that $a_0 \le a_1$ iff for every def-use edge $s1 \rightarrow s2$ in $a_0$ there is an edge
- $s1' \rightarrow s2'$ where $s1 \equiv s1'$, $s2 \equiv s2'$ and the two edges of either both forward or both backward.
- \end{definition}
-
- The def-use edges in the concrete schedule are as expected. There is a def use edge from statement $s1$ that defines $\x[\underline{i},\underline{j},\underline{k}]$
- to statement $s2$ that uses $\x[\underline{i},\underline{j},\underline{k}]$ if $s1$ is scheduled ahead of $s2$ in the linear schedule. We note that the underlined
- indices, e.g., $\underline{i}$, refer to fixed values, not iterative or vectorized dimensions since in the concrete schedule all induction variables are expanded. 
- E.g., there is a def-use edge from the statement that defines $\x[0,1,2]$ and a statement that uses $\x[0,1,2]$.
-
-\paragraph{Theorems}
-
-\begin{theorem} $a_0 \le a_1 \Rightarrow \gamma(a_0) \subseteq \gamma(a_1)$.
-\end{theorem}
-
-\begin{theorem} Let $a_0$ be the iterative MPC Source and let $a_1$ be the vectorized MPC Source computed by the vectorization algorithm. 
-We have that $a_0 \le a_1$. 
-\end{theorem}
-
-\ana{Write the proof sketch and final argument, etc.}
-
-\subsection{Extension of Basic Vectorization with Array Writes}
-
-\subsubsection{Removal of Infeasible Edges}
-
-Array writes limit vectorization as they sometimes introduce infeasible loop-carried dependencies. Consider the following example: \ana{Have to add citation to Aiken's paper}
-
-{\small
-\begin{verbatim}
-for i in range(N):
-  A[i] = B[i] + 10;
-  B[i] = A[i] * D[i-1];
-  C[i] = A[i] * D[i-1];
-  D[i] = B[i] * C[i];
-\end{verbatim}
-}
-
-In Cytron's SSA this code (roughly) translates into
-
-{\small
-\begin{verbatim}
-for i in range(N):
-1. A_1 = PHI(A_0,A_2)
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. A_2 = update(A_1, i, B_1[i] + 10); 
-6. B_2 = update(B_1, i, A_2[i] * D_1[i-1]);
-7. C_2 = update(C_1, i, A_2[i] * D_1[i-1]);
-8. D_2 = update(D_1, i, B_2[i] * C_2[i]);
-\end{verbatim}
-}
-
-\begin{comment}
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE \hspace{0.25cm} A[i] = B[i] + 10;
-\STATE \hspace{0.25cm} B[i] = A[i] * D[i-1];
-\STATE \hspace{0.25cm} C[i] = A[i] * D[i-1];
-\STATE \hspace{0.25cm} D[i] = B[i] * C[i];
-\end{algorithmic}
-
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{A}_0$ = $\phi(\texttt{A},\texttt{A}_1)$
-\STATE 2. $\texttt{B}_0$ = $\phi(\texttt{B},\texttt{B}_1)$
-\STATE 3. $\texttt{C}_0$ = $\phi(\texttt{C},\texttt{C}_1)$
-\STATE 4. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$
-\STATE 5. $\texttt{A}_1$ = $\texttt{A}_0$; $\texttt{A}_1$[i] = $\texttt{B}_0$[i] + 10; \COMMENT{equiv. to Update}
-\STATE 6. $\texttt{B}_1$ = $\texttt{B}_0$; $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 7. $\texttt{C}_1$ = $\texttt{C}_0$; $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 8. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
-\end{algorithmic}
-\end{comment}
-
-There is a cycle around \texttt{B\_1 = PHI(B\_0,B\_2)} that includes statement 
-\texttt{A\_1 = update(A\_0, i, B\_1[i] + 10)} and that statement won't be vectorized even 
-though in fact there is no loop-carried dependency from the write of \texttt{B\_1[i]} at 6 to the
-read of \texttt{ ... = B\_1[i]} at 8.
-
-The following algorithm removes certain infeasible loop-carried dependencies that are due to array writes. Consider a loop with index $0 \le j < J$
-nested at $i,j,k$. Here $i$ represents the enclosing loops of $j$ and $k$ represents the enclosed loops in $j$.
-
-\begin{algorithmic}
-\FOR {each array $\texttt{A}$ written in loop $j$}
-\STATE \COMMENT { including enclosed loops in $j$ }
-\STATE dep = False
-\FOR {each pair def: $\texttt{A}_m[f(i,j,k)] = ... $, and use: $ ... = \texttt{A}_n[f'(i,j,k)]$ in loop $j$}
-\IF {$\exists \underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$, s.t.
-$0\!
-\le\!
-\underline{i}\!
-<\!
-I,\: 0\!
-\le\!
-\underline{j}, \underline{j}' \!
-<\!
-J,\: 0\!
-\le\!
-\underline{k}, \underline{k}' \!
-<\!
-K,\: \underline{j}<\underline{j}'$, and $f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$}
-\STATE dep = True
-\ENDIF
-\ENDFOR
-\IF {dep == False}
-\STATE remove back edge into $\texttt{A}$'s $\phi$-node in loop $j$.
-\ENDIF
-\ENDFOR
-\end{algorithmic}
-
-Consider a loop $j$ enclosed in some fixed $\underline{i}$. Only if an update (definition) $\texttt{A}_m[f(i,j,k)] = ... $ at some iteration $\underline{j}$
-references the \emph{same} array element as a use $ ... = \texttt{A}_n[f'(i,j,k)]$ at some later iteration $\underline{j}'$,
-we may have a loop-carried dependence for $\texttt{A}$ due to this def-use pair. (In contrast, Cytron's algorithm inserts a loop-carried dependency every time there is an array update.)
-The algorithm above examines all def-use pairs in loop $j$, including defs and uses in nested loops, searching for values $\underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$ that satisfy
-$f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$. If such values exist for some def-use pair, then there is a potential
-loop-carried dependence on $\texttt{A}$; otherwise there is not and we can remove the spurious backward edge thus ``freeing up'' statements for vectorization. %\ana{Double check this!!!}
-
-Consider the earlier example.
-There is a single loop, $i$.
-Clearly, there is no pair $\underline{i}$ and $\underline{i}'$, where $\underline{i} < \underline{i}'$ that make $\underline{i} = \underline{i}'$
-due to the def-use pairs of \texttt{A} 5-6 and 5-7.
-Therefore, we remove the back edge from 5 to the phi-node 1.
-Analogously, we remove the back edges from 6 to 2 and from 7 to 3.
-However, there are many values $\underline{i} < \underline{i'}$ that make $\underline{i} = \underline{i'}-1$ and the back edge from 8 to 4 remains (def-use pairs for \texttt{D}).
-As a result of removing these spurious edges, Basic Vectorization will find that statement 5 is vectorizable.
-Statements 6, 7 and 8 will correctly appear in the FOR loop.
-
-Note however, that this step renders some array phi-nodes target-less. We handle target-less phi-nodes with a minor extension of Basic Vectorization (Phase 2). 
-First, we merge closures that update the same array. This simplifies handling of array $\phi$-nodes: if each closure is turned into a separate loop
-each loop will need to have its own array phi-node to account for the update and this would complicate the analysis.
-Second, we add the target-less node of array \texttt{A} back to the closure that updates \texttt{A} ---
-the intuition is, even if there is no loop-carried dependence from writes to reads on \texttt{A}, \texttt{A} is written and the write (i.e., update) cannot be vectorized; 
-therefore, the updated array has to carry to the next iteration of the loop. Third, in cases when the phi-node remains target-less, i.e., cases when the
-array write can be vectorized, we have to properly remove the phi-node replacing uses of the left-hand side of the phi-node with its arguments.
-
-%\ana{Commented out previous writeup. Removal/handling of targetless phi-nodes will go into the Extension to Basic vectorization.}
-
-\begin{comment}
-The final step is to remove array phi-nodes that are left without incoming back edges after removal of infeasible edges.
-Such a phi-node is redundant when there are no loop-carried dependences that are due to writes to the corresponding array.
-In our running example, the phi-nodes for \texttt{A}, \texttt{B}, and \texttt{C}, are redundant.
-
-Removing redundant phi-nodes proceeds from \emph{innermost to outermost} loops as follows.
-Denote a redundant phi-node by $\texttt{A}_1 = \phi(\texttt{A}_0,\texttt{A}_k)$. $\texttt{A}_0$ is the array at the beginning of the loop
-and $\texttt{A}_k$ is the array at the end of an iteration (including the last iteration). We consider two cases:
-\begin{enumerate}
-\item[Case 1] Inner-to-outer or same-level def-use edge
-\begin{itemize}
-\item If use is an array read $ ... = \texttt{A}_1[f(i)]$, then replace $\texttt{A}_1$ with $\texttt{A}_0$. This is correct, i.e., it is guaranteed that the use
-references $\texttt{A}_0$; had the use been referencing an array modified in an earlier iteration, the back-edge to the phi-node would
-not have been removed and the phi-node would not have been classified as redundant.
-\item If use is an array write (update) $\texttt{A}_2 = \texttt{A}_1; \texttt{A}_2[i] = ... $ equiv. to $\texttt{A}_2 = \mathit{update}(\texttt{A}_1, i, ...)$ then handling becomes more tricky.
-The semantics of this update is that when $i$ is 0, then $\texttt{A}_0$ is being read.
-But when $i > 0$, then $\texttt{A}_k$ is being read.
-If we are at this case, this means that the def-use edge is same-level and we are removing a redundant node in the loop enclosing the update.
-The removal of the original phi-node entails pushing the phi-node at the update:
-$\texttt{A}_2 = \phi(\texttt{A}_0,\texttt{A}_k); \texttt{A}_2[i] = ... $ equiv. to $\texttt{A}_2 = \mathit{update}(\phi(\texttt{A}_0,\texttt{A}_k), i, ...)$.
-\item If use is a phi-node $... = \phi(\texttt{A}_1,...)$, that means we have an outer-to-inner edge. The
-
-\end{itemize}
-\end{enumerate}
-For each outer-to-inner or same-level def-use edge
-that originates at the node, replace $\texttt{A}_1$ at the use node by $\texttt{A}_0$.
-For each inner-to-outer def-use edge, replace $\texttt{A}_1$ by $\texttt{A}_k$. In this case, the use references the array
-that results from the computation in the loop.
-
-Our running example is transformed as follows. Removing the phi-nodes for \texttt{A}, \texttt{B}, and \texttt{C} transforms the
-replaces the references to $\texttt{A}_0$, $\texttt{B}_0$, and $\texttt{C}_0$, with references to $\texttt{A}$, $\texttt{B}$, and $\texttt{C}$
-respectively, i.e., the arrays before the loop.
-
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$
-\STATE 2. $\texttt{A}_1$ = $\texttt{A}$; $\texttt{A}_1$[i] = $\texttt{B}$[i] + 10;
-\STATE 3. $\texttt{B}_1$ = $\texttt{B}$; $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 4. $\texttt{C}_1$ = $\texttt{C}$; $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
-\end{algorithmic}
-\end{comment}
-
-\begin{comment}
-\subsubsection{Array MUX refinement}
-
-\ana{TODO: WIP, get to BV without MUX refinement for now? For now, skip MUX refinement.}
-
-Next, the algorithm refines array MUX statements.
-MPC-source after Cytron's SSA may result in statements $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$, which imply that any index
-of $A$ can be written at this point and therefore there is a loop-carried dependency. In some cases the MUX can be refined to just a single
-index or a pair of indices, e.g., $\texttt{A}_j[i] = \emph{MUX}(c, \texttt{A}_k[i], \texttt{A}_l[i])$.
-
-%\ana{Clarify why this helps. Still work in progress.}
-This is to reduce the dimensionality of simd-ified computation. Technically, $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$ is
-a simdified operation that can be carried out in parallel "in one round". However, particularly when \texttt{A} is a multi-dimensional array, there is
-substantial increase in the size of the arrays (vectors) we send to SIMD operations. Refining to an update to a specific index would reduce
-the size of those vectors. Note that this is a heuristic that handles a common case, but not all cases of array updates.
-%\ana{TODO: Simplify this algorithm, taking into account the restriction to canonical updates. It should be handling \_all\_ cases.}
-
-\begin{algorithmic}
-\FOR {each $\mathit{stmt}$: $\texttt{A}_j = \emph{MUX}(c, \texttt{A}_k, \texttt{A}_l)$ in the MPC-source seq.}
-\STATE $i_1 = \mathit{find\_update}(A_k)$ \COMMENT { Is null when $A_k = \phi(...)$ }
-\STATE $i_2 = \mathit{find\_update}(A_l)$ \COMMENT { Is null when $A_l = \phi(...)$ }
-\IF {$i_1 == i_2$ or $i_1$ is null or $i_2$ is null} \STATE \COMMENT{ With our restrictions on writes we must have $i_1 = i_2$. }
-\STATE replace $\mathit{stmt}$ with \\
-\STATE $\texttt{A}_j = \texttt{A}_{j-1}$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
-%\ELSIF {$i_1 \ncong i_2$}
-%\STATE replace $\mathit{stmt}$ with \\
-%\STATE $\texttt{A}_j = \texttt{A}_k$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
-\ELSE
-\STATE $\mathit{stmt}$ stays as is
-\ENDIF
-\ENDFOR
-\end{algorithmic}
-\end{comment}
-
-\subsubsection{Restricting Array Writes}
-
-For now, we restrict array updates to \emph{canonical updates}. Assume (for simplicity) a two-dimensional array \texttt{A[I,J]}. A canonical update is the following:
-\begin{verbatim}
-for i in range(I):
-  for j in range(J):
-    ...
-    A[i,j] = ...
-    ...
-\end{verbatim}
-
-The update \texttt{A[i,j]} can be nested into an inner loop and there may be multiple updates, i.e., writes to \texttt{A[i,j]}.
-However, update such as \texttt{A[i-1,j] = ...} or \texttt{A[i-1,j-1] = ...}, etc., is not allowed.
-Additionally, while there could be several different loops that perform canonical updates, they must be of the same dimensionality, i.e., an update of higher or lower dimension, e.g., \texttt{A[i,j,k] = ...} is not allowed.
-We compute the \emph{canonical dimensionality} of each write array by examining the array writes in the original program and rejecting programs that violate the canonical write restriction.
-This restriction simplifies reasoning in this early stage of the compiler; we will look to relax the restriction in future work.
-
-Another restriction/assumption is that we assume the output array is given as input with initial values, and it is of size consistent with its canonical dimensionality. 
-%\ana{This restriction will require that we update our benchmarks that come with output arrays.}
-
-Reads through an arbitrary formula, such as \texttt{A[i-1]} for example, are allowed; currently, the projection function returns dummy values if the read formula is out of bounds; we assume the programmer ensures that the program still computes correct output in this case.
-
-\subsubsection{Changes to Basic Vectorization}
-
-\begin{comment}
-There are two changes to Basic Vectorization to account for write arrays. We can account for them during processing of raise dimension/drop dimension in Phase 1 of Basic Vectorization.
-
-One change to Basic vectorization is the expansion of dimension if the array write or read occurs in a nested loop.
-That is, if there is an update $\texttt{A}[i,j] = ...$ that occurs in loop dimensionality $i,j,k$, $\texttt{A}[i,j]$ will be rewritten into $\texttt{A}[i,j,k]$.
-Similarly, a read $\texttt{A}[f(i,j)]$ will be rewritten into $\texttt{A}[f(i,j),k]$.
-($\texttt{A}[f(i,j),k]$ will be reshaped as $f(i,j)*I*J+k$ during Phase 1 of Basic Vectorization.
-\ana{TODO: Double check.})
-\end{comment}
-
-In addition to the changes for the handling of target-less phi-nodes, Basic Vectorization has to handle def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable.
-The definition can be an update \texttt{A\_2 = update(A\_1,i,...)}, a pseudo $\phi$-node \texttt{A\_2 = PHI(A\_0,A\_1)}, etc..
-Note that $\phi$-nodes for arrays have no subscript operations the way there are subscript operations in analysis-introduced arrays representing scalars.
-While there are variations, the most intuitive implementation will perform Basic Vectorization Phase 1 as is, inserting
-$\mathit{raise\_dim}$ and $\mathit{drop\_dim}$ in the same way.
-However, the implementation of raise dimension and drop dimension will be adapted because the dimension cannot be raised or dropped to a dimension lower than the canonical one.
-Consider a def-use edge $X \rightarrow Y$ for an array $\texttt{A}$.
-
-\begin{enumerate}
-
-\item same-level $X \rightarrow Y$. Do nothing, propagate the array, which happens to be of the right dimension. 
-%\ana{There might be some opportunities to do copy propagation optimization and save some cycles, but let's leave this for later.}
-
-\item inner-to-outer $X \rightarrow Y$ triggers the addition of $\mathit{drop\_dim}$.
-However, the dimensionality cannot be dropped below the canonical dimensionality of the array.
-E.g., if the dimensionality of the loop enclosure $X$ is already at the canonical one, then $\mathit{drop\_dim}$ has no effect.
-
-\item outer-to-inner $X \rightarrow Y$ riggers  $\mathit{raise\_dim}$.
-Again, if the dimensionality of the loop enclosure of $Y$ is smaller or same as the canonical dimensionality of the array, then it has no effect, otherwise, if dimensionality is greater than the canonical dimensionality, $\mathit{raise\_dim(...)}$ (at $X$) is the same as in Basic Vectorization.
-
-\item "mixed" $X \rightarrow Y$. We assume that the mixed edge is transformed into an inner-to-outer followed by outer-to-inner edge before we perform vectorization, just as with Basic vectorization.
-
-\end{enumerate}
-
-If the use of the array is a read $\texttt{A}[f(i,j,k)]$ different than a canonical read $\texttt{A}[i,j,k]$, then we need to add a reshape operation as 
-all arrays are $\texttt{A}[i,j,k]$. It can be added after raise\_dim/drop\_dim or incorporated in these operations.
-The bulk of the change is in Phase 2 of Basic Vectorization as outlined earlier.
-
-\subsubsection{Examples with Array Writes}
-
-\paragraph{Example 1}
-
-First, the canonical dimensionality of all \texttt{A,B,C} and \texttt{D} is 1. 
-After Phase 1 of Basic Vectorization the Aiken's array write example will be (roughly) as follows:
-
-{\small
-\begin{verbatim}
-for i in range(N):
-1. A_1 = PHI(A_0,A_2)
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. A_2 = update(A_1, I, B_1[I] + 10); 
-6. B_2 = update(B_1, I, A_2[I] * D_1[I-1]);
-7. C_2 = update(C_1, I, A_2[I] * D_1[I-1]);
-8. D_2 = update(D_1, I, B_2[I] * C_2[I]);
-\end{verbatim}
-}
-
-Note that since all def-uses are same-level (i.e., reads and writes of the array elements) 
-no raise dimension or drop dimension happens.
-
-\begin{comment}
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{A}_1$ = $\phi(\texttt{A}_0,\texttt{A}_2)$
-\STATE 2. $\texttt{B}_1$ = $\phi(\texttt{B}_0,\texttt{B}_2)$
-\STATE 3. $\texttt{C}_1$ = $\phi(\texttt{C}_0,\texttt{C}_2)$
-\STATE 4. $\texttt{D}_1$ = $\phi(\texttt{D}_0,\texttt{D}_2)$
-\STATE 5. $\texttt{A}_2 = update(\texttt{A}_1, \stackrel{\rightarrow}{i}, \texttt{B}_1[\stackrel{\rightarrow}{i}] + 10)$;
-\STATE 6. $\texttt{B}_2 = update(\texttt{B}_1, \stackrel{\rightarrow}{i}, \texttt{A}_2[\stackrel{\rightarrow}{i}] * \texttt{D}_1[\stackrel{\rightarrow}{i-1}])$;
-\STATE 7. $\texttt{C}_2 = update(\texttt{C}_1, \stackrel{\rightarrow}{i}, \texttt{A}_2[\stackrel{\rightarrow}{i}] * \texttt{D}_1[\stackrel{\rightarrow}{i-1}])$;
-\STATE 8. $\texttt{D}_2 = update(\texttt{D}_1; \stackrel{\rightarrow}{i}, \texttt{B}_2[\stackrel{\rightarrow}{i}] * \texttt{C}_2[\stackrel{\rightarrow}{i}])$;
-\end{algorithmic}
-\end{comment}
-
-Phase 2 computes the closure of 4; $cl = \{4,6,7, 8\}$ 
-while 5 is vectorizable. Recall that 1,2, and 3 are target-less phi-nodes. Since the closure $cl$ includes updates to \texttt{B} and \texttt{C},
-the corresponding phi-nodes are added back to the closure and the def-use edges are added back to the target-less nodes. The uses of \texttt{A\_1} and \texttt{B\_1} in the vectorized statement
-turn into uses of \texttt{A\_0} and \texttt{B\_0} respectively; this is done for all original target-less phi-node.
-(But note that \texttt{A\_0} is irrelevant; the update writes into array \texttt{A\_2} in parallel.)
-Finally, the target-less phi-node for \texttt{A} is discarded.
-
-{\small
-\begin{verbatim}
-1. A_2 = update(A_0, I, ADD_SIMD(B_0[I],10));
-   equiv. to A_2[I] = ADD_SIMD(B_0[I],10)
-FOR i=0; i<N; i++; // MOTION loop
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. B_2 = update(B_1, i, A_2[i] * D_1[i-1]);
-   equiv. to B_2 = B_1; B_2[i] = A_2[i] * D_1[i-1];
-6. C_2 = update(C_1, i, A_2[i] * D_1[i-1]);
-7. D_2 = update(D_1, i, B_2[i] * C_2[i]);
-\end{verbatim}
-}
-
-\begin{comment}
-\begin{algorithmic}
-\STATE 1. $\texttt{A}_2 = update(\texttt{A}_0; \stackrel{\rightarrow}{i}, \mathit{ADD\_SIMD}(\texttt{B}_0[\stackrel{\rightarrow}i],[10,...])$ \COMMENT{Fully vectorized, size N.}
-\STATE FOR i=0; i<N; i++; \COMMENT{ MOTION loop }
-\STATE 2. $\texttt{B}_1$ = $\phi(\texttt{B}_0,\texttt{B}_2)$
-\STATE 3. $\texttt{C}_1$ = $\phi(\texttt{C}_0,\texttt{C}_2)$
-\STATE 4. $\texttt{D}_1$ = $\phi(\texttt{D}_0,\texttt{D}_2)$
-\STATE 5. $\texttt{B}_2 = update(\texttt{B}_1, i, \mathit{MUL}(\texttt{A}_2[i], \texttt{D}_1[{i-1}])$
-\STATE 6. $\texttt{C}_2 = update(\texttt{C}_1, i, \mathit{MUL}(\texttt{A}_2[i], \texttt{D}_1[{i-1}])$
-\STATE 7. $\texttt{D}_2 = update(\texttt{D}_1, i, \mathit{MUL}(\texttt{B}_2[i], \texttt{C}_1[i])$
-\end{algorithmic}
-\end{comment}
-
-\paragraph{Example 2} Now consider the MPC Source of Histogram:
-
-{\small
-\begin{verbatim}
-for i in range(0, num_bins):
-   res1 = PHI(res, res2)
-   for j in range(0, N):
-      res2 = PHI(res1, res3)
-      tmp1 = (A[j] == i)
-      tmp2 = (res2[i] + B[j])
-      tmp3 = MUX(tmp1, res2[i], tmp2)
-      res3 = Update(res2, i, tmp3)     
-return res1
-\end{verbatim}
-}
-
-The canonical dimensionality of \texttt{res} is 1.
-Also, the phi-node \texttt{res1 = PHI(res, res2)} is a target-less phi-node (the implication being that the inner for loop can be vectorized across $i$).
-After Phase 1, Basic vectorization produces the following code (statements are implicitly vectorized along $i$ and $j$).
-In a vectorized update statement, we can ignore the incoming array, \texttt{res2} in this case.
-The update writes (in parallel) all locations of the 2-dimensional array, in this case it sets up each \texttt{res3[i,j] = tmp3[i,j]}.
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I = raise_dim(i, ((i:num_bins),(j:N)))
-for i in range(0, num_bins):
-   res1 = PHI(res, res2') // target-less phi-node
-   res1' = raise_dim(res1, (j:N))
-   for j in range(0, N):
-      res2 = PHI(res1', res3)
-      tmp1 = (A1 == I)
-      tmp2 = (res2 + B1)
-      tmp3 = MUX(tmp1, res2, tmp2)
-      res3 = Update(res2, (I,J), tmp3) 
-   res2' = drop_dim(res2)    
-res1'' = drop_dim(res1)
-return res1''
-\end{verbatim}
-}
-
-Processing the inner loop in Phase 2 vectorizes \texttt{tmp1 = (A1 == I)} along the $j$ dimension but leaves the rest of the statements in a MOTION loop.
-Processing the outer loop is interesting.
-This is because the PHI node is a target-less node, and therefore, there are no closures!
-Several things happen.
-(1) Everything can be vectorized along the $i$ dimension.
-(2) We remove the target-less PHI node, however, we must update uses of \texttt{res1} appropriately: the use at raise\_dim goes to the first argument of the PHI function and the use at drop\_dim goes to the second argument.
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I1 = raise_dim(i, ((i:num_bins),(j:N)))
-
-tmp1[I,J] = (A1[I,J] == I1[I,J])
-
-res1' = raise_dim(res, (j:N)) // replacing res1 with res, 1st arg
-for j in range(0, N):
-   res2 = PHI(res1', res3)
-   tmp2[I,j] = (res2[I,j] + B1[I,j])
-   tmp3[I,j] = MUX(tmp1[I,j], res2[I,j], tmp2[I,j])
-   res3 = Update(res2, (I,j), tmp3)
-   equiv. to res3 = res2; res3[I,j] = tmp3[I,j]
-res2' = drop_dim(res2)   
-res1 = drop_dim(res2') // replacing with res2', 2nd arg. NOOP
-return res1
-\end{verbatim}
-}
-
-\begin{comment}
-The cleanup phase can then get rid of $j$ in the \texttt{res} variables and \texttt{tmp} variables, so the loop operates over the \texttt{num\_bins}-size vectors. 
-This will lead to dropping raise\_dim and drop\_dim as well:
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I1 = raise_dim(i, ((i:num_bins),(j:N)))
-
-tmp1[I,J] = (A1[I,J] == I1[I,J])
-
-for j in range(0, N):
-   res2 = PHI(res1, res3) // now 1-dim array
-   tmp2[I] = ADD_SIMD(res2[I],B1[I,j])
-   tmp3[I] = MUX_SIMD(tmp1[I,j], res2[I], tmp2[I])
-   res3 = Update(res2, I, tmp3[I])
-return res2
-\end{verbatim}
-}
-\end{comment}
-
-\begin{comment}
-\paragraph{Example 2}
-
-Consider MPC-source of Histogram after removal of infeasible edges and redundant phi-nodes, and MUX refinement:
-
-\begin{algorithmic}
-
-\STATE res!1 = []
-\STATE for i: plaintext in range(0, num\_bins):
-\STATE \hspace{0.25cm} res!2 = $\phi$(res!1, res!3)
-\STATE  \hspace{0.25cm} ...
-\STATE res!4 = res!2 \COMMENT{Added due to the "mixed" def-use edge.}
-\STATE for i: plaintext in range(0, num\_bins):
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE \hspace{0.25cm}        for j: plaintext in range(0, N):
-\STATE \hspace{0.5cm}            res!5 = $\phi$(res!4, res!7)
-\STATE \hspace{0.5cm}            !2!3 = (A[j] == i)
-\STATE \hspace{0.5cm}            !3!3 = (res!5[i] + B[j])
-\STATE \hspace{0.5cm}            res!6 = res!5; res!6[i] = !3!3
-\STATE \hspace{0.5cm}            res!7 = res!6; res!7[i] = MUX(!2!3, res!6[i], res!5[i])
-
-\end{algorithmic}
-
-After Phase 1 of Basic Vectorization:
-
-\begin{algorithmic}
-\STATE res!1 = []
-\STATE for i: plaintext in range(0, num\_bins):
-\STATE \hspace{0.25cm} res!2 = $\phi$(res!1, res!3)
-\STATE  \hspace{0.25cm} ...
-\STATE res!4 = res!2 \COMMENT{Added due to the "mixed" def-use edge. Because the canonical dimensionality of res is 1, we DO NOT drop dimension of res!2.}
-\STATE A[$i,j$] = $\mathit{raise\_dim}$(A[$j$],$i$) \COMMENT{Copying row A num\_bin times. Reusing name.}
-\STATE B[$i,j$] = $\mathit{raise\_dim}$(B[$i$],$j$) \COMMENT{Copying column B N times. Reusing name.}
-\STATE res!4[$i,j$] = $\mathit{raise\_dim}$(res!4[$i$],$j$) \COMMENT{Reusing name.}
-\STATE for i: plaintext in range(0, num\_bins):
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE \hspace{0.25cm}        for j: plaintext in range(0, N):
-\STATE \hspace{0.5cm}            res!5 = $\phi$(res!4, res!7) \COMMENT{Do nothing. res!4 has been raised to proper dimension.}
-\STATE \hspace{0.5cm}            !2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{EQ}$(A[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], [[0,0...], [1,1...],...[num\_bins-1,num\_bins-1...]])
-\STATE \hspace{0.5cm}            !3!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{ADD}$(res!5[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$],B[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$])
-\STATE \hspace{0.5cm}            res!6 = res!5 \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.5cm}            res!6[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = !3!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$]
-\STATE \hspace{0.5cm}            res!7 = res!6
-\STATE \hspace{0.5cm}            res!7[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{MUX}$(!2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], res!6[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], res!5[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$])
-\end{algorithmic}
-
-After Phase 2 and Phase 3. The EQ operation is vectorizable across both dimensions, and the rest of the computation is vectorizable across the $i$-dimension.
-
-\begin{algorithmic}
-\STATE ...
-\STATE !2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{EQ\_SIMD}$(A[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], [[0,0...], [1,1...],...[num\_bins-1,num\_bins-1...]])
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE FOR j=0; j<N; j++; \COMMENT{MOTION loop}
-\STATE \hspace{0.25cm}            res!5 = $\phi$(res!4, res!7) \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.25cm}            !3!3[$\stackrel{\rightarrow}i,j$] = $\mathit{ADD\_SIMD}$(res!5[$\stackrel{\rightarrow}i,j$],B[$\stackrel{\rightarrow}i,j$])
-\STATE \hspace{0.25cm}            res!6 = res!5 \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.25cm}            res!6[$\stackrel{\rightarrow}i,j$] = !3!3[$\stackrel{\rightarrow}i,j$]
-\STATE \hspace{0.25cm}            res!7 = res!6
-\STATE \hspace{0.25cm}            res!7[$\stackrel{\rightarrow}i,j$] = $\mathit{MUX\_SIMD}$(!2!3[$\stackrel{\rightarrow}i,j$], res!6[$\stackrel{\rightarrow}i,j$], res!5[$\stackrel{\rightarrow}i,j$])
-\end{algorithmic}
-\end{comment}
-
-%\section{Divide-and-Conquer}
-%\label{sec:div_and_conq}
-%\ana{TODO: Now that we have broken FOR loops into smaller chunks, we can add Divide-and-conquer reasoning with Z3 and implement this additional transform.}
-
-\section{Implementation and Evaluation}
-\label{sec:implementation}
-
-\section{Related Work}
-\label{sec:related}
-
-\paragraph{MPC languages and compilers.}
-
-Languages and compilers for secure computation have seen significant attention and advances in recent years. The early MPC compilers Fairplay~\cite{CCS:BenNisPin08}, 
-and Sharemind~\cite{ESORICS:BogLauWil08} were followed by PICCO~\cite{CCS:ZhaSteBla13}, Obliv-C~\cite{Zahur:2015}, TinyGarble~\cite{SP:SHSSK15}, 
-Wystiria~\cite{SP:RasHamHic14}, and others. A new generation of MPC compilers 
-includes SPDZ/SCALE-MAMBA/MP-SPDZ~\cite{Keller:2020} and the ABY/HyCC/MOTION~\cite{NDSS:DemSchZoh15,CCS:BDKKS18,Braun:2022} frameworks. 
-These two families are the state-of-the art and
-are actively developed. Another recent development is Viaduct, a functional language and compiler that supports a range of secure computation 
-frameworks, including MPC and ZKP. Hastings et el. present a review of compiler frameworks~\cite{Hastings:2019}.
-
-While each of these languages and compilers brings in new ideas and advances, none addresses the problem of ``circuit independent'' intermediate 
-representation and optimization. We envision a classical compiler structure: (1) a Wysteria, Viaduct, Obliv-C, or IMP Source front end, including rich type systems and 
-AST-level semantic analysis, compile into the MPC Source IR, (2) MPC Source-level optimizations take place, followed by (3) back-end compilers into circuits. 
-Our focus is at the intermediate level.
-
-%{\begin{center}
-%\includegraphics[width=0.5\linewidth]{figs/focus.png}
-%\end{center}
-%}
-
-Many works focus on the implementation of MPC protocols exposing an API to the programmer. For example, the ABY/MOTION line of 
-compiler frameworks provides a library of MPC primitives; the programmer writes MPC programs in C++ on top of the library. These back ends implement 
-different protocols and allow for mixing, but notably, they leave it to the programer to assign different protocols to different parts of the 
-computation and perform share conversion accordingly. In addition, MOTION provides SIMD primitives, which allows for efficient execution 
-of MPC operations, but again, using SIMD primitives is the responsibility of the programer. There is interest in frameworks for automatic mixing, 
-e.g., \cite{CCS:BDKKS18,Ishaq:2019,Fang:2022}.
-
-Other works, e.g., Obliv-C~\cite{Zahur:2015}, Wystiria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021} focus on higher-level language design, particularly information-flow systems that 
-restrict flow between secure and insecure parts of the program. 
-%Key ideas such as ${\sf insecure} <: {\sf secure}$ qualifier systems are similar to the ideas we 
-%present in~\Secref{sec:interprocedural} and~\Secref{sec:zkp} and we intend to draw on these work. While we will design a high-level language expanding our 
-%intraprocedural IMP-MPC syntax and semantics into the \emph{interprocedural} case, our interest and focus lie primarily in the ``machine-independent'' optimizations 
-%thrust of the proposal.
-
-%There is significant development in ZKP languages as well. ~\ana{Alex Ozdemir's talk, ZoKrates, Cairo, others} Our work focuses on ZKP via MPC and our key idea
-%is the separation of the higher-level ``backend-independent'' compilation and optimization and the lower-level compilation into circuits.
-
-\paragraph{Classical HPC compilers.}
-Automatic vectorization is a longstanding problem in high-performance computing (HPC). 
-There are thousands of works in this area reflecting over 40 years of research. We presented a vectorization
-algorithm for MPC Source, essentially extending classical loop vectorization \cite{Allen:1987}. In HPC vectorization, conditional control flow 
-presents a challenge --- one cannot estimate the cost of a schedule or vectorize branches in a straightforward manner --- in contrast to 
-MPC Source vectorization. 
-We view Karrenberg's work on Whole function vectorization~\cite{Karrenberg:2015} as most closely related to ours --- it linearizes the program and vectorizes 
-both branches of a conditional applying masking to avoid execution of the branch-not-taken code, and selection (similar to MUX) to select the correct value based on the 
-result of the condition at runtime. 
-The problem is that masking and selection, or more generally, handling control predicates~\cite{Benabderrahmane:2010,Karrenberg:2015}, 
-can lead to \emph{slowdown}. 
-%As a result, vectorizing compilers and polyhedral compilers tend to focus on straight-line code due to the unpredictable 
-%overhead conditionals may introduce. \ana{add Louis noel pochet's benchmarks. citation on vectorizing compilers???}
-
-We argue that vectorization over linear MPC Source is a different problem, one that warrants a new look, while drawing from 
-results in HPC.
-Since both branches of the conditional and the multiplexer \emph{always} execute, not only can we apply aggressive vectorization on linear code, but (perhaps more importantly) 
-we can also build analytical models that accurately predict execution time. These models in turn would drive optimizations such as vectorization, protocol mixing, and others. 
-Vectorization meshes in with those additional optimizations in non-trivial ways.
-
-Furthermore, extensions of classical loop vectorization with array writes, arbitrary indexing, including non-affine indexing, and interaction with SSA are non-trivial 
-and present novel challenges and opportunities for contribution. Polyhedral parallelization~\cite{Benabderrahmane:2010} considers a higher-level source (typically AST) 
-representation, while our work takes advantage of linear MPC Source and SSA form. The work by Karrenberg~\cite{Karrenberg:2015} is rare in that space, in the sense that it considers 
-vectorization over SSA form, which has similarities to MPC Source. We consider different array representation, notion of dependence, and reasoning 
-about dependence, which we conjecture is more suitable for MPC Source. Buscher~\cite{Buscher:2018} considers SIMD-vectorization 
-at the level of source code, which then combines with circuit-level optimizations in the TinyGarble compiler. 
-He proposes using an off-the-shelf polyhedral compiler, however, application is limited to only two routines, 
-essentially just inner product and euclidian distance; it is unclear how effective the off-the-shelf compiler is. 
-In contrast, we consider vectorization at the level of MPC Source separating ``backend-independent'' vectorization
-and circuit-level amortization (done by MOTION). We apply our compiler on a wide range of routines.
-
-%\section{Future Work}
-%\label{sec:implementation_and_benchmarks}
-%\input{../sections/implementation_and_benchmarks.tex}
-%\input{../sections/evaluation.tex}
-
-\section{Conclusion and Future Work}
-\label{sec:conclusion}
-%\input{../sections/conclusion.tex}
-
-%\begin{acks}
-%    \ishaq{TODO}
-%\end{acks}
-
-%\printbibliography
-
-\bibliographystyle{plain}
-%\bibliography{c:/trink/Doc/MyTexFiles/texrefs}
-%\bibliography{mybib,library,library1,library2,library3,library_ana}
-\bibliography{library_ana,crypto_crossref}
-
-% that's all folks
-\end{document}

### 2022-04-28 overleaf/related.tex
Split paper into sections.

--- 2022-04-27 overleaf/paper_SIMD.tex
@@
-% !TeX TXS-program:compile = txs:///pdflatex/[--shell-escape]
-%
-% acm templates v1.57 (sigconf template)
-%
-%\documentclass[sigconf, screen, review, anonymous, natbib=false]{acmart}
-\documentclass[sigconf, screen, natbib=false, dvipsnames, table]{acmart}
-\settopmatter{printacmref=true}
-% mandatory for CCS'19
-\usepackage{balance}
-\usepackage{mathtools}
-% for creating a balanced last page (usually last page with references)
-
-% % defining the \BibTeX command - from Oren Patashnik's original BibTeX documentation.
-% \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08emT\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-
-% Rights management information.
-% This information is sent to you when you complete the rights form.
-% These commands have SAMPLE values in them; it is your responsibility as an author to replace
-% the commands and values with those provided to you when you complete the rights form.
-%
-% These commands are for a PROCEEDINGS abstract or paper.
-
-\copyrightyear{2019}
-\acmYear{2019}
-\setcopyright{acmcopyright}
-\acmConference[CCS '19]{2019 ACM SIGSAC Conference on Computer and Communications Security}{November 11--15, 2019}{London, United Kingdom}
-\acmBooktitle{2019 ACM SIGSAC Conference on Computer and Communications Security (CCS '19), November 11--15, 2019, London, United Kingdom}
-\acmPrice{15.00}
-\acmDOI{10.1145/3319535.3339818}
-\acmISBN{978-1-4503-6747-9/19/11}
-
-%
-% These commands are for a JOURNAL article.
-%\setcopyright{acmcopyright}
-%\acmJournal{TOG}
-%\acmYear{2018}\acmVolume{37}\acmNumber{4}\acmArticle{111}\acmMonth{8}
-%\acmDOI{10.1145/1122445.1122456}
-
-%
-% Submission ID.
-% Use this when submitting an article to a sponsored event. You'll receive a unique submission ID from the organizers
-% of the event, and this ID should be used as the parameter to this command.
-%\acmSubmissionID{123-A56-BU3}
-
-%
-% The majority of ACM publications use numbered citations and references. If you are preparing content for an event
-% sponsored by ACM SIGGRAPH, you must use the "author year" style of citations and references. Uncommenting
-% the next command will enable that style.
-%\citestyle{acmauthoryear}
-
- \makeatletter
- % \renewcommand{\section}{\abovedisplayskip 10\p@ \@plus3\p@ \@minus1\p@%
- %                       \belowdisplayskip 5\p@ \@plus3\p@ \@minus1\p@%
- %                       \abovedisplayshortskip 0pt \@plus2\p@%
- %                       \belowdisplayshortskip 0pt \@plus2\p@ \@minus0\p@%
- %                       \@startsection{section}{1}{\z@}%
- %                        {-17\p@ \@plus -4\p@ \@minus -4\p@}%
- %                        {6\p@ \@plus 4\p@ \@minus 4\p@}%
- %                        {\normalfont\large\bfseries\boldmath
- %                         \rightskip=\z@ \@plus 8em\pretolerance=10000 }}
- \renewcommand{\subsection}{\@startsection{subsection}{2}{\z@}%
-                        {-8\p@ \@plus -4\p@ \@minus -4\p@}%
-                        {5\p@ \@plus 2\p@ \@minus 2\p@}%
-                        {\normalfont\Large\bfseries\boldmath
-                         \rightskip=\z@ \@plus 3em\pretolerance=10000 }}
-  \renewcommand{\subsubsection}{\@startsection{subsubsection}{3}{\z@}%
-                        {-6\p@ \@plus -4\p@ \@minus -4\p@}%
-                        {1\p@ \@plus 1\p@ \@minus 0\p@}%
-                        {\normalfont\normalsize\bfseries\boldmath}}
-% \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\z@}%
-%                       {-8\p@ \@plus -4\p@ \@minus -4\p@}%
-%                       {-2\p@ \@plus -0.22em \@minus -0.1em}%
-%                       {\normalfont\normalsize\bfseries}}
- \makeatother
-
-% *** GRAPHICS RELATED PACKAGES ***
-\usepackage{graphicx}
-\usepackage{comment}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{listings}
-% declare the path(s) where your graphic files are
-\graphicspath{{../figures/}}
-% and their extensions so you won't have to specify these with
-% every instance of \includegraphics
-\DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-
-% Minted is used for source code highlighting
-%\usepackage[outputdir=/Volumes/ramdisk/]{minted}
-% \usepackage{minted}
-% *** MATH PACKAGES ***
-\usepackage{amsmath}
-
-% *** SPECIALIZED LIST PACKAGES ***
-%\usepackage{algorithm2e}
-\usepackage{algorithmic}
-
-% *** ALIGNMENT PACKAGES ***
-\usepackage{array}
-
-\usepackage{url}
-
-\usepackage[utf8]{inputenc}
-\usepackage[american]{babel}
-%\usepackage[backend=biber,style=ACM-Reference-Format]{biblatex}
-%\addbibresource{./library.bib}
-%\addbibresource{./cryptobib/abbrev2.bib}
-%\addbibresource{./cryptobib/crypto_crossref.bib}
-
-% Used for Theorems and Definitions
-\usepackage{amsthm}
-\newtheorem{definition}{Definition}
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}{Lemma}
-\theoremstyle{definition}
-\newtheorem{remark}{Remark}
-\newtheorem{assumption}{Assumption}
-
-\usepackage{xcolor}
-\usepackage{xspace}
-\usepackage{xstring} % required by IfEqCase
-
-\usepackage{booktabs} % better tables
-\usepackage{algorithm}
-\usepackage{algorithmic}
-
-% for references, use it like this: \cref{some_label}
-% it will automatically put Appendix, Figure, Section, etc before the reference
-\usepackage[capitalise]{cleveref}
-
-% ishaq: adding our macros
-\input{macros.tex}
-
-\sloppy
-\begin{document}
-\fancyhead{}
-% do not delete this code.
-
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
-\title[Compilation and Backend-Independent Vectorization for Multi-Party Computation]{Compilation and Backend-Independent Vectorization for Multi-Party Computation}
-
-%
-% The "author" command and its associated commands are used to define the authors and their affiliations.
-% Of note is the shared affiliation of the first two authors, and the "authornote" and "authornotemark" commands
-% used to denote shared contribution to the research.
-
-\author{Benjamin Levy}
-\email{levyb3@rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Benjamin Sherman}
-\email{shermb@rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Muhammad Ishaq}
-%\authornote{This work was done in part while the author was at RPI.}
-\email{ishaqm@purdue.edu}
-%\orcid{1234-5678-9012}
-%\author{G.K.M. Tobin}
-%\authornotemark[1]
-%\email{webmaster@marysville-ohio.com}
-\affiliation{%
-    \institution{Purdue University}
-    %\streetaddress{P.O. Box 1212}
-    \city{West Lafayette}
-    \state{Indiana}
-%    \postcode{43017-6221}
-}
-
-\author{Lindsey Kennard}
-\email{fire.elemental@gmail.com}
-\affiliation{%
-    \institution{STR}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Boston}
-    \state{Massachusets}
-    %\country{Iceland}
-}
-
-\author{Ana Milanova}
-\email{milanova@cs.rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Vassilis Zikas}
-%\authornote{This work was done in part while the author was visiting UCLA and supported in part by DARPA and SPAWAR under contract N66001-15-C-4065 and by a SICSA Cyber Nexus Research Exchanges grant.}
-\email{vzikas@purdue.edu}
-%\orcid{1234-5678-9012}
-%\author{G.K.M. Tobin}
-%\authornotemark[1]
-%\email{webmaster@marysville-ohio.com}
-\affiliation{%
-    \institution{Purdue University}
-    %\streetaddress{P.O. Box 1212}
-    \city{West Lafayette}
-    \state{Indiana}
-    %    \postcode{43017-6221}
-}
-
-%
-% By default, the full list of authors will be used in the page headers. Often, this list is too long, and will overlap
-% other information printed in the page headers. This command allows the author to define a more concise list
-% of authors' names for this purpose.
-%\renewcommand{\shortauthors}{Trovato and Tobin, et al.}
-
-\renewcommand{\A}{{\sf A}}
-\renewcommand{\B}{{\sf B}}
-\renewcommand{\C}{{\sf C}}
-\renewcommand{\D}{{\sf D}}
-
-% As a general rule, do not put math, special symbols or citations
-% in the abstract
-\begin{abstract}
-
-\ana{We need an abstract.}
-
-\end{abstract}
-
-%
-% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-% Please copy and paste the code instead of the example below.
-%
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10003752.10010124.10010138.10010143</concept_id>
-<concept_desc>Theory of computation~Program analysis</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10003752.10003777.10003789</concept_id>
-<concept_desc>Theory of computation~Cryptographic protocols</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10002978.10002979</concept_id>
-<concept_desc>Security and privacy~Cryptography</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Theory of computation~Program analysis}
-\ccsdesc[500]{Theory of computation~Cryptographic protocols}
-\ccsdesc[300]{Security and privacy~Cryptography}
-
-%
-% Keywords. The author(s) should pick words that accurately describe the work being
-% presented. Separate the keywords with commas.
-\keywords{multiparty computation; compilers; cryptography}
-
-\begin{comment}
-%
-% A "teaser" image appears between the author and affiliation information and the body
-% of the document, and typically spans the page.
-\begin{teaserfigure}
-    \includegraphics[width=\textwidth]{sampleteaser}
-    \caption{Seattle Mariners at Spring Training, 2010.}
-    \Description{Enjoying the baseball game from the third-base seats. Ichiro Suzuki preparing to bat.}
-    \label{fig:teaser}
-\end{teaserfigure}
-\end{comment}
-%
-% This command processes the author and affiliation and title information and builds
-% the first part of the formatted document.
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-
-Multi-party computation (MPC) allows $N$ parties $p_1,\dots, p_N$ to perform a computation on their private inputs securely.
-Informally, security means that the secure computation protocol computes the correct output (correctness) and it does not leak any information about the individual party inputs, other than what can be deduced from the output (privacy).
-
-% The first results from the early 80's~\cite{FOCS:Yao82b,STOC:GolMicWig87,STOC:BenGolWig88,STOC:ChaCreDam88} introduced the problem and formally defined its security, and investigate feasibility in different models---e.g., passive or malicious adversaries, unconditional or computational security, with or without a broadcast channel.
-This spawned a rich theoretical literature of beautiful works improving and extending in various manners the original ideas.
-
-MPC theory dates back to the early 1980-ies~\cite{FOCS:Yao82b,STOC:GolMicWig87,STOC:BenGolWig88,STOC:ChaCreDam88}. 
-Long the realm of theoretical cryptography, MPC has seen significant advances in programming technology in recent years. 
-These advances bring MPC closer to practice and wider applicability --- 
-MPC technology has been employed in real-world scenarios such as auctions~\cite{FC:BCDGJK09}, biometric identification~\cite{ESORICS:BlaGas11}, 
-and privacy-preserving machine learning~\cite{SP:MohZha17,CCS:MohRin18}. 
-The goal is to bring the technology to a level where programmers can write \emph{secure} and \emph{efficient} programs without 
-commanding extensive knowledge of cryptographic primitives.
-
-The problem, therefore, is to build a high-level programming language and a compiler, and there has been significant advance in 
-this space, e.g., ~\cite{CCS:BenNisPin08,ESORICS:BogLauWil08,CCS:ZhaSteBla13,Hastings:2019,Keller:2020,Acay:2021,Braun:2022} among other work. 
-Current research largely falls at the two ends 
-of the classical compiler: (1) work on \emph{front-end} language design and (2) work on \emph{back-end} protocol implementation. Work on language design 
-focuses on high-level constructs necessary to express multiple parties, computation by different parties, and information flow from one party 
-to another~\cite{SP:RasHamHic14, Acay:2021}. On the other end, work on protocol implementation focuses on cryptographic foundations and their efficient circuit-level
-implementation~\cite{NDSS:DemSchZoh15, CCS:ABFKLO18, Braun:2022}, e.g., implementation of operations (e.g., MUL, ADD) using different sharing protocols 
-(Boolean or Arithmetic GMW~\cite{STOC:GolMicWig87} or Yao's garbled circuits~\cite{FOCS:Yao82b}), as well as efficient share conversion from one representation 
-to another.
-
-Earlier compilers did both back-end and front-end translation without a specific focus on either, as their aim was to demonstrate applicability of MPC on real-world
-programming problems. As the field advanced, works have focused more closely on front-end language design (e.g., Wysteria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021}) 
-or back-end ``circuit-level'' design and implementation (e.g., MOTION~\cite{Braun:2022}).
-
-In this work we focus on an intermediate language and what we call \emph{backend-independent optimizations}, in a close analogy to \emph{machine-independent} optimizations
-in the classical compiler. The following figure summarizes our key idea:
-
-{\begin{center}
-\includegraphics[width=0.8\linewidth]{figs_paper_SIMD/focus.png}
-\end{center}
-}
-
-We emphasize the MPC Source intermediate representation and optimization over MPC Source. As in classical compilers, we envision different front ends (e.g., Wysteria,
-our front end IMP-MPC) compiling into MPC Source. MPC Source is particularly suitable for optimizations such as protocol mixing~\cite{Ishaq:2019, Fang:2022} and SIMD-vectorization, 
-which takes advantage of amortization at the circuit level. The MPC Source IR exposes the linear structure of MPC programs, which simplifies program analysis; this is
-in contrast to source, which has if-then-else constructs. In the same time, MPC Source is sufficiently ``high-level''  to support analyses and optimizations that take into account 
-control and data flow in a specific program. MPC Source is small in size and analysis is tractable, as opposed to analysis over an unrolled circuit~\cite{Ishaq:2019}. Again as
-in classical compilers, we envision translation of MPC Source (optimized or unoptimized) into MOTION, SPDZ, or other back-end code.
-
-%In contrast, the high-level program contains if-then-else and array writes, or the low-level circuit . 
-%The compilation problem, i.e., translation from a high-level language to circuits, is difficult and therefore, breaking the problem into sub-problems can  tractable.
-
-\subsection{Our Contribution} In this paper, we develop a compiler framework that takes a Python-like routine and produces MOTION code: we describe (a) the IMP-MPC language, its syntax and semantic restrictions, (b) translation from IMP-MPC into MPC Source, (c) a specific backend-independent optimization: novel SIMD-vectorization on MPC Source, and (d) translation from MPC Source into MOTION code.
-%\ana{I think we have to add more on MOTION, specifically mixing and amortization.}
-
-We focus on the MOTION framework as our back-end for several reasons.
-First, it is the state of the art in terms of performance~\cite{Braun:2022}.
-Second, it provides an API over efficient implementation for a wide variety of cryptographic operations in three different protocols ---  Arithmetic GMW, Boolean GMW, and BMR --- which allows for protocol mixing~\cite{Ishaq:2019, Fang:2022}, a known backend-independent optimization.
-Third, MOTION provides API for SIMD-level operations, which amortize cost and lead to significant improvement in memory footprint and
-throughput~\cite{NDSS:DemSchZoh15, CCS:ABFKLO18, Braun:2022}. It also enables MPC Source-level optimizations such SIMD vectorization.
-
-Our second contribution is an analytical model for cost estimation of amortized schedules. Originally, we hoped that optimal scheduling (under our model) was tractable, as the problem
-appeared simpler than the classical scheduling problem. Unfortunately, we show that optimal scheduling is NP-hard via a reduction to the Shortest Common Supersequence (SCS) problem. 
-Cost estimation is important as it drives not only vectorization but other optimizations such as protocol mixing optimizations and scheduling as well~\cite{Ishaq:2019, Fang:2022}.
-
-Our most important contribution is the implementation and evaluation of the compiler framework.
-We demonstrate expressivity of the source language by running the compiler on \ana{X} programs; these include classical MPC benchmarks such as PSI and Biometric matching, as well as kMeans, Histogram, and others.
-Vectorization leads to \ana{X}x improvement in number of gates on average, \ana{Y}x speedup in circuit generation time, \ana{Z}x speedup in the LAN setting, and \ana{Z}x speedup in the WAN setting using the Boolean GMW protocol.
-Results form BMR are similar.
-
-Our results emphasize the importance of backend-independent optimizations --- vectorization (described in this work) and protocol mixing (tackled in previous works~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022}) are two optimizations readily available at the level of MPC Source.
-We believe that our work can lead to future work on backend-independent compilation and optimization, bringing in new MPC optimizations and combinations of optimizations, much in the vein of standard compilers, and bring MPC programming technology closer to practice and wider applicability.
-
-\subsection{Outline}
-
-The rest of the paper is organized as follows.
-\secref{sec:overview} presents an overview of the compiler.
-\secref{sec:model} describes our model for cost estimation and argues NP-hardness of optimal scheduling and \secref{sec:compiler} describes the compiler.
-\secref{sec:experiments} presents implementation and evaluation.
-\secref{sec:related} discusses related work and \secref{sec:conclusion} concludes.
-
-\begin{comment}
-\begin{itemize}
-\item We define the scheduling problem for MPC. 
-We present an analytical model to reason about cost of schedules
-and show that scheduling is NP-hard via a reduction to the shortest 
-common supersequence problem.
-
-\item We present a compiler that takes an IMP-like high-level 
-program and produces amortized (i.e., vectorized) low-level cryptographic 
-code in the MOTION framework. Central contributions are 1) a novel compiler framework, 
-2) a vectorization algorithm that produces optimal schedules for a large 
-number of MPC programs, and 3) reasoning over output arrays without Array SSA; we remove 
-infeasible loop-carried dependences introduced by standard SSA to improve vectorization.
-
-\item We present an implementation and evaluation in the MOTION framework. 
-\ana{Fill in with final results. Mention benchmarks (standard + new ones + HyCC). }
-
-\end{itemize}
-\end{comment}
-
-\section{Overview}
-\label{sec:overview}
-
-\begin{table*}
-\begin{tabular}{ccc}
-\begin{minipage}{0.275\textwidth}
-{\small
-\begin{pythonn}
-def biometric(C: shared[list[int]], D: int, 
-      S: shared[list[int]], N: int) -> 
-      shared[tuple[int,int]]:
-   min_sum : int = MAX_INT
-   min_idx : int = 0  
-   for i in range(N): 
-      sum : int = 0
-      for j in range(D): 
-         # d = S[i,j] - C[j]
-         d : int = S[i * D + j] - C[j] 
-         p : int = d * d
-         sum = sum + p
-      if sum < min_sum:
-         min_sum : int = sum
-         min_idx : int = i
-   return (min_sum, min_idx)
-\end{pythonn}
-}
-\end{minipage}
-
-&
-
-\begin{minipage}{0.33\textwidth}
-{\tiny
-\begin{pythonn}
-min_sum!1 = MAX_INT
-min_idx!1 = 0
-for i in range(0, N):
-   min_sum!2 = PHI(min_sum!1, min_sum!4)
-   min_idx!2 = PHI(min_idx!1, min_idx!4)
-   sum!2 = 0
-   for j in range(0, D):
-      sum!3 = PHI(sum!2, sum!4)
-      d = SUB(S[((i * D) + j)],C[j]) 
-      p = MUL(d,d) 
-      sum!4 = ADD(sum!3,p) 
-   t = CMP(sum!3,min_sum!2) 
-   min_sum!3 = sum!3
-   min_idx!3 = i
-   min_sum!4 = MUX(t, min_sum!3, min_sum!2) 
-   min_idx!4 = MUX(t, min_idx!3, min_idx!2) 
-return (min_sum!2, min_idx!2)   
-\end{pythonn}
-}
-\end{minipage}
-
-&
-
-\begin{minipage}{0.33\textwidth}
-{\small
-\begin{pythonn}
-min_sum!1 = MAX_INT
-min_idx!1 = 0
-# S^ is same as S. C^ replicates C N times: 
-S^ = raise_dim(S, ((i * D) + j), (i:N,j:D)) #S^[i,j] = S[i,j]
-C^ = raise_dim(C, j, (i:N,j:D)) #C^[i,j] = C[j]
-
-sum!2[I] = [0,..,0]
-# computes _all_  "at once"
-d[I,J] = SUB_SIMD(S^[I,J],C^[I,J])
-p[I,J] = MUL_SIMD(d[I,J],d[I,J])
-
-for j in range(0, D):
-   # sum!2[I], sum!3[I], sum!4[I] are size-N vectors 
-   # computes N intermediate sums "at once"
-   sum!3[I] = PHI(sum!2[I], sum!4[I])       
-   sum!4[I] = ADD_SIMD(sum!3[I],p[I,j])
-
-min_idx!3[I] = [0,1,...N-1]   
-for i in range(0, N):
-   min_sum!2 = PHI(min_sum!1, min_sum!4) 
-   t[i] = CMP(sum!3[i],min_sum!2)
-   min_sum!4 = MUX(t[i], sum!3[i], min_sum!2)  
-for i in range(0, N):
-   min_idx!2 = PHI(min_idx!1, min_idx!4)    
-   min_idx!4 = MUX(t[i], min_idx!3[i], min_idx!2)
-return (min_sum!2, min_idx!2)   
-\end{pythonn}
-}
-\end{minipage}
-
-\\
-
-(a) IMP Source & (b) MPC Source & (c) Optimized MPC Source
-\end{tabular}
-\caption{Biometric Matching: From IMP Source to MPC Source to Optimized MPC Source. }
-\label{tab:source_and_MPC_source_and_optimized_MPC_source}
-\end{table*}
-
-\subsection{Source}
-
-As a running example, consider Biometric matching, a standard MPC benchmark.
-An intuitive (and naive) implementation is as shown in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(a).
-Array \python{C} is the feature vector of \texttt{D} features that we wish to match and \python{S} 
-is the database of \python{N} size-\python{D} vectors that we match against.
-
-\begin{comment}
-{\small
-\begin{verbatim}
-def biometric(C: shared[list[int]], D: int, 
-              S: shared[list[int]], N: int) -> 
-                      tuple[shared[int], shared[int]]:
-    min_sum = 10000
-    min_index = 0  
-    for i in range(N): #loop over database
-        sum = 0
-        for j in range(D): #loop over features
-            d = S[i * D + j] - C[j] #i.e., d = S[i,j] - C[j]
-            p = d * d
-            sum = sum + p
-        if sum < min_sum:
-            min_sum = sum
-            min_index = i
-    return (min_sum, min_index)
-\end{verbatim}
-}
-\end{comment}
-
-Our compiler takes essentially standard IMP~\ana{Need citation here.} 
-syntax and imposes certain semantic restrictions. (We detail the restrictions in the following sections.)
-The programmer writes an iterative program and annotate certain inputs 
-and outputs as \emph{shared}. In the example arrays \python{C} and \python{S}
-are shared, meaning that they store shares, however, the array sizes \python{D} and
-\python{N} respectively are plaintext. The code iterates over the entries in the 
-database and computes the Euclidean distance (its square actually) of the current 
-entry \python{S[i]} and \python{C}. The program returns the index of the vector that gives 
-the best match plus the corresponding sum of squares.
-
-\subsection{MPC Source and Cost of Schedule}
-
-Our compiler generates an intermediate representation, MPC Source. MPC Source 
-for Biometric Matching is shown in 
-Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(b). First, MPC Source is an SSA form. 
-Second, it is linear. The conditional in lines 13-14 in IMP Source turns into the linear code in lines 12-16 in MPC Source. 
-The test turns into a CMP operation (\python{t = CMP(sum!3,min\_sum!2)}), followed by the 
-true-branch sequence, followed by the MUX operations. The first MUX operation selects the value
-of \python{min\_sum}: if \python{t} is true, then \python{min\_sum} gets the value of the second multiplexer
-argument,  \python{min\_sum!3}, otherwise it takes the value of the third argument, \python{min\_sum!2}.  
-Third, MPC Source is a special form of SSA. The SSA $\phi$-nodes at the if-then-else (lines 13-15) turn into 
-MUX operations, while the $\phi$-nodes at for loops turn into what we call pseudo PHI nodes with a straightforward semantics.
-
-\begin{comment}
-{\small
-\begin{pythonn}
-1. min_sum!1 = 10000
-2. min_index!1 = 0
-3. for i in range(0, N!0):
-4.    min_sum!2 = PHI(min_sum!1, min_sum!4)
-5.    min_index!2 = PHI(min_index!1, min_index!4)
-6.    sum!2 = 0
-7.    for j in range(0, D!0):
-8.      sum!3 = PHI(sum!2, sum!4)
-9.      d!3 = (S!0[((i * D!0) + j)] - C!0[j]) // MPC
-10.     p!3 = (d!3 * d!3) // MPC
-11.     sum!4 = (sum!3 + p!3) // MPC
-12.  !1!2 = (sum!3 < min_sum!2) // MPC
-13.  min_sum!3 = sum!3
-14.  min_index!3 = i
-15.  min_sum!4 = MUX(!1!2, min_sum!3, min_sum!2) // MPC
-16.  min_index!4 = MUX(!1!2, min_index!3, min_index!2) // MPC
-17.!2!1 = (min_sum!2, min_index!2)   
-\end{pythonn}
-}
-\end{comment}
-
-%The compiler linearizes the source turning conditionals into MUX statements. The PHI nodes
-%are remnants of the SSA IR; the compiler generates code that picks the correct value 
-%when producing MOTION output; the MOTION framework in turn linearizes 
-%loops when it generates the circuit.
-
-We turn to our analytical model to compute the \emph{cost} of the program. Assuming fixed 
-cost $\beta$ for a local MPC operation (essentially just ADD) and cost $\alpha$ for a remote
-MPC operation (e.g., MUX, CMP, and remaining operations), the cost of the iterative schedule 
-will be $ND(2\alpha+\beta) + N(3\alpha$).
-
-A key contribution is the vectorizing transformation. We can compute all $N*D$ 
-subtraction operations (line 9) in a single SIMD instruction; similarly we can compute 
-all multiplication operations (line 10) in a single SIMD instruction. And while we cannot
-vectorize computation of the N individual sums, we can compute the N sums in parallel. 
-Our compiler \emph{automatically detects these opportunities and transforms the program}. 
-It is standard that MPC researchers write vectorized versions of the Biometric
-program by hand; we are the first (to the best of our knowledge) to automatically 
-transform a naive, iterative MPC program into an unintuitive vectorized one.
-
-\subsection{Vectorized MPC Source and Cost of Schedule}
-
-Our compiler produces the vectorized program shown in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(c). 
-Note that this is still our intermediate representation, Optimized MPC Source. Subsequently,
-the compiler turns this code into MOTION variables, loops and SIMD primitives, which MOTION then 
-uses to generate the circuit.
-
-The compiler determines that SUB and MUL in ``naive'' MPC Source (lines 9 and 10 in (b))
-can be fully vectorized into the SIMD SUB and MUL in optimized MPC Source (lines 9 and 10 in (c)). 
-Notation \python{p[I,J]} denotes a 2-dimensional array with fully vectorized dimensions.
-The computation of sum (line 11 in (b))
-is sequential across the $j$-dimension, but it is parallel across the $i$-dimension. 
-The loop in lines 12-16 in (c) illustrates; here \python{p[I,j]} refers to the $j$-th column in \python{p}. 
-Unfortunately, CMP and MUX remain sequential.
-
-\begin{comment}
-{\small
-\begin{verbatim}
-min_sum!1 = 10000
-min_index!1 = 0
-// S!0^ is same as S!0. C!0^ replicates C!0 N-times: 
-S!0^ = raise_dim(S!0, ((i * D!0) + j), (i:N!0,j:D!0))
-C!0^ = raise_dim(C!0, j, (i:N!0,j:D!0))
-
-sum!2 = [0,..,0]
-// Computes all differences and all products "at once"
-d!3[I,J] = SUB_SIMD(S!0^[I,J],C!0^[I,J])
-p!3[I,J] = MUL_SIMD(d!3[I,J],d!3[I,J])
-
-for j in range(0, D!0):
-  // sum!2[I], sum!3[I], sum!4[I] are vectors of size N  
-  // Computes N intermediate sums "at once"
-  sum!3[I] = PHI(sum!2[I], sum!4[I])       
-  sum!4[I] = ADD_SIMD(sum!3[I],p!3[I,j])
-
-min_index!3 = [0,1,...N!0-1]
-
-for i in range(0, N!0):
-  min_sum!2 = PHI(min_sum!1, min_sum!4) 
-  !1!2[i] = CMP(sum!3[i],min_sum!2)
-  min_sum!4 = MUX(!1!2[i], sum!3[i], min_sum!2)
-
-for i in range(0, N!0):
-  min_index!2 = PHI(min_index!1, min_index!4)  
-  min_index!4 = MUX(!1!2[i], min_index!3[i], min_index!2)
-!2!1 = (min_sum!2, min_index!2)   
-\end{verbatim}
-}
-\end{comment}
-
-In MPC backends, executing $n$ operations ``at once'' in a single SIMD operation costs a lot less than executing those $n$ operations one by one.
-This is particularly important when there is communication (i.e., in remote), since many 1-bit values are sent at once rather than sequentially. 
-%In MPC compilers a vectorized operation that computing M operations "at once" costs 
-%essentially the same ($\alpha$ or $\beta$) as an individual operation. 
-We elaborate on the cost model in the following section but for now consider that 
-each operation has a \emph{fixed} portion (does benefits from amortization) and 
-a \emph{variable} portion (does not benefit from amortization): $\alpha = \alpha_\mathit{fix} + \alpha_\mathit{var}$. 
-This gives rise to the following formula for amortized cost: $f(n) = \alpha_\mathit{fix} + n\alpha_\mathit{var}$, 
-as opposed to unamortized cost $g(n) = n\alpha_\mathit{fix} + n\alpha_\mathit{var}$. (We extend the same reasoning to 
-$\beta$-instructions.)
-
-Thus, the fixed cost of the vectorized program amounts to
-$2\alpha_\mathit{fix}$ + $D \beta_\mathit{fix} + N(3\alpha_\mathit{fix})$. (The variable cost is the same in both the vectorized and non-vectorized programs.) 
-The first term in the sum corresponds to the vectorized
-subtraction and multiplication, the second term corresponds to the FOR loop on $j$ and the third 
-one corresponds to the remaining FOR loops on $i$.
-Clearly, $2\alpha_\mathit{fix} + D\beta_\mathit{fix} + N(3\alpha_\mathit{fix}) << ND(2\alpha_\mathit{fix}+\beta_\mathit{fix}) + N3\alpha_\mathit{fix}$.
-Empirically, we observe 10x to 50x improvement over un-vectorized Biometric Matching in circuit generation time and setup time and 5x to 30x improvement in online time.
-Additionally, the un-vectorized version runs out of memory for $N=256$, while the vectorized one runs with the standard maximal input size $N=4,096$.
-\ana{Edit numbers with final experiments.}
-
-\section{Analytical Model}
-\label{sec:model}
-
-\ana{We need an intro to section here.}
-
-\subsection{Scheduling in MPC}
-\label{sec:mpc}
-
-%\ana{Here we need more on MPC before jumping to scheduling. MPC-source, basic assumptions about static loop bounds, etc.}
-
-For this treatment we make the following simplifying assumptions:
-
-\begin{enumerate}
-\item All statements in the program execute using the same protocol (sharing). That is, there is no share conversion. 
-%\ishaq{This is not an assumption, this is our setting. We work in the single protocol world.}
-\item There are two tiers of MPC instructions, local and remote. A local instruction (essentially just ADD) has cost $\beta$ 
-and a remote instruction (e.g., MUX, MUL, SHL, etc.) has cost $\alpha$, where $\alpha >> \beta$. We assume that all remote 
-instructions have the same cost.
-%\ishaq{Don't we need to distinguish between local (cheap) vs.
-interactive (expensive) instructions?
-I don't see how we could assume this.
--- Past suggestion from Vassilis: Ideally we should assume cost to be a convex function.}\ana{After discussion 1/12: replace single costs 1 with two levels of costs: $\alpha$ for non-local operations, e.g., MUL, and $\beta$ for local ops, e.g., ADD.}
-\item  In MPC frameworks, executing $n$ operations ``at once'' in a single SIMD operation costs a lot less than executing those $n$ operations one by one.
-Following Amdahl's law~\ana{citation!}, we write $\alpha = \frac{1}{s}p\alpha + (1-p)\alpha$, where $p$ is the fraction of execution time that benefits from amortization and $(1-p)$ 
-is the fraction that does not, and $s$ is the available resource. Thus, $n\alpha = \frac{n}{s}p\alpha + n(1-p)\alpha$. 
-For the purpose of the model we assume that $s$ is large enough and the term $\frac{n}{s}p\alpha$ amounts to a \emph{fixed cost} incurred regardless of
-whether $n$ is $10,000$ or just $1$. (This models the cost of sending a packet from party A to party B for example.) The bottom line is that amortized execution 
-of $n$ operations is $f(n) = \alpha_\mathit{fix} + n\alpha_{var}$ in contrast to unamortized execution $g(n) = n\alpha_\mathit{fix} + n\alpha_{var}$ and, of course,
-$\alpha_\mathit{fix} << n\alpha_\mathit{fix}$.
-%We assume infinite parallel capacity---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely $\alpha$ or $\beta$. 
-%This is a standard assumption in Cryptographic Parallel RAM. ABY presents empirical support for this assumption~\ana{Add citations. PRAM, ABY.}
-%\ishaq{Comment from Vassilis: We assume there are infinite parallel capacities, this is the standard assumption in PRAM (Cryptographic Parallel RAM).}\ana{PRAM assumption stays.
-We have to replace the unit cost of 1 with a suitable amortization function $f(n)$, which I don't think changes much in the cost modeling and analysis.}
-\item MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption,
-2 MUL instructions can be amortized in a single SIMD instruction that costs $\alpha_\mathit{fix} + 2\alpha_\mathit{var}$, however a MUL and a MUX instruction 
-still cost $2\alpha_\mathit{fix} + 2\alpha_\mathit{var}$ even when scheduled ``in parallel''. 
-%MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption,
-%2 MUL instructions scheduled in parallel benefit from amortization and cost $\alpha$, however a MUL and a MUX instructions scheduled
-%in parallel still cost $2\alpha$. 
-%\ishaq{We need to reword this assumption to something like "$K$ parallel MUL costs much less than $K$ (because they get amortized), but any mix of $K$  MUL and MUX still cost roughly $K$.
-Specifically, we should take away the low constants (2 and 1) because we know for low constants this is not true.} \ana{Again, core assumption that MUL and MUX don't benefit from amortization stays.
-We have to replace constant costs with functions, as in (3).}
-\end{enumerate}
-
-\input{prob_stmt_np_hardness.tex}
-
-\begin{figure*}
-  \includegraphics[width=1.0\linewidth]{figs_paper_SIMD/compiler_framework.png}
-  \caption{Compiler Framework. \ana{Change figure since we are not doing div-and-conquer.}}
-  \label{fig:compiler_framework}
-\end{figure*}
-
-\section{\bf Compiler Framework}
-\label{sec:preliminaries}
-
-%We assume arbitrarily nested loops in the MPC-source IR and read-only arrays.
-We assume that loops range from 0 to some constant N.
-Arrays are linearized (row-major order as in MOTION) and accesses are via functions of the induction variables of the enclosing loops.
-
-\figref{fig:compiler_framework} presents an overview of our compiler. This section describes the front-end phases of the compiler
-and Section~\secref{sec:vectorization} describes our vectorization algorithm. \ana{We'll need a section on MOTION code gen too.}
-%\ana{Need to state this more precisely.}
-
-\subsection{Syntax and Semantic Restrictions}
-\label{sec:syntax}
-
-Source syntax is essentially standard IMP syntax as shown below:
-%\begin{figure}[t]
-{\small
-\[
-\begin{array}{l@{~~~~~}l}
-  \begin{array}{l@{~}l@{~~~}l}
- e & ::= e \; \mathit{op} \; e \mid e \mid \x \mid  \texttt{const} \mid \texttt{A[$e$]} & \mathit{expression} \\
- % s &::= t \;\f & \mathit{field} \\
- %\md &::= t\;{\m}(t \;{\this}, \; t\;\x) \; \{\; \ol{t\;\y}\;\s;\;\return\;\y \; \} 
- %                 & \mathit{method} \\ 
-  s & ::= s;s \mid  & \mathit{sequence} \\
-  & \x = e \mid \texttt{A[$e$]} = e \mid & \mathit{assignment \; stmt} \\
-  & \texttt{for $i$ in range($I$)}: s \mid & \mathit{for \; stmt} \\
-  & \texttt{if $e$: $s$ else: $s$} & \mathit{if \; stmt} \\ 
-%   t &::= q \; {\sf C} & \mathit{qualified \; type}\\
-%   q &::= \high \mid \poly \mid \low & \mathit{FlowCFL \; qualifier}\\
-%    t' &::= q' \; \C & \mathit{field \; qualified \; type}\\
-%   q' &::= \tainted \mid \poly & \mathit{field \; qualifier}\\
-
-  \end{array}
-\end{array}
-\]
-}
-%\end{figure}
-
-The syntax allows for array accesses, arbitrarily nested loops, and if-then-else control flow. 
-Expressions are typed $q \; \tau$, where qualifier $q$ and base type $\tau$ are:
-
-{\small
-\[
-\begin{array}{l@{~~~~~}l}
-  \begin{array}{l@{~}l@{~~~}l}
- \tau & ::= \texttt{int}  \mid \texttt{bool} \mid \texttt{list[int]} \mid  \texttt{list[bool]} & \mathit{base \; types} \\
-  q & ::= \texttt{shared} \mid \texttt{plain} & \mathit{qualifiers} \\
-  \end{array}
-\end{array}
-\]
-}
-
-The type system is mostly standard, and in our experience, a sweet spot between readability and expressivity.
-The \texttt{shared} qualifier denotes shared values, i.e., ones shared among the parties and computed upon 
-under secure computation protocols; the \texttt{plain} qualifier denotes plaintext values. Shared lists denote 
-shared elements, but plaintext length. Subtyping is again as expected: $q_1 \; \tau <: q_2 \; \tau$ 
-iff $q_1 <: q_2$ where $\texttt{plain} <: \texttt{shared}$. Thus, we can convert a plaintext value into a shared one, 
-but not vice versa.
-
-Our compiler imposes certain semantic restrictions that it enforces throughout the various 
-phases of compilation. We note that in some cases, the restrictions 
-can be easily lifted and we plan to do so in future iterations of the work.
-
-\begin{enumerate}
-\item Loops are of the form $0 <= i < I$ and bounds are fixed at compile time. 
-It is a standard restriction in MPC that the bounds must be known at circuit-generation time.
-\item Arrays are one-dimensional. N-dimensional arrays are linearized and accessed 
-in row-major order and at this point the programmer is responsible for linearization
-and access. (This restriction can be easily lifted.)
-\item Array subscrpts are plaintext values as specified by the rule:
-\begin{semantics}
-    \ntyperule{Array Access}{
-%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
-  \Gamma \vdash e : \texttt{plain int} \quad \Gamma \vdash \texttt{A} : q \; \texttt{list}[\tau] \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
-}{
-  \Gamma \vdash \texttt{A[$e$]} : q \; \tau
-}
-\end{semantics}
-The subscript $e$ is a function of the indices of the enclosing loops. 
-For read access, the compiler allows an arbitrary such function. 
-However, it restricts write access to \emph{canonical writes}, i.e., \texttt{A[i,j,k]} = ...
-where \texttt{i}, \texttt{j} and \texttt{k} loop over the three dimensions of \texttt{A}.  
-Write access such as for example \texttt{A[i,j+2]} = ... is not allowed. 
-%\item \ana{Add rules for logical ops}
-%\item \ana{Add rules for arithmetic ops}
-\item The final restriction involves MUX as expressed by the rule: 
-\begin{semantics}
-    \ntyperule{MUX}{
-%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
-  \Gamma \vdash e_1 : q_1 \; \texttt{bool} \quad \Gamma \vdash e_2 : q_2 \; \tau \quad \Gamma \vdash e_2 : q_2 \; \tau \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
-}{
-  \Gamma \vdash \texttt{MUX($e_1$,$e_2$,$e_3$)} : q_1 \vee q_2 \vee q_3 \; \tau
-}
-\end{semantics}
-\end{enumerate}
-Importantly, the arguments of MUX are restricted to base types, as passing array arguments 
-is not valid MOTION MPC. \ana{Check this.} This causes a slight inconvenience as we could not write 
-\begin{pythonn}
-if e: A[i] = val
-\end{pythonn}
-Instead we had to write
-\begin{pythonn}
-if not(e): val = A[i]
-A[i] = val
-\end{pythonn}
-
-We note that while the programmer writes 
-annotations at the level of IMP Source (recall \ana{ref to overview section}), the 
-annotations propagate through the transformations and are checked at the 
-level of MPC Source.
-
-%\ana{Disadvantage of MUX}
-
-For the rest of this section we write $i,j,k$ to denote the loop nest: $i$ is the outermost loop, $j$, is immediately nested in $i$, and so on until $k$
-and we use $I,J,K$ to denote the corresponding upper bounds. We write $A[i,j,k]$ to denote canonical access 
-to an array element. In the program, canonical access is achieved via the standard row-major order formula: $(J*K)*i + K*j + k$.
-To simplify the presentation we describe our algorithms in terms of three-element tuples $i,j,k$, however, discussion easily generalizes to 
-arbitrarily large loop nests.
-
-\subsection{From IMP Source to SSA}
-
-Our compiler translates from Source to SSA as follows:
-
-\begin{enumerate}
-  \item
-        \textbf{Parsing}:
-        Use Python's \texttt{ast} module to parse the input source code to a Python AST.
-  \item
-        \textbf{Syntax checking}:
-        Ensure that the AST matches the restricted subset defined in Section~\secref{sec:syntax}.
-        This step outputs an instance of the \texttt{restricted\_ast.Function} class, which represents our restricted subset of the Python AST.
-  \item
-        \textbf{3-address CFG conversion}:
-        %\benjamin{TODO: Is this a good amount of detail?} \ana{This is good amount of detail for the Tech report. In the actual paper we'll shorten a lot and add pointers to the Tech report.}
-
-        Convert the restricted-syntax AST to a three-address control-flow graph.
-        To do this, first,
-        add an empty basic block to the CFG and mark it as current.
-        Next, for each statement in the restricted AST's function body,
-        process the statement.
-        Statements can either be for-loops, if-statements, or assignments (as in~\secref{sec:syntax}).
-        Rules for processing each kind of statement are given below:
-        \begin{enumerate}
-          \item
-                \textbf{For-loops}:
-                Create new basic blocks for
-                the loop condition
-                (the \emph{condition-block}),
-                the loop body
-                (the \emph{body-block}),
-                and the code after the loop
-                (the \emph{after-block}).
-                Insert a jump from the end of the current block to the condition-block.
-                Then, mark the condition-block as the current block.
-                Insert a for-instruction at the end of the current block with the loop counter variable and bounds from the AST.
-                Next, add an edge from the current block to the after-block labeled ``FALSE'' and
-                an edge from the current block to the
-                body-block labeled ``TRUE''.
-                Then, set the body-block to be the current block
-                and process all statements in the AST's loop body.
-                Finally, insert a jump to the condition-block and set the after-block as current.
-          \item
-                \textbf{If-statements}:
-                Create new basic blocks for
-                the ``then'' statements of the if-statement
-                (the \emph{then-block}),
-                the ``else'' statements of the if-statement
-                (the \emph{else-block}),
-                and the code after the if-statement
-                (the \emph{after-block}).
-                At the end of the current block,
-                insert a conditional jump to the then-block or else-block
-                depending on the if-statement condition in the AST.
-                Next,
-                mark the then-block as current,
-                process all then-statements,
-                and add a jump to the after-block.
-                Similarly,
-                mark the else-block as current,
-                process all else-statements,
-                and add a jump to the after-block.
-                Finally, set the after-block to be the current block,
-                and give it a \emph{merge condition} property equal to the condition of the if-statement.
-          \item
-                \textbf{Assignments}:
-                In the restricted-syntax AST,
-                the left-hand side of assignments
-                can be a variable or an array subscript.
-                If it is an array subscript, e.g., \texttt{A[i] = x},
-                change the statement to \texttt{A = Update(A, i, x)}.
-                If the statement is not already three-address code,
-                for each sub-expression in the right-hand side of the assignment,
-                insert an assignment to a temporary variable.
-        \end{enumerate}
-  \item
-        \textbf{SSA conversion}:
-        Convert the 3-address CFG to SSA with Cytron's algorithm.
-\end{enumerate}
-
-\subsection{From SSA to MPC Source} %: MUX Nodes and Pseudo $\phi$-nodes}
-
-Once the compiler converts the code to SSA,
-it transforms $\phi$-nodes that correspond to if-statements into MUX nodes.
-From the 3-address CFG conversion step,
-$\phi$-nodes corresponding to if-statements will be in a basic block
-with the merge condition property.
-For example, if $X_3 = \phi(X_1,X_2)$ is in a block with merge condition $C$,
-the compiler transforms it into $X_3 = \mathit{MUX}(C, X_1, X_2)$.
-Next, the compiler runs the dead code elimination algorithm from Cytron's SSA paper.
-
-Next, the control-flow graph is \emph{linearized} into MPC Source,
-which has loops but no if-then-else-statements.
-This means that both branches of all if-statements are executed,
-and the MUX nodes determine whether to use results from the then-block or from the else-block.
-The compiler linearizes the control-flow graph with a variation of breadth-first search.
-Blocks with the ``merge condition'' property are only considered
-the second time they are visited,
-since that will be after both branches of the if-statement are visited. 
-(The Python AST naturally gives rise to a translation where each conditional has exactly two targets, 
-and each ``merge condition'' block has exactly two incoming edges, a TRUE and a FALSE edge.
-Thus, each $\phi$-node has exactly two multiplexer arguments, which dovetails into MUX.
-This is in contrast with Cytron's algorithm which operates at the level of the CFG and allows for 
-$\phi$-nodes with multiple arguments.)
-Each time the compiler visits a block,
-it adds the block's statements to the MPC source.
-If the block ends in a for-instruction,
-the compiler recursively converts the body and code after the loop to MPC source
-and adds the for-loop and code after the loop to the main MPC source.
-If the block does not end in a for-instruction,
-the compiler recursively converts all successor branches to MPC source and
-appends these to the main MPC source.
-
-Now, the remaining $\phi$-nodes in MPC source are the loop header nodes. We call these nodes \emph{pseudo} $\phi$-nodes 
-and we write \python{PHI} in MPC Source. A pseudo $\phi$-node $X_1 = $\python{PHI}$(X_0,X_2)$ in a loop header is evaluated 
-during circuit generation. If it is the 0-th iteration, then the $\phi$-node evaluates to $X_0$, otherwise, it evaluates to $X_2$.
-
-%\ana{We need to double check that the following captures all cases of def-use edges.}
-
-\subsection{Dependence Analysis}
-
-The next steps construct the dependence graph and \ana{Add an intro here.}
-
-\subsubsection{Def-use Edges}
-
-The dependence graph has the following def-use edges:
-
-\begin{itemize}
-\item same-level edge $X\rightarrow Y$ where $X$ and $Y$ are in the same loop nest, say $i,j,k$.
-E.g., the def-use edge 9 to 10 in the Biometric MPC Source in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source} is a same-level edge.
-A same-level edge can be a back-edge in which case a $\phi$ node is the target of the edge.
-E.g., 15 to 4 in Biometric is a same-level back-edge.
-\item outer-to-inner $X\rightarrow Y$ where $X$ is in an outer loop nest, say $i$, and $Y$ is in an inner one, say $i,j,k$. E.g., 1 to 4 in Biometric forms is an outer-to-inner edge.
-%A $\phi$-node can be a source or a target of an outer-to-inner forward edge.
-\item inner-to-outer $X\rightarrow Y$ where $X$ is a $phi$-node in an inner loop nest, $i,j,k$, and $Y$ is in the enclosing loop nest $i,j$.
-E.g., the def-use from 8 to 12 gives rise to an inner-to-outer edge.
-%E.g. $\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$ is an inner-to-outer edge. 
-An inner-to-outer edge can be a back-edge as well, in which case both $X$ and $Y$ are phi-nodes with the source $X$ in a loop nested into $Y$'s loop (not necessarily immediately).
-
-%Note that the source is \emph{always} a $\phi$-node of the $i,j,k$ loop and the target is in the immediately enclosing loop.
-The interpretation of this edge is that the use node $Y$ uses the definition made in the last iteration of the inner loop.
-%\ben{The current representation of pseudo $\phi$-nodes shows them attached to the loop header (i.e. in the loop).  We may want to clarify that they are also evaluated at the loop's termination.}
-%\item same-level back-edge $X\rightarrow Y$.
-$Y$ is a phi-node in the header of the loop and $X$ is a definition of the variable in the loop body.
-E.g., $\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{sum}_1,\texttt{min}_1)$ to $\texttt{min}_0 = \phi(\texttt{min}_1, 10000)$ in Biometric is a same-level back-edge.
-%\item inner-to-outer back-edge $X\rightarrow Y$: $X$ and $Y$ are both $\phi$-nodes for some variable. The source $X$ is in a loop nested into $Y$'s loop (not necessarily immediately).
-
-\item mixed forward edge $X\rightarrow Y$.
-$X$ is in some loop $i,j,k$ and $Y$ is in a loop nested into $i,j,k'$.
-We transform mixed forward edges as follows.
-Let $\x$ be the variable defined at $X$.
-We add a variable and assignment $\x' = \x$ immediately after the $i,j,k$ loop.
-Then we replace the use of $\x$ at $Y$ with $\x'$.
-This transforms a mixed forward edge into an "inner-to-outer" forward edge followed by an outer-to-inner forward edge.
-Thus, Basic Vectorization handles one of "same-level", "inner-to-outer", or "outer-to-inner" def-use edges.
- \end{itemize}
-
-\subsubsection{Closures}
-
-We define $\mathit{closure}(n)$ where $n$ is a phi-node. Intuitively, it computes the set of nodes (i.e., statements) that form a dependence cycle with $n$. The closure of $n$ is defined as follows:
-\begin{itemize}
-\item $n$ is in $\mathit{closure}(n)$
-\item $X$ is in $\mathit{closure}(n)$ if there is a same-level path from $n$ to $X$, and $X \rightarrow n$ is a same-level back-edge.
-\item $Y$ is in $\mathit{closure}(n)$ if there is a same-level path from $n$ to $Y$ and there is a same-level path from $Y$ to some $X$ in $\mathit{closure}(n)$.
-\end{itemize}
-
-\subsection{Taint Analysis}
-
-We require that all inputs are marked as either shared or plaintext, however, we do not require any other qualifier annotations in the program.
-Then we infer if intermediate variables and expressions are shared through taint analysis with "taintedness" referring to the shared attribute.
-Specifically, the compiler follows the following rules, which are standard in positive-negative qualifier systems (here \texttt{shared} is the positive qualifier and \texttt{plain} is the negative one):
-\begin{enumerate}
-\item Loop counters and array upper bounds are always \texttt{plain}.
-\item If any variable on the right-hand side of an assignment is shared, then the assigned variable is \texttt{shared}.
-%\item If all variables on the right-hand side of an assignment are plaintext, then the assigned variable is plaintext \ana{Will be covered by the last rule.}
-\item Any variables that cannot be determined as shared via the above rules are \texttt{plain}.
-\end{enumerate}
-
-%The first two rules are standard for taint analysis, and the third rule follows from the MPC problem statement. \ben{Is the explanation for the third rule correct?}  \ana{Yes, that is exactly it!}
-The final rule is needed in examples like the following. In the below snippet \texttt{sum!2} and \texttt{sum!3} form a dependency cycle and there is no \texttt{shared} value that flows to either one. 
-They are inferred as plaintext. 
-%cannot be marked as plaintext through simple taint analysis:
-{\small
-\begin{verbatim}
-plaintext_array = [0, 1, 2, ...]
-sum!1 = 0
-for i in range(0, N):
-    sum!2 = PHI(sum!1, sum!3)
-    sum!3 = sum!2 + plaintext_array[i]
-\end{verbatim}
-}
-%\ben{I think the above example is unnecessary and could be replaced by an explanation of how "untainted" variables are implicitly plaintext, but I couldn't think of a way to phrase that.}\ana{I left the example, we can leave it in the TR, remove from paper.}
-
-When converting to MOTION code, any plaintext value used in the right-hand side of a shared assignment is converted to a shared value for that expression. 
-%\ben{Is this necessary to include?} \ana{Yes, that's important.}
-
-Importantly, the taint analysis works on MPC Source, which lacks if-then-else control flow. 
-This significantly simplifies the treatment as there is no need to handle conditionals and implicit flow.
-
-\subsection{From (Optimized) MPC Source to MOTION}
-\label{sec:MOTION}
-\ana{Ben, add the MOTION section here.}
-
-\section{Vectorization}
-\label{sec:vectorization}
-
-An important component of our algorithm is the scalar expansion to the corresponding loop dimensionality. 
-For example, \texttt{d = S[i * D + j] - C[j]} equiv. to \texttt{d = S[i,j] - C[j]}, which gave rise to \texttt{N*D} subtraction operations in the sequential schedule, 
-is lifted. The argument arrays \texttt{S} and \texttt{C} are lifted and the scaler \texttt{d} is lifted: \texttt{d[i,j] = S[i,j] - C[i,j]}.
-The algorithm then detects that the statement can be vectorized.
-
-There are three kinds of arrays (for now all kept internally as one-dimensional arrays, but that's under discussion).
-\begin{itemize}
-
-\item Scalars: These are scalar variables we lift into arrays for the purposes of vectorization. 
-For those, all writes are canonical writes and all reads are canonical reads. We will apply raise dimension when 
-a scalar is used in an inner loop (e.g., \texttt{sum0} in line 6 of the MPC source code will be raised to a 1-dimensional 
-array since \texttt{sum0} is used in the inner $j$-loop). Drop dimension applies as well; this happens when a scalar
-written in an inner loop is used outside of the loop (e.g., \texttt{sum0} for which the lifted inner loop computes 
-\texttt{D} values, but the outer loop only needs the last one.)
-
-\item Read-only input arrays: Read-only inputs. There are NO writes, while we may have non-canonical reads, $f(i,j,k)$. 
-Phase 1 of Basic vectorization will add raise dimension operation at the beginning of the function to lift these arrays, 
-and raise dimension may \emph{reshape} arrays. If there are multiple "views" of the input array, there would be multiple raise dimension statements to create
-each one of these views. The invariant is that at reads in loops, the reads of "views" of the original input array are canonical.
-Only raise dimension applies to these arrays, and only in the beginning of the program. For example, 1-dimensional array 
-\texttt{C} is lifted into 2-dimensional array \texttt{N x D} by copying the row \texttt{N} times.
-
- \item Read-write output arrays: Writes are canonical (by restriction) but reads can be non-canonical. 
- Dependence analysis limits vectorization when non-canonical read access refers to array writes in
- previous iterations, thus creating loop-carried dependences. We may apply both raise and drop dimension, however, 
- they respect the fixed dimensionality of the output array. The array cannot be raised to a dimension lower than
- its canonical (fixed) dimensionality and it cannot be dropped lower. In addition, non-canonical reads may 
- require lifting (i.e., reshaping) of the array after the most recent write, rather than in the beginning of the program, 
- in order to reduce a non-canonical read to a canonical one.
-
-\end{itemize}
-
-In the sections below we detail the $\mathit{raise\_dim}$ (raise dimension) and $\mathit{drop\_dim}$ (drop dimension) operations,
-followed by our vectorization algorithm.
-
-%\ana{Add back-edges into Phase 1.
-A back-edge from a non-phi-node in loop $i$ to a phi-node in loop $i$'s header is a same-level edge.
-The only difference with the handling of normal same-level forward def-use is that the operand index will become $i-1$.
-A back edge from a phi-node in loop $j$ to a phi-node in loop $i$, where $j$ is nested in $i$ is an inner-to-outer edge and will require dropping dimension.
-We can show these are the only kinds of back-edges that may occur.}
-
-%\subsection{Algorithm (Still work in progress.)}
-
-\subsection{Raise Dimension and Drop Dimension}
-
-There are two conceptual versions of $\mathit{raise\_dim}$. One applies on read-only input arrays and reshapes those arrays 
-when necessary to ensure canonical read access in the corresponding loop. The signature of $\mathit{raise\_dim}$
-is as follows. It takes the original array \texttt{C}, the access pattern function $f(i,j,k)$ in loop nest $i,j,k$ and the loop bounds
-\texttt{((i:I),(j:J),(k:K))}: 
-\[ \mathit{raise\_dim}(\texttt{A}, f(i,j,k), \texttt{((i:I),(j:J),(k:K))}) \]
-It produces a new 3-dimensional array \texttt{A'} by iterating over $i,j,k$ and setting each element of  \texttt{A'} as follows:
-\[ \texttt{A'[i,j,k] =  A[f(i,j,k)]} \]
-The end result is that uses of \texttt{[A[f(i,jk)]} in loop nest $i,j,k$ are replaced with canonical read-accesses to \texttt{A'[i,j,k]}
-that can be vectorized. In the running Biometric example, \texttt{C'} = $\mathit{raise\_dim}$\texttt{(C, j, (i:N,j:D))} lifts the 
-1-dimensional array \texttt{C} into a 2-dimensional array. The $i,j$ loop now accesses \texttt{C'} in the canonical way, \texttt{C'[i,j]}. 
-Similarly, \texttt{S'} = $\mathit{raise\_dim}$\texttt{(S, i*D+j, (i:N,j:D))} tries to lift \texttt{S}, but the operation turns into a no-op 
-because \texttt{S} is already a 2-dimensional array and the read access is canonical.
-
-The other version of $\mathit{raise\_dim}$ applies on scalars and read-write arrays. It lifts a lower-dimension array into a 
-higher-dimension for access in a nested loop. Here \texttt{A} is an $i$ array and raise dimension adds two additional dimensions:
-\[ \mathit{raise\_dim}(\texttt{A}, \texttt{(j:J,k:K)}) \]
-This version is reduced to the above version by adding the access pattern function, which is just $i$: 
-\[ \mathit{raise\_dim}(\texttt{A}, i, \texttt{(j:J,k:K)}) \]
-
-The corresponding $\mathit{drop\_dim}$ is carried out when an array written in an inner loop is used in an enclosing loop. 
-It takes a higher dimensional array, say $i,j,k$ and removes trailing dimensions, say $j,k$:
-\[ \mathit{drop\_dim}(\texttt{A}, \texttt{(j:J,k:K)}) \]
-It iterates over $i$ and takes the result at the maximal index of $j$ and $k$, i.e., the result at the last iterations of $j$ and $k$:
-\[ \texttt{A'[i] = A[i,J-1,K-1]} \]
-
-\begin{comment}
-We define the $\mathit{raise\_dim}$ (raise dimensions) and $\mathit{drop\_dim}$ (drop dimension) functions.
-Raise dimension "lifts" a lower-dimension array (\ana{The right term here is tensor, I am pretty sure!}) into a higher dimension one.
-This is necessary when a lower-dimensional array is used in a higher dimensional loop and, essentially, is just copying of values.
-For example, Biometric contains the statement $d = \mathit{SUB}(S[i, j], C[j])$, in the $j$-loop which is nested into the $i$-loop.
-Here $C$ is a one-dimensional array, however, to vectorize across both loops it is necessary to turn it into a two-dimensional array: $C[i,j]$ becomes $[C[0],C[1],...C[J],C[0],C[1],...C[J], ...
-C[0],C[1],...C[J]]$, which turns the row into a matrix of $I$ identical rows.
-(We use capital letters to denote the upper bounds of loops, e.g., $J$ is the upper bound of the $j$-loop and $I$ is the upper bound of the $i$-loop.)
-
-$\mathit{raise\_dim}(A[i_1,...,i_k],i_j)$ is defined as follows.
-It results in a new {k+1}-dimensional array $A'$ where for every $0 \le i_{k} < I_{k+1}$, $A'[i_1, \ldots, i_j, \ldots, i_k] = A[i_1, \ldots, i_k]$.
-Adding $n$ dimensions is trivially extended as a composition of $n$ $\mathit{raise\_dim}$ that each adds a single dimension.
-
-As expected, drop dimension turns a higher-dimensional array into a lower-dimensional one.
-The key use case is an inner-to-outer def-use edge.
-The code may define a variable, e.g., $\x$ in an inner loop, say $j$, then use this variable in an enclosing loop, say $i$.
-Our algorithm may vectorize the computation of $\x$ in the $j$-loop thus producing a vector $\x[j]$ where $\x[1]$ is the value of \x after the 1st iteration and so on.
-Drop dimension states that the outer loop will use the value of the variable at the last iteration.
-
-$\mathit{drop\_dim}(A[i_1,\ldots, i_j, \ldots,i_k], i_j)$ produces a k-dimensional array $A'$ where $A'[i_1,\ldots,i_k] = A'[i_1,\ldots, I_j-1, \ldots,i_k]$.
-In our analysis, dimensions are always dropped at the end, and again, one can define dropping $n$ dimensions as a composition of $n$ $\mathit{drop\_dim}$.
-
-\ben{I'm not sure if I wrote up the above paragraphs correctly -- the idea is that we're just copying values to extend over dimension $j$ or only retain its last elements.
-The current writeup implies that $1 < j < k$ though, which isn't always true.
-Python implementations of $\mathit{raise\_dim}()$ and $\mathit{drop\_dim}()$ can be found here: \url{https://github.com/milana2/ParallelizationForMPC/blob/master/compiler/compiler/motion_backend/reference_implementations.py}}
-
-\ana{This section is still informal. Needs work to make more precise.}
-\end{comment}
-
-\subsection{Basic Vectorization}
-
-\begin{algorithmic}
-
-\STATE \COMMENT { Phase 1: Raise dimension of scalar variables to corresponding loop nest. We can traverse stmts linearly in MPC-source. }
-
-\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i,j,k$}
-%\FOR {each $Y_n$}
-\FOR {each argument $Y_n$} %def-use edge $\mathit{stmt}' (\mbox{def of } Y) \rightarrow \mathit{stmt} (\mbox{def of } X)$}
-\STATE {case def-use edge $\mathit{stmt}' (\mbox{def of } Y_n) \rightarrow \mathit{stmt} (\mbox{def of } X)$ of} \\
-\STATE \hspace{0.25cm} {same-level: $Y'_n$ is $Y_n$} \\
-\STATE \hspace{0.25cm} {outer-to-inner: add $Y'_n[i,j,k] = \mathit{raise\_dim}(Y_n)$ at $\mathit{stmt'}$}
-\STATE \hspace{0.25cm} {(more precisely, right after $\mathit{stmt'}$)}
-\STATE \hspace{0.25cm} {inner-to-outer: add $Y'_n[i,j,k] = \mathit{drop\_dim}(Y_n)$ at $\mathit{stmt}$}
-\STATE \hspace{0.25cm} {(more precisely, in loop of $\mathit{stmt}$ right after loop of $\mathit{stmt'}$)}
-\ENDFOR
-
-\COMMENT { Optimistically vectorize all. $I$ means vectorized dimension. }
-%\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i,j,k$}
-\STATE {change to $X[I,J,K] = \mathit{Op}(Y'_1[I,J,K],Y'_2[I,J,K])$ }
-\ENDFOR
-
-\COMMENT { Phase 2: Recreating FOR loops for cycles; vectorizable statements hoisted up. }
-
-\FOR {each dimension $d$ from highest to 0}
-\FOR {each $\phi$-node $n$ in loop $i_1,...,i_d$}
-\STATE {compute $\mathit{closure}(n)$}
-\ENDFOR
-
-\COMMENT {{\color{red} $cl_1$ and $cl_2$ intersect if they have common statement or update same array; "intersect" definition can be expanded}}
-
-\WHILE {there are closure $cl_1$ and $cl_2$ that intersect}
-\STATE {merge $cl_1$ and $cl_2$}
-\ENDWHILE
-\FOR {each closure $cl$ (after merge)}
-\STATE {create FOR $i_d = 0; ...$ loop}
-\STATE {add $\phi$-nodes in $cl$ to header block}
-\STATE {{\color{red}{add target-less $\phi$-node for A if $cl$ updates array A}}}
-\STATE {add statements in $cl$ to loop body in some order of dependences}
-\STATE \COMMENT { Dimension is not vectorizable: }
-\STATE {change ${I_d}$ to $i_d$ in all statements in loop}
-\STATE {treat FOR loop as monolith node: some def-use edges become same-level.}
-\ENDFOR
-{
-\color{red}
-\FOR {each target-less $\phi$-node $A_1 = \phi(A_0,A_k)$}
-\STATE {in vectorizable stmts, replace use of $A_1$ with $A_0$}
-\STATE {discard $\phi$-node if not used in any $cl$, replacing $A_1$ with $A_0$ or $A_k$ appropriately}
-\ENDFOR
-}
-\ENDFOR
-
-\COMMENT { Phase 3: Remove unnecessary dimensionality. }
-
-\COMMENT { A dimension $i$ is dead on exit from stmt $X[...i...] = ...$ if all def-uses with targets outside of the enclosing 
-FOR $i=0 ...$ MOTION loop end at target (use) $X' = \mathit{drop\_dim}(X, i)$. } 
-\FOR {each stmt and dimension $X[...i...] = ...$}
-\STATE {if $i$ is a dead dimension on exit from stmt $X[...i...] = ...$, remove $i$ from $X$ (all defs and uses)} 
-%\COMMENT{ A stmt v[...i] is a "dead dimension i" if all edges out of FOR i = 0; ... loop end at drop\_dim(v, i)}
-\ENDFOR
-
-\COMMENT { Now clean up drop\_dim and raise\_dim }
-\FOR {each $X' =  \mathit{drop\_dim(X,i)}$ }
-\STATE {replace with $X' = X$ if $i$ is dead in $X$.}
-\ENDFOR
-\STATE {do (1) (extended) constant propagation, (2) copy propagation and (3) dead code elimination to get rid of redundant variables and raise and drop dimension statements}
-
-\COMMENT { Phase 4: }
-\STATE {add SIMD for simdified dimensions}
-
-\end{algorithmic}
-
-\input{biometric_vectorization_example.tex}
-
-\begin{comment}
-\subsection{Example: Biometric}
-
-\subsubsection{MPC-source}
-
-Below is the MPC-source resulting from Benjamin's analysis. It gives rise to a corresponding dependence graph. (Not shown.)
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0 = \phi(-1, \texttt{index}_1)$}
-\STATE {$\texttt{min}_0 = \phi(10000,\texttt{min}_1)$}
-
-\STATE \COMMENT {Begin of inner loop $j$.}
-\STATE {$\texttt{sum}_0 = \phi(0, \texttt{sum}_1)$}
-\STATE {$\texttt{d} = \mathit{SUB}(\texttt{S}[i,j], \texttt{C}[j])$}
-\STATE {$\texttt{p} = \mathit{MUL}(\texttt{d},\texttt{d})$}
-\STATE {$\texttt{sum}_1 = \texttt{ADD}(\texttt{sum}_0,\texttt{p})$}
-\STATE \COMMENT {End of inner loop $j$.}
-\STATE {$\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$}
-\STATE {$\texttt{index}_1 = \mathit{MUX}(\texttt{c},\texttt{index}_0,i)$}
-\STATE {$\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{min}_0,\texttt{sum}_0)$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-\subsubsection{Phase 1 of Basic Vectorization:}
-
-The transformation preserves the dependence edges.
-It raises the dimensions of scalars and optimistically vectorizes all operations.
-The next phase discovers loop-carried dependences and removes affected vectorization.
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$}
-\STATE {$\texttt{min}_0[i] = \phi(10000, \texttt{min}_1[i-1])$}
-
-\STATE \COMMENT {Begin of inner loop $j$.}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...], \texttt{sum}_1[i,j-1])$}
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{sum}_1[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE \COMMENT {End of inner loop $j$.}
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {$\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$}
-\STATE {$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$}
-\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-\ishaq{At least for input variables: when raising dimensions, we should keep some meta data around so we can later write code that could use 1 share if it knows the other N shares are just duplicates of it.}
-
-\subsubsection{Phase 2}
-
-This phase analyzes statements from the innermost loop to the outermost. The key point is to discover loop-carried dependencies and re-introduce loops whenever dependencies make this necessary.
-
-Starting at the inner phi-node $\texttt{sum}_0[i,j] = \phi([0,0,...],\texttt{sum}_1[i,j-1])$, the algorithm first computes its closure.
-The closure amounts to the phi-node itself and $\texttt{sum}_1[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$, accounting for the loop-carried dependency of the computation of \texttt{sum}.
-The algorithm replaces this closure with a FOR loop on $j$ removing vectoriaztion on $j$.
-Note that the SUB and MUL computations remain outside of the loop as they do not depend on phi-nodes that are part of cycles.
-The algorithm adds same-level edges, one from $\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = ...$ to the monolithic FOR-loop node, and one from the FOR-loop node to $\texttt{c}[\stackrel{\rightarrow}{i}] = ...$.
-
-\ishaq{For our radar: Such dependencies will not involve phi nodes only. e.g. $a[i] = a[i] + a[i-1]$}
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0[i] = \phi(-1, \texttt{index}_1[i-1])$}
-\STATE {$\texttt{min}_0[i] = \phi(10000, \texttt{min}_1[i-1])$}
-
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-
-\STATE \COMMENT {Begin of inner loop $j$. Loop is now much "shorter"!}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...], \texttt{sum}_1[i,j-1])$}
-\STATE {FOR j=0; j<D; j++} \COMMENT{ Will turn into a MOTION loop.}
-\STATE \hspace{0.25cm}{$\texttt{sum}_1[\stackrel{\rightarrow}{i},{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},{j}],\texttt{p}[\stackrel{\rightarrow}{i},{j}])$}
-\STATE \COMMENT {End of inner loop $j$.}
-
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {$\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$}
-\STATE {$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$}
-\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-Next, analysis moves to outer dimension $i$. There are two phi-nodes, $\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1])$ and $\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$.
-The closure of the first is
-\[
-\begin{array}{l}
-\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1]) \\
-\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
-\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])
-\end{array}
-\]
-and the closure of the second one is
-\[
-\begin{array}{l}
-\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1]), \\
-%\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
-\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])
-\end{array}
-\]
-
-Since the two closures \emph{do not} intersect, we have two distinct FOR-loops on $i$.
-The first FOR loop (on \texttt{min}) is scheduled first, because of the dependence edge from $\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$ to
-$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$.
-We can now rewrite those statements, getting rid of the vectorization on $i$ in the two FOR loops.
-Notably, the SUB and MUL computations are fully vectorizable and the ADD computation is vectorizable across the $i$-dimension.
-
-\begin{algorithmic}
-
-\STATE
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB\_SIMD}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL\_SIMD}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$} \\
-\STATE {}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...],\texttt{sum}_1[i,j-1])$}
-\STATE {FOR j=0; j<D; j++} \COMMENT{ Turns into a MOTION loop, was $j$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{sum}_1[\stackrel{\rightarrow}{i},{j}] = \mathit{ADD\_SIMD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},{j}],\texttt{p}[\stackrel{\rightarrow}{i},{j}])$}
-\STATE {}
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {}
-\STATE {$\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1])$}
-\STATE {FOR i=0; i<N; i++} \COMMENT{ Turns into a MOTION loop, was $i$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{c}[{i}] = \mathit{CMP}(\texttt{sum}'_0[{i}],\texttt{min}_0[{i}])$}
-\STATE \hspace{0.25cm}{$\texttt{min}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{min}_0[{i}],\texttt{sum}'_1[{i}])$}
-\STATE {}
-\STATE {$\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$}
-\STATE {FOR i=0; i<N; i++} \COMMENT{ Turns into a MOTION loop, was $i$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{index}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{index}_0[{i}],i)$}
-
-\ben{The above code implicitly stores the final values of min\_0 and index\_0 in their last indices, but there is no primitive like drop\_dim to make it clear.
-This will have to be added in the SSA code generation (though it might not be necessary to include in the paper?).
-It also is pretty wishy-washy about the size and layout of each vector (again more of an implementation issue), but maybe we should have a different syntax for array accesses for read/write (e.g.
-the ADD\_SIMD steps) vs the array definitions (e.g.
-the SUB\_SIMD and raise/drop\_dim statements)}
-
-\end{algorithmic}
-\end{comment}
-
-\subsection{Correctness Argument}
-
-We build a correctness argument that loosely follows the theory of Abstract Interpretation. 
-We define the syntax of MPC Source programs. The domain of MPC Source programs
-expressible in the syntax (with certain semantic restrictions) is the abstract domain $A$. 
-We then define the \emph{linearization} of an MPC Source program as an interpretation 
-over the syntax. The linearization, which is a \emph{schedule}, is the concrete domain $C$. 
-Since we reason over def-use graphs in $A$ we define a partial order relation over elements of $A$ 
-in terms of def-use relations. We define a partial order over elements of $C$ as well,
-in terms of def-use relations in the concrete domain $C$. We prove two theorems
-that state (informally) that the schedule corresponding to the original program computes the same 
-result as the schedule corresponding to the vectorized program.
-
-\begin{figure}%[!b]
-%\small
-\[
-\begin{array}{l@{~~~~~}l}
-  \begin{array}{l@{~}l@{~~~}l}
-  s &::= s;s & sequence \\
-     & \mid \x[i,J,k] = \y[i,J,k] \; \mathit{op\_SIMD} \; \z[i,J,k] & operation \\ 
-     & \mid \x[i,J,k]  = \texttt{PHI}(\x_1[i,J,k],\x_2[i,J,k\!-\!1]) & \mathit{phi} \; \mathit{node} \\
-     & \mid \x[i,J,k] = \mathit{raise\_dim}(\x'[i],(J\!:\!J,k\!:\!K)) & \mathit{raise} \; \mathit{dimension(s)} \\
-     & \mid \x[i,J] = \mathit{drop\_dim}(\x'[i,J,k],k)  & \mathit{drop} \; \mathit{dimension(s)} \\
-     & \mid \x = \y & \mathit{propagation} \\
-     & \mid \texttt{FOR} \; 0 \le i < I: s  & \mathit{loop} \\
-  \end{array}    \\
-\end{array}
-\]
-\vspace{-3mm}
-\caption{MPC Source Syntax}\label{fig:MPC_Source_syntax}
-\vspace{-5mm}
-\end{figure}
-
-\paragraph{MPC Source Syntax}
-
-\figref{fig:MPC_Source_syntax} defines the syntax for our intermediate representation, MPC Source. 
-There are semantics restrictions over the syntax as well: a variable $\x[i,j,k]$ is a 3-dimensional array 
-$(i:I,j:J,k:K)$ and also, a statement $\x[i,J,k] = ...$ is enclosed in loops over $i$ and $k$ as shown below. 
-Thus, $i$ and $k$ are in scope.
-\begin{verbatim}
-FOR 0 <= i < I:
-    ...
-    FOR 0 <= k < K:
-        x[i,J,k] = ...
-\end{verbatim}
-
-Statements $\mathit{operation, phi, raise \; dimension(s), drop \; dimension(s)}$ are \emph{base} statements, 
-and $\mathit{sequence, loop}$ are \emph{compound statements}.
-
-\paragraph{Linearization}
-
-Linearization is the concretization operation, which, as we mentioned earlier computes a schedule.
-The concretization function $\gamma: A \rightarrow C$ is defined as an interpretation of
-MPC Source syntax, as it is standard. The concretization of each one 
-of the base statements is as follows: 
-\[\begin{array}{l}
-\gamma(\x[i,J,k] = \mathit{op\_SIMD}(\y[i,J,k], \z[i,J,k])) = \\ 
-~~\x[i,0,k] = \y[i,0,k] \; \mathit{op} \; \z[i,0,k] \;\; || \\
-~~~~\x[i,1,k] = \y[i,1,k] \; \mathit{op} \; \z[i,1,k] \;\; || \; ...\; || \\
-~~~~~~\x[i,I-1,k] = \y[i,I-1,k] \; \mathit{op} \; \z[i,I-1,k]
-\end{array}\]
-meaning that the vectorized dimension(s) are expanded into parallel statements. || introduces SIMD (parallel) execution.
-
-The concretization of the FOR statement is as follows:
-\[ \gamma(\texttt{FOR} \; 0 \le i < I: s) = \gamma(s)[0/i] \; ; \; \gamma(s)[1/i] \; ; \; ... \gamma(s)[I-1/i] \]
-$\gamma$ simply unrolls the loop substituting $i$ with 0, 1, etc. Here ; denotes sequential execution.
-
-\paragraph{Partial Orders}
-
-For each MPC Source program $a$ we compute the def-use edges in the standard way: if base statement $s1 \in a$ defines variable \x, e.g., $\x[i,j,k] = ...$, 
-and base statement $s2 \in a$ uses \x, e.g., $... = ... \x[i,j,k]$ and there is a path in the trivial CFG from $s1$ to $s2$, then there is a def-use edge from
- $s1$ to $s2$. We extend the dimensionality of a statement into $s1[i,j,k]$ where $s1[i,j,k]$ inherits the dimensionality of the left-hand-side of the assignment.
-
- Let $a_0, a_1$ be two MPC Source programs in $A$. 
- Two base statements, $s_0 \in a_0$ and $s_1 \in a_1$ are \emph{same}, written $s_0 \equiv s_1$ if they are of the same operation and they operate on 
- the same variables: same variable name and same dimensionality. Recall that dimensions in MPC Source are either iterative, lower case, or vectorized, upper case. 
- Two statements are same even if one operates on an iterative dimension and the other one operates on a vectorized one, e.g., $s_0[i,j,k] \equiv s_1[I,j,K]$.
-
- \begin{definition} Let $a_0, a_1 \in A$. We say that $a_0 \le a_1$ iff for every def-use edge $s1 \rightarrow s2$ in $a_0$ there is an edge
- $s1' \rightarrow s2'$ where $s1 \equiv s1'$, $s2 \equiv s2'$ and the two edges of either both forward or both backward.
- \end{definition}
-
- The def-use edges in the concrete schedule are as expected. There is a def use edge from statement $s1$ that defines $\x[\underline{i},\underline{j},\underline{k}]$
- to statement $s2$ that uses $\x[\underline{i},\underline{j},\underline{k}]$ if $s1$ is scheduled ahead of $s2$ in the linear schedule. We note that the underlined
- indices, e.g., $\underline{i}$, refer to fixed values, not iterative or vectorized dimensions since in the concrete schedule all induction variables are expanded. 
- E.g., there is a def-use edge from the statement that defines $\x[0,1,2]$ and a statement that uses $\x[0,1,2]$.
-
-\paragraph{Theorems}
-
-\begin{theorem} $a_0 \le a_1 \Rightarrow \gamma(a_0) \subseteq \gamma(a_1)$.
-\end{theorem}
-
-\begin{theorem} Let $a_0$ be the iterative MPC Source and let $a_1$ be the vectorized MPC Source computed by the vectorization algorithm. 
-We have that $a_0 \le a_1$. 
-\end{theorem}
-
-\ana{Write the proof sketch and final argument, etc.}
-
-\subsection{Extension of Basic Vectorization with Array Writes}
-
-\subsubsection{Removal of Infeasible Edges}
-
-Array writes limit vectorization as they sometimes introduce infeasible loop-carried dependencies. Consider the following example: \ana{Have to add citation to Aiken's paper}
-
-{\small
-\begin{verbatim}
-for i in range(N):
-  A[i] = B[i] + 10;
-  B[i] = A[i] * D[i-1];
-  C[i] = A[i] * D[i-1];
-  D[i] = B[i] * C[i];
-\end{verbatim}
-}
-
-In Cytron's SSA this code (roughly) translates into
-
-{\small
-\begin{verbatim}
-for i in range(N):
-1. A_1 = PHI(A_0,A_2)
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. A_2 = update(A_1, i, B_1[i] + 10); 
-6. B_2 = update(B_1, i, A_2[i] * D_1[i-1]);
-7. C_2 = update(C_1, i, A_2[i] * D_1[i-1]);
-8. D_2 = update(D_1, i, B_2[i] * C_2[i]);
-\end{verbatim}
-}
-
-\begin{comment}
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE \hspace{0.25cm} A[i] = B[i] + 10;
-\STATE \hspace{0.25cm} B[i] = A[i] * D[i-1];
-\STATE \hspace{0.25cm} C[i] = A[i] * D[i-1];
-\STATE \hspace{0.25cm} D[i] = B[i] * C[i];
-\end{algorithmic}
-
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{A}_0$ = $\phi(\texttt{A},\texttt{A}_1)$
-\STATE 2. $\texttt{B}_0$ = $\phi(\texttt{B},\texttt{B}_1)$
-\STATE 3. $\texttt{C}_0$ = $\phi(\texttt{C},\texttt{C}_1)$
-\STATE 4. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$
-\STATE 5. $\texttt{A}_1$ = $\texttt{A}_0$; $\texttt{A}_1$[i] = $\texttt{B}_0$[i] + 10; \COMMENT{equiv. to Update}
-\STATE 6. $\texttt{B}_1$ = $\texttt{B}_0$; $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 7. $\texttt{C}_1$ = $\texttt{C}_0$; $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 8. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
-\end{algorithmic}
-\end{comment}
-
-There is a cycle around \texttt{B\_1 = PHI(B\_0,B\_2)} that includes statement 
-\texttt{A\_1 = update(A\_0, i, B\_1[i] + 10)} and that statement won't be vectorized even 
-though in fact there is no loop-carried dependency from the write of \texttt{B\_1[i]} at 6 to the
-read of \texttt{ ... = B\_1[i]} at 8.
-
-The following algorithm removes certain infeasible loop-carried dependencies that are due to array writes. Consider a loop with index $0 \le j < J$
-nested at $i,j,k$. Here $i$ represents the enclosing loops of $j$ and $k$ represents the enclosed loops in $j$.
-
-\begin{algorithmic}
-\FOR {each array $\texttt{A}$ written in loop $j$}
-\STATE \COMMENT { including enclosed loops in $j$ }
-\STATE dep = False
-\FOR {each pair def: $\texttt{A}_m[f(i,j,k)] = ... $, and use: $ ... = \texttt{A}_n[f'(i,j,k)]$ in loop $j$}
-\IF {$\exists \underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$, s.t.
-$0\!
-\le\!
-\underline{i}\!
-<\!
-I,\: 0\!
-\le\!
-\underline{j}, \underline{j}' \!
-<\!
-J,\: 0\!
-\le\!
-\underline{k}, \underline{k}' \!
-<\!
-K,\: \underline{j}<\underline{j}'$, and $f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$}
-\STATE dep = True
-\ENDIF
-\ENDFOR
-\IF {dep == False}
-\STATE remove back edge into $\texttt{A}$'s $\phi$-node in loop $j$.
-\ENDIF
-\ENDFOR
-\end{algorithmic}
-
-Consider a loop $j$ enclosed in some fixed $\underline{i}$. Only if an update (definition) $\texttt{A}_m[f(i,j,k)] = ... $ at some iteration $\underline{j}$
-references the \emph{same} array element as a use $ ... = \texttt{A}_n[f'(i,j,k)]$ at some later iteration $\underline{j}'$,
-we may have a loop-carried dependence for $\texttt{A}$ due to this def-use pair. (In contrast, Cytron's algorithm inserts a loop-carried dependency every time there is an array update.)
-The algorithm above examines all def-use pairs in loop $j$, including defs and uses in nested loops, searching for values $\underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$ that satisfy
-$f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$. If such values exist for some def-use pair, then there is a potential
-loop-carried dependence on $\texttt{A}$; otherwise there is not and we can remove the spurious backward edge thus ``freeing up'' statements for vectorization. %\ana{Double check this!!!}
-
-Consider the earlier example.
-There is a single loop, $i$.
-Clearly, there is no pair $\underline{i}$ and $\underline{i}'$, where $\underline{i} < \underline{i}'$ that make $\underline{i} = \underline{i}'$
-due to the def-use pairs of \texttt{A} 5-6 and 5-7.
-Therefore, we remove the back edge from 5 to the phi-node 1.
-Analogously, we remove the back edges from 6 to 2 and from 7 to 3.
-However, there are many values $\underline{i} < \underline{i'}$ that make $\underline{i} = \underline{i'}-1$ and the back edge from 8 to 4 remains (def-use pairs for \texttt{D}).
-As a result of removing these spurious edges, Basic Vectorization will find that statement 5 is vectorizable.
-Statements 6, 7 and 8 will correctly appear in the FOR loop.
-
-Note however, that this step renders some array phi-nodes target-less. We handle target-less phi-nodes with a minor extension of Basic Vectorization (Phase 2). 
-First, we merge closures that update the same array. This simplifies handling of array $\phi$-nodes: if each closure is turned into a separate loop
-each loop will need to have its own array phi-node to account for the update and this would complicate the analysis.
-Second, we add the target-less node of array \texttt{A} back to the closure that updates \texttt{A} ---
-the intuition is, even if there is no loop-carried dependence from writes to reads on \texttt{A}, \texttt{A} is written and the write (i.e., update) cannot be vectorized; 
-therefore, the updated array has to carry to the next iteration of the loop. Third, in cases when the phi-node remains target-less, i.e., cases when the
-array write can be vectorized, we have to properly remove the phi-node replacing uses of the left-hand side of the phi-node with its arguments.
-
-%\ana{Commented out previous writeup. Removal/handling of targetless phi-nodes will go into the Extension to Basic vectorization.}
-
-\begin{comment}
-The final step is to remove array phi-nodes that are left without incoming back edges after removal of infeasible edges.
-Such a phi-node is redundant when there are no loop-carried dependences that are due to writes to the corresponding array.
-In our running example, the phi-nodes for \texttt{A}, \texttt{B}, and \texttt{C}, are redundant.
-
-Removing redundant phi-nodes proceeds from \emph{innermost to outermost} loops as follows.
-Denote a redundant phi-node by $\texttt{A}_1 = \phi(\texttt{A}_0,\texttt{A}_k)$. $\texttt{A}_0$ is the array at the beginning of the loop
-and $\texttt{A}_k$ is the array at the end of an iteration (including the last iteration). We consider two cases:
-\begin{enumerate}
-\item[Case 1] Inner-to-outer or same-level def-use edge
-\begin{itemize}
-\item If use is an array read $ ... = \texttt{A}_1[f(i)]$, then replace $\texttt{A}_1$ with $\texttt{A}_0$. This is correct, i.e., it is guaranteed that the use
-references $\texttt{A}_0$; had the use been referencing an array modified in an earlier iteration, the back-edge to the phi-node would
-not have been removed and the phi-node would not have been classified as redundant.
-\item If use is an array write (update) $\texttt{A}_2 = \texttt{A}_1; \texttt{A}_2[i] = ... $ equiv. to $\texttt{A}_2 = \mathit{update}(\texttt{A}_1, i, ...)$ then handling becomes more tricky.
-The semantics of this update is that when $i$ is 0, then $\texttt{A}_0$ is being read.
-But when $i > 0$, then $\texttt{A}_k$ is being read.
-If we are at this case, this means that the def-use edge is same-level and we are removing a redundant node in the loop enclosing the update.
-The removal of the original phi-node entails pushing the phi-node at the update:
-$\texttt{A}_2 = \phi(\texttt{A}_0,\texttt{A}_k); \texttt{A}_2[i] = ... $ equiv. to $\texttt{A}_2 = \mathit{update}(\phi(\texttt{A}_0,\texttt{A}_k), i, ...)$.
-\item If use is a phi-node $... = \phi(\texttt{A}_1,...)$, that means we have an outer-to-inner edge. The
-
-\end{itemize}
-\end{enumerate}
-For each outer-to-inner or same-level def-use edge
-that originates at the node, replace $\texttt{A}_1$ at the use node by $\texttt{A}_0$.
-For each inner-to-outer def-use edge, replace $\texttt{A}_1$ by $\texttt{A}_k$. In this case, the use references the array
-that results from the computation in the loop.
-
-Our running example is transformed as follows. Removing the phi-nodes for \texttt{A}, \texttt{B}, and \texttt{C} transforms the
-replaces the references to $\texttt{A}_0$, $\texttt{B}_0$, and $\texttt{C}_0$, with references to $\texttt{A}$, $\texttt{B}$, and $\texttt{C}$
-respectively, i.e., the arrays before the loop.
-
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$
-\STATE 2. $\texttt{A}_1$ = $\texttt{A}$; $\texttt{A}_1$[i] = $\texttt{B}$[i] + 10;
-\STATE 3. $\texttt{B}_1$ = $\texttt{B}$; $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 4. $\texttt{C}_1$ = $\texttt{C}$; $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
-\end{algorithmic}
-\end{comment}
-
-\begin{comment}
-\subsubsection{Array MUX refinement}
-
-\ana{TODO: WIP, get to BV without MUX refinement for now? For now, skip MUX refinement.}
-
-Next, the algorithm refines array MUX statements.
-MPC-source after Cytron's SSA may result in statements $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$, which imply that any index
-of $A$ can be written at this point and therefore there is a loop-carried dependency. In some cases the MUX can be refined to just a single
-index or a pair of indices, e.g., $\texttt{A}_j[i] = \emph{MUX}(c, \texttt{A}_k[i], \texttt{A}_l[i])$.
-
-%\ana{Clarify why this helps. Still work in progress.}
-This is to reduce the dimensionality of simd-ified computation. Technically, $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$ is
-a simdified operation that can be carried out in parallel "in one round". However, particularly when \texttt{A} is a multi-dimensional array, there is
-substantial increase in the size of the arrays (vectors) we send to SIMD operations. Refining to an update to a specific index would reduce
-the size of those vectors. Note that this is a heuristic that handles a common case, but not all cases of array updates.
-%\ana{TODO: Simplify this algorithm, taking into account the restriction to canonical updates. It should be handling \_all\_ cases.}
-
-\begin{algorithmic}
-\FOR {each $\mathit{stmt}$: $\texttt{A}_j = \emph{MUX}(c, \texttt{A}_k, \texttt{A}_l)$ in the MPC-source seq.}
-\STATE $i_1 = \mathit{find\_update}(A_k)$ \COMMENT { Is null when $A_k = \phi(...)$ }
-\STATE $i_2 = \mathit{find\_update}(A_l)$ \COMMENT { Is null when $A_l = \phi(...)$ }
-\IF {$i_1 == i_2$ or $i_1$ is null or $i_2$ is null} \STATE \COMMENT{ With our restrictions on writes we must have $i_1 = i_2$. }
-\STATE replace $\mathit{stmt}$ with \\
-\STATE $\texttt{A}_j = \texttt{A}_{j-1}$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
-%\ELSIF {$i_1 \ncong i_2$}
-%\STATE replace $\mathit{stmt}$ with \\
-%\STATE $\texttt{A}_j = \texttt{A}_k$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
-\ELSE
-\STATE $\mathit{stmt}$ stays as is
-\ENDIF
-\ENDFOR
-\end{algorithmic}
-\end{comment}
-
-\subsubsection{Restricting Array Writes}
-
-For now, we restrict array updates to \emph{canonical updates}. Assume (for simplicity) a two-dimensional array \texttt{A[I,J]}. A canonical update is the following:
-\begin{verbatim}
-for i in range(I):
-  for j in range(J):
-    ...
-    A[i,j] = ...
-    ...
-\end{verbatim}
-
-The update \texttt{A[i,j]} can be nested into an inner loop and there may be multiple updates, i.e., writes to \texttt{A[i,j]}.
-However, update such as \texttt{A[i-1,j] = ...} or \texttt{A[i-1,j-1] = ...}, etc., is not allowed.
-Additionally, while there could be several different loops that perform canonical updates, they must be of the same dimensionality, i.e., an update of higher or lower dimension, e.g., \texttt{A[i,j,k] = ...} is not allowed.
-We compute the \emph{canonical dimensionality} of each write array by examining the array writes in the original program and rejecting programs that violate the canonical write restriction.
-This restriction simplifies reasoning in this early stage of the compiler; we will look to relax the restriction in future work.
-
-Another restriction/assumption is that we assume the output array is given as input with initial values, and it is of size consistent with its canonical dimensionality. 
-%\ana{This restriction will require that we update our benchmarks that come with output arrays.}
-
-Reads through an arbitrary formula, such as \texttt{A[i-1]} for example, are allowed; currently, the projection function returns dummy values if the read formula is out of bounds; we assume the programmer ensures that the program still computes correct output in this case.
-
-\subsubsection{Changes to Basic Vectorization}
-
-\begin{comment}
-There are two changes to Basic Vectorization to account for write arrays. We can account for them during processing of raise dimension/drop dimension in Phase 1 of Basic Vectorization.
-
-One change to Basic vectorization is the expansion of dimension if the array write or read occurs in a nested loop.
-That is, if there is an update $\texttt{A}[i,j] = ...$ that occurs in loop dimensionality $i,j,k$, $\texttt{A}[i,j]$ will be rewritten into $\texttt{A}[i,j,k]$.
-Similarly, a read $\texttt{A}[f(i,j)]$ will be rewritten into $\texttt{A}[f(i,j),k]$.
-($\texttt{A}[f(i,j),k]$ will be reshaped as $f(i,j)*I*J+k$ during Phase 1 of Basic Vectorization.
-\ana{TODO: Double check.})
-\end{comment}
-
-In addition to the changes for the handling of target-less phi-nodes, Basic Vectorization has to handle def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable.
-The definition can be an update \texttt{A\_2 = update(A\_1,i,...)}, a pseudo $\phi$-node \texttt{A\_2 = PHI(A\_0,A\_1)}, etc..
-Note that $\phi$-nodes for arrays have no subscript operations the way there are subscript operations in analysis-introduced arrays representing scalars.
-While there are variations, the most intuitive implementation will perform Basic Vectorization Phase 1 as is, inserting
-$\mathit{raise\_dim}$ and $\mathit{drop\_dim}$ in the same way.
-However, the implementation of raise dimension and drop dimension will be adapted because the dimension cannot be raised or dropped to a dimension lower than the canonical one.
-Consider a def-use edge $X \rightarrow Y$ for an array $\texttt{A}$.
-
-\begin{enumerate}
-
-\item same-level $X \rightarrow Y$. Do nothing, propagate the array, which happens to be of the right dimension. 
-%\ana{There might be some opportunities to do copy propagation optimization and save some cycles, but let's leave this for later.}
-
-\item inner-to-outer $X \rightarrow Y$ triggers the addition of $\mathit{drop\_dim}$.
-However, the dimensionality cannot be dropped below the canonical dimensionality of the array.
-E.g., if the dimensionality of the loop enclosure $X$ is already at the canonical one, then $\mathit{drop\_dim}$ has no effect.
-
-\item outer-to-inner $X \rightarrow Y$ riggers  $\mathit{raise\_dim}$.
-Again, if the dimensionality of the loop enclosure of $Y$ is smaller or same as the canonical dimensionality of the array, then it has no effect, otherwise, if dimensionality is greater than the canonical dimensionality, $\mathit{raise\_dim(...)}$ (at $X$) is the same as in Basic Vectorization.
-
-\item "mixed" $X \rightarrow Y$. We assume that the mixed edge is transformed into an inner-to-outer followed by outer-to-inner edge before we perform vectorization, just as with Basic vectorization.
-
-\end{enumerate}
-
-If the use of the array is a read $\texttt{A}[f(i,j,k)]$ different than a canonical read $\texttt{A}[i,j,k]$, then we need to add a reshape operation as 
-all arrays are $\texttt{A}[i,j,k]$. It can be added after raise\_dim/drop\_dim or incorporated in these operations.
-The bulk of the change is in Phase 2 of Basic Vectorization as outlined earlier.
-
-\subsubsection{Examples with Array Writes}
-
-\paragraph{Example 1}
-
-First, the canonical dimensionality of all \texttt{A,B,C} and \texttt{D} is 1. 
-After Phase 1 of Basic Vectorization the Aiken's array write example will be (roughly) as follows:
-
-{\small
-\begin{verbatim}
-for i in range(N):
-1. A_1 = PHI(A_0,A_2)
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. A_2 = update(A_1, I, B_1[I] + 10); 
-6. B_2 = update(B_1, I, A_2[I] * D_1[I-1]);
-7. C_2 = update(C_1, I, A_2[I] * D_1[I-1]);
-8. D_2 = update(D_1, I, B_2[I] * C_2[I]);
-\end{verbatim}
-}
-
-Note that since all def-uses are same-level (i.e., reads and writes of the array elements) 
-no raise dimension or drop dimension happens.
-
-\begin{comment}
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{A}_1$ = $\phi(\texttt{A}_0,\texttt{A}_2)$
-\STATE 2. $\texttt{B}_1$ = $\phi(\texttt{B}_0,\texttt{B}_2)$
-\STATE 3. $\texttt{C}_1$ = $\phi(\texttt{C}_0,\texttt{C}_2)$
-\STATE 4. $\texttt{D}_1$ = $\phi(\texttt{D}_0,\texttt{D}_2)$
-\STATE 5. $\texttt{A}_2 = update(\texttt{A}_1, \stackrel{\rightarrow}{i}, \texttt{B}_1[\stackrel{\rightarrow}{i}] + 10)$;
-\STATE 6. $\texttt{B}_2 = update(\texttt{B}_1, \stackrel{\rightarrow}{i}, \texttt{A}_2[\stackrel{\rightarrow}{i}] * \texttt{D}_1[\stackrel{\rightarrow}{i-1}])$;
-\STATE 7. $\texttt{C}_2 = update(\texttt{C}_1, \stackrel{\rightarrow}{i}, \texttt{A}_2[\stackrel{\rightarrow}{i}] * \texttt{D}_1[\stackrel{\rightarrow}{i-1}])$;
-\STATE 8. $\texttt{D}_2 = update(\texttt{D}_1; \stackrel{\rightarrow}{i}, \texttt{B}_2[\stackrel{\rightarrow}{i}] * \texttt{C}_2[\stackrel{\rightarrow}{i}])$;
-\end{algorithmic}
-\end{comment}
-
-Phase 2 computes the closure of 4; $cl = \{4,6,7, 8\}$ 
-while 5 is vectorizable. Recall that 1,2, and 3 are target-less phi-nodes. Since the closure $cl$ includes updates to \texttt{B} and \texttt{C},
-the corresponding phi-nodes are added back to the closure and the def-use edges are added back to the target-less nodes. The uses of \texttt{A\_1} and \texttt{B\_1} in the vectorized statement
-turn into uses of \texttt{A\_0} and \texttt{B\_0} respectively; this is done for all original target-less phi-node.
-(But note that \texttt{A\_0} is irrelevant; the update writes into array \texttt{A\_2} in parallel.)
-Finally, the target-less phi-node for \texttt{A} is discarded.
-
-{\small
-\begin{verbatim}
-1. A_2 = update(A_0, I, ADD_SIMD(B_0[I],10));
-   equiv. to A_2[I] = ADD_SIMD(B_0[I],10)
-FOR i=0; i<N; i++; // MOTION loop
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. B_2 = update(B_1, i, A_2[i] * D_1[i-1]);
-   equiv. to B_2 = B_1; B_2[i] = A_2[i] * D_1[i-1];
-6. C_2 = update(C_1, i, A_2[i] * D_1[i-1]);
-7. D_2 = update(D_1, i, B_2[i] * C_2[i]);
-\end{verbatim}
-}
-
-\begin{comment}
-\begin{algorithmic}
-\STATE 1. $\texttt{A}_2 = update(\texttt{A}_0; \stackrel{\rightarrow}{i}, \mathit{ADD\_SIMD}(\texttt{B}_0[\stackrel{\rightarrow}i],[10,...])$ \COMMENT{Fully vectorized, size N.}
-\STATE FOR i=0; i<N; i++; \COMMENT{ MOTION loop }
-\STATE 2. $\texttt{B}_1$ = $\phi(\texttt{B}_0,\texttt{B}_2)$
-\STATE 3. $\texttt{C}_1$ = $\phi(\texttt{C}_0,\texttt{C}_2)$
-\STATE 4. $\texttt{D}_1$ = $\phi(\texttt{D}_0,\texttt{D}_2)$
-\STATE 5. $\texttt{B}_2 = update(\texttt{B}_1, i, \mathit{MUL}(\texttt{A}_2[i], \texttt{D}_1[{i-1}])$
-\STATE 6. $\texttt{C}_2 = update(\texttt{C}_1, i, \mathit{MUL}(\texttt{A}_2[i], \texttt{D}_1[{i-1}])$
-\STATE 7. $\texttt{D}_2 = update(\texttt{D}_1, i, \mathit{MUL}(\texttt{B}_2[i], \texttt{C}_1[i])$
-\end{algorithmic}
-\end{comment}
-
-\paragraph{Example 2} Now consider the MPC Source of Histogram:
-
-{\small
-\begin{verbatim}
-for i in range(0, num_bins):
-   res1 = PHI(res, res2)
-   for j in range(0, N):
-      res2 = PHI(res1, res3)
-      tmp1 = (A[j] == i)
-      tmp2 = (res2[i] + B[j])
-      tmp3 = MUX(tmp1, res2[i], tmp2)
-      res3 = Update(res2, i, tmp3)     
-return res1
-\end{verbatim}
-}
-
-The canonical dimensionality of \texttt{res} is 1.
-Also, the phi-node \texttt{res1 = PHI(res, res2)} is a target-less phi-node (the implication being that the inner for loop can be vectorized across $i$).
-After Phase 1, Basic vectorization produces the following code (statements are implicitly vectorized along $i$ and $j$).
-In a vectorized update statement, we can ignore the incoming array, \texttt{res2} in this case.
-The update writes (in parallel) all locations of the 2-dimensional array, in this case it sets up each \texttt{res3[i,j] = tmp3[i,j]}.
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I = raise_dim(i, ((i:num_bins),(j:N)))
-for i in range(0, num_bins):
-   res1 = PHI(res, res2') // target-less phi-node
-   res1' = raise_dim(res1, (j:N))
-   for j in range(0, N):
-      res2 = PHI(res1', res3)
-      tmp1 = (A1 == I)
-      tmp2 = (res2 + B1)
-      tmp3 = MUX(tmp1, res2, tmp2)
-      res3 = Update(res2, (I,J), tmp3) 
-   res2' = drop_dim(res2)    
-res1'' = drop_dim(res1)
-return res1''
-\end{verbatim}
-}
-
-Processing the inner loop in Phase 2 vectorizes \texttt{tmp1 = (A1 == I)} along the $j$ dimension but leaves the rest of the statements in a MOTION loop.
-Processing the outer loop is interesting.
-This is because the PHI node is a target-less node, and therefore, there are no closures!
-Several things happen.
-(1) Everything can be vectorized along the $i$ dimension.
-(2) We remove the target-less PHI node, however, we must update uses of \texttt{res1} appropriately: the use at raise\_dim goes to the first argument of the PHI function and the use at drop\_dim goes to the second argument.
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I1 = raise_dim(i, ((i:num_bins),(j:N)))
-
-tmp1[I,J] = (A1[I,J] == I1[I,J])
-
-res1' = raise_dim(res, (j:N)) // replacing res1 with res, 1st arg
-for j in range(0, N):
-   res2 = PHI(res1', res3)
-   tmp2[I,j] = (res2[I,j] + B1[I,j])
-   tmp3[I,j] = MUX(tmp1[I,j], res2[I,j], tmp2[I,j])
-   res3 = Update(res2, (I,j), tmp3)
-   equiv. to res3 = res2; res3[I,j] = tmp3[I,j]
-res2' = drop_dim(res2)   
-res1 = drop_dim(res2') // replacing with res2', 2nd arg. NOOP
-return res1
-\end{verbatim}
-}
-
-\begin{comment}
-The cleanup phase can then get rid of $j$ in the \texttt{res} variables and \texttt{tmp} variables, so the loop operates over the \texttt{num\_bins}-size vectors. 
-This will lead to dropping raise\_dim and drop\_dim as well:
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I1 = raise_dim(i, ((i:num_bins),(j:N)))
-
-tmp1[I,J] = (A1[I,J] == I1[I,J])
-
-for j in range(0, N):
-   res2 = PHI(res1, res3) // now 1-dim array
-   tmp2[I] = ADD_SIMD(res2[I],B1[I,j])
-   tmp3[I] = MUX_SIMD(tmp1[I,j], res2[I], tmp2[I])
-   res3 = Update(res2, I, tmp3[I])
-return res2
-\end{verbatim}
-}
-\end{comment}
-
-\begin{comment}
-\paragraph{Example 2}
-
-Consider MPC-source of Histogram after removal of infeasible edges and redundant phi-nodes, and MUX refinement:
-
-\begin{algorithmic}
-
-\STATE res!1 = []
-\STATE for i: plaintext in range(0, num\_bins):
-\STATE \hspace{0.25cm} res!2 = $\phi$(res!1, res!3)
-\STATE  \hspace{0.25cm} ...
-\STATE res!4 = res!2 \COMMENT{Added due to the "mixed" def-use edge.}
-\STATE for i: plaintext in range(0, num\_bins):
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE \hspace{0.25cm}        for j: plaintext in range(0, N):
-\STATE \hspace{0.5cm}            res!5 = $\phi$(res!4, res!7)
-\STATE \hspace{0.5cm}            !2!3 = (A[j] == i)
-\STATE \hspace{0.5cm}            !3!3 = (res!5[i] + B[j])
-\STATE \hspace{0.5cm}            res!6 = res!5; res!6[i] = !3!3
-\STATE \hspace{0.5cm}            res!7 = res!6; res!7[i] = MUX(!2!3, res!6[i], res!5[i])
-
-\end{algorithmic}
-
-After Phase 1 of Basic Vectorization:
-
-\begin{algorithmic}
-\STATE res!1 = []
-\STATE for i: plaintext in range(0, num\_bins):
-\STATE \hspace{0.25cm} res!2 = $\phi$(res!1, res!3)
-\STATE  \hspace{0.25cm} ...
-\STATE res!4 = res!2 \COMMENT{Added due to the "mixed" def-use edge. Because the canonical dimensionality of res is 1, we DO NOT drop dimension of res!2.}
-\STATE A[$i,j$] = $\mathit{raise\_dim}$(A[$j$],$i$) \COMMENT{Copying row A num\_bin times. Reusing name.}
-\STATE B[$i,j$] = $\mathit{raise\_dim}$(B[$i$],$j$) \COMMENT{Copying column B N times. Reusing name.}
-\STATE res!4[$i,j$] = $\mathit{raise\_dim}$(res!4[$i$],$j$) \COMMENT{Reusing name.}
-\STATE for i: plaintext in range(0, num\_bins):
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE \hspace{0.25cm}        for j: plaintext in range(0, N):
-\STATE \hspace{0.5cm}            res!5 = $\phi$(res!4, res!7) \COMMENT{Do nothing. res!4 has been raised to proper dimension.}
-\STATE \hspace{0.5cm}            !2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{EQ}$(A[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], [[0,0...], [1,1...],...[num\_bins-1,num\_bins-1...]])
-\STATE \hspace{0.5cm}            !3!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{ADD}$(res!5[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$],B[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$])
-\STATE \hspace{0.5cm}            res!6 = res!5 \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.5cm}            res!6[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = !3!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$]
-\STATE \hspace{0.5cm}            res!7 = res!6
-\STATE \hspace{0.5cm}            res!7[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{MUX}$(!2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], res!6[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], res!5[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$])
-\end{algorithmic}
-
-After Phase 2 and Phase 3. The EQ operation is vectorizable across both dimensions, and the rest of the computation is vectorizable across the $i$-dimension.
-
-\begin{algorithmic}
-\STATE ...
-\STATE !2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{EQ\_SIMD}$(A[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], [[0,0...], [1,1...],...[num\_bins-1,num\_bins-1...]])
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE FOR j=0; j<N; j++; \COMMENT{MOTION loop}
-\STATE \hspace{0.25cm}            res!5 = $\phi$(res!4, res!7) \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.25cm}            !3!3[$\stackrel{\rightarrow}i,j$] = $\mathit{ADD\_SIMD}$(res!5[$\stackrel{\rightarrow}i,j$],B[$\stackrel{\rightarrow}i,j$])
-\STATE \hspace{0.25cm}            res!6 = res!5 \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.25cm}            res!6[$\stackrel{\rightarrow}i,j$] = !3!3[$\stackrel{\rightarrow}i,j$]
-\STATE \hspace{0.25cm}            res!7 = res!6
-\STATE \hspace{0.25cm}            res!7[$\stackrel{\rightarrow}i,j$] = $\mathit{MUX\_SIMD}$(!2!3[$\stackrel{\rightarrow}i,j$], res!6[$\stackrel{\rightarrow}i,j$], res!5[$\stackrel{\rightarrow}i,j$])
-\end{algorithmic}
-\end{comment}
-
-%\section{Divide-and-Conquer}
-%\label{sec:div_and_conq}
-%\ana{TODO: Now that we have broken FOR loops into smaller chunks, we can add Divide-and-conquer reasoning with Z3 and implement this additional transform.}
-
-\section{Implementation and Evaluation}
-\label{sec:implementation}
-
 \section{Related Work}
@@
 and circuit-level amortization (done by MOTION). We apply our compiler on a wide range of routines.
-
-%\section{Future Work}
-%\label{sec:implementation_and_benchmarks}
-%\input{../sections/implementation_and_benchmarks.tex}
-%\input{../sections/evaluation.tex}
-
-\section{Conclusion and Future Work}
-\label{sec:conclusion}
-%\input{../sections/conclusion.tex}
-
-%\begin{acks}
-%    \ishaq{TODO}
-%\end{acks}
-
-%\printbibliography
-
-\bibliographystyle{plain}
-%\bibliography{c:/trink/Doc/MyTexFiles/texrefs}
-%\bibliography{mybib,library,library1,library2,library3,library_ana}
-\bibliography{library_ana,crypto_crossref}
-
-% that's all folks
-\end{document}

### 2022-04-28 overleaf/results.tex
Split paper into sections.

--- 2021-02-06 chapterfour/results.tex
@@
 \label{sec:results}
-
-\input{table1.tex}
-
-We compared 49 benchmarks, 23 of which are loops taken from
-the test code in HyCC \cite{Buscher2018}, while the other 26 are part of a micro benchmark suite cases 
-we developed for this purpose. \tabref{table:resultstablehycc} shows our results from the HyCC loops 
-and \tabref{table:resultstablemicro} shows our results on our own benchmark suite. We found that
-our assumptions held for all the test cases. 
-The majority of our programs benefited from our analysis with $\nicefrac{40}{49}$ programs showing
-a run time improvement.
-
-\input{table2.tex}
-
-As stated at the end of \secref{sec:loopscheduling} most of the loops we analyzed from HyCC were manipulations
-of vectorized multi-dimensional arrays into output arrays or values. Since the multi-dimensional arrays
-were represented in single dimensional format, indexes on these arrays were almost always very simple operations.
-\figref{fig:MNIST2HyCC} shows a typical example of this; the arrays {\sf kernel} and {\sf image} are used to modify the 
-loop output variable {\sf tmp}. The only across-loop dependency
-is {\sf tmp} on itself which can be vectorized much like the inner-product example in 
-\secref{sec:mpcamortization} using divide-and-conquer. In many cases we also found that 
-as parallelization of independent operations increased, the cost most often 
-decreased from $O(n)$ to either $O(log_{2}(n))$ or $O(1)$. The results
-for this are  shown in \tabref{table:runtimeresults}. Not all problems showed changed;
-For example, simple assignments incur no cost meaning that any loop body that consists of 
-\emph{only} assignments shows no improvement. This is the case for DB_JOIN_TWO1 (shown in \figref{fig:DBJOINTWO1HyCC}).
-In general our hand crafted benchmarks were significantly more difficult than those of HyCC. 
-Consider Test6 (shown in \figref{fig:Example6}). The use of many different
-conditional statements and complex dependencies lead to the high vectorized cost. 
-Whereas in the HyCC benchmarks we found that most dependencies were quite simple.
-An important note is that none of the the number of sequential opereations in our tests 
-ever increased, they only decreased or remained costant. 
-This means that existing techniques for loop optimizations only benefit from our work.
-
-Any schedule of SCC components that consists of one SCC node of any complexity flowing into a fully 
-parallelizable node (performing $n$ operations simultaneously) is optimal when executed linearly. We found this
-to be the case in the majority of the HyCC loop bodies. All of the benchmarks in the HyCC suite were found to be optimal.
-On the other hand any, as we stated in \secref{sec:theoreticalguarentees}, any SCC DAG that contains edges 
-that connect a non constant rate node to another non-constant rate node node is not known to be optimal. 
-Again, this is because our analysis does not provide a way to optimally schedule the statements within the 
-individual iterations in parallel. Node pairs of this type have unique dependencies, and running them linearly may miss 
-places where statements from one node can be run in parallel with statements from the other
-node. We did find that the majority of our tests were indeed optimal, this is shown in
-\tabref{table:optimalclassification}. This interleaving of statements between SCC nodes is most certainly an area 
-for future work, and possible proofs for optimality. Despite a lack of a formal optimality result we saw
-that almost all of our test cases showed a drastic improvement in run time complexity.
-
-\input{table3.tex}
-
-\begin{figure}[h]
-\centering
-\begin{minipage}{0.70\textwidth}
-\begin{javacode}
-...
-for(wx = 0; wx < window_size; wx++) {
-    int convPos = wx+wy*window_size;
-    tmp += kernel[convPos] 
-    	* image[(y*stride + wy) 
-    	* image_width 
-    	+ (x*stride + wx)];
-}
-...
-\end{javacode}
-\end{minipage}
-\caption{Example taken from MNIST2 in HyCC. %\ana{This is O(logN), no? Just a readonly array and a sum.}
-%\lindsey{yes! I reference it in the second paragraph of \secref{sec:results} to illustrate the fact that most loops are manipulating vectorized multi-dimensional arrays}
-}
-\label{fig:MNIST2HyCC}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\begin{minipage}{0.70\textwidth}
-\begin{javacode}
-...
-for(int j = from; j < n; j++) {
-    int tmp = m[from*n+j];
-    m[from*n+j] = m[to*n+j];
-    m[to*n+j] = tmp;
-}
-...
-\end{javacode}
-\end{minipage}
-\caption{Example taken from DB_JOIN_TWO in HyCC. %\ana{This is just 0, no? It's local computation since it's just}
-%\lindsey{referenced above to show that not all programs change since this is just a cost of 0}
-}
-\label{fig:DBJOINTWO1HyCC}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\begin{minipage}{0.70\textwidth}
-\begin{javacode}
-...
-k = i;
-z = z + i;
-a[i] = i;
-a[k] = i + a[i];
-int j = i - 1;
-if(a[i] > 3) {
-    a[k] = p * 10;
-    if(p > 20) {
-        a[i] = p * 20;
-    } else {
-        a[k] = p * 30;
-    }
-    z = z + a[i];
-} else {
-    a[i] = p * 40;
-}
-q = q + a[i];
-i++;
-...
-\end{javacode}
-\end{minipage}
-\caption{Test6 from our handcrafted benchmarks.}
-\label{fig:Example6}
-\end{figure}

### 2022-04-28 overleaf/vectorization.tex
Split paper into sections.

--- 2022-04-28 overleaf/paper_SIMD.tex
@@
-% !TeX TXS-program:compile = txs:///pdflatex/[--shell-escape]
-%
-% acm templates v1.57 (sigconf template)
-%
-%\documentclass[sigconf, screen, review, anonymous, natbib=false]{acmart}
-\documentclass[sigconf, screen, natbib=false, dvipsnames, table]{acmart}
-\settopmatter{printacmref=true}
-% mandatory for CCS'19
-\usepackage{balance}
-\usepackage{mathtools}
-% for creating a balanced last page (usually last page with references)
-
-% % defining the \BibTeX command - from Oren Patashnik's original BibTeX documentation.
-% \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08emT\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-
-% Rights management information.
-% This information is sent to you when you complete the rights form.
-% These commands have SAMPLE values in them; it is your responsibility as an author to replace
-% the commands and values with those provided to you when you complete the rights form.
-%
-% These commands are for a PROCEEDINGS abstract or paper.
-
-\copyrightyear{2019}
-\acmYear{2019}
-\setcopyright{acmcopyright}
-\acmConference[CCS '19]{2019 ACM SIGSAC Conference on Computer and Communications Security}{November 11--15, 2019}{London, United Kingdom}
-\acmBooktitle{2019 ACM SIGSAC Conference on Computer and Communications Security (CCS '19), November 11--15, 2019, London, United Kingdom}
-\acmPrice{15.00}
-\acmDOI{10.1145/3319535.3339818}
-\acmISBN{978-1-4503-6747-9/19/11}
-
-%
-% These commands are for a JOURNAL article.
-%\setcopyright{acmcopyright}
-%\acmJournal{TOG}
-%\acmYear{2018}\acmVolume{37}\acmNumber{4}\acmArticle{111}\acmMonth{8}
-%\acmDOI{10.1145/1122445.1122456}
-
-%
-% Submission ID.
-% Use this when submitting an article to a sponsored event. You'll receive a unique submission ID from the organizers
-% of the event, and this ID should be used as the parameter to this command.
-%\acmSubmissionID{123-A56-BU3}
-
-%
-% The majority of ACM publications use numbered citations and references. If you are preparing content for an event
-% sponsored by ACM SIGGRAPH, you must use the "author year" style of citations and references. Uncommenting
-% the next command will enable that style.
-%\citestyle{acmauthoryear}
-
- \makeatletter
- % \renewcommand{\section}{\abovedisplayskip 10\p@ \@plus3\p@ \@minus1\p@%
- %                       \belowdisplayskip 5\p@ \@plus3\p@ \@minus1\p@%
- %                       \abovedisplayshortskip 0pt \@plus2\p@%
- %                       \belowdisplayshortskip 0pt \@plus2\p@ \@minus0\p@%
- %                       \@startsection{section}{1}{\z@}%
- %                        {-17\p@ \@plus -4\p@ \@minus -4\p@}%
- %                        {6\p@ \@plus 4\p@ \@minus 4\p@}%
- %                        {\normalfont\large\bfseries\boldmath
- %                         \rightskip=\z@ \@plus 8em\pretolerance=10000 }}
- \renewcommand{\subsection}{\@startsection{subsection}{2}{\z@}%
-                        {-8\p@ \@plus -4\p@ \@minus -4\p@}%
-                        {5\p@ \@plus 2\p@ \@minus 2\p@}%
-                        {\normalfont\Large\bfseries\boldmath
-                         \rightskip=\z@ \@plus 3em\pretolerance=10000 }}
-  \renewcommand{\subsubsection}{\@startsection{subsubsection}{3}{\z@}%
-                        {-6\p@ \@plus -4\p@ \@minus -4\p@}%
-                        {1\p@ \@plus 1\p@ \@minus 0\p@}%
-                        {\normalfont\normalsize\bfseries\boldmath}}
-% \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\z@}%
-%                       {-8\p@ \@plus -4\p@ \@minus -4\p@}%
-%                       {-2\p@ \@plus -0.22em \@minus -0.1em}%
-%                       {\normalfont\normalsize\bfseries}}
- \makeatother
-
-% *** GRAPHICS RELATED PACKAGES ***
-\usepackage{graphicx}
-\usepackage{comment}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{listings}
-% declare the path(s) where your graphic files are
-\graphicspath{{../figures/}}
-% and their extensions so you won't have to specify these with
-% every instance of \includegraphics
-\DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-
-% Minted is used for source code highlighting
-%\usepackage[outputdir=/Volumes/ramdisk/]{minted}
-% \usepackage{minted}
-% *** MATH PACKAGES ***
-\usepackage{amsmath}
-
-% *** SPECIALIZED LIST PACKAGES ***
-%\usepackage{algorithm2e}
-\usepackage{algorithmic}
-
-% *** ALIGNMENT PACKAGES ***
-\usepackage{array}
-
-\usepackage{url}
-
-\usepackage[utf8]{inputenc}
-\usepackage[american]{babel}
-%\usepackage[backend=biber,style=ACM-Reference-Format]{biblatex}
-%\addbibresource{./library.bib}
-%\addbibresource{./cryptobib/abbrev2.bib}
-%\addbibresource{./cryptobib/crypto_crossref.bib}
-
-% Used for Theorems and Definitions
-\usepackage{amsthm}
-\newtheorem{definition}{Definition}
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}{Lemma}
-\theoremstyle{definition}
-\newtheorem{remark}{Remark}
-\newtheorem{assumption}{Assumption}
-
-\usepackage{xcolor}
-\usepackage{xspace}
-\usepackage{xstring} % required by IfEqCase
-
-\usepackage{booktabs} % better tables
-\usepackage{algorithm}
-\usepackage{algorithmic}
-
-% for references, use it like this: \cref{some_label}
-% it will automatically put Appendix, Figure, Section, etc before the reference
-\usepackage[capitalise]{cleveref}
-
-% ishaq: adding our macros
-\input{macros.tex}
-
-\sloppy
-\begin{document}
-\fancyhead{}
-% do not delete this code.
-
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
-\title[Compilation and Backend-Independent Vectorization for Multi-Party Computation]{Compilation and Backend-Independent Vectorization for Multi-Party Computation}
-
-%
-% The "author" command and its associated commands are used to define the authors and their affiliations.
-% Of note is the shared affiliation of the first two authors, and the "authornote" and "authornotemark" commands
-% used to denote shared contribution to the research.
-
-\author{Benjamin Levy}
-\email{levyb3@rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Benjamin Sherman}
-\email{shermb@rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Muhammad Ishaq}
-%\authornote{This work was done in part while the author was at RPI.}
-\email{ishaqm@purdue.edu}
-%\orcid{1234-5678-9012}
-%\author{G.K.M. Tobin}
-%\authornotemark[1]
-%\email{webmaster@marysville-ohio.com}
-\affiliation{%
-    \institution{Purdue University}
-    %\streetaddress{P.O. Box 1212}
-    \city{West Lafayette}
-    \state{Indiana}
-%    \postcode{43017-6221}
-}
-
-\author{Lindsey Kennard}
-\email{fire.elemental@gmail.com}
-\affiliation{%
-    \institution{STR}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Boston}
-    \state{Massachusets}
-    %\country{Iceland}
-}
-
-\author{Ana Milanova}
-\email{milanova@cs.rpi.edu}
-\affiliation{%
-    \institution{Rensselaer Polytechnic Institute}
-    %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
-    \city{Troy}
-    \state{New York}
-    %\country{Iceland}
-}
-
-\author{Vassilis Zikas}
-%\authornote{This work was done in part while the author was visiting UCLA and supported in part by DARPA and SPAWAR under contract N66001-15-C-4065 and by a SICSA Cyber Nexus Research Exchanges grant.}
-\email{vzikas@purdue.edu}
-%\orcid{1234-5678-9012}
-%\author{G.K.M. Tobin}
-%\authornotemark[1]
-%\email{webmaster@marysville-ohio.com}
-\affiliation{%
-    \institution{Purdue University}
-    %\streetaddress{P.O. Box 1212}
-    \city{West Lafayette}
-    \state{Indiana}
-    %    \postcode{43017-6221}
-}
-
-%
-% By default, the full list of authors will be used in the page headers. Often, this list is too long, and will overlap
-% other information printed in the page headers. This command allows the author to define a more concise list
-% of authors' names for this purpose.
-%\renewcommand{\shortauthors}{Trovato and Tobin, et al.}
-
-\renewcommand{\A}{{\sf A}}
-\renewcommand{\B}{{\sf B}}
-\renewcommand{\C}{{\sf C}}
-\renewcommand{\D}{{\sf D}}
-
-% As a general rule, do not put math, special symbols or citations
-% in the abstract
-\begin{abstract}
-
-\ana{We need an abstract.}
-
-\end{abstract}
-
-%
-% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-% Please copy and paste the code instead of the example below.
-%
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10003752.10010124.10010138.10010143</concept_id>
-<concept_desc>Theory of computation~Program analysis</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10003752.10003777.10003789</concept_id>
-<concept_desc>Theory of computation~Cryptographic protocols</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10002978.10002979</concept_id>
-<concept_desc>Security and privacy~Cryptography</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Theory of computation~Program analysis}
-\ccsdesc[500]{Theory of computation~Cryptographic protocols}
-\ccsdesc[300]{Security and privacy~Cryptography}
-
-%
-% Keywords. The author(s) should pick words that accurately describe the work being
-% presented. Separate the keywords with commas.
-\keywords{multiparty computation; compilers; cryptography}
-
-\begin{comment}
-%
-% A "teaser" image appears between the author and affiliation information and the body
-% of the document, and typically spans the page.
-\begin{teaserfigure}
-    \includegraphics[width=\textwidth]{sampleteaser}
-    \caption{Seattle Mariners at Spring Training, 2010.}
-    \Description{Enjoying the baseball game from the third-base seats. Ichiro Suzuki preparing to bat.}
-    \label{fig:teaser}
-\end{teaserfigure}
-\end{comment}
-%
-% This command processes the author and affiliation and title information and builds
-% the first part of the formatted document.
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-
-Multi-party computation (MPC) allows $N$ parties $p_1,\dots, p_N$ to perform a computation on their private inputs securely.
-Informally, security means that the secure computation protocol computes the correct output (correctness) and it does not leak any information about the individual party inputs, other than what can be deduced from the output (privacy).
-
-% The first results from the early 80's~\cite{FOCS:Yao82b,STOC:GolMicWig87,STOC:BenGolWig88,STOC:ChaCreDam88} introduced the problem and formally defined its security, and investigate feasibility in different models---e.g., passive or malicious adversaries, unconditional or computational security, with or without a broadcast channel.
-This spawned a rich theoretical literature of beautiful works improving and extending in various manners the original ideas.
-
-MPC theory dates back to the early 1980-ies~\cite{FOCS:Yao82b,STOC:GolMicWig87,STOC:BenGolWig88,STOC:ChaCreDam88}. 
-Long the realm of theoretical cryptography, MPC has seen significant advances in programming technology in recent years. 
-These advances bring MPC closer to practice and wider applicability --- 
-MPC technology has been employed in real-world scenarios such as auctions~\cite{FC:BCDGJK09}, biometric identification~\cite{ESORICS:BlaGas11}, 
-and privacy-preserving machine learning~\cite{SP:MohZha17,CCS:MohRin18}. 
-The goal is to bring the technology to a level where programmers can write \emph{secure} and \emph{efficient} programs without 
-commanding extensive knowledge of cryptographic primitives.
-
-The problem, therefore, is to build a high-level programming language and a compiler, and there has been significant advance in 
-this space, e.g., ~\cite{CCS:BenNisPin08,ESORICS:BogLauWil08,CCS:ZhaSteBla13,Hastings:2019,Keller:2020,Acay:2021,Braun:2022} among other work. 
-Current research largely falls at the two ends 
-of the classical compiler: (1) work on \emph{front-end} language design and (2) work on \emph{back-end} protocol implementation. Work on language design 
-focuses on high-level constructs necessary to express multiple parties, computation by different parties, and information flow from one party 
-to another~\cite{SP:RasHamHic14, Acay:2021}. On the other end, work on protocol implementation focuses on cryptographic foundations and their efficient circuit-level
-implementation~\cite{NDSS:DemSchZoh15, CCS:ABFKLO18, Braun:2022}, e.g., implementation of operations (e.g., MUL, ADD) using different sharing protocols 
-(Boolean or Arithmetic GMW~\cite{STOC:GolMicWig87} or Yao's garbled circuits~\cite{FOCS:Yao82b}), as well as efficient share conversion from one representation 
-to another.
-
-Earlier compilers did both back-end and front-end translation without a specific focus on either, as their aim was to demonstrate applicability of MPC on real-world
-programming problems. As the field advanced, works have focused more closely on front-end language design (e.g., Wysteria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021}) 
-or back-end ``circuit-level'' design and implementation (e.g., MOTION~\cite{Braun:2022}).
-
-In this work we focus on an intermediate language and what we call \emph{backend-independent optimizations}, in a close analogy to \emph{machine-independent} optimizations
-in the classical compiler. The following figure summarizes our key idea:
-
-{\begin{center}
-\includegraphics[width=0.8\linewidth]{figs_paper_SIMD/focus.png}
-\end{center}
-}
-
-We emphasize the MPC Source intermediate representation and optimization over MPC Source. As in classical compilers, we envision different front ends (e.g., Wysteria,
-our front end IMP-MPC) compiling into MPC Source. MPC Source is particularly suitable for optimizations such as protocol mixing~\cite{Ishaq:2019, Fang:2022} and SIMD-vectorization, 
-which takes advantage of amortization at the circuit level. The MPC Source IR exposes the linear structure of MPC programs, which simplifies program analysis; this is
-in contrast to source, which has if-then-else constructs. In the same time, MPC Source is sufficiently ``high-level''  to support analyses and optimizations that take into account 
-control and data flow in a specific program. MPC Source is small in size and analysis is tractable, as opposed to analysis over an unrolled circuit~\cite{Ishaq:2019}. Again as
-in classical compilers, we envision translation of MPC Source (optimized or unoptimized) into MOTION, SPDZ, or other back-end code.
-
-%In contrast, the high-level program contains if-then-else and array writes, or the low-level circuit . 
-%The compilation problem, i.e., translation from a high-level language to circuits, is difficult and therefore, breaking the problem into sub-problems can  tractable.
-
-\subsection{Our Contribution} In this paper, we develop a compiler framework that takes a Python-like routine and produces MOTION code: we describe (a) the IMP-MPC language, its syntax and semantic restrictions, (b) translation from IMP-MPC into MPC Source, (c) a specific backend-independent optimization: novel SIMD-vectorization on MPC Source, and (d) translation from MPC Source into MOTION code.
-%\ana{I think we have to add more on MOTION, specifically mixing and amortization.}
-
-We focus on the MOTION framework as our back-end for several reasons.
-First, it is the state of the art in terms of performance~\cite{Braun:2022}.
-Second, it provides an API over efficient implementation for a wide variety of cryptographic operations in three different protocols ---  Arithmetic GMW, Boolean GMW, and BMR --- which allows for protocol mixing~\cite{Ishaq:2019, Fang:2022}, a known backend-independent optimization.
-Third, MOTION provides API for SIMD-level operations, which amortize cost and lead to significant improvement in memory footprint and
-throughput~\cite{NDSS:DemSchZoh15, CCS:ABFKLO18, Braun:2022}. It also enables MPC Source-level optimizations such SIMD vectorization.
-
-Our second contribution is an analytical model for cost estimation of amortized schedules. Originally, we hoped that optimal scheduling (under our model) was tractable, as the problem
-appeared simpler than the classical scheduling problem. Unfortunately, we show that optimal scheduling is NP-hard via a reduction to the Shortest Common Supersequence (SCS) problem. 
-Cost estimation is important as it drives not only vectorization but other optimizations such as protocol mixing optimizations and scheduling as well~\cite{Ishaq:2019, Fang:2022}.
-
-Our most important contribution is the implementation and evaluation of the compiler framework.
-We demonstrate expressivity of the source language by running the compiler on \ana{X} programs; these include classical MPC benchmarks such as PSI and Biometric matching, as well as kMeans, Histogram, and others.
-Vectorization leads to \ana{X}x improvement in number of gates on average, \ana{Y}x speedup in circuit generation time, \ana{Z}x speedup in the LAN setting, and \ana{Z}x speedup in the WAN setting using the Boolean GMW protocol.
-Results form BMR are similar.
-
-Our results emphasize the importance of backend-independent optimizations --- vectorization (described in this work) and protocol mixing (tackled in previous works~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022}) are two optimizations readily available at the level of MPC Source.
-We believe that our work can lead to future work on backend-independent compilation and optimization, bringing in new MPC optimizations and combinations of optimizations, much in the vein of standard compilers, and bring MPC programming technology closer to practice and wider applicability.
-
-\subsection{Outline}
-
-The rest of the paper is organized as follows.
-\secref{sec:overview} presents an overview of the compiler.
-\secref{sec:model} describes our model for cost estimation and argues NP-hardness of optimal scheduling and \secref{sec:compiler} describes the compiler.
-\secref{sec:experiments} presents implementation and evaluation.
-\secref{sec:related} discusses related work and \secref{sec:conclusion} concludes.
-
-\begin{comment}
-\begin{itemize}
-\item We define the scheduling problem for MPC. 
-We present an analytical model to reason about cost of schedules
-and show that scheduling is NP-hard via a reduction to the shortest 
-common supersequence problem.
-
-\item We present a compiler that takes an IMP-like high-level 
-program and produces amortized (i.e., vectorized) low-level cryptographic 
-code in the MOTION framework. Central contributions are 1) a novel compiler framework, 
-2) a vectorization algorithm that produces optimal schedules for a large 
-number of MPC programs, and 3) reasoning over output arrays without Array SSA; we remove 
-infeasible loop-carried dependences introduced by standard SSA to improve vectorization.
-
-\item We present an implementation and evaluation in the MOTION framework. 
-\ana{Fill in with final results. Mention benchmarks (standard + new ones + HyCC). }
-
-\end{itemize}
-\end{comment}
-
-\section{Overview}
-\label{sec:overview}
-
-\begin{table*}
-\begin{tabular}{ccc}
-\begin{minipage}{0.275\textwidth}
-{\small
-\begin{pythonn}
-def biometric(C: shared[list[int]], D: int, 
-      S: shared[list[int]], N: int) -> 
-      shared[tuple[int,int]]:
-   min_sum : int = MAX_INT
-   min_idx : int = 0  
-   for i in range(N): 
-      sum : int = 0
-      for j in range(D): 
-         # d = S[i,j] - C[j]
-         d : int = S[i * D + j] - C[j] 
-         p : int = d * d
-         sum = sum + p
-      if sum < min_sum:
-         min_sum : int = sum
-         min_idx : int = i
-   return (min_sum, min_idx)
-\end{pythonn}
-}
-\end{minipage}
-
-&
-
-\begin{minipage}{0.33\textwidth}
-{\tiny
-\begin{pythonn}
-min_sum!1 = MAX_INT
-min_idx!1 = 0
-for i in range(0, N):
-   min_sum!2 = PHI(min_sum!1, min_sum!4)
-   min_idx!2 = PHI(min_idx!1, min_idx!4)
-   sum!2 = 0
-   for j in range(0, D):
-      sum!3 = PHI(sum!2, sum!4)
-      d = SUB(S[((i * D) + j)],C[j]) 
-      p = MUL(d,d) 
-      sum!4 = ADD(sum!3,p) 
-   t = CMP(sum!3,min_sum!2) 
-   min_sum!3 = sum!3
-   min_idx!3 = i
-   min_sum!4 = MUX(t, min_sum!3, min_sum!2) 
-   min_idx!4 = MUX(t, min_idx!3, min_idx!2) 
-return (min_sum!2, min_idx!2)   
-\end{pythonn}
-}
-\end{minipage}
-
-&
-
-\begin{minipage}{0.33\textwidth}
-{\small
-\begin{pythonn}
-min_sum!1 = MAX_INT
-min_idx!1 = 0
-# S^ is same as S. C^ replicates C N times: 
-S^ = raise_dim(S, ((i * D) + j), (i:N,j:D)) #S^[i,j] = S[i,j]
-C^ = raise_dim(C, j, (i:N,j:D)) #C^[i,j] = C[j]
-
-sum!2[I] = [0,..,0]
-# computes _all_  "at once"
-d[I,J] = SUB_SIMD(S^[I,J],C^[I,J])
-p[I,J] = MUL_SIMD(d[I,J],d[I,J])
-
-for j in range(0, D):
-   # sum!2[I], sum!3[I], sum!4[I] are size-N vectors 
-   # computes N intermediate sums "at once"
-   sum!3[I] = PHI(sum!2[I], sum!4[I])       
-   sum!4[I] = ADD_SIMD(sum!3[I],p[I,j])
-
-min_idx!3[I] = [0,1,...N-1]   
-for i in range(0, N):
-   min_sum!2 = PHI(min_sum!1, min_sum!4) 
-   t[i] = CMP(sum!3[i],min_sum!2)
-   min_sum!4 = MUX(t[i], sum!3[i], min_sum!2)  
-for i in range(0, N):
-   min_idx!2 = PHI(min_idx!1, min_idx!4)    
-   min_idx!4 = MUX(t[i], min_idx!3[i], min_idx!2)
-return (min_sum!2, min_idx!2)   
-\end{pythonn}
-}
-\end{minipage}
-
-\\
-
-(a) IMP Source & (b) MPC Source & (c) Optimized MPC Source
-\end{tabular}
-\caption{Biometric Matching: From IMP Source to MPC Source to Optimized MPC Source. }
-\label{tab:source_and_MPC_source_and_optimized_MPC_source}
-\end{table*}
-
-\subsection{Source}
-
-As a running example, consider Biometric matching, a standard MPC benchmark.
-An intuitive (and naive) implementation is as shown in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(a).
-Array \python{C} is the feature vector of \texttt{D} features that we wish to match and \python{S} 
-is the database of \python{N} size-\python{D} vectors that we match against.
-
-\begin{comment}
-{\small
-\begin{verbatim}
-def biometric(C: shared[list[int]], D: int, 
-              S: shared[list[int]], N: int) -> 
-                      tuple[shared[int], shared[int]]:
-    min_sum = 10000
-    min_index = 0  
-    for i in range(N): #loop over database
-        sum = 0
-        for j in range(D): #loop over features
-            d = S[i * D + j] - C[j] #i.e., d = S[i,j] - C[j]
-            p = d * d
-            sum = sum + p
-        if sum < min_sum:
-            min_sum = sum
-            min_index = i
-    return (min_sum, min_index)
-\end{verbatim}
-}
-\end{comment}
-
-Our compiler takes essentially standard IMP~\ana{Need citation here.} 
-syntax and imposes certain semantic restrictions. (We detail the restrictions in the following sections.)
-The programmer writes an iterative program and annotate certain inputs 
-and outputs as \emph{shared}. In the example arrays \python{C} and \python{S}
-are shared, meaning that they store shares, however, the array sizes \python{D} and
-\python{N} respectively are plaintext. The code iterates over the entries in the 
-database and computes the Euclidean distance (its square actually) of the current 
-entry \python{S[i]} and \python{C}. The program returns the index of the vector that gives 
-the best match plus the corresponding sum of squares.
-
-\subsection{MPC Source and Cost of Schedule}
-
-Our compiler generates an intermediate representation, MPC Source. MPC Source 
-for Biometric Matching is shown in 
-Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(b). First, MPC Source is an SSA form. 
-Second, it is linear. The conditional in lines 13-14 in IMP Source turns into the linear code in lines 12-16 in MPC Source. 
-The test turns into a CMP operation (\python{t = CMP(sum!3,min\_sum!2)}), followed by the 
-true-branch sequence, followed by the MUX operations. The first MUX operation selects the value
-of \python{min\_sum}: if \python{t} is true, then \python{min\_sum} gets the value of the second multiplexer
-argument,  \python{min\_sum!3}, otherwise it takes the value of the third argument, \python{min\_sum!2}.  
-Third, MPC Source is a special form of SSA. The SSA $\phi$-nodes at the if-then-else (lines 13-15) turn into 
-MUX operations, while the $\phi$-nodes at for loops turn into what we call pseudo PHI nodes with a straightforward semantics.
-
-\begin{comment}
-{\small
-\begin{pythonn}
-1. min_sum!1 = 10000
-2. min_index!1 = 0
-3. for i in range(0, N!0):
-4.    min_sum!2 = PHI(min_sum!1, min_sum!4)
-5.    min_index!2 = PHI(min_index!1, min_index!4)
-6.    sum!2 = 0
-7.    for j in range(0, D!0):
-8.      sum!3 = PHI(sum!2, sum!4)
-9.      d!3 = (S!0[((i * D!0) + j)] - C!0[j]) // MPC
-10.     p!3 = (d!3 * d!3) // MPC
-11.     sum!4 = (sum!3 + p!3) // MPC
-12.  !1!2 = (sum!3 < min_sum!2) // MPC
-13.  min_sum!3 = sum!3
-14.  min_index!3 = i
-15.  min_sum!4 = MUX(!1!2, min_sum!3, min_sum!2) // MPC
-16.  min_index!4 = MUX(!1!2, min_index!3, min_index!2) // MPC
-17.!2!1 = (min_sum!2, min_index!2)   
-\end{pythonn}
-}
-\end{comment}
-
-%The compiler linearizes the source turning conditionals into MUX statements. The PHI nodes
-%are remnants of the SSA IR; the compiler generates code that picks the correct value 
-%when producing MOTION output; the MOTION framework in turn linearizes 
-%loops when it generates the circuit.
-
-We turn to our analytical model to compute the \emph{cost} of the program. Assuming fixed 
-cost $\beta$ for a local MPC operation (essentially just ADD) and cost $\alpha$ for a remote
-MPC operation (e.g., MUX, CMP, and remaining operations), the cost of the iterative schedule 
-will be $ND(2\alpha+\beta) + N(3\alpha$).
-
-A key contribution is the vectorizing transformation. We can compute all $N*D$ 
-subtraction operations (line 9) in a single SIMD instruction; similarly we can compute 
-all multiplication operations (line 10) in a single SIMD instruction. And while we cannot
-vectorize computation of the N individual sums, we can compute the N sums in parallel. 
-Our compiler \emph{automatically detects these opportunities and transforms the program}. 
-It is standard that MPC researchers write vectorized versions of the Biometric
-program by hand; we are the first (to the best of our knowledge) to automatically 
-transform a naive, iterative MPC program into an unintuitive vectorized one.
-
-\subsection{Vectorized MPC Source and Cost of Schedule}
-
-Our compiler produces the vectorized program shown in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(c). 
-Note that this is still our intermediate representation, Optimized MPC Source. Subsequently,
-the compiler turns this code into MOTION variables, loops and SIMD primitives, which MOTION then 
-uses to generate the circuit.
-
-The compiler determines that SUB and MUL in ``naive'' MPC Source (lines 9 and 10 in (b))
-can be fully vectorized into the SIMD SUB and MUL in optimized MPC Source (lines 9 and 10 in (c)). 
-Notation \python{p[I,J]} denotes a 2-dimensional array with fully vectorized dimensions.
-The computation of sum (line 11 in (b))
-is sequential across the $j$-dimension, but it is parallel across the $i$-dimension. 
-The loop in lines 12-16 in (c) illustrates; here \python{p[I,j]} refers to the $j$-th column in \python{p}. 
-Unfortunately, CMP and MUX remain sequential.
-
-\begin{comment}
-{\small
-\begin{verbatim}
-min_sum!1 = 10000
-min_index!1 = 0
-// S!0^ is same as S!0. C!0^ replicates C!0 N-times: 
-S!0^ = raise_dim(S!0, ((i * D!0) + j), (i:N!0,j:D!0))
-C!0^ = raise_dim(C!0, j, (i:N!0,j:D!0))
-
-sum!2 = [0,..,0]
-// Computes all differences and all products "at once"
-d!3[I,J] = SUB_SIMD(S!0^[I,J],C!0^[I,J])
-p!3[I,J] = MUL_SIMD(d!3[I,J],d!3[I,J])
-
-for j in range(0, D!0):
-  // sum!2[I], sum!3[I], sum!4[I] are vectors of size N  
-  // Computes N intermediate sums "at once"
-  sum!3[I] = PHI(sum!2[I], sum!4[I])       
-  sum!4[I] = ADD_SIMD(sum!3[I],p!3[I,j])
-
-min_index!3 = [0,1,...N!0-1]
-
-for i in range(0, N!0):
-  min_sum!2 = PHI(min_sum!1, min_sum!4) 
-  !1!2[i] = CMP(sum!3[i],min_sum!2)
-  min_sum!4 = MUX(!1!2[i], sum!3[i], min_sum!2)
-
-for i in range(0, N!0):
-  min_index!2 = PHI(min_index!1, min_index!4)  
-  min_index!4 = MUX(!1!2[i], min_index!3[i], min_index!2)
-!2!1 = (min_sum!2, min_index!2)   
-\end{verbatim}
-}
-\end{comment}
-
-In MPC backends, executing $n$ operations ``at once'' in a single SIMD operation costs a lot less than executing those $n$ operations one by one.
-This is particularly important when there is communication (i.e., in remote), since many 1-bit values are sent at once rather than sequentially. 
-%In MPC compilers a vectorized operation that computing M operations "at once" costs 
-%essentially the same ($\alpha$ or $\beta$) as an individual operation. 
-We elaborate on the cost model in the following section but for now consider that 
-each operation has a \emph{fixed} portion (does benefits from amortization) and 
-a \emph{variable} portion (does not benefit from amortization): $\alpha = \alpha_\mathit{fix} + \alpha_\mathit{var}$. 
-This gives rise to the following formula for amortized cost: $f(n) = \alpha_\mathit{fix} + n\alpha_\mathit{var}$, 
-as opposed to unamortized cost $g(n) = n\alpha_\mathit{fix} + n\alpha_\mathit{var}$. (We extend the same reasoning to 
-$\beta$-instructions.)
-
-Thus, the fixed cost of the vectorized program amounts to
-$2\alpha_\mathit{fix}$ + $D \beta_\mathit{fix} + N(3\alpha_\mathit{fix})$. (The variable cost is the same in both the vectorized and non-vectorized programs.) 
-The first term in the sum corresponds to the vectorized
-subtraction and multiplication, the second term corresponds to the FOR loop on $j$ and the third 
-one corresponds to the remaining FOR loops on $i$.
-Clearly, $2\alpha_\mathit{fix} + D\beta_\mathit{fix} + N(3\alpha_\mathit{fix}) << ND(2\alpha_\mathit{fix}+\beta_\mathit{fix}) + N3\alpha_\mathit{fix}$.
-Empirically, we observe 10x to 50x improvement over un-vectorized Biometric Matching in circuit generation time and setup time and 5x to 30x improvement in online time.
-Additionally, the un-vectorized version runs out of memory for $N=256$, while the vectorized one runs with the standard maximal input size $N=4,096$.
-\ana{Edit numbers with final experiments.}
-
-\section{Analytical Model}
-\label{sec:model}
-
-\ana{We need an intro to section here.}
-
-\subsection{Scheduling in MPC}
-\label{sec:mpc}
-
-%\ana{Here we need more on MPC before jumping to scheduling. MPC-source, basic assumptions about static loop bounds, etc.}
-
-For this treatment we make the following simplifying assumptions:
-
-\begin{enumerate}
-\item All statements in the program execute using the same protocol (sharing). That is, there is no share conversion. 
-%\ishaq{This is not an assumption, this is our setting. We work in the single protocol world.}
-\item There are two tiers of MPC instructions, local and remote. A local instruction (essentially just ADD) has cost $\beta$ 
-and a remote instruction (e.g., MUX, MUL, SHL, etc.) has cost $\alpha$, where $\alpha >> \beta$. We assume that all remote 
-instructions have the same cost.
-%\ishaq{Don't we need to distinguish between local (cheap) vs.
-interactive (expensive) instructions?
-I don't see how we could assume this.
--- Past suggestion from Vassilis: Ideally we should assume cost to be a convex function.}\ana{After discussion 1/12: replace single costs 1 with two levels of costs: $\alpha$ for non-local operations, e.g., MUL, and $\beta$ for local ops, e.g., ADD.}
-\item  In MPC frameworks, executing $n$ operations ``at once'' in a single SIMD operation costs a lot less than executing those $n$ operations one by one.
-Following Amdahl's law~\ana{citation!}, we write $\alpha = \frac{1}{s}p\alpha + (1-p)\alpha$, where $p$ is the fraction of execution time that benefits from amortization and $(1-p)$ 
-is the fraction that does not, and $s$ is the available resource. Thus, $n\alpha = \frac{n}{s}p\alpha + n(1-p)\alpha$. 
-For the purpose of the model we assume that $s$ is large enough and the term $\frac{n}{s}p\alpha$ amounts to a \emph{fixed cost} incurred regardless of
-whether $n$ is $10,000$ or just $1$. (This models the cost of sending a packet from party A to party B for example.) The bottom line is that amortized execution 
-of $n$ operations is $f(n) = \alpha_\mathit{fix} + n\alpha_{var}$ in contrast to unamortized execution $g(n) = n\alpha_\mathit{fix} + n\alpha_{var}$ and, of course,
-$\alpha_\mathit{fix} << n\alpha_\mathit{fix}$.
-%We assume infinite parallel capacity---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely $\alpha$ or $\beta$. 
-%This is a standard assumption in Cryptographic Parallel RAM. ABY presents empirical support for this assumption~\ana{Add citations. PRAM, ABY.}
-%\ishaq{Comment from Vassilis: We assume there are infinite parallel capacities, this is the standard assumption in PRAM (Cryptographic Parallel RAM).}\ana{PRAM assumption stays.
-We have to replace the unit cost of 1 with a suitable amortization function $f(n)$, which I don't think changes much in the cost modeling and analysis.}
-\item MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption,
-2 MUL instructions can be amortized in a single SIMD instruction that costs $\alpha_\mathit{fix} + 2\alpha_\mathit{var}$, however a MUL and a MUX instruction 
-still cost $2\alpha_\mathit{fix} + 2\alpha_\mathit{var}$ even when scheduled ``in parallel''. 
-%MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption,
-%2 MUL instructions scheduled in parallel benefit from amortization and cost $\alpha$, however a MUL and a MUX instructions scheduled
-%in parallel still cost $2\alpha$. 
-%\ishaq{We need to reword this assumption to something like "$K$ parallel MUL costs much less than $K$ (because they get amortized), but any mix of $K$  MUL and MUX still cost roughly $K$.
-Specifically, we should take away the low constants (2 and 1) because we know for low constants this is not true.} \ana{Again, core assumption that MUL and MUX don't benefit from amortization stays.
-We have to replace constant costs with functions, as in (3).}
-\end{enumerate}
-
-\input{prob_stmt_np_hardness.tex}
-
-\begin{figure*}
-  \includegraphics[width=1.0\linewidth]{figs_paper_SIMD/compiler_framework.png}
-  \caption{Compiler Framework. \ana{Change figure since we are not doing div-and-conquer.}}
-  \label{fig:compiler_framework}
-\end{figure*}
-
-\section{\bf Compiler Framework}
-\label{sec:preliminaries}
-
-\ana{Convention: use \texttt{shared, plain}, {\sf code}, $i$ and $I$ for counters outside of code.}
-
-%We assume arbitrarily nested loops in the MPC-source IR and read-only arrays.
-We assume that loops range from 0 to some constant N.
-Arrays are linearized (row-major order as in MOTION) and accesses are via functions of the induction variables of the enclosing loops.
-
-\figref{fig:compiler_framework} presents an overview of our compiler. This section describes the front-end phases of the compiler
-and Section~\secref{sec:vectorization} describes our vectorization algorithm. \ana{We'll need a section on MOTION code gen too.}
-%\ana{Need to state this more precisely.}
-
-\subsection{Syntax and Semantic Restrictions}
-\label{sec:syntax}
-
-Source syntax is essentially standard IMP syntax as shown below:
-%\begin{figure}[t]
-{\small
-\[
-\begin{array}{l@{~~~~~}l}
-  \begin{array}{l@{~}l@{~~~}l}
- e & ::= e \; \mathit{op} \; e \mid e \mid \x \mid  \texttt{const} \mid \texttt{A[$e$]} & \mathit{expression} \\
- % s &::= t \;\f & \mathit{field} \\
- %\md &::= t\;{\m}(t \;{\this}, \; t\;\x) \; \{\; \ol{t\;\y}\;\s;\;\return\;\y \; \} 
- %                 & \mathit{method} \\ 
-  s & ::= s;s \mid  & \mathit{sequence} \\
-  & \x = e \mid \texttt{A[$e$]} = e \mid & \mathit{assignment \; stmt} \\
-  & \texttt{for $i$ in range($I$)}: s \mid & \mathit{for \; stmt} \\
-  & \texttt{if $e$: $s$ else: $s$} & \mathit{if \; stmt} \\ 
-%   t &::= q \; {\sf C} & \mathit{qualified \; type}\\
-%   q &::= \high \mid \poly \mid \low & \mathit{FlowCFL \; qualifier}\\
-%    t' &::= q' \; \C & \mathit{field \; qualified \; type}\\
-%   q' &::= \tainted \mid \poly & \mathit{field \; qualifier}\\
-
-  \end{array}
-\end{array}
-\]
-}
-%\end{figure}
-
-The syntax allows for array accesses, arbitrarily nested loops, and if-then-else control flow. 
-Expressions are typed $q \; \tau$, where qualifier $q$ and base type $\tau$ are:
-
-{\small
-\[
-\begin{array}{l@{~~~~~}l}
-  \begin{array}{l@{~}l@{~~~}l}
- \tau & ::= \texttt{int}  \mid \texttt{bool} \mid \texttt{list[int]} \mid  \texttt{list[bool]} & \mathit{base \; types} \\
-  q & ::= \texttt{shared} \mid \texttt{plain} & \mathit{qualifiers} \\
-  \end{array}
-\end{array}
-\]
-}
-
-The type system is mostly standard, and in our experience, a sweet spot between readability and expressivity.
-The \texttt{shared} qualifier denotes shared values, i.e., ones shared among the parties and computed upon 
-under secure computation protocols; the \texttt{plain} qualifier denotes plaintext values. Shared lists denote 
-shared elements, but plaintext length. Subtyping is again as expected: $q_1 \; \tau <: q_2 \; \tau$ 
-iff $q_1 <: q_2$ where $\texttt{plain} <: \texttt{shared}$. Thus, we can convert a plaintext value into a shared one, 
-but not vice versa.
-
-Our compiler imposes certain semantic restrictions that it enforces throughout the various 
-phases of compilation. We note that in some cases, the restrictions 
-can be easily lifted and we plan to do so in future iterations of the work.
-
-\begin{enumerate}
-\item Loops are of the form $0 <= i < I$ and bounds are fixed at compile time. 
-It is a standard restriction in MPC that the bounds must be known at circuit-generation time.
-\item Arrays are one-dimensional. N-dimensional arrays are linearized and accessed 
-in row-major order and at this point the programmer is responsible for linearization
-and access. (This restriction can be easily lifted.)
-\item Array subscrpts are plaintext values as specified by the rule:
-\begin{semantics}
-    \ntyperule{Array Access}{
-%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
-  \Gamma \vdash e : \texttt{plain int} \quad \Gamma \vdash \texttt{A} : q \; \texttt{list}[\tau] \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
-}{
-  \Gamma \vdash \texttt{A[$e$]} : q \; \tau
-}
-\end{semantics}
-The subscript $e$ is a function of the indices of the enclosing loops. 
-For read access, the compiler allows an arbitrary such function. 
-However, it restricts write access to \emph{canonical writes}, i.e., {\sf A[i,j,k] = ...}
-where $i$, $j$ and $k$ loop over the three dimensions of {\sf A}.  
-Write access such as for example {\sf A[i,j+2] = ...} is not allowed. 
-%\item \ana{Add rules for logical ops}
-%\item \ana{Add rules for arithmetic ops}
-\item The final restriction involves MUX as expressed by the rule: 
-\begin{semantics}
-    \ntyperule{MUX}{
-%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
-  \Gamma \vdash e_1 : q_1 \; \texttt{bool} \quad \Gamma \vdash e_2 : q_2 \; \tau \quad \Gamma \vdash e_2 : q_2 \; \tau \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
-}{
-  \Gamma \vdash \texttt{MUX($e_1$,$e_2$,$e_3$)} : q_1 \vee q_2 \vee q_3 \; \tau
-}
-\end{semantics}
-\end{enumerate}
-Importantly, the arguments of MUX are restricted to base types, as passing array arguments 
-is not valid MOTION MPC. \ana{Check this.} This causes a slight inconvenience as we could not write 
-\begin{pythonn}
-if e: A[i] = val
-\end{pythonn}
-Instead we had to write
-\begin{pythonn}
-if not(e): val = A[i]
-A[i] = val
-\end{pythonn}
-
-We note that while the programmer writes 
-annotations at the level of IMP Source (recall \ana{ref to overview section}), the 
-annotations propagate through the transformations and are checked at the 
-level of MPC Source.
-
-%\ana{Disadvantage of MUX}
-
-For the rest of this section we write $i,j,k$ to denote the loop nest: $i$ is the outermost loop, $j$, is immediately nested in $i$, and so on until $k$
-and we use $I,J,K$ to denote the corresponding upper bounds. We write $A[i,j,k]$ to denote canonical access 
-to an array element. In the program, canonical access is achieved via the standard row-major order formula: $(J*K)*i + K*j + k$.
-To simplify the presentation we describe our algorithms in terms of three-element tuples $i,j,k$, however, discussion easily generalizes to 
-arbitrarily large loop nests.
-
-\subsection{From IMP Source to SSA}
-
-Our compiler translates from Source to SSA as follows:
-
-\begin{enumerate}
-  \item
-        \textbf{Parsing}:
-        Use Python's \texttt{ast} module to parse the input source code to a Python AST.
-  \item
-        \textbf{Syntax checking}:
-        Ensure that the AST matches the restricted subset defined in Section~\secref{sec:syntax}.
-        This step outputs an instance of the \texttt{restricted\_ast.Function} class, which represents our restricted subset of the Python AST.
-  \item
-        \textbf{3-address CFG conversion}:
-        %\benjamin{TODO: Is this a good amount of detail?} \ana{This is good amount of detail for the Tech report. In the actual paper we'll shorten a lot and add pointers to the Tech report.}
-
-        Convert the restricted-syntax AST to a three-address control-flow graph.
-        To do this, first,
-        add an empty basic block to the CFG and mark it as current.
-        Next, for each statement in the restricted AST's function body,
-        process the statement.
-        Statements can either be for-loops, if-statements, or assignments (as in~\secref{sec:syntax}).
-        Rules for processing each kind of statement are given below:
-        \begin{enumerate}
-          \item
-                \textbf{For-loops}:
-                Create new basic blocks for
-                the loop condition
-                (the \emph{condition-block}),
-                the loop body
-                (the \emph{body-block}),
-                and the code after the loop
-                (the \emph{after-block}).
-                Insert a jump from the end of the current block to the condition-block.
-                Then, mark the condition-block as the current block.
-                Insert a for-instruction at the end of the current block with the loop counter variable and bounds from the AST.
-                Next, add an edge from the current block to the after-block labeled ``FALSE'' and
-                an edge from the current block to the
-                body-block labeled ``TRUE''.
-                Then, set the body-block to be the current block
-                and process all statements in the AST's loop body.
-                Finally, insert a jump to the condition-block and set the after-block as current.
-          \item
-                \textbf{If-statements}:
-                Create new basic blocks for
-                the ``then'' statements of the if-statement
-                (the \emph{then-block}),
-                the ``else'' statements of the if-statement
-                (the \emph{else-block}),
-                and the code after the if-statement
-                (the \emph{after-block}).
-                At the end of the current block,
-                insert a conditional jump to the then-block or else-block
-                depending on the if-statement condition in the AST.
-                Next,
-                mark the then-block as current,
-                process all then-statements,
-                and add a jump to the after-block.
-                Similarly,
-                mark the else-block as current,
-                process all else-statements,
-                and add a jump to the after-block.
-                Finally, set the after-block to be the current block,
-                and give it a \emph{merge condition} property equal to the condition of the if-statement.
-          \item
-                \textbf{Assignments}:
-                In the restricted-syntax AST,
-                the left-hand side of assignments
-                can be a variable or an array subscript.
-                If it is an array subscript, e.g., {\sf A[i] = x},
-                change the statement to {\sf A = Update(A, i, x)}.
-                If the statement is not already three-address code,
-                for each sub-expression in the right-hand side of the assignment,
-                insert an assignment to a temporary variable.
-        \end{enumerate}
-  \item
-        \textbf{SSA conversion}:
-        Convert the 3-address CFG to SSA with Cytron's algorithm.
-\end{enumerate}
-
-\subsection{From SSA to MPC Source} %: MUX Nodes and Pseudo $\phi$-nodes}
-
-Once the compiler converts the code to SSA,
-it transforms $\phi$-nodes that correspond to if-statements into MUX nodes.
-From the 3-address CFG conversion step,
-$\phi$-nodes corresponding to if-statements will be in a basic block
-with the merge condition property.
-For example, if $X_3 = \phi(X_1,X_2)$ is in a block with merge condition $C$,
-the compiler transforms it into $X_3 = \mathit{MUX}(C, X_1, X_2)$.
-Next, the compiler runs the dead code elimination algorithm from Cytron's SSA paper.
-
-Next, the control-flow graph is \emph{linearized} into MPC Source,
-which has loops but no if-then-else-statements.
-This means that both branches of all if-statements are executed,
-and the MUX nodes determine whether to use results from the then-block or from the else-block.
-The compiler linearizes the control-flow graph with a variation of breadth-first search.
-Blocks with the ``merge condition'' property are only considered
-the second time they are visited,
-since that will be after both branches of the if-statement are visited. 
-(The Python AST naturally gives rise to a translation where each conditional has exactly two targets, 
-and each ``merge condition'' block has exactly two incoming edges, a TRUE and a FALSE edge.
-Thus, each $\phi$-node has exactly two multiplexer arguments, which dovetails into MUX.
-This is in contrast with Cytron's algorithm which operates at the level of the CFG and allows for 
-$\phi$-nodes with multiple arguments.)
-Each time the compiler visits a block,
-it adds the block's statements to the MPC source.
-If the block ends in a for-instruction,
-the compiler recursively converts the body and code after the loop to MPC source
-and adds the for-loop and code after the loop to the main MPC source.
-If the block does not end in a for-instruction,
-the compiler recursively converts all successor branches to MPC source and
-appends these to the main MPC source.
-
-Now, the remaining $\phi$-nodes in MPC source are the loop header nodes. We call these nodes \emph{pseudo} $\phi$-nodes 
-and we write \python{PHI} in MPC Source. A pseudo $\phi$-node $X_1 = $\python{PHI}$(X_0,X_2)$ in a loop header is evaluated 
-during circuit generation. If it is the 0-th iteration, then the $\phi$-node evaluates to $X_0$, otherwise, it evaluates to $X_2$.
-
-%\ana{We need to double check that the following captures all cases of def-use edges.}
-
-\subsection{Dependence Analysis}
-
-The next steps construct the dependence graph and \ana{Add an intro here.}
-
-\subsubsection{Def-use Edges}
-
-The dependence graph has the following def-use edges:
-
-\begin{itemize}
-\item same-level edge $X\rightarrow Y$ where $X$ and $Y$ are in the same loop nest, say $i,j,k$.
-E.g., the def-use edge 9 to 10 in the Biometric MPC Source in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source} is a same-level edge.
-A same-level edge can be a back-edge in which case a $\phi$ node is the target of the edge.
-E.g., 15 to 4 in Biometric is a same-level back-edge.
-\item outer-to-inner $X\rightarrow Y$ where $X$ is in an outer loop nest, say $i$, and $Y$ is in an inner one, say $i,j,k$. E.g., 1 to 4 in Biometric forms is an outer-to-inner edge.
-%A $\phi$-node can be a source or a target of an outer-to-inner forward edge.
-\item inner-to-outer $X\rightarrow Y$ where $X$ is a $phi$-node in an inner loop nest, $i,j,k$, and $Y$ is in the enclosing loop nest $i,j$.
-E.g., the def-use from 8 to 12 gives rise to an inner-to-outer edge.
-%E.g. $\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$ is an inner-to-outer edge. 
-An inner-to-outer edge can be a back-edge as well, in which case both $X$ and $Y$ are phi-nodes with the source $X$ in a loop nested into $Y$'s loop (not necessarily immediately).
-
-%Note that the source is \emph{always} a $\phi$-node of the $i,j,k$ loop and the target is in the immediately enclosing loop.
-The interpretation of this edge is that the use node $Y$ uses the definition made in the last iteration of the inner loop.
-%\ben{The current representation of pseudo $\phi$-nodes shows them attached to the loop header (i.e. in the loop).  We may want to clarify that they are also evaluated at the loop's termination.}
-%\item same-level back-edge $X\rightarrow Y$.
-$Y$ is a phi-node in the header of the loop and $X$ is a definition of the variable in the loop body.
-E.g., $\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{sum}_1,\texttt{min}_1)$ to $\texttt{min}_0 = \phi(\texttt{min}_1, 10000)$ in Biometric is a same-level back-edge.
-%\item inner-to-outer back-edge $X\rightarrow Y$: $X$ and $Y$ are both $\phi$-nodes for some variable. The source $X$ is in a loop nested into $Y$'s loop (not necessarily immediately).
-
-\item mixed forward edge $X\rightarrow Y$.
-$X$ is in some loop $i,j,k$ and $Y$ is in a loop nested into $i,j,k'$.
-We transform mixed forward edges as follows.
-Let $\x$ be the variable defined at $X$.
-We add a variable and assignment $\x' = \x$ immediately after the $i,j,k$ loop.
-Then we replace the use of $\x$ at $Y$ with $\x'$.
-This transforms a mixed forward edge into an "inner-to-outer" forward edge followed by an outer-to-inner forward edge.
-Thus, Basic Vectorization handles one of "same-level", "inner-to-outer", or "outer-to-inner" def-use edges.
- \end{itemize}
-
-\subsubsection{Closures}
-
-We define $\mathit{closure}(n)$ where $n$ is a phi-node. Intuitively, it computes the set of nodes (i.e., statements) that form a dependence cycle with $n$. The closure of $n$ is defined as follows:
-\begin{itemize}
-\item $n$ is in $\mathit{closure}(n)$
-\item $X$ is in $\mathit{closure}(n)$ if there is a same-level path from $n$ to $X$, and $X \rightarrow n$ is a same-level back-edge.
-\item $Y$ is in $\mathit{closure}(n)$ if there is a same-level path from $n$ to $Y$ and there is a same-level path from $Y$ to some $X$ in $\mathit{closure}(n)$.
-\end{itemize}
-
-\subsection{Taint Analysis}
-
-We require that all inputs are marked as either shared or plaintext, however, we do not require any other qualifier annotations in the program.
-Then we infer if intermediate variables and expressions are shared through taint analysis with "taintedness" referring to the shared attribute.
-Specifically, the compiler follows the following rules, which are standard in positive-negative qualifier systems (here \texttt{shared} is the positive qualifier and \texttt{plain} is the negative one):
-\begin{enumerate}
-\item Loop counters and array upper bounds are always \texttt{plain}.
-\item If any variable on the right-hand side of an assignment is shared, then the assigned variable is \texttt{shared}.
-%\item If all variables on the right-hand side of an assignment are plaintext, then the assigned variable is plaintext \ana{Will be covered by the last rule.}
-\item Any variables that cannot be determined as shared via the above rules are \texttt{plain}.
-\end{enumerate}
-
-%The first two rules are standard for taint analysis, and the third rule follows from the MPC problem statement. \ben{Is the explanation for the third rule correct?}  \ana{Yes, that is exactly it!}
-The final rule is needed in examples like the following. In the below snippet \texttt{sum!2} and \texttt{sum!3} form a dependency cycle and there is no \texttt{shared} value that flows to either one. 
-They are inferred as plaintext. 
-%cannot be marked as plaintext through simple taint analysis:
-{\small
-\begin{pythonn}
-plaintext_array = [0, 1, 2, ...]
-sum!1 = 0
-for i in range(0, N):
-    sum!2 = PHI(sum!1, sum!3)
-    sum!3 = sum!2 + plaintext_array[i]
-\end{pythonn}
-}
-%\ben{I think the above example is unnecessary and could be replaced by an explanation of how "untainted" variables are implicitly plaintext, but I couldn't think of a way to phrase that.}\ana{I left the example, we can leave it in the TR, remove from paper.}
-
-When converting to MOTION code, any plaintext value used in the right-hand side of a shared assignment is converted to a shared value for that expression. 
-%\ben{Is this necessary to include?} \ana{Yes, that's important.}
-
-Importantly, the taint analysis works on MPC Source, which lacks if-then-else control flow. 
-This significantly simplifies treatment as there is no need to handle conditionals and implicit flow.
-
-\subsection{From (Optimized) MPC Source to MOTION}
-\label{sec:MOTION}
-\ana{Ben, add the MOTION section here.}
-
 \section{Vectorization}
@@
 followed by our vectorization algorithm.
-
-%\ana{Add back-edges into Phase 1.
-A back-edge from a non-phi-node in loop $i$ to a phi-node in loop $i$'s header is a same-level edge.
-The only difference with the handling of normal same-level forward def-use is that the operand index will become $i-1$.
-A back edge from a phi-node in loop $j$ to a phi-node in loop $i$, where $j$ is nested in $i$ is an inner-to-outer edge and will require dropping dimension.
-We can show these are the only kinds of back-edges that may occur.}
-
-%\subsection{Algorithm (Still work in progress.)}
 
@@
 
-\begin{comment}
-We define the $\mathit{raise\_dim}$ (raise dimensions) and $\mathit{drop\_dim}$ (drop dimension) functions.
-Raise dimension "lifts" a lower-dimension array (\ana{The right term here is tensor, I am pretty sure!}) into a higher dimension one.
-This is necessary when a lower-dimensional array is used in a higher dimensional loop and, essentially, is just copying of values.
-For example, Biometric contains the statement $d = \mathit{SUB}(S[i, j], C[j])$, in the $j$-loop which is nested into the $i$-loop.
-Here $C$ is a one-dimensional array, however, to vectorize across both loops it is necessary to turn it into a two-dimensional array: $C[i,j]$ becomes $[C[0],C[1],...C[J],C[0],C[1],...C[J], ...
-C[0],C[1],...C[J]]$, which turns the row into a matrix of $I$ identical rows.
-(We use capital letters to denote the upper bounds of loops, e.g., $J$ is the upper bound of the $j$-loop and $I$ is the upper bound of the $i$-loop.)
-
-$\mathit{raise\_dim}(A[i_1,...,i_k],i_j)$ is defined as follows.
-It results in a new {k+1}-dimensional array $A'$ where for every $0 \le i_{k} < I_{k+1}$, $A'[i_1, \ldots, i_j, \ldots, i_k] = A[i_1, \ldots, i_k]$.
-Adding $n$ dimensions is trivially extended as a composition of $n$ $\mathit{raise\_dim}$ that each adds a single dimension.
-
-As expected, drop dimension turns a higher-dimensional array into a lower-dimensional one.
-The key use case is an inner-to-outer def-use edge.
-The code may define a variable, e.g., $\x$ in an inner loop, say $j$, then use this variable in an enclosing loop, say $i$.
-Our algorithm may vectorize the computation of $\x$ in the $j$-loop thus producing a vector $\x[j]$ where $\x[1]$ is the value of \x after the 1st iteration and so on.
-Drop dimension states that the outer loop will use the value of the variable at the last iteration.
-
-$\mathit{drop\_dim}(A[i_1,\ldots, i_j, \ldots,i_k], i_j)$ produces a k-dimensional array $A'$ where $A'[i_1,\ldots,i_k] = A'[i_1,\ldots, I_j-1, \ldots,i_k]$.
-In our analysis, dimensions are always dropped at the end, and again, one can define dropping $n$ dimensions as a composition of $n$ $\mathit{drop\_dim}$.
-
-\ben{I'm not sure if I wrote up the above paragraphs correctly -- the idea is that we're just copying values to extend over dimension $j$ or only retain its last elements.
-The current writeup implies that $1 < j < k$ though, which isn't always true.
-Python implementations of $\mathit{raise\_dim}()$ and $\mathit{drop\_dim}()$ can be found here: \url{https://github.com/milana2/ParallelizationForMPC/blob/master/compiler/compiler/motion_backend/reference_implementations.py}}
-
-\ana{This section is still informal. Needs work to make more precise.}
-\end{comment}
-
 \subsection{Basic Vectorization}
@@
 \input{biometric_vectorization_example.tex}
-
-\begin{comment}
-\subsection{Example: Biometric}
-
-\subsubsection{MPC-source}
-
-Below is the MPC-source resulting from Benjamin's analysis. It gives rise to a corresponding dependence graph. (Not shown.)
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0 = \phi(-1, \texttt{index}_1)$}
-\STATE {$\texttt{min}_0 = \phi(10000,\texttt{min}_1)$}
-
-\STATE \COMMENT {Begin of inner loop $j$.}
-\STATE {$\texttt{sum}_0 = \phi(0, \texttt{sum}_1)$}
-\STATE {$\texttt{d} = \mathit{SUB}(\texttt{S}[i,j], \texttt{C}[j])$}
-\STATE {$\texttt{p} = \mathit{MUL}(\texttt{d},\texttt{d})$}
-\STATE {$\texttt{sum}_1 = \texttt{ADD}(\texttt{sum}_0,\texttt{p})$}
-\STATE \COMMENT {End of inner loop $j$.}
-\STATE {$\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$}
-\STATE {$\texttt{index}_1 = \mathit{MUX}(\texttt{c},\texttt{index}_0,i)$}
-\STATE {$\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{min}_0,\texttt{sum}_0)$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-\subsubsection{Phase 1 of Basic Vectorization:}
-
-The transformation preserves the dependence edges.
-It raises the dimensions of scalars and optimistically vectorizes all operations.
-The next phase discovers loop-carried dependences and removes affected vectorization.
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$}
-\STATE {$\texttt{min}_0[i] = \phi(10000, \texttt{min}_1[i-1])$}
-
-\STATE \COMMENT {Begin of inner loop $j$.}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...], \texttt{sum}_1[i,j-1])$}
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{sum}_1[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE \COMMENT {End of inner loop $j$.}
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {$\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$}
-\STATE {$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$}
-\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-\ishaq{At least for input variables: when raising dimensions, we should keep some meta data around so we can later write code that could use 1 share if it knows the other N shares are just duplicates of it.}
-
-\subsubsection{Phase 2}
-
-This phase analyzes statements from the innermost loop to the outermost. The key point is to discover loop-carried dependencies and re-introduce loops whenever dependencies make this necessary.
-
-Starting at the inner phi-node $\texttt{sum}_0[i,j] = \phi([0,0,...],\texttt{sum}_1[i,j-1])$, the algorithm first computes its closure.
-The closure amounts to the phi-node itself and $\texttt{sum}_1[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$, accounting for the loop-carried dependency of the computation of \texttt{sum}.
-The algorithm replaces this closure with a FOR loop on $j$ removing vectoriaztion on $j$.
-Note that the SUB and MUL computations remain outside of the loop as they do not depend on phi-nodes that are part of cycles.
-The algorithm adds same-level edges, one from $\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = ...$ to the monolithic FOR-loop node, and one from the FOR-loop node to $\texttt{c}[\stackrel{\rightarrow}{i}] = ...$.
-
-\ishaq{For our radar: Such dependencies will not involve phi nodes only. e.g. $a[i] = a[i] + a[i-1]$}
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0[i] = \phi(-1, \texttt{index}_1[i-1])$}
-\STATE {$\texttt{min}_0[i] = \phi(10000, \texttt{min}_1[i-1])$}
-
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-
-\STATE \COMMENT {Begin of inner loop $j$. Loop is now much "shorter"!}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...], \texttt{sum}_1[i,j-1])$}
-\STATE {FOR j=0; j<D; j++} \COMMENT{ Will turn into a MOTION loop.}
-\STATE \hspace{0.25cm}{$\texttt{sum}_1[\stackrel{\rightarrow}{i},{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},{j}],\texttt{p}[\stackrel{\rightarrow}{i},{j}])$}
-\STATE \COMMENT {End of inner loop $j$.}
-
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {$\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$}
-\STATE {$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$}
-\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-Next, analysis moves to outer dimension $i$. There are two phi-nodes, $\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1])$ and $\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$.
-The closure of the first is
-\[
-\begin{array}{l}
-\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1]) \\
-\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
-\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])
-\end{array}
-\]
-and the closure of the second one is
-\[
-\begin{array}{l}
-\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1]), \\
-%\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
-\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])
-\end{array}
-\]
-
-Since the two closures \emph{do not} intersect, we have two distinct FOR-loops on $i$.
-The first FOR loop (on \texttt{min}) is scheduled first, because of the dependence edge from $\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$ to
-$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$.
-We can now rewrite those statements, getting rid of the vectorization on $i$ in the two FOR loops.
-Notably, the SUB and MUL computations are fully vectorizable and the ADD computation is vectorizable across the $i$-dimension.
-
-\begin{algorithmic}
-
-\STATE
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB\_SIMD}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL\_SIMD}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$} \\
-\STATE {}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...],\texttt{sum}_1[i,j-1])$}
-\STATE {FOR j=0; j<D; j++} \COMMENT{ Turns into a MOTION loop, was $j$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{sum}_1[\stackrel{\rightarrow}{i},{j}] = \mathit{ADD\_SIMD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},{j}],\texttt{p}[\stackrel{\rightarrow}{i},{j}])$}
-\STATE {}
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {}
-\STATE {$\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1])$}
-\STATE {FOR i=0; i<N; i++} \COMMENT{ Turns into a MOTION loop, was $i$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{c}[{i}] = \mathit{CMP}(\texttt{sum}'_0[{i}],\texttt{min}_0[{i}])$}
-\STATE \hspace{0.25cm}{$\texttt{min}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{min}_0[{i}],\texttt{sum}'_1[{i}])$}
-\STATE {}
-\STATE {$\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$}
-\STATE {FOR i=0; i<N; i++} \COMMENT{ Turns into a MOTION loop, was $i$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{index}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{index}_0[{i}],i)$}
-
-\ben{The above code implicitly stores the final values of min\_0 and index\_0 in their last indices, but there is no primitive like drop\_dim to make it clear.
-This will have to be added in the SSA code generation (though it might not be necessary to include in the paper?).
-It also is pretty wishy-washy about the size and layout of each vector (again more of an implementation issue), but maybe we should have a different syntax for array accesses for read/write (e.g.
-the ADD\_SIMD steps) vs the array definitions (e.g.
-the SUB\_SIMD and raise/drop\_dim statements)}
-
-\end{algorithmic}
-\end{comment}
 
@@
 }
-
-\begin{comment}
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE \hspace{0.25cm} A[i] = B[i] + 10;
-\STATE \hspace{0.25cm} B[i] = A[i] * D[i-1];
-\STATE \hspace{0.25cm} C[i] = A[i] * D[i-1];
-\STATE \hspace{0.25cm} D[i] = B[i] * C[i];
-\end{algorithmic}
-
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{A}_0$ = $\phi(\texttt{A},\texttt{A}_1)$
-\STATE 2. $\texttt{B}_0$ = $\phi(\texttt{B},\texttt{B}_1)$
-\STATE 3. $\texttt{C}_0$ = $\phi(\texttt{C},\texttt{C}_1)$
-\STATE 4. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$
-\STATE 5. $\texttt{A}_1$ = $\texttt{A}_0$; $\texttt{A}_1$[i] = $\texttt{B}_0$[i] + 10; \COMMENT{equiv. to Update}
-\STATE 6. $\texttt{B}_1$ = $\texttt{B}_0$; $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 7. $\texttt{C}_1$ = $\texttt{C}_0$; $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 8. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
-\end{algorithmic}
-\end{comment}
 
@@
 \begin{comment}
-The final step is to remove array phi-nodes that are left without incoming back edges after removal of infeasible edges.
-Such a phi-node is redundant when there are no loop-carried dependences that are due to writes to the corresponding array.
-In our running example, the phi-nodes for \texttt{A}, \texttt{B}, and \texttt{C}, are redundant.
-
-Removing redundant phi-nodes proceeds from \emph{innermost to outermost} loops as follows.
-Denote a redundant phi-node by $\texttt{A}_1 = \phi(\texttt{A}_0,\texttt{A}_k)$. $\texttt{A}_0$ is the array at the beginning of the loop
-and $\texttt{A}_k$ is the array at the end of an iteration (including the last iteration). We consider two cases:
-\begin{enumerate}
-\item[Case 1] Inner-to-outer or same-level def-use edge
-\begin{itemize}
-\item If use is an array read $ ... = \texttt{A}_1[f(i)]$, then replace $\texttt{A}_1$ with $\texttt{A}_0$. This is correct, i.e., it is guaranteed that the use
-references $\texttt{A}_0$; had the use been referencing an array modified in an earlier iteration, the back-edge to the phi-node would
-not have been removed and the phi-node would not have been classified as redundant.
-\item If use is an array write (update) $\texttt{A}_2 = \texttt{A}_1; \texttt{A}_2[i] = ... $ equiv. to $\texttt{A}_2 = \mathit{update}(\texttt{A}_1, i, ...)$ then handling becomes more tricky.
-The semantics of this update is that when $i$ is 0, then $\texttt{A}_0$ is being read.
-But when $i > 0$, then $\texttt{A}_k$ is being read.
-If we are at this case, this means that the def-use edge is same-level and we are removing a redundant node in the loop enclosing the update.
-The removal of the original phi-node entails pushing the phi-node at the update:
-$\texttt{A}_2 = \phi(\texttt{A}_0,\texttt{A}_k); \texttt{A}_2[i] = ... $ equiv. to $\texttt{A}_2 = \mathit{update}(\phi(\texttt{A}_0,\texttt{A}_k), i, ...)$.
-\item If use is a phi-node $... = \phi(\texttt{A}_1,...)$, that means we have an outer-to-inner edge. The
-
-\end{itemize}
-\end{enumerate}
-For each outer-to-inner or same-level def-use edge
-that originates at the node, replace $\texttt{A}_1$ at the use node by $\texttt{A}_0$.
-For each inner-to-outer def-use edge, replace $\texttt{A}_1$ by $\texttt{A}_k$. In this case, the use references the array
-that results from the computation in the loop.
-
-Our running example is transformed as follows. Removing the phi-nodes for \texttt{A}, \texttt{B}, and \texttt{C} transforms the
-replaces the references to $\texttt{A}_0$, $\texttt{B}_0$, and $\texttt{C}_0$, with references to $\texttt{A}$, $\texttt{B}$, and $\texttt{C}$
-respectively, i.e., the arrays before the loop.
-
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$
-\STATE 2. $\texttt{A}_1$ = $\texttt{A}$; $\texttt{A}_1$[i] = $\texttt{B}$[i] + 10;
-\STATE 3. $\texttt{B}_1$ = $\texttt{B}$; $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 4. $\texttt{C}_1$ = $\texttt{C}$; $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
-\end{algorithmic}
-\end{comment}
-
-\begin{comment}
 \subsubsection{Array MUX refinement}
@@
 
-\begin{comment}
-There are two changes to Basic Vectorization to account for write arrays. We can account for them during processing of raise dimension/drop dimension in Phase 1 of Basic Vectorization.
-
-One change to Basic vectorization is the expansion of dimension if the array write or read occurs in a nested loop.
-That is, if there is an update $\texttt{A}[i,j] = ...$ that occurs in loop dimensionality $i,j,k$, $\texttt{A}[i,j]$ will be rewritten into $\texttt{A}[i,j,k]$.
-Similarly, a read $\texttt{A}[f(i,j)]$ will be rewritten into $\texttt{A}[f(i,j),k]$.
-($\texttt{A}[f(i,j),k]$ will be reshaped as $f(i,j)*I*J+k$ during Phase 1 of Basic Vectorization.
-\ana{TODO: Double check.})
-\end{comment}
-
 In addition to the changes for the handling of target-less phi-nodes, Basic Vectorization has to handle def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable.
@@
 no raise dimension or drop dimension happens.
-
-\begin{comment}
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{A}_1$ = $\phi(\texttt{A}_0,\texttt{A}_2)$
-\STATE 2. $\texttt{B}_1$ = $\phi(\texttt{B}_0,\texttt{B}_2)$
-\STATE 3. $\texttt{C}_1$ = $\phi(\texttt{C}_0,\texttt{C}_2)$
-\STATE 4. $\texttt{D}_1$ = $\phi(\texttt{D}_0,\texttt{D}_2)$
-\STATE 5. $\texttt{A}_2 = update(\texttt{A}_1, \stackrel{\rightarrow}{i}, \texttt{B}_1[\stackrel{\rightarrow}{i}] + 10)$;
-\STATE 6. $\texttt{B}_2 = update(\texttt{B}_1, \stackrel{\rightarrow}{i}, \texttt{A}_2[\stackrel{\rightarrow}{i}] * \texttt{D}_1[\stackrel{\rightarrow}{i-1}])$;
-\STATE 7. $\texttt{C}_2 = update(\texttt{C}_1, \stackrel{\rightarrow}{i}, \texttt{A}_2[\stackrel{\rightarrow}{i}] * \texttt{D}_1[\stackrel{\rightarrow}{i-1}])$;
-\STATE 8. $\texttt{D}_2 = update(\texttt{D}_1; \stackrel{\rightarrow}{i}, \texttt{B}_2[\stackrel{\rightarrow}{i}] * \texttt{C}_2[\stackrel{\rightarrow}{i}])$;
-\end{algorithmic}
-\end{comment}
 
@@
 }
-
-\begin{comment}
-The cleanup phase can then get rid of $j$ in the \texttt{res} variables and \texttt{tmp} variables, so the loop operates over the \texttt{num\_bins}-size vectors. 
-This will lead to dropping raise\_dim and drop\_dim as well:
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I1 = raise_dim(i, ((i:num_bins),(j:N)))
-
-tmp1[I,J] = (A1[I,J] == I1[I,J])
-
-for j in range(0, N):
-   res2 = PHI(res1, res3) // now 1-dim array
-   tmp2[I] = ADD_SIMD(res2[I],B1[I,j])
-   tmp3[I] = MUX_SIMD(tmp1[I,j], res2[I], tmp2[I])
-   res3 = Update(res2, I, tmp3[I])
-return res2
-\end{verbatim}
-}
-\end{comment}
-
-\begin{comment}
-\paragraph{Example 2}
-
-Consider MPC-source of Histogram after removal of infeasible edges and redundant phi-nodes, and MUX refinement:
-
-\begin{algorithmic}
-
-\STATE res!1 = []
-\STATE for i: plaintext in range(0, num\_bins):
-\STATE \hspace{0.25cm} res!2 = $\phi$(res!1, res!3)
-\STATE  \hspace{0.25cm} ...
-\STATE res!4 = res!2 \COMMENT{Added due to the "mixed" def-use edge.}
-\STATE for i: plaintext in range(0, num\_bins):
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE \hspace{0.25cm}        for j: plaintext in range(0, N):
-\STATE \hspace{0.5cm}            res!5 = $\phi$(res!4, res!7)
-\STATE \hspace{0.5cm}            !2!3 = (A[j] == i)
-\STATE \hspace{0.5cm}            !3!3 = (res!5[i] + B[j])
-\STATE \hspace{0.5cm}            res!6 = res!5; res!6[i] = !3!3
-\STATE \hspace{0.5cm}            res!7 = res!6; res!7[i] = MUX(!2!3, res!6[i], res!5[i])
-
-\end{algorithmic}
-
-After Phase 1 of Basic Vectorization:
-
-\begin{algorithmic}
-\STATE res!1 = []
-\STATE for i: plaintext in range(0, num\_bins):
-\STATE \hspace{0.25cm} res!2 = $\phi$(res!1, res!3)
-\STATE  \hspace{0.25cm} ...
-\STATE res!4 = res!2 \COMMENT{Added due to the "mixed" def-use edge. Because the canonical dimensionality of res is 1, we DO NOT drop dimension of res!2.}
-\STATE A[$i,j$] = $\mathit{raise\_dim}$(A[$j$],$i$) \COMMENT{Copying row A num\_bin times. Reusing name.}
-\STATE B[$i,j$] = $\mathit{raise\_dim}$(B[$i$],$j$) \COMMENT{Copying column B N times. Reusing name.}
-\STATE res!4[$i,j$] = $\mathit{raise\_dim}$(res!4[$i$],$j$) \COMMENT{Reusing name.}
-\STATE for i: plaintext in range(0, num\_bins):
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE \hspace{0.25cm}        for j: plaintext in range(0, N):
-\STATE \hspace{0.5cm}            res!5 = $\phi$(res!4, res!7) \COMMENT{Do nothing. res!4 has been raised to proper dimension.}
-\STATE \hspace{0.5cm}            !2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{EQ}$(A[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], [[0,0...], [1,1...],...[num\_bins-1,num\_bins-1...]])
-\STATE \hspace{0.5cm}            !3!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{ADD}$(res!5[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$],B[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$])
-\STATE \hspace{0.5cm}            res!6 = res!5 \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.5cm}            res!6[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = !3!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$]
-\STATE \hspace{0.5cm}            res!7 = res!6
-\STATE \hspace{0.5cm}            res!7[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{MUX}$(!2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], res!6[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], res!5[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$])
-\end{algorithmic}
-
-After Phase 2 and Phase 3. The EQ operation is vectorizable across both dimensions, and the rest of the computation is vectorizable across the $i$-dimension.
-
-\begin{algorithmic}
-\STATE ...
-\STATE !2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{EQ\_SIMD}$(A[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], [[0,0...], [1,1...],...[num\_bins-1,num\_bins-1...]])
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE FOR j=0; j<N; j++; \COMMENT{MOTION loop}
-\STATE \hspace{0.25cm}            res!5 = $\phi$(res!4, res!7) \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.25cm}            !3!3[$\stackrel{\rightarrow}i,j$] = $\mathit{ADD\_SIMD}$(res!5[$\stackrel{\rightarrow}i,j$],B[$\stackrel{\rightarrow}i,j$])
-\STATE \hspace{0.25cm}            res!6 = res!5 \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.25cm}            res!6[$\stackrel{\rightarrow}i,j$] = !3!3[$\stackrel{\rightarrow}i,j$]
-\STATE \hspace{0.25cm}            res!7 = res!6
-\STATE \hspace{0.25cm}            res!7[$\stackrel{\rightarrow}i,j$] = $\mathit{MUX\_SIMD}$(!2!3[$\stackrel{\rightarrow}i,j$], res!6[$\stackrel{\rightarrow}i,j$], res!5[$\stackrel{\rightarrow}i,j$])
-\end{algorithmic}
-\end{comment}
-
-%\section{Divide-and-Conquer}
-%\label{sec:div_and_conq}
-%\ana{TODO: Now that we have broken FOR loops into smaller chunks, we can add Divide-and-conquer reasoning with Z3 and implement this additional transform.}
-
-\section{Implementation and Evaluation}
-\label{sec:implementation}
-
-\section{Related Work}
-\label{sec:related}
-
-\paragraph{MPC languages and compilers.}
-
-Languages and compilers for secure computation have seen significant attention and advances in recent years. The early MPC compilers Fairplay~\cite{CCS:BenNisPin08}, 
-and Sharemind~\cite{ESORICS:BogLauWil08} were followed by PICCO~\cite{CCS:ZhaSteBla13}, Obliv-C~\cite{Zahur:2015}, TinyGarble~\cite{SP:SHSSK15}, 
-Wystiria~\cite{SP:RasHamHic14}, and others. A new generation of MPC compilers 
-includes SPDZ/SCALE-MAMBA/MP-SPDZ~\cite{Keller:2020} and the ABY/HyCC/MOTION~\cite{NDSS:DemSchZoh15,CCS:BDKKS18,Braun:2022} frameworks. 
-These two families are the state-of-the art and
-are actively developed. Another recent development is Viaduct, a functional language and compiler that supports a range of secure computation 
-frameworks, including MPC and ZKP. Hastings et el. present a review of compiler frameworks~\cite{Hastings:2019}.
-
-While each of these languages and compilers brings in new ideas and advances, none addresses the problem of ``circuit independent'' intermediate 
-representation and optimization. We envision a classical compiler structure: (1) a Wysteria, Viaduct, Obliv-C, or IMP Source front end, including rich type systems and 
-AST-level semantic analysis, compile into the MPC Source IR, (2) MPC Source-level optimizations take place, followed by (3) back-end compilers into circuits. 
-Our focus is at the intermediate level.
-
-%{\begin{center}
-%\includegraphics[width=0.5\linewidth]{figs/focus.png}
-%\end{center}
-%}
-
-Many works focus on the implementation of MPC protocols exposing an API to the programmer. For example, the ABY/MOTION line of 
-compiler frameworks provides a library of MPC primitives; the programmer writes MPC programs in C++ on top of the library. These back ends implement 
-different protocols and allow for mixing, but notably, they leave it to the programer to assign different protocols to different parts of the 
-computation and perform share conversion accordingly. In addition, MOTION provides SIMD primitives, which allows for efficient execution 
-of MPC operations, but again, using SIMD primitives is the responsibility of the programer. There is interest in frameworks for automatic mixing, 
-e.g., \cite{CCS:BDKKS18,Ishaq:2019,Fang:2022}.
-
-Other works, e.g., Obliv-C~\cite{Zahur:2015}, Wystiria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021} focus on higher-level language design, particularly information-flow systems that 
-restrict flow between secure and insecure parts of the program. 
-%Key ideas such as ${\sf insecure} <: {\sf secure}$ qualifier systems are similar to the ideas we 
-%present in~\Secref{sec:interprocedural} and~\Secref{sec:zkp} and we intend to draw on these work. While we will design a high-level language expanding our 
-%intraprocedural IMP-MPC syntax and semantics into the \emph{interprocedural} case, our interest and focus lie primarily in the ``machine-independent'' optimizations 
-%thrust of the proposal.
-
-%There is significant development in ZKP languages as well. ~\ana{Alex Ozdemir's talk, ZoKrates, Cairo, others} Our work focuses on ZKP via MPC and our key idea
-%is the separation of the higher-level ``backend-independent'' compilation and optimization and the lower-level compilation into circuits.
-
-\paragraph{Classical HPC compilers.}
-Automatic vectorization is a longstanding problem in high-performance computing (HPC). 
-There are thousands of works in this area reflecting over 40 years of research. We presented a vectorization
-algorithm for MPC Source, essentially extending classical loop vectorization \cite{Allen:1987}. In HPC vectorization, conditional control flow 
-presents a challenge --- one cannot estimate the cost of a schedule or vectorize branches in a straightforward manner --- in contrast to 
-MPC Source vectorization. 
-We view Karrenberg's work on Whole function vectorization~\cite{Karrenberg:2015} as most closely related to ours --- it linearizes the program and vectorizes 
-both branches of a conditional applying masking to avoid execution of the branch-not-taken code, and selection (similar to MUX) to select the correct value based on the 
-result of the condition at runtime. 
-The problem is that masking and selection, or more generally, handling control predicates~\cite{Benabderrahmane:2010,Karrenberg:2015}, 
-can lead to \emph{slowdown}. 
-%As a result, vectorizing compilers and polyhedral compilers tend to focus on straight-line code due to the unpredictable 
-%overhead conditionals may introduce. \ana{add Louis noel pochet's benchmarks. citation on vectorizing compilers???}
-
-We argue that vectorization over linear MPC Source is a different problem, one that warrants a new look, while drawing from 
-results in HPC.
-Since both branches of the conditional and the multiplexer \emph{always} execute, not only can we apply aggressive vectorization on linear code, but (perhaps more importantly) 
-we can also build analytical models that accurately predict execution time. These models in turn would drive optimizations such as vectorization, protocol mixing, and others. 
-Vectorization meshes in with those additional optimizations in non-trivial ways.
-
-Furthermore, extensions of classical loop vectorization with array writes, arbitrary indexing, including non-affine indexing, and interaction with SSA are non-trivial 
-and present novel challenges and opportunities for contribution. Polyhedral parallelization~\cite{Benabderrahmane:2010} considers a higher-level source (typically AST) 
-representation, while our work takes advantage of linear MPC Source and SSA form. The work by Karrenberg~\cite{Karrenberg:2015} is rare in that space, in the sense that it considers 
-vectorization over SSA form, which has similarities to MPC Source. We consider different array representation, notion of dependence, and reasoning 
-about dependence, which we conjecture is more suitable for MPC Source. Buscher~\cite{Buscher:2018} considers SIMD-vectorization 
-at the level of source code, which then combines with circuit-level optimizations in the TinyGarble compiler. 
-He proposes using an off-the-shelf polyhedral compiler, however, application is limited to only two routines, 
-essentially just inner product and euclidian distance; it is unclear how effective the off-the-shelf compiler is. 
-In contrast, we consider vectorization at the level of MPC Source separating ``backend-independent'' vectorization
-and circuit-level amortization (done by MOTION). We apply our compiler on a wide range of routines.
-
-%\section{Future Work}
-%\label{sec:implementation_and_benchmarks}
-%\input{../sections/implementation_and_benchmarks.tex}
-%\input{../sections/evaluation.tex}
-
-\section{Conclusion and Future Work}
-\label{sec:conclusion}
-%\input{../sections/conclusion.tex}
-
-%\begin{acks}
-%    \ishaq{TODO}
-%\end{acks}
-
-%\printbibliography
-
-\bibliographystyle{plain}
-%\bibliography{c:/trink/Doc/MyTexFiles/texrefs}
-%\bibliography{mybib,library,library1,library2,library3,library_ana}
-\bibliography{library_ana,crypto_crossref}
-
-% that's all folks
-\end{document}

### 2022-04-28 overleaf/introduction.tex
Edits to 1 and 2.

@@
 Third, MOTION provides API for SIMD-level operations, which amortize cost and lead to significant improvement in memory footprint and
-throughput~\cite{NDSS:DemSchZoh15, CCS:ABFKLO18, Braun:2022}. It also enables MPC Source-level optimizations such SIMD vectorization.
+throughput~\cite{NDSS:DemSchZoh15, CCS:ABFKLO18, Braun:2022}. It enables MPC Source-level vectorization, a key focus of this paper.
 
@@
 appeared simpler than the classical scheduling problem. Unfortunately, we show that optimal scheduling is NP-hard via a reduction to the Shortest Common Supersequence (SCS) problem.
-Cost estimation is important as it drives not only vectorization but other optimizations such as protocol mixing optimizations and scheduling as well~\cite{Ishaq:2019, Fang:2022}.
+Cost modeling is important as it drives not only vectorization but optimizations such as protocol mixing and scheduling as well~\cite{Ishaq:2019, Fang:2022}.
 
@@
 Our results emphasize the importance of backend-independent optimizations --- vectorization (described in this work) and protocol mixing (tackled in previous works~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022}) are two optimizations readily available at the level of MPC Source.
-We believe that our work can lead to future work on backend-independent compilation and optimization, bringing in new MPC optimizations and combinations of optimizations, much in the vein of standard compilers, and bring MPC programming technology closer to practice and wider applicability.
+We believe that our work can lead to future work on backend-independent compilation and optimization, ushering new MPC optimizations and combinations of optimizations in the vein of standard compilers, and thus bringing MPC programming technology closer to practice and wider applicability.
 
@@
 The rest of the paper is organized as follows.
-\secref{sec:overview} presents an overview of the compiler.
-\secref{sec:model} describes our model for cost estimation and argues NP-hardness of optimal scheduling and \secref{sec:compiler} describes the compiler.
-\secref{sec:experiments} presents implementation and evaluation.
-\secref{sec:related} discusses related work and \secref{sec:conclusion} concludes.
+Section \secref{sec:overview} presents an overview of the compiler.
+Section \secref{sec:model} describes our model for cost estimation and argues NP-hardness of optimal scheduling.
+Section \secref{sec:compiler} details the phases of the compiler and Section \secref{sec:vectorization} focuses in on vectorization.
+\secref{sec:results} presents the experimental evaluation.
+Section \secref{sec:related} discusses related work and Section \secref{sec:conclusion} concludes.
 

### 2022-04-28 overleaf/overview.tex
Edits to 1 and 2.

@@
 An intuitive (and naive) implementation is as shown in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(a).
-Array \python{C} is the feature vector of \texttt{D} features that we wish to match and \python{S}
-is the database of \python{N} size-\python{D} vectors that we match against.
+Array {\sf C} is the feature vector of {\sf D} features that we wish to match and {\sf S}
+is the database of {\sf N} size-{\sf D} vectors that we match against.
 
@@
 syntax and imposes certain semantic restrictions. (We detail the restrictions in the following sections.)
-The programmer writes an iterative program and annotate certain inputs
-and outputs as \emph{shared}. In the example arrays \python{C} and \python{S}
-are shared, meaning that they store shares, however, the array sizes \python{D} and
-\python{N} respectively are plaintext. The code iterates over the entries in the
-database and computes the Euclidean distance (its square actually) of the current
-entry \python{S[i]} and \python{C}. The program returns the index of the vector that gives
+The programmer writes an iterative program and annotates certain inputs
+and outputs as \emph{shared}. In the example arrays {\sf C} and {\sf S}
+are \texttt{shared}, meaning that they store shares, however, the array sizes {\sf D} and
+{\sf N} respectively are plaintext. The code iterates over the entries in the
+database and computes the Euclidean distance of the current
+entry {\sf S[i]} and {\sf C} (its square actually). The program returns the index of the vector that gives
 the best match plus the corresponding sum of squares.
@@
 Second, it is linear. The conditional in lines 13-14 in IMP Source turns into the linear code in lines 12-16 in MPC Source.
-The test turns into a CMP operation (\python{t = CMP(sum!3,min\_sum!2)}), followed by the
+The test turns into the CMP operation {\sf t = CMP(sum!3,min\_sum!2)}, followed by the
 true-branch sequence, followed by the MUX operations. The first MUX operation selects the value
-of \python{min\_sum}: if \python{t} is true, then \python{min\_sum} gets the value of the second multiplexer
-argument,  \python{min\_sum!3}, otherwise it takes the value of the third argument, \python{min\_sum!2}.
+of {\sf min\_sum}: if {\sf t} is true, then {\sf min\_sum} gets the value of the second multiplexer
+argument,  {\sf min\_sum!3}, otherwise it takes the value of the third argument, {\sf min\_sum!2}.
 Third, MPC Source is a special form of SSA. The SSA $\phi$-nodes at the if-then-else (lines 13-15) turn into
-MUX operations, while the $\phi$-nodes at for loops turn into what we call pseudo PHI nodes with a straightforward semantics.
+MUX operations, while the $\phi$-nodes at for-loops turn into \emph{pseudo} PHI nodes with a straightforward semantics.
 
-We turn to our analytical model to compute the \emph{cost} of the program. Assuming fixed
+We turn to our analytical model to compute the \emph{cost} of the iterative program. Assuming
 cost $\beta$ for a local MPC operation (essentially just ADD) and cost $\alpha$ for a remote
@@
 A key contribution is the vectorizing transformation. We can compute all $N*D$
-subtraction operations (line 9) in a single SIMD instruction; similarly we can compute
-all multiplication operations (line 10) in a single SIMD instruction. And while we cannot
-vectorize computation of the N individual sums, we can compute the N sums in parallel.
+subtraction operations (line 9 in (b)) in a single SIMD instruction; similarly we can compute
+all multiplication operations (line 10) in a single SIMD instruction. And while computation
+of an individual sum remains sequential, we can compute the $N$ sums in parallel.
 Our compiler \emph{automatically detects these opportunities and transforms the program}.
@@
 can be fully vectorized into the SIMD SUB and MUL in optimized MPC Source (lines 9 and 10 in (c)).
-Notation \python{p[I,J]} denotes a 2-dimensional array with fully vectorized dimensions.
+Notation {\sf p[I,J]} denotes a 2-dimensional array with fully vectorized dimensions.
 The computation of sum (line 11 in (b))
 is sequential across the $j$-dimension, but it is parallel across the $i$-dimension.
-The loop in lines 12-16 in (c) illustrates; here \python{p[I,j]} refers to the $j$-th column in \python{p}.
+The loop in lines 12-16 in (c) illustrates; here {\sf p[I,j]} refers to the $j$-th column in {\sf p}.
 Unfortunately, CMP and MUX remain sequential.
@@
 %essentially the same ($\alpha$ or $\beta$) as an individual operation.
-We elaborate on the cost model in the following section but for now consider that
+We elaborate on the cost model in Section~\secref{sec:model} but for now consider that
 each operation has a \emph{fixed} portion (does benefits from amortization) and
@@
 The first term in the sum corresponds to the vectorized
-subtraction and multiplication, the second term corresponds to the FOR loop on $j$ and the third
-one corresponds to the remaining FOR loops on $i$.
+subtraction and multiplication (lines 9-10 in (c)), the second term corresponds to the FOR loop on $j$ (lines 12-16) and the third
+one corresponds to the remaining FOR loops on $i$ (lines 19-25). 
 Clearly, $2\alpha_\mathit{fix} + D\beta_\mathit{fix} + N(3\alpha_\mathit{fix}) << ND(2\alpha_\mathit{fix}+\beta_\mathit{fix}) + N3\alpha_\mathit{fix}$.

### 2022-04-28 overleaf/paper_SIMD.tex
Edits to 1 and 2.

--- 2022-04-27 overleaf/paper_SIMD.tex
@@
 
-\section{Introduction}
-\label{sec:introduction}
+% Intro Section
+\input{introduction.tex}
 
-Multi-party computation (MPC) allows $N$ parties $p_1,\dots, p_N$ to perform a computation on their private inputs securely.
-Informally, security means that the secure computation protocol computes the correct output (correctness) and it does not leak any information about the individual party inputs, other than what can be deduced from the output (privacy).
+% Overview Section
+\input{overview.tex}
 
-% The first results from the early 80's~\cite{FOCS:Yao82b,STOC:GolMicWig87,STOC:BenGolWig88,STOC:ChaCreDam88} introduced the problem and formally defined its security, and investigate feasibility in different models---e.g., passive or malicious adversaries, unconditional or computational security, with or without a broadcast channel.
-This spawned a rich theoretical literature of beautiful works improving and extending in various manners the original ideas.
+% Analytical Model Section
+\input{model.tex}
 
-MPC theory dates back to the early 1980-ies~\cite{FOCS:Yao82b,STOC:GolMicWig87,STOC:BenGolWig88,STOC:ChaCreDam88}. 
-Long the realm of theoretical cryptography, MPC has seen significant advances in programming technology in recent years. 
-These advances bring MPC closer to practice and wider applicability --- 
-MPC technology has been employed in real-world scenarios such as auctions~\cite{FC:BCDGJK09}, biometric identification~\cite{ESORICS:BlaGas11}, 
-and privacy-preserving machine learning~\cite{SP:MohZha17,CCS:MohRin18}. 
-The goal is to bring the technology to a level where programmers can write \emph{secure} and \emph{efficient} programs without 
-commanding extensive knowledge of cryptographic primitives.
+% Compiler Framework Section
+\input{compiler.tex}
 
-The problem, therefore, is to build a high-level programming language and a compiler, and there has been significant advance in 
-this space, e.g., ~\cite{CCS:BenNisPin08,ESORICS:BogLauWil08,CCS:ZhaSteBla13,Hastings:2019,Keller:2020,Acay:2021,Braun:2022} among other work. 
-Current research largely falls at the two ends 
-of the classical compiler: (1) work on \emph{front-end} language design and (2) work on \emph{back-end} protocol implementation. Work on language design 
-focuses on high-level constructs necessary to express multiple parties, computation by different parties, and information flow from one party 
-to another~\cite{SP:RasHamHic14, Acay:2021}. On the other end, work on protocol implementation focuses on cryptographic foundations and their efficient circuit-level
-implementation~\cite{NDSS:DemSchZoh15, CCS:ABFKLO18, Braun:2022}, e.g., implementation of operations (e.g., MUL, ADD) using different sharing protocols 
-(Boolean or Arithmetic GMW~\cite{STOC:GolMicWig87} or Yao's garbled circuits~\cite{FOCS:Yao82b}), as well as efficient share conversion from one representation 
-to another.
+% Vectorization Section
+\input{vectorization.tex}
 
-Earlier compilers did both back-end and front-end translation without a specific focus on either, as their aim was to demonstrate applicability of MPC on real-world
-programming problems. As the field advanced, works have focused more closely on front-end language design (e.g., Wysteria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021}) 
-or back-end ``circuit-level'' design and implementation (e.g., MOTION~\cite{Braun:2022}).
+% Results Section
+\input{results.tex}
 
-In this work we focus on an intermediate language and what we call \emph{backend-independent optimizations}, in a close analogy to \emph{machine-independent} optimizations
-in the classical compiler. The following figure summarizes our key idea:
-
-{\begin{center}
-\includegraphics[width=0.8\linewidth]{figs_paper_SIMD/focus.png}
-\end{center}
-}
-
-We emphasize the MPC Source intermediate representation and optimization over MPC Source. As in classical compilers, we envision different front ends (e.g., Wysteria,
-our front end IMP-MPC) compiling into MPC Source. MPC Source is particularly suitable for optimizations such as protocol mixing~\cite{Ishaq:2019, Fang:2022} and SIMD-vectorization, 
-which takes advantage of amortization at the circuit level. The MPC Source IR exposes the linear structure of MPC programs, which simplifies program analysis; this is
-in contrast to source, which has if-then-else constructs. In the same time, MPC Source is sufficiently ``high-level''  to support analyses and optimizations that take into account 
-control and data flow in a specific program. MPC Source is small in size and analysis is tractable, as opposed to analysis over an unrolled circuit~\cite{Ishaq:2019}. Again as
-in classical compilers, we envision translation of MPC Source (optimized or unoptimized) into MOTION, SPDZ, or other back-end code.
-
-%In contrast, the high-level program contains if-then-else and array writes, or the low-level circuit . 
-%The compilation problem, i.e., translation from a high-level language to circuits, is difficult and therefore, breaking the problem into sub-problems can  tractable.
-
-\subsection{Our Contribution} In this paper, we develop a compiler framework that takes a Python-like routine and produces MOTION code: we describe (a) the IMP-MPC language, its syntax and semantic restrictions, (b) translation from IMP-MPC into MPC Source, (c) a specific backend-independent optimization: novel SIMD-vectorization on MPC Source, and (d) translation from MPC Source into MOTION code.
-%\ana{I think we have to add more on MOTION, specifically mixing and amortization.}
-
-We focus on the MOTION framework as our back-end for several reasons.
-First, it is the state of the art in terms of performance~\cite{Braun:2022}.
-Second, it provides an API over efficient implementation for a wide variety of cryptographic operations in three different protocols ---  Arithmetic GMW, Boolean GMW, and BMR --- which allows for protocol mixing~\cite{Ishaq:2019, Fang:2022}, a known backend-independent optimization.
-Third, MOTION provides API for SIMD-level operations, which amortize cost and lead to significant improvement in memory footprint and
-throughput~\cite{NDSS:DemSchZoh15, CCS:ABFKLO18, Braun:2022}. It also enables MPC Source-level optimizations such SIMD vectorization.
-
-Our second contribution is an analytical model for cost estimation of amortized schedules. Originally, we hoped that optimal scheduling (under our model) was tractable, as the problem
-appeared simpler than the classical scheduling problem. Unfortunately, we show that optimal scheduling is NP-hard via a reduction to the Shortest Common Supersequence (SCS) problem. 
-Cost estimation is important as it drives not only vectorization but other optimizations such as protocol mixing optimizations and scheduling as well~\cite{Ishaq:2019, Fang:2022}.
-
-Our most important contribution is the implementation and evaluation of the compiler framework.
-We demonstrate expressivity of the source language by running the compiler on \ana{X} programs; these include classical MPC benchmarks such as PSI and Biometric matching, as well as kMeans, Histogram, and others.
-Vectorization leads to \ana{X}x improvement in number of gates on average, \ana{Y}x speedup in circuit generation time, \ana{Z}x speedup in the LAN setting, and \ana{Z}x speedup in the WAN setting using the Boolean GMW protocol.
-Results form BMR are similar.
-
-Our results emphasize the importance of backend-independent optimizations --- vectorization (described in this work) and protocol mixing (tackled in previous works~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022}) are two optimizations readily available at the level of MPC Source.
-We believe that our work can lead to future work on backend-independent compilation and optimization, bringing in new MPC optimizations and combinations of optimizations, much in the vein of standard compilers, and bring MPC programming technology closer to practice and wider applicability.
-
-\subsection{Outline}
-
-The rest of the paper is organized as follows.
-\secref{sec:overview} presents an overview of the compiler.
-\secref{sec:model} describes our model for cost estimation and argues NP-hardness of optimal scheduling and \secref{sec:compiler} describes the compiler.
-\secref{sec:experiments} presents implementation and evaluation.
-\secref{sec:related} discusses related work and \secref{sec:conclusion} concludes.
-
-\begin{comment}
-\begin{itemize}
-\item We define the scheduling problem for MPC. 
-We present an analytical model to reason about cost of schedules
-and show that scheduling is NP-hard via a reduction to the shortest 
-common supersequence problem.
-
-\item We present a compiler that takes an IMP-like high-level 
-program and produces amortized (i.e., vectorized) low-level cryptographic 
-code in the MOTION framework. Central contributions are 1) a novel compiler framework, 
-2) a vectorization algorithm that produces optimal schedules for a large 
-number of MPC programs, and 3) reasoning over output arrays without Array SSA; we remove 
-infeasible loop-carried dependences introduced by standard SSA to improve vectorization.
-
-\item We present an implementation and evaluation in the MOTION framework. 
-\ana{Fill in with final results. Mention benchmarks (standard + new ones + HyCC). }
-
-\end{itemize}
-\end{comment}
-
-\section{Overview}
-\label{sec:overview}
-
-\begin{table*}
-\begin{tabular}{ccc}
-\begin{minipage}{0.275\textwidth}
-{\small
-\begin{pythonn}
-def biometric(C: shared[list[int]], D: int, 
-      S: shared[list[int]], N: int) -> 
-      shared[tuple[int,int]]:
-   min_sum : int = MAX_INT
-   min_idx : int = 0  
-   for i in range(N): 
-      sum : int = 0
-      for j in range(D): 
-         # d = S[i,j] - C[j]
-         d : int = S[i * D + j] - C[j] 
-         p : int = d * d
-         sum = sum + p
-      if sum < min_sum:
-         min_sum : int = sum
-         min_idx : int = i
-   return (min_sum, min_idx)
-\end{pythonn}
-}
-\end{minipage}
-
-&
-
-\begin{minipage}{0.33\textwidth}
-{\tiny
-\begin{pythonn}
-min_sum!1 = MAX_INT
-min_idx!1 = 0
-for i in range(0, N):
-   min_sum!2 = PHI(min_sum!1, min_sum!4)
-   min_idx!2 = PHI(min_idx!1, min_idx!4)
-   sum!2 = 0
-   for j in range(0, D):
-      sum!3 = PHI(sum!2, sum!4)
-      d = SUB(S[((i * D) + j)],C[j]) 
-      p = MUL(d,d) 
-      sum!4 = ADD(sum!3,p) 
-   t = CMP(sum!3,min_sum!2) 
-   min_sum!3 = sum!3
-   min_idx!3 = i
-   min_sum!4 = MUX(t, min_sum!3, min_sum!2) 
-   min_idx!4 = MUX(t, min_idx!3, min_idx!2) 
-return (min_sum!2, min_idx!2)   
-\end{pythonn}
-}
-\end{minipage}
-
-&
-
-\begin{minipage}{0.33\textwidth}
-{\small
-\begin{pythonn}
-min_sum!1 = MAX_INT
-min_idx!1 = 0
-# S^ is same as S. C^ replicates C N times: 
-S^ = raise_dim(S, ((i * D) + j), (i:N,j:D)) #S^[i,j] = S[i,j]
-C^ = raise_dim(C, j, (i:N,j:D)) #C^[i,j] = C[j]
-
-sum!2[I] = [0,..,0]
-# computes _all_  "at once"
-d[I,J] = SUB_SIMD(S^[I,J],C^[I,J])
-p[I,J] = MUL_SIMD(d[I,J],d[I,J])
-
-for j in range(0, D):
-   # sum!2[I], sum!3[I], sum!4[I] are size-N vectors 
-   # computes N intermediate sums "at once"
-   sum!3[I] = PHI(sum!2[I], sum!4[I])       
-   sum!4[I] = ADD_SIMD(sum!3[I],p[I,j])
-
-min_idx!3[I] = [0,1,...N-1]   
-for i in range(0, N):
-   min_sum!2 = PHI(min_sum!1, min_sum!4) 
-   t[i] = CMP(sum!3[i],min_sum!2)
-   min_sum!4 = MUX(t[i], sum!3[i], min_sum!2)  
-for i in range(0, N):
-   min_idx!2 = PHI(min_idx!1, min_idx!4)    
-   min_idx!4 = MUX(t[i], min_idx!3[i], min_idx!2)
-return (min_sum!2, min_idx!2)   
-\end{pythonn}
-}
-\end{minipage}
-
-\\
-
-(a) IMP Source & (b) MPC Source & (c) Optimized MPC Source
-\end{tabular}
-\caption{Biometric Matching: From IMP Source to MPC Source to Optimized MPC Source. }
-\label{tab:source_and_MPC_source_and_optimized_MPC_source}
-\end{table*}
-
-\subsection{Source}
-
-As a running example, consider Biometric matching, a standard MPC benchmark.
-An intuitive (and naive) implementation is as shown in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(a).
-Array \python{C} is the feature vector of \texttt{D} features that we wish to match and \python{S} 
-is the database of \python{N} size-\python{D} vectors that we match against.
-
-\begin{comment}
-{\small
-\begin{verbatim}
-def biometric(C: shared[list[int]], D: int, 
-              S: shared[list[int]], N: int) -> 
-                      tuple[shared[int], shared[int]]:
-    min_sum = 10000
-    min_index = 0  
-    for i in range(N): #loop over database
-        sum = 0
-        for j in range(D): #loop over features
-            d = S[i * D + j] - C[j] #i.e., d = S[i,j] - C[j]
-            p = d * d
-            sum = sum + p
-        if sum < min_sum:
-            min_sum = sum
-            min_index = i
-    return (min_sum, min_index)
-\end{verbatim}
-}
-\end{comment}
-
-Our compiler takes essentially standard IMP~\ana{Need citation here.} 
-syntax and imposes certain semantic restrictions. (We detail the restrictions in the following sections.)
-The programmer writes an iterative program and annotate certain inputs 
-and outputs as \emph{shared}. In the example arrays \python{C} and \python{S}
-are shared, meaning that they store shares, however, the array sizes \python{D} and
-\python{N} respectively are plaintext. The code iterates over the entries in the 
-database and computes the Euclidean distance (its square actually) of the current 
-entry \python{S[i]} and \python{C}. The program returns the index of the vector that gives 
-the best match plus the corresponding sum of squares.
-
-\subsection{MPC Source and Cost of Schedule}
-
-Our compiler generates an intermediate representation, MPC Source. MPC Source 
-for Biometric Matching is shown in 
-Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(b). First, MPC Source is an SSA form. 
-Second, it is linear. The conditional in lines 13-14 in IMP Source turns into the linear code in lines 12-16 in MPC Source. 
-The test turns into a CMP operation (\python{t = CMP(sum!3,min\_sum!2)}), followed by the 
-true-branch sequence, followed by the MUX operations. The first MUX operation selects the value
-of \python{min\_sum}: if \python{t} is true, then \python{min\_sum} gets the value of the second multiplexer
-argument,  \python{min\_sum!3}, otherwise it takes the value of the third argument, \python{min\_sum!2}.  
-Third, MPC Source is a special form of SSA. The SSA $\phi$-nodes at the if-then-else (lines 13-15) turn into 
-MUX operations, while the $\phi$-nodes at for loops turn into what we call pseudo PHI nodes with a straightforward semantics.
-
-\begin{comment}
-{\small
-\begin{pythonn}
-1. min_sum!1 = 10000
-2. min_index!1 = 0
-3. for i in range(0, N!0):
-4.    min_sum!2 = PHI(min_sum!1, min_sum!4)
-5.    min_index!2 = PHI(min_index!1, min_index!4)
-6.    sum!2 = 0
-7.    for j in range(0, D!0):
-8.      sum!3 = PHI(sum!2, sum!4)
-9.      d!3 = (S!0[((i * D!0) + j)] - C!0[j]) // MPC
-10.     p!3 = (d!3 * d!3) // MPC
-11.     sum!4 = (sum!3 + p!3) // MPC
-12.  !1!2 = (sum!3 < min_sum!2) // MPC
-13.  min_sum!3 = sum!3
-14.  min_index!3 = i
-15.  min_sum!4 = MUX(!1!2, min_sum!3, min_sum!2) // MPC
-16.  min_index!4 = MUX(!1!2, min_index!3, min_index!2) // MPC
-17.!2!1 = (min_sum!2, min_index!2)   
-\end{pythonn}
-}
-\end{comment}
-
-%The compiler linearizes the source turning conditionals into MUX statements. The PHI nodes
-%are remnants of the SSA IR; the compiler generates code that picks the correct value 
-%when producing MOTION output; the MOTION framework in turn linearizes 
-%loops when it generates the circuit.
-
-We turn to our analytical model to compute the \emph{cost} of the program. Assuming fixed 
-cost $\beta$ for a local MPC operation (essentially just ADD) and cost $\alpha$ for a remote
-MPC operation (e.g., MUX, CMP, and remaining operations), the cost of the iterative schedule 
-will be $ND(2\alpha+\beta) + N(3\alpha$).
-
-A key contribution is the vectorizing transformation. We can compute all $N*D$ 
-subtraction operations (line 9) in a single SIMD instruction; similarly we can compute 
-all multiplication operations (line 10) in a single SIMD instruction. And while we cannot
-vectorize computation of the N individual sums, we can compute the N sums in parallel. 
-Our compiler \emph{automatically detects these opportunities and transforms the program}. 
-It is standard that MPC researchers write vectorized versions of the Biometric
-program by hand; we are the first (to the best of our knowledge) to automatically 
-transform a naive, iterative MPC program into an unintuitive vectorized one.
-
-\subsection{Vectorized MPC Source and Cost of Schedule}
-
-Our compiler produces the vectorized program shown in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(c). 
-Note that this is still our intermediate representation, Optimized MPC Source. Subsequently,
-the compiler turns this code into MOTION variables, loops and SIMD primitives, which MOTION then 
-uses to generate the circuit.
-
-The compiler determines that SUB and MUL in ``naive'' MPC Source (lines 9 and 10 in (b))
-can be fully vectorized into the SIMD SUB and MUL in optimized MPC Source (lines 9 and 10 in (c)). 
-Notation \python{p[I,J]} denotes a 2-dimensional array with fully vectorized dimensions.
-The computation of sum (line 11 in (b))
-is sequential across the $j$-dimension, but it is parallel across the $i$-dimension. 
-The loop in lines 12-16 in (c) illustrates; here \python{p[I,j]} refers to the $j$-th column in \python{p}. 
-Unfortunately, CMP and MUX remain sequential.
-
-\begin{comment}
-{\small
-\begin{verbatim}
-min_sum!1 = 10000
-min_index!1 = 0
-// S!0^ is same as S!0. C!0^ replicates C!0 N-times: 
-S!0^ = raise_dim(S!0, ((i * D!0) + j), (i:N!0,j:D!0))
-C!0^ = raise_dim(C!0, j, (i:N!0,j:D!0))
-
-sum!2 = [0,..,0]
-// Computes all differences and all products "at once"
-d!3[I,J] = SUB_SIMD(S!0^[I,J],C!0^[I,J])
-p!3[I,J] = MUL_SIMD(d!3[I,J],d!3[I,J])
-
-for j in range(0, D!0):
-  // sum!2[I], sum!3[I], sum!4[I] are vectors of size N  
-  // Computes N intermediate sums "at once"
-  sum!3[I] = PHI(sum!2[I], sum!4[I])       
-  sum!4[I] = ADD_SIMD(sum!3[I],p!3[I,j])
-
-min_index!3 = [0,1,...N!0-1]
-
-for i in range(0, N!0):
-  min_sum!2 = PHI(min_sum!1, min_sum!4) 
-  !1!2[i] = CMP(sum!3[i],min_sum!2)
-  min_sum!4 = MUX(!1!2[i], sum!3[i], min_sum!2)
-
-for i in range(0, N!0):
-  min_index!2 = PHI(min_index!1, min_index!4)  
-  min_index!4 = MUX(!1!2[i], min_index!3[i], min_index!2)
-!2!1 = (min_sum!2, min_index!2)   
-\end{verbatim}
-}
-\end{comment}
-
-In MPC backends, executing $n$ operations ``at once'' in a single SIMD operation costs a lot less than executing those $n$ operations one by one.
-This is particularly important when there is communication (i.e., in remote), since many 1-bit values are sent at once rather than sequentially. 
-%In MPC compilers a vectorized operation that computing M operations "at once" costs 
-%essentially the same ($\alpha$ or $\beta$) as an individual operation. 
-We elaborate on the cost model in the following section but for now consider that 
-each operation has a \emph{fixed} portion (does benefits from amortization) and 
-a \emph{variable} portion (does not benefit from amortization): $\alpha = \alpha_\mathit{fix} + \alpha_\mathit{var}$. 
-This gives rise to the following formula for amortized cost: $f(n) = \alpha_\mathit{fix} + n\alpha_\mathit{var}$, 
-as opposed to unamortized cost $g(n) = n\alpha_\mathit{fix} + n\alpha_\mathit{var}$. (We extend the same reasoning to 
-$\beta$-instructions.)
-
-Thus, the fixed cost of the vectorized program amounts to
-$2\alpha_\mathit{fix}$ + $D \beta_\mathit{fix} + N(3\alpha_\mathit{fix})$. (The variable cost is the same in both the vectorized and non-vectorized programs.) 
-The first term in the sum corresponds to the vectorized
-subtraction and multiplication, the second term corresponds to the FOR loop on $j$ and the third 
-one corresponds to the remaining FOR loops on $i$.
-Clearly, $2\alpha_\mathit{fix} + D\beta_\mathit{fix} + N(3\alpha_\mathit{fix}) << ND(2\alpha_\mathit{fix}+\beta_\mathit{fix}) + N3\alpha_\mathit{fix}$.
-Empirically, we observe 10x to 50x improvement over un-vectorized Biometric Matching in circuit generation time and setup time and 5x to 30x improvement in online time.
-Additionally, the un-vectorized version runs out of memory for $N=256$, while the vectorized one runs with the standard maximal input size $N=4,096$.
-\ana{Edit numbers with final experiments.}
-
-\section{Analytical Model}
-\label{sec:model}
-
-\ana{We need an intro to section here.}
-
-\subsection{Scheduling in MPC}
-\label{sec:mpc}
-
-%\ana{Here we need more on MPC before jumping to scheduling. MPC-source, basic assumptions about static loop bounds, etc.}
-
-For this treatment we make the following simplifying assumptions:
-
-\begin{enumerate}
-\item All statements in the program execute using the same protocol (sharing). That is, there is no share conversion. 
-%\ishaq{This is not an assumption, this is our setting. We work in the single protocol world.}
-\item There are two tiers of MPC instructions, local and remote. A local instruction (essentially just ADD) has cost $\beta$ 
-and a remote instruction (e.g., MUX, MUL, SHL, etc.) has cost $\alpha$, where $\alpha >> \beta$. We assume that all remote 
-instructions have the same cost.
-%\ishaq{Don't we need to distinguish between local (cheap) vs.
-interactive (expensive) instructions?
-I don't see how we could assume this.
--- Past suggestion from Vassilis: Ideally we should assume cost to be a convex function.}\ana{After discussion 1/12: replace single costs 1 with two levels of costs: $\alpha$ for non-local operations, e.g., MUL, and $\beta$ for local ops, e.g., ADD.}
-\item  In MPC frameworks, executing $n$ operations ``at once'' in a single SIMD operation costs a lot less than executing those $n$ operations one by one.
-Following Amdahl's law~\ana{citation!}, we write $\alpha = \frac{1}{s}p\alpha + (1-p)\alpha$, where $p$ is the fraction of execution time that benefits from amortization and $(1-p)$ 
-is the fraction that does not, and $s$ is the available resource. Thus, $n\alpha = \frac{n}{s}p\alpha + n(1-p)\alpha$. 
-For the purpose of the model we assume that $s$ is large enough and the term $\frac{n}{s}p\alpha$ amounts to a \emph{fixed cost} incurred regardless of
-whether $n$ is $10,000$ or just $1$. (This models the cost of sending a packet from party A to party B for example.) The bottom line is that amortized execution 
-of $n$ operations is $f(n) = \alpha_\mathit{fix} + n\alpha_{var}$ in contrast to unamortized execution $g(n) = n\alpha_\mathit{fix} + n\alpha_{var}$ and, of course,
-$\alpha_\mathit{fix} << n\alpha_\mathit{fix}$.
-%We assume infinite parallel capacity---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely $\alpha$ or $\beta$. 
-%This is a standard assumption in Cryptographic Parallel RAM. ABY presents empirical support for this assumption~\ana{Add citations. PRAM, ABY.}
-%\ishaq{Comment from Vassilis: We assume there are infinite parallel capacities, this is the standard assumption in PRAM (Cryptographic Parallel RAM).}\ana{PRAM assumption stays.
-We have to replace the unit cost of 1 with a suitable amortization function $f(n)$, which I don't think changes much in the cost modeling and analysis.}
-\item MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption,
-2 MUL instructions can be amortized in a single SIMD instruction that costs $\alpha_\mathit{fix} + 2\alpha_\mathit{var}$, however a MUL and a MUX instruction 
-still cost $2\alpha_\mathit{fix} + 2\alpha_\mathit{var}$ even when scheduled ``in parallel''. 
-%MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption,
-%2 MUL instructions scheduled in parallel benefit from amortization and cost $\alpha$, however a MUL and a MUX instructions scheduled
-%in parallel still cost $2\alpha$. 
-%\ishaq{We need to reword this assumption to something like "$K$ parallel MUL costs much less than $K$ (because they get amortized), but any mix of $K$  MUL and MUX still cost roughly $K$.
-Specifically, we should take away the low constants (2 and 1) because we know for low constants this is not true.} \ana{Again, core assumption that MUL and MUX don't benefit from amortization stays.
-We have to replace constant costs with functions, as in (3).}
-\end{enumerate}
-
-\input{prob_stmt_np_hardness.tex}
-
-\begin{figure*}
-  \includegraphics[width=1.0\linewidth]{figs_paper_SIMD/compiler_framework.png}
-  \caption{Compiler Framework. \ana{Change figure since we are not doing div-and-conquer.}}
-  \label{fig:compiler_framework}
-\end{figure*}
-
-\section{\bf Compiler Framework}
-\label{sec:preliminaries}
-
-%We assume arbitrarily nested loops in the MPC-source IR and read-only arrays.
-We assume that loops range from 0 to some constant N.
-Arrays are linearized (row-major order as in MOTION) and accesses are via functions of the induction variables of the enclosing loops.
-
-\figref{fig:compiler_framework} presents an overview of our compiler. This section describes the front-end phases of the compiler
-and Section~\secref{sec:vectorization} describes our vectorization algorithm. \ana{We'll need a section on MOTION code gen too.}
-%\ana{Need to state this more precisely.}
-
-\subsection{Syntax and Semantic Restrictions}
-\label{sec:syntax}
-
-Source syntax is essentially standard IMP syntax as shown below:
-%\begin{figure}[t]
-{\small
-\[
-\begin{array}{l@{~~~~~}l}
-  \begin{array}{l@{~}l@{~~~}l}
- e & ::= e \; \mathit{op} \; e \mid e \mid \x \mid  \texttt{const} \mid \texttt{A[$e$]} & \mathit{expression} \\
- % s &::= t \;\f & \mathit{field} \\
- %\md &::= t\;{\m}(t \;{\this}, \; t\;\x) \; \{\; \ol{t\;\y}\;\s;\;\return\;\y \; \} 
- %                 & \mathit{method} \\ 
-  s & ::= s;s \mid  & \mathit{sequence} \\
-  & \x = e \mid \texttt{A[$e$]} = e \mid & \mathit{assignment \; stmt} \\
-  & \texttt{for $i$ in range($I$)}: s \mid & \mathit{for \; stmt} \\
-  & \texttt{if $e$: $s$ else: $s$} & \mathit{if \; stmt} \\ 
-%   t &::= q \; {\sf C} & \mathit{qualified \; type}\\
-%   q &::= \high \mid \poly \mid \low & \mathit{FlowCFL \; qualifier}\\
-%    t' &::= q' \; \C & \mathit{field \; qualified \; type}\\
-%   q' &::= \tainted \mid \poly & \mathit{field \; qualifier}\\
-
-  \end{array}
-\end{array}
-\]
-}
-%\end{figure}
-
-The syntax allows for array accesses, arbitrarily nested loops, and if-then-else control flow. 
-Expressions are typed $q \; \tau$, where qualifier $q$ and base type $\tau$ are:
-
-{\small
-\[
-\begin{array}{l@{~~~~~}l}
-  \begin{array}{l@{~}l@{~~~}l}
- \tau & ::= \texttt{int}  \mid \texttt{bool} \mid \texttt{list[int]} \mid  \texttt{list[bool]} & \mathit{base \; types} \\
-  q & ::= \texttt{shared} \mid \texttt{plain} & \mathit{qualifiers} \\
-  \end{array}
-\end{array}
-\]
-}
-
-The type system is mostly standard, and in our experience, a sweet spot between readability and expressivity.
-The \texttt{shared} qualifier denotes shared values, i.e., ones shared among the parties and computed upon 
-under secure computation protocols; the \texttt{plain} qualifier denotes plaintext values. Shared lists denote 
-shared elements, but plaintext length. Subtyping is again as expected: $q_1 \; \tau <: q_2 \; \tau$ 
-iff $q_1 <: q_2$ where $\texttt{plain} <: \texttt{shared}$. Thus, we can convert a plaintext value into a shared one, 
-but not vice versa.
-
-Our compiler imposes certain semantic restrictions that it enforces throughout the various 
-phases of compilation. We note that in some cases, the restrictions 
-can be easily lifted and we plan to do so in future iterations of the work.
-
-\begin{enumerate}
-\item Loops are of the form $0 <= i < I$ and bounds are fixed at compile time. 
-It is a standard restriction in MPC that the bounds must be known at circuit-generation time.
-\item Arrays are one-dimensional. N-dimensional arrays are linearized and accessed 
-in row-major order and at this point the programmer is responsible for linearization
-and access. (This restriction can be easily lifted.)
-\item Array subscrpts are plaintext values as specified by the rule:
-\begin{semantics}
-    \ntyperule{Array Access}{
-%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
-  \Gamma \vdash e : \texttt{plain int} \quad \Gamma \vdash \texttt{A} : q \; \texttt{list}[\tau] \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
-}{
-  \Gamma \vdash \texttt{A[$e$]} : q \; \tau
-}
-\end{semantics}
-The subscript $e$ is a function of the indices of the enclosing loops. 
-For read access, the compiler allows an arbitrary such function. 
-However, it restricts write access to \emph{canonical writes}, i.e., \texttt{A[i,j,k]} = ...
-where \texttt{i}, \texttt{j} and \texttt{k} loop over the three dimensions of \texttt{A}.  
-Write access such as for example \texttt{A[i,j+2]} = ... is not allowed. 
-%\item \ana{Add rules for logical ops}
-%\item \ana{Add rules for arithmetic ops}
-\item The final restriction involves MUX as expressed by the rule: 
-\begin{semantics}
-    \ntyperule{MUX}{
-%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
-  \Gamma \vdash e_1 : q_1 \; \texttt{bool} \quad \Gamma \vdash e_2 : q_2 \; \tau \quad \Gamma \vdash e_2 : q_2 \; \tau \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
-}{
-  \Gamma \vdash \texttt{MUX($e_1$,$e_2$,$e_3$)} : q_1 \vee q_2 \vee q_3 \; \tau
-}
-\end{semantics}
-\end{enumerate}
-Importantly, the arguments of MUX are restricted to base types, as passing array arguments 
-is not valid MOTION MPC. \ana{Check this.} This causes a slight inconvenience as we could not write 
-\begin{pythonn}
-if e: A[i] = val
-\end{pythonn}
-Instead we had to write
-\begin{pythonn}
-if not(e): val = A[i]
-A[i] = val
-\end{pythonn}
-
-We note that while the programmer writes 
-annotations at the level of IMP Source (recall \ana{ref to overview section}), the 
-annotations propagate through the transformations and are checked at the 
-level of MPC Source.
-
-%\ana{Disadvantage of MUX}
-
-For the rest of this section we write $i,j,k$ to denote the loop nest: $i$ is the outermost loop, $j$, is immediately nested in $i$, and so on until $k$
-and we use $I,J,K$ to denote the corresponding upper bounds. We write $A[i,j,k]$ to denote canonical access 
-to an array element. In the program, canonical access is achieved via the standard row-major order formula: $(J*K)*i + K*j + k$.
-To simplify the presentation we describe our algorithms in terms of three-element tuples $i,j,k$, however, discussion easily generalizes to 
-arbitrarily large loop nests.
-
-\subsection{From IMP Source to SSA}
-
-Our compiler translates from Source to SSA as follows:
-
-\begin{enumerate}
-  \item
-        \textbf{Parsing}:
-        Use Python's \texttt{ast} module to parse the input source code to a Python AST.
-  \item
-        \textbf{Syntax checking}:
-        Ensure that the AST matches the restricted subset defined in Section~\secref{sec:syntax}.
-        This step outputs an instance of the \texttt{restricted\_ast.Function} class, which represents our restricted subset of the Python AST.
-  \item
-        \textbf{3-address CFG conversion}:
-        %\benjamin{TODO: Is this a good amount of detail?} \ana{This is good amount of detail for the Tech report. In the actual paper we'll shorten a lot and add pointers to the Tech report.}
-
-        Convert the restricted-syntax AST to a three-address control-flow graph.
-        To do this, first,
-        add an empty basic block to the CFG and mark it as current.
-        Next, for each statement in the restricted AST's function body,
-        process the statement.
-        Statements can either be for-loops, if-statements, or assignments (as in~\secref{sec:syntax}).
-        Rules for processing each kind of statement are given below:
-        \begin{enumerate}
-          \item
-                \textbf{For-loops}:
-                Create new basic blocks for
-                the loop condition
-                (the \emph{condition-block}),
-                the loop body
-                (the \emph{body-block}),
-                and the code after the loop
-                (the \emph{after-block}).
-                Insert a jump from the end of the current block to the condition-block.
-                Then, mark the condition-block as the current block.
-                Insert a for-instruction at the end of the current block with the loop counter variable and bounds from the AST.
-                Next, add an edge from the current block to the after-block labeled ``FALSE'' and
-                an edge from the current block to the
-                body-block labeled ``TRUE''.
-                Then, set the body-block to be the current block
-                and process all statements in the AST's loop body.
-                Finally, insert a jump to the condition-block and set the after-block as current.
-          \item
-                \textbf{If-statements}:
-                Create new basic blocks for
-                the ``then'' statements of the if-statement
-                (the \emph{then-block}),
-                the ``else'' statements of the if-statement
-                (the \emph{else-block}),
-                and the code after the if-statement
-                (the \emph{after-block}).
-                At the end of the current block,
-                insert a conditional jump to the then-block or else-block
-                depending on the if-statement condition in the AST.
-                Next,
-                mark the then-block as current,
-                process all then-statements,
-                and add a jump to the after-block.
-                Similarly,
-                mark the else-block as current,
-                process all else-statements,
-                and add a jump to the after-block.
-                Finally, set the after-block to be the current block,
-                and give it a \emph{merge condition} property equal to the condition of the if-statement.
-          \item
-                \textbf{Assignments}:
-                In the restricted-syntax AST,
-                the left-hand side of assignments
-                can be a variable or an array subscript.
-                If it is an array subscript, e.g., \texttt{A[i] = x},
-                change the statement to \texttt{A = Update(A, i, x)}.
-                If the statement is not already three-address code,
-                for each sub-expression in the right-hand side of the assignment,
-                insert an assignment to a temporary variable.
-        \end{enumerate}
-  \item
-        \textbf{SSA conversion}:
-        Convert the 3-address CFG to SSA with Cytron's algorithm.
-\end{enumerate}
-
-\subsection{From SSA to MPC Source} %: MUX Nodes and Pseudo $\phi$-nodes}
-
-Once the compiler converts the code to SSA,
-it transforms $\phi$-nodes that correspond to if-statements into MUX nodes.
-From the 3-address CFG conversion step,
-$\phi$-nodes corresponding to if-statements will be in a basic block
-with the merge condition property.
-For example, if $X_3 = \phi(X_1,X_2)$ is in a block with merge condition $C$,
-the compiler transforms it into $X_3 = \mathit{MUX}(C, X_1, X_2)$.
-Next, the compiler runs the dead code elimination algorithm from Cytron's SSA paper.
-
-Next, the control-flow graph is \emph{linearized} into MPC Source,
-which has loops but no if-then-else-statements.
-This means that both branches of all if-statements are executed,
-and the MUX nodes determine whether to use results from the then-block or from the else-block.
-The compiler linearizes the control-flow graph with a variation of breadth-first search.
-Blocks with the ``merge condition'' property are only considered
-the second time they are visited,
-since that will be after both branches of the if-statement are visited. 
-(The Python AST naturally gives rise to a translation where each conditional has exactly two targets, 
-and each ``merge condition'' block has exactly two incoming edges, a TRUE and a FALSE edge.
-Thus, each $\phi$-node has exactly two multiplexer arguments, which dovetails into MUX.
-This is in contrast with Cytron's algorithm which operates at the level of the CFG and allows for 
-$\phi$-nodes with multiple arguments.)
-Each time the compiler visits a block,
-it adds the block's statements to the MPC source.
-If the block ends in a for-instruction,
-the compiler recursively converts the body and code after the loop to MPC source
-and adds the for-loop and code after the loop to the main MPC source.
-If the block does not end in a for-instruction,
-the compiler recursively converts all successor branches to MPC source and
-appends these to the main MPC source.
-
-Now, the remaining $\phi$-nodes in MPC source are the loop header nodes. We call these nodes \emph{pseudo} $\phi$-nodes 
-and we write \python{PHI} in MPC Source. A pseudo $\phi$-node $X_1 = $\python{PHI}$(X_0,X_2)$ in a loop header is evaluated 
-during circuit generation. If it is the 0-th iteration, then the $\phi$-node evaluates to $X_0$, otherwise, it evaluates to $X_2$.
-
-%\ana{We need to double check that the following captures all cases of def-use edges.}
-
-\subsection{Dependence Analysis}
-
-The next steps construct the dependence graph and \ana{Add an intro here.}
-
-\subsubsection{Def-use Edges}
-
-The dependence graph has the following def-use edges:
-
-\begin{itemize}
-\item same-level edge $X\rightarrow Y$ where $X$ and $Y$ are in the same loop nest, say $i,j,k$.
-E.g., the def-use edge 9 to 10 in the Biometric MPC Source in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source} is a same-level edge.
-A same-level edge can be a back-edge in which case a $\phi$ node is the target of the edge.
-E.g., 15 to 4 in Biometric is a same-level back-edge.
-\item outer-to-inner $X\rightarrow Y$ where $X$ is in an outer loop nest, say $i$, and $Y$ is in an inner one, say $i,j,k$. E.g., 1 to 4 in Biometric forms is an outer-to-inner edge.
-%A $\phi$-node can be a source or a target of an outer-to-inner forward edge.
-\item inner-to-outer $X\rightarrow Y$ where $X$ is a $phi$-node in an inner loop nest, $i,j,k$, and $Y$ is in the enclosing loop nest $i,j$.
-E.g., the def-use from 8 to 12 gives rise to an inner-to-outer edge.
-%E.g. $\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$ is an inner-to-outer edge. 
-An inner-to-outer edge can be a back-edge as well, in which case both $X$ and $Y$ are phi-nodes with the source $X$ in a loop nested into $Y$'s loop (not necessarily immediately).
-
-%Note that the source is \emph{always} a $\phi$-node of the $i,j,k$ loop and the target is in the immediately enclosing loop.
-The interpretation of this edge is that the use node $Y$ uses the definition made in the last iteration of the inner loop.
-%\ben{The current representation of pseudo $\phi$-nodes shows them attached to the loop header (i.e. in the loop).  We may want to clarify that they are also evaluated at the loop's termination.}
-%\item same-level back-edge $X\rightarrow Y$.
-$Y$ is a phi-node in the header of the loop and $X$ is a definition of the variable in the loop body.
-E.g., $\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{sum}_1,\texttt{min}_1)$ to $\texttt{min}_0 = \phi(\texttt{min}_1, 10000)$ in Biometric is a same-level back-edge.
-%\item inner-to-outer back-edge $X\rightarrow Y$: $X$ and $Y$ are both $\phi$-nodes for some variable. The source $X$ is in a loop nested into $Y$'s loop (not necessarily immediately).
-
-\item mixed forward edge $X\rightarrow Y$.
-$X$ is in some loop $i,j,k$ and $Y$ is in a loop nested into $i,j,k'$.
-We transform mixed forward edges as follows.
-Let $\x$ be the variable defined at $X$.
-We add a variable and assignment $\x' = \x$ immediately after the $i,j,k$ loop.
-Then we replace the use of $\x$ at $Y$ with $\x'$.
-This transforms a mixed forward edge into an "inner-to-outer" forward edge followed by an outer-to-inner forward edge.
-Thus, Basic Vectorization handles one of "same-level", "inner-to-outer", or "outer-to-inner" def-use edges.
- \end{itemize}
-
-\subsubsection{Closures}
-
-We define $\mathit{closure}(n)$ where $n$ is a phi-node. Intuitively, it computes the set of nodes (i.e., statements) that form a dependence cycle with $n$. The closure of $n$ is defined as follows:
-\begin{itemize}
-\item $n$ is in $\mathit{closure}(n)$
-\item $X$ is in $\mathit{closure}(n)$ if there is a same-level path from $n$ to $X$, and $X \rightarrow n$ is a same-level back-edge.
-\item $Y$ is in $\mathit{closure}(n)$ if there is a same-level path from $n$ to $Y$ and there is a same-level path from $Y$ to some $X$ in $\mathit{closure}(n)$.
-\end{itemize}
-
-\subsection{Taint Analysis}
-
-We require that all inputs are marked as either shared or plaintext, however, we do not require any other qualifier annotations in the program.
-Then we infer if intermediate variables and expressions are shared through taint analysis with "taintedness" referring to the shared attribute.
-Specifically, the compiler follows the following rules, which are standard in positive-negative qualifier systems (here \texttt{shared} is the positive qualifier and \texttt{plain} is the negative one):
-\begin{enumerate}
-\item Loop counters and array upper bounds are always \texttt{plain}.
-\item If any variable on the right-hand side of an assignment is shared, then the assigned variable is \texttt{shared}.
-%\item If all variables on the right-hand side of an assignment are plaintext, then the assigned variable is plaintext \ana{Will be covered by the last rule.}
-\item Any variables that cannot be determined as shared via the above rules are \texttt{plain}.
-\end{enumerate}
-
-%The first two rules are standard for taint analysis, and the third rule follows from the MPC problem statement. \ben{Is the explanation for the third rule correct?}  \ana{Yes, that is exactly it!}
-The final rule is needed in examples like the following. In the below snippet \texttt{sum!2} and \texttt{sum!3} form a dependency cycle and there is no \texttt{shared} value that flows to either one. 
-They are inferred as plaintext. 
-%cannot be marked as plaintext through simple taint analysis:
-{\small
-\begin{verbatim}
-plaintext_array = [0, 1, 2, ...]
-sum!1 = 0
-for i in range(0, N):
-    sum!2 = PHI(sum!1, sum!3)
-    sum!3 = sum!2 + plaintext_array[i]
-\end{verbatim}
-}
-%\ben{I think the above example is unnecessary and could be replaced by an explanation of how "untainted" variables are implicitly plaintext, but I couldn't think of a way to phrase that.}\ana{I left the example, we can leave it in the TR, remove from paper.}
-
-When converting to MOTION code, any plaintext value used in the right-hand side of a shared assignment is converted to a shared value for that expression. 
-%\ben{Is this necessary to include?} \ana{Yes, that's important.}
-
-Importantly, the taint analysis works on MPC Source, which lacks if-then-else control flow. 
-This significantly simplifies the treatment as there is no need to handle conditionals and implicit flow.
-
-\subsection{From (Optimized) MPC Source to MOTION}
-\label{sec:MOTION}
-\ana{Ben, add the MOTION section here.}
-
-\section{Vectorization}
-\label{sec:vectorization}
-
-An important component of our algorithm is the scalar expansion to the corresponding loop dimensionality. 
-For example, \texttt{d = S[i * D + j] - C[j]} equiv. to \texttt{d = S[i,j] - C[j]}, which gave rise to \texttt{N*D} subtraction operations in the sequential schedule, 
-is lifted. The argument arrays \texttt{S} and \texttt{C} are lifted and the scaler \texttt{d} is lifted: \texttt{d[i,j] = S[i,j] - C[i,j]}.
-The algorithm then detects that the statement can be vectorized.
-
-There are three kinds of arrays (for now all kept internally as one-dimensional arrays, but that's under discussion).
-\begin{itemize}
-
-\item Scalars: These are scalar variables we lift into arrays for the purposes of vectorization. 
-For those, all writes are canonical writes and all reads are canonical reads. We will apply raise dimension when 
-a scalar is used in an inner loop (e.g., \texttt{sum0} in line 6 of the MPC source code will be raised to a 1-dimensional 
-array since \texttt{sum0} is used in the inner $j$-loop). Drop dimension applies as well; this happens when a scalar
-written in an inner loop is used outside of the loop (e.g., \texttt{sum0} for which the lifted inner loop computes 
-\texttt{D} values, but the outer loop only needs the last one.)
-
-\item Read-only input arrays: Read-only inputs. There are NO writes, while we may have non-canonical reads, $f(i,j,k)$. 
-Phase 1 of Basic vectorization will add raise dimension operation at the beginning of the function to lift these arrays, 
-and raise dimension may \emph{reshape} arrays. If there are multiple "views" of the input array, there would be multiple raise dimension statements to create
-each one of these views. The invariant is that at reads in loops, the reads of "views" of the original input array are canonical.
-Only raise dimension applies to these arrays, and only in the beginning of the program. For example, 1-dimensional array 
-\texttt{C} is lifted into 2-dimensional array \texttt{N x D} by copying the row \texttt{N} times.
-
- \item Read-write output arrays: Writes are canonical (by restriction) but reads can be non-canonical. 
- Dependence analysis limits vectorization when non-canonical read access refers to array writes in
- previous iterations, thus creating loop-carried dependences. We may apply both raise and drop dimension, however, 
- they respect the fixed dimensionality of the output array. The array cannot be raised to a dimension lower than
- its canonical (fixed) dimensionality and it cannot be dropped lower. In addition, non-canonical reads may 
- require lifting (i.e., reshaping) of the array after the most recent write, rather than in the beginning of the program, 
- in order to reduce a non-canonical read to a canonical one.
-
-\end{itemize}
-
-In the sections below we detail the $\mathit{raise\_dim}$ (raise dimension) and $\mathit{drop\_dim}$ (drop dimension) operations,
-followed by our vectorization algorithm.
-
-%\ana{Add back-edges into Phase 1.
-A back-edge from a non-phi-node in loop $i$ to a phi-node in loop $i$'s header is a same-level edge.
-The only difference with the handling of normal same-level forward def-use is that the operand index will become $i-1$.
-A back edge from a phi-node in loop $j$ to a phi-node in loop $i$, where $j$ is nested in $i$ is an inner-to-outer edge and will require dropping dimension.
-We can show these are the only kinds of back-edges that may occur.}
-
-%\subsection{Algorithm (Still work in progress.)}
-
-\subsection{Raise Dimension and Drop Dimension}
-
-There are two conceptual versions of $\mathit{raise\_dim}$. One applies on read-only input arrays and reshapes those arrays 
-when necessary to ensure canonical read access in the corresponding loop. The signature of $\mathit{raise\_dim}$
-is as follows. It takes the original array \texttt{C}, the access pattern function $f(i,j,k)$ in loop nest $i,j,k$ and the loop bounds
-\texttt{((i:I),(j:J),(k:K))}: 
-\[ \mathit{raise\_dim}(\texttt{A}, f(i,j,k), \texttt{((i:I),(j:J),(k:K))}) \]
-It produces a new 3-dimensional array \texttt{A'} by iterating over $i,j,k$ and setting each element of  \texttt{A'} as follows:
-\[ \texttt{A'[i,j,k] =  A[f(i,j,k)]} \]
-The end result is that uses of \texttt{[A[f(i,jk)]} in loop nest $i,j,k$ are replaced with canonical read-accesses to \texttt{A'[i,j,k]}
-that can be vectorized. In the running Biometric example, \texttt{C'} = $\mathit{raise\_dim}$\texttt{(C, j, (i:N,j:D))} lifts the 
-1-dimensional array \texttt{C} into a 2-dimensional array. The $i,j$ loop now accesses \texttt{C'} in the canonical way, \texttt{C'[i,j]}. 
-Similarly, \texttt{S'} = $\mathit{raise\_dim}$\texttt{(S, i*D+j, (i:N,j:D))} tries to lift \texttt{S}, but the operation turns into a no-op 
-because \texttt{S} is already a 2-dimensional array and the read access is canonical.
-
-The other version of $\mathit{raise\_dim}$ applies on scalars and read-write arrays. It lifts a lower-dimension array into a 
-higher-dimension for access in a nested loop. Here \texttt{A} is an $i$ array and raise dimension adds two additional dimensions:
-\[ \mathit{raise\_dim}(\texttt{A}, \texttt{(j:J,k:K)}) \]
-This version is reduced to the above version by adding the access pattern function, which is just $i$: 
-\[ \mathit{raise\_dim}(\texttt{A}, i, \texttt{(j:J,k:K)}) \]
-
-The corresponding $\mathit{drop\_dim}$ is carried out when an array written in an inner loop is used in an enclosing loop. 
-It takes a higher dimensional array, say $i,j,k$ and removes trailing dimensions, say $j,k$:
-\[ \mathit{drop\_dim}(\texttt{A}, \texttt{(j:J,k:K)}) \]
-It iterates over $i$ and takes the result at the maximal index of $j$ and $k$, i.e., the result at the last iterations of $j$ and $k$:
-\[ \texttt{A'[i] = A[i,J-1,K-1]} \]
-
-\begin{comment}
-We define the $\mathit{raise\_dim}$ (raise dimensions) and $\mathit{drop\_dim}$ (drop dimension) functions.
-Raise dimension "lifts" a lower-dimension array (\ana{The right term here is tensor, I am pretty sure!}) into a higher dimension one.
-This is necessary when a lower-dimensional array is used in a higher dimensional loop and, essentially, is just copying of values.
-For example, Biometric contains the statement $d = \mathit{SUB}(S[i, j], C[j])$, in the $j$-loop which is nested into the $i$-loop.
-Here $C$ is a one-dimensional array, however, to vectorize across both loops it is necessary to turn it into a two-dimensional array: $C[i,j]$ becomes $[C[0],C[1],...C[J],C[0],C[1],...C[J], ...
-C[0],C[1],...C[J]]$, which turns the row into a matrix of $I$ identical rows.
-(We use capital letters to denote the upper bounds of loops, e.g., $J$ is the upper bound of the $j$-loop and $I$ is the upper bound of the $i$-loop.)
-
-$\mathit{raise\_dim}(A[i_1,...,i_k],i_j)$ is defined as follows.
-It results in a new {k+1}-dimensional array $A'$ where for every $0 \le i_{k} < I_{k+1}$, $A'[i_1, \ldots, i_j, \ldots, i_k] = A[i_1, \ldots, i_k]$.
-Adding $n$ dimensions is trivially extended as a composition of $n$ $\mathit{raise\_dim}$ that each adds a single dimension.
-
-As expected, drop dimension turns a higher-dimensional array into a lower-dimensional one.
-The key use case is an inner-to-outer def-use edge.
-The code may define a variable, e.g., $\x$ in an inner loop, say $j$, then use this variable in an enclosing loop, say $i$.
-Our algorithm may vectorize the computation of $\x$ in the $j$-loop thus producing a vector $\x[j]$ where $\x[1]$ is the value of \x after the 1st iteration and so on.
-Drop dimension states that the outer loop will use the value of the variable at the last iteration.
-
-$\mathit{drop\_dim}(A[i_1,\ldots, i_j, \ldots,i_k], i_j)$ produces a k-dimensional array $A'$ where $A'[i_1,\ldots,i_k] = A'[i_1,\ldots, I_j-1, \ldots,i_k]$.
-In our analysis, dimensions are always dropped at the end, and again, one can define dropping $n$ dimensions as a composition of $n$ $\mathit{drop\_dim}$.
-
-\ben{I'm not sure if I wrote up the above paragraphs correctly -- the idea is that we're just copying values to extend over dimension $j$ or only retain its last elements.
-The current writeup implies that $1 < j < k$ though, which isn't always true.
-Python implementations of $\mathit{raise\_dim}()$ and $\mathit{drop\_dim}()$ can be found here: \url{https://github.com/milana2/ParallelizationForMPC/blob/master/compiler/compiler/motion_backend/reference_implementations.py}}
-
-\ana{This section is still informal. Needs work to make more precise.}
-\end{comment}
-
-\subsection{Basic Vectorization}
-
-\begin{algorithmic}
-
-\STATE \COMMENT { Phase 1: Raise dimension of scalar variables to corresponding loop nest. We can traverse stmts linearly in MPC-source. }
-
-\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i,j,k$}
-%\FOR {each $Y_n$}
-\FOR {each argument $Y_n$} %def-use edge $\mathit{stmt}' (\mbox{def of } Y) \rightarrow \mathit{stmt} (\mbox{def of } X)$}
-\STATE {case def-use edge $\mathit{stmt}' (\mbox{def of } Y_n) \rightarrow \mathit{stmt} (\mbox{def of } X)$ of} \\
-\STATE \hspace{0.25cm} {same-level: $Y'_n$ is $Y_n$} \\
-\STATE \hspace{0.25cm} {outer-to-inner: add $Y'_n[i,j,k] = \mathit{raise\_dim}(Y_n)$ at $\mathit{stmt'}$}
-\STATE \hspace{0.25cm} {(more precisely, right after $\mathit{stmt'}$)}
-\STATE \hspace{0.25cm} {inner-to-outer: add $Y'_n[i,j,k] = \mathit{drop\_dim}(Y_n)$ at $\mathit{stmt}$}
-\STATE \hspace{0.25cm} {(more precisely, in loop of $\mathit{stmt}$ right after loop of $\mathit{stmt'}$)}
-\ENDFOR
-
-\COMMENT { Optimistically vectorize all. $I$ means vectorized dimension. }
-%\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i,j,k$}
-\STATE {change to $X[I,J,K] = \mathit{Op}(Y'_1[I,J,K],Y'_2[I,J,K])$ }
-\ENDFOR
-
-\COMMENT { Phase 2: Recreating FOR loops for cycles; vectorizable statements hoisted up. }
-
-\FOR {each dimension $d$ from highest to 0}
-\FOR {each $\phi$-node $n$ in loop $i_1,...,i_d$}
-\STATE {compute $\mathit{closure}(n)$}
-\ENDFOR
-
-\COMMENT {{\color{red} $cl_1$ and $cl_2$ intersect if they have common statement or update same array; "intersect" definition can be expanded}}
-
-\WHILE {there are closure $cl_1$ and $cl_2$ that intersect}
-\STATE {merge $cl_1$ and $cl_2$}
-\ENDWHILE
-\FOR {each closure $cl$ (after merge)}
-\STATE {create FOR $i_d = 0; ...$ loop}
-\STATE {add $\phi$-nodes in $cl$ to header block}
-\STATE {{\color{red}{add target-less $\phi$-node for A if $cl$ updates array A}}}
-\STATE {add statements in $cl$ to loop body in some order of dependences}
-\STATE \COMMENT { Dimension is not vectorizable: }
-\STATE {change ${I_d}$ to $i_d$ in all statements in loop}
-\STATE {treat FOR loop as monolith node: some def-use edges become same-level.}
-\ENDFOR
-{
-\color{red}
-\FOR {each target-less $\phi$-node $A_1 = \phi(A_0,A_k)$}
-\STATE {in vectorizable stmts, replace use of $A_1$ with $A_0$}
-\STATE {discard $\phi$-node if not used in any $cl$, replacing $A_1$ with $A_0$ or $A_k$ appropriately}
-\ENDFOR
-}
-\ENDFOR
-
-\COMMENT { Phase 3: Remove unnecessary dimensionality. }
-
-\COMMENT { A dimension $i$ is dead on exit from stmt $X[...i...] = ...$ if all def-uses with targets outside of the enclosing 
-FOR $i=0 ...$ MOTION loop end at target (use) $X' = \mathit{drop\_dim}(X, i)$. } 
-\FOR {each stmt and dimension $X[...i...] = ...$}
-\STATE {if $i$ is a dead dimension on exit from stmt $X[...i...] = ...$, remove $i$ from $X$ (all defs and uses)} 
-%\COMMENT{ A stmt v[...i] is a "dead dimension i" if all edges out of FOR i = 0; ... loop end at drop\_dim(v, i)}
-\ENDFOR
-
-\COMMENT { Now clean up drop\_dim and raise\_dim }
-\FOR {each $X' =  \mathit{drop\_dim(X,i)}$ }
-\STATE {replace with $X' = X$ if $i$ is dead in $X$.}
-\ENDFOR
-\STATE {do (1) (extended) constant propagation, (2) copy propagation and (3) dead code elimination to get rid of redundant variables and raise and drop dimension statements}
-
-\COMMENT { Phase 4: }
-\STATE {add SIMD for simdified dimensions}
-
-\end{algorithmic}
-
-\input{biometric_vectorization_example.tex}
-
-\begin{comment}
-\subsection{Example: Biometric}
-
-\subsubsection{MPC-source}
-
-Below is the MPC-source resulting from Benjamin's analysis. It gives rise to a corresponding dependence graph. (Not shown.)
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0 = \phi(-1, \texttt{index}_1)$}
-\STATE {$\texttt{min}_0 = \phi(10000,\texttt{min}_1)$}
-
-\STATE \COMMENT {Begin of inner loop $j$.}
-\STATE {$\texttt{sum}_0 = \phi(0, \texttt{sum}_1)$}
-\STATE {$\texttt{d} = \mathit{SUB}(\texttt{S}[i,j], \texttt{C}[j])$}
-\STATE {$\texttt{p} = \mathit{MUL}(\texttt{d},\texttt{d})$}
-\STATE {$\texttt{sum}_1 = \texttt{ADD}(\texttt{sum}_0,\texttt{p})$}
-\STATE \COMMENT {End of inner loop $j$.}
-\STATE {$\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$}
-\STATE {$\texttt{index}_1 = \mathit{MUX}(\texttt{c},\texttt{index}_0,i)$}
-\STATE {$\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{min}_0,\texttt{sum}_0)$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-\subsubsection{Phase 1 of Basic Vectorization:}
-
-The transformation preserves the dependence edges.
-It raises the dimensions of scalars and optimistically vectorizes all operations.
-The next phase discovers loop-carried dependences and removes affected vectorization.
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$}
-\STATE {$\texttt{min}_0[i] = \phi(10000, \texttt{min}_1[i-1])$}
-
-\STATE \COMMENT {Begin of inner loop $j$.}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...], \texttt{sum}_1[i,j-1])$}
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{sum}_1[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE \COMMENT {End of inner loop $j$.}
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {$\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$}
-\STATE {$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$}
-\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-\ishaq{At least for input variables: when raising dimensions, we should keep some meta data around so we can later write code that could use 1 share if it knows the other N shares are just duplicates of it.}
-
-\subsubsection{Phase 2}
-
-This phase analyzes statements from the innermost loop to the outermost. The key point is to discover loop-carried dependencies and re-introduce loops whenever dependencies make this necessary.
-
-Starting at the inner phi-node $\texttt{sum}_0[i,j] = \phi([0,0,...],\texttt{sum}_1[i,j-1])$, the algorithm first computes its closure.
-The closure amounts to the phi-node itself and $\texttt{sum}_1[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$, accounting for the loop-carried dependency of the computation of \texttt{sum}.
-The algorithm replaces this closure with a FOR loop on $j$ removing vectoriaztion on $j$.
-Note that the SUB and MUL computations remain outside of the loop as they do not depend on phi-nodes that are part of cycles.
-The algorithm adds same-level edges, one from $\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = ...$ to the monolithic FOR-loop node, and one from the FOR-loop node to $\texttt{c}[\stackrel{\rightarrow}{i}] = ...$.
-
-\ishaq{For our radar: Such dependencies will not involve phi nodes only. e.g. $a[i] = a[i] + a[i-1]$}
-
-\begin{algorithmic}
-
-\STATE \COMMENT {Begin of outer loop $i$.}
-\STATE {$\texttt{index}_0[i] = \phi(-1, \texttt{index}_1[i-1])$}
-\STATE {$\texttt{min}_0[i] = \phi(10000, \texttt{min}_1[i-1])$}
-
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-
-\STATE \COMMENT {Begin of inner loop $j$. Loop is now much "shorter"!}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...], \texttt{sum}_1[i,j-1])$}
-\STATE {FOR j=0; j<D; j++} \COMMENT{ Will turn into a MOTION loop.}
-\STATE \hspace{0.25cm}{$\texttt{sum}_1[\stackrel{\rightarrow}{i},{j}] = \mathit{ADD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},{j}],\texttt{p}[\stackrel{\rightarrow}{i},{j}])$}
-\STATE \COMMENT {End of inner loop $j$.}
-
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {$\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$}
-\STATE {$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$}
-\STATE {$\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])$}
-\STATE \COMMENT {End of outer loop $i$.}
-
-\end{algorithmic}
-
-Next, analysis moves to outer dimension $i$. There are two phi-nodes, $\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1])$ and $\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$.
-The closure of the first is
-\[
-\begin{array}{l}
-\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1]) \\
-\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
-\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_0[\stackrel{\rightarrow}{i}])
-\end{array}
-\]
-and the closure of the second one is
-\[
-\begin{array}{l}
-\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1]), \\
-%\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
-\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])
-\end{array}
-\]
-
-Since the two closures \emph{do not} intersect, we have two distinct FOR-loops on $i$.
-The first FOR loop (on \texttt{min}) is scheduled first, because of the dependence edge from $\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_0[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}])$ to
-$\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N])$.
-We can now rewrite those statements, getting rid of the vectorization on $i$ in the two FOR loops.
-Notably, the SUB and MUL computations are fully vectorizable and the ADD computation is vectorizable across the $i$-dimension.
-
-\begin{algorithmic}
-
-\STATE
-\STATE {$\texttt{C}'[i,j] = \mathit{raise\_dim}(\texttt{C}[j],i)$}
-\STATE {$\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{SUB\_SIMD}(\texttt{S}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}], \texttt{C}'[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$}
-\STATE {$\texttt{p}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}] = \mathit{MUL\_SIMD}(\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}],\texttt{d}[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j}])$} \\
-\STATE {}
-\STATE {$\texttt{sum}_0[i,j] = \phi([0,0,...],\texttt{sum}_1[i,j-1])$}
-\STATE {FOR j=0; j<D; j++} \COMMENT{ Turns into a MOTION loop, was $j$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{sum}_1[\stackrel{\rightarrow}{i},{j}] = \mathit{ADD\_SIMD}(\texttt{sum}_0[\stackrel{\rightarrow}{i},{j}],\texttt{p}[\stackrel{\rightarrow}{i},{j}])$}
-\STATE {}
-\STATE {$\texttt{sum}'_0[i] = \mathit{drop\_dim}(\texttt{sum}_0[i,j],j)$}
-\STATE {}
-\STATE {$\texttt{min}_0[i] = \phi(10000,\texttt{min}_1[i-1])$}
-\STATE {FOR i=0; i<N; i++} \COMMENT{ Turns into a MOTION loop, was $i$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{c}[{i}] = \mathit{CMP}(\texttt{sum}'_0[{i}],\texttt{min}_0[{i}])$}
-\STATE \hspace{0.25cm}{$\texttt{min}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{min}_0[{i}],\texttt{sum}'_1[{i}])$}
-\STATE {}
-\STATE {$\texttt{index}_0[i] = \phi(-1,\texttt{index}_1[i-1])$}
-\STATE {FOR i=0; i<N; i++} \COMMENT{ Turns into a MOTION loop, was $i$-loop.}
-\STATE \hspace{0.25cm}{$\texttt{index}_1[{i}] = \mathit{MUX}(\texttt{c}[{i}],\texttt{index}_0[{i}],i)$}
-
-\ben{The above code implicitly stores the final values of min\_0 and index\_0 in their last indices, but there is no primitive like drop\_dim to make it clear.
-This will have to be added in the SSA code generation (though it might not be necessary to include in the paper?).
-It also is pretty wishy-washy about the size and layout of each vector (again more of an implementation issue), but maybe we should have a different syntax for array accesses for read/write (e.g.
-the ADD\_SIMD steps) vs the array definitions (e.g.
-the SUB\_SIMD and raise/drop\_dim statements)}
-
-\end{algorithmic}
-\end{comment}
-
-\subsection{Correctness Argument}
-
-We build a correctness argument that loosely follows the theory of Abstract Interpretation. 
-We define the syntax of MPC Source programs. The domain of MPC Source programs
-expressible in the syntax (with certain semantic restrictions) is the abstract domain $A$. 
-We then define the \emph{linearization} of an MPC Source program as an interpretation 
-over the syntax. The linearization, which is a \emph{schedule}, is the concrete domain $C$. 
-Since we reason over def-use graphs in $A$ we define a partial order relation over elements of $A$ 
-in terms of def-use relations. We define a partial order over elements of $C$ as well,
-in terms of def-use relations in the concrete domain $C$. We prove two theorems
-that state (informally) that the schedule corresponding to the original program computes the same 
-result as the schedule corresponding to the vectorized program.
-
-\begin{figure}%[!b]
-%\small
-\[
-\begin{array}{l@{~~~~~}l}
-  \begin{array}{l@{~}l@{~~~}l}
-  s &::= s;s & sequence \\
-     & \mid \x[i,J,k] = \y[i,J,k] \; \mathit{op\_SIMD} \; \z[i,J,k] & operation \\ 
-     & \mid \x[i,J,k]  = \texttt{PHI}(\x_1[i,J,k],\x_2[i,J,k\!-\!1]) & \mathit{phi} \; \mathit{node} \\
-     & \mid \x[i,J,k] = \mathit{raise\_dim}(\x'[i],(J\!:\!J,k\!:\!K)) & \mathit{raise} \; \mathit{dimension(s)} \\
-     & \mid \x[i,J] = \mathit{drop\_dim}(\x'[i,J,k],k)  & \mathit{drop} \; \mathit{dimension(s)} \\
-     & \mid \x = \y & \mathit{propagation} \\
-     & \mid \texttt{FOR} \; 0 \le i < I: s  & \mathit{loop} \\
-  \end{array}    \\
-\end{array}
-\]
-\vspace{-3mm}
-\caption{MPC Source Syntax}\label{fig:MPC_Source_syntax}
-\vspace{-5mm}
-\end{figure}
-
-\paragraph{MPC Source Syntax}
-
-\figref{fig:MPC_Source_syntax} defines the syntax for our intermediate representation, MPC Source. 
-There are semantics restrictions over the syntax as well: a variable $\x[i,j,k]$ is a 3-dimensional array 
-$(i:I,j:J,k:K)$ and also, a statement $\x[i,J,k] = ...$ is enclosed in loops over $i$ and $k$ as shown below. 
-Thus, $i$ and $k$ are in scope.
-\begin{verbatim}
-FOR 0 <= i < I:
-    ...
-    FOR 0 <= k < K:
-        x[i,J,k] = ...
-\end{verbatim}
-
-Statements $\mathit{operation, phi, raise \; dimension(s), drop \; dimension(s)}$ are \emph{base} statements, 
-and $\mathit{sequence, loop}$ are \emph{compound statements}.
-
-\paragraph{Linearization}
-
-Linearization is the concretization operation, which, as we mentioned earlier computes a schedule.
-The concretization function $\gamma: A \rightarrow C$ is defined as an interpretation of
-MPC Source syntax, as it is standard. The concretization of each one 
-of the base statements is as follows: 
-\[\begin{array}{l}
-\gamma(\x[i,J,k] = \mathit{op\_SIMD}(\y[i,J,k], \z[i,J,k])) = \\ 
-~~\x[i,0,k] = \y[i,0,k] \; \mathit{op} \; \z[i,0,k] \;\; || \\
-~~~~\x[i,1,k] = \y[i,1,k] \; \mathit{op} \; \z[i,1,k] \;\; || \; ...\; || \\
-~~~~~~\x[i,I-1,k] = \y[i,I-1,k] \; \mathit{op} \; \z[i,I-1,k]
-\end{array}\]
-meaning that the vectorized dimension(s) are expanded into parallel statements. || introduces SIMD (parallel) execution.
-
-The concretization of the FOR statement is as follows:
-\[ \gamma(\texttt{FOR} \; 0 \le i < I: s) = \gamma(s)[0/i] \; ; \; \gamma(s)[1/i] \; ; \; ... \gamma(s)[I-1/i] \]
-$\gamma$ simply unrolls the loop substituting $i$ with 0, 1, etc. Here ; denotes sequential execution.
-
-\paragraph{Partial Orders}
-
-For each MPC Source program $a$ we compute the def-use edges in the standard way: if base statement $s1 \in a$ defines variable \x, e.g., $\x[i,j,k] = ...$, 
-and base statement $s2 \in a$ uses \x, e.g., $... = ... \x[i,j,k]$ and there is a path in the trivial CFG from $s1$ to $s2$, then there is a def-use edge from
- $s1$ to $s2$. We extend the dimensionality of a statement into $s1[i,j,k]$ where $s1[i,j,k]$ inherits the dimensionality of the left-hand-side of the assignment.
-
- Let $a_0, a_1$ be two MPC Source programs in $A$. 
- Two base statements, $s_0 \in a_0$ and $s_1 \in a_1$ are \emph{same}, written $s_0 \equiv s_1$ if they are of the same operation and they operate on 
- the same variables: same variable name and same dimensionality. Recall that dimensions in MPC Source are either iterative, lower case, or vectorized, upper case. 
- Two statements are same even if one operates on an iterative dimension and the other one operates on a vectorized one, e.g., $s_0[i,j,k] \equiv s_1[I,j,K]$.
-
- \begin{definition} Let $a_0, a_1 \in A$. We say that $a_0 \le a_1$ iff for every def-use edge $s1 \rightarrow s2$ in $a_0$ there is an edge
- $s1' \rightarrow s2'$ where $s1 \equiv s1'$, $s2 \equiv s2'$ and the two edges of either both forward or both backward.
- \end{definition}
-
- The def-use edges in the concrete schedule are as expected. There is a def use edge from statement $s1$ that defines $\x[\underline{i},\underline{j},\underline{k}]$
- to statement $s2$ that uses $\x[\underline{i},\underline{j},\underline{k}]$ if $s1$ is scheduled ahead of $s2$ in the linear schedule. We note that the underlined
- indices, e.g., $\underline{i}$, refer to fixed values, not iterative or vectorized dimensions since in the concrete schedule all induction variables are expanded. 
- E.g., there is a def-use edge from the statement that defines $\x[0,1,2]$ and a statement that uses $\x[0,1,2]$.
-
-\paragraph{Theorems}
-
-\begin{theorem} $a_0 \le a_1 \Rightarrow \gamma(a_0) \subseteq \gamma(a_1)$.
-\end{theorem}
-
-\begin{theorem} Let $a_0$ be the iterative MPC Source and let $a_1$ be the vectorized MPC Source computed by the vectorization algorithm. 
-We have that $a_0 \le a_1$. 
-\end{theorem}
-
-\ana{Write the proof sketch and final argument, etc.}
-
-\subsection{Extension of Basic Vectorization with Array Writes}
-
-\subsubsection{Removal of Infeasible Edges}
-
-Array writes limit vectorization as they sometimes introduce infeasible loop-carried dependencies. Consider the following example: \ana{Have to add citation to Aiken's paper}
-
-{\small
-\begin{verbatim}
-for i in range(N):
-  A[i] = B[i] + 10;
-  B[i] = A[i] * D[i-1];
-  C[i] = A[i] * D[i-1];
-  D[i] = B[i] * C[i];
-\end{verbatim}
-}
-
-In Cytron's SSA this code (roughly) translates into
-
-{\small
-\begin{verbatim}
-for i in range(N):
-1. A_1 = PHI(A_0,A_2)
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. A_2 = update(A_1, i, B_1[i] + 10); 
-6. B_2 = update(B_1, i, A_2[i] * D_1[i-1]);
-7. C_2 = update(C_1, i, A_2[i] * D_1[i-1]);
-8. D_2 = update(D_1, i, B_2[i] * C_2[i]);
-\end{verbatim}
-}
-
-\begin{comment}
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE \hspace{0.25cm} A[i] = B[i] + 10;
-\STATE \hspace{0.25cm} B[i] = A[i] * D[i-1];
-\STATE \hspace{0.25cm} C[i] = A[i] * D[i-1];
-\STATE \hspace{0.25cm} D[i] = B[i] * C[i];
-\end{algorithmic}
-
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{A}_0$ = $\phi(\texttt{A},\texttt{A}_1)$
-\STATE 2. $\texttt{B}_0$ = $\phi(\texttt{B},\texttt{B}_1)$
-\STATE 3. $\texttt{C}_0$ = $\phi(\texttt{C},\texttt{C}_1)$
-\STATE 4. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$
-\STATE 5. $\texttt{A}_1$ = $\texttt{A}_0$; $\texttt{A}_1$[i] = $\texttt{B}_0$[i] + 10; \COMMENT{equiv. to Update}
-\STATE 6. $\texttt{B}_1$ = $\texttt{B}_0$; $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 7. $\texttt{C}_1$ = $\texttt{C}_0$; $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 8. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
-\end{algorithmic}
-\end{comment}
-
-There is a cycle around \texttt{B\_1 = PHI(B\_0,B\_2)} that includes statement 
-\texttt{A\_1 = update(A\_0, i, B\_1[i] + 10)} and that statement won't be vectorized even 
-though in fact there is no loop-carried dependency from the write of \texttt{B\_1[i]} at 6 to the
-read of \texttt{ ... = B\_1[i]} at 8.
-
-The following algorithm removes certain infeasible loop-carried dependencies that are due to array writes. Consider a loop with index $0 \le j < J$
-nested at $i,j,k$. Here $i$ represents the enclosing loops of $j$ and $k$ represents the enclosed loops in $j$.
-
-\begin{algorithmic}
-\FOR {each array $\texttt{A}$ written in loop $j$}
-\STATE \COMMENT { including enclosed loops in $j$ }
-\STATE dep = False
-\FOR {each pair def: $\texttt{A}_m[f(i,j,k)] = ... $, and use: $ ... = \texttt{A}_n[f'(i,j,k)]$ in loop $j$}
-\IF {$\exists \underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$, s.t.
-$0\!
-\le\!
-\underline{i}\!
-<\!
-I,\: 0\!
-\le\!
-\underline{j}, \underline{j}' \!
-<\!
-J,\: 0\!
-\le\!
-\underline{k}, \underline{k}' \!
-<\!
-K,\: \underline{j}<\underline{j}'$, and $f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$}
-\STATE dep = True
-\ENDIF
-\ENDFOR
-\IF {dep == False}
-\STATE remove back edge into $\texttt{A}$'s $\phi$-node in loop $j$.
-\ENDIF
-\ENDFOR
-\end{algorithmic}
-
-Consider a loop $j$ enclosed in some fixed $\underline{i}$. Only if an update (definition) $\texttt{A}_m[f(i,j,k)] = ... $ at some iteration $\underline{j}$
-references the \emph{same} array element as a use $ ... = \texttt{A}_n[f'(i,j,k)]$ at some later iteration $\underline{j}'$,
-we may have a loop-carried dependence for $\texttt{A}$ due to this def-use pair. (In contrast, Cytron's algorithm inserts a loop-carried dependency every time there is an array update.)
-The algorithm above examines all def-use pairs in loop $j$, including defs and uses in nested loops, searching for values $\underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$ that satisfy
-$f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$. If such values exist for some def-use pair, then there is a potential
-loop-carried dependence on $\texttt{A}$; otherwise there is not and we can remove the spurious backward edge thus ``freeing up'' statements for vectorization. %\ana{Double check this!!!}
-
-Consider the earlier example.
-There is a single loop, $i$.
-Clearly, there is no pair $\underline{i}$ and $\underline{i}'$, where $\underline{i} < \underline{i}'$ that make $\underline{i} = \underline{i}'$
-due to the def-use pairs of \texttt{A} 5-6 and 5-7.
-Therefore, we remove the back edge from 5 to the phi-node 1.
-Analogously, we remove the back edges from 6 to 2 and from 7 to 3.
-However, there are many values $\underline{i} < \underline{i'}$ that make $\underline{i} = \underline{i'}-1$ and the back edge from 8 to 4 remains (def-use pairs for \texttt{D}).
-As a result of removing these spurious edges, Basic Vectorization will find that statement 5 is vectorizable.
-Statements 6, 7 and 8 will correctly appear in the FOR loop.
-
-Note however, that this step renders some array phi-nodes target-less. We handle target-less phi-nodes with a minor extension of Basic Vectorization (Phase 2). 
-First, we merge closures that update the same array. This simplifies handling of array $\phi$-nodes: if each closure is turned into a separate loop
-each loop will need to have its own array phi-node to account for the update and this would complicate the analysis.
-Second, we add the target-less node of array \texttt{A} back to the closure that updates \texttt{A} ---
-the intuition is, even if there is no loop-carried dependence from writes to reads on \texttt{A}, \texttt{A} is written and the write (i.e., update) cannot be vectorized; 
-therefore, the updated array has to carry to the next iteration of the loop. Third, in cases when the phi-node remains target-less, i.e., cases when the
-array write can be vectorized, we have to properly remove the phi-node replacing uses of the left-hand side of the phi-node with its arguments.
-
-%\ana{Commented out previous writeup. Removal/handling of targetless phi-nodes will go into the Extension to Basic vectorization.}
-
-\begin{comment}
-The final step is to remove array phi-nodes that are left without incoming back edges after removal of infeasible edges.
-Such a phi-node is redundant when there are no loop-carried dependences that are due to writes to the corresponding array.
-In our running example, the phi-nodes for \texttt{A}, \texttt{B}, and \texttt{C}, are redundant.
-
-Removing redundant phi-nodes proceeds from \emph{innermost to outermost} loops as follows.
-Denote a redundant phi-node by $\texttt{A}_1 = \phi(\texttt{A}_0,\texttt{A}_k)$. $\texttt{A}_0$ is the array at the beginning of the loop
-and $\texttt{A}_k$ is the array at the end of an iteration (including the last iteration). We consider two cases:
-\begin{enumerate}
-\item[Case 1] Inner-to-outer or same-level def-use edge
-\begin{itemize}
-\item If use is an array read $ ... = \texttt{A}_1[f(i)]$, then replace $\texttt{A}_1$ with $\texttt{A}_0$. This is correct, i.e., it is guaranteed that the use
-references $\texttt{A}_0$; had the use been referencing an array modified in an earlier iteration, the back-edge to the phi-node would
-not have been removed and the phi-node would not have been classified as redundant.
-\item If use is an array write (update) $\texttt{A}_2 = \texttt{A}_1; \texttt{A}_2[i] = ... $ equiv. to $\texttt{A}_2 = \mathit{update}(\texttt{A}_1, i, ...)$ then handling becomes more tricky.
-The semantics of this update is that when $i$ is 0, then $\texttt{A}_0$ is being read.
-But when $i > 0$, then $\texttt{A}_k$ is being read.
-If we are at this case, this means that the def-use edge is same-level and we are removing a redundant node in the loop enclosing the update.
-The removal of the original phi-node entails pushing the phi-node at the update:
-$\texttt{A}_2 = \phi(\texttt{A}_0,\texttt{A}_k); \texttt{A}_2[i] = ... $ equiv. to $\texttt{A}_2 = \mathit{update}(\phi(\texttt{A}_0,\texttt{A}_k), i, ...)$.
-\item If use is a phi-node $... = \phi(\texttt{A}_1,...)$, that means we have an outer-to-inner edge. The
-
-\end{itemize}
-\end{enumerate}
-For each outer-to-inner or same-level def-use edge
-that originates at the node, replace $\texttt{A}_1$ at the use node by $\texttt{A}_0$.
-For each inner-to-outer def-use edge, replace $\texttt{A}_1$ by $\texttt{A}_k$. In this case, the use references the array
-that results from the computation in the loop.
-
-Our running example is transformed as follows. Removing the phi-nodes for \texttt{A}, \texttt{B}, and \texttt{C} transforms the
-replaces the references to $\texttt{A}_0$, $\texttt{B}_0$, and $\texttt{C}_0$, with references to $\texttt{A}$, $\texttt{B}$, and $\texttt{C}$
-respectively, i.e., the arrays before the loop.
-
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$
-\STATE 2. $\texttt{A}_1$ = $\texttt{A}$; $\texttt{A}_1$[i] = $\texttt{B}$[i] + 10;
-\STATE 3. $\texttt{B}_1$ = $\texttt{B}$; $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 4. $\texttt{C}_1$ = $\texttt{C}$; $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
-\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
-\end{algorithmic}
-\end{comment}
-
-\begin{comment}
-\subsubsection{Array MUX refinement}
-
-\ana{TODO: WIP, get to BV without MUX refinement for now? For now, skip MUX refinement.}
-
-Next, the algorithm refines array MUX statements.
-MPC-source after Cytron's SSA may result in statements $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$, which imply that any index
-of $A$ can be written at this point and therefore there is a loop-carried dependency. In some cases the MUX can be refined to just a single
-index or a pair of indices, e.g., $\texttt{A}_j[i] = \emph{MUX}(c, \texttt{A}_k[i], \texttt{A}_l[i])$.
-
-%\ana{Clarify why this helps. Still work in progress.}
-This is to reduce the dimensionality of simd-ified computation. Technically, $\texttt{A}_j = \emph{MUX}(..., \texttt{A}_k, \texttt{A}_l)$ is
-a simdified operation that can be carried out in parallel "in one round". However, particularly when \texttt{A} is a multi-dimensional array, there is
-substantial increase in the size of the arrays (vectors) we send to SIMD operations. Refining to an update to a specific index would reduce
-the size of those vectors. Note that this is a heuristic that handles a common case, but not all cases of array updates.
-%\ana{TODO: Simplify this algorithm, taking into account the restriction to canonical updates. It should be handling \_all\_ cases.}
-
-\begin{algorithmic}
-\FOR {each $\mathit{stmt}$: $\texttt{A}_j = \emph{MUX}(c, \texttt{A}_k, \texttt{A}_l)$ in the MPC-source seq.}
-\STATE $i_1 = \mathit{find\_update}(A_k)$ \COMMENT { Is null when $A_k = \phi(...)$ }
-\STATE $i_2 = \mathit{find\_update}(A_l)$ \COMMENT { Is null when $A_l = \phi(...)$ }
-\IF {$i_1 == i_2$ or $i_1$ is null or $i_2$ is null} \STATE \COMMENT{ With our restrictions on writes we must have $i_1 = i_2$. }
-\STATE replace $\mathit{stmt}$ with \\
-\STATE $\texttt{A}_j = \texttt{A}_{j-1}$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
-%\ELSIF {$i_1 \ncong i_2$}
-%\STATE replace $\mathit{stmt}$ with \\
-%\STATE $\texttt{A}_j = \texttt{A}_k$; $\texttt{A}_j[i_1] = \emph{MUX}(c, \texttt{A}_k[i_1], \texttt{A}_l[i_1])$
-\ELSE
-\STATE $\mathit{stmt}$ stays as is
-\ENDIF
-\ENDFOR
-\end{algorithmic}
-\end{comment}
-
-\subsubsection{Restricting Array Writes}
-
-For now, we restrict array updates to \emph{canonical updates}. Assume (for simplicity) a two-dimensional array \texttt{A[I,J]}. A canonical update is the following:
-\begin{verbatim}
-for i in range(I):
-  for j in range(J):
-    ...
-    A[i,j] = ...
-    ...
-\end{verbatim}
-
-The update \texttt{A[i,j]} can be nested into an inner loop and there may be multiple updates, i.e., writes to \texttt{A[i,j]}.
-However, update such as \texttt{A[i-1,j] = ...} or \texttt{A[i-1,j-1] = ...}, etc., is not allowed.
-Additionally, while there could be several different loops that perform canonical updates, they must be of the same dimensionality, i.e., an update of higher or lower dimension, e.g., \texttt{A[i,j,k] = ...} is not allowed.
-We compute the \emph{canonical dimensionality} of each write array by examining the array writes in the original program and rejecting programs that violate the canonical write restriction.
-This restriction simplifies reasoning in this early stage of the compiler; we will look to relax the restriction in future work.
-
-Another restriction/assumption is that we assume the output array is given as input with initial values, and it is of size consistent with its canonical dimensionality. 
-%\ana{This restriction will require that we update our benchmarks that come with output arrays.}
-
-Reads through an arbitrary formula, such as \texttt{A[i-1]} for example, are allowed; currently, the projection function returns dummy values if the read formula is out of bounds; we assume the programmer ensures that the program still computes correct output in this case.
-
-\subsubsection{Changes to Basic Vectorization}
-
-\begin{comment}
-There are two changes to Basic Vectorization to account for write arrays. We can account for them during processing of raise dimension/drop dimension in Phase 1 of Basic Vectorization.
-
-One change to Basic vectorization is the expansion of dimension if the array write or read occurs in a nested loop.
-That is, if there is an update $\texttt{A}[i,j] = ...$ that occurs in loop dimensionality $i,j,k$, $\texttt{A}[i,j]$ will be rewritten into $\texttt{A}[i,j,k]$.
-Similarly, a read $\texttt{A}[f(i,j)]$ will be rewritten into $\texttt{A}[f(i,j),k]$.
-($\texttt{A}[f(i,j),k]$ will be reshaped as $f(i,j)*I*J+k$ during Phase 1 of Basic Vectorization.
-\ana{TODO: Double check.})
-\end{comment}
-
-In addition to the changes for the handling of target-less phi-nodes, Basic Vectorization has to handle def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable.
-The definition can be an update \texttt{A\_2 = update(A\_1,i,...)}, a pseudo $\phi$-node \texttt{A\_2 = PHI(A\_0,A\_1)}, etc..
-Note that $\phi$-nodes for arrays have no subscript operations the way there are subscript operations in analysis-introduced arrays representing scalars.
-While there are variations, the most intuitive implementation will perform Basic Vectorization Phase 1 as is, inserting
-$\mathit{raise\_dim}$ and $\mathit{drop\_dim}$ in the same way.
-However, the implementation of raise dimension and drop dimension will be adapted because the dimension cannot be raised or dropped to a dimension lower than the canonical one.
-Consider a def-use edge $X \rightarrow Y$ for an array $\texttt{A}$.
-
-\begin{enumerate}
-
-\item same-level $X \rightarrow Y$. Do nothing, propagate the array, which happens to be of the right dimension. 
-%\ana{There might be some opportunities to do copy propagation optimization and save some cycles, but let's leave this for later.}
-
-\item inner-to-outer $X \rightarrow Y$ triggers the addition of $\mathit{drop\_dim}$.
-However, the dimensionality cannot be dropped below the canonical dimensionality of the array.
-E.g., if the dimensionality of the loop enclosure $X$ is already at the canonical one, then $\mathit{drop\_dim}$ has no effect.
-
-\item outer-to-inner $X \rightarrow Y$ riggers  $\mathit{raise\_dim}$.
-Again, if the dimensionality of the loop enclosure of $Y$ is smaller or same as the canonical dimensionality of the array, then it has no effect, otherwise, if dimensionality is greater than the canonical dimensionality, $\mathit{raise\_dim(...)}$ (at $X$) is the same as in Basic Vectorization.
-
-\item "mixed" $X \rightarrow Y$. We assume that the mixed edge is transformed into an inner-to-outer followed by outer-to-inner edge before we perform vectorization, just as with Basic vectorization.
-
-\end{enumerate}
-
-If the use of the array is a read $\texttt{A}[f(i,j,k)]$ different than a canonical read $\texttt{A}[i,j,k]$, then we need to add a reshape operation as 
-all arrays are $\texttt{A}[i,j,k]$. It can be added after raise\_dim/drop\_dim or incorporated in these operations.
-The bulk of the change is in Phase 2 of Basic Vectorization as outlined earlier.
-
-\subsubsection{Examples with Array Writes}
-
-\paragraph{Example 1}
-
-First, the canonical dimensionality of all \texttt{A,B,C} and \texttt{D} is 1. 
-After Phase 1 of Basic Vectorization the Aiken's array write example will be (roughly) as follows:
-
-{\small
-\begin{verbatim}
-for i in range(N):
-1. A_1 = PHI(A_0,A_2)
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. A_2 = update(A_1, I, B_1[I] + 10); 
-6. B_2 = update(B_1, I, A_2[I] * D_1[I-1]);
-7. C_2 = update(C_1, I, A_2[I] * D_1[I-1]);
-8. D_2 = update(D_1, I, B_2[I] * C_2[I]);
-\end{verbatim}
-}
-
-Note that since all def-uses are same-level (i.e., reads and writes of the array elements) 
-no raise dimension or drop dimension happens.
-
-\begin{comment}
-\begin{algorithmic}
-\STATE for i in range(N):
-\STATE 1. $\texttt{A}_1$ = $\phi(\texttt{A}_0,\texttt{A}_2)$
-\STATE 2. $\texttt{B}_1$ = $\phi(\texttt{B}_0,\texttt{B}_2)$
-\STATE 3. $\texttt{C}_1$ = $\phi(\texttt{C}_0,\texttt{C}_2)$
-\STATE 4. $\texttt{D}_1$ = $\phi(\texttt{D}_0,\texttt{D}_2)$
-\STATE 5. $\texttt{A}_2 = update(\texttt{A}_1, \stackrel{\rightarrow}{i}, \texttt{B}_1[\stackrel{\rightarrow}{i}] + 10)$;
-\STATE 6. $\texttt{B}_2 = update(\texttt{B}_1, \stackrel{\rightarrow}{i}, \texttt{A}_2[\stackrel{\rightarrow}{i}] * \texttt{D}_1[\stackrel{\rightarrow}{i-1}])$;
-\STATE 7. $\texttt{C}_2 = update(\texttt{C}_1, \stackrel{\rightarrow}{i}, \texttt{A}_2[\stackrel{\rightarrow}{i}] * \texttt{D}_1[\stackrel{\rightarrow}{i-1}])$;
-\STATE 8. $\texttt{D}_2 = update(\texttt{D}_1; \stackrel{\rightarrow}{i}, \texttt{B}_2[\stackrel{\rightarrow}{i}] * \texttt{C}_2[\stackrel{\rightarrow}{i}])$;
-\end{algorithmic}
-\end{comment}
-
-Phase 2 computes the closure of 4; $cl = \{4,6,7, 8\}$ 
-while 5 is vectorizable. Recall that 1,2, and 3 are target-less phi-nodes. Since the closure $cl$ includes updates to \texttt{B} and \texttt{C},
-the corresponding phi-nodes are added back to the closure and the def-use edges are added back to the target-less nodes. The uses of \texttt{A\_1} and \texttt{B\_1} in the vectorized statement
-turn into uses of \texttt{A\_0} and \texttt{B\_0} respectively; this is done for all original target-less phi-node.
-(But note that \texttt{A\_0} is irrelevant; the update writes into array \texttt{A\_2} in parallel.)
-Finally, the target-less phi-node for \texttt{A} is discarded.
-
-{\small
-\begin{verbatim}
-1. A_2 = update(A_0, I, ADD_SIMD(B_0[I],10));
-   equiv. to A_2[I] = ADD_SIMD(B_0[I],10)
-FOR i=0; i<N; i++; // MOTION loop
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. B_2 = update(B_1, i, A_2[i] * D_1[i-1]);
-   equiv. to B_2 = B_1; B_2[i] = A_2[i] * D_1[i-1];
-6. C_2 = update(C_1, i, A_2[i] * D_1[i-1]);
-7. D_2 = update(D_1, i, B_2[i] * C_2[i]);
-\end{verbatim}
-}
-
-\begin{comment}
-\begin{algorithmic}
-\STATE 1. $\texttt{A}_2 = update(\texttt{A}_0; \stackrel{\rightarrow}{i}, \mathit{ADD\_SIMD}(\texttt{B}_0[\stackrel{\rightarrow}i],[10,...])$ \COMMENT{Fully vectorized, size N.}
-\STATE FOR i=0; i<N; i++; \COMMENT{ MOTION loop }
-\STATE 2. $\texttt{B}_1$ = $\phi(\texttt{B}_0,\texttt{B}_2)$
-\STATE 3. $\texttt{C}_1$ = $\phi(\texttt{C}_0,\texttt{C}_2)$
-\STATE 4. $\texttt{D}_1$ = $\phi(\texttt{D}_0,\texttt{D}_2)$
-\STATE 5. $\texttt{B}_2 = update(\texttt{B}_1, i, \mathit{MUL}(\texttt{A}_2[i], \texttt{D}_1[{i-1}])$
-\STATE 6. $\texttt{C}_2 = update(\texttt{C}_1, i, \mathit{MUL}(\texttt{A}_2[i], \texttt{D}_1[{i-1}])$
-\STATE 7. $\texttt{D}_2 = update(\texttt{D}_1, i, \mathit{MUL}(\texttt{B}_2[i], \texttt{C}_1[i])$
-\end{algorithmic}
-\end{comment}
-
-\paragraph{Example 2} Now consider the MPC Source of Histogram:
-
-{\small
-\begin{verbatim}
-for i in range(0, num_bins):
-   res1 = PHI(res, res2)
-   for j in range(0, N):
-      res2 = PHI(res1, res3)
-      tmp1 = (A[j] == i)
-      tmp2 = (res2[i] + B[j])
-      tmp3 = MUX(tmp1, res2[i], tmp2)
-      res3 = Update(res2, i, tmp3)     
-return res1
-\end{verbatim}
-}
-
-The canonical dimensionality of \texttt{res} is 1.
-Also, the phi-node \texttt{res1 = PHI(res, res2)} is a target-less phi-node (the implication being that the inner for loop can be vectorized across $i$).
-After Phase 1, Basic vectorization produces the following code (statements are implicitly vectorized along $i$ and $j$).
-In a vectorized update statement, we can ignore the incoming array, \texttt{res2} in this case.
-The update writes (in parallel) all locations of the 2-dimensional array, in this case it sets up each \texttt{res3[i,j] = tmp3[i,j]}.
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I = raise_dim(i, ((i:num_bins),(j:N)))
-for i in range(0, num_bins):
-   res1 = PHI(res, res2') // target-less phi-node
-   res1' = raise_dim(res1, (j:N))
-   for j in range(0, N):
-      res2 = PHI(res1', res3)
-      tmp1 = (A1 == I)
-      tmp2 = (res2 + B1)
-      tmp3 = MUX(tmp1, res2, tmp2)
-      res3 = Update(res2, (I,J), tmp3) 
-   res2' = drop_dim(res2)    
-res1'' = drop_dim(res1)
-return res1''
-\end{verbatim}
-}
-
-Processing the inner loop in Phase 2 vectorizes \texttt{tmp1 = (A1 == I)} along the $j$ dimension but leaves the rest of the statements in a MOTION loop.
-Processing the outer loop is interesting.
-This is because the PHI node is a target-less node, and therefore, there are no closures!
-Several things happen.
-(1) Everything can be vectorized along the $i$ dimension.
-(2) We remove the target-less PHI node, however, we must update uses of \texttt{res1} appropriately: the use at raise\_dim goes to the first argument of the PHI function and the use at drop\_dim goes to the second argument.
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I1 = raise_dim(i, ((i:num_bins),(j:N)))
-
-tmp1[I,J] = (A1[I,J] == I1[I,J])
-
-res1' = raise_dim(res, (j:N)) // replacing res1 with res, 1st arg
-for j in range(0, N):
-   res2 = PHI(res1', res3)
-   tmp2[I,j] = (res2[I,j] + B1[I,j])
-   tmp3[I,j] = MUX(tmp1[I,j], res2[I,j], tmp2[I,j])
-   res3 = Update(res2, (I,j), tmp3)
-   equiv. to res3 = res2; res3[I,j] = tmp3[I,j]
-res2' = drop_dim(res2)   
-res1 = drop_dim(res2') // replacing with res2', 2nd arg. NOOP
-return res1
-\end{verbatim}
-}
-
-\begin{comment}
-The cleanup phase can then get rid of $j$ in the \texttt{res} variables and \texttt{tmp} variables, so the loop operates over the \texttt{num\_bins}-size vectors. 
-This will lead to dropping raise\_dim and drop\_dim as well:
-
-{\small
-\begin{verbatim}
-A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
-B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
-I1 = raise_dim(i, ((i:num_bins),(j:N)))
-
-tmp1[I,J] = (A1[I,J] == I1[I,J])
-
-for j in range(0, N):
-   res2 = PHI(res1, res3) // now 1-dim array
-   tmp2[I] = ADD_SIMD(res2[I],B1[I,j])
-   tmp3[I] = MUX_SIMD(tmp1[I,j], res2[I], tmp2[I])
-   res3 = Update(res2, I, tmp3[I])
-return res2
-\end{verbatim}
-}
-\end{comment}
-
-\begin{comment}
-\paragraph{Example 2}
-
-Consider MPC-source of Histogram after removal of infeasible edges and redundant phi-nodes, and MUX refinement:
-
-\begin{algorithmic}
-
-\STATE res!1 = []
-\STATE for i: plaintext in range(0, num\_bins):
-\STATE \hspace{0.25cm} res!2 = $\phi$(res!1, res!3)
-\STATE  \hspace{0.25cm} ...
-\STATE res!4 = res!2 \COMMENT{Added due to the "mixed" def-use edge.}
-\STATE for i: plaintext in range(0, num\_bins):
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE \hspace{0.25cm}        for j: plaintext in range(0, N):
-\STATE \hspace{0.5cm}            res!5 = $\phi$(res!4, res!7)
-\STATE \hspace{0.5cm}            !2!3 = (A[j] == i)
-\STATE \hspace{0.5cm}            !3!3 = (res!5[i] + B[j])
-\STATE \hspace{0.5cm}            res!6 = res!5; res!6[i] = !3!3
-\STATE \hspace{0.5cm}            res!7 = res!6; res!7[i] = MUX(!2!3, res!6[i], res!5[i])
-
-\end{algorithmic}
-
-After Phase 1 of Basic Vectorization:
-
-\begin{algorithmic}
-\STATE res!1 = []
-\STATE for i: plaintext in range(0, num\_bins):
-\STATE \hspace{0.25cm} res!2 = $\phi$(res!1, res!3)
-\STATE  \hspace{0.25cm} ...
-\STATE res!4 = res!2 \COMMENT{Added due to the "mixed" def-use edge. Because the canonical dimensionality of res is 1, we DO NOT drop dimension of res!2.}
-\STATE A[$i,j$] = $\mathit{raise\_dim}$(A[$j$],$i$) \COMMENT{Copying row A num\_bin times. Reusing name.}
-\STATE B[$i,j$] = $\mathit{raise\_dim}$(B[$i$],$j$) \COMMENT{Copying column B N times. Reusing name.}
-\STATE res!4[$i,j$] = $\mathit{raise\_dim}$(res!4[$i$],$j$) \COMMENT{Reusing name.}
-\STATE for i: plaintext in range(0, num\_bins):
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE \hspace{0.25cm}        for j: plaintext in range(0, N):
-\STATE \hspace{0.5cm}            res!5 = $\phi$(res!4, res!7) \COMMENT{Do nothing. res!4 has been raised to proper dimension.}
-\STATE \hspace{0.5cm}            !2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{EQ}$(A[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], [[0,0...], [1,1...],...[num\_bins-1,num\_bins-1...]])
-\STATE \hspace{0.5cm}            !3!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{ADD}$(res!5[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$],B[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$])
-\STATE \hspace{0.5cm}            res!6 = res!5 \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.5cm}            res!6[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = !3!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$]
-\STATE \hspace{0.5cm}            res!7 = res!6
-\STATE \hspace{0.5cm}            res!7[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{MUX}$(!2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], res!6[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], res!5[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$])
-\end{algorithmic}
-
-After Phase 2 and Phase 3. The EQ operation is vectorizable across both dimensions, and the rest of the computation is vectorizable across the $i$-dimension.
-
-\begin{algorithmic}
-\STATE ...
-\STATE !2!3[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$] = $\mathit{EQ\_SIMD}$(A[$\stackrel{\rightarrow}i,\stackrel{\rightarrow}j$], [[0,0...], [1,1...],...[num\_bins-1,num\_bins-1...]])
-%\STATE \hspace{0.25cm}        result!4 = $\phi$(result!2, result!5)
-\STATE FOR j=0; j<N; j++; \COMMENT{MOTION loop}
-\STATE \hspace{0.25cm}            res!5 = $\phi$(res!4, res!7) \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.25cm}            !3!3[$\stackrel{\rightarrow}i,j$] = $\mathit{ADD\_SIMD}$(res!5[$\stackrel{\rightarrow}i,j$],B[$\stackrel{\rightarrow}i,j$])
-\STATE \hspace{0.25cm}            res!6 = res!5 \COMMENT{Do nothing, just copy.}
-\STATE \hspace{0.25cm}            res!6[$\stackrel{\rightarrow}i,j$] = !3!3[$\stackrel{\rightarrow}i,j$]
-\STATE \hspace{0.25cm}            res!7 = res!6
-\STATE \hspace{0.25cm}            res!7[$\stackrel{\rightarrow}i,j$] = $\mathit{MUX\_SIMD}$(!2!3[$\stackrel{\rightarrow}i,j$], res!6[$\stackrel{\rightarrow}i,j$], res!5[$\stackrel{\rightarrow}i,j$])
-\end{algorithmic}
-\end{comment}
-
-%\section{Divide-and-Conquer}
-%\label{sec:div_and_conq}
-%\ana{TODO: Now that we have broken FOR loops into smaller chunks, we can add Divide-and-conquer reasoning with Z3 and implement this additional transform.}
-
-\section{Implementation and Evaluation}
-\label{sec:implementation}
-
-\section{Related Work}
-\label{sec:related}
-
-\paragraph{MPC languages and compilers.}
-
-Languages and compilers for secure computation have seen significant attention and advances in recent years. The early MPC compilers Fairplay~\cite{CCS:BenNisPin08}, 
-and Sharemind~\cite{ESORICS:BogLauWil08} were followed by PICCO~\cite{CCS:ZhaSteBla13}, Obliv-C~\cite{Zahur:2015}, TinyGarble~\cite{SP:SHSSK15}, 
-Wystiria~\cite{SP:RasHamHic14}, and others. A new generation of MPC compilers 
-includes SPDZ/SCALE-MAMBA/MP-SPDZ~\cite{Keller:2020} and the ABY/HyCC/MOTION~\cite{NDSS:DemSchZoh15,CCS:BDKKS18,Braun:2022} frameworks. 
-These two families are the state-of-the art and
-are actively developed. Another recent development is Viaduct, a functional language and compiler that supports a range of secure computation 
-frameworks, including MPC and ZKP. Hastings et el. present a review of compiler frameworks~\cite{Hastings:2019}.
-
-While each of these languages and compilers brings in new ideas and advances, none addresses the problem of ``circuit independent'' intermediate 
-representation and optimization. We envision a classical compiler structure: (1) a Wysteria, Viaduct, Obliv-C, or IMP Source front end, including rich type systems and 
-AST-level semantic analysis, compile into the MPC Source IR, (2) MPC Source-level optimizations take place, followed by (3) back-end compilers into circuits. 
-Our focus is at the intermediate level.
-
-%{\begin{center}
-%\includegraphics[width=0.5\linewidth]{figs/focus.png}
-%\end{center}
-%}
-
-Many works focus on the implementation of MPC protocols exposing an API to the programmer. For example, the ABY/MOTION line of 
-compiler frameworks provides a library of MPC primitives; the programmer writes MPC programs in C++ on top of the library. These back ends implement 
-different protocols and allow for mixing, but notably, they leave it to the programer to assign different protocols to different parts of the 
-computation and perform share conversion accordingly. In addition, MOTION provides SIMD primitives, which allows for efficient execution 
-of MPC operations, but again, using SIMD primitives is the responsibility of the programer. There is interest in frameworks for automatic mixing, 
-e.g., \cite{CCS:BDKKS18,Ishaq:2019,Fang:2022}.
-
-Other works, e.g., Obliv-C~\cite{Zahur:2015}, Wystiria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021} focus on higher-level language design, particularly information-flow systems that 
-restrict flow between secure and insecure parts of the program. 
-%Key ideas such as ${\sf insecure} <: {\sf secure}$ qualifier systems are similar to the ideas we 
-%present in~\Secref{sec:interprocedural} and~\Secref{sec:zkp} and we intend to draw on these work. While we will design a high-level language expanding our 
-%intraprocedural IMP-MPC syntax and semantics into the \emph{interprocedural} case, our interest and focus lie primarily in the ``machine-independent'' optimizations 
-%thrust of the proposal.
-
-%There is significant development in ZKP languages as well. ~\ana{Alex Ozdemir's talk, ZoKrates, Cairo, others} Our work focuses on ZKP via MPC and our key idea
-%is the separation of the higher-level ``backend-independent'' compilation and optimization and the lower-level compilation into circuits.
-
-\paragraph{Classical HPC compilers.}
-Automatic vectorization is a longstanding problem in high-performance computing (HPC). 
-There are thousands of works in this area reflecting over 40 years of research. We presented a vectorization
-algorithm for MPC Source, essentially extending classical loop vectorization \cite{Allen:1987}. In HPC vectorization, conditional control flow 
-presents a challenge --- one cannot estimate the cost of a schedule or vectorize branches in a straightforward manner --- in contrast to 
-MPC Source vectorization. 
-We view Karrenberg's work on Whole function vectorization~\cite{Karrenberg:2015} as most closely related to ours --- it linearizes the program and vectorizes 
-both branches of a conditional applying masking to avoid execution of the branch-not-taken code, and selection (similar to MUX) to select the correct value based on the 
-result of the condition at runtime. 
-The problem is that masking and selection, or more generally, handling control predicates~\cite{Benabderrahmane:2010,Karrenberg:2015}, 
-can lead to \emph{slowdown}. 
-%As a result, vectorizing compilers and polyhedral compilers tend to focus on straight-line code due to the unpredictable 
-%overhead conditionals may introduce. \ana{add Louis noel pochet's benchmarks. citation on vectorizing compilers???}
-
-We argue that vectorization over linear MPC Source is a different problem, one that warrants a new look, while drawing from 
-results in HPC.
-Since both branches of the conditional and the multiplexer \emph{always} execute, not only can we apply aggressive vectorization on linear code, but (perhaps more importantly) 
-we can also build analytical models that accurately predict execution time. These models in turn would drive optimizations such as vectorization, protocol mixing, and others. 
-Vectorization meshes in with those additional optimizations in non-trivial ways.
-
-Furthermore, extensions of classical loop vectorization with array writes, arbitrary indexing, including non-affine indexing, and interaction with SSA are non-trivial 
-and present novel challenges and opportunities for contribution. Polyhedral parallelization~\cite{Benabderrahmane:2010} considers a higher-level source (typically AST) 
-representation, while our work takes advantage of linear MPC Source and SSA form. The work by Karrenberg~\cite{Karrenberg:2015} is rare in that space, in the sense that it considers 
-vectorization over SSA form, which has similarities to MPC Source. We consider different array representation, notion of dependence, and reasoning 
-about dependence, which we conjecture is more suitable for MPC Source. Buscher~\cite{Buscher:2018} considers SIMD-vectorization 
-at the level of source code, which then combines with circuit-level optimizations in the TinyGarble compiler. 
-He proposes using an off-the-shelf polyhedral compiler, however, application is limited to only two routines, 
-essentially just inner product and euclidian distance; it is unclear how effective the off-the-shelf compiler is. 
-In contrast, we consider vectorization at the level of MPC Source separating ``backend-independent'' vectorization
-and circuit-level amortization (done by MOTION). We apply our compiler on a wide range of routines.
+% Related Work Section
+\input{related.tex}
 

### 2022-04-28 overleaf/introduction.tex
Edits to 1-3. Changed label in 4.

@@
 The rest of the paper is organized as follows.
-Section \secref{sec:overview} presents an overview of the compiler.
-Section \secref{sec:model} describes our model for cost estimation and argues NP-hardness of optimal scheduling.
-Section \secref{sec:compiler} details the phases of the compiler and Section \secref{sec:vectorization} focuses in on vectorization.
-\secref{sec:results} presents the experimental evaluation.
-Section \secref{sec:related} discusses related work and Section \secref{sec:conclusion} concludes.
+\secref{sec:overview} presents an overview of the compiler.
+\secref{sec:model} describes our model for cost estimation and argues NP-hardness of optimal scheduling.
+\secref{sec:compiler} details the phases of the compiler and \secref{sec:vectorization} focuses in on vectorization.
+\secref{sec:results} presents the experimental evaluation.\secref{sec:related} discusses related work and \secref{sec:conclusion} concludes.
 

### 2022-04-28 overleaf/model.tex
Edits to 1-3. Changed label in 4.

@@
 
-\ana{We need an intro to section here.}
+%\ana{We need an intro to section here.}
+\ana{Terminology: "operation" or "instruction" or "gate"?}
+
+This section presents a model to reason about the cost of execution of MPC programs, including accounting for amortization.
+We define the assumptions and setting in~\secref{sec:mpc}. We proceed to define the scheduling problem in~\secref{sec:problem}, 
+which we expected to be able to solve optimally. \secref{sec:np} show that the problem is NP-hard via a reduction to the 
+Shortest Common Supersequence (SCS) problem. Despite the negative general result, we expect the formulation in terms of SCS to 
+be useful as sequences are short and few in practice.
 
@@
 For the purpose of the model we assume that $s$ is large enough and the term $\frac{n}{s}p\alpha$ amounts to a \emph{fixed cost} incurred regardless of
-whether $n$ is $10,000$ or just $1$. (This models the cost of sending a packet from party A to party B for example.) The bottom line is that amortized execution
+whether $n$ is $10,000$ or just $1$. (This models the cost of preparing and sending a packet from party A to party B for example.) The bottom line is that amortized execution
 of $n$ operations is $f(n) = \alpha_\mathit{fix} + n\alpha_{var}$ in contrast to unamortized execution $g(n) = n\alpha_\mathit{fix} + n\alpha_{var}$ and, of course,

### 2022-04-28 overleaf/model.tex
Finished edits to 3.

@@
 We define the assumptions and setting in~\secref{sec:mpc}. We proceed to define the scheduling problem in~\secref{sec:problem}, 
-which we expected to be able to solve optimally. \secref{sec:np} show that the problem is NP-hard via a reduction to the 
+which we expected to be able to solve optimally. \secref{sec:np} shows that the problem is NP-hard via a reduction to the 
 Shortest Common Supersequence (SCS) problem. Despite the negative general result, we expect the formulation in terms of SCS to 
@@
 For the purpose of the model we assume that $s$ is large enough and the term $\frac{n}{s}p\alpha$ amounts to a \emph{fixed cost} incurred regardless of
-whether $n$ is $10,000$ or just $1$. (This models the cost of preparing and sending a packet from party A to party B for example.) The bottom line is that amortized execution
-of $n$ operations is $f(n) = \alpha_\mathit{fix} + n\alpha_{var}$ in contrast to unamortized execution $g(n) = n\alpha_\mathit{fix} + n\alpha_{var}$ and, of course,
-$\alpha_\mathit{fix} << n\alpha_\mathit{fix}$.
+whether $n$ is $10,000$ or just $1$. (This models the cost of preparing and sending a packet from party A to party B for example.) Therefore, amortized execution
+of $n$ operations is $f(n) = \alpha_\mathit{fix} + n\alpha_{var}$ in contrast to unamortized execution $g(n) = n\alpha_\mathit{fix} + n\alpha_{var}$ and, of course.
+Clearly $\alpha_\mathit{fix} << n\alpha_\mathit{fix}$ and since fixed cost dominates variable cost (particularly for remote operations), we have $f(n) << g(n)$.
 %We assume infinite parallel capacity---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely $\alpha$ or $\beta$.
@@
 2 MUL instructions can be amortized in a single SIMD instruction that costs $\alpha_\mathit{fix} + 2\alpha_\mathit{var}$, however a MUL and a MUX instruction
-still cost $2\alpha_\mathit{fix} + 2\alpha_\mathit{var}$ even when scheduled ``in parallel''.
+still cost $2\alpha_\mathit{fix} + 2\alpha_\mathit{var}$ even when scheduled ``in parallel''.\footnote{This is not strictly true, but assuming it, e.g. as in 
+\cite{Ishaq2019, Demmler2015ABYA, Mohassel2018}, helps simplify the problem.}
 %MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption,

### 2022-04-28 overleaf/prob_stmt_np_hardness.tex
Finished edits to 3.

@@
 \begin{enumerate}
-    \item All $P$'s consist of MPC instructions of the same kind, e.g., all MUL, or ADD, etc. 
+    \item All $P_i$'s consist of MPC instructions of the same kind, e.g., all MUL, MUX, ADD, etc. 
     \item Def-use dependencies of the graph $G(V, E)$ should be preserved i.e.
@@
 
-In order to benefit from parallelization/amortization, we must schedule two or more $A$-instructions in the same parallel node (or two or more $M$-instructions in the same parallel node).
+%In order to benefit from parallelization/amortization, we must schedule two or more $A$-instructions in the same parallel node $P_i$ (or two or more $M$-instructions in the same parallel node).
 Recall that we also assume that scheduling $A$-instructions in parallel with $M$-instruction does not benefit from amortization\footnote{this is not strictly true, but assuming it, e.g.
@@
 \begin{equation}
-    \mathit{cost}(S) = \sum_{i=1}^n \mathit{cost}(S_i)
+    \mathit{cost}(S) = \sum_{i=1}^n \mathit{cost}(S_i) = L_\alpha \alpha_\mathit{fix} + L_\beta \beta_\mathit{fix} + L_\alpha \alpha_\mathit{var} + L_\beta \beta_\mathit{var}
 \end{equation}
 
-where $cost(S_i) = \alpha$ or $\beta$. %if $S_i$ is an $A$-instruction and $cost(S_i) = \beta$ if $S_i$ is an $M$-instruction. 
-Similarly, the cost of schedule $P$ is
+where $L_\alpha$ is the number of $\alpha$-instructions and $L_\beta$ is the number of $\beta$ ones.
+(We used this formula to compute the cost of the unrolled MPC Source program in~\secref{sec:overview}.)
+%$cost(S_i) = \alpha$ or $\beta$. if $S_i$ is an $A$-instruction and $cost(S_i) = \beta$ if $S_i$ is an $M$-instruction. 
+The cost of schedule $P$ is more interesting:
 
@@
 Each $P_i$ may contain multiple instructions, and $\mathit{cost}(P_i)$ is amortized.
-Thus, according to our model $\mathit{cost}(P_i) = \alpha_\mathit{fix} + |P_i|\alpha_\mathit{var}$ if $P_i$ stores $\alpha$-instructions, 
+Thus, according to our model $\mathit{cost}(P_i) = \alpha_\mathit{fix} + |P_i|\alpha_\mathit{var}$ if $P_i$ stores $|P_i|$ $\alpha$-instructions, 
 or $\mathit{cost}(P_i) = \beta_\mathit{fix} + |P_i|\beta_\mathit{var}$ if it stores $\beta$-instructions. 
+(Similarly, we used this formula to compute the cost of the Optimized MPC Source program in~\secref{sec:overview}.)
 %if $P_i$ consists of $A$-instructions only, it is $\beta$ if $P_i$ consists of $M$-instructions only, and it is $(\alpha + \beta)$ if $P_i$ mixes $A$-instructions and $M$-instructions.
+
 Our goal is to construct a parallel schedule $P$ that reduces the program cost (when compared to cost of $S$), possibly an optimal schedule. 
@@
 
-Note that we consider a linearized MPC schedule $S$ above for ease of exposition only.
+%Note that we consider a linearized MPC schedule $S$ above for ease of exposition only.
 In our tool-chain we use an MPC Source control flow graph (CFG) $G'(V', E')$ along with def-use graph $G(V,E)$ to construct $P$.
@@
 
-To prove that optimal scheduling is an NP-Hard problem, we consider the following convenient representation.
-An MPC program is represented as a set of sequences $S = \{S_1, \dots, S_n\}$.
-Each element $S_i \in S$ is a tuple.
-The items of the tuple $S_i$ are operations, i.e.
-$A$ or $M$ instructions, that have to be executed in order (operations depend on previous operations i.e.
-$S_i[j], j > 1$ depends $S_i[j-1]$).
-However, the sequences $S_i, 1 \leq i \leq n$ themselves, can overlap each other in any way i.e.
+To prove that optimal scheduling is an NP-Hard problem, we consider the following convenient representation. An MPC program is represented as a set of sequences $\{s_1, \dots, s_n\}$ of operations. 
+%Each element $S_i \in S$ is a tuple. The items of the tuple $S_i$ are operations, i.e. $A$ or $M$ instructions, that have to be executed in order (
+In each sequence $s_i$ operations depend on previous operations via a def-use i.e. $s_i[j], j > 1$ depends on $s_i[j-1]$. 
+%However, the sequences $S_i, 1 \leq i \leq n$ themselves, can overlap each other in any way i.e.
 distinct sequences can be executed in parallel.
@@
 
-As an example, consider the MPC program consisting of the following three sequences, all made up of two distinct $\alpha$-instructions $M_1$ and $M_2$.
+As an example, consider the MPC program consisting of the following three sequences, all made up of two distinct $\alpha$-instructions $M_1$ and $M_2$, e.g., a $M_1$ is MUL and $M_2$ is MUX.
 The right arrow indicates a def-use \emph{dependence}, meaning that the source node must execute before the target node:
@@
 The problem is to find a schedule $P$ with \emph{minimal cost}. For example, a schedule with minimal cost for the sequences above is 
-
-$$ 
+\[ 
 M_1(1), M_1(2) \; ;\; M_1(2) \; ; \; M_2(1), M_2(3) \; ; \; M_1(1), M_1(2), M_1(3) \; ; \;  M_2(3)
-$$
-
+\]
 The parentheses above indicate the sequence where the instruction comes from: (1), (2), or (3). 
@@
 The shortest common supersequence problem is as follows: {\it given two or more sequences find the the shortest sequence that contains all of the original sequences.} This can be solved in $O(n^k)$ time, where $n$ is the cardinality of the longest sequence and $k$ is the number of sequences.
-For our problem $n$ is the maximum length of a sequence and $k$ is the number of total number of sequences.
+For our problem $n$ is the maximum length of a sequence and $k$ is the number of instructions of sequences.
 We can immediately see that the optimal schedule is the shortest schedule, since the shortest schedule minimizes the fixed cost while the variable cost remains the same.
@@
 %We can derive a schedule $P'$ with the same cost as $P$, by mapping each mixed node $P_i \in P$ to two consecutive nodes in $P'$: an $A$-instruction node followed by an $M$-instruction node. 
-Clearly, $P$, which now is a sequence of $M_1$ and $M_2$ nodes, is a supersequence of each sequence $S_i$, i.e., $P$ is a common supersequence of $S_1 \dots S_n$.
-It is also a shortest common supersequence.
+Clearly $P$ is a supersequence of each sequence $s_i$, i.e., $P$ is a common supersequence of $s_1 \dots s_n$. It is also a shortest common supersequence. 
 %To see this, let $X$ and $Y$ denote, respectively, the number of $A$ and $M$ nodes in $P'$. 
@@
 Now suppose, there exist a shorter common supersequence $P'$ of length $L'$.
-$\mathit{cost}(P') < \mathit{cost}(P)$ since $L'\alpha_\mathit{var} + N\alpha_\mathit{var} < L\alpha_\mathit{var} + N\alpha_\mathit{var}$, however, this is a contradiction to the assumption that $P$ has the lowest cost.
+$\mathit{cost}(P') < \mathit{cost}(P)$ since $L'\alpha_\mathit{var} + N\alpha_\mathit{var} < L\alpha_\mathit{var} + N\alpha_\mathit{var}$, contradicting the assumption that $P$ has the lowest cost.
 \qed

### 2022-04-28 overleaf/compiler.tex
Edits to 1-3. Changed label in 4.
WIP on 4. Moved sections into 5.

@@
-\begin{figure*}
-  \includegraphics[width=1.0\linewidth]{figs_paper_SIMD/compiler_framework.png}
-  \caption{Compiler Framework. \ana{Change figure since we are not doing div-and-conquer.}}
-  \label{fig:compiler_framework}
-\end{figure*}
-
 \section{\bf Compiler Framework}
-\label{sec:preliminaries}
+\label{sec:compiler}
 
@@
 
-\figref{fig:compiler_framework} presents an overview of our compiler. This section describes the front-end phases of the compiler
-and Section~\secref{sec:vectorization} describes our vectorization algorithm. \ana{We'll need a section on MOTION code gen too.}
+\begin{figure*}
+  \includegraphics[width=0.9\linewidth]{figs_paper_SIMD/compiler_framework.png}
+  \caption{Compiler Framework. \ana{Change figure since we are not doing div-and-conquer.}}
+  \label{fig:compiler_framework}
+\end{figure*}
+
+\figref{fig:compiler_framework} presents an overview of our compiler. 
+We start the section with a brief description of the top-level syntax and semantic restrictions.
+We proceed to describe several phases of the compiler:~\secref{sec:imp_to_ssa} 
+describes translation from IMP Source to SSA, and~\secref{sec:ssa_to_MPC_Source}
+describes translation form SSA into MPC Source, i.e., the front-end of our compiler. 
+Finally,~\secref{sec:MOTION} describes translation of MPC Source, un-optimized and 
+optimized into MOTION, i.e., this is the back-end of our compiler.
+
+Analysis on MPC Source and vectorization comes in a separate section, Section~\secref{sec:vectorization}. 
 %\ana{Need to state this more precisely.}
@@
 The syntax allows for array accesses, arbitrarily nested loops, and if-then-else control flow.
-Expressions are typed $q \; \tau$, where qualifier $q$ and base type $\tau$ are:
+Expressions are typed $\langle q \; \tau\rangle$, where qualifier $q$ and type $\tau$ are:
 
@@
 The \texttt{shared} qualifier denotes shared values, i.e., ones shared among the parties and computed upon
-under secure computation protocols; the \texttt{plain} qualifier denotes plaintext values. Shared lists denote
-shared elements, but plaintext length. Subtyping is again as expected: $q_1 \; \tau <: q_2 \; \tau$
-iff $q_1 <: q_2$ where $\texttt{plain} <: \texttt{shared}$. Thus, we can convert a plaintext value into a shared one,
-but not vice versa.
+under secure computation protocols; the \texttt{plain} qualifier denotes plaintext values. 
+Shared lists denote shared elements, but plaintext length. Subtyping is 
+$\texttt{plain} <: \texttt{shared}$, meaning that we can convert a plaintext value into a shared one,
+but not vice versa. Subtyping on qualified types is again as expected, it is covariant in the qualifier  
+and invariant in the type: $\langle q_1 \; \tau_1 \rangle <: \langle q_2 \; \tau_2 \rangle$
+iff $q_1 <: q_2$ and  $\tau_1 = \tau_2$.
 
@@
 It is a standard restriction in MPC that the bounds must be known at circuit-generation time.
-\item Arrays are one-dimensional. N-dimensional arrays are linearized and accessed
+\item Arrays are one-dimensional. $N$-dimensional arrays are linearized and accessed
 in row-major order and at this point the programmer is responsible for linearization
@@
 %  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
-  \Gamma \vdash e : \texttt{plain int} \quad \Gamma \vdash \texttt{A} : q \; \texttt{list}[\tau] \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
+  \Gamma \vdash e : \langle \texttt{plain int} \rangle \quad \Gamma \vdash \texttt{A} : \langle q \; \texttt{list}[\tau] \rangle \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
 }{
-  \Gamma \vdash \texttt{A[$e$]} : q \; \tau
+  \Gamma \vdash \texttt{A[$e$]} : \langle q \; \tau \rangle
 }
@@
 %  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
-  \Gamma \vdash e_1 : q_1 \; \texttt{bool} \quad \Gamma \vdash e_2 : q_2 \; \tau \quad \Gamma \vdash e_2 : q_2 \; \tau \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
+  \Gamma \vdash e_1 : \langle q_1 \; \texttt{bool} \rangle \quad \Gamma \vdash e_2 : \langle q_2 \; \tau \rangle \quad \Gamma \vdash e_2 : \langle q_2 \; \tau \rangle \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
 }{
-  \Gamma \vdash \texttt{MUX($e_1$,$e_2$,$e_3$)} : q_1 \vee q_2 \vee q_3 \; \tau
+  \Gamma \vdash \texttt{MUX($e_1$,$e_2$,$e_3$)} : \langle q_1 \vee q_2 \vee q_3  \; \tau \rangle
 }
@@
 \subsection{From IMP Source to SSA}
+\label{sec:imp_to_ssa}
 
@@
 \subsection{From SSA to MPC Source} %: MUX Nodes and Pseudo $\phi$-nodes}
+\label{sec:ssa_to_MPC_Source}
 
@@
 
-\subsection{Dependence Analysis}
-
-The next steps construct the dependence graph and \ana{Add an intro here.}
-
-\subsubsection{Def-use Edges}
-
-The dependence graph has the following def-use edges:
-
-\begin{itemize}
-\item same-level edge $X\rightarrow Y$ where $X$ and $Y$ are in the same loop nest, say $i,j,k$.
-E.g., the def-use edge 9 to 10 in the Biometric MPC Source in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source} is a same-level edge.
-A same-level edge can be a back-edge in which case a $\phi$ node is the target of the edge.
-E.g., 15 to 4 in Biometric is a same-level back-edge.
-\item outer-to-inner $X\rightarrow Y$ where $X$ is in an outer loop nest, say $i$, and $Y$ is in an inner one, say $i,j,k$. E.g., 1 to 4 in Biometric forms is an outer-to-inner edge.
-%A $\phi$-node can be a source or a target of an outer-to-inner forward edge.
-\item inner-to-outer $X\rightarrow Y$ where $X$ is a $phi$-node in an inner loop nest, $i,j,k$, and $Y$ is in the enclosing loop nest $i,j$.
-E.g., the def-use from 8 to 12 gives rise to an inner-to-outer edge.
-%E.g. $\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$ is an inner-to-outer edge.
-An inner-to-outer edge can be a back-edge as well, in which case both $X$ and $Y$ are phi-nodes with the source $X$ in a loop nested into $Y$'s loop (not necessarily immediately).
-
-%Note that the source is \emph{always} a $\phi$-node of the $i,j,k$ loop and the target is in the immediately enclosing loop.
-The interpretation of this edge is that the use node $Y$ uses the definition made in the last iteration of the inner loop.
-%\ben{The current representation of pseudo $\phi$-nodes shows them attached to the loop header (i.e. in the loop).  We may want to clarify that they are also evaluated at the loop's termination.}
-%\item same-level back-edge $X\rightarrow Y$.
-$Y$ is a phi-node in the header of the loop and $X$ is a definition of the variable in the loop body.
-E.g., $\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{sum}_1,\texttt{min}_1)$ to $\texttt{min}_0 = \phi(\texttt{min}_1, 10000)$ in Biometric is a same-level back-edge.
-%\item inner-to-outer back-edge $X\rightarrow Y$: $X$ and $Y$ are both $\phi$-nodes for some variable. The source $X$ is in a loop nested into $Y$'s loop (not necessarily immediately).
-
-\item mixed forward edge $X\rightarrow Y$.
-$X$ is in some loop $i,j,k$ and $Y$ is in a loop nested into $i,j,k'$.
-We transform mixed forward edges as follows.
-Let $\x$ be the variable defined at $X$.
-We add a variable and assignment $\x' = \x$ immediately after the $i,j,k$ loop.
-Then we replace the use of $\x$ at $Y$ with $\x'$.
-This transforms a mixed forward edge into an "inner-to-outer" forward edge followed by an outer-to-inner forward edge.
-Thus, Basic Vectorization handles one of "same-level", "inner-to-outer", or "outer-to-inner" def-use edges.
- \end{itemize}
-
-\subsubsection{Closures}
-
-We define $\mathit{closure}(n)$ where $n$ is a phi-node. Intuitively, it computes the set of nodes (i.e., statements) that form a dependence cycle with $n$. The closure of $n$ is defined as follows:
-\begin{itemize}
-\item $n$ is in $\mathit{closure}(n)$
-\item $X$ is in $\mathit{closure}(n)$ if there is a same-level path from $n$ to $X$, and $X \rightarrow n$ is a same-level back-edge.
-\item $Y$ is in $\mathit{closure}(n)$ if there is a same-level path from $n$ to $Y$ and there is a same-level path from $Y$ to some $X$ in $\mathit{closure}(n)$.
-\end{itemize}
-
-\subsection{Taint Analysis}
-
-We require that all inputs are marked as either shared or plaintext, however, we do not require any other qualifier annotations in the program.
-Then we infer if intermediate variables and expressions are shared through taint analysis with "taintedness" referring to the shared attribute.
-Specifically, the compiler follows the following rules, which are standard in positive-negative qualifier systems (here \texttt{shared} is the positive qualifier and \texttt{plain} is the negative one):
-\begin{enumerate}
-\item Loop counters and array upper bounds are always \texttt{plain}.
-\item If any variable on the right-hand side of an assignment is shared, then the assigned variable is \texttt{shared}.
-%\item If all variables on the right-hand side of an assignment are plaintext, then the assigned variable is plaintext \ana{Will be covered by the last rule.}
-\item Any variables that cannot be determined as shared via the above rules are \texttt{plain}.
-\end{enumerate}
-
-%The first two rules are standard for taint analysis, and the third rule follows from the MPC problem statement. \ben{Is the explanation for the third rule correct?}  \ana{Yes, that is exactly it!}
-The final rule is needed in examples like the following. In the below snippet \texttt{sum!2} and \texttt{sum!3} form a dependency cycle and there is no \texttt{shared} value that flows to either one.
-They are inferred as plaintext.
-%cannot be marked as plaintext through simple taint analysis:
-{\small
-\begin{pythonn}
-plaintext_array = [0, 1, 2, ...]
-sum!1 = 0
-for i in range(0, N):
-    sum!2 = PHI(sum!1, sum!3)
-    sum!3 = sum!2 + plaintext_array[i]
-\end{pythonn}
-}
-%\ben{I think the above example is unnecessary and could be replaced by an explanation of how "untainted" variables are implicitly plaintext, but I couldn't think of a way to phrase that.}\ana{I left the example, we can leave it in the TR, remove from paper.}
-
-When converting to MOTION code, any plaintext value used in the right-hand side of a shared assignment is converted to a shared value for that expression.
-%\ben{Is this necessary to include?} \ana{Yes, that's important.}
-
-Importantly, the taint analysis works on MPC Source, which lacks if-then-else control flow.
-This significantly simplifies treatment as there is no need to handle conditionals and implicit flow.
-
 \subsection{From (Optimized) MPC Source to MOTION}
@@
 % Variable naming
-Our generated C++ uses the following variable-naming scheme: shared variables are named the same as in the MPC source with the \texttt{!} replaced with an underscore (e.g.
-\texttt{sum!2} would be translated to \texttt{sum\_2}).
+Our generated C++ uses the following variable-naming scheme: shared variables are named the same as in the MPC Source with the {\sf !} replaced with an underscore (e.g.
+{\sf sum!2} would be translated to {\sf sum\_2}).
 Plaintext variables follow the same naming convention as shared variables but are prefixed with \texttt{\_MPC\_PLAINTEXT\_}.

### 2022-04-28 overleaf/vectorization.tex
WIP on 4. Moved sections into 5.

@@
 \ana{Need an intro here!}
+
+\subsection{Dependence Analysis}
+\label{sec:dependence}
+
+The next steps construct the dependence graph and \ana{Add an intro here.}
+
+\subsubsection{Def-use Edges}
+
+The dependence graph has the following def-use edges:
+
+\begin{itemize}
+\item same-level edge $X\rightarrow Y$ where $X$ and $Y$ are in the same loop nest, say $i,j,k$.
+E.g., the def-use edge 9 to 10 in the Biometric MPC Source in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source} is a same-level edge.
+A same-level edge can be a back-edge in which case a $\phi$ node is the target of the edge.
+E.g., 15 to 4 in Biometric is a same-level back-edge.
+\item outer-to-inner $X\rightarrow Y$ where $X$ is in an outer loop nest, say $i$, and $Y$ is in an inner one, say $i,j,k$. E.g., 1 to 4 in Biometric forms is an outer-to-inner edge.
+%A $\phi$-node can be a source or a target of an outer-to-inner forward edge.
+\item inner-to-outer $X\rightarrow Y$ where $X$ is a $phi$-node in an inner loop nest, $i,j,k$, and $Y$ is in the enclosing loop nest $i,j$.
+E.g., the def-use from 8 to 12 gives rise to an inner-to-outer edge.
+%E.g. $\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$ is an inner-to-outer edge.
+An inner-to-outer edge can be a back-edge as well, in which case both $X$ and $Y$ are phi-nodes with the source $X$ in a loop nested into $Y$'s loop (not necessarily immediately).
+
+%Note that the source is \emph{always} a $\phi$-node of the $i,j,k$ loop and the target is in the immediately enclosing loop.
+The interpretation of this edge is that the use node $Y$ uses the definition made in the last iteration of the inner loop.
+%\ben{The current representation of pseudo $\phi$-nodes shows them attached to the loop header (i.e. in the loop).  We may want to clarify that they are also evaluated at the loop's termination.}
+%\item same-level back-edge $X\rightarrow Y$.
+$Y$ is a phi-node in the header of the loop and $X$ is a definition of the variable in the loop body.
+E.g., $\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{sum}_1,\texttt{min}_1)$ to $\texttt{min}_0 = \phi(\texttt{min}_1, 10000)$ in Biometric is a same-level back-edge.
+%\item inner-to-outer back-edge $X\rightarrow Y$: $X$ and $Y$ are both $\phi$-nodes for some variable. The source $X$ is in a loop nested into $Y$'s loop (not necessarily immediately).
+
+\item mixed forward edge $X\rightarrow Y$.
+$X$ is in some loop $i,j,k$ and $Y$ is in a loop nested into $i,j,k'$.
+We transform mixed forward edges as follows.
+Let $\x$ be the variable defined at $X$.
+We add a variable and assignment $\x' = \x$ immediately after the $i,j,k$ loop.
+Then we replace the use of $\x$ at $Y$ with $\x'$.
+This transforms a mixed forward edge into an "inner-to-outer" forward edge followed by an outer-to-inner forward edge.
+Thus, Basic Vectorization handles one of "same-level", "inner-to-outer", or "outer-to-inner" def-use edges.
+ \end{itemize}
+
+\subsubsection{Closures}
+
+We define $\mathit{closure}(n)$ where $n$ is a phi-node. Intuitively, it computes the set of nodes (i.e., statements) that form a dependence cycle with $n$. The closure of $n$ is defined as follows:
+\begin{itemize}
+\item $n$ is in $\mathit{closure}(n)$
+\item $X$ is in $\mathit{closure}(n)$ if there is a same-level path from $n$ to $X$, and $X \rightarrow n$ is a same-level back-edge.
+\item $Y$ is in $\mathit{closure}(n)$ if there is a same-level path from $n$ to $Y$ and there is a same-level path from $Y$ to some $X$ in $\mathit{closure}(n)$.
+\end{itemize}
+
+\subsection{Taint Analysis}
+\label{sec:taint}
+
+We require that all inputs are marked as either shared or plaintext, however, we do not require any other qualifier annotations in the program.
+Then we infer if intermediate variables and expressions are shared through taint analysis with "taintedness" referring to the shared attribute.
+Specifically, the compiler follows the following rules, which are standard in positive-negative qualifier systems (here \texttt{shared} is the positive qualifier and \texttt{plain} is the negative one):
+\begin{enumerate}
+\item Loop counters and array upper bounds are always \texttt{plain}.
+\item If any variable on the right-hand side of an assignment is shared, then the assigned variable is \texttt{shared}.
+%\item If all variables on the right-hand side of an assignment are plaintext, then the assigned variable is plaintext \ana{Will be covered by the last rule.}
+\item Any variables that cannot be determined as shared via the above rules are \texttt{plain}.
+\end{enumerate}
+
+%The first two rules are standard for taint analysis, and the third rule follows from the MPC problem statement. \ben{Is the explanation for the third rule correct?}  \ana{Yes, that is exactly it!}
+The final rule is needed in examples like the following. In the below snippet \texttt{sum!2} and \texttt{sum!3} form a dependency cycle and there is no \texttt{shared} value that flows to either one.
+They are inferred as plaintext.
+%cannot be marked as plaintext through simple taint analysis:
+{\small
+\begin{pythonn}
+plaintext_array = [0, 1, 2, ...]
+sum!1 = 0
+for i in range(0, N):
+    sum!2 = PHI(sum!1, sum!3)
+    sum!3 = sum!2 + plaintext_array[i]
+\end{pythonn}
+}
+%\ben{I think the above example is unnecessary and could be replaced by an explanation of how "untainted" variables are implicitly plaintext, but I couldn't think of a way to phrase that.}\ana{I left the example, we can leave it in the TR, remove from paper.}
+
+When converting to MOTION code, any plaintext value used in the right-hand side of a shared assignment is converted to a shared value for that expression.
+%\ben{Is this necessary to include?} \ana{Yes, that's important.}
+
+Importantly, the taint analysis works on MPC Source, which lacks if-then-else control flow.
+This significantly simplifies treatment as there is no need to handle conditionals and implicit flow.
 

### 2022-04-28 compiler/paper_benchmarks.py
fixed minor issues, added support for basic stats

@@ -90,20 +90,6 @@ def get_biometric_inputs() -> tuple[list[InputArgs], int]:
         args.extend(list(map(str, C)))
         args.append("--S")
         args.extend(list(map(str, S)))
-
-        two_C = [2 * C[i] for i in range(D)]
-        args.append("--two_C")
-        args.extend(list(map(str, two_C)))
-        C_sqr_sum = sum(val * val for val in C)
-        args.append("--C_sqr_sum")
-        args.extend(list(map(str, C_sqr_sum)))
-        S_sqr_sum = [sum(S[i * D + j] * S[i * D + j] for j in range(D)) for i in range(N)]
-        args.append("--S_sqr_sum")
-        args.extend(list(map(str, S_sqr_sum)))
-        differences = [0] * D
-        args.append("--differences")
-        args.extend(list(map(str, differences)))
-
         label = "D: {}, N: {}".format(D, N)
         all_args.append(InputArgs(label, args))
     return (all_args, non_vec_up_to)
@@ -124,6 +110,20 @@ def get_biometric_fast_inputs() -> tuple[list[InputArgs], int]:
         args.extend(list(map(str, C)))
         args.append("--S")
         args.extend(list(map(str, S)))
+
+        two_C = [2 * C[i] for i in range(D)]
+        args.append("--two_C")
+        args.extend(list(map(str, two_C)))
+        C_sqr_sum = sum(val * val for val in C)
+        args.append("--C_sqr_sum")
+        args.append(str(C_sqr_sum))
+        S_sqr_sum = [sum(S[i * D + j] * S[i * D + j] for j in range(D)) for i in range(N)]
+        args.append("--S_sqr_sum")
+        args.extend(list(map(str, S_sqr_sum)))
+        differences = [0] * D
+        args.append("--differences")
+        args.extend(list(map(str, differences)))
+
         label = "D: {}, N: {}".format(D, N)
         all_args.append(InputArgs(label, args))
     return (all_args, non_vec_up_to)
@@ -747,11 +747,8 @@ def run_client_role(address):
                             accum_p0 = p0
                             accum_p1 = p1
                         else:
-                            accum_p0 = BenchmarkOutput.sum(accum_p0, p0)
-                            accum_p1 = BenchmarkOutput.sum(accum_p1, p1)
-
-                    accum_p0 = BenchmarkOutput.div(accum_p0, NUM_ITERS)
[…]

### 2022-04-28 compiler/tests/benchmark.py

@@ -36,20 +36,16 @@ class BenchmarkOutput:
                 }
 
     @classmethod
-    def sum(cls, a, b):
+    def by_accumulating_readings(cls, a, b):
         return cls(
            name = a.name,
            output = a.output,
-           timing_stats = statistics.TimingStatistics.sum(a.timing_stats, b.timing_stats),
-           circuit_stats = statistics.CircuitStatistics.sum(a.circuit_stats, b.circuit_stats)
-        )
-    @classmethod
-    def div(cls, obj, d):
-        return cls(
-            name = obj.name,
-            output = obj.output,
-            timing_stats = statistics.TimingStatistics.div(obj.timing_stats, d),
-            circuit_stats = statistics.CircuitStatistics.div(obj.circuit_stats, d)   
+           timing_stats = statistics.TimingStatistics.by_accumulating_readings(
+            a.timing_stats, b.timing_stats
+            ),
+           circuit_stats = statistics.CircuitStatistics.by_accumulating_readings(
+            a.circuit_stats, b.circuit_stats
+            )
         )
 
 

### 2022-04-28 compiler/tests/statistics.py

@@ -4,6 +4,25 @@ from datetime import datetime
 from enum import Enum, unique
 from typing import Optional
 
+def median(lst: list[float]) -> float:
+    assert len(lst) > 1, lst
+    sorted_lst = sorted(lst)
+    mid = len(sorted_lst) // 2
+    median = (sorted_lst[mid] + sorted_lst[~mid]) / 2
+    return median
+
+def mean(lst: list[float]) -> float:
+    assert len(lst) > 1, lst
+    mean = sum(lst) / len(lst)
+    return mean
+
+def stddev(lst: list[float]) -> float:
+    assert len(lst) > 1, lst
+    mean = sum(lst) / len(lst)
+    variance = sum([((x - mean) ** 2) for x in lst]) / len(lst)
+    stddev = variance ** 0.5
+    return stddev
+
 
 @dataclass
 class TimingDatapoint:
@@ -13,6 +32,17 @@ class TimingDatapoint:
     median: float
     stddev: float
     # unit: str = 'ms'
+    readings: list[float]
+
+    def __init__(self, datapoint_name: str, mean: float, median: float, 
+        stddev: float, readings: list[float] = []):
+        self.datapoint_name = datapoint_name
+        self.mean = mean
+        self.median = median
+        self.stddev = stddev
+        self.readings = readings
+        if len(self.readings) == 0:
+            self.readings = [self.mean]
 
     @classmethod
     def from_dictionary(cls, params):
@@ -20,42 +50,37 @@ class TimingDatapoint:
         mean = params['mean']
         median = params['median']
         stddev = params['stddev']
+        readings = params['readings']
         return cls(
             datapoint_name=datapoint_name,
             mean=mean,
             median=median,
-            stddev=stddev
+            stddev=stddev,
+            readings=readings
             )
 
     def to_dictionary(self):
         return {'datapoint_name': self.datapoint_name,
                 'mean': self.mean,
                 'median': self.median,
-                'stddev': self.stddev
+                'stddev': self.stddev,
+                'readings': self.readings
                 }
 
     @classmethod
-    def sum(cls, a, b):
-        """
-        In our results, mean is the only statistic that has a value, and we only run 1 
-        iteration, so it is not really a mean, it is just a number. Therefore, it makes
-        sense to perform arithmetic on it like this
-        """
[…]

### 2022-04-28 compiler/paper_benchmarks.py
set iteration count to 10

@@ -24,7 +24,7 @@ SERVER_PORT = 42142
 CONNECTION_TIMEOUT = 3000
 MPC_PARTY_SERVER_ID = "0"
 MPC_PARTY_CLIENT_ID = "1"
-NUM_ITERS = 5
+NUM_ITERS = 10
 
 GMW_PROTOCOL = "BooleanGmw"
 BMR_PROTOCOL = "Bmr"
@@ -439,14 +439,14 @@ def get_psi_inputs()-> tuple[list[InputArgs], int]:
 def get_inputs(name: str) -> tuple[list[InputArgs], int]:
     # if name == "biometric":
     #     return get_biometric_inputs()
-    if name == "biometric_fast":
-        return get_biometric_fast_inputs()
+    # if name == "biometric_fast":
+    #     return get_biometric_fast_inputs()
     # if name == "chapterfour_figure_12": # millionaire's problem, not interesting
     #     return get_chapterfour_figure_12_inputs()
     # if name == "convex_hull" or name == "minimal_points":
     #     return get_convex_hull_inputs()
-    # if name == "count_102" or name == "longest_102":
-    #     return get_count_102_inputs()
+    if name == "longest_102": #name == "count_102" or 
+        return get_count_102_inputs()
     # if name == "count_10s":
     #     return get_count_10s_inputs()
     # if name == "count_123":

### 2022-04-28 compiler/paper_benchmarks.py
bug fix (not running vectorized)

@@ -709,7 +709,7 @@ def run_client_role(address):
             outputs = [];
             for protocol in [GMW_PROTOCOL, BMR_PROTOCOL]:
                 for vectorized in [False, True]:
-                    if i > non_vec_up_to:
+                    if i > non_vec_up_to and vectorized is False:
                         pair = (None, None)
                         outputs.append(pair)
                         continue

### 2022-04-28 overleaf/compiler.tex
Edits to Section 4.

@@
 
-\begin{enumerate}
-  \item
-        \textbf{Parsing}:
+%\begin{enumerate}
+%  \item
+        \paragraph{Parsing:}
         Use Python's \texttt{ast} module to parse the input source code to a Python AST.
-  \item
-        \textbf{Syntax checking}:
+%  \item
+        \paragraph{Syntax checking:}
         Ensure that the AST matches the restricted subset defined in Section~\secref{sec:syntax}.
         This step outputs an instance of the \texttt{restricted\_ast.Function} class, which represents our restricted subset of the Python AST.
-  \item
-        \textbf{3-address CFG conversion}:
+%  \item
+        \paragraph{3-address CFG conversion:}
         %\benjamin{TODO: Is this a good amount of detail?} \ana{This is good amount of detail for the Tech report. In the actual paper we'll shorten a lot and add pointers to the Tech report.}
@@
         \end{enumerate}
-  \item
-        \textbf{SSA conversion}:
+  %\item
+        \paragraph{SSA conversion}:
         Convert the 3-address CFG to SSA with Cytron's algorithm.
-\end{enumerate}
+%\end{enumerate}
 
@@
 with the merge condition property.
-For example, if $X_3 = \phi(X_1,X_2)$ is in a block with merge condition $C$,
-the compiler transforms it into $X_3 = \mathit{MUX}(C, X_1, X_2)$.
+For example, if {\sf X!3 = $\phi$(X!1,X!2)} is in a block with merge condition {\sf C},
+the compiler transforms it into {\sf X!3 = MUX(C, X!1, X!2)}.
 Next, the compiler runs the dead code elimination algorithm from Cytron's SSA paper.
@@
 Now, the remaining $\phi$-nodes in MPC source are the loop header nodes. We call these nodes \emph{pseudo} $\phi$-nodes
-and we write \python{PHI} in MPC Source. A pseudo $\phi$-node $X_1 = $\python{PHI}$(X_0,X_2)$ in a loop header is evaluated
-during circuit generation. If it is the 0-th iteration, then the $\phi$-node evaluates to $X_0$, otherwise, it evaluates to $X_2$.
+and we write {\sf PHI} in MPC Source. A pseudo $\phi$-node {\sf X!1 = PHI(X!0,X!2)} in a loop header is evaluated
+during circuit generation. If it is the 0-th iteration, then the $\phi$-node evaluates to {\sf X!0}, otherwise, it evaluates to {\sf X!2}.
 
@@
 % Variable naming
+\paragraph{Variable declarations:}
+Our compiler performs taint analysis on MPC Source based on the type system in~\secref{sec:syntax}. The analysis is described in the following section;
+for the most part, it is a standard taint analysis that propagates that \texttt{shared} qualifiers throughout the routine. 
+Taint analysis assigns \texttt{shared} or \texttt{plain} qualifiers to all local variables, which is important for code generation.
 Our generated C++ uses the following variable-naming scheme: shared variables are named the same as in the MPC Source with the {\sf !} replaced with an underscore (e.g.
 {\sf sum!2} would be translated to {\sf sum\_2}).
-Plaintext variables follow the same naming convention as shared variables but are prefixed with \texttt{\_MPC\_PLAINTEXT\_}.
-The shared representation of constants are named \texttt{\_MPC\_CONSTANT\_} followed by the literal constant (e.g.
-the shared constant \texttt{0} would be named \texttt{\_MPC\_CONSTANT\_0}).
+Plaintext variables follow the same naming convention as shared variables but are prefixed with {\sf \_MPC\_PLAINTEXT\_}.
+The shared representation of constants are named {\sf \_MPC\_CONSTANT\_} followed by the literal constant (e.g.
+the shared constant 0 would be named {\sf \_MPC\_CONSTANT\_0}).
 
@@
 Next, all constant values which are used as part of shared expressions are initialized as a shared input from party 0.
-Finally, all plaintext parameters are converted used as shared inputs from party 0 to initialize their shared counterparts.
+Finally, plaintext parameters are converted used as shared inputs from party 0 to initialize their shared counterparts.
+(Importantly, as we mention, they are converted only if they are needed as an argument to a shared operation; one can think of it as an upcast from \texttt{plain} to \texttt{share}.)
 
 % Non-vectorized assignments and returns
-Once the function preamble is complete, the MPC source is translated into C++ one statement at a time.
-If there is no vectorization present in a statement, translation to C++ is straightforward: outside of Mux statements and array updates, non-vectorized assignments, expressions, and returns directly translated into their C++ equivalents.
-Unvectorized Mux statements are converted to MOTION's Mux member function on the condition variable.
-Array updates are translated into two C++ assignments: one to update the value in the original array and one to assign the new array.
+\paragraph{Code generation:}
+Once the function preamble is complete, the MPC Source is translated into C++ one statement at a time.
+The linear structure of MPC Source enables translation.
+If there is no vectorization present in a statement, translation to C++ is straightforward: outside of MUX statements and array updates, non-vectorized assignments, expressions, and returns directly translate into their C++ equivalents.
+Non-vectorized MUX statements are converted to MOTION's MUX member function on the condition variable.
+Array updates are translated into two C++ assignments: one to update the value in the original array and one to assign the new array as shown in Listing~\ref{tab:motion_translation_array_updates}.
 
@@
 % Phi nodes and FOR loops
-Phi nodes are broken into two components: the ``false'' branch which assigns the initial value of the phi node and the ``true'' branch which assigns the phi node's backedge.
-The assignment of the false branch occurs right before the Phi node's enclosing loop.
-Inside of the Phi node's enclosing loop, a C++ \texttt{if} statement is inserted to only assign the true branch of the phi node after the first iteration.
+Pseudo PHI nodes are broken into two components: the ``FALSE'' branch which assigns the initial value of the PHI node and the ``TRUE'' branch which assigns the PHI node's back-edge.
+The assignment of the false branch occurs right before the PHI node's enclosing loop.
+Inside of the PHI node's enclosing loop, a C++ \texttt{if} statement is inserted to only assign the true branch of the PHI node after the first iteration.
 MPC FOR loops are converted to C++ FOR loops which iterate the loop counter over the specified range.
-The loop counter is initialized before any phi nodes in the loop as they could rely on it; there is no other special handling of FOR loops.
+The loop counter is initialized before any PHI nodes in the loop as they could rely on it; there is no other special handling of FOR loops.
+\ana{Add the reference to the corresponding figure somewhere in this paragraph.}
 
@@
 \begin{tabular}{cc}
-\begin{minipage}{0.5\textwidth}
+\begin{minipage}{0.33\textwidth}
 {\small
@@
 
-\begin{minipage}{0.5\textwidth}
+\begin{minipage}{0.66\textwidth}
 {\small
@@
 All SIMD values are stored in non-vectorized form as 1-dimensional \texttt{std::vector}s in row-major order.
-Whenever a SIMD value is used in an expression, the utility function \texttt{vectorized\_access()} takes the unvectorized representation of a SIMD value, along with the size of each dimension and the requested slice's indices, and converts that slice to a MOTION SIMD value.
-Because MOTION supports SIMD operations using the same C++ operators as non-SIMD operations, we do not need to perform any other transformations to the expression.
-Therefore the translation of an expression containing SIMD values is identical to that of expressions without SIMD values once vectorized accesses are inserted.
+\ana{Does MOTION support just 1-dim SIMD arrays?
+If yes, mention it here.} Whenever a SIMD value is used in an expression, the utility function \texttt{vectorized\_access()} takes the multi-dimentional representation of a SIMD value, along with the size of each dimension and the requested slice's indices, and converts that slice to a MOTION SIMD value.
+\ana{Add the reference to the ADD example figure here.} Because MOTION supports SIMD operations using the same C++ operators as non-SIMD operations, we do not need to perform any other transformations to the expression.
+\ana{Is it correct to say that the operators are overloaded?
+If yes, use that term?} Therefore, the translation of an expression containing SIMD values is identical to that of expressions without SIMD values.
 
@@
 \ben{Is the next sentence needed?  They can always check the implementation of vectorized\_assign() if they want to see specifics.}
-The function takes the array being assigned to as a reference to update it in place so that any existing values are not lost.  SIMD values are never directly assigned to non-SIMD values.
+The function takes the array being assigned to as a reference to update it in place so that any existing values are not lost.
+SIMD values are never directly assigned to non-SIMD values.
+\ana{Yes, this is implementation-specific, you can remove.}
 
@@
 \begin{cppp}
-vectorized_assign(sum_4, {_MPC_PLAINTEXT_N}, {true}, {}, vectorized_access(sum_3, {_MPC_PLAINTEXT_N}, {true}, {}) + vectorized_access(p, {_MPC_PLAINTEXT_N, _MPC_PLAINTEXT_D}, {true, false}, {_MPC_PLAINTEXT_j}));
+vectorized_assign(sum_4, {_MPC_PLAINTEXT_N}, {true}, {}, 
+      vectorized_access(sum_3, {_MPC_PLAINTEXT_N}, {true}, {}) + 
+      vectorized_access(p, {_MPC_PLAINTEXT_N, _MPC_PLAINTEXT_D}, {true, false}, 
+           {_MPC_PLAINTEXT_j}));
 \end{cppp}
@@
 \ben{Technically this is semantically the same as non-vectorized update but we use a utility function to combine the steps into a single C++ statement - should it still be included?}
+\ana{I would leave this in the TR but it will go from the paper. Lots of stuff in Secs. 4 and 5 will go from the paper.} 
 Updating SIMD arrays is also implemented differently from updating non-vectorized arrays.
@@
 % raise_dim() implementation
-\ben{The compiler and C++ internally uses ``lift'' as the terminology for raise\_dim(), but the paper does not :(...}
-Raising the dimensions of an expression uses the \texttt{lift()} utility function which takes a lambda for the raised expression and the dimensions of the output.
+
+\ben{The compiler and C++ internally uses ``lift'' as the terminology for raise\_dim(), but the paper does not :(...} \ana{Yes, that's true.
+To make matters worse, most of the explanation on this comes in the next section.
+But I'll leave as is for now.}
+\paragraph{Reshaping and raising dimensions:}
+Raising the dimensions of a scalar or array uses the \texttt{lift()} utility function which takes a lambda for the raised expression and the dimensions of the output.
 The lambda accepts an array of integers representing the index along each of the dimensions being raised, and the translation of the expression which is being raised replaces each of the dimension index variables with the relevant subscript of this array.
@@
 In this case, the array is simply extended along the new dimensions (if the array is already correctly sized it is returned as is).
+\ana{Add the ref.
+to the figure.
+Also, try to rewrite.
+I understand what the paragraph is saying but it is too technical and does not convey the higher-level goal of these.
+Try first to state what the goal is, and then talk about the details such as the lambda funcs.}
 
@@
 
-\begin{minipage}{0.5\textwidth}
+\begin{minipage}{0.66\textwidth}
 {\small
 \begin{cppp}
-lift(std::function([&](const std::vector<std::uint32_t> &idxs){return idxs[0] + idxs[1];}), {_MPC_PLAINTEXT_N, _MPX_PLAINTEXT_M})
+lift(std::function([&](const std::vector<std::uint32_t > &idxs) {return idxs[0] + idxs[1];}), 
+      {_MPC_PLAINTEXT_N, _MPX_PLAINTEXT_M})
 \end{cppp}
@@
 % drop_dim() implementation
+\paragraph{Dropping dimensions:}
 Dropping dimensions use the \texttt{drop\_dim()} and \texttt{drop\_dim\_monoreturn()} utility functions.
@@
 These functions take the non-vectorized representation of an array, along with the dimensions of that array, and return the array with the final dimension dropped.
-\ben{drop\_dim() taking the true branch of phi node values doesn't need to be covered here, right?}
+\ben{drop\_dim() taking the true branch of phi node values doesn't need to be covered here, right?} \ana{Right, I wouldn't add it here.}
 
@@
 \ben{Ishaq or Vassilis: please check over this paragraph... I don't want to make some mistake about MOTION's capabilities.}
-Currently, our compiler only supports the \texttt{Bmr} and \texttt{BooleanGMW} protocols as MOTION does not implement all operations for other protocols \ben{Should that go here?}.
+Currently, our compiler only supports the \texttt{Bmr} and \texttt{BooleanGMW} protocols as MOTION does not implement all operations for other protocols \ben{Should that go here?} \ana{Yes, I think it's good}.
 MOTION does not support publicly-known constants for these protocols, so all conversions from plaintext values to shares are performed by providing the plaintext value as a shared input from party 0.
 Due to this limitation, our translation to MOTION code attempts to minimize the number of conversions into a plaintext value.
-This is accomplished by creating a shared copy of each plaintext variable and updating that copy in lock-step with the plaintext variable.
+\ana{Typo: conversions from plaintext values, right?} This is accomplished by creating a shared copy of each plaintext variable and updating that copy in lock-step with the plaintext variable.
 Since variables are often initialized to a common constant value (e.g.
-0), this approach decreases the number of input gates by only creating a shared input for each constant.
-Loop counters must still be converted to a shared value on each iteration that they are used, however we only output this conversion when necessary to prevent loops where the counter is not used in a shared expression from increasing the number of input gates.
+0), this approach decreases the number of input gates by only creating a shared input for each initialization constant.
+Loop counters must still be converted to a shared value on each iteration that they are used, however we only generate this conversion when necessary, i.e., when the counter flows to a shared computation.
+This is to prevent unnecessary increase in the number of input gates.
 
@@
 These copies do not incur a runtime cost as the arrays simply hold pointers to the underlying shares, so no new shares or gates are created as a result of this copying.
+Cost in MPC programs is dominated by shares and computation on shares.
 \ben{Is there anything else to say here?
-Should this be placed somewhere else?}
+Should this be placed somewhere else?} \ana{It is clear to me!
+Will stay in the TR but likely will be cut from the paper.}

### 2022-04-28 overleaf/compiler.tex
Respond to motion translation comments

@@
 \ana{Does MOTION support just 1-dim SIMD arrays?
-If yes, mention it here.} Whenever a SIMD value is used in an expression, the utility function \texttt{vectorized\_access()} takes the multi-dimentional representation of a SIMD value, along with the size of each dimension and the requested slice's indices, and converts that slice to a MOTION SIMD value.
+If yes, mention it here.} \ben{MOTION does only support 1-dim SIMD arrays, but since we use helper functions to create and access SIMD arrays this restriction doesn't apply to our representation.
+There is no reason why these arrays need to be 1-dimensional other than that it slightly simplifies the code generation and avoids some nasty templates on the utility functions.} Whenever a SIMD value is used in an expression, the utility function \texttt{vectorized\_access()} takes the multi-dimentional representation of a SIMD value, along with the size of each dimension and the requested slice's indices, and converts that slice to a MOTION SIMD value.
 \ana{Add the reference to the ADD example figure here.} Because MOTION supports SIMD operations using the same C++ operators as non-SIMD operations, we do not need to perform any other transformations to the expression.
 \ana{Is it correct to say that the operators are overloaded?
-If yes, use that term?} Therefore, the translation of an expression containing SIMD values is identical to that of expressions without SIMD values.
+If yes, use that term?} \ben{I don't think that's the right term since SIMD and scalar MOTION values use the same C++ class, so the operator is the same.} Therefore, the translation of an expression containing SIMD values is identical to that of expressions without SIMD values.
 
@@
 This operation cannot be done with a simple subscript as SIMD assignments will update a range of values in the non-vectorized representation.
-\ben{Is the next sentence needed?  They can always check the implementation of vectorized\_assign() if they want to see specifics.}
-The function takes the array being assigned to as a reference to update it in place so that any existing values are not lost.
+%\ben{Is the next sentence needed?  They can always check the implementation of vectorized\_assign() if they want to see specifics.}
+%The function takes the array being assigned to as a reference to update it in place so that any existing values are not lost.
 SIMD values are never directly assigned to non-SIMD values.
@@
 Raising the dimensions of a scalar or array uses the \texttt{lift()} utility function which takes a lambda for the raised expression and the dimensions of the output.
+This function evaluates the expression for each permutation of indices along the dimensions and returns the resulting array in row-major order.
 The lambda accepts an array of integers representing the index along each of the dimensions being raised, and the translation of the expression which is being raised replaces each of the dimension index variables with the relevant subscript of this array.
 There is also a special case of the \texttt{lift()} function which occurs when we are raising an array.
-This case only occurs when the dimensionality of the array is equal to the first dimensions of the array (any dimension which is being lifted to which is not in the array is after every dimension in the array) \ben{I don't know how to phrase the previous sentence more concisely}.
+In this case, instead of concatenating the array for each index, we extend the array along all dimensions being raised which are not present in the array already.
+For example, when raising an array with dimensions $N \times M$ to an array with dimensions $N \times M \times D$, the input array will simply be extended along the $D$ dimension.
+If input array is already correctly sized it will be returned as-is.
+%This case only occurs when the dimensionality of the array is equal to the first dimensions of the array (any dimension which is being lifted to which is not in the array is after every dimension in the array) \ben{I don't know how to phrase the previous sentence more concisely}.
 In this case, the array is simply extended along the new dimensions (if the array is already correctly sized it is returned as is).
@@
 I understand what the paragraph is saying but it is too technical and does not convey the higher-level goal of these.
-Try first to state what the goal is, and then talk about the details such as the lambda funcs.}
+Try first to state what the goal is, and then talk about the details such as the lambda funcs.} \ben{Rewrote the above paragraph to be less confusing/implementation-specific, please review.}
 
@@
 These functions take the non-vectorized representation of an array, along with the dimensions of that array, and return the array with the final dimension dropped.
-\ben{drop\_dim() taking the true branch of phi node values doesn't need to be covered here, right?} \ana{Right, I wouldn't add it here.}
 
@@
 \ben{Ishaq or Vassilis: please check over this paragraph... I don't want to make some mistake about MOTION's capabilities.}
-Currently, our compiler only supports the \texttt{Bmr} and \texttt{BooleanGMW} protocols as MOTION does not implement all operations for other protocols \ben{Should that go here?} \ana{Yes, I think it's good}.
+Currently, our compiler only supports the \texttt{Bmr} and \texttt{BooleanGMW} protocols as MOTION does not implement all operations for other protocols.
 MOTION does not support publicly-known constants for these protocols, so all conversions from plaintext values to shares are performed by providing the plaintext value as a shared input from party 0.
-Due to this limitation, our translation to MOTION code attempts to minimize the number of conversions into a plaintext value.
-\ana{Typo: conversions from plaintext values, right?} This is accomplished by creating a shared copy of each plaintext variable and updating that copy in lock-step with the plaintext variable.
+Due to this limitation, our translation to MOTION code attempts to minimize the number of conversions from a plaintext value.
+This is accomplished by creating a shared copy of each plaintext variable and updating that copy in lock-step with the plaintext variable.
 Since variables are often initialized to a common constant value (e.g.

### 2022-04-28 compiler/paper_benchmarks.py
lan/wan files are read only when needed

@@ -737,7 +737,7 @@ def run_client_role(address):
 
                         if p0 is None or p1 is None:
                             log.error("Run Failed! p0 is None: {} - p1 is None: {}".format(p0 is None, p1 is None))
-                            exit()
+                            continue
 
                         log.info("Output {}".format(p0.output.strip()))
                         assert p0.output.strip() == p1.output.strip(), \
@@ -1028,22 +1028,23 @@ def generate_graphs(lan, wan):
         os.path.isfile(os.path.join(WAN_DIR, f)) and f.endswith('.json')]
     
     lan_stats = []
-    for f in lan_files:
-        with open(f, "r", encoding='utf-8') as f:
-            json_str = f.read()
-            file_stats = json_deserialize(json_str)
-            lan_stats.append(file_stats)
-
     wan_stats = []
-    for f in wan_files:
-        with open(f, "r", encoding='utf-8') as f:
-            json_str = f.read()
-            file_stats = json_deserialize(json_str)
-            wan_stats.append(file_stats)
 
     if lan is True:
+        for f in lan_files:
+            with open(f, "r", encoding='utf-8') as f:
+                json_str = f.read()
+                file_stats = json_deserialize(json_str)
+                lan_stats.append(file_stats)
         generate_single_network_graphs(lan_stats, LAN_GRAPHS_DIR)
+
     if wan is True:
+        for f in wan_files:
+            log.info("Loading: {}".format(f))
+            with open(f, "r", encoding='utf-8') as f:
+                json_str = f.read()
+                file_stats = json_deserialize(json_str)
+                wan_stats.append(file_stats)
         generate_single_network_graphs(wan_stats, WAN_GRAPHS_DIR)
 
     if lan is True and wan is True:

### 2022-04-28 compiler/paper_benchmarks.py
input configuration for final benchmarks

@@ -76,8 +76,9 @@ def get_rand_ints(n, min=1, max=100):
 
 def get_biometric_inputs() -> tuple[list[InputArgs], int]:
     all_args = []
-    non_vec_up_to = 6 # Only run non-vectorized benchmark upto this index
-    for config in [[4, 4]]:#, [4, 8], [4, 16], [4, 32], [4, 64], [4, 128], [4, 256], [4, 512], [4, 1024], [4, 2048], [4, 4096]]:
+    non_vec_up_to = 0 #6# Only run non-vectorized benchmark upto this index
+    # for config in [[4, 4], [4, 8], [4, 16], [4, 32], [4, 64], [4, 128], [4, 256], [4, 512], [4, 1024], [4, 2048], [4, 4096]]:
+    for config in [[4, 128], [4, 4096]]:
         D = config[0]
         N = config[1]
         args = [
@@ -96,8 +97,9 @@ def get_biometric_inputs() -> tuple[list[InputArgs], int]:
 
 def get_biometric_fast_inputs() -> tuple[list[InputArgs], int]:
     all_args = []
-    non_vec_up_to = 6 # Only run non-vectorized benchmark upto this index
-    for config in [[4, 4]]:#, [4, 8], [4, 16], [4, 32], [4, 64], [4, 128], [4, 256], [4, 512], [4, 1024], [4, 2048], [4, 4096]]:
+    non_vec_up_to = 0 #6 # Only run non-vectorized benchmark upto this index
+    # for config in [[4, 4], [4, 8], [4, 16], [4, 32], [4, 64], [4, 128], [4, 256], [4, 512], [4, 1024], [4, 2048], [4, 4096]]:
+    for config in [[4, 128], [4, 4096]]:
         D = config[0]
         N = config[1]
         args = [
@@ -143,8 +145,9 @@ def get_chapterfour_figure_12_inputs() -> tuple[list[InputArgs], int]:
 
 def get_convex_hull_inputs():
     all_args = []
-    non_vec_up_to = 100
-    for N in [4, 8, 16, 32, 64, 128, 256, 512, 1024, 2048, 4096]:
+    non_vec_up_to = 0
+    #for N in [4, 8, 16, 32, 64, 128, 256, 512, 1024, 2048, 4096]:
+    for N in [32, 256]:
         args = [
         "--N", "{}".format(N),
         ]
@@ -166,8 +169,9 @@ def get_convex_hull_inputs():
 
[…]

### 2022-04-28 compiler/paper_benchmarks.py
commenting millionaire's problem

@@ -457,8 +457,8 @@ def get_inputs(name: str) -> tuple[list[InputArgs], int]:
         return get_biometric_inputs()
     if name == "biometric_fast":
         return get_biometric_fast_inputs()
-    if name == "chapterfour_figure_12": # millionaire's problem, not interesting
-        return get_chapterfour_figure_12_inputs()
+    # if name == "chapterfour_figure_12": # millionaire's problem, not interesting
+    #     return get_chapterfour_figure_12_inputs()
     if name == "convex_hull" or name == "minimal_points":
         return get_convex_hull_inputs()
     if name ==  name == "count_102" or "longest_102":

### 2022-04-28 compiler/paper_benchmarks.py
bug fix only gmw non-vec was running

@@ -739,14 +739,14 @@ def run_client_role(address):
                             party1_mpc_addr=mpc_party_client,
                             cmd_args=args.args,
                             benchmark_name=test_case_dir.name,
-                            protocol=GMW_PROTOCOL,
-                            vectorized=False
+                            protocol=protocol,
+                            vectorized=vectorized
                             )
 
                         write_message(server_sock, request)          
                         p1 = run_benchmark_for_party(
                             MPC_PARTY_CLIENT_ID, mpc_party_server, mpc_party_client, test_case_dir.name, 
-                            test_case_dir.path, GMW_PROTOCOL, False, None, args.args
+                            test_case_dir.path, protocol, vectorized, None, args.args
                         )
 
                         p0 = read_message(server_sock)

### 2022-04-28 overleaf/compiler.tex
Edits to MOTION section.

@@
 Our compiler performs taint analysis on MPC Source based on the type system in~\secref{sec:syntax}. The analysis is described in the following section;
-for the most part, it is a standard taint analysis that propagates that \texttt{shared} qualifiers throughout the routine.
+for the most part, it is a standard taint analysis that propagates the declared \texttt{shared} qualifiers throughout the routine.
 Taint analysis assigns \texttt{shared} or \texttt{plain} qualifiers to all local variables, which is important for code generation.
@@
 MPC FOR loops are converted to C++ FOR loops which iterate the loop counter over the specified range.
+Listing~\ref{tab:motion_translation_for_loop} illustrates.
 The loop counter is initialized before any PHI nodes in the loop as they could rely on it; there is no other special handling of FOR loops.
-\ana{Add the reference to the corresponding figure somewhere in this paragraph.}
+%\ana{Add the reference to the corresponding figure somewhere in this paragraph.}
 
@@
 Vectorization is handled with utility functions to manage accessing and updating slices of arrays.
-All SIMD values are stored in non-vectorized form as 1-dimensional \texttt{std::vector}s in row-major order.
-\ana{Does MOTION support just 1-dim SIMD arrays?
+All SIMD values are stored in non-vectorized form as 1-dimensional {\sf std::vector}s in row-major order.
+%\ana{Does MOTION support just 1-dim SIMD arrays?
 If yes, mention it here.} \ben{MOTION does only support 1-dim SIMD arrays, but since we use helper functions to create and access SIMD arrays this restriction doesn't apply to our representation.
-There is no reason why these arrays need to be 1-dimensional other than that it slightly simplifies the code generation and avoids some nasty templates on the utility functions.} Whenever a SIMD value is used in an expression, the utility function \texttt{vectorized\_access()} takes the multi-dimentional representation of a SIMD value, along with the size of each dimension and the requested slice's indices, and converts that slice to a MOTION SIMD value.
-\ana{Add the reference to the ADD example figure here.} Because MOTION supports SIMD operations using the same C++ operators as non-SIMD operations, we do not need to perform any other transformations to the expression.
-\ana{Is it correct to say that the operators are overloaded?
-If yes, use that term?} \ben{I don't think that's the right term since SIMD and scalar MOTION values use the same C++ class, so the operator is the same.} Therefore, the translation of an expression containing SIMD values is identical to that of expressions without SIMD values.
+There is no reason why these arrays need to be 1-dimensional other than that it slightly simplifies the code generation and avoids some nasty templates on the utility functions.} \ana{Ok!
+Leave as is.}
+Whenever a SIMD value is used in an expression, the utility function {\sf vectorized\_access()} takes the multi-dimentional representation of a SIMD value, along with the size of each dimension and the requested slice's indices, and converts that slice to a MOTION SIMD value.
+Because MOTION supports SIMD operations using the same C++ operators as non-SIMD operations, we do not need to perform any other transformations to the expression.
+%\ana{Is it correct to say that the operators are overloaded?
+If yes, use that term?} \ben{I don't think that's the right term since SIMD and scalar MOTION values use the same C++ class, so the operator is the same.} \ana{Ok, leave as is.}
+Therefore, the translation of an expression containing SIMD values is identical to that of expressions without SIMD values.
 
-Similarly, the \texttt{vectorized\_assign()} function assigns a (potentially SIMD) value to a slice of a vectorized array.
+Similarly, the {\sf vectorized\_assign()} function assigns a (potentially SIMD) value to a slice of a vectorized array.
 This operation cannot be done with a simple subscript as SIMD assignments will update a range of values in the non-vectorized representation.
@@
 % vectorized_update()
-\ben{Technically this is semantically the same as non-vectorized update but we use a utility function to combine the steps into a single C++ statement - should it still be included?}
-\ana{I would leave this in the TR but it will go from the paper. Lots of stuff in Secs. 4 and 5 will go from the paper.}
+%\ben{Technically this is semantically the same as non-vectorized update but we use a utility function to combine the steps into a single C++ statement - should it still be included?}
+%\ana{I would leave this in the TR but it will go from the paper. Lots of stuff in Secs. 4 and 5 will go from the paper.} \ana{Leaving in TR, will shorten in paper.}
 Updating SIMD arrays is also implemented differently from updating non-vectorized arrays.
-Instead of separating the array update from the assignment of the new array, these steps are combined with the \texttt{vectorized\_update()} utility function.
-This function operates identically to \texttt{vectorized\_assign()}, however it additionally returns the array after the assignment occurs.
+Instead of separating the array update from the assignment of the new array, these steps are combined with the {\sf vectorized\_update()} utility function.
+This function operates identically to {\sf vectorized\_assign()}, however it additionally returns the array after the assignment occurs.
 This value is then used for the assignment to the new variable.
+Listing~\ref{tab:motion_translation_simd_assignment} illustrates {\sf vectorized\_assign()} and {\sf vectorized\_update()} on the Biometric example.
 
@@
 
-\ben{The compiler and C++ internally uses ``lift'' as the terminology for raise\_dim(), but the paper does not :(...} \ana{Yes, that's true.
+\ben{The compiler and C++ internally uses ``lift'' as the terminology for raise\_dim(), but the paper does not :(...} \ana{True.
 To make matters worse, most of the explanation on this comes in the next section.
-But I'll leave as is for now.}
+But I'll leave as is for now but leave the comments too because it needs work.}
 \paragraph{Reshaping and raising dimensions:}
-Raising the dimensions of a scalar or array uses the \texttt{lift()} utility function which takes a lambda for the raised expression and the dimensions of the output.
+Raising the dimensions of a scalar or array uses the {\sf lift()} utility function which takes a lambda for the raised expression and the dimensions of the output.
 This function evaluates the expression for each permutation of indices along the dimensions and returns the resulting array in row-major order.
 The lambda accepts an array of integers representing the index along each of the dimensions being raised, and the translation of the expression which is being raised replaces each of the dimension index variables with the relevant subscript of this array.
-There is also a special case of the \texttt{lift()} function which occurs when we are raising an array.
+There is also a special case of the {\sf lift()} function which occurs when we are raising an array.
 In this case, instead of concatenating the array for each index, we extend the array along all dimensions being raised which are not present in the array already.
-For example, when raising an array with dimensions $N \times M$ to an array with dimensions $N \times M \times D$, the input array will simply be extended along the $D$ dimension.
-If input array is already correctly sized it will be returned as-is.
+For example, when raising an array with dimensions $N \times M$ to an array with dimensions $N \times M \times D$, the input array will simply be extended along the $D$ dimension: {\sf A'[n,m,d] = A[n,m]} for every $d$.
+If the input array is already correctly sized it will be returned as-is.
 %This case only occurs when the dimensionality of the array is equal to the first dimensions of the array (any dimension which is being lifted to which is not in the array is after every dimension in the array) \ben{I don't know how to phrase the previous sentence more concisely}.
 In this case, the array is simply extended along the new dimensions (if the array is already correctly sized it is returned as is).
-\ana{Add the ref.
+%\ana{Add the ref.
 to the figure.
@@
 I understand what the paragraph is saying but it is too technical and does not convey the higher-level goal of these.
-Try first to state what the goal is, and then talk about the details such as the lambda funcs.} \ben{Rewrote the above paragraph to be less confusing/implementation-specific, please review.}
+Try first to state what the goal is, and then talk about the details such as the lambda funcs.} \ben{Rewrote the above paragraph to be less confusing/implementation-specific, please review.} \ana{Better, removing comment.}
 
@@
 \paragraph{Dropping dimensions:}
-Dropping dimensions use the \texttt{drop\_dim()} and \texttt{drop\_dim\_monoreturn()} utility functions.
+Dropping dimensions use the {\sf drop\_dim()} and {\sf drop\_dim\_monoreturn()} utility functions.
 They function identically but the latter returns a scalar for the case when the final dimension of an array is dropped.
@@
 % Tons of copies != runtime costs
-Due to the SSA translation phase as well as the various conversions to and from SIMD values which our utility functions perform, our generated MOTION code often includes many copies of arrays and scalar values.
-These copies do not incur a runtime cost as the arrays simply hold pointers to the underlying shares, so no new shares or gates are created as a result of this copying.
+Due to the SSA translation phase as well as the conversions to and from SIMD values which our utility functions perform, our generated vectorized MOTION code often includes multiple copies of arrays and scalar values.
+These copies do not incur a runtime cost as the arrays simply hold \emph{pointers} to the underlying shares, so no new shares or gates are created as a result of this copying.
 Cost in MPC programs is dominated by shares and computation on shares.
-\ben{Is there anything else to say here?
+%\ben{Is there anything else to say here?
 Should this be placed somewhere else?} \ana{It is clear to me!

### 2022-04-28 overleaf/compiler.tex
Move 'Code Generation' paragraph header

@@
 % Function preamble
+\paragraph{Code generation:}
 The generated MOTION code begins with the declaration of all variables used in the function, including loop counters.
@@
 % Non-vectorized assignments and returns
-\paragraph{Code generation:}
 Once the function preamble is complete, the MPC Source is translated into C++ one statement at a time.

### 2022-04-29 overleaf/backend.tex
Reorganized sections again...

--- 2022-04-28 overleaf/compiler.tex
@@
-\section{\bf Compiler Framework}
-\label{sec:compiler}
+\section{Compiler Back End}
+\label{sec:backend}
 
-\ana{Convention: use \texttt{shared, plain}, {\sf code}, $i$ and $I$ for counters outside of code.}
+\subsection{Taint Analysis}
+\label{sec:taint}
 
-%We assume arbitrarily nested loops in the MPC-source IR and read-only arrays.
-We assume that loops range from 0 to some constant N.
-Arrays are linearized (row-major order as in MOTION) and accesses are via functions of the induction variables of the enclosing loops.
+We require that all inputs are marked as either shared or plaintext, however, we do not require any other qualifier annotations in the program.
+Then we infer if intermediate variables and expressions are shared through taint analysis with "taintedness" referring to the shared attribute.
+Specifically, the compiler follows the following rules, which are standard in positive-negative qualifier systems (here \texttt{shared} is the positive qualifier and \texttt{plain} is the negative one):
+\begin{enumerate}
+\item Loop counters and array upper bounds are always \texttt{plain}.
+\item If any variable on the right-hand side of an assignment is shared, then the assigned variable is \texttt{shared}.
+%\item If all variables on the right-hand side of an assignment are plaintext, then the assigned variable is plaintext \ana{Will be covered by the last rule.}
+\item Any variables that cannot be determined as shared via the above rules are \texttt{plain}.
+\end{enumerate}
 
-\begin{figure*}
-  \includegraphics[width=0.9\linewidth]{figs_paper_SIMD/compiler_framework.png}
-  \caption{Compiler Framework. \ana{Change figure since we are not doing div-and-conquer.}}
-  \label{fig:compiler_framework}
-\end{figure*}
+%The first two rules are standard for taint analysis, and the third rule follows from the MPC problem statement. \ben{Is the explanation for the third rule correct?}  \ana{Yes, that is exactly it!}
+The final rule is needed in examples like the following. In the below snippet \texttt{sum!2} and \texttt{sum!3} form a dependency cycle and there is no \texttt{shared} value that flows to either one.
+They are inferred as plaintext.
+%cannot be marked as plaintext through simple taint analysis:
+{\small
+\begin{pythonn}
+plaintext_array = [0, 1, 2, ...]
+sum!1 = 0
+for i in range(0, N):
+    sum!2 = PHI(sum!1, sum!3)
+    sum!3 = sum!2 + plaintext_array[i]
+\end{pythonn}
+}
+%\ben{I think the above example is unnecessary and could be replaced by an explanation of how "untainted" variables are implicitly plaintext, but I couldn't think of a way to phrase that.}\ana{I left the example, we can leave it in the TR, remove from paper.}
 
-\figref{fig:compiler_framework} presents an overview of our compiler.
-We start the section with a brief description of the top-level syntax and semantic restrictions.
-We proceed to describe several phases of the compiler:~\secref{sec:imp_to_ssa}
-describes translation from IMP Source to SSA, and~\secref{sec:ssa_to_MPC_Source}
-describes translation form SSA into MPC Source, i.e., the front-end of our compiler.
-Finally,~\secref{sec:MOTION} describes translation of MPC Source, un-optimized and
-optimized into MOTION, i.e., this is the back-end of our compiler.
+When converting to MOTION code, any plaintext value used in the right-hand side of a shared assignment is converted to a shared value for that expression.
+%\ben{Is this necessary to include?} \ana{Yes, that's important.}
 
-Analysis on MPC Source and vectorization comes in a separate section, Section~\secref{sec:vectorization}.
-%\ana{Need to state this more precisely.}
-
-\subsection{Syntax and Semantic Restrictions}
-\label{sec:syntax}
-
-Source syntax is essentially standard IMP syntax as shown below:
-%\begin{figure}[t]
-{\small
-\[
-\begin{array}{l@{~~~~~}l}
-  \begin{array}{l@{~}l@{~~~}l}
- e & ::= e \; \mathit{op} \; e \mid e \mid \x \mid  \texttt{const} \mid \texttt{A[$e$]} & \mathit{expression} \\
- % s &::= t \;\f & \mathit{field} \\
- %\md &::= t\;{\m}(t \;{\this}, \; t\;\x) \; \{\; \ol{t\;\y}\;\s;\;\return\;\y \; \}
- %                 & \mathit{method} \\
-  s & ::= s;s \mid  & \mathit{sequence} \\
-  & \x = e \mid \texttt{A[$e$]} = e \mid & \mathit{assignment \; stmt} \\
-  & \texttt{for $i$ in range($I$)}: s \mid & \mathit{for \; stmt} \\
-  & \texttt{if $e$: $s$ else: $s$} & \mathit{if \; stmt} \\
-%   t &::= q \; {\sf C} & \mathit{qualified \; type}\\
-%   q &::= \high \mid \poly \mid \low & \mathit{FlowCFL \; qualifier}\\
-%    t' &::= q' \; \C & \mathit{field \; qualified \; type}\\
-%   q' &::= \tainted \mid \poly & \mathit{field \; qualifier}\\
-
-  \end{array}
-\end{array}
-\]
-}
-%\end{figure}
-
-The syntax allows for array accesses, arbitrarily nested loops, and if-then-else control flow.
-Expressions are typed $\langle q \; \tau\rangle$, where qualifier $q$ and type $\tau$ are:
-
-{\small
-\[
-\begin{array}{l@{~~~~~}l}
-  \begin{array}{l@{~}l@{~~~}l}
- \tau & ::= \texttt{int}  \mid \texttt{bool} \mid \texttt{list[int]} \mid  \texttt{list[bool]} & \mathit{base \; types} \\
-  q & ::= \texttt{shared} \mid \texttt{plain} & \mathit{qualifiers} \\
-  \end{array}
-\end{array}
-\]
-}
-
-The type system is mostly standard, and in our experience, a sweet spot between readability and expressivity.
-The \texttt{shared} qualifier denotes shared values, i.e., ones shared among the parties and computed upon
-under secure computation protocols; the \texttt{plain} qualifier denotes plaintext values.
-Shared lists denote shared elements, but plaintext length. Subtyping is
-$\texttt{plain} <: \texttt{shared}$, meaning that we can convert a plaintext value into a shared one,
-but not vice versa. Subtyping on qualified types is again as expected, it is covariant in the qualifier
-and invariant in the type: $\langle q_1 \; \tau_1 \rangle <: \langle q_2 \; \tau_2 \rangle$
-iff $q_1 <: q_2$ and  $\tau_1 = \tau_2$.
-
-Our compiler imposes certain semantic restrictions that it enforces throughout the various
-phases of compilation. We note that in some cases, the restrictions
-can be easily lifted and we plan to do so in future iterations of the work.
-
-\begin{enumerate}
-\item Loops are of the form $0 <= i < I$ and bounds are fixed at compile time.
-It is a standard restriction in MPC that the bounds must be known at circuit-generation time.
-\item Arrays are one-dimensional. $N$-dimensional arrays are linearized and accessed
-in row-major order and at this point the programmer is responsible for linearization
-and access. (This restriction can be easily lifted.)
-\item Array subscrpts are plaintext values as specified by the rule:
-\begin{semantics}
-    \ntyperule{Array Access}{
-%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
-  \Gamma \vdash e : \langle \texttt{plain int} \rangle \quad \Gamma \vdash \texttt{A} : \langle q \; \texttt{list}[\tau] \rangle \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
-}{
-  \Gamma \vdash \texttt{A[$e$]} : \langle q \; \tau \rangle
-}
-\end{semantics}
-The subscript $e$ is a function of the indices of the enclosing loops.
-For read access, the compiler allows an arbitrary such function.
-However, it restricts write access to \emph{canonical writes}, i.e., {\sf A[i,j,k] = ...}
-where $i$, $j$ and $k$ loop over the three dimensions of {\sf A}.
-Write access such as for example {\sf A[i,j+2] = ...} is not allowed.
-%\item \ana{Add rules for logical ops}
-%\item \ana{Add rules for arithmetic ops}
-\item The final restriction involves MUX as expressed by the rule:
-\begin{semantics}
-    \ntyperule{MUX}{
-%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
-  \Gamma \vdash e_1 : \langle q_1 \; \texttt{bool} \rangle \quad \Gamma \vdash e_2 : \langle q_2 \; \tau \rangle \quad \Gamma \vdash e_2 : \langle q_2 \; \tau \rangle \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
-}{
-  \Gamma \vdash \texttt{MUX($e_1$,$e_2$,$e_3$)} : \langle q_1 \vee q_2 \vee q_3  \; \tau \rangle
-}
-\end{semantics}
-\end{enumerate}
-Importantly, the arguments of MUX are restricted to base types, as passing array arguments
-is not valid MOTION MPC. \ana{Check this.} This causes a slight inconvenience as we could not write
-\begin{pythonn}
-if e: A[i] = val
-\end{pythonn}
-Instead we had to write
-\begin{pythonn}
-if not(e): val = A[i]
-A[i] = val
-\end{pythonn}
-
-We note that while the programmer writes
-annotations at the level of IMP Source (recall \ana{ref to overview section}), the
-annotations propagate through the transformations and are checked at the
-level of MPC Source.
-
-%\ana{Disadvantage of MUX}
-
-For the rest of this section we write $i,j,k$ to denote the loop nest: $i$ is the outermost loop, $j$, is immediately nested in $i$, and so on until $k$
-and we use $I,J,K$ to denote the corresponding upper bounds. We write $A[i,j,k]$ to denote canonical access
-to an array element. In the program, canonical access is achieved via the standard row-major order formula: $(J*K)*i + K*j + k$.
-To simplify the presentation we describe our algorithms in terms of three-element tuples $i,j,k$, however, discussion easily generalizes to
-arbitrarily large loop nests.
-
-\subsection{From IMP Source to SSA}
-\label{sec:imp_to_ssa}
-
-Our compiler translates from Source to SSA as follows:
-
-%\begin{enumerate}
-%  \item
-        \paragraph{Parsing:}
-        Use Python's \texttt{ast} module to parse the input source code to a Python AST.
-%  \item
-        \paragraph{Syntax checking:}
-        Ensure that the AST matches the restricted subset defined in Section~\secref{sec:syntax}.
-        This step outputs an instance of the \texttt{restricted\_ast.Function} class, which represents our restricted subset of the Python AST.
-%  \item
-        \paragraph{3-address CFG conversion:}
-        %\benjamin{TODO: Is this a good amount of detail?} \ana{This is good amount of detail for the Tech report. In the actual paper we'll shorten a lot and add pointers to the Tech report.}
-
-        Convert the restricted-syntax AST to a three-address control-flow graph.
-        To do this, first,
-        add an empty basic block to the CFG and mark it as current.
-        Next, for each statement in the restricted AST's function body,
-        process the statement.
-        Statements can either be for-loops, if-statements, or assignments (as in~\secref{sec:syntax}).
-        Rules for processing each kind of statement are given below:
-        \begin{enumerate}
-          \item
-                \textbf{For-loops}:
-                Create new basic blocks for
-                the loop condition
-                (the \emph{condition-block}),
-                the loop body
-                (the \emph{body-block}),
-                and the code after the loop
-                (the \emph{after-block}).
-                Insert a jump from the end of the current block to the condition-block.
-                Then, mark the condition-block as the current block.
-                Insert a for-instruction at the end of the current block with the loop counter variable and bounds from the AST.
-                Next, add an edge from the current block to the after-block labeled ``FALSE'' and
-                an edge from the current block to the
-                body-block labeled ``TRUE''.
-                Then, set the body-block to be the current block
-                and process all statements in the AST's loop body.
-                Finally, insert a jump to the condition-block and set the after-block as current.
-          \item
-                \textbf{If-statements}:
-                Create new basic blocks for
-                the ``then'' statements of the if-statement
-                (the \emph{then-block}),
-                the ``else'' statements of the if-statement
-                (the \emph{else-block}),
-                and the code after the if-statement
-                (the \emph{after-block}).
-                At the end of the current block,
-                insert a conditional jump to the then-block or else-block
-                depending on the if-statement condition in the AST.
-                Next,
-                mark the then-block as current,
-                process all then-statements,
-                and add a jump to the after-block.
-                Similarly,
-                mark the else-block as current,
-                process all else-statements,
-                and add a jump to the after-block.
-                Finally, set the after-block to be the current block,
-                and give it a \emph{merge condition} property equal to the condition of the if-statement.
-          \item
-                \textbf{Assignments}:
-                In the restricted-syntax AST,
-                the left-hand side of assignments
-                can be a variable or an array subscript.
-                If it is an array subscript, e.g., {\sf A[i] = x},
-                change the statement to {\sf A = Update(A, i, x)}.
-                If the statement is not already three-address code,
-                for each sub-expression in the right-hand side of the assignment,
-                insert an assignment to a temporary variable.
-        \end{enumerate}
-  %\item
-        \paragraph{SSA conversion}:
-        Convert the 3-address CFG to SSA with Cytron's algorithm.
-%\end{enumerate}
-
-\subsection{From SSA to MPC Source} %: MUX Nodes and Pseudo $\phi$-nodes}
-\label{sec:ssa_to_MPC_Source}
-
-Once the compiler converts the code to SSA,
-it transforms $\phi$-nodes that correspond to if-statements into MUX nodes.
-From the 3-address CFG conversion step,
-$\phi$-nodes corresponding to if-statements will be in a basic block
-with the merge condition property.
-For example, if {\sf X!3 = $\phi$(X!1,X!2)} is in a block with merge condition {\sf C},
-the compiler transforms it into {\sf X!3 = MUX(C, X!1, X!2)}.
-Next, the compiler runs the dead code elimination algorithm from Cytron's SSA paper.
-
-Next, the control-flow graph is \emph{linearized} into MPC Source,
-which has loops but no if-then-else-statements.
-This means that both branches of all if-statements are executed,
-and the MUX nodes determine whether to use results from the then-block or from the else-block.
-The compiler linearizes the control-flow graph with a variation of breadth-first search.
-Blocks with the ``merge condition'' property are only considered
-the second time they are visited,
-since that will be after both branches of the if-statement are visited.
-(The Python AST naturally gives rise to a translation where each conditional has exactly two targets,
-and each ``merge condition'' block has exactly two incoming edges, a TRUE and a FALSE edge.
-Thus, each $\phi$-node has exactly two multiplexer arguments, which dovetails into MUX.
-This is in contrast with Cytron's algorithm which operates at the level of the CFG and allows for
-$\phi$-nodes with multiple arguments.)
-Each time the compiler visits a block,
-it adds the block's statements to the MPC source.
-If the block ends in a for-instruction,
-the compiler recursively converts the body and code after the loop to MPC source
-and adds the for-loop and code after the loop to the main MPC source.
-If the block does not end in a for-instruction,
-the compiler recursively converts all successor branches to MPC source and
-appends these to the main MPC source.
-
-Now, the remaining $\phi$-nodes in MPC source are the loop header nodes. We call these nodes \emph{pseudo} $\phi$-nodes
-and we write {\sf PHI} in MPC Source. A pseudo $\phi$-node {\sf X!1 = PHI(X!0,X!2)} in a loop header is evaluated
-during circuit generation. If it is the 0-th iteration, then the $\phi$-node evaluates to {\sf X!0}, otherwise, it evaluates to {\sf X!2}.
-
-%\ana{We need to double check that the following captures all cases of def-use edges.}
+Importantly, the taint analysis works on MPC Source, which lacks if-then-else control flow.
+This significantly simplifies treatment as there is no need to handle conditionals and implicit flow.
 

### 2022-04-29 overleaf/vectorization.tex
Reorganized sections again...

@@
-\section{Vectorization}
+\section{Backend-Independent Vectorization}
 \label{sec:vectorization}
@@
 \end{itemize}
-
-\subsection{Taint Analysis}
-\label{sec:taint}
-
-We require that all inputs are marked as either shared or plaintext, however, we do not require any other qualifier annotations in the program.
-Then we infer if intermediate variables and expressions are shared through taint analysis with "taintedness" referring to the shared attribute.
-Specifically, the compiler follows the following rules, which are standard in positive-negative qualifier systems (here \texttt{shared} is the positive qualifier and \texttt{plain} is the negative one):
-\begin{enumerate}
-\item Loop counters and array upper bounds are always \texttt{plain}.
-\item If any variable on the right-hand side of an assignment is shared, then the assigned variable is \texttt{shared}.
-%\item If all variables on the right-hand side of an assignment are plaintext, then the assigned variable is plaintext \ana{Will be covered by the last rule.}
-\item Any variables that cannot be determined as shared via the above rules are \texttt{plain}.
-\end{enumerate}
-
-%The first two rules are standard for taint analysis, and the third rule follows from the MPC problem statement. \ben{Is the explanation for the third rule correct?}  \ana{Yes, that is exactly it!}
-The final rule is needed in examples like the following. In the below snippet \texttt{sum!2} and \texttt{sum!3} form a dependency cycle and there is no \texttt{shared} value that flows to either one.
-They are inferred as plaintext.
-%cannot be marked as plaintext through simple taint analysis:
-{\small
-\begin{pythonn}
-plaintext_array = [0, 1, 2, ...]
-sum!1 = 0
-for i in range(0, N):
-    sum!2 = PHI(sum!1, sum!3)
-    sum!3 = sum!2 + plaintext_array[i]
-\end{pythonn}
-}
-%\ben{I think the above example is unnecessary and could be replaced by an explanation of how "untainted" variables are implicitly plaintext, but I couldn't think of a way to phrase that.}\ana{I left the example, we can leave it in the TR, remove from paper.}
-
-When converting to MOTION code, any plaintext value used in the right-hand side of a shared assignment is converted to a shared value for that expression.
-%\ben{Is this necessary to include?} \ana{Yes, that's important.}
-
-Importantly, the taint analysis works on MPC Source, which lacks if-then-else control flow.
-This significantly simplifies treatment as there is no need to handle conditionals and implicit flow.
 

### 2022-04-29 overleaf/compiler.tex
Minor edits in Intro and Front End

@@
-\section{\bf Compiler Framework}
+\section{\bf Compiler Front End}
 \label{sec:compiler}
@@
 \figref{fig:compiler_framework} presents an overview of our compiler.
-We start the section with a brief description of the top-level syntax and semantic restrictions.
-We proceed to describe several phases of the compiler:~\secref{sec:imp_to_ssa}
+We start the section with a brief description of the top-level syntax and semantic restrictions in~\secref{sec:syntax}.
+We proceed to describe the front end of the compiler:~\secref{sec:imp_to_ssa}
 describes translation from IMP Source to SSA, and~\secref{sec:ssa_to_MPC_Source}
-describes translation form SSA into MPC Source, i.e., the front-end of our compiler.
-Finally,~\secref{sec:MOTION} describes translation of MPC Source, un-optimized and
-optimized into MOTION, i.e., this is the back-end of our compiler.
+describes translation form SSA into MPC Source.
 
-Analysis on MPC Source and vectorization comes in a separate section, Section~\secref{sec:vectorization}.
+We describe analysis on MPC Source and backend-independent vectorization in~\secref{sec:vectorization}
+and translation into MOTION code in~\secref{sec:backend}.
 %\ana{Need to state this more precisely.}
@@
 
+\begin{comment}
 \subsection{From (Optimized) MPC Source to MOTION}
@@
 Will stay in the TR but likely will be cut from the paper.}
+\end{comment}

### 2022-04-29 overleaf/introduction.tex
Minor edits in Intro and Front End

@@
 \secref{sec:model} describes our model for cost estimation and argues NP-hardness of optimal scheduling.
-\secref{sec:compiler} details the phases of the compiler and \secref{sec:vectorization} focuses in on vectorization.
+\secref{sec:compiler} details the front-end phases of the compiler, \secref{sec:vectorization} focuses in on backend-independent vectorization, and \secref{sec:backend} describes translation into MOTION.
 \secref{sec:results} presents the experimental evaluation.\secref{sec:related} discusses related work and \secref{sec:conclusion} concludes.

### 2022-04-29 overleaf/biometric_vectorization_example.tex
Did a pass on Section 5.

@@
 
-\subsection{Example: Biometric}
+\subsubsection{Example: Biometric}
 
-We start from MPC Source:
+We now demonstrate the workings of the basic algorithm on our running example. Recall the MPC Source for Biometric:
 
@@
 
-\subsubsection{Phase 1 of Basic Vectorization}
+\paragraph{Phase 1 of Vectorization Algorithm}
 
@@
 
-\subsubsection{Phase 2 of Basic Vectorization}
+\paragraph{Phase 2 of Vectorization Algorithm}
 
@@
 
-\subsubsection{Phase 3 of Basic Vectorization}
+\paragraph{Phase 3 of Vectorization Algorithm}
 
@@
 
-// i is redundant for min_sum!2, min_sum!4 but not for !12![i]!
+// i is redundant for min_sum!2, min_sum!4 but not for !12![i]
 for i in range(0, N!0):
@@
 
-\subsubsection{Phase 4 of Basic Vectorization}
+\paragraph{Phase 4 of Basic Vectorization}
 
-And this phase adds SIMD operations:
+This phase adds SIMD operations:
 

### 2022-04-29 overleaf/vectorization.tex
Did a pass on Section 5.

@@
 
-\ana{Need an intro here!}
+\ana{Need an intro/overview here!}
+\ana{Need to define notion of canonical access!}
+\ana{Talk about MPC Source, opportunities for vectorization, algorithm unique to MPC Source.}
 
@@
 
+\subsubsection{Overview}
+
 An important component of our algorithm is the scalar expansion to the corresponding loop dimensionality.
@@
 
-There are three kinds of arrays (for now all kept internally as one-dimensional arrays, but that's under discussion).
+There are three kinds of arrays.
 \begin{itemize}
 
-\item Scalars: These are scalar variables we lift into arrays for the purposes of vectorization.
-For those, all writes are canonical writes and all reads are canonical reads. We will apply raise dimension when
-a scalar is used in an inner loop (e.g., {\sf sum!2} in line 6 of the MPC source code will be raised to a 1-dimensional
-array since {\sf sum!2} is used in the inner $j$-loop). Drop dimension applies as well; this happens when a scalar
-written in an inner loop is used outside of the loop (e.g., {\sf sum!2} for which the lifted inner loop computes
+\item Scalars: These are scalar variables we expand into arrays for the purposes of vectorization.
+For those, all writes are canonical writes and all reads are canonical reads. We will \emph{raise dimension} when
+a scalar gives rise to an outer-to-inner dependence edge (e.g., {\sf sum!2} in line 6 of the MPC source code will be raised to a 1-dimensional
+array since {\sf sum!2} is used in the inner $j$-loop). We will \emph{drop dimension} when a scalar
+gives rise to an inner-to-outer dependence edge (e.g., {\sf sum!2} for which the lifted inner loop computes
 {\sf D} values, but the outer loop only needs the last one.)
 
-\item Read-only input arrays: Read-only inputs. There are NO writes, while we may have non-canonical reads, $f(i,j,k)$.
-Phase 1 of Basic vectorization will add raise dimension operation at the beginning of the function to lift these arrays,
-and raise dimension may \emph{reshape} arrays. If there are multiple "views" of the input array, there would be multiple raise dimension statements to create
+\item Read-only input arrays: There are no writes, while we may have non-canonical reads, $f(i,j,k)$.
+Vectorization adds raise dimension operations at the beginning of the function to lift these arrays to 
+the dimensionality of the loop where they are used, possibly \emph{reshaping} the arrays. 
+If there are multiple ``views'' of the input array, there would be multiple raise dimension statements to create
 each one of these views. The invariant is that at reads in loops, the reads of "views" of the original input array are canonical.
-Only raise dimension applies to these arrays, and only in the beginning of the program. For example, 1-dimensional array
-\python{C} is lifted into 2-dimensional array {\sf N x D} by copying the row {\sf N} times.
+%Only raise dimension applies to these arrays, and only in the beginning of the program. For example, 1-dimensional array
+%\python{C} is lifted into 2-dimensional array {\sf N x D} by copying the row {\sf N} times.
 
  \item Read-write output arrays: Writes are canonical (by restriction) but reads can be non-canonical.
- Dependence analysis limits vectorization when non-canonical read access refers to array writes in
+ Dependences limit vectorization when non-canonical read access refers to array writes in
  previous iterations, thus creating loop-carried dependences. We may apply both raise and drop dimension, however,
  they respect the fixed dimensionality of the output array. The array cannot be raised to a dimension lower than
- its canonical (fixed) dimensionality and it cannot be dropped lower. In addition, non-canonical reads may
- require lifting (i.e., reshaping) of the array after the most recent write, rather than in the beginning of the program,
+ its canonical (fixed) dimensionality and it cannot be dropped to lower dimension. In addition, non-canonical reads may
+ require lifting (i.e., reshaping) of the array after the most recent write rather than in the beginning of the program
  in order to reduce a non-canonical read to a canonical one.
@@
 
-\subsection{Raise Dimension and Drop Dimension}
+\subsubsection{Raise Dimension and Drop Dimension}
 
-There are two conceptual versions of $\mathit{raise\_dim}$. One applies on read-only input arrays and reshapes those arrays
-when necessary to ensure canonical read access in the corresponding loop. The signature of $\mathit{raise\_dim}$
-is as follows. It takes the original array \texttt{C}, the access pattern function $f(i,j,k)$ in loop nest $i,j,k$ and the loop bounds
+There are two conceptual versions of $\mathit{raise\_dim}$. One reshapes an arbitrary access into a canonical read access in the corresponding loop. 
+%The signature of $\mathit{raise\_dim}$ is as follows. 
+It takes the original array, the access pattern function $f(i,j,k)$ in loop nest $i,j,k$ and the loop bounds
 $((i:I),(j:J),(k:K))$:
-\[ \mathit{raise\_dim}(\texttt{A}, f(i,j,k), \texttt{((i:I),(j:J),(k:K))}) \]
-It produces a new 3-dimensional array \texttt{A'} by iterating over $i,j,k$ and setting each element of  \texttt{A'} as follows:
-\[ \texttt{A'[i,j,k] =  A[f(i,j,k)]} \]
-The end result is that uses of \texttt{[A[f(i,jk)]} in loop nest $i,j,k$ are replaced with canonical read-accesses to \texttt{A'[i,j,k]}
-that can be vectorized. In the running Biometric example, \texttt{C'} = $\mathit{raise\_dim}$\texttt{(C, j, (i:N,j:D))} lifts the
-1-dimensional array \texttt{C} into a 2-dimensional array. The $i,j$ loop now accesses \texttt{C'} in the canonical way, \texttt{C'[i,j]}.
-Similarly, \texttt{S'} = $\mathit{raise\_dim}$\texttt{(S, i*D+j, (i:N,j:D))} tries to lift \texttt{S}, but the operation turns into a no-op
-because \texttt{S} is already a 2-dimensional array and the read access is canonical.
+\[ \mathit{raise\_dim}({\sf A, f(i,j,k), ((i:I),(j:J),(k:K))}) \]
+It produces a new 3-dimensional array {\sf A'} by iterating over $i,j,k$ and setting each element of  {\sf A'} as follows:
+\[ {\sf A'[i,j,k] =  A[f(i,j,k)]} \]
+The end result is that uses of {\sf [A[f(i,j,k)]} in loop nest $i,j,k$ are replaced with canonical read-accesses to {\sf A'[i,j,k]}
+that can be vectorized. In the running Biometric example, {\sf C'} = $\mathit{raise\_dim}${\sf (C, j, (i:N,j:D))} lifts the
+1-dimensional array {\sf C} into a 2-dimensional array. The $i,j$ loop now accesses {\sf C'} in the canonical way, {\sf C'[i,j]}.
+Similarly, {\sf S'} = $\mathit{raise\_dim}${\sf (S, i*D+j, (i:N,j:D))} tries to lift {\sf S}, but the operation turns into a no-op
+because {\sf S} is already a 2-dimensional array and the read access is canonical.
 
-The other version of $\mathit{raise\_dim}$ applies on scalars and read-write arrays. It lifts a lower-dimension array into a
-higher-dimension for access in a nested loop. Here \texttt{A} is an $i$ array and raise dimension adds two additional dimensions:
-\[ \mathit{raise\_dim}(\texttt{A}, \texttt{(j:J,k:K)}) \]
+The other version of $\mathit{raise\_dim}$ lifts a lower-dimension array into a
+higher-dimension for access in a nested loop. It is necessary when processing outer-to-inner dependences. 
+Here {\sf A} is an $i$-array and raise dimension adds two additional dimensions:
+\[ \mathit{raise\_dim}({\sf A}, {\sf (j:J,k:K)}) \]
 This version is reduced to the above version by adding the access pattern function, which is just $i$:
-\[ \mathit{raise\_dim}(\texttt{A}, i, \texttt{(j:J,k:K)}) \]
+\[ \mathit{raise\_dim}({\sf A, i, (j:J,k:K)}) \]
 
@@
 It takes a higher dimensional array, say $i,j,k$ and removes trailing dimensions, say $j,k$:
-\[ \mathit{drop\_dim}(\texttt{A}, \texttt{(j:J,k:K)}) \]
+\[ \mathit{drop\_dim}({\sf A, (j:J,k:K)}) \]
 It iterates over $i$ and takes the result at the maximal index of $j$ and $k$, i.e., the result at the last iterations of $j$ and $k$:
-\[ \texttt{A'[i] = A[i,J-1,K-1]} \]
+\[ {\sf A'[i] = A[i,J-1,K-1]} \]
 
 \subsection{Basic Vectorization}
+
+We present our vectorization algorithm in detail followed by a detailed example on Biometric Matching.
+
+\subsubsection{Algorithm}
+
+There are two key phases of the algorithm. Phase 1 applies dependence analysis and inserts
+raise dimension and drop dimension operations as necessary. After this phase operations work
+on arrays of the corresponding dimensionality and we optimistically vectorize all arrays.
+
+Phase 2 takes into account dependence analysis. It proceeds from the inner-most towards the outer-most 
+loop. For each loop it anchors dependence cycles (closures) around pseudo PHI nodes, removing vectorization 
+from the dimension of the loop.
+
+There are several important points in this phase. We may break a loop into smaller loops which would discover opportunities 
+for optimization in intermediate statements in the loop. We remove redundant back edges into PHI nodes that are 
+due entirely to arrays, which opens up opportunities as well. The pseudo-code in red deals with this extension
+of the basic algorithm. We determine that some array pseudo-PHIs are \emph{target-less}, meaning
+that there is no loop-carried dependence for that array and back-edges into these nodes can be safely removed. 
+This opens up opportunities for vectorization across the write array. Handling of target-less nodes is non-trivial, 
+as in certain cases the node cannot be removed. \secref{sec:extension_with_writes} details our algorithm.
+
+\ana{Add an example from histogram here.}
+
+Phases 3 and 4 are less important --- phase 3 cleans up local arrays of references and phase 4 turns operations into MOTION operations.
 
@@
 \STATE {change ${I_d}$ to $i_d$ in all statements in loop}
-\STATE {treat FOR loop as monolith node: some def-use edges become same-level.}
+\STATE {treat FOR loop as monolith node for def-uses: e.g., some def-use edges become same-level.}
 \ENDFOR
@@
 
-We build a correctness argument that loosely follows the theory of Abstract Interpretation.
-We define the syntax of MPC Source programs. The domain of MPC Source programs
+We build a correctness argument that loosely follows Abstract Interpretation.
+First we define the MPC Source syntax. The domain of MPC Source programs
 expressible in the syntax (with certain semantic restrictions) is the abstract domain $A$.
 We then define the \emph{linearization} of an MPC Source program as an interpretation
-over the syntax. The linearization, which is a \emph{schedule}, is the concrete domain $C$.
+over the syntax. The linearization, which is a \emph{schedule} (as in~\secref{sec:overview}), is the concrete domain $C$.
 Since we reason over def-use graphs in $A$ we define a partial order relation over elements of $A$
@@
 
+\begin{comment}
 \begin{figure}%[!b]
@@
      & \mid \x[i,J] = \mathit{drop\_dim}(\x'[i,J,k],k)  & \mathit{drop} \; \mathit{dimension(s)} \\
-     & \mid \x = \y & \mathit{propagation} \\
+%     & \mid \x = \y & \mathit{propagation} \\
      & \mid \texttt{FOR} \; 0 \le i < I: s  & \mathit{loop} \\
@@
 \vspace{-3mm}
-\caption{MPC Source Syntax}\label{fig:MPC_Source_syntax}
+\caption{MPC Source Syntax. \ana{Add more to caption.}}\label{fig:MPC_Source_syntax}
 \vspace{-5mm}
 \end{figure}
+\end{comment}
+
+\begin{figure*}%[!b]
+\[
+\begin{array}{l@{~~~~~}l}
+  \begin{array}{l@{~~~~~~~~~}l@{~~~~~~~~~}l@{~~~~~~~~~}l}
+  s &::= s_1;s_2 & \gamma(s) = \gamma(s_1) \; ; \; \gamma(s_2) & \mathit{sequence} \\
+     & \mid \x[i,J,k] = {\sf op\_SIMD}(\y[i,J,k], \z[i,J,k]) & \gamma(\x[i,J,k] = {\sf op\_SIMD}(\y[i,J,k], \z[i,J,k])) = & \mathit{operation} \\ 
+     &                                                                               & ~~\x[i,0,k] = \y[i,0,k] \; {\sf op} \; \z[i,0,k] \;\; || &  \\
+     &                                                                               & ~~~~\x[i,1,k] = \y[i,1,k] \; {\sf op} \; \z[i,1,k] \;\; || \; ...\; || & \\
+     &                                                                               & ~~~~~~\x[i,J\!-\!1,k] = \y[i,J\!-\!1,k] \; {\sf op} \; \z[i,J\!-\!1,k] & \\
+     & \mid \x[i,J,k] = {\sf const} & \mbox{analogous} & \mathit{constant} \\
+     & \mid \x[i,J,k]  = {\sf PHI}(\x_1[i,J,k],\x_2[i,J,k\!-\!1]) & & \mathit{pseudo} \; \mathit{PHI} \\
+     & \mid \x[i,J,k] = {\sf raise\_dim}(\x'[i],(J\!:\!J,k\!:\!K)) & & \mathit{raise} \; \mathit{dimension(s)} \\
+     & \mid \x[i,J] = {\sf drop\_dim}(\x'[i,J,k],k) &  & \mathit{drop} \; \mathit{dimension(s)} \\
+%     & \mid \x = \y & & \mathit{propagation} \\
+     & \mid {\sf for} \; i \; {\sf in} \; {\sf range}(I): s  & \gamma({\sf for} \; i \; {\sf in} \; {\sf range}(I): s) = & \mathit{loop} \\
+     & & ~~\gamma(s)[0/i] \; ; \; \gamma(s)[1/i] \; ; \; ... \; ; \; \gamma(s)[I\!-\!1/i] & \\
+  \end{array}    \\
+\end{array}
+\]
+\vspace{-3mm}
+\caption{MPC Source Syntax and Semantics.
+$\gamma$ defines the semantics of MPC source which is a linearization of MPC Source.
+A SIMD operation parallelizes operations across the vectorized $J$ dimension.
+$||$ denotes parallel execution, which is standard.
+$\gamma$ of a for loop unrolls the loop.
+$;$ denotes sequential execution.
+Iterative MPC Source trivially extends to non-vectorized dimensions over the enclosing loops.}\label{fig:MPC_Source_syntax}
+\label{fig:mpc_syntax}
+\vspace{-5mm}
+\end{figure*}
 
@@
 
+\figref{fig:mpc_syntax} states the syntax and linearization semantics of MPC Source. For simiplicity,
+we consider only scalars and read-only arrays, however, the treatment extends to write arrays as well.
+$\x[i,J,k]$ denotes the value of scalar variable \x{} at loop nest $i,j,k$. 
+%(We use 3 dimensions, $i,j,k$, however, the treatment generalizes to arbitrary dimensions.) 
+Upper case $J$ denotes a vectorized dimension and lower case $i,k$ denote iterative dimensions. Our compiler imposes 
+semantic restrictions over the syntax: (1) $\x$ is treated as a 3-dimensional array and (2) $\x[i,J,k]$ must be enclosed
+into for-loops on non-vectorized dimensions $i$ and $k$:
+\begin{pythonn}
+for i in range(I):
+   ...
+   for k in range(K):
+      ... x[i,J,k] ...
+\end{pythonn}
+
+\begin{comment}
 \figref{fig:MPC_Source_syntax} defines the syntax for our intermediate representation, MPC Source.
-There are semantics restrictions over the syntax as well: a variable $\x[i,j,k]$ is a 3-dimensional array
-$(i:I,j:J,k:K)$ and also, a statement $\x[i,J,k] = ...$ is enclosed in loops over $i$ and $k$ as shown below.
-Thus, $i$ and $k$ are in scope.
-\begin{verbatim}
-FOR 0 <= i < I:
+Note that this is a generalization of MPC Source to include vectorized MPC Source.
+There are semantic restrictions over the syntax as well: a variable $\x[i,j,k]$ is a 3-dimensional array
+$(i:I,j:J,k:K)$ and also, a statement $\x[i,J,k] = ...$ is enclosed in loops over the non-vectorized dimensions 
+$i$ and $k$ as shown below. Thus, $i$ and $k$ are in scope.
+\begin{pythonn}
+for i in range(I):
     ...
-    FOR 0 <= k < K:
+    for k in range(K):
         x[i,J,k] = ...
-\end{verbatim}
+\end{pythonn}
 
-Statements $\mathit{operation, phi, raise \; dimension(s), drop \; dimension(s)}$ are \emph{base} statements,
-and $\mathit{sequence, loop}$ are \emph{compound statements}.
+%Statements $\mathit{operation, phi, raise \; dimension(s), drop \; dimension(s)}$ are \emph{base} statements,
+%and $\mathit{sequence, loop}$ are \emph{compound statements}.
 
@@
 $\gamma$ simply unrolls the loop substituting $i$ with 0, 1, etc. Here ; denotes sequential execution.
+\end{comment}
 
+\begin{comment}
 \paragraph{Partial Orders}
@@
 \ana{Write the proof sketch and final argument, etc.}
+\end{comment}
 
-\subsection{Extension of Basic Vectorization with Array Writes}
+\paragraph{Partial Orders}
+
+For each MPC Source program $a$ we compute the def-use edges in the standard way: if base statement $s1 \in a$ defines variable \x, e.g., $\x[i,j,k] = ...$, 
+and base statement $s2 \in a$ uses \x, e.g., $... = ... \x[i,j,k]$ and there is a path in CFG from $s1$ to $s2$, then there is a def-use edge from
+ $s1$ to $s2$. We extend the dimensionality of a statement into $s1[i,j,k]$ where $s1[i,j,k]$ inherits the dimensionality of the left-hand-side of the assignment.
+
+ \ana{Add on the linearization semantics?}
+
+ Let $a_0, a_1$ be two MPC Source programs in $A$. 
+ Two statements, $s \in a_0$ and $s' \in a_1$ are \emph{same}, written $s \equiv s'$ if they are of the same operation and they operate on 
+ the same variables: same variable name and same dimensionality. Recall that dimensions in MPC Source are either iterative (lower case), or vectorized (upper case). 
+ Two statements are same even if one operates on an iterative dimension and the other one operates on a vectorized one, e.g., $s[i,j,k] \equiv s'[I,j,K]$.
+ We extend the definition to def-use edges in the obvious way: a def-use edge $s_0 \rightarrow s_1$ in $a_0$ and an edge $s'_0 \rightarrow s'_1$ in $a_1$
+ are \emph{same}, written $s_0 \rightarrow s_1 \equiv s'_0 \rightarrow s'_1$, if and only if $s_0 \equiv s'_0$, $s_1 \equiv s'_1$, and the two edges are both either 
+ forward or backward.
+
+\begin{definition} Let $a_0, a_1 \in A$. We say that $a_0 \le a_1$ iff for every def-use edge $e$ in $a_0$ there is an edge
+$e'$ in $a_1$ such that $e \equiv e'$.
+\end{definition}
+
+ The def-use edges in the concrete schedule are as expected: there is a def-use edge from statement $s1$ that defines $\x[\underline{i},\underline{j},\underline{k}]$
+ to statement $s2$ that uses $\x[\underline{i},\underline{j},\underline{k}]$ if $s1$ is scheduled ahead of $s2$ in the linear schedule. We note that the underlined
+ indices, e.g., $\underline{i}$, refer to fixed values, not iterative or vectorized dimensions since in the concrete schedule all induction variables are expanded. 
+ E.g., there is a def-use edge from the statement that defines $\x[0,1,2]$ and a statement that uses $\x[0,1,2]$.
+
+\paragraph{Theorems} The two theorems arising from the Basic vectorization optimization are as follows:
+
+\begin{theorem} $a_0 \le a_1 \Rightarrow \gamma(a_0) \subseteq \gamma(a_1)$.
+\end{theorem}
+
+\begin{theorem} Let $a_0$ be the iterative MPC Source and let $a_1$ be the vectorized MPC Source computed by Basic vectorization. 
+We have that $a_0 \le a_1$. 
+\end{theorem}
+
+Theorem 1 states that a def-use edge in MPC Source $a_0$ gives rise to a set of def-use edges in $\gamma(a_0)$
+and a ``same'' edge in $a_1$ gives rise to a set of def-use edges in $\gamma(a_1)$ and the former set is included in the latter one.
+This means that the computation of $\gamma(a_0)$ is carried out by $\gamma(a_1)$ as well, computing the same result. 
+The theorem allows that $a_1$ and $\gamma(a_1)$ have larger state that $a_0$ and $\gamma(a_0)$, i.e., doing additional 
+computation into additional memory locations. Theorem 2 establishes the correctness of vectorization --- 
+clearly, vectorization preserves all statements and def-use edges in the original MPC Source, and therefore the linearization computes the 
+same result as the original MPC Source. (In this case we have an equivalence, $a_0 = a_1$ according to our partial order.)
+
+\subsection{Extension with Array Writes}
+\label{sec:extension_with_writes}
 
@@
 {\small
-\begin{verbatim}
+\begin{pythonn}
 for i in range(N):
@@
   D[i] = B[i] * C[i];
-\end{verbatim}
+\end{pythonn}
 }
@@
 {\small
-\begin{verbatim}
+\begin{pythonn}
 for i in range(N):
-1. A_1 = PHI(A_0,A_2)
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. A_2 = update(A_1, i, B_1[i] + 10);
-6. B_2 = update(B_1, i, A_2[i] * D_1[i-1]);
-7. C_2 = update(C_1, i, A_2[i] * D_1[i-1]);
-8. D_2 = update(D_1, i, B_2[i] * C_2[i]);
-\end{verbatim}
+  A_1 = PHI(A_0,A_2)
+  B_1 = PHI(B_0,B_2)
+  C_1 = PHI(C_0,C_2)
+  D_1 = PHI(D_0,D_2)
+  A_2 = update(A_1, i, B_1[i] + 10);
+  B_2 = update(B_1, i, A_2[i] * D_1[i-1]);
+  C_2 = update(C_1, i, A_2[i] * D_1[i-1]);
+  D_2 = update(D_1, i, B_2[i] * C_2[i]);
+\end{pythonn}
 }
 
-There is a cycle around \texttt{B\_1 = PHI(B\_0,B\_2)} that includes statement
-\texttt{A\_1 = update(A\_0, i, B\_1[i] + 10)} and that statement won't be vectorized even
-though in fact there is no loop-carried dependency from the write of \texttt{B\_1[i]} at 6 to the
-read of \texttt{ ... = B\_1[i]} at 8.
+There is a cycle around {\sf B\_1 = PHI(B\_0,B\_2)} that includes statement
+{\sf A\_1 = update(A\_0, i, B\_1[i] + 10)} and that statement won't be vectorized even
+though in fact there is no loop-carried dependency from the write of {\sf B\_1[i]} at 7 to the
+read of {\sf ... = B\_1[i]} at 6.
 
@@
 \begin{algorithmic}
-\FOR {each array $\texttt{A}$ written in loop $j$}
+\FOR {each array ${\sf A}$ written in loop $j$}
 \STATE \COMMENT { including enclosed loops in $j$ }
 \STATE dep = False
-\FOR {each pair def: $\texttt{A}_m[f(i,j,k)] = ... $, and use: $ ... = \texttt{A}_n[f'(i,j,k)]$ in loop $j$}
+\FOR {each pair def: ${\sf A}_m[f(i,j,k)] = ... $, and use: $ ... = {\sf A}_n[f'(i,j,k)]$ in loop $j$}
 \IF {$\exists \underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$, s.t.
@@
 \IF {dep == False}
-\STATE remove back edge into $\texttt{A}$'s $\phi$-node in loop $j$.
+\STATE remove back edge into ${\sf A}$'s $\phi$-node in loop $j$.
 \ENDIF
@@
 
-Consider a loop $j$ enclosed in some fixed $\underline{i}$. Only if an update (definition) $\texttt{A}_m[f(i,j,k)] = ... $ at some iteration $\underline{j}$
-references the \emph{same} array element as a use $ ... = \texttt{A}_n[f'(i,j,k)]$ at some later iteration $\underline{j}'$,
-we may have a loop-carried dependence for $\texttt{A}$ due to this def-use pair. (In contrast, Cytron's algorithm inserts a loop-carried dependency every time there is an array update.)
+Consider a loop $j$ enclosed in some fixed $\underline{i}$. Only if an update (definition) ${\sf A}_m[f(i,j,k)] = ... $ at some iteration $\underline{j}$
+references the \emph{same} array element as a use $ ... = {\sf A}_n[f'(i,j,k)]$ at some later iteration $\underline{j}'$,
+we may have a loop-carried dependence for ${\sf A}$ due to this def-use pair. (In contrast, Cytron's algorithm inserts a loop-carried dependency every time there is an array update.)
 The algorithm above examines all def-use pairs in loop $j$, including defs and uses in nested loops, searching for values $\underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$ that satisfy
 $f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$. If such values exist for some def-use pair, then there is a potential
-loop-carried dependence on $\texttt{A}$; otherwise there is not and we can remove the spurious backward edge thus ``freeing up'' statements for vectorization. %\ana{Double check this!!!}
+loop-carried dependence on ${\sf A}$; otherwise there is not and we can remove the spurious backward edge thus ``freeing up'' statements for vectorization. %\ana{Double check this!!!}
 
@@
 Clearly, there is no pair $\underline{i}$ and $\underline{i}'$, where $\underline{i} < \underline{i}'$ that make $\underline{i} = \underline{i}'$
-due to the def-use pairs of \texttt{A} 5-6 and 5-7.
-Therefore, we remove the back edge from 5 to the phi-node 1.
-Analogously, we remove the back edges from 6 to 2 and from 7 to 3.
-However, there are many values $\underline{i} < \underline{i'}$ that make $\underline{i} = \underline{i'}-1$ and the back edge from 8 to 4 remains (def-use pairs for \texttt{D}).
-As a result of removing these spurious edges, Basic Vectorization will find that statement 5 is vectorizable.
-Statements 6, 7 and 8 will correctly appear in the FOR loop.
+due to the def-use pairs of {\sf A} 6-7 and 6-8.
+Therefore, we remove the back edge from 6 to the phi-node 2.
+Analogously, we remove the back edges from 7 to 3 and from 8 to 4.
+However, there are many values $\underline{i} < \underline{i'}$ that make $\underline{i} = \underline{i'}-1$ and the back edge from 9 to 5 remains (def-use pairs for {\sf D}).
+As a result of removing these spurious edges, Vectorization will find that statement 6 is vectorizable.
+Statements 7, 8 and 9 will correctly appear in the FOR loop.
 
-Note however, that this step renders some array phi-nodes target-less. We handle target-less phi-nodes with a minor extension of Basic Vectorization (Phase 2).
+Note however, that this step renders some array phi-nodes target-less. We handle target-less phi-nodes with a minor extension of Vectorization (Phase 2).
 First, we merge closures that update the same array. This simplifies handling of array $\phi$-nodes: if each closure is turned into a separate loop
 each loop will need to have its own array phi-node to account for the update and this would complicate the analysis.
-Second, we add the target-less node of array \texttt{A} back to the closure that updates \texttt{A} ---
-the intuition is, even if there is no loop-carried dependence from writes to reads on \texttt{A}, \texttt{A} is written and the write (i.e., update) cannot be vectorized;
+Second, we add the target-less node of array {\sf A} back to the closure that updates {\sf A} ---
+the intuition is, even if there is no loop-carried dependence from writes to reads on {\sf A}, {\sf A} is written and the write (i.e., update) cannot be vectorized;
 therefore, the updated array has to carry to the next iteration of the loop. Third, in cases when the phi-node remains target-less, i.e., cases when the
@@
 
-For now, we restrict array updates to \emph{canonical updates}. Assume (for simplicity) a two-dimensional array \texttt{A[I,J]}. A canonical update is the following:
-\begin{verbatim}
+For now, we restrict array updates to \emph{canonical updates}. Assume (for simplicity) a two-dimensional array {\sf A[I,J]}. A canonical update is the following:
+\begin{pythonn}
 for i in range(I):
@@
     ...
-\end{verbatim}
+\end{pythonn}
 
-The update \texttt{A[i,j]} can be nested into an inner loop and there may be multiple updates, i.e., writes to \texttt{A[i,j]}.
-However, update such as \texttt{A[i-1,j] = ...} or \texttt{A[i-1,j-1] = ...}, etc., is not allowed.
-Additionally, while there could be several different loops that perform canonical updates, they must be of the same dimensionality, i.e., an update of higher or lower dimension, e.g., \texttt{A[i,j,k] = ...} is not allowed.
+The update {\sf A[i,j]} can be nested into an inner loop and there may be multiple updates, i.e., writes to {\sf A[i,j]}.
+However, update such as {\sf A[i-1,j] = ...} or {\sf A[i-1,j-1] = ...}, etc., is not allowed.
+Additionally, while there could be several different loops that perform canonical updates, they must be of the same dimensionality, i.e., an update of higher or lower dimension, e.g., {\sf A[i,j,k] = ...} is not allowed.
 We compute the \emph{canonical dimensionality} of each write array by examining the array writes in the original program and rejecting programs that violate the canonical write restriction.
@@
 
-Reads through an arbitrary formula, such as \texttt{A[i-1]} for example, are allowed; currently, the projection function returns dummy values if the read formula is out of bounds; we assume the programmer ensures that the program still computes correct output in this case.
+Reads through an arbitrary formula, such as {\sf A[i-1]} for example, are allowed; currently, the projection function returns dummy values if the read formula is out of bounds; we assume the programmer ensures that the program still computes correct output in this case.
 
@@
 In addition to the changes for the handling of target-less phi-nodes, Basic Vectorization has to handle def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable.
-The definition can be an update \texttt{A\_2 = update(A\_1,i,...)}, a pseudo $\phi$-node \texttt{A\_2 = PHI(A\_0,A\_1)}, etc..
+The definition can be an update {\sf A\_2 = update(A\_1,i,...)}, a pseudo $\phi$-node {\sf A\_2 = PHI(A\_0,A\_1)}, etc..
 Note that $\phi$-nodes for arrays have no subscript operations the way there are subscript operations in analysis-introduced arrays representing scalars.
@@
 However, the implementation of raise dimension and drop dimension will be adapted because the dimension cannot be raised or dropped to a dimension lower than the canonical one.
-Consider a def-use edge $X \rightarrow Y$ for an array $\texttt{A}$.
+Consider a def-use edge $X \rightarrow Y$ for an array ${\sf A}$.
 
@@
 
-If the use of the array is a read $\texttt{A}[f(i,j,k)]$ different than a canonical read $\texttt{A}[i,j,k]$, then we need to add a reshape operation as
-all arrays are $\texttt{A}[i,j,k]$. It can be added after raise\_dim/drop\_dim or incorporated in these operations.
-The bulk of the change is in Phase 2 of Basic Vectorization as outlined earlier.
+If the use of the array is a read ${\sf A}[f(i,j,k)]$ different than a canonical read ${\sf A}[i,j,k]$, then we need to add a reshape operation as
+all arrays are ${\sf A}[i,j,k]$. It can be added after raise\_dim/drop\_dim or incorporated in these operations.
+The bulk of the change is in Phase 2 of Vectorization as outlined earlier.
 
@@
 
-First, the canonical dimensionality of all \texttt{A,B,C} and \texttt{D} is 1.
-After Phase 1 of Basic Vectorization the Aiken's array write example will be (roughly) as follows:
+First, the canonical dimensionality of all {\sf A,B,C} and {\sf D} is 1.
+After Phase 1 of Vectorization the Aiken's array write example will be (roughly) as follows:
 
 {\small
-\begin{verbatim}
+\begin{pythonn}
 for i in range(N):
-1. A_1 = PHI(A_0,A_2)
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. A_2 = update(A_1, I, B_1[I] + 10);
-6. B_2 = update(B_1, I, A_2[I] * D_1[I-1]);
-7. C_2 = update(C_1, I, A_2[I] * D_1[I-1]);
-8. D_2 = update(D_1, I, B_2[I] * C_2[I]);
-\end{verbatim}
+  A_1 = PHI(A_0,A_2)
+  B_1 = PHI(B_0,B_2)
+  C_1 = PHI(C_0,C_2)
+  D_1 = PHI(D_0,D_2)
+  A_2 = update(A_1, I, B_1[I] + 10);
+  B_2 = update(B_1, I, A_2[I] * D_1[I-1]);
+  C_2 = update(C_1, I, A_2[I] * D_1[I-1]);
+  D_2 = update(D_1, I, B_2[I] * C_2[I]);
+\end{pythonn}
 }
@@
 
-Phase 2 computes the closure of 4; $cl = \{4,6,7, 8\}$
-while 5 is vectorizable. Recall that 1,2, and 3 are target-less phi-nodes. Since the closure $cl$ includes updates to \texttt{B} and \texttt{C},
-the corresponding phi-nodes are added back to the closure and the def-use edges are added back to the target-less nodes. The uses of \texttt{A\_1} and \texttt{B\_1} in the vectorized statement
-turn into uses of \texttt{A\_0} and \texttt{B\_0} respectively; this is done for all original target-less phi-node.
-(But note that \texttt{A\_0} is irrelevant; the update writes into array \texttt{A\_2} in parallel.)
-Finally, the target-less phi-node for \texttt{A} is discarded.
+Phase 2 computes the closure of 5; $cl = \{5,7,8, 9\}$
+while 6 is vectorizable. Recall that 2,3, and 4 are target-less phi-nodes. Since the closure $cl$ includes updates to {\sf B} and {\sf C},
+the corresponding phi-nodes are added back to the closure and the def-use edges are added back to the target-less nodes. The uses of {\sf A\_1} and {\sf B\_1} in the vectorized statement
+turn into uses of {\sf A\_0} and {\sf B\_0} respectively; this is done for all original target-less phi-node.
+(But note that {\sf A\_0} is irrelevant; the update writes into array {\sf A\_2} in parallel.)
+Finally, the target-less phi-node for {\sf A} is discarded.
 
 {\small
-\begin{verbatim}
-1. A_2 = update(A_0, I, ADD_SIMD(B_0[I],10));
+\begin{pythonn}
+A_2 = update(A_0, I, ADD_SIMD(B_0[I],10));
    equiv. to A_2[I] = ADD_SIMD(B_0[I],10)
-FOR i=0; i<N; i++; // MOTION loop
-2. B_1 = PHI(B_0,B_2)
-3. C_1 = PHI(C_0,C_2)
-4. D_1 = PHI(D_0,D_2)
-5. B_2 = update(B_1, i, A_2[i] * D_1[i-1]);
+for i in range(N): // MOTION loop
+  B_1 = PHI(B_0,B_2)
+  C_1 = PHI(C_0,C_2)
+  D_1 = PHI(D_0,D_2)
+  B_2 = update(B_1, i, A_2[i] * D_1[i-1]);
      equiv. to B_2 = B_1; B_2[i] = A_2[i] * D_1[i-1];
-6. C_2 = update(C_1, i, A_2[i] * D_1[i-1]);
-7. D_2 = update(D_1, i, B_2[i] * C_2[i]);
-\end{verbatim}
+  C_2 = update(C_1, i, A_2[i] * D_1[i-1]);
+  D_2 = update(D_1, i, B_2[i] * C_2[i]);
+\end{pythonn}
 }
@@
 {\small
-\begin{verbatim}
+\begin{pythonn}
 for i in range(0, num_bins):
@@
 return res1
-\end{verbatim}
+\end{pythonn}
 }
 
-The canonical dimensionality of \texttt{res} is 1.
-Also, the phi-node \texttt{res1 = PHI(res, res2)} is a target-less phi-node (the implication being that the inner for loop can be vectorized across $i$).
-After Phase 1, Basic vectorization produces the following code (statements are implicitly vectorized along $i$ and $j$).
-In a vectorized update statement, we can ignore the incoming array, \texttt{res2} in this case.
-The update writes (in parallel) all locations of the 2-dimensional array, in this case it sets up each \texttt{res3[i,j] = tmp3[i,j]}.
+The canonical dimensionality of {\sf res} is 1.
+Also, the phi-node {\sf res1 = PHI(res, res2)} is a target-less phi-node (the implication being that the inner for loop can be vectorized across $i$).
+After Phase 1, Vectorization produces the following code (statements are implicitly vectorized along $i$ and $j$).
+In a vectorized update statement, we can ignore the incoming array, {\sf res2} in this case.
+The update writes (in parallel) all locations of the 2-dimensional array, in this case it sets up each {\sf res3[i,j] = tmp3[i,j]}.
 
 {\small
-\begin{verbatim}
+\begin{pythonn}
 A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
@@
 for i in range(0, num_bins):
-   res1 = PHI(res, res2') // target-less phi-node
-   res1' = raise_dim(res1, (j:N))
+   res1 = PHI(res, res2^) # target-less phi-node
+   res1^ = raise_dim(res1, (j:N))
    for j in range(0, N):
-      res2 = PHI(res1', res3)
+     res2 = PHI(res1^, res3)
      tmp1 = (A1 == I)
@@
      res3 = Update(res2, (I,J), tmp3)
-   res2' = drop_dim(res2)
+   res2^ = drop_dim(res2)
 res1'' = drop_dim(res1)
 return res1''
-\end{verbatim}
+\end{pythonn}
 }
 
-Processing the inner loop in Phase 2 vectorizes \texttt{tmp1 = (A1 == I)} along the $j$ dimension but leaves the rest of the statements in a MOTION loop.
+Processing the inner loop in Phase 2 vectorizes {\sf tmp1 = (A1 == I)} along the $j$ dimension but leaves the rest of the statements in a MOTION loop.
 Processing the outer loop is interesting.
@@
 (1) Everything can be vectorized along the $i$ dimension.
-(2) We remove the target-less PHI node, however, we must update uses of \texttt{res1} appropriately: the use at raise\_dim goes to the first argument of the PHI function and the use at drop\_dim goes to the second argument.
+(2) We remove the target-less PHI node, however, we must update uses of {\sf res1} appropriately: the use at $\mathit{raise\_dim}$ goes to the first argument of the PHI function and the use at $\mathit{drop\_dim}$ goes to the second argument.
 
 {\small
-\begin{verbatim}
+\begin{pythonn}
 A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
@@
 
-res1' = raise_dim(res, (j:N)) // replacing res1 with res, 1st arg
+res1^ = raise_dim(res, (j:N)) // replacing res1 with res, 1st arg
 for j in range(0, N):
-   res2 = PHI(res1', res3)
+   res2 = PHI(res1^, res3)
    tmp2[I,j] = (res2[I,j] + B1[I,j])
@@
    equiv. to res3 = res2; res3[I,j] = tmp3[I,j]
-res2' = drop_dim(res2)
-res1 = drop_dim(res2') // replacing with res2', 2nd arg. NOOP
+res2^ = drop_dim(res2)
+res1 = drop_dim(res2^) // replacing with res2^, 2nd arg. NOOP
 return res1
-\end{verbatim}
+\end{pythonn}
 }

### 2022-04-30 overleaf/compiler.tex
Edits across several sections.

@@
 For read access, the compiler allows an arbitrary such function.
-However, it restricts write access to \emph{canonical writes}, i.e., {\sf A[i,j,k] = ...}
+However, it restricts write access to \emph{canonical writes}, i.e., {\sf A[$i,j,k$] = ...}
 where $i$, $j$ and $k$ loop over the three dimensions of {\sf A}.
-Write access such as for example {\sf A[i,j+2] = ...} is not allowed.
+Write access such as for example {\sf A[$i,j+2$] = ...} is not allowed.
 %\item \ana{Add rules for logical ops}
@@
 We note that while the programmer writes
-annotations at the level of IMP Source (recall \ana{ref to overview section}), the
+annotations at the level of IMP Source (as in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(a)), the
 annotations propagate through the transformations and are checked at the 
-level of MPC Source.
+level of MPC Source. Recall that the programmer annotates only inputs
 

### 2022-04-30 overleaf/prob_stmt_np_hardness.tex
Edits across several sections.

@@
 
-As an example, consider the MPC program consisting of the following three sequences, all made up of two distinct $\alpha$-instructions $M_1$ and $M_2$, e.g., a $M_1$ is MUL and $M_2$ is MUX.
+As an example, consider the MPC program consisting of the following three sequences, all made up of two distinct $\alpha$-instructions $M_1$ and $M_2$, e.g., $M_1$ is MUL and $M_2$ is MUX.
 The right arrow indicates a def-use \emph{dependence}, meaning that the source node must execute before the target node:
@@
 The shortest common supersequence problem is as follows: {\it given two or more sequences find the the shortest sequence that contains all of the original sequences.} This can be solved in $O(n^k)$ time, where $n$ is the cardinality of the longest sequence and $k$ is the number of sequences.
-For our problem $n$ is the maximum length of a sequence and $k$ is the number of instructions of sequences.
+%For our problem $n$ is the maximum length of a sequence and $k$ is the number of instructions of sequences.
 We can immediately see that the optimal schedule is the shortest schedule, since the shortest schedule minimizes the fixed cost while the variable cost remains the same.

### 2022-04-30 overleaf/vectorization.tex
Edits across several sections.

@@
 %E.g. $\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$ is an inner-to-outer edge.
-An inner-to-outer edge can be a back-edge as well, in which case both $X$ and $Y$ are phi-nodes with the source $X$ in a loop nested into $Y$'s loop (not necessarily immediately).
+An inner-to-outer edge can be a back-edge as well, in which case both $X$ and $Y$ are $\phi$-nodes with the source $X$ in a loop nested into $Y$'s loop (not necessarily immediately).
 
@@
 
-\subsubsection{Overview}
+%\subsubsection{Overview}
 
-An important component of our algorithm is the scalar expansion to the corresponding loop dimensionality.
+An important component of our algorithm is the scalar expansion to the corresponding loop dimensionality, 
+which is necessary to expose opportunities for vectorization.
 For example, {\sf d = S[i*D+j] - C[j]} equiv. to {\sf d = S[i,j] - C[j]}, which gave rise to {\sf N*D} subtraction operations in the sequential schedule,
@@
 
-There are three kinds of arrays.
+\paragraph{Arrays} There are three kinds of arrays.
 \begin{itemize}
@@
 
-\subsubsection{Raise Dimension and Drop Dimension}
+%\subsubsection{Raise Dimension and Drop Dimension}
 
+\paragraph{Raise dimension} 
 There are two conceptual versions of $\mathit{raise\_dim}$. One reshapes an arbitrary access into a canonical read access in the corresponding loop. 
@@
 
+\paragraph{Drop dimension} 
 The other version of $\mathit{raise\_dim}$ lifts a lower-dimension array into a

### 2022-04-30 overleaf/biometric_vectorization_example.tex
Edits to Section 5.

@@
 min_sum!1 = MAX_INT
-min_index!1 = 0
-for i in range(0, N!0):
+min_idx!1 = 0
+for i in range(0, N):
    min_sum!2 = PHI(min_sum!1, min_sum!4)
-   min_index!2 = PHI(min_index!1, min_index!4)
+   min_idx!2 = PHI(min_idx!1, min_idx!4)
    sum!2 = 0
-   for j in range(0, D!0):
+   for j in range(0, D):
       sum!3 = PHI(sum!2, sum!4)
-     d!3 = (S!0[((i * D!0) + j)] - C!0[j])
-     p!3 = (d!3 * d!3)
-     sum!4 = (sum!3 + p!3)
-   !1!2 = (sum!3 < min_sum!2)
+      d = SUB(S[((i * D) + j)],C[j])
+      p = MUL(d,d)
+      sum!4 = ADD(sum!3,p)
+   t = CMP(sum!3,min_sum!2)
    min_sum!3 = sum!3
-   min_index!3 = i
-   min_sum!4 = MUX(!1!2, min_sum!3, min_sum!2)
-   min_index!4 = MUX(!1!2, min_index!3, min_index!2)
-!2!1 = (min_sum!2, min_index!2)   
+   min_idx!3 = i
+   min_sum!4 = MUX(t, min_sum!3, min_sum!2)
+   min_idx!4 = MUX(t, min_idx!3, min_idx!2)
+return (min_sum!2, min_idx!2)
+
 \end{pythonn}
@@
 
-In the code below statements (e.g., {\sf min\_sum!3 = i}) are implicitly vectorized. 
+In the code below statements (e.g., {\sf min\_sum!3 = sum!3}) are implicitly vectorized. 
 %$\mathit{raise\_dim}$ and $\mathit{drop\_dim}$ statements have slightly different interpretation. %\ana{This is vague, may need elaboration.}
-The example illustrates the two different versions of $\mathit{raise\_dim}$. {\sf raise\_dim(C!0, j, (i:N!0,j:D!0))} reshapes
-the read-only input array, while {\sf drop\_dim(sum!3)} removes the $j$ dimension of {\sf sum!3}.
+The example illustrates the two different versions of $\mathit{raise\_dim}$. E.g., {\sf raise\_dim(C, j, (i:N,j:D))} reshapes
+the read-only input array. {\sf drop\_dim(sum!3)} removes the $j$ dimension of {\sf sum!3}.
 %\ana{More here! One just adds a least-significant dimension. The other one may reshape an input array.}
@@
 min_sum!1 = MAX_INT
-min_sum!1^ = raise_dim(min_sum!1, (i:N!0))
-min_index!1 = 0
-min_index!1^ = raise_dim(min_index!1, (i:N!0))
-S!0^ = raise_dim(S!0, ((i * D!0) + j), (i:N!0,j:D!0))
-C!0^ = raise_dim(C!0, j, (i:N!0,j:D!0))
-for i in range(0, N!0):
+min_sum!1^ = raise_dim(min_sum!1, (i:N))
+min_idx!1 = 0
+min_idx!1^ = raise_dim(min_idx!1, (i:N))
+S^ = raise_dim(S, ((i * D) + j), (i:N,j:D))
+C^ = raise_dim(C, j, (i:N,j:D))
+for i in range(0, N):
    min_sum!2 = PHI(min_sum!1^, min_sum!4)
-   min_index!2 = PHI(min_index!1^, min_index!4) 
+   min_idx!2 = PHI(min_idx!1^, min_idx!4) 
    sum!2 = 0 // Will lift, when hoisted
-   sum!2^ = raise_dim(sum!2, (j:D!0)) 
-   for j in range(0, D!0):
+   sum!2^ = raise_dim(sum!2, (j:D)) 
+   for j in range(0, D):
       sum!3 = PHI(sum!2^, sum!4)
-      d!3 = S!0^ - C!0^
-      p!3 = (d!3 * d!3) 
-      sum!4 = (sum!3 + p!3)
+      d = SUB(S^,C^)
+      p = MUL(d,d) 
+      sum!4 = ADD(sum!3,p)
    sum!3^ = drop_dim(sum!3)     
-   !1!2 = (sum!3^ < min_sum!2)
+   t = CMP(sum!3^,min_sum!2)
    min_sum!3 = sum!3^
-   min_index!3 = i  // Same-level, will lift when hoisted
-   min_sum!4 = MUX(!1!2, min_sum!3, min_sum!2)
-   min_index!4 = MUX(!1!2, min_index!3, min_index!2)
+   min_idx!3 = i  // Same-level, will lift when hoisted
+   min_sum!4 = MUX(t, min_sum!3, min_sum!2)
+   min_idx!4 = MUX(t, min_idx!3, min_idx!2)
 min_sum!2^ = drop_dim(min_sum!2)
-min_index!2^ = drop_dim(min_index!2)   
-!2!1 = (min_sum!2^, min_index!2^)     
+min_idx!2^ = drop_dim(min_idx!2)   
+return (min_sum!2^, min_idx!2^)     
 \end{pythonn}
@@
 Starting at the inner phi-node {\sf sum!3 = PHI(...)}, the algorithm first computes its closure.
-The closure amounts to the phi-node itself and the addition node {\sf sum!4 = (sum!3 + p!3)}, accounting for the loop-carried dependency of the computation of {\sf sum}.
-The algorithm replaces this closure with a FOR loop on $j$ removing vectoriaztion on $j$.
-Note that the SUB and MUL computations remain outside of the loop as they do not depend on phi-nodes that are part of cycles.
-The dependences are from {\sf p!3[I,J] = (d!3[I,J] * d!3[I,J])} to the monolithic FOR loop and from the FOR loop to {\sf sum!3\^ = drop\_dim(sum!3)}.
+The closure amounts to the phi-node itself and the addition node {\sf sum!4 = ADD(sum!3,p!3)}, accounting for the loop-carried dependency of the computation of {\sf sum}.
+The algorithm replaces this closure with a for-loop on $j$ removing vectorization on $j$.
+Note that the SUB and MUL computations remain outside of the loop as they do not depend on PHI-nodes that are part of cycles.
+The dependences are from {\sf p[I,J] = MUL(d[I,J],d[I,J])} to the monolithic for-loop and from the for-loop to {\sf sum!3\^ = drop\_dim(sum!3)}.
 Lower case index, e.g., \texttt{i}, indicates non-vectorized dimension, while uppercase index, e.g., \texttt{I} indicates vectorized dimension.
 
-After processing inner loop code becomes:
+After processing the inner loop code becomes:
 
@@
 min_sum!1^ = raise_dim(min_sum!1, (i:N!0))
-min_index!1 = 0
-min_index!1^ = raise_dim(min_index!1, (i:N!0))
-S!0^ = raise_dim(S!0, ((i * D!0) + j), (i:N!0,j:D!0))
-C!0^ = raise_dim(C!0, j, (i:N!0,j:D!0))
-for i in range(0, N!0):
+min_idx!1 = 0
+min_idx!1^ = raise_dim(min_idx!1, (i:N))
+S^ = raise_dim(S, ((i * D) + j), (i:N,j:D))
+C^ = raise_dim(C, j, (i:N,j:D))
+for i in range(0, N):
   min_sum!2[I] = PHI(min_sum!1^[I], min_sum!4[I]) 
-  min_index!2[I] = PHI(min_index!1^[I], min_index!4[I])  
+  min_idx!2[I] = PHI(min_idx!1^[I], min_idx!4[I])  
   sum!2 = [0,..,0] 
-  sum!2^ = raise_dim(sum!2, (j:D!0))
-  d!3[I,J] = S!0^[I,J] - C!0^[I,J]
-  p!3[I,J] = (d!3[I,J] * d!3[I,J])
-  for j in range(0, D!0):
+  sum!2^ = raise_dim(sum!2, (j:D))
+  d[I,J] = SUB(S^[I,J],C^[I,J])
+  p[I,J] = MUL(d[I,J],d[I,J])
+  for j in range(0, D):
     sum!3[I,j] = PHI(sum!2^[I,j], sum!4[I,j-1])       
-    sum!4[I,j] = (sum!3[I,j] + p!3[I,j])
+    sum!4[I,j] = ADD(sum!3[I,j],p[I,j])
   sum!3^ = drop_dim(sum!3)     
-  !1!2[I] = (sum!3^[I] < min_sum!2[I])
+  t[I] = CMP(sum!3^[I],min_sum!2[I])
   min_sum!3 = sum!3^
-  min_index!3 = i
-  min_sum!4[I] = MUX(!1!2[I], min_sum!3[I], min_sum!2[I])
-   min_index!4[I] = MUX(!1!2[I], min_index!3[I], min_index!2[I])
+  min_idx!3 = i
+  min_sum!4[I] = MUX(t[I], min_sum!3[I], min_sum!2[I])
+  min_idx!4[I] = MUX(t[I], min_idx!3[I], min_idx!2[I])
 min_sum!2^ = drop_dim(min_sum!2)
-min_index!2^ = drop_dim(min_index!2)   
-!2!1 = (min_sum!2^, min_index!2^)
+min_idx!2^ = drop_dim(min_idx!2)   
+return (min_sum!2^, min_idx!2^)
 \end{pythonn}
@@
 When processing the outer loop two closures arise, one for {\sf min\_sum!2[I] = PHI(...)} and one 
-for {\sf min\_index!2[I] = PHI(...)}. Since the two closures \emph{do not} intersect, we have two distinct FOR-loops on $i$:
+for {\sf min\_idx!2[I] = PHI(...)}. Since the two closures \emph{do not} intersect, we have two distinct for-loops on $i$:
 
@@
 min_sum!1 = MAX_INT
-min_sum!1^ = raise_dim(min_sum!1, (i:N!0))
-min_index!1 = 0
-min_index!1^ = raise_dim(min_index!1, (i:N!0))
-S!0^ = raise_dim(S!0, ((i * D!0) + j), (i:N!0,j:D!0))
-C!0^ = raise_dim(C!0, j, (i:N!0,j:D!0))
+min_sum!1^ = raise_dim(min_sum!1, (i:N))
+min_idx!1 = 0
+min_idx!1^ = raise_dim(min_idx!1, (i:N))
+S^ = raise_dim(S, ((i * D) + j), (i:N,j:D))
+C^ = raise_dim(C, j, (i:N,j:D))
 
 sum!2 = [0,..,0]
-sum!2^ = raise_dim(sum!2, (j:D!0))
-d!3[I,J] = S!0^[I,J] - C!0^[I,J]
-p!3[I,J] = (d!3[I,J] * d!3[I,J])
+sum!2^ = raise_dim(sum!2, (j:D))
+d[I,J] = SUB(S^[I,J], C^[I,J])
+p[I,J] = MUL(d[I,J], d[I,J])
 
-for j in range(0, D!0):
+for j in range(0, D):
   sum!3[I,j] = PHI(sum!2^[I,j], sum!4[I,j-1])       
-  sum!4[I,j] = (sum!3[I,j] + p!3[I,j])
+  sum!4[I,j] = ADD(sum!3[I,j],p[I,j])
 
 sum!3^ = drop_dim(sum!3)     
-min_index!3 = [0,1,2,...N!0-1] // or min_index!3 = [i, (i:N!0)]
+min_idx!3 = [0,1,2,...N-1] // i.e., min_idx!3 = [i, (i:N)]
 min_sum!3 = sum!3^
 
-for i in range(0, N!0):
+for i in range(0, N):
   min_sum!2[i] = PHI(min_sum!1^[i], min_sum!4[i-1]) 
-  !1!2[i] = (sum!3^[i] < min_sum!2[i])
-  min_sum!4[i] = MUX(!1!2[i], min_sum!3[i], min_sum!2[i])
+  t[i] = CMP(sum!3^[i], min_sum!2[i])
+  min_sum!4[i] = MUX(t[i], min_sum!3[i], min_sum!2[i])
 
-for i in range(0, N!0):
-  min_index!2[i] = PHI(min_index!1^[i], min_index!4[i-1])  
-   min_index!4[i] = MUX(!1!2[i], min_index!3[i], min_index!2[i])
+for i in range(0, N):
+  min_idx!2[i] = PHI(min_idx!1^[i], min_idx!4[i-1])  
+  min_idx!4[i] = MUX(t[i], min_idx!3[i], min_idx!2[i])
 
 min_sum!2^ = drop_dim(min_sum!2)
-min_index!2^ = drop_dim(min_index!2)   
-!2!1 = (min_sum!2^, min_index!2^)
+min_idx!2^ = drop_dim(min_idx!2)   
+return (min_sum!2^, min_idx!2^)
 \end{pythonn}
@@
 min_sum!1 = MAX_INT
-min_index!1 = 0
-S!0^ = raise_dim(S!0, ((i * D!0) + j), (i:N!0,j:D!0))
-C!0^ = raise_dim(C!0, j, (i:N!0,j:D!0))
+min_idx!1 = 0
+S^ = raise_dim(S, ((i * D) + j), (i:N,j:D))
+C^ = raise_dim(C, j, (i:N,j:D))
 
 sum!2 = [0,..,0]
-d!3[I,J] = S!0^[I,J] - C!0^[I,J]
-p!3[I,J] = (d!3[I,J] * d!3[I,J])
+d[I,J] = SUB(S^[I,J],C^[I,J])
+p[I,J] = MUL(d[I,J],d[I,J])
 
 // j is redundant for sum!3 and sum!4
-for j in range(0, D!0):
+for j in range(0, D):
   sum!3[I] = PHI(sum!2[I], sum!4[I])       
-  sum!4[I] = (sum!3[I] + p!3[I,j])
+  sum!4[I] = ADD(sum!3[I], p[I,j])
 
@@
 // then copy propagation and dead code elimination 
-min_index!3 = [0,1,2,...N!0-1] // or min_index!3 = [i, (i:N!0)]
+min_idx!3 = [0,1,2,...N-1] // i.e., min_idx!3 = [i, (i:N)]
 
-// i is redundant for min_sum!2, min_sum!4 but not for !12![i]
-for i in range(0, N!0):
+// i is redundant for min_sum!2, min_sum!4 but not for t[i]
+for i in range(0, N):
   min_sum!2 = PHI(min_sum!1, min_sum!4) 
-  !1!2[i] = (sum!3[i] < min_sum!2)
-  min_sum!4 = MUX(!1!2[i], sum!3[i], min_sum!2)
+  t[i] = CMP(sum!3[i],min_sum!2)
+  min_sum!4 = MUX(t[i], sum!3[i], min_sum!2)
 
-// same, i is redundant for min_index!2, min_index!4
-for i in range(0, N!0):
-  min_index!2 = PHI(min_index!1, min_index!4)  
-   min_index!4 = MUX(!1!2[i], min_index!3[i], min_index!2)
+// same, i is redundant for min_idx!2, min_idx!4
+for i in range(0, N):
+  min_idx!2 = PHI(min_idx!1, min_idx!4)  
+  min_idx!4 = MUX(t[i], min_idx!3[i], min_idx!2)
 
 // drop_dim becomes redundant
-!2!1 = (min_sum!2, min_index!2)
+return (min_sum!2, min_idx!2)
 \end{pythonn}
@@
 min_sum!1 = MAX_INT
-min_index!1 = 0
-S!0^ = raise_dim(S!0, ((i * D!0) + j), (i:N!0,j:D!0))
-C!0^ = raise_dim(C!0, j, (i:N!0,j:D!0))
+min_idx!1 = 0
+S^ = raise_dim(S, ((i * D) + j), (i:N,j:D))
+C^ = raise_dim(C, j, (i:N,j:D))
 
 sum!2 = [0,..,0]
-d!3[I,J] = SUB_SIMD(S!0^[I,J],C!0^[I,J])
-p!3[I,J] = MUL_SIMD(d!3[I,J], d!3[I,J])
+d[I,J] = SUB_SIMD(S^[I,J],C^[I,J])
+p[I,J] = MUL_SIMD(d[I,J], d[I,J])
 
-for j in range(0, D!0):
+for j in range(0, D):
   // I dim is a noop. sum is already a one-dimensional vector
   sum!3[I] = PHI(sum!2[I], sum!4[I])       
-  sum!4[I] = ADD_SIMD(sum!3[I],p!3[I,j])
+  sum!4[I] = ADD_SIMD(sum!3[I],p[I,j])
 
-min_index!3 = [0,1,...N!0-1]
+min_idx!3 = [0,1,...N-1]
 
-for i in range(0, N!0):
+for i in range(0, N):
   min_sum!2 = PHI(min_sum!1, min_sum!4) 
-  !1!2[i] = CMP(sum!3[i],min_sum!2)
-  min_sum!4 = MUX(!1!2[i], sum!3[i], min_sum!2)
+  t[i] = CMP(sum!3[i],min_sum!2)
+  min_sum!4 = MUX(t[i], sum!3[i], min_sum!2)
 
-for i in range(0, N!0):
-  min_index!2 = PHI(min_index!1, min_index!4)  
-  min_index!4 = MUX(!1!2[i], min_index!3[i], min_index!2)
+for i in range(0, N):
+  min_idx!2 = PHI(min_idx!1, min_idx!4)  
+  min_idx!4 = MUX(t[i], min_idx!3[i], min_idx!2)
 
-!2!1 = (min_sum!2, min_index!2)   
+return (min_sum!2, min_idx!2)   
 \end{pythonn}

### 2022-04-30 overleaf/vectorization.tex
Edits to Section 5.

@@
 
-\ana{Need an intro/overview here!}
-\ana{Need to define notion of canonical access!}
-\ana{Talk about MPC Source, opportunities for vectorization, algorithm unique to MPC Source.}
+%\ana{Need an intro/overview here!}
+%\ana{Need to define notion of canonical access!}
+%\ana{Talk about MPC Source, opportunities for vectorization, algorithm unique to MPC Source.}
+
+This section describes our vectorization algorithm. While vectorization is a longstanding problem, 
+and we build upon existing work on scalar expansion and classical loop vectorization~\cite{Allen:1987}, 
+our algorithm is unique as it works on the MPC Source SSA-form representation. We posit that vectorization 
+over MPC Source is a new problem that warrants a new look, in part because of MPC's unique linear 
+structure and in part because vectorization meshes in with other MPC-specific optimizations in non-trivial ways
+(other works have explored manual vectorization and protocol mixing in an ad-hoc way, e.g.,~\cite{NDSS:DemSchZoh15,CCS:BDKKS18,Ishaq:2019}).
+
+\secref{sec:dependence} describes our dependence analysis and ~\secref{sec:expansion} describes scalar expansion,
+which lifts scalars (and arrays) to the corresponding loop dimensionality to create opportunities for vectorization. 
+\secref{sec:basic_vectorization} describes our core vectorization algorithm and~\secref{sec:correctness} argues correctness of the vectorization transformation.
+\secref{sec:array_writes} extends vectorization with array writes.
 
@@
 
-The next steps construct the dependence graph and \ana{Add an intro here.}
+We build a dependence graph where the nodes are the MPC Source statements and the edges represent the def-use relations.
 
-\subsubsection{Def-use Edges}
-
-The dependence graph has the following def-use edges:
+\paragraph{Def-use Edges} We distinguish the following def-use edges:
 
@@
 
-\subsubsection{Closures}
-
-We define $\mathit{closure}(n)$ where $n$ is a phi-node. Intuitively, it computes the set of nodes (i.e., statements) that form a dependence cycle with $n$. The closure of $n$ is defined as follows:
+\paragraph{Closures}
+We define $\mathit{closure}(n)$ where $n$ is a PHI-node. Intuitively, it computes the set of nodes (i.e., statements) that form a dependence cycle with $n$. The closure of $n$ is defined as follows:
 \begin{itemize}
@@
 \subsection{Scalar Expansion}
+\label{sec:expansion}
 
@@
 which is necessary to expose opportunities for vectorization.
-For example, {\sf d = S[i*D+j] - C[j]} equiv. to {\sf d = S[i,j] - C[j]}, which gave rise to {\sf N*D} subtraction operations in the sequential schedule,
-is lifted. The argument arrays {\sf S} and {\sf C} are lifted and the scaler {\sf d} is lifted: {\sf d[i,j] = S[i,j] - C[i,j]}.
+In the Biometric example, {\sf d = S[i*D+j] - C[j]} equiv. to {\sf d = S[i,j] - C[j]}, which gave rise to $N*D$ subtraction operations in the sequential schedule,
+is lifted. The argument arrays {\sf S} and {\sf C} are lifted and the scalar {\sf d} is lifted: {\sf d[i,j] = S[i,j] - C[i,j]}.
 The algorithm then detects that the statement can be vectorized.
 
-\paragraph{Arrays} There are three kinds of arrays.
+\paragraph{Arrays} Conceptually, we treat all variables as arrays. There are three kinds of arrays. 
+%\ana{Arrays is not a good name here.}
+
 \begin{itemize}
-
 \item Scalars: These are scalar variables we expand into arrays for the purposes of vectorization.
@@
 
-In the sections below we detail the $\mathit{raise\_dim}$ (raise dimension) and $\mathit{drop\_dim}$ (drop dimension) operations,
-followed by our vectorization algorithm.
+%In the sections below we detail the $\mathit{raise\_dim}$ (raise dimension) and $\mathit{drop\_dim}$ (drop dimension) operations,
+%followed by our vectorization algorithm.
 
@@
 \paragraph{Raise dimension} 
+The $\mathit{raise\_dim}$ function expands a scalar (or array). 
 There are two conceptual versions of $\mathit{raise\_dim}$. One reshapes an arbitrary access into a canonical read access in the corresponding loop. 
@@
 $((i:I),(j:J),(k:K))$:
-\[ \mathit{raise\_dim}({\sf A, f(i,j,k), ((i:I),(j:J),(k:K))}) \]
-It produces a new 3-dimensional array {\sf A'} by iterating over $i,j,k$ and setting each element of  {\sf A'} as follows:
-\[ {\sf A'[i,j,k] =  A[f(i,j,k)]} \]
-The end result is that uses of {\sf [A[f(i,j,k)]} in loop nest $i,j,k$ are replaced with canonical read-accesses to {\sf A'[i,j,k]}
-that can be vectorized. In the running Biometric example, {\sf C'} = $\mathit{raise\_dim}${\sf (C, j, (i:N,j:D))} lifts the
-1-dimensional array {\sf C} into a 2-dimensional array. The $i,j$ loop now accesses {\sf C'} in the canonical way, {\sf C'[i,j]}.
-Similarly, {\sf S'} = $\mathit{raise\_dim}${\sf (S, i*D+j, (i:N,j:D))} tries to lift {\sf S}, but the operation turns into a no-op
+\[ \mathit{raise\_dim}(A, f(i,j,k), ((i:I),(j:J),(k:K))) \]
+It produces a new 3-dimensional array ${\sf A'}$ by iterating over $i,j,k$ and setting each element of  ${\sf A'}$ as follows:
+\[ {\sf A'}[i,j,k] =  {\sf A}[f(i,j,k)] \]
+The end result is that uses of ${\sf A}[f(i,j,k)]$ in loop nest $i,j,k$ are replaced with canonical read-accesses to ${\sf A'}[i,j,k]$
+that can be vectorized. In the running Biometric example, ${\sf C'} = \mathit{raise\_dim}({\sf C}, j, (i:N,j:D))$ lifts the
+1-dimensional array {\sf C} into a 2-dimensional array. The $i,j$ loop now accesses ${\sf C'}$ in the canonical way, ${\sf C'}[i,j]$.
+Similarly, ${\sf S'} = \mathit{raise\_dim}({\sf S}, i*D+j, (i:N,j:D))$ tries to lift {\sf S}, but the operation turns into a no-op
 because {\sf S} is already a 2-dimensional array and the read access is canonical.
 
-\paragraph{Drop dimension} 
 The other version of $\mathit{raise\_dim}$ lifts a lower-dimension array into a
@@
 Here {\sf A} is an $i$-array and raise dimension adds two additional dimensions:
-\[ \mathit{raise\_dim}({\sf A}, {\sf (j:J,k:K)}) \]
+\[ \mathit{raise\_dim}({\sf A}, (j:J,k:K)) \]
 This version is reduced to the above version by adding the access pattern function, which is just $i$:
-\[ \mathit{raise\_dim}({\sf A, i, (j:J,k:K)}) \]
+\[ \mathit{raise\_dim}({\sf A}, i, (j:J,k:K)) \]
 
+\paragraph{Drop dimension}  
 The corresponding $\mathit{drop\_dim}$ is carried out when an array written in an inner loop is used in an enclosing loop.
 It takes a higher dimensional array, say $i,j,k$ and removes trailing dimensions, say $j,k$:
-\[ \mathit{drop\_dim}({\sf A, (j:J,k:K)}) \]
+\[ \mathit{drop\_dim}({\sf A}, (j:J,k:K)) \]
 It iterates over $i$ and takes the result at the maximal index of $j$ and $k$, i.e., the result at the last iterations of $j$ and $k$:
-\[ {\sf A'[i] = A[i,J-1,K-1]} \]
+\[ {\sf A'}[i] = A[i,J-1,K-1] \]
 
 \subsection{Basic Vectorization}
+\label{sec:basic_vectorization}
 
-We present our vectorization algorithm in detail followed by a detailed example on Biometric Matching.
+We present our algorithm followed by an example on Biometric.
 
@@
 
-\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i,j,k$}
+\FOR {each MPC $\mathit{stmt}: \x = \mathit{Op}(\y_1,\y_2)$ in loop $i,j,k$}
 %\FOR {each $Y_n$}
-\FOR {each argument $Y_n$} %def-use edge $\mathit{stmt}' (\mbox{def of } Y) \rightarrow \mathit{stmt} (\mbox{def of } X)$}
-\STATE {case def-use edge $\mathit{stmt}' (\mbox{def of } Y_n) \rightarrow \mathit{stmt} (\mbox{def of } X)$ of} \\
-\STATE \hspace{0.25cm} {same-level: $Y'_n$ is $Y_n$} \\
-\STATE \hspace{0.25cm} {outer-to-inner: add $Y'_n[i,j,k] = \mathit{raise\_dim}(Y_n)$ at $\mathit{stmt'}$}
+\FOR {each argument $\y_n$} %def-use edge $\mathit{stmt}' (\mbox{def of } Y) \rightarrow \mathit{stmt} (\mbox{def of } X)$}
+\STATE {case def-use edge $\mathit{stmt}' (\mbox{def of } \y_n) \rightarrow \mathit{stmt} (\mbox{def of } \x)$ of} \\
+\STATE \hspace{0.25cm} {same-level: $\y'_n$ is $\y_n$} \\
+\STATE \hspace{0.25cm} {outer-to-inner: add $\y'_n[i,j,k] = \mathit{raise\_dim}(\y_n)$ at $\mathit{stmt'}$}
 \STATE \hspace{0.25cm} {(more precisely, right after $\mathit{stmt'}$)}
-\STATE \hspace{0.25cm} {inner-to-outer: add $Y'_n[i,j,k] = \mathit{drop\_dim}(Y_n)$ at $\mathit{stmt}$}
+\STATE \hspace{0.25cm} {inner-to-outer: add $\y'_n[i,j,k] = \mathit{drop\_dim}(\y_n)$ at $\mathit{stmt}$}
 \STATE \hspace{0.25cm} {(more precisely, in loop of $\mathit{stmt}$ right after loop of $\mathit{stmt'}$)}
@@
 %\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i,j,k$}
-\STATE {change to $X[I,J,K] = \mathit{Op}(Y'_1[I,J,K],Y'_2[I,J,K])$ }
+\STATE {change to $\x[I,J,K] = \mathit{Op}(\y'_1[I,J,K],\y'_2[I,J,K])$ }
 \ENDFOR
 
-\COMMENT { Phase 2: Recreating FOR loops for cycles; vectorizable statements hoisted up. }
+\COMMENT { Phase 2: Recreating for-loops for cycles; vectorizable stmts hoisted up. }
 
 \FOR {each dimension $d$ from highest to 0}
-\FOR {each $\phi$-node $n$ in loop $i_1,...,i_d$}
+\FOR {each PHI-node $n$ in loop $i_1,...,i_d$}
 \STATE {compute $\mathit{closure}(n)$}
@@
 \FOR {each closure $cl$ (after merge)}
-\STATE {create FOR $i_d = 0; ...$ loop}
-\STATE {add $\phi$-nodes in $cl$ to header block}
-\STATE {{\color{red}{add target-less $\phi$-node for A if $cl$ updates array A}}}
-\STATE {add statements in $cl$ to loop body in some order of dependences}
+\STATE {create \python{for} $i_d$ \python{in} ... loop}
+\STATE {add PHI-nodes in $cl$ to header block}
+\STATE {{\color{red}{add target-less PHI-node for A if $cl$ updates array A}}}
+\STATE {add statements in $cl$ to loop in some order of dependences}
 \STATE \COMMENT { Dimension is not vectorizable: }
 \STATE {change ${I_d}$ to $i_d$ in all statements in loop}
-\STATE {treat FOR loop as monolith node for def-uses: e.g., some def-use edges become same-level.}
+\STATE {treat \python{for}-loop as monolith node for def-uses: e.g., some def-use edges become same-level.}
 \ENDFOR
@@
 \color{red}
-\FOR {each target-less $\phi$-node $A_1 = \phi(A_0,A_k)$}
-\STATE {in vectorizable stmts, replace use of $A_1$ with $A_0$}
-\STATE {discard $\phi$-node if not used in any $cl$, replacing $A_1$ with $A_0$ or $A_k$ appropriately}
+\FOR {each target-less PHI-node {\sf A!1 = PHI(A!0,A!k)}}
+\STATE {in vectorizable stmts, replace use of {\sf A!1} with {\sf A!0}}
+\STATE {discard PHI-node if not used in any $cl$, replacing {\sf A!1} with {\sf A!0} or {\sf A!k} appropriately}
 \ENDFOR
@@
 
-\COMMENT { A dimension $i$ is dead on exit from stmt $X[...i...] = ...$ if all def-uses with targets outside of the enclosing
-FOR $i=0 ...$ MOTION loop end at target (use) $X' = \mathit{drop\_dim}(X, i)$. }
-\FOR {each stmt and dimension $X[...i...] = ...$}
-\STATE {if $i$ is a dead dimension on exit from stmt $X[...i...] = ...$, remove $i$ from $X$ (all defs and uses)}
+\COMMENT { A dimension $i$ is dead on exit from stmt $\x[...i...] = ...$ if all def-uses with targets outside of the enclosing
+\python{for} $i$ ... MOTION loop end at target (use) $\x' = \mathit{drop\_dim}(\x, i)$. }
+\FOR {each stmt and dimension $\x[...i...] = ...$}
+\STATE {if $i$ is a dead dimension on exit from stmt $\x[...i...] = ...$, remove $i$ from $\x$ (all defs and uses)}
 %\COMMENT{ A stmt v[...i] is a "dead dimension i" if all edges out of FOR i = 0; ... loop end at drop\_dim(v, i)}
@@
 \COMMENT { Now clean up drop\_dim and raise\_dim }
-\FOR {each $X' =  \mathit{drop\_dim(X,i)}$ }
-\STATE {replace with $X' = X$ if $i$ is dead in $X$.}
+\FOR {each $\x' =  \mathit{drop\_dim(\x,i)}$ }
+\STATE {replace with $\x' = \x$ if $i$ is dead in $\x$.}
 \ENDFOR
@@
 \subsection{Correctness Argument}
+\label{sec:correctness}
 
@@
 \subsection{Extension with Array Writes}
-\label{sec:extension_with_writes}
+\label{sec:array_writes}
 

### 2022-04-30 overleaf/vectorization.tex
Edits to Sec. 5.

@@
 
-There are two key phases of the algorithm. Phase 1 applies dependence analysis and inserts
-raise dimension and drop dimension operations as necessary. After this phase operations work
+There are two key phases of the algorithm. Phase 1 inserts raise dimension and drop dimension 
+operations according to def-uses. E.g., if there is an inner-to-outer dependence, it inserts $\mathit{raise\_dim}$,
+and similarly, if there is an outer-to-inner dependence, it inserts $\mathit{drop\_dim}$. After this phase operations work
 on arrays of the corresponding dimensionality and we optimistically vectorize all arrays.
 
-Phase 2 takes into account dependence analysis. It proceeds from the inner-most towards the outer-most 
-loop. For each loop it anchors dependence cycles (closures) around pseudo PHI nodes, removing vectorization 
-from the dimension of the loop.
+Phase 2 proceeds from the inner-most towards the outer-most loop. For each loop it anchors dependence cycles (closures) 
+around pseudo PHI nodes then removes vectorization from the dimension of that loop. There are two important points in this 
+phase. First, it may break a loop into smaller loops which would discover opportunities 
+for vectorization in intermediate statements in the loop. Second, it handles write arrays. 
+It creates opportunities for vectorization in the presence of write arrays, even though 
+Cytron's SSA adds a back-edge to the array PHI-node, thus killing vectorization of statements
+that read and write that array.
 
-There are several important points in this phase. We may break a loop into smaller loops which would discover opportunities 
-for optimization in intermediate statements in the loop. We remove redundant back edges into PHI nodes that are 
-due entirely to arrays, which opens up opportunities as well. The pseudo-code in red deals with this extension
-of the basic algorithm. We determine that some array pseudo-PHIs are \emph{target-less}, meaning
-that there is no loop-carried dependence for that array and back-edges into these nodes can be safely removed. 
-This opens up opportunities for vectorization across the write array. Handling of target-less nodes is non-trivial, 
-as in certain cases the node cannot be removed. \secref{sec:extension_with_writes} details our algorithm.
+The excerpts in {\color{red} red} color in the pseudo code below highlight the extension 
+with array writes. We advise the reader to omit the extension and assume just read-only
+arrays. We explain the extension in \secref{sec:array_writes} details our algorithm. 
+(As many of our benchmarks include array writes, it plays an important role.)
 
-\ana{Add an example from histogram here.}
+%The pseudo-code in red deals with this extension
+%of the basic algorithm. We determine that some array pseudo-PHIs are \emph{target-less}, meaning
+%that there is no loop-carried dependence for that array and back-edges into these nodes can be safely removed. 
+%This opens up opportunities for vectorization across the write array. Handling of target-less nodes is non-trivial, 
+%as in certain cases the node cannot be removed. \secref{sec:extension_with_writes} details our algorithm.
 
-Phases 3 and 4 are less important --- phase 3 cleans up local arrays of references and phase 4 turns operations into MOTION operations.
+Phases 3 cleans up local arrays of references (this is an optional phase and our current implementation does not include it)
+and Phase 4 explicitly turns operations into MOTION SIMD operations.
 
@@
   s &::= s_1;s_2 & \gamma(s) = \gamma(s_1) \; ; \; \gamma(s_2) & \mathit{sequence} \\
-     & \mid \x[i,J,k] = {\sf op\_SIMD}(\y[i,J,k], \z[i,J,k]) & \gamma(\x[i,J,k] = {\sf op\_SIMD}(\y[i,J,k], \z[i,J,k])) = & \mathit{operation} \\ 
-     &                                                                               & ~~\x[i,0,k] = \y[i,0,k] \; {\sf op} \; \z[i,0,k] \;\; || &  \\
-     &                                                                               & ~~~~\x[i,1,k] = \y[i,1,k] \; {\sf op} \; \z[i,1,k] \;\; || \; ...\; || & \\
-     &                                                                               & ~~~~~~\x[i,J\!-\!1,k] = \y[i,J\!-\!1,k] \; {\sf op} \; \z[i,J\!-\!1,k] & \\
+     & \mid \x[i,J,k] = {\sf op\_SIMD}(\y_1[i,J,k], \y_2[i,J,k]) & \gamma(\x[i,J,k] = {\sf op\_SIMD}(\y_1[i,J,k], \y_2[i,J,k])) = & \mathit{operation} \\ 
+     &                                                                               & ~~\x[i,0,k] = \y_1[i,0,k] \; {\sf op} \; \y_2[i,0,k] \;\; || &  \\
+     &                                                                               & ~~~~\x[i,1,k] = \y_1[i,1,k] \; {\sf op} \; \y_2[i,1,k] \;\; || \; ...\; || & \\
+     &                                                                               & ~~~~~~\x[i,J\!-\!1,k] = \y_1[i,J\!-\!1,k] \; {\sf op} \; \y_2[i,J\!-\!1,k] & \\
      & \mid \x[i,J,k] = {\sf const} & \mbox{analogous} & \mathit{constant} \\
@@
 
-\figref{fig:mpc_syntax} states the syntax and linearization semantics of MPC Source. For simiplicity,
-we consider only scalars and read-only arrays, however, the treatment extends to write arrays as well.
+\figref{fig:mpc_syntax} states the syntax and linearization semantics of MPC Source. 
+Although notation is heavy, the linearization simply produces schedules as discussed in~\secref{sec:model}; 
+the iterative MPC Source gives rise to what we called sequential schedule where loops are unrolled 
+and MPC Source with vectorized dimensions gives rise to what we called parallel schedule. 
+For simplicity, we consider only scalars and read-only arrays, however, the treatment extends to write arrays as well.
 $\x[i,J,k]$ denotes the value of scalar variable \x{} at loop nest $i,j,k$. 
@@
 
-For each MPC Source program $a$ we compute the def-use edges in the standard way: if base statement $s1 \in a$ defines variable \x, e.g., $\x[i,j,k] = ...$, 
-and base statement $s2 \in a$ uses \x, e.g., $... = ... \x[i,j,k]$ and there is a path in CFG from $s1$ to $s2$, then there is a def-use edge from
+For each MPC Source program $a$ we compute the def-use edges in the standard way: if statement $s1 \in a$ defines variable \x, e.g., $\x[i,j,k] = ...$, 
+and statement $s2 \in a$ uses \x, e.g., $... = ... \x[i,j,k]$ and there is a path in CFG from $s1$ to $s2$, then there is a def-use edge from
  $s1$ to $s2$. We extend the dimensionality of a statement into $s1[i,j,k]$ where $s1[i,j,k]$ inherits the dimensionality of the left-hand-side of the assignment.
-
- \ana{Add on the linearization semantics?}
 
@@
 
-Array writes limit vectorization as they sometimes introduce infeasible loop-carried dependencies. Consider the following example: \ana{Have to add citation to Aiken's paper}
+Array writes limit vectorization as they sometimes introduce infeasible loop-carried dependencies. Consider the classical example from~\cite{Aiken:1988}:
 

### 2022-04-30 compiler/paper_benchmarks.py
fixed out of range error

@@ -905,7 +905,7 @@ def generate_cumulative_graph_for_attr(all_stats, get_val, y_label, dir):
         x = 1
         for task_stat in all_stats:
             label = get_x_label_for_benchmark(task_stat.label)
-            for i in range(len(task_stat.input_configs)-1, 0, -1):
+            for i in range(len(task_stat.input_configs)-1, -1, -1):
                 input = task_stat.input_configs[i]
                 if input.gmw_p0 is not None and input.gmw_vec_p0 is not None and \
                     input.bmr_p0 is not None and input.bmr_vec_p0 is not None:
@@ -956,6 +956,7 @@ def generate_single_network_graphs(all_stats, dir):
     generate_graph_for_attr(all_stats, get_send_size, 'Communication (MiB)', dir)
     generate_cumulative_graph_for_attr(all_stats, get_send_size, 'Communication (MiB)', dir)
 
+    # Not needed anymore
     # get_recv_size = lambda x: x.timing_stats.communication.recv_size if x is not None else 0
     # generate_graph_for_attr(all_stats, get_recv_size, 'Received Data (MiB)', dir)  
     # generate_cumulative_graph_for_attr(all_stats, get_recv_size, 'Received Data (MiB)', dir)

### 2022-04-30 overleaf/backend.tex
Edits to section 6

@@
 \begin{enumerate}
-\item Loop counters and array upper bounds are always \texttt{plain}.
+\item Loop counters and bounds are always \texttt{plain}.
 \item If any variable on the right-hand side of an assignment is shared, then the assigned variable is \texttt{shared}.
@@
 MOTION supports FOR loops and SIMD operations, so translation from MPC source to MOTION C++ code is relatively straightforward.
+\paragraph{Variable declarations:}
 % Variable naming
-\paragraph{Variable declarations:}
-Our compiler performs taint analysis on MPC Source based on the type system in~\secref{sec:syntax}. The analysis is described in the following section;
-for the most part, it is a standard taint analysis that propagates the declared \texttt{shared} qualifiers throughout the routine.
-Taint analysis assigns \texttt{shared} or \texttt{plain} qualifiers to all local variables, which is important for code generation.
 Our generated C++ uses the following variable-naming scheme: shared variables are named the same as in the MPC Source with the {\sf !} replaced with an underscore (e.g.
@@
 % Function preamble
-\paragraph{Code generation:}
 The generated MOTION code begins with the declaration of all variables used in the function, including loop counters.
@@
 Finally, plaintext parameters are converted used as shared inputs from party 0 to initialize their shared counterparts.
-(Importantly, as we mention, they are converted only if they are needed as an argument to a shared operation; one can think of it as an upcast from \texttt{plain} to \texttt{share}.)
+(Importantly, as we mention, they are converted only if they are needed as an argument to a shared operation; one can think of it as an upcast from \texttt{plain} to \texttt{share}.) \ben{The note in parentheses is not true: we currently assume that all plaintext variables other than loop counters will be used in a shared expression.
+We initialize a shared version of them in the preamble (using the \texttt{\_MPC\_CONSTANT} variables or the plaintext parameter, depending on how the plaintext variable is initialized) and update that shared variable whenever the plaintext variable is updated.
+This is done to minimize the number of input gates (e.g.
+if a plaintext variable is upcast to shared in a loop, that would insert N input gates).
+With protocols that allow publicly known shared constants, this would not be necessary.
+This is described in the final subsection of this section.}
 
+\paragraph{Code generation:}
 % Non-vectorized assignments and returns
 Once the function preamble is complete, the MPC Source is translated into C++ one statement at a time.
-The linear structure of MPC Source enables translation.
+The linear structure of MPC Source enables this approach to translation.
 If there is no vectorization present in a statement, translation to C++ is straightforward: outside of MUX statements and array updates, non-vectorized assignments, expressions, and returns directly translate into their C++ equivalents.
@@
 % Phi nodes and FOR loops
+MPC FOR loops are converted to C++ FOR loops which iterate the loop counter over the specified range.
 Pseudo PHI nodes are broken into two components: the ``FALSE'' branch which assigns the initial value of the PHI node and the ``TRUE'' branch which assigns the PHI node's back-edge.
 The assignment of the false branch occurs right before the PHI node's enclosing loop.
+As these assignments may rely on the loop counter, the loop counter is initialized before these statements.
 Inside of the PHI node's enclosing loop, a C++ \texttt{if} statement is inserted to only assign the true branch of the PHI node after the first iteration.
-MPC FOR loops are converted to C++ FOR loops which iterate the loop counter over the specified range.
-Listing~\ref{tab:motion_translation_for_loop} illustrates.
-The loop counter is initialized before any PHI nodes in the loop as they could rely on it; there is no other special handling of FOR loops.
-%\ana{Add the reference to the corresponding figure somewhere in this paragraph.}
+Listing~\ref{tab:motion_translation_for_loop} illustrates this translation.%\ana{Add the reference to the corresponding figure somewhere in this paragraph.}
 
@@
 % SIMD representation
+\paragraph{Vectorization and SIMD operations:}
 Vectorization is handled with utility functions to manage accessing and updating slices of arrays.
@@
 If yes, use that term?} \ben{I don't think that's the right term since SIMD and scalar MOTION values use the same C++ class, so the operator is the same.} \ana{Ok, leave as is.}
-Therefore, the translation of an expression containing SIMD values is identical to that of expressions without SIMD values.
+Therefore, once vectorized accesses are inserted the translation of an expression containing SIMD values is identical to that of expressions without SIMD values.
 
@@
 To make matters worse, most of the explanation on this comes in the next section.
-But I'll leave as is for now but leave the comments too because it needs work.}
+But I'll leave as is for now but leave the comments too because it needs work.} \ben{Attempted to explain the naming, please review.}
 \paragraph{Reshaping and raising dimensions:}
 Raising the dimensions of a scalar or array uses the {\sf lift()} utility function which takes a lambda for the raised expression and the dimensions of the output.
+This function is also used for the scalar expansion of values which have been lifted out of FOR loops as described in \ref{sec:expansion}.
 This function evaluates the expression for each permutation of indices along the dimensions and returns the resulting array in row-major order.
@@
 % drop_dim() implementation
-\paragraph{Dropping dimensions:}
 Dropping dimensions use the {\sf drop\_dim()} and {\sf drop\_dim\_monoreturn()} utility functions.
@@
 % Shared/plaintext conversions
-\ben{Ishaq or Vassilis: please check over this paragraph... I don't want to make some mistake about MOTION's capabilities.}
+\paragraph{Upcasting from plaintext to shared:}
 Currently, our compiler only supports the \texttt{Bmr} and \texttt{BooleanGMW} protocols as MOTION does not implement all operations for other protocols.
@@
 Loop counters must still be converted to a shared value on each iteration that they are used, however we only generate this conversion when necessary, i.e., when the counter flows to a shared computation.
-This is to prevent unnecessary increase in the number of input gates.
+This is to prevent unnecessary increase in the number of input gates when loop counters are only used as plaintext.
+\ben{Ishaq or Vassilis: please check over this paragraph... I don't want to make some mistake about MOTION's capabilities.}
 

### 2022-04-30 overleaf/backend.tex
Minor changes to 6.

@@
 \label{sec:backend}
+
+MOTION code generation requires that variables are marked as \texttt{plain} or \texttt{shared} 
+following the type system in~\secref{sec:syntax}. We require that all inputs are marked as either shared or plaintext, 
+however, we infer qualifiers for the rest of the variables. \secref{sec:taint} briefly describes the taint analysis
+and \secref{sec:MOTION} describes MOTION code generation.
 
@@
 
-We require that all inputs are marked as either shared or plaintext, however, we do not require any other qualifier annotations in the program.
-Then we infer if intermediate variables and expressions are shared through taint analysis with "taintedness" referring to the shared attribute.
-Specifically, the compiler follows the following rules, which are standard in positive-negative qualifier systems (here \texttt{shared} is the positive qualifier and \texttt{plain} is the negative one):
+The taint analysis works on MPC Source, which lacks if-then-else control flow.
+This significantly simplifies treatment as there is no need to handle conditionals and implicit flow.
+%taint analysis with "taintedness" referring to the shared attribute.  
+Specifically, the compiler uses the following rules, which are standard in positive-negative qualifier systems (here \texttt{shared} is the positive qualifier and \texttt{plain} is the negative one):
 \begin{enumerate}
-\item Loop counters and bounds are always \texttt{plain}.
-\item If any variable on the right-hand side of an assignment is shared, then the assigned variable is \texttt{shared}.
+\item Loop counters and array bounds are always \texttt{plain}.
+\item If any variable on the right-hand side {\sf rhs} of an assignment is shared, then the assigned variable {\sf lhs} is \texttt{shared} following subtyping rule ${\sf rhs} <: {\sf lhs}$.
 %\item If all variables on the right-hand side of an assignment are plaintext, then the assigned variable is plaintext \ana{Will be covered by the last rule.}
@@
 %The first two rules are standard for taint analysis, and the third rule follows from the MPC problem statement. \ben{Is the explanation for the third rule correct?}  \ana{Yes, that is exactly it!}
-The final rule is needed in examples like the following. In the below snippet \texttt{sum!2} and \texttt{sum!3} form a dependency cycle and there is no \texttt{shared} value that flows to either one.
+The final rule is needed in examples like the following. In the below snippet {\sf sum!2} and {\sf sum!3} form a dependency cycle and there is no \texttt{shared} value that flows to either one.
 They are inferred as plaintext.
@@
 %\ben{Is this necessary to include?} \ana{Yes, that's important.}
-
-Importantly, the taint analysis works on MPC Source, which lacks if-then-else control flow.
-This significantly simplifies treatment as there is no need to handle conditionals and implicit flow.
 
@@
 Finally, plaintext parameters are converted used as shared inputs from party 0 to initialize their shared counterparts.
-(Importantly, as we mention, they are converted only if they are needed as an argument to a shared operation; one can think of it as an upcast from \texttt{plain} to \texttt{share}.) \ben{The note in parentheses is not true: we currently assume that all plaintext variables other than loop counters will be used in a shared expression.
+\ben{The note in parentheses is not true: we currently assume that all plaintext variables other than loop counters will be used in a shared expression.
 We initialize a shared version of them in the preamble (using the \texttt{\_MPC\_CONSTANT} variables or the plaintext parameter, depending on how the plaintext variable is initialized) and update that shared variable whenever the plaintext variable is updated.
@@
 With protocols that allow publicly known shared constants, this would not be necessary.
-This is described in the final subsection of this section.}
+This is described in the final subsection of this section.}\ana{Got it, thanks!
+I think it is good now.}
 
@@
 Pseudo PHI nodes are broken into two components: the ``FALSE'' branch which assigns the initial value of the PHI node and the ``TRUE'' branch which assigns the PHI node's back-edge.
-The assignment of the false branch occurs right before the PHI node's enclosing loop.
+The assignment of the ``FALSE'' branch occurs right before the PHI node's enclosing loop.
 As these assignments may rely on the loop counter, the loop counter is initialized before these statements.
@@
 Similarly, the {\sf vectorized\_assign()} function assigns a (potentially SIMD) value to a slice of a vectorized array.
-This operation cannot be done with a simple subscript as SIMD assignments will update a range of values in the non-vectorized representation.
+This operation cannot be done with a simple subscript as SIMD assignments will update a range of values in the underlying array representation.
+\ana{Changed non-vectorized -> array representation.
+Check if correct.}
 %\ben{Is the next sentence needed?  They can always check the implementation of vectorized\_assign() if they want to see specifics.}
@@
 Raising the dimensions of a scalar or array uses the {\sf lift()} utility function which takes a lambda for the raised expression and the dimensions of the output.
-This function is also used for the scalar expansion of values which have been lifted out of FOR loops as described in \ref{sec:expansion}.
+This function is also used for the scalar expansion of values which have been lifted out of FOR loops as described in \secref{sec:expansion}.
 This function evaluates the expression for each permutation of indices along the dimensions and returns the resulting array in row-major order.
@@
 There is also a special case of the {\sf lift()} function which occurs when we are raising an array.
-In this case, instead of concatenating the array for each index, we extend the array along all dimensions being raised which are not present in the array already.
-For example, when raising an array with dimensions $N \times M$ to an array with dimensions $N \times M \times D$, the input array will simply be extended along the $D$ dimension: {\sf A'[n,m,d] = A[n,m]} for every $d$.
+In this case, instead of concatenating the array for each index, \ana{can we remove "instead of concatenating the array"} we extend the array along all dimensions being raised which are not present in the array already.
+For example, when raising an array with dimensions $N \times M$ to an array with dimensions $N \times M \times D$, the input array will simply be extended along the $D$ dimension: ${\sf A'}[n,m,d] = {\sf A}[n,m]$ for every $d$.
 If the input array is already correctly sized it will be returned as-is.

### 2022-04-30 overleaf/backend.tex
Edits to 6

@@
 \begin{enumerate}
-\item Loop counters and array bounds are always \texttt{plain}.
+\item Loop counters are always \texttt{plain}. \ben{I removed the reference to array bounds since those are not variables defined in MPC source}
 \item If any variable on the right-hand side {\sf rhs} of an assignment is shared, then the assigned variable {\sf lhs} is \texttt{shared} following subtyping rule ${\sf rhs} <: {\sf lhs}$.
@@
 Finally, plaintext parameters are converted used as shared inputs from party 0 to initialize their shared counterparts.
-\ben{The note in parentheses is not true: we currently assume that all plaintext variables other than loop counters will be used in a shared expression.
-We initialize a shared version of them in the preamble (using the \texttt{\_MPC\_CONSTANT} variables or the plaintext parameter, depending on how the plaintext variable is initialized) and update that shared variable whenever the plaintext variable is updated.
-This is done to minimize the number of input gates (e.g.
-if a plaintext variable is upcast to shared in a loop, that would insert N input gates).
-With protocols that allow publicly known shared constants, this would not be necessary.
-This is described in the final subsection of this section.}\ana{Got it, thanks!
-I think it is good now.}
 
@@
 This operation cannot be done with a simple subscript as SIMD assignments will update a range of values in the underlying array representation.
-\ana{Changed non-vectorized -> array representation.
-Check if correct.}
 %\ben{Is the next sentence needed?  They can always check the implementation of vectorized\_assign() if they want to see specifics.}
@@
 There is also a special case of the {\sf lift()} function which occurs when we are raising an array.
-In this case, instead of concatenating the array for each index, \ana{can we remove "instead of concatenating the array"} we extend the array along all dimensions being raised which are not present in the array already.
+In this case, instead of concatenating the array for each index, \ana{can we remove "instead of concatenating the array"} \ben{I included it to show why the standard implementation of lift() doesn't work, if you don't think that's necessary then it can be removed.} we extend the array along all dimensions being raised which are not present in the array already.
 For example, when raising an array with dimensions $N \times M$ to an array with dimensions $N \times M \times D$, the input array will simply be extended along the $D$ dimension: ${\sf A'}[n,m,d] = {\sf A}[n,m]$ for every $d$.

### 2022-04-30 overleaf/results.tex
WIP evaluation section

@@
 \label{sec:results}
+
+\subsection{Experiment Setup}
+We tested our framework with several benchmarks.
+For the multiparty computation (MPC), we restriced our evaluation to 2 party computation (2PC) setting because it requires fewer computing resources.
+We stress that there is no such inherent restriction in our framework.
+We use hardware resources provided by CloudLab\cite{DuplyakinATC19} and consider two network settings, namely Local Area Netowrk (LAN) and Wide Area Network (WAN).
+In the LAN setting, we use \texttt{c6525-25g} machines for both parties.
+These machines are equipped with 16-core AMD 7302P 3.0GHz processors and 128GB of RAM.
+The connection between these machines had 10Gbps bandwidth and sub-millisecond latency.
+This setting reflects typical LAN usecase considering that 10Gbps LAN is increasingly common in business networks and is now available even in some home networks.
+For WAN setting, we again used a \texttt{c6525-25g} machine (located in Utah, US) for the first party and a \texttt{c220g1} machine (located in Wisconsin, US) for the second.
+The \texttt{c220g1} machine is equiped with two Intel E5-2630 8-core 2.40GHz processors and 128GB of RAM.
+We measured the connection bandwidth between these machines to be 560Mbps and average round trip time (RTT) to be 38ms.
+At the time of this writing, all major internet providers in the US offer 1Gbps connections to home consumers, therefore this setting reasonably reflects the typical WAN usecase.
+
+\subsection{Results and Analysis}
+TODO

### 2022-04-30 overleaf/results.tex
wip experiments section

@@
 
-\subsection{Experiment Setup}
+\subsection{Experiment Setup}\label{sec:experiment_setup}
 We tested our framework with several benchmarks.
@@
 
-\subsection{Results and Analysis}
+We repeated all experiments (except one) five times and report average and standard deviation values of various metrics.
+The one experiment where we measure metrics across a range of input configurations (from small to large), was run only once.
+Note however that due to the very low standard deviation (more on this below) seen in the multiple-run experiments, the accuracy of results of this experiments is not unreasonably affected.
+
+\subsection{Results and Analysis}\label{sec:result_analysis}
 TODO

### 2022-04-30 overleaf/backend.tex

@@
 \begin{enumerate}
-\item Loop counters are always \texttt{plain}. \ben{I removed the reference to array bounds since those are not variables defined in MPC source}
+\item Loop counters are always \texttt{plain}. %\ben{I removed the reference to array bounds since those are not variables defined in MPC source}\ana{Ok, resolved.}
 \item If any variable on the right-hand side {\sf rhs} of an assignment is shared, then the assigned variable {\sf lhs} is \texttt{shared} following subtyping rule ${\sf rhs} <: {\sf lhs}$.
@@
 %The first two rules are standard for taint analysis, and the third rule follows from the MPC problem statement. \ben{Is the explanation for the third rule correct?}  \ana{Yes, that is exactly it!}
-The final rule is needed in examples like the following. In the below snippet {\sf sum!2} and {\sf sum!3} form a dependency cycle and there is no \texttt{shared} value that flows to either one.
+%The final rule is needed in examples like the following. 
+In the below snippet {\sf sum!2} and {\sf sum!3} form a dependency cycle and there is no \texttt{shared} value that flows to either one.
 They are inferred as plaintext.
@@
 
-\ben{The compiler and C++ internally uses ``lift'' as the terminology for raise\_dim(), but the paper does not :(...} \ana{True.
+%\ben{The compiler and C++ internally uses ``lift'' as the terminology for raise\_dim(), but the paper does not :(...} \ana{True.
 To make matters worse, most of the explanation on this comes in the next section.
-But I'll leave as is for now but leave the comments too because it needs work.} \ben{Attempted to explain the naming, please review.}
+But I'll leave as is for now but leave the comments too because it needs work.} \ben{Attempted to explain the naming, please review.} \ana{Marked as resolved.}
 \paragraph{Reshaping and raising dimensions:}
@@
 There is also a special case of the {\sf lift()} function which occurs when we are raising an array.
-In this case, instead of concatenating the array for each index, \ana{can we remove "instead of concatenating the array"} \ben{I included it to show why the standard implementation of lift() doesn't work, if you don't think that's necessary then it can be removed.} we extend the array along all dimensions being raised which are not present in the array already.
+In this case, instead of concatenating the array for each index, %\ana{can we remove "instead of concatenating the array"} \ben{I included it to show why the standard implementation of lift() doesn't work, if you don't think that's necessary then it can be removed.} \ana{ok, left it in, marked as resolved.}
+we extend the array along all dimensions being raised which are not present in the array already.
 For example, when raising an array with dimensions $N \times M$ to an array with dimensions $N \times M \times D$, the input array will simply be extended along the $D$ dimension: ${\sf A'}[n,m,d] = {\sf A}[n,m]$ for every $d$.

### 2022-04-30 overleaf/introduction.tex
Made a minor edit.
Edits to 1-5.

@@
 
-We emphasize the MPC Source intermediate representation and optimization over MPC Source. As in classical compilers, we envision different front ends (e.g., Wysteria,
-our front end IMP-MPC) compiling into MPC Source. MPC Source is particularly suitable for optimizations such as protocol mixing~\cite{Ishaq:2019, Fang:2022} and SIMD-vectorization,
+We use the MPC Source~\cite{Ishaq:2019} intermediate representation and emphasize optimization over MPC Source.
+As in classical compilers, we envision different front ends (e.g., our front end IMP Source, other) compiling into MPC Source.
+MPC Source is particularly suitable for optimizations such as protocol mixing~\cite{Ishaq:2019, Fang:2022} and SIMD-vectorization,
 which takes advantage of amortization at the circuit level. The MPC Source IR exposes the linear structure of MPC programs, which simplifies program analysis; this is
@@
 
-\subsection{Our Contribution} In this paper, we develop a compiler framework that takes a Python-like routine and produces MOTION code: we describe (a) the IMP-MPC language, its syntax and semantic restrictions, (b) translation from IMP-MPC into MPC Source, (c) a specific backend-independent optimization: novel SIMD-vectorization on MPC Source, and (d) translation from MPC Source into MOTION code.
+\subsection{Our Contribution} In this paper, we develop a compiler framework that takes a Python-like routine and produces MOTION code: we describe (a) the IMP Source language, its syntax and semantic restrictions, (b) translation from IMP Source into MPC Source, (c) a specific backend-independent optimization: novel SIMD-vectorization on MPC Source, and (d) translation from MPC Source into MOTION code.
 %\ana{I think we have to add more on MOTION, specifically mixing and amortization.}
@@
 We demonstrate expressivity of the source language by running the compiler on \ana{X} programs; these include classical MPC benchmarks such as PSI and Biometric matching, as well as kMeans, Histogram, and others.
-Vectorization leads to \ana{X}x improvement in number of gates on average, \ana{Y}x speedup in circuit generation time, \ana{Z}x speedup in the LAN setting, and \ana{Z}x speedup in the WAN setting using the Boolean GMW protocol.
-Results form BMR are similar.
+\ana{Summarize results.}
 

### 2022-04-30 overleaf/vectorization.tex
Edits to 1-5.

@@
 phase. First, it may break a loop into smaller loops which would discover opportunities 
-for vectorization in intermediate statements in the loop. Second, it handles write arrays. 
+for vectorization in intermediate statements in the loop. Second, it handles writes arrays. 
 It creates opportunities for vectorization in the presence of write arrays, even though 
@@
 The excerpts in {\color{red} red} color in the pseudo code below highlight the extension 
-with array writes. We advise the reader to omit the extension and assume just read-only
-arrays. We explain the extension in \secref{sec:array_writes} details our algorithm. 
-(As many of our benchmarks include array writes, it plays an important role.)
+with array writes. We advise the reader to omit the extension for now and consider just read-only
+arrays. We explain the extension in \secref{sec:array_writes}. 
+(As many of our benchmarks include write arrays, it plays an important role.)
 

### 2022-04-30 overleaf/overview.tex
Edits across several sections.

@@
 \end{tabular}
-\caption{Biometric Matching: From IMP Source to MPC Source to Optimized MPC Source. }
+\caption{Biometric Matching: ==== From (a) IMP Source to (b) MPC Source: 
+First, MPC Source is an SSA form.
+Second, it is linear. The conditional in lines 13-14 in IMP Source turns into the linear code in lines 12-16 in MPC Source.
+The test turns into the CMP operation {\sf t = CMP(sum!3,min\_sum!2)}, followed by the
+true-branch sequence, followed by the MUX operations. The first MUX operation selects the value
+of {\sf min\_sum}: if {\sf t} is true, then {\sf min\_sum} gets the value of the second multiplexer
+argument,  {\sf min\_sum!3}, otherwise it takes the value of the third argument, {\sf min\_sum!2}.
+Third, MPC Source is a special form of SSA. The SSA $\phi$-nodes at the if-then-else (lines 13-15) turn into
+MUX operations, while the $\phi$-nodes at for-loops turn into \emph{pseudo} PHI nodes with a straightforward semantics.
+==== From (b) MPC Source to (c) Optimized MPC Source: 
+The compiler determines that SUB and MUL in ``naive'' MPC Source (lines 9 and 10 in (b))
+can be fully vectorized into the SIMD SUB and MUL in optimized MPC Source (lines 9 and 10 in (c)).
+Notation {\sf p[I,J]} denotes a 2-dimensional array with fully vectorized dimensions.
+The computation of sum (line 11 in (b))
+is sequential across the $j$-dimension, but it is parallel across the $i$-dimension.
+The loop in lines 12-16 in (c) illustrates; here {\sf p[I,j]} refers to the $j$-th column in {\sf p}.
+Unfortunately, CMP and MUX remain sequential.
+% \ana{Add explanation to caption, this is hard to see.}
+}
 \label{tab:source_and_MPC_source_and_optimized_MPC_source}
@@
 database and computes the Euclidean distance of the current
-entry {\sf S[i]} and {\sf C} (its square actually). The program returns the index of the vector that gives
+entry ${\sf S}[i]$ and {\sf C} (its square actually). The program returns the index of the vector that gives
 the best match plus the corresponding sum of squares.
@@
 
-Our compiler generates an intermediate representation, MPC Source. MPC Source
-for Biometric Matching is shown in
-Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(b). First, MPC Source is an SSA form.
+Our compiler generates an intermediate representation, MPC Source. 
+MPC Source is a \emph{linear} SSA form.
+MPC Source for Biometric Matching is shown and described in detail in
+Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(b).
+
+\begin{comment}
+First, MPC Source is an SSA form.
 Second, it is linear. The conditional in lines 13-14 in IMP Source turns into the linear code in lines 12-16 in MPC Source.
@@
 MUX operations, while the $\phi$-nodes at for-loops turn into \emph{pseudo} PHI nodes with a straightforward semantics.
+\end{comment}
 
@@
 
-Our compiler produces the vectorized program shown in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(c).
+Our compiler produces the vectorized program shown and described in 
+Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(c).
 Note that this is still our intermediate representation, Optimized MPC Source. Subsequently,
@@
 uses to generate the circuit.
-
+\begin{comment}
 The compiler determines that SUB and MUL in ``naive'' MPC Source (lines 9 and 10 in (b))
@@
 Unfortunately, CMP and MUX remain sequential.
+\end{comment}
 
-In MPC backends, executing $n$ operations ``at once'' in a single SIMD operation costs a lot less than executing those $n$ operations one by one.
+In MPC back ends, executing $n$ operations ``at once'' in a single SIMD operation costs a lot less than executing those $n$ operations one by one.
 This is particularly important when there is communication (i.e., in remote), since many 1-bit values are sent at once rather than sequentially.
@@
 The first term in the sum corresponds to the vectorized
-subtraction and multiplication (lines 9-10 in (c)), the second term corresponds to the FOR loop on $j$ (lines 12-16) and the third
-one corresponds to the remaining FOR loops on $i$ (lines 19-25). 
+subtraction and multiplication (lines 9-10 in (c)), the second term corresponds to the for-loop on $j$ (lines 12-16) and the third
+one corresponds to the remaining for-loops on $i$ (lines 19-25). 
 Clearly, $2\alpha_\mathit{fix} + D\beta_\mathit{fix} + N(3\alpha_\mathit{fix}) << ND(2\alpha_\mathit{fix}+\beta_\mathit{fix}) + N3\alpha_\mathit{fix}$.

### 2022-05-01 overleaf/results.tex
added LAN benchmarks to the paper
add benchmark descriptions

--- 2022-04-30 overleaf/results.tex
@@
 We use hardware resources provided by CloudLab\cite{DuplyakinATC19} and consider two network settings, namely Local Area Netowrk (LAN) and Wide Area Network (WAN).
-In the LAN setting, we use \texttt{c6525-25g} machines for both parties.
+In the LAN setting, we use {\tt c6525-25g} machines for both parties.
 These machines are equipped with 16-core AMD 7302P 3.0GHz processors and 128GB of RAM.
@@
 This setting reflects typical LAN usecase considering that 10Gbps LAN is increasingly common in business networks and is now available even in some home networks.
-For WAN setting, we again used a \texttt{c6525-25g} machine (located in Utah, US) for the first party and a \texttt{c220g1} machine (located in Wisconsin, US) for the second.
-The \texttt{c220g1} machine is equiped with two Intel E5-2630 8-core 2.40GHz processors and 128GB of RAM.
+For WAN setting, we again used a {\tt c6525-25g} machine (located in Utah, US) for the first party and a {\tt c220g1} machine (located in Wisconsin, US) for the second.
+The {\tt c220g1} machine is equiped with two Intel E5-2630 8-core 2.40GHz processors and 128GB of RAM.
 We measured the connection bandwidth between these machines to be 560Mbps and average round trip time (RTT) to be 38ms.
@@
 
-\subsection{Results and Analysis}
-TODO
+We run all experiments 5 times and report average values of various metrics.
+Note that standard deviation in all observations was at most 1\% of the mean value, therefore the accuracy of the results is not effected by the relatively fewer runs.
+
+\subsection{Benchmarks}\label{sec:benchmarks_description}
+In the following, we say {\em both} for an experiment in which we execute both non-vectorized and vectorized protocols and {\em vec} for the vectorized only experiment.
+We used the following benchmarks in our evaluation:
+
+\ishaq{We don't have space to show individual benchmark graphs, and vec data is only plotted in individual benchmark graphs, so specifying what we use for vec experiment in the following text is unnecessary.
+(unless, we change the graph)}
+
+\ishaq{TODO: go through description again, and add why each benchmark may be important from MPC/private-learning/distributed-computing perspective}
+\begin{enumerate}
+    \item {\em Biometric Matching} Server has a database S of {\tt N} records, each record's dimension is {\tt D}.
+Client submits a query C, client and server compute the closest record to C in an MPC.
+We use {\tt N=128} for {\em both} and {\tt N=4096} for {\em vec}.
+{\tt D} is fixed at 4.
+
+    \item {\em Biometric Matching (Fast)} is the faster instantiation taken from \ishaq{cite ABY}. Parameters are same as above.
+
+    \item {\em Convex Hull} given a polygon of {\tt N} vertices (split between Alice and Bob), convex hull is computed in an MPC.
+adapted from \ishaq{cite: https://www.cs.toronto.edu/~azadeh/resources/papers/pldi21-full.pdf} We use {\tt N=32} for {\em both} experiment and {\tt N=256} for {\em vec} experiment.
+
+    \item {\em Count 102} Alice has a string of length {\tt N} of symbols, Bob has a regular expression of the form 1(0*)2, together they compute number of substrings that match the regular expression.
+adopted from \ishaq{cite: same as above}.
+We use {\tt N=1024} for {\em both} and {\tt N=4096} for {\em vec}.
+
+    \item {\em Count 10} Same as {\em Count 102} except now the regular expression is of the form 1(0+). Parameters are same as above.
+
+    \item {\em Cryptonets Max Pooling} Given a matrix of $\mathtt{rows}\times\mathtt{cols}$ elements that are split between Alice and Bob, they compute the max pooling subroutine of the cryptonet benchmark\ishaq{cite: https://www.microsoft.com/en-us/research/wp-content/uploads/2016/04/CryptonetsTechReport.pdf}.
+We use {\tt rows=64, cols=64} for {\em both} experiment.
+
+    \item {\em Database Join} given two databases with {\tt A} and {\tt B} 2-element records, compute cross join. We use {\tt A=B=32} for {\em both} and {\tt A=B=64} for {\em vec}.
+
+    \item {\em Database Variance} given a database of {\tt len} records, compute variance.
+\ishaq{Ben/Ana: Please correct the description, I don't know what this benchmark is} We use {\tt len=512} for {\em both} and {\tt len=4096} for {\em vec}.
+
+    \item {\em Histogram} given {\tt N} 5-star ratings, compute their histogram, taken from: \ishaq{Ana, cite OPA?
+or this one: https://www.cs.toronto.edu/~azadeh/resources/papers/pldi21-full.pdf ?} We use {\tt N=512} for {\em both} and {\tt N=4096} for {\em vec}.
+
+    \item {\em Inner Product} given two vectors, each of {\tt N} elements, compute their inner product. We use {\tt N=512} for {\em both} and {\tt N=4096} for {\em vec}.
+
+    \item {\em k-means Iteration} performs the iteration subroutine of k-means database clustering operation \cite{https://dl.acm.org/doi/10.1145/1081870.1081942, https://dl.acm.org/doi/10.1145/956750.956776}.
+Here {\tt len1} is the size of input data, and {\tt len2} is the number of clusters.
+We use {\tt len1=32, len2=5} for {\em both} and {\tt len1=256, len2=8} for {\em vec}.
+\ishaq{Ana/Ben: please update description}
+
+    \item {\em Longest 102} Similar to {\em Count 102} except that it computes the largest substring matching the regular expression. We use same parameters as {\em Count 102}.
+
+    \item {\em Max Distance b/w Symbols} Alice has a string of {\tt N} symbols and Bob has some symbol {\tt 0}.
+The MPC computes the maximum distance between {\tt 0}s in the string.
+adopted from {\ishaq, cite here too}.
+We use {\tt N=1024} for {\em both} and {\tt N=2048} for {\em vec}.
+
+    \item {\em Minimal Points} Given a set of {\tt N} points (split between Alice and Bob), a set of minimal points is computed i.e.
+there is no other point that has both a lower x and y coordinate.
+We use {\tt N=32} for {\em both} and {\tt N=64} for {\em vec}.
+
+    \item {\em MNIST ReLU} given an input of ${\tt outer}\times{\tt inner}$ elements, executes the MNIST ReLU subroutine.
+We use {\tt inner=512} for {\em both} and {\tt inner=2048} for {\em vec}.
+{\tt outer} is fixed at 16.
+\ishaq{Ben/Ana, please update}
+
+    \item {\em Private Set Intersection (PSI)} Alice holds set S1 with size {\tt SA}, Bob holds set S2 with size {\tt SB}, together they compute intersection of their sets.
+We use {\tt SA=SB=128} for {\em both} and {\tt SA=SB=1024} for {\em vec}.
+\end{enumerate}
+
+\subsection{Results and Analysis}\label{sec:result_analysis}
+
+- Discuss Overall LAN results (Time, Communication Size)
+- Discuss inner product and biometric in detail on LAN
+- Discuss their comparison (Time) with WAN
+- Discuss hand simdified code (native MOTION)
+- Put error bars
+
+\begin{figure*}[htbp]
+\centering
+\include{graphs/all-hist-OnlineSetupTimesec}
+\caption{Circuit Evaluation Time (Setup + Online) of Benchmarks}
+\label{fig:graph_all_online_setup}
+\end{figure*}
+
+\begin{figure*}[htbp]
+\centering
+\include{graphs/all-hist-CommunicationMiB}
+\caption{Communication Size of Benchmarks}
+\label{fig:graph_comm_size_time}
+\end{figure*}
+
+\begin{figure*}[htbp]
+\centering
+\include{graphs/all-hist-CircuitGenerationTimesec}
+\caption{Circuit Generation Time of Benchmarks}
+\label{fig:graph_circ_gen_time}
+\end{figure*}
+
+\begin{figure*}[htbp]
+\centering
+\include{graphs/all-hist-TotalGates}
+\caption{Number of Gates of Benchmarks}
+\label{fig:graph_total_gates}
+\end{figure*}
+
+\begin{figure*}[htbp]
+\centering
+\include{graphs/all-hist-OnlineTimesec}
+\caption{Online Time of Benchmarks}
+\label{fig:graph_online_time}
+\end{figure*}
+
+\begin{figure*}[htbp]
+\centering
+\include{graphs/all-hist-SetupTimesec}
+\caption{Setup Time of Benchmarks}
+\label{fig:graph_setup_time}
+\end{figure*}

### 2022-05-01 overleaf/results.tex
add updated graphs
add error bars to the graphs

@@
 We run all experiments 5 times and report average values of various metrics.
-Note that standard deviation in all observations was at most 1\% of the mean value, therefore the accuracy of the results is not effected by the relatively fewer runs.
+Note that standard deviation--shown as errorbar on top of the bars in the graphs--in all observations was at most 4.5\% of the mean value, therefore the accuracy of the results is not effected by the relatively fewer runs.
 
@@
 \centering
-\include{graphs/all-hist-OnlineSetupTimesec}
+\include{graphs/lan/all-hist-OnlineSetupTimesec}
 \caption{Circuit Evaluation Time (Setup + Online) of Benchmarks}
@@
 \centering
-\include{graphs/all-hist-CommunicationMiB}
+\include{graphs/lan/all-hist-CommunicationMiB}
 \caption{Communication Size of Benchmarks}
@@
 \centering
-\include{graphs/all-hist-CircuitGenerationTimesec}
+\include{graphs/lan/all-hist-CircuitGenerationTimesec}
 \caption{Circuit Generation Time of Benchmarks}
@@
 \centering
-\include{graphs/all-hist-TotalGates}
+\include{graphs/lan/all-hist-TotalGates}
 \caption{Number of Gates of Benchmarks}
@@
 \centering
-\include{graphs/all-hist-OnlineTimesec}
+\include{graphs/lan/all-hist-OnlineTimesec}
 \caption{Online Time of Benchmarks}
@@
 \centering
-\include{graphs/all-hist-SetupTimesec}
+\include{graphs/lan/all-hist-SetupTimesec}
 \caption{Setup Time of Benchmarks}

### 2022-05-01 overleaf/results.tex
wip experiments section

@@
 \ishaq{TODO: go through description again, and add why each benchmark may be important from MPC/private-learning/distributed-computing perspective}
+
 \begin{enumerate}
@@
 
-- Discuss Overall LAN results (Time, Communication Size)
-- Discuss inner product and biometric in detail on LAN
-- Discuss their comparison (Time) with WAN
-- Discuss hand simdified code (native MOTION)
-- Put error bars
-
 \begin{figure*}[htbp]
@@
 \caption{Circuit Evaluation Time (Setup + Online) of Benchmarks}
-\label{fig:graph_all_online_setup}
+\label{fig:graph_all_eval_time}
 \end{figure*}
+
+We show circuit evaluation times in \cref{fig:graph_all_eval_time}.
+In terms of amenability to vectorization, we divide benchmarks into 3 categories: 1) {\it High:} these include convex hull, cryptonets max pooling, minimal points and private set intersection.
+These benchmarks are highly parallelizable and see 25x to 70x speedup in BMR, and 30x to 55x in GMW protocol.
+2) {\it Medium:} these include biometric matching, DB Variance, histogram, inner product, k-means iteration and MNIST ReLU.
+These benchmarks have non-parallelizable phases e.g.
+the summing phase of inner product and biometric matching.
+Still, most computation is parallelizable and it results in speedup from 5x to 25x in BMR, and 2x to 25x in GMW protocol.
+3) {\it Low:} these include the Database Join and the regular expression benchmarks (count 102, count 10, longest 102 and max distance between symbols).
+There is very little parallelizable computation in these programs, thus the speedup is lower.
+We see a speedup from 1.1x to 2x in BMR.
+In GMW, DB Join, Count 102 and Count 10s see speedup from 1.1x to 1.3x.
+However, longest 102 and max distance between symbols suffer a slowdown of 0.5x.
+The structure of these benchmarks is such that transformation to vectorized code increases multiplicative depth and, the negative effect of increased depth is more noticeable in a round-based protocol like GMW.
+We fix this via a simple heuristic where if the transformation increases circuit depth beyond some threshold (e.g.
+more than 10\% of the original circuit), we reject the transformation.
+Nevertheless, we show these graphs here for the sake of completeness and to highlight that vectorization is not always reduce run time.
+Note that in some settings it may still be desirable to vectorize e.g.
+in data constrained environments.
+As shown in \cref{fig:graph_comm_size_time}, vectorization results in reduced communication (i.e.
+fewer bits are transferred).
+
+We present evaluation for communication size in \cref{fig:graph_comm_size_time}, circuit generation time in \cref{fig:graph_circ_gen_time}, number of gates in \cref{fig:graph_total_gates} and online time and setup time in \cref{fig:graph_online_time} and \cref{fig:graph_setup_time} respectively.
+\ishaq{Ana: Let me know if I should add more substance.}
+
+- Discuss inner product and biometric in detail on LAN
+- Discuss their comparison (Time) with WAN
 

### 2022-05-01 compiler/paper_benchmarks.py
benchdata ready (png graphs)

@@ -797,9 +797,9 @@ def run_client_role(address):
 
 def get_x_label_for_benchmark(name):
     if name == "biometric":
-        return "Biometric Distance"
+        return "Biometric Matching"
     if name == "biometric_fast":
-        return "Biometric Distance (Fast)"
+        return "Biometric Matching (Fast)"
     if name == "convex_hull":
         return "Convex Hull"
     if name == "count_102":
@@ -811,7 +811,7 @@ def get_x_label_for_benchmark(name):
     if name == "cryptonets_max_pooling":
         return "Cryptonets (Max Pooling)"
     if name == "db_cross_join_trivial":
-        return "DB Cross Join (Trivial)"
+        return "Database Join"
     if name == "db_variance":
         return "Database Variance"
     if name == "histogram":
@@ -854,7 +854,7 @@ def ratio(a, b):
     return 0
 
 
-def generate_graph_for_attr(all_stats, get_val, y_label, dir):
+def generate_graph_for_attr(all_stats, get_val, get_sd, y_label, dir):
     y_label_simple = ''.join(c for c in y_label if c.isalnum())
     for task_stat in all_stats:
         fname = "{}-{}.txt".format(task_stat.label, y_label_simple)
@@ -862,8 +862,28 @@ def generate_graph_for_attr(all_stats, get_val, y_label, dir):
         line_graph_file = "{}-line-{}{}".format(task_stat.label, y_label_simple, GRAPH_EXT)
         file_path = os.path.join(dir, fname)
         with open(file_path, mode='w', encoding='utf-8') as f:
-            f.write("x\tInput\t\"GMW\"\t\"GMW (Vectorized)\"\t\"BMR\"" \
-                "\t\"BMR (Vectorized)\"\t\"GMW Improvement\"\t\"BMR Improvement\"\n")
+            line = '\t'.join(['x', 
+                '"Benchmark"', 
+                '"GMW"', 
+                '"GMW (Vectorized)"',
+                '"BMR"',
+                '"BMR (Vectorized)"', 
+                '"GMW Improvement"', 
+                '"BMR Improvement"'])
+            if get_sd is not None:
+                line = '\t'.join(['x', 
+                    '"Benchmark"', 
+                    '"GMW"', 
+                    '"SD"',
+                    '"GMW (Vectorized)"', 
[…]

### 2022-05-01 compiler/paper_benchmarks.py
plots for tex ready

@@ -36,7 +36,7 @@ LAN_GRAPHS_DIR = os.path.join(LAN_DIR, "graphs")
 WAN_DIR = os.path.join(BENCHDATA_DIR, "wan")
 WAN_GRAPHS_DIR = os.path.join(WAN_DIR, "graphs")
 COMPARISON_GRAPHS_DIR = os.path.join(BENCHDATA_DIR, "comparison-graphs")
-GRAPH_EXT = ".png"
+GRAPH_EXT = ".tex"
 
 os.makedirs(BENCHDATA_DIR, exist_ok=True)
 os.makedirs(LAN_DIR, exist_ok=True)

### 2022-05-01 overleaf/results.tex
add inner product and biometric graphs

@@
 \centering
-\include{graphs/lan/all-hist-OnlineSetupTimesec}
+%\resizebox{7in}{!}{\input{graphs/lan/all-hist-OnlineSetupTimesec}}
+\input{graphs/lan/all-hist-OnlineSetupTimesec}
 \caption{Circuit Evaluation Time (Setup + Online) of Benchmarks}
@@
 - Discuss inner product and biometric in detail on LAN
+
+\begin{figure}[htbp]
+\centering
+\resizebox{3.6in}{!}{\input{graphs/lan/biometric-hist-OnlineSetupTimesec}}
+\caption{Circuit Evaluation Time for Biometric Matching, x-axis lists database size}
+\label{fig:graph_biometric_eval_time}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\resizebox{3.6in}{!}{\input{graphs/lan/inner_product-hist-OnlineSetupTimesec}}
+\caption{Circuit Evaluation Time for Inner Product, x-axis lists database size}
+\label{fig:graph_inner_product_eval_time}
+\end{figure}
+
 - Discuss their comparison (Time) with WAN
@@
 \centering
-\include{graphs/lan/all-hist-CommunicationMiB}
+\input{graphs/lan/all-hist-CommunicationMiB}
 \caption{Communication Size of Benchmarks}
@@
 \centering
-\include{graphs/lan/all-hist-CircuitGenerationTimesec}
+\input{graphs/lan/all-hist-CircuitGenerationTimesec}
 \caption{Circuit Generation Time of Benchmarks}
@@
 \centering
-\include{graphs/lan/all-hist-TotalGates}
+\input{graphs/lan/all-hist-TotalGates}
 \caption{Number of Gates of Benchmarks}
@@
 \centering
-\include{graphs/lan/all-hist-OnlineTimesec}
+\input{graphs/lan/all-hist-OnlineTimesec}
 \caption{Online Time of Benchmarks}
@@
 \centering
-\include{graphs/lan/all-hist-SetupTimesec}
+\input{graphs/lan/all-hist-SetupTimesec}
 \caption{Setup Time of Benchmarks}

### 2022-05-01 overleaf/results.tex
wip evaluation section

@@
 in data constrained environments.
-As shown in \cref{fig:graph_comm_size_time}, vectorization results in reduced communication (i.e.
-fewer bits are transferred).
+As shown in \cref{fig:graph_comm_size}, vectorization results in reduced communication (fewer bits are transferred).
 
@@
 
-- Discuss inner product and biometric in detail on LAN
-
 \begin{figure}[htbp]
@@
 \resizebox{3.6in}{!}{\input{graphs/lan/biometric-hist-OnlineSetupTimesec}}
-\caption{Circuit Evaluation Time for Biometric Matching, x-axis lists database size}
+\caption{Biometric Matching Circuit Evaluation Time, x-axis lists database size}
 \label{fig:graph_biometric_eval_time}
@@
 \centering
-\resizebox{3.6in}{!}{\input{graphs/lan/inner_product-hist-OnlineSetupTimesec}}
-\caption{Circuit Evaluation Time for Inner Product, x-axis lists database size}
-\label{fig:graph_inner_product_eval_time}
+\resizebox{3.6in}{!}{\input{graphs/lan/biometric-hist-CommunicationMiB}}
+\caption{Biometric Matching Communication Size, x-axis lists database size}
+\label{fig:graph_biometic_comm_size}
 \end{figure}
 
-- Discuss their comparison (Time) with WAN
+\begin{figure}[htbp]
+\centering
+\resizebox{3.6in}{!}{\input{graphs/lan/biometric-hist-CircuitGenerationTimesec}}
+\caption{Biometric Matching Circuit Generation Time, x-axis lists database size}
+\label{fig:graph_biometic_circ_gen_time}
+\end{figure}
+
+Next, we look closely at circuit evaluation \cref{fig:graph_biometric_eval_time}, communication size \cref{fig:graph_biometic_comm_size} and circuit generation time \cref{fig:graph_biometic_circ_gen_time} for biometric matching benchmark.
+For input size beyond {\tt N=128} the memory usage exceeds available memory and prevents circuit generation.
+Consequently, non-vectorized bars are missing beyond this threshold.
+Notice that vectorization improves all metrics.
+A database of size {\tt N=128} is smaller than any real world usecase, therefore lets use it as baseline for improvement.
+Comparing performance improvement between BMR and GMW, we see more speedup for BMR (23x vs 10x), GMW gets more communication size reduction (10x vs 2.5x) and circuit generation sees a speedup of 35x and 45x for BMR and GMW respectively.
+
+\begin{figure}[htbp]
+\centering
+\resizebox{3.6in}{!}{\input{graphs/comparison-hist-OnlineSetupTimesec}}
+\caption{LAN vs. WAN: Circuit Evaluation Time Comparison}
+\label{fig:graph_comparison_eval_time}
+\end{figure}
+
+Since our vectorization framework is network agnostic, it will produce the same circuit for both LAN and WAN settings.
+This means that the number of gates and communication size remain the same.
+Moreover, time for circuit generation, which is a local operation, also remains unchanged.
+Setup and Online times, however, increase due to lower bandwidth and higher latency of the WAN.
+Indeed, this is what we observe in \cref{fig:graph_comparison_eval_time}.
 
@@
 \caption{Communication Size of Benchmarks}
-\label{fig:graph_comm_size_time}
+\label{fig:graph_comm_size}
 \end{figure*}

### 2022-05-01 overleaf/results.tex
first pass at evaluation section

@@
 
+\ishaq{TODO: add a large table for LAN, incorporating as much data as possible}
+
 \begin{enumerate}
@@
 
-    \item {\em Biometric Matching (Fast)} is the faster instantiation taken from \ishaq{cite ABY}. Parameters are same as above.
+    \item {\em Biometric Matching (Fast)} is the faster instantiation taken from \cite{Demmler:2015}. Parameters are same as above.
 
     \item {\em Convex Hull} given a polygon of {\tt N} vertices (split between Alice and Bob), convex hull is computed in an MPC.
-adapted from \ishaq{cite: https://www.cs.toronto.edu/~azadeh/resources/papers/pldi21-full.pdf} We use {\tt N=32} for {\em both} experiment and {\tt N=256} for {\em vec} experiment.
+adapted from \cite{Farzan:2021}.
+We use {\tt N=32} for {\em both} experiment and {\tt N=256} for {\em vec} experiment.
 
     \item {\em Count 102} Alice has a string of length {\tt N} of symbols, Bob has a regular expression of the form 1(0*)2, together they compute number of substrings that match the regular expression.
-adopted from \ishaq{cite: same as above}.
+adapted from \cite{Farzan:2021}.
 We use {\tt N=1024} for {\em both} and {\tt N=4096} for {\em vec}.
@@
 
-    \item {\em Cryptonets Max Pooling} Given a matrix of $\mathtt{rows}\times\mathtt{cols}$ elements that are split between Alice and Bob, they compute the max pooling subroutine of the cryptonet benchmark\ishaq{cite: https://www.microsoft.com/en-us/research/wp-content/uploads/2016/04/CryptonetsTechReport.pdf}.
+    \item {\em Cryptonets Max Pooling} Given a matrix of $\mathtt{rows}\times\mathtt{cols}$ elements that are split between Alice and Bob, they compute the max pooling subroutine of the cryptonet benchmark\cite{Dowlin:2016}.
 We use {\tt rows=64, cols=64} for {\em both} experiment.
@@
 
-    \item {\em Histogram} given {\tt N} 5-star ratings, compute their histogram, taken from: \ishaq{Ana, cite OPA?
-or this one: https://www.cs.toronto.edu/~azadeh/resources/papers/pldi21-full.pdf ?} We use {\tt N=512} for {\em both} and {\tt N=4096} for {\em vec}.
+    \item {\em Histogram} given {\tt N} 5-star ratings, compute their histogram, taken from \cite{Ishaq:2019, Farzan:2021}. We use {\tt N=512} for {\em both} and {\tt N=4096} for {\em vec}.
 
@@
 
-    \item {\em k-means Iteration} performs the iteration subroutine of k-means database clustering operation \cite{https://dl.acm.org/doi/10.1145/1081870.1081942, https://dl.acm.org/doi/10.1145/956750.956776}.
+    \item {\em k-means Iteration} performs the iteration subroutine of k-means database clustering operation \cite{Jagannathan:2005, Vaidya:2003}.
 Here {\tt len1} is the size of input data, and {\tt len2} is the number of clusters.
@@
 
-    \item {\em Longest 102} Similar to {\em Count 102} except that it computes the largest substring matching the regular expression. We use same parameters as {\em Count 102}.
+    \item {\em Longest 102} Similar to {\em Count 102} except that it computes the largest substring matching the regular expression.
+We use same parameters as {\em Count 102}, adapted from \cite{Farzan:2021}.
 
@@
 The MPC computes the maximum distance between {\tt 0}s in the string.
-adopted from {\ishaq, cite here too}.
+We adapted it from \cite{Farzan:2021}.
 We use {\tt N=1024} for {\em both} and {\tt N=2048} for {\em vec}.
@@
     \item {\em Minimal Points} Given a set of {\tt N} points (split between Alice and Bob), a set of minimal points is computed i.e.
-there is no other point that has both a lower x and y coordinate.
+there is no other point that has both a lower x and y coordinate, adapted from \cite{Farzan:2021}.
 We use {\tt N=32} for {\em both} and {\tt N=64} for {\em vec}.
@@
 
-Since our vectorization framework is network agnostic, it will produce the same circuit for both LAN and WAN settings.
+Since our vectorization framework is network agnostic, it produces the same circuit for both LAN and WAN.
 This means that the number of gates and communication size remain the same.

### 2022-05-01 compiler/paper_benchmarks.py
updated comparison graph, latex graphs ready

@@ -11,6 +11,7 @@ import random
 import json
 import socket
 
+
 import compiler
 
 from tests import context as test_context
@@ -1015,22 +1016,22 @@ def generate_single_network_graphs(all_stats, dir):
     generate_graph_for_attr(all_stats, get_num_gates, None, 'Total Gates', dir)
     generate_cumulative_graph_for_attr(all_stats, get_num_gates, None, 'Total Gates', dir)
     
-    get_circ_gen_time = lambda x: int(x.circuit_stats.circuit_gen_time/1000) if x is not None else 0
+    get_circ_gen_time = lambda x: (x.circuit_stats.circuit_gen_time/1000) if x is not None else 0
     generate_graph_for_attr(all_stats, get_circ_gen_time, None, 'Circuit Generation Time (sec)', dir)
     generate_cumulative_graph_for_attr(all_stats, get_circ_gen_time, None, 'Circuit Generation Time (sec)', dir)
     
-    get_online_time = lambda x: int(x.timing_stats.gates_online.mean/1000) if x is not None else 0
-    get_online_time_sd = lambda x: int(x.timing_stats.gates_online.stddev/1000) if x is not None else 0
+    get_online_time = lambda x: (x.timing_stats.gates_online.mean/1000) if x is not None else 0
+    get_online_time_sd = lambda x: (x.timing_stats.gates_online.stddev/1000) if x is not None else 0
     generate_graph_for_attr(all_stats, get_online_time, get_online_time_sd, 'Online Time (sec)', dir)  
     generate_cumulative_graph_for_attr(all_stats, get_online_time, get_online_time_sd, 'Online Time (sec)', dir)
 
-    get_setup_time = lambda x: int((x.timing_stats.gates_setup.mean + x.timing_stats.preprocess_total.mean)/1000) if x is not None else 0
-    get_setup_time_sd = lambda x: int((x.timing_stats.gates_setup.stddev + x.timing_stats.preprocess_total.stddev)/1000) if x is not None else 0
+    get_setup_time = lambda x: ((x.timing_stats.gates_setup.mean + x.timing_stats.preprocess_total.mean)/1000) if x is not None else 0
+    get_setup_time_sd = lambda x: ((x.timing_stats.gates_setup.stddev + x.timing_stats.preprocess_total.stddev)/1000) if x is not None else 0
     generate_graph_for_attr(all_stats, get_setup_time, get_setup_time_sd, 'Setup Time (sec)', dir)    
[…]

### 2022-05-01 overleaf/introduction.tex
Minor edits in these secs.

@@
 
-Earlier compilers did both back-end and front-end translation without a specific focus on either, as their aim was to demonstrate applicability of MPC on real-world
+Earlier compilers did both back-end and front-end translation as their aim was to demonstrate applicability of MPC on real-world
 programming problems. As the field advanced, works have focused more closely on front-end language design (e.g., Wysteria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021})
@@
 
-We use the MPC Source~\cite{Ishaq:2019} intermediate representation and emphasize optimization over MPC Source.
+We formalize the MPC Source~\cite{Ishaq:2019} intermediate representation and emphasize optimization over MPC Source.
 As in classical compilers, we envision different front ends (e.g., our front end IMP Source, other) compiling into MPC Source.
 MPC Source is particularly suitable for optimizations such as protocol mixing~\cite{Ishaq:2019, Fang:2022} and SIMD-vectorization,
-which takes advantage of amortization at the circuit level. The MPC Source IR exposes the linear structure of MPC programs, which simplifies program analysis; this is
-in contrast to source, which has if-then-else constructs. In the same time, MPC Source is sufficiently ``high-level''  to support analyses and optimizations that take into account
-control and data flow in a specific program. MPC Source is small in size and analysis is tractable, as opposed to analysis over an unrolled circuit~\cite{Ishaq:2019}. Again as
-in classical compilers, we envision translation of MPC Source (optimized or unoptimized) into MOTION, SPDZ, or other back-end code.
+which takes advantage of amortization at the circuit level. The MPC Source IR exposes the \emph{linear structure} of MPC programs, which simplifies program analysis; this is
+in contrast to source, which has if-then-else constructs. In the same time, MPC Source is sufficiently ``high-level''  to support analysis and optimizations that take into account
+control and data flow in a specific program. MPC Source facilitates cost modeling. 
+%MPC Source is small in size and analysis is tractable, as opposed to analysis over an unrolled circuit~\cite{Ishaq:2019}. 
+Again as in classical compilers, we envision translation of MPC Source (optimized or unoptimized) into MOTION, SPDZ, or other back-end code.
 

### 2022-05-01 overleaf/results.tex
Minor edits in these secs.

@@
 This setting reflects typical LAN usecase considering that 10Gbps LAN is increasingly common in business networks and is now available even in some home networks.
-For WAN setting, we again used a {\tt c6525-25g} machine (located in Utah, US) for the first party and a {\tt c220g1} machine (located in Wisconsin, US) for the second.
+In the WAN setting we again used a {\tt c6525-25g} machine (located in Utah, US) for the first party and a {\tt c220g1} machine (located in Wisconsin, US) for the second.
 The {\tt c220g1} machine is equiped with two Intel E5-2630 8-core 2.40GHz processors and 128GB of RAM.
@@
 We run all experiments 5 times and report average values of various metrics.
-Note that standard deviation--shown as errorbar on top of the bars in the graphs--in all observations was at most 4.5\% of the mean value, therefore the accuracy of the results is not effected by the relatively fewer runs.
+Note that standard deviation --- shown as errorbar on top of the bars in the graphs --- in all observations was at most 4.5\% of the mean value, therefore the accuracy of the results is not effected by the relatively fewer runs.
 

### 2022-05-01 overleaf/introduction.tex
Added comments.

@@
 
-Our second contribution is an analytical model for cost estimation of amortized schedules. Originally, we hoped that optimal scheduling (under our model) was tractable, as the problem
-appeared simpler than the classical scheduling problem. Unfortunately, we show that optimal scheduling is NP-hard via a reduction to the Shortest Common Supersequence (SCS) problem.
+Our second contribution is an analytical model for cost estimation of amortized schedules.
+Originally, we hoped that optimal scheduling (under our model, which essentially minimizes the length of the schedule) was tractable, as the problem appeared simpler than the classical scheduling problem.
+Unfortunately, we show that optimal scheduling is NP-hard via a reduction to the Shortest Common Supersequence (SCS) problem.
 Cost modeling is important as it drives not only vectorization but optimizations such as protocol mixing and scheduling as well~\cite{Ishaq:2019, Fang:2022}.
@@
 Our most important contribution is the implementation and evaluation of the compiler framework.
-We demonstrate expressivity of the source language by running the compiler on \ana{X} programs; these include classical MPC benchmarks such as PSI and Biometric matching, as well as kMeans, Histogram, and others.
+We demonstrate expressivity of the source language by running the compiler on \ana{X} programs with interleaved if- and for-statements; these include classical MPC benchmarks such as PSI and Biometric matching, as well as kMeans, Histogram, and others.
 \ana{Summarize results.}

### 2022-05-01 overleaf/results.tex
Added comments.
evaluation section, added the big table

--- 2022-05-01 overleaf/results.tex
@@
 
+   \begin{table*}[htbp]
+       \centering
+       \caption{Vectorized vs Non-Vectorized Comparison, times in seconds (in LAN setting where applicable), Communication in MiB, Numbers in 1000s, values rounded to nearest integer, benchmark names ending in {\em V} are vectorized.}
+       \label{table:metrics}
+       \rowcolors{5}{}{gray!10}
+       \tabcolsep3pt
+       \begin{tabular}{lrrrrrrrrrrrr}
+           \toprule
+           {} & \multicolumn{6}{c}{\textbf{GMW}} & \multicolumn{6}{c}{\textbf{BMR}} \\
+            \cmidrule(r){2-7} \cmidrule(r){8-13} \\
+            {\bf Benchmark} & Online & Setup & \# Gates & Circ Gen & \# Msgs & Comm. &  Online & Setup & \# Gates & Circ Gen & \# Msgs & Comm. \\
+          \midrule
+            Biometric Matching & 146 & 16 & 1,784 & 119 & 1,413 & 140 & 89 & 263 & 1,595 & 139 & 2,716 & 312\\
+            Biometric Matching (V) & 12 & 4 & 34 & 2 & 28 & 14 & 2 & 13 & 30 & 4 & 61 & 130\\
+            \midrule
+            Convex Hull & 48 & 6 & 551 & 40 & 516 & 51 & 28 & 72 & 494 & 39 & 695 & 80\\
+            Convex Hull (V) & 0 & 1 & 2 & 0 & 1 & 4 & 0 & 2 & 1 & 1 & 2 & 32\\
+            \midrule
+            Count 102 & 79 & 6 & 418 & 35 & 525 & 52 & 15 & 62 & 269 & 33 & 785 & 92\\
+            Count 102 (V) & 71 & 5 & 316 & 24 & 332 & 34 & 11 & 30 & 167 & 16 & 304 & 59\\
+            \midrule
+            Count 10s & 79 & 6 & 419 & 35 & 525 & 52 & 14 & 62 & 270 & 33 & 785 & 92\\
+            Count 10s (V) & 71 & 4 & 316 & 24 & 332 & 34 & 11 & 29 & 167 & 16 & 304 & 59\\
+            \midrule
+            Cryptonets (Max Pooling) & 50 & 11 & 688 & 46 & 554 & 55 & 36 & 89 & 608 & 51 & 898 & 110\\
+            Cryptonets (Max Pooling) (V) & 1 & 1 & 7 & 1 & 2 & 5 & 2 & 4 & 7 & 2 & 12 & 49\\
+            \midrule
+            Database Join & 70 & 8 & 433 & 48 & 790 & 80 & 19 & 229 & 458 & 119 & 3,518 & 427\\
+            Database Join (V) & 54 & 6 & 320 & 35 & 575 & 61 & 16 & 112 & 320 & 57 & 1,457 & 285\\
+            \midrule
+            Database Variance & 166 & 18 & 2,009 & 135 & 1,639 & 163 & 95 & 269 & 1,708 & 145 & 2,795 & 320\\
+            Database Variance (V) & 37 & 6 & 321 & 24 & 334 & 43 & 10 & 30 & 170 & 13 & 178 & 141\\
+            \midrule
+            Histogram & 94 & 10 & 862 & 68 & 979 & 97 & 27 & 94 & 491 & 51 & 1,132 & 135\\
+            Histogram (V) & 33 & 5 & 166 & 16 & 164 & 23 & 7 & 17 & 92 & 13 & 154 & 68\\
+            \midrule
+            Inner Product & 127 & 15 & 1,675 & 108 & 1,308 & 130 & 83 & 250 & 1,526 & 134 & 2,623 & 301\\
+            Inner Product (V) & 16 & 5 & 158 & 12 & 165 & 25 & 6 & 18 & 83 & 7 & 86 & 127\\
+            \midrule
+            k-means & 108 & 12 & 1,333 & 88 & 1,090 & 108 & 63 & 185 & 1,141 & 99 & 1,958 & 225\\
+            k-means (V) & 6 & 3 & 47 & 4 & 43 & 12 & 2 & 11 & 32 & 4 & 54 & 95\\
+            \midrule
+            Longest 102 & 93 & 7 & 650 & 52 & 713 & 71 & 26 & 93 & 475 & 49 & 1,091 & 128\\
+            Longest 102 (V) & 169 & 6 & 544 & 41 & 519 & 53 & 25 & 60 & 369 & 33 & 605 & 95\\
+            \midrule
+            Max. Dist. b/w Symbols & 71 & 8 & 572 & 43 & 576 & 57 & 24 & 69 & 397 & 38 & 748 & 89\\
+            Max. Dist. b/w Symbols (V) & 166 & 7 & 538 & 39 & 512 & 51 & 24 & 57 & 363 & 32 & 589 & 78\\
+            \midrule
+            Minimal Points & 35 & 5 & 458 & 31 & 369 & 37 & 24 & 46 & 401 & 26 & 347 & 40\\
+            Minimal Points (V) & 0 & 1 & 1 & 0 & 1 & 3 & 0 & 1 & 1 & 0 & 1 & 16\\
+            \midrule
+            MNIST ReLU & 132 & 31 & 1,843 & 126 & 1,483 & 152 & 98 & 247 & 1,630 & 135 & 2,401 & 298\\
+            MNIST ReLU (V) & 3 & 3 & 25 & 3 & 9 & 17 & 5 & 11 & 25 & 5 & 33 & 136\\
+            \midrule
+            Private Set Intersection & 95 & 9 & 558 & 59 & 1,049 & 104 & 22 & 186 & 591 & 96 & 2,639 & 302\\
+            Private Set Intersection (V) & 1 & 2 & 1 & 2 & 1 & 8 & 1 & 8 & 2 & 4 & 2 & 122\\
+          \bottomrule
+       \end{tabular}
+       \label{table:metrics}
+   \end{table*}
+
 \begin{figure*}[htbp]
@@
 
+A detailed summery of the effects of vectorization on various benchmarks is presented in \cref{table:metrics}.
 We show circuit evaluation times in \cref{fig:graph_all_eval_time}.

### 2022-05-01 overleaf/related.tex
Minor edits in these secs.
A comment in Related work.
Added comments.

@@
 restrict flow between secure and insecure parts of the program.
+
+%\ana{Add discussion on HyCC/Buscher here. 1. Formalization, reasoning about correctness of transformation. Source-to-source cannot handle if-then-else vs. MPC Source.}
 %Key ideas such as ${\sf insecure} <: {\sf secure}$ qualifier systems are similar to the ideas we
@@
 %thrust of the proposal.
+
+HyCC~\cite{CCS:BDKKS18} is a compiler from C Source into ABY/MOTION circuits. It does source-to-source compilation with the key goal to decompose the 
+program into modules and then assign protocols to modules. In contrast, we focus on MPC Source-level optimizations, specifically vectorization, although we envision 
+future optimizations as well. We formalize MPC Source and reasoning about transformations, which we conjecture is more tractable than reasoning over
+the higher-level AST. On the other hand, HyCC does inter-procedural optimizations, while our analysis is intra-procedural. We will explore context-sensitive inter-procedural 
+analysis over MPC Source in future work. HyCC, similarly to Buscher~\cite{Buscher:2018} uses an of-the-shelf source-to-source polyhedral compiler~\footnote{We believe HyCC uses 
+Par4All~\ana{add citation!}, however, it is not appear to be included with the publicly available distribution of HyCC.}
+to perform vectorization at the level of source code. The disadvantage of using an of-the-shelf source-to-source compiler is that it solves a more general 
+problem than what MPC presents and may forgo opportunities for optimization --- specifically, it is well-known that vectorization and polyhedral compilation 
+does not work well with conditionals~\cite{Benabderrahmane:2010,Karrenberg:2015}. In contrast, we consider vectorization at the level of MPC Source which 
+linearizes conditionals; we are able to handle programs with interleaved if-statements and for-loops and achieve significant speedup.
 
@@
 vectorization over SSA form, which has similarities to MPC Source. We consider different array representation, notion of dependence, and reasoning
-about dependence, which we conjecture is more suitable for MPC Source. Buscher~\cite{Buscher:2018} considers SIMD-vectorization
-at the level of source code, which then combines with circuit-level optimizations in the TinyGarble compiler.
-He proposes using an off-the-shelf polyhedral compiler, however, application is limited to only two routines,
-essentially just inner product and euclidian distance; it is unclear how effective the off-the-shelf compiler is.
-In contrast, we consider vectorization at the level of MPC Source separating ``backend-independent'' vectorization
-and circuit-level amortization (done by MOTION). We apply our compiler on a wide range of routines.
+about dependence, which we conjecture is more suitable for MPC Source. 
+%Buscher~\cite{Buscher:2018} considers SIMD-vectorization
+%at the level of source code, which then combines with circuit-level optimizations in the TinyGarble compiler.
+%He proposes using an off-the-shelf polyhedral compiler, however, application is limited to only two routines,
+%essentially just inner product and euclidian distance; it is unclear how effective the off-the-shelf compiler is.
+%In contrast, we consider vectorization at the level of MPC Source separating ``backend-independent'' vectorization
+%and circuit-level amortization (done by MOTION). We apply our compiler on a wide range of routines.

### 2022-05-01 overleaf/results.tex

@@
 \end{figure*}
+
+\subsection{Discussion}
+\ana{Ana todo: Discuss why we are slower than HyCC --- because of ad-hoc div-and-conquer optimizations. Have to summarize what Ben and I figured out yesterday.}
+
+\ana{If we can add that our cost model works great (it does!), that would be great but probably won't have time...}

### 2022-05-01 overleaf/results.tex
evaluation section nuking references to biometric_fast
comment update

@@
 
-\ishaq{We don't have space to show individual benchmark graphs, and vec data is only plotted in individual benchmark graphs, so specifying what we use for vec experiment in the following text is unnecessary.
-(unless, we change the graph)}
+\ishaq{@Ana: I don't know what to do about this: In the following list of benchmarks, I specify paramters for vectorized experiments too.
+but we don't have enough space to show the vectorized only experiments (we could replace the graphs).
+Not sure if I should take it out}
 
 \ishaq{TODO: go through description again, and add why each benchmark may be important from MPC/private-learning/distributed-computing perspective}
-
-\ishaq{TODO: add a large table for LAN, incorporating as much data as possible}
 
@@
 
-    \item {\em Biometric Matching (Fast)} is the faster instantiation taken from \cite{Demmler:2015}. Parameters are same as above.
+    % \item {\em Biometric Matching (Fast)} is the faster instantiation taken from \cite{Demmler:2015}. Parameters are same as above.
 

### 2022-05-01 compiler/paper_benchmarks.py
not generating biometric_fast graph/data anymore
It is exactly the same as biometric, I think I executed the same benchmark twice (it is unlikely that I missed it, but could have happened if some internal configuration was wrong).

@@ -1048,7 +1048,7 @@ def generate_single_network_graphs(all_stats, dir):
 def get_benchmark_stats(all_stats, benchmark_name, input_label):
     for task_stat in all_stats:
         if task_stat.label == benchmark_name:
-            for i in range(len(task_stat.input_configs)-1, 0, -1):
+            for i in range(len(task_stat.input_configs)-1, -1, -1):
                 input = task_stat.input_configs[i]
                 if input.label == input_label:
                     if input.gmw_vec_p0 is not None and input.bmr_vec_p0 is not None:
@@ -1070,7 +1070,7 @@ def generate_comparison_graphs(lan_stats, wan_stats, get_val, y_label, dir):
         for lan_task_stat in lan_stats:
             label = get_x_label_for_benchmark(lan_task_stat.label)
 
-            for i in range(len(lan_task_stat.input_configs)-1, 0, -1):
+            for i in range(len(lan_task_stat.input_configs)-1, -1, -1):
                 lan_input = lan_task_stat.input_configs[i]
                 if lan_input.gmw_vec_p0 is not None and lan_input.bmr_vec_p0 is not None:
                     wan_input = get_benchmark_stats(wan_stats, lan_task_stat.label, lan_input.label)
@@ -1102,6 +1102,67 @@ def generate_comparison_graphs(lan_stats, wan_stats, get_val, y_label, dir):
     # run_gnuplot(hist_script, bmrfile_path, bmr_hist_graph_file, bmr_title, y_label, dir, other_args = [])
     # run_gnuplot(line_script, file_path, line_graph_file, title, y_label, dir, other_args = [])
 
+def gen_latex_table(all_stats):
+    print('\n\n')
+    for task_stats in all_stats:
+        if(task_stats.label == 'biometric_fast'):
+            continue
+        label = get_x_label_for_benchmark(task_stats.label)
+        for i in range(len(task_stats.input_configs)-1, -1, -1):
+            input = task_stats.input_configs[i]
+            if input.gmw_p0 is not None and input.gmw_vec_p0 is not None and \
+                input.bmr_p0 is not None and input.bmr_vec_p0 is not None:
+                
+                a_ts = input.gmw_p0.timing_stats
+                a_cs = input.gmw_p0.circuit_stats
+                b_ts = input.bmr_p0.timing_stats
+                b_cs = input.bmr_p0.circuit_stats
+
+                vals1 = [label,
[…]

### 2022-05-02 overleaf/results.tex
Edits to Results.

@@
 We tested our framework with several benchmarks.
-For the multiparty computation (MPC), we restriced our evaluation to 2 party computation (2PC) setting because it requires fewer computing resources.
+For the multiparty computation (MPC), we restricted our evaluation to 2 party computation (2PC) setting because it requires fewer computing resources.
 We stress that there is no such inherent restriction in our framework.
-We use hardware resources provided by CloudLab\cite{DuplyakinATC19} and consider two network settings, namely Local Area Netowrk (LAN) and Wide Area Network (WAN).
+We use hardware resources provided by CloudLab\cite{DuplyakinATC19} and consider two network settings, namely Local Area Network (LAN) and Wide Area Network (WAN).
 In the LAN setting, we use {\tt c6525-25g} machines for both parties.
@@
 The connection between these machines had 10Gbps bandwidth and sub-millisecond latency.
-This setting reflects typical LAN usecase considering that 10Gbps LAN is increasingly common in business networks and is now available even in some home networks.
+This setting reflects typical LAN use case considering that 10Gbps LAN is increasingly common in business networks and is now available even in some home networks.
 In the WAN setting we again used a {\tt c6525-25g} machine (located in Utah, US) for the first party and a {\tt c220g1} machine (located in Wisconsin, US) for the second.
-The {\tt c220g1} machine is equiped with two Intel E5-2630 8-core 2.40GHz processors and 128GB of RAM.
+The {\tt c220g1} machine is equipped with two Intel E5-2630 8-core 2.40GHz processors and 128GB of RAM.
 We measured the connection bandwidth between these machines to be 560Mbps and average round trip time (RTT) to be 38ms.
-At the time of this writing, all major internet providers in the US offer 1Gbps connections to home consumers, therefore this setting reasonably reflects the typical WAN usecase.
+At the time of this writing, all major internet providers in the US offer 1Gbps connections to home consumers, therefore this setting reasonably reflects the typical WAN use case.
 
 We run all experiments 5 times and report average values of various metrics.
-Note that standard deviation --- shown as errorbar on top of the bars in the graphs --- in all observations was at most 4.5\% of the mean value, therefore the accuracy of the results is not effected by the relatively fewer runs.
+Note that standard deviation --- shown as error bar on top of the bars in the graphs --- in all observations was at most 4.5\% of the mean value, therefore the accuracy of the results is not effected by the relatively fewer runs.
 
@@
 In the following, we say {\em both} for an experiment in which we execute both non-vectorized and vectorized protocols and {\em vec} for the vectorized only experiment. 
+In a preliminary experiment we ran {\tt N=2, N=4}, all the way to {\tt N=4096}; the non-vectorized version ran out of memory at the value {\em both} and we fixed this value for these experiments.
+The vectorized one completed for {\tt N=4096} for nearly all benchmarks (numbers are not shown for space reasons; however, they times are largely consistent, e.g., if {\em both} is $N = 2^k$ and it completes in $X$ seconds for vectorized, then $N = 2^{12}$ completes in $(12-k)X$ seconds).
 We used the following benchmarks in our evaluation:
 
-\ishaq{@Ana: I don't know what to do about this: In the following list of benchmarks, I specify paramters for vectorized experiments too.
+%\ishaq{@Ana: I don't know what to do about this: In the following list of benchmarks, I specify paramters for vectorized experiments too.
 but we don't have enough space to show the vectorized only experiments (we could replace the graphs).
-Not sure if I should take it out}
+Not sure if I should take it out} \ana{I think I got this.}
 
-\ishaq{TODO: go through description again, and add why each benchmark may be important from MPC/private-learning/distributed-computing perspective}
+%\ishaq{TODO: go through description again, and add why each benchmark may be important from MPC/private-learning/distributed-computing perspective}
 
 \begin{enumerate}
-    \item {\em Biometric Matching} Server has a database S of {\tt N} records, each record's dimension is {\tt D}.
+    \item {\em Biometric Matching:} Server has a database S of {\tt N} records, each record's dimension is {\tt D}.
 Client submits a query C, client and server compute the closest record to C in an MPC.
@@
 
-    \item {\em Convex Hull} given a polygon of {\tt N} vertices (split between Alice and Bob), convex hull is computed in an MPC.
-adapted from \cite{Farzan:2021}.
+    \item {\em Convex Hull:} Given a polygon of {\tt N} vertices (split between Alice and Bob), convex hull is computed in an MPC.
+It is adapted from \cite{Farzan:2021}.
 We use {\tt N=32} for {\em both} experiment and {\tt N=256} for {\em vec} experiment.
 
-    \item {\em Count 102} Alice has a string of length {\tt N} of symbols, Bob has a regular expression of the form 1(0*)2, together they compute number of substrings that match the regular expression.
-adapted from \cite{Farzan:2021}.
+    \item {\em Count 102:} Alice has a string of length {\tt N} of symbols, Bob has a regular expression of the form 1(0*)2, together they compute number of substrings that match the regular expression.
+It is adapted from \cite{Farzan:2021}.
 We use {\tt N=1024} for {\em both} and {\tt N=4096} for {\em vec}.
 
-    \item {\em Count 10} Same as {\em Count 102} except now the regular expression is of the form 1(0+). Parameters are same as above.
+    \item {\em Count 10:} Same as {\em Count 102} except now the regular expression is of the form 1(0+). Parameters are same as above.
 
-    \item {\em Cryptonets Max Pooling} Given a matrix of $\mathtt{rows}\times\mathtt{cols}$ elements that are split between Alice and Bob, they compute the max pooling subroutine of the cryptonet benchmark\cite{Dowlin:2016}.
+    \item {\em Cryptonets Max Pooling:} Given a matrix of $\mathtt{rows}\times\mathtt{cols}$ elements that are split between Alice and Bob, they compute the max pooling subroutine of the cryptonet benchmark\cite{Dowlin:2016}.
 We use {\tt rows=64, cols=64} for {\em both} experiment.
 
-    \item {\em Database Join} given two databases with {\tt A} and {\tt B} 2-element records, compute cross join. We use {\tt A=B=32} for {\em both} and {\tt A=B=64} for {\em vec}.
+    \item {\em Database Join:} given two databases with {\tt A} and {\tt B} 2-element records, compute cross join. We use {\tt A=B=32} for {\em both} and {\tt A=B=64} for {\em vec}.
 
-    \item {\em Database Variance} given a database of {\tt len} records, compute variance.
-\ishaq{Ben/Ana: Please correct the description, I don't know what this benchmark is} We use {\tt len=512} for {\em both} and {\tt len=4096} for {\em vec}.
+    \item {\em Database Variance} given a database of {\tt len} records, compute variance. %\ishaq{Ben/Ana: Please correct the description, I don't know what this benchmark is} 
+    We use {\tt len=512} for {\em both} and {\tt len=4096} for {\em vec}.
 
-    \item {\em Histogram} given {\tt N} 5-star ratings, compute their histogram, taken from \cite{Ishaq:2019, Farzan:2021}. We use {\tt N=512} for {\em both} and {\tt N=4096} for {\em vec}.
+    \item {\em Histogram:} Given {\tt N} 5-star ratings, compute their histogram, taken from \cite{Ishaq:2019, Farzan:2021}. We use {\tt N=512} for {\em both} and {\tt N=4096} for {\em vec}.
 
-    \item {\em Inner Product} given two vectors, each of {\tt N} elements, compute their inner product. We use {\tt N=512} for {\em both} and {\tt N=4096} for {\em vec}.
+    \item {\em Inner Product:} given two vectors, each of {\tt N} elements, compute their inner product. We use {\tt N=512} for {\em both} and {\tt N=4096} for {\em vec}.
 
-    \item {\em k-means Iteration} performs the iteration subroutine of k-means database clustering operation \cite{Jagannathan:2005, Vaidya:2003}.
+    \item {\em k-means Iteration:} performs the iteration subroutine of k-means database clustering operation \cite{Jagannathan:2005, Vaidya:2003}.
 Here {\tt len1} is the size of input data, and {\tt len2} is the number of clusters.
 We use {\tt len1=32, len2=5} for {\em both} and {\tt len1=256, len2=8} for {\em vec}.
-\ishaq{Ana/Ben: please update description}
+%\ishaq{Ana/Ben: please update description} \ana{I think that's it?}
 
-    \item {\em Longest 102} Similar to {\em Count 102} except that it computes the largest substring matching the regular expression.
+    \item {\em Longest 102:} Similar to {\em Count 102} except that it computes the largest substring matching the regular expression.
 We use same parameters as {\em Count 102}, adapted from \cite{Farzan:2021}.
@@
 {\tt outer} is fixed at 16.
-\ishaq{Ben/Ana, please update}
+%\ishaq{Ben/Ana, please update} \ana{Don't know what to do...}
 
@@
 
-A detailed summery of the effects of vectorization on various benchmarks is presented in \cref{table:metrics}.
+A detailed summary of the effects of vectorization on various benchmarks is presented in \cref{table:metrics}.
 We show circuit evaluation times in \cref{fig:graph_all_eval_time}.
@@
 3) {\it Low:} these include the Database Join and the regular expression benchmarks (count 102, count 10, longest 102 and max distance between symbols).
-There is very little parallelizable computation in these programs, thus the speedup is lower.
+There is less parallelizable computation in these programs, thus the speedup is lower.
 We see a speedup from 1.1x to 2x in BMR.
 In GMW, DB Join, Count 102 and Count 10s see speedup from 1.1x to 1.3x.
-However, longest 102 and max distance between symbols suffer a slowdown of 0.5x.
-The structure of these benchmarks is such that transformation to vectorized code increases multiplicative depth and, the negative effect of increased depth is more noticeable in a round-based protocol like GMW.
-We fix this via a simple heuristic where if the transformation increases circuit depth beyond some threshold (e.g.
+However, longest102 and max distance between symbols suffer a slowdown of 0.5x.
+There is opportunity for vectorization in these benchmarks according to our analytical model, particularly, a large EQ operation is vectorized, though a large portion of the loop cannot be vectorized.
+We observed that transformation to vectorized code increased multiplicative depth and, the negative effect of increased depth is more noticeable in a round-based protocol like GMW.
+The cause of the increase is not clear --- we conjecture that MOTION performs optimizations over the non-vectorized loop body that decreases depth; also, EQ is relatively inexpensive in Boolean GMW and BMR compared to ADD and MUL, which also de-emphasizes the benefit of vectorization.
+We propose a simple heuristic: if the transformation increases circuit depth beyond some threshold (e.g.
 more than 10\% of the original circuit), we reject the transformation.
-Nevertheless, we show these graphs here for the sake of completeness and to highlight that vectorization is not always reduce run time.
+%Nevertheless, we show these graphs here for the sake of completeness and to highlight that vectorization is not always reduce run time.
 Note that in some settings it may still be desirable to vectorize e.g.
@@
 
-We present evaluation for communication size in \cref{fig:graph_comm_size_time}, circuit generation time in \cref{fig:graph_circ_gen_time}, number of gates in \cref{fig:graph_total_gates} and online time and setup time in \cref{fig:graph_online_time} and \cref{fig:graph_setup_time} respectively.
-\ishaq{Ana: Let me know if I should add more substance.}
+We present evaluation for communication size in \cref{fig:graph_comm_size}, circuit generation time in \cref{fig:graph_circ_gen_time}, number of gates in \cref{fig:graph_total_gates} and online time and setup time in \cref{fig:graph_online_time} and \cref{fig:graph_setup_time} respectively.
+%\ishaq{Ana: Let me know if I should add more substance.} \ana{It's good!}
 
@@
 
-Next, we look closely at circuit evaluation \cref{fig:graph_biometric_eval_time}, communication size \cref{fig:graph_biometic_comm_size} and circuit generation time \cref{fig:graph_biometic_circ_gen_time} for biometric matching benchmark.
+We look closely at the Biometric Matching benchmark: circuit evaluation in~\cref{fig:graph_biometric_eval_time}, communication size in \cref{fig:graph_biometic_comm_size} and circuit generation time in~\cref{fig:graph_biometic_circ_gen_time}.
 For input size beyond {\tt N=128} the memory usage exceeds available memory and prevents circuit generation.
@@
 Notice that vectorization improves all metrics.
-A database of size {\tt N=128} is smaller than any real world usecase, therefore lets use it as baseline for improvement.
+%A database of size {\tt N=128} is smaller than any real world use case, therefore lets use it as baseline for improvement.
 Comparing performance improvement between BMR and GMW, we see more speedup for BMR (23x vs 10x), GMW gets more communication size reduction (10x vs 2.5x) and circuit generation sees a speedup of 35x and 45x for BMR and GMW respectively.

### 2022-05-02 overleaf/related.tex
Changes to results and related work.

@@
 
-HyCC~\cite{CCS:BDKKS18} is a compiler from C Source into ABY/MOTION circuits. It does source-to-source compilation with the key goal to decompose the 
+HyCC~\cite{CCS:BDKKS18} is a compiler from C Source into ABY circuits. It does source-to-source compilation with the key goal to decompose the 
 program into modules and then assign protocols to modules. In contrast, we focus on MPC Source-level optimizations, specifically vectorization, although we envision 
@@
 to perform vectorization at the level of source code. The disadvantage of using an of-the-shelf source-to-source compiler is that it solves a more general 
-problem than what MPC presents and may forgo opportunities for optimization --- specifically, it is well-known that vectorization and polyhedral compilation 
-does not work well with conditionals~\cite{Benabderrahmane:2010,Karrenberg:2015}. In contrast, we consider vectorization at the level of MPC Source which 
+problem than what MPC presents and may forgo opportunities for optimization --- concretely, it is well-known that vectorization and polyhedral compilation 
+do not work well with conditionals~\cite{Benabderrahmane:2010,Karrenberg:2015}. In contrast, we consider vectorization at the level of MPC Source which 
 linearizes conditionals; we are able to handle programs with interleaved if-statements and for-loops and achieve significant speedup.

### 2022-05-02 overleaf/results.tex
Changes to results and related work.

@@
 
-\subsection{Discussion}
-\ana{Ana todo: Discuss why we are slower than HyCC --- because of ad-hoc div-and-conquer optimizations. Have to summarize what Ben and I figured out yesterday.}
+\subsection{Comparison with MOTION-native Inner Product and Discussion}
 
-\ana{If we can add that our cost model works great (it does!), that would be great but probably won't have time...}
+%We conclude the . 
+Finally, we compared our automatically generated routine for Inner Product with the manually SIMD-ified MOTION-native routine in the distribution. 
+We were surprised that we were an order of magnitude slower in Boolean GMW as our circuit ran a significantly larger 
+number of communication rounds. Upon investigation, it turned out that the vectorized multiplications were essentially the same, 
+however, our addition loop incurred significant cost (ADD is non-local and expensive in Boolean GMW). The MOTION-native loop ran 
+{\sf result += mult\_unsimdified[i];} while our loop generated and ran {\sf result[i] = result[i] + mult\_unsimdified[i];}. 
+We rewrote the accumulation (manually, for testing purposes) and that led to the comparable speedup!
+
+Therefore, MOTION's compiler performs analysis that informs circuit generation. In the above example, 
+MOTION does the standard divide-and-conquer accumulation. The example also illustrates the limitation of higher-level AST analysis --- while 
+the analysis was able to detect the associative accumulation when operator {\sf += } was used, it was unable to do so when an equivalent 
+code was used. We conjecture that MPC source, as a more straight-forward representation will
+not only enable detection of general associative loops, but also allow for program synthesis to increase opportunities for divide-and-conquer
+parallelization~\cite{Farzan:2021}; we leave this as future work.
+
+The most closely related work, HyCC~\cite{CCS:BDKKS18}, is a mixing compiler that focuses on mixed protocols, while we run vectorization within a single protocol. 
+The paper only provides data for Boolean and Yao for a version of Biometric matching on $N=1000$; it does provide a lot of data on mixed protocol executions.
+Again, we estimate we are about an order of magnitude slower. This is likely due to the same issue as Inner Product --- the computation of min can be optimized.
+
+%\ana{Ana todo: Discuss why we are slower than HyCC --- because of ad-hoc div-and-conquer optimizations. Have to summarize what Ben and I figured out yesterday.}
+
+%\ana{If we can add that our cost model works great (it does!), that would be great but probably won't have time...}

### 2022-05-02 overleaf/results.tex
Results minor change.

@@
 
-Therefore, MOTION's compiler performs analysis that informs circuit generation. In the above example, 
-MOTION does the standard divide-and-conquer accumulation. The example also illustrates the limitation of higher-level AST analysis --- while 
+Therefore, MOTION's compiler performs analysis that informs circuit generation and the example illustrates the power of the analysis. In the above example, 
+MOTION does the standard divide-and-conquer accumulation. It also illustrates the limitation of higher-level AST analysis --- while 
 the analysis was able to detect the associative accumulation when operator {\sf += } was used, it was unable to do so when an equivalent 

### 2022-05-02 overleaf/results.tex
Made a note.

@@
 however, our addition loop incurred significant cost (ADD is non-local and expensive in Boolean GMW). The MOTION-native loop ran 
-{\sf result += mult\_unsimdified[i];} while our loop generated and ran {\sf result[i] = result[i] + mult\_unsimdified[i];}. 
+{\sf result += mult\_unsimdified[i];} while our loop generated and ran {\sf result[i] = prev + mult\_unsimdified[i];}. 
 We rewrote the accumulation (manually, for testing purposes) and that led to the comparable speedup!
@@
 the analysis was able to detect the associative accumulation when operator {\sf += } was used, it was unable to do so when an equivalent 
-code was used. We conjecture that MPC source, as a more straight-forward representation will
+code was used. \ana{This is not true, have to tone down.} We conjecture that MPC source, as a more straight-forward representation will
 not only enable detection of general associative loops, but also allow for program synthesis to increase opportunities for divide-and-conquer

### 2022-05-02 overleaf/results.tex
Rewrote results paragraph.

--- 2022-05-02 overleaf/results.tex
@@
 however, our addition loop incurred significant cost (ADD is non-local and expensive in Boolean GMW). The MOTION-native loop ran 
-{\sf result += mult\_unsimdified[i];} while our loop generated and ran {\sf result[i] = result[i] + mult\_unsimdified[i];}. 
-We rewrote the accumulation (manually, for testing purposes) and that led to the comparable speedup!
+{\sf result += mult\_unsimdified[i];} while our loop generated and ran {\sf result[i] = result[i-1] + mult\_unsimdified[i];} (recall that the array computation
+is an artifact of our vectorization). We rewrote the accumulation (manually, for testing purposes) and that led to the comparable speedup!
 
 Therefore, MOTION's compiler performs analysis that informs circuit generation and the example illustrates the power of the analysis. In the above example, 
-MOTION does the standard divide-and-conquer accumulation. It also illustrates the limitation of higher-level AST analysis --- while 
-the analysis was able to detect the associative accumulation when operator {\sf += } was used, it was unable to do so when an equivalent 
-code was used. We conjecture that MPC source, as a more straight-forward representation will
-not only enable detection of general associative loops, but also allow for program synthesis to increase opportunities for divide-and-conquer
-parallelization~\cite{Farzan:2021}; we leave this as future work.
+MOTION does the standard divide-and-conquer accumulation. The poor performance of our loop is not a limitation of our approach, but of our current 
+implementation. Recall that the next phase of Vectorization, in~\secref{sec:vectorization}, which we have not implemented yet, gets rid of redundant 
+dimensions and generates {\sf result = result\_prev + mult\_unsimdified[i]}. This could have benefited from MOTION's divide-and-conquer optimization
+but it might lead to confusion of the static analysis as analysis on the AST is difficult (unfortunately, we ran out of time and did not experiment further).
+%Interestingly, the non-vectorized version, which is the literal translation of inner product does not appear to benefit from any optimizations.
+
+We conjecture that MPC source, as a more straight-forward representation will not only enable detection of general associative loops, but also allow 
+for program synthesis to increase opportunities for divide-and-conquer parallelization~\cite{Farzan:2021}; we leave this as future work.
 

### 2022-05-02 overleaf/results.tex
shrank the graphs

@@
 %\resizebox{7in}{!}{\input{graphs/lan/all-hist-OnlineSetupTimesec}}
-\input{graphs/lan/all-hist-OnlineSetupTimesec}
+\resizebox{0.7\textwidth}{!}{\input{graphs/lan/all-hist-OnlineSetupTimesec}}
 \caption{Circuit Evaluation Time (Setup + Online) of Benchmarks}
@@
 \centering
-\resizebox{3.6in}{!}{\input{graphs/lan/biometric-hist-OnlineSetupTimesec}}
+\resizebox{0.7\columnwidth}{!}{\input{graphs/lan/biometric-hist-OnlineSetupTimesec}}
 \caption{Biometric Matching Circuit Evaluation Time, x-axis lists database size}
@@
 \centering
-\resizebox{3.6in}{!}{\input{graphs/lan/biometric-hist-CommunicationMiB}}
+\resizebox{0.7\columnwidth}{!}{\input{graphs/lan/biometric-hist-CommunicationMiB}}
 \caption{Biometric Matching Communication Size, x-axis lists database size}
@@
 \centering
-\resizebox{3.6in}{!}{\input{graphs/lan/biometric-hist-CircuitGenerationTimesec}}
+\resizebox{0.7\columnwidth}{!}{\input{graphs/lan/biometric-hist-CircuitGenerationTimesec}}
 \caption{Biometric Matching Circuit Generation Time, x-axis lists database size}
@@
 \centering
-\resizebox{3.6in}{!}{\input{graphs/comparison-hist-OnlineSetupTimesec}}
+\resizebox{0.7\columnwidth}{!}{\input{graphs/comparison-hist-OnlineSetupTimesec}}
 \caption{LAN vs. WAN: Circuit Evaluation Time Comparison}
@@
 \centering
-\input{graphs/lan/all-hist-CommunicationMiB}
+\resizebox{0.7\textwidth}{!}{\input{graphs/lan/all-hist-CommunicationMiB}}
 \caption{Communication Size of Benchmarks}
@@
 \centering
-\input{graphs/lan/all-hist-CircuitGenerationTimesec}
+\resizebox{0.7\textwidth}{!}{\input{graphs/lan/all-hist-CircuitGenerationTimesec}}
 \caption{Circuit Generation Time of Benchmarks}
@@
 \centering
-\input{graphs/lan/all-hist-TotalGates}
+\resizebox{0.7\textwidth}{!}{\input{graphs/lan/all-hist-TotalGates}}
 \caption{Number of Gates of Benchmarks}
@@
 \centering
-\input{graphs/lan/all-hist-OnlineTimesec}
+\resizebox{0.7\textwidth}{!}{\input{graphs/lan/all-hist-OnlineTimesec}}
 \caption{Online Time of Benchmarks}
@@
 \centering
-\input{graphs/lan/all-hist-SetupTimesec}
+\resizebox{0.7\textwidth}{!}{\input{graphs/lan/all-hist-SetupTimesec}}
 \caption{Setup Time of Benchmarks}
@@
 
-\subsection{Comparison with MOTION-native Inner Product and Discussion}
+\subsection{Comparison with MOTION-native Inner Product}
 

### 2022-05-02 overleaf/paper_SIMD.tex
Reorganized sections again...
Changed Lindsey's email.
added LAN benchmarks to the paper
add updated graphs
wip evaluation section
Edits to Results.
Changes to results and related work.
red line numbers return
Removed Conf. info.

@@
 %
-%\documentclass[sigconf, screen, review, anonymous, natbib=false]{acmart}
-\documentclass[sigconf, screen, natbib=false, dvipsnames, table]{acmart}
-\settopmatter{printacmref=true}
+\documentclass[sigconf, screen, review, anonymous, natbib=false, dvipsnames, table]{acmart}
+%\documentclass[sigconf, screen, natbib=false, dvipsnames, table]{acmart}
+\settopmatter{printacmref=false}
+\renewcommand\footnotetextcopyrightpermission[1]{}
 % mandatory for CCS'19
@@
 
-\copyrightyear{2019}
-\acmYear{2019}
-\setcopyright{acmcopyright}
-\acmConference[CCS '19]{2019 ACM SIGSAC Conference on Computer and Communications Security}{November 11--15, 2019}{London, United Kingdom}
-\acmBooktitle{2019 ACM SIGSAC Conference on Computer and Communications Security (CCS '19), November 11--15, 2019, London, United Kingdom}
-\acmPrice{15.00}
-\acmDOI{10.1145/3319535.3339818}
-\acmISBN{978-1-4503-6747-9/19/11}
+%\copyrightyear{2019}
+%\acmYear{2019}
+%\setcopyright{acmcopyright}
+%\acmConference[CCS '19]{2019 ACM SIGSAC Conference on Computer and Communications Security}{November 11--15, 2019}{London, United Kingdom}
+%\acmBooktitle{2019 ACM SIGSAC Conference on Computer and Communications Security (CCS '19), November 11--15, 2019, London, United Kingdom}
+%\acmPrice{15.00}
+%\acmDOI{10.1145/3319535.3339818}
+%\acmISBN{978-1-4503-6747-9/19/11}
 
@@
 % declare the path(s) where your graphic files are
-\graphicspath{{../figures/}}
+\graphicspath{{../figures/}{graphs/}{graphs/lan/}}
 % and their extensions so you won't have to specify these with
 % every instance of \includegraphics
-\DeclareGraphicsExtensions{.pdf,.jpeg,.png}
+\DeclareGraphicsExtensions{.pdf,.jpeg,.png, .eps}
 
@@
 \begin{document}
-\fancyhead{}
+% \fancyhead{}
 % do not delete this code.
@@
 \author{Lindsey Kennard}
-\email{fire.elemental@gmail.com}
+\email{fireelemental.ne@gmail.com}
 \affiliation{%
@@
 %
+\begin{comment}
 \begin{CCSXML}
@@
 \keywords{multiparty computation; compilers; cryptography}
+\end{comment}
 
@@
 \maketitle
+\thispagestyle{empty}
 
@@
 
-% Compiler Framework Section
+% Compiler Front End
 \input{compiler.tex}
@@
 \input{vectorization.tex}
+
+% Backend
+\input{backend.tex}
 
@@
 \label{sec:conclusion}
+
+\ana{Need a conclusion here!}
+
 %\input{../sections/conclusion.tex}

### 2022-05-02 overleaf/model.tex
Final edits. Will freeze TR.

@@
 %\ana{We need an intro to section here.}
-\ana{Terminology: "operation" or "instruction" or "gate"?}
+%\ana{Terminology: "operation" or "instruction" or "gate"?}
 
@@
 %\ishaq{This is not an assumption, this is our setting. We work in the single protocol world.}
-\item There are two tiers of MPC instructions, local and remote. A local instruction (essentially just ADD) has cost $\beta$
-and a remote instruction (e.g., MUX, MUL, SHL, etc.) has cost $\alpha$, where $\alpha >> \beta$. We assume that all remote
+\item There are two tiers of MPC instructions, local and remote. A local instruction (e.g., ADD in Arithmetic, XOR in Boolean) 
+has cost $\beta$ and a remote instruction (e.g., MUX, MUL, SHL, etc.) has cost $\alpha$, where $\alpha >> \beta$. We assume that all remote
 instructions have the same cost.
@@
 \item  In MPC frameworks, executing $n$ operations ``at once'' in a single SIMD operation costs a lot less than executing those $n$ operations one by one.
-Following Amdahl's law~\ana{citation!}, we write $\alpha = \frac{1}{s}p\alpha + (1-p)\alpha$, where $p$ is the fraction of execution time that benefits from amortization and $(1-p)$
+Following Amdahl's law, we write $\alpha = \frac{1}{s}p\alpha + (1-p)\alpha$, where $p$ is the fraction of execution time that benefits from amortization and $(1-p)$
 is the fraction that does not, and $s$ is the available resource. Thus, $n\alpha = \frac{n}{s}p\alpha + n(1-p)\alpha$.
@@
 whether $n$ is $10,000$ or just $1$. (This models the cost of preparing and sending a packet from party A to party B for example.) Therefore, amortized execution
-of $n$ operations is $f(n) = \alpha_\mathit{fix} + n\alpha_{var}$ in contrast to unamortized execution $g(n) = n\alpha_\mathit{fix} + n\alpha_{var}$ and, of course.
-Clearly $\alpha_\mathit{fix} << n\alpha_\mathit{fix}$ and since fixed cost dominates variable cost (particularly for remote operations), we have $f(n) << g(n)$.
+of $n$ operations is $f(n) = \alpha_\mathit{fix} + n\alpha_{var}$ in contrast to unamortized execution $g(n) = n\alpha_\mathit{fix} + n\alpha_{var}$.
+We have $\alpha_\mathit{fix} << n\alpha_\mathit{fix}$ and since fixed cost dominates variable cost (particularly for remote operations), we have $f(n) << g(n)$.
 %We assume infinite parallel capacity---i.e., a single MPC-instruction costs as much as $N$ amortized instructions, namely $\alpha$ or $\beta$.
@@
 still cost $2\alpha_\mathit{fix} + 2\alpha_\mathit{var}$ even when scheduled ``in parallel''.\footnote{This is not strictly true, but assuming it, e.g. as in 
-\cite{Ishaq2019, Demmler2015ABYA, Mohassel2018}, helps simplify the problem.}
+\cite{Ishaq:2019, NDSS:DemSchZoh15, Mohassel:2018}, helps simplify the problem.}
 %MPC instructions scheduled in parallel benefit from amortization \emph{only if} they are the same instruction. Given our previous assumption,

### 2022-05-02 overleaf/paper_SIMD.tex
Final edits. Will freeze TR.

@@
 
-\ana{We need an abstract.}
+Research on MPC programming technology largely falls at the two ends of the classical compiler: (1) work on front-end language design (e.g., Wysteria, Viaduct) and (2) work on back-end protocol implementation (e.g., ABY, MOTION).
+
+In this work, we formalize the MPC Source intermediate language and advance the idea of what we call backend-independent optimizations, in a close analogy to machine-independent optimizations in the classical compiler.
+We present a compiler framework that takes a Python-like routine and produces MOTION code.
+We focus on a specific backend-independent optimization: novel SIMD-vectorization on MPC Source, which we show leads to significant speedup in circuit generation time and running time, as well as significant reduction in communication and number of gates over the corresponding iterative schedule.
+%Our results emphasize the importance of backend-independent optimizations --- vectorization (this work) and protocol mixing (previous work) are two optimizations available at the level of MPC Source.
+We envision future work on backend-independent optimization: new MPC optimizations and combinations of optimizations in the vein of the classical compiler.
 
@@
 
-\ana{Need a conclusion here!}
+We presented a formalization of the MPC Source intermediate language followed by a specific back-end optimization at the level of MPC Source: novel SIMD-vectorization.
+We demonstrated that vectorization has significant impact on performance.
+We are excited about the opportunities for future work --- integration with protocol mixing, divide-and-conquer reasoning and parallelization, as well as inter-procedural context-sensitive analysis at the level of MPC Source will improve MPC programmability and efficiency.
 

### 2022-05-02 overleaf/related.tex
Final edits. Will freeze TR.

@@
 analysis over MPC Source in future work. HyCC, similarly to Buscher~\cite{Buscher:2018} uses an of-the-shelf source-to-source polyhedral compiler~\footnote{We believe HyCC uses 
-Par4All~\ana{add citation!}, however, it is not appear to be included with the publicly available distribution of HyCC.}
+Par4All (\url{https://github.com/Par4All/par4all}), however, does not appear to be included with the publicly available distribution of HyCC.}
 to perform vectorization at the level of source code. The disadvantage of using an of-the-shelf source-to-source compiler is that it solves a more general 

### 2022-05-02 overleaf/results.tex
Final edits. Will freeze TR.

@@
 However, longest102 and max distance between symbols suffer a slowdown of 0.5x.
-There is opportunity for vectorization in these benchmarks according to our analytical model, particularly, a large EQ operation is vectorized, though a large portion of the loop cannot be vectorized.
+There is opportunity for vectorization in these benchmarks according to our analytical model, particularly, a there is a large EQ that is vectorized, although a large portion of the loop cannot be vectorized.
 We observed that transformation to vectorized code increased multiplicative depth and, the negative effect of increased depth is more noticeable in a round-based protocol like GMW.
 The cause of the increase is not clear --- we conjecture that MOTION performs optimizations over the non-vectorized loop body that decreases depth; also, EQ is relatively inexpensive in Boolean GMW and BMR compared to ADD and MUL, which also de-emphasizes the benefit of vectorization.
-We propose a simple heuristic: if the transformation increases circuit depth beyond some threshold (e.g.
-more than 10\% of the original circuit), we reject the transformation.
+We propose a simple heuristic although we do leave all the benchmarks in the table: if the transformation increases circuit depth beyond some threshold (e.g.
+more than 10\% of the original circuit), we can reject the transformation.
 %Nevertheless, we show these graphs here for the sake of completeness and to highlight that vectorization is not always reduce run time.
@@
 however, our addition loop incurred significant cost (ADD is non-local and expensive in Boolean GMW). The MOTION-native loop ran 
-{\sf result += mult\_unsimdified[i];} while our loop generated and ran {\sf result[i] = result[i-1] + mult\_unsimdified[i];} (recall that the array computation
-is an artifact of our vectorization). We rewrote the accumulation (manually, for testing purposes) and that led to the comparable speedup!
+{\sf result += mult\_unsimdified[i];} while our loop generated and ran {\sf result[i] = result[i-1] + mult\_unsimdified[i];} recall that the scalar expansion
+is an artifact of our vectorization. We rewrote the accumulation (manually, for testing purposes) and that led to the comparable speedup!
 
-Therefore, MOTION's compiler performs analysis that informs circuit generation and the example illustrates the power of the analysis. In the above example, 
-MOTION does the standard divide-and-conquer accumulation. The poor performance of our loop is not a limitation of our approach, but of our current 
-implementation. Recall that the next phase of Vectorization, in~\secref{sec:vectorization}, which we have not implemented yet, gets rid of redundant 
-dimensions and generates {\sf result = result\_prev + mult\_unsimdified[i]}. This could have benefited from MOTION's divide-and-conquer optimization
-but it might lead to confusion of the static analysis as analysis on the AST is difficult (unfortunately, we ran out of time and did not experiment further).
+MOTION's compiler performs analysis that informs circuit generation and the example illustrates the power of the analysis. In the above example, 
+MOTION does the standard divide-and-conquer accumulation. The poor performance of our loop is not a limitation of our approach, but it could be a 
+limitation of our current implementation. Recall that the next phase of Vectorization, in~\secref{sec:vectorization}, which we have not implemented yet, gets rid of redundant 
+dimensions and generates {\sf if (i!=0) { result\_pref = result; } result = result\_prev + mult\_unsimdified[i]}. And while it is unrealistic to expect that MOTION's static analysis 
+will detect the associative accumulation in the scalar expansion code, it is realistic to expect that it will in the above only slightly more complex code. 
+It still might lead to confusion in the static analysis as analysis on the AST is difficult (unfortunately, we ran out of time and did not experiment further).
 %Interestingly, the non-vectorized version, which is the literal translation of inner product does not appear to benefit from any optimizations.
 
-We conjecture that MPC source, as a more straight-forward representation will not only enable detection of general associative loops, but also allow 
+We conjecture that MPC Source, a straight-forward representation, will not only enable detection of general associative loops, but also allow 
 for program synthesis to increase opportunities for divide-and-conquer parallelization~\cite{Farzan:2021}; we leave this as future work.
@@
 The paper only provides data for Boolean and Yao for a version of Biometric matching on $N=1000$; it does provide a lot of data on mixed protocol executions.
-Again, we estimate we are about an order of magnitude slower. This is likely due to the same issue as Inner Product --- the computation of min can be optimized.
+Again, we estimate we are about an order of magnitude slower, which is likely due to the same issue as Inner Product --- the computation of $\mathit{min}$ can be optimized.
 

### 2022-05-02 overleaf/tech_report/introduction.tex
Freezing tech report.

--- 2022-05-01 overleaf/introduction.tex
@@
 and privacy-preserving machine learning~\cite{SP:MohZha17,CCS:MohRin18}.
-The goal is to bring the technology to a level where programmers can write \emph{secure} and \emph{efficient} programs without
+The goal is to improve technology so that programmers can write \emph{secure} and \emph{efficient} programs without
 commanding extensive knowledge of cryptographic primitives.
@@
 As in classical compilers, we envision different front ends (e.g., our front end IMP Source, other) compiling into MPC Source.
-MPC Source is particularly suitable for optimizations such as protocol mixing~\cite{Ishaq:2019, Fang:2022} and SIMD-vectorization,
+MPC Source is particularly suitable for optimizations such as protocol mixing~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022} and SIMD-vectorization,
 which takes advantage of amortization at the circuit level. The MPC Source IR exposes the \emph{linear structure} of MPC programs, which simplifies program analysis; this is
@@
 
-\subsection{Our Contribution} In this paper, we develop a compiler framework that takes a Python-like routine and produces MOTION code: we describe (a) the IMP Source language, its syntax and semantic restrictions, (b) translation from IMP Source into MPC Source, (c) a specific backend-independent optimization: novel SIMD-vectorization on MPC Source, and (d) translation from MPC Source into MOTION code.
+\subsection{Our Contribution} In this paper, we develop a compiler framework that takes a Python-like routine and produces MOTION code: we describe (a) the IMP Source language, its syntax and semantic restrictions, (b) translation into MPC Source, (c) a specific backend-independent optimization: novel SIMD-vectorization on MPC Source, and (d) translation from MPC Source into MOTION code.
 %\ana{I think we have to add more on MOTION, specifically mixing and amortization.}
@@
 First, it is the state of the art in terms of performance~\cite{Braun:2022}.
-Second, it provides an API over efficient implementation for a wide variety of cryptographic operations in three different protocols ---  Arithmetic GMW, Boolean GMW, and BMR --- which allows for protocol mixing~\cite{Ishaq:2019, Fang:2022}, a known backend-independent optimization.
+Second, it provides an API over efficient implementation for a wide variety of cryptographic operations in three different protocols ---  Arithmetic GMW, Boolean GMW, and BMR --- which allows for protocol mixing~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022}, a known backend-independent optimization.
 Third, MOTION provides API for SIMD-level operations, which amortize cost and lead to significant improvement in memory footprint and
@@
 Unfortunately, we show that optimal scheduling is NP-hard via a reduction to the Shortest Common Supersequence (SCS) problem.
-Cost modeling is important as it drives not only vectorization but optimizations such as protocol mixing and scheduling as well~\cite{Ishaq:2019, Fang:2022}.
+Cost modeling is important as it drives not only vectorization but optimizations such as protocol mixing and scheduling as well~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022}.
 
 Our most important contribution is the implementation and evaluation of the compiler framework.
-We demonstrate expressivity of the source language by running the compiler on \ana{X} programs with interleaved if- and for-statements; these include classical MPC benchmarks such as PSI and Biometric matching, as well as kMeans, Histogram, and others.
-\ana{Summarize results.}
+We demonstrate expressivity of the source language by running the compiler on 16 programs with interleaved if- and for-statements; these include classical MPC benchmarks such as PSI and Biometric matching, as well as kMeans, Histogram, and other examples from the literature.
+Our compiler takes the routine and generated \emph{non-vectorized} MOTION code (from MPC Source on the picture above). It then optimizes MPC Source and generates 
+\emph{vectorized} MOTION code (from Optimized MPC Source). We then run the two versions using Boolean GMW and the BMR protocols. (MOTION, which is designed for 
+protocol mixing, supports Arithmetic GMW, however, it does not implement Comparison (CMP) and Multiplexing (MUX) as they would be rather inefficient.)
+In our experiments vectorized code exhibits 24x improvement on average in circuit generation time for Boolean GMW (20x for BMR), 7x reduction 
+in communication (2x), 97x reduction in number of gates (91x), 4x improvement in setup time (23x) and 21x improvement in online time (18x).
+%\ana{Summarize results.}
 
@@
 \secref{sec:results} presents the experimental evaluation.\secref{sec:related} discusses related work and \secref{sec:conclusion} concludes.
+
+All our code, including benchmark Python-like code, compilation phases, and generated MOTION code is available on Github. We do not provide the link only to preserve anonymity, 
+however, we will gladly make it available upon request from reviewers. The Github setup generates graphs and intermediate code for each program along each compiler phase and runs MOTION on 
+small inputs to generate tables of data (the experiments in the paper run on real LAN and WAN). We plan to release the link and code, which we believe will be useful to researchers.
 

### 2022-05-02 overleaf/tech_report/overview.tex
Freezing tech report.

--- 2022-04-30 overleaf/overview.tex
@@
 
-Our compiler takes essentially standard IMP~\ana{Need citation here.}
+Our compiler takes essentially standard IMP~\cite{Nipkow2014}
 syntax and imposes certain semantic restrictions. (We detail the restrictions in the following sections.)
@@
 
-We turn to our analytical model to compute the \emph{cost} of the iterative program. Assuming
-cost $\beta$ for a local MPC operation (essentially just ADD) and cost $\alpha$ for a remote
-MPC operation (e.g., MUX, CMP, and remaining operations), the cost of the iterative schedule
-will be $ND(2\alpha+\beta) + N(3\alpha$).
+We turn to our analytical model to compute the \emph{cost} of the iterative program. Assume
+cost $\beta$ for a local MPC operation (e.g., ADD in Arithmetic sharing) and cost $\alpha$ for a remote
+MPC operation (e.g., MUX, CMP, etc.). Assuming that ADD is $\alpha$ and SUM, CMP and MUX are $\beta$, 
+the MPC Source in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(b) gives 
+rise to an iterative schedule with cost $ND(2\alpha+\beta) + N(3\alpha$).
 
@@
 of an individual sum remains sequential, we can compute the $N$ sums in parallel.
-Our compiler \emph{automatically detects these opportunities and transforms the program}.
-It is standard that MPC researchers write vectorized versions of the Biometric
-program by hand; we are the first (to the best of our knowledge) to automatically
-transform a naive, iterative MPC program into an unintuitive vectorized one.
+%Our compiler \emph{automatically detects these opportunities and transforms the program}.
+%It is standard that MPC researchers write vectorized versions of the Biometric
+%program by hand; we are the first (to the best of our knowledge) to automatically
+%transform a naive, iterative MPC program into an unintuitive vectorized one.
 
@@
 Clearly, $2\alpha_\mathit{fix} + D\beta_\mathit{fix} + N(3\alpha_\mathit{fix}) << ND(2\alpha_\mathit{fix}+\beta_\mathit{fix}) + N3\alpha_\mathit{fix}$.
-Empirically, we observe 10x to 50x improvement over un-vectorized Biometric Matching in circuit generation time and setup time and 5x to 30x improvement in online time.
-Additionally, the un-vectorized version runs out of memory for $N=256$, while the vectorized one runs with the standard maximal input size $N=4,096$.
-\ana{Edit numbers with final experiments.}
+Empirically, we observe that $\alpha_\mathit{var} \approx 0$ and orders of magnitude improvement (e.g., 12x in online time in GMW). 
+Additionally, the un-vectorized version runs out of memory for $N=256$, while the vectorized one runs with the standard maximal input 
+size $N=4,096$.
+%\ana{Edit numbers with final experiments.}

### 2022-05-02 overleaf/introduction.tex
Edits to shorten.

@@
 
-Earlier compilers did both back-end and front-end translation as their aim was to demonstrate applicability of MPC on real-world
-programming problems. As the field advanced, works have focused more closely on front-end language design (e.g., Wysteria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021})
-or back-end ``circuit-level'' design and implementation (e.g., MOTION~\cite{Braun:2022}).
+%Earlier compilers did both back-end and front-end translation as their aim was to demonstrate applicability of MPC on real-world
+%programming problems. As the field advanced, works have focused more closely on front-end language design (e.g., Wysteria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021})
+%or back-end ``circuit-level'' design and implementation (e.g., MOTION~\cite{Braun:2022}).
 
@@
 We formalize the MPC Source~\cite{Ishaq:2019} intermediate representation and emphasize optimization over MPC Source.
-As in classical compilers, we envision different front ends (e.g., our front end IMP Source, other) compiling into MPC Source.
-MPC Source is particularly suitable for optimizations such as protocol mixing~\cite{Ishaq:2019, Fang:2022} and SIMD-vectorization,
+As in classical compilers, we envision different front ends (e.g., our front end IMP Source) compiling into MPC Source.
+MPC Source is particularly suitable for optimizations such as protocol mixing~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022} and SIMD-vectorization,
 which takes advantage of amortization at the circuit level. The MPC Source IR exposes the \emph{linear structure} of MPC programs, which simplifies program analysis; this is
 in contrast to source, which has if-then-else constructs. In the same time, MPC Source is sufficiently ``high-level''  to support analysis and optimizations that take into account
-control and data flow in a specific program. MPC Source facilitates cost modeling. 
+control and data flow in a specific program. %MPC Source facilitates cost modeling. 
 %MPC Source is small in size and analysis is tractable, as opposed to analysis over an unrolled circuit~\cite{Ishaq:2019}. 
@@
 We focus on the MOTION framework as our back-end for several reasons.
-First, it is the state of the art in terms of performance~\cite{Braun:2022}.
-Second, it provides an API over efficient implementation for a wide variety of cryptographic operations in three different protocols ---  Arithmetic GMW, Boolean GMW, and BMR --- which allows for protocol mixing~\cite{Ishaq:2019, Fang:2022}, a known backend-independent optimization.
-Third, MOTION provides API for SIMD-level operations, which amortize cost and lead to significant improvement in memory footprint and
-throughput~\cite{NDSS:DemSchZoh15, CCS:ABFKLO18, Braun:2022}. It enables MPC Source-level vectorization, a key focus of this paper.
+First, it demonstrates high performance~\cite{Braun:2022}.
+Second, it provides an API over efficient implementation for a wide variety of cryptographic operations (e.g., MUL, CMP, etc.) in three different protocols ---  Arithmetic GMW, Boolean GMW, and BMR --- which allows for protocol mixing~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022}, a known backend-independent optimization.
+Third, MOTION provides API for SIMD-level operations (e.g., MUL\_SIMD), which amortize cost and lead to significant improvement in memory footprint and throughput~\cite{NDSS:DemSchZoh15, CCS:ABFKLO18, Braun:2022}.
+It enables MPC Source-level vectorization, a key focus of this paper.
 
@@
 \secref{sec:results} presents the experimental evaluation.\secref{sec:related} discusses related work and \secref{sec:conclusion} concludes.
+
+All our code, including benchmark Python-like code, compilation phases, and generated MOTION code is available on Github. We omit the link for anonymity, 
+however, we will gladly make it available upon request from reviewers. The Github setup generates graphs and intermediate code for each program along each compiler phase, compiles with 
+MOTION and runs the circuits on small inputs to generate tables of data (the experiments we present later run on real LAN and WAN).
+We plan to release the link and code, which we believe will be useful to researchers.
 

### 2022-05-02 overleaf/model.tex
Edits to shorten.

@@
 \begin{enumerate}
+
 \item All statements in the program execute using the same protocol (sharing). That is, there is no share conversion.
@@
 -- Past suggestion from Vassilis: Ideally we should assume cost to be a convex function.}\ana{After discussion 1/12: replace single costs 1 with two levels of costs: $\alpha$ for non-local operations, e.g., MUL, and $\beta$ for local ops, e.g., ADD.}
-\item  In MPC frameworks, executing $n$ operations ``at once'' in a single SIMD operation costs a lot less than executing those $n$ operations one by one.
+\item  %In MPC frameworks, executing $n$ operations ``at once'' in a single SIMD operation costs a lot less than executing those $n$ operations one by one.
 Following Amdahl's law, we write $\alpha = \frac{1}{s}p\alpha + (1-p)\alpha$, where $p$ is the fraction of execution time that benefits from amortization and $(1-p)$

### 2022-05-02 overleaf/overview.tex
Edits to shorten.

@@
 
-Our compiler takes essentially standard IMP~\ana{Need citation here.}
-syntax and imposes certain semantic restrictions. (We detail the restrictions in the following sections.)
+Our compiler takes essentially standard IMP~\cite{Nipkow2014}
+syntax and imposes certain semantic restrictions. %(We detail the restrictions in the following sections.)
 The programmer writes an iterative program and annotates certain inputs
@@
 are \texttt{shared}, meaning that they store shares, however, the array sizes {\sf D} and
-{\sf N} respectively are plaintext. The code iterates over the entries in the
-database and computes the Euclidean distance of the current
-entry ${\sf S}[i]$ and {\sf C} (its square actually). The program returns the index of the vector that gives
-the best match plus the corresponding sum of squares.
+{\sf N} respectively are plaintext. 
+%The code iterates over the entries in the
+%database and computes the Euclidean distance of the current
+%entry ${\sf S}[i]$ and {\sf C} (its square actually). The program returns the index of the vector that gives
+%the best match plus the corresponding sum of squares.
 
@@
 
-We turn to our analytical model to compute the \emph{cost} of the iterative program. Assuming
-cost $\beta$ for a local MPC operation (essentially just ADD) and cost $\alpha$ for a remote
-MPC operation (e.g., MUX, CMP, and remaining operations), the cost of the iterative schedule
-will be $ND(2\alpha+\beta) + N(3\alpha$).
+We turn to our analytical model to compute the \emph{cost} of the iterative program. Assume
+cost $\beta$ for a local MPC operation (e.g., ADD in Arithmetic sharing) and cost $\alpha$ for a remote
+MPC operation (e.g., MUX, CMP, etc.). Assuming that ADD is $\beta$ and SUB, CMP and MUX are $\alpha$, 
+the MPC Source in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(b) gives 
+rise to an iterative schedule with cost $ND(2\alpha+\beta) + N(3\alpha$).
 
 A key contribution is the vectorizing transformation. We can compute all $N*D$
-subtraction operations (line 9 in (b)) in a single SIMD instruction; similarly we can compute
-all multiplication operations (line 10) in a single SIMD instruction. And while computation
+subtractions (line 9 in (b)) in a single SIMD instruction; similarly we can compute
+all multiplications (line 10) in a single SIMD instruction. And while computation
 of an individual sum remains sequential, we can compute the $N$ sums in parallel.
-Our compiler \emph{automatically detects these opportunities and transforms the program}.
-It is standard that MPC researchers write vectorized versions of the Biometric
-program by hand; we are the first (to the best of our knowledge) to automatically
-transform a naive, iterative MPC program into an unintuitive vectorized one.
+%Our compiler \emph{automatically detects these opportunities and transforms the program}.
+%It is standard that MPC researchers write vectorized versions of the Biometric
+%program by hand; we are the first (to the best of our knowledge) to automatically
+%transform a naive, iterative MPC program into an unintuitive vectorized one.
 
@@
 This gives rise to the following formula for amortized cost: $f(n) = \alpha_\mathit{fix} + n\alpha_\mathit{var}$,
-as opposed to unamortized cost $g(n) = n\alpha_\mathit{fix} + n\alpha_\mathit{var}$. (We extend the same reasoning to
-$\beta$-instructions.)
+as opposed to unamortized cost $g(n) = n\alpha_\mathit{fix} + n\alpha_\mathit{var}$. We extend the same reasoning to
+$\beta$-instructions.
 
@@
 Clearly, $2\alpha_\mathit{fix} + D\beta_\mathit{fix} + N(3\alpha_\mathit{fix}) << ND(2\alpha_\mathit{fix}+\beta_\mathit{fix}) + N3\alpha_\mathit{fix}$.
-Empirically, we observe 10x to 50x improvement over un-vectorized Biometric Matching in circuit generation time and setup time and 5x to 30x improvement in online time.
-Additionally, the un-vectorized version runs out of memory for $N=256$, while the vectorized one runs with the standard maximal input size $N=4,096$.
-\ana{Edit numbers with final experiments.}
+Empirically, we observe that (1) $\alpha_\mathit{var} \approx 0$ and (2) there is orders of magnitude improvement in running time and memory. 
+E.g., we see about 12x improvement in online time in GMW for $N=128$. Additionally, the non-vectorized version runs out of memory for $N=256$, 
+while the vectorized one runs with the standard maximal input size $N=4,096$.
+%\ana{Edit numbers with final experiments.}

### 2022-05-02 overleaf/paper_SIMD.tex
Edits to shorten.

@@
 
-In this work, we formalize the MPC Source intermediate language and advance the idea of what we call backend-independent optimizations, in a close analogy to machine-independent optimizations in the classical compiler.
+In this work, we formalize the MPC Source intermediate language and advance what we call \emph{backend-independent} optimizations, in a close analogy to machine-independent optimizations in the classical compiler.
 We present a compiler framework that takes a Python-like routine and produces MOTION code.
-We focus on a specific backend-independent optimization: novel SIMD-vectorization on MPC Source, which we show leads to significant speedup in circuit generation time and running time, as well as significant reduction in communication and number of gates over the corresponding iterative schedule.
+We focus on a specific backend-independent optimization: novel SIMD-vectorization on MPC Source, which we show leads to significant improvement in circuit generation time, running time, and communication over the corresponding iterative schedule.
 %Our results emphasize the importance of backend-independent optimizations --- vectorization (this work) and protocol mixing (previous work) are two optimizations available at the level of MPC Source.

### 2022-05-02 overleaf/prob_stmt_np_hardness.tex
Edits to shorten.

@@
 
-As mentioned earlier, at the lowest level, we have two types of MPC instructions (also called \emph{gates} in similar works) 1) local/non-interactive instruction (i.e.
-an addition instruction $A$) and 2) remote/interactive instruction (i.e.
-a multiplication instruction $M$).
+As mentioned earlier, at the lowest level, we have two types of MPC instructions (also called \emph{gates} or \emph{operations} in similar works) 1) local/non-interactive (e.g., an addition instruction $A$) and 2) remote/interactive (e.g., a multiplication instruction $M$).
 %Each instruction in the program is either an $A$-instruction or an $M$-instruction.
@@
     \item All $P_i$'s consist of MPC instructions of the same kind, e.g., all MUL, MUX, ADD, etc. 
-    \item Def-use dependencies of the graph $G(V, E)$ should be preserved i.e.
+    \item Def-use dependencies of the graph $G(V, E)$ are preserved i.e.
 if instructions $S_i, S_j, i < j$ form a def-use i.e.
@@
 %$cost(S_i) = \alpha$ or $\beta$. if $S_i$ is an $A$-instruction and $cost(S_i) = \beta$ if $S_i$ is an $M$-instruction. 
-The cost of schedule $P$ is more interesting:
+The cost of $P$ is more interesting:
 
@@
 
-The problem of finding a schedule $P$ with a minimal $cost(P)$ is shown to be NP-Hard problem, as it can be reduced to the problem of finding a \emph{shortest common supersequence}, a known NP-Hard problem\cite{Maier1978, Vazirani2010}.
-The shortest common supersequence problem is as follows: {\it given two or more sequences find the the shortest sequence that contains all of the original sequences.} This can be solved in $O(n^k)$ time, where $n$ is the cardinality of the longest sequence and $k$ is the number of sequences.
+%The problem of finding a schedule $P$ with a minimal $cost(P)$ is shown to be NP-Hard problem, as it can be reduced to the problem of finding a \emph{shortest common supersequence}, a known NP-Hard problem~\cite{Vazirani:2010}.
+The shortest common supersequence problem~\cite{Vazirani:2010} is as follows: {\it given two or more sequences find the the shortest sequence that contains all of the original sequences.} This can be solved in $O(n^k)$ time, where $n$ is the cardinality of the longest sequence and $k$ is the number of sequences.
 %For our problem $n$ is the maximum length of a sequence and $k$ is the number of instructions of sequences.
-We can immediately see that the optimal schedule is the shortest schedule, since the shortest schedule minimizes the fixed cost while the variable cost remains the same.
+We can see that the optimal schedule is the shortest schedule, since the shortest schedule minimizes the fixed cost while the variable cost remains the same.
 

### 2022-05-02 overleaf/vectorization.tex
Edits to shorten.

@@
 
+In this section, we briefly describe the analysis. Details, including edge cases and examples, can be found in the supplementary extended version~\cite{Anon_TR}.
 \secref{sec:dependence} describes our dependence analysis and ~\secref{sec:expansion} describes scalar expansion,
@@
 \item same-level edge $X\rightarrow Y$ where $X$ and $Y$ are in the same loop nest, say $i,j,k$.
-E.g., the def-use edge 9 to 10 in the Biometric MPC Source in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source} is a same-level edge.
-A same-level edge can be a back-edge in which case a $\phi$ node is the target of the edge.
-E.g., 15 to 4 in Biometric is a same-level back-edge.
+E.g., the def-use edge 9 to 10 in the Biometric MPC Source in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(b) is a same-level edge.
+%A same-level edge can be a back-edge in which case a $\phi$ node is the target of the edge. E.g., 15 to 4 in Biometric is a same-level back-edge.
 \item outer-to-inner $X\rightarrow Y$ where $X$ is in an outer loop nest, say $i$, and $Y$ is in an inner one, say $i,j,k$. E.g., 1 to 4 in Biometric forms is an outer-to-inner edge.
@@
 %E.g. $\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$ is an inner-to-outer edge.
-An inner-to-outer edge can be a back-edge as well, in which case both $X$ and $Y$ are $\phi$-nodes with the source $X$ in a loop nested into $Y$'s loop (not necessarily immediately).
+%An inner-to-outer edge can be a back-edge as well, in which case both $X$ and $Y$ are $\phi$-nodes with the source $X$ in a loop nested into $Y$'s loop (not necessarily immediately).
 
@@
 This transforms a mixed forward edge into an "inner-to-outer" forward edge followed by an outer-to-inner forward edge.
-Thus, Basic Vectorization handles one of "same-level", "inner-to-outer", or "outer-to-inner" def-use edges.
+%Thus, Basic Vectorization handles one of "same-level", "inner-to-outer", or "outer-to-inner" def-use edges.
  \end{itemize}
@@
 
-An important component of our algorithm is the scalar expansion to the corresponding loop dimensionality, 
+An important component of our algorithm is expansion of scalars and arrays to the corresponding loop dimensionality, 
 which is necessary to expose opportunities for vectorization.
@@
 
+Expansion (and also reduction) is done with the $\mathit{raise\_dim}$ and $\mathit{drop\_dim}$ functions, 
+which we believe are standard. $\mathit{raise\_dim}$ takes the original array, the access function 
+$f(i,j,k)$ in loop nest $i,j,k$ and the loop bounds $((i\!:\!I),(j\!:\!J),(k\!:\!K))$ and produces a new 3-dimensional array 
+${\sf A'}$ by iterating over $i,j,k$ and setting each element of  ${\sf A'}$:
+\[ \mathit{raise\_dim}(A, f(i,j,k), ((i\!:\!I),(j\!:\!J),(k\!:\!K))): {\sf A'}[i,j,k] =  {\sf A}[f(i,j,k)] \]
+Raise dimension applies when reshaping input arrays and also, at outer-to-inner def-use edges.
+Analogously $\mathit{drop\_dim}$ applies at inner-to-outer def-use edges.
+It takes a higher dimensional array, say $i,j,k$ and removes trailing dimensions, say $j,k$. 
+It iterates over $i$ and takes the result at the maximal index of $j$ and $k$, i.e., the result at the last iterations of $j$ and $k$:
+\[ \mathit{drop\_dim}({\sf A}, (j\!:\!J,k\!:\!K)): {\sf A'}[i] = A[i,J-1,K-1] \] 
+%It iterates over $i$ and takes the result at the maximal index of $j$ and $k$, i.e., the result at the last iterations of $j$ and $k$:
+%\[ {\sf A'}[i] = A[i,J-1,K-1] \]
+
+\begin{comment}
 \paragraph{Arrays} Conceptually, we treat all variables as arrays. There are three kinds of arrays. 
@@
 \[ {\sf A'}[i] = A[i,J-1,K-1] \]
+\end{comment}
 
@@
 
-We present our algorithm followed by an example on Biometric.
-
-\subsubsection{Algorithm}
+%\subsubsection{Algorithm}
 
@@
 and Phase 4 explicitly turns operations into MOTION SIMD operations.
+
+\paragraph{Pseudocode:}
 
@@
 \FOR {each closure $cl$ (after merge)}
-\STATE {create \python{for} $i_d$ \python{in} ... loop}
+\STATE {create \python{for} $i_d$ \python{in} ... loop} \COMMENT{ i.e., MOTION loop }
 \STATE {add PHI-nodes in $cl$ to header block} 

### 2022-05-02 overleaf/tech_report/overview.tex
Edits to overview.

@@
 cost $\beta$ for a local MPC operation (e.g., ADD in Arithmetic sharing) and cost $\alpha$ for a remote
-MPC operation (e.g., MUX, CMP, etc.). Assuming that ADD is $\alpha$ and SUM, CMP and MUX are $\beta$, 
+MPC operation (e.g., MUX, CMP, etc.). Assuming that ADD is $\beta$ and SUM, CMP and MUX are $\alpha$, 
 the MPC Source in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(b) gives 

### 2022-05-02 overleaf/vectorization.tex
Took a pass cutting on Vectorization.

@@
 and similarly, if there is an outer-to-inner dependence, it inserts $\mathit{drop\_dim}$. After this phase operations work
-on arrays of the corresponding dimensionality and we optimistically vectorize all arrays.
+on arrays of the corresponding dimensionality and we optimistically vectorize all arrays. For example, 
+lines 4 and 5 in Biometric~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(c) reshape the input arrays
 
 Phase 2 proceeds from the inner-most towards the outer-most loop. For each loop it anchors dependence cycles (closures) 
-around pseudo PHI nodes then removes vectorization from the dimension of that loop. There are two important points in this 
+around pseudo PHI nodes then removes vectorization from the dimension of that loop. In the 
+Biometric example~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(b), the iteration on the inner loop (7-11)
+determines that $j$ is not vectorizable. The next iteration on the outer loop determines that $i$ is vectorizable. It also 
+splits the remainder of the loop anchoring each loop around the corresponding PHI node.
+
+There are two important points in this 
 phase. First, it may break a loop into smaller loops which would discover opportunities 
@@
 
-\subsection{Correctness Argument}
-\label{sec:correctness}
-
-We build a correctness argument that loosely follows Abstract Interpretation.
-First we define the MPC Source syntax. The domain of MPC Source programs
-expressible in the syntax (with certain semantic restrictions) is the abstract domain $A$.
-We then define the \emph{linearization} of an MPC Source program as an interpretation
-over the syntax. The linearization, which is a \emph{schedule} (as in~\secref{sec:overview}), is the concrete domain $C$.
-Since we reason over def-use graphs in $A$ we define a partial order relation over elements of $A$
-in terms of def-use relations. We define a partial order over elements of $C$ as well,
-in terms of def-use relations in the concrete domain $C$. We prove two theorems
-that state (informally) that the schedule corresponding to the original program computes the same
-result as the schedule corresponding to the vectorized program.
-
 \begin{comment}
@@
 \end{figure*}
+
+\subsection{Correctness Argument}
+\label{sec:correctness}
+
+We build a correctness argument that loosely follows Abstract Interpretation.
+First we define the MPC Source syntax. The domain of MPC Source programs
+expressible in the syntax (with certain semantic restrictions) is the abstract domain $A$.
+We then define the \emph{linearization} of an MPC Source program as an interpretation
+over the syntax. The linearization, which is a \emph{schedule} (as in~\secref{sec:overview}), is the concrete domain $C$.
+Since we reason over def-use graphs in $A$ we define a partial order relation over elements of $A$
+in terms of def-use relations. We define a partial order over elements of $C$ as well,
+in terms of def-use relations in the concrete domain $C$. We state two theorems
+that state (informally) that the schedule corresponding to the original program computes the same
+result as the schedule corresponding to the vectorized program.
 
@@
 
-Theorem 1 states that a def-use edge in MPC Source $a_0$ gives rise to a set of def-use edges in $\gamma(a_0)$
-and a ``same'' edge in $a_1$ gives rise to a set of def-use edges in $\gamma(a_1)$ and the former set is included in the latter one.
-This means that the computation of $\gamma(a_0)$ is carried out by $\gamma(a_1)$ as well, computing the same result. 
-The theorem allows that $a_1$ and $\gamma(a_1)$ have larger state that $a_0$ and $\gamma(a_0)$, i.e., doing additional 
-computation into additional memory locations. Theorem 2 establishes the correctness of vectorization --- 
-clearly, vectorization preserves all statements and def-use edges in the original MPC Source, and therefore the linearization computes the 
-same result as the original MPC Source. (In this case we have an equivalence, $a_0 = a_1$ according to our partial order.)
+These theorems simply state that the transformation preserves def-use relations which is an invariant of the algorithm.
+
+%Theorem 1 states that a def-use edge in MPC Source $a_0$ gives rise to a set of def-use edges in $\gamma(a_0)$
+%and a ``same'' edge in $a_1$ gives rise to a set of def-use edges in $\gamma(a_1)$ and the former set is included in the latter one.
+%This means that the computation of $\gamma(a_0)$ is carried out by $\gamma(a_1)$ as well, computing the same result. 
+%The theorem allows that $a_1$ and $\gamma(a_1)$ have larger state that $a_0$ and $\gamma(a_0)$, i.e., doing additional 
+%computation into additional memory locations. Theorem 2 establishes the correctness of vectorization --- 
+%clearly, vectorization preserves all statements and def-use edges in the original MPC Source, and therefore the linearization computes the 
+%same result as the original MPC Source. (In this case we have an equivalence, $a_0 = a_1$ according to our partial order.)
 
@@
 
-\subsubsection{Removal of Infeasible Edges}
+%\subsubsection{Removal of Infeasible Edges}
 
-Array writes limit vectorization as they sometimes introduce infeasible loop-carried dependencies. Consider the classical example from~\cite{Aiken:1988}:
+Array writes limit vectorization as they sometimes introduce infeasible loop-carried dependencies. For example, consider 
+one of our benchmarks, Histogram. The source takes two arrays {\sf A} and {\sf B} of size {\sf N}, where ${\sf A}[j]$ stores the bin number and
+${\sf B}[j]$ stores the corresponding value that we need to add. The MPC Source is:
+
+{\small
+\begin{pythonn}
+for i in range(0, num_bins):
+  res1 = PHI(res, res2)
+  for j in range(0, N):
+    res2 = PHI(res1, res3)
+    tmp1 = (A[j] == i)
+    tmp2 = (res2[i] + B[j])
+    tmp3 = MUX(tmp1, res2[i], tmp2)
+    res3 = Update(res2, i, tmp3)
+return res1
+\end{pythonn}
+}
+
+There is loop-carried dependence from 8 to 2 for write array {\sf res} even though the loop can be vectorized across $i$.
+
+\begin{comment}
+Consider the classical example from~\cite{Aiken:1988}:
 
@@
 read of {\sf ... = B\_1[i]} at 6.
+\end{comment}
 
 The following algorithm removes certain infeasible loop-carried dependencies that are due to array writes. Consider a loop with index $0 \le j < J$
-nested at $i,j,k$. Here $i$ represents the enclosing loops of $j$ and $k$ represents the enclosed loops in $j$.
-
+nested at $i,j,k$. 
+%Here $i$ represents the enclosing loops of $j$ and $k$ represents the enclosed loops in $j$.
+\begin{comment}
 \begin{algorithmic}
@@
 \end{algorithmic}
-
+\end{comment}
 Consider a loop $j$ enclosed in some fixed $\underline{i}$. Only if an update (definition) ${\sf A}_m[f(i,j,k)] = ... $ at some iteration $\underline{j}$
@@
 we may have a loop-carried dependence for ${\sf A}$ due to this def-use pair. (In contrast, Cytron's algorithm inserts a loop-carried dependency every time there is an array update.)
-The algorithm above examines all def-use pairs in loop $j$, including defs and uses in nested loops, searching for values $\underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$ that satisfy
-$f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$. If such values exist for some def-use pair, then there is a potential
-loop-carried dependence on ${\sf A}$; otherwise there is not and we can remove the spurious backward edge thus ``freeing up'' statements for vectorization. %\ana{Double check this!!!}
+The algorithm, shown in~\cite{Anon_TR}, examines all def-use pairs in loop $j$, including defs and uses in nested loops, 
+searching for values $\underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$ that satisfy
+$f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$ (using Z3). If such values exist for some def-use pair, then there is a potential
+loop-carried dependence on ${\sf A}$; otherwise there is not and we can remove the spurious backward edge to the array PHI-node 
+thus ``freeing up'' statements for vectorization. %\ana{Double check this!!!}
 
+In the Histogram example, this removes the back edge from the update at 8 to the PHI node at 2. Clearly there does not exist $i < i'$ that will make 
+$i$ at 9 and $i'$ at 6 and 7 equal. The back-edge from 9 to 4 stays because for every $j < j'$ 9 and 6 and 7 access the same location in {\sf res}. %\ana{Can't wrap my head around this...}
+
+\begin{comment}
 Consider the earlier example.
@@
 Statements 7, 8 and 9 will correctly appear in the FOR loop.
+\end{comment}
 
@@
 array write can be vectorized, we have to properly remove the phi-node replacing uses of the left-hand side of the phi-node with its arguments.
+Additional detail, as well as detailed treatment of the minor changes to $\mathit{raise\_dim}$ and $\mathit{drop\_dim}$ required to do expansion of arrays 
+in addition to scalars, appear in~\cite{Anon_TR}.
 
@@
 
+\begin{comment}
 \subsubsection{Restricting Array Writes}
@@
 
+\end{comment}
+
+\begin{comment}
 \subsubsection{Examples with Array Writes}
@@
 
+\end{comment}
+
 \begin{comment}
@@
 
+\begin{comment}
 \paragraph{Example 2} Now consider the MPC Source of Histogram:
@@
 }
+\end{comment}
 
-The canonical dimensionality of {\sf res} is 1.
+Returning to the histogram example, the canonical (i.e., declared) dimensionality of {\sf res} is 1.
 Also, the phi-node {\sf res1 = PHI(res, res2)} is a target-less phi-node (the implication being that the inner for loop can be vectorized across $i$).
@@
 Processing the outer loop is interesting.
-This is because the PHI node is a target-less node, and therefore, there are no closures!
-Several things happen.
+This is because the PHI node is a target-less node, and therefore, there are no closures.
 (1) Everything can be vectorized along the $i$ dimension.

### 2022-05-02 overleaf/related.tex
Committing changes to Related work and Results.

@@
 These two families are the state-of-the art and
-are actively developed. Another recent development is Viaduct, a functional language and compiler that supports a range of secure computation
-frameworks, including MPC and ZKP. Hastings et el. present a review of compiler frameworks~\cite{Hastings:2019}.
+are actively developed. Another recent development is Viaduct~\cite{Acay:2021}, a functional language and compiler that supports a range of secure computation
+frameworks, including MPC and ZKP. Hastings et el. present a review of compiler frameworks~\cite{Hastings:2019}. We believe that our work is unique as 
+as a facade between the high-level language and lower-level circuit protocols.
 
+\begin{comment}
 While each of these languages and compilers brings in new ideas and advances, none addresses the problem of ``circuit independent'' intermediate
@@
 restrict flow between secure and insecure parts of the program.
+\end{comment}
 
@@
 \paragraph{Classical HPC compilers.}
-Automatic vectorization is a longstanding problem in high-performance computing (HPC).
-There are thousands of works in this area reflecting over 40 years of research. We presented a vectorization
+Automatic vectorization is a longstanding problem in high-performance computing and
+there are thousands of works in this area reflecting over 40 years of research. We presented a vectorization
 algorithm for MPC Source, essentially extending classical loop vectorization \cite{Allen:1987}. In HPC vectorization, conditional control flow
@@
 results in HPC.
-Since both branches of the conditional and the multiplexer \emph{always} execute, not only can we apply aggressive vectorization on linear code, but (perhaps more importantly)
-we can also build analytical models that accurately predict execution time. These models in turn would drive optimizations such as vectorization, protocol mixing, and others.
-Vectorization meshes in with those additional optimizations in non-trivial ways.
-
-Furthermore, extensions of classical loop vectorization with array writes, arbitrary indexing, including non-affine indexing, and interaction with SSA are non-trivial
-and present novel challenges and opportunities for contribution. Polyhedral parallelization~\cite{Benabderrahmane:2010} considers a higher-level source (typically AST)
+%Since both branches of the conditional and the multiplexer \emph{always} execute, not only can we apply aggressive vectorization on linear code, but (perhaps more importantly)
+%we can also build analytical models that accurately predict execution time. These models in turn would drive optimizations such as vectorization, protocol mixing, and others.
+%Vectorization meshes in with those additional optimizations in non-trivial ways.
+%Furthermore, extensions of classical loop vectorization with array writes, arbitrary indexing, including non-affine indexing, and interaction with SSA are non-trivial
+%and present novel challenges and opportunities for contribution. 
+Polyhedral parallelization~\cite{Benabderrahmane:2010} considers a higher-level source (typically AST)
 representation, while our work takes advantage of linear MPC Source and SSA form. The work by Karrenberg~\cite{Karrenberg:2015} is rare in that space, in the sense that it considers

### 2022-05-02 overleaf/results.tex
Committing changes to Related work and Results.

@@
 \subsection{Results and Analysis}\label{sec:result_analysis}
-
+{\small
    \begin{table*}[htbp]
@@
    \end{table*}
-
+}
 \begin{figure*}[htbp]
@@
 Note that in some settings it may still be desirable to vectorize e.g.
-in data constrained environments.
-As shown in \cref{fig:graph_comm_size}, vectorization results in reduced communication (fewer bits are transferred).
+in data constrained environments as communication is reduced for all benchmarks.
+(The report presents plots of communication, circuit generation time, number of gates, and separate online and setup time~\cite{Anon_TR}.)%As shown in \cref{fig:graph_comm_size}, vectorization results in reduced communication (fewer bits are transferred).
 
-We present evaluation for communication size in \cref{fig:graph_comm_size}, circuit generation time in \cref{fig:graph_circ_gen_time}, number of gates in \cref{fig:graph_total_gates} and online time and setup time in \cref{fig:graph_online_time} and \cref{fig:graph_setup_time} respectively.
+%We present evaluation for communication size in \cref{fig:graph_comm_size}, circuit generation time in \cref{fig:graph_circ_gen_time}, number of gates in \cref{fig:graph_total_gates} and online time and setup time in \cref{fig:graph_online_time} and \cref{fig:graph_setup_time} respectively.
 %\ishaq{Ana: Let me know if I should add more substance.} \ana{It's good!}
@@
 
+\begin{comment}
 \begin{figure*}[htbp]
@@
 \end{figure*}
+\end{comment}
 
@@
 however, our addition loop incurred significant cost (ADD is non-local and expensive in Boolean GMW). The MOTION-native loop ran 
-{\sf result += mult\_unsimdified[i];} while our loop generated and ran {\sf result[i] = result[i-1] + mult\_unsimdified[i];} recall that the scalar expansion
-is an artifact of our vectorization. We rewrote the accumulation (manually, for testing purposes) and that led to the comparable speedup!
+{\sf result += mult\_unsimdified[i];} while our loop ran auto generated {\sf result[i] = result\_prev[i] + mult\_unsimdified[i];} (recall that scalar expansion
+is an artifact of our vectorization). We rewrote the accumulation (manually, for testing purposes) and that led to the comparable speedup!
 
 MOTION's compiler performs analysis that informs circuit generation and the example illustrates the power of the analysis. In the above example, 
-MOTION does the standard divide-and-conquer accumulation. The poor performance of our loop is not a limitation of our approach, but it could be a 
-limitation of our current implementation. Recall that the next phase of Vectorization, in~\secref{sec:vectorization}, which we have not implemented yet, gets rid of redundant 
-dimensions and generates {\sf if (i!=0) { result\_pref = result; } result = result\_prev + mult\_unsimdified[i]}. And while it is unrealistic to expect that MOTION's static analysis 
-will detect the associative accumulation in the scalar expansion code, it is realistic to expect that it will in the above only slightly more complex code. 
-It still might lead to confusion in the static analysis as analysis on the AST is difficult (unfortunately, we ran out of time and did not experiment further).
+MOTION does the standard divide-and-conquer accumulation. We conjectured that poor performance of our loop was due to a
+limitation of our current implementation. Recall that the Phase 3 of Vectorization, in~\secref{sec:vectorization}, which we have not implemented yet, gets rid of redundant 
+dimensions and generates {\sf if (i!=0) \{ result\_prev = result; \} result = result\_prev + mult\_unsimdified[i]}. And while it is unrealistic to expect that MOTION's static analysis 
+will detect the associative accumulation in the scalar expansion code, one might expect that it will in the above slightly more complex code. 
+Our experience indicates the limitation of AST-based analysis; because AST analysis is difficult it does not appear to detect the associative accumulation.
+%(unfortunately, we ran out of time and did not experiment further).
 %Interestingly, the non-vectorized version, which is the literal translation of inner product does not appear to benefit from any optimizations.
 
-We conjecture that MPC Source, a straight-forward representation, will not only enable detection of general associative loops, but also allow 
+We conjecture that MPC Source, a straight-forward representation suitable for static analysis, will not only enable detection of general associative loops, but also allow 
 for program synthesis to increase opportunities for divide-and-conquer parallelization~\cite{Farzan:2021}; we leave this as future work.

### 2022-05-02 overleaf/results.tex
Minor edits to results.

@@
 We conjecture that MPC Source, a straight-forward representation suitable for static analysis, will not only enable detection of general associative loops, but also allow 
-for program synthesis to increase opportunities for divide-and-conquer parallelization~\cite{Farzan:2021}; we leave this as future work.
+for program synthesis to increase opportunities for divide-and-conquer parallelization~\cite{Farzan:2021}; we leave this and other optimizations as future work.
 

### 2022-05-02 overleaf/backend.tex
Final edits. Will freeze TR.
Shortening sections. Still way too long...

@@
 following the type system in~\secref{sec:syntax}. We require that all inputs are marked as either shared or plaintext,
-however, we infer qualifiers for the rest of the variables. \secref{sec:taint} briefly describes the taint analysis
+however, we infer qualifiers for the rest of the variables. Type inference is done on MPC Source 
+and amounts to a standard positive-negative qualifier system (\texttt{shared} is positive and \texttt{plain} is negative). 
+Translation from MPC Source to MOTION C++ code is relatively straightforward. The extended version~\cite{Anon_TR} details the taint analysis and code generation.
+
+\begin{comment}
+\secref{sec:taint} briefly describes the taint analysis
 and \secref{sec:MOTION} describes MOTION code generation.
@@
 %\ben{Is this necessary to include?} \ana{Yes, that's important.}
+\end{comment}
 
-\subsection{From (Optimized) MPC Source to MOTION}
-\label{sec:MOTION}
+%\subsection{From (Optimized) MPC Source to MOTION}
+%\label{sec:MOTION}
 
-MOTION supports FOR loops and SIMD operations, so translation from MPC source to MOTION C++ code is relatively straightforward.
 \paragraph{Variable declarations:}
 % Variable naming
-Our generated C++ uses the following variable-naming scheme: shared variables are named the same as in the MPC Source with the {\sf !} replaced with an underscore (e.g.
+%Our generated C++ uses the following variable-naming scheme: shared variables are named the same as in the MPC Source with the {\sf !} replaced with an underscore (e.g.
 {\sf sum!2} would be translated to {\sf sum\_2}).
@@
 Next, all constant values which are used as part of shared expressions are initialized as a shared input from party 0.
-Finally, plaintext parameters are converted used as shared inputs from party 0 to initialize their shared counterparts.
+Finally, plaintext parameters are converted as shared inputs from party 0 to initialize their shared counterparts.
+(We note that inputs are read-only and
+the conversion currently incurs 2-3 redundant input gates per application; this can be easily fixed.)
 
@@
 % Non-vectorized assignments and returns
-Once the function preamble is complete, the MPC Source is translated into C++ one statement at a time.
-The linear structure of MPC Source enables this approach to translation.
-If there is no vectorization present in a statement, translation to C++ is straightforward: outside of MUX statements and array updates, non-vectorized assignments, expressions, and returns directly translate into their C++ equivalents.
+Once the function preamble is complete, the MPC Source is translated into C++ one statement at a time. The linear structure of MPC Source enables this approach to translation. 
+E.g., MPC Source for-loops are converted to C++ for-loops which iterate the loop counter over the specified range. The psuedo-PHI node has the expected semantics as detailed in~\cite{Anon_TR}.
+%If there is no vectorization present in a statement, translation to C++ is straightforward: outside of MUX statements and array updates, non-vectorized assignments, expressions, and returns directly translate into their C++ equivalents.
 Non-vectorized MUX statements are converted to MOTION's MUX member function on the condition variable.
@@
 
+\begin{comment}
 \begin{table*}
@@
 \end{table*}
+\end{comment}
 
 % Phi nodes and FOR loops
-MPC FOR loops are converted to C++ FOR loops which iterate the loop counter over the specified range.
+%MPC FOR loops are converted to C++ FOR loops which iterate the loop counter over the specified range.
 Pseudo PHI nodes are broken into two components: the ``FALSE'' branch which assigns the initial value of the PHI node and the ``TRUE'' branch which assigns the PHI node's back-edge.
@@
 \end{table*}
+\end{comment}
 
@@
 Leave as is.}
-Whenever a SIMD value is used in an expression, the utility function {\sf vectorized\_access()} takes the multi-dimentional representation of a SIMD value, along with the size of each dimension and the requested slice's indices, and converts that slice to a MOTION SIMD value.
+Whenever a SIMD value is used in an expression, the utility function {\sf vectorized\_access()} (see Listing~\ref{tab:motion_translation_simd_assignment}) takes the multi-dimentional representation of a SIMD value, along with the size of each dimension and the requested slice's indices, and converts that slice to a MOTION SIMD value.
 Because MOTION supports SIMD operations using the same C++ operators as non-SIMD operations, we do not need to perform any other transformations to the expression.
@@
 Therefore, once vectorized accesses are inserted the translation of an expression containing SIMD values is identical to that of expressions without SIMD values.
-
-Similarly, the {\sf vectorized\_assign()} function assigns a (potentially SIMD) value to a slice of a vectorized array.
+Similarly, {\sf vectorized\_assign()} assigns a (potentially SIMD) value to a slice of a vectorized array.
 This operation cannot be done with a simple subscript as SIMD assignments will update a range of values in the underlying array representation.
+Listing~\ref{tab:motion_translation_simd_assignment} illustrates the ADD in Biometric that is vectorized along the $i$ dimension.
 %\ben{Is the next sentence needed?  They can always check the implementation of vectorized\_assign() if they want to see specifics.}
@@
 %\ana{I would leave this in the TR but it will go from the paper. Lots of stuff in Secs. 4 and 5 will go from the paper.} \ana{Leaving in TR, will shorten in paper.}
-Updating SIMD arrays is also implemented differently from updating non-vectorized arrays.
+%Updating SIMD arrays is also implemented differently from updating non-vectorized arrays.
 Instead of separating the array update from the assignment of the new array, these steps are combined with the {\sf vectorized\_update()} utility function.
@@
 
+\begin{comment}
 %\ben{The compiler and C++ internally uses ``lift'' as the terminology for raise\_dim(), but the paper does not :(...} \ana{True.
@@
 
+\end{comment}
+
 % Shared/plaintext conversions
@@
 MOTION does not support publicly-known constants for these protocols, so all conversions from plaintext values to shares are performed by providing the plaintext value as a shared input from party 0.
-Due to this limitation, our translation to MOTION code attempts to minimize the number of conversions from a plaintext value.
-This is accomplished by creating a shared copy of each plaintext variable and updating that copy in lock-step with the plaintext variable.
-Since variables are often initialized to a common constant value (e.g.
+%Due to this limitation, our translation to MOTION code attempts to minimize the number of conversions from a plaintext value.
+This is accomplished by
+To minimize conversions we create a shared copy of each plaintext variable and update that copy in lock-step with the plaintext variable.
+%Since variables are often initialized to a common constant value (e.g.
 0), this approach decreases the number of input gates by only creating a shared input for each initialization constant.
-Loop counters must still be converted to a shared value on each iteration that they are used, however we only generate this conversion when necessary, i.e., when the counter flows to a shared computation.
-This is to prevent unnecessary increase in the number of input gates when loop counters are only used as plaintext.
-\ben{Ishaq or Vassilis: please check over this paragraph... I don't want to make some mistake about MOTION's capabilities.}
+Loop counters are the one major case that trigger such update (and thus many incur runtime cost as they must be converted to a shared value on each iteration); thus, we only generate this conversion when necessary, i.e., when the counter flows to a shared computation.
+%This is to prevent unnecessary increase in the number of input gates when loop counters are only used as plaintext.
+\ana{shorten} %\ben{Ishaq or Vassilis: please check over this paragraph...
+I don't want to make some mistake about MOTION's capabilities.}
 
 % Tons of copies != runtime costs
-Due to the SSA translation phase as well as the conversions to and from SIMD values which our utility functions perform, our generated vectorized MOTION code often includes multiple copies of arrays and scalar values.
+Due to SSA translation and scalar expansion our generated vectorized MOTION code often includes multiple copies of arrays (typically expanded scalar values).
 These copies do not incur a runtime cost as the arrays simply hold \emph{pointers} to the underlying shares, so no new shares or gates are created as a result of this copying.
-Cost in MPC programs is dominated by shares and computation on shares.
+Cost in MPC is dominated by shares and computation on shares.
 %\ben{Is there anything else to say here?

### 2022-05-02 overleaf/compiler.tex
Fixed figure in Sec. 4.
Changes to Sec. 4.
Shortening sections. Still way too long...

@@
 
-\ana{Convention: use \texttt{shared, plain}, {\sf code}, $i$ and $I$ for counters outside of code.}
+%\ana{Convention: use \texttt{shared, plain}, {\sf code}, $i$ and $I$ for counters outside of code.}
 
@@
   \includegraphics[width=0.9\linewidth]{figs_paper_SIMD/compiler_framework.png}
-  \caption{Compiler Framework. \ana{Change figure since we are not doing div-and-conquer.}}
+  \caption{Compiler Framework.}
   \label{fig:compiler_framework}
@@
 
-Source syntax is essentially standard IMP syntax as shown below:
+Source syntax is essentially standard IMP syntax but with for-loops:
 %\begin{figure}[t]
@@
   \begin{array}{l@{~}l@{~~~}l}
- e & ::= e \; \mathit{op} \; e \mid e \mid \x \mid  \texttt{const} \mid \texttt{A[$e$]} & \mathit{expression} \\
+ e & ::= e \; \mathit{op} \; e \mid \x \mid  \texttt{const} \mid \texttt{A[$e$]} & \mathit{expression} \\
  % s &::= t \;\f & \mathit{field} \\
@@
 under secure computation protocols; the \texttt{plain} qualifier denotes plaintext values.
-Shared lists denote shared elements, but plaintext length. Subtyping is
-$\texttt{plain} <: \texttt{shared}$, meaning that we can convert a plaintext value into a shared one,
+Subtyping is $\texttt{plain} <: \texttt{shared}$, meaning that we can convert a plaintext value into a shared one,
 but not vice versa. Subtyping on qualified types is again as expected, it is covariant in the qualifier
@@
 Write access such as for example {\sf A[$i,j+2$] = ...} is not allowed.
+(This is again a restriction we imposed for convenience in our current implementation; we plan to extend the compiler with arbitrary write access.)
 %\item \ana{Add rules for logical ops}
@@
 \end{enumerate}
-Importantly, the arguments of MUX are restricted to base types, as passing array arguments
-is not valid MOTION MPC. \ana{Check this.} This causes a slight inconvenience as we could not write
+The arguments of MUX are restricted to base types. 
+(Again, this is just a restriction of our current implementation that will be lifted.)
+This causes an inconvenience as we could not write
 \begin{pythonn}
@@
 annotations at the level of IMP Source (as in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(a)), the
-annotations propagate through the transformations and are checked at the 
-level of MPC Source. Recall that the programmer annotates only inputs
+annotations propagate through the transformations; annotations are inferred and checked with a taint analysis at the 
+level of MPC Source. Tthe programmer annotates only inputs.
 
@@
 
-Our compiler translates from Source to SSA as follows:
+Our compiler translates from Source to SSA as outlined below. Full details are in the extended version~\cite{Anon_TR}.
 
@@
         Ensure that the AST matches the restricted subset defined in Section~\secref{sec:syntax}.
-        This step outputs an instance of the \texttt{restricted\_ast.Function} class, which represents our restricted subset of the Python AST.
+        %This step outputs an instance of the \texttt{restricted\_ast.Function} class, which represents our restricted subset of the Python AST.
 %  \item
@@
 
-        Convert the restricted-syntax AST to a three-address control-flow graph.
+        Convert the restricted-syntax AST to a three-address control-flow graph. The step processes for-loops, if-statements and assignments as restricted by the syntax.
+\begin{comment}
         To do this, first,
@@
         \end{enumerate}
+\end{comment}        
   %\item
-        \paragraph{SSA conversion}:
-        Convert the 3-address CFG to SSA with Cytron's algorithm.
+        \paragraph{SSA conversion:}
+        Convert 3-address CFG to SSA with Cytron's algorithm.
 %\end{enumerate}

### 2022-05-02 overleaf/compiler.tex
More edits.

@@
 We describe analysis on MPC Source and backend-independent vectorization in~\secref{sec:vectorization}
-and translation into MOTION code in~\secref{sec:backend}.
+and translation into MOTION code in~\secref{sec:backend}. Full details on all parts appear in~\cite{Anon_TR}.
 %\ana{Need to state this more precisely.}
@@
 %\item \ana{Add rules for arithmetic ops}
+
+\begin{comment}
 \item The final restriction involves MUX as expressed by the rule:
@@
 \end{semantics}
+\end{comment}
+
 \end{enumerate}
+
+\begin{comment}
 The arguments of MUX are restricted to base types. 
@@
 
+%\end{comment}
+
 We note that while the programmer writes
@@
 annotations propagate through the transformations; annotations are inferred and checked with a taint analysis at the 
-level of MPC Source. Tthe programmer annotates only inputs.
+level of MPC Source. The programmer annotates only inputs.
+\end{comment}
 
@@
 
-Our compiler translates from Source to SSA as outlined below. Full details are in the extended version~\cite{Anon_TR}.
+Our compiler translates from Source to SSA as sketched below. %Full details are in the extended version~\cite{Anon_TR}.
 
@@
         \paragraph{SSA conversion:}
-        Convert 3-address CFG to SSA with Cytron's algorithm.
+        Convert 3-address CFG to SSA with Cytron's algorithm~\cite{Cytron:1991}.
 %\end{enumerate}
@@
 the compiler recursively converts all successor branches to MPC source and
-appends these to the main MPC source.
+appends to the main MPC source.
 

### 2022-05-02 overleaf/results.tex
More edits.

@@
 The connection between these machines had 10Gbps bandwidth and sub-millisecond latency.
-This setting reflects typical LAN use case considering that 10Gbps LAN is increasingly common in business networks and is now available even in some home networks.
+%This setting reflects typical LAN use case considering that 10Gbps LAN is increasingly common in business networks and is now available even in some home networks.
 In the WAN setting we again used a {\tt c6525-25g} machine (located in Utah, US) for the first party and a {\tt c220g1} machine (located in Wisconsin, US) for the second.
@@
 We measured the connection bandwidth between these machines to be 560Mbps and average round trip time (RTT) to be 38ms.
-At the time of this writing, all major internet providers in the US offer 1Gbps connections to home consumers, therefore this setting reasonably reflects the typical WAN use case.
+%At the time of this writing, all major internet providers in the US offer 1Gbps connections to home consumers, therefore this setting reasonably reflects the typical WAN use case.
 
 We run all experiments 5 times and report average values of various metrics.
-Note that standard deviation --- shown as error bar on top of the bars in the graphs --- in all observations was at most 4.5\% of the mean value, therefore the accuracy of the results is not effected by the relatively fewer runs.
+Note that standard deviation --- shown as error bar on top of the bars in the graphs --- in all observations was at most 4.5\% of the mean value.
+%,  therefore the accuracy of the results is not effected by the relatively fewer runs.
 
@@
 In a preliminary experiment we ran {\tt N=2, N=4}, all the way to {\tt N=4096}; the non-vectorized version ran out of memory at the value {\em both} and we fixed this value for these experiments.
-The vectorized one completed for {\tt N=4096} for nearly all benchmarks (numbers are not shown for space reasons; however, they times are largely consistent, e.g., if {\em both} is $N = 2^k$ and it completes in $X$ seconds for vectorized, then $N = 2^{12}$ completes in $(12-k)X$ seconds).
+The vectorized one completed for {\tt N=4096} for nearly all benchmarks (numbers are not shown for space reasons).
+% however, they times are largely consistent, e.g., if {\em both} is $N = 2^k$ and it completes in $X$ seconds for vectorized, then $N = 2^{12}$ completes in $(12-k)X$ seconds).
 We used the following benchmarks in our evaluation:
@@
 
+\begin{comment}
 \begin{enumerate}
@@
 \end{enumerate}
+\end{comment}
 
@@
 We show circuit evaluation times in \cref{fig:graph_all_eval_time}.
-In terms of amenability to vectorization, we divide benchmarks into 3 categories: 1) {\it High:} these include convex hull, cryptonets max pooling, minimal points and private set intersection.
-These benchmarks are highly parallelizable and see 25x to 70x speedup in BMR, and 30x to 55x in GMW protocol.
-2) {\it Medium:} these include biometric matching, DB Variance, histogram, inner product, k-means iteration and MNIST ReLU.
+In terms of observed benefits from vectorization, we divide benchmarks into 3 categories: 1) {\it High:} these include Convex Hull, Cryptonets Max Pooling, Minimal Points and Private Set Intersection.
+%These benchmarks are highly parallelizable and see 25x to 70x speedup in BMR, and 30x to 55x in GMW protocol. 
+2) {\it Medium:} these include Biometric Matching, DB Variance, Histogram, Inner Product, k-means Iteration and MNIST ReLU.
 These benchmarks have non-parallelizable phases e.g.
-the summing phase of inner product and biometric matching.
+the summing phase of Inner Product.
 Still, most computation is parallelizable and it results in speedup from 5x to 25x in BMR, and 2x to 25x in GMW protocol.
-3) {\it Low:} these include the Database Join and the regular expression benchmarks (count 102, count 10, longest 102 and max distance between symbols).
+3) {\it Low:} these include the Database Join and the regular expression benchmarks (Count 102, Count 10, Longest 102 and Max distance between symbols).
 There is less parallelizable computation in these programs, thus the speedup is lower.
@@
 In GMW, DB Join, Count 102 and Count 10s see speedup from 1.1x to 1.3x.
-However, longest102 and max distance between symbols suffer a slowdown of 0.5x.
+However, Longest 102 and Max distance between symbols suffer a slowdown of 0.5x.
 There is opportunity for vectorization in these benchmarks according to our analytical model, particularly, a there is a large EQ that is vectorized, although a large portion of the loop cannot be vectorized.
@@
 Note that in some settings it may still be desirable to vectorize e.g.
-in data constrained environments as communication is reduced for all benchmarks.
-(The report presents plots of communication, circuit generation time, number of gates, and separate online and setup time~\cite{Anon_TR}.)%As shown in \cref{fig:graph_comm_size}, vectorization results in reduced communication (fewer bits are transferred).
+in data constrained environments as communication as well as all other metrics, is reduced for all benchmarks.
+(The report presents plots of for all metrics~\cite{Anon_TR}.)%As shown in \cref{fig:graph_comm_size}, vectorization results in reduced communication (fewer bits are transferred).
 
@@
 
+\begin{comment}
 \begin{figure}[htbp]
@@
 \end{figure}
+\end{comment}
 
-We look closely at the Biometric Matching benchmark: circuit evaluation in~\cref{fig:graph_biometric_eval_time}, communication size in \cref{fig:graph_biometic_comm_size} and circuit generation time in~\cref{fig:graph_biometic_circ_gen_time}.
+We look closely at the Biometric Matching benchmark: circuit evaluation in~\cref{fig:graph_biometric_eval_time}, communication size in~\cite{Anon_TR} and circuit generation time in~\cite{Anon_TR}.
+%\cref{fig:graph_biometic_comm_size} and circuit generation time in~\cref{fig:graph_biometic_circ_gen_time}. 
 For input size beyond {\tt N=128} the memory usage exceeds available memory and prevents circuit generation.
@@
 \resizebox{0.7\columnwidth}{!}{\input{graphs/comparison-hist-OnlineSetupTimesec}}
-\caption{LAN vs. WAN: Circuit Evaluation Time Comparison}
+\caption{LAN vs. WAN: Circuit Evaluation Time}
 \label{fig:graph_comparison_eval_time}
@@
 
-\subsection{Comparison with MOTION-native Inner Product}
+\subsection{Comparison with MOTION-native Inner Product and Discussion}
 

### 2022-05-02 overleaf/results.tex
More cuts... Still 1/2 page over.

@@
 \subsection{Benchmarks}\label{sec:benchmarks_description}
+
+We used benchmarks from ABY and OPA~\cite{NDSS:DemSchZoh15,Ishaq:2019} (Biometric Matching, Histogram, PSI), HyCC~\cite{CCS:BDKKS18} (Database Join, Database Variance, Inner Product, Cryptonets Max Pooling, MNIST ReLU, k-means Iteration) and Parsynth~\cite{Farzan:2021} (Convex Hull, Minimal Points, Count 102, Count 10, Longest 102, Max distance between symbols).
+We did not use Gcd and Modular Exponentiation from~\cite{Ishaq:2019} as they do not benefit from vectorization according to our model. 
+We ran as many routines from~\cite{CCS:BDKKS18} and~\cite{Farzan:2021} as possible. Gaussian decomposition cannot be expressed 
+in our syntax at the moment and the compiler errors on Cryptonets Convolution at the time of writing, although we expect to fix the issue soon. 
+The extended version describes what the benchmarks do in detail~\cite{Anon_TR}.
+
 In the following, we say {\em both} for an experiment in which we execute both non-vectorized and vectorized protocols and {\em vec} for the vectorized only experiment. 
-In a preliminary experiment we ran {\tt N=2, N=4}, all the way to {\tt N=4096}; the non-vectorized version ran out of memory at the value {\em both} and we fixed this value for these experiments.
+In a preliminary experiment we ran {\tt N=2, N=4}, all the way to {\tt N=4096}; the non-vectorized version ran out of memory at the value for {\em both} and we fixed this value for these experiments.
 The vectorized one completed for {\tt N=4096} for nearly all benchmarks (numbers are not shown for space reasons).
 % however, they times are largely consistent, e.g., if {\em both} is $N = 2^k$ and it completes in $X$ seconds for vectorized, then $N = 2^{12}$ completes in $(12-k)X$ seconds).
-We used the following benchmarks in our evaluation:
 

### 2022-05-02 overleaf/compiler.tex
Cut more.

@@
 \figref{fig:compiler_framework} presents an overview of our compiler.
-We start the section with a brief description of the top-level syntax and semantic restrictions in~\secref{sec:syntax}.
-We proceed to describe the front end of the compiler:~\secref{sec:imp_to_ssa}
-describes translation from IMP Source to SSA, and~\secref{sec:ssa_to_MPC_Source}
-describes translation form SSA into MPC Source.
-
-We describe analysis on MPC Source and backend-independent vectorization in~\secref{sec:vectorization}
-and translation into MOTION code in~\secref{sec:backend}. Full details on all parts appear in~\cite{Anon_TR}.
+%We start the section with a brief description of the top-level syntax and semantic restrictions in~\secref{sec:syntax}.
+%We proceed to describe the front end of the compiler:~\secref{sec:imp_to_ssa}
+%describes translation from IMP Source to SSA, and~\secref{sec:ssa_to_MPC_Source}
+%describes translation form SSA into MPC Source.
+This section outlines the front end, the following section describes analysis on MPC Source and backend vectorization 
+and~\secref{sec:backend} outlines the back end. 
+%We describe analysis on MPC Source and backend-independent vectorization in~\secref{sec:vectorization}
+%and translation into MOTION code in~\secref{sec:backend}. 
+Details on all parts appear in~\cite{Anon_TR}.
 %\ana{Need to state this more precisely.}
@@
 %  \item
+        \squeeze
         \paragraph{Parsing:}
@@
 %  \item
+        \squeeze
         \paragraph{Syntax checking:}
@@
 %  \item
+        \squeeze
         \paragraph{3-address CFG conversion:}
         %\benjamin{TODO: Is this a good amount of detail?} \ana{This is good amount of detail for the Tech report. In the actual paper we'll shorten a lot and add pointers to the Tech report.}
-
-        Convert the restricted-syntax AST to a three-address control-flow graph. The step processes for-loops, if-statements and assignments as restricted by the syntax.
+        Convert the restricted-syntax AST to a three-address control-flow graph (CFG). The step processes for-loops, if-statements and assignments as restricted by the syntax.
 \begin{comment}
@@
   %\item
+        \squeeze
         \paragraph{SSA conversion:}
@@
 $\phi$-nodes corresponding to if-statements will be in a basic block
-with the merge condition property.
-For example, if {\sf X!3 = $\phi$(X!1,X!2)} is in a block with merge condition {\sf C},
+with the ``merge condition'' property.
+E.g., if {\sf X!3 = $\phi$(X!1,X!2)} is in a block with merge condition {\sf C},
 the compiler transforms it into {\sf X!3 = MUX(C, X!1, X!2)}.
-Next, the compiler runs the dead code elimination algorithm from Cytron's SSA paper.
+Next, the compiler runs the dead code elimination algorithm from~\cite{Cytron:1991}. %Cytron's SSA paper.
 
@@
 The compiler linearizes the control-flow graph with a variation of breadth-first search.
+%\begin{comment}
 Blocks with the ``merge condition'' property are only considered
@@
 appends to the main MPC source.
+%\end{comment}
 

### 2022-05-02 overleaf/vectorization.tex
Cut more.

@@
 In this section, we briefly describe the analysis. Details, including edge cases and examples, can be found in the supplementary extended version~\cite{Anon_TR}.
-\secref{sec:dependence} describes our dependence analysis and ~\secref{sec:expansion} describes scalar expansion,
-which lifts scalars (and arrays) to the corresponding loop dimensionality to create opportunities for vectorization. 
-\secref{sec:basic_vectorization} describes our core vectorization algorithm and~\secref{sec:correctness} argues correctness of the vectorization transformation.
-\secref{sec:array_writes} extends vectorization with array writes.
+%\secref{sec:dependence} describes our dependence analysis and ~\secref{sec:expansion} describes scalar expansion,
+%which lifts scalars (and arrays) to the corresponding loop dimensionality to create opportunities for vectorization. 
+%\secref{sec:basic_vectorization} describes our core vectorization algorithm and~\secref{sec:correctness} argues correctness of the vectorization transformation.
+%\secref{sec:array_writes} extends vectorization with array writes.
 
@@
 on arrays of the corresponding dimensionality and we optimistically vectorize all arrays. For example, 
-lines 4 and 5 in Biometric~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(c) reshape the input arrays
+lines 4,5 in Biometric~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(c) reshape the input arrays.
 
@@
 
-In the Histogram example, this removes the back edge from the update at 8 to the PHI node at 2. Clearly there does not exist $i < i'$ that will make 
+In Histogram, this removes the back edge from the update at 8 to the PHI node at 2. Clearly there does not exist $i < i'$ that makes
 $i$ at 9 and $i'$ at 6 and 7 equal. The back-edge from 9 to 4 stays because for every $j < j'$ 9 and 6 and 7 access the same location in {\sf res}. %\ana{Can't wrap my head around this...}
@@
 In a vectorized update statement, we can ignore the incoming array, {\sf res2} in this case.
-The update writes (in parallel) all locations of the 2-dimensional array, in this case it sets up each {\sf res3[i,j] = tmp3[i,j]}.
+The update writes (in parallel) all locations of the 2-d array, in this case it sets up each {\sf res3[i,j] = tmp3[i,j]}.
 

### 2022-05-02 overleaf/related.tex
Cut more.
More cuts.

@@
 \paragraph{MPC languages and compilers.}
-
 Languages and compilers for secure computation have seen significant attention and advances in recent years. The early MPC compilers Fairplay~\cite{CCS:BenNisPin08},
@@
 %is the separation of the higher-level ``backend-independent'' compilation and optimization and the lower-level compilation into circuits.
-
+\squeeze
 \paragraph{Classical HPC compilers.}
@@
 The problem is that masking and selection, or more generally, handling control predicates~\cite{Benabderrahmane:2010,Karrenberg:2015},
-can lead to \emph{slowdown}.
+can lead to slowdown.
 %As a result, vectorizing compilers and polyhedral compilers tend to focus on straight-line code due to the unpredictable
@@
 
-We argue that vectorization over linear MPC Source is a different problem, one that warrants a new look, while drawing from
+We believe that vectorization over linear MPC Source warrants a new look, while drawing from
 results in HPC.

### 2022-05-02 overleaf/paper_SIMD.tex
Cut more.
Finally

@@
 \usepackage[capitalise]{cleveref}
+
+\newcommand{\squeeze}{\vspace{-0.2cm}}
 
@@
 We demonstrated that vectorization has significant impact on performance.
-We are excited about the opportunities for future work --- integration with protocol mixing, divide-and-conquer reasoning and parallelization, as well as inter-procedural context-sensitive analysis at the level of MPC Source will improve MPC programmability and efficiency.
+We are excited about many opportunities for future work --- integration with protocol mixing, divide-and-conquer reasoning and parallelization, as well as inter-procedural context-sensitive analysis for  MPC Source.
+%will improve MPC programmability and efficiency.
 

### 2022-05-02 overleaf/related.tex
Finally

@@
 \paragraph{MPC languages and compilers.}
-Languages and compilers for secure computation have seen significant attention and advances in recent years. The early MPC compilers Fairplay~\cite{CCS:BenNisPin08},
+Languages and compilers for secure computation have seen significant advances in recent years. The early MPC compilers Fairplay~\cite{CCS:BenNisPin08},
 and Sharemind~\cite{ESORICS:BogLauWil08} were followed by PICCO~\cite{CCS:ZhaSteBla13}, Obliv-C~\cite{Zahur:2015}, TinyGarble~\cite{SP:SHSSK15},
@@
 frameworks, including MPC and ZKP. Hastings et el. present a review of compiler frameworks~\cite{Hastings:2019}. We believe that our work is unique as 
-as a facade between the high-level language and lower-level circuit protocols.
+as it falls between the high-level language and lower-level circuit protocols.
 
@@
 do not work well with conditionals~\cite{Benabderrahmane:2010,Karrenberg:2015}. In contrast, we consider vectorization at the level of MPC Source which 
-linearizes conditionals; we are able to handle programs with interleaved if-statements and for-loops and achieve significant speedup.
+linearizes conditionals; we are able to handle programs with interleaved if-statements and for-loops. %and achieve significant speedup.
 
@@
 Automatic vectorization is a longstanding problem in high-performance computing and
-there are thousands of works in this area reflecting over 40 years of research. We presented a vectorization
+there are thousands of works reflecting many years of research. We presented a vectorization
 algorithm for MPC Source, essentially extending classical loop vectorization \cite{Allen:1987}. In HPC vectorization, conditional control flow
@@
 We view Karrenberg's work on Whole function vectorization~\cite{Karrenberg:2015} as most closely related to ours --- it linearizes the program and vectorizes
-both branches of a conditional applying masking to avoid execution of the branch-not-taken code, and selection (similar to MUX) to select the correct value based on the
-result of the condition at runtime.
-The problem is that masking and selection, or more generally, handling control predicates~\cite{Benabderrahmane:2010,Karrenberg:2015},
-can lead to slowdown.
+both branches of a conditional applying masking to avoid execution of the branch-not-taken code as well as selection (similar to MUX). 
+%to select the correct value based on the
+%result of the condition at runtime.
+%The problem is that masking and selection, or more generally, handling control predicates~\cite{Benabderrahmane:2010,Karrenberg:2015},
+%can lead to slowdown.
 %As a result, vectorizing compilers and polyhedral compilers tend to focus on straight-line code due to the unpredictable
@@
 representation, while our work takes advantage of linear MPC Source and SSA form. The work by Karrenberg~\cite{Karrenberg:2015} is rare in that space, in the sense that it considers
-vectorization over SSA form, which has similarities to MPC Source. We consider different array representation, notion of dependence, and reasoning
-about dependence, which we conjecture is more suitable for MPC Source. 
+vectorization over SSA form. %, which has similarities to MPC Source. 
+We consider different array representation, notion of dependence, and reasoning
+about dependence, which is more suitable for MPC Source. 
 %Buscher~\cite{Buscher:2018} considers SIMD-vectorization

### 2022-05-02 overleaf/results.tex
More cuts.
Finally

@@
 The {\tt c220g1} machine is equipped with two Intel E5-2630 8-core 2.40GHz processors and 128GB of RAM.
-We measured the connection bandwidth between these machines to be 560Mbps and average round trip time (RTT) to be 38ms.
+We measured the connection bandwidth to be 560Mbps and average round trip time (RTT) to be 38ms.
 %At the time of this writing, all major internet providers in the US offer 1Gbps connections to home consumers, therefore this setting reasonably reflects the typical WAN use case.
@@
 We run all experiments 5 times and report average values of various metrics.
-Note that standard deviation --- shown as error bar on top of the bars in the graphs --- in all observations was at most 4.5\% of the mean value.
+Note that standard deviation in all observations was at most 4.5\% of the mean value.
 %,  therefore the accuracy of the results is not effected by the relatively fewer runs.
@@
 We ran as many routines from~\cite{CCS:BDKKS18} and~\cite{Farzan:2021} as possible. Gaussian decomposition cannot be expressed 
-in our syntax at the moment and the compiler errors on Cryptonets Convolution at the time of writing, although we expect to fix the issue soon. 
-The extended version describes what the benchmarks do in detail~\cite{Anon_TR}.
+in our syntax and the compiler errors on Cryptonets Convolution at the time of writing, although we expect to fix the issue soon. 
+\cite{Anon_TR} describes what the benchmarks do.
 
-In the following, we say {\em both} for an experiment in which we execute both non-vectorized and vectorized protocols and {\em vec} for the vectorized only experiment. 
-In a preliminary experiment we ran {\tt N=2, N=4}, all the way to {\tt N=4096}; the non-vectorized version ran out of memory at the value for {\em both} and we fixed this value for these experiments.
+To compare we run {\em both} versions with the maximal value of $N$ for which the non-vectorized version succeeds (typically {\tt N=128}) and we run {\em vec} with typically {\tt N=4096}.
+Again, due to space constraints \cite{Anon_TR} gives the exact values.
+% In the following, we say {\em both} for an experiment in which we execute both non-vectorized and vectorized protocols and {\em vec} for the vectorized only experiment. 
+%In a preliminary experiment we ran {\tt N=2, N=4}, all the way to {\tt N=4096}; the non-vectorized version ran out of memory at the value for {\em both} and we fixed this value for these experiments.
 The vectorized one completed for {\tt N=4096} for nearly all benchmarks (numbers are not shown for space reasons).
@@
 \resizebox{0.7\columnwidth}{!}{\input{graphs/lan/biometric-hist-OnlineSetupTimesec}}
-\caption{Biometric Matching Circuit Evaluation Time, x-axis lists database size}
+\caption{Biometric Matching Circuit Evaluation Time}
 \label{fig:graph_biometric_eval_time}
@@
 however, our addition loop incurred significant cost (ADD is non-local and expensive in Boolean GMW). The MOTION-native loop ran 
-{\sf result += mult\_unsimdified[i];} while our loop ran auto generated {\sf result[i] = result\_prev[i] + mult\_unsimdified[i];} (recall that scalar expansion
-is an artifact of our vectorization). We rewrote the accumulation (manually, for testing purposes) and that led to the comparable speedup!
+{\sf result += mult\_unsimdified[i];} while our loop ran auto generated {\sf result[i] = result\_prev[i] + mult\_unsimdified[i];} (as scalar expansion
+is an artifact of our vectorization). We rewrote the accumulation manually and that led to the comparable speedup!
 
 MOTION's compiler performs analysis that informs circuit generation and the example illustrates the power of the analysis. In the above example, 
-MOTION does the standard divide-and-conquer accumulation. We conjectured that poor performance of our loop was due to a
-limitation of our current implementation. Recall that the Phase 3 of Vectorization, in~\secref{sec:vectorization}, which we have not implemented yet, gets rid of redundant 
-dimensions and generates {\sf if (i!=0) \{ result\_prev = result; \} result = result\_prev + mult\_unsimdified[i]}. And while it is unrealistic to expect that MOTION's static analysis 
+MOTION does the standard divide-and-conquer accumulation. We conjectured that poor performance of our loop was due to
+limitations of our implementation. Recall that Phase 3 of Vectorization, in~\secref{sec:vectorization}, which we have not implemented yet, gets rid of redundant 
+dimensions and generates {\sf if (i!=0) \{ result\_prev = result; \} result = result\_prev + mult\_unsimdified[i]}. And while it is unrealistic to expect that static analysis 
 will detect the associative accumulation in the scalar expansion code, one might expect that it will in the above slightly more complex code. 
-Our experience indicates the limitation of AST-based analysis; because AST analysis is difficult it does not appear to detect the associative accumulation.
+However, because AST analysis is difficult it does not appear to detect the accumulation.
 %(unfortunately, we ran out of time and did not experiment further).
@@
 
-The most closely related work, HyCC~\cite{CCS:BDKKS18}, is a mixing compiler that focuses on mixed protocols, while we run vectorization within a single protocol. 
-The paper only provides data for Boolean and Yao for a version of Biometric matching on $N=1000$; it does provide a lot of data on mixed protocol executions.
-Again, we estimate we are about an order of magnitude slower, which is likely due to the same issue as Inner Product --- the computation of $\mathit{min}$ can be optimized.
+HyCC~\cite{CCS:BDKKS18} is a mixing compiler that focuses on mixed protocols, while we run vectorization within a single protocol. 
+The paper only provides data for Boolean and Yao for a version of Biometric matching on $N=1000$ as its focus is mixing. %; it does provide a lot of data on mixed protocol executions.
+We estimate we are about an order of magnitude slower, which is likely due to the same issue --- the computation of $\mathit{min}$ can be optimized.
 

### 2022-05-02 overleaf/vectorization.tex
Finally

@@
 
-Phases 3 cleans up local arrays of references (this is an optional phase and our current implementation does not include it)
-and Phase 4 explicitly turns operations into MOTION SIMD operations.
+Phases 3 cleans up local arrays of references (this is an optional phase and our current implementation does not include it).
+%and Phase 4 explicitly turns operations into MOTION SIMD operations.
 
@@
 
-\COMMENT { Phase 4: }
-\STATE {add SIMD for simdified dimensions}
+%\COMMENT { Phase 4: }
+%\STATE {add SIMD for simdified dimensions}
 
@@
 
+\squeeze
 \paragraph{MPC Source Syntax}
@@
 for i in range(I):
-   ...
    for k in range(K):
@@
 
+\squeeze
 \paragraph{Partial Orders}
@@
 
+\squeeze 
 \paragraph{Theorems} The two theorems arising from the Basic vectorization optimization are as follows:

### 2022-05-03 overleaf/introduction.tex
Go over 1-3 to fix typos.

@@
 
-\subsection{Our Contribution} In this paper, we develop a compiler framework that takes a Python-like routine and produces MOTION code: we describe (a) the IMP Source language, its syntax and semantic restrictions, (b) translation into MPC Source, (c) a specific backend-independent optimization: novel SIMD-vectorization on MPC Source, and (d) translation from MPC Source into MOTION code.
+\subsection{Our Contribution} In this paper, we develop a compiler framework that takes a Python-like routine and produces MOTION code.
+We describe: (a) the IMP Source language, its syntax and semantic restrictions, (b) translation into MPC Source, (c) a specific backend-independent optimization: novel SIMD-vectorization on MPC Source, and (d) translation from MPC Source into MOTION code.
 %\ana{I think we have to add more on MOTION, specifically mixing and amortization.}
@@
 protocol mixing, supports Arithmetic GMW, however, it does not implement Comparison (CMP) and Multiplexing (MUX) as they would be rather inefficient.)
-In our experiments vectorized code exhibits 24x improvement on average in circuit generation time for Boolean GMW (20x for BMR), 7x reduction 
+In our LAN experiments vectorized code exhibits 24x improvement on average in circuit generation time for Boolean GMW (20x for BMR), 7x reduction 
 in communication (2x), 97x reduction in number of gates (91x), 4x improvement in setup time (23x) and 21x improvement in online time (18x).
@@
 All our code, including benchmark Python-like code, compilation phases, and generated MOTION code is available on Github. We omit the link for anonymity, 
-however, we will gladly make it available upon request from reviewers. The Github setup generates graphs and intermediate code for each program along each compiler phase, compiles with 
+however, we will gladly make it available upon request from reviewers. The Github setup generates graphs and intermediate code for each benchmark along each compiler phase, it compiles with 
 MOTION and runs the circuits on small inputs to generate tables of data (the experiments we present later run on real LAN and WAN).

### 2022-05-03 overleaf/model.tex
Go over 1-3 to fix typos.

@@
 For the purpose of the model we assume that $s$ is large enough and the term $\frac{n}{s}p\alpha$ amounts to a \emph{fixed cost} incurred regardless of
-whether $n$ is $10,000$ or just $1$. (This models the cost of preparing and sending a packet from party A to party B for example.) Therefore, amortized execution
+whether $n$ is $10,000$ or just $1$. (This models the cost of preparing and sending a packet from party 0 to party 1 for example.) Therefore, amortized execution
 of $n$ operations is $f(n) = \alpha_\mathit{fix} + n\alpha_{var}$ in contrast to unamortized execution $g(n) = n\alpha_\mathit{fix} + n\alpha_{var}$.

### 2022-05-03 overleaf/compiler.tex
More cuts.
Finally
Proofread 4.

@@
 %describes translation form SSA into MPC Source.
-This section outlines the front end, the following section describes analysis on MPC Source and backend vectorization 
+This section outlines the front end, \secref{sec:vectorization} describes analysis on MPC Source and backend-independent vectorization 
 and~\secref{sec:backend} outlines the back end. 
@@
 \item Loops are of the form $0 <= i < I$ and bounds are fixed at compile time.
-It is a standard restriction in MPC that the bounds must be known at circuit-generation time.
+It is a standard restriction in MPC that computation is bounded~\cite{NDSS:DemSchZoh15,CCS:BDKKS18}.
 \item Arrays are one-dimensional. $N$-dimensional arrays are linearized and accessed
@@
   %\item
+        \squeeze
         \squeeze

### 2022-05-03 overleaf/vectorization.tex
Proofread 4.

@@
 structure and in part because vectorization meshes in with other MPC-specific optimizations in non-trivial ways
-(other works have explored manual vectorization and protocol mixing in an ad-hoc way, e.g.,~\cite{NDSS:DemSchZoh15,CCS:BDKKS18,Ishaq:2019}).
+(other works have explored manual vectorization and protocol mixing in ad-hoc ways, e.g.,~\cite{NDSS:DemSchZoh15,CCS:BDKKS18,Ishaq:2019}).
 
-In this section, we briefly describe the analysis. Details, including edge cases and examples, can be found in the supplementary extended version~\cite{Anon_TR}.
+In this section, we briefly describe the analysis. Details, including edge cases and examples, can be found in the extended version~\cite{Anon_TR}.
 %\secref{sec:dependence} describes our dependence analysis and ~\secref{sec:expansion} describes scalar expansion,

### 2022-05-03 overleaf/vectorization.tex
Made a pass.

@@
 %A $\phi$-node can be a source or a target of an outer-to-inner forward edge.
-\item inner-to-outer $X\rightarrow Y$ where $X$ is a $phi$-node in an inner loop nest, $i,j,k$, and $Y$ is in the enclosing loop nest $i,j$.
-E.g., the def-use from 8 to 12 gives rise to an inner-to-outer edge.
+\item inner-to-outer $X\rightarrow Y$ where $X$ is in an inner loop nest, $i,j,k$, and $Y$ is in the enclosing loop nest $i,j$. E.g., the def-use from 8 to 12 gives rise to an inner-to-outer edge.
 %E.g. $\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$ is an inner-to-outer edge.
@@
 operations according to def-uses. E.g., if there is an inner-to-outer dependence, it inserts $\mathit{raise\_dim}$,
-and similarly, if there is an outer-to-inner dependence, it inserts $\mathit{drop\_dim}$. After this phase operations work
-on arrays of the corresponding dimensionality and we optimistically vectorize all arrays. For example, 
-lines 4,5 in Biometric~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(c) reshape the input arrays.
+and similarly, if there is an outer-to-inner dependence, it inserts $\mathit{drop\_dim}$. 
+For example, lines 4,5 in Biometric~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(c) reshape the input arrays. 
+After this phase, operations work on arrays of the corresponding dimensionality and we optimistically vectorize all arrays.
 
@@
 
-Phases 3 cleans up local arrays of references (this is an optional phase and our current implementation does not include it).
+Phases 3 reverses scalar expansion (this is an optional phase and our current implementation does not include it).
 %and Phase 4 explicitly turns operations into MOTION SIMD operations.
@@
 in terms of def-use relations. We define a partial order over elements of $C$ as well,
-in terms of def-use relations in the concrete domain $C$. We state two theorems
-that state (informally) that the schedule corresponding to the original program computes the same
-result as the schedule corresponding to the vectorized program.
+in terms of def-use relations in the concrete domain $C$. The theorems state 
+(informally) that the vectorized program preserves def-uses and thus computes the same 
+result as the original program.
 

### 2022-05-03 overleaf/vectorization.tex
Done with backend.

@@
 for i in range(0, num_bins):
-  res1 = PHI(res, res2)
+  res1 = PHI(res, res2) # res is the result array
   for j in range(0, N):
@@
 
-There is loop-carried dependence from 8 to 2 for write array {\sf res} even though the loop can be vectorized across $i$.
+There is a def-use edge to 2 due to the update at 8 of array {\sf res} even though the loop can be vectorized across $i$.
 
@@
 loop-carried dependence on ${\sf A}$; otherwise there is not and we can remove the spurious backward edge to the array PHI-node 
-thus ``freeing up'' statements for vectorization. %\ana{Double check this!!!}
+for loop $j$, thus ``freeing up'' statements for vectorization. %\ana{Double check this!!!}
 
-In Histogram, this removes the back edge from the update at 8 to the PHI node at 2. Clearly there does not exist $i < i'$ that makes
+In Histogram, this removes the back edge from 4 to the PHI node at 2. Clearly there does not exist $i < i'$ that makes
 $i$ at 9 and $i'$ at 6 and 7 equal. The back-edge from 9 to 4 stays because for every $j < j'$ 9 and 6 and 7 access the same location in {\sf res}. %\ana{Can't wrap my head around this...}
@@
 
-Note however, that this step renders some array phi-nodes target-less. We handle target-less phi-nodes with a minor extension of Vectorization (Phase 2).
+Removing infeasible edges renders some array phi-nodes \emph{target-less}. We handle target-less phi-nodes with a minor extension of Vectorization (Phase 2).
 First, we merge closures that update the same array. This simplifies handling of array $\phi$-nodes: if each closure is turned into a separate loop
@@
 array write can be vectorized, we have to properly remove the phi-node replacing uses of the left-hand side of the phi-node with its arguments.
-Additional detail, as well as detailed treatment of the minor changes to $\mathit{raise\_dim}$ and $\mathit{drop\_dim}$ required to do expansion of arrays 
-in addition to scalars, appear in~\cite{Anon_TR}.
+Additional detail on the minor changes to $\mathit{raise\_dim}$ and $\mathit{drop\_dim}$ required to do expansion of arrays appear in~\cite{Anon_TR}.
 
@@
 
-res1^ = raise_dim(res, (j:N)) // replacing res1 with res, 1st arg
+res1^ = raise_dim(res, (j:N)) # replacing res1 with res, 1st arg
 for j in range(0, N):
@@
 res2^ = drop_dim(res2)
-res1 = drop_dim(res2^) // replacing with res2^, 2nd arg. NOOP
+res1 = drop_dim(res2^) # replacing with res2^, 2nd arg. NOOP
 return res1

### 2022-05-03 overleaf/results.tex
Edits to results.

@@
 
-We used benchmarks from ABY and OPA~\cite{NDSS:DemSchZoh15,Ishaq:2019} (Biometric Matching, Histogram, PSI), HyCC~\cite{CCS:BDKKS18} (Database Join, Database Variance, Inner Product, Cryptonets Max Pooling, MNIST ReLU, k-means Iteration) and Parsynth~\cite{Farzan:2021} (Convex Hull, Minimal Points, Count 102, Count 10, Longest 102, Max distance between symbols).
-We did not use Gcd and Modular Exponentiation from~\cite{Ishaq:2019} as they do not benefit from vectorization according to our model. 
+We used benchmarks from ABY and OPA~\cite{NDSS:DemSchZoh15,Ishaq:2019} (Biometric Matching, Histogram, PSI), HyCC~\cite{CCS:BDKKS18} (Database Join, Database Variance, Inner Product, Cryptonets Max Pooling, MNIST ReLU, k-means Iteration) and Parsynt~\cite{Farzan:2021} (Convex Hull, Minimal Points, Count 102, Count 10, Longest 102, Max distance between symbols).
+We did not use Gcd~\cite{Ishaq:2019} and Modular Exponentiation from~\cite{NDSS:DemSchZoh15} as they do not benefit from vectorization according to our model. 
 We ran as many routines from~\cite{CCS:BDKKS18} and~\cite{Farzan:2021} as possible. Gaussian decomposition cannot be expressed 
@@
 
-To compare we run {\em both} versions with the maximal value of $N$ for which the non-vectorized version succeeds (typically {\tt N=128}) and we run {\em vec} with typically {\tt N=4096}.
-Again, due to space constraints \cite{Anon_TR} gives the exact values.
+To compare we run {\em both} versions with the maximal value of $N$ for which the non-vectorized version succeeds (typically {\tt N=128}). We also run {\em vec} with larger values, 
+typically {\tt N=4096}. Again, due to space constraints \cite{Anon_TR} gives the exact values.
 % In the following, we say {\em both} for an experiment in which we execute both non-vectorized and vectorized protocols and {\em vec} for the vectorized only experiment. 
@@
 However, Longest 102 and Max distance between symbols suffer a slowdown of 0.5x.
-There is opportunity for vectorization in these benchmarks according to our analytical model, particularly, a there is a large EQ that is vectorized, although a large portion of the loop cannot be vectorized.
+There is opportunity for vectorization in these benchmarks according to our analytical model, particularly, there is a large EQ that is vectorized, although a large portion of the loop cannot be vectorized.
 We observed that transformation to vectorized code increased multiplicative depth and, the negative effect of increased depth is more noticeable in a round-based protocol like GMW.
-The cause of the increase is not clear --- we conjecture that MOTION performs optimizations over the non-vectorized loop body that decreases depth; also, EQ is relatively inexpensive in Boolean GMW and BMR compared to ADD and MUL, which also de-emphasizes the benefit of vectorization.
-We propose a simple heuristic although we do leave all the benchmarks in the table: if the transformation increases circuit depth beyond some threshold (e.g.
+The cause of the increase is not clear --- we conjecture that MOTION performs optimization over the non-vectorized loop body that decreases depth; also, EQ is relatively inexpensive in Boolean GMW and BMR compared to ADD and MUL, which also de-emphasizes the benefit of vectorization.
+We propose a simple heuristic: if the transformation increases circuit depth beyond some threshold (e.g.
 more than 10\% of the original circuit), we can reject the transformation.
-%Nevertheless, we show these graphs here for the sake of completeness and to highlight that vectorization is not always reduce run time.
+Nevertheless, we show these benchmarks here for the sake of completeness and to highlight that vectorization does not always reduce run time.
 Note that in some settings it may still be desirable to vectorize e.g.
 in data constrained environments as communication as well as all other metrics, is reduced for all benchmarks.
-(The report presents plots of for all metrics~\cite{Anon_TR}.)%As shown in \cref{fig:graph_comm_size}, vectorization results in reduced communication (fewer bits are transferred).
+(The extended version presents plots of for all metrics~\cite{Anon_TR}.)%As shown in \cref{fig:graph_comm_size}, vectorization results in reduced communication (fewer bits are transferred).
 
@@
 MOTION does the standard divide-and-conquer accumulation. We conjectured that poor performance of our loop was due to
-limitations of our implementation. Recall that Phase 3 of Vectorization, in~\secref{sec:vectorization}, which we have not implemented yet, gets rid of redundant 
+limitations of our implementation. Recall that Phase 3 of Vectorization in~\secref{sec:vectorization}, which we have not implemented yet, gets rid of redundant 
 dimensions and generates {\sf if (i!=0) \{ result\_prev = result; \} result = result\_prev + mult\_unsimdified[i]}. And while it is unrealistic to expect that static analysis 
-will detect the associative accumulation in the scalar expansion code, one might expect that it will in the above slightly more complex code. 
-However, because AST analysis is difficult it does not appear to detect the accumulation.
+will detect the associative accumulation in the scalar expansion code, one might expect that it will in the above simpler code. 
+However, because AST analysis is difficult, it does not appear to detect the accumulation.
 %(unfortunately, we ran out of time and did not experiment further).
@@
 HyCC~\cite{CCS:BDKKS18} is a mixing compiler that focuses on mixed protocols, while we run vectorization within a single protocol. 
-The paper only provides data for Boolean and Yao for a version of Biometric matching on $N=1000$ as its focus is mixing. %; it does provide a lot of data on mixed protocol executions.
+The paper only provides data for Boolean and Yao for a version of Biometric matching on $N=1000$ (as its focus is mixing). %; it does provide a lot of data on mixed protocol executions.
 We estimate we are about an order of magnitude slower, which is likely due to the same issue --- the computation of $\mathit{min}$ can be optimized.

### 2022-05-03 overleaf/related.tex
Fixed refs.

@@
 \paragraph{MPC languages and compilers.}
-Languages and compilers for secure computation have seen significant advances in recent years. The early MPC compilers Fairplay~\cite{CCS:BenNisPin08},
-and Sharemind~\cite{ESORICS:BogLauWil08} were followed by PICCO~\cite{CCS:ZhaSteBla13}, Obliv-C~\cite{Zahur:2015}, TinyGarble~\cite{SP:SHSSK15},
-Wystiria~\cite{SP:RasHamHic14}, and others. A new generation of MPC compilers
-includes SPDZ/SCALE-MAMBA/MP-SPDZ~\cite{Keller:2020} and the ABY/HyCC/MOTION~\cite{NDSS:DemSchZoh15,CCS:BDKKS18,Braun:2022} frameworks.
+Languages and compilers for secure computation have seen significant advances in recent years. The early MPC compilers Fairplay~\cite{Ben-David:2008},
+and Sharemind~\cite{Bogdanov:2008} were followed by PICCO~\cite{Zhang:2013}, Obliv-C~\cite{Zahur:2015}, TinyGarble~\cite{Songhori:2015},
+Wystiria~\cite{Rastogi:2014}, and others. A new generation of MPC compilers
+includes SPDZ/SCALE-MAMBA/MP-SPDZ~\cite{Keller:2020} and the ABY/HyCC/MOTION~\cite{Demmler:2015,Buscher:2018b,Braun:2022} frameworks.
 These two families are the state-of-the art and
@@
 
-HyCC~\cite{CCS:BDKKS18} is a compiler from C Source into ABY circuits. It does source-to-source compilation with the key goal to decompose the 
+HyCC~\cite{Buscher:2018b} is a compiler from C Source into ABY circuits. It does source-to-source compilation with the key goal to decompose the 
 program into modules and then assign protocols to modules. In contrast, we focus on MPC Source-level optimizations, specifically vectorization, although we envision 

### 2022-05-03 overleaf/results.tex
Fixed refs.

@@
 
-We used benchmarks from ABY and OPA~\cite{NDSS:DemSchZoh15,Ishaq:2019} (Biometric Matching, Histogram, PSI), HyCC~\cite{CCS:BDKKS18} (Database Join, Database Variance, Inner Product, Cryptonets Max Pooling, MNIST ReLU, k-means Iteration) and Parsynt~\cite{Farzan:2021} (Convex Hull, Minimal Points, Count 102, Count 10, Longest 102, Max distance between symbols).
-We did not use Gcd~\cite{Ishaq:2019} and Modular Exponentiation from~\cite{NDSS:DemSchZoh15} as they do not benefit from vectorization according to our model. 
-We ran as many routines from~\cite{CCS:BDKKS18} and~\cite{Farzan:2021} as possible. Gaussian decomposition cannot be expressed 
+We used benchmarks from ABY and OPA~\cite{Demmler:2015,Ishaq:2019} (Biometric Matching, Histogram, PSI), HyCC~\cite{Buscher:2018b} (Database Join, Database Variance, Inner Product, Cryptonets Max Pooling, MNIST ReLU, k-means Iteration) and Parsynt~\cite{Farzan:2021} (Convex Hull, Minimal Points, Count 102, Count 10, Longest 102, Max distance between symbols).
+We did not use Gcd~\cite{Ishaq:2019} and Modular Exponentiation from~\cite{Demmler:2015} as they do not benefit from vectorization according to our model. 
+We ran as many routines from~\cite{Buscher:2018} and~\cite{Farzan:2021} as possible. Gaussian decomposition cannot be expressed 
 in our syntax and the compiler errors on Cryptonets Convolution at the time of writing, although we expect to fix the issue soon. 
@@
 
-HyCC~\cite{CCS:BDKKS18} is a mixing compiler that focuses on mixed protocols, while we run vectorization within a single protocol. 
+HyCC~\cite{Buscher:2018b} is a mixing compiler that focuses on mixed protocols, while we run vectorization within a single protocol. 
 The paper only provides data for Boolean and Yao for a version of Biometric matching on $N=1000$ (as its focus is mixing). %; it does provide a lot of data on mixed protocol executions.

### 2022-05-03 overleaf/vectorization.tex
Fixed refs.

@@
 structure and in part because vectorization meshes in with other MPC-specific optimizations in non-trivial ways
-(other works have explored manual vectorization and protocol mixing in ad-hoc ways, e.g.,~\cite{NDSS:DemSchZoh15,CCS:BDKKS18,Ishaq:2019}).
+(other works have explored manual vectorization and protocol mixing in ad-hoc ways, e.g.,~\cite{Ishaq:2019}).
 

### 2022-05-25 notes/notes.tex
Commit before clean up.

@@
 
+\section{Array SSA Algorithm}
+
+\begin{verbatim}
+// initialize the downwardExposed map for entryBlock with the original A's.
+// ANA: I think this will be the phi-node block when extending with nested loops
+add entryBlock to WL
+BFS(WL)
+
+BFS(worklist) // bfs traversal of loop body CFG
+  block <- remove next basic block from WL queue
+  process(block) 
+  for each block1 in succ(block), add block1 to WL if not already there.
+
+process(basicBlock) 
+  if basicBlock is a merge block 
+    B1 <- predecessor of basicBlock
+    B2 <- predecessor of basicBlock
+    for each array A
+       currentA_B1 <- downwardExposed.get(B1).get(A) 
+       currentA_B2 <- downwardExposed.get(B2).get(A) 
+       if currentA_B1 != currentA_B2 
+       // if there were modifications in either B1 or B2 will differ
+          newA <- phi(currentAB1, currentAB2) // creates phi node for array A
+          downwardExposed.get(basicBlock).put(A,newA) 
+  else // it is not a merge block
+    Block <- predecessor of basicBlock // single pred, not a merge node
+    for each array A
+       currentA <- downwardExposed.get(Block).get(A)
+       downwardExposed.get(basicBlock).put(A,currentA)
+
+  // now start the traversal of statements in basicBlock:
+  for (each stmt in statements of basicBlock)
+     case stmt is array write A[f(i)] = ...
+         newA <- create a new version of A
+         downwardExposed.get(basicBlock).put(A,newA) 
+         // have to add the new stmt
+     case stmt is array read ... = A[f(i)]
+         currentA <- downwardExposed.get(basicBlock).get(A)
+         change stmt to ... = currentA[f(i)]
+
+\end{verbatim}
+
 \end{document}

### 2022-05-25 paper.tex
Commit before clean up.

@@
 
+\ana{We need an intro here to give a roadmap of what we are going to be doing.}
+
 \begin{comment}
@@
   & \mid {\sf for} \; (i = 0; i \le \mathit{Const}; \mathit{i\!+\!+}) \; \{ \; s \; \} \\
-  & \mid {\sf if} \; (\x \; \cop \; \y) \; \{ \; s \; \} \; {\sf else} \; \{ \; s \; \} \\
+  & \mid {\sf if} \; (\x \; \cop \; \y) \; {\sf then} \; \{ \; s \; \} \; {\sf else} \; \{ \; s \; \} \\
  % \bool &::= \x \; \cop \; \y & \mathit{comparison} \\
@@
   }  else {
-    z = -1;
+    z = 2;
   }
@@
 creates a $\phi$ node at 10: $z_3 = \phi(z_0, z_1, z_2)$. There is no information retained
-that $z_0$ is triggered when $x \le 0$, $z_1$ is triggered when $x > 0 \; \wedge \; y > 0$, 
-and $z_2$ is triggered when $x > 0 \; \wedge \; y \le 0$. The Multiplexer at the join point
-needs this information however in order to assign the correct value to $z_3$ at runtime.
+that $z_0$ is triggered under path condition $x \le 0$, $z_1$ is triggered under path condition 
+$x > 0 \; \wedge \; y > 0$, and $z_2$ is triggered when $x > 0 \; \wedge \; y \le 0$. The 
+Multiplexer at the join point needs this information however in order to assign the correct 
+value to $z_3$ at runtime.
 
@@
 \begin{algorithm}
-\caption{Determine Graph edges}
+\caption{From $\phi$-nodes to MUX nodes}
 \begin{algorithmic}[1]
@@
 \ENDIF   
-     \STATE{// $Y$ is not done if $Y$ has an incomplete MUX predecessor or an original CFG predecessor \ana{Better?}}
+     \STATE{// $Y$ is not done if $Y$ has an incomplete MUX predecessor or an original CFG predecessor}
       \IF {$Y$ has $\phi$-assignments and $Y$ is not done} 
@@
 
-So far we focused on the classical SSA analysis and the extension to create MUX nodes. 
+So far we focused on the classical SSA analysis and the extension necessary to create MUX nodes. 
 Classical SSA treats arrays as scalars, i.e., an array write, e.g., {\sf A[i] = x+y} is considered
-a modification of {\sf A} and warrants its own indexed variable ${\sf A}_k$. A later use of {\sf A}, 
+a modification of {\sf A} and warrants its own indexed variable ${\sf A}_k$. A subsequent use of {\sf A}, 
 even one such as {\sf x = A[i+1]} where there is no def-use link from {\sf A[i] = x+y}, will be 
@@
 
-To mitigate against this, we propose a technique, similar to Array-SSA though much simpler and 
-less general, which breaks certain dependence edges and allows for better vectorization. 
+To mitigate against this, we propose a technique, similar to Array-SSA though much simpler, 
+which breaks certain dependence edges and allows for better vectorization. 
 \ana{Have to make a claim how 
@@
 
-\paragraph {Equivalence and Non-equivalence} Let $f(i)$ and $f'(i)$ be two index functions on induction variables $i$, 
+\paragraph {Equivalence and Non-equivalence} Let $f(i)$ and $f'(i)$ be two index functions on induction variable $i$, 
 e.g., we have array accesses ${\A}[f(i)]$ and ${\A}[f'(i)]$. We say that $f(i) \cong f'(i)$ iff for every $1 \le i \le N$,
@@
 
+\ana{This needs explanation and examples. The context is single loops where multidimensional arrays are turned 
+into one dimensional ones. The index function becomes complex.}
+
 \paragraph{Array Well-formedness Conditions} Let ${\A}$ be an array mentioned in the loop body. 
-We will treat array accesses ${\A}[f(i)]$ as distinct scalar variables if we can show that the loop body is
+We treat array accesses ${\A}[f(i)]$ as distinct scalar variables if we can show that the loop body is
 well-formed with respect to \A. Otherwise, we treat array {\A} as a single variable, thus defaulting to the 
@@
 %a unique loop exit node) on a forward path, and there is  $n_2$ on a forward path, 
-we have either $f(i-d) \cong f'(i)$ or $f(i-d) \ncong f'(i)$. \ana{FIX!}
+we have $f(i-d) \cong f'(i)$ for at most one $d$ such that $1 \le d \le N-1$ and 
+we have $f(i-d') \ncong f'(i)$ for every $d' \neq d$.
 
@@
 \ana{I need explanations of (2) and (3).}
+Now consider rule (3). If there exist a $d$ such that $f(i-d) \cong f'(i)$, that means that the array write $j: \A[f(i)] = ... $ that
+happened $d$ iterations before array read $k: ... = \A[f'(i)]$ is being used at that read. In addition, this holds for every $i$, 
+i.e., for every iteration, the array read at $k$ references the array write at $j$ that happened $d$ iterations ago. Also,
+since for every other $d'$ the array accesses are non-equivalent, this means, that the read at $k$ never references  
+the write at $j$ in a different iteration.
+%any other previous iteration, just $I-d$, where $I$ is the current iteration.
 
@@
 
-\ana{Where do we put this? Theorem: simple rule is not restrictive, and Theorem 2: program equivalent to transformed one where 
-each A[f(i)] is replaced with scalar.}
-
-\subsubsection{Constructing G}
+\subsubsection{Constructing MPC G}
 
@@
 
-\item We run \ref{alg:algorithm1} to compute SSA of the loop body with MUX nodes. Equivalent accesses get their own indexed versions.
+\item We run Algorithm~\ref{alg:algorithm1} to compute SSA of the loop body with MUX nodes. Equivalent accesses get their own indexed versions.
 
@@
 \ana{Another example, with conditional?}
+
+\ana{Here we need a theorem: program equivalent to transformed one in which every A[f(i)] is replaced with scalar. 
+Prove that original def-use graph and one with scalars are the same?}
 
@@
 Without loss of generality, we may assume that each cycle has the form $n_1 \rightarrow n_2 \rightarrow ... \rightarrow n_k \stackrel{-d}{\dasharrow} n_1$, 
-or in other words, there is a single backward edge in the entire cycle. Even though cycles with $n \ge 2$ backward 
+or in other words, there is a single backward edge in the entire cycle. Even though cycles with more than $2$ backward 
 edges are possible, such cycles can be transformed into $n$ disjoint cycles of the above form. The $-d$ on the
@@
 as we showed earlier, however, given a small SCC, one can brute force an optimal $S$. For the running example in 
-in~\figref{fig:MPCexample}(b), assuming that nodes 3 and 4 perform the same operation, we work with schedule 
-$S: \{3,4\} \rightarrow \{5\}$
+~\figref{fig:MPCexample}(b), assuming that nodes 3 and 4 perform the same operation, we work with schedule 
+$S: \{3,4\} \rightarrow \{5\}$. Importantly, due to our treatment of arrays we have the same schedule for 
+each iteration of the loop, i.e., for each iteration, we either have a dependence edge or we do not have a dependence.
 
@@
 this is the schedule across all iterations of the loop. For each $P_i$ in $S$, there are $\lceil{\frac{N}{d(SCC)}}\rceil$
-nodes in $\mathcal{A}_{\mathit{SCC}}$. $P_i(1),P_i(2),...P_i(d(SCC))$ go into node $P^1_i$, $P_i(d(SCC)+1),...P_i(2\cdot d(SCC))$ 
-go into node $P^2_i$, etc. The parenthesized number denotes the iteration, e.g., $P_i(1)$ denotes $P_i$ at the first iteration 
-of the loop. There are dependence edges as follows:
+nodes in $\mathcal{A}_{\mathit{SCC}}$:
+\begin{itemize}
+\item[] $P_i(1),P_i(2),...P_i(d(SCC))$ go into node $P^1_i$, 
+\item[] $P_i(d(SCC)+1),...P_i(2\cdot d(SCC))$ go into node $P^2_i$, etc. 
+\end{itemize}
+The parenthesized number denotes the iteration, e.g., $P_i(1)$ denotes $P_i$ at the first iteration 
+of the loop.
+
+There are dependence edges as follows:
 \begin{enumerate}
@@
 
-\item User gives an SCC schedule $SCC_1 \rightarrow SCC_2 \rightarrow ... \rightarrow SCC_m$. Again, an optimal schedule 
-is NP-complete. The final schedule is $\mathcal{A} = \mathcal{A}_{\mathit{SCC}_1} \rightarrow \mathcal{A}_{\mathit{SCC}_2} ... \rightarrow \mathcal{A}_{\mathit{SCC}_m}$.
-In our running example, there is only one possible ordering of the two components, and the final schedule is as depicted in~\figref{fig:MPCexample}(d).
-%dependence edges $P(n)_1 \dasharrow P(n)_2 ... \dasharrow P(n)_{\lceil{\frac{N}{d(n)}}\rceil}$, reflecting the across-loop dependence 
-%edges we saw earlier. The rest of the edges for every edge $n_1 \rightarrow n_2 \in G$ and every $1 \le i \le N$, add an edge 
-%$G'(n_1(i)) \rightarrow G'(n_2(i))$ to $G'$ where $G'(n(i))$ returns the $P$ node in $G'$ that contains $n(i)$.
+\item Computing an optimal schedule for the loop, even when we assume optimal schedules 
+for the SCCs, remains NP-complete. 
+One possible schedule can be computed by ordering the strongly connected components 
+in a straight line: $SCC_1 \rightarrow SCC_2 \rightarrow ... \rightarrow SCC_m$, such that
+if $SCC_i$ precedes $SCC_j$ in the SCC DAG, then $SCC_i$ precedes $SCC_j$ in the
+above straight line. The final schedule is $\mathcal{A} = \mathcal{A}_{\mathit{SCC}_1} 
+\rightarrow \mathcal{A}_{\mathit{SCC}_2} ... \rightarrow \mathcal{A}_{\mathit{SCC}_m}$.
+In our running example, there is only one possible ordering of the two components, and 
+the final schedule is as depicted in~\figref{fig:MPCexample}(d).
 
@@
 
-\ana{Have to make some kind of claim that certain \% of bench falls under this assumptions or otherwise our results don't count for much.}
-
-\subsection{Theoretical Guarantees}
+\subsection{Theoretical Guarantees on Correctness and Optimality}
 \label{sec:results}
@@
 Unfortunately, finding an optimal across-loop amortized schedule is NP-complete, just as finding an optimal loop-body schedule is.
-Our key premise is that one can still compute an optimal across-loop amortized schedule for real-world loops. Given certain 
-properties of the loop, the amortized schedule we compute in the previous section turns out to be correct and optimal.
+Our key premise is that the schedule we compute in~\secref{sec:scc} (1) improves upon the original schedule $\mathcal{S}$, and 
+(2) is an optimal across-loop amortized schedule in many cases of real-world loop examples. \ana{Talk about decomposition.}
 
@@
 
+\ana{1. Theorem: Each SCC schedule is optimal}
+
+\ana{2. Theorem: An optimal across-loop schedule (even under assumptions about optimality of individual SCC schedules) is NP-complete.}
+
+\ana{Conditions under which schedule is optimal. Exploit the fact that all SCCs are either d=1 or d=N. Scheduling SCCs in parallel and computing the schedule.
+Also, ???}
+
 %\ana{Correctness theorem should come here...}
 
-\begin{theorem}(Optimality)
-Let $G$ be a constant-rate graph that entails a DAG $D$, where the nodes of $D$ are the strongly connected components of $G$. 
-If $D$ is linear, and each $SCC \in D$ entails a loop-body schedule $S: P_1 \rightarrow P_2 ... P_k$, s.t. (1) $S$ is optimal 
-and (2) for every $P_i, i<k$, for every $n_1 \in P_i, \exists n_2 \in P_{i+1}$, s.t. $n_1 \rightarrow n_2 \in G$, then $\mathcal{A}$ is optimal.
-\end{theorem}
-
-\begin{proof} (Sketch)
-Each SCC schedule is optimal because 1) dependences are the same across the loop iterations, and 2) there is dependence between 
-iteration $i$ and iteration $i+1$, no node in $i+1$ can be computed before the last node, i.e., $P_k(i)$ has been computed, as all 
-nodes in $i+1$ depend on $P_k(i)$. \ana{Expand here.}
-\end{proof}
+%\begin{theorem}(Optimality)
+%Let $G$ be a constant-rate graph that entails a DAG $D$, where the nodes of $D$ are the strongly connected components of $G$. 
+%If $D$ is linear, and each $SCC \in D$ entails a loop-body schedule $S: P_1 \rightarrow P_2 ... P_k$, s.t. (1) $S$ is optimal 
+%and (2) for every $P_i, i<k$, for every $n_1 \in P_i, \exists n_2 \in P_{i+1}$, s.t. $n_1 \rightarrow n_2 \in G$, then $\mathcal{A}$ is optimal.
+%\end{theorem}
+
+%\begin{proof} (Sketch)
+%Each SCC schedule is optimal because 1) dependences are the same across the loop iterations, and 2) there is dependence between 
+%iteration $i$ and iteration $i+1$, no node in $i+1$ can be computed before the last node, i.e., $P_k(i)$ has been computed, as all 
+%nodes in $i+1$ depend on $P_k(i)$. \ana{Expand here.}
+%\end{proof}
 

### 2022-05-25 overleaf/introduction.tex

@@
 
-MPC theory dates back to the early 1980-ies~\cite{FOCS:Yao82b,STOC:GolMicWig87,STOC:BenGolWig88,STOC:ChaCreDam88}.
+MPC theory dates back to the early 1980-ies~\cite{Yao:1982,Goldreich:1987,Ben-Or:1988,Chaum:1988}.
 Long the realm of theoretical cryptography, MPC has seen significant advances in programming technology in recent years.
 These advances bring MPC closer to practice and wider applicability ---
-MPC technology has been employed in real-world scenarios such as auctions~\cite{FC:BCDGJK09}, biometric identification~\cite{ESORICS:BlaGas11},
-and privacy-preserving machine learning~\cite{SP:MohZha17,CCS:MohRin18}.
+MPC technology has been employed in real-world scenarios such as auctions~\cite{Bogetoft:2009}, biometric identification~\cite{Blanton:2011},
+and privacy-preserving machine learning~\cite{Mohassel:2017,Mohassel:2018}.
 The goal is to improve technology so that programmers can write \emph{secure} and \emph{efficient} programs without
@@
 The problem, therefore, is to build a high-level programming language and a compiler, and there has been significant advance in
-this space, e.g., ~\cite{CCS:BenNisPin08,ESORICS:BogLauWil08,CCS:ZhaSteBla13,Hastings:2019,Keller:2020,Acay:2021,Braun:2022} among other work.
+this space, e.g., ~\cite{Ben-David:2008,Bogdanov:2008,Zhang:2013,Hastings:2019,Keller:2020,Acay:2021,Braun:2022} among other work.
 Current research largely falls at the two ends
@@
 focuses on high-level constructs necessary to express multiple parties, computation by different parties, and information flow from one party
-to another~\cite{SP:RasHamHic14, Acay:2021}. On the other end, work on protocol implementation focuses on cryptographic foundations and their efficient circuit-level
-implementation~\cite{NDSS:DemSchZoh15, CCS:ABFKLO18, Braun:2022}, e.g., implementation of operations (e.g., MUL, ADD) using different sharing protocols
-(Boolean or Arithmetic GMW~\cite{STOC:GolMicWig87} or Yao's garbled circuits~\cite{FOCS:Yao82b}), as well as efficient share conversion from one representation
+to another~\cite{Rastogi:2014, Acay:2021}. On the other end, work on protocol implementation focuses on cryptographic foundations and their efficient circuit-level
+implementation~\cite{Demmler:2015, Araki:2018, Braun:2022}, e.g., implementation of operations (e.g., MUL, ADD) using different sharing protocols
+(Boolean or Arithmetic GMW~\cite{Goldreich:1987} or Yao's garbled circuits~\cite{Yao:1982}), as well as efficient share conversion from one representation
 to another.
@@
 As in classical compilers, we envision different front ends (e.g., our front end IMP Source) compiling into MPC Source.
-MPC Source is particularly suitable for optimizations such as protocol mixing~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022} and SIMD-vectorization,
+MPC Source is particularly suitable for optimizations such as protocol mixing~\cite{Buscher:2018b,Ishaq:2019, Fang:2022} and SIMD-vectorization,
 which takes advantage of amortization at the circuit level. The MPC Source IR exposes the \emph{linear structure} of MPC programs, which simplifies program analysis; this is
@@
 First, it demonstrates high performance~\cite{Braun:2022}.
-Second, it provides an API over efficient implementation for a wide variety of cryptographic operations (e.g., MUL, CMP, etc.) in three different protocols ---  Arithmetic GMW, Boolean GMW, and BMR --- which allows for protocol mixing~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022}, a known backend-independent optimization.
-Third, MOTION provides API for SIMD-level operations (e.g., MUL\_SIMD), which amortize cost and lead to significant improvement in memory footprint and throughput~\cite{NDSS:DemSchZoh15, CCS:ABFKLO18, Braun:2022}.
+Second, it provides an API over efficient implementation for a wide variety of cryptographic operations (e.g., MUL, CMP, etc.) in three different protocols ---  Arithmetic GMW, Boolean GMW, and BMR --- which allows for protocol mixing~\cite{Buscher:2018b,Ishaq:2019, Fang:2022}, a known backend-independent optimization.
+Third, MOTION provides API for SIMD-level operations (e.g., MUL\_SIMD), which amortize cost and lead to significant improvement in memory footprint and throughput~\cite{Demmler:2015, Araki:2018, Braun:2022}.
 It enables MPC Source-level vectorization, a key focus of this paper.
@@
 Unfortunately, we show that optimal scheduling is NP-hard via a reduction to the Shortest Common Supersequence (SCS) problem.
-Cost modeling is important as it drives not only vectorization but optimizations such as protocol mixing and scheduling as well~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022}.
+Cost modeling is important as it drives not only vectorization but optimizations such as protocol mixing and scheduling as well~\cite{Buscher:2018b,Ishaq:2019, Fang:2022}.
 
@@
 
-Our results emphasize the importance of backend-independent optimizations --- vectorization (described in this work) and protocol mixing (tackled in previous works~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022}) are two optimizations readily available at the level of MPC Source.
+Our results emphasize the importance of backend-independent optimizations --- vectorization (described in this work) and protocol mixing (tackled in previous works~\cite{Buscher:2018b,Ishaq:2019, Fang:2022}) are two optimizations readily available at the level of MPC Source.
 We believe that our work can lead to future work on backend-independent compilation and optimization, ushering new MPC optimizations and combinations of optimizations in the vein of standard compilers, and thus bringing MPC programming technology closer to practice and wider applicability.

### 2022-05-25 overleaf/overview.tex

@@
 First, MPC Source is an SSA form.
-Second, it is linear. The conditional in lines 13-14 in IMP Source turns into the linear code in lines 12-16 in MPC Source.
+Second, it is linear. The conditional in lines 13-15 in IMP Source turns into the linear code in lines 12-16 in MPC Source.
 The test turns into the CMP operation {\sf t = CMP(sum!3,min\_sum!2)}, followed by the
@@
 E.g., we see about 12x improvement in online time in GMW for $N=128$. Additionally, the non-vectorized version runs out of memory for $N=256$, 
-while the vectorized one runs with the standard maximal input size $N=4,096$.
+while the vectorized one runs with the standard maximal input size $N=4,096$ (and perhaps larger $N$).
 %\ana{Edit numbers with final experiments.}

### 2022-05-25 overleaf/prob_stmt_np_hardness.tex

@@
 Our goal is to construct a parallel schedule $P$ that reduces the program cost (when compared to cost of $S$), possibly an optimal schedule. 
-Originally we hoped that the problem is simper and computation of the optimal schedule is tractable. Unfortunately, the optimal schedule turns out to be NP-hard via a reduction 
+Originally we hoped that the problem is simper and computation of the optimal schedule was tractable. Unfortunately, the optimal schedule turns out to be NP-hard via a reduction 
 to the Shortest Common Supersequence problem.
@@
 %The problem of finding a schedule $P$ with a minimal $cost(P)$ is shown to be NP-Hard problem, as it can be reduced to the problem of finding a \emph{shortest common supersequence}, a known NP-Hard problem~\cite{Vazirani:2010}.
-The shortest common supersequence problem~\cite{Vazirani:2010} is as follows: {\it given two or more sequences find the the shortest sequence that contains all of the original sequences.} This can be solved in $O(n^k)$ time, where $n$ is the cardinality of the longest sequence and $k$ is the number of sequences.
+The shortest common supersequence problem~\cite{Vazirani:2010} is as follows: {\it given two or more sequences find the shortest sequence that contains all of the original sequences.} This can be solved in $O(n^k)$ time, where $n$ is the cardinality of the longest sequence and $k$ is the number of sequences.
 %For our problem $n$ is the maximum length of a sequence and $k$ is the number of instructions of sequences.

### 2022-05-25 overleaf/tech_report/compiler.tex

--- 2022-05-02 overleaf/compiler.tex
@@
 We describe analysis on MPC Source and backend-independent vectorization in~\secref{sec:vectorization}
-and translation into MOTION code in~\secref{sec:backend}. Full details on all parts appear in~\cite{Anon_TR}.
+and translation into MOTION code in~\secref{sec:backend}.
 %\ana{Need to state this more precisely.}
@@
 %\item \ana{Add rules for arithmetic ops}
-
-\begin{comment}
 \item The final restriction involves MUX as expressed by the rule:
@@
 \end{semantics}
-\end{comment}
-
 \end{enumerate}
-
-\begin{comment}
 The arguments of MUX are restricted to base types. 
@@
 
-%\end{comment}
-
 We note that while the programmer writes
@@
 annotations propagate through the transformations; annotations are inferred and checked with a taint analysis at the 
-level of MPC Source. The programmer annotates only inputs.
-\end{comment}
+level of MPC Source. Tthe programmer annotates only inputs.
 
@@
 
-Our compiler translates from Source to SSA as sketched below. %Full details are in the extended version~\cite{Anon_TR}.
+Our compiler translates from Source to SSA as follows:
 
@@
         Ensure that the AST matches the restricted subset defined in Section~\secref{sec:syntax}.
-        %This step outputs an instance of the \texttt{restricted\_ast.Function} class, which represents our restricted subset of the Python AST.
+        This step outputs an instance of the \texttt{restricted\_ast.Function} class, which represents our restricted subset of the Python AST.
 %  \item
@@
 
-        Convert the restricted-syntax AST to a three-address control-flow graph. The step processes for-loops, if-statements and assignments as restricted by the syntax.
-\begin{comment}
+        Convert the restricted-syntax AST to a three-address control-flow graph.
         To do this, first,
@@
         \end{enumerate}
-\end{comment}        
   %\item
         \paragraph{SSA conversion:}
-        Convert 3-address CFG to SSA with Cytron's algorithm~\cite{Cytron:1991}.
+        Convert the 3-address CFG to SSA with Cytron's algorithm.
 %\end{enumerate}
@@
 the compiler recursively converts all successor branches to MPC source and
-appends to the main MPC source.
+appends these to the main MPC source.
 

### 2022-05-25 overleaf/tech_report/prob_stmt_np_hardness.tex

--- 2022-05-02 overleaf/prob_stmt_np_hardness.tex
@@
 %$cost(S_i) = \alpha$ or $\beta$. if $S_i$ is an $A$-instruction and $cost(S_i) = \beta$ if $S_i$ is an $M$-instruction. 
-The cost of $P$ is more interesting:
+The cost of schedule $P$ is more interesting:
 
@@
 
-%The problem of finding a schedule $P$ with a minimal $cost(P)$ is shown to be NP-Hard problem, as it can be reduced to the problem of finding a \emph{shortest common supersequence}, a known NP-Hard problem~\cite{Vazirani:2010}.
-The shortest common supersequence problem~\cite{Vazirani:2010} is as follows: {\it given two or more sequences find the the shortest sequence that contains all of the original sequences.} This can be solved in $O(n^k)$ time, where $n$ is the cardinality of the longest sequence and $k$ is the number of sequences.
+The problem of finding a schedule $P$ with a minimal $cost(P)$ is shown to be NP-Hard problem, as it can be reduced to the problem of finding a \emph{shortest common supersequence}, a known NP-Hard problem~\cite{Vazirani:2010}.
+The shortest common supersequence problem is as follows: {\it given two or more sequences find the the shortest sequence that contains all of the original sequences.} This can be solved in $O(n^k)$ time, where $n$ is the cardinality of the longest sequence and $k$ is the number of sequences.
 %For our problem $n$ is the maximum length of a sequence and $k$ is the number of instructions of sequences.

### 2022-05-25 overleaf/tech_report/results.tex

--- 2022-05-02 overleaf/results.tex
@@
 The connection between these machines had 10Gbps bandwidth and sub-millisecond latency.
-%This setting reflects typical LAN use case considering that 10Gbps LAN is increasingly common in business networks and is now available even in some home networks.
+This setting reflects typical LAN use case considering that 10Gbps LAN is increasingly common in business networks and is now available even in some home networks.
 In the WAN setting we again used a {\tt c6525-25g} machine (located in Utah, US) for the first party and a {\tt c220g1} machine (located in Wisconsin, US) for the second.
 The {\tt c220g1} machine is equipped with two Intel E5-2630 8-core 2.40GHz processors and 128GB of RAM.
-We measured the connection bandwidth to be 560Mbps and average round trip time (RTT) to be 38ms.
-%At the time of this writing, all major internet providers in the US offer 1Gbps connections to home consumers, therefore this setting reasonably reflects the typical WAN use case.
+We measured the connection bandwidth between these machines to be 560Mbps and average round trip time (RTT) to be 38ms.
+At the time of this writing, all major internet providers in the US offer 1Gbps connections to home consumers, therefore this setting reasonably reflects the typical WAN use case.
 
 We run all experiments 5 times and report average values of various metrics.
-Note that standard deviation in all observations was at most 4.5\% of the mean value.
-%,  therefore the accuracy of the results is not effected by the relatively fewer runs.
+Note that standard deviation --- shown as error bar on top of the bars in the graphs --- in all observations was at most 4.5\% of the mean value, therefore the accuracy of the results is not effected by the relatively fewer runs.
 
 \subsection{Benchmarks}\label{sec:benchmarks_description}
-
-We used benchmarks from ABY and OPA~\cite{NDSS:DemSchZoh15,Ishaq:2019} (Biometric Matching, Histogram, PSI), HyCC~\cite{CCS:BDKKS18} (Database Join, Database Variance, Inner Product, Cryptonets Max Pooling, MNIST ReLU, k-means Iteration) and Parsynth~\cite{Farzan:2021} (Convex Hull, Minimal Points, Count 102, Count 10, Longest 102, Max distance between symbols).
-We did not use Gcd and Modular Exponentiation from~\cite{Ishaq:2019} as they do not benefit from vectorization according to our model. 
-We ran as many routines from~\cite{CCS:BDKKS18} and~\cite{Farzan:2021} as possible. Gaussian decomposition cannot be expressed 
-in our syntax and the compiler errors on Cryptonets Convolution at the time of writing, although we expect to fix the issue soon. 
-\cite{Anon_TR} describes what the benchmarks do.
-
-To compare we run {\em both} versions with the maximal value of $N$ for which the non-vectorized version succeeds (typically {\tt N=128}) and we run {\em vec} with typically {\tt N=4096}.
-Again, due to space constraints \cite{Anon_TR} gives the exact values.
-% In the following, we say {\em both} for an experiment in which we execute both non-vectorized and vectorized protocols and {\em vec} for the vectorized only experiment. 
-%In a preliminary experiment we ran {\tt N=2, N=4}, all the way to {\tt N=4096}; the non-vectorized version ran out of memory at the value for {\em both} and we fixed this value for these experiments.
-The vectorized one completed for {\tt N=4096} for nearly all benchmarks (numbers are not shown for space reasons).
-% however, they times are largely consistent, e.g., if {\em both} is $N = 2^k$ and it completes in $X$ seconds for vectorized, then $N = 2^{12}$ completes in $(12-k)X$ seconds).
+In the following, we say {\em both} for an experiment in which we execute both non-vectorized and vectorized protocols and {\em vec} for the vectorized only experiment. 
+In a preliminary experiment we ran {\tt N=2, N=4}, all the way to {\tt N=4096}; the non-vectorized version ran out of memory at the value {\em both} and we fixed this value for these experiments.
+The vectorized one completed for {\tt N=4096} for nearly all benchmarks (numbers are not shown for space reasons; however, they times are largely consistent, e.g., if {\em both} is $N = 2^k$ and it completes in $X$ seconds for vectorized, then $N = 2^{12}$ completes in $(12-k)X$ seconds).
+We used the following benchmarks in our evaluation:
 
@@
 
-\begin{comment}
 \begin{enumerate}
@@
 \end{enumerate}
-\end{comment}
 
 \subsection{Results and Analysis}\label{sec:result_analysis}
-{\small
+
    \begin{table*}[htbp]
@@
    \end{table*}
-}
+
 \begin{figure*}[htbp]
@@
 We show circuit evaluation times in \cref{fig:graph_all_eval_time}.
-In terms of observed benefits from vectorization, we divide benchmarks into 3 categories: 1) {\it High:} these include Convex Hull, Cryptonets Max Pooling, Minimal Points and Private Set Intersection.
-%These benchmarks are highly parallelizable and see 25x to 70x speedup in BMR, and 30x to 55x in GMW protocol. 
-2) {\it Medium:} these include Biometric Matching, DB Variance, Histogram, Inner Product, k-means Iteration and MNIST ReLU.
+In terms of amenability to vectorization, we divide benchmarks into 3 categories: 1) {\it High:} these include convex hull, cryptonets max pooling, minimal points and private set intersection.
+These benchmarks are highly parallelizable and see 25x to 70x speedup in BMR, and 30x to 55x in GMW protocol.
+2) {\it Medium:} these include biometric matching, DB Variance, histogram, inner product, k-means iteration and MNIST ReLU.
 These benchmarks have non-parallelizable phases e.g.
-the summing phase of Inner Product.
+the summing phase of inner product and biometric matching.
 Still, most computation is parallelizable and it results in speedup from 5x to 25x in BMR, and 2x to 25x in GMW protocol.
-3) {\it Low:} these include the Database Join and the regular expression benchmarks (Count 102, Count 10, Longest 102 and Max distance between symbols).
+3) {\it Low:} these include the Database Join and the regular expression benchmarks (count 102, count 10, longest 102 and max distance between symbols).
 There is less parallelizable computation in these programs, thus the speedup is lower.
@@
 In GMW, DB Join, Count 102 and Count 10s see speedup from 1.1x to 1.3x.
-However, Longest 102 and Max distance between symbols suffer a slowdown of 0.5x.
+However, longest102 and max distance between symbols suffer a slowdown of 0.5x.
 There is opportunity for vectorization in these benchmarks according to our analytical model, particularly, a there is a large EQ that is vectorized, although a large portion of the loop cannot be vectorized.
@@
 Note that in some settings it may still be desirable to vectorize e.g.
-in data constrained environments as communication as well as all other metrics, is reduced for all benchmarks.
-(The report presents plots of for all metrics~\cite{Anon_TR}.)%As shown in \cref{fig:graph_comm_size}, vectorization results in reduced communication (fewer bits are transferred).
+in data constrained environments.
+As shown in \cref{fig:graph_comm_size}, vectorization results in reduced communication (fewer bits are transferred).
 
-%We present evaluation for communication size in \cref{fig:graph_comm_size}, circuit generation time in \cref{fig:graph_circ_gen_time}, number of gates in \cref{fig:graph_total_gates} and online time and setup time in \cref{fig:graph_online_time} and \cref{fig:graph_setup_time} respectively.
+We present evaluation for communication size in \cref{fig:graph_comm_size}, circuit generation time in \cref{fig:graph_circ_gen_time}, number of gates in \cref{fig:graph_total_gates} and online time and setup time in \cref{fig:graph_online_time} and \cref{fig:graph_setup_time} respectively.
 %\ishaq{Ana: Let me know if I should add more substance.} \ana{It's good!}
@@
 \resizebox{0.7\columnwidth}{!}{\input{graphs/lan/biometric-hist-OnlineSetupTimesec}}
-\caption{Biometric Matching Circuit Evaluation Time}
+\caption{Biometric Matching Circuit Evaluation Time, x-axis lists database size}
 \label{fig:graph_biometric_eval_time}
@@
 
-\begin{comment}
 \begin{figure}[htbp]
@@
 \end{figure}
-\end{comment}
 
-We look closely at the Biometric Matching benchmark: circuit evaluation in~\cref{fig:graph_biometric_eval_time}, communication size in~\cite{Anon_TR} and circuit generation time in~\cite{Anon_TR}.
-%\cref{fig:graph_biometic_comm_size} and circuit generation time in~\cref{fig:graph_biometic_circ_gen_time}. 
+We look closely at the Biometric Matching benchmark: circuit evaluation in~\cref{fig:graph_biometric_eval_time}, communication size in \cref{fig:graph_biometic_comm_size} and circuit generation time in~\cref{fig:graph_biometic_circ_gen_time}.
 For input size beyond {\tt N=128} the memory usage exceeds available memory and prevents circuit generation.
@@
 \resizebox{0.7\columnwidth}{!}{\input{graphs/comparison-hist-OnlineSetupTimesec}}
-\caption{LAN vs. WAN: Circuit Evaluation Time}
+\caption{LAN vs. WAN: Circuit Evaluation Time Comparison}
 \label{fig:graph_comparison_eval_time}
@@
 
-\begin{comment}
 \begin{figure*}[htbp]
@@
 \end{figure*}
-\end{comment}
 
-\subsection{Comparison with MOTION-native Inner Product and Discussion}
+\subsection{Comparison with MOTION-native Inner Product}
 
@@
 however, our addition loop incurred significant cost (ADD is non-local and expensive in Boolean GMW). The MOTION-native loop ran 
-{\sf result += mult\_unsimdified[i];} while our loop ran auto generated {\sf result[i] = result\_prev[i] + mult\_unsimdified[i];} (as scalar expansion
-is an artifact of our vectorization). We rewrote the accumulation manually and that led to the comparable speedup!
+{\sf result += mult\_unsimdified[i];} while our loop generated and ran {\sf result[i] = result[i-1] + mult\_unsimdified[i];} recall that the scalar expansion
+is an artifact of our vectorization. We rewrote the accumulation (manually, for testing purposes) and that led to the comparable speedup!
 
 MOTION's compiler performs analysis that informs circuit generation and the example illustrates the power of the analysis. In the above example, 
-MOTION does the standard divide-and-conquer accumulation. We conjectured that poor performance of our loop was due to
-limitations of our implementation. Recall that Phase 3 of Vectorization, in~\secref{sec:vectorization}, which we have not implemented yet, gets rid of redundant 
-dimensions and generates {\sf if (i!=0) \{ result\_prev = result; \} result = result\_prev + mult\_unsimdified[i]}. And while it is unrealistic to expect that static analysis 
-will detect the associative accumulation in the scalar expansion code, one might expect that it will in the above slightly more complex code. 
-However, because AST analysis is difficult it does not appear to detect the accumulation.
-%(unfortunately, we ran out of time and did not experiment further).
+MOTION does the standard divide-and-conquer accumulation. The poor performance of our loop is not a limitation of our approach, but it could be a 
+limitation of our current implementation. Recall that the next phase of Vectorization, in~\secref{sec:vectorization}, which we have not implemented yet, gets rid of redundant 
+dimensions and generates {\sf if (i!=0) \{ result\_prev = result; \} result = result\_prev + mult\_unsimdified[i]}. And while it is unrealistic to expect that MOTION's static analysis 
+will detect the associative accumulation in the scalar expansion code, it is realistic to expect that it will in the above only slightly more complex code. 
+It still might lead to confusion in the static analysis as analysis on the AST is difficult (unfortunately, we ran out of time and did not experiment further).
 %Interestingly, the non-vectorized version, which is the literal translation of inner product does not appear to benefit from any optimizations.
 
-We conjecture that MPC Source, a straight-forward representation suitable for static analysis, will not only enable detection of general associative loops, but also allow 
-for program synthesis to increase opportunities for divide-and-conquer parallelization~\cite{Farzan:2021}; we leave this and other optimizations as future work.
+We conjecture that MPC Source, a straight-forward representation, will not only enable detection of general associative loops, but also allow 
+for program synthesis to increase opportunities for divide-and-conquer parallelization~\cite{Farzan:2021}; we leave this as future work.
 
-HyCC~\cite{CCS:BDKKS18} is a mixing compiler that focuses on mixed protocols, while we run vectorization within a single protocol. 
-The paper only provides data for Boolean and Yao for a version of Biometric matching on $N=1000$ as its focus is mixing. %; it does provide a lot of data on mixed protocol executions.
-We estimate we are about an order of magnitude slower, which is likely due to the same issue --- the computation of $\mathit{min}$ can be optimized.
+The most closely related work, HyCC~\cite{CCS:BDKKS18}, is a mixing compiler that focuses on mixed protocols, while we run vectorization within a single protocol. 
+The paper only provides data for Boolean and Yao for a version of Biometric matching on $N=1000$; it does provide a lot of data on mixed protocol executions.
+Again, we estimate we are about an order of magnitude slower, which is likely due to the same issue as Inner Product --- the computation of $\mathit{min}$ can be optimized.
 

### 2022-05-26 overleaf/tech_report/introduction.tex
Fixed refs and added new related work.

@@
 Earlier compilers did both back-end and front-end translation as their aim was to demonstrate applicability of MPC on real-world
-programming problems. As the field advanced, works have focused more closely on front-end language design (e.g., Wysteria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021})
+programming problems. As the field advanced, works have focused more closely on front-end language design (e.g., Wysteria~\cite{Rastogi:2014} and Viaduct~\cite{Acay:2021})
 or back-end ``circuit-level'' design and implementation (e.g., MOTION~\cite{Braun:2022}).
@@
 As in classical compilers, we envision different front ends (e.g., our front end IMP Source, other) compiling into MPC Source.
-MPC Source is particularly suitable for optimizations such as protocol mixing~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022} and SIMD-vectorization,
+MPC Source is particularly suitable for optimizations such as protocol mixing~\cite{Buscher:2018b,Ishaq:2019, Fang:2022} and SIMD-vectorization,
 which takes advantage of amortization at the circuit level. The MPC Source IR exposes the \emph{linear structure} of MPC programs, which simplifies program analysis; this is
@@
 First, it is the state of the art in terms of performance~\cite{Braun:2022}.
-Second, it provides an API over efficient implementation for a wide variety of cryptographic operations in three different protocols ---  Arithmetic GMW, Boolean GMW, and BMR --- which allows for protocol mixing~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022}, a known backend-independent optimization.
+Second, it provides an API over efficient implementation for a wide variety of cryptographic operations in three different protocols ---  Arithmetic GMW, Boolean GMW, and BMR --- which allows for protocol mixing~\cite{Buscher:2018b,Ishaq:2019, Fang:2022}, a known backend-independent optimization.
 Third, MOTION provides API for SIMD-level operations, which amortize cost and lead to significant improvement in memory footprint and
-throughput~\cite{NDSS:DemSchZoh15, CCS:ABFKLO18, Braun:2022}. It enables MPC Source-level vectorization, a key focus of this paper.
+throughput~\cite{Demmler:2015, Araki:2018, Braun:2022}. It enables MPC Source-level vectorization, a key focus of this paper.
 
@@
 Unfortunately, we show that optimal scheduling is NP-hard via a reduction to the Shortest Common Supersequence (SCS) problem.
-Cost modeling is important as it drives not only vectorization but optimizations such as protocol mixing and scheduling as well~\cite{CCS:BDKKS18,Ishaq:2019, Fang:2022}.
+Cost modeling is important as it drives not only vectorization but optimizations such as protocol mixing and scheduling as well~\cite{Buscher:2018b}.
 

### 2022-05-26 overleaf/tech_report/related.tex
Fixed refs and added new related work.

--- 2022-05-02 overleaf/related.tex
@@
 \paragraph{MPC languages and compilers.}
-Languages and compilers for secure computation have seen significant attention and advances in recent years. The early MPC compilers Fairplay~\cite{CCS:BenNisPin08},
-and Sharemind~\cite{ESORICS:BogLauWil08} were followed by PICCO~\cite{CCS:ZhaSteBla13}, Obliv-C~\cite{Zahur:2015}, TinyGarble~\cite{SP:SHSSK15},
-Wystiria~\cite{SP:RasHamHic14}, and others. A new generation of MPC compilers
-includes SPDZ/SCALE-MAMBA/MP-SPDZ~\cite{Keller:2020} and the ABY/HyCC/MOTION~\cite{NDSS:DemSchZoh15,CCS:BDKKS18,Braun:2022} frameworks.
+
+Languages and compilers for secure computation have seen significant attention and advances in recent years. The early MPC compilers Fairplay~\cite{Ben-David:2008},
+and Sharemind~\cite{Bogdanov:2008} were followed by PICCO~\cite{Zhang:2013}, Obliv-C~\cite{Zahur:2015}, TinyGarble~\cite{Songhori:2015},
+Wystiria~\cite{Rastogi:2014}, and others. A new generation of MPC compilers
+includes SPDZ/SCALE-MAMBA/MP-SPDZ~\cite{Keller:2020} and the ABY/HyCC/MOTION~\cite{Demmler:2015,Buscher:2018b,Braun:2022} frameworks.
 These two families are the state-of-the art and
-are actively developed. Another recent development is Viaduct~\cite{Acay:2021}, a functional language and compiler that supports a range of secure computation
-frameworks, including MPC and ZKP. Hastings et el. present a review of compiler frameworks~\cite{Hastings:2019}. We believe that our work is unique as 
-as a facade between the high-level language and lower-level circuit protocols.
+are actively developed. Another recent development is Viaduct, a functional language and compiler that supports a range of secure computation
+frameworks, including MPC and ZKP. Hastings et el. present a review of compiler frameworks~\cite{Hastings:2019}.
 
-\begin{comment}
 While each of these languages and compilers brings in new ideas and advances, none addresses the problem of ``circuit independent'' intermediate
@@
 of MPC operations, but again, using SIMD primitives is the responsibility of the programer. There is interest in frameworks for automatic mixing,
-e.g., \cite{CCS:BDKKS18,Ishaq:2019,Fang:2022}.
+e.g., \cite{Buscher:2018b,Ishaq:2019,Fang:2022}.
 
-Other works, e.g., Obliv-C~\cite{Zahur:2015}, Wystiria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021} focus on higher-level language design, particularly information-flow systems that
+Other works, e.g., Obliv-C~\cite{Zahur:2015}, Wystiria~\cite{Rastogi:2014} and Viaduct~\cite{Acay:2021} focus on higher-level language design, particularly information-flow systems that
 restrict flow between secure and insecure parts of the program.
-\end{comment}
+ObliVM~\cite{Liu:2015} has similar goals to ours --- it optimizes oblivious (i.e., secure) source code into an efficient garbled circuit.
+Our work complements ObliVM, in the sense that while ObliVM relies on programmer annotations such as map-reduce constructs, we automatically detect opportunities for optimization at 
+an intermediate level of representation. We have implemented vectorization, however, divide-and-conquer is a natural next step. Ozdemir et al.~\cite{Ozdemir:2020} develop CirC, an IR with  
+backends into zero-knowledge proof primitives as well as SMT primitives.
+Our work focuses on MPC, which CirC does not support yet.
+MPC Source is a higher level of representation than CirC, e.g., it does not unroll loops, thus presenting a smaller-size input program suitable for vectorization and scalable program analysis.
+Overall, we view our work and CirC as orthogonal --- we are focusing on analysis and optimization while CirC focuses on infrastructure.
+% \ana{Add oblivm here plus the CicC work.}
 
@@
 %is the separation of the higher-level ``backend-independent'' compilation and optimization and the lower-level compilation into circuits.
-\squeeze
+
 \paragraph{Classical HPC compilers.}
-Automatic vectorization is a longstanding problem in high-performance computing and
-there are thousands of works in this area reflecting over 40 years of research. We presented a vectorization
+Automatic vectorization is a longstanding problem in high-performance computing (HPC).
+There are thousands of works in this area reflecting over 40 years of research. We presented a vectorization
 algorithm for MPC Source, essentially extending classical loop vectorization \cite{Allen:1987}. In HPC vectorization, conditional control flow
@@
 The problem is that masking and selection, or more generally, handling control predicates~\cite{Benabderrahmane:2010,Karrenberg:2015},
-can lead to slowdown.
+can lead to \emph{slowdown}.
 %As a result, vectorizing compilers and polyhedral compilers tend to focus on straight-line code due to the unpredictable
@@
 
-We believe that vectorization over linear MPC Source warrants a new look, while drawing from
+We argue that vectorization over linear MPC Source is a different problem, one that warrants a new look, while drawing from
 results in HPC.
-%Since both branches of the conditional and the multiplexer \emph{always} execute, not only can we apply aggressive vectorization on linear code, but (perhaps more importantly)
-%we can also build analytical models that accurately predict execution time. These models in turn would drive optimizations such as vectorization, protocol mixing, and others.
-%Vectorization meshes in with those additional optimizations in non-trivial ways.
-%Furthermore, extensions of classical loop vectorization with array writes, arbitrary indexing, including non-affine indexing, and interaction with SSA are non-trivial
-%and present novel challenges and opportunities for contribution. 
-Polyhedral parallelization~\cite{Benabderrahmane:2010} considers a higher-level source (typically AST)
+Since both branches of the conditional and the multiplexer \emph{always} execute, not only can we apply aggressive vectorization on linear code, but (perhaps more importantly)
+we can also build analytical models that accurately predict execution time. These models in turn would drive optimizations such as vectorization, protocol mixing, and others.
+Vectorization meshes in with those additional optimizations in non-trivial ways.
+
+Furthermore, extensions of classical loop vectorization with array writes, arbitrary indexing, including non-affine indexing, and interaction with SSA are non-trivial
+and present novel challenges and opportunities for contribution. Polyhedral parallelization~\cite{Benabderrahmane:2010} considers a higher-level source (typically AST)
 representation, while our work takes advantage of linear MPC Source and SSA form. The work by Karrenberg~\cite{Karrenberg:2015} is rare in that space, in the sense that it considers

### 2022-05-26 overleaf/tech_report/results.tex
Fixed refs and added new related work.

@@
 
-The most closely related work, HyCC~\cite{CCS:BDKKS18}, is a mixing compiler that focuses on mixed protocols, while we run vectorization within a single protocol. 
-The paper only provides data for Boolean and Yao for a version of Biometric matching on $N=1000$; it does provide a lot of data on mixed protocol executions.
-Again, we estimate we are about an order of magnitude slower, which is likely due to the same issue as Inner Product --- the computation of $\mathit{min}$ can be optimized.
+%The most closely related work, HyCC~\cite{Buscher:2018}, is a mixing compiler that focuses on mixed protocols, while we run vectorization within a single protocol. 
+%The paper only provides data for Boolean and Yao for a version of Biometric matching on $N=1000$; it does provide a lot of data on mixed protocol executions.
+%Again, we estimate we are about an order of magnitude slower, which is likely due to the same issue as Inner Product --- the computation of $\mathit{min}$ can be optimized.
 

### 2022-05-26 overleaf/tech_report/vectorization.tex
Fixed refs and added new related work.

--- 2022-05-03 overleaf/vectorization.tex
@@
 structure and in part because vectorization meshes in with other MPC-specific optimizations in non-trivial ways
-(other works have explored manual vectorization and protocol mixing in ad-hoc ways, e.g.,~\cite{NDSS:DemSchZoh15,CCS:BDKKS18,Ishaq:2019}).
+(other works have explored manual vectorization and protocol mixing in an ad-hoc way, e.g.,~\cite{Demmler:2015,Buscher:2018,Ishaq:2019}).
 
-In this section, we briefly describe the analysis. Details, including edge cases and examples, can be found in the extended version~\cite{Anon_TR}.
-%\secref{sec:dependence} describes our dependence analysis and ~\secref{sec:expansion} describes scalar expansion,
-%which lifts scalars (and arrays) to the corresponding loop dimensionality to create opportunities for vectorization. 
-%\secref{sec:basic_vectorization} describes our core vectorization algorithm and~\secref{sec:correctness} argues correctness of the vectorization transformation.
-%\secref{sec:array_writes} extends vectorization with array writes.
+\secref{sec:dependence} describes our dependence analysis and ~\secref{sec:expansion} describes scalar expansion,
+which lifts scalars (and arrays) to the corresponding loop dimensionality to create opportunities for vectorization. 
+\secref{sec:basic_vectorization} describes our core vectorization algorithm and~\secref{sec:correctness} argues correctness of the vectorization transformation.
+\secref{sec:array_writes} extends vectorization with array writes.
 
@@
 \item same-level edge $X\rightarrow Y$ where $X$ and $Y$ are in the same loop nest, say $i,j,k$.
-E.g., the def-use edge 9 to 10 in the Biometric MPC Source in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(b) is a same-level edge.
-%A same-level edge can be a back-edge in which case a $\phi$ node is the target of the edge. E.g., 15 to 4 in Biometric is a same-level back-edge.
+E.g., the def-use edge 9 to 10 in the Biometric MPC Source in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source} is a same-level edge.
+A same-level edge can be a back-edge in which case a $\phi$ node is the target of the edge.
+E.g., 15 to 4 in Biometric is a same-level back-edge.
 \item outer-to-inner $X\rightarrow Y$ where $X$ is in an outer loop nest, say $i$, and $Y$ is in an inner one, say $i,j,k$. E.g., 1 to 4 in Biometric forms is an outer-to-inner edge.
 %A $\phi$-node can be a source or a target of an outer-to-inner forward edge.
-\item inner-to-outer $X\rightarrow Y$ where $X$ is in an inner loop nest, $i,j,k$, and $Y$ is in the enclosing loop nest $i,j$. E.g., the def-use from 8 to 12 gives rise to an inner-to-outer edge.
+\item inner-to-outer $X\rightarrow Y$ where $X$ is a $phi$-node in an inner loop nest, $i,j,k$, and $Y$ is in the enclosing loop nest $i,j$.
+E.g., the def-use from 8 to 12 gives rise to an inner-to-outer edge.
 %E.g. $\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$ is an inner-to-outer edge.
-%An inner-to-outer edge can be a back-edge as well, in which case both $X$ and $Y$ are $\phi$-nodes with the source $X$ in a loop nested into $Y$'s loop (not necessarily immediately).
+An inner-to-outer edge can be a back-edge as well, in which case both $X$ and $Y$ are $\phi$-nodes with the source $X$ in a loop nested into $Y$'s loop (not necessarily immediately).
 
@@
 This transforms a mixed forward edge into an "inner-to-outer" forward edge followed by an outer-to-inner forward edge.
-%Thus, Basic Vectorization handles one of "same-level", "inner-to-outer", or "outer-to-inner" def-use edges.
+Thus, Basic Vectorization handles one of "same-level", "inner-to-outer", or "outer-to-inner" def-use edges.
  \end{itemize}
@@
 
-An important component of our algorithm is expansion of scalars and arrays to the corresponding loop dimensionality, 
+An important component of our algorithm is the scalar expansion to the corresponding loop dimensionality, 
 which is necessary to expose opportunities for vectorization.
@@
 
-Expansion (and also reduction) is done with the $\mathit{raise\_dim}$ and $\mathit{drop\_dim}$ functions, 
-which we believe are standard. $\mathit{raise\_dim}$ takes the original array, the access function 
-$f(i,j,k)$ in loop nest $i,j,k$ and the loop bounds $((i\!:\!I),(j\!:\!J),(k\!:\!K))$ and produces a new 3-dimensional array 
-${\sf A'}$ by iterating over $i,j,k$ and setting each element of  ${\sf A'}$:
-\[ \mathit{raise\_dim}(A, f(i,j,k), ((i\!:\!I),(j\!:\!J),(k\!:\!K))): {\sf A'}[i,j,k] =  {\sf A}[f(i,j,k)] \]
-Raise dimension applies when reshaping input arrays and also, at outer-to-inner def-use edges.
-Analogously $\mathit{drop\_dim}$ applies at inner-to-outer def-use edges.
-It takes a higher dimensional array, say $i,j,k$ and removes trailing dimensions, say $j,k$. 
-It iterates over $i$ and takes the result at the maximal index of $j$ and $k$, i.e., the result at the last iterations of $j$ and $k$:
-\[ \mathit{drop\_dim}({\sf A}, (j\!:\!J,k\!:\!K)): {\sf A'}[i] = A[i,J-1,K-1] \] 
-%It iterates over $i$ and takes the result at the maximal index of $j$ and $k$, i.e., the result at the last iterations of $j$ and $k$:
-%\[ {\sf A'}[i] = A[i,J-1,K-1] \]
-
-\begin{comment}
 \paragraph{Arrays} Conceptually, we treat all variables as arrays. There are three kinds of arrays. 
@@
 \[ {\sf A'}[i] = A[i,J-1,K-1] \]
-\end{comment}
 
@@
 
-%\subsubsection{Algorithm}
+We present our algorithm followed by an example on Biometric.
+
+\subsubsection{Algorithm}
 
@@
 operations according to def-uses. E.g., if there is an inner-to-outer dependence, it inserts $\mathit{raise\_dim}$,
-and similarly, if there is an outer-to-inner dependence, it inserts $\mathit{drop\_dim}$. 
-For example, lines 4,5 in Biometric~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(c) reshape the input arrays. 
-After this phase, operations work on arrays of the corresponding dimensionality and we optimistically vectorize all arrays.
+and similarly, if there is an outer-to-inner dependence, it inserts $\mathit{drop\_dim}$. After this phase operations work
+on arrays of the corresponding dimensionality and we optimistically vectorize all arrays.
 
 Phase 2 proceeds from the inner-most towards the outer-most loop. For each loop it anchors dependence cycles (closures) 
-around pseudo PHI nodes then removes vectorization from the dimension of that loop. In the 
-Biometric example~\ref{tab:source_and_MPC_source_and_optimized_MPC_source}(b), the iteration on the inner loop (7-11)
-determines that $j$ is not vectorizable. The next iteration on the outer loop determines that $i$ is vectorizable. It also 
-splits the remainder of the loop anchoring each loop around the corresponding PHI node.
-
-There are two important points in this 
+around pseudo PHI nodes then removes vectorization from the dimension of that loop. There are two important points in this 
 phase. First, it may break a loop into smaller loops which would discover opportunities 
@@
 
-Phases 3 reverses scalar expansion (this is an optional phase and our current implementation does not include it).
-%and Phase 4 explicitly turns operations into MOTION SIMD operations.
-
-\paragraph{Pseudocode:}
+Phases 3 cleans up local arrays of references (this is an optional phase and our current implementation does not include it)
+and Phase 4 explicitly turns operations into MOTION SIMD operations.
 
@@
 \FOR {each closure $cl$ (after merge)}
-\STATE {create \python{for} $i_d$ \python{in} ... loop} \COMMENT{ i.e., MOTION loop }
+\STATE {create \python{for} $i_d$ \python{in} ... loop}
 \STATE {add PHI-nodes in $cl$ to header block}
@@
 
-%\COMMENT { Phase 4: }
-%\STATE {add SIMD for simdified dimensions}
+\COMMENT { Phase 4: }
+\STATE {add SIMD for simdified dimensions}
 
@@
 \input{biometric_vectorization_example.tex}
+
+\subsection{Correctness Argument}
+\label{sec:correctness}
+
+We build a correctness argument that loosely follows Abstract Interpretation.
+First we define the MPC Source syntax. The domain of MPC Source programs
+expressible in the syntax (with certain semantic restrictions) is the abstract domain $A$.
+We then define the \emph{linearization} of an MPC Source program as an interpretation
+over the syntax. The linearization, which is a \emph{schedule} (as in~\secref{sec:overview}), is the concrete domain $C$.
+Since we reason over def-use graphs in $A$ we define a partial order relation over elements of $A$
+in terms of def-use relations. We define a partial order over elements of $C$ as well,
+in terms of def-use relations in the concrete domain $C$. We prove two theorems
+that state (informally) that the schedule corresponding to the original program computes the same
+result as the schedule corresponding to the vectorized program.
 
@@
 
-\subsection{Correctness Argument}
-\label{sec:correctness}
-
-We build a correctness argument that loosely follows Abstract Interpretation.
-First we define the MPC Source syntax. The domain of MPC Source programs
-expressible in the syntax (with certain semantic restrictions) is the abstract domain $A$.
-We then define the \emph{linearization} of an MPC Source program as an interpretation
-over the syntax. The linearization, which is a \emph{schedule} (as in~\secref{sec:overview}), is the concrete domain $C$.
-Since we reason over def-use graphs in $A$ we define a partial order relation over elements of $A$
-in terms of def-use relations. We define a partial order over elements of $C$ as well,
-in terms of def-use relations in the concrete domain $C$. The theorems state 
-(informally) that the vectorized program preserves def-uses and thus computes the same 
-result as the original program.
-
-\squeeze
 \paragraph{MPC Source Syntax}
@@
 for i in range(I):
+   ...
    for k in range(K):
@@
 
-\squeeze
 \paragraph{Partial Orders}
@@
 
-\squeeze 
 \paragraph{Theorems} The two theorems arising from the Basic vectorization optimization are as follows:
@@
 
-These theorems simply state that the transformation preserves def-use relations which is an invariant of the algorithm.
-
-%Theorem 1 states that a def-use edge in MPC Source $a_0$ gives rise to a set of def-use edges in $\gamma(a_0)$
-%and a ``same'' edge in $a_1$ gives rise to a set of def-use edges in $\gamma(a_1)$ and the former set is included in the latter one.
-%This means that the computation of $\gamma(a_0)$ is carried out by $\gamma(a_1)$ as well, computing the same result. 
-%The theorem allows that $a_1$ and $\gamma(a_1)$ have larger state that $a_0$ and $\gamma(a_0)$, i.e., doing additional 
-%computation into additional memory locations. Theorem 2 establishes the correctness of vectorization --- 
-%clearly, vectorization preserves all statements and def-use edges in the original MPC Source, and therefore the linearization computes the 
-%same result as the original MPC Source. (In this case we have an equivalence, $a_0 = a_1$ according to our partial order.)
+Theorem 1 states that a def-use edge in MPC Source $a_0$ gives rise to a set of def-use edges in $\gamma(a_0)$
+and a ``same'' edge in $a_1$ gives rise to a set of def-use edges in $\gamma(a_1)$ and the former set is included in the latter one.
+This means that the computation of $\gamma(a_0)$ is carried out by $\gamma(a_1)$ as well, computing the same result. 
+The theorem allows that $a_1$ and $\gamma(a_1)$ have larger state that $a_0$ and $\gamma(a_0)$, i.e., doing additional 
+computation into additional memory locations. Theorem 2 establishes the correctness of vectorization --- 
+clearly, vectorization preserves all statements and def-use edges in the original MPC Source, and therefore the linearization computes the 
+same result as the original MPC Source. (In this case we have an equivalence, $a_0 = a_1$ according to our partial order.)
 
@@
 
-%\subsubsection{Removal of Infeasible Edges}
+\subsubsection{Removal of Infeasible Edges}
 
-Array writes limit vectorization as they sometimes introduce infeasible loop-carried dependencies. For example, consider 
-one of our benchmarks, Histogram. The source takes two arrays {\sf A} and {\sf B} of size {\sf N}, where ${\sf A}[j]$ stores the bin number and
-${\sf B}[j]$ stores the corresponding value that we need to add. The MPC Source is:
-
-{\small
-\begin{pythonn}
-for i in range(0, num_bins):
-  res1 = PHI(res, res2)
-  for j in range(0, N):
-    res2 = PHI(res1, res3)
-    tmp1 = (A[j] == i)
-    tmp2 = (res2[i] + B[j])
-    tmp3 = MUX(tmp1, res2[i], tmp2)
-    res3 = Update(res2, i, tmp3)
-return res1
-\end{pythonn}
-}
-
-There is loop-carried dependence from 8 to 2 for write array {\sf res} even though the loop can be vectorized across $i$.
-
-\begin{comment}
-Consider the classical example from~\cite{Aiken:1988}:
+Array writes limit vectorization as they sometimes introduce infeasible loop-carried dependencies. Consider the classical example from~\cite{Aiken:1988}:
 
@@
 read of {\sf ... = B\_1[i]} at 6.
-\end{comment}
 
 The following algorithm removes certain infeasible loop-carried dependencies that are due to array writes. Consider a loop with index $0 \le j < J$
-nested at $i,j,k$. 
-%Here $i$ represents the enclosing loops of $j$ and $k$ represents the enclosed loops in $j$.
-\begin{comment}
+nested at $i,j,k$. Here $i$ represents the enclosing loops of $j$ and $k$ represents the enclosed loops in $j$.
+
 \begin{algorithmic}
@@
 \end{algorithmic}
-\end{comment}
+
 Consider a loop $j$ enclosed in some fixed $\underline{i}$. Only if an update (definition) ${\sf A}_m[f(i,j,k)] = ... $ at some iteration $\underline{j}$
@@
 we may have a loop-carried dependence for ${\sf A}$ due to this def-use pair. (In contrast, Cytron's algorithm inserts a loop-carried dependency every time there is an array update.)
-The algorithm, shown in~\cite{Anon_TR}, examines all def-use pairs in loop $j$, including defs and uses in nested loops, 
-searching for values $\underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$ that satisfy
-$f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$ (using Z3). If such values exist for some def-use pair, then there is a potential
-loop-carried dependence on ${\sf A}$; otherwise there is not and we can remove the spurious backward edge to the array PHI-node 
-thus ``freeing up'' statements for vectorization. %\ana{Double check this!!!}
+The algorithm above examines all def-use pairs in loop $j$, including defs and uses in nested loops, searching for values $\underline{i}, \underline{j}, \underline{j}', \underline{k}, \underline{k}'$ that satisfy
+$f(\underline{i},\underline{j},\underline{k}) = f'(\underline{i},\underline{j}',\underline{k}')$. If such values exist for some def-use pair, then there is a potential
+loop-carried dependence on ${\sf A}$; otherwise there is not and we can remove the spurious backward edge thus ``freeing up'' statements for vectorization. %\ana{Double check this!!!}
 
-In Histogram, this removes the back edge from the update at 8 to the PHI node at 2. Clearly there does not exist $i < i'$ that makes
-$i$ at 9 and $i'$ at 6 and 7 equal. The back-edge from 9 to 4 stays because for every $j < j'$ 9 and 6 and 7 access the same location in {\sf res}. %\ana{Can't wrap my head around this...}
-
-\begin{comment}
 Consider the earlier example.
@@
 Statements 7, 8 and 9 will correctly appear in the FOR loop.
-\end{comment}
 
@@
 array write can be vectorized, we have to properly remove the phi-node replacing uses of the left-hand side of the phi-node with its arguments.
-Additional detail, as well as detailed treatment of the minor changes to $\mathit{raise\_dim}$ and $\mathit{drop\_dim}$ required to do expansion of arrays 
-in addition to scalars, appear in~\cite{Anon_TR}.
 
@@
 
-\begin{comment}
 \subsubsection{Restricting Array Writes}
@@
 
-\end{comment}
-
-\begin{comment}
 \subsubsection{Examples with Array Writes}
@@
 
-\end{comment}
-
 \begin{comment}
@@
 
-\begin{comment}
 \paragraph{Example 2} Now consider the MPC Source of Histogram:
@@
 }
-\end{comment}
 
-Returning to the histogram example, the canonical (i.e., declared) dimensionality of {\sf res} is 1.
+The canonical dimensionality of {\sf res} is 1.
 Also, the phi-node {\sf res1 = PHI(res, res2)} is a target-less phi-node (the implication being that the inner for loop can be vectorized across $i$).
@@
 In a vectorized update statement, we can ignore the incoming array, {\sf res2} in this case.
-The update writes (in parallel) all locations of the 2-d array, in this case it sets up each {\sf res3[i,j] = tmp3[i,j]}.
+The update writes (in parallel) all locations of the 2-dimensional array, in this case it sets up each {\sf res3[i,j] = tmp3[i,j]}.
 
@@
 Processing the outer loop is interesting.
-This is because the PHI node is a target-less node, and therefore, there are no closures.
+This is because the PHI node is a target-less node, and therefore, there are no closures!
+Several things happen.
 (1) Everything can be vectorized along the $i$ dimension.

### 2022-05-26 overleaf/tech_report/introduction.tex
Added link.

@@
 In our experiments vectorized code exhibits 24x improvement on average in circuit generation time for Boolean GMW (20x for BMR), 7x reduction 
-in communication (2x), 97x reduction in number of gates (91x), 4x improvement in setup time (23x) and 21x improvement in online time (18x).
+in communication (2x), 97x reduction in number of gates (91x), 4x improvement in setup time in the LAN setting (23x) and 21x improvement in online time, again in the LAN setting (18x).
 %\ana{Summarize results.}
+
+The compiler can be found at~\url{https://github.com/milana2/ParallelizationForMPC} where~\url{https://github.com/milana2/ParallelizationForMPC/tree/gh-pages} shows experiments on small input sizes and details the different stages of the compiler.
 
